### 2015-09-02 data_explore.Rmd
init commit

---
title: "Data Exploration: IUCN and AquaMaps"
author: "Jamie Afflerbach"
date: "9/2/2015"
output: html_document
---

```{r,echo=F,warning=FALSE,message=F}
library(ggplot2)
library(maps)
library(readr)
library(raster)
library(tidyr)
library(dplyr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_anx      <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
```

```{r,echo=F,warning=FALSE,message=F,results='hide'}
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells.csv'), col_types = 'idiii', progress = TRUE)

am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells.csv'), progress = TRUE)

#species list
spp_list <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)

#data for raster (used to plot species)

loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
loiczid_raster       <- raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'
```

```{r,echo=F,warning=FALSE,message=F}

### Species Map Function ###
# This function takes a single species scientific name as input, then grabs all 
# occurrence cells and associated Aquamaps probability and/or IUCN proportional area
# per cell
get_spp_map <- function(species){
  spp_id <- spp_list %>%
    filter(sciname == species & is.na(parent_sid)) %>%
    dplyr::select(am_sid, iucn_sid, sciname) %>%
    unique()
  iucn_spp_map <- iucn_spp_cells %>%
    filter(iucn_sid == spp_id$iucn_sid) %>%
    dplyr::select(-parent_sid, -subpop_sid, loiczid = iucn_loiczid) %>%
    group_by(loiczid, iucn_sid) %>%        
    summarize(iucn_area = max(iucn_area))
  ### The group_by() and summarize() are to collapse cells with overlapped polygons, fixing the duplicate 'by' issue
  am_spp_map   <- am_spp_cells %>%
    filter(am_sid == spp_id$am_sid) %>%
    rename(loiczid = am_loiczid)
  spp_map <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid') %>%
    as.data.frame()

  return(spp_map)
}
```
***

```{r,echo=F,eval=FALSE,warning=FALSE,message=F}
spp_list = spp_list%>%
              mutate(am_area = NA,
                     iucn_area=NA,
                     perc = NA,
                     am_perc=NA,
                     iucn_perc=NA)

for (i in 1:nrow(spp_list)){

  print(i)

  spp <- spp_list[i,]$sciname

  spp_map <- get_spp_map(spp)

  am_spp_map = spp_map%>%
                dplyr::select(loiczid,am_sid)%>% #add in am_prob when we want to look at probability
                  unique()

#   r_am_spp  <-  subs(loiczid_raster, 
#                      am_spp_map[ , c('loiczid', 'am_sid')], 
#                      by = 'loiczid', 
#                      which = 'am_sid', 
#                      subsWithNA = TRUE)
#   
#   spp_list[i,]$am_area = area(r_am_spp,na.rm=T)%>%
#               cellStats(.,stat='sum')
#

  iucn_map <- spp_map %>%
    mutate(iucn_pres = iucn_area > 0)%>%
      dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
        unique()

  #TODO: incorporate the iucn_area (area of cell that species overlaps with...I think thats what it is)

   if(TRUE %in% duplicated(iucn_map$loiczid))next()
#   
#   r_iucn_spp <- subs(loiczid_raster, 
#                      iucn_map[ , c('loiczid', 'iucn_pres')], 
#                     by = 'loiczid', 
#                      which = 'iucn_pres', 
#                      subsWithNA = TRUE)
#   
#   spp_list[i,]$iucn_area = area(r_iucn_spp,na.rm=T)%>%
#                 cellStats(.,stat='sum')

  # get percent overlap of map

  #iucn map cells
  cells_iucn<-iucn_map%>%
              filter(iucn_pres==TRUE)%>%
              .$loiczid
  cells_am <- am_spp_map%>%
                filter(!is.na(am_sid))%>%
                .$loiczid

  cells_total = unique(spp_map$loiczid)
  cells_overlap = intersect(cells_iucn,cells_am)

  #percent overlap
  perc = (length(cells_overlap)/length(cells_total))*100
  spp_list[i,]$perc = perc

  am_only = setdiff(cells_am,cells_iucn)
  iucn_only = setdiff(cells_iucn,cells_am)

  perc_am = (length(am_only)/length(cells_total))*100
  perc_iucn = (length(iucn_only)/length(cells_total))*100

  spp_list[i,]$am_perc = perc_am
  spp_list[i,]$iucn_perc = perc_iucn

cat(perc,perc_am,perc_iucn)

  }

# for Conus arenatus, the IUCN map gives 4 duplicate cells, but each of the entries has a different iucn_area...which doesn't make sense

#row 985 for Acanthurus leucopareius - AM is giving duplicate loiczids but with different am_prob...which is so strange. Looked
# at the raw data and it seems to be in the raw data... (JA on 8/31/15)

write.csv(spp_list,file='spp_list_w_area_percent.csv')
```

###(1) Number of species in both datasets split up by group

```{r,echo=F,warning=F,message=F,warning=FALSE,message=F}
# look at number of species per group

spp_list = read.csv('spp_list_w_area_percent.csv')

df = spp_list%>%
      mutate(diff = iucn_area - am_area)

df <- within(df,spp_group <- factor(spp_group,levels=names(sort(table(spp_group),increasing=TRUE))))

g<-ggplot(df,aes(x=spp_group))+
  geom_bar(stat='bin')+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))+ coord_flip()+
  labs(x='Species Group', y='Number of Species',title='Number of species in both AquaMaps and IUCN datasets')

print(g)
```

***

###(2) Looking at the total difference in area (km2) for each species range map in both datasets

```{r,echo=F,warning=F,message=F,warning=FALSE,message=F}

dot = ggplot(df,aes(iucn_area,am_area,color=category))+ geom_point() + geom_abline()+
      labs(x='Area (km2) from IUCN dataset',y='Area (km2) from AquaMaps dataset',title='Comparing total area of species range maps')
print(dot)
```

There are large differences so to get a better sense, here is the same comparison but with area log transformed.
````{r echo=F,warning=F,message=F,warning=FALSE,message=F}

df = df%>%
      mutate(am_log = log(am_area),
             iucn_log = log(iucn_area))

all_status <- ggplot(df,aes(iucn_log,am_log,color=category)) + 
  geom_point() + geom_abline() +
  labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
        title= 'Comparing area of range maps\n (all area values are km2 log transformed)',fill='IUCN Category')

print(all_status)
```

Data split into species groups and by IUCN category
```{r,echo=F,warning=F,message=F}

groups <- ggplot(df, aes(iucn_log, am_log,color=category)) + geom_point() +facet_wrap(~spp_group) + 
        geom_abline() + labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
                             title='Comparing differences in area across species groups', fill='IUCN Category')
print(groups)
```
***

###(3) Looking at the overlap between species range maps from IUCN and AquaMaps

This plot shows the total percent of a species range map that is (1) only covered by IUCN cells, (2) only covered by AquaMaps cells and (3) covered by both IUCN and AquaMaps cells.

####How this breaks down between AquaMaps and IUCN

Here is an example of a single species, the yellow speckled chromis (*Chromis alpha*)

```{r,echo=F,warning=FALSE,message=F}

spp = 'Chromis alpha'
map <- get_spp_map(spp)

 am_map <- map%>%
            filter(!is.na(am_prob))

    r_am_spp  <-  subs(loiczid_raster, 
                       am_map[ , c('loiczid', 'am_prob')], 
                       by = 'loiczid', 
                       which = 'am_prob', 
                       subsWithNA = TRUE)

    r_am_spp[!is.na(r_am_spp)]<-1

    iucn_map <- map %>%
                  mutate(iucn_pres = iucn_area > 0)

    r_iucn_spp <- subs(loiczid_raster, 
                       iucn_map[ , c('loiczid', 'iucn_pres')], 
                       by = 'loiczid', 
                       which = 'iucn_pres', 
                       subsWithNA = TRUE)

#plot
  plot(r_iucn_spp,col='mediumpurple',useRaster=FALSE,axes=FALSE,box=FALSE,legend=FALSE)
  plot(r_am_spp,col='orange',main="Yellow-speckled Chromis", useRaster=FALSE,add=T,axes=FALSE,box=FALSE,legend=FALSE)
  map('world', col = 'gray95', fill = T, border = 'gray80', add = TRUE)
  legend("topright",legend=c("AquaMaps","IUCN"),fill=c('orange','mediumpurple'))

```

```{r,echo=F,warning=FALSE,message=F}

  df <- gather(spp_list,perc.type,percent, perc,am_perc,iucn_perc)%>%
          dplyr::select(sciname,spp_group,perc.type,percent)%>%
            unique()

  perc_c = df%>%
            filter(sciname==spp)

  p1 = perc_c%>%dplyr::select(percent)
  paste0(round(p1[1,],2),'% of the total area in this map is covered by both IUCN and AquaMaps')
  paste0(round(p1[2,],2),'% of the total area in this map is covered just by AquaMaps')
  paste0(round(p1[3,],2),'% of the total area in this map is covered just by IUCN')
```

####Now looking across all species, here is the breakdown of how much overlap there is between the two datasets

This plot shows that the majority of species range maps have only ~30% of their full map covered by both maps, and almost no species have more than 50% of their two ranges (IUCN and AquaMaps) overlapping.

```{r,echo=F,warning=FALSE,message=F}

#look at density plot

all<-ggplot(df,aes(percent,fill=perc.type))+geom_histogram(binwidth=1,alpha=.3)+scale_fill_discrete(name="Legend",labels=c("Overlap", "AquaMaps only", "IUCN only"))+labs(x='Percent of total range', y='Number of Species')

print(all)
```

Split up by species - not sure if this is useful or not
```{r,echo=F}

sp <- ggplot(df,aes(percent,fill=perc.type))+geom_histogram(binwidth=1,alpha=.3) +facet_wrap(~spp_group) +
  scale_fill_discrete(name="Legend",labels=c("Overlap", "AquaMaps only", "IUCN only"))+labs(x='Percent of total range', y='Number of Species')

print(sp)

```

### 2015-09-02 data_explore.R
moved to .Rmd

@@ -1,169 +1,0 @@
-# Exploring aquamaps/iucn comparisons
-library(ggplot2)
-
-server_fxns.R
-
-dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-           'Darwin'  = '/Volumes/data_edit',
-           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-dir_anx      <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-
-
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells.csv'), col_types = 'idiii', progress = TRUE)
-
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells.csv'), progress = TRUE)
-
-#species list
-spp_list <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
-
-
-#data for raster (used to plot species)
-
-loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
-loiczid_raster       <- raster(loiczid_raster_file)
-names(loiczid_raster) <- 'loiczid'
-
-
-### Species Map Function ###
-# This function takes a single species scientific name as input, then grabs all 
-# occurrence cells and associated Aquamaps probability and/or IUCN proportional area
-# per cell
-get_spp_map <- function(species){
-  spp_id <- spp_list %>%
-    filter(sciname == species & is.na(parent_sid)) %>%
-    dplyr::select(am_sid, iucn_sid, sciname) %>%
-    unique()
-  iucn_spp_map <- iucn_spp_cells %>%
-    filter(iucn_sid == spp_id$iucn_sid) %>%
-    dplyr::select(-parent_sid, -subpop_sid, loiczid = iucn_loiczid) %>%
-    group_by(loiczid, iucn_sid) %>%        
-    summarize(iucn_area = max(iucn_area))
-  ### The group_by() and summarize() are to collapse cells with overlapped polygons, fixing the duplicate 'by' issue
-  am_spp_map   <- am_spp_cells %>%
-    filter(am_sid == spp_id$am_sid) %>%
-    rename(loiczid = am_loiczid)
-  spp_map <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid') %>%
-    as.data.frame()
-  
-  return(spp_map)
-}
-
-
[…]

### 2015-09-04 data_explore.Rmd
fixed incorrect math for finding smaller range overlap with larger range

@@
 title: "Data Exploration: IUCN and AquaMaps"
-author: "Jamie Afflerbach"
-date: "9/2/2015"
 output: html_document
@@
 ```{r,echo=F,warning=FALSE,message=F}
+
+# Librarires and Paths
+
 library(ggplot2)
@@
 ```{r,echo=F,warning=FALSE,message=F,results='hide'}
+
+# Data files
+
 iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells.csv'), col_types = 'idiii', progress = TRUE)
@@
 ```{r,echo=F,eval=FALSE,warning=FALSE,message=F}
+
 spp_list = spp_list%>%
@@
                      am_perc=NA,
-                     iucn_perc=NA)
+                     iucn_perc=NA,
+                     sm_perc = NA, #filled with % of smaller range that is within the larger range
+                     sm_range = NA) #what range is smaller
 
+#there's likely a better way to do this than a for loop (lapply or sapply?)
 for (i in 1:nrow(spp_list)){
@@
 
-  spp <- spp_list[i,]$sciname
+  spp <- spp_list[i,]$sciname #species name
 
-  spp_map <- get_spp_map(spp)
+  spp_map <- get_spp_map(spp) #get species map
 
+  #get just the aquamaps cells
   am_spp_map = spp_map%>%
-                dplyr::select(loiczid,am_sid)%>% #add in am_prob when we want to look at probability
+                  filter(!is.na(am_prob))%>%
+                dplyr::select(loiczid,am_sid,am_prob)%>% #add in am_prob when we want to look at probability
                   unique()
-
-#   r_am_spp  <-  subs(loiczid_raster, 
-#                      am_spp_map[ , c('loiczid', 'am_sid')], 
-#                      by = 'loiczid', 
-#                      which = 'am_sid', 
-#                      subsWithNA = TRUE)
+  #rasterize aquamaps map
+  r_am_spp  <-  subs(loiczid_raster, 
+                     am_spp_map[ , c('loiczid', 'am_sid')], 
+                     by = 'loiczid', 
+                     which = 'am_sid', 
+                     subsWithNA = TRUE)
 #   
@@
 
+  #get the iucn cells
   iucn_map <- spp_map %>%
@@
 
+  #this is incorporated because there are some duplications for loiczids with IUCN cells. I'm guessing this is because of the parent/subpopulations?
+
    if(TRUE %in% duplicated(iucn_map$loiczid))next()
 #   
-#   r_iucn_spp <- subs(loiczid_raster, 
-#                      iucn_map[ , c('loiczid', 'iucn_pres')], 
-#                     by = 'loiczid', 
-#                      which = 'iucn_pres', 
-#                      subsWithNA = TRUE)
+#  rasterize IUCN map
+  r_iucn_spp <- subs(loiczid_raster, 
+                     iucn_map[ , c('loiczid', 'iucn_pres')], 
+                    by = 'loiczid', 
+                     which = 'iucn_pres', 
+                     subsWithNA = TRUE)
 #   
@@
 
+   r_iucn_spp <- subs(loiczid_raster, 
+                      iucn_map[ , c('loiczid', 'iucn_pres')], 
+                     by = 'loiczid', 
+                      which = 'iucn_pres', 
+                      subsWithNA = TRUE)
+
   #percent overlap
@@
 
+  spp_list[i,]$sm_range = ifelse(length(cells_am)<length(cells_iucn),'AM','IUCN') #what dataset has the smaller range
+  spp_list[i,]$sm_perc = length(cells_overlap)/min(length(cells_am),length(cells_iucn))*100 #this is the total percent of smaller range encompassed in the larger range
 cat(perc,perc_am,perc_iucn)
@@
 ```
+
+Looking at the percent of smaller species range that is within the larger range.
+
+```{r,echo=F}
+
+d <- spp_list%>%
+  dplyr::select(sciname,sm_perc,sm_range)%>%
+      unique()
+
+ggplot(d,aes(sm_perc,fill=sm_range))+geom_histogram(binwidth=1,alpha=.7)+
+  scale_fill_discrete(name="Dataset with smaller range",labels=c("AquaMaps", "IUCN"))+labs(x='Percent of smalle range that falls within the larger range', y='Number of Species')
+
+```

### 2015-09-04 data_explore.Rmd
added regression analysis

@@
 spp_list = spp_list%>%
-              mutate(am_area = NA,
+              mutate(total_area=NA,
+                     am_area = NA,
                      iucn_area=NA,
@@
   #get just the aquamaps cells
-  am_spp_map = spp_map%>%
-                  filter(!is.na(am_prob))%>%
-                dplyr::select(loiczid,am_sid,am_prob)%>% #add in am_prob when we want to look at probability
-                  unique()
+#   am_spp_map = spp_map%>%
+#                   filter(!is.na(am_prob))%>%
+#                 dplyr::select(loiczid,am_sid,am_prob)%>% #add in am_prob when we want to look at probability
+#                   unique()
+
+#   if(TRUE %in% duplicated(am_spp_map$loiczid)){
+#                   am_spp_map = am_spp_map%>%
+#                                 dplyr::select(loiczid,am_sid)%>%
+#                                   unique()}
+#   
   #rasterize aquamaps map
-  r_am_spp  <-  subs(loiczid_raster, 
-                     am_spp_map[ , c('loiczid', 'am_sid')], 
-                     by = 'loiczid', 
-                     which = 'am_sid', 
-                     subsWithNA = TRUE)
+#   r_am_spp  <-  subs(loiczid_raster, 
+#                      am_spp_map[ , c('loiczid', 'am_sid')], 
+#                      by = 'loiczid', 
+#                      which = 'am_sid', 
+#                      subsWithNA = TRUE)
 #   
@@
   #get the iucn cells
-  iucn_map <- spp_map %>%
-    mutate(iucn_pres = iucn_area > 0)%>%
-      dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
-        unique()
+#   iucn_map <- spp_map %>%
+#     mutate(iucn_pres = iucn_area > 0)%>%
+#       dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
+#         unique()
 
@@
 
-   if(TRUE %in% duplicated(iucn_map$loiczid))next()
-#   
+#    if(TRUE %in% duplicated(iucn_map$loiczid))next()
+# #   
 #  rasterize IUCN map
-  r_iucn_spp <- subs(loiczid_raster, 
-                     iucn_map[ , c('loiczid', 'iucn_pres')], 
-                    by = 'loiczid', 
-                     which = 'iucn_pres', 
-                     subsWithNA = TRUE)
-#   
+#   r_iucn_spp <- subs(loiczid_raster, 
+#                      iucn_map[ , c('loiczid', 'iucn_pres')], 
+#                     by = 'loiczid', 
+#                      which = 'iucn_pres', 
+#                      subsWithNA = TRUE)
+# #   
 #   spp_list[i,]$iucn_area = area(r_iucn_spp,na.rm=T)%>%
@@
 
+  # Get total area (IUCN and AM combined)
+  allCells <- spp_map%>%
+                dplyr::select(loiczid)%>%
+                  mutate(value=1)%>%
+                  unique()
+
+  fullRange <- subs(loiczid_raster, 
+                     allCells[ , c('loiczid', 'value')], 
+                    by = 'loiczid', 
+                     which = 'value', 
+                     subsWithNA = TRUE)
+
+  spp_list[i,]$total_area = area(fullRange,na.rm=T)%>%
+                              cellStats(.,stat='sum')
+
   # get percent overlap of map
 
-  #iucn map cells
-  cells_iucn<-iucn_map%>%
-              filter(iucn_pres==TRUE)%>%
-              .$loiczid
-  cells_am <- am_spp_map%>%
-                filter(!is.na(am_sid))%>%
-                .$loiczid
-
-  cells_total = unique(spp_map$loiczid)
-  cells_overlap = intersect(cells_iucn,cells_am)
-
-   r_iucn_spp <- subs(loiczid_raster, 
-                      iucn_map[ , c('loiczid', 'iucn_pres')], 
-                     by = 'loiczid', 
-                      which = 'iucn_pres', 
-                      subsWithNA = TRUE)
-
+#   #iucn map cells
+#   cells_iucn<-iucn_map%>%
+#               filter(iucn_pres==TRUE)%>%
+#               .$loiczid
+#   cells_am <- am_spp_map%>%
+#                 filter(!is.na(am_sid))%>%
+#                 .$loiczid
+#   
+#   cells_total = unique(spp_map$loiczid)
+#   cells_overlap = intersect(cells_iucn,cells_am)
+#   
+#    r_iucn_spp <- subs(loiczid_raster, 
+#                       iucn_map[ , c('loiczid', 'iucn_pres')], 
+#                      by = 'loiczid', 
+#                       which = 'iucn_pres', 
+#                       subsWithNA = TRUE)
+#   
   #percent overlap
-  perc = (length(cells_overlap)/length(cells_total))*100
-  spp_list[i,]$perc = perc
-
-  am_only = setdiff(cells_am,cells_iucn)
-  iucn_only = setdiff(cells_iucn,cells_am)
-
-  perc_am = (length(am_only)/length(cells_total))*100
-  perc_iucn = (length(iucn_only)/length(cells_total))*100
-
-  spp_list[i,]$am_perc = perc_am
-  spp_list[i,]$iucn_perc = perc_iucn
-
-  spp_list[i,]$sm_range = ifelse(length(cells_am)<length(cells_iucn),'AM','IUCN') #what dataset has the smaller range
-  spp_list[i,]$sm_perc = length(cells_overlap)/min(length(cells_am),length(cells_iucn))*100 #this is the total percent of smaller range encompassed in the larger range
-cat(perc,perc_am,perc_iucn)
+#   perc = (length(cells_overlap)/length(cells_total))*100
+#   spp_list[i,]$perc = perc
+#   
+#   am_only = setdiff(cells_am,cells_iucn)
+#   iucn_only = setdiff(cells_iucn,cells_am)
+#   
+#   perc_am = (length(am_only)/length(cells_total))*100
+#   perc_iucn = (length(iucn_only)/length(cells_total))*100
+#   
+#   spp_list[i,]$am_perc = perc_am
+#   spp_list[i,]$iucn_perc = perc_iucn
+# 
+#   spp_list[i,]$sm_range = ifelse(length(cells_am)<length(cells_iucn),'AM','IUCN') #what dataset has the smaller range
+#   spp_list[i,]$sm_perc = length(cells_overlap)/min(length(cells_am),length(cells_iucn))*100 #this is the total percent of smaller range encompassed in the larger range
+# cat(perc,perc_am,perc_iucn)
+#
 
@@
 
+reg <- ggplot(spp_list, aes(sm_perc,total_area,color=sm_range)) + geom_point() + geom_abline() + labs(y='Total Area', x='Percent overlap smaller range in larger', title='Comparing differences in area across species groups', fill='IUCN Category')
+
+print(reg)
+
+model = lm(spp_list$total_area~spp_list$sm_perc)
+summary(model)
+
 ```

### 2015-09-04 data_explore.Rmd
added regression

@@
 
-reg <- ggplot(spp_list, aes(sm_perc,total_area,color=sm_range)) + geom_point() + geom_abline() + labs(y='Total Area', x='Percent overlap smaller range in larger', title='Comparing differences in area across species groups', fill='IUCN Category')
+reg <- ggplot(spp_list, aes(sm_perc,total_area,color=sm_range)) + geom_point() + geom_abline() + labs(y='Total Area (km2)', x='Percent overlap smaller range in larger', title='Size of species range vs.
+percent overlap of \nsmaller range with larger')
 
@@
 
-model = lm(spp_list$total_area~spp_list$sm_perc)
-summary(model)
+# remove all area greater than 100000000
+
+l = spp_list%>%filter(total_area <100000000)
+
+r <- ggplot(l, aes(sm_perc,total_area,color=sm_range)) + geom_point() + labs(y='Total Area (km2)', x='Percent overlap smaller range in larger', title='Size of species range vs.
+percent overlap of \nsmaller range with larger')
+
+print(r)
+
+#-----
+
+# Regression
+
+lm_eqn <- function(df){
+    m <- lm(total_area ~ sm_perc, spp_list);
+    eq <- substitute(italic('Total Area') == a + b %.% italic('Percent')*","~~italic(r)^2~"="~r2, 
+         list(a = format(coef(m)[1], digits = 2), 
+              b = format(coef(m)[2], digits = 2), 
+             r2 = format(summary(m)$r.squared, digits = 3)))
+    as.character(as.expression(eq));                 
+}
+
+p1 <- reg + geom_text(x = 50, y = 80000000, label = lm_eqn(spp_list), parse = TRUE)
 

### 2015-09-04 data_explore.Rmd
Tried a few things... not looking super for area correlation...

--- 2015-09-04 data_explore.Rmd
@@
 
-spp_list = spp_list%>%
-              mutate(total_area=NA,
+spp_list = spp_list %>%
+              mutate(total_area = NA,
                      am_area    = NA,
-                     iucn_area=NA,
+                     iucn_area  = NA,
                      perc       = NA,
-                     am_perc=NA,
-                     iucn_perc=NA,
+                     am_perc    = NA,
+                     iucn_perc  = NA,
                      sm_perc    = NA, #filled with % of smaller range that is within the larger range
@@
 
-  spp <- spp_list[i,]$sciname #species name
+  spp     <- spp_list[i, ]$sciname #species name
 
@@
 
-  df <- gather(spp_list,perc.type,percent, perc,am_perc,iucn_perc)%>%
-          dplyr::select(sciname,spp_group,perc.type,percent)%>%
+  df <- gather(spp_list, perc.type, percent, perc, am_perc, iucn_perc) %>%
+          dplyr::select(sciname, spp_group, perc.type, percent) %>%
             unique()
 
-  perc_c = df%>%
-            filter(sciname==spp)
+  perc_c = df %>%
+            filter(sciname == spp)
 
-  p1 = perc_c%>%dplyr::select(percent)
-  paste0(round(p1[1,],2),'% of the total area in this map is covered by both IUCN and AquaMaps')
-  paste0(round(p1[2,],2),'% of the total area in this map is covered just by AquaMaps')
-  paste0(round(p1[3,],2),'% of the total area in this map is covered just by IUCN')
+  p1 = perc_c %>% dplyr::select(percent)
+  paste0(round(p1[1, ], 2), '% of the total area in this map is covered by both IUCN and AquaMaps')
+  paste0(round(p1[2, ], 2), '% of the total area in this map is covered just by AquaMaps')
+  paste0(round(p1[3, ], 2), '% of the total area in this map is covered just by IUCN')
 ```
@@
 
-all<-ggplot(df,aes(percent,fill=perc.type))+geom_histogram(binwidth=1,alpha=.3)+scale_fill_discrete(name="Legend",labels=c("Overlap", "AquaMaps only", "IUCN only"))+labs(x='Percent of total range', y='Number of Species')
+all<-ggplot(df,aes(percent,fill=perc.type))+
+    geom_histogram(binwidth=1,alpha=.3)+
+    scale_fill_discrete(name="Legend",labels=c("Overlap", "AquaMaps only", "IUCN only"))+
+    labs(x='Percent of total range', y='Number of Species')
 
@@
 
-sp <- ggplot(df,aes(percent,fill=perc.type))+geom_histogram(binwidth=1,alpha=.3) +facet_wrap(~spp_group) +
-  scale_fill_discrete(name="Legend",labels=c("Overlap", "AquaMaps only", "IUCN only"))+labs(x='Percent of total range', y='Number of Species')
+sp <- ggplot(df,aes(percent,fill=perc.type))+geom_histogram(binwidth=1,alpha=.3) +
+  facet_wrap(~spp_group) +
+  scale_fill_discrete(name="Legend",labels=c("Overlap", "AquaMaps only", "IUCN only"))+
+  labs(x='Percent of total range', y='Number of Species')
 
@@
 
-```{r,echo=F}
+```{r pct small within large, echo=F}
 
-d <- spp_list%>%
-  dplyr::select(sciname,sm_perc,sm_range)%>%
+d <- spp_list %>%
+  dplyr::select(sciname, sm_perc, sm_range) %>%
       unique()
 
-ggplot(d,aes(sm_perc,fill=sm_range))+geom_histogram(binwidth=1,alpha=.7)+
-  scale_fill_discrete(name="Dataset with smaller range",labels=c("AquaMaps", "IUCN"))+labs(x='Percent of smalle range that falls within the larger range', y='Number of Species')
+ggplot(d, aes(sm_perc, fill = sm_range)) + 
+  geom_histogram(binwidth = 1, alpha = .7) +
+  scale_fill_discrete(name = "Dataset with smaller range", labels = c("AquaMaps", "IUCN")) + 
+  labs(x = 'Percent of smaller range that falls within the larger range', y = 'Number of Species')
 
-reg <- ggplot(spp_list, aes(sm_perc,total_area,color=sm_range)) + geom_point() + geom_abline() + labs(y='Total Area', x='Percent overlap smaller range in larger', title='Comparing differences in area across species groups', fill='IUCN Category')
+```
+
+``` {r pct overlap vs area, echo = FALSE}
+
+reg <- ggplot(spp_list, aes(total_area, sm_perc, color = sm_range)) + 
+  geom_point() + 
+  geom_abline() + 
+  labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
+       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
 
@@
 
-model = lm(spp_list$total_area~spp_list$sm_perc)
-summary(model)
+# remove all area greater than 100000000
+
+l <- spp_list %>% filter(total_area < 100000000)
+
+r <- ggplot(l, aes(total_area, sm_perc, color = sm_range)) + 
+  geom_point() + 
+  labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
+       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
+
+print(r)
+
+# Log transform area
+
+logA_vs_pct <- ggplot(spp_list, aes(log_total_area, sm_perc, color = sm_range)) + 
+  geom_point() + 
+  labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
+       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
+
+print(logA_vs_pct)
+```
+
+``` {r examine regressions, echo = FALSE}
+
+A_mdl      <- lm(total_area ~ sm_perc, spp_list)
+#                Estimate Std. Error t value Pr(>|t|)    
+#   (Intercept) -16285410    2858388  -5.697  1.4e-08 ***
+#   sm_perc        533370      41702  12.790  < 2e-16 ***
+#   Multiple R-squared:  0.07732,	Adjusted R-squared:  0.07685
+
+A_mdl_clip <- lm(total_area ~ sm_perc, l)
+#               Estimate Std. Error t value Pr(>|t|)    
+#   (Intercept)  2993939     810560   3.694 0.000227 ***
+#   sm_perc       153523      11952  12.844  < 2e-16 ***
+#   Multiple R-squared:  0.07955,	Adjusted R-squared:  0.07906 
+logA_mdl   <- lm(log_total_area ~ sm_perc, spp_list)
+#                Estimate Std. Error t value Pr(>|t|)    
+#   (Intercept) 14.775325   0.096446   153.2   <2e-16 ***
+#   sm_perc      0.016885   0.001407    12.0   <2e-16 ***
+#   Multiple R-squared:  0.0687,	Adjusted R-squared:  0.06823 
+```
+
+``` {r examine quartiles, echo = FALSE}
+
+summary(spp_list$total_area)
+#      Min.   1st Qu.    Median      Mean   3rd Qu.      Max.
+#     55760   3032000  10790000  18500000  22670000 372100000
+
+qtiles <- c(quantile(spp_list$total_area, 0.0),
+            quantile(spp_list$total_area, .25),
+            quantile(spp_list$total_area, .50),
+            quantile(spp_list$total_area, .75),
+            quantile(spp_list$total_area, 1.0))
+qtile_mdl <- list()
+for (i in 1:4) {
+  area_clip <- spp_list %>%
+    filter(total_area >= qtiles[i] & total_area < qtiles[i + 1])
+  qtile_mdl[[i]] <- lm(total_area ~ sm_perc, area_clip)
+
+  qtile_plot <- ggplot(area_clip, aes(total_area, sm_perc, color = sm_range)) + 
+  geom_point() + 
+  labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
+       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
+
+print(qtile_plot)
+}
+# First quartile regression:
+#                 Estimate Std. Error t value Pr(>|t|)    
+#     (Intercept)  1423917     101053  14.091   <2e-16 ***
+#     sm_perc         1188       1555   0.764    0.445    
+#     Multiple R-squared:  0.001198,	Adjusted R-squared:  -0.0008529 
+#
+# second quartile:
+#                 Estimate Std. Error t value Pr(>|t|)    
+#     (Intercept)  7011028     275943  25.408  < 2e-16 ***
+#     sm_perc       -17725       4228  -4.192 3.28e-05 ***
+#     Multiple R-squared:  0.03483,	Adjusted R-squared:  0.03285 
+#
+# third quartile:
+#                 Estimate Std. Error t value Pr(>|t|)    
+#     (Intercept) 12813279     507856  25.230  < 2e-16 ***
+#     sm_perc        64710       8159   7.931  1.5e-14 ***
+#     Multiple R-squared:  0.1146,	Adjusted R-squared:  0.1128 
+#
+# fourth quartile:
+#                   Estimate Std. Error t value Pr(>|t|)    
+#     (Intercept) -101736351   17259544  -5.894 7.04e-09 ***
+#     sm_perc        1911286     215216   8.881  < 2e-16 ***
+#     Multiple R-squared:  0.1399,	Adjusted R-squared:  0.1381
+
+# This avenue of exploration looks like a bust.
 
 ```
+
+``` {r, echo = FALSE}
+
+# Regression
+
+lm_eqn <- function(df){
+    m  <- lm(total_area ~ sm_perc, spp_list)
+    eq <- substitute(italic('Total Area') == a + b %.% italic('Percent')*","~~italic(r)^2~"="~r2, 
+          list(a  = format(coef(m)[1], digits = 2), 
+               b  = format(coef(m)[2], digits = 2), 
+               r2 = format(summary(m)$r.squared, digits = 3)))
+    as.character(as.expression(eq))             
+}
+
+p1 <- reg + geom_text(x = 50, y = 80000000, label = lm_eqn(spp_list), parse = TRUE)
+
+```

### 2015-09-08 data_explore.Rmd
Fixed plot titles...

@@
 
+Exploring possible correlation between area and percent coverage:
 ``` {r pct overlap vs area, echo = FALSE}
@@
 
+```
+
+Perent coverage as function of area, clipping to area below 100,000,000 km^2:
+``` {r, echo = F}
 # remove all area greater than 100000000
@@
 
+```
+
+Percent coverage as a function of log(area)
+``` {r, echo = F}
 # Log transform area
@@
   geom_point() + 
-  labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
-       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
+  labs(x = 'Log(total Area)', y = 'Percent overlap smaller range in larger', 
+       title = 'Size of species range (log) vs. percent overlap of \nsmaller range with larger')
 
@@
 
+Examining data broken into quartiles to see if smaller patterns emerge
 ``` {r examine quartiles, echo = FALSE}
@@
   labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
-       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
+       title = sprintf('Size of species range (quartile %s) vs. percent overlap of \nsmaller range with larger', i))
 

### 2015-09-10 data_explore.Rmd
adding additional chunk for aquamaps only exploration

--- 2015-09-04 data_explore.Rmd
@@
 
-# Librarires and Paths
+# Libraries and Paths
 
@@
     m <- lm(total_area ~ sm_perc, spp_list);
-    eq <- substitute(italic('Total Area') == a + b %.% italic('Percent')*","~~italic(r)^2~"="~r2, 
+    eq <- substitute(italic('Y') == a + b %.% italic('X')*","~~italic(r)^2~"="~r2, 
          list(a = format(coef(m)[1], digits = 2), 
@@
 
-p1 <- reg + geom_text(x = 50, y = 80000000, label = lm_eqn(spp_list), parse = TRUE)
+p1 <- reg + geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)
+
+print(p1)
+```
+***
+
+Looking at how many and what type (threat category) of species we are losing by setting a threshold of 0.4 on the aquamaps species.
+
+```{r, warning=F,message=F}
+
+#source mapping aquamaps.R to read in all aquamaps raw data. This takes a while so rendering will take a while. 
+source('mapping aquamaps.R')
 

### 2015-09-11 data_explore.Rmd
added aquamaps threshold exploration

@@
 library(raster)
+library(data.table)
 library(tidyr)
@@
    p1 = perc_c %>% dplyr::select(percent)
+
   paste0(round(p1[1, ], 2), '% of the total area in this map is covered by both IUCN and AquaMaps')
@@
 
-This plot shows that the majority of species range maps have only ~30% of their full map covered by both maps, and almost no species have more than 50% of their two ranges (IUCN and AquaMaps) overlapping.
+This plot shows that the majority of species range maps have only ~30% of their full map covered by both maps, and very few species have more than 50% of their two ranges (IUCN and AquaMaps) overlapping.
 
@@
 
-all<-ggplot(df,aes(percent,fill=perc.type))+
-    geom_histogram(binwidth=1,alpha=.3)+
-    scale_fill_discrete(name="Legend",labels=c("Overlap", "AquaMaps only", "IUCN only"))+
-    labs(x='Percent of total range', y='Number of Species')
+overlap = df%>%dplyr::filter(perc.type=='perc')
+
+all<-ggplot(overlap,aes(percent))+
+    geom_histogram(aes(y=..count../1926),binwidth=5,alpha=.8)+
+    labs(x='Percent of total range that is shared between both datasets', y='Proportion of Total Species')
 
@@
 
-sp <- ggplot(df,aes(percent,fill=perc.type))+geom_histogram(binwidth=1,alpha=.3) +
+sp <- ggplot(overlap,aes(percent))+geom_histogram(binwidth=5,alpha=.8) +
   facet_wrap(~spp_group) +
-  scale_fill_discrete(name="Legend",labels=c("Overlap", "AquaMaps only", "IUCN only"))+
-  labs(x='Percent of total range', y='Number of Species')
+  labs(x='Percent of total range that is shared between both datasets', y='Number of Total Species')
 
@@
 
-ggplot(d, aes(sm_perc, fill = sm_range)) + 
-  geom_histogram(binwidth = 1, alpha = .7) +
+sm_plot <- ggplot(d, aes(sm_perc, fill = sm_range)) + 
+       geom_histogram(aes(y=..count../sum(..count..)),binwidth = 4, alpha = .8) +
         scale_fill_discrete(name = "Dataset with smaller range", labels = c("AquaMaps", "IUCN")) + 
-  labs(x = 'Percent of smaller range that falls within the larger range', y = 'Number of Species')
+        labs(x = 'Percent of smaller range that falls within the larger range', y = 'Proportion of all species (n=1962)')
+
+print(sm_plot)
 
@@
 
-Exploring possible correlation between area and percent coverage:
+Exploring possible correlation between total species range area and percent overlap of smaller range within the larger:
 ``` {r pct overlap vs area, echo = FALSE}
 
-reg <- ggplot(spp_list, aes(total_area, sm_perc, color = sm_range)) + 
+reg <- ggplot(spp_list, aes(sm_perc,total_area, color = sm_range)) + 
   geom_point() + 
   geom_abline() + 
-  labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
+  labs(y = 'Total Area (km2)', x = 'Percent overlap smaller range in larger', 
        title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
@@
 ```
+
+``` {r, echo = FALSE}
+
+# Regression
+
+lm_eqn <- function(df){
+    m  <- lm(total_area~sm_perc, spp_list)
+    eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
+          list(a  = format(coef(m)[1], digits = 2), 
+               b  = format(coef(m)[2], digits = 2), 
+               r2 = format(summary(m)$r.squared, digits = 3)))
+    as.character(as.expression(eq))             
+}
+
+p1 <- reg + geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)
+
+print(p1)
+
+summary(lm(total_area~sm_perc,spp_list))
+```
+
+***
 
@@
 
-``` {r, echo = FALSE}
-
-# Regression
-
-lm_eqn <- function(df){
-    m  <- lm(total_area ~ sm_perc, spp_list)
-    eq <- substitute(italic('Total Area') == a + b %.% italic('Percent')*","~~italic(r)^2~"="~r2, 
-          list(a  = format(coef(m)[1], digits = 2), 
-               b  = format(coef(m)[2], digits = 2), 
-               r2 = format(summary(m)$r.squared, digits = 3)))
-    as.character(as.expression(eq))             
-}
-
-p1 <- reg + geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)
-
-print(p1)
-```
-***
-
 Looking at how many and what type (threat category) of species we are losing by setting a threshold of 0.4 on the aquamaps species.
 
-```{r, warning=F,message=F}
+```{r aquamaps probabilities, warning=F,message=F,cache.lazy=FALSE,cache=T,verbose=F,results=F,echo=F}
 
-#source mapping aquamaps.R to read in all aquamaps raw data. This takes a while so rendering will take a while. 
-source('mapping aquamaps.R')
+# get spp_cells raw data for aquamaps
+
+dir_N = c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
+          'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
+
+dir_aquamaps = file.path(dir_N,"git-annex/globalprep/SPP_ICO/raw/aquamaps_2014")
+
+spp_cells = fread(input=file.path(dir_aquamaps,"tables/ohi_hcaf_species_native.csv"),verbose=F,header=T,
+                  colClasses=c('NULL','character',NA,'numeric','NULL','NULL')) #1 min 49 sec
+
+setnames(spp_cells,c('SpeciesID','CsquareCode','probability'),c("species_id","csquare_code","probability"))
+
+spp_cells = as.data.frame(spp_cells)
+
+spp       = fread(input=file.path(dir_aquamaps,"tables/ohi_speciesoccursum.csv"),verbose=F,header=T)%>%
+              mutate(scientific = paste(Genus,Species,sep=" "))
+spp = as.data.frame(spp)
 
 ```
+
+```{r,warning=F,message=F, echo=F}
+
+#create dataframe that gives me all species, their total number of cells, total number of cells >=0.4 prob, differnece between the two and threat
+df = spp_cells%>%
+      group_by(species_id)%>%
+        summarize(cells_0.4 = sum(probability>=0.4),
+                  cells_t = sum(probability>=0))%>%
+          mutate(prop = cells_0.4/cells_t)%>%
+            left_join(spp,by=c('species_id'='SPECIESID'))%>%
+          select(scientific,species_id,cells_0.4,cells_t,prop,iucn_code)
+
+hist <- ggplot(df,aes(prop,fill=iucn_code))+geom_histogram(aes(y=..count../sum(..count..)),binwidth = 0.05, alpha = .7)+labs(x='Proportion of range area included when selecting a threshold of 0.4',y='Proportion of all aquamaps species (17,348 total)')
+print(hist)
+
+```

### 2015-09-18 ms_outline.Rmd
init commit

---
title: "A comparative analysis of AquaMaps and IUCN species ranges"
author: "JA, CO, BH"
output:
  html_document:
    toc: true
    theme: "spacelab"
    highlight: "haddock" 
---

##Section One: Understanding each dataset separately

####**IUCN**

**Data Type**  
Shapefiles for individual species. Spatial files include subpopulation information, presence (extinct/extant etc.) and seasonality information (breeding/migration/origin etc.).

**INSERT MAP**

**Methods**  
From [IUCN site](http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources):

> The species distribution maps, sometimes referred to as 'limits of distribution' or 'field guide' maps, aim to provide the current known distribution of the species within its native range. The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range. In most cases the range is depicted as polygons. *A polygon displaying the limits of a species distribution is essentially meant to communicate that the species probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon*.

####AquaMaps

[Site](http://www.aquamaps.org/)

Over 17,000 species

Computer model using 6 parameters (SST, salinity, temperature, depth, distance to land, sea ice cover)

***  
##Section Two: Comparison of IUCN and AquaMaps

![](./figures/sp_groups_barchart.png)
![](./figures/CompareTotalArea_AM-IUCN_scatter.png)
![](./figures/CompareTotalAreaLOG_AM-IUCN_scatter.png)
![](./figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png)
![](./figures/SpeciesMap.png)
![](./figures/Overlap-PercTotalRange.png)
![](./figures/Overlap-PercTotalRange_spGroups.png)
![](./figures/PercSmallRangeInLarger.png)
![](./figures/AreavsPercOverlap.png)
![](./figures/AreavsPercOverlap_2.png)
![](./figures/AreavsOverlap_quantile_1.png)
![](./figures/AreavsOverlap_quantile_2.png)
![](./figures/AreavsOverlap_quantile_3.png)
![](./figures/AreavsOverlap_quantile_4.png)
![](./figures/AMThreshold.png)

***  
##Section Three: Application to OHI

***

##Section Four: Implications of data choices
- could talk about MPA gap analysis
- Highlights tudies that have used one and not the other. Could this have changed the results?

***

##References

### 2015-09-18 data_explore.Rmd
added ggsave for plots throughout

@@
 
+ggsave('figures/sp_groups_barchart.png',plot=g,width=8, height=4, dpi=100)
 print(g)
+
 ```
@@
       labs(x='Area (km2) from IUCN dataset',y='Area (km2) from AquaMaps dataset',title='Comparing total area of species range maps')
+ggsave('figures/CompareTotalArea_AM-IUCN_scatter.png',plot=dot,width=8, height=4, dpi=100)
 print(dot)
@@
         title= 'Comparing area of range maps\n (all area values are km2 log transformed)',fill='IUCN Category')
-
+ggsave('figures/CompareTotalAreaLOG_AM-IUCN_scatter.png',plot=all_status,width=8,height=4,dpi=100)
 print(all_status)
@@
                              title='Comparing differences in area across species groups', fill='IUCN Category')
+ggsave('figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png',plot=groups,width=8,height=6,dpi=100)
 print(groups)
@@
     geom_histogram(aes(y=..count../1926),binwidth=5,alpha=.8)+
-    labs(x='Percent of total range that is shared between both datasets', y='Proportion of Total Species')
-
+    labs(x='% total range that is shared b/w both datasets', y='Proportion of Total Species')
+ggsave('figures/Overlap-PercTotalRange.png')
 print(all)
@@
   facet_wrap(~spp_group) +
-  labs(x='Percent of total range that is shared between both datasets', y='Number of Total Species')
+  labs(x='Percent of total range that is shared between both datasets', y='Number of Total Species')+ 
+       theme(strip.text.x = element_text(size = 8))
 
+ggsave('figures/Overlap-PercTotalRange_spGroups.png')
 print(sp)
@@
         labs(x = 'Percent of smaller range that falls within the larger range', y = 'Proportion of all species (n=1962)')
+ggsave('figures/PercSmallRangeInLarger.png',width=6,height=4)
 
@@
        title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
+
+ggsave('figures/AreavsPercOverlap.png',width=6,height=4)
 
@@
 
+ggsave('figures/CorrelationAreavsPercOverlap.png')
+
 print(p1)
@@
 
+ggsave('figures/AreavsPercOverlap_2.png',width=6,height=4)
+
 print(r)
@@
        title = 'Size of species range (log) vs. percent overlap of \nsmaller range with larger')
+
+ggsave('AreavsPercOverlap_log.png',width=6,height=4)
 
@@
        title = sprintf('Size of species range (quartile %s) vs. percent overlap of \nsmaller range with larger', i))
-
+ggsave(paste0('figures/AreavsOverlap_quantile_',i,'.png'),width=6,height=4)
 print(qtile_plot)
@@
 hist <- ggplot(df,aes(prop,fill=iucn_code))+geom_histogram(aes(y=..count../sum(..count..)),binwidth = 0.05, alpha = .7)+labs(x='Proportion of range area included when selecting a threshold of 0.4',y='Proportion of all aquamaps species (17,348 total)')
+ggsave('figures/AMThreshold.png',width=6,height=4)
 print(hist)

### 2015-10-13 data_explore.Rmd
fixed a ggsave path
adding box plots

@@
 
-ggsave('AreavsPercOverlap_log.png',width=6,height=4)
+ggsave('figures/AreavsPercOverlap_log.png',width=6,height=4)
 
@@
 ```
+
+## Changes in SPP for scenarios
+
+Comparing three scenarios to the current (v2015) global model for SPP:
+
+* Baseline is: preferring IUCN rangemaps over Aquamaps rangemaps; Aquamaps "presence" threshold is set at 40% or greater.
+* Scenario 1: preferring IUCN rangemaps over Aquamaps; AM presence threshold set at >= 1% (effectively, all occurrences).
+* Scenario 2: preferring AM maps over IUCN; AM presence >= 40%.
+* Scenario 3: preferring AM over IUCN; AM presence >= 1%.
+
+Boxplots:
+
+![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
+
+![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)

### 2015-10-13 data_explore.Rmd
fixed html formatting, fixed column names from rerun of spp_list

--- 2015-09-18 data_explore.Rmd
@@
 # Data files
-
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells.csv'), col_types = 'idiii', progress = TRUE)
+#iucn spp cells WITHOUT extinct locations of IUCN species
+iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), col_types = 'cdddd', progress = TRUE)%>%as.data.frame()%>%filter(presence!=5)
 
@@
   iucn_spp_map <- iucn_spp_cells %>%
-    filter(iucn_sid == spp_id$iucn_sid) %>%
-    dplyr::select(-parent_sid, -subpop_sid, loiczid = iucn_loiczid) %>%
-    group_by(loiczid, iucn_sid) %>%        
-    summarize(iucn_area = max(iucn_area))
+    filter(sciname == species) %>%
+    group_by(loiczid) %>%        
+    summarize(iucn_area = max(prop_area))
   ### The group_by() and summarize() are to collapse cells with overlapped polygons, fixing the duplicate 'by' issue
@@
     filter(am_sid == spp_id$am_sid) %>%
-    rename(loiczid = am_loiczid)
-  spp_map <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid') %>%
+    rename(am_prob = prob)
+  spp_map <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid')%>%
     as.data.frame()
@@
 
-  print(i)
-
   spp     <- spp_list[i, ]$sciname #species name
@@
 
-  #get just the aquamaps cells
-#   am_spp_map = spp_map%>%
-#                   filter(!is.na(am_prob))%>%
-#                 dplyr::select(loiczid,am_sid,am_prob)%>% #add in am_prob when we want to look at probability
-#                   unique()
+###get just the aquamaps cells
+   am_spp_map = spp_map%>%
+                   filter(!is.na(am_prob))%>%
+                    dplyr::select(loiczid,am_sid,am_prob)%>% #add in am_prob when we want to look at probability
+                     unique()
 
-#   if(TRUE %in% duplicated(am_spp_map$loiczid)){
-#                   am_spp_map = am_spp_map%>%
-#                                 dplyr::select(loiczid,am_sid)%>%
-#                                   unique()}
-#   
-  #rasterize aquamaps map
-#   r_am_spp  <-  subs(loiczid_raster, 
-#                      am_spp_map[ , c('loiczid', 'am_sid')], 
-#                      by = 'loiczid', 
-#                      which = 'am_sid', 
-#                      subsWithNA = TRUE)
-#   
-#   spp_list[i,]$am_area = area(r_am_spp,na.rm=T)%>%
-#               cellStats(.,stat='sum')
-#
+  if(TRUE %in% duplicated(am_spp_map$loiczid)){
+                  am_spp_map = am_spp_map%>%
+                                dplyr::select(loiczid,am_sid)%>%
+                                  unique()}
 
-  #get the iucn cells
-#   iucn_map <- spp_map %>%
-#     mutate(iucn_pres = iucn_area > 0)%>%
-#       dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
-#         unique()
+###  rasterize aquamaps map
+  r_am_spp  <-  subs(loiczid_raster, 
+                     am_spp_map[ , c('loiczid', 'am_sid')], 
+                     by = 'loiczid', 
+                     which = 'am_sid', 
+                     subsWithNA = TRUE)
+
+  spp_list[i,]$am_area = area(r_am_spp,na.rm=T)%>%
+              cellStats(.,stat='sum')
+
+###  get the iucn cells
+  iucn_map <- spp_map %>%
+    mutate(iucn_pres = iucn_area > 0)%>%
+      dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
+        unique()
 
@@
 
-  #this is incorporated because there are some duplications for loiczids with IUCN cells. I'm guessing this is because of the parent/subpopulations?
+###this is incorporated because there are some duplications for loiczids with IUCN cells. I'm guessing this is because of the parent/subpopulations?
 
-#    if(TRUE %in% duplicated(iucn_map$loiczid))next()
-# #   
-#  rasterize IUCN map
-#   r_iucn_spp <- subs(loiczid_raster, 
-#                      iucn_map[ , c('loiczid', 'iucn_pres')], 
-#                     by = 'loiczid', 
-#                      which = 'iucn_pres', 
-#                      subsWithNA = TRUE)
-# #   
-#   spp_list[i,]$iucn_area = area(r_iucn_spp,na.rm=T)%>%
-#                 cellStats(.,stat='sum')
+   if(TRUE %in% duplicated(iucn_map$loiczid))next()
 
-  # Get total area (IUCN and AM combined)
+### rasterize IUCN map
+  r_iucn_spp <- subs(loiczid_raster, 
+                     iucn_map[ , c('loiczid', 'iucn_pres')], 
+                    by = 'loiczid', 
+                     which = 'iucn_pres', 
+                     subsWithNA = TRUE)
+
+  spp_list[i,]$iucn_area = area(r_iucn_spp,na.rm=T)%>%
+                cellStats(.,stat='sum')
+
+### Get total area (IUCN and AM combined)
   allCells <- spp_map%>%
@@
 
-  # get percent overlap of map
+### get percent overlap of map
 
-#   #iucn map cells
-#   cells_iucn<-iucn_map%>%
-#               filter(iucn_pres==TRUE)%>%
-#               .$loiczid
-#   cells_am <- am_spp_map%>%
-#                 filter(!is.na(am_sid))%>%
-#                 .$loiczid
-#   
-#   cells_total = unique(spp_map$loiczid)
-#   cells_overlap = intersect(cells_iucn,cells_am)
-#   
+   #iucn map cells
+  cells_iucn<-iucn_map%>%
+              filter(iucn_pres==TRUE)%>%
+              .$loiczid
+  cells_am <- am_spp_map%>%
+                filter(!is.na(am_sid))%>%
+                .$loiczid
+
+  cells_total = unique(spp_map$loiczid)
+  cells_overlap = intersect(cells_iucn,cells_am)
+
+  #I THINK THIS IS A DUPLICATE? r_iucn_spp raster created above... (JA 10/12/15)
 #    r_iucn_spp <- subs(loiczid_raster, 
@@
   #percent overlap
-#   perc = (length(cells_overlap)/length(cells_total))*100
-#   spp_list[i,]$perc = perc
-#   
-#   am_only = setdiff(cells_am,cells_iucn)
-#   iucn_only = setdiff(cells_iucn,cells_am)
-#   
-#   perc_am = (length(am_only)/length(cells_total))*100
-#   perc_iucn = (length(iucn_only)/length(cells_total))*100
-#   
-#   spp_list[i,]$am_perc = perc_am
-#   spp_list[i,]$iucn_perc = perc_iucn
-# 
-#   spp_list[i,]$sm_range = ifelse(length(cells_am)<length(cells_iucn),'AM','IUCN') #what dataset has the smaller range
-#   spp_list[i,]$sm_perc = length(cells_overlap)/min(length(cells_am),length(cells_iucn))*100 #this is the total percent of smaller range encompassed in the larger range
-# cat(perc,perc_am,perc_iucn)
-#
+  perc = (length(cells_overlap)/length(cells_total))*100
+  spp_list[i,]$perc = perc
+
+  am_only = setdiff(cells_am,cells_iucn)
+  iucn_only = setdiff(cells_iucn,cells_am)
+
+  perc_am = (length(am_only)/length(cells_total))*100
+  perc_iucn = (length(iucn_only)/length(cells_total))*100
+
+  spp_list[i,]$am_perc = perc_am
+  spp_list[i,]$iucn_perc = perc_iucn
+
+  spp_list[i,]$sm_range = ifelse(length(cells_am)<length(cells_iucn),'AM','IUCN') #what dataset has the smaller range
+  spp_list[i,]$sm_perc = length(cells_overlap)/min(length(cells_am),length(cells_iucn))*100 #this is the total percent of smaller range encompassed in the larger range
+cat(sprintf('%s: For species %s, %.2f%% of both datasets overlap, %.2f%% is just AquaMaps and %.2f%% is just IUCN\n',i,spp,perc,perc_am,perc_iucn))
 
@@
 
-write.csv(spp_list,file='spp_list_w_area_percent.csv')
+#remove non IUCN species and log total area (for use later on)
+spp_list = spp_list%>%
+            filter(spp_group != 'hagfishes' & spp_group != 'non-homalopsids')%>%
+             mutate(log_total_area = log(total_area))
+
+write.csv(spp_list,file='spp_list_w_area_percent_noextinct.csv')
 ```
@@
 
-spp_list = read.csv('spp_list_w_area_percent.csv')
+spp_list = read.csv('spp_list_w_area_percent_noextinct.csv')
 
@@
 
-dot = ggplot(df,aes(iucn_area,am_area,color=category))+ geom_point() + geom_abline()+
+dot = ggplot(df,aes(iucn_area,am_area,color=iucn_category))+ geom_point() + geom_abline()+
       labs(x='Area (km2) from IUCN dataset',y='Area (km2) from AquaMaps dataset',title='Comparing total area of species range maps')
@@
 
-all_status <- ggplot(df,aes(iucn_log,am_log,color=category)) + 
+all_status <- ggplot(df,aes(iucn_log,am_log,color=iucn_category)) + 
   geom_point() + geom_abline() +
@@
 
-groups <- ggplot(df, aes(iucn_log, am_log,color=category)) + geom_point() +facet_wrap(~spp_group) + 
+groups <- ggplot(df, aes(iucn_log, am_log,color=iucn_category)) + geom_point() +facet_wrap(~spp_group) + 
         geom_abline() + labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
                              title='Comparing differences in area across species groups', fill='IUCN Category')
+
 ggsave('figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png',plot=groups,width=8,height=6,dpi=100)
@@
 
-  paste0(round(p1[1, ], 2), '% of the total area in this map is covered by both IUCN and AquaMaps')
-  paste0(round(p1[2, ], 2), '% of the total area in this map is covered just by AquaMaps')
-  paste0(round(p1[3, ], 2), '% of the total area in this map is covered just by IUCN')
 ```
+
+For `r spp`, `r round(pi[1], 2)`% of the total area is covered by both IUCN and AquaMaps, `r round(p1[2,],2)`% is covered just by AquaMaps and `r round(p1[3,], 2)`% is covered just by IUCN.
+
+***
 
@@
 Split up by species - not sure if this is useful or not
-```{r,echo=F}
+```{r,echo=F, message = F, warning = F}
 
@@
 
-Looking at the percent of smaller species range that is within the larger range.
+###(4) Looking at the percent of smaller species range that is within the larger range.
 
-```{r pct small within large, echo=F}
+For each range
+
+```{r pct small within large, echo=F, warning= F}
 
@@
 Exploring possible correlation between total species range area and percent overlap of smaller range within the larger:
-``` {r pct overlap vs area, echo = FALSE}
+``` {r pct overlap vs area, echo = FALSE, warning = F, message=F}
 
@@
 
-``` {r, echo = FALSE}
+``` {r, echo = FALSE, warning = F, message=F}
 
@@
 Perent coverage as function of area, clipping to area below 100,000,000 km^2:
-``` {r, echo = F}
+``` {r, echo = F, warning = F, message = F}
 # remove all area greater than 100000000
@@
 Percent coverage as a function of log(area)
-``` {r, echo = F}
+``` {r, echo = F, warning = F, message = F}
 # Log transform area
@@
 
-``` {r examine regressions, echo = FALSE}
+``` {r examine regressions, echo = FALSE, warning = F}
 
@@
 
-Examining data broken into quartiles to see if smaller patterns emerge
-``` {r examine quartiles, echo = FALSE}
+***
+
+###(5) Examining data broken into quartiles to see if smaller patterns emerge
+``` {r examine quartiles, echo = FALSE, warning = F}
 
@@
 
-qtiles <- c(quantile(spp_list$total_area, 0.0),
-            quantile(spp_list$total_area, .25),
-            quantile(spp_list$total_area, .50),
-            quantile(spp_list$total_area, .75),
-            quantile(spp_list$total_area, 1.0))
+qtiles <- c(quantile(spp_list$total_area, 0.0,na.rm=T),
+            quantile(spp_list$total_area, .25,na.rm=T),
+            quantile(spp_list$total_area, .50,na.rm=T),
+            quantile(spp_list$total_area, .75,na.rm=T),
+            quantile(spp_list$total_area, 1.0,na.rm=T))
+
 qtile_mdl <- list()
@@
 
-Looking at how many and what type (threat category) of species we are losing by setting a threshold of 0.4 on the aquamaps species.
+***
 
-```{r aquamaps probabilities, warning=F,message=F,cache.lazy=FALSE,cache=T,verbose=F,results=F,echo=F}
+###(6) Looking at how many and what type (threat category) of species we are losing by setting a threshold of 0.4 on the aquamaps species.
+
+```{r aquamaps probabilities, warning=F,message=F,cache.lazy=FALSE,cache=T,verbose=F,results=F,echo=F,include=F}
 

### 2015-10-14 ms_outline.Rmd
updating with text (section three)

@@
 ---
+
+##Main message (25 words or less):
+
+With the increasing numbers of open datasets, it is important to understand the implications of using certain data over others.
+
+##Working Abstract
 
@@
 
+The global Ocean Health Index (Halpern et al.
+2012), an index made up of 10 goals, utilizes both of these datasets to inform the  Biodiversity goal.
+The biodiversity goal is made up of two parts; Species and Habitats.
+The species subgoal uses the IUCN range map data and threat status to calculate average species status in global ocean regions around the world.
+Since IUCN only has range maps for X% of total assessed species, AquaMaps provided the data in instances where we have a threat status but no range map.
+
 ***

### 2015-10-30 ms_outline.Rmd
added text

@@
 
-With the increasing numbers of open datasets, it is important to understand the implications of using certain data over others.
+A comparison of two open source, spatial datasets and the implications of making decisions about what data to use as applied to the Ocean Health Index.
 
@@
 
+***
+
 ##Section One: Understanding each dataset separately
 
-####**IUCN**
+![](./figures/SpeciesMap.png)
+
+**Figure 1:** Species range map for *Chromis alpha* from the AquaMaps (orange) and IUCN (purple) datasets.
+
+###IUCN
 
@@
 Shapefiles for individual species. Spatial files include subpopulation information, presence (extinct/extant etc.) and seasonality information (breeding/migration/origin etc.).
-
-**INSERT MAP**
 
@@
 
-####AquaMaps
+###AquaMaps
 
@@
 
-Over 17,000 species
+AquaMaps (Kirschner et al.
+2013) was developed to create species range maps based on a computer model that takes up to six parameters to determine where marine species are located.
+The model uses known species preferences for sea surface temperature, salinity, depth, productivity, and in some instances distance to land and sea ice cover to predict species ranges.
+Range maps are based on a 0.5 degree global grid where values in each cell is equal to the probability of occurrence that the species is there.
+These maps can then be tweaked by experts altering the parameters and regenerating the modeled map.
+As of October 2015, 17,348 species have been mapped.
+See (Kirschner et al.
+2013) for more detailed methods.
 
-Computer model using 6 parameters (SST, salinity, temperature, depth, distance to land, sea ice cover)
+The main differences between these two sources of marine species ranges is how the maps are developed, and their output.
+When a species range is developed during an IUCN assessment, experts determine the boundaries of the species range based on their expertise which could be using known occurrences, habitat preferences, suitable habitat, or other knowledge.
+The range is depicted as a polygon where the boundaries are *"meant to communicate that the species probably occurs within this polygon, but it does not mean that it is distributed equally within the polygon or occurs everywhere within the polygon."*
+
+These two clearly differ in one aspect - AquaMaps provides higher resolution information about where species do and do not occur globally whereas the IUCN range maps are unable to provide specific information about where within a polygon a species occurs.
+This is understandable given the mobile nature of many marine species and the difficulty in observing and tracking known occurrences.
 
@@
 ![](./figures/sp_groups_barchart.png)
-![](./figures/CompareTotalArea_AM-IUCN_scatter.png)
 ![](./figures/CompareTotalAreaLOG_AM-IUCN_scatter.png)
 ![](./figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png)
-![](./figures/SpeciesMap.png)
 ![](./figures/Overlap-PercTotalRange.png)
@@
 ![](./figures/PercSmallRangeInLarger.png)
-![](./figures/AreavsPercOverlap.png)
 ![](./figures/AreavsPercOverlap_2.png)
-![](./figures/AreavsOverlap_quantile_1.png)
-![](./figures/AreavsOverlap_quantile_2.png)
-![](./figures/AreavsOverlap_quantile_3.png)
-![](./figures/AreavsOverlap_quantile_4.png)
 ![](./figures/AMThreshold.png)
 
+Figure 1 is an example of very large differences in a species range map from the IUCN and AquaMaps datasets.
+By considering the entire range of the species to be every cell that is in either dataset, only 3.14% of the map is in both.
+3.18% is covered just by AquaMaps and 77.3% is covered just by IUCN.
+Figure 2
+
+This is not always the case as some species range maps align better, especially for marine mammals.
+
 ***
+
 ##Section Three: Application to OHI
@@
 The species subgoal uses the IUCN range map data and threat status to calculate average species status in global ocean regions around the world.
-Since IUCN only has range maps for X% of total assessed species, AquaMaps provided the data in instances where we have a threat status but no range map.
+Spatial range maps are developed by experts for species that are assessed.
+The assessment process happens in phases and only [X] number of marine species have been given spatial range maps.
+AquaMaps has spatial range maps for over 17,000 species.
+For those species that are missing spatial information from IUCN, AquaMaps data was used.
+If a species was assigned an IUCN threat status but did not have spatial range information in either dataset, it was excluded.
+
+This decision to use IUCN data preferentially to AquaMaps data was made based on the fact that experts are developing the IUCN maps and AquaMaps data is based off a computer model, which in some cases is then consulted by an expert and appropriately altered.
+
+Between the two databases, there are 1,923 species that have a range map in both.
+Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
+
+![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
+
+![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)
+
+Scenario 1 shows change in score if IUCN maps were still preferentially used over AquaMaps, but the AquaMaps threshold was set to 1% rather than 40% probability of occurrence.
+Scenario 2 shows the change in score if we switched preferences for AquaMaps over IUCN data while keeping the 40% probability of occurrence threshold.
+The final Scenario is the same as scenario 2 but without a probability of occurrence threshold.
+
+It is important to remember that although only [X (1,923/total number of species)]% of all species included in the species goal are found in both datasets, there are still significant differences when electing to choose one dataset over the other.
 

### 2015-10-30 ms_outline.Rmd
added text

--- 2015-10-14 ms_outline.Rmd
@@
 
-##Main message (25 words or less):
+##Main message:
 
-With the increasing numbers of open datasets, it is important to understand the implications of using certain data over others.
+Understanding the differences in marine species range maps from IUCN and AquaMaps, and the implications of using one or both.
 
 ##Working Abstract
+
+Species range maps are integral for spatial analyses across multiple disciplines including geography, conservation, ecology and policy.
+Two main sources for marine species are the IUCN and AquaMaps.
+These two datasets differ in development and outcomes, and therefore it is pertinent to understand the implications of using one over the other.
+AquaMaps uses a computer model based on preferential environmental parameters to predict species locations and probability of occurrence.
+IUCN uses expert insight to draw polygons around known locations of marine species.
+Approximately 1,900 species are found in both datasets.
+This comparative analysis draws attention to the similarities and differences between the two datasets, and highlights the implications of using one or the other in the context of the Ocean Health Index.
 
@@
 Computer model using 6 parameters (SST, salinity, temperature, depth, distance to land, sea ice cover)
+
+AquaMaps uses both occurrence data and environmental parameters to predict relative species occurrence.
+This approach attempts to reduce sampling bias and fill gaps where data is not available.
+Six environmental predictors are used to estimate where a species could be found; sea surface temperature, bottom depth, salinity, primary production, sea ice concentration and distance to land.
+For more information on methods see Kaschner et al.
+(2007) and online at http://www.aquamaps.org/main/AboutAquaMaps.php.
+
+There are inherent issues with AquaMaps. It can't be expected to accurately predict all marine species ranges. Expert insight is sought and incorporated when available.
 

### 2015-11-13 data_explore.Rmd
generating new plots for document: amount of overlap vs ratio of area, etc

@@
 
-```{r,echo=F,warning=FALSE,message=F}
+```{r, echo = FALSE, warning = FALSE, message = FALSE}
 
@@
 
-```{r,echo=F,warning=FALSE,message=F,results='hide'}
+```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
 
@@
 #iucn spp cells WITHOUT extinct locations of IUCN species
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), col_types = 'cdddd', progress = TRUE)%>%as.data.frame()%>%filter(presence!=5)
+iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), 
+                           col_types = 'cdddd', progress = TRUE) %>%
+  as.data.frame() %>%
+  filter(presence != 5)
 
@@
 
-```{r,echo=F,warning=FALSE,message=F}
+```{r species_map_function, echo = FALSE, warning = FALSE, message = FALSE}
 
@@
 
-```{r,echo=F,eval=FALSE,warning=FALSE,message=F}
+```{r create_species_list_w_area, echo = FALSE, eval = FALSE, warning = FALSE, message = FALSE}
 
-spp_list = spp_list %>%
+spp_list <- spp_list %>%
               mutate(total_area = NA,
@@
 for (i in 1:nrow(spp_list)){
-
   spp     <- spp_list[i, ]$sciname #species name
-
   spp_map <- get_spp_map(spp) #get species map
@@
 ###get just the aquamaps cells
-   am_spp_map = spp_map%>%
+   am_spp_map <- spp_map%>%
                    filter(!is.na(am_prob))%>%
-                    dplyr::select(loiczid,am_sid,am_prob)%>% #add in am_prob when we want to look at probability
+                    dplyr::select(loiczid, am_sid, am_prob)%>% #add in am_prob when we want to look at probability
                      unique()
 
-  if(TRUE %in% duplicated(am_spp_map$loiczid)){
-                  am_spp_map = am_spp_map%>%
-                                dplyr::select(loiczid,am_sid)%>%
-                                  unique()}
+  if(any(duplicated(am_spp_map$loiczid))){
+    am_spp_map <- am_spp_map %>%
+      dplyr::select(loiczid, am_sid) %>% 
+      unique()
+    }
 
@@
 
-  spp_list[i,]$am_area = area(r_am_spp,na.rm=T)%>%
-              cellStats(.,stat='sum')
+  spp_list[i, ]$am_area <- area(r_am_spp, na.rm = TRUE)%>%
+              cellStats(., stat='sum')
 
@@
     mutate(iucn_pres = iucn_area > 0)%>%
-      dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
+      dplyr::select(loiczid, iucn_area, am_sid, am_prob, iucn_pres)%>%
         unique()
@@
 
-  spp_list[i,]$iucn_area = area(r_iucn_spp,na.rm=T)%>%
-                cellStats(.,stat='sum')
+  spp_list[i, ]$iucn_area <- area(r_iucn_spp, na.rm = TRUE)%>%
+                cellStats(., stat='sum')
 
@@
 
-  spp_list[i,]$total_area = area(fullRange,na.rm=T)%>%
-                              cellStats(.,stat='sum')
+  spp_list[i, ]$total_area <- area(fullRange, na.rm = TRUE)%>%
+                              cellStats(., stat='sum')
 
@@
   cells_iucn<-iucn_map%>%
-              filter(iucn_pres==TRUE)%>%
+              filter(iucn_pres == TRUE)%>%
               .$loiczid
@@
 
-  cells_total = unique(spp_map$loiczid)
-  cells_overlap = intersect(cells_iucn,cells_am)
+  cells_total <- unique(spp_map$loiczid)
+  cells_overlap <- intersect(cells_iucn, cells_am)
 
@@
   #percent overlap
-  perc = (length(cells_overlap)/length(cells_total))*100
-  spp_list[i,]$perc = perc
+  perc <- (length(cells_overlap) / length(cells_total)) * 100
+  spp_list[i, ]$perc <- perc
 
-  am_only = setdiff(cells_am,cells_iucn)
-  iucn_only = setdiff(cells_iucn,cells_am)
+  am_only <- setdiff(cells_am, cells_iucn)
+  iucn_only <- setdiff(cells_iucn, cells_am)
 
-  perc_am = (length(am_only)/length(cells_total))*100
-  perc_iucn = (length(iucn_only)/length(cells_total))*100
+  perc_am <- (length(am_only) / length(cells_total)) * 100
+  perc_iucn <- (length(iucn_only) / length(cells_total)) * 100
 
-  spp_list[i,]$am_perc = perc_am
-  spp_list[i,]$iucn_perc = perc_iucn
+  spp_list[i, ]$am_perc <- perc_am
+  spp_list[i, ]$iucn_perc <- perc_iucn
 
-  spp_list[i,]$sm_range = ifelse(length(cells_am)<length(cells_iucn),'AM','IUCN') #what dataset has the smaller range
-  spp_list[i,]$sm_perc = length(cells_overlap)/min(length(cells_am),length(cells_iucn))*100 #this is the total percent of smaller range encompassed in the larger range
-cat(sprintf('%s: For species %s, %.2f%% of both datasets overlap, %.2f%% is just AquaMaps and %.2f%% is just IUCN\n',i,spp,perc,perc_am,perc_iucn))
+  spp_list[i, ]$sm_range <- ifelse(length(cells_am)<length(cells_iucn), 'AM', 'IUCN') #what dataset has the smaller range
+  spp_list[i, ]$sm_perc <- length(cells_overlap)/min(length(cells_am), length(cells_iucn)) * 100 #this is the total percent of smaller range encompassed in the larger range
+cat(sprintf('%s: For species %s, %.2f%% of both datasets overlap, %.2f%% is just AquaMaps and %.2f%% is just IUCN\n', i, spp, perc, perc_am, perc_iucn))
 
@@
 #remove non IUCN species and log total area (for use later on)
-spp_list = spp_list%>%
+spp_list <- spp_list%>%
             filter(spp_group != 'hagfishes' & spp_group != 'non-homalopsids')%>%
@@
 
-write.csv(spp_list,file='spp_list_w_area_percent_noextinct.csv')
+write.csv(spp_list, file='spp_list_w_area_percent_noextinct.csv')
 ```
@@
 
-```{r,echo=F,warning=F,message=F,warning=FALSE,message=F}
+```{r, echo = FALSE, warning = FALSE, message = FALSE}
 # look at number of species per group
 
-spp_list = read.csv('spp_list_w_area_percent_noextinct.csv')
+spp_list <- read.csv('spp_list_w_area_percent_noextinct.csv')
 
-df = spp_list%>%
+spp_df <- spp_list %>%
   mutate(diff = iucn_area - am_area)
 
-df <- within(df,spp_group <- factor(spp_group,levels=names(sort(table(spp_group),increasing=TRUE))))
+spp_df <- within(spp_df, spp_group <- factor(spp_group, levels = names(sort(table(spp_group), increasing = TRUE))))
 
-g<-ggplot(df,aes(x=spp_group))+
-  geom_bar(stat='bin')+
-  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))+ coord_flip()+
-  labs(x='Species Group', y='Number of Species',title='Number of species in both AquaMaps and IUCN datasets')
+barchart_sp_groups <- ggplot(spp_df, aes(x = spp_group)) +
+  geom_bar(stat = 'bin') +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + 
+  coord_flip() +
+  labs(x = 'Species Group', 
+       y = 'Number of Species', 
+       title = 'Number of species in both AquaMaps and IUCN datasets')
 
-ggsave('figures/sp_groups_barchart.png',plot=g,width=8, height=4, dpi=100)
-print(g)
+ggsave('figures/sp_groups_barchart.png', 
+       plot = barchart_sp_groups, 
+       width = 8, height = 4, dpi = 100)
+
+print(barchart_sp_groups)
 
@@
 
-```{r,echo=F,warning=F,message=F,warning=FALSE,message=F}
+```{r, echo = FALSE, warning = FALSE, message = FALSE, warning = FALSE, message = FALSE}
 
-dot = ggplot(df,aes(iucn_area,am_area,color=iucn_category))+ geom_point() + geom_abline()+
-      labs(x='Area (km2) from IUCN dataset',y='Area (km2) from AquaMaps dataset',title='Comparing total area of species range maps')
-ggsave('figures/CompareTotalArea_AM-IUCN_scatter.png',plot=dot,width=8, height=4, dpi=100)
-print(dot)
+scatter_am_iucn_area <- ggplot(spp_df, aes(iucn_area, am_area, color = iucn_category)) + 
+  geom_point() + 
+  geom_abline(slope = 1, intercept = 0, color = 'red') +
+  labs(x = 'Area (km2) from IUCN dataset', 
+       y = 'Area (km2) from AquaMaps dataset', 
+       title = 'Comparing total area of species range maps')
+
+ggsave('figures/CompareTotalArea_AM-IUCN_scatter.png', 
+       plot = scatter_am_iucn_area, 
+       width = 8, height = 4, dpi = 100)
+
+print(scatter_am_iucn_area)
 ```
@@
 There are large differences so to get a better sense, here is the same comparison but with area log transformed.
-````{r echo=F,warning=F,message=F,warning=FALSE,message=F}
+``` {r echo = FALSE, warning = FALSE, message = FALSE}
 
-df = df%>%
+spp_df <- spp_df %>%
   mutate(am_log = log(am_area),
@@
 
-all_status <- ggplot(df,aes(iucn_log,am_log,color=iucn_category)) + 
-  geom_point() + geom_abline() +
-  labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
-        title= 'Comparing area of range maps\n (all area values are km2 log transformed)',fill='IUCN Category')
-ggsave('figures/CompareTotalAreaLOG_AM-IUCN_scatter.png',plot=all_status,width=8,height=4,dpi=100)
-print(all_status)
+scatter_am_iucn_area_log <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
+  geom_point() + 
+  geom_abline(slope = 1, intercept = 0, color = 'red') +
+  labs(x = 'IUCN Area (log transformed)', 
+       y = 'AquaMaps Area (log transformed)',
+       title = 'Comparing area of range maps\n (all area values are km2 log transformed)', 
+       fill = 'IUCN Category')
+
+ggsave('figures/CompareTotalAreaLOG_AM-IUCN_scatter.png', 
+       plot = scatter_am_iucn_area_log, 
+       width = 8, height = 4, dpi = 100)
+
+print(scatter_am_iucn_area_log)
 ```
@@
 Data split into species groups and by IUCN category
-```{r,echo=F,warning=F,message=F}
+```{r, echo = FALSE, warning = FALSE, message = FALSE}
 
-groups <- ggplot(df, aes(iucn_log, am_log,color=iucn_category)) + geom_point() +facet_wrap(~spp_group) + 
-        geom_abline() + labs(x='IUCN Area (log transformed)', y='AquaMaps Area (log transformed)',
-                             title='Comparing differences in area across species groups', fill='IUCN Category')
+scatter_am_iucn_alog_gp <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
+  geom_point() +
+  facet_wrap(~spp_group) + 
+  geom_abline() + 
+  labs(x = 'IUCN Area (log transformed)', 
+       y = 'AquaMaps Area (log transformed)',
+       title = 'Comparing differences in area across species groups', 
+       fill = 'IUCN Category')
 
-ggsave('figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png',plot=groups,width=8,height=6,dpi=100)
-print(groups)
+ggsave('figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png', 
+       plot = scatter_am_iucn_alog_gp, 
+       width = 8, height = 6, dpi = 100)
+
+print(scatter_am_iucn_alog_gp)
 ```
@@
 
-```{r,echo=F,warning=FALSE,message=F}
+```{r, echo = FALSE, warning = FALSE, message = FALSE}
 
-spp = 'Chromis alpha'
+spp <- 'Chromis alpha'
 map <- get_spp_map(spp)
 
- am_map <- map%>%
+am_map <- map %>%
   filter(!is.na(am_prob))
@@
 
-    r_am_spp[!is.na(r_am_spp)]<-1
+r_am_spp[!is.na(r_am_spp)] <- 1
 
@@
 
-#plot
-    plot(r_iucn_spp,col='mediumpurple',useRaster=FALSE,axes=FALSE,box=FALSE,legend=FALSE)
-  plot(r_am_spp,col='orange',main="Yellow-speckled Chromis", useRaster=FALSE,add=T,axes=FALSE,box=FALSE,legend=FALSE)
-  map('world', col = 'gray95', fill = T, border = 'gray80', add = TRUE)
-  legend("topright",legend=c("AquaMaps","IUCN"),fill=c('orange','mediumpurple'))
+### plot raster of species: IUCN map, AM map, world map, legend
+plot(r_iucn_spp, col = 'mediumpurple', 
+     useRaster = FALSE, 
+     axes = FALSE, 
+     box = FALSE, 
+     legend = FALSE)
+
+plot(r_am_spp, col = 'orange', 
+     main = "Yellow-speckled Chromis", 
+     useRaster = FALSE, 
+     add = TRUE, 
+     axes = FALSE, 
+     box = FALSE, 
+     legend = FALSE)
+
+map('world', col = 'gray95', 
+    fill = TRUE, 
+    border = 'gray80', 
+    add = TRUE)
+
+legend("topright", 
+       legend = c("AquaMaps", "IUCN"), 
+       fill = c('orange', 'mediumpurple'))
 
@@
 
-```{r,echo=F,warning=FALSE,message=F}
+```{r, echo = FALSE, warning = FALSE, message = FALSE}
 
-  df <- gather(spp_list, perc.type, percent, perc, am_perc, iucn_perc) %>%
+  spp_df2 <- gather(spp_list, perc.type, percent, perc, am_perc, iucn_perc) %>%
     dplyr::select(sciname, spp_group, perc.type, percent) %>%
@@
 
-  perc_c = df %>%
+  perc_c <- spp_df2 %>%
     filter(sciname == spp)
 
-   p1 = perc_c %>% dplyr::select(percent)
+  p1 <- perc_c %>% 
+    dplyr::select(percent)
 
@@
 
-For `r spp`, `r round(pi[1], 2)`% of the total area is covered by both IUCN and AquaMaps, `r round(p1[2,],2)`% is covered just by AquaMaps and `r round(p1[3,], 2)`% is covered just by IUCN.
+For `r spp`, `r round(pi[1], 2)`% of the total area is covered by both IUCN and AquaMaps, `r round(p1[2, ], 2)`% is covered just by AquaMaps and `r round(p1[3, ], 2)`% is covered just by IUCN.
 
@@
 
-```{r,echo=F,warning=FALSE,message=F}
+```{r, echo = FALSE, warning = FALSE, message = FALSE}
 
-#look at density plot
+### look at density plot
 
-overlap = df%>%dplyr::filter(perc.type=='perc')
+overlap <- spp_df2 %>%
+  dplyr::filter(perc.type == 'perc')
 
-all<-ggplot(overlap,aes(percent))+
-    geom_histogram(aes(y=..count../1926),binwidth=5,alpha=.8)+
-    labs(x='% total range that is shared b/w both datasets', y='Proportion of Total Species')
+hist_overlap_pct <- ggplot(overlap, aes(percent)) +
+  geom_histogram(aes(y = ..count../1926), binwidth = 5, alpha = .8) +
+  labs(x = '% total range that is shared b/w both datasets', 
+       y = 'Proportion of Total Species')
+
 ggsave('figures/Overlap-PercTotalRange.png')
-print(all)
+
+print(hist_overlap_pct)
 ```
@@
 Split up by species - not sure if this is useful or not
-```{r,echo=F, message = F, warning = F}
+```{r, echo = FALSE, message = FALSE, warning = FALSE}
 
-sp <- ggplot(overlap,aes(percent))+geom_histogram(binwidth=5,alpha=.8) +
+hist_overlap_pct_gp <- ggplot(overlap, aes(percent)) + 
+  geom_histogram(binwidth = 5, alpha = .8) +
   facet_wrap(~spp_group) +
-  labs(x='Percent of total range that is shared between both datasets', y='Number of Total Species')+ 
+  labs(x = 'Percent of total range that is shared between both datasets', 
+       y = 'Number of Total Species') + 
   theme(strip.text.x = element_text(size = 8))
@@
 ggsave('figures/Overlap-PercTotalRange_spGroups.png')
-print(sp)
+print(hist_overlap_pct_gp)
 
@@
 
-```{r pct small within large, echo=F, warning= F}
+```{r pct small within large, echo = FALSE, warning = FALSE}
 
-d <- spp_list %>%
+spp_df4 <- spp_list %>%
   dplyr::select(sciname, sm_perc, sm_range) %>%
@@
 
-sm_plot <- ggplot(d, aes(sm_perc, fill = sm_range)) + 
-       geom_histogram(aes(y=..count../sum(..count..)),binwidth = 4, alpha = .8) +
+sm_plot <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
+  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
   scale_fill_discrete(name = "Dataset with smaller range", labels = c("AquaMaps", "IUCN")) + 
-        labs(x = 'Percent of smaller range that falls within the larger range', y = 'Proportion of all species (n=1962)')
-ggsave('figures/PercSmallRangeInLarger.png',width=6,height=4)
+  labs(x = 'Percent of smaller range that falls within the larger range', 
+       y = 'Proportion of all species (n = 1962)')
+
+ggsave('figures/PercSmallRangeInLarger.png', width = 6, height = 4)
 
@@
 Exploring possible correlation between total species range area and percent overlap of smaller range within the larger:
-``` {r pct overlap vs area, echo = FALSE, warning = F, message=F}
+``` {r pct overlap vs area, echo = FALSE, warning = FALSE, message = FALSE}
 
-reg <- ggplot(spp_list, aes(sm_perc,total_area, color = sm_range)) + 
+reg <- ggplot(spp_list, aes(sm_perc, total_area, color = sm_range)) + 
   geom_point() + 
   geom_abline() + 
-  labs(y = 'Total Area (km2)', x = 'Percent overlap smaller range in larger', 
+  labs(y = 'Total Area (km2)', 
+       x = 'Percent overlap smaller range in larger', 
        title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
 
-ggsave('figures/AreavsPercOverlap.png',width=6,height=4)
+ggsave('figures/AreavsPercOverlap.png', width = 6, height = 4)
 
@@
 
-``` {r, echo = FALSE, warning = F, message=F}
+``` {r, echo = FALSE, warning = FALSE, message = FALSE}
 
@@
 
-p1 <- reg + geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)
+p1 <- reg + 
+  geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)
 
@@
 
-summary(lm(total_area~sm_perc,spp_list))
+summary(lm(total_area~sm_perc, spp_list))
 ```
@@
 Perent coverage as function of area, clipping to area below 100,000,000 km^2:
-``` {r, echo = F, warning = F, message = F}
+``` {r, echo = FALSE, warning = FALSE, message = FALSE}
 # remove all area greater than 100000000
 
-l <- spp_list %>% filter(total_area < 100000000)
+l <- spp_list %>% 
+  filter(total_area < 100000000)
 
@@
   geom_point() + 
-  labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
+  labs(x = 'Total Area (km2)', 
+       y = 'Percent overlap smaller range in larger', 
        title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
 
-ggsave('figures/AreavsPercOverlap_2.png',width=6,height=4)
+ggsave('figures/AreavsPercOverlap_2.png', width = 6, height = 4)
 
@@
 Percent coverage as a function of log(area)
-``` {r, echo = F, warning = F, message = F}
+``` {r, echo = FALSE, warning = FALSE, message = FALSE}
 # Log transform area
@@
   geom_point() + 
-  labs(x = 'Log(total Area)', y = 'Percent overlap smaller range in larger', 
+  labs(x = 'Log(total Area)', 
+       y = 'Percent overlap smaller range in larger', 
        title = 'Size of species range (log) vs. percent overlap of \nsmaller range with larger')
 
-ggsave('figures/AreavsPercOverlap_log.png',width=6,height=4)
+ggsave('figures/AreavsPercOverlap_log.png', width=6, height=4)
 
@@
 
-``` {r examine regressions, echo = FALSE, warning = F}
+``` {r examine regressions, echo = FALSE, warning = FALSE}
 
@@
 ###(5) Examining data broken into quartiles to see if smaller patterns emerge
-``` {r examine quartiles, echo = FALSE, warning = F}
+``` {r examine quartiles, echo = FALSE, warning = FALSE}
 
@@
 
-qtiles <- c(quantile(spp_list$total_area, 0.0,na.rm=T),
-            quantile(spp_list$total_area, .25,na.rm=T),
-            quantile(spp_list$total_area, .50,na.rm=T),
-            quantile(spp_list$total_area, .75,na.rm=T),
-            quantile(spp_list$total_area, 1.0,na.rm=T))
+qtiles <- c(quantile(spp_list$total_area, 0.0, na.rm = TRUE),
+            quantile(spp_list$total_area, .25, na.rm = TRUE),
+            quantile(spp_list$total_area, .50, na.rm = TRUE),
+            quantile(spp_list$total_area, .75, na.rm = TRUE),
+            quantile(spp_list$total_area, 1.0, na.rm = TRUE))
 
@@
     geom_point() + 
-  labs(x = 'Total Area (km2)', y = 'Percent overlap smaller range in larger', 
+    labs(x = 'Total Area (km2)', 
+         y = 'Percent overlap smaller range in larger', 
          title = sprintf('Size of species range (quartile %s) vs. percent overlap of \nsmaller range with larger', i))
-ggsave(paste0('figures/AreavsOverlap_quantile_',i,'.png'),width=6,height=4)
+
+  ggsave(paste0('figures/AreavsOverlap_quantile_', i, '.png'), width = 6, height = 4)
   print(qtile_plot)
@@
 
-```{r aquamaps probabilities, warning=F,message=F,cache.lazy=FALSE,cache=T,verbose=F,results=F,echo=F,include=F}
+```{r aquamaps probabilities, warning = FALSE, message = FALSE, cache.lazy = FALSE, cache = TRUE, verbose = FALSE, results = FALSE, echo = FALSE, include = FALSE}
 
@@
 
-dir_N = c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
+dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
 
-dir_aquamaps = file.path(dir_N,"git-annex/globalprep/SPP_ICO/raw/aquamaps_2014")
+dir_aquamaps <- file.path(dir_N, "git-annex/globalprep/SPP_ICO/raw/aquamaps_2014")
 
-spp_cells = fread(input=file.path(dir_aquamaps,"tables/ohi_hcaf_species_native.csv"),verbose=F,header=T,
-                  colClasses=c('NULL','character',NA,'numeric','NULL','NULL')) #1 min 49 sec
+spp_cells <- fread(input = file.path(dir_aquamaps, "tables/ohi_hcaf_species_native.csv"), verbose = FALSE, header = TRUE, 
+                  colClasses = c('NULL', 'character', NA, 'numeric', 'NULL', 'NULL')) #1 min 49 sec
 
-setnames(spp_cells,c('SpeciesID','CsquareCode','probability'),c("species_id","csquare_code","probability"))
+setnames(spp_cells, c('SpeciesID', 'CsquareCode', 'probability'), c("species_id", "csquare_code", "probability"))
 
-spp_cells = as.data.frame(spp_cells)
+spp_cells <- as.data.frame(spp_cells)
 
-spp       = fread(input=file.path(dir_aquamaps,"tables/ohi_speciesoccursum.csv"),verbose=F,header=T)%>%
-              mutate(scientific = paste(Genus,Species,sep=" "))
-spp = as.data.frame(spp)
+spp       <- fread(input = file.path(dir_aquamaps, "tables/ohi_speciesoccursum.csv"), 
+                  verbose = FALSE, header = TRUE) %>%
+              mutate(scientific = paste(Genus, Species, sep = " "))
+spp <- as.data.frame(spp)
 
@@
 
-```{r,warning=F,message=F, echo=F}
+```{r, warning = FALSE, message = FALSE, echo = FALSE}
 
-#create dataframe that gives me all species, their total number of cells, total number of cells >=0.4 prob, differnece between the two and threat
-df = spp_cells%>%
-      group_by(species_id)%>%
-        summarize(cells_0.4 = sum(probability>=0.4),
-                  cells_t = sum(probability>=0))%>%
-          mutate(prop = cells_0.4/cells_t)%>%
-            left_join(spp,by=c('species_id'='SPECIESID'))%>%
-          select(scientific,species_id,cells_0.4,cells_t,prop,iucn_code)
+#create dataframe that gives me all species, their total number of cells, total number of cells >= 0.4 prob, differnece between the two and threat
+spp_df3 <- spp_cells %>%
+  group_by(species_id) %>%
+  summarize(cells_0.4 = sum(probability >= 0.4),
+            cells_t = sum(probability >= 0)) %>%
+  mutate(prop = cells_0.4 / cells_t) %>%
+  left_join(spp, by = c('species_id' = 'SPECIESID')) %>%
+  select(scientific, species_id, cells_0.4, cells_t, prop, iucn_code)
 
-hist <- ggplot(df,aes(prop,fill=iucn_code))+geom_histogram(aes(y=..count../sum(..count..)),binwidth = 0.05, alpha = .7)+labs(x='Proportion of range area included when selecting a threshold of 0.4',y='Proportion of all aquamaps species (17,348 total)')
-ggsave('figures/AMThreshold.png',width=6,height=4)
+hist <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
+  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
+  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', y = 'Proportion of all aquamaps species (17,348 total)')
+
+ggsave('figures/AMThreshold.png', width = 6, height = 4)
+
 print(hist)

### 2015-11-14 data_explore.Rmd
still updating to get to the point where I can do that damn graph: percent overlap vs area ratio.

@@
 #species list 
-spp_list <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
+spp_list_base <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
 
@@
 get_spp_map <- function(species){
-  spp_id <- spp_list %>%
+  spp_id <- spp_list_base %>%
     filter(sciname == species & is.na(parent_sid)) %>%
@@
 
-spp_list <- spp_list %>%
+spp_list <- spp_list_base %>%
               mutate(total_area = NA,
@@
 ###get just the aquamaps cells
-   am_spp_map <- spp_map%>%
+  am_spp_map <- spp_map %>%
     filter(!is.na(am_prob))%>%
-                    dplyr::select(loiczid, am_sid, am_prob)%>% #add in am_prob when we want to look at probability
+    dplyr::select(loiczid, am_sid, am_prob) %>% #add in am_prob when we want to look at probability
     unique()
@@
   spp_list[i, ]$sm_range <- ifelse(length(cells_am)<length(cells_iucn), 'AM', 'IUCN') #what dataset has the smaller range
-  spp_list[i, ]$sm_perc <- length(cells_overlap)/min(length(cells_am), length(cells_iucn)) * 100 #this is the total percent of smaller range encompassed in the larger range
-cat(sprintf('%s: For species %s, %.2f%% of both datasets overlap, %.2f%% is just AquaMaps and %.2f%% is just IUCN\n', i, spp, perc, perc_am, perc_iucn))
+  spp_list[i, ]$sm_perc <- length(cells_overlap)/min(length(cells_am), length(cells_iucn)) * 100 
+    #this is the total percent of smaller range encompassed in the larger range
 
+  cat(sprintf('%s: For species %s, %.2f%% of both datasets overlap, %.2f%% is just AquaMaps and %.2f%% is just IUCN\n', 
+              i, spp, perc, perc_am, perc_iucn))
   }
@@
 #remove non IUCN species and log total area (for use later on)
-spp_list <- spp_list%>%
-            filter(spp_group != 'hagfishes' & spp_group != 'non-homalopsids')%>%
+spp_list <- spp_list %>%
+            filter(spp_group != 'hagfishes' & spp_group != 'non-homalopsids') %>%
               mutate(log_total_area = log(total_area))
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, warning = FALSE, message = FALSE}
+```{r, echo = FALSE, warning = FALSE, message = FALSE}
 
-scatter_am_iucn_area <- ggplot(spp_df, aes(iucn_area, am_area, color = iucn_category)) + 
+scatter_am_iucn_area <- ggplot(spp_df, aes(x = iucn_area, y = am_area, color = iucn_category)) + 
   geom_point() + 
@@
 
-sm_plot <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
+hist_sm_within_lg <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
   geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
@@
 
-print(sm_plot)
+print(hist_sm_within_lg)
 
@@
 
-reg <- ggplot(spp_list, aes(sm_perc, total_area, color = sm_range)) + 
+scatter_area_vs_overlap <- ggplot(spp_list, aes(sm_perc, total_area, color = sm_range)) + 
   geom_point() + 
@@
 
-print(reg)
+print(scatter_area_vs_overlap)
 
@@
 lm_eqn <- function(df){
-  m  <- lm(total_area~sm_perc, spp_list)
+  m  <- lm(total_area ~ sm_perc, spp_list)
   eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
@@
 
-p1 <- reg + 
+scatter_area_vs_overlap <- scatter_area_vs_overlap + 
   geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)
@@
 
-print(p1)
+print(scatter_area_vs_overlap)
 
-summary(lm(total_area~sm_perc, spp_list))
+summary(lm(total_area ~ sm_perc, spp_list))
 ```
@@
 
-Perent coverage as function of area, clipping to area below 100,000,000 km^2:
+Percent coverage as function of area, clipping to area below 100,000,000 km^2:
 ``` {r, echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-l <- spp_list %>% 
+spp_small_df <- spp_list %>% 
   filter(total_area < 100000000)
 
-r <- ggplot(l, aes(total_area, sm_perc, color = sm_range)) + 
+scatter_a_vs_overlap_small <- ggplot(spp_small_df, aes(total_area, sm_perc, color = sm_range)) + 
   geom_point() + 
@@
 
-print(r)
+print(scatter_a_vs_overlap_small)
 
@@
 
-logA_vs_pct <- ggplot(spp_list, aes(log_total_area, sm_perc, color = sm_range)) + 
+scatter_logA_vs_overlap <- ggplot(spp_list, aes(log_total_area, sm_perc, color = sm_range)) + 
   geom_point() + 
@@
 
-print(logA_vs_pct)
+print(scatter_logA_vs_overlap)
 ```
 
-``` {r examine regressions, echo = FALSE, warning = FALSE}
+``` {r examine regressions, eval = FALSE, echo = FALSE, warning = FALSE}
 
@@
 ###(5) Examining data broken into quartiles to see if smaller patterns emerge
-``` {r examine quartiles, echo = FALSE, warning = FALSE}
+``` {r examine quartiles, eval = FALSE, echo = FALSE, warning = FALSE}
 
@@
 
-###(6) Looking at how many and what type (threat category) of species we are losing by setting a threshold of 0.4 on the aquamaps species.
+###(6) Looking at how many and what type (threat category) of species we are losing by setting a threshold of 0.4 on the aquamaps species
+
+This looks at the full Aquamaps dataset... should it be limited to species with valid categories?
 
@@
 
-hist <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
+hist_area_vs_threshold <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
   geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
@@
 
-print(hist)
+print(hist_area_vs_threshold)
 

### 2015-11-16 ms_outline.Rmd
added more txt

--- 2015-10-30 ms_outline.Rmd
@@
 
-##Main message (25 words or less):
+##Main message
 
-A comparison of two open source, spatial datasets and the implications of making decisions about what data to use as applied to the Ocean Health Index.
+Understanding the differences in marine species range maps from IUCN and AquaMaps, and the implications of using one or both.
 
@@
 
+- spatial distribution efforts range from modeled to expert input
+- help identify hotspots for global biodiversity 
+- priorities for conservation
+- marine spatial planning
+- assessing biodiversity 
+- monitor impacts (human, natural) overtime
+-
+
+Species range maps are useful in conservation and management, as well as understanding the spatial distribution of movement and overlap of species in different habitats globally.
+Marine species prove more difficult to map based on their inherent nature of being difficult to observe and track.
+Two main sources for marine species are the IUCN and AquaMaps.
+These two datasets differ in methodology and final products, yet the differences between the two are not well documented.
+AquaMaps uses a computer model based on preferential environmental parameters to predict species locations and probability of occurrence, while the IUCN uses expert insight to assign spatial boundaries around known locations of marine species.
+Approximately 1,900 species are found in both datasets and there are very large differences between these maps.
+Depending on data usage, choosing one of these datasets over the other can lead to potentially large differences in results as exemplified here by the Ocean Health Index as an example.
+
 ***
+
+##Introduction
+
+There have been multiple efforts to map marine species based on both observational data and known life history and habitat preferences. AquaMaps and IUCN are two of these efforts.
 
@@
 ###IUCN
+
+The IUCN conducts global assessments for species [detail assessment process - one paragraph].
+
+Spatial designation of the species distribution is left to the experts with strict guidance for how to assign the designation.
+Using known occurrence data, experts depict the species distribution by polygons using a systematic methodology laid out by the [IUCN](http://www.amphibians.org/wp-content/uploads/2013/09/Red-List-Mapping-standards-ARLA-Jan2014-web-version.pdf).
+These maps are meant to act as 'limits of distibution' and should be interpreted as the species 'probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon'.
 
@@
 
-These two clearly differ in one aspect - AquaMaps provides higher resolution information about where species do and do not occur globally whereas the IUCN range maps are unable to provide specific information about where within a polygon a species occurs.
+These two differ in spatial resolution - AquaMaps provides higher resolution information about where species do and do not occur globally whereas the IUCN range maps are unable to provide specific information about where within a polygon a species occurs.
 This is understandable given the mobile nature of many marine species and the difficulty in observing and tracking known occurrences.
+
+AquaMaps uses both occurrence data and environmental parameters to predict relative species occurrence.
+This approach attempts to reduce sampling bias and fill gaps where data is not available.
+Six environmental predictors are used to estimate where a species could be found; sea surface temperature, bottom depth, salinity, primary production, sea ice concentration and distance to land.
+For more information on methods see Kaschner et al.
+(2007) and online at http://www.aquamaps.org/main/AboutAquaMaps.php.
+
+There are inherent issues with AquaMaps. It can't be expected to accurately predict all marine species ranges. Expert insight is sought and incorporated when available.
 
@@
 
+To compare these two datasets for all species found in each (1920), the IUCN maps were rasterized to the same cell resolution as AquaMaps, 0.5 degrees.
+For every grid cell where the polygon overlapped, a value of 1 was assigned.
+To allow for comparison, the probability of occurrence values in the AquaMaps data were ignored and all cells where there was a value greater than 0 were assigned a 1.
+This way we have comparable maps for total species range.
+
+Using predifined species groups by the IUCN, it is clear to see that tropical species - specifically wrasses, corals, blennies and groupers, make up the majority of species found in both.
+This is driven more by the IUCN dataset since the number of species with spatial data is directly attributed to the species that have assessments.
+
 ![](./figures/sp_groups_barchart.png)
+
+For each species map, the total area was calculated for all cells where the species was determined as 'possibly present' by both datasets.
+Due to large areas and differences, the total area per species was log transformed and then compared to the range map of the same species between datasets.
+There are differences, mainly driven by differeneces in the wrasses and combtooth blennies, but this could be due to there being more species in both of those species groups.
+
 ![](./figures/CompareTotalAreaLOG_AM-IUCN_scatter.png)
 ![](./figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png)
+
+For each species range map, the total area was calculated for all cells that were included in either species distributions, regardless of source. The number of cells that were found in both
+
 ![](./figures/Overlap-PercTotalRange.png)

### 2015-11-17 data_explore.Rmd
adding plot of area ratio to overlap, and representative species maps

@@
 ---
-title: "Data Exploration: IUCN and AquaMaps"
+title: 'Data Exploration: IUCN and AquaMaps'
 output: html_document
@@
     filter(!is.na(am_prob))%>%
-    dplyr::select(loiczid, am_sid, am_prob) %>% #add in am_prob when we want to look at probability
+    dplyr::select(loiczid, am_sid, am_prob) %>% 
+      ### add in am_prob when we want to look at probability
     unique()
@@
    #iucn map cells
-  cells_iucn<-iucn_map%>%
-              filter(iucn_pres == TRUE)%>%
+  cells_iucn<-iucn_map %>%
+    filter(iucn_pres == TRUE) %>%
     .$loiczid
-  cells_am <- am_spp_map%>%
-                filter(!is.na(am_sid))%>%
+  cells_am <- am_spp_map %>%
+    filter(!is.na(am_sid)) %>%
     .$loiczid
@@
 
-  spp_list[i, ]$sm_range <- ifelse(length(cells_am)<length(cells_iucn), 'AM', 'IUCN') #what dataset has the smaller range
+  spp_list[i, ]$sm_range <- ifelse(length(cells_am)<length(cells_iucn), 'AM', 'IUCN') 
+    ### what dataset has the smaller range
   spp_list[i, ]$sm_perc <- length(cells_overlap)/min(length(cells_am), length(cells_iucn)) * 100 
-    #this is the total percent of smaller range encompassed in the larger range
+    ### this is the total percent of smaller range encompassed in the larger range
 
@@
 ### plot raster of species: IUCN map, AM map, world map, legend
-plot(r_iucn_spp, col = 'mediumpurple', 
+plot(r_iucn_spp, 
+     col       = 'mediumpurple', 
      useRaster = FALSE, 
@@
 
-plot(r_am_spp, col = 'orange', 
-     main = "Yellow-speckled Chromis", 
+plot(r_am_spp, 
+     col       = 'orange', 
+     main      = 'Yellow-speckled Chromis', 
      useRaster = FALSE, 
@@
 
-map('world', col = 'gray95', 
+map('world', 
+    col    = 'gray95', 
     fill   = TRUE, 
@@
 
-legend("topright", 
-       legend = c("AquaMaps", "IUCN"), 
+legend('topright', 
+       legend = c('AquaMaps', 'IUCN'), 
        fill   = c('orange', 'mediumpurple'))
@@
   geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
-  scale_fill_discrete(name = "Dataset with smaller range", labels = c("AquaMaps", "IUCN")) + 
+  scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
   labs(x = 'Percent of smaller range that falls within the larger range', 
@@
 print(scatter_area_vs_overlap)
+
+```
+
+``` {r pct overlap vs ratio of area, echo = FALSE, warning = FALSE, message = FALSE}
+
+spp_list <- spp_list %>%
+  mutate(area_ratio = am_area / iucn_area,
+         area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
+         area_ratio = area_ratio * 100)
+
+### Inspect each quadrant to find representative species
+quadrant1_spp <- spp_list %>%
+  filter(area_ratio > 90 & sm_perc > 90)
+quadrant2_spp <- spp_list %>%
+  filter(area_ratio > 25 & area_ratio < 30) %>%
+  filter(sm_perc > 90)
+quadrant4_spp <- spp_list %>%
+  filter(sm_perc > 25 & sm_perc < 30) %>%
+  filter(area_ratio > 90)
+quadrant3_spp <- spp_list %>%
+  filter(area_ratio < 25 & sm_perc < 25)
+
+spp_quads <- c('Kajikia albida', 'Galaxea cryptoramosa', 'Acanthurus nigroris', 'Metanephrops australiensis', 'Conus magnificus')
+
+spp_list_label <- spp_list %>%
+  filter(sciname %in% spp_quads)
+
+scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(area_ratio, sm_perc, color = sm_range)) + 
+  geom_point() + 
+  geom_abline() + 
+  geom_point(data = spp_list_label, color = 'black', size = 4) + 
+  geom_text(data = spp_list_label, 
+            aes(label = sciname), 
+            color = 'black',
+            size = 4,
+            vjust = 1.5) +
+  labs(x = 'Range area ratio (smaller:larger) * 100%', 
+       y = 'Percent overlap smaller range in larger', 
+       title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
+
+ggsave('figures/AreaRatiovsPercOverlap.png', width = 6, height = 4)
+
+print(scatter_arearatio_vs_overlap)
+
+```
+
+```{r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE}
+
+for (spp in spp_quads) { # spp = 'Kajikia albida'
+  map <- get_spp_map(spp)
+
+am_map <- map %>%
+  filter(!is.na(am_prob))
+
+r_am_spp  <-  subs(loiczid_raster, 
+                   am_map[ , c('loiczid', 'am_prob')], 
+                   by = 'loiczid', 
+                   which = 'am_prob', 
+                   subsWithNA = TRUE)
+
+r_am_spp[!is.na(r_am_spp)] <- 1
+
+iucn_map <- map %>%
+  mutate(iucn_pres = iucn_area > 0)
+
+r_iucn_spp <- subs(loiczid_raster, 
+                   iucn_map[ , c('loiczid', 'iucn_pres')], 
+                   by = 'loiczid', 
+                   which = 'iucn_pres', 
+                   subsWithNA = TRUE)
+
+spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
+  full_join(as.data.frame(rasterToPoints(r_am_spp)),
+            by = c('x', 'y')) %>%
+
+spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
+  geom_raster(aes(fill = iucn_pres), alpha = .5) +
+  geom_raster(aes(fill = am_prob), alpha = .5) + 
+  borders('world', color='gray40', fill='gray45') + # create a layer of borders
+
+### plot raster of species: IUCN map, AM map, world map, legend
+plot(r_iucn_spp, 
+     col       = '#ff6666', 
+     main      = spp,
+     alpha     = 1,
+     useRaster = FALSE, 
+     axes      = FALSE, 
+     box       = FALSE, 
+     legend    = FALSE)
+
+plot(r_am_spp, 
+     col       = '#6666ff', 
+     alpha     = .4,
+     useRaster = FALSE, 
+     add       = TRUE, 
+     axes      = FALSE, 
+     box       = FALSE, 
+     legend    = FALSE)
+
+map('world', 
+    col    = 'gray95', 
+    fill   = TRUE, 
+    border = 'gray80', 
+    add    = TRUE)
+
+legend('topright', 
+       legend = c('AquaMaps', 'IUCN'), 
+       fill   = c('#6666ff', '#ff6666'))
+}
 
@@
 
-dir_aquamaps <- file.path(dir_N, "git-annex/globalprep/SPP_ICO/raw/aquamaps_2014")
+dir_aquamaps <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO/raw/aquamaps_2014')
 
-spp_cells <- fread(input = file.path(dir_aquamaps, "tables/ohi_hcaf_species_native.csv"), verbose = FALSE, header = TRUE, 
+spp_cells <- fread(input = file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), verbose = FALSE, header = TRUE, 
                   colClasses = c('NULL', 'character', NA, 'numeric', 'NULL', 'NULL')) #1 min 49 sec
 
-setnames(spp_cells, c('SpeciesID', 'CsquareCode', 'probability'), c("species_id", "csquare_code", "probability"))
+setnames(spp_cells, c('SpeciesID', 'CsquareCode', 'probability'), c('species_id', 'csquare_code', 'probability'))
 
@@
 
-spp       <- fread(input = file.path(dir_aquamaps, "tables/ohi_speciesoccursum.csv"), 
+spp       <- fread(input = file.path(dir_aquamaps, 'tables/ohi_speciesoccursum.csv'), 
                   verbose = FALSE, header = TRUE) %>%
-              mutate(scientific = paste(Genus, Species, sep = " "))
+              mutate(scientific = paste(Genus, Species, sep = ' '))
 spp <- as.data.frame(spp)
@@
 
-* Baseline is: preferring IUCN rangemaps over Aquamaps rangemaps; Aquamaps "presence" threshold is set at 40% or greater.
+* Baseline is: preferring IUCN rangemaps over Aquamaps rangemaps; Aquamaps 'presence' threshold is set at 40% or greater.
 * Scenario 1: preferring IUCN rangemaps over Aquamaps; AM presence threshold set at >= 1% (effectively, all occurrences).

### 2015-11-17 data_explore.Rmd
updating figures with theme

@@
 library(dplyr)
+library(stringr)
 
@@
 # Data files
-#iucn spp cells WITHOUT extinct locations of IUCN species
+#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
 iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), 
@@
   as.data.frame() %>%
-  filter(presence != 5)
+  filter(presence != 5) ### removes extinct subpopulations
 
@@
 
-#species list 
+### species info list - scinames, IUCN species ID, AM species ID, category etc
 spp_list_base <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
 
-#data for raster (used to plot species)
-
+### LOICZID cell ID numbers - used in plotting rasters
 loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
@@
 
-```{r species_map_function, echo = FALSE, warning = FALSE, message = FALSE}
+``` {r ggplot_theme_setups, echo = FALSE}
+
+### generic theme for all plots
+ggtheme_plot <- theme(axis.ticks = element_blank(),
+                    text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
+                    plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
+                    legend.position = 'right')
+
+# ggtheme_map <- theme(axis.ticks = element_blank(), axis.text = element_blank(),
+#                     text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
+#                     plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
+#                     legend.position = 'right')
+
+### theme for species range maps - based on generic plot theme
+ggtheme_map <- ggtheme_plot +
+  theme(axis.text = element_blank())
+
+```
+
+``` {r species_map_functions, echo = FALSE, warning = FALSE, message = FALSE}
 
@@
 }
+
+### Plot Species Map function
+### uses ggplot to create a formatted map of species ranges, including
+### Aquamaps, IUCN, and overlapped ranges; returns the plot object.
+plot_rangemap <- function(spp) {
+  map <- get_spp_map(spp)
+
+  am_map <- map %>%
+    filter(!is.na(am_prob))
+
+  r_am_spp  <-  subs(loiczid_raster, 
+                     am_map[ , c('loiczid', 'am_prob')], 
+                     by = 'loiczid', 
+                     which = 'am_prob', 
+                     subsWithNA = TRUE)
+  r_am_spp[!is.na(r_am_spp)] <- 1
+
+  iucn_map <- map %>%
+    mutate(iucn_pres = iucn_area > 0)
+  r_iucn_spp <- subs(loiczid_raster, 
+                     iucn_map[ , c('loiczid', 'iucn_pres')], 
+                     by = 'loiczid', 
+                     which = 'iucn_pres', 
+                     subsWithNA = TRUE)
+
+  spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
+    full_join(as.data.frame(rasterToPoints(r_am_spp)),
+              by = c('x', 'y')) %>%
+    mutate(presence = ifelse(am_prob & is.na(iucn_pres), 'am',
+                             ifelse(iucn_pres & is.na(am_prob), 'iucn',
+                                    ifelse(iucn_pres & am_prob, 'both', NA))))
+
+  spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
+    ggtheme_map + 
+    geom_raster(aes(fill = presence), alpha = .5) +
+    scale_fill_manual(values = c('iucn' = 'blue', 'am' = 'red', 'both' = 'mediumorchid')) +
+    borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
+    labs(title = spp, x = NULL, y = NULL) 
+}
+
 ```
@@
 barchart_sp_groups <- ggplot(spp_df, aes(x = spp_group)) +
+  ggtheme_plot + 
   geom_bar(stat = 'bin') +
@@
 
-###(2) Looking at the total difference in area (km2) for each species range map in both datasets
+###(2) Looking at the total difference in area (km^2) for each species range map in both datasets
 
@@
 scatter_am_iucn_area <- ggplot(spp_df, aes(x = iucn_area, y = am_area, color = iucn_category)) + 
+  ggtheme_plot + 
   geom_point() + 
   geom_abline(slope = 1, intercept = 0, color = 'red') +
-  labs(x = 'Area (km2) from IUCN dataset', 
-       y = 'Area (km2) from AquaMaps dataset', 
+  labs(x = 'Area (km^2) from IUCN dataset', 
+       y = 'Area (km^2) from AquaMaps dataset', 
        title = 'Comparing total area of species range maps')
@@
 scatter_am_iucn_area_log <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
+  ggtheme_plot + 
   geom_point() + 
@@
        y = 'AquaMaps Area (log transformed)',
-       title = 'Comparing area of range maps\n (all area values are km2 log transformed)', 
+       title = 'Comparing area of range maps\n (all area values are km^2 log transformed)', 
        fill = 'IUCN Category')
@@
 scatter_am_iucn_alog_gp <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
+  ggtheme_plot + 
   geom_point() +
@@
 spp <- 'Chromis alpha'
-map <- get_spp_map(spp)
 
-am_map <- map %>%
-  filter(!is.na(am_prob))
+spp_map <- plot_rangemap(spp)
 
-r_am_spp  <-  subs(loiczid_raster, 
-                   am_map[ , c('loiczid', 'am_prob')], 
-                   by = 'loiczid', 
-                   which = 'am_prob', 
-                   subsWithNA = TRUE)
-
-r_am_spp[!is.na(r_am_spp)] <- 1
-
-iucn_map <- map %>%
-  mutate(iucn_pres = iucn_area > 0)
-
-r_iucn_spp <- subs(loiczid_raster, 
-                   iucn_map[ , c('loiczid', 'iucn_pres')], 
-                   by = 'loiczid', 
-                   which = 'iucn_pres', 
-                   subsWithNA = TRUE)
-
-### plot raster of species: IUCN map, AM map, world map, legend
-plot(r_iucn_spp, 
-     col       = 'mediumpurple', 
-     useRaster = FALSE, 
-     axes      = FALSE, 
-     box       = FALSE, 
-     legend    = FALSE)
-
-plot(r_am_spp, 
-     col       = 'orange', 
-     main      = 'Yellow-speckled Chromis', 
-     useRaster = FALSE, 
-     add       = TRUE, 
-     axes      = FALSE, 
-     box       = FALSE, 
-     legend    = FALSE)
-
-map('world', 
-    col    = 'gray95', 
-    fill   = TRUE, 
-    border = 'gray80', 
-    add    = TRUE)
-
-legend('topright', 
-       legend = c('AquaMaps', 'IUCN'), 
-       fill   = c('orange', 'mediumpurple'))
-
+print(spp_map)
+ggsave(plot = spp_map,
+       filename = sprintf('figures/rangemap_%s.png', spp),
+       width = 8, height = 4, dpi = 200)
 ```
@@
 hist_overlap_pct <- ggplot(overlap, aes(percent)) +
+  ggtheme_plot + 
   geom_histogram(aes(y = ..count../1926), binwidth = 5, alpha = .8) +
-  labs(x = '% total range that is shared b/w both datasets', 
+  labs(title = 'Percent of total range shared between datasets',
+       x = '% total range that is shared b/w both datasets', 
        y = 'Proportion of Total Species')
@@
 hist_overlap_pct_gp <- ggplot(overlap, aes(percent)) + 
+  ggtheme_plot + 
   geom_histogram(binwidth = 5, alpha = .8) +
   facet_wrap(~spp_group) +
-  labs(x = 'Percent of total range that is shared between both datasets', 
+  labs(title = 'Percent of total range shared, by species group',
+       x = 'Percent of total range that is shared between both datasets', 
        y = 'Number of Total Species') + 
@@
 hist_sm_within_lg <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
+  ggtheme_plot + 
   geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
   scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
-  labs(x = 'Percent of smaller range that falls within the larger range', 
-       y = 'Proportion of all species (n = 1962)')
+  labs(title = 'Percent of smaller range within larger range',
+       x = 'Percent of smaller range that falls within the larger range', 
+       y = sprintf('Proportion of all species (n = %s)', nrow(spp_df4)))
 
@@
 scatter_area_vs_overlap <- ggplot(spp_list, aes(sm_perc, total_area, color = sm_range)) + 
+  ggtheme_plot + 
   geom_point() + 
   geom_abline() + 
-  labs(y = 'Total Area (km2)', 
+  labs(y = 'Total Area (km^2)', 
        x = 'Percent overlap smaller range in larger', 
@@
 ### Inspect each quadrant to find representative species
-quadrant1_spp <- spp_list %>%
-  filter(area_ratio > 90 & sm_perc > 90)
-quadrant2_spp <- spp_list %>%
-  filter(area_ratio > 25 & area_ratio < 30) %>%
-  filter(sm_perc > 90)
-quadrant4_spp <- spp_list %>%
-  filter(sm_perc > 25 & sm_perc < 30) %>%
-  filter(area_ratio > 90)
-quadrant3_spp <- spp_list %>%
-  filter(area_ratio < 25 & sm_perc < 25)
+# quadrant1_spp <- spp_list %>%
+#   filter(area_ratio > 90 & sm_perc > 90)
+# quadrant2_spp <- spp_list %>%
+#   filter(area_ratio > 25 & area_ratio < 30) %>%
+#   filter(sm_perc > 90)
+# quadrant4_spp <- spp_list %>%
+#   filter(sm_perc > 25 & sm_perc < 30) %>%
+#   filter(area_ratio > 90)
+# quadrant3_spp <- spp_list %>%
+#   filter(area_ratio < 25 & sm_perc < 25)
 
-spp_quads <- c('Kajikia albida', 'Galaxea cryptoramosa', 'Acanthurus nigroris', 'Metanephrops australiensis', 'Conus magnificus')
+spp_quads <- c('q1' = 'Kajikia albida', 
+               'q2' = 'Galaxea cryptoramosa', 
+               'q3' = 'Acanthurus nigroris', 
+               'q3a' = 'Metanephrops australiensis', 
+               'q4' = 'Conus magnificus')
 
 spp_list_label <- spp_list %>%
-  filter(sciname %in% spp_quads)
+  filter(sciname %in% spp_quads) %>%
+  separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
+  mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
+  select(-g, -s)
 
 scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(area_ratio, sm_perc, color = sm_range)) + 
+  ggtheme_plot + 
   geom_point() + 
   geom_abline() + 
-  geom_point(data = spp_list_label, color = 'black', size = 4) + 
+  geom_point(data = spp_list_label, color = 'grey30') + 
   geom_text(data = spp_list_label, 
-            aes(label = sciname), 
-            color = 'black',
-            size = 4,
+            aes(label = sciname_label), 
+            color = 'grey30',
+            fontface = 3,
+            size = 2,
+            hjust = .25,
             vjust = 1.5) +
@@
 
-for (spp in spp_quads) { # spp = 'Kajikia albida'
-  map <- get_spp_map(spp)
+for (i in 1:length(spp_quads)) { # i = 1
+  spp <- spp_quads[i]
+  spp_quad <- names(spp_quads)[i]
 
-am_map <- map %>%
-  filter(!is.na(am_prob))
+  spp_plot <- plot_rangemap(spp)
+  print(spp_plot)
 
-r_am_spp  <-  subs(loiczid_raster, 
-                   am_map[ , c('loiczid', 'am_prob')], 
-                   by = 'loiczid', 
-                   which = 'am_prob', 
-                   subsWithNA = TRUE)
-
-r_am_spp[!is.na(r_am_spp)] <- 1
-
-iucn_map <- map %>%
-  mutate(iucn_pres = iucn_area > 0)
-
-r_iucn_spp <- subs(loiczid_raster, 
-                   iucn_map[ , c('loiczid', 'iucn_pres')], 
-                   by = 'loiczid', 
-                   which = 'iucn_pres', 
-                   subsWithNA = TRUE)
-
-spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
-  full_join(as.data.frame(rasterToPoints(r_am_spp)),
-            by = c('x', 'y')) %>%
-
-spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
-  geom_raster(aes(fill = iucn_pres), alpha = .5) +
-  geom_raster(aes(fill = am_prob), alpha = .5) + 
-  borders('world', color='gray40', fill='gray45') + # create a layer of borders
-
-### plot raster of species: IUCN map, AM map, world map, legend
-plot(r_iucn_spp, 
-     col       = '#ff6666', 
-     main      = spp,
-     alpha     = 1,
-     useRaster = FALSE, 
-     axes      = FALSE, 
-     box       = FALSE, 
-     legend    = FALSE)
-
-plot(r_am_spp, 
-     col       = '#6666ff', 
-     alpha     = .4,
-     useRaster = FALSE, 
-     add       = TRUE, 
-     axes      = FALSE, 
-     box       = FALSE, 
-     legend    = FALSE)
-
-map('world', 
-    col    = 'gray95', 
-    fill   = TRUE, 
-    border = 'gray80', 
-    add    = TRUE)
-
-legend('topright', 
-       legend = c('AquaMaps', 'IUCN'), 
-       fill   = c('#6666ff', '#ff6666'))
+  ggsave(filename = sprintf('figures/rangemap_%s_%s.png', spp_quad, str_replace(tolower(spp), ' ', '_')),
+         plot = spp_plot, 
+         width = 8, height = 4, dpi = 200)
 }
@@
 scatter_a_vs_overlap_small <- ggplot(spp_small_df, aes(total_area, sm_perc, color = sm_range)) + 
+  ggtheme_plot + 
   geom_point() + 
-  labs(x = 'Total Area (km2)', 
+  labs(x = 'Total Area (km^2)', 
        y = 'Percent overlap smaller range in larger', 
@@
 scatter_logA_vs_overlap <- ggplot(spp_list, aes(log_total_area, sm_perc, color = sm_range)) + 
+  ggtheme_plot + 
   geom_point() + 
@@
   qtile_plot <- ggplot(area_clip, aes(total_area, sm_perc, color = sm_range)) + 
+    ggtheme_plot + 
     geom_point() + 
-    labs(x = 'Total Area (km2)', 
+    labs(x = 'Total Area (km^2)', 
          y = 'Percent overlap smaller range in larger', 
@@
 # get spp_cells raw data for aquamaps
-
-dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
 dir_aquamaps <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO/raw/aquamaps_2014')
 
-spp_cells <- fread(input = file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), verbose = FALSE, header = TRUE, 
-                  colClasses = c('NULL', 'character', NA, 'numeric', 'NULL', 'NULL')) #1 min 49 sec
+spp_cells <- read_csv(file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), 
+                      col_types = '_ccd__', progress = TRUE)
+
+# spp_cells <- fread(input = file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), verbose = FALSE, header = TRUE, 
+#                   colClasses = c('NULL', 'character', NA, 'numeric', 'NULL', 'NULL')) #1 min 49 sec
 
@@
              verbose = FALSE, header = TRUE) %>%
-              mutate(scientific = paste(Genus, Species, sep = ' '))
+  mutate(sciname = paste(Genus, Species, sep = ' '))
 spp <- as.data.frame(spp)
@@
 
-#create dataframe that gives me all species, their total number of cells, total number of cells >= 0.4 prob, differnece between the two and threat
+### create dataframe that gives me all species, their total number of cells, 
+### total number of cells >= 0.4 prob, difference between the two and threat
 spp_df3 <- spp_cells %>%
@@
   left_join(spp, by = c('species_id' = 'SPECIESID')) %>%
-  select(scientific, species_id, cells_0.4, cells_t, prop, iucn_code)
+  select(sciname, species_id, cells_0.4, cells_t, prop, iucn_code)
 
 hist_area_vs_threshold <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
+  ggtheme_plot + 
   geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 

### 2015-11-18 data_explore.Rmd
uploading new figures etc

@@
 
-### Inspect each quadrant to find representative species
-# quadrant1_spp <- spp_list %>%
-#   filter(area_ratio > 90 & sm_perc > 90)
-# quadrant2_spp <- spp_list %>%
-#   filter(area_ratio > 25 & area_ratio < 30) %>%
-#   filter(sm_perc > 90)
-# quadrant4_spp <- spp_list %>%
-#   filter(sm_perc > 25 & sm_perc < 30) %>%
-#   filter(area_ratio > 90)
-# quadrant3_spp <- spp_list %>%
-#   filter(area_ratio < 25 & sm_perc < 25)
+area_ratio_quartiles <- quantile(spp_list$area_ratio, na.rm = TRUE)[2:4]
+pct_overlap_quartiles <- quantile(spp_list$sm_perc, na.rm = TRUE)[2:4]
 
@@
                'q2' = 'Galaxea cryptoramosa', 
-               'q3' = 'Acanthurus nigroris', 
-               'q3a' = 'Metanephrops australiensis', 
-               'q4' = 'Conus magnificus')
+               'q4' = 'Acanthurus nigroris', 
+               'q4a' = 'Metanephrops australiensis', 
+               'q3' = 'Conus magnificus')
 
@@
 
-scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(area_ratio, sm_perc, color = sm_range)) + 
+scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = log_total_area)) + 
   ggtheme_plot + 
@@
   geom_abline() + 
-  geom_point(data = spp_list_label, color = 'grey30') + 
+  geom_vline(x = area_ratio_quartiles,
+             color = 'grey50', linetype = 'dashed') +
+  geom_hline(y = pct_overlap_quartiles,
+             color = 'grey50', linetype = 'dashed') +
+  geom_point(data = spp_list_label, color = 'tomato2') + 
   geom_text(data = spp_list_label, 
             aes(label = sciname_label), 
-            color = 'grey30',
-            fontface = 3,
+            color = 'tomato2',
+            fontface = 'bold.italic',
             size = 2,
@@
 print(scatter_arearatio_vs_overlap)
+
+scatter_arearatio_vs_overlap2 <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = sm_range)) + 
+  ggtheme_plot + 
+  geom_point() + 
+  geom_abline() + 
+  geom_vline(x = area_ratio_quartiles,
+             color = 'grey50', linetype = 'dashed') +
+  geom_hline(y = pct_overlap_quartiles,
+             color = 'grey50', linetype = 'dashed') +
+  geom_point(data = spp_list_label, color = 'grey30') + 
+  geom_text(data = spp_list_label, 
+            aes(label = sciname_label), 
+            color = 'grey30',
+            fontface = 'bold.italic',
+            size = 2,
+            hjust = .25,
+            vjust = 1.5) +
+  labs(x = 'Range area ratio (smaller:larger) * 100%', 
+       y = 'Percent overlap smaller range in larger', 
+       title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
+
+ggsave('figures/AreaRatiovsPercOverlap2.png', width = 6, height = 4)
+
+print(scatter_arearatio_vs_overlap2)
+
+```
+
+``` {r examine_spp_gp_quartiles}
+
+spp_list1 <- spp_list %>%
+  group_by(spp_group) %>%
+  mutate(n_spp = n()) %>%
+  ungroup()
+
+spp_gp_q1 <- spp_list1 %>%
+  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE)) %>%
+  group_by(spp_group) %>%
+  summarize(n_spp = first(n_spp),
+            n_spp_q1 = n())
+spp_gp_q2 <- spp_list1 %>%
+  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE)) %>%
+  group_by(spp_group) %>%
+  summarize(n_spp = first(n_spp),
+            n_spp_q2 = n())
+spp_gp_q4 <- spp_list1 %>%
+  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE)) %>%
+  group_by(spp_group) %>%
+  summarize(n_spp = first(n_spp),
+            n_spp_q4 = n())
+spp_gp_q3 <- spp_list1 %>%
+  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE)) %>%
+  group_by(spp_group) %>%
+  summarize(n_spp = first(n_spp),
+            n_spp_q3 = n())
+
+spp_gp_quadrants <- spp_gp_q1 %>%
+  left_join(spp_gp_q2, by = c('spp_group', 'n_spp')) %>%
+  left_join(spp_gp_q3, by = c('spp_group', 'n_spp')) %>%
+  left_join(spp_gp_q4, by = c('spp_group', 'n_spp')) %>%
+  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
+  mutate(quad = str_replace(quad, 'n_spp_', ''),
+         pct_quad = n_quad/n_spp)
+
+spp_gp_quadrants <- spp_gp_quadrants %>%
+  left_join(spp_gp_quadrants %>% 
+              filter(quad == 'q1') %>%
+              select(spp_group, pct_q1 = pct_quad),
+            by = 'spp_group')
+
+barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
+                                  transform(spp_group = reorder(spp_group, pct_q1)), 
+                                aes(x = spp_group, fill = quad, weight = pct_quad)) +
+  ggtheme_plot + 
+  geom_bar(stat = 'bin') +
+  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0) +
+  coord_flip() +
+  labs(x = 'Species Group', 
+       y = 'Relative number of species per quadrant', 
+       title = 'Species group breakdown by quadrant')
 
@@
              verbose = FALSE, header = TRUE) %>%
-  mutate(sciname = paste(Genus, Species, sep = ' '))
+  mutate(sciname = paste(Genus, Species, sep = ' '),
+         iucn_code = str_replace(toupper(iucn_code), 'LR/', ''),
+         iucn_code = str_replace(iucn_code, 'N.E.', 'NE'))
 spp <- as.data.frame(spp)
@@
 
-```{r, warning = FALSE, message = FALSE, echo = FALSE}
+```{r hist_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
 
@@
   mutate(prop = cells_0.4 / cells_t) %>%
-  left_join(spp, by = c('species_id' = 'SPECIESID')) %>%
+  left_join(spp %>%
+#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')), 
+              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
+            by = c('species_id' = 'SPECIESID')) %>%
   select(sciname, species_id, cells_0.4, cells_t, prop, iucn_code)
@@
   geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
-  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', y = 'Proportion of all aquamaps species (17,348 total)')
+  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
+       y = 'Proportion of all aquamaps species (17,348 total)')
 
@@
 print(hist_area_vs_threshold)
+
+```
+
+```{r scatter_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
+
+### create dataframe that gives me all species, their total number of cells, 
+### total number of cells >= 0.4 prob, difference between the two and threat
+
+threshold_df <- data.frame()
+for (i in seq(0, 1, .05)) { # i = .05
+  temp_df <- spp_cells %>%
+    group_by(species_id) %>%
+    summarize(thresh       = i,
+              cells_thresh = sum(probability >= thresh),
+              cells_total  = sum(probability >= 0),
+              proportion   = cells_thresh/cells_total) %>%
+    left_join(spp %>%
+#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')), 
+              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
+            by = c('species_id' = 'SPECIESID')) %>%
+    select(sciname, species_id, thresh, cells_thresh, cells_total, proportion, iucn_code)
+
+  threshold_df <- threshold_df %>%
+    bind_rows(temp_df)
+}
+
+thresh_df_summary <- threshold_df %>%
+  group_by(thresh) %>%
+  summarize(mean_prop   = mean(proportion, na.rm = TRUE),
+            median_prop = median(proportion, na.rm = TRUE))
+
+scatter_area_vs_threshold <- ggplot(thresh_df_summary, aes(x = thresh, y = mean_prop)) + 
+  ggtheme_plot + 
+  geom_point(aes(y = median_prop), color = 'grey60') +
+  geom_point() + 
+  geom_abline(slope = -1, intercept = 1, color = 'red') + 
+  geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
+  ylim(0, 1) +
+  labs(title = 'Mean (median) species range after presence threshold cutoff',
+       x = 'Presence threshold cutoff', 
+       y = 'Mean (median) relative species range')
+
+ggsave('figures/scatter_AMThreshold.png', width = 6, height = 4)
+
+print(scatter_area_vs_threshold)
 

### 2015-11-18 data_explore.Rmd
update data_explore to plot new graphs; show regressions; etc

@@
 ggtheme_map <- ggtheme_plot +
-  theme(axis.text = element_blank())
+  theme(axis.text = element_blank()) +       
+  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
+  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2))
 
@@
 
-###(2) Looking at the total difference in area (km^2) for each species range map in both datasets
+###(2) Looking at the total difference in area (km^2^) for each species range map in both datasets
 
@@
 
-####Now looking across all species, here is the breakdown of how much overlap there is between the two datasets
+#### Now looking across all species, here is the breakdown of how much overlap there is between the two datasets
 
@@
 
-Split up by species - not sure if this is useful or not
+Split up by species groups
+
 ```{r, echo = FALSE, message = FALSE, warning = FALSE}
@@
 
-ggsave('figures/AreavsPercOverlap.png', width = 6, height = 4)
+lm_eqn <- function(df){
+  m  <- lm(total_area ~ sm_perc, spp_list)
+  eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
+        list(a  = format(coef(m)[1], digits = 2), 
+             b  = format(coef(m)[2], digits = 2), 
+             r2 = format(summary(m)$r.squared, digits = 3)))
+  as.character(as.expression(eq))             
+}
+
+scatter_area_vs_overlap <- scatter_area_vs_overlap + 
+  geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)
 
 print(scatter_area_vs_overlap)
+
+summary(lm(total_area ~ sm_perc, spp_list))
+
+ggsave('figures/AreavsPercOverlap.png', width = 6, height = 4)
 
@@
 
+lm_eqn2 <- function(df){
+  m  <- lm(area_ratio ~ sm_perc, spp_list)
+  eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
+        list(a  = format(coef(m)[1], digits = 2), 
+             b  = format(coef(m)[2], digits = 2), 
+             r2 = format(summary(m)$r.squared, digits = 3)))
+  as.character(as.expression(eq))             
+}
+
 scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = log_total_area)) + 
@@
        title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
+
+scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap + 
+  geom_text(x = 15, y = 1, label = lm_eqn2(spp_list), parse = TRUE, color = 'tomato2')
 
@@
 
-ggsave('figures/AreaRatiovsPercOverlap2.png', width = 6, height = 4)
+scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 + 
+  geom_text(x = 15, y = 1, label = lm_eqn2(spp_list), parse = TRUE, color = 'grey30')
+
+ggsave('figures/AreaRatiovsPercOverlap2.png', 
+       width = 6, height = 4)
 
@@
 
+summary(lm(area_ratio ~ sm_perc, spp_list))
+
 ```
 
-``` {r examine_spp_gp_quartiles}
+``` {r examine_spp_gp_quartiles, echo = FALSE}
 
@@
        title = 'Species group breakdown by quadrant')
+
+print(barchart_spp_gp_quads)
+ggsave(plot = barchart_spp_gp_quads,
+       filename = 'figures/barchart_spp_gp_quads.png',
+       width = 6, height = 4)
 

### 2015-11-18 data_explore.Rmd
updating plots and such

@@
 ggtheme_map <- ggtheme_plot +
-  theme(axis.text = element_blank()) +       
-  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
-  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2))
+  theme(axis.text = element_blank())
 
@@
     borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
+    scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
+    scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
     labs(title = spp, x = NULL, y = NULL) 
@@
 
-summary(lm(total_area ~ sm_perc, spp_list))
+# summary(lm(total_area ~ sm_perc, spp_list))
 
@@
   ggtheme_plot + 
+  ### color the quadrant backgrounds:
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
+                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
+                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
+                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
+                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
   geom_point() + 
-  geom_abline() + 
-  geom_vline(x = area_ratio_quartiles,
+  geom_vline(x = area_ratio_quartiles[c(1, 3)],
              color = 'grey50', linetype = 'dashed') +
-  geom_hline(y = pct_overlap_quartiles,
+  geom_hline(y = pct_overlap_quartiles[c(1, 3)],
              color = 'grey50', linetype = 'dashed') +
@@
 scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap + 
-  geom_text(x = 15, y = 1, label = lm_eqn2(spp_list), parse = TRUE, color = 'tomato2')
+  geom_text(x = 40, y = 5, label = lm_eqn2(spp_list), parse = TRUE, color = 'tomato2')
 
@@
 
-scatter_arearatio_vs_overlap2 <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline() + 
-  geom_vline(x = area_ratio_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_hline(y = pct_overlap_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_point(data = spp_list_label, color = 'grey30') + 
-  geom_text(data = spp_list_label, 
-            aes(label = sciname_label), 
-            color = 'grey30',
-            fontface = 'bold.italic',
-            size = 2,
-            hjust = .25,
-            vjust = 1.5) +
-  labs(x = 'Range area ratio (smaller:larger) * 100%', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
+# scatter_arearatio_vs_overlap2 <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = sm_range)) + 
+#   ggtheme_plot + 
+#   geom_point() + 
+#   geom_vline(x = area_ratio_quartiles,
+#              color = 'grey50', linetype = 'dashed') +
+#   geom_hline(y = pct_overlap_quartiles,
+#              color = 'grey50', linetype = 'dashed') +
+#   geom_point(data = spp_list_label, color = 'grey30') + 
+#   geom_text(data = spp_list_label, 
+#             aes(label = sciname_label), 
+#             color = 'grey30',
+#             fontface = 'bold.italic',
+#             size = 2,
+#             hjust = .25,
+#             vjust = 1.5) +
+#   labs(x = 'Range area ratio (smaller:larger) * 100%', 
+#        y = 'Percent overlap smaller range in larger', 
+#        title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
+# 
+# scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 + 
+#   geom_text(x = 40, y = 5, label = lm_eqn2(spp_list), parse = TRUE, color = 'grey30')
+# 
+# ggsave('figures/AreaRatiovsPercOverlap2.png', 
+#        width = 6, height = 4)
+# 
+# print(scatter_arearatio_vs_overlap2)
 
-scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 + 
-  geom_text(x = 15, y = 1, label = lm_eqn2(spp_list), parse = TRUE, color = 'grey30')
-
-ggsave('figures/AreaRatiovsPercOverlap2.png', 
-       width = 6, height = 4)
-
-print(scatter_arearatio_vs_overlap2)
-
-summary(lm(area_ratio ~ sm_perc, spp_list))
+# summary(lm(area_ratio ~ sm_perc, spp_list))
 
@@
   ggtheme_plot + 
-  geom_bar(stat = 'bin') +
+  geom_bar(stat = 'bin', alpha = .5) +
+  scale_fill_manual(values = c('q1' = 'green4', 'q2' = 'greenyellow', 'q3' = 'orange', 'q4' = 'orangered2')) +
   geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0) +
@@
 
-``` {r, echo = FALSE, warning = FALSE, message = FALSE}
-
-# Regression
-
-lm_eqn <- function(df){
-  m  <- lm(total_area ~ sm_perc, spp_list)
-  eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
-        list(a  = format(coef(m)[1], digits = 2), 
-             b  = format(coef(m)[2], digits = 2), 
-             r2 = format(summary(m)$r.squared, digits = 3)))
-  as.character(as.expression(eq))             
-}
-
-scatter_area_vs_overlap <- scatter_area_vs_overlap + 
-  geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)
-
-ggsave('figures/CorrelationAreavsPercOverlap.png')
-
-print(scatter_area_vs_overlap)
-
-summary(lm(total_area ~ sm_perc, spp_list))
-```
-
 ***
 
-Percent coverage as function of area, clipping to area below 100,000,000 km^2:
+Percent coverage as function of area, clipping to area below 100,000,000 km^2^:
+
 ``` {r, echo = FALSE, warning = FALSE, message = FALSE}
@@
             cells_t = sum(probability >= 0)) %>%
-  mutate(prop = cells_0.4 / cells_t) %>%
-  left_join(spp %>%
-#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')), 
-              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
-            by = c('species_id' = 'SPECIESID')) %>%
+  left_join(spp, by = c('species_id' = 'SPECIESID')) %>%
+  mutate(prop = cells_0.4 / cells_t,
+         iucn_code = ifelse(is.na(iucn_code), 'NE', iucn_code)) %>%
+  filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')) %>%
+  # filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')) %>%
   select(sciname, species_id, cells_0.4, cells_t, prop, iucn_code)
+
+iucn_levels = c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')
+spp_df3 <- spp_df3 %>% 
+  transform(iucn_code = factor(iucn_code, levels = iucn_levels))
 
@@
   geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
+  scale_fill_manual(values = c('LC' = 'green4', 
+                               'NT' = 'greenyellow', 
+                               'VU' = 'orange', 
+                               'EN' = 'orangered2', 
+                               'CR' = 'red4',
+                               'DD' = 'grey60',
+                               'NE' = 'grey70')) +
   labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
-       y = 'Proportion of all aquamaps species (17,348 total)')
+       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3)))
 
@@
 print(hist_area_vs_threshold)
+
+spp_df3a <- spp_df3 %>% 
+  filter(!iucn_code %in% c('DD', 'NE'))
+
+hist_area_vs_threshold2 <- ggplot(spp_df3a, aes(prop, fill = iucn_code)) + 
+  ggtheme_plot + 
+  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
+  scale_fill_manual(values = c('LC' = 'green4', 
+                               'NT' = 'greenyellow', 
+                               'VU' = 'orange', 
+                               'EN' = 'orangered2', 
+                               'CR' = 'red4',
+                               'DD' = 'grey60',
+                               'NE' = 'grey70')) +
+  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
+       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3a)))
+
+ggsave('figures/AMThreshold2.png', width = 6, height = 4)
+
+print(hist_area_vs_threshold2)
 

### 2015-11-21 R/data_explore_am_thresh.Rmd
trying out new maps by quadrant to see if there are common characteristics...

--- 2015-11-18 data_explore.Rmd
@@
----
-title: 'Data Exploration: IUCN and AquaMaps'
-output: html_document
----
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-# Libraries and Paths
-
-library(ggplot2)
-library(maps)
-library(readr)
-library(raster)
-library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-
-dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-           'Darwin'  = '/Volumes/data_edit',
-           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-```
-
-```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
-
-# Data files
-#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), 
-                           col_types = 'cdddd', progress = TRUE) %>%
-  as.data.frame() %>%
-  filter(presence != 5) ### removes extinct subpopulations
-
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells.csv'), progress = TRUE)
-
-### species info list - scinames, IUCN species ID, AM species ID, category etc
-spp_list_base <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
-
-### LOICZID cell ID numbers - used in plotting rasters
-loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
-loiczid_raster       <- raster(loiczid_raster_file)
-names(loiczid_raster) <- 'loiczid'
-```
-
-``` {r ggplot_theme_setups, echo = FALSE}
-
-### generic theme for all plots
-ggtheme_plot <- theme(axis.ticks = element_blank(),
-                    text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-                    plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-                    legend.position = 'right')
-
-# ggtheme_map <- theme(axis.ticks = element_blank(), axis.text = element_blank(),
-#                     text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-#                     plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-#                     legend.position = 'right')
-
-### theme for species range maps - based on generic plot theme
-ggtheme_map <- ggtheme_plot +
-  theme(axis.text = element_blank())
-
-```
-
-``` {r species_map_functions, echo = FALSE, warning = FALSE, message = FALSE}
-
-### Species Map Function ###
-# This function takes a single species scientific name as input, then grabs all 
-# occurrence cells and associated Aquamaps probability and/or IUCN proportional area
-# per cell
-get_spp_map <- function(species){
-  spp_id <- spp_list_base %>%
-    filter(sciname == species & is.na(parent_sid)) %>%
-    dplyr::select(am_sid, iucn_sid, sciname) %>%
-    unique()
-  iucn_spp_map <- iucn_spp_cells %>%
-    filter(sciname == species) %>%
-    group_by(loiczid) %>%        
-    summarize(iucn_area = max(prop_area))
-  ### The group_by() and summarize() are to collapse cells with overlapped polygons, fixing the duplicate 'by' issue
-  am_spp_map   <- am_spp_cells %>%
-    filter(am_sid == spp_id$am_sid) %>%
-    rename(am_prob = prob)
-  spp_map <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid')%>%
-    as.data.frame()
-
-  return(spp_map)
-}
-
-### Plot Species Map function
-### uses ggplot to create a formatted map of species ranges, including
-### Aquamaps, IUCN, and overlapped ranges; returns the plot object.
-plot_rangemap <- function(spp) {
-  map <- get_spp_map(spp)
-
-  am_map <- map %>%
-    filter(!is.na(am_prob))
-
-  r_am_spp  <-  subs(loiczid_raster, 
-                     am_map[ , c('loiczid', 'am_prob')], 
-                     by = 'loiczid', 
-                     which = 'am_prob', 
-                     subsWithNA = TRUE)
-  r_am_spp[!is.na(r_am_spp)] <- 1
-
-  iucn_map <- map %>%
-    mutate(iucn_pres = iucn_area > 0)
-  r_iucn_spp <- subs(loiczid_raster, 
-                     iucn_map[ , c('loiczid', 'iucn_pres')], 
-                     by = 'loiczid', 
-                     which = 'iucn_pres', 
-                     subsWithNA = TRUE)
-
-  spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
-    full_join(as.data.frame(rasterToPoints(r_am_spp)),
-              by = c('x', 'y')) %>%
-    mutate(presence = ifelse(am_prob & is.na(iucn_pres), 'am',
-                             ifelse(iucn_pres & is.na(am_prob), 'iucn',
-                                    ifelse(iucn_pres & am_prob, 'both', NA))))
-
-  spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
-    ggtheme_map + 
-    geom_raster(aes(fill = presence), alpha = .5) +
-    scale_fill_manual(values = c('iucn' = 'blue', 'am' = 'red', 'both' = 'mediumorchid')) +
-    borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
-    scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
-    scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
-    labs(title = spp, x = NULL, y = NULL) 
-}
-
-```
-***
-
-```{r create_species_list_w_area, echo = FALSE, eval = FALSE, warning = FALSE, message = FALSE}
-
-spp_list <- spp_list_base %>%
-              mutate(total_area = NA,
-                     am_area    = NA,
-                     iucn_area  = NA,
-                     perc       = NA,
-                     am_perc    = NA,
-                     iucn_perc  = NA,
-                     sm_perc    = NA, #filled with % of smaller range that is within the larger range
-                     sm_range   = NA) #what range is smaller
-
-#there's likely a better way to do this than a for loop (lapply or sapply?)
-for (i in 1:nrow(spp_list)){
-  spp     <- spp_list[i, ]$sciname #species name
-  spp_map <- get_spp_map(spp) #get species map
-
-###get just the aquamaps cells
-  am_spp_map <- spp_map %>%
-    filter(!is.na(am_prob))%>%
-    dplyr::select(loiczid, am_sid, am_prob) %>% 
-      ### add in am_prob when we want to look at probability
-    unique()
-
-  if(any(duplicated(am_spp_map$loiczid))){
-    am_spp_map <- am_spp_map %>%
-      dplyr::select(loiczid, am_sid) %>% 
-      unique()
-    }
-
-### rasterize aquamaps map
-  r_am_spp <- subs(loiczid_raster, 
-                   am_spp_map[ , c('loiczid', 'am_sid')], 
-                   by = 'loiczid', 
-                   which = 'am_sid', 
-                   subsWithNA = TRUE)
-
-  spp_list[i, ]$am_area <- area(r_am_spp, na.rm = TRUE)%>%
-              cellStats(., stat='sum')
-
-###  get the iucn cells
-  iucn_map <- spp_map %>%
-    mutate(iucn_pres = iucn_area > 0)%>%
-      dplyr::select(loiczid, iucn_area, am_sid, am_prob, iucn_pres)%>%
-        unique()
-
-  #TODO: incorporate the iucn_area (area of cell that species overlaps with...I think thats what it is)
-
-###this is incorporated because there are some duplications for loiczids with IUCN cells. I'm guessing this is because of the parent/subpopulations?
-
-   if(TRUE %in% duplicated(iucn_map$loiczid))next()
-
-### rasterize IUCN map
-  r_iucn_spp <- subs(loiczid_raster, 
-                     iucn_map[ , c('loiczid', 'iucn_pres')], 
-                     by = 'loiczid', 
-                     which = 'iucn_pres', 
-                     subsWithNA = TRUE)
-
-  spp_list[i, ]$iucn_area <- area(r_iucn_spp, na.rm = TRUE)%>%
-                cellStats(., stat='sum')
-
-### Get total area (IUCN and AM combined)
-  allCells <- spp_map%>%
-                dplyr::select(loiczid)%>%
-                  mutate(value=1)%>%
-                  unique()
-
-  fullRange <- subs(loiczid_raster, 
-                    allCells[ , c('loiczid', 'value')], 
-                    by = 'loiczid', 
-                    which = 'value', 
-                    subsWithNA = TRUE)
-
-  spp_list[i, ]$total_area <- area(fullRange, na.rm = TRUE)%>%
-                              cellStats(., stat='sum')
-
-### get percent overlap of map
-
-   #iucn map cells
-  cells_iucn<-iucn_map %>%
-    filter(iucn_pres == TRUE) %>%
-    .$loiczid
-  cells_am <- am_spp_map %>%
-    filter(!is.na(am_sid)) %>%
-    .$loiczid
-
-  cells_total <- unique(spp_map$loiczid)
-  cells_overlap <- intersect(cells_iucn, cells_am)
-
-  #I THINK THIS IS A DUPLICATE? r_iucn_spp raster created above... (JA 10/12/15)
-#    r_iucn_spp <- subs(loiczid_raster, 
-#                       iucn_map[ , c('loiczid', 'iucn_pres')], 
-#                      by = 'loiczid', 
-#                       which = 'iucn_pres', 
-#                       subsWithNA = TRUE)
-#   
-  #percent overlap
-  perc <- (length(cells_overlap) / length(cells_total)) * 100
-  spp_list[i, ]$perc <- perc
-
-  am_only <- setdiff(cells_am, cells_iucn)
-  iucn_only <- setdiff(cells_iucn, cells_am)
-
-  perc_am <- (length(am_only) / length(cells_total)) * 100
-  perc_iucn <- (length(iucn_only) / length(cells_total)) * 100
-
-  spp_list[i, ]$am_perc <- perc_am
-  spp_list[i, ]$iucn_perc <- perc_iucn
-
-  spp_list[i, ]$sm_range <- ifelse(length(cells_am)<length(cells_iucn), 'AM', 'IUCN') 
-    ### what dataset has the smaller range
-  spp_list[i, ]$sm_perc <- length(cells_overlap)/min(length(cells_am), length(cells_iucn)) * 100 
-    ### this is the total percent of smaller range encompassed in the larger range
-
-  cat(sprintf('%s: For species %s, %.2f%% of both datasets overlap, %.2f%% is just AquaMaps and %.2f%% is just IUCN\n', 
-              i, spp, perc, perc_am, perc_iucn))
-  }
-
-# for Conus arenatus, the IUCN map gives 4 duplicate cells, but each of the entries has a different iucn_area...which doesn't make sense
-
-#row 985 for Acanthurus leucopareius - AM is giving duplicate loiczids but with different am_prob...which is so strange. Looked
-# at the raw data and it seems to be in the raw data... (JA on 8/31/15)
-
-#remove non IUCN species and log total area (for use later on)
-spp_list <- spp_list %>%
-            filter(spp_group != 'hagfishes' & spp_group != 'non-homalopsids') %>%
-              mutate(log_total_area = log(total_area))
-
-write.csv(spp_list, file='spp_list_w_area_percent_noextinct.csv')
-```
-
-###(1) Number of species in both datasets split up by group
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-# look at number of species per group
-
-spp_list <- read.csv('spp_list_w_area_percent_noextinct.csv')
-
-spp_df <- spp_list %>%
-  mutate(diff = iucn_area - am_area)
-
-spp_df <- within(spp_df, spp_group <- factor(spp_group, levels = names(sort(table(spp_group), increasing = TRUE))))
-
-barchart_sp_groups <- ggplot(spp_df, aes(x = spp_group)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'bin') +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + 
-  coord_flip() +
-  labs(x = 'Species Group', 
-       y = 'Number of Species', 
-       title = 'Number of species in both AquaMaps and IUCN datasets')
-
-ggsave('figures/sp_groups_barchart.png', 
-       plot = barchart_sp_groups, 
-       width = 8, height = 4, dpi = 100)
-
-print(barchart_sp_groups)
-
-```
-
-***
-
-###(2) Looking at the total difference in area (km^2^) for each species range map in both datasets
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-scatter_am_iucn_area <- ggplot(spp_df, aes(x = iucn_area, y = am_area, color = iucn_category)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline(slope = 1, intercept = 0, color = 'red') +
-  labs(x = 'Area (km^2) from IUCN dataset', 
-       y = 'Area (km^2) from AquaMaps dataset', 
-       title = 'Comparing total area of species range maps')
-
-ggsave('figures/CompareTotalArea_AM-IUCN_scatter.png', 
-       plot = scatter_am_iucn_area, 
-       width = 8, height = 4, dpi = 100)
-
-print(scatter_am_iucn_area)
-```
-
-There are large differences so to get a better sense, here is the same comparison but with area log transformed.
-``` {r echo = FALSE, warning = FALSE, message = FALSE}
-
-spp_df <- spp_df %>%
-  mutate(am_log = log(am_area),
-         iucn_log = log(iucn_area))
-
-scatter_am_iucn_area_log <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline(slope = 1, intercept = 0, color = 'red') +
-  labs(x = 'IUCN Area (log transformed)', 
-       y = 'AquaMaps Area (log transformed)',
-       title = 'Comparing area of range maps\n (all area values are km^2 log transformed)', 
-       fill = 'IUCN Category')
-
-ggsave('figures/CompareTotalAreaLOG_AM-IUCN_scatter.png', 
-       plot = scatter_am_iucn_area_log, 
-       width = 8, height = 4, dpi = 100)
-
-print(scatter_am_iucn_area_log)
-```
-
-Data split into species groups and by IUCN category
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-scatter_am_iucn_alog_gp <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
-  ggtheme_plot + 
-  geom_point() +
-  facet_wrap(~spp_group) + 
-  geom_abline() + 
-  labs(x = 'IUCN Area (log transformed)', 
-       y = 'AquaMaps Area (log transformed)',
-       title = 'Comparing differences in area across species groups', 
-       fill = 'IUCN Category')
-
-ggsave('figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png', 
-       plot = scatter_am_iucn_alog_gp, 
-       width = 8, height = 6, dpi = 100)
-
-print(scatter_am_iucn_alog_gp)
-```
-***
-
-###(3) Looking at the overlap between species range maps from IUCN and AquaMaps
-
-This plot shows the total percent of a species range map that is (1) only covered by IUCN cells, (2) only covered by AquaMaps cells and (3) covered by both IUCN and AquaMaps cells.
-
-####How this breaks down between AquaMaps and IUCN
-
-Here is an example of a single species, the yellow speckled chromis (*Chromis alpha*)
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-spp <- 'Chromis alpha'
-
-spp_map <- plot_rangemap(spp)
-
-print(spp_map)
-ggsave(plot = spp_map,
-       filename = sprintf('figures/rangemap_%s.png', spp),
-       width = 8, height = 4, dpi = 200)
-```
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-  spp_df2 <- gather(spp_list, perc.type, percent, perc, am_perc, iucn_perc) %>%
-    dplyr::select(sciname, spp_group, perc.type, percent) %>%
-    unique()
-
-  perc_c <- spp_df2 %>%
-    filter(sciname == spp)
-
-  p1 <- perc_c %>% 
-    dplyr::select(percent)
-
-```
-
-For `r spp`, `r round(pi[1], 2)`% of the total area is covered by both IUCN and AquaMaps, `r round(p1[2, ], 2)`% is covered just by AquaMaps and `r round(p1[3, ], 2)`% is covered just by IUCN.
-
-***
-
-#### Now looking across all species, here is the breakdown of how much overlap there is between the two datasets
-
-This plot shows that the majority of species range maps have only ~30% of their full map covered by both maps, and very few species have more than 50% of their two ranges (IUCN and AquaMaps) overlapping.
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-### look at density plot
-
-overlap <- spp_df2 %>%
-  dplyr::filter(perc.type == 'perc')
-
-hist_overlap_pct <- ggplot(overlap, aes(percent)) +
-  ggtheme_plot + 
-  geom_histogram(aes(y = ..count../1926), binwidth = 5, alpha = .8) +
-  labs(title = 'Percent of total range shared between datasets',
-       x = '% total range that is shared b/w both datasets', 
-       y = 'Proportion of Total Species')
-
-ggsave('figures/Overlap-PercTotalRange.png')
-
-print(hist_overlap_pct)
-```
-
-Split up by species groups
-
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
-
-hist_overlap_pct_gp <- ggplot(overlap, aes(percent)) + 
-  ggtheme_plot + 
-  geom_histogram(binwidth = 5, alpha = .8) +
-  facet_wrap(~spp_group) +
-  labs(title = 'Percent of total range shared, by species group',
-       x = 'Percent of total range that is shared between both datasets', 
-       y = 'Number of Total Species') + 
-  theme(strip.text.x = element_text(size = 8))
-
-ggsave('figures/Overlap-PercTotalRange_spGroups.png')
-print(hist_overlap_pct_gp)
-
-```
-
-###(4) Looking at the percent of smaller species range that is within the larger range.
-
-For each range
-
-```{r pct small within large, echo = FALSE, warning = FALSE}
-
-spp_df4 <- spp_list %>%
-  dplyr::select(sciname, sm_perc, sm_range) %>%
-  unique()
-
-hist_sm_within_lg <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
-  ggtheme_plot + 
-  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
-  scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
-  labs(title = 'Percent of smaller range within larger range',
-       x = 'Percent of smaller range that falls within the larger range', 
-       y = sprintf('Proportion of all species (n = %s)', nrow(spp_df4)))
-
-ggsave('figures/PercSmallRangeInLarger.png', width = 6, height = 4)
-
-print(hist_sm_within_lg)
-
-```
-
-Exploring possible correlation between total species range area and percent overlap of smaller range within the larger:
-``` {r pct overlap vs area, echo = FALSE, warning = FALSE, message = FALSE}
-
-scatter_area_vs_overlap <- ggplot(spp_list, aes(sm_perc, total_area, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline() + 
-  labs(y = 'Total Area (km^2)', 
-       x = 'Percent overlap smaller range in larger', 
-       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
-
-lm_eqn <- function(df){
-  m  <- lm(total_area ~ sm_perc, spp_list)
-  eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
-        list(a  = format(coef(m)[1], digits = 2), 
-             b  = format(coef(m)[2], digits = 2), 
-             r2 = format(summary(m)$r.squared, digits = 3)))
-  as.character(as.expression(eq))             
-}
-
-scatter_area_vs_overlap <- scatter_area_vs_overlap + 
-  geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)
-
-print(scatter_area_vs_overlap)
-
-# summary(lm(total_area ~ sm_perc, spp_list))
-
-ggsave('figures/AreavsPercOverlap.png', width = 6, height = 4)
-
-```
-
-``` {r pct overlap vs ratio of area, echo = FALSE, warning = FALSE, message = FALSE}
-
-spp_list <- spp_list %>%
-  mutate(area_ratio = am_area / iucn_area,
-         area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
-         area_ratio = area_ratio * 100)
-
-area_ratio_quartiles <- quantile(spp_list$area_ratio, na.rm = TRUE)[2:4]
-pct_overlap_quartiles <- quantile(spp_list$sm_perc, na.rm = TRUE)[2:4]
-
-spp_quads <- c('q1' = 'Kajikia albida', 
-               'q2' = 'Galaxea cryptoramosa', 
-               'q4' = 'Acanthurus nigroris', 
-               'q4a' = 'Metanephrops australiensis', 
-               'q3' = 'Conus magnificus')
-
-spp_list_label <- spp_list %>%
-  filter(sciname %in% spp_quads) %>%
-  separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
-  mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
-  select(-g, -s)
-
-lm_eqn2 <- function(df){
-  m  <- lm(area_ratio ~ sm_perc, spp_list)
-  eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
-        list(a  = format(coef(m)[1], digits = 2), 
-             b  = format(coef(m)[2], digits = 2), 
-             r2 = format(summary(m)$r.squared, digits = 3)))
-  as.character(as.expression(eq))             
-}
-
-scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = log_total_area)) + 
-  ggtheme_plot + 
-  ### color the quadrant backgrounds:
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
-  geom_point() + 
-  geom_vline(x = area_ratio_quartiles[c(1, 3)],
-             color = 'grey50', linetype = 'dashed') +
-  geom_hline(y = pct_overlap_quartiles[c(1, 3)],
-             color = 'grey50', linetype = 'dashed') +
-  geom_point(data = spp_list_label, color = 'tomato2') + 
-  geom_text(data = spp_list_label, 
-            aes(label = sciname_label), 
-            color = 'tomato2',
-            fontface = 'bold.italic',
-            size = 2,
-            hjust = .25,
-            vjust = 1.5) +
-  labs(x = 'Range area ratio (smaller:larger) * 100%', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
-
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap + 
-  geom_text(x = 40, y = 5, label = lm_eqn2(spp_list), parse = TRUE, color = 'tomato2')
-
-ggsave('figures/AreaRatiovsPercOverlap.png', width = 6, height = 4)
-
-print(scatter_arearatio_vs_overlap)
-
-# scatter_arearatio_vs_overlap2 <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = sm_range)) + 
-#   ggtheme_plot + 
-#   geom_point() + 
-#   geom_vline(x = area_ratio_quartiles,
-#              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = pct_overlap_quartiles,
-#              color = 'grey50', linetype = 'dashed') +
-#   geom_point(data = spp_list_label, color = 'grey30') + 
-#   geom_text(data = spp_list_label, 
-#             aes(label = sciname_label), 
-#             color = 'grey30',
-#             fontface = 'bold.italic',
-#             size = 2,
-#             hjust = .25,
-#             vjust = 1.5) +
-#   labs(x = 'Range area ratio (smaller:larger) * 100%', 
-#        y = 'Percent overlap smaller range in larger', 
-#        title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
-# 
-# scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 + 
-#   geom_text(x = 40, y = 5, label = lm_eqn2(spp_list), parse = TRUE, color = 'grey30')
-# 
-# ggsave('figures/AreaRatiovsPercOverlap2.png', 
-#        width = 6, height = 4)
-# 
-# print(scatter_arearatio_vs_overlap2)
-
-# summary(lm(area_ratio ~ sm_perc, spp_list))
-
-```
-
-``` {r examine_spp_gp_quartiles, echo = FALSE}
-
-spp_list1 <- spp_list %>%
-  group_by(spp_group) %>%
-  mutate(n_spp = n()) %>%
-  ungroup()
-
-spp_gp_q1 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q1 = n())
-spp_gp_q2 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q2 = n())
-spp_gp_q4 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q4 = n())
-spp_gp_q3 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q3 = n())
-
-spp_gp_quadrants <- spp_gp_q1 %>%
-  left_join(spp_gp_q2, by = c('spp_group', 'n_spp')) %>%
-  left_join(spp_gp_q3, by = c('spp_group', 'n_spp')) %>%
-  left_join(spp_gp_q4, by = c('spp_group', 'n_spp')) %>%
-  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
-  mutate(quad = str_replace(quad, 'n_spp_', ''),
-         pct_quad = n_quad/n_spp)
-
-spp_gp_quadrants <- spp_gp_quadrants %>%
-  left_join(spp_gp_quadrants %>% 
-              filter(quad == 'q1') %>%
-              select(spp_group, pct_q1 = pct_quad),
-            by = 'spp_group')
-
-barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
-                                  transform(spp_group = reorder(spp_group, pct_q1)), 
-                                aes(x = spp_group, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'bin', alpha = .5) +
-  scale_fill_manual(values = c('q1' = 'green4', 'q2' = 'greenyellow', 'q3' = 'orange', 'q4' = 'orangered2')) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0) +
-  coord_flip() +
-  labs(x = 'Species Group', 
-       y = 'Relative number of species per quadrant', 
-       title = 'Species group breakdown by quadrant')
-
-print(barchart_spp_gp_quads)
-ggsave(plot = barchart_spp_gp_quads,
-       filename = 'figures/barchart_spp_gp_quads.png',
-       width = 6, height = 4)
-
-```
-
-```{r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE}
-
-for (i in 1:length(spp_quads)) { # i = 1
-  spp <- spp_quads[i]
-  spp_quad <- names(spp_quads)[i]
-
-  spp_plot <- plot_rangemap(spp)
-  print(spp_plot)
-
-  ggsave(filename = sprintf('figures/rangemap_%s_%s.png', spp_quad, str_replace(tolower(spp), ' ', '_')),
-         plot = spp_plot, 
-         width = 8, height = 4, dpi = 200)
-}
-
-```
-
-***
-
-Percent coverage as function of area, clipping to area below 100,000,000 km^2^:
-
-``` {r, echo = FALSE, warning = FALSE, message = FALSE}
-# remove all area greater than 100000000
-
-spp_small_df <- spp_list %>% 
-  filter(total_area < 100000000)
-
-scatter_a_vs_overlap_small <- ggplot(spp_small_df, aes(total_area, sm_perc, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  labs(x = 'Total Area (km^2)', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreavsPercOverlap_2.png', width = 6, height = 4)
-
-print(scatter_a_vs_overlap_small)
-
-```
-
-Percent coverage as a function of log(area)
-``` {r, echo = FALSE, warning = FALSE, message = FALSE}
-# Log transform area
-
-scatter_logA_vs_overlap <- ggplot(spp_list, aes(log_total_area, sm_perc, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  labs(x = 'Log(total Area)', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Size of species range (log) vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreavsPercOverlap_log.png', width=6, height=4)
-
-print(scatter_logA_vs_overlap)
-```
-
-``` {r examine regressions, eval = FALSE, echo = FALSE, warning = FALSE}
-
-A_mdl      <- lm(total_area ~ sm_perc, spp_list)
-#                Estimate Std. Error t value Pr(>|t|)    
-#   (Intercept) -16285410    2858388  -5.697  1.4e-08 ***
-#   sm_perc        533370      41702  12.790  < 2e-16 ***
-#   Multiple R-squared:  0.07732,	Adjusted R-squared:  0.07685
-
-A_mdl_clip <- lm(total_area ~ sm_perc, l)
-#               Estimate Std. Error t value Pr(>|t|)    
-#   (Intercept)  2993939     810560   3.694 0.000227 ***
-#   sm_perc       153523      11952  12.844  < 2e-16 ***
-#   Multiple R-squared:  0.07955,	Adjusted R-squared:  0.07906 
-logA_mdl   <- lm(log_total_area ~ sm_perc, spp_list)
-#                Estimate Std. Error t value Pr(>|t|)    
-#   (Intercept) 14.775325   0.096446   153.2   <2e-16 ***
-#   sm_perc      0.016885   0.001407    12.0   <2e-16 ***
-#   Multiple R-squared:  0.0687,	Adjusted R-squared:  0.06823 
-```
-
-***
-
-###(5) Examining data broken into quartiles to see if smaller patterns emerge
-``` {r examine quartiles, eval = FALSE, echo = FALSE, warning = FALSE}
-
-summary(spp_list$total_area)
-#      Min.   1st Qu.    Median      Mean   3rd Qu.      Max.
-#     55760   3032000  10790000  18500000  22670000 372100000
-
-qtiles <- c(quantile(spp_list$total_area, 0.0, na.rm = TRUE),
-            quantile(spp_list$total_area, .25, na.rm = TRUE),
-            quantile(spp_list$total_area, .50, na.rm = TRUE),
-            quantile(spp_list$total_area, .75, na.rm = TRUE),
-            quantile(spp_list$total_area, 1.0, na.rm = TRUE))
-
-qtile_mdl <- list()
-for (i in 1:4) {
-  area_clip <- spp_list %>%
-    filter(total_area >= qtiles[i] & total_area < qtiles[i + 1])
-  qtile_mdl[[i]] <- lm(total_area ~ sm_perc, area_clip)
-
-  qtile_plot <- ggplot(area_clip, aes(total_area, sm_perc, color = sm_range)) + 
-    ggtheme_plot + 
-    geom_point() + 
-    labs(x = 'Total Area (km^2)', 
-         y = 'Percent overlap smaller range in larger', 
-         title = sprintf('Size of species range (quartile %s) vs. percent overlap of \nsmaller range with larger', i))
-
-  ggsave(paste0('figures/AreavsOverlap_quantile_', i, '.png'), width = 6, height = 4)
-  print(qtile_plot)
-}
-# First quartile regression:
-#                 Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept)  1423917     101053  14.091   <2e-16 ***
-#     sm_perc         1188       1555   0.764    0.445    
-#     Multiple R-squared:  0.001198,	Adjusted R-squared:  -0.0008529 
-#
-# second quartile:
-#                 Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept)  7011028     275943  25.408  < 2e-16 ***
-#     sm_perc       -17725       4228  -4.192 3.28e-05 ***
-#     Multiple R-squared:  0.03483,	Adjusted R-squared:  0.03285 
-#
-# third quartile:
-#                 Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept) 12813279     507856  25.230  < 2e-16 ***
-#     sm_perc        64710       8159   7.931  1.5e-14 ***
-#     Multiple R-squared:  0.1146,	Adjusted R-squared:  0.1128 
-#
-# fourth quartile:
-#                   Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept) -101736351   17259544  -5.894 7.04e-09 ***
-#     sm_perc        1911286     215216   8.881  < 2e-16 ***
-#     Multiple R-squared:  0.1399,	Adjusted R-squared:  0.1381
-
-# This avenue of exploration looks like a bust.
-
-```
-
-***
-
 ###(6) Looking at how many and what type (threat category) of species we are losing by setting a threshold of 0.4 on the aquamaps species
@@
 
-ggsave('figures/AMThreshold.png', width = 6, height = 4)
+ggsave('../figures/AMThreshold.png', width = 6, height = 4)
 
@@
 
-ggsave('figures/AMThreshold2.png', width = 6, height = 4)
+ggsave('../figures/AMThreshold2.png', width = 6, height = 4)
 
@@
 
-ggsave('figures/scatter_AMThreshold.png', width = 6, height = 4)
+ggsave('../figures/scatter_AMThreshold.png', width = 6, height = 4)
 
@@
 ```
-
-## Changes in SPP for scenarios
-
-Comparing three scenarios to the current (v2015) global model for SPP:
-
-* Baseline is: preferring IUCN rangemaps over Aquamaps rangemaps; Aquamaps 'presence' threshold is set at 40% or greater.
-* Scenario 1: preferring IUCN rangemaps over Aquamaps; AM presence threshold set at >= 1% (effectively, all occurrences).
-* Scenario 2: preferring AM maps over IUCN; AM presence >= 40%.
-* Scenario 3: preferring AM over IUCN; AM presence >= 1%.
-
-Boxplots:
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
-
-![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)

### 2015-11-21 R/data_explore_pt1.Rmd
trying out new maps by quadrant to see if there are common characteristics...

--- 2015-11-18 data_explore.Rmd
@@
----
-title: 'Data Exploration: IUCN and AquaMaps'
-output: html_document
----
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-# Libraries and Paths
-
-library(ggplot2)
-library(maps)
-library(readr)
-library(raster)
-library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-
-dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-           'Darwin'  = '/Volumes/data_edit',
-           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-```
-
-```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
-
-# Data files
-#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), 
-                           col_types = 'cdddd', progress = TRUE) %>%
-  as.data.frame() %>%
-  filter(presence != 5) ### removes extinct subpopulations
-
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells.csv'), progress = TRUE)
-
-### species info list - scinames, IUCN species ID, AM species ID, category etc
-spp_list_base <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
-
-### LOICZID cell ID numbers - used in plotting rasters
-loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
-loiczid_raster       <- raster(loiczid_raster_file)
-names(loiczid_raster) <- 'loiczid'
-```
-
-``` {r ggplot_theme_setups, echo = FALSE}
-
-### generic theme for all plots
-ggtheme_plot <- theme(axis.ticks = element_blank(),
-                    text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-                    plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-                    legend.position = 'right')
-
-# ggtheme_map <- theme(axis.ticks = element_blank(), axis.text = element_blank(),
-#                     text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-#                     plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-#                     legend.position = 'right')
-
-### theme for species range maps - based on generic plot theme
-ggtheme_map <- ggtheme_plot +
-  theme(axis.text = element_blank())
-
-```
-
-``` {r species_map_functions, echo = FALSE, warning = FALSE, message = FALSE}
-
-### Species Map Function ###
-# This function takes a single species scientific name as input, then grabs all 
-# occurrence cells and associated Aquamaps probability and/or IUCN proportional area
-# per cell
-get_spp_map <- function(species){
-  spp_id <- spp_list_base %>%
-    filter(sciname == species & is.na(parent_sid)) %>%
-    dplyr::select(am_sid, iucn_sid, sciname) %>%
-    unique()
-  iucn_spp_map <- iucn_spp_cells %>%
-    filter(sciname == species) %>%
-    group_by(loiczid) %>%        
-    summarize(iucn_area = max(prop_area))
-  ### The group_by() and summarize() are to collapse cells with overlapped polygons, fixing the duplicate 'by' issue
-  am_spp_map   <- am_spp_cells %>%
-    filter(am_sid == spp_id$am_sid) %>%
-    rename(am_prob = prob)
-  spp_map <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid')%>%
-    as.data.frame()
-
-  return(spp_map)
-}
-
-### Plot Species Map function
-### uses ggplot to create a formatted map of species ranges, including
-### Aquamaps, IUCN, and overlapped ranges; returns the plot object.
-plot_rangemap <- function(spp) {
-  map <- get_spp_map(spp)
-
-  am_map <- map %>%
-    filter(!is.na(am_prob))
-
-  r_am_spp  <-  subs(loiczid_raster, 
-                     am_map[ , c('loiczid', 'am_prob')], 
-                     by = 'loiczid', 
-                     which = 'am_prob', 
-                     subsWithNA = TRUE)
-  r_am_spp[!is.na(r_am_spp)] <- 1
-
-  iucn_map <- map %>%
-    mutate(iucn_pres = iucn_area > 0)
-  r_iucn_spp <- subs(loiczid_raster, 
-                     iucn_map[ , c('loiczid', 'iucn_pres')], 
-                     by = 'loiczid', 
-                     which = 'iucn_pres', 
-                     subsWithNA = TRUE)
-
-  spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
-    full_join(as.data.frame(rasterToPoints(r_am_spp)),
-              by = c('x', 'y')) %>%
-    mutate(presence = ifelse(am_prob & is.na(iucn_pres), 'am',
-                             ifelse(iucn_pres & is.na(am_prob), 'iucn',
-                                    ifelse(iucn_pres & am_prob, 'both', NA))))
-
-  spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
-    ggtheme_map + 
-    geom_raster(aes(fill = presence), alpha = .5) +
-    scale_fill_manual(values = c('iucn' = 'blue', 'am' = 'red', 'both' = 'mediumorchid')) +
-    borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
-    labs(title = spp, x = NULL, y = NULL) 
-}
-
-```
-***
-
-```{r create_species_list_w_area, echo = FALSE, eval = FALSE, warning = FALSE, message = FALSE}
-
-spp_list <- spp_list_base %>%
-              mutate(total_area = NA,
-                     am_area    = NA,
-                     iucn_area  = NA,
-                     perc       = NA,
-                     am_perc    = NA,
-                     iucn_perc  = NA,
-                     sm_perc    = NA, #filled with % of smaller range that is within the larger range
-                     sm_range   = NA) #what range is smaller
-
-#there's likely a better way to do this than a for loop (lapply or sapply?)
-for (i in 1:nrow(spp_list)){
-  spp     <- spp_list[i, ]$sciname #species name
-  spp_map <- get_spp_map(spp) #get species map
-
-###get just the aquamaps cells
-  am_spp_map <- spp_map %>%
-    filter(!is.na(am_prob))%>%
-    dplyr::select(loiczid, am_sid, am_prob) %>% 
-      ### add in am_prob when we want to look at probability
-    unique()
-
-  if(any(duplicated(am_spp_map$loiczid))){
-    am_spp_map <- am_spp_map %>%
-      dplyr::select(loiczid, am_sid) %>% 
-      unique()
-    }
-
-### rasterize aquamaps map
-  r_am_spp <- subs(loiczid_raster, 
-                   am_spp_map[ , c('loiczid', 'am_sid')], 
-                   by = 'loiczid', 
-                   which = 'am_sid', 
-                   subsWithNA = TRUE)
-
-  spp_list[i, ]$am_area <- area(r_am_spp, na.rm = TRUE)%>%
-              cellStats(., stat='sum')
-
-###  get the iucn cells
-  iucn_map <- spp_map %>%
-    mutate(iucn_pres = iucn_area > 0)%>%
-      dplyr::select(loiczid, iucn_area, am_sid, am_prob, iucn_pres)%>%
-        unique()
-
-  #TODO: incorporate the iucn_area (area of cell that species overlaps with...I think thats what it is)
-
-###this is incorporated because there are some duplications for loiczids with IUCN cells. I'm guessing this is because of the parent/subpopulations?
-
-   if(TRUE %in% duplicated(iucn_map$loiczid))next()
-
-### rasterize IUCN map
-  r_iucn_spp <- subs(loiczid_raster, 
-                     iucn_map[ , c('loiczid', 'iucn_pres')], 
-                     by = 'loiczid', 
-                     which = 'iucn_pres', 
-                     subsWithNA = TRUE)
-
-  spp_list[i, ]$iucn_area <- area(r_iucn_spp, na.rm = TRUE)%>%
-                cellStats(., stat='sum')
-
-### Get total area (IUCN and AM combined)
-  allCells <- spp_map%>%
-                dplyr::select(loiczid)%>%
-                  mutate(value=1)%>%
-                  unique()
-
-  fullRange <- subs(loiczid_raster, 
-                    allCells[ , c('loiczid', 'value')], 
-                    by = 'loiczid', 
-                    which = 'value', 
-                    subsWithNA = TRUE)
-
-  spp_list[i, ]$total_area <- area(fullRange, na.rm = TRUE)%>%
-                              cellStats(., stat='sum')
-
-### get percent overlap of map
-
-   #iucn map cells
-  cells_iucn<-iucn_map %>%
-    filter(iucn_pres == TRUE) %>%
-    .$loiczid
-  cells_am <- am_spp_map %>%
-    filter(!is.na(am_sid)) %>%
-    .$loiczid
-
-  cells_total <- unique(spp_map$loiczid)
-  cells_overlap <- intersect(cells_iucn, cells_am)
-
-  #I THINK THIS IS A DUPLICATE? r_iucn_spp raster created above... (JA 10/12/15)
-#    r_iucn_spp <- subs(loiczid_raster, 
-#                       iucn_map[ , c('loiczid', 'iucn_pres')], 
-#                      by = 'loiczid', 
-#                       which = 'iucn_pres', 
-#                       subsWithNA = TRUE)
-#   
-  #percent overlap
-  perc <- (length(cells_overlap) / length(cells_total)) * 100
-  spp_list[i, ]$perc <- perc
-
-  am_only <- setdiff(cells_am, cells_iucn)
-  iucn_only <- setdiff(cells_iucn, cells_am)
-
-  perc_am <- (length(am_only) / length(cells_total)) * 100
-  perc_iucn <- (length(iucn_only) / length(cells_total)) * 100
-
-  spp_list[i, ]$am_perc <- perc_am
-  spp_list[i, ]$iucn_perc <- perc_iucn
-
-  spp_list[i, ]$sm_range <- ifelse(length(cells_am)<length(cells_iucn), 'AM', 'IUCN') 
-    ### what dataset has the smaller range
-  spp_list[i, ]$sm_perc <- length(cells_overlap)/min(length(cells_am), length(cells_iucn)) * 100 
-    ### this is the total percent of smaller range encompassed in the larger range
-
-  cat(sprintf('%s: For species %s, %.2f%% of both datasets overlap, %.2f%% is just AquaMaps and %.2f%% is just IUCN\n', 
-              i, spp, perc, perc_am, perc_iucn))
-  }
-
-# for Conus arenatus, the IUCN map gives 4 duplicate cells, but each of the entries has a different iucn_area...which doesn't make sense
-
-#row 985 for Acanthurus leucopareius - AM is giving duplicate loiczids but with different am_prob...which is so strange. Looked
-# at the raw data and it seems to be in the raw data... (JA on 8/31/15)
-
-#remove non IUCN species and log total area (for use later on)
-spp_list <- spp_list %>%
-            filter(spp_group != 'hagfishes' & spp_group != 'non-homalopsids') %>%
-              mutate(log_total_area = log(total_area))
-
-write.csv(spp_list, file='spp_list_w_area_percent_noextinct.csv')
-```
-
 ###(1) Number of species in both datasets split up by group
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-# look at number of species per group
-
-spp_list <- read.csv('spp_list_w_area_percent_noextinct.csv')
-
+``` {r}
 spp_df <- spp_list %>%
@@
 spp_df <- within(spp_df, spp_group <- factor(spp_group, levels = names(sort(table(spp_group), increasing = TRUE))))
+
+```
+
+```{r number of spp per group, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+
+### look at number of species per group
 
@@
 
-ggsave('figures/sp_groups_barchart.png', 
+ggsave(file.path(dir_fig, 'sp_groups_barchart.png'), 
        plot = barchart_sp_groups, 
@@
 
-###(2) Looking at the total difference in area (km^2) for each species range map in both datasets
+###(2) Looking at the total difference in area (km^2^) for each species range map in both datasets
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
+```{r area compare 1, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
@@
 
-ggsave('figures/CompareTotalArea_AM-IUCN_scatter.png', 
+ggsave(file.path(dir_fig, 'CompareTotalArea_AM-IUCN_scatter.png'), 
        plot = scatter_am_iucn_area, 
@@
 There are large differences so to get a better sense, here is the same comparison but with area log transformed.
-``` {r echo = FALSE, warning = FALSE, message = FALSE}
+``` {r area compare log, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 
-ggsave('figures/CompareTotalAreaLOG_AM-IUCN_scatter.png', 
+ggsave(file.path(dir_fig, 'CompareTotalAreaLOG_AM-IUCN_scatter.png'), 
        plot = scatter_am_iucn_area_log, 
@@
 Data split into species groups and by IUCN category
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
+```{r area compare by spp group, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 
-ggsave('figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png', 
+ggsave(file.path(dir_fig, 'CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png'), 
        plot = scatter_am_iucn_alog_gp, 
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
+```{r map chromis alpha, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
@@
 ggsave(plot = spp_map,
-       filename = sprintf('figures/rangemap_%s.png', spp),
+       filename = file.path(dir_fig, sprintf('rangemap_%s.png', spp)),
        width = 8, height = 4, dpi = 200)
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
+```{r, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
@@
 
-####Now looking across all species, here is the breakdown of how much overlap there is between the two datasets
+  #### Now looking across all species, here is the breakdown of how much overlap there is between the two datasets
 
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
+```{r hist overlap vs total range, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 
-ggsave('figures/Overlap-PercTotalRange.png')
+ggsave(file.path(dir_fig, 'Overlap-PercTotalRange.png'))
 
@@
 
-Split up by species - not sure if this is useful or not
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
+Split up by species groups
+
+```{r facet by spp, echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE}
 
@@
 
-ggsave('figures/Overlap-PercTotalRange_spGroups.png')
+ggsave(file.path(dir_fig, 'Overlap-PercTotalRange_spGroups.png'))
 print(hist_overlap_pct_gp)
@@
 ```
-
-###(4) Looking at the percent of smaller species range that is within the larger range.
-
-For each range
-
-```{r pct small within large, echo = FALSE, warning = FALSE}
-
-spp_df4 <- spp_list %>%
-  dplyr::select(sciname, sm_perc, sm_range) %>%
-  unique()
-
-hist_sm_within_lg <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
-  ggtheme_plot + 
-  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
-  scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
-  labs(title = 'Percent of smaller range within larger range',
-       x = 'Percent of smaller range that falls within the larger range', 
-       y = sprintf('Proportion of all species (n = %s)', nrow(spp_df4)))
-
-ggsave('figures/PercSmallRangeInLarger.png', width = 6, height = 4)
-
-print(hist_sm_within_lg)
-
-```
-
-Exploring possible correlation between total species range area and percent overlap of smaller range within the larger:
-``` {r pct overlap vs area, echo = FALSE, warning = FALSE, message = FALSE}
-
-scatter_area_vs_overlap <- ggplot(spp_list, aes(sm_perc, total_area, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline() + 
-  labs(y = 'Total Area (km^2)', 
-       x = 'Percent overlap smaller range in larger', 
-       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreavsPercOverlap.png', width = 6, height = 4)
-
-print(scatter_area_vs_overlap)
-
-```
-
-``` {r pct overlap vs ratio of area, echo = FALSE, warning = FALSE, message = FALSE}
-
-spp_list <- spp_list %>%
-  mutate(area_ratio = am_area / iucn_area,
-         area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
-         area_ratio = area_ratio * 100)
-
-area_ratio_quartiles <- quantile(spp_list$area_ratio, na.rm = TRUE)[2:4]
-pct_overlap_quartiles <- quantile(spp_list$sm_perc, na.rm = TRUE)[2:4]
-
-spp_quads <- c('q1' = 'Kajikia albida', 
-               'q2' = 'Galaxea cryptoramosa', 
-               'q4' = 'Acanthurus nigroris', 
-               'q4a' = 'Metanephrops australiensis', 
-               'q3' = 'Conus magnificus')
-
-spp_list_label <- spp_list %>%
-  filter(sciname %in% spp_quads) %>%
-  separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
-  mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
-  select(-g, -s)
-
-scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = log_total_area)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline() + 
-  geom_vline(x = area_ratio_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_hline(y = pct_overlap_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_point(data = spp_list_label, color = 'tomato2') + 
-  geom_text(data = spp_list_label, 
-            aes(label = sciname_label), 
-            color = 'tomato2',
-            fontface = 'bold.italic',
-            size = 2,
-            hjust = .25,
-            vjust = 1.5) +
-  labs(x = 'Range area ratio (smaller:larger) * 100%', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreaRatiovsPercOverlap.png', width = 6, height = 4)
-
-print(scatter_arearatio_vs_overlap)
-
-scatter_arearatio_vs_overlap2 <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline() + 
-  geom_vline(x = area_ratio_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_hline(y = pct_overlap_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_point(data = spp_list_label, color = 'grey30') + 
-  geom_text(data = spp_list_label, 
-            aes(label = sciname_label), 
-            color = 'grey30',
-            fontface = 'bold.italic',
-            size = 2,
-            hjust = .25,
-            vjust = 1.5) +
-  labs(x = 'Range area ratio (smaller:larger) * 100%', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreaRatiovsPercOverlap2.png', width = 6, height = 4)
-
-print(scatter_arearatio_vs_overlap2)
-
-```
-
-``` {r examine_spp_gp_quartiles}
-
-spp_list1 <- spp_list %>%
-  group_by(spp_group) %>%
-  mutate(n_spp = n()) %>%
-  ungroup()
-
-spp_gp_q1 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q1 = n())
-spp_gp_q2 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q2 = n())
-spp_gp_q4 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q4 = n())
-spp_gp_q3 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q3 = n())
-
-spp_gp_quadrants <- spp_gp_q1 %>%
-  left_join(spp_gp_q2, by = c('spp_group', 'n_spp')) %>%
-  left_join(spp_gp_q3, by = c('spp_group', 'n_spp')) %>%
-  left_join(spp_gp_q4, by = c('spp_group', 'n_spp')) %>%
-  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
-  mutate(quad = str_replace(quad, 'n_spp_', ''),
-         pct_quad = n_quad/n_spp)
-
-spp_gp_quadrants <- spp_gp_quadrants %>%
-  left_join(spp_gp_quadrants %>% 
-              filter(quad == 'q1') %>%
-              select(spp_group, pct_q1 = pct_quad),
-            by = 'spp_group')
-
-barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
-                                  transform(spp_group = reorder(spp_group, pct_q1)), 
-                                aes(x = spp_group, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'bin') +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0) +
-  coord_flip() +
-  labs(x = 'Species Group', 
-       y = 'Relative number of species per quadrant', 
-       title = 'Species group breakdown by quadrant')
-
-```
-
-```{r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE}
-
-for (i in 1:length(spp_quads)) { # i = 1
-  spp <- spp_quads[i]
-  spp_quad <- names(spp_quads)[i]
-
-  spp_plot <- plot_rangemap(spp)
-  print(spp_plot)
-
-  ggsave(filename = sprintf('figures/rangemap_%s_%s.png', spp_quad, str_replace(tolower(spp), ' ', '_')),
-         plot = spp_plot, 
-         width = 8, height = 4, dpi = 200)
-}
-
-```
-
-``` {r, echo = FALSE, warning = FALSE, message = FALSE}
-
-# Regression
-
-lm_eqn <- function(df){
-  m  <- lm(total_area ~ sm_perc, spp_list)
-  eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
-        list(a  = format(coef(m)[1], digits = 2), 
-             b  = format(coef(m)[2], digits = 2), 
-             r2 = format(summary(m)$r.squared, digits = 3)))
-  as.character(as.expression(eq))             
-}
-
-scatter_area_vs_overlap <- scatter_area_vs_overlap + 
-  geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)
-
-ggsave('figures/CorrelationAreavsPercOverlap.png')
-
-print(scatter_area_vs_overlap)
-
-summary(lm(total_area ~ sm_perc, spp_list))
-```
-
-***
-
-Percent coverage as function of area, clipping to area below 100,000,000 km^2:
-``` {r, echo = FALSE, warning = FALSE, message = FALSE}
-# remove all area greater than 100000000
-
-spp_small_df <- spp_list %>% 
-  filter(total_area < 100000000)
-
-scatter_a_vs_overlap_small <- ggplot(spp_small_df, aes(total_area, sm_perc, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  labs(x = 'Total Area (km^2)', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreavsPercOverlap_2.png', width = 6, height = 4)
-
-print(scatter_a_vs_overlap_small)
-
-```
-
-Percent coverage as a function of log(area)
-``` {r, echo = FALSE, warning = FALSE, message = FALSE}
-# Log transform area
-
-scatter_logA_vs_overlap <- ggplot(spp_list, aes(log_total_area, sm_perc, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  labs(x = 'Log(total Area)', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Size of species range (log) vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreavsPercOverlap_log.png', width=6, height=4)
-
-print(scatter_logA_vs_overlap)
-```
-
-``` {r examine regressions, eval = FALSE, echo = FALSE, warning = FALSE}
-
-A_mdl      <- lm(total_area ~ sm_perc, spp_list)
-#                Estimate Std. Error t value Pr(>|t|)    
-#   (Intercept) -16285410    2858388  -5.697  1.4e-08 ***
-#   sm_perc        533370      41702  12.790  < 2e-16 ***
-#   Multiple R-squared:  0.07732,	Adjusted R-squared:  0.07685
-
-A_mdl_clip <- lm(total_area ~ sm_perc, l)
-#               Estimate Std. Error t value Pr(>|t|)    
-#   (Intercept)  2993939     810560   3.694 0.000227 ***
-#   sm_perc       153523      11952  12.844  < 2e-16 ***
-#   Multiple R-squared:  0.07955,	Adjusted R-squared:  0.07906 
-logA_mdl   <- lm(log_total_area ~ sm_perc, spp_list)
-#                Estimate Std. Error t value Pr(>|t|)    
-#   (Intercept) 14.775325   0.096446   153.2   <2e-16 ***
-#   sm_perc      0.016885   0.001407    12.0   <2e-16 ***
-#   Multiple R-squared:  0.0687,	Adjusted R-squared:  0.06823 
-```
-
-***
-
-###(5) Examining data broken into quartiles to see if smaller patterns emerge
-``` {r examine quartiles, eval = FALSE, echo = FALSE, warning = FALSE}
-
-summary(spp_list$total_area)
-#      Min.   1st Qu.    Median      Mean   3rd Qu.      Max.
-#     55760   3032000  10790000  18500000  22670000 372100000
-
-qtiles <- c(quantile(spp_list$total_area, 0.0, na.rm = TRUE),
-            quantile(spp_list$total_area, .25, na.rm = TRUE),
-            quantile(spp_list$total_area, .50, na.rm = TRUE),
-            quantile(spp_list$total_area, .75, na.rm = TRUE),
-            quantile(spp_list$total_area, 1.0, na.rm = TRUE))
-
-qtile_mdl <- list()
-for (i in 1:4) {
-  area_clip <- spp_list %>%
-    filter(total_area >= qtiles[i] & total_area < qtiles[i + 1])
-  qtile_mdl[[i]] <- lm(total_area ~ sm_perc, area_clip)
-
-  qtile_plot <- ggplot(area_clip, aes(total_area, sm_perc, color = sm_range)) + 
-    ggtheme_plot + 
-    geom_point() + 
-    labs(x = 'Total Area (km^2)', 
-         y = 'Percent overlap smaller range in larger', 
-         title = sprintf('Size of species range (quartile %s) vs. percent overlap of \nsmaller range with larger', i))
-
-  ggsave(paste0('figures/AreavsOverlap_quantile_', i, '.png'), width = 6, height = 4)
-  print(qtile_plot)
-}
-# First quartile regression:
-#                 Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept)  1423917     101053  14.091   <2e-16 ***
-#     sm_perc         1188       1555   0.764    0.445    
-#     Multiple R-squared:  0.001198,	Adjusted R-squared:  -0.0008529 
-#
-# second quartile:
-#                 Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept)  7011028     275943  25.408  < 2e-16 ***
-#     sm_perc       -17725       4228  -4.192 3.28e-05 ***
-#     Multiple R-squared:  0.03483,	Adjusted R-squared:  0.03285 
-#
-# third quartile:
-#                 Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept) 12813279     507856  25.230  < 2e-16 ***
-#     sm_perc        64710       8159   7.931  1.5e-14 ***
-#     Multiple R-squared:  0.1146,	Adjusted R-squared:  0.1128 
-#
-# fourth quartile:
-#                   Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept) -101736351   17259544  -5.894 7.04e-09 ***
-#     sm_perc        1911286     215216   8.881  < 2e-16 ***
-#     Multiple R-squared:  0.1399,	Adjusted R-squared:  0.1381
-
-# This avenue of exploration looks like a bust.
-
-```
-
-***
-
-###(6) Looking at how many and what type (threat category) of species we are losing by setting a threshold of 0.4 on the aquamaps species
-
-This looks at the full Aquamaps dataset... should it be limited to species with valid categories?
-
-```{r aquamaps probabilities, warning = FALSE, message = FALSE, cache.lazy = FALSE, cache = TRUE, verbose = FALSE, results = FALSE, echo = FALSE, include = FALSE}
-
-# get spp_cells raw data for aquamaps
-dir_aquamaps <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO/raw/aquamaps_2014')
-
-spp_cells <- read_csv(file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), 
-                      col_types = '_ccd__', progress = TRUE)
-
-# spp_cells <- fread(input = file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), verbose = FALSE, header = TRUE, 
-#                   colClasses = c('NULL', 'character', NA, 'numeric', 'NULL', 'NULL')) #1 min 49 sec
-
-setnames(spp_cells, c('SpeciesID', 'CsquareCode', 'probability'), c('species_id', 'csquare_code', 'probability'))
-
-spp_cells <- as.data.frame(spp_cells)
-
-spp <- fread(input = file.path(dir_aquamaps, 'tables/ohi_speciesoccursum.csv'), 
-             verbose = FALSE, header = TRUE) %>%
-  mutate(sciname = paste(Genus, Species, sep = ' '),
-         iucn_code = str_replace(toupper(iucn_code), 'LR/', ''),
-         iucn_code = str_replace(iucn_code, 'N.E.', 'NE'))
-spp <- as.data.frame(spp)
-
-```
-
-```{r hist_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
-
-### create dataframe that gives me all species, their total number of cells, 
-### total number of cells >= 0.4 prob, difference between the two and threat
-spp_df3 <- spp_cells %>%
-  group_by(species_id) %>%
-  summarize(cells_0.4 = sum(probability >= 0.4),
-            cells_t = sum(probability >= 0)) %>%
-  mutate(prop = cells_0.4 / cells_t) %>%
-  left_join(spp %>%
-#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')), 
-              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
-            by = c('species_id' = 'SPECIESID')) %>%
-  select(sciname, species_id, cells_0.4, cells_t, prop, iucn_code)
-
-hist_area_vs_threshold <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
-  ggtheme_plot + 
-  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
-  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
-       y = 'Proportion of all aquamaps species (17,348 total)')
-
-ggsave('figures/AMThreshold.png', width = 6, height = 4)
-
-print(hist_area_vs_threshold)
-
-```
-
-```{r scatter_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
-
-### create dataframe that gives me all species, their total number of cells, 
-### total number of cells >= 0.4 prob, difference between the two and threat
-
-threshold_df <- data.frame()
-for (i in seq(0, 1, .05)) { # i = .05
-  temp_df <- spp_cells %>%
-    group_by(species_id) %>%
-    summarize(thresh       = i,
-              cells_thresh = sum(probability >= thresh),
-              cells_total  = sum(probability >= 0),
-              proportion   = cells_thresh/cells_total) %>%
-    left_join(spp %>%
-#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')), 
-              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
-            by = c('species_id' = 'SPECIESID')) %>%
-    select(sciname, species_id, thresh, cells_thresh, cells_total, proportion, iucn_code)
-
-  threshold_df <- threshold_df %>%
-    bind_rows(temp_df)
-}
-
-thresh_df_summary <- threshold_df %>%
-  group_by(thresh) %>%
-  summarize(mean_prop   = mean(proportion, na.rm = TRUE),
-            median_prop = median(proportion, na.rm = TRUE))
-
-scatter_area_vs_threshold <- ggplot(thresh_df_summary, aes(x = thresh, y = mean_prop)) + 
-  ggtheme_plot + 
-  geom_point(aes(y = median_prop), color = 'grey60') +
-  geom_point() + 
-  geom_abline(slope = -1, intercept = 1, color = 'red') + 
-  geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
-  ylim(0, 1) +
-  labs(title = 'Mean (median) species range after presence threshold cutoff',
-       x = 'Presence threshold cutoff', 
-       y = 'Mean (median) relative species range')
-
-ggsave('figures/scatter_AMThreshold.png', width = 6, height = 4)
-
-print(scatter_area_vs_threshold)
-
-```
-
-## Changes in SPP for scenarios
-
-Comparing three scenarios to the current (v2015) global model for SPP:
-
-* Baseline is: preferring IUCN rangemaps over Aquamaps rangemaps; Aquamaps 'presence' threshold is set at 40% or greater.
-* Scenario 1: preferring IUCN rangemaps over Aquamaps; AM presence threshold set at >= 1% (effectively, all occurrences).
-* Scenario 2: preferring AM maps over IUCN; AM presence >= 40%.
-* Scenario 3: preferring AM over IUCN; AM presence >= 1%.
-
-Boxplots:
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
-
-![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)

### 2015-11-21 R/data_explore_pt2.Rmd
trying out new maps by quadrant to see if there are common characteristics...

--- 2015-11-18 data_explore.Rmd
@@
----
-title: 'Data Exploration: IUCN and AquaMaps'
-output: html_document
----
+Exploring possible correlation between total species range area and percent overlap of smaller range within the larger:
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-# Libraries and Paths
-
-library(ggplot2)
-library(maps)
-library(readr)
-library(raster)
-library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-
-dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-           'Darwin'  = '/Volumes/data_edit',
-           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-```
-
-```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
-
-# Data files
-#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), 
-                           col_types = 'cdddd', progress = TRUE) %>%
-  as.data.frame() %>%
-  filter(presence != 5) ### removes extinct subpopulations
-
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells.csv'), progress = TRUE)
-
-### species info list - scinames, IUCN species ID, AM species ID, category etc
-spp_list_base <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
-
-### LOICZID cell ID numbers - used in plotting rasters
-loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
-loiczid_raster       <- raster(loiczid_raster_file)
-names(loiczid_raster) <- 'loiczid'
-```
-
-``` {r ggplot_theme_setups, echo = FALSE}
-
-### generic theme for all plots
-ggtheme_plot <- theme(axis.ticks = element_blank(),
-                    text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-                    plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-                    legend.position = 'right')
-
-# ggtheme_map <- theme(axis.ticks = element_blank(), axis.text = element_blank(),
-#                     text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-#                     plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-#                     legend.position = 'right')
-
-### theme for species range maps - based on generic plot theme
-ggtheme_map <- ggtheme_plot +
-  theme(axis.text = element_blank())
-
-```
-
-``` {r species_map_functions, echo = FALSE, warning = FALSE, message = FALSE}
-
-### Species Map Function ###
-# This function takes a single species scientific name as input, then grabs all 
-# occurrence cells and associated Aquamaps probability and/or IUCN proportional area
-# per cell
-get_spp_map <- function(species){
-  spp_id <- spp_list_base %>%
-    filter(sciname == species & is.na(parent_sid)) %>%
-    dplyr::select(am_sid, iucn_sid, sciname) %>%
-    unique()
-  iucn_spp_map <- iucn_spp_cells %>%
-    filter(sciname == species) %>%
-    group_by(loiczid) %>%        
-    summarize(iucn_area = max(prop_area))
-  ### The group_by() and summarize() are to collapse cells with overlapped polygons, fixing the duplicate 'by' issue
-  am_spp_map   <- am_spp_cells %>%
-    filter(am_sid == spp_id$am_sid) %>%
-    rename(am_prob = prob)
-  spp_map <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid')%>%
-    as.data.frame()
-
-  return(spp_map)
-}
-
-### Plot Species Map function
-### uses ggplot to create a formatted map of species ranges, including
-### Aquamaps, IUCN, and overlapped ranges; returns the plot object.
-plot_rangemap <- function(spp) {
-  map <- get_spp_map(spp)
-
-  am_map <- map %>%
-    filter(!is.na(am_prob))
-
-  r_am_spp  <-  subs(loiczid_raster, 
-                     am_map[ , c('loiczid', 'am_prob')], 
-                     by = 'loiczid', 
-                     which = 'am_prob', 
-                     subsWithNA = TRUE)
-  r_am_spp[!is.na(r_am_spp)] <- 1
-
-  iucn_map <- map %>%
-    mutate(iucn_pres = iucn_area > 0)
-  r_iucn_spp <- subs(loiczid_raster, 
-                     iucn_map[ , c('loiczid', 'iucn_pres')], 
-                     by = 'loiczid', 
-                     which = 'iucn_pres', 
-                     subsWithNA = TRUE)
-
-  spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
-    full_join(as.data.frame(rasterToPoints(r_am_spp)),
-              by = c('x', 'y')) %>%
-    mutate(presence = ifelse(am_prob & is.na(iucn_pres), 'am',
-                             ifelse(iucn_pres & is.na(am_prob), 'iucn',
-                                    ifelse(iucn_pres & am_prob, 'both', NA))))
-
-  spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
-    ggtheme_map + 
-    geom_raster(aes(fill = presence), alpha = .5) +
-    scale_fill_manual(values = c('iucn' = 'blue', 'am' = 'red', 'both' = 'mediumorchid')) +
-    borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
-    labs(title = spp, x = NULL, y = NULL) 
-}
-
-```
-***
-
-```{r create_species_list_w_area, echo = FALSE, eval = FALSE, warning = FALSE, message = FALSE}
-
-spp_list <- spp_list_base %>%
-              mutate(total_area = NA,
-                     am_area    = NA,
-                     iucn_area  = NA,
-                     perc       = NA,
-                     am_perc    = NA,
-                     iucn_perc  = NA,
-                     sm_perc    = NA, #filled with % of smaller range that is within the larger range
-                     sm_range   = NA) #what range is smaller
-
-#there's likely a better way to do this than a for loop (lapply or sapply?)
-for (i in 1:nrow(spp_list)){
-  spp     <- spp_list[i, ]$sciname #species name
-  spp_map <- get_spp_map(spp) #get species map
-
-###get just the aquamaps cells
-  am_spp_map <- spp_map %>%
-    filter(!is.na(am_prob))%>%
-    dplyr::select(loiczid, am_sid, am_prob) %>% 
-      ### add in am_prob when we want to look at probability
-    unique()
-
-  if(any(duplicated(am_spp_map$loiczid))){
-    am_spp_map <- am_spp_map %>%
-      dplyr::select(loiczid, am_sid) %>% 
-      unique()
-    }
-
-### rasterize aquamaps map
-  r_am_spp <- subs(loiczid_raster, 
-                   am_spp_map[ , c('loiczid', 'am_sid')], 
-                   by = 'loiczid', 
-                   which = 'am_sid', 
-                   subsWithNA = TRUE)
-
-  spp_list[i, ]$am_area <- area(r_am_spp, na.rm = TRUE)%>%
-              cellStats(., stat='sum')
-
-###  get the iucn cells
-  iucn_map <- spp_map %>%
-    mutate(iucn_pres = iucn_area > 0)%>%
-      dplyr::select(loiczid, iucn_area, am_sid, am_prob, iucn_pres)%>%
-        unique()
-
-  #TODO: incorporate the iucn_area (area of cell that species overlaps with...I think thats what it is)
-
-###this is incorporated because there are some duplications for loiczids with IUCN cells. I'm guessing this is because of the parent/subpopulations?
-
-   if(TRUE %in% duplicated(iucn_map$loiczid))next()
-
-### rasterize IUCN map
-  r_iucn_spp <- subs(loiczid_raster, 
-                     iucn_map[ , c('loiczid', 'iucn_pres')], 
-                     by = 'loiczid', 
-                     which = 'iucn_pres', 
-                     subsWithNA = TRUE)
-
-  spp_list[i, ]$iucn_area <- area(r_iucn_spp, na.rm = TRUE)%>%
-                cellStats(., stat='sum')
-
-### Get total area (IUCN and AM combined)
-  allCells <- spp_map%>%
-                dplyr::select(loiczid)%>%
-                  mutate(value=1)%>%
-                  unique()
-
-  fullRange <- subs(loiczid_raster, 
-                    allCells[ , c('loiczid', 'value')], 
-                    by = 'loiczid', 
-                    which = 'value', 
-                    subsWithNA = TRUE)
-
-  spp_list[i, ]$total_area <- area(fullRange, na.rm = TRUE)%>%
-                              cellStats(., stat='sum')
-
-### get percent overlap of map
-
-   #iucn map cells
-  cells_iucn<-iucn_map %>%
-    filter(iucn_pres == TRUE) %>%
-    .$loiczid
-  cells_am <- am_spp_map %>%
-    filter(!is.na(am_sid)) %>%
-    .$loiczid
-
-  cells_total <- unique(spp_map$loiczid)
-  cells_overlap <- intersect(cells_iucn, cells_am)
-
-  #I THINK THIS IS A DUPLICATE? r_iucn_spp raster created above... (JA 10/12/15)
-#    r_iucn_spp <- subs(loiczid_raster, 
-#                       iucn_map[ , c('loiczid', 'iucn_pres')], 
-#                      by = 'loiczid', 
-#                       which = 'iucn_pres', 
-#                       subsWithNA = TRUE)
-#   
-  #percent overlap
-  perc <- (length(cells_overlap) / length(cells_total)) * 100
-  spp_list[i, ]$perc <- perc
-
-  am_only <- setdiff(cells_am, cells_iucn)
-  iucn_only <- setdiff(cells_iucn, cells_am)
-
-  perc_am <- (length(am_only) / length(cells_total)) * 100
-  perc_iucn <- (length(iucn_only) / length(cells_total)) * 100
-
-  spp_list[i, ]$am_perc <- perc_am
-  spp_list[i, ]$iucn_perc <- perc_iucn
-
-  spp_list[i, ]$sm_range <- ifelse(length(cells_am)<length(cells_iucn), 'AM', 'IUCN') 
-    ### what dataset has the smaller range
-  spp_list[i, ]$sm_perc <- length(cells_overlap)/min(length(cells_am), length(cells_iucn)) * 100 
-    ### this is the total percent of smaller range encompassed in the larger range
-
-  cat(sprintf('%s: For species %s, %.2f%% of both datasets overlap, %.2f%% is just AquaMaps and %.2f%% is just IUCN\n', 
-              i, spp, perc, perc_am, perc_iucn))
-  }
-
-# for Conus arenatus, the IUCN map gives 4 duplicate cells, but each of the entries has a different iucn_area...which doesn't make sense
-
-#row 985 for Acanthurus leucopareius - AM is giving duplicate loiczids but with different am_prob...which is so strange. Looked
-# at the raw data and it seems to be in the raw data... (JA on 8/31/15)
-
-#remove non IUCN species and log total area (for use later on)
-spp_list <- spp_list %>%
-            filter(spp_group != 'hagfishes' & spp_group != 'non-homalopsids') %>%
-              mutate(log_total_area = log(total_area))
-
-write.csv(spp_list, file='spp_list_w_area_percent_noextinct.csv')
-```
-
-###(1) Number of species in both datasets split up by group
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-# look at number of species per group
-
-spp_list <- read.csv('spp_list_w_area_percent_noextinct.csv')
-
-spp_df <- spp_list %>%
-  mutate(diff = iucn_area - am_area)
-
-spp_df <- within(spp_df, spp_group <- factor(spp_group, levels = names(sort(table(spp_group), increasing = TRUE))))
-
-barchart_sp_groups <- ggplot(spp_df, aes(x = spp_group)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'bin') +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + 
-  coord_flip() +
-  labs(x = 'Species Group', 
-       y = 'Number of Species', 
-       title = 'Number of species in both AquaMaps and IUCN datasets')
-
-ggsave('figures/sp_groups_barchart.png', 
-       plot = barchart_sp_groups, 
-       width = 8, height = 4, dpi = 100)
-
-print(barchart_sp_groups)
-
-```
-
-***
-
-###(2) Looking at the total difference in area (km^2) for each species range map in both datasets
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-scatter_am_iucn_area <- ggplot(spp_df, aes(x = iucn_area, y = am_area, color = iucn_category)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline(slope = 1, intercept = 0, color = 'red') +
-  labs(x = 'Area (km^2) from IUCN dataset', 
-       y = 'Area (km^2) from AquaMaps dataset', 
-       title = 'Comparing total area of species range maps')
-
-ggsave('figures/CompareTotalArea_AM-IUCN_scatter.png', 
-       plot = scatter_am_iucn_area, 
-       width = 8, height = 4, dpi = 100)
-
-print(scatter_am_iucn_area)
-```
-
-There are large differences so to get a better sense, here is the same comparison but with area log transformed.
-``` {r echo = FALSE, warning = FALSE, message = FALSE}
-
-spp_df <- spp_df %>%
-  mutate(am_log = log(am_area),
-         iucn_log = log(iucn_area))
-
-scatter_am_iucn_area_log <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline(slope = 1, intercept = 0, color = 'red') +
-  labs(x = 'IUCN Area (log transformed)', 
-       y = 'AquaMaps Area (log transformed)',
-       title = 'Comparing area of range maps\n (all area values are km^2 log transformed)', 
-       fill = 'IUCN Category')
-
-ggsave('figures/CompareTotalAreaLOG_AM-IUCN_scatter.png', 
-       plot = scatter_am_iucn_area_log, 
-       width = 8, height = 4, dpi = 100)
-
-print(scatter_am_iucn_area_log)
-```
-
-Data split into species groups and by IUCN category
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-scatter_am_iucn_alog_gp <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
-  ggtheme_plot + 
-  geom_point() +
-  facet_wrap(~spp_group) + 
-  geom_abline() + 
-  labs(x = 'IUCN Area (log transformed)', 
-       y = 'AquaMaps Area (log transformed)',
-       title = 'Comparing differences in area across species groups', 
-       fill = 'IUCN Category')
-
-ggsave('figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png', 
-       plot = scatter_am_iucn_alog_gp, 
-       width = 8, height = 6, dpi = 100)
-
-print(scatter_am_iucn_alog_gp)
-```
-***
-
-###(3) Looking at the overlap between species range maps from IUCN and AquaMaps
-
-This plot shows the total percent of a species range map that is (1) only covered by IUCN cells, (2) only covered by AquaMaps cells and (3) covered by both IUCN and AquaMaps cells.
-
-####How this breaks down between AquaMaps and IUCN
-
-Here is an example of a single species, the yellow speckled chromis (*Chromis alpha*)
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-spp <- 'Chromis alpha'
-
-spp_map <- plot_rangemap(spp)
-
-print(spp_map)
-ggsave(plot = spp_map,
-       filename = sprintf('figures/rangemap_%s.png', spp),
-       width = 8, height = 4, dpi = 200)
-```
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-  spp_df2 <- gather(spp_list, perc.type, percent, perc, am_perc, iucn_perc) %>%
-    dplyr::select(sciname, spp_group, perc.type, percent) %>%
-    unique()
-
-  perc_c <- spp_df2 %>%
-    filter(sciname == spp)
-
-  p1 <- perc_c %>% 
-    dplyr::select(percent)
-
-```
-
-For `r spp`, `r round(pi[1], 2)`% of the total area is covered by both IUCN and AquaMaps, `r round(p1[2, ], 2)`% is covered just by AquaMaps and `r round(p1[3, ], 2)`% is covered just by IUCN.
-
-***
-
-####Now looking across all species, here is the breakdown of how much overlap there is between the two datasets
-
-This plot shows that the majority of species range maps have only ~30% of their full map covered by both maps, and very few species have more than 50% of their two ranges (IUCN and AquaMaps) overlapping.
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-### look at density plot
-
-overlap <- spp_df2 %>%
-  dplyr::filter(perc.type == 'perc')
-
-hist_overlap_pct <- ggplot(overlap, aes(percent)) +
-  ggtheme_plot + 
-  geom_histogram(aes(y = ..count../1926), binwidth = 5, alpha = .8) +
-  labs(title = 'Percent of total range shared between datasets',
-       x = '% total range that is shared b/w both datasets', 
-       y = 'Proportion of Total Species')
-
-ggsave('figures/Overlap-PercTotalRange.png')
-
-print(hist_overlap_pct)
-```
-
-Split up by species - not sure if this is useful or not
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
-
-hist_overlap_pct_gp <- ggplot(overlap, aes(percent)) + 
-  ggtheme_plot + 
-  geom_histogram(binwidth = 5, alpha = .8) +
-  facet_wrap(~spp_group) +
-  labs(title = 'Percent of total range shared, by species group',
-       x = 'Percent of total range that is shared between both datasets', 
-       y = 'Number of Total Species') + 
-  theme(strip.text.x = element_text(size = 8))
-
-ggsave('figures/Overlap-PercTotalRange_spGroups.png')
-print(hist_overlap_pct_gp)
-
-```
-
-###(4) Looking at the percent of smaller species range that is within the larger range.
-
-For each range
-
-```{r pct small within large, echo = FALSE, warning = FALSE}
-
-spp_df4 <- spp_list %>%
-  dplyr::select(sciname, sm_perc, sm_range) %>%
-  unique()
-
-hist_sm_within_lg <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
-  ggtheme_plot + 
-  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
-  scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
-  labs(title = 'Percent of smaller range within larger range',
-       x = 'Percent of smaller range that falls within the larger range', 
-       y = sprintf('Proportion of all species (n = %s)', nrow(spp_df4)))
-
-ggsave('figures/PercSmallRangeInLarger.png', width = 6, height = 4)
-
-print(hist_sm_within_lg)
-
-```
-
-Exploring possible correlation between total species range area and percent overlap of smaller range within the larger:
 ``` {r pct overlap vs area, echo = FALSE, warning = FALSE, message = FALSE}
@@
        title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreavsPercOverlap.png', width = 6, height = 4)
-
-print(scatter_area_vs_overlap)
-
-```
-
-``` {r pct overlap vs ratio of area, echo = FALSE, warning = FALSE, message = FALSE}
-
-spp_list <- spp_list %>%
-  mutate(area_ratio = am_area / iucn_area,
-         area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
-         area_ratio = area_ratio * 100)
-
-area_ratio_quartiles <- quantile(spp_list$area_ratio, na.rm = TRUE)[2:4]
-pct_overlap_quartiles <- quantile(spp_list$sm_perc, na.rm = TRUE)[2:4]
-
-spp_quads <- c('q1' = 'Kajikia albida', 
-               'q2' = 'Galaxea cryptoramosa', 
-               'q4' = 'Acanthurus nigroris', 
-               'q4a' = 'Metanephrops australiensis', 
-               'q3' = 'Conus magnificus')
-
-spp_list_label <- spp_list %>%
-  filter(sciname %in% spp_quads) %>%
-  separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
-  mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
-  select(-g, -s)
-
-scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = log_total_area)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline() + 
-  geom_vline(x = area_ratio_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_hline(y = pct_overlap_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_point(data = spp_list_label, color = 'tomato2') + 
-  geom_text(data = spp_list_label, 
-            aes(label = sciname_label), 
-            color = 'tomato2',
-            fontface = 'bold.italic',
-            size = 2,
-            hjust = .25,
-            vjust = 1.5) +
-  labs(x = 'Range area ratio (smaller:larger) * 100%', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreaRatiovsPercOverlap.png', width = 6, height = 4)
-
-print(scatter_arearatio_vs_overlap)
-
-scatter_arearatio_vs_overlap2 <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline() + 
-  geom_vline(x = area_ratio_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_hline(y = pct_overlap_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_point(data = spp_list_label, color = 'grey30') + 
-  geom_text(data = spp_list_label, 
-            aes(label = sciname_label), 
-            color = 'grey30',
-            fontface = 'bold.italic',
-            size = 2,
-            hjust = .25,
-            vjust = 1.5) +
-  labs(x = 'Range area ratio (smaller:larger) * 100%', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreaRatiovsPercOverlap2.png', width = 6, height = 4)
-
-print(scatter_arearatio_vs_overlap2)
-
-```
-
-``` {r examine_spp_gp_quartiles}
-
-spp_list1 <- spp_list %>%
-  group_by(spp_group) %>%
-  mutate(n_spp = n()) %>%
-  ungroup()
-
-spp_gp_q1 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q1 = n())
-spp_gp_q2 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q2 = n())
-spp_gp_q4 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q4 = n())
-spp_gp_q3 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q3 = n())
-
-spp_gp_quadrants <- spp_gp_q1 %>%
-  left_join(spp_gp_q2, by = c('spp_group', 'n_spp')) %>%
-  left_join(spp_gp_q3, by = c('spp_group', 'n_spp')) %>%
-  left_join(spp_gp_q4, by = c('spp_group', 'n_spp')) %>%
-  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
-  mutate(quad = str_replace(quad, 'n_spp_', ''),
-         pct_quad = n_quad/n_spp)
-
-spp_gp_quadrants <- spp_gp_quadrants %>%
-  left_join(spp_gp_quadrants %>% 
-              filter(quad == 'q1') %>%
-              select(spp_group, pct_q1 = pct_quad),
-            by = 'spp_group')
-
-barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
-                                  transform(spp_group = reorder(spp_group, pct_q1)), 
-                                aes(x = spp_group, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'bin') +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0) +
-  coord_flip() +
-  labs(x = 'Species Group', 
-       y = 'Relative number of species per quadrant', 
-       title = 'Species group breakdown by quadrant')
-
-```
-
-```{r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE}
-
-for (i in 1:length(spp_quads)) { # i = 1
-  spp <- spp_quads[i]
-  spp_quad <- names(spp_quads)[i]
-
-  spp_plot <- plot_rangemap(spp)
-  print(spp_plot)
-
-  ggsave(filename = sprintf('figures/rangemap_%s_%s.png', spp_quad, str_replace(tolower(spp), ' ', '_')),
-         plot = spp_plot, 
-         width = 8, height = 4, dpi = 200)
-}
-
-```
-
-``` {r, echo = FALSE, warning = FALSE, message = FALSE}
-
-# Regression
 
@@
 
-ggsave('figures/CorrelationAreavsPercOverlap.png')
-
 print(scatter_area_vs_overlap)
 
-summary(lm(total_area ~ sm_perc, spp_list))
+# summary(lm(total_area ~ sm_perc, spp_list))
+
+ggsave(file.path(dir_fig, 'AreavsPercOverlap.png'), width = 6, height = 4)
+
 ```
@@
 
-Percent coverage as function of area, clipping to area below 100,000,000 km^2:
+Percent coverage as function of area, clipping to area below 100,000,000 km^2^:
+
 ``` {r, echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-ggsave('figures/AreavsPercOverlap_2.png', width = 6, height = 4)
+ggsave(file.path(dir_fig, 'AreavsPercOverlap_2.png'), width = 6, height = 4)
 
@@
 
-ggsave('figures/AreavsPercOverlap_log.png', width=6, height=4)
+ggsave(file.path(dir_fig, 'AreavsPercOverlap_log.png'), width=6, height=4)
 
@@
 
-  ggsave(paste0('figures/AreavsOverlap_quantile_', i, '.png'), width = 6, height = 4)
+  ggsave(file.path(dir_fig, paste0('AreavsOverlap_quantile_', i, '.png')), width = 6, height = 4)
   print(qtile_plot)
@@
 ***
-
-###(6) Looking at how many and what type (threat category) of species we are losing by setting a threshold of 0.4 on the aquamaps species
-
-This looks at the full Aquamaps dataset... should it be limited to species with valid categories?
-
-```{r aquamaps probabilities, warning = FALSE, message = FALSE, cache.lazy = FALSE, cache = TRUE, verbose = FALSE, results = FALSE, echo = FALSE, include = FALSE}
-
-# get spp_cells raw data for aquamaps
-dir_aquamaps <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO/raw/aquamaps_2014')
-
-spp_cells <- read_csv(file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), 
-                      col_types = '_ccd__', progress = TRUE)
-
-# spp_cells <- fread(input = file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), verbose = FALSE, header = TRUE, 
-#                   colClasses = c('NULL', 'character', NA, 'numeric', 'NULL', 'NULL')) #1 min 49 sec
-
-setnames(spp_cells, c('SpeciesID', 'CsquareCode', 'probability'), c('species_id', 'csquare_code', 'probability'))
-
-spp_cells <- as.data.frame(spp_cells)
-
-spp <- fread(input = file.path(dir_aquamaps, 'tables/ohi_speciesoccursum.csv'), 
-             verbose = FALSE, header = TRUE) %>%
-  mutate(sciname = paste(Genus, Species, sep = ' '),
-         iucn_code = str_replace(toupper(iucn_code), 'LR/', ''),
-         iucn_code = str_replace(iucn_code, 'N.E.', 'NE'))
-spp <- as.data.frame(spp)
-
-```
-
-```{r hist_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
-
-### create dataframe that gives me all species, their total number of cells, 
-### total number of cells >= 0.4 prob, difference between the two and threat
-spp_df3 <- spp_cells %>%
-  group_by(species_id) %>%
-  summarize(cells_0.4 = sum(probability >= 0.4),
-            cells_t = sum(probability >= 0)) %>%
-  mutate(prop = cells_0.4 / cells_t) %>%
-  left_join(spp %>%
-#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')), 
-              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
-            by = c('species_id' = 'SPECIESID')) %>%
-  select(sciname, species_id, cells_0.4, cells_t, prop, iucn_code)
-
-hist_area_vs_threshold <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
-  ggtheme_plot + 
-  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
-  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
-       y = 'Proportion of all aquamaps species (17,348 total)')
-
-ggsave('figures/AMThreshold.png', width = 6, height = 4)
-
-print(hist_area_vs_threshold)
-
-```
-
-```{r scatter_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
-
-### create dataframe that gives me all species, their total number of cells, 
-### total number of cells >= 0.4 prob, difference between the two and threat
-
-threshold_df <- data.frame()
-for (i in seq(0, 1, .05)) { # i = .05
-  temp_df <- spp_cells %>%
-    group_by(species_id) %>%
-    summarize(thresh       = i,
-              cells_thresh = sum(probability >= thresh),
-              cells_total  = sum(probability >= 0),
-              proportion   = cells_thresh/cells_total) %>%
-    left_join(spp %>%
-#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')), 
-              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
-            by = c('species_id' = 'SPECIESID')) %>%
-    select(sciname, species_id, thresh, cells_thresh, cells_total, proportion, iucn_code)
-
-  threshold_df <- threshold_df %>%
-    bind_rows(temp_df)
-}
-
-thresh_df_summary <- threshold_df %>%
-  group_by(thresh) %>%
-  summarize(mean_prop   = mean(proportion, na.rm = TRUE),
-            median_prop = median(proportion, na.rm = TRUE))
-
-scatter_area_vs_threshold <- ggplot(thresh_df_summary, aes(x = thresh, y = mean_prop)) + 
-  ggtheme_plot + 
-  geom_point(aes(y = median_prop), color = 'grey60') +
-  geom_point() + 
-  geom_abline(slope = -1, intercept = 1, color = 'red') + 
-  geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
-  ylim(0, 1) +
-  labs(title = 'Mean (median) species range after presence threshold cutoff',
-       x = 'Presence threshold cutoff', 
-       y = 'Mean (median) relative species range')
-
-ggsave('figures/scatter_AMThreshold.png', width = 6, height = 4)
-
-print(scatter_area_vs_threshold)
-
-```
-
-## Changes in SPP for scenarios
-
-Comparing three scenarios to the current (v2015) global model for SPP:
-
-* Baseline is: preferring IUCN rangemaps over Aquamaps rangemaps; Aquamaps 'presence' threshold is set at 40% or greater.
-* Scenario 1: preferring IUCN rangemaps over Aquamaps; AM presence threshold set at >= 1% (effectively, all occurrences).
-* Scenario 2: preferring AM maps over IUCN; AM presence >= 40%.
-* Scenario 3: preferring AM over IUCN; AM presence >= 1%.
-
-Boxplots:
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
-
-![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)

### 2015-11-21 data_explore.Rmd
trying out new maps by quadrant to see if there are common characteristics...

--- 2015-11-18 data_explore.Rmd
@@
 dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
+dir_fig <- file.path(getwd(), 'figures')
+
+source('R/data_explore_fxns.R')
 ```
@@
 
-### species info list - scinames, IUCN species ID, AM species ID, category etc
-spp_list_base <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
-
 ### LOICZID cell ID numbers - used in plotting rasters
@@
 
-``` {r ggplot_theme_setups, echo = FALSE}
+```{r get spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+data_file <- 'data/spp_list_w_area_percent_noextinct.csv'
 
-### generic theme for all plots
-ggtheme_plot <- theme(axis.ticks = element_blank(),
-                    text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-                    plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-                    legend.position = 'right')
+if(!file.exists(data_file)) {
+  ### species info list - scinames, IUCN species ID, AM species ID, category etc
+  spp_list_base <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
 
-# ggtheme_map <- theme(axis.ticks = element_blank(), axis.text = element_blank(),
-#                     text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-#                     plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-#                     legend.position = 'right')
+  create_spp_list(spp_list_base, data_file)
+}
 
-### theme for species range maps - based on generic plot theme
-ggtheme_map <- ggtheme_plot +
-  theme(axis.text = element_blank())
-
+spp_list <- read.csv(data_file, stringsAsFactors = FALSE)
 ```
 
-``` {r species_map_functions, echo = FALSE, warning = FALSE, message = FALSE}
-
-### Species Map Function ###
-# This function takes a single species scientific name as input, then grabs all 
-# occurrence cells and associated Aquamaps probability and/or IUCN proportional area
-# per cell
-get_spp_map <- function(species){
-  spp_id <- spp_list_base %>%
-    filter(sciname == species & is.na(parent_sid)) %>%
-    dplyr::select(am_sid, iucn_sid, sciname) %>%
-    unique()
-  iucn_spp_map <- iucn_spp_cells %>%
-    filter(sciname == species) %>%
-    group_by(loiczid) %>%        
-    summarize(iucn_area = max(prop_area))
-  ### The group_by() and summarize() are to collapse cells with overlapped polygons, fixing the duplicate 'by' issue
-  am_spp_map   <- am_spp_cells %>%
-    filter(am_sid == spp_id$am_sid) %>%
-    rename(am_prob = prob)
-  spp_map <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid')%>%
-    as.data.frame()
-
-  return(spp_map)
-}
-
-### Plot Species Map function
-### uses ggplot to create a formatted map of species ranges, including
-### Aquamaps, IUCN, and overlapped ranges; returns the plot object.
-plot_rangemap <- function(spp) {
-  map <- get_spp_map(spp)
-
-  am_map <- map %>%
-    filter(!is.na(am_prob))
-
-  r_am_spp  <-  subs(loiczid_raster, 
-                     am_map[ , c('loiczid', 'am_prob')], 
-                     by = 'loiczid', 
-                     which = 'am_prob', 
-                     subsWithNA = TRUE)
-  r_am_spp[!is.na(r_am_spp)] <- 1
-
-  iucn_map <- map %>%
-    mutate(iucn_pres = iucn_area > 0)
-  r_iucn_spp <- subs(loiczid_raster, 
-                     iucn_map[ , c('loiczid', 'iucn_pres')], 
-                     by = 'loiczid', 
-                     which = 'iucn_pres', 
-                     subsWithNA = TRUE)
-
-  spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
-    full_join(as.data.frame(rasterToPoints(r_am_spp)),
-              by = c('x', 'y')) %>%
-    mutate(presence = ifelse(am_prob & is.na(iucn_pres), 'am',
-                             ifelse(iucn_pres & is.na(am_prob), 'iucn',
-                                    ifelse(iucn_pres & am_prob, 'both', NA))))
-
-  spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
-    ggtheme_map + 
-    geom_raster(aes(fill = presence), alpha = .5) +
-    scale_fill_manual(values = c('iucn' = 'blue', 'am' = 'red', 'both' = 'mediumorchid')) +
-    borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
-    labs(title = spp, x = NULL, y = NULL) 
-}
-
-```
-***
-
-```{r create_species_list_w_area, echo = FALSE, eval = FALSE, warning = FALSE, message = FALSE}
-
-spp_list <- spp_list_base %>%
-              mutate(total_area = NA,
-                     am_area    = NA,
-                     iucn_area  = NA,
-                     perc       = NA,
-                     am_perc    = NA,
-                     iucn_perc  = NA,
-                     sm_perc    = NA, #filled with % of smaller range that is within the larger range
-                     sm_range   = NA) #what range is smaller
-
-#there's likely a better way to do this than a for loop (lapply or sapply?)
-for (i in 1:nrow(spp_list)){
-  spp     <- spp_list[i, ]$sciname #species name
-  spp_map <- get_spp_map(spp) #get species map
-
-###get just the aquamaps cells
-  am_spp_map <- spp_map %>%
-    filter(!is.na(am_prob))%>%
-    dplyr::select(loiczid, am_sid, am_prob) %>% 
-      ### add in am_prob when we want to look at probability
-    unique()
-
-  if(any(duplicated(am_spp_map$loiczid))){
-    am_spp_map <- am_spp_map %>%
-      dplyr::select(loiczid, am_sid) %>% 
-      unique()
-    }
-
-### rasterize aquamaps map
-  r_am_spp <- subs(loiczid_raster, 
-                   am_spp_map[ , c('loiczid', 'am_sid')], 
-                   by = 'loiczid', 
-                   which = 'am_sid', 
-                   subsWithNA = TRUE)
-
-  spp_list[i, ]$am_area <- area(r_am_spp, na.rm = TRUE)%>%
-              cellStats(., stat='sum')
-
-###  get the iucn cells
-  iucn_map <- spp_map %>%
-    mutate(iucn_pres = iucn_area > 0)%>%
-      dplyr::select(loiczid, iucn_area, am_sid, am_prob, iucn_pres)%>%
-        unique()
-
-  #TODO: incorporate the iucn_area (area of cell that species overlaps with...I think thats what it is)
-
-###this is incorporated because there are some duplications for loiczids with IUCN cells. I'm guessing this is because of the parent/subpopulations?
-
-   if(TRUE %in% duplicated(iucn_map$loiczid))next()
-
-### rasterize IUCN map
-  r_iucn_spp <- subs(loiczid_raster, 
-                     iucn_map[ , c('loiczid', 'iucn_pres')], 
-                     by = 'loiczid', 
-                     which = 'iucn_pres', 
-                     subsWithNA = TRUE)
-
-  spp_list[i, ]$iucn_area <- area(r_iucn_spp, na.rm = TRUE)%>%
-                cellStats(., stat='sum')
-
-### Get total area (IUCN and AM combined)
-  allCells <- spp_map%>%
-                dplyr::select(loiczid)%>%
-                  mutate(value=1)%>%
-                  unique()
-
-  fullRange <- subs(loiczid_raster, 
-                    allCells[ , c('loiczid', 'value')], 
-                    by = 'loiczid', 
-                    which = 'value', 
-                    subsWithNA = TRUE)
-
-  spp_list[i, ]$total_area <- area(fullRange, na.rm = TRUE)%>%
-                              cellStats(., stat='sum')
-
-### get percent overlap of map
-
-   #iucn map cells
-  cells_iucn<-iucn_map %>%
-    filter(iucn_pres == TRUE) %>%
-    .$loiczid
-  cells_am <- am_spp_map %>%
-    filter(!is.na(am_sid)) %>%
-    .$loiczid
-
-  cells_total <- unique(spp_map$loiczid)
-  cells_overlap <- intersect(cells_iucn, cells_am)
-
-  #I THINK THIS IS A DUPLICATE? r_iucn_spp raster created above... (JA 10/12/15)
-#    r_iucn_spp <- subs(loiczid_raster, 
-#                       iucn_map[ , c('loiczid', 'iucn_pres')], 
-#                      by = 'loiczid', 
-#                       which = 'iucn_pres', 
-#                       subsWithNA = TRUE)
-#   
-  #percent overlap
-  perc <- (length(cells_overlap) / length(cells_total)) * 100
-  spp_list[i, ]$perc <- perc
-
-  am_only <- setdiff(cells_am, cells_iucn)
-  iucn_only <- setdiff(cells_iucn, cells_am)
-
-  perc_am <- (length(am_only) / length(cells_total)) * 100
-  perc_iucn <- (length(iucn_only) / length(cells_total)) * 100
-
-  spp_list[i, ]$am_perc <- perc_am
-  spp_list[i, ]$iucn_perc <- perc_iucn
-
-  spp_list[i, ]$sm_range <- ifelse(length(cells_am)<length(cells_iucn), 'AM', 'IUCN') 
-    ### what dataset has the smaller range
-  spp_list[i, ]$sm_perc <- length(cells_overlap)/min(length(cells_am), length(cells_iucn)) * 100 
-    ### this is the total percent of smaller range encompassed in the larger range
-
-  cat(sprintf('%s: For species %s, %.2f%% of both datasets overlap, %.2f%% is just AquaMaps and %.2f%% is just IUCN\n', 
-              i, spp, perc, perc_am, perc_iucn))
-  }
-
-# for Conus arenatus, the IUCN map gives 4 duplicate cells, but each of the entries has a different iucn_area...which doesn't make sense
-
-#row 985 for Acanthurus leucopareius - AM is giving duplicate loiczids but with different am_prob...which is so strange. Looked
-# at the raw data and it seems to be in the raw data... (JA on 8/31/15)
-
-#remove non IUCN species and log total area (for use later on)
-spp_list <- spp_list %>%
-            filter(spp_group != 'hagfishes' & spp_group != 'non-homalopsids') %>%
-              mutate(log_total_area = log(total_area))
-
-write.csv(spp_list, file='spp_list_w_area_percent_noextinct.csv')
+``` {r child = 'R/data_explore_pt1.Rmd', eval = FALSE}
 ```
 
-###(1) Number of species in both datasets split up by group
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-# look at number of species per group
-
-spp_list <- read.csv('spp_list_w_area_percent_noextinct.csv')
-
-spp_df <- spp_list %>%
-  mutate(diff = iucn_area - am_area)
-
-spp_df <- within(spp_df, spp_group <- factor(spp_group, levels = names(sort(table(spp_group), increasing = TRUE))))
-
-barchart_sp_groups <- ggplot(spp_df, aes(x = spp_group)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'bin') +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) + 
-  coord_flip() +
-  labs(x = 'Species Group', 
-       y = 'Number of Species', 
-       title = 'Number of species in both AquaMaps and IUCN datasets')
-
-ggsave('figures/sp_groups_barchart.png', 
-       plot = barchart_sp_groups, 
-       width = 8, height = 4, dpi = 100)
-
-print(barchart_sp_groups)
-
-```
-
-***
-
-###(2) Looking at the total difference in area (km^2) for each species range map in both datasets
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-scatter_am_iucn_area <- ggplot(spp_df, aes(x = iucn_area, y = am_area, color = iucn_category)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline(slope = 1, intercept = 0, color = 'red') +
-  labs(x = 'Area (km^2) from IUCN dataset', 
-       y = 'Area (km^2) from AquaMaps dataset', 
-       title = 'Comparing total area of species range maps')
-
-ggsave('figures/CompareTotalArea_AM-IUCN_scatter.png', 
-       plot = scatter_am_iucn_area, 
-       width = 8, height = 4, dpi = 100)
-
-print(scatter_am_iucn_area)
-```
-
-There are large differences so to get a better sense, here is the same comparison but with area log transformed.
-``` {r echo = FALSE, warning = FALSE, message = FALSE}
-
-spp_df <- spp_df %>%
-  mutate(am_log = log(am_area),
-         iucn_log = log(iucn_area))
-
-scatter_am_iucn_area_log <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline(slope = 1, intercept = 0, color = 'red') +
-  labs(x = 'IUCN Area (log transformed)', 
-       y = 'AquaMaps Area (log transformed)',
-       title = 'Comparing area of range maps\n (all area values are km^2 log transformed)', 
-       fill = 'IUCN Category')
-
-ggsave('figures/CompareTotalAreaLOG_AM-IUCN_scatter.png', 
-       plot = scatter_am_iucn_area_log, 
-       width = 8, height = 4, dpi = 100)
-
-print(scatter_am_iucn_area_log)
-```
-
-Data split into species groups and by IUCN category
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-scatter_am_iucn_alog_gp <- ggplot(spp_df, aes(iucn_log, am_log, color = iucn_category)) + 
-  ggtheme_plot + 
-  geom_point() +
-  facet_wrap(~spp_group) + 
-  geom_abline() + 
-  labs(x = 'IUCN Area (log transformed)', 
-       y = 'AquaMaps Area (log transformed)',
-       title = 'Comparing differences in area across species groups', 
-       fill = 'IUCN Category')
-
-ggsave('figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png', 
-       plot = scatter_am_iucn_alog_gp, 
-       width = 8, height = 6, dpi = 100)
-
-print(scatter_am_iucn_alog_gp)
-```
-***
-
-###(3) Looking at the overlap between species range maps from IUCN and AquaMaps
-
-This plot shows the total percent of a species range map that is (1) only covered by IUCN cells, (2) only covered by AquaMaps cells and (3) covered by both IUCN and AquaMaps cells.
-
-####How this breaks down between AquaMaps and IUCN
-
-Here is an example of a single species, the yellow speckled chromis (*Chromis alpha*)
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-spp <- 'Chromis alpha'
-
-spp_map <- plot_rangemap(spp)
-
-print(spp_map)
-ggsave(plot = spp_map,
-       filename = sprintf('figures/rangemap_%s.png', spp),
-       width = 8, height = 4, dpi = 200)
-```
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-  spp_df2 <- gather(spp_list, perc.type, percent, perc, am_perc, iucn_perc) %>%
-    dplyr::select(sciname, spp_group, perc.type, percent) %>%
-    unique()
-
-  perc_c <- spp_df2 %>%
-    filter(sciname == spp)
-
-  p1 <- perc_c %>% 
-    dplyr::select(percent)
-
-```
-
-For `r spp`, `r round(pi[1], 2)`% of the total area is covered by both IUCN and AquaMaps, `r round(p1[2, ], 2)`% is covered just by AquaMaps and `r round(p1[3, ], 2)`% is covered just by IUCN.
-
-***
-
-####Now looking across all species, here is the breakdown of how much overlap there is between the two datasets
-
-This plot shows that the majority of species range maps have only ~30% of their full map covered by both maps, and very few species have more than 50% of their two ranges (IUCN and AquaMaps) overlapping.
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
-
-### look at density plot
-
-overlap <- spp_df2 %>%
-  dplyr::filter(perc.type == 'perc')
-
-hist_overlap_pct <- ggplot(overlap, aes(percent)) +
-  ggtheme_plot + 
-  geom_histogram(aes(y = ..count../1926), binwidth = 5, alpha = .8) +
-  labs(title = 'Percent of total range shared between datasets',
-       x = '% total range that is shared b/w both datasets', 
-       y = 'Proportion of Total Species')
-
-ggsave('figures/Overlap-PercTotalRange.png')
-
-print(hist_overlap_pct)
-```
-
-Split up by species - not sure if this is useful or not
-```{r, echo = FALSE, message = FALSE, warning = FALSE}
-
-hist_overlap_pct_gp <- ggplot(overlap, aes(percent)) + 
-  ggtheme_plot + 
-  geom_histogram(binwidth = 5, alpha = .8) +
-  facet_wrap(~spp_group) +
-  labs(title = 'Percent of total range shared, by species group',
-       x = 'Percent of total range that is shared between both datasets', 
-       y = 'Number of Total Species') + 
-  theme(strip.text.x = element_text(size = 8))
-
-ggsave('figures/Overlap-PercTotalRange_spGroups.png')
-print(hist_overlap_pct_gp)
-
+``` {r, child = 'R/data_explore_pt2.Rmd', eval = FALSE}
 ```
@@
 
-```{r pct small within large, echo = FALSE, warning = FALSE}
+```{r pct small within large, echo = FALSE, warning = FALSE, eval = TRUE}
 
@@
 
-ggsave('figures/PercSmallRangeInLarger.png', width = 6, height = 4)
+ggsave(file.path(dir_fig, 'PercSmallRangeInLarger.png'), width = 6, height = 4)
 
@@
 
-Exploring possible correlation between total species range area and percent overlap of smaller range within the larger:
-``` {r pct overlap vs area, echo = FALSE, warning = FALSE, message = FALSE}
-
-scatter_area_vs_overlap <- ggplot(spp_list, aes(sm_perc, total_area, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline() + 
-  labs(y = 'Total Area (km^2)', 
-       x = 'Percent overlap smaller range in larger', 
-       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreavsPercOverlap.png', width = 6, height = 4)
-
-print(scatter_area_vs_overlap)
-
-```
-
-``` {r pct overlap vs ratio of area, echo = FALSE, warning = FALSE, message = FALSE}
+``` {r pct overlap vs ratio of area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
@@
 
+lm_eqn2 <- function(df){
+  m  <- lm(area_ratio ~ sm_perc, spp_list)
+  eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
+        list(a  = format(coef(m)[1], digits = 2), 
+             b  = format(coef(m)[2], digits = 2), 
+             r2 = format(summary(m)$r.squared, digits = 3)))
+  as.character(as.expression(eq))             
+}
+
 scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = log_total_area)) + 
   ggtheme_plot + 
+  ### color the quadrant backgrounds:
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
+                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
+                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
+                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
+                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
   geom_point() + 
-  geom_abline() + 
-  geom_vline(x = area_ratio_quartiles,
+  geom_vline(x = area_ratio_quartiles[c(1, 3)],
              color = 'grey50', linetype = 'dashed') +
-  geom_hline(y = pct_overlap_quartiles,
+  geom_hline(y = pct_overlap_quartiles[c(1, 3)],
              color = 'grey50', linetype = 'dashed') +
@@
             vjust = 1.5) +
-  labs(x = 'Range area ratio (smaller:larger) * 100%', 
+  labs(x = 'Area match: \n(range area smaller : range area larger) * 100%', 
        y = 'Percent overlap smaller range in larger', 
-       title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
+       title = 'Distribution match: \n % overlap of smaller range w/ larger')
 
-ggsave('figures/AreaRatiovsPercOverlap.png', width = 6, height = 4)
+ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 6, height = 4)
 
@@
 
-scatter_arearatio_vs_overlap2 <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  geom_abline() + 
-  geom_vline(x = area_ratio_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_hline(y = pct_overlap_quartiles,
-             color = 'grey50', linetype = 'dashed') +
-  geom_point(data = spp_list_label, color = 'grey30') + 
-  geom_text(data = spp_list_label, 
-            aes(label = sciname_label), 
-            color = 'grey30',
-            fontface = 'bold.italic',
-            size = 2,
-            hjust = .25,
-            vjust = 1.5) +
-  labs(x = 'Range area ratio (smaller:larger) * 100%', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Ratio of species range vs. percent overlap of \nsmaller range with larger')
+```
 
-ggsave('figures/AreaRatiovsPercOverlap2.png', width = 6, height = 4)
+```{r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
-print(scatter_arearatio_vs_overlap2)
+for (i in 1:length(spp_quads)) { # i = 1
+  spp <- spp_quads[i]
+  spp_quad <- names(spp_quads)[i]
+
+  spp_plot <- plot_rangemap(spp)
+  print(spp_plot)
+
+  ggsave(filename = file.path(dir_fig, sprintf('rangemap_%s_%s.png', spp_quad, str_replace(tolower(spp), ' ', '_'))),
+         plot = spp_plot, 
+         width = 8, height = 4, dpi = 200)
+}
 
@@
 
-``` {r examine_spp_gp_quartiles}
+``` {r examine_spp_gp_quartiles, echo = FALSE, eval = TRUE}
 
@@
 
-spp_gp_q1 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE)) %>%
+spp_list_q1 <- spp_list1 %>%
+  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE))
+spp_list_q2 <- spp_list1 %>%
+  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE))
+spp_list_q3 <- spp_list1 %>%
+  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE))
+spp_list_q4 <- spp_list1 %>%
+  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE))
+
+spp_gp_q1 <- spp_list_q1 %>%
   group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q1 = n())
-spp_gp_q2 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE)) %>%
+  summarize(n_spp = first(n_spp), n_spp_q1 = n())
+spp_gp_q2 <- spp_list_q2 %>%
   group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q2 = n())
-spp_gp_q4 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE)) %>%
+  summarize(n_spp = first(n_spp), n_spp_q2 = n())
+spp_gp_q3 <- spp_list_q3 %>%
   group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q4 = n())
-spp_gp_q3 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE)) %>%
+  summarize(n_spp = first(n_spp), n_spp_q3 = n())
+spp_gp_q4 <- spp_list_q4 %>%
   group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp),
-            n_spp_q3 = n())
+  summarize(n_spp = first(n_spp), n_spp_q4 = n())
 
@@
               select(spp_group, pct_q1 = pct_quad),
-            by = 'spp_group')
+            by = 'spp_group') %>%
+  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
+         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
+         quad = str_replace(quad, 'q3', 'q3 area-matched'),
+         quad = str_replace(quad, 'q4', 'q4 poorly matched'))
 
@@
   ggtheme_plot + 
-  geom_bar(stat = 'bin') +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0) +
+  geom_bar(stat = 'bin', alpha = .5) +
+  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
+                               'q2 dist-matched'   = 'greenyellow', 
+                               'q3 area-matched'   = 'orange', 
+                               'q4 poorly matched' = 'orangered2')) +
+  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 4) +
   coord_flip() +
@@
 
+print(barchart_spp_gp_quads)
+ggsave(plot = barchart_spp_gp_quads,
+       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
+       width = 6, height = 4)
+
 ```
 
-```{r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE}
+Plotting species from q1 (well-matched)
+```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
-for (i in 1:length(spp_quads)) { # i = 1
-  spp <- spp_quads[i]
-  spp_quad <- names(spp_quads)[i]
+x <- 20
+y <- nrow(spp_list_q1)
 
+spp_names_q1 <- spp_list_q1$sciname[seq(1, y, y/x)]
+
+for (spp in spp_names_q1) { # i = 1
   spp_plot <- plot_rangemap(spp)
@@
 
-  ggsave(filename = sprintf('figures/rangemap_%s_%s.png', spp_quad, str_replace(tolower(spp), ' ', '_')),
+  ggsave(filename = file.path(dir_fig, sprintf('rangemap_q1_%s.png', str_replace(tolower(spp), ' ', '_'))),
          plot = spp_plot, 
@@
 
-``` {r, echo = FALSE, warning = FALSE, message = FALSE}
+Plotting species from q2 (distribution-matched)
+```{r plot_spp_from_q2, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
-# Regression
+x <- 20
+y <- nrow(spp_list_q2)
 
-lm_eqn <- function(df){
-  m  <- lm(total_area ~ sm_perc, spp_list)
-  eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
-        list(a  = format(coef(m)[1], digits = 2), 
-             b  = format(coef(m)[2], digits = 2), 
-             r2 = format(summary(m)$r.squared, digits = 3)))
-  as.character(as.expression(eq))             
+spp_names_q2 <- spp_list_q2$sciname[seq(1, y, y/x)]
+
+for (spp in spp_names_q2) { # i = 1
+  spp_plot <- plot_rangemap(spp)
+  print(spp_plot)
+
+  ggsave(filename = file.path(dir_fig, sprintf('rangemap_q2_%s.png', str_replace(tolower(spp), ' ', '_'))),
+         plot = spp_plot, 
+         width = 8, height = 4, dpi = 200)
 }
-
-scatter_area_vs_overlap <- scatter_area_vs_overlap + 
-  geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)
-
-ggsave('figures/CorrelationAreavsPercOverlap.png')
-
-print(scatter_area_vs_overlap)
-
-summary(lm(total_area ~ sm_perc, spp_list))
-```
-
-***
-
-Percent coverage as function of area, clipping to area below 100,000,000 km^2:
-``` {r, echo = FALSE, warning = FALSE, message = FALSE}
-# remove all area greater than 100000000
-
-spp_small_df <- spp_list %>% 
-  filter(total_area < 100000000)
-
-scatter_a_vs_overlap_small <- ggplot(spp_small_df, aes(total_area, sm_perc, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  labs(x = 'Total Area (km^2)', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')
-
-ggsave('figures/AreavsPercOverlap_2.png', width = 6, height = 4)
-
-print(scatter_a_vs_overlap_small)
 
@@
 
-Percent coverage as a function of log(area)
-``` {r, echo = FALSE, warning = FALSE, message = FALSE}
-# Log transform area
+Plotting species from q3 (area-matched)
+```{r plot_spp_from_q3, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
-scatter_logA_vs_overlap <- ggplot(spp_list, aes(log_total_area, sm_perc, color = sm_range)) + 
-  ggtheme_plot + 
-  geom_point() + 
-  labs(x = 'Log(total Area)', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Size of species range (log) vs. percent overlap of \nsmaller range with larger')
+x <- 20
+y <- nrow(spp_list_q3)
 
-ggsave('figures/AreavsPercOverlap_log.png', width=6, height=4)
+spp_names_q3 <- spp_list_q3$sciname[seq(1, y, y/x)]
 
-print(scatter_logA_vs_overlap)
-```
+for (spp in spp_names_q3) { # i = 1
+  spp_plot <- plot_rangemap(spp)
+  print(spp_plot)
 
-``` {r examine regressions, eval = FALSE, echo = FALSE, warning = FALSE}
-
-A_mdl      <- lm(total_area ~ sm_perc, spp_list)
-#                Estimate Std. Error t value Pr(>|t|)    
-#   (Intercept) -16285410    2858388  -5.697  1.4e-08 ***
-#   sm_perc        533370      41702  12.790  < 2e-16 ***
-#   Multiple R-squared:  0.07732,	Adjusted R-squared:  0.07685
-
-A_mdl_clip <- lm(total_area ~ sm_perc, l)
-#               Estimate Std. Error t value Pr(>|t|)    
-#   (Intercept)  2993939     810560   3.694 0.000227 ***
-#   sm_perc       153523      11952  12.844  < 2e-16 ***
-#   Multiple R-squared:  0.07955,	Adjusted R-squared:  0.07906 
-logA_mdl   <- lm(log_total_area ~ sm_perc, spp_list)
-#                Estimate Std. Error t value Pr(>|t|)    
-#   (Intercept) 14.775325   0.096446   153.2   <2e-16 ***
-#   sm_perc      0.016885   0.001407    12.0   <2e-16 ***
-#   Multiple R-squared:  0.0687,	Adjusted R-squared:  0.06823 
-```
-
-***
-
-###(5) Examining data broken into quartiles to see if smaller patterns emerge
-``` {r examine quartiles, eval = FALSE, echo = FALSE, warning = FALSE}
-
-summary(spp_list$total_area)
-#      Min.   1st Qu.    Median      Mean   3rd Qu.      Max.
-#     55760   3032000  10790000  18500000  22670000 372100000
-
-qtiles <- c(quantile(spp_list$total_area, 0.0, na.rm = TRUE),
-            quantile(spp_list$total_area, .25, na.rm = TRUE),
-            quantile(spp_list$total_area, .50, na.rm = TRUE),
-            quantile(spp_list$total_area, .75, na.rm = TRUE),
-            quantile(spp_list$total_area, 1.0, na.rm = TRUE))
-
-qtile_mdl <- list()
-for (i in 1:4) {
-  area_clip <- spp_list %>%
-    filter(total_area >= qtiles[i] & total_area < qtiles[i + 1])
-  qtile_mdl[[i]] <- lm(total_area ~ sm_perc, area_clip)
-
-  qtile_plot <- ggplot(area_clip, aes(total_area, sm_perc, color = sm_range)) + 
-    ggtheme_plot + 
-    geom_point() + 
-    labs(x = 'Total Area (km^2)', 
-         y = 'Percent overlap smaller range in larger', 
-         title = sprintf('Size of species range (quartile %s) vs. percent overlap of \nsmaller range with larger', i))
-
-  ggsave(paste0('figures/AreavsOverlap_quantile_', i, '.png'), width = 6, height = 4)
-  print(qtile_plot)
+  ggsave(filename = file.path(dir_fig, sprintf('rangemap_q3_%s.png', str_replace(tolower(spp), ' ', '_'))),
+         plot = spp_plot, 
+         width = 8, height = 4, dpi = 200)
 }
-# First quartile regression:
-#                 Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept)  1423917     101053  14.091   <2e-16 ***
-#     sm_perc         1188       1555   0.764    0.445    
-#     Multiple R-squared:  0.001198,	Adjusted R-squared:  -0.0008529 
-#
-# second quartile:
-#                 Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept)  7011028     275943  25.408  < 2e-16 ***
-#     sm_perc       -17725       4228  -4.192 3.28e-05 ***
-#     Multiple R-squared:  0.03483,	Adjusted R-squared:  0.03285 
-#
-# third quartile:
-#                 Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept) 12813279     507856  25.230  < 2e-16 ***
-#     sm_perc        64710       8159   7.931  1.5e-14 ***
-#     Multiple R-squared:  0.1146,	Adjusted R-squared:  0.1128 
-#
-# fourth quartile:
-#                   Estimate Std. Error t value Pr(>|t|)    
-#     (Intercept) -101736351   17259544  -5.894 7.04e-09 ***
-#     sm_perc        1911286     215216   8.881  < 2e-16 ***
-#     Multiple R-squared:  0.1399,	Adjusted R-squared:  0.1381
-
-# This avenue of exploration looks like a bust.
 
@@
 
-***
+Plotting species from q4 (poorly matched)
+```{r plot_spp_from_q4, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
-###(6) Looking at how many and what type (threat category) of species we are losing by setting a threshold of 0.4 on the aquamaps species
+x <- 20
+y <- nrow(spp_list_q4)
 
-This looks at the full Aquamaps dataset... should it be limited to species with valid categories?
+spp_names_q4 <- spp_list_q4$sciname[seq(1, y, y/x)]
 
-```{r aquamaps probabilities, warning = FALSE, message = FALSE, cache.lazy = FALSE, cache = TRUE, verbose = FALSE, results = FALSE, echo = FALSE, include = FALSE}
+for (spp in spp_names_q4) { # i = 1
+  spp_plot <- plot_rangemap(spp)
+  print(spp_plot)
 
-# get spp_cells raw data for aquamaps
-dir_aquamaps <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO/raw/aquamaps_2014')
-
-spp_cells <- read_csv(file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), 
-                      col_types = '_ccd__', progress = TRUE)
-
-# spp_cells <- fread(input = file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), verbose = FALSE, header = TRUE, 
-#                   colClasses = c('NULL', 'character', NA, 'numeric', 'NULL', 'NULL')) #1 min 49 sec
-
-setnames(spp_cells, c('SpeciesID', 'CsquareCode', 'probability'), c('species_id', 'csquare_code', 'probability'))
-
-spp_cells <- as.data.frame(spp_cells)
-
-spp <- fread(input = file.path(dir_aquamaps, 'tables/ohi_speciesoccursum.csv'), 
-             verbose = FALSE, header = TRUE) %>%
-  mutate(sciname = paste(Genus, Species, sep = ' '),
-         iucn_code = str_replace(toupper(iucn_code), 'LR/', ''),
-         iucn_code = str_replace(iucn_code, 'N.E.', 'NE'))
-spp <- as.data.frame(spp)
+  ggsave(filename = file.path(dir_fig, sprintf('rangemap_q4_%s.png', str_replace(tolower(spp), ' ', '_'))),
+         plot = spp_plot, 
+         width = 8, height = 4, dpi = 200)
+}
 
@@
 
-```{r hist_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
-
-### create dataframe that gives me all species, their total number of cells, 
-### total number of cells >= 0.4 prob, difference between the two and threat
-spp_df3 <- spp_cells %>%
-  group_by(species_id) %>%
-  summarize(cells_0.4 = sum(probability >= 0.4),
-            cells_t = sum(probability >= 0)) %>%
-  mutate(prop = cells_0.4 / cells_t) %>%
-  left_join(spp %>%
-#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')), 
-              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
-            by = c('species_id' = 'SPECIESID')) %>%
-  select(sciname, species_id, cells_0.4, cells_t, prop, iucn_code)
-
-hist_area_vs_threshold <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
-  ggtheme_plot + 
-  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
-  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
-       y = 'Proportion of all aquamaps species (17,348 total)')
-
-ggsave('figures/AMThreshold.png', width = 6, height = 4)
-
-print(hist_area_vs_threshold)
-
-```
-
-```{r scatter_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
-
-### create dataframe that gives me all species, their total number of cells, 
-### total number of cells >= 0.4 prob, difference between the two and threat
-
-threshold_df <- data.frame()
-for (i in seq(0, 1, .05)) { # i = .05
-  temp_df <- spp_cells %>%
-    group_by(species_id) %>%
-    summarize(thresh       = i,
-              cells_thresh = sum(probability >= thresh),
-              cells_total  = sum(probability >= 0),
-              proportion   = cells_thresh/cells_total) %>%
-    left_join(spp %>%
-#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')), 
-              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
-            by = c('species_id' = 'SPECIESID')) %>%
-    select(sciname, species_id, thresh, cells_thresh, cells_total, proportion, iucn_code)
-
-  threshold_df <- threshold_df %>%
-    bind_rows(temp_df)
-}
-
-thresh_df_summary <- threshold_df %>%
-  group_by(thresh) %>%
-  summarize(mean_prop   = mean(proportion, na.rm = TRUE),
-            median_prop = median(proportion, na.rm = TRUE))
-
-scatter_area_vs_threshold <- ggplot(thresh_df_summary, aes(x = thresh, y = mean_prop)) + 
-  ggtheme_plot + 
-  geom_point(aes(y = median_prop), color = 'grey60') +
-  geom_point() + 
-  geom_abline(slope = -1, intercept = 1, color = 'red') + 
-  geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
-  ylim(0, 1) +
-  labs(title = 'Mean (median) species range after presence threshold cutoff',
-       x = 'Presence threshold cutoff', 
-       y = 'Mean (median) relative species range')
-
-ggsave('figures/scatter_AMThreshold.png', width = 6, height = 4)
-
-print(scatter_area_vs_threshold)
-
+``` {r, child = 'R/data_explore_am_thresh.Rmd', eval = FALSE}
 ```

### 2015-11-21 R/data_explore_fxns.R

@@ -1,0 +1,219 @@
+### data_explore_fxns.R
+
+### generic theme for all plots
+ggtheme_plot <- theme(axis.ticks = element_blank(),
+                      text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
+                      plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
+                      legend.position = 'right')
+
+# ggtheme_map <- theme(axis.ticks = element_blank(), axis.text = element_blank(),
+#                     text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
+#                     plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
+#                     legend.position = 'right')
+
+### theme for species range maps - based on generic plot theme
+ggtheme_map <- ggtheme_plot +
+  theme(axis.text = element_blank())
+
+### Function to create a species list from scratch... use when new data is available
+create_spp_list <- function(spp_list_base, data_file) {
+  spp_list <- spp_list_base %>%
+    mutate(total_area = NA,
+           am_area    = NA,
+           iucn_area  = NA,
+           perc       = NA,
+           am_perc    = NA,
+           iucn_perc  = NA,
+           sm_perc    = NA, #filled with % of smaller range that is within the larger range
+           sm_range   = NA) #what range is smaller
+  
+  
+  #there's likely a better way to do this than a for loop (lapply or sapply?)
+  for (i in 1:nrow(spp_list)){
+    spp     <- spp_list[i, ]$sciname #species name
+    spp_map <- get_spp_map(spp) #get species map
+    
+    ###get just the aquamaps cells
+    am_spp_map <- spp_map %>%
+      filter(!is.na(am_prob))%>%
+      dplyr::select(loiczid, am_sid, am_prob) %>% 
+      ### add in am_prob when we want to look at probability
+      unique()
+    
+    if(any(duplicated(am_spp_map$loiczid))){
+      am_spp_map <- am_spp_map %>%
+        dplyr::select(loiczid, am_sid) %>% 
+        unique()
+    }
+    
+    ### rasterize aquamaps map
+    r_am_spp <- subs(loiczid_raster, 
+                     am_spp_map[ , c('loiczid', 'am_sid')], 
[…]

### 2015-11-21 data_explore.Rmd
plotting more maps per quadrant

@@
   spp_plot <- plot_rangemap(spp)
+  spp_solo <- spp_list %>% filter(sciname == spp)
+  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
+  spp_plot <- spp_plot +
+    labs(title = sprintf('%s, quadrant 1: %s', spp, text_tag))
+
   print(spp_plot)
@@
   spp_plot <- plot_rangemap(spp)
+  spp_solo <- spp_list %>% filter(sciname == spp)
+  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
+  spp_plot <- spp_plot +
+    labs(title = sprintf('%s, quadrant 2: %s', spp, text_tag))
   print(spp_plot)
@@
   spp_plot <- plot_rangemap(spp)
+  spp_solo <- spp_list %>% filter(sciname == spp)
+  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
+  spp_plot <- spp_plot +
+    labs(title = sprintf('%s, quadrant 3: %s', spp, text_tag))
   print(spp_plot)
@@
   spp_plot <- plot_rangemap(spp)
+  spp_solo <- spp_list %>% filter(sciname == spp)
+  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
+  spp_plot <- spp_plot +
+    labs(title = sprintf('%s, quadrant 4: %s', spp, text_tag))
   print(spp_plot)

### 2015-11-21 data_explore.Rmd
adding map of all species to look at global representation across latitudes

--- 2015-11-18 data_explore.Rmd
@@
 
+library(rasterVis)
 library(ggplot2)
@@
 # per cell
+
 get_spp_map <- function(species){
@@
 ```
+***
+
+###Global representation of species
+
+By counting the number of unique species are found in each of the half degree cells we can then create a single global raster of number of species.
+
+```{r map, echo=F, warning=F, message=F}
+
+# Number of aquamaps species per cell
+
+am_n_per_cell = filter(am_spp_cells,am_sid %in% unique(spp_list$am_sid))%>%
+                  group_by(loiczid)%>%
+                   summarize(count.am=n())%>%
+                    as.data.frame()
+
+am_sp_map = subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'count.am')], 
+                     by = 'loiczid', 
+                     which = 'count.am', 
+                     subsWithNA = TRUE)
+
+#number of IUCN species per cell
+iucn_n_per_cell = iucn_spp_cells%>%
+                   group_by(loiczid)%>%
+                    summarize(count.iucn=n())%>%
+                    as.data.frame()
+
+iucn_sp_map = subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'count.iucn')], 
+                     by = 'loiczid', 
+                     which = 'count.iucn', 
+                     subsWithNA = TRUE)
+
+#Plot both maps
+myTheme=rasterTheme(region=brewer.pal('Blues', n=9))
+
+levelplot(am_sp_map,par.settings=myTheme,main='Aquamaps Dataset',colorkey=list( at=seq(0, 1063, length=15)))
+
+levelplot(iucn_sp_map,par.settings=myTheme,main='IUCN Dataset',colorkey=list(at=seq(0, 1063, length=15)))
+```
+
+Now look at all aquamaps species
+
+```{r all aquamaps,eval=F}
+
+am_spp_cells_big = fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/ohi_hcaf_species_native.csv'))
+#am_spp_cell_big is missing loiczid info so bring in hcaf to match
+
+hcaf <- read_csv(file.path(dir_anx,'raw/aquamaps_2014/tables/hcaf.csv')) #bring this in just to get loiczid
+
+all_am = am_spp_cells_big%>%
+          group_by(loiczid)%>%
+          summarize(count=n())%>%
+          as.data.frame()
+
+all_am_map = subs(loiczid_raster, all_am[ , c('loiczid', 'count')], 
+                     by = 'loiczid', 
+                     which = 'count', 
+                     subsWithNA = TRUE)
+
+levelplot(all_am_map,par.settings=myTheme,main='Entire Aquamaps Dataset',colorkey=list(at=seq(0, 1063, length=15)))
+```
 

### 2015-11-23 data_explore.Rmd
added maps of species from iucn and aquamaps datasets

--- 2015-11-21 data_explore.Rmd
@@
 ```
+***
+
+###Global representation of species
+
+By counting the number of unique species are found in each of the half degree cells we can then create a single global raster of number of species.
+
+```{r map, echo=F, warning=F, message=F}
+
+# Number of aquamaps species per cell
+
+am_n_per_cell = filter(am_spp_cells,am_sid %in% unique(spp_list$am_sid))%>%
+                  group_by(loiczid)%>%
+                   summarize(count.am=n())%>%
+                    as.data.frame()
+
+am_sp_map = subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'count.am')], 
+                     by = 'loiczid', 
+                     which = 'count.am', 
+                     subsWithNA = TRUE)
+
+#number of IUCN species per cell
+iucn_n_per_cell = iucn_spp_cells%>%
+                   group_by(loiczid)%>%
+                    summarize(count.iucn=n())%>%
+                    as.data.frame()
+
+iucn_sp_map = subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'count.iucn')], 
+                     by = 'loiczid', 
+                     which = 'count.iucn', 
+                     subsWithNA = TRUE)
+
+#Plot both maps
+myTheme=rasterTheme(region=brewer.pal('Blues', n=9))
+
+levelplot(am_sp_map,par.settings=myTheme,main='Aquamaps Dataset',colorkey=list( at=seq(0, 1063, length=15)))
+
+levelplot(iucn_sp_map,par.settings=myTheme,main='IUCN Dataset',colorkey=list(at=seq(0, 1063, length=15)))
+```
+
+Now look at all aquamaps species
+
+```{r all aquamaps,eval=F, echo=F}
+
+am_spp_cells_big = fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/ohi_hcaf_species_native.csv'))
+#am_spp_cell_big is missing loiczid info so bring in hcaf to match
+
+hcaf <- fread(file.path(dir_anx,'raw/aquamaps_2014/tables/hcaf.csv'))%>% #bring this in just to get loiczid
+          as.data.frame()%>%
+            select(CsquareCode,LOICZID)%>%
+              rename(loiczid=LOICZID)
+
+all_am = am_spp_cells_big%>%
+          select(SpeciesID,CsquareCode)%>%
+          group_by(CsquareCode)%>%
+          summarize(count=n())%>%
+          as.data.frame()
+
+all = left_join(all_am,hcaf)
+
+all_am_map = subs(loiczid_raster, all[ , c('loiczid', 'count')], 
+                     by = 'loiczid', 
+                     which = 'count', 
+                     subsWithNA = TRUE)
+
+levelplot(all_am_map,par.settings=myTheme,main='Entire Aquamaps Dataset')
+```
+
+![image](./figures/all_aquamaps_map.png)
 

### 2015-11-25 data_explore.Rmd
cleaning up data_explore and graphs

@@
 
-lm_eqn2 <- function(df){
-  m  <- lm(area_ratio ~ sm_perc, spp_list)
-  eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
-        list(a  = format(coef(m)[1], digits = 2), 
-             b  = format(coef(m)[2], digits = 2), 
-             r2 = format(summary(m)$r.squared, digits = 3)))
-  as.character(as.expression(eq))             
-}
-
 scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = log_total_area)) + 
@@
 
-  ggsave(filename = file.path(dir_fig, sprintf('rangemap_%s_%s.png', spp_quad, str_replace(tolower(spp), ' ', '_'))),
-         plot = spp_plot, 
-         width = 8, height = 4, dpi = 200)
+  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_%s_%s.png', spp_quad,
+  #                                              str_replace(tolower(spp), ' ', '_'))),
+  #        plot = spp_plot, 
+  #        width = 8, height = 4, dpi = 200)
 }
@@
 
-Plotting species from q1 (well-matched)
-```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+Plotting species from q1 (well-matched) (see data_explore_quadmaps.html for plots)
+```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 
-  ggsave(filename = file.path(dir_fig, sprintf('rangemap_q1_%s.png', str_replace(tolower(spp), ' ', '_'))),
-         plot = spp_plot, 
-         width = 8, height = 4, dpi = 200)
+  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q1_%s.png', str_replace(tolower(spp), ' ', '_'))),
+  #        plot = spp_plot, 
+  #        width = 8, height = 4, dpi = 200)
 }
@@
 Plotting species from q2 (distribution-matched)
-```{r plot_spp_from_q2, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+```{r plot_spp_from_q2, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 
-  ggsave(filename = file.path(dir_fig, sprintf('rangemap_q2_%s.png', str_replace(tolower(spp), ' ', '_'))),
-         plot = spp_plot, 
-         width = 8, height = 4, dpi = 200)
+  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q2_%s.png', str_replace(tolower(spp), ' ', '_'))),
+  #        plot = spp_plot, 
+  #        width = 8, height = 4, dpi = 200)
 }
@@
 ```
+
+Plotting species from q3 (area-matched)
+```{r plot_spp_from_q3, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+
+x <- 20
+y <- nrow(spp_list_q3)
+
+spp_names_q3 <- spp_list_q3$sciname[seq(1, y, y/x)]
+
+for (spp in spp_names_q3) { # i = 1
+  spp_plot <- plot_rangemap(spp)
+  spp_solo <- spp_list %>% filter(sciname == spp)
+  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
+  spp_plot <- spp_plot +
+    labs(title = sprintf('%s, quadrant 3: %s', spp, text_tag))
+  print(spp_plot)
+
+  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q3_%s.png', str_replace(tolower(spp), ' ', '_'))),
+  #        plot = spp_plot, 
+  #        width = 8, height = 4, dpi = 200)
+}
+
+```
+
+Plotting species from q4 (poorly matched)
+```{r plot_spp_from_q4, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+
+x <- 20
+y <- nrow(spp_list_q4)
+
+spp_names_q4 <- spp_list_q4$sciname[seq(1, y, y/x)]
+
+for (spp in spp_names_q4) { # i = 1
+  spp_plot <- plot_rangemap(spp)
+  spp_solo <- spp_list %>% filter(sciname == spp)
+  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
+  spp_plot <- spp_plot +
+    labs(title = sprintf('%s, quadrant 4: %s', spp, text_tag))
+  print(spp_plot)
+
+  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q4_%s.png', str_replace(tolower(spp), ' ', '_'))),
+  #        plot = spp_plot, 
+  #        width = 8, height = 4, dpi = 200)
+}
+
+```
+
 ***
@@
 #Plot both maps
-myTheme=rasterTheme(region=brewer.pal('Blues', n=9))
+myTheme = rasterTheme(region = brewer.pal('Blues', n = 9))
 
-levelplot(am_sp_map,par.settings=myTheme,main='Aquamaps Dataset',colorkey=list( at=seq(0, 1063, length=15)))
+levelplot(am_sp_map, par.settings = myTheme, main = 'Aquamaps Dataset', colorkey = list( at = seq(0, 1063, length = 15)))
 
-levelplot(iucn_sp_map,par.settings=myTheme,main='IUCN Dataset',colorkey=list(at=seq(0, 1063, length=15)))
+levelplot(iucn_sp_map, par.settings = myTheme, main = 'IUCN Dataset', colorkey = list(at = seq(0, 1063, length = 15)))
 ```
@@
 
-hcaf <- fread(file.path(dir_anx,'raw/aquamaps_2014/tables/hcaf.csv'))%>% #bring this in just to get loiczid
-          as.data.frame()%>%
-            select(CsquareCode,LOICZID)%>%
-              rename(loiczid=LOICZID)
+hcaf <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/hcaf.csv')) %>% #bring this in just to get loiczid
+          as.data.frame() %>%
+            select(CsquareCode, LOICZID) %>%
+              rename(loiczid = LOICZID)
 
-all_am = am_spp_cells_big%>%
-          select(SpeciesID,CsquareCode)%>%
-          group_by(CsquareCode)%>%
-          summarize(count=n())%>%
+all_am = am_spp_cells_big %>%
+          select(SpeciesID,CsquareCode) %>%
+          group_by(CsquareCode) %>%
+          summarize(count=n()) %>%
           as.data.frame()
 
-all = left_join(all_am,hcaf)
+all = left_join(all_am, hcaf)
 
@@
 
-levelplot(all_am_map,par.settings=myTheme,main='Entire Aquamaps Dataset')
+levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
 ```
@@
 ![image](./figures/all_aquamaps_map.png)
-
-Plotting species from q3 (area-matched)
-```{r plot_spp_from_q3, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-
-x <- 20
-y <- nrow(spp_list_q3)
-
-spp_names_q3 <- spp_list_q3$sciname[seq(1, y, y/x)]
-
-for (spp in spp_names_q3) { # i = 1
-  spp_plot <- plot_rangemap(spp)
-  spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
-  spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 3: %s', spp, text_tag))
-  print(spp_plot)
-
-  ggsave(filename = file.path(dir_fig, sprintf('rangemap_q3_%s.png', str_replace(tolower(spp), ' ', '_'))),
-         plot = spp_plot, 
-         width = 8, height = 4, dpi = 200)
-}
-
-```
-
-Plotting species from q4 (poorly matched)
-```{r plot_spp_from_q4, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-
-x <- 20
-y <- nrow(spp_list_q4)
-
-spp_names_q4 <- spp_list_q4$sciname[seq(1, y, y/x)]
-
-for (spp in spp_names_q4) { # i = 1
-  spp_plot <- plot_rangemap(spp)
-  spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
-  spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 4: %s', spp, text_tag))
-  print(spp_plot)
-
-  ggsave(filename = file.path(dir_fig, sprintf('rangemap_q4_%s.png', str_replace(tolower(spp), ' ', '_'))),
-         plot = spp_plot, 
-         width = 8, height = 4, dpi = 200)
-}
-
-```
 

### 2015-11-25 data_explore.Rmd
moved quad maps to new .html, added ext risk as color/size for main quad plot

@@
 
-scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = log_total_area)) + 
+scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = popn_category, size = category_score)) + 
   ggtheme_plot + 
@@
   geom_point() + 
+  scale_colour_manual(values = c('LC'   = 'green4', 
+                                 'NT'   = 'greenyellow', 
+                                 'VU'   = 'yellow', 
+                                 'EN'   = 'orange', 
+                                 'CR' = 'orangered2')) +
   geom_vline(x = area_ratio_quartiles[c(1, 3)],
@@
              color = 'grey50', linetype = 'dashed') +
-  geom_point(data = spp_list_label, color = 'tomato2') + 
+  geom_point(data = spp_list_label, aes(size = category_score), color = 'grey30') + 
   geom_text(data = spp_list_label, 
             aes(label = sciname_label), 
-            color = 'tomato2',
+            color = 'grey30',
             fontface = 'bold.italic',

### 2015-11-25 data_explore.Rmd

--- 2015-11-25 data_explore.Rmd
@@
 
+### define 25% and 75% quartiles for both x and y axes
 area_ratio_quartiles <- quantile(spp_list$area_ratio, na.rm = TRUE)[2:4]
@@
 
+### define representative species from each quadrant, for over-plotting
 spp_quads <- c('q1' = 'Kajikia albida', 
@@
 
+### truncate the species names for the representative species
 spp_list_label <- spp_list %>%
@@
 
-scatter_arearatio_vs_overlap <- ggplot(spp_list, aes(x = area_ratio, y = sm_perc, color = log_total_area)) + 
+### mongo plot time
+scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
+                                         transform(popn_category = reorder(popn_category, category_score)),
+                                       aes(x = area_ratio, y = sm_perc, 
+                                           color = popn_category,
+                                           size = popn_category
+                                           )) + 
   ggtheme_plot + 
@@
   geom_point() + 
+  scale_colour_manual(values = c('LC' = 'green4', 
+                                 'NT' = 'greenyellow', 
+                                 'VU' = 'yellow', 
+                                 'EN' = 'orange', 
+                                 'CR' = 'orangered2')) +
   geom_vline(x = area_ratio_quartiles[c(1, 3)],
@@
              color = 'grey50', linetype = 'dashed') +
-  geom_point(data = spp_list_label, color = 'tomato2') + 
+  geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
   geom_text(data = spp_list_label, 
             aes(label = sciname_label), 
-            color = 'tomato2',
+            color = 'grey30',
             fontface = 'bold.italic',
@@
   labs(x = 'Area match: \n(range area smaller : range area larger) * 100%', 
-       y = 'Percent overlap smaller range in larger', 
-       title = 'Distribution match: \n % overlap of smaller range w/ larger')
+       y = 'Distribution match: \n % overlap of smaller range w/ larger', 
+       title = 'Area match vs distribution match',
+       color = 'Risk', size = 'Risk')
 
@@
 
+### Create species lists by quadrant
 spp_list_q1 <- spp_list1 %>%
@@
 
+### Summarize quadrant species list by species group (e.g. CORALS1)
 spp_gp_q1 <- spp_list_q1 %>%
@@
 
+### Join species group by quadrant into a single dataframe
 spp_gp_quadrants <- spp_gp_q1 %>%
@@
 
+### Plot the bar chart
 barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 

### 2015-11-26 data_explore.Rmd
notes added to manuscript; tried a map of difference in frequency between the two datasets but that still needs work

@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE}
+```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
 
@@
 
-```{r get spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 data_file <- 'data/spp_list_w_area_percent_noextinct.csv'
@@
 ``` {r child = 'R/data_explore_pt1.Rmd', eval = FALSE}
+# bar chart: number of spp per group
+# scatter plot: area compare IUCN vs AM, including log plot and plot by species group
+# rangemap: Chromis alpha 
 ```
@@
 
-```{r pct small within large, echo = FALSE, warning = FALSE, eval = TRUE}
+```{r pct_small_within_large, echo = FALSE, warning = FALSE, eval = TRUE}
 
@@
 
-``` {r pct overlap vs ratio of area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
@@
 
-```{r map, echo=F, warning=F, message=F}
+```{r global_dist_map, echo = FALSE, warning = FALSE, message = FALSE}
 
-# Number of aquamaps species per cell
+### find total species in each dataset for later percent calcs
+am_n_tot   <- am_spp_cells %>% 
+  filter(am_sid %in% unique(spp_list$am_sid)) %>%
+  select(am_sid) %>% 
+  unique() %>% nrow()
+iucn_n_tot <- iucn_spp_cells %>% 
+  filter(sciname %in% unique(spp_list$sciname)) %>%
+  select(sciname) %>% 
+  unique() %>% nrow()
+### good - the same number
 
-am_n_per_cell = filter(am_spp_cells,am_sid %in% unique(spp_list$am_sid))%>%
-                  group_by(loiczid)%>%
-                   summarize(count.am=n())%>%
+### Number of aquamaps species per cell
+am_n_per_cell <- am_spp_cells %>%
+  filter(am_sid %in% unique(spp_list$am_sid)) %>%
+  group_by(loiczid) %>%
+  summarize(count_am = n(),
+            pct_am = count_am/am_n_tot) %>%
   as.data.frame()
 
-am_sp_map = subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'count.am')], 
+am_sp_map <- subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'count_am')], 
                   by = 'loiczid', 
-                     which = 'count.am', 
+                  which = 'count_am', 
                   subsWithNA = TRUE)
 
-#number of IUCN species per cell
-iucn_n_per_cell = iucn_spp_cells%>%
-                   group_by(loiczid)%>%
-                    summarize(count.iucn=n())%>%
+### Number of IUCN species per cell
+iucn_n_per_cell <- iucn_spp_cells %>%
+  group_by(loiczid) %>%
+  summarize(count_iucn = n(),
+            pct_iucn = count_iucn/iucn_n_tot) %>%
   as.data.frame()
 
-iucn_sp_map = subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'count.iucn')], 
+iucn_sp_map <- subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'count_iucn')], 
                     by = 'loiczid', 
-                     which = 'count.iucn', 
+                    which = 'count_iucn', 
                     subsWithNA = TRUE)
 
-#Plot both maps
-myTheme = rasterTheme(region = brewer.pal('Blues', n = 9))
+### Plot both maps
+myTheme <- rasterTheme(region = brewer.pal('Blues', n = 9))
 
-levelplot(am_sp_map, par.settings = myTheme, main = 'Aquamaps Dataset', colorkey = list( at = seq(0, 1063, length = 15)))
+levelplot(am_sp_map, par.settings = myTheme, 
+          main = 'Aquamaps Dataset', 
+          colorkey = list( at = seq(0, 1063, length = 15)))
 
-levelplot(iucn_sp_map, par.settings = myTheme, main = 'IUCN Dataset', colorkey = list(at = seq(0, 1063, length = 15)))
+levelplot(iucn_sp_map, par.settings = myTheme, 
+          main = 'IUCN Dataset', 
+          colorkey = list(at = seq(0, 1063, length = 15)))
+
 ```
@@
 
-```{r all aquamaps,eval=F, echo=F}
+```{r all_aquamaps, eval = FALSE, echo = FALSE}
 
-am_spp_cells_big = fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/ohi_hcaf_species_native.csv'))
-#am_spp_cell_big is missing loiczid info so bring in hcaf to match
+am_spp_cells_big <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/ohi_hcaf_species_native.csv')) %>%
+  as.data.frame()
+### am_spp_cell_big is missing loiczid info so bring in hcaf to match
 
@@
 
-all_am = am_spp_cells_big %>%
-          select(SpeciesID,CsquareCode) %>%
+### find total spp to allow for percent calcs later
+am_all_n_tot <- am_spp_cells_big %>% 
+  select(SpeciesID) %>% 
+  unique() %>% nrow()
+
+all_am <- am_spp_cells_big %>%
+  select(SpeciesID, CsquareCode) %>%
   group_by(CsquareCode) %>%
-          summarize(count=n()) %>%
+  summarize(all_count = n(),
+            all_pct   = all_count/am_all_n_tot) %>%
   as.data.frame()
 
-all = left_join(all_am, hcaf)
+all <- all_am %>%
+  left_join(hcaf, by = 'CsquareCode')
 
-all_am_map = subs(loiczid_raster, all[ , c('loiczid', 'count')], 
+all_am_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_count')], 
                    by = 'loiczid', 
-                     which = 'count', 
+                   which = 'all_count', 
                    subsWithNA = TRUE)
@@
 
-![image](./figures/all_aquamaps_map.png)
+![image](figures/all_aquamaps_map.png)
+
+``` {r variations_on_gl_dist_maps, echo = FALSE, eval = FALSE}
+### find distribution by percent of total species
+am_sp_pct_map <- subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'pct_am')], 
+                      by = 'loiczid', 
+                      which = 'pct_am', 
+                      subsWithNA = TRUE)
+iucn_sp_pct_map <- subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'pct_iucn')], 
+                        by = 'loiczid', 
+                        which = 'pct_iucn', 
+                        subsWithNA = TRUE)
+### find difference in percent between the two distribution maps; should be small,
+### if species are well-matched, since it's all the same species
+diff_pct_map    <- iucn_sp_pct_map - am_sp_pct_map
+
+### plot maps with distributions as percentages.  Should it be 0 - 1, or 
+### 0 - max value?
+levelplot(am_sp_pct_map, par.settings = myTheme, 
+          main = 'Aquamaps Dataset pct', 
+          colorkey = list( at = seq(0, 1, length = 15)))
+levelplot(iucn_sp_pct_map, par.settings = myTheme, 
+          main = 'IUCN Dataset pct', 
+          colorkey = list(at = seq(0, 1, length = 15)))
+### this is the difference one. Should it be min value to max value?
+levelplot(diff_pct_map, par.settings = myTheme, 
+          main = 'IUCN pct - AM pct', 
+          colorkey = list(at = seq(-1, 1, length = 15)))
+
+### Do similar, but with the full aquamaps list (should also do this
+### against the full IUCN list...)
+all_am_pct_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_pct')], 
+                   by = 'loiczid', 
+                   which = 'all_pct', 
+                   subsWithNA = TRUE)
+
+diff_pct_map2    <- iucn_sp_pct_map - all_am_pct_map
+levelplot(all_am_pct_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset, pct')
+levelplot(diff_pct_map2, par.settings = myTheme, main = 'iucn (overlap only) - aquamaps (all)')
+
+```
 

### 2015-11-30 R/data_explore_pt2.Rmd
still more ideas? new comparison of area vs overlap...

@@
 
-``` {r pct overlap vs area, echo = FALSE, warning = FALSE, message = FALSE}
+``` {r pct overlap vs tot area, echo = FALSE, warning = FALSE, message = FALSE}
 
@@
 ggsave(file.path(dir_fig, 'AreavsPercOverlap.png'), width = 6, height = 4)
+
+```
+
+``` {r pct overlap vs lg area, echo = FALSE, warning = FALSE, message = FALSE}
+
+spp_list <- spp_list %>%
+  mutate(lg_area = ifelse(am_area > iucn_area, am_area, iucn_area),
+         sm_area = ifelse(am_area < iucn_area, am_area, iucn_area),
+         lg_area_pct = 100 * lg_area / max(iucn_area, am_area, na.rm = TRUE)) ### use largest data set as approx for total ocean range
+
+scatter_lg_area_vs_overlap <- ggplot(spp_list, aes(lg_area_pct, sm_perc, color = sm_range)) + 
+  ggtheme_plot + 
+  geom_point() + 
+  geom_abline() + 
+  labs(x = 'Larger range (% total ocean)', 
+       y = 'Percent overlap smaller range in larger', 
+       title = 'Percent overlap vs. larger range')
+
+print(scatter_lg_area_vs_overlap)
+
+# summary(lm(total_area ~ sm_perc, spp_list))
+
+ggsave(file.path(dir_fig, 'PercOverlapvsLgArea.png'), width = 6, height = 4)
 

### 2015-12-01 R/data_explore_pt1.Rmd

@@
 ggsave(plot = spp_map,
-       filename = file.path(dir_fig, sprintf('rangemap_%s.png', spp)),
+       filename = file.path(dir_fig, sprintf('rangemap_%s.png', str_replace(tolower(spp), ' ', '_'))),
        width = 8, height = 4, dpi = 200)

### 2015-12-01 R/data_explore_pt2.Rmd

@@
          sm_area = ifelse(am_area < iucn_area, am_area, iucn_area),
-         lg_area_pct = 100 * lg_area / max(iucn_area, am_area, na.rm = TRUE)) ### use largest data set as approx for total ocean range
+         lg_area_pct = 100 * lg_area / max(total_area, na.rm = TRUE)) ### use largest data set as approx for total ocean range
 

### 2015-12-01 ms_outline.Rmd

@@
 
-- spatial distribution efforts range from modeled to expert input
-- help identify hotspots for global biodiversity 
-- priorities for conservation
-- marine spatial planning
-- assessing biodiversity 
-- monitor impacts (human, natural) overtime
--
+* Importance of understanding spatial distribution of species
+    * help identify hotspots for global biodiversity 
+    * priorities for conservation
+    * marine spatial planning
+    * assessing biodiversity 
+    * monitor impacts (human, natural) over time (should this be in here?)
+* Challenges
+    * difficulty of monitoring marine species in space
+    * even more difficult to monitor over time
+        * can Aquamaps, as a bioclimatic envelope model, be used to predict changes or used as proxy for species shifts if envelopes are known to have shifted?
+__can we use this as trend?
+or does all that count as pressures?
+seems like human factors can drastically change the trend that wouldn't be captured here__
+        * Does that additional data in Aquamaps include predictions for 2050, 2100 (bioclimatic envelopes) - how is that reflected in the species distributions?
+* IUCN vs Aquamaps
+    * spatial distribution efforts range from modeled to expert input
+    * distribution data for 1900+ species found in both data sets
+    * differences in predictions of species presence in both distribution and area
+        * Aquamaps validation and benefits from _Ready et al. 2010_:
+          -  We present and evaluate AquaMaps, a presence-only species distribution modelling system that allows the incorporation of expert knowledge about habitat usage and was designed for maximum output of standardized species range maps at the global scale.
+In the marine environment there is a significant challenge to the production of range maps due to large biases in the amount and location of occurrence data for most species.
+AquaMaps is compared with traditional presence-only species distribution modelling methods to determine the quality of outputs under equivalently automated conditions.
+_The effect of the inclusion of expert knowledge to AquaMaps is also investigated._
+          - Model outputs were tested internally, through data partitioning, and externally against independent survey data to determine the ability of models to predict presence versus absence.
+Models were also tested externally by assessing correlation with independent survey estimates of relative species abundance.
+AquaMaps outputs compare well to the existing methods tested, and inclusion of expert knowledge results in a general improvement in model outputs.
+The transparency, speed and adaptability of the AquaMaps system, as well as the existing online framework which allows expert review to compensate for sampling biases and thus improve model predictions are proposed as additional benefits for public and research use alike.
+
+* Implications
+    * some species need more attention?
+    * which data sets are more useful for which uses?
+        * who uses IUCN? who uses AquaMaps?  Who uses both?
+    * how can data decisions affect results of research or conservation efforts?
+        * using OHI as an example
 
@@
 Two main sources for marine species are the IUCN and AquaMaps.
-These two datasets differ in methodology and final products, yet the differences between the two are not well documented.
+These two datasets differ in methodology and final products, yet the differences between the two are not well documented *(though Aquamaps, and probably IUCN, each probably have some validation studies)*.
 AquaMaps uses a computer model based on preferential environmental parameters to predict species locations and probability of occurrence, while the IUCN uses expert insight to assign spatial boundaries around known locations of marine species.
@@
 
-![](./figures/SpeciesMap.png)
+![](./figures/rangemap_chromis_alpha.png)
 
 **Figure 1:** Species range map for *Chromis alpha* from the AquaMaps (orange) and IUCN (purple) datasets.
+***Do we want to have this way up here?  needs explanation?***
 
@@
 Shapefiles for individual species. Spatial files include subpopulation information, presence (extinct/extant etc.) and seasonality information (breeding/migration/origin etc.).  
+- ***Include other columns typically within each shapefile***
 
@@
 From [IUCN site](http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources):  
-
-> The species distribution maps, sometimes referred to as 'limits of distribution' or 'field guide' maps, aim to provide the current known distribution of the species within its native range.
+    - The species distribution maps, sometimes referred to as 'limits of distribution' or 'field guide' maps, aim to provide the current known distribution of the species within its native range.
 The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range.
@@
 *A polygon displaying the limits of a species distribution is essentially meant to communicate that the species probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon*.
+
+- ***Spatial files for IUCN species groups attempt to include a broad sample across the group, to avoid sampling bias (maybe?)*** - but only some species groups have had spatial data released, so some taxa are not represented at all
+    - from http://www.iucnredlist.org/technical-documents/spatial-data: *"This spatial data provided below is for comprehensively assessed taxonomic groups."  "...
+the data is divided by taxonomic groups.
+Groups that have been comprehensively assessed are those containing >90% species evaluated within the described taxon group."*
 
@@
 
-AquaMaps (Kirschner et al.
-2013) was developed to create species range maps based on a computer model that takes up to six parameters to determine where marine species are located.
+AquaMaps (Kaschner et al.
+2013) was developed to create species range maps based on a computer model ***Bioclimatic envelope model*** that takes up to six parameters to determine where marine species are located.
 The model uses known species preferences for sea surface temperature, salinity, depth, productivity, and in some instances distance to land and sea ice cover to predict species ranges.
-Range maps are based on a 0.5 degree global grid where values in each cell is equal to the probability of occurrence that the species is there.
-These maps can then be tweaked by experts altering the parameters and regenerating the modeled map.
-As of October 2015, 17,348 species have been mapped.
-See (Kirschner et al.
+Range maps are based on a 0.5 degree global grid where value in each cell is equal to the ***probability of occurrence*** (or ***relative environmental suitability?*** (see Kaschner et al 2006 and 2011)) of that species within that cell These maps can then be tweaked by experts altering the parameters and regenerating the modeled map.
+As of December 2015, ***22,000*** species have been mapped.
+See (Kaschner et al.
 2013) for more detailed methods.
-
-The main differences between these two sources of marine species ranges is how the maps are developed, and their output.
-When a species range is developed during an IUCN assessment, experts determine the boundaries of the species range based on their expertise which could be using known occurrences, habitat preferences, suitable habitat, or other knowledge.
-The range is depicted as a polygon where the boundaries are *"meant to communicate that the species probably occurs within this polygon, but it does not mean that it is distributed equally within the polygon or occurs everywhere within the polygon."*
-
-These two differ in spatial resolution - AquaMaps provides higher resolution information about where species do and do not occur globally whereas the IUCN range maps are unable to provide specific information about where within a polygon a species occurs.
-This is understandable given the mobile nature of many marine species and the difficulty in observing and tracking known occurrences.
 
@@
 
+- from Kaschner 2011: The RES model generates an index of species-specific *relative environmental suitability (RES)* of each individual half degree grid cell by scoring how well its physical attributes matched the known aspects of species’ habitat use.
+RES values range between 0 (not suitable) to 1 (highly suitable) and represent the product of the suitability scores assigned for the individual environmental attributes (bottom depth, SST, sea ice concentration, and distance from land in some cases), which were calculated using pre-defined trapezoidal functional response curves
+- but from Aquamaps website: Relative likelihood values range between 0.00 - 1.00 and represent the product of the relative likelihood scores assigned for the individual environmental attributes, which are in turn calculated based on the environmental envelopes described above.
+***so they just changed the phrase; I like the original better...***
+
 There are inherent issues with AquaMaps. It can't be expected to accurately predict all marine species ranges. Expert insight is sought and incorporated when available. 
+- ***how is expert insight used?
+how frequently?
+new data set includes "reviewed" field*** expert insight seems to be used to generate the environmental suitability envelope for species, to provide sightings info that use used in training the model, and to exclude species from regions of known absence; but the model generates the maps.
+However, some species maps are reviewed by experts (approved?
+changed?
+what happens in the review)
+
+### Comparison between AM and IUCN
+
+#### How map is developed (maybe in above paragraphs?)
+
+- Problem of expert-defined range maps (over-predicting) vs sightings (under-predicting): 
+from **Kaschner et al 2011**:  *Similar to the trade-offs of different habitat prediction modeling approaches [11], these two methods lie on ___opposite ends of a spectrum from potentially overpredicting expert-derived (range maps) to potentially underpredicting (empirical sighting surveys) range sizes___*
+
+The main differences between these two sources of marine species ranges is how the maps are developed, and their output.
+When a species range is developed during an IUCN assessment, experts determine the boundaries of the species range based on their expertise which could be using known occurrences, habitat preferences, suitable habitat, or other knowledge.
+The range is depicted as a polygon where the boundaries are *"meant to communicate that the species probably occurs within this polygon, but it does not mean that it is distributed equally within the polygon or occurs everywhere within the polygon."*
+
+#### Spatial resolution
+These two differ in spatial resolution - AquaMaps provides higher resolution information about where species do and do not occur globally whereas the IUCN range maps are unable to provide specific information about where within a polygon a species occurs.
+This is understandable given the mobile nature of many marine species and the difficulty in observing and tracking known occurrences.
+
+### Defining alignment between the two data sets
+
+* Here we can discuss the idea of multiple "dimensions" of alignment - distributional and area alignment.  Well-aligned species will show high alignment in both dimensions.  
+* Maybe a diagram here?
+* What can alignment tell us about the individual species, the species groups, and the data sets?
+* Note that alignment does not necessarily mean accuracy or precision - a poorly-understood species could appear to be well-aligned.
 
 ***
-##Section Two: Comparison of IUCN and AquaMaps
 
-To compare these two datasets for all species found in each (1920), the IUCN maps were rasterized to the same cell resolution as AquaMaps, 0.5 degrees.
-For every grid cell where the polygon overlapped, a value of 1 was assigned.
-To allow for comparison, the probability of occurrence values in the AquaMaps data were ignored and all cells where there was a value greater than 0 were assigned a 1.
+##Section Two: Comparison of IUCN and AquaMaps ranges
+
+### Determining "presence" for each data set
+
+To compare these two datasets for all species found in each (n = ***1920***?), we rasterized the IUCN polygons to the same cell resolution as AquaMaps: 0.5 degrees in both latitude and longitude.
+For every grid cell where the polygon overlapped ***the center point?
+or is it based on ANY overlap, e.g.
+using weights and including any non-zero weight?
+This could boost IUCN presence***, a value of 1 was assigned.
+For the purposes of comparison, the probability of occurrence values in the AquaMaps data were ignored and all cells where there was a value greater than 0 were assigned a 1.
 This way we have comparable maps for total species range.
+- ***How do we justify a zero-threshold here?*** 
+    - need to understand the "probability" - is it really ***relative environmental suitability***   From Kaschner et al 2011: "Patterns based on a presence threshold RES.0.6 (Fig.
+1) were most strongly supported by empirical species richness data (see ‘Validation of species richness’, below)."
+    - since IUCN "presence" seems to be pretty loose, set a wide tolerance on AM?
+
+### Global distribution between each data set
 
@@
 
-![](./figures/sp_groups_barchart.png)
+![](figures/sp_groups_barchart.png)
+***Maybe also include the species distribution maps here - maybe a map with the difference in coverage between IUCN and Aquamaps? where is IUCN under-represented? and in what species groups?***
 
+![](figures/all_aquamaps_map.png)
+
+<!-- Not sure this is super helpful any more - this info comes out in the quadrant graph
 For each species map, the total area was calculated for all cells where the species was determined as 'possibly present' by both datasets.
@@
 
-![](./figures/CompareTotalAreaLOG_AM-IUCN_scatter.png)
-![](./figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png)
+![](figures/CompareTotalAreaLOG_AM-IUCN_scatter.png)
+![](figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png)
+-->
 
-For each species range map, the total area was calculated for all cells that were included in either species distributions, regardless of source. The number of cells that were found in both
+***This shows differences area maybe?*** What's the story of this plot...
+![](figures/CompareTotalAreaLOG_AM-IUCN_scatter.png)
 
-![](./figures/Overlap-PercTotalRange.png)
-![](./figures/Overlap-PercTotalRange_spGroups.png)
-![](./figures/PercSmallRangeInLarger.png)
-![](./figures/AreavsPercOverlap_2.png)
-![](./figures/AMThreshold.png)
+### Comparing alignment
+
+#### Comparing alignment by species
+
+For each species range map, the total area was calculated for all cells that were included in either species distributions, regardless of source.
+The number of cells that were found in both *** <- this implies area may just be cell count?***
 
@@
 Figure 2
+***focus on the percent overlap of small within large*** - describe this species in terms of alignment
 
-This is not always the case as some species range maps align better, especially for marine mammals.
+*** Is the area calculated based on cell area, or just number of cells? ***
+
+![](figures/AreaRatiovsPercOverlap.png)
+
+***Walk through the quadrant plot*** - describe meaning of each axis and each quadrant
+
+<!-- These are also essentially included within the quadrant plot?
+![](./figures/Overlap-PercTotalRange.png)
+![](./figures/Overlap-PercTotalRange_spGroups.png)
+![](./figures/PercSmallRangeInLarger.png)
+![](./figures/AreavsPercOverlap_2.png)
+-->
+
+#### Comparing alignment by species group
+
+![](figures/barchart_spp_gp_quads.png)
+
+Which species groups are better aligned?  (save discussion until later?)
+
+### Effect of changing threshold on area representation
+
+![](figures/AMThreshold.png)
+
+***Where does the 40% threshold come from?  Is it commonly used?  Justification?*** Show effect of using a 40% threshold vs a 0% threshold.
+
+![](figures/scatter_AMThreshold.png)
+Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
 
@@
 
+<!--
 ![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)
+-->
 
@@
 
-It is important to remember that although only [X (1,923/total number of species)]% of all species included in the species goal are found in both datasets, there are still significant differences when electing to choose one dataset over the other.
+It is important to remember that although only ***[X (1,923/total number of species)]%*** of all species included in the species goal are found in both datasets, there are still significant differences when electing to choose one dataset over the other.
 
@@
 ##Section Four: Implications of data choices
+
+***need more info on these options? they both sound interesting though***
 - could talk about MPA gap analysis
-- Highlights tudies that have used one and not the other. Could this have changed the results?
+- Highlight studies that have used one and not the other. Could this have changed the results?
 
@@
 ##References
+
+**Key AquaMaps publications:**
+
+Kaschner, K., D.P. Tittensor, J. Ready, T Gerrodette and B. Worm (2011). Current and Future Patterns of Global Marine Mammal Biodiversity. PLoS ONE 6(5): e19653. PDF
+
+Ready, J., K.
+Kaschner, A.B.
+South, P.D Eastwood, T.
+Rees, J.
+Rius, E.
+Agbayani, S.
+Kullander and R.
+Froese (2010).
+Predicting the distributions of marine organisms at the global scale.
+Ecological Modelling 221(3): 467-478.
+PDF
+
+**Papers based on AquaMaps:** ***Use these for section 4?***
+
+Jones, M.C., S.R.
+Dyeb, J.K.
+Pinnegar and W.W.L.
+Cheung (2012).
+Modelling commercial fish distributions: Prediction and assessment using different approaches.
+Ecological Modelling 225(2012): 133-145.
+PDF
+
+Coll, M., C. Piroddi, J. Steenbeek, K. Kaschner, F. Ben Rais Lasram et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. PDF
+
+Martin C.S., Fletcher R., Jones M.C., Kaschner K., Sullivan E., Tittensor D.P., Mcowen C., Geffert J.L., van Bochove J.W., Thomas H., Blyth S., Ravillious C., Tolley M., Stanwell-Smith D.
+(2014).
+Manual of marine nd coastal datasets of biodiversity importance.
+May 2014 release.
+Cambridge (UK): UNEP World Conservation Monitoring Centre.
+28 pp.
+(+ 4 annexes totalling 174 pp.
+and one e-supplement).
+PDF

### 2015-12-01 data_explore_2015_data.Rmd
updated species list including 2015 aquamaps data

--- 2015-11-26 data_explore.Rmd
@@
 #iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), 
+iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both_2015.csv'), 
                            col_types = 'cdddd', progress = TRUE) %>%
@@
 
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells.csv'), progress = TRUE)
+am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), progress = TRUE)
 
@@
 ```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-data_file <- 'data/spp_list_w_area_percent_noextinct.csv'
+data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
 
@@
   ### species info list - scinames, IUCN species ID, AM species ID, category etc
-  spp_list_base <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
+  spp_list_base <- read.csv('shiny/data/spp_am_v_iucn_2015.csv', stringsAsFactors = FALSE)
 
@@
 
-  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_%s_%s.png', spp_quad,
-  #                                              str_replace(tolower(spp), ' ', '_'))),
-  #        plot = spp_plot, 
-  #        width = 8, height = 4, dpi = 200)
+  ggsave(filename = file.path(dir_fig, sprintf('rangemap_%s_%s.png', spp_quad,
+                                               str_replace(tolower(spp), ' ', '_'))),
+         plot = spp_plot, 
+         width = 8, height = 4, dpi = 200)
 }

### 2015-12-02 data_explore_2015_data.Rmd
script uses new data

@@
 #iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both_2015.csv'), 
+iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                            col_types = 'cdddd', progress = TRUE) %>%
@@
 
-if(!file.exists(data_file)) {
+if (!file.exists(data_file)) {
   ### species info list - scinames, IUCN species ID, AM species ID, category etc
-  spp_list_base <- read.csv('shiny/data/spp_am_v_iucn_2015.csv', stringsAsFactors = FALSE)
+  spp_list_base <- read.csv('data/spp_am_v_iucn_2015_inclDD.csv', stringsAsFactors = FALSE)
 
@@
 
-spp_list <- read.csv(data_file, stringsAsFactors = FALSE)
+spp_list1 <- read.csv(data_file, stringsAsFactors = FALSE)
+```
+
+Note: At this point the species list (`spp_list1`) still contains data deficient species.
+274 species show zero area for IUCN; most of these are DD.
+The next chunk filters for `!is.na(sm_perc)` and `popn_category != 'DD'`
+
+``` {r eliminate DD and NA areas, echo = FALSE, warning = FALSE}
+spp_list <- spp_list1 %>% 
+  filter(!is.na(sm_perc))
+
+DT::datatable(spp_list %>%
+                 select(sciname, am_sid, iucn_sid, am_review = reviewed, 
+                        cat = popn_category, trend = popn_trend, spp_group, 
+                        A_tot = total_area, A_am = am_area, A_iucn = iucn_area, 
+                        sm_perc, sm_range, area_ratio, lg_area_pct))
+
 ```
@@
 ``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-
-spp_list <- spp_list %>%
-  mutate(area_ratio = am_area / iucn_area,
-         area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
-         area_ratio = area_ratio * 100)
 

### 2015-12-02 R/data_explore_fxns.R

@@ -29,7 +29,7 @@ create_spp_list <- function(spp_list_base, data_file) {
   
   
   #there's likely a better way to do this than a for loop (lapply or sapply?)
-  for (i in 1:nrow(spp_list)){
+  for (i in 1:nrow(spp_list)){ # i = 1
     spp     <- spp_list[i, ]$sciname #species name
     spp_map <- get_spp_map(spp) #get species map
     
@@ -53,21 +53,21 @@ create_spp_list <- function(spp_list_base, data_file) {
                      which = 'am_sid', 
                      subsWithNA = TRUE)
     
-    spp_list[i, ]$am_area <- area(r_am_spp, na.rm = TRUE)%>%
+    spp_list[i, ]$am_area <- area(r_am_spp, na.rm = TRUE) %>%
       cellStats(., stat='sum')
     
     
     ###  get the iucn cells
     iucn_map <- spp_map %>%
-      mutate(iucn_pres = iucn_area > 0)%>%
-      dplyr::select(loiczid, iucn_area, am_sid, am_prob, iucn_pres)%>%
+      mutate(iucn_pres = iucn_area > 0) %>%
+      dplyr::select(loiczid, iucn_area, am_sid, am_prob, iucn_pres) %>%
       unique()
     
     #TODO: incorporate the iucn_area (area of cell that species overlaps with...I think thats what it is)
     
     ###this is incorporated because there are some duplications for loiczids with IUCN cells. I'm guessing this is because of the parent/subpopulations?
     
-    if(TRUE %in% duplicated(iucn_map$loiczid))next()
+    if(TRUE %in% duplicated(iucn_map$loiczid)) next()
     
     ### rasterize IUCN map
     r_iucn_spp <- subs(loiczid_raster, 
@@ -76,14 +76,14 @@ create_spp_list <- function(spp_list_base, data_file) {
                        which = 'iucn_pres', 
                        subsWithNA = TRUE)
     
-    spp_list[i, ]$iucn_area <- area(r_iucn_spp, na.rm = TRUE)%>%
-      cellStats(., stat='sum')
+    spp_list[i, ]$iucn_area <- area(r_iucn_spp, na.rm = TRUE) %>%
+      cellStats(., stat = 'sum')
     
     
     ### Get total area (IUCN and AM combined)
-    allCells <- spp_map%>%
-      dplyr::select(loiczid)%>%
-      mutate(value=1)%>%
+    allCells <- spp_map %>%
+      dplyr::select(loiczid) %>%
[…]

### 2015-12-03 data_explore_2015_data.Rmd
figs using 2015 am data

@@
 dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-dir_fig <- file.path(getwd(), 'figures')
+dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
+dir_fig <- file.path(getwd(), 'figs2015')
 
@@
   scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
-  labs(title = 'Percent of smaller range within larger range',
+  labs(title = 'Percent of smaller range within larger range 2015 data',
        x = 'Percent of smaller range that falls within the larger range', 
@@
                                  'EN' = 'orange', 
-                                 'CR' = 'orangered2')) +
+                                 'CR' = 'orangered2',
+                                 'DD' = 'grey50')) +
   geom_vline(x = area_ratio_quartiles[c(1, 3)],
@@
        y = 'Relative number of species per quadrant', 
-       title = 'Species group breakdown by quadrant')
+       title = 'Species group breakdown by quadrant 2015 data')
 
@@
 
-levelplot(am_sp_map, par.settings = myTheme, 
-          main = 'Aquamaps Dataset', 
+am_lvlplot <- levelplot(am_sp_map, par.settings = myTheme, 
+          main = 'Aquamaps Dataset 2015', 
           colorkey = list( at = seq(0, 1063, length = 15)))
 
-levelplot(iucn_sp_map, par.settings = myTheme, 
+png(file.path(dir_fig, 'levelplot_am.png'))
+print(am_lvlplot)
+dev.off()
+
+iucn_lvlplot <- levelplot(iucn_sp_map, par.settings = myTheme, 
           main = 'IUCN Dataset', 
           colorkey = list(at = seq(0, 1063, length = 15)))
+
+png(file.path(dir_fig, 'levelplot_iucn.png'))
+print(iucn_lvlplot)
+dev.off()
 
@@
 
-am_spp_cells_big <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/ohi_hcaf_species_native.csv')) %>%
+am_spp_cells_big <- fread(file.path(dir_data, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv')) %>%
   as.data.frame()
-### am_spp_cell_big is missing loiczid info so bring in hcaf to match
-
-hcaf <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/hcaf.csv')) %>% #bring this in just to get loiczid
-  as.data.frame() %>%
-  select(CsquareCode, LOICZID) %>%
-  rename(loiczid = LOICZID)
 
@@
 am_all_n_tot <- am_spp_cells_big %>% 
-  select(SpeciesID) %>% 
-  unique() %>% nrow()
+  select(speciesid) %>% 
+  unique() %>% 
+  nrow()
 
 all_am <- am_spp_cells_big %>%
-  select(SpeciesID, CsquareCode) %>%
-  group_by(CsquareCode) %>%
+  select(speciesid, loiczid) %>%
+  group_by(loiczid) %>%
   summarize(all_count = n(),
@@
 
-all <- all_am %>%
-  left_join(hcaf, by = 'CsquareCode')
-
-all_am_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_count')], 
+all_am_map <- subs(loiczid_raster, all_am[ , c('loiczid', 'all_count')], 
                    by = 'loiczid', 
@@
 
-levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
+am_lvlplot_all <- levelplot(all_am_map, par.settings = myTheme, 
+                            main = 'Entire Aquamaps Dataset 2015')
+
+png(file.path(dir_fig, 'levelplot_am_all.png'))
+print(am_lvlplot_all)
+dev.off()
+
 ```
 
-![image](figures/all_aquamaps_map.png)
+![image](figs2015/all_aquamaps_map.png)
 
@@
 
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
+![Absolute change in SPP score](figs2015/SPP_scenarios/boxplot_abs_diff.png)
 
-![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)
+![Relative change in SPP score](figs2015/SPP_scenarios/boxplot_rel_diff.png)

### 2015-12-08 data_explore_2015_data.Rmd
Working on manuscript

@@
 
-library(rasterVis)
+# library(rasterVis)
 library(ggplot2)
@@
 ``` {r child = 'R/data_explore_pt1.Rmd', eval = FALSE}
+```
 # bar chart: number of spp per group
@@
 # rangemap: Chromis alpha
-```
 
@@
 
-```{r global_dist_map, echo = FALSE, warning = FALSE, message = FALSE}
+```{r global_dist_map, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 

### 2015-12-08 ms_outline.Rmd
Working on manuscript

@@
     * assessing biodiversity 
-    * monitor impacts (human, natural) over time (should this be in here?)
+    * monitor impacts (human, natural)
+    * global-scale assessments
+    * poor overlap of species lists between the two data sets - sometimes helpful to use both
 * Challenges
@@
 
-##Introduction
+#Introduction
 
@@
 
-##Section One: Understanding each dataset separately
+#Section One: Understanding each dataset separately
 
@@
 
-The IUCN conducts global assessments for species [detail assessment process - one paragraph].
+The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups.
+Species spatial distributions are determined by experts based upon <known occurrence, suitable habitat, and blah>.
+The IUCN presents these maps as 'limits of distibution' with the caveat that the species 'probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon'.<cite?>  IUCN only publishes spatial distribution map sets for taxonomic groups in which at least 90% of the species within the taxon group have been evaluated.
+While this helps prevent sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
+As of December 2015, IUCN species distribution maps were available for XXX marine species (habitat = "marine" on IUCN species information pages).
 
+*something about format - polygons, fields, etc? presence/no presence*
+
+*presence field, and seasonality info*
+
+*terrestrial and marine both*
+
+https://www.conservationtraining.org/course/view.php?id=217&lang=en
+
+<!--
 Spatial designation of the species distribution is left to the experts with strict guidance for how to assign the designation.
-Using known occurrence data, experts depict the species distribution by polygons using a systematic methodology laid out by the [IUCN](http://www.amphibians.org/wp-content/uploads/2013/09/Red-List-Mapping-standards-ARLA-Jan2014-web-version.pdf).
-These maps are meant to act as 'limits of distibution' and should be interpreted as the species 'probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon'.
+Using known occurrence data, experts depict the species distribution by polygons using a systematic methodology laid out by the [IUCN](http://www.amphibians.org/wp-content/uploads/2013/09/Red-List-Mapping-standards-ARLA-Jan2014-web-version.pdf) (this needs a better source).
 
@@
 Groups that have been comprehensively assessed are those containing >90% species evaluated within the described taxon group."*
+    -->
 
@@
 AquaMaps (Kaschner et al.
-2013) was developed to create species range maps based on a computer model ***Bioclimatic envelope model*** that takes up to six parameters to determine where marine species are located.
-The model uses known species preferences for sea surface temperature, salinity, depth, productivity, and in some instances distance to land and sea ice cover to predict species ranges.
-Range maps are based on a 0.5 degree global grid where value in each cell is equal to the ***probability of occurrence*** (or ***relative environmental suitability?*** (see Kaschner et al 2006 and 2011)) of that species within that cell These maps can then be tweaked by experts altering the parameters and regenerating the modeled map.
+2013) offers species range maps based on modeled relative environmental suitability.
+For each assessed species, data from observations and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, and where applicable, sea ice cover and/or distance to land.
+The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of ***probability of occurrence*** (or ***relative environmental suitability?*** (see Kaschner et al 2006 and 2011)), with cell values ranging from 0 (unsuitable) to 1.00 (highly suitable).
+These maps can then be tweaked by experts altering the parameters and regenerating the modeled map.
 As of December 2015, ***22,000*** species have been mapped.
-See (Kaschner et al.
+***Expert mode to further refine models?*** See (Kaschner et al.
 2013) for more detailed methods.
 
-AquaMaps uses both occurrence data and environmental parameters to predict relative species occurrence.
-This approach attempts to reduce sampling bias and fill gaps where data is not available.
-Six environmental predictors are used to estimate where a species could be found; sea surface temperature, bottom depth, salinity, primary production, sea ice concentration and distance to land.
-For more information on methods see Kaschner et al.
-(2007) and online at http://www.aquamaps.org/main/AboutAquaMaps.php.
-
-- from Kaschner 2011: The RES model generates an index of species-specific *relative environmental suitability (RES)* of each individual half degree grid cell by scoring how well its physical attributes matched the known aspects of species’ habitat use.
-RES values range between 0 (not suitable) to 1 (highly suitable) and represent the product of the suitability scores assigned for the individual environmental attributes (bottom depth, SST, sea ice concentration, and distance from land in some cases), which were calculated using pre-defined trapezoidal functional response curves
-- but from Aquamaps website: Relative likelihood values range between 0.00 - 1.00 and represent the product of the relative likelihood scores assigned for the individual environmental attributes, which are in turn calculated based on the environmental envelopes described above.
-***so they just changed the phrase; I like the original better...***
-
-There are inherent issues with AquaMaps. It can't be expected to accurately predict all marine species ranges. Expert insight is sought and incorporated when available. 
 - ***how is expert insight used?
@@
 
-#### How map is developed (maybe in above paragraphs?)
-
 - Problem of expert-defined range maps (over-predicting) vs sightings (under-predicting): 
 from **Kaschner et al 2011**:  *Similar to the trade-offs of different habitat prediction modeling approaches [11], these two methods lie on ___opposite ends of a spectrum from potentially overpredicting expert-derived (range maps) to potentially underpredicting (empirical sighting surveys) range sizes___*
+- are there other sources (non-IUCN and AM sources) that can back this up?
 
-The main differences between these two sources of marine species ranges is how the maps are developed, and their output.
-When a species range is developed during an IUCN assessment, experts determine the boundaries of the species range based on their expertise which could be using known occurrences, habitat preferences, suitable habitat, or other knowledge.
-The range is depicted as a polygon where the boundaries are *"meant to communicate that the species probably occurs within this polygon, but it does not mean that it is distributed equally within the polygon or occurs everywhere within the polygon."*
+### Spatial resolution - not sure we need this, it is not really *resolution* but more about nuances of distribution... difference already explained above?
 
-#### Spatial resolution
 These two differ in spatial resolution - AquaMaps provides higher resolution information about where species do and do not occur globally whereas the IUCN range maps are unable to provide specific information about where within a polygon a species occurs.
-This is understandable given the mobile nature of many marine species and the difficulty in observing and tracking known occurrences.
+
+### Poor overlap between assessed species
+
+***
+
+# Section Two: In what ways do the two data sets differ
 
@@
 
+<!--
 * Here we can discuss the idea of multiple "dimensions" of alignment - distributional and area alignment.  Well-aligned species will show high alignment in both dimensions.  
@@
 * Note that alignment does not necessarily mean accuracy or precision - a poorly-understood species could appear to be well-aligned.  
+-->
 
-***
+AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to differences, we would expect to see strong similarities between range maps for any given species.
 
-##Section Two: Comparison of IUCN and AquaMaps ranges
+To facilitate our analysis, we selected the subset of marine species that have range maps in both IUCN and AquaMaps (n = ???), matched by Genus and species names, and then rasterized the IUCN shapefiles to match the Aquamaps half-cell raster.
+We flattened each map by assigning to each cell a simple "presence/absence" value:
 
+* For IUCN data, any cell with non-zero overlap with a range map polygon was considered "presence."  
+* For AquaMaps data, any cell with non-zero probability (or RES) was considered "presence."  
+* In both cases, these criteria for presence will maximize the number of cells included.
+Using these criteria, we found that for 72% ???
+of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
+
+Overlaying distribution maps from the two data sets for a given species, we calculated two dimensions of spatial alignment: distribution match, which we defined as the ratio of the smaller range that overlapped with the larger range; and area match, which we defined as the ratio of the smaller range to the larger range.
+
+$$Match_{dist} = \frac{A_{overlap}}{A_{larger} * 100%$$
+
+$$Match_{area} = \frac{A_{smaller}}{A_{larger} * 100%$$
+
+For a species with well-matched distribution maps between the two data sets, we would expect to see a value near 100% for each dimension, indicating near-total inclusion of the smaller range within the larger (distribution alignment) and very similar extent of range (area alignment).
+
+<!--
 ### Determining "presence" for each data set
@@
     - since IUCN "presence" seems to be pretty loose, set a wide tolerance on AM?
+-->
 
-### Global distribution between each data set
+### Differences in global distribution between data sets
 
@@
 ![](figures/sp_groups_barchart.png)
-***Maybe also include the species distribution maps here - maybe a map with the difference in coverage between IUCN and Aquamaps? where is IUCN under-represented? and in what species groups?***
+
+***Total species occurrences for each data set: all AM, all IUCN, and overlap (one? the other? occurrences of both?)***
 
@@
 
-<!-- Not sure this is super helpful any more - this info comes out in the quadrant graph
-For each species map, the total area was calculated for all cells where the species was determined as 'possibly present' by both datasets.
-Due to large areas and differences, the total area per species was log transformed and then compared to the range map of the same species between datasets.
-There are differences, mainly driven by differeneces in the wrasses and combtooth blennies, but this could be due to there being more species in both of those species groups.
-
-![](figures/CompareTotalAreaLOG_AM-IUCN_scatter.png)
-![](figures/CompareTotalAreaLOG_AM-IUCN_spgroups_scatter.png)
--->
-
-***This shows differences area maybe?*** What's the story of this plot...
-![](figures/CompareTotalAreaLOG_AM-IUCN_scatter.png)
+***Relevance of this*** - unequal distribution of assessed species - strengths and weaknesses of each; benefits of using both maybe?
 
@@
 
-<!-- These are also essentially included within the quadrant plot?
-![](./figures/Overlap-PercTotalRange.png)
-![](./figures/Overlap-PercTotalRange_spGroups.png)
-![](./figures/PercSmallRangeInLarger.png)
-![](./figures/AreavsPercOverlap_2.png)
--->
-
 #### Comparing alignment by species group
@@
 
-##Section Three: Application to OHI
+##Section Three:
+
+# Section Four: Implications
+
+## Application to OHI
 

### 2015-12-09 ms_outline.Rmd
adding Jamie's intro and refs

@@
 
+Mapping species abundances and distributions is a critical component to understanding global biodiversity and detecting patterns and trends through time.
+There have been multiple efforts to map terrestrial, freshwater and marine species from local to global scales.
+These pieces of information inform our understanding of taxonomic assemblages, biogeographical limits to species ranges, suitable habitat for species.
+
+These two spatial datasets have been used to analyze species status globally (Halpern et al.
+2012, Selig et al.
+2013), global biodiversity patterns and trends, predicting range shifts (Molinos et al.
+2015),and setting conservation priorities (Klein et al.
+2015).
+Yet there have been no comprehensive comparisons of these two datasets to identify the differences and provide insight into how using one or the other can influence the outcome of an analysis.
+With impactful conclusions being drawn from these large-scale studies on global biodiversity and conservation, it's imperative that the data used in these analyses are appropriate and the data decisions are informed.
+
+The IUCN Red List of Threatened Species has assessed **X** number of marine species, but only (**Y**) have associated spatial information.
+AquaMaps has developed range maps for over 22,800 species.
+There are only 2,100 species that have maps in both datasets.
+This gap in overlapping species between these two datasets further highlights the need for understanding each of these datasets.
+Both IUCN and AquaMaps have gaps in marine species information and a significant portion of available data may be overlooked if limited to just one of these datasets.
+
+Species range maps provided by AquaMaps (Kaschner et al.
+2013) are model-based, relying on environmental parameters in addition to known occurrences and expert input.
+AquaMaps use information about individual species environmental tolerances for salinity, temperature, depth, productivity, and where applicable, sea ice and distance from land, to create a presence-only gridded range map at a resolution of 0.5 degree cells with cell values ranging from 0 to 1.00.
+These values represent the *relative environmental suitability* for the species, with 0 representing unsuitable habitat and 1.00 being the most suitable habitat.
+Once produced, each range map maintains a dynamic element that allows individuals to adjust parameters based on their expert knowledge to improve the model-generated map.
+
+The maps provided by the IUCN for each assessed species are created by one or a few experts on that species.
+There are general guidelines to how the map should be created.
+Specifically, the final mapped polygons are meant to indicate that a species *probably only occurs within the polygon* but that the species *is not* evenly distributed everywhere within the polygon.
+The guidelines vary slightly for each type of species (terrestrial, freshwater, marine).
+These maps are dependent on records of observations from databases such as GBIF.
+
+IUCN only publishes spatial distribution map sets for taxonomic groups in which at least 90% of the species within the taxon group have been evaluated.
+While this helps prevent sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
+As of December 2015, IUCN species distribution maps were available for XXX marine species (habitat = "marine" on IUCN species information pages).
+
+<!--
 There have been multiple efforts to map marine species based on both observational data and known life history and habitat preferences. AquaMaps and IUCN are two of these efforts.
+-->
 
@@
 PDF
+
+#References
+
+García Molinos, Jorge, Benjamin S.
+Halpern, David S.
+Schoeman, Christopher J.
+Brown, Wolfgang Kiessling, Pippa J.
+Moore, John M.
+Pandolfi, Elvira S.
+Poloczanska, Anthony J.
+Richardson, and Michael T.
+Burrows.
+“Climate Velocity and the Future Global Redistribution of Marine Biodiversity.” Nature Climate Change advance online publication (August 31, 2015).
+doi:10.1038/nclimate2769.
+
+Halpern, Benjamin S., Catherine Longo, Darren Hardy, Karen L.
+McLeod, Jameal F.
+Samhouri, Steven K.
+Katona, Kristin Kleisner, et al.
+“An Index to Assess the Health and Benefits of the Global Ocean.” Nature 488, no.
+7413 (August 30, 2012): 615–20.
+doi:10.1038/nature11397.
+
+Halpin, P.N., A.J.
+Read, E.
+Fujioka, B.D.
+Best, B.
+Donnelly, L.J.
+Hazen, C.
+Kot, K.
+Urian, E.
+LaBrecque, A.
+Dimatteo, J.
+Cleary, C.
+Good, L.B.
+Crowder, and K.D.
+Hyrenbach.
+2009.
+OBIS-SEAMAP: The world data center for marine mammal, sea bird, and sea turtle distributions.
+Oceanography 22(2):104-115
+
+Klein, Carissa J., Christopher J.
+Brown, Benjamin S.
+Halpern, Daniel B.
+Segan, Jennifer McGowan, Maria Beger, and James E.M.
+Watson.
+“Shortfalls in the Global Protected Area Network at Representing Marine Biodiversity.” Scientific Reports 5 (December 3, 2015): 17539.
+doi:10.1038/srep17539.
+
+Selig, Elizabeth R., Catherine Longo, Benjamin S.
+Halpern, Benjamin D.
+Best, Darren Hardy, Cristiane T.
+Elfes, Courtney Scarborough, Kristin M.
+Kleisner, and Steven K.
+Katona.
+“Assessing Global Marine Biodiversity Status within a Coupled Socio-Ecological Perspective.” PLoS ONE 8, no.
+4 (April 11, 2013): e60284.
+doi:10.1371/journal.pone.0060284.

### 2015-12-09 ms_outline.Rmd
tweaking the intro with some IUCN numbers

@@
 
-The IUCN Red List of Threatened Species has assessed **X** number of marine species, but only (**Y**) have associated spatial information.
+The IUCN Red List of Threatened Species includes more than 19,000 assessed marine species, 
+<!--define what "assessed" means?
+include DD?
+keeping those in for now--> but only **4463** <!--11975 maps in those shapefiles, but fewer than 4500 correspond to assessed marine species...
+and only 2500 of those overall correpond to AquaMaps species--> of these have associated spatial information.
 AquaMaps has developed range maps for over 22,800 species.
-There are only 2,100 species that have maps in both datasets.
+Only 2,100 species have been mapped in both datasets.
 This gap in overlapping species between these two datasets further highlights the need for understanding each of these datasets.
@@
 
-The maps provided by the IUCN for each assessed species are created by one or a few experts on that species.
+The maps provided by the IUCN for each assessed species are created by one or more experts on that species.
 There are general guidelines to how the map should be created.
-Specifically, the final mapped polygons are meant to indicate that a species *probably only occurs within the polygon* but that the species *is not* evenly distributed everywhere within the polygon.
+Specifically, the final mapped polygons are meant to indicate that a species *probably only occurs within the polygon* but are *not* meant to indicate that the species is evenly distributed everywhere within the polygon.
 The guidelines vary slightly for each type of species (terrestrial, freshwater, marine).
-These maps are dependent on records of observations from databases such as GBIF.
+These maps are typically dependent on observation records from databases such as GBIF.
 

### 2015-12-09 ms_outline.Rmd
body of the text

--- 2015-12-08 ms_outline.Rmd
@@
     toc: true
+    number_sections: true
     theme: "spacelab"
@@
 ##Working Abstract  
-
+<!--
 * Importance of understanding spatial distribution of species
@@
         * Aquamaps validation and benefits from _Ready et al. 2010_:
-          -  We present and evaluate AquaMaps, a presence-only species distribution modelling system that allows the incorporation of expert knowledge about habitat usage and was designed for maximum output of standardized species range maps at the global scale.
+          *  We present and evaluate AquaMaps, a presence-only species distribution modelling system that allows the incorporation of expert knowledge about habitat usage and was designed for maximum output of standardized species range maps at the global scale.
 In the marine environment there is a significant challenge to the production of range maps due to large biases in the amount and location of occurrence data for most species.
@@
 _The effect of the inclusion of expert knowledge to AquaMaps is also investigated._
-          - Model outputs were tested internally, through data partitioning, and externally against independent survey data to determine the ability of models to predict presence versus absence.
+          * Model outputs were tested internally, through data partitioning, and externally against independent survey data to determine the ability of models to predict presence versus absence.
 Models were also tested externally by assessing correlation with independent survey estimates of relative species abundance.
@@
 The transparency, speed and adaptability of the AquaMaps system, as well as the existing online framework which allows expert review to compensate for sampling biases and thus improve model predictions are proposed as additional benefits for public and research use alike.
-
-* Implications
-    * some species need more attention?
-    * which data sets are more useful for which uses?
-        * who uses IUCN? who uses AquaMaps?  Who uses both?
-    * how can data decisions affect results of research or conservation efforts?
-        * using OHI as an example
+-->
 
@@
 Two main sources for marine species are the IUCN and AquaMaps.
-These two datasets differ in methodology and final products, yet the differences between the two are not well documented *(though Aquamaps, and probably IUCN, each probably have some validation studies)*.
+These two datasets differ in methodology and final products, yet the differences between the two are not well documented _(though Aquamaps, and probably IUCN, each probably have some validation studies)_.
 AquaMaps uses a computer model based on preferential environmental parameters to predict species locations and probability of occurrence, while the IUCN uses expert insight to assign spatial boundaries around known locations of marine species.
@@
 
-***
+-----
 
@@
 
-#Section One: Understanding each dataset separately
-
-![](./figures/rangemap_chromis_alpha.png)
-
-**Figure 1:** Species range map for *Chromis alpha* from the AquaMaps (orange) and IUCN (purple) datasets.
-***Do we want to have this way up here?  needs explanation?***
+# Understanding each dataset separately
 
@@
 
-*something about format - polygons, fields, etc? presence/no presence*
-
-*presence field, and seasonality info*
-
-*terrestrial and marine both*
-
-https://www.conservationtraining.org/course/view.php?id=217&lang=en
+* _something about format - polygons, fields, etc? presence/no presence_
+* _presence field, and seasonality info_
+* _terrestrial and marine both_
+* _use as a reference:_ https://www.conservationtraining.org/course/view.php?id=217&lang=en
 
@@
 
-**Data Type**  
+__Data Type__  
 Shapefiles for individual species. Spatial files include subpopulation information, presence (extinct/extant etc.) and seasonality information (breeding/migration/origin etc.).  
-- ***Include other columns typically within each shapefile***
+* ___Include other columns typically within each shapefile___
 
-**Methods**  
+__Methods__  
 From [IUCN site](http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources):  
-    - The species distribution maps, sometimes referred to as 'limits of distribution' or 'field guide' maps, aim to provide the current known distribution of the species within its native range.
+    * The species distribution maps, sometimes referred to as 'limits of distribution' or 'field guide' maps, aim to provide the current known distribution of the species within its native range.
 The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range.
 In most cases the range is depicted as polygons.
-*A polygon displaying the limits of a species distribution is essentially meant to communicate that the species probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon*.
+_A polygon displaying the limits of a species distribution is essentially meant to communicate that the species probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon_.
 
-- ***Spatial files for IUCN species groups attempt to include a broad sample across the group, to avoid sampling bias (maybe?)*** - but only some species groups have had spatial data released, so some taxa are not represented at all
-    - from http://www.iucnredlist.org/technical-documents/spatial-data: *"This spatial data provided below is for comprehensively assessed taxonomic groups."  "...
+* ___Spatial files for IUCN species groups attempt to include a broad sample across the group, to avoid sampling bias (maybe?)___ - but only some species groups have had spatial data released, so some taxa are not represented at all
+    * from http://www.iucnredlist.org/technical-documents/spatial-data: _"This spatial data provided below is for comprehensively assessed taxonomic groups."  "...
 the data is divided by taxonomic groups.
-Groups that have been comprehensively assessed are those containing >90% species evaluated within the described taxon group."*
+Groups that have been comprehensively assessed are those containing >90% species evaluated within the described taxon group."_
     -->
 
-###AquaMaps
+### AquaMaps
 
@@
 For each assessed species, data from observations and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, and where applicable, sea ice cover and/or distance to land.
-The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of ***probability of occurrence*** (or ***relative environmental suitability?*** (see Kaschner et al 2006 and 2011)), with cell values ranging from 0 (unsuitable) to 1.00 (highly suitable).
+The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of ___probability of occurrence___ (or ___relative environmental suitability?___ (see Kaschner et al 2006 and 2011)), with cell values ranging from 0 (unsuitable) to 1.00 (highly suitable).
 These maps can then be tweaked by experts altering the parameters and regenerating the modeled map.
-As of December 2015, ***22,000*** species have been mapped.
-***Expert mode to further refine models?*** See (Kaschner et al.
+As of December 2015, ___22,000___ species have been mapped.
+___Expert mode to further refine models?___ See (Kaschner et al.
 2013) for more detailed methods.
 
-- ***how is expert insight used?
+* ___how is expert insight used?
 how frequently?
-new data set includes "reviewed" field*** expert insight seems to be used to generate the environmental suitability envelope for species, to provide sightings info that use used in training the model, and to exclude species from regions of known absence; but the model generates the maps.
+new data set includes "reviewed" field___ expert insight seems to be used to generate the environmental suitability envelope for species, to provide sightings info that use used in training the model, and to exclude species from regions of known absence; but the model generates the maps.
 However, some species maps are reviewed by experts (approved?
@@
 
-- Problem of expert-defined range maps (over-predicting) vs sightings (under-predicting): 
-from **Kaschner et al 2011**:  *Similar to the trade-offs of different habitat prediction modeling approaches [11], these two methods lie on ___opposite ends of a spectrum from potentially overpredicting expert-derived (range maps) to potentially underpredicting (empirical sighting surveys) range sizes___*
-- are there other sources (non-IUCN and AM sources) that can back this up?
+* Differences in intent: capture all observations and be inclusive as possible, vs model suitable habitats
+* Differences in methodology - expert-defined range maps (over-predicting) vs sightings (under-predicting) vs models: 
+from __Kaschner et al 2011__:  _Similar to the trade-offs of different habitat prediction modeling approaches [11], these two methods lie on ___opposite ends of a spectrum from potentially overpredicting expert-derived (range maps) to potentially underpredicting (empirical sighting surveys) range sizes____
+    * are there other sources (non-IUCN and AM sources) that can back this up?
+* Differences in the interpretation: probability/RES, vs. presence "probably within this polygon, but not everywhere within" - apples to oranges.
 
-### Spatial resolution - not sure we need this, it is not really *resolution* but more about nuances of distribution... difference already explained above?
+<!--
+### Spatial resolution - not sure we need this, it is not really _resolution_ but more about nuances of distribution... difference already explained above?
 
 These two differ in spatial resolution - AquaMaps provides higher resolution information about where species do and do not occur globally whereas the IUCN range maps are unable to provide specific information about where within a polygon a species occurs.
+-->
 
-### Poor overlap between assessed species
+#### Poor overlap between assessed species
 
-***
+___number of species in each data set, and the number that overlap___
 
-# Section Two: In what ways do the two data sets differ
+* this highlights that each provides valuable and unduplicated spatial range information.
 
-### Defining alignment between the two data sets
+#### Summarize this section?
+
+Researchers have two large and useful data sets at their disposal, each with its own utility.  Being able to use both simultaneously makes many more species available, but here there be challenges.
+
+-----
+
+# In what ways do the two data sets differ
+
+<!--
+![](./figures/rangemap_chromis_alpha.png)
+
+__Figure 1:__ Species range map for _Chromis alpha_ from the AquaMaps (orange) and IUCN (purple) datasets.
+___is this the right map to use here? find a good one with cool differences___
+-->
+
+### Defining and comparing spatial alignment between the two data sets
+
+___Main messages:___
+
+* Definition of "alignment" in two dimensions, incl "good" vs "poor"
+* basic methodology (raster, overlay, calculate)? or should there be a separate methodology section?
 
@@
 
-$$Match_{dist} = \frac{A_{overlap}}{A_{larger} * 100%$$
+$$Match_{dist} = \frac{A_{overlap}}{A_{larger}} * 100\%$$
 
-$$Match_{area} = \frac{A_{smaller}}{A_{larger} * 100%$$
+$$Match_{area} = \frac{A_{smaller}}{A_{larger}} * 100\%$$
 
-For a species with well-matched distribution maps between the two data sets, we would expect to see a value near 100% for each dimension, indicating near-total inclusion of the smaller range within the larger (distribution alignment) and very similar extent of range (area alignment).
+For a species whose distribution is well understood, and whose range is described in both data sets, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (distribution alignment) and very similar extent of range (area alignment).
 
@@
 
-To compare these two datasets for all species found in each (n = ***1920***?), we rasterized the IUCN polygons to the same cell resolution as AquaMaps: 0.5 degrees in both latitude and longitude.
-For every grid cell where the polygon overlapped ***the center point?
+To compare these two datasets for all species found in each (n = ___1920___?), we rasterized the IUCN polygons to the same cell resolution as AquaMaps: 0.5 degrees in both latitude and longitude.
+For every grid cell where the polygon overlapped ___the center point?
 or is it based on ANY overlap, e.g.
 using weights and including any non-zero weight?
-This could boost IUCN presence***, a value of 1 was assigned.
+This could boost IUCN presence___, a value of 1 was assigned.
 For the purposes of comparison, the probability of occurrence values in the AquaMaps data were ignored and all cells where there was a value greater than 0 were assigned a 1.
 This way we have comparable maps for total species range.
-- ***How do we justify a zero-threshold here?*** 
-    - need to understand the "probability" - is it really ***relative environmental suitability***   From Kaschner et al 2011: "Patterns based on a presence threshold RES.0.6 (Fig.
+- ___How do we justify a zero-threshold here?___ 
+    - need to understand the "probability" - is it really ___relative environmental suitability___   From Kaschner et al 2011: "Patterns based on a presence threshold RES.0.6 (Fig.
 1) were most strongly supported by empirical species richness data (see ‘Validation of species richness’, below)."
@@
 
+___Main messages:___
+
+* breaking down alignment differences by quadrant provides useful information.
+    * q1 = well-matched (well understood? not necessarily, could be luck or over-generous environmental envelopes)
+    * other quadrants have defining characteristics that can highlight areas where we should look for better data and better understanding
+
+<!--
+Figure 1 shows four representative species: Kajikia Albida, with a high degree of alignment in both area and distribution; Galaxea cryptoramosa, well-matched in spatial alignment but poorly matched in areal alignment; Conus magnificus, well-matched in area but poorly matched in distribution; and Acanthurus nigrosis, poorly matched in both area and distribution.
+
+![](figures/four_maps.png)
+
+___Figure 1: Representative species with distribution mapped in both AquaMaps and IUCN data sets.___  
+-->
+
+For each of the ___2200___ species whose range is described by both data sets, we calculated the distribution match and the area ratio, and plotted the ___2200___ species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
+
+![](figures/AreaRatiovsPercOverlap.png)
+
+___Figure 1: Distribution alignment vs area alignment for ___2200___ species mapped in both Aquamaps and IUCN species distribution maps.___
+
+We divided the plot into quadrants based on the median values for each axis to differentiate between four different qualities of alignment:
+
+| Description              |     Distribution match    |    Area match     |
+| :----------------------: | :-----------------------: | :---------------: |
+| q1: Well-aligned         | Distribution match > XX % | Area match > YY % |
+| q2: Distribution-aligned | Distribution match < XX % | Area match > YY % |
+| q3: Area-aligned         | Distribution match > XX % | Area match < YY % |
+| q4: Poorly-aligned       | Distribution match < XX % | Area match < YY % |
+
+Clearly the upper right quadrant (quadrant 1) is the place to be.
+Species in this quadrant generally agree in spatial distribution between the two distribution maps, and the area of described ranges will be similar; expert-drawn range map and bioclimatic envelope model produce similar predictions of species distribution.
+(n = XXX; YYY %)
+
+The other quadrants each suggest a different story.
+Species map pairs that fall within the upper left quadrant (quadrant 2: well-matched in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
+(n = XXX; YYY %)
+
+The lower right quadrant (quadrant 3: well-matched in area but poorly matched in distribution) includes species for which the two maps generally agree in range area, but disagree in where those ranges occur.
+General disagreement in distribution seems to indicate a greater issue than general disagreement in area extent, so we categorized map pairs in this quadrant as being more problematic than map pairs in either of the upper quadrants.
+(n = XXX; YYY %)
+
+The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. (n = XXX; YYY %)
+
+<!--
+(Looking at some representative species maps)
+The range described by AquaMaps map covers XXX km^2^; the range described by IUCN map covers YYY km^2^; the overlap between the two maps covers ZZZ km^2^.
+Distribution match = ZZZ km^2^ / XXX km^2^; area match = XXX km^2^ / YYY km^2^.
+
+* Single species map: ___Do we want this as a separate figure?___ _it helps to explain what the hell we're talking about; but we could also just jump to the quadrant map?_ 
+* ___OR: plot four representative maps FIRST, to show differences in the two dimensions of alignment, THEN show the quadrant map with those species indicated.___ 
+    * _This is backwards from how we did the analysis, but I think might flow better for the reader._
+-->
+
+![](figures/barchart_spp_gp_quads.png)
+
+Figure 2 breaks the ___2200___ species map pairs into taxonomic groups as classified by the published IUCN range maps.
+We were not surprised to find that map pairs for marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2); many of these species are global-ranging pelagic organisms, with some ranges encompassing nearly the entire ocean.
+Mangroves are also not a surprise, being immobile and easily mapped by remote sensing methods.
+
+_Map pairs for corals, most of which fall into quadrant 2, ***hypothesis...*** generally agree well in distribution, but in many cases IUCN ranges include buffers that extend much farther from coastlines than AquaMaps modeled ranges, which use seafloor depth as a limiting factor._
+
+___Other hypotheses/case studies to explain mechanistic differences?___
+
+-----
+
 ### Differences in global distribution between data sets
+
+___main messages:___
+
+* IUCN seems to be weighted toward tropics; AM might be more evenly distributed?
+* This difference in distribution might come into play when selecting a data set for a more regionally-focused study
+* ... or for global studies, it might be desirable to use both data sets to get better global distribution?
+* Small number of overlapping species indicates that each offers data on distinct spp
+* This section is really supporting the idea that it is critical to understanding the differences, as both data sets support one another
 
@@
 This is driven more by the IUCN dataset since the number of species with spatial data is directly attributed to the species that have assessments.
+___This could be shown using the n = XXX on the species group quadrant breakdown chart too.___
 
-![](figures/sp_groups_barchart.png)
-
-***Total species occurrences for each data set: all AM, all IUCN, and overlap (one? the other? occurrences of both?)***
+_Total species occurrences for each data set: all AM, all IUCN, and overlap (one? the other? occurrences of both?)_
 
@@
 
-***Relevance of this*** - unequal distribution of assessed species - strengths and weaknesses of each; benefits of using both maybe?
-
-### Comparing alignment
-
-#### Comparing alignment by species
-
-For each species range map, the total area was calculated for all cells that were included in either species distributions, regardless of source.
-The number of cells that were found in both *** <- this implies area may just be cell count?***
-
-Figure 1 is an example of very large differences in a species range map from the IUCN and AquaMaps datasets.
-By considering the entire range of the species to be every cell that is in either dataset, only 3.14% of the map is in both.
-3.18% is covered just by AquaMaps and 77.3% is covered just by IUCN.
-Figure 2
-***focus on the percent overlap of small within large*** - describe this species in terms of alignment
-
-*** Is the area calculated based on cell area, or just number of cells? ***
-
-![](figures/AreaRatiovsPercOverlap.png)
-
-***Walk through the quadrant plot*** - describe meaning of each axis and each quadrant
-
-#### Comparing alignment by species group
-
-![](figures/barchart_spp_gp_quads.png)
-
-Which species groups are better aligned?  (save discussion until later?)
+_Relevance of this_ - unequal distribution of assessed species - strengths and weaknesses of each; benefits of using both maybe?  save for discussion
 
@@
 
-***Where does the 40% threshold come from?  Is it commonly used?  Justification?*** Show effect of using a 40% threshold vs a 0% threshold.
+___Where does the 40% threshold come from?  Is it commonly used?  Justification?___ Show effect of using a 40% threshold vs a 0% threshold.
 
@@
 
-***
+-----
 
-##Section Three:
+# Methodology? discussion?
 
-# Section Four: Implications
+# Implications
 
@@
 
-It is important to remember that although only ***[X (1,923/total number of species)]%*** of all species included in the species goal are found in both datasets, there are still significant differences when electing to choose one dataset over the other.
+It is important to remember that although only ___[X (1,923/total number of species)]%___ of all species included in the species goal are found in both datasets, there are still significant differences when electing to choose one dataset over the other.
 
-***
+-----
 
-##Section Four: Implications of data choices
+## Application to MPA Gap Analysis
 
-***need more info on these options? they both sound interesting though***
+___need more info on these options? they both sound interesting though___
 - could talk about MPA gap analysis
-- Highlight studies that have used one and not the other. Could this have changed the results?
 
-***
+# Conclusion
 
-##References
+-----
 
-**Key AquaMaps publications:**
+#References
+
+__Key AquaMaps publications:__
 
@@
 
-**Papers based on AquaMaps:** ***Use these for section 4?***
+__Papers based on AquaMaps:__ ___Use these for section 4?___
 

### 2015-12-09 ms_outline.Rmd

@@
 
+Mapping species abundances and distributions is a critical component to understanding global biodiversity and detecting patterns and trends through time.
+There have been multiple efforts to map terrestrial, freshwater and marine species from local to global scales.
+These pieces of information inform our understanding of taxonomic assemblages, biogeographical limits to species ranges, suitable habitat for species.
+
+These two spatial datasets have been used to analyze species status globally (Halpern et al.
+2012, Selig et al.
+2013), global biodiversity patterns and trends, predicting range shifts (Molinos et al.
+2015),and setting conservation priorities (Klein et al.
+2015).
+Yet there have been no comprehensive comparisons of these two datasets to identify the differences and provide insight into how using one or the other can influence the outcome of an analysis.
+With impactful conclusions being drawn from these large-scale studies on global biodiversity and conservation, it's imperative that the data used in these analyses are appropriate and the data decisions are informed.
+
+The IUCN Red List of Threatened Species includes more than 19,000 assessed marine species, 
+<!--define what "assessed" means?
+include DD?
+keeping those in for now--> but only **4463** <!--11975 maps in those shapefiles, but fewer than 4500 correspond to assessed marine species...
+and only 2500 of those overall correpond to AquaMaps species--> of these have associated spatial information.
+AquaMaps has developed range maps for over 22,800 species.
+Only 2,100 species have been mapped in both datasets.
+This gap in overlapping species between these two datasets further highlights the need for understanding each of these datasets.
+Both IUCN and AquaMaps have gaps in marine species information and a significant portion of available data may be overlooked if limited to just one of these datasets.
+
+Species range maps provided by AquaMaps (Kaschner et al.
+2013) are model-based, relying on environmental parameters in addition to known occurrences and expert input.
+AquaMaps use information about individual species environmental tolerances for salinity, temperature, depth, productivity, and where applicable, sea ice and distance from land, to create a presence-only gridded range map at a resolution of 0.5 degree cells with cell values ranging from 0 to 1.00.
+These values represent the *relative environmental suitability* for the species, with 0 representing unsuitable habitat and 1.00 being the most suitable habitat.
+Once produced, each range map maintains a dynamic element that allows individuals to adjust parameters based on their expert knowledge to improve the model-generated map.
+
+The maps provided by the IUCN for each assessed species are created by one or more experts on that species.
+There are general guidelines to how the map should be created.
+Specifically, the final mapped polygons are meant to indicate that a species *probably only occurs within the polygon* but are *not* meant to indicate that the species is evenly distributed everywhere within the polygon.
+The guidelines vary slightly for each type of species (terrestrial, freshwater, marine).
+These maps are typically dependent on observation records from databases such as GBIF.
+
+IUCN only publishes spatial distribution map sets for taxonomic groups in which at least 90% of the species within the taxon group have been evaluated.
+While this helps prevent sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
+As of December 2015, IUCN species distribution maps were available for XXX marine species (habitat = "marine" on IUCN species information pages).
+
+<!--
 There have been multiple efforts to map marine species based on both observational data and known life history and habitat preferences. AquaMaps and IUCN are two of these efforts.
+-->
 

### 2015-12-09 data_explore_2015_data.Rmd
added R Markdown codes in manuscript to calc numbers etc

@@
 dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
-dir_fig <- file.path(getwd(), 'figs2015')
+dir_fig <- file.path(getwd(), 'figures')
 

### 2015-12-09 ms_outline.Rmd
added R Markdown codes in manuscript to calc numbers etc

@@
 
-##Main message
+``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
+
+# library(ggplot2)
+# library(maps)
+# library(readr)
+# library(raster)
+# library(data.table)
+library(tidyr)
+library(dplyr)
+library(stringr)
+
+# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
+#            'Darwin'  = '/Volumes/data_edit',
+#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
+
+# dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
+# dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
+# dir_fig <- file.path(getwd(), 'figures')
+```
+
+``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
+### use data_explore.Rmd to create this file, if it does not yet exist.
+
+data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
+spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
+  filter(!is.na(popn_category)) %>% ### filter out unassessed species
+  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
+
+### generate other lists
+
+spp_all_file <- 'data/spp_all_2015.csv'
+spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
+
+iucn_maps_file <- 'data/spp_iucn_maps.csv'
+spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE)
+
+iucn_marine_file <- 'data/spp_iucn_marine.csv'
+spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
+
+spp_am_file <- 'data/spp_am.csv'
+spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
+
+```
+## Main message
 
@@
 
+<!--
 ##Working Abstract  
-<!--
 * Importance of understanding spatial distribution of species
@@
 
-#Introduction
+# Introduction
 
@@
 include DD?
-keeping those in for now--> but only **4463** <!--11975 maps in those shapefiles, but fewer than 4500 correspond to assessed marine species...
+keeping those in for now--> but fewer than 4,500 <!--11975 maps in those shapefiles, but fewer than 4500 correspond to assessed marine species...
 and only 2500 of those overall correpond to AquaMaps species--> of these have associated spatial information.
 AquaMaps has developed range maps for over 22,800 species.
-Only 2,100 species have been mapped in both datasets.
+Only 2,200 species are mapped in both datasets.
 This gap in overlapping species between these two datasets further highlights the need for understanding each of these datasets.
@@
 While this helps prevent sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN species distribution maps were available for XXX marine species (habitat = "marine" on IUCN species information pages).
+As of December 2015, IUCN species distribution maps were available for XXX marine species in YYY taxonomic groups.
+However, only YYY of these species have been formally assessed under the IUCN Red List of Threatened Species (listed on that spreadsheet, and habitat = "marine" on IUCN species information pages).
 
@@
 The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups.
-Species spatial distributions are determined by experts based upon <known occurrence, suitable habitat, and blah>.
+Species spatial distributions are determined by experts based upon ___known occurrence, suitable habitat, and blah___.
 The IUCN presents these maps as 'limits of distibution' with the caveat that the species 'probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon'.<cite?>  IUCN only publishes spatial distribution map sets for taxonomic groups in which at least 90% of the species within the taxon group have been evaluated.
 While this helps prevent sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN species distribution maps were available for XXX marine species (habitat = "marine" on IUCN species information pages).
+As of December 2015, IUCN had published species distribution maps for nearly 12,000 marine species across 26 taxonomic groups, though fewer than 4,500 of these have been formally assessed under the IUCN Red List of Threatened Species.
 
@@
 These maps can then be tweaked by experts altering the parameters and regenerating the modeled map.
-As of December 2015, ___22,000___ species have been mapped.
+As of December 2015, nearly 23,000 species have been mapped, though only 5,600 of these have been evaluated for the IUCN Red List.
 ___Expert mode to further refine models?___ See (Kaschner et al.
@@
 
-___number of species in each data set, and the number that overlap___
+| Map set   | n~total~                | n~Red_List~                                                               |
+| :-------: | ----------------------: | ------------------------------------------------------------------------: |
+| AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))`         |
+| IUCN      | `r nrow(spp_iucn_maps)` | `r nrow(spp_iucn_maps %>% filter(binomial %in% spp_iucn_marine$sciname))` |
+| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$binomial))` | `r nrow(spp_map_pairs)`          |
 
@@
 
-Researchers have two large and useful data sets at their disposal, each with its own utility.  Being able to use both simultaneously makes many more species available, but here there be challenges.
+Researchers have two large and useful data sets at their disposal, each with its own utility.
+Being able to use both simultaneously makes many more species available, but beware: here there be challenges.
 
@@
 
+<!--
 ___Main messages:___
@@
 * basic methodology (raster, overlay, calculate)? or should there be a separate methodology section?
+-->
 
 <!--
-* Here we can discuss the idea of multiple "dimensions" of alignment - distributional and area alignment.  Well-aligned species will show high alignment in both dimensions.  
-* Maybe a diagram here?
-* What can alignment tell us about the individual species, the species groups, and the data sets?
 * Note that alignment does not necessarily mean accuracy or precision - a poorly-understood species could appear to be well-aligned.  
@@
 
-To facilitate our analysis, we selected the subset of marine species that have range maps in both IUCN and AquaMaps (n = ???), matched by Genus and species names, and then rasterized the IUCN shapefiles to match the Aquamaps half-cell raster.
+To facilitate our analysis, we selected the subset of marine species that have range maps in both IUCN (filtering out ranges identified as extinct) and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`), and used genus and species binomials to identify paired maps.
+We then rasterized the IUCN shapefiles to match the Aquamaps half-cell raster.
 We flattened each map by assigning to each cell a simple "presence/absence" value:
 
-* For IUCN data, any cell with non-zero overlap with a range map polygon was considered "presence."  
-* For AquaMaps data, any cell with non-zero probability (or RES) was considered "presence."  
-* In both cases, these criteria for presence will maximize the number of cells included.
-Using these criteria, we found that for 72% ???
+* For IUCN data, we defined "presence" within a cell as any non-zero overlap with a range map polygon.
+* For AquaMaps data, we defined presence as any cell with non-zero probability ___(or RES?)___.  
+* In both cases, these criteria for presence will maximize the number of cells included.  Using these criteria, we found that for 
+`r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`%
 of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
@@
 
-For each of the ___2200___ species whose range is described by both data sets, we calculated the distribution match and the area ratio, and plotted the ___2200___ species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
+For each of the `r nrow(spp_map_pairs)` species whose range is described by both data sets, we calculated the distribution match and the area ratio, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
 
@@
 
-___Figure 1: Distribution alignment vs area alignment for ___2200___ species mapped in both Aquamaps and IUCN species distribution maps.___
+___Figure 1: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
 
@@
 
+``` {r calc medians and quadrant counts, echo = FALSE}
+dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
+area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
+q1_count <- sum(spp_map_pairs$sm_perc > dist_median  & spp_map_pairs$area_ratio >  area_median)
+q2_count <- sum(spp_map_pairs$sm_perc <= dist_median & spp_map_pairs$area_ratio >  area_median)
+q3_count <- sum(spp_map_pairs$sm_perc > dist_median  & spp_map_pairs$area_ratio <= area_median)
+q4_count <- sum(spp_map_pairs$sm_perc <= dist_median & spp_map_pairs$area_ratio <= area_median)
+```
+
 | Description              |   Distribution match |      Area match      |
-| :----------------------: | :-----------------------: | :---------------: |
-| q1: Well-aligned         | Distribution match > XX % | Area match > YY % |
-| q2: Distribution-aligned | Distribution match < XX % | Area match > YY % |
-| q3: Area-aligned         | Distribution match > XX % | Area match < YY % |
-| q4: Poorly-aligned       | Distribution match < XX % | Area match < YY % |
+| :----------------------: | :------------------: | :------------------: |
+| q1: Well-aligned         |  > `r dist_median` % |  > `r area_median` % |
+| q2: Distribution-aligned |  < `r dist_median` % |  > `r area_median` % |
+| q3: Area-aligned         |  > `r dist_median` % |  < `r area_median` % |
+| q4: Poorly-aligned       |  < `r dist_median` % |  < `r area_median` % |
 
-Clearly the upper right quadrant (quadrant 1) is the place to be.
-Species in this quadrant generally agree in spatial distribution between the two distribution maps, and the area of described ranges will be similar; expert-drawn range map and bioclimatic envelope model produce similar predictions of species distribution.
-(n = XXX; YYY %)
+The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
+For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
+(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
 
@@
 Species map pairs that fall within the upper left quadrant (quadrant 2: well-matched in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
-(n = XXX; YYY %)
+(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
 
 The lower right quadrant (quadrant 3: well-matched in area but poorly matched in distribution) includes species for which the two maps generally agree in range area, but disagree in where those ranges occur.
-General disagreement in distribution seems to indicate a greater issue than general disagreement in area extent, so we categorized map pairs in this quadrant as being more problematic than map pairs in either of the upper quadrants.
-(n = XXX; YYY %)
+Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being more problematic than map pairs in either of the upper quadrants.
+(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
 
-The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. (n = XXX; YYY %)
+The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
+(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
 
@@
 
-Figure 2 breaks the ___2200___ species map pairs into taxonomic groups as classified by the published IUCN range maps.
-We were not surprised to find that map pairs for marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2); many of these species are global-ranging pelagic organisms, with some ranges encompassing nearly the entire ocean.
+Figure 2 divides the `r nrow(spp_map_pairs)` species map pairs into taxonomic groups as classified by the published IUCN range maps, and examines the breakdown of each taxonomic group by quadrant.
+
+___move to discussion___ We were not surprised to find that map pairs for marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2); many of these species are global-ranging pelagic organisms, with some ranges encompassing nearly the entire ocean.
 Mangroves are also not a surprise, being immobile and easily mapped by remote sensing methods.
@@
 Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
+Do regression - slope indicates that sensitivity is less than 1:1
 
@@
 
+<!--
 Halpin, P.N., A.J.
@@
 Oceanography 22(2):104-115
+-->
 

### 2015-12-11 R/data_explore_pt2.Rmd
script uses new data
I think the distribution by taxonomic group code is working now...

@@
 ``` {r pct overlap vs lg area, echo = FALSE, warning = FALSE, message = FALSE}
-
-spp_list <- spp_list %>%
-  mutate(lg_area = ifelse(am_area > iucn_area, am_area, iucn_area),
-         sm_area = ifelse(am_area < iucn_area, am_area, iucn_area),
-         lg_area_pct = 100 * lg_area / max(total_area, na.rm = TRUE)) ### use largest data set as approx for total ocean range
 
@@
 ***
+
+###(4) Looking at the percent of smaller species range that is within the larger range.
+
+For each range
+
+```{r pct_small_within_large, echo = FALSE, warning = FALSE, eval = TRUE}
+
+spp_df4 <- spp_list %>%
+  dplyr::select(sciname, sm_perc, sm_range) %>%
+  unique()
+
+hist_sm_within_lg <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
+  ggtheme_plot + 
+  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
+  scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
+  labs(title = 'Percent of smaller range within larger range',
+       x = 'Percent of smaller range that falls within the larger range', 
+       y = sprintf('Proportion of all species (n = %s)', nrow(spp_df4)))
+
+ggsave(file.path(dir_fig, 'PercSmallRangeInLarger.png'), width = 6, height = 4)
+
+print(hist_sm_within_lg)
+
+```

### 2015-12-11 data_explore.Rmd
I think the distribution by taxonomic group code is working now...

@@
 ``` {r, child = 'R/data_explore_pt2.Rmd', eval = FALSE}
-```
-
-###(4) Looking at the percent of smaller species range that is within the larger range.
-
-For each range
-
-```{r pct_small_within_large, echo = FALSE, warning = FALSE, eval = TRUE}
-
-spp_df4 <- spp_list %>%
-  dplyr::select(sciname, sm_perc, sm_range) %>%
-  unique()
-
-hist_sm_within_lg <- ggplot(spp_df4, aes(sm_perc, fill = sm_range)) + 
-  ggtheme_plot + 
-  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 4, alpha = .8) +
-  scale_fill_discrete(name = 'Dataset with smaller range', labels = c('AquaMaps', 'IUCN')) + 
-  labs(title = 'Percent of smaller range within larger range',
-       x = 'Percent of smaller range that falls within the larger range', 
-       y = sprintf('Proportion of all species (n = %s)', nrow(spp_df4)))
-
-ggsave(file.path(dir_fig, 'PercSmallRangeInLarger.png'), width = 6, height = 4)
-
-print(hist_sm_within_lg)
-
 ```
@@
 
-Plotting species from q1 (well-matched) (see data_explore_quadmaps.html for plots)
-```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
-
-x <- 20
-y <- nrow(spp_list_q1)
-
-spp_names_q1 <- spp_list_q1$sciname[seq(1, y, y/x)]
-
-for (spp in spp_names_q1) { # i = 1
-  spp_plot <- plot_rangemap(spp)
-  spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
-  spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 1: %s', spp, text_tag))
-
-  print(spp_plot)
-
-  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q1_%s.png', str_replace(tolower(spp), ' ', '_'))),
-  #        plot = spp_plot, 
-  #        width = 8, height = 4, dpi = 200)
-}
-
-```
-
-Plotting species from q2 (distribution-matched)
-```{r plot_spp_from_q2, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
-
-x <- 20
-y <- nrow(spp_list_q2)
-
-spp_names_q2 <- spp_list_q2$sciname[seq(1, y, y/x)]
-
-for (spp in spp_names_q2) { # i = 1
-  spp_plot <- plot_rangemap(spp)
-  spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
-  spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 2: %s', spp, text_tag))
-  print(spp_plot)
-
-  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q2_%s.png', str_replace(tolower(spp), ' ', '_'))),
-  #        plot = spp_plot, 
-  #        width = 8, height = 4, dpi = 200)
-}
-
-```
-
-Plotting species from q3 (area-matched)
-```{r plot_spp_from_q3, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
-
-x <- 20
-y <- nrow(spp_list_q3)
-
-spp_names_q3 <- spp_list_q3$sciname[seq(1, y, y/x)]
-
-for (spp in spp_names_q3) { # i = 1
-  spp_plot <- plot_rangemap(spp)
-  spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
-  spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 3: %s', spp, text_tag))
-  print(spp_plot)
-
-  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q3_%s.png', str_replace(tolower(spp), ' ', '_'))),
-  #        plot = spp_plot, 
-  #        width = 8, height = 4, dpi = 200)
-}
-
-```
-
-Plotting species from q4 (poorly matched)
-```{r plot_spp_from_q4, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
-
-x <- 20
-y <- nrow(spp_list_q4)
-
-spp_names_q4 <- spp_list_q4$sciname[seq(1, y, y/x)]
-
-for (spp in spp_names_q4) { # i = 1
-  spp_plot <- plot_rangemap(spp)
-  spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
-  spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 4: %s', spp, text_tag))
-  print(spp_plot)
-
-  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q4_%s.png', str_replace(tolower(spp), ' ', '_'))),
-  #        plot = spp_plot, 
-  #        width = 8, height = 4, dpi = 200)
-}
-
+``` {r child = 'R/data_explore_pt3.Rmd', eval = FALSE}
+# Multiple species maps per quadrant
 ```

### 2015-12-11 ms_outline.Rmd
I think the distribution by taxonomic group code is working now...

@@
 ---
-title: "A comparative analysis of AquaMaps and IUCN species ranges"
+title: "A comparative analysis of AquaMaps and IUCN species distribution maps"
 author: "JA, CO, BH"
@@
 
-iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE)
-
 iucn_marine_file <- 'data/spp_iucn_marine.csv'
 spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
+
+iucn_maps_file <- 'data/spp_iucn_maps.csv'
+spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
+  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
+  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
 
@@
 
-The maps provided by the IUCN for each assessed species are created by one or more experts on that species.
-There are general guidelines to how the map should be created.
-Specifically, the final mapped polygons are meant to indicate that a species *probably only occurs within the polygon* but are *not* meant to indicate that the species is evenly distributed everywhere within the polygon.
-The guidelines vary slightly for each type of species (terrestrial, freshwater, marine).
-These maps are typically dependent on observation records from databases such as GBIF.
+Three comparisons presented here:
 
-IUCN only publishes spatial distribution map sets for taxonomic groups in which at least 90% of the species within the taxon group have been evaluated.
-While this helps prevent sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN species distribution maps were available for XXX marine species in YYY taxonomic groups.
-However, only YYY of these species have been formally assessed under the IUCN Red List of Threatened Species (listed on that spreadsheet, and habitat = "marine" on IUCN species information pages).
-
-<!--
-There have been multiple efforts to map marine species based on both observational data and known life history and habitat preferences. AquaMaps and IUCN are two of these efforts.
--->
+* Differences in overall species representation - by taxa and by spatial distribution - how well does each set of maps represent the global distribution of species, the global distribution of biodiversity
+* alignment between maps for species in common - how well do the maps align, when compared at a species-by-species level
+* sensitivity analysis to AquaMaps presence threshold - diff reports use diff thresholds; what is the likely effect? is there a "correct" threshold?
+    * this seems like a sub-analysis to justify choice of 0% threshold etc.  
+    * ___That graph should be box plots at each threshold, not simply mean and/or median.___
 
@@
 
-The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups.
-Species spatial distributions are determined by experts based upon ___known occurrence, suitable habitat, and blah___.
-The IUCN presents these maps as 'limits of distibution' with the caveat that the species 'probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon'.<cite?>  IUCN only publishes spatial distribution map sets for taxonomic groups in which at least 90% of the species within the taxon group have been evaluated.
-While this helps prevent sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
+The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.
+Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.
+These maps are typically dependent on observation records from databases ___such as GBIF other examples?___, and IUCN provides general guidelines to facilitate consistency in how the maps are generated.
+
+The IUCN presents these maps as 'limits of distibution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed everywhere within the polygon.
+While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
+
+IUCN publishes spatial distribution map sets only for taxonomic groups in which at least 90% of the species within the taxon group have been evaluated.
+While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
 As of December 2015, IUCN had published species distribution maps for nearly 12,000 marine species across 26 taxonomic groups, though fewer than 4,500 of these have been formally assessed under the IUCN Red List of Threatened Species.
 
-* _something about format - polygons, fields, etc? presence/no presence_
-* _presence field, and seasonality info_
-* _terrestrial and marine both_
-* _use as a reference:_ https://www.conservationtraining.org/course/view.php?id=217&lang=en
-
-<!--
-Spatial designation of the species distribution is left to the experts with strict guidance for how to assign the designation.
-Using known occurrence data, experts depict the species distribution by polygons using a systematic methodology laid out by the [IUCN](http://www.amphibians.org/wp-content/uploads/2013/09/Red-List-Mapping-standards-ARLA-Jan2014-web-version.pdf) (this needs a better source).
-
-__Data Type__  
-Shapefiles for individual species. Spatial files include subpopulation information, presence (extinct/extant etc.) and seasonality information (breeding/migration/origin etc.).  
-* ___Include other columns typically within each shapefile___
-
-__Methods__  
-From [IUCN site](http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources):  
-    * The species distribution maps, sometimes referred to as 'limits of distribution' or 'field guide' maps, aim to provide the current known distribution of the species within its native range.
-The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range.
-In most cases the range is depicted as polygons.
-_A polygon displaying the limits of a species distribution is essentially meant to communicate that the species probably occurs within this polygon, but it does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon_.
-
-* ___Spatial files for IUCN species groups attempt to include a broad sample across the group, to avoid sampling bias (maybe?)___ - but only some species groups have had spatial data released, so some taxa are not represented at all
-    * from http://www.iucnredlist.org/technical-documents/spatial-data: _"This spatial data provided below is for comprehensively assessed taxonomic groups."  "...
-the data is divided by taxonomic groups.
-Groups that have been comprehensively assessed are those containing >90% species evaluated within the described taxon group."_
-    -->
+* _use as a reference:_ 
+    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
+    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources
 
 ### AquaMaps
-
-[Site](http://www.aquamaps.org/)
 
@@
 These maps can then be tweaked by experts altering the parameters and regenerating the modeled map.
-As of December 2015, nearly 23,000 species have been mapped, though only 5,600 of these have been evaluated for the IUCN Red List.
+As of December 2015, nearly 23,000 species have been mapped, though only 5,600 of these species have been evaluated for the IUCN Red List.
 ___Expert mode to further refine models?___ See (Kaschner et al.
@@
 
-<!--
-### Spatial resolution - not sure we need this, it is not really _resolution_ but more about nuances of distribution... difference already explained above?
-
-These two differ in spatial resolution - AquaMaps provides higher resolution information about where species do and do not occur globally whereas the IUCN range maps are unable to provide specific information about where within a polygon a species occurs.
--->
-
-#### Poor overlap between assessed species
+### Overlap between assessed species
 
@@
 | AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))`         |
-| IUCN      | `r nrow(spp_iucn_maps)` | `r nrow(spp_iucn_maps %>% filter(binomial %in% spp_iucn_marine$sciname))` |
-| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$binomial))` | `r nrow(spp_map_pairs)`          |
-
-* this highlights that each provides valuable and unduplicated spatial range information.
-
-#### Summarize this section?
+| IUCN      | `r nrow(spp_iucn_maps)` | `r nrow(spp_iucn_maps %>% filter(sciname %in% spp_iucn_marine$sciname))`  |
+| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))` | `r nrow(spp_map_pairs)`           |
 
@@
 
-# In what ways do the two data sets differ
+# Analysis of taxonomic and regional inclusion
 
-<!--
-![](./figures/rangemap_chromis_alpha.png)
+Focus on IUCN red-list evaluated species (incl or not incl DD)? or all species?
 
-__Figure 1:__ Species range map for _Chromis alpha_ from the AquaMaps (orange) and IUCN (purple) datasets.
-___is this the right map to use here? find a good one with cool differences___
--->
+## Taxonomic distribution between data sets
 
-### Defining and comparing spatial alignment between the two data sets
+Examine the overlapping and non-overlapping species by taxonomic group (order? family?) - are there differences in taxonomic group coverage between the two data sets?
 
-<!--
-___Main messages:___
+IUCN - translate the taxonomic group into families - is there a way to automate this? http://www.itis.gov/downloads/?
+AM - access taxa group information directly from AM species list - order? family?
 
-* Definition of "alignment" in two dimensions, incl "good" vs "poor"
-* basic methodology (raster, overlay, calculate)? or should there be a separate methodology section?
--->
+``` {r breakdown by taxa, echo = FALSE}
 
-<!--
-* Note that alignment does not necessarily mean accuracy or precision - a poorly-understood species could appear to be well-aligned.  
--->
+sciname_lookup <- read.csv('sciname_lookups/scinames.csv', stringsAsFactors = FALSE)
+
+iucn_maps_class <- spp_iucn_maps %>% 
+  ### add sciname list, then jump through hoops to delete dupes
+  ### NOTE: this also ditches ones with no sciname match...
+  separate(sciname, 'genus', sep = ' ', remove = FALSE, extra = 'drop') %>%
+  mutate(genus = tolower(genus)) %>%
+  left_join(sciname_lookup, by = 'genus') %>% 
+  filter(!(str_detect(tolower(spp_group), 'grass')    & !str_detect(kingdom, 'plant'))) %>%
+  filter(!(str_detect(tolower(spp_group), 'mangrove') & !str_detect(kingdom, 'plant'))) %>%
+  filter(!(str_detect(tolower(spp_group), 'coral')    & !str_detect(kingdom, 'animal'))) %>%
+  filter(!(str_detect(tolower(spp_group), 'reptile')  & !str_detect(kingdom, 'animal'))) %>%
+  filter(!(str_detect(tolower(spp_group), 'seabream') & !str_detect(kingdom, 'animal'))) %>%
+  filter(!(str_detect(tolower(spp_group), 'wrass')    & !str_detect(kingdom, 'animal'))) %>%
+  filter(!(str_detect(tolower(spp_group), 'mamm')     & !str_detect(kingdom, 'animal'))) %>%
+  filter(!is.na(phylum)) %>%
+  select(-kingdom, -phylum, -family) %>%
+  unique()
+
+taxa_iucn_all <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
+  filter(!is.na(class)) %>% ### don't count the unknowns
+  mutate(class = str_to_title(class),
+         n_iucn_tot = n()) %>%        ### length of whole maps list
+  group_by(class) %>%
+  summarize(n_iucn_class_all  = n(),
+            p_iucn_class_all  = round((n_iucn_class_all/first(n_iucn_tot) * 100), 2))
+
+taxa_iucn_solo <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
+  filter(!is.na(class)) %>% ### don't count the unknowns
+  filter(!sciname %in% spp_am$sciname) %>%  ### filter list to IUCN only
+  mutate(class = str_to_title(class),
+         n_iucn_solo = n()) %>%       ### length of IUCN only list
+  group_by(class) %>%
+  summarize(n_iucn_class_solo = n(),
+            p_iucn_class_solo = round((n_iucn_class_solo/first(n_iucn_solo) * 100), 2))
+
+taxa_am_all <- spp_am %>%
+  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
+  mutate(class = str_to_title(class),
+         n_am_tot = n()) %>%        ### length of whole AM maps list
+  group_by(class) %>%
+  summarize(n_am_class_all  = n(),
+            p_am_class_all  = round((n_am_class_all/first(n_am_tot) * 100), 2))
+
+taxa_am_solo <- spp_am %>%
+  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
+  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
+  mutate(class = str_to_title(class),
+         n_am_solo = n()) %>%       ### length of AM only list
+  group_by(class) %>%
+  summarize(n_am_class_solo = n(),
+            p_am_class_solo = round((n_am_class_solo/first(n_am_solo) * 100), 2))
+
+taxa_both <- spp_am %>%
+  filter(sciname %in% spp_iucn_maps$sciname) %>%
+  filter(!class %in% c('Not assigned', 'null')) %>%
+  mutate(class = str_to_title(class),
+         n_both_tot = n()) %>%
+  group_by(class) %>%
+  summarize(n_class_both = n(),
+            p_class_both = round(n_class_both/first(n_both_tot) * 100, 2))
+
+taxa_df <- taxa_am_all %>%
+  full_join(taxa_am_solo,   by = 'class') %>%
+  full_join(taxa_iucn_all,  by = 'class') %>%
+  full_join(taxa_iucn_solo, by = 'class') %>%
+  full_join(taxa_both,      by = 'class') %>%
+  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))
+
+knitr::kable(taxa_df %>%
+               select(class, am_perc = p_am_class_all, iucn_perc = p_iucn_class_all, both_perc = p_class_both))
+```
+
+## Spatial global distribution between data sets
+
+___main messages:___
+
+* IUCN seems to be weighted toward tropics; AM might be more evenly distributed?
+* This difference in distribution might come into play when selecting a data set for a more regionally-focused study
+* ... or for global studies, it might be desirable to use both data sets to get better global distribution?
+* Small number of overlapping species indicates that each offers data on distinct spp
+* This section is really supporting the idea that it is critical to understanding the differences, as both data sets support one another
+
+Using predifined species groups by the IUCN, it is clear to see that tropical species - specifically wrasses, corals, blennies and groupers, make up the majority of species found in both.
+This is driven more by the IUCN dataset since the number of species with spatial data is directly attributed to the species that have assessments.
+___This could be shown using the n = XXX on the species group quadrant breakdown chart too.___
+
+_Total species occurrences for each data set: all AM, all IUCN, and overlap (one? the other? occurrences of both?)_
+
+![](figures/all_aquamaps_map.png)
+
+_Relevance of this_ - unequal distribution of assessed species - strengths and weaknesses of each; benefits of using both maybe?  save for discussion
+
+-----
+
+# Analysis of spatial alignment by species
+
+## Defining spatial alignment between the two data sets
 
@@
 
-<!--
-### Determining "presence" for each data set
-
-To compare these two datasets for all species found in each (n = ___1920___?), we rasterized the IUCN polygons to the same cell resolution as AquaMaps: 0.5 degrees in both latitude and longitude.
-For every grid cell where the polygon overlapped ___the center point?
-or is it based on ANY overlap, e.g.
-using weights and including any non-zero weight?
-This could boost IUCN presence___, a value of 1 was assigned.
-For the purposes of comparison, the probability of occurrence values in the AquaMaps data were ignored and all cells where there was a value greater than 0 were assigned a 1.
-This way we have comparable maps for total species range.
-- ___How do we justify a zero-threshold here?___ 
-    - need to understand the "probability" - is it really ___relative environmental suitability___   From Kaschner et al 2011: "Patterns based on a presence threshold RES.0.6 (Fig.
-1) were most strongly supported by empirical species richness data (see ‘Validation of species richness’, below)."
-    - since IUCN "presence" seems to be pretty loose, set a wide tolerance on AM?
--->
-
-___Main messages:___
-
-* breaking down alignment differences by quadrant provides useful information.
-    * q1 = well-matched (well understood? not necessarily, could be luck or over-generous environmental envelopes)
-    * other quadrants have defining characteristics that can highlight areas where we should look for better data and better understanding
-
-<!--
-Figure 1 shows four representative species: Kajikia Albida, with a high degree of alignment in both area and distribution; Galaxea cryptoramosa, well-matched in spatial alignment but poorly matched in areal alignment; Conus magnificus, well-matched in area but poorly matched in distribution; and Acanthurus nigrosis, poorly matched in both area and distribution.
-
-![](figures/four_maps.png)
-
-___Figure 1: Representative species with distribution mapped in both AquaMaps and IUCN data sets.___  
--->
+## Comparing spatial alignment between the two data sets
 
@@
 
-___Figure 1: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
+___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
+
+___ditch the species identifiers here?  or highlight ones that are used in discussion of possible mechanistic explanations for differences?___
 
@@
 
-<!--
-(Looking at some representative species maps)
-The range described by AquaMaps map covers XXX km^2^; the range described by IUCN map covers YYY km^2^; the overlap between the two maps covers ZZZ km^2^.
-Distribution match = ZZZ km^2^ / XXX km^2^; area match = XXX km^2^ / YYY km^2^.
-
-* Single species map: ___Do we want this as a separate figure?___ _it helps to explain what the hell we're talking about; but we could also just jump to the quadrant map?_ 
-* ___OR: plot four representative maps FIRST, to show differences in the two dimensions of alignment, THEN show the quadrant map with those species indicated.___ 
-    * _This is backwards from how we did the analysis, but I think might flow better for the reader._
--->
-
 ![](figures/barchart_spp_gp_quads.png)
@@
 
-___move to discussion___ We were not surprised to find that map pairs for marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2); many of these species are global-ranging pelagic organisms, with some ranges encompassing nearly the entire ocean.
+___move to discussion?___ We were not surprised to find that map pairs for marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2); many of these species are global-ranging pelagic organisms, with some ranges encompassing nearly the entire ocean.
 Mangroves are also not a surprise, being immobile and easily mapped by remote sensing methods.
@@
 
-___Other hypotheses/case studies to explain mechanistic differences?___
+___Other hypotheses/case studies to highlight possible mechanistic differences?___
+
+# AquaMaps: Effect of changing "presence" threshold on area representation
+
+![](figures/AMThreshold.png) ___add a mean/median vline on this?___ do we need this figure?
+it looks cool though...
+why broken down by risk category?
+Ditch NE category since most studies probably care about the species health as well as distribution?
+
+___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
+
+![](figures/scatter_AMThreshold_mockup.png)
+Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
+Do regression - shallow slope indicates that sensitivity is less than 1:1.
 
 -----
-
-### Differences in global distribution between data sets
-
-___main messages:___
-
-* IUCN seems to be weighted toward tropics; AM might be more evenly distributed?
-* This difference in distribution might come into play when selecting a data set for a more regionally-focused study
-* ... or for global studies, it might be desirable to use both data sets to get better global distribution?
-* Small number of overlapping species indicates that each offers data on distinct spp
-* This section is really supporting the idea that it is critical to understanding the differences, as both data sets support one another
-
-Using predifined species groups by the IUCN, it is clear to see that tropical species - specifically wrasses, corals, blennies and groupers, make up the majority of species found in both.
-This is driven more by the IUCN dataset since the number of species with spatial data is directly attributed to the species that have assessments.
-___This could be shown using the n = XXX on the species group quadrant breakdown chart too.___
-
-_Total species occurrences for each data set: all AM, all IUCN, and overlap (one? the other? occurrences of both?)_
-
-![](figures/all_aquamaps_map.png)
-
-_Relevance of this_ - unequal distribution of assessed species - strengths and weaknesses of each; benefits of using both maybe?  save for discussion
-
-### Effect of changing threshold on area representation
-
-![](figures/AMThreshold.png)
-
-___Where does the 40% threshold come from?  Is it commonly used?  Justification?___ Show effect of using a 40% threshold vs a 0% threshold.
-
-![](figures/scatter_AMThreshold.png)
-Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
-Do regression - slope indicates that sensitivity is less than 1:1
-
------
-
-# Methodology? discussion?
 

### 2015-12-11 sciname_lookups/process_scinames.R

@@ -1,0 +1,34 @@
+library(tidyr)
+library(dplyr)
+library(stringr)
+
+files_list <- list.files('sciname_lookups')
+
+all_df <- data.frame()
+
+for (i in 1:length(files_list)) { # i = 1
+  x <- scan(file.path('sciname_lookups', files_list[i]), what = 'character', sep = '\n', skip = 3)
+  
+  df <- data.frame('raw' = str_trim(x), stringsAsFactors = FALSE) %>%
+    mutate(raw = str_replace(raw, '  ', ' ')) %>%
+    separate(raw, c('level', 'name'), sep = ' ', remove = FALSE, extra = 'drop') %>%
+    mutate(level = tolower(str_replace(level, ':', '')),
+           name = tolower(name)) %>%
+    mutate(sub_infra = str_detect(level, '^sub')  | str_detect(level, '^infra')) %>%
+    mutate(kingdom = ifelse(str_detect(level, 'kingdom'), name, NA),
+           phylum  = ifelse(str_detect(level, 'phylum'),  name, NA),
+           class   = ifelse(str_detect(level, 'class'),   name, NA),
+           order   = ifelse(str_detect(level, 'order'),   name, NA),
+           family  = ifelse(str_detect(level, 'family'),  name, NA),
+           genus   = ifelse(str_detect(level, 'genus'),   name, NA))
+  
+  df1 <- df %>%
+    filter(!sub_infra) %>% 
+    fill(kingdom:family) %>%
+    filter(!is.na(genus)) %>%
+    select(kingdom:genus)
+  
+  all_df <- rbind(all_df, df1)
+}
+
+write.csv(all_df, 'sciname_lookups/scinames.csv', row.names = FALSE)

### 2015-12-12 R/data_explore_global_maps.Rmd
more on manuscript...

--- 2015-12-11 data_explore.Rmd
@@
----
-title: 'Data Exploration: IUCN and AquaMaps'
-output: html_document
----
+###Global representation of species
 
-```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
+By counting the number of unique species are found in each of the half degree cells we can then create a single global raster of number of species.
 
-# Libraries and Paths
-
+``` {r setup, message = FALSE, eval = FALSE}
 library(rasterVis)
@@
 library(maps)
-library(readr)
 library(raster)
@@
 library(stringr)
+library(readr)
 
@@
 dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
+dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')
 dir_fig <- file.path(getwd(), 'figures')
@@
 source('R/data_explore_fxns.R')
+
+data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
+
+spp_list <- read.csv(data_file, stringsAsFactors = FALSE) %>%
+  filter(!is.na(sm_perc)) ### will ditch the ones where no IUCN polygon exists
+
 ```
 
-```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
-
-# Data files
-#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), 
+``` {r load data sets, echo = FALSE, message = FALSE}
+iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                            col_types = 'cdddd', progress = TRUE) %>%
@@
 
-```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-data_file <- 'data/spp_list_w_area_percent_noextinct.csv'
-
-if(!file.exists(data_file)) {
-  ### species info list - scinames, IUCN species ID, AM species ID, category etc
-  spp_list_base <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
-
-  create_spp_list(spp_list_base, data_file)
-}
-
-spp_list <- read.csv(data_file, stringsAsFactors = FALSE)
-```
-
-``` {r child = 'R/data_explore_pt1.Rmd', eval = FALSE}
-# bar chart: number of spp per group
-# scatter plot: area compare IUCN vs AM, including log plot and plot by species group
-# rangemap: Chromis alpha 
-```
-
-``` {r, child = 'R/data_explore_pt2.Rmd', eval = FALSE}
-```
-
-``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-
-spp_list <- spp_list %>%
-  mutate(area_ratio = am_area / iucn_area,
-         area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
-         area_ratio = area_ratio * 100)
-
-### define 25% and 75% quartiles for both x and y axes
-area_ratio_quartiles <- quantile(spp_list$area_ratio, na.rm = TRUE)[2:4]
-pct_overlap_quartiles <- quantile(spp_list$sm_perc, na.rm = TRUE)[2:4]
-
-### define representative species from each quadrant, for over-plotting
-spp_quads <- c('q1' = 'Kajikia albida', 
-               'q2' = 'Galaxea cryptoramosa', 
-               'q4' = 'Acanthurus nigroris', 
-               'q4a' = 'Metanephrops australiensis', 
-               'q3' = 'Conus magnificus')
-
-### truncate the species names for the representative species
-spp_list_label <- spp_list %>%
-  filter(sciname %in% spp_quads) %>%
-  separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
-  mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
-  select(-g, -s)
-
-### mongo plot time
-scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
-                                         transform(popn_category = reorder(popn_category, category_score)),
-                                       aes(x = area_ratio, y = sm_perc, 
-                                           color = popn_category,
-                                           size = popn_category
-                                           )) + 
-  ggtheme_plot + 
-  ### color the quadrant backgrounds:
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
-  geom_point() + 
-  scale_colour_manual(values = c('LC' = 'green4', 
-                                 'NT' = 'greenyellow', 
-                                 'VU' = 'yellow', 
-                                 'EN' = 'orange', 
-                                 'CR' = 'orangered2')) +
-  geom_vline(x = area_ratio_quartiles[c(1, 3)],
-             color = 'grey50', linetype = 'dashed') +
-  geom_hline(y = pct_overlap_quartiles[c(1, 3)],
-             color = 'grey50', linetype = 'dashed') +
-  geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
-  geom_text(data = spp_list_label, 
-            aes(label = sciname_label), 
-            color = 'grey30',
-            fontface = 'bold.italic',
-            size = 2,
-            hjust = .25,
-            vjust = 1.5) +
-  labs(x = 'Area match: \n(range area smaller : range area larger) * 100%', 
-       y = 'Distribution match: \n % overlap of smaller range w/ larger', 
-       title = 'Area match vs distribution match',
-       color = 'Risk', size = 'Risk')
-
-ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 6, height = 4)
-
-print(scatter_arearatio_vs_overlap)
-
-```
-
-```{r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-
-for (i in 1:length(spp_quads)) { # i = 1
-  spp <- spp_quads[i]
-  spp_quad <- names(spp_quads)[i]
-
-  spp_plot <- plot_rangemap(spp)
-  print(spp_plot)
-
-  ggsave(filename = file.path(dir_fig, sprintf('rangemap_%s_%s.png', spp_quad,
-                                               str_replace(tolower(spp), ' ', '_'))),
-         plot = spp_plot, 
-         width = 8, height = 4, dpi = 200)
-}
-
-```
-
-``` {r examine_spp_gp_quartiles, echo = FALSE, eval = TRUE}
-
-spp_list1 <- spp_list %>%
-  group_by(spp_group) %>%
-  mutate(n_spp = n()) %>%
-  ungroup()
-
-### Create species lists by quadrant
-spp_list_q1 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE))
-spp_list_q2 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE))
-spp_list_q3 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE))
-spp_list_q4 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE))
-
-### Summarize quadrant species list by species group (e.g. CORALS1)
-spp_gp_q1 <- spp_list_q1 %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp), n_spp_q1 = n())
-spp_gp_q2 <- spp_list_q2 %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp), n_spp_q2 = n())
-spp_gp_q3 <- spp_list_q3 %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp), n_spp_q3 = n())
-spp_gp_q4 <- spp_list_q4 %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = first(n_spp), n_spp_q4 = n())
-
-### Join species group by quadrant into a single dataframe
-spp_gp_quadrants <- spp_gp_q1 %>%
-  left_join(spp_gp_q2, by = c('spp_group', 'n_spp')) %>%
-  left_join(spp_gp_q3, by = c('spp_group', 'n_spp')) %>%
-  left_join(spp_gp_q4, by = c('spp_group', 'n_spp')) %>%
-  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
-  mutate(quad = str_replace(quad, 'n_spp_', ''),
-         pct_quad = n_quad/n_spp)
-
-spp_gp_quadrants <- spp_gp_quadrants %>%
-  left_join(spp_gp_quadrants %>% 
-              filter(quad == 'q1') %>%
-              select(spp_group, pct_q1 = pct_quad),
-            by = 'spp_group') %>%
-  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
-         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
-         quad = str_replace(quad, 'q3', 'q3 area-matched'),
-         quad = str_replace(quad, 'q4', 'q4 poorly matched'))
-
-### Plot the bar chart
-barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
-                                  transform(spp_group = reorder(spp_group, pct_q1)), 
-                                aes(x = spp_group, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'bin', alpha = .5) +
-  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
-                               'q2 dist-matched'   = 'greenyellow', 
-                               'q3 area-matched'   = 'orange', 
-                               'q4 poorly matched' = 'orangered2')) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 4) +
-  coord_flip() +
-  labs(x = 'Species Group', 
-       y = 'Relative number of species per quadrant', 
-       title = 'Species group breakdown by quadrant')
-
-print(barchart_spp_gp_quads)
-ggsave(plot = barchart_spp_gp_quads,
-       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
-       width = 6, height = 4)
-
-```
-
-``` {r child = 'R/data_explore_pt3.Rmd', eval = FALSE}
-# Multiple species maps per quadrant
-```
-
-***
-
-###Global representation of species
-
-By counting the number of unique species are found in each of the half degree cells we can then create a single global raster of number of species.
-
-```{r global_dist_map, echo = FALSE, warning = FALSE, message = FALSE}
+```{r global map matched species, echo = FALSE, warning = FALSE, message = FALSE}
 
 ### find total species in each dataset for later percent calcs
-am_n_tot   <- am_spp_cells %>% 
-  filter(am_sid %in% unique(spp_list$am_sid)) %>%
-  select(am_sid) %>% 
-  unique() %>% nrow()
-iucn_n_tot <- iucn_spp_cells %>% 
-  filter(sciname %in% unique(spp_list$sciname)) %>%
-  select(sciname) %>% 
-  unique() %>% nrow()
+# am_n_tot   <- am_spp_cells %>% 
+#   filter(am_sid %in% unique(spp_list$am_sid)) %>%
+#   dplyr::select(am_sid) %>% 
+#   unique() %>% nrow()
+# iucn_n_tot <- iucn_spp_cells %>% 
+#   filter(sciname %in% unique(spp_list$sciname)) %>%
+#   dplyr::select(sciname) %>% 
+#   unique() %>% nrow()
 ### good - the same number
@@
   group_by(loiczid) %>%
-  summarize(count_am = n(),
-            pct_am = count_am/am_n_tot) %>%
+  summarize(count_am = n()) %>%
+#   summarize(count_am = n(),
+#             pct_am = count_am/am_n_tot) %>%
   as.data.frame()
@@
   group_by(loiczid) %>%
-  summarize(count_iucn = n(),
-            pct_iucn = count_iucn/iucn_n_tot) %>%
+  summarize(count_iucn = n()) %>%
+#   summarize(count_iucn = n(),
+#             pct_iucn = count_iucn/iucn_n_tot) %>%
   as.data.frame()
@@
 
-Now look at all aquamaps species
+Now look at full aquamaps species
 
@@
 
-am_spp_cells_big <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/ohi_hcaf_species_native.csv')) %>%
+am_spp_cells_big <- fread(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) %>%
   as.data.frame()
-### am_spp_cell_big is missing loiczid info so bring in hcaf to match
-
-hcaf <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/hcaf.csv')) %>% #bring this in just to get loiczid
-  as.data.frame() %>%
-  select(CsquareCode, LOICZID) %>%
-  rename(loiczid = LOICZID)
+# ### am_spp_cell_big is missing loiczid info so bring in hcaf to match
+# 
+# hcaf <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/hcaf.csv')) %>% #bring this in just to get loiczid
+#   as.data.frame() %>%
+#   dplyr::select(CsquareCode, LOICZID) %>%
+#   rename(loiczid = LOICZID)
 
@@
 am_all_n_tot <- am_spp_cells_big %>% 
-  select(SpeciesID) %>% 
+  dplyr::select(SpeciesID) %>% 
   unique() %>% nrow()
@@
 all_am <- am_spp_cells_big %>%
-  select(SpeciesID, CsquareCode) %>%
+  dplyr::select(SpeciesID, CsquareCode) %>%
   group_by(CsquareCode) %>%
@@
 ```
-
-``` {r, child = 'R/data_explore_am_thresh.Rmd', eval = FALSE}
-```
-
-## Changes in SPP for scenarios
-
-Comparing three scenarios to the current (v2015) global model for SPP:
-
-* Baseline is: preferring IUCN rangemaps over Aquamaps rangemaps; Aquamaps 'presence' threshold is set at 40% or greater.
-* Scenario 1: preferring IUCN rangemaps over Aquamaps; AM presence threshold set at >= 1% (effectively, all occurrences).
-* Scenario 2: preferring AM maps over IUCN; AM presence >= 40%.
-* Scenario 3: preferring AM over IUCN; AM presence >= 1%.
-
-Boxplots:
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
-
-![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)

### 2015-12-12 data_explore.Rmd
more on manuscript...

@@
 
-library(rasterVis)
 library(ggplot2)
@@
 library(readr)
-library(raster)
 library(data.table)
@@
 #iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_both.csv'), 
+iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                            col_types = 'cdddd', progress = TRUE) %>%
@@
 ```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-data_file <- 'data/spp_list_w_area_percent_noextinct.csv'
+data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
 
-if(!file.exists(data_file)) {
+if (!file.exists(data_file)) {
   ### species info list - scinames, IUCN species ID, AM species ID, category etc
-  spp_list_base <- read.csv('shiny/data/spp_am_v_iucn.csv', stringsAsFactors = FALSE)
+  spp_list_base <- read.csv('data/spp_am_v_iucn_2015_inclDD.csv', stringsAsFactors = FALSE)
 
@@
 
-spp_list <- read.csv(data_file, stringsAsFactors = FALSE)
+spp_list <- read.csv(data_file, stringsAsFactors = FALSE) %>%
+  filter(!is.na(sm_perc)) ### will ditch the ones where no IUCN polygon exists
+
 ```
@@
 ``` {r child = 'R/data_explore_pt1.Rmd', eval = FALSE}
-# bar chart: number of spp per group
+```
+<!-- # bar chart: number of spp per group
 # scatter plot: area compare IUCN vs AM, including log plot and plot by species group
-# rangemap: Chromis alpha 
-```
+# rangemap: Chromis alpha -->
 
@@
 ### define representative species from each quadrant, for over-plotting
-spp_quads <- c('q1' = 'Kajikia albida', 
-               'q2' = 'Galaxea cryptoramosa', 
-               'q4' = 'Acanthurus nigroris', 
-               'q4a' = 'Metanephrops australiensis', 
-               'q3' = 'Conus magnificus')
+# spp_quads <- c('q1' = 'Kajikia albida', 
+#                'q2' = 'Galaxea cryptoramosa', 
+#                'q4' = 'Acanthurus nigroris', 
+#                'q4a' = 'Metanephrops australiensis', 
+#                'q3' = 'Conus magnificus')
 
 ### truncate the species names for the representative species
-spp_list_label <- spp_list %>%
-  filter(sciname %in% spp_quads) %>%
-  separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
-  mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
-  select(-g, -s)
+# spp_list_label <- spp_list %>%
+#   filter(sciname %in% spp_quads) %>%
+#   separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
+#   mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
+#   select(-g, -s)
 
@@
                                            color = popn_category,
-                                           size = popn_category
-                                           )) + 
+                                           size = popn_category)) + 
   ggtheme_plot + 
@@
                                  'EN' = 'orange', 
-                                 'CR' = 'orangered2')) +
+                                 'CR' = 'orangered2',
+                                 'DD' = 'grey30')) +
   geom_vline(x = area_ratio_quartiles[c(1, 3)],
@@
              color = 'grey50', linetype = 'dashed') +
-  geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
-  geom_text(data = spp_list_label, 
-            aes(label = sciname_label), 
-            color = 'grey30',
-            fontface = 'bold.italic',
-            size = 2,
-            hjust = .25,
-            vjust = 1.5) +
+#   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
+#   geom_text(data = spp_list_label, 
+#             aes(label = sciname_label), 
+#             color = 'grey30',
+#             fontface = 'bold.italic',
+#             size = 2,
+#             hjust = .25,
+#             vjust = 1.5) +
   labs(x = 'Area match: \n(range area smaller : range area larger) * 100%', 
@@
 
-```{r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+```{r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 spp_list_q1 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE))
+  filter(area_ratio >= median(area_ratio) & sm_perc >= median(sm_perc))
 spp_list_q2 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc >= median(sm_perc, na.rm = TRUE))
+  filter(area_ratio < median(area_ratio) & sm_perc >= median(sm_perc))
 spp_list_q3 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE))
+  filter(area_ratio >= median(area_ratio) & sm_perc < median(sm_perc))
 spp_list_q4 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio, na.rm = TRUE) & sm_perc < median(sm_perc, na.rm = TRUE))
+  filter(area_ratio < median(area_ratio) & sm_perc < median(sm_perc))
 
@@
 spp_gp_quadrants <- spp_gp_q1 %>%
-  left_join(spp_gp_q2, by = c('spp_group', 'n_spp')) %>%
-  left_join(spp_gp_q3, by = c('spp_group', 'n_spp')) %>%
-  left_join(spp_gp_q4, by = c('spp_group', 'n_spp')) %>%
+  full_join(spp_gp_q2, by = c('spp_group', 'n_spp')) %>%
+  full_join(spp_gp_q3, by = c('spp_group', 'n_spp')) %>%
+  full_join(spp_gp_q4, by = c('spp_group', 'n_spp')) %>%
   gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
   mutate(quad = str_replace(quad, 'n_spp_', ''),
-         pct_quad = n_quad/n_spp)
+         pct_quad = n_quad/n_spp,
+         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
+         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))
 
@@
               filter(quad == 'q1') %>%
-              select(spp_group, pct_q1 = pct_quad),
+              dplyr::select(spp_group, pct_q1 = pct_quad),
             by = 'spp_group') %>%
@@
        y = 'Relative number of species per quadrant', 
-       title = 'Species group breakdown by quadrant')
+       title = 'Species group breakdown by quadrant',
+       fill = 'Match quadrant')
 
@@
 ``` {r child = 'R/data_explore_pt3.Rmd', eval = FALSE}
-# Multiple species maps per quadrant
+```
+<!-- # Multiple species maps per quadrant -->
+
+***
+
+Breakdown maps by taxa
+
+``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
+### use data_explore.Rmd to create this file, if it does not yet exist.
+
+data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
+spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
+  filter(!is.na(popn_category)) %>% ### filter out unassessed species
+  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
+
+### generate other lists
+
+spp_all_file <- 'data/spp_all_2015.csv'
+spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
+
+iucn_marine_file <- 'data/spp_iucn_marine.csv'
+spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
+
+iucn_maps_file <- 'data/spp_iucn_maps.csv'
+spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
+  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
+  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
+
+spp_am_file <- 'data/spp_am.csv'
+spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
+
+```
+
+``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE}
+
+sciname_lookup <- read.csv('sciname_lookups/spp_iucn_all.csv', stringsAsFactors = FALSE)
+
+iucn_maps_class <- spp_iucn_maps %>% 
+  left_join(sciname_lookup %>% 
+              select(sciname, class) %>%
+              unique(), 
+            by = 'sciname')
+
+taxa_iucn_all <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
+  filter(!is.na(class)) %>% ### don't count the unknowns
+  mutate(class = str_to_title(class),
+         n_iucn_tot = n()) %>%        ### length of whole maps list
+  group_by(class) %>%
+  summarize(n_iucn_class_all  = n(),
+            p_iucn_class_all  = round((n_iucn_class_all/first(n_iucn_tot) * 100), 2))
+
+taxa_iucn_solo <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
+  filter(!is.na(class)) %>% ### don't count the unknowns
+  filter(!sciname %in% spp_am$sciname) %>%  ### filter list to IUCN only
+  mutate(class = str_to_title(class),
+         n_iucn_solo = n()) %>%       ### length of IUCN only list
+  group_by(class) %>%
+  summarize(n_iucn_class_solo = n(),
+            p_iucn_class_solo = round((n_iucn_class_solo/first(n_iucn_solo) * 100), 2))
+
+taxa_am_all <- spp_am %>%
+  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
+  mutate(class = str_to_title(class),
+         n_am_tot = n()) %>%        ### length of whole AM maps list
+  group_by(class) %>%
+  summarize(n_am_class_all  = n(),
+            p_am_class_all  = round((n_am_class_all/first(n_am_tot) * 100), 2))
+
+taxa_am_solo <- spp_am %>%
+  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
+  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
+  mutate(class = str_to_title(class),
+         n_am_solo = n()) %>%       ### length of AM only list
+  group_by(class) %>%
+  summarize(n_am_class_solo = n(),
+            p_am_class_solo = round((n_am_class_solo/first(n_am_solo) * 100), 2))
+
+taxa_both <- spp_am %>%
+  filter(sciname %in% spp_iucn_maps$sciname) %>%
+  filter(!class %in% c('Not assigned', 'null')) %>%
+  mutate(class = str_to_title(class),
+         n_both_tot = n()) %>%
+  group_by(class) %>%
+  summarize(n_class_both = n(),
+            p_class_both = round(n_class_both/first(n_both_tot) * 100, 2))
+
+taxa_am_all_assessed <- spp_am %>%
+  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
+  filter(sciname %in% spp_iucn_marine$sciname) %>%
+  mutate(class = str_to_title(class),
+         n_am_tot = n()) %>%        ### length of whole AM maps list
+  group_by(class) %>%
+  summarize(n_am_class_ass  = n(),
+            p_am_class_ass  = round((n_am_class_ass/first(n_am_tot) * 100), 2))
+
+taxa_am_solo_assessed <- spp_am %>%
+  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
+  filter(sciname %in% spp_iucn_marine$sciname) %>%
+  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
+  mutate(class = str_to_title(class),
+         n_am_solo = n()) %>%       ### length of AM only list
+  group_by(class) %>%
+  summarize(n_am_class_solo_ass = n(),
+            p_am_class_solo_ass = round((n_am_class_solo_ass/first(n_am_solo) * 100), 2))
+
+taxa_df <- taxa_am_all %>%
+  full_join(taxa_am_solo,   by = 'class') %>%
+  full_join(taxa_am_all_assessed,  by = 'class') %>%
+  full_join(taxa_am_solo_assessed, by = 'class') %>%
+  full_join(taxa_iucn_all,  by = 'class') %>%
+  full_join(taxa_iucn_solo, by = 'class') %>%
+  full_join(taxa_both,      by = 'class') %>%
+  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))
+
+knitr::kable(taxa_df %>%
+               select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both))
+```
+
+``` {r plot bar charts by taxa, echo = FALSE, message = FALSE, warning = FALSE}
+### bar charts comparing number of species in each class for each data set
+
+taxa_df_long1 <- taxa_df %>%
+  select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both) %>%
+  gather(dataset, n_spp, AquaMaps:Both)
+bar_taxa1 <- ggplot(taxa_df_long1, aes(x = class, y = n_spp)) +
+  ggtheme_plot + 
+  geom_bar(stat = 'identity') +
+  facet_grid(dataset ~ ., scales = 'free_y') +
+  theme(axis.text.x = element_text(angle = 75, hjust = 1))
+print(bar_taxa1)
+
+taxa_df_long2 <- taxa_df %>%
+  select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
+  gather(dataset, n_spp, AquaMaps:Both) %>%
+  group_by(class) %>%
+  mutate(n_total = sum(n_spp, na.rm = TRUE),
+         percent = n_spp/n_total * 100)
+bar_taxa2 <- ggplot(taxa_df_long2, aes(x = class, y = percent, fill = dataset)) +
+  ggtheme_plot + 
+  geom_bar(stat = 'identity', alpha = .5) +
+  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
+  geom_text(aes(label = sprintf('n = %s', n_total), y = .05), hjust = 0, size = 4, angle = 90, hjust = 1) +
+  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
+  labs(x = 'Class', y = 'Percent of mapped species',
+       title = 'Distribution of species maps by taxonomic group (class)')
+print(bar_taxa2)
+ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), height = 4, width = 8)
+
+taxa_df_long3 <- taxa_df %>%
+  select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
+  gather(dataset, n_spp, AquaMaps:Both) %>%
+  group_by(class) %>%
+  mutate(n_total = sum(n_spp, na.rm = TRUE),
+         percent = n_spp/n_total * 100)
+bar_taxa3 <- ggplot(taxa_df_long3, aes(x = class, y = percent, fill = dataset)) +
+  ggtheme_plot + 
+  geom_bar(stat = 'identity', alpha = .5) +
+  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
+  geom_text(aes(label = sprintf('n = %s', n_total), y = .05), hjust = 0, size = 4, angle = 90, hjust = 1) +
+  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
+  labs(x = 'Class', y = 'Percent of mapped and assessed species',
+       title = 'Distribution of species maps by taxonomic group (class)')
+print(bar_taxa3)
+ggsave(filename = file.path(dir_fig, 'bar_by_taxa_assessed.png'), height = 4, width = 8)
+
 ```
@@
 
-###Global representation of species
-
-By counting the number of unique species are found in each of the half degree cells we can then create a single global raster of number of species.
-
-```{r global_dist_map, echo = FALSE, warning = FALSE, message = FALSE}
-
-### find total species in each dataset for later percent calcs
-am_n_tot   <- am_spp_cells %>% 
-  filter(am_sid %in% unique(spp_list$am_sid)) %>%
-  select(am_sid) %>% 
-  unique() %>% nrow()
-iucn_n_tot <- iucn_spp_cells %>% 
-  filter(sciname %in% unique(spp_list$sciname)) %>%
-  select(sciname) %>% 
-  unique() %>% nrow()
-### good - the same number
-
-### Number of aquamaps species per cell
-am_n_per_cell <- am_spp_cells %>%
-  filter(am_sid %in% unique(spp_list$am_sid)) %>%
-  group_by(loiczid) %>%
-  summarize(count_am = n(),
-            pct_am = count_am/am_n_tot) %>%
-  as.data.frame()
-
-am_sp_map <- subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'count_am')], 
-                  by = 'loiczid', 
-                  which = 'count_am', 
-                  subsWithNA = TRUE)
-
-### Number of IUCN species per cell
-iucn_n_per_cell <- iucn_spp_cells %>%
-  group_by(loiczid) %>%
-  summarize(count_iucn = n(),
-            pct_iucn = count_iucn/iucn_n_tot) %>%
-  as.data.frame()
-
-iucn_sp_map <- subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'count_iucn')], 
-                    by = 'loiczid', 
-                    which = 'count_iucn', 
-                    subsWithNA = TRUE)
-
-### Plot both maps
-myTheme <- rasterTheme(region = brewer.pal('Blues', n = 9))
-
-levelplot(am_sp_map, par.settings = myTheme, 
-          main = 'Aquamaps Dataset', 
-          colorkey = list( at = seq(0, 1063, length = 15)))
-
-levelplot(iucn_sp_map, par.settings = myTheme, 
-          main = 'IUCN Dataset', 
-          colorkey = list(at = seq(0, 1063, length = 15)))
-
-```
-
-Now look at all aquamaps species
-
-```{r all_aquamaps, eval = FALSE, echo = FALSE}
-
-am_spp_cells_big <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/ohi_hcaf_species_native.csv')) %>%
-  as.data.frame()
-### am_spp_cell_big is missing loiczid info so bring in hcaf to match
-
-hcaf <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/hcaf.csv')) %>% #bring this in just to get loiczid
-  as.data.frame() %>%
-  select(CsquareCode, LOICZID) %>%
-  rename(loiczid = LOICZID)
-
-### find total spp to allow for percent calcs later
-am_all_n_tot <- am_spp_cells_big %>% 
-  select(SpeciesID) %>% 
-  unique() %>% nrow()
-
-all_am <- am_spp_cells_big %>%
-  select(SpeciesID, CsquareCode) %>%
-  group_by(CsquareCode) %>%
-  summarize(all_count = n(),
-            all_pct   = all_count/am_all_n_tot) %>%
-  as.data.frame()
-
-all <- all_am %>%
-  left_join(hcaf, by = 'CsquareCode')
-
-all_am_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_count')], 
-                   by = 'loiczid', 
-                   which = 'all_count', 
-                   subsWithNA = TRUE)
-
-levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
-```
-
-![image](figures/all_aquamaps_map.png)
-
-``` {r variations_on_gl_dist_maps, echo = FALSE, eval = FALSE}
-### find distribution by percent of total species
-am_sp_pct_map <- subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'pct_am')], 
-                      by = 'loiczid', 
-                      which = 'pct_am', 
-                      subsWithNA = TRUE)
-iucn_sp_pct_map <- subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'pct_iucn')], 
-                        by = 'loiczid', 
-                        which = 'pct_iucn', 
-                        subsWithNA = TRUE)
-### find difference in percent between the two distribution maps; should be small,
-### if species are well-matched, since it's all the same species
-diff_pct_map    <- iucn_sp_pct_map - am_sp_pct_map
-
-### plot maps with distributions as percentages.  Should it be 0 - 1, or 
-### 0 - max value?
-levelplot(am_sp_pct_map, par.settings = myTheme, 
-          main = 'Aquamaps Dataset pct', 
-          colorkey = list( at = seq(0, 1, length = 15)))
-levelplot(iucn_sp_pct_map, par.settings = myTheme, 
-          main = 'IUCN Dataset pct', 
-          colorkey = list(at = seq(0, 1, length = 15)))
-### this is the difference one. Should it be min value to max value?
-levelplot(diff_pct_map, par.settings = myTheme, 
-          main = 'IUCN pct - AM pct', 
-          colorkey = list(at = seq(-1, 1, length = 15)))
-
-### Do similar, but with the full aquamaps list (should also do this
-### against the full IUCN list...)
-all_am_pct_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_pct')], 
-                   by = 'loiczid', 
-                   which = 'all_pct', 
-                   subsWithNA = TRUE)
-
-diff_pct_map2    <- iucn_sp_pct_map - all_am_pct_map
-levelplot(all_am_pct_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset, pct')
-levelplot(diff_pct_map2, par.settings = myTheme, main = 'iucn (overlap only) - aquamaps (all)')
-
+``` {r, child = 'R/data_explore_global_maps.Rmd', eval = FALSE}
 ```

### 2015-12-12 data_explore_2015_data.Rmd
more on manuscript...

@@
 
-Plotting species from q1 (well-matched) (see data_explore_quadmaps.html for plots)
-```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
+### use data_explore.Rmd to create this file, if it does not yet exist.
 
-x <- 20
-y <- nrow(spp_list_q1)
+data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
+spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
+  filter(!is.na(popn_category)) %>% ### filter out unassessed species
+  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
 
-spp_names_q1 <- spp_list_q1$sciname[seq(1, y, y/x)]
+### generate other lists
 
-for (spp in spp_names_q1) { # i = 1
-  spp_plot <- plot_rangemap(spp)
-  spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
-  spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 1: %s', spp, text_tag))
+spp_all_file <- 'data/spp_all_2015.csv'
+spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
 
-  print(spp_plot)
+iucn_marine_file <- 'data/spp_iucn_marine.csv'
+spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
 
-  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q1_%s.png', str_replace(tolower(spp), ' ', '_'))),
-  #        plot = spp_plot, 
-  #        width = 8, height = 4, dpi = 200)
-}
+iucn_maps_file <- 'data/spp_iucn_maps.csv'
+spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
+  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
+  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
+
+spp_am_file <- 'data/spp_am.csv'
+spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
 
@@
 
-Plotting species from q2 (distribution-matched)
-```{r plot_spp_from_q2, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+``` {r breakdown by taxa, echo = FALSE}
 
-x <- 20
-y <- nrow(spp_list_q2)
+sciname_lookup <- read.csv('sciname_lookups/spp_iucn_all.csv', stringsAsFactors = FALSE)
 
-spp_names_q2 <- spp_list_q2$sciname[seq(1, y, y/x)]
+iucn_maps_class <- spp_iucn_maps %>% 
+  left_join(sciname_lookup %>% 
+              select(sciname, class) %>%
+              unique(), 
+            by = 'sciname')
 
-for (spp in spp_names_q2) { # i = 1
-  spp_plot <- plot_rangemap(spp)
-  spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
-  spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 2: %s', spp, text_tag))
-  print(spp_plot)
+taxa_iucn_all <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
+  filter(!is.na(class)) %>% ### don't count the unknowns
+  mutate(class = str_to_title(class),
+         n_iucn_tot = n()) %>%        ### length of whole maps list
+  group_by(class) %>%
+  summarize(n_iucn_class_all  = n(),
+            p_iucn_class_all  = round((n_iucn_class_all/first(n_iucn_tot) * 100), 2))
 
-  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q2_%s.png', str_replace(tolower(spp), ' ', '_'))),
-  #        plot = spp_plot, 
-  #        width = 8, height = 4, dpi = 200)
-}
+taxa_iucn_solo <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
+  filter(!is.na(class)) %>% ### don't count the unknowns
+  filter(!sciname %in% spp_am$sciname) %>%  ### filter list to IUCN only
+  mutate(class = str_to_title(class),
+         n_iucn_solo = n()) %>%       ### length of IUCN only list
+  group_by(class) %>%
+  summarize(n_iucn_class_solo = n(),
+            p_iucn_class_solo = round((n_iucn_class_solo/first(n_iucn_solo) * 100), 2))
+
+taxa_am_all <- spp_am %>%
+  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
+  mutate(class = str_to_title(class),
+         n_am_tot = n()) %>%        ### length of whole AM maps list
+  group_by(class) %>%
+  summarize(n_am_class_all  = n(),
+            p_am_class_all  = round((n_am_class_all/first(n_am_tot) * 100), 2))
+
+taxa_am_solo <- spp_am %>%
+  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
+  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
+  mutate(class = str_to_title(class),
+         n_am_solo = n()) %>%       ### length of AM only list
+  group_by(class) %>%
+  summarize(n_am_class_solo = n(),
+            p_am_class_solo = round((n_am_class_solo/first(n_am_solo) * 100), 2))
+
+taxa_both <- spp_am %>%
+  filter(sciname %in% spp_iucn_maps$sciname) %>%
+  filter(!class %in% c('Not assigned', 'null')) %>%
+  mutate(class = str_to_title(class),
+         n_both_tot = n()) %>%
+  group_by(class) %>%
+  summarize(n_class_both = n(),
+            p_class_both = round(n_class_both/first(n_both_tot) * 100, 2))
+
+taxa_df <- taxa_am_all %>%
+  full_join(taxa_am_solo,   by = 'class') %>%
+  full_join(taxa_iucn_all,  by = 'class') %>%
+  full_join(taxa_iucn_solo, by = 'class') %>%
+  full_join(taxa_both,      by = 'class') %>%
+  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))
+
+knitr::kable(taxa_df %>%
+               select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both))
+
+### bar charts comparing number of species in each class for each data set
+
+taxa_df_long1 <- taxa_df %>%
+  select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both) %>%
+  gather(dataset, n_spp, AquaMaps:Both)
+bar_taxa1 <- ggplot(taxa_df_long1, aes(x = class, y = n_spp)) +
+  geom_bar(stat = 'identity') +
+  facet_grid(dataset ~ ., scales = 'free_y') +
+  theme(axis.text.x = element_text(angle = 75, hjust = 1))
+print(bar_taxa1)
+
+taxa_df_long2 <- taxa_df %>%
+  select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
+  gather(dataset, n_spp, AquaMaps:Both) %>%
+  group_by(class) %>%
+  mutate(n_total = sum(n_spp, na.rm = TRUE),
+         percent = n_spp/n_total * 100)
+bar_taxa2 <- ggplot(taxa_df_long2, aes(x = class, y = percent, fill = dataset)) +
+  geom_bar(stat = 'identity') +
+  geom_text(aes(label = sprintf('n = %s', n_total), y = .05), hjust = 0, size = 4, angle = 90, hjust = 1) +
+  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
+  labs(x = 'Class', y = 'Percent of mapped species',
+       title = 'Distribution of species maps by taxonomic group (class)')
+print(bar_taxa2)
 
@@
 
-Plotting species from q3 (area-matched)
-```{r plot_spp_from_q3, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
-
-x <- 20
-y <- nrow(spp_list_q3)
-
-spp_names_q3 <- spp_list_q3$sciname[seq(1, y, y/x)]
-
-for (spp in spp_names_q3) { # i = 1
-  spp_plot <- plot_rangemap(spp)
-  spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
-  spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 3: %s', spp, text_tag))
-  print(spp_plot)
-
-  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q3_%s.png', str_replace(tolower(spp), ' ', '_'))),
-  #        plot = spp_plot, 
-  #        width = 8, height = 4, dpi = 200)
-}
-
-```
-
-Plotting species from q4 (poorly matched)
-```{r plot_spp_from_q4, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
-
-x <- 20
-y <- nrow(spp_list_q4)
-
-spp_names_q4 <- spp_list_q4$sciname[seq(1, y, y/x)]
-
-for (spp in spp_names_q4) { # i = 1
-  spp_plot <- plot_rangemap(spp)
-  spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
-  spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 4: %s', spp, text_tag))
-  print(spp_plot)
-
-  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q4_%s.png', str_replace(tolower(spp), ' ', '_'))),
-  #        plot = spp_plot, 
-  #        width = 8, height = 4, dpi = 200)
-}
-
+``` {r child = 'R/data_explore_pt3.Rmd', eval = FALSE}
+# Multiple species maps per quadrant
 ```

### 2015-12-12 ms_outline.Rmd
more on manuscript...

@@
 spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
+  rename(sciname = binomial) %>%
   filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
@@
 ```
+
 ## Main message
@@
 
-###IUCN
+## IUCN species distribution maps
 
@@
 
-IUCN publishes spatial distribution map sets only for taxonomic groups in which at least 90% of the species within the taxon group have been evaluated.
+IUCN publishes spatial distribution map sets only for taxonomic groups whjich have been "comprehensively assessed," i.e.
+in which at least 90% of the species within the taxon group have been evaluated.
 While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
@@
 
-### AquaMaps
+## AquaMaps standardized range maps of marine species
 
@@
 
-### Comparison between AM and IUCN
+## Comparison between AM and IUCN
 
@@
 
-### Overlap between assessed species
-
-| Map set   | n~total~                | n~Red_List~                                                               |
-| :-------: | ----------------------: | ------------------------------------------------------------------------: |
-| AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))`         |
-| IUCN      | `r nrow(spp_iucn_maps)` | `r nrow(spp_iucn_maps %>% filter(sciname %in% spp_iucn_marine$sciname))`  |
-| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))` | `r nrow(spp_map_pairs)`           |
-
-Researchers have two large and useful data sets at their disposal, each with its own utility.
-Being able to use both simultaneously makes many more species available, but beware: here there be challenges.
-
 -----
@@
 
-Focus on IUCN red-list evaluated species (incl or not incl DD)? or all species?
+## Overlap between assessed species
+
+Due to differences in both methodology and intent, there are significant differences in the coverage afforded by the species distribution maps published by AquaMaps and those published by the IUCN.
+This is immediately clear from an examination of the number of species maps offered by each source, and the small proportion of species with maps from both sources.
+
+| Map set   | n~total~                | n~Red_List~                                                       |
+| :-------: | ----------------------: | :---------------------------------------------------------------: | 
+| AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` | 
+| IUCN      | `r nrow(spp_iucn_maps)` | all                                                               | 
+| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))` | all                       |
+
+___Should we break it down by "reviewed" Aquamaps as well?___
+
+The AquaMaps species list includes all species available as of December 2015.
+The IUCN species list includes all species maps available directly from the IUCN Red List website (http://www.iucnredlist.org/technical-documents/spatial-data) as of December 2015, filtered to include only species with habitat listed as "marine."  For the purposes of this analysis, we did not include IUCN range maps for bird species, as those range maps are hosted separately by BirdLife International.
 
@@
 
-Examine the overlapping and non-overlapping species by taxonomic group (order? family?) - are there differences in taxonomic group coverage between the two data sets?
+![](figures/bar_by_taxa_all.png)
 
-IUCN - translate the taxonomic group into families - is there a way to automate this? http://www.itis.gov/downloads/?
-AM - access taxa group information directly from AM species list - order? family?
+![](figures/bar_by_taxa_assessed.png)
 
-``` {r breakdown by taxa, echo = FALSE}
+As IUCN only releases map sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed, spatial data for many taxonomic classes are not available, and within a class, the assessed sub-groups may not represent the entire class.
+For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes.
+However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
 
-sciname_lookup <- read.csv('sciname_lookups/scinames.csv', stringsAsFactors = FALSE)
-
-iucn_maps_class <- spp_iucn_maps %>% 
-  ### add sciname list, then jump through hoops to delete dupes
-  ### NOTE: this also ditches ones with no sciname match...
-  separate(sciname, 'genus', sep = ' ', remove = FALSE, extra = 'drop') %>%
-  mutate(genus = tolower(genus)) %>%
-  left_join(sciname_lookup, by = 'genus') %>% 
-  filter(!(str_detect(tolower(spp_group), 'grass')    & !str_detect(kingdom, 'plant'))) %>%
-  filter(!(str_detect(tolower(spp_group), 'mangrove') & !str_detect(kingdom, 'plant'))) %>%
-  filter(!(str_detect(tolower(spp_group), 'coral')    & !str_detect(kingdom, 'animal'))) %>%
-  filter(!(str_detect(tolower(spp_group), 'reptile')  & !str_detect(kingdom, 'animal'))) %>%
-  filter(!(str_detect(tolower(spp_group), 'seabream') & !str_detect(kingdom, 'animal'))) %>%
-  filter(!(str_detect(tolower(spp_group), 'wrass')    & !str_detect(kingdom, 'animal'))) %>%
-  filter(!(str_detect(tolower(spp_group), 'mamm')     & !str_detect(kingdom, 'animal'))) %>%
-  filter(!is.na(phylum)) %>%
-  select(-kingdom, -phylum, -family) %>%
-  unique()
-
-taxa_iucn_all <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
-  filter(!is.na(class)) %>% ### don't count the unknowns
-  mutate(class = str_to_title(class),
-         n_iucn_tot = n()) %>%        ### length of whole maps list
-  group_by(class) %>%
-  summarize(n_iucn_class_all  = n(),
-            p_iucn_class_all  = round((n_iucn_class_all/first(n_iucn_tot) * 100), 2))
-
-taxa_iucn_solo <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
-  filter(!is.na(class)) %>% ### don't count the unknowns
-  filter(!sciname %in% spp_am$sciname) %>%  ### filter list to IUCN only
-  mutate(class = str_to_title(class),
-         n_iucn_solo = n()) %>%       ### length of IUCN only list
-  group_by(class) %>%
-  summarize(n_iucn_class_solo = n(),
-            p_iucn_class_solo = round((n_iucn_class_solo/first(n_iucn_solo) * 100), 2))
-
-taxa_am_all <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  mutate(class = str_to_title(class),
-         n_am_tot = n()) %>%        ### length of whole AM maps list
-  group_by(class) %>%
-  summarize(n_am_class_all  = n(),
-            p_am_class_all  = round((n_am_class_all/first(n_am_tot) * 100), 2))
-
-taxa_am_solo <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
-  mutate(class = str_to_title(class),
-         n_am_solo = n()) %>%       ### length of AM only list
-  group_by(class) %>%
-  summarize(n_am_class_solo = n(),
-            p_am_class_solo = round((n_am_class_solo/first(n_am_solo) * 100), 2))
-
-taxa_both <- spp_am %>%
-  filter(sciname %in% spp_iucn_maps$sciname) %>%
-  filter(!class %in% c('Not assigned', 'null')) %>%
-  mutate(class = str_to_title(class),
-         n_both_tot = n()) %>%
-  group_by(class) %>%
-  summarize(n_class_both = n(),
-            p_class_both = round(n_class_both/first(n_both_tot) * 100, 2))
-
-taxa_df <- taxa_am_all %>%
-  full_join(taxa_am_solo,   by = 'class') %>%
-  full_join(taxa_iucn_all,  by = 'class') %>%
-  full_join(taxa_iucn_solo, by = 'class') %>%
-  full_join(taxa_both,      by = 'class') %>%
-  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))
-
-knitr::kable(taxa_df %>%
-               select(class, am_perc = p_am_class_all, iucn_perc = p_iucn_class_all, both_perc = p_class_both))
-```
+The release of AquaMaps distribution maps is not limited in the same way, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
+___How do they choose their species to assess?___
 
@@
 
-Using predifined species groups by the IUCN, it is clear to see that tropical species - specifically wrasses, corals, blennies and groupers, make up the majority of species found in both.
+Using predefined species groups by the IUCN, it is clear to see that tropical species - specifically wrasses, corals, blennies and groupers, make up the majority of species found in both.
 This is driven more by the IUCN dataset since the number of species with spatial data is directly attributed to the species that have assessments.
@@
 
-_Total species occurrences for each data set: all AM, all IUCN, and overlap (one? the other? occurrences of both?)_
+___Total species occurrences for each data set: all AM, all IUCN, and overlap (one? the other? occurrences of both?)___
 
@@
 _Relevance of this_ - unequal distribution of assessed species - strengths and weaknesses of each; benefits of using both maybe?  save for discussion
+
+Each data set offers spatial distribution information for large numbers of species.
+However, the data sets vary in terms of taxonomic coverage and regional coverage.
+For spatial assessments of biodiversity, the choice of one data set over the other is likely to create significantly different results.
+For studies confined to a narrow range of taxa or to a narrow spatial scale, one data set may offer an advantage over the other in the number of species maps available.
+For global scale biodiversity studies, however, the selection of one data set over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
+___find a reference that describes what makes a "good" data set for global biodiversity, e.g.
+OHI___
 
@@
 * For AquaMaps data, we defined presence as any cell with non-zero probability ___(or RES?)___.  
-* In both cases, these criteria for presence will maximize the number of cells included.  Using these criteria, we found that for 
-`r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`%
-of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
+* In both cases, these criteria for presence will maximize the number of cells included.
+
+Using these criteria, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
 
@@
 area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc > dist_median  & spp_map_pairs$area_ratio >  area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <= dist_median & spp_map_pairs$area_ratio >  area_median)
-q3_count <- sum(spp_map_pairs$sm_perc > dist_median  & spp_map_pairs$area_ratio <= area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <= dist_median & spp_map_pairs$area_ratio <= area_median)
+q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
+q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
+q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
+q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
 ```
@@
 | :----------------------: | :------------------: | :------------------: |
-| q1: Well-aligned         |  > `r dist_median` % |  > `r area_median` % |
-| q2: Distribution-aligned |  < `r dist_median` % |  > `r area_median` % |
-| q3: Area-aligned         |  > `r dist_median` % |  < `r area_median` % |
+| q1: Well-aligned         | >= `r dist_median` % | >= `r area_median` % |
+| q2: Distribution-aligned |  < `r dist_median` % | >= `r area_median` % |
+| q3: Area-aligned         | >= `r dist_median` % |  < `r area_median` % |
 | q4: Poorly-aligned       |  < `r dist_median` % |  < `r area_median` % |

### 2015-12-12 R/data_explore_global_maps.Rmd
created all-IUCN level plot

@@
 am_spp_cells_big <- fread(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) %>%
-  as.data.frame()
-# ### am_spp_cell_big is missing loiczid info so bring in hcaf to match
-# 
-# hcaf <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/hcaf.csv')) %>% #bring this in just to get loiczid
-#   as.data.frame() %>%
-#   dplyr::select(CsquareCode, LOICZID) %>%
-#   rename(loiczid = LOICZID)
+  as.data.frame(stringsAsFactors = FALSE)
 
 ### find total spp to allow for percent calcs later
-am_all_n_tot <- am_spp_cells_big %>% 
-  dplyr::select(SpeciesID) %>% 
-  unique() %>% nrow()
+# am_all_n_tot <- am_spp_cells_big %>% 
+#   dplyr::select(SpeciesID) %>% 
+#   unique() %>% nrow()
 
 all_am <- am_spp_cells_big %>%
-  dplyr::select(SpeciesID, CsquareCode) %>%
-  group_by(CsquareCode) %>%
-  summarize(all_count = n(),
-            all_pct   = all_count/am_all_n_tot) %>%
+  group_by(loiczid) %>%
+  summarize(all_count = n()) %>%
+#   summarize(all_count = n(),
+#             all_pct   = all_count/am_all_n_tot) %>%
   as.data.frame()
 
-all <- all_am %>%
-  left_join(hcaf, by = 'CsquareCode')
-
-all_am_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_count')], 
+all_am_map <- subs(loiczid_raster, all_am[ , c('loiczid', 'all_count')], 
                    by = 'loiczid', 
@@
 levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
+```
+
+Now look at full IUCN species
+
+```{r all_aquamaps, eval = FALSE, echo = FALSE}
+
+iucn_cells_file    <- file.path(dir_data_iucn, 'iucn_cells_2015.csv')
+iucn_spp_cells_all <- fread(iucn_cells_file) %>%
+      as.data.frame(stringsAsFactors = FALSE)
+
+### find total spp to allow for percent calcs later
+# am_all_n_tot <- am_spp_cells_big %>% 
+#   dplyr::select(SpeciesID) %>% 
+#   unique() %>% nrow()
+
+all_iucn <- iucn_spp_cells_all %>%
+  group_by(loiczid) %>%
+  summarize(all_count = n()) %>%
+#   summarize(all_count = n(),
+#             all_pct   = all_count/am_all_n_tot) %>%
+  as.data.frame()
+
+all_iucn_map <- subs(loiczid_raster, all_iucn[ , c('loiczid', 'all_count')], 
+                   by = 'loiczid', 
+                   which = 'all_count', 
+                   subsWithNA = TRUE)
+
+levelplot(all_iucn_map, par.settings = myTheme, main = 'Entire IUCN Dataset')
 ```

### 2015-12-14 ms_outline.Rmd
added iucn level plot
what changes did I make here?  nooo!!!

@@
 
-___Total species occurrences for each data set: all AM, all IUCN, and overlap (one? the other? occurrences of both?)___
+___N Species by cell for each data set: all AM, all IUCN, and overlap (one? the other? occurrences of both?)___
 
 ![](figures/all_aquamaps_map.png)
+![](figures/all_iucn_map.png)
+
+___Note that these level plots show the mean() of the cell value for each row and each column.___  For a difference plot, maybe there's another function that can show us that?
 
@@
 
-AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to differences, we would expect to see strong similarities between range maps for any given species.
+AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, we would expect to see some spatial correlation.
 

### 2015-12-14 R/data_explore_global_maps.Rmd
testing out AM vs IUCN difference level plot

@@
 ### Number of aquamaps species per cell
+iucn_pair_rast_file <- 'rasters/iucn_pair_rast.tif'
+am_pair_rast_file   <- 'rasters/am_pair_rast.tif'
+
+if(!file.exists(iucn_pair_rast_file) | !file.exists(am_pair_rast_file)) {
   am_n_per_cell <- am_spp_cells %>%
@@
 
+  writeRaster(iucn_sp_map, filename = iucn_pair_rast_file, overwrite = TRUE)
+  writeRaster(am_sp_map, filename = am_pair_rast_file, overwrite = TRUE)
+} else {
+  am_sp_map   <- raster(am_pair_rast_file)
+  iucn_sp_map <- raster(iucn_pair_rast_file)
+}
+
 ### Plot both maps
@@
 
+all_am_rast_file   <- 'rasters/am_all_rast.tif'
+
+if(!file.exists(am_all_rast_file)) {
   am_spp_cells_big <- fread(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) %>%
@@
 
+  writeRaster(all_am_map, filename = am_all_rast_file, overwrite = TRUE)
+
+} else {
+  all_am_map   <- raster(all_am_rast_file)
+}
+
 levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
@@
 
+all_iucn_rast_file   <- 'rasters/am_all_rast.tif'
+
+if(!file.exists(iucn_all_rast_file)) {
   iucn_cells_file    <- file.path(dir_data_iucn, 'iucn_cells_2015.csv')
@@
 
+  writeRaster(all_iucn_map, filename = all_iucn_rast_file, overwrite = TRUE)
+
+} else {
+  all_iucn_map   <- raster(iucn_all_rast_file)
+}
+
 levelplot(all_iucn_map, par.settings = myTheme, main = 'Entire IUCN Dataset')
@@
 ``` {r variations_on_gl_dist_maps, echo = FALSE, eval = FALSE}
-### find distribution by percent of total species
-am_sp_pct_map <- subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'pct_am')], 
-                      by = 'loiczid', 
-                      which = 'pct_am', 
-                      subsWithNA = TRUE)
-iucn_sp_pct_map <- subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'pct_iucn')], 
-                        by = 'loiczid', 
-                        which = 'pct_iucn', 
-                        subsWithNA = TRUE)
-### find difference in percent between the two distribution maps; should be small,
-### if species are well-matched, since it's all the same species
-diff_pct_map    <- iucn_sp_pct_map - am_sp_pct_map
+### rescale values relative to maximum cell value
+
+am_map_max <- max(all_am_map, na.rm = TRUE)
+all_am_pct_map <- all_am_map/am_map_max
+
+iucn_map_max <- max(all_iucn_map, na.rm = TRUE)
+all_iucn_pct_map <- all_iucn_map/iucn_map_max
+
+diff_pct_map    <- all_am_pct_map - all_iucn_pct_map
 

### 2015-12-14 ms_outline.Rmd
testing out AM vs IUCN difference level plot

@@
 output:
+  pdf_document:
+    number_sections: yes
+    toc: no
   html_document:
-    toc: true
-    number_sections: true
-    theme: "spacelab"
-    highlight: "haddock" 
+    highlight: haddock
+    number_sections: yes
+    theme: spacelab
+    toc: yes
 ---
@@
 
+_can we set up a compelling story line here?_
+
 Mapping species abundances and distributions is a critical component to understanding global biodiversity and detecting patterns and trends through time.
@@
 Yet there have been no comprehensive comparisons of these two datasets to identify the differences and provide insight into how using one or the other can influence the outcome of an analysis.
-With impactful conclusions being drawn from these large-scale studies on global biodiversity and conservation, it's imperative that the data used in these analyses are appropriate and the data decisions are informed.
 
-The IUCN Red List of Threatened Species includes more than 19,000 assessed marine species, 
-<!--define what "assessed" means?
-include DD?
-keeping those in for now--> but fewer than 4,500 <!--11975 maps in those shapefiles, but fewer than 4500 correspond to assessed marine species...
-and only 2500 of those overall correpond to AquaMaps species--> of these have associated spatial information.
+_Differences in methodology (model-driven vs expert knowledge-driven); differences in intent (what does it communicate); differences in represented species and regions_
+
+<!--
+The IUCN Red List of Threatened Species includes more than 19,000 assessed marine species, but fewer than 4,500 of these have associated spatial information.
 AquaMaps has developed range maps for over 22,800 species.
@@
 Once produced, each range map maintains a dynamic element that allows individuals to adjust parameters based on their expert knowledge to improve the model-generated map.
+-->
 
-Three comparisons presented here:
+To better understand the similarities and differences between the two data sets, we examined three fundamental questions:
 
-* Differences in overall species representation - by taxa and by spatial distribution - how well does each set of maps represent the global distribution of species, the global distribution of biodiversity
-* alignment between maps for species in common - how well do the maps align, when compared at a species-by-species level
+* How do the two data sets compare in their representation of global spatial distribution of species diversity? 
+* How do the two data sets compare in their representation of species diversity across taxonomic groups?
+* For species represented in both data sets, how closely do the spatial distribution maps align?
+
+We examined the implications of alternate data use scenarios on two large-scale studies of global biodiversity and conservation:  the Ocean Health Index and MPA Gap Analysis study.
+
 * sensitivity analysis to AquaMaps presence threshold - diff reports use diff thresholds; what is the likely effect? is there a "correct" threshold?
     * this seems like a sub-analysis to justify choice of 0% threshold etc.  
-    * ___That graph should be box plots at each threshold, not simply mean and/or median.___
+    * Should this just be in supplemental info?  This could be important though...
+
+<!--
+With impactful conclusions being drawn from these large-scale studies on global biodiversity and conservation, it's imperative that the data used in these analyses are appropriate and the data decisions are informed.
+-->
 
@@
 
-The IUCN presents these maps as 'limits of distibution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed everywhere within the polygon.
+The IUCN presents these maps as 'limits of distribution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed within the polygon.
 While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
 
-IUCN publishes spatial distribution map sets only for taxonomic groups whjich have been "comprehensively assessed," i.e.
-in which at least 90% of the species within the taxon group have been evaluated.
+IUCN publishes spatial distribution map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
+in which at least 90% of the species within the taxonomic group have been evaluated.
 While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
@@
 The AquaMaps species list includes all species available as of December 2015.
-The IUCN species list includes all species maps available directly from the IUCN Red List website (http://www.iucnredlist.org/technical-documents/spatial-data) as of December 2015, filtered to include only species with habitat listed as "marine."  For the purposes of this analysis, we did not include IUCN range maps for bird species, as those range maps are hosted separately by BirdLife International.
+The IUCN species list includes all species maps available directly from the IUCN Red List website (http://www.iucnredlist.org/technical-documents/spatial-data) as of December 2015, filtered to include only species with habitat listed as "marine."  For this analysis, we did not include IUCN range maps for bird species, as those data are hosted separately by BirdLife International.
 
@@
 
-As IUCN only releases map sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed, spatial data for many taxonomic classes are not available, and within a class, the assessed sub-groups may not represent the entire class.
+The IUCN only releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed.
+As such, spatial data for many taxonomic classes are not available, and within a class, the assessed sub-groups may not represent the entire class.
 For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes.
@@
 ___How do they choose their species to assess?___
+
+Some studies of species diversity, such as the Ocean Health Index, require both spatial distribution information and non-spatial information such as extinction risk, population trend, or native/non-native status provided by IUCN Red List assessments.
+While all species represented in IUCN distribution maps have also been evaluated for Red List inclusion, we found that only `r round(nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))/nrow(spp_am) * 100, 1)`% (`r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` of `r nrow(spp_am)`) of AquaMaps species have been evaluated, severely limiting the catalog of available species maps for diversity studies.
 
@@
 
-Using predefined species groups by the IUCN, it is clear to see that tropical species - specifically wrasses, corals, blennies and groupers, make up the majority of species found in both.
+Using predefined taxonomic groups by the IUCN, it is clear to see that tropical species - specifically wrasses, corals, blennies and groupers, make up the majority of species found in both.
 This is driven more by the IUCN dataset since the number of species with spatial data is directly attributed to the species that have assessments.
-___This could be shown using the n = XXX on the species group quadrant breakdown chart too.___
+___This could be shown using the n = XXX on the taxonomic group quadrant breakdown chart too.___
 

### 2015-12-14 R/data_explore_global_maps.Rmd
continuing to work on manuscript.  Added figure showing difference in spatial coverage

@@
 
-if(!file.exists(am_all_rast_file)) {
+if(!file.exists(all_am_rast_file)) {
   am_spp_cells_big <- fread(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) %>%
@@
 
-  writeRaster(all_am_map, filename = am_all_rast_file, overwrite = TRUE)
+  writeRaster(all_am_map, filename = all_am_rast_file, overwrite = TRUE)
 
@@
 
-all_iucn_rast_file   <- 'rasters/am_all_rast.tif'
+all_iucn_rast_file   <- 'rasters/iucn_all_rast.tif'
 
-if(!file.exists(iucn_all_rast_file)) {
+if(!file.exists(all_iucn_rast_file)) {
   iucn_cells_file    <- file.path(dir_data_iucn, 'iucn_cells_2015.csv')
@@
 } else {
-  all_iucn_map   <- raster(iucn_all_rast_file)
+  all_iucn_map   <- raster(all_iucn_rast_file)
 }
@@
 
-am_map_max <- max(all_am_map, na.rm = TRUE)
+am_map_max <- maxValue(all_am_map)
 all_am_pct_map <- all_am_map/am_map_max
 
-iucn_map_max <- max(all_iucn_map, na.rm = TRUE)
+iucn_map_max <- maxValue(all_iucn_map)
 all_iucn_pct_map <- all_iucn_map/iucn_map_max
@@
 ### 0 - max value?
-levelplot(am_sp_pct_map, par.settings = myTheme, 
-          main = 'Aquamaps Dataset pct', 
+levelplot(all_am_pct_map, par.settings = myTheme, 
+          main = 'Aquamaps Dataset percent of max', 
           colorkey = list( at = seq(0, 1, length = 15)))
-levelplot(iucn_sp_pct_map, par.settings = myTheme, 
-          main = 'IUCN Dataset pct', 
+levelplot(all_iucn_pct_map, par.settings = myTheme, 
+          main = 'IUCN Dataset percent of max', 
           colorkey = list(at = seq(0, 1, length = 15)))
@@
 levelplot(diff_pct_map, par.settings = myTheme, 
-          main = 'IUCN pct - AM pct', 
+          main = 'Difference: AquaMaps - IUCN by percent', 
           colorkey = list(at = seq(-1, 1, length = 15)))
 
-### Do similar, but with the full aquamaps list (should also do this
-### against the full IUCN list...)
-all_am_pct_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_pct')], 
-                   by = 'loiczid', 
-                   which = 'all_pct', 
-                   subsWithNA = TRUE)
-
-diff_pct_map2    <- iucn_sp_pct_map - all_am_pct_map
-levelplot(all_am_pct_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset, pct')
-levelplot(diff_pct_map2, par.settings = myTheme, main = 'iucn (overlap only) - aquamaps (all)')
-
 ```

### 2015-12-14 ms_outline.Rmd
continuing to work on manuscript.  Added figure showing difference in spatial coverage

@@
 
-___Note that these level plots show the mean() of the cell value for each row and each column.___  For a difference plot, maybe there's another function that can show us that?
+![](figures/all_am-all_iucn_map.png)
+
+___Note: I believe these level plots show the mean() of the cell value for each row and each column.___  For a difference plot, maybe there's another function that can show us that?
 

### 2015-12-15 ms_outline.Rmd
Still editing it... work in progress!

@@
 
-Some studies of species diversity, such as the Ocean Health Index, require both spatial distribution information and non-spatial information such as extinction risk, population trend, or native/non-native status provided by IUCN Red List assessments.
-While all species represented in IUCN distribution maps have also been evaluated for Red List inclusion, we found that only `r round(nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))/nrow(spp_am) * 100, 1)`% (`r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` of `r nrow(spp_am)`) of AquaMaps species have been evaluated, severely limiting the catalog of available species maps for diversity studies.
+Some global-scale studies of biodiversity, such as the Ocean Health Index, require both spatial distribution information and non-spatial information such as extinction risk, population trend, or native/non-native status provided by IUCN Red List assessments.
+While all species represented in IUCN distribution maps have also been evaluated for Red List inclusion, we found that only `r round(nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))/nrow(spp_am) * 100, 1)`% (`r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` of `r nrow(spp_am)`) of AquaMaps species have been similarly evaluated, severely limiting the catalog of available species maps for studies of species diversity and ecosystem health.
+
+-----
 
 ## Spatial global distribution between data sets
-
-___main messages:___
-
-* IUCN seems to be weighted toward tropics; AM might be more evenly distributed?
-* This difference in distribution might come into play when selecting a data set for a more regionally-focused study
-* ... or for global studies, it might be desirable to use both data sets to get better global distribution?
-* Small number of overlapping species indicates that each offers data on distinct spp
-* This section is really supporting the idea that it is critical to understanding the differences, as both data sets support one another
-
-Using predefined taxonomic groups by the IUCN, it is clear to see that tropical species - specifically wrasses, corals, blennies and groupers, make up the majority of species found in both.
-This is driven more by the IUCN dataset since the number of species with spatial data is directly attributed to the species that have assessments.
-___This could be shown using the n = XXX on the taxonomic group quadrant breakdown chart too.___
-
-___N Species by cell for each data set: all AM, all IUCN, and overlap (one? the other? occurrences of both?)___
 
@@
 
-___Note: I believe these level plots show the mean() of the cell value for each row and each column.___  For a difference plot, maybe there's another function that can show us that?
+The balance of species represented in IUCN spatial data skews toward tropical latitudes compared to the distribution represented in AquaMaps distribution maps.
+(The tropical skew of IUCN species is also evident from an examination of the taxonomic groups included in the IUCN map set; see Fig.
+XXX).
+While both data sets indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps data set shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN data set.
 
-_Relevance of this_ - unequal distribution of assessed species - strengths and weaknesses of each; benefits of using both maybe?  save for discussion
+-----
 
@@
 ___find a reference that describes what makes a "good" data set for global biodiversity, e.g.
-OHI___
+OHI___ _or maybe species richness vs diversity vs "health" or whatnot_
 
@@
 
-___move to discussion?___ We were not surprised to find that map pairs for marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2); many of these species are global-ranging pelagic organisms, with some ranges encompassing nearly the entire ocean.
-Mangroves are also not a surprise, being immobile and easily mapped by remote sensing methods.
+### Mechanistic explanations for some of the variation?
 
-_Map pairs for corals, most of which fall into quadrant 2, ***hypothesis...*** generally agree well in distribution, but in many cases IUCN ranges include buffers that extend much farther from coastlines than AquaMaps modeled ranges, which use seafloor depth as a limiting factor._
+We were not surprised to find that map pairs for marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2); many of these species are global-ranging pelagic organisms, with some ranges encompassing nearly the entire ocean.
+
+_Map pairs for corals, most of which fall into quadrant 2, ***hypothesis...
+check a bunch of maps to see if this is a real trend*** generally agree well in distribution, but in many cases IUCN ranges include buffers that extend much farther from coastlines than AquaMaps modeled ranges, which use seafloor depth as a limiting factor._
 
 ___Other hypotheses/case studies to highlight possible mechanistic differences?___
+
+-----
 
@@
 The biodiversity goal is made up of two parts; Species and Habitats.
-The species subgoal uses the IUCN range map data and threat status to calculate average species status in global ocean regions around the world.
-Spatial range maps are developed by experts for species that are assessed.
-The assessment process happens in phases and only [X] number of marine species have been given spatial range maps.
-AquaMaps has spatial range maps for over 17,000 species.
-For those species that are missing spatial information from IUCN, AquaMaps data was used.
-If a species was assigned an IUCN threat status but did not have spatial range information in either dataset, it was excluded.
+As it is currently calculated, the Species subgoal uses the IUCN range map data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
+For those species that are are represented within the IUCN range maps, AquaMaps distribution maps were used.
+Species with no spatial data in either data set were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
 
 This decision to use IUCN data preferentially to AquaMaps data was made based on the fact that experts are developing the IUCN maps and AquaMaps data is based off a computer model, which in some cases is then consulted by an expert and appropriately altered.
+___is this really the reasoning behind the decision?
+let's get the real reason in here___
 
-Between the two databases, there are 1,923 species that have a range map in both.
 Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
@@
 
-<!--
-![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)
--->
+| Scenario              | Priority data source | AquaMaps presence threshold |
+| --------------------- | :------------------: | :-------------------------: |
+| Scenario 0 (current)  | IUCN                 |  >= 40%                     |
+| Scenario 1            | IUCN                 |   > 0%                      |
+| Scenario 2            | AquaMaps             |  >= 40%                     |
+| Scenario 3            | AquaMaps             |   > 0%                      |
 
-Scenario 1 shows change in score if IUCN maps were still preferentially used over AquaMaps, but the AquaMaps threshold was set to 1% rather than 40% probability of occurrence.
-Scenario 2 shows the change in score if we switched preferences for AquaMaps over IUCN data while keeping the 40% probability of occurrence threshold.
-The final Scenario is the same as scenario 2 but without a probability of occurrence threshold.
+Scenario 1 shows the effect of reducing the threshold for AquaMaps presence.
+This will always increase the apparent range of any AquaMaps species in the analysis.
+The slight decrease in average score suggests increased spatial representation of species at greater risk of extinction.
 
-It is important to remember that although only ___[X (1,923/total number of species)]%___ of all species included in the species goal are found in both datasets, there are still significant differences when electing to choose one dataset over the other.
+Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.
+This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so we could predict an average decrease in range for these overlapping species.
+_The slight bump in mean score indicates that a slight increased representation of low-risk species counterbalances a decreased representation of high-risk species._ ___can we really make any generalizations here?
+not really anything useful without further analysis___
+
+Scenario 3 shows the combination of the two previous scenarios.  The larger decrease in score indicates that the ___hmm gotta think about this a little more - come back to it___
 
@@
 
-___need more info on these options? they both sound interesting though___
-- could talk about MPA gap analysis
+___yeah still gotta do something on this one___
+
+-----
 

### 2015-12-17 data_explore.Rmd
updated draft to send to Ben

@@
 spp_list <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(sm_perc)) ### will ditch the ones where no IUCN polygon exists
+  filter(!is.na(sm_perc)) 
+  ### will ditch the ones where no IUCN polygon exists - or are these species
+  ### whose ranges only occur in cells beyond the LOICZID grid?
+spp_list <- spp_list %>%
+  left_join(read.csv('sciname_lookups/spp_groupnames.csv', stringsAsFactors = FALSE),
+            by = 'spp_group')
 
@@
 
+# spp_rando <- c("Lagocephalus lagocephalus",
+#                "Makaira nigricans",
+#                "Pentacheles laevis",
+#                "Stereomastis suhmi",
+#                "Stereomastis surda") 
+# 
 ### truncate the species names for the representative species
 # spp_list_label <- spp_list %>%
-#   filter(sciname %in% spp_quads) %>%
+#   filter(sciname %in% spp_rando) %>%
 #   separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
 #   mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
-#   select(-g, -s)
+#   dplyr::select(-g, -s)
 
@@
                                            color = popn_category,
-                                           size = popn_category)) + 
+                                           size = lg_area_pct)) + 
   ggtheme_plot + 
@@
        title = 'Area match vs distribution match',
-       color = 'Risk', size = 'Risk')
+       color = 'Risk', size = '% ocean range')
 
@@
 spp_list1 <- spp_list %>%
-  group_by(spp_group) %>%
+  group_by(spp_group_text) %>%
   mutate(n_spp = n()) %>%
@@
 spp_gp_q1 <- spp_list_q1 %>%
-  group_by(spp_group) %>%
+  group_by(spp_group_text) %>%
   summarize(n_spp = first(n_spp), n_spp_q1 = n())
 spp_gp_q2 <- spp_list_q2 %>%
-  group_by(spp_group) %>%
+  group_by(spp_group_text) %>%
   summarize(n_spp = first(n_spp), n_spp_q2 = n())
 spp_gp_q3 <- spp_list_q3 %>%
-  group_by(spp_group) %>%
+  group_by(spp_group_text) %>%
   summarize(n_spp = first(n_spp), n_spp_q3 = n())
 spp_gp_q4 <- spp_list_q4 %>%
-  group_by(spp_group) %>%
+  group_by(spp_group_text) %>%
   summarize(n_spp = first(n_spp), n_spp_q4 = n())
@@
 spp_gp_quadrants <- spp_gp_q1 %>%
-  full_join(spp_gp_q2, by = c('spp_group', 'n_spp')) %>%
-  full_join(spp_gp_q3, by = c('spp_group', 'n_spp')) %>%
-  full_join(spp_gp_q4, by = c('spp_group', 'n_spp')) %>%
+  full_join(spp_gp_q2, by = c('spp_group_text', 'n_spp')) %>%
+  full_join(spp_gp_q3, by = c('spp_group_text', 'n_spp')) %>%
+  full_join(spp_gp_q4, by = c('spp_group_text', 'n_spp')) %>%
   gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
@@
               filter(quad == 'q1') %>%
-              dplyr::select(spp_group, pct_q1 = pct_quad),
-            by = 'spp_group') %>%
+              dplyr::select(spp_group_text, pct_q1 = pct_quad),
+            by = 'spp_group_text') %>%
   mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
@@
 barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
-                                  transform(spp_group = reorder(spp_group, pct_q1)), 
-                                aes(x = spp_group, fill = quad, weight = pct_quad)) +
+                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
+                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
   ggtheme_plot + 
@@
        filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
+       width = 6, height = 4)
+
+```
+
+``` {r examine_risk_quartiles, echo = FALSE, eval = TRUE}
+
+spp_list1 <- spp_list %>%
+  group_by(popn_category) %>%
+  mutate(n_spp = n()) %>%
+  ungroup()
+
+### Create species lists by quadrant
+spp_list_q1 <- spp_list1 %>%
+  filter(area_ratio >= median(area_ratio) & sm_perc >= median(sm_perc))
+spp_list_q2 <- spp_list1 %>%
+  filter(area_ratio < median(area_ratio) & sm_perc >= median(sm_perc))
+spp_list_q3 <- spp_list1 %>%
+  filter(area_ratio >= median(area_ratio) & sm_perc < median(sm_perc))
+spp_list_q4 <- spp_list1 %>%
+  filter(area_ratio < median(area_ratio) & sm_perc < median(sm_perc))
+
+### Summarize quadrant species list by species group (e.g. CORALS1)
+spp_risk_q1 <- spp_list_q1 %>%
+  group_by(popn_category) %>%
+  summarize(n_spp = first(n_spp), n_spp_q1 = n())
+spp_risk_q2 <- spp_list_q2 %>%
+  group_by(popn_category) %>%
+  summarize(n_spp = first(n_spp), n_spp_q2 = n())
+spp_risk_q3 <- spp_list_q3 %>%
+  group_by(popn_category) %>%
+  summarize(n_spp = first(n_spp), n_spp_q3 = n())
+spp_risk_q4 <- spp_list_q4 %>%
+  group_by(popn_category) %>%
+  summarize(n_spp = first(n_spp), n_spp_q4 = n())
+
+### Join species group by quadrant into a single dataframe
+spp_risk_quadrants <- spp_risk_q1 %>%
+  full_join(spp_risk_q2, by = c('popn_category', 'n_spp')) %>%
+  full_join(spp_risk_q3, by = c('popn_category', 'n_spp')) %>%
+  full_join(spp_risk_q4, by = c('popn_category', 'n_spp')) %>%
+  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
+  mutate(quad = str_replace(quad, 'n_spp_', ''),
+         pct_quad = n_quad/n_spp,
+         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
+         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))
+
+spp_risk_quadrants <- spp_risk_quadrants %>%
+  left_join(spp_risk_quadrants %>% 
+              filter(quad == 'q1') %>%
+              dplyr::select(popn_category, pct_q1 = pct_quad),
+            by = 'popn_category') %>%
+  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
+         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
+         quad = str_replace(quad, 'q3', 'q3 area-matched'),
+         quad = str_replace(quad, 'q4', 'q4 poorly matched')) %>%
+  left_join(spp_list %>%
+              dplyr::select(popn_category, cat_rank = category_score) %>%
+              unique() %>%
+              mutate(cat_rank = ifelse(popn_category == 'DD', -1, cat_rank)),
+            by = 'popn_category')
+
+### Plot the bar chart
+barchart_spp_risk_quads <- ggplot(spp_risk_quadrants %>% 
+                                  transform(popn_category = reorder(popn_category, cat_rank)), 
+                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
+  ggtheme_plot + 
+  geom_bar(stat = 'bin', alpha = .5) +
+  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
+                               'q2 dist-matched'   = 'greenyellow', 
+                               'q3 area-matched'   = 'orange', 
+                               'q4 poorly matched' = 'orangered2')) +
+  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 4) +
+  coord_flip() +
+  labs(x = 'Risk category', 
+       y = 'Relative number of species per quadrant', 
+       title = 'Risk category breakdown by quadrant',
+       fill = 'Match quadrant')
+
+print(barchart_spp_risk_quads)
+ggsave(plot = barchart_spp_risk_quads,
+       filename = file.path(dir_fig, 'barchart_spp_risk_quads.png'),
        width = 6, height = 4)

### 2015-12-17 ms_outline2.Rmd
updated draft to send to Ben

--- 2015-12-15 ms_outline.Rmd
@@
 output:
+  word_document: null
   pdf_document:
@@
 
-## Main message
+## Significance
 
-Understanding the differences in marine species range maps from IUCN and AquaMaps, and the implications of using one or both.
-
-<!--
-##Working Abstract  
-* Importance of understanding spatial distribution of species
-    * help identify hotspots for global biodiversity 
-    * priorities for conservation
-    * marine spatial planning
-    * assessing biodiversity 
-    * monitor impacts (human, natural)
-    * global-scale assessments
-    * poor overlap of species lists between the two data sets - sometimes helpful to use both
-* Challenges
-    * difficulty of monitoring marine species in space
-    * even more difficult to monitor over time
-        * can Aquamaps, as a bioclimatic envelope model, be used to predict changes or used as proxy for species shifts if envelopes are known to have shifted?
-__can we use this as trend?
-or does all that count as pressures?
-seems like human factors can drastically change the trend that wouldn't be captured here__
-        * Does that additional data in Aquamaps include predictions for 2050, 2100 (bioclimatic envelopes) - how is that reflected in the species distributions?
-* IUCN vs Aquamaps
-    * spatial distribution efforts range from modeled to expert input
-    * distribution data for 1900+ species found in both data sets
-    * differences in predictions of species presence in both distribution and area
-        * Aquamaps validation and benefits from _Ready et al. 2010_:
-          *  We present and evaluate AquaMaps, a presence-only species distribution modelling system that allows the incorporation of expert knowledge about habitat usage and was designed for maximum output of standardized species range maps at the global scale.
-In the marine environment there is a significant challenge to the production of range maps due to large biases in the amount and location of occurrence data for most species.
-AquaMaps is compared with traditional presence-only species distribution modelling methods to determine the quality of outputs under equivalently automated conditions.
-_The effect of the inclusion of expert knowledge to AquaMaps is also investigated._
-          * Model outputs were tested internally, through data partitioning, and externally against independent survey data to determine the ability of models to predict presence versus absence.
-Models were also tested externally by assessing correlation with independent survey estimates of relative species abundance.
-AquaMaps outputs compare well to the existing methods tested, and inclusion of expert knowledge results in a general improvement in model outputs.
-The transparency, speed and adaptability of the AquaMaps system, as well as the existing online framework which allows expert review to compensate for sampling biases and thus improve model predictions are proposed as additional benefits for public and research use alike.
--->
-
-Species range maps are useful in conservation and management, as well as understanding the spatial distribution of movement and overlap of species in different habitats globally.
-Marine species prove more difficult to map based on their inherent nature of being difficult to observe and track.
-Two main sources for marine species are the IUCN and AquaMaps.
-These two datasets differ in methodology and final products, yet the differences between the two are not well documented _(though Aquamaps, and probably IUCN, each probably have some validation studies)_.
-AquaMaps uses a computer model based on preferential environmental parameters to predict species locations and probability of occurrence, while the IUCN uses expert insight to assign spatial boundaries around known locations of marine species.
-Approximately 1,900 species are found in both datasets and there are very large differences between these maps.
-Depending on data usage, choosing one of these datasets over the other can lead to potentially large differences in results as exemplified here by the Ocean Health Index as an example.
+_This is important because of some reason or another...!!!_
 
@@
 
-_can we set up a compelling story line here?_
+<!--By 2050 ___some threshold year?___ our oceans will __in some way, find a study that predicts big changes?__ be very different from the oceans we enjoy today.--> Climate change, ocean temperature rise, ocean acidification, sea ice loss, human impacts from fishing and introduction of nonnative species are likely to significantly alter the global distribution of marine species - but how well do we really understand the current distribution?
+Understanding species abundances and distributions is a critical component to assessing global biodiversity and detecting patterns and trends through time.
+<!--Many efforts to map terrestrial, freshwater and marine species from local to global scales inform our understanding of taxonomic assemblages, biogeographical limits to species ranges, suitable habitat for species.-->
 
-Mapping species abundances and distributions is a critical component to understanding global biodiversity and detecting patterns and trends through time.
-There have been multiple efforts to map terrestrial, freshwater and marine species from local to global scales.
-These pieces of information inform our understanding of taxonomic assemblages, biogeographical limits to species ranges, suitable habitat for species.
+AquaMaps and IUCN marine species distribution maps are both global-scale spatial data collections that have been used to analyze marine species status (Halpern et al.
+2012, Selig et al.
+2013), global biodiversity patterns and trends (other references), predicting range shifts (Molinos et al.
+2015), and setting conservation priorities (Klein et al.
+2015).
+The two data collections ostensibly describe the same information, but significant differences in methodology, intent, and taxonomic and geographic sampling could lead to dramatically different understandings of our marine ecosystems.
 
-These two spatial datasets have been used to analyze species status globally (Halpern et al.
-2012, Selig et al.
-2013), global biodiversity patterns and trends, predicting range shifts (Molinos et al.
-2015),and setting conservation priorities (Klein et al.
-2015).
-Yet there have been no comprehensive comparisons of these two datasets to identify the differences and provide insight into how using one or the other can influence the outcome of an analysis.
+<!--_Differences in methodology (model-driven vs expert knowledge-driven); differences in intent (what does it communicate); differences in represented species and regions_-->
 
-_Differences in methodology (model-driven vs expert knowledge-driven); differences in intent (what does it communicate); differences in represented species and regions_
+To better understand the similarities and differences between the two data collections, we examined three fundamental questions:
 
-<!--
-The IUCN Red List of Threatened Species includes more than 19,000 assessed marine species, but fewer than 4,500 of these have associated spatial information.
-AquaMaps has developed range maps for over 22,800 species.
-Only 2,200 species are mapped in both datasets.
-This gap in overlapping species between these two datasets further highlights the need for understanding each of these datasets.
-Both IUCN and AquaMaps have gaps in marine species information and a significant portion of available data may be overlooked if limited to just one of these datasets.
+* How do the two spatial data collections compare in their representation of global spatial distribution of species diversity? 
+* How do the two spatial data collections compare in their representation of species diversity across taxonomic groups?
+* For species represented in both collections, how closely do the spatial distribution maps align?
 
-Species range maps provided by AquaMaps (Kaschner et al.
-2013) are model-based, relying on environmental parameters in addition to known occurrences and expert input.
-AquaMaps use information about individual species environmental tolerances for salinity, temperature, depth, productivity, and where applicable, sea ice and distance from land, to create a presence-only gridded range map at a resolution of 0.5 degree cells with cell values ranging from 0 to 1.00.
-These values represent the *relative environmental suitability* for the species, with 0 representing unsuitable habitat and 1.00 being the most suitable habitat.
-Once produced, each range map maintains a dynamic element that allows individuals to adjust parameters based on their expert knowledge to improve the model-generated map.
--->
-
-To better understand the similarities and differences between the two data sets, we examined three fundamental questions:
-
-* How do the two data sets compare in their representation of global spatial distribution of species diversity? 
-* How do the two data sets compare in their representation of species diversity across taxonomic groups?
-* For species represented in both data sets, how closely do the spatial distribution maps align?
-
-We examined the implications of alternate data use scenarios on two large-scale studies of global biodiversity and conservation:  the Ocean Health Index and MPA Gap Analysis study.
-
-* sensitivity analysis to AquaMaps presence threshold - diff reports use diff thresholds; what is the likely effect? is there a "correct" threshold?
-    * this seems like a sub-analysis to justify choice of 0% threshold etc.  
-    * Should this just be in supplemental info?  This could be important though...
-
-<!--
-With impactful conclusions being drawn from these large-scale studies on global biodiversity and conservation, it's imperative that the data used in these analyses are appropriate and the data decisions are informed.
--->
+The Ocean Health Index and MPA Gap Analysis study both provide case studies examined the implications of alternate data use scenarios on two large-scale studies of global biodiversity and conservation.
 
@@
 
-## IUCN species distribution maps
-
 The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.
 Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.
-These maps are typically dependent on observation records from databases ___such as GBIF other examples?___, and IUCN provides general guidelines to facilitate consistency in how the maps are generated.
-
+These maps are typically based on species occurrence records from databases such as ___OBIS___ and ___GBIF___; IUCN provides general guidelines to facilitate consistency in how the maps are generated.
 The IUCN presents these maps as 'limits of distribution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed within the polygon.
@@
 While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN had published species distribution maps for nearly 12,000 marine species across 26 taxonomic groups, though fewer than 4,500 of these have been formally assessed under the IUCN Red List of Threatened Species.
-
-* _use as a reference:_ 
-    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
-    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources
-
-## AquaMaps standardized range maps of marine species
+As of December 2015, IUCN had published species distribution maps for nearly 4500 marine species across 24 taxonomic groups.
+(For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.)
 
@@
 2013) offers species range maps based on modeled relative environmental suitability.
-For each assessed species, data from observations and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, and where applicable, sea ice cover and/or distance to land.
-The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of ___probability of occurrence___ (or ___relative environmental suitability?___ (see Kaschner et al 2006 and 2011)), with cell values ranging from 0 (unsuitable) to 1.00 (highly suitable).
-These maps can then be tweaked by experts altering the parameters and regenerating the modeled map.
-As of December 2015, nearly 23,000 species have been mapped, though only 5,600 of these species have been evaluated for the IUCN Red List.
-___Expert mode to further refine models?___ See (Kaschner et al.
-2013) for more detailed methods.
-
-* ___how is expert insight used?
-how frequently?
-new data set includes "reviewed" field___ expert insight seems to be used to generate the environmental suitability envelope for species, to provide sightings info that use used in training the model, and to exclude species from regions of known absence; but the model generates the maps.
-However, some species maps are reviewed by experts (approved?
-changed?
-what happens in the review)
-
-## Comparison between AM and IUCN
-
-* Differences in intent: capture all observations and be inclusive as possible, vs model suitable habitats
-* Differences in methodology - expert-defined range maps (over-predicting) vs sightings (under-predicting) vs models: 
-from __Kaschner et al 2011__:  _Similar to the trade-offs of different habitat prediction modeling approaches [11], these two methods lie on ___opposite ends of a spectrum from potentially overpredicting expert-derived (range maps) to potentially underpredicting (empirical sighting surveys) range sizes____
-    * are there other sources (non-IUCN and AM sources) that can back this up?
-* Differences in the interpretation: probability/RES, vs. presence "probably within this polygon, but not everywhere within" - apples to oranges.
+For each assessed species, data from occurrence records such as OBIS and GBIF, published species databases such as FishBase, and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, sea ice concentration, and distance to land.
+The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence, with cell values ranging from 0 to 1.00.
+These maps can then be reviewed by experts to fine-tune the parameters to better match known occurrences.
+As of December 2015, AquaMaps current native distribution maps have been produced for nearly 23,000 species, though only 5,600 of these species have been evaluated for the IUCN Red List.
+_See (Kaschner et al.
+2013) for more detailed methods._
 
@@
 
-# Analysis of taxonomic and regional inclusion
+# Methods
 
-## Overlap between assessed species
+## Analysis of taxonomic and regional inclusion
+
+### Overlap between assessed species
 
@@
 
-___Should we break it down by "reviewed" Aquamaps as well?___
+### Taxonomic distribution between data collections
 
-The AquaMaps species list includes all species available as of December 2015.
-The IUCN species list includes all species maps available directly from the IUCN Red List website (http://www.iucnredlist.org/technical-documents/spatial-data) as of December 2015, filtered to include only species with habitat listed as "marine."  For this analysis, we did not include IUCN range maps for bird species, as those data are hosted separately by BirdLife International.
-
-## Taxonomic distribution between data sets
+Studies of species richness and biodiversity depend on sampling a representative cross section of ecosystems, and thus benefit from a broad representation across taxonomic groups.
+To examine the overall taxonomic distribution across the spatial data collections, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
+Since biodiversity studies often rely on both spatial distribution and indicators of species health, we further divided the data by whether species had been evaluated for the IUCN Red List of Threatened Species.
 
@@
 The IUCN only releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed.
-As such, spatial data for many taxonomic classes are not available, and within a class, the assessed sub-groups may not represent the entire class.
-For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes.
+As such, spatial data for many taxonomic classes are unavailable, and within a class, the assessed sub-groups may not represent the entire class.
+For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
 However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
 
-The release of AquaMaps distribution maps is not limited in the same way, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
-___How do they choose their species to assess?___
+The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
 
 Some global-scale studies of biodiversity, such as the Ocean Health Index, require both spatial distribution information and non-spatial information such as extinction risk, population trend, or native/non-native status provided by IUCN Red List assessments.
-While all species represented in IUCN distribution maps have also been evaluated for Red List inclusion, we found that only `r round(nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))/nrow(spp_am) * 100, 1)`% (`r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` of `r nrow(spp_am)`) of AquaMaps species have been similarly evaluated, severely limiting the catalog of available species maps for studies of species diversity and ecosystem health.
+While all species represented in IUCN distribution maps have also been evaluated for Red List inclusion, we found that only `r round(nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))/nrow(spp_am) * 100, 1)`% (`r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` of `r nrow(spp_am)`) of AquaMaps species have been similarly evaluated, severely limiting the catalog of available species maps for studies of biodiversity and ecosystem health.
 
------
-
-## Spatial global distribution between data sets
+Broad taxonomic inclusion is critical for understanding distribution and health of ecosystems at any spatial scale, but global and large regional studies of species diversity also require representative sampling across regions.
+To compare the spatial representation of the two data collections directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
+For the AquaMaps data collection, we determined per-cell species richness by summing all species with non-zero probability of occurrence.
+We represented relative distribution of species richness for each data collection by plotting average species count against latitude and longitude.
+To highlight the differences between the two data collections, we also examined the relative difference in per-cell species richness against latitude and longitude.
 
 ![](figures/all_aquamaps_map.png)
+
 ![](figures/all_iucn_map.png)
 
-![](figures/all_am-all_iucn_map.png)
+The balance of species represented in IUCN spatial data skews toward tropical latitudes compared to the distribution represented in AquaMaps distribution maps.
+This may be driven by the fact that the taxonomic distribution of IUCN distribution maps focuses heavily on coral reef-associated species (see fig.
+XXX).
+While both data collections indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps data collection shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN data collection.
 
-The balance of species represented in IUCN spatial data skews toward tropical latitudes compared to the distribution represented in AquaMaps distribution maps.
-(The tropical skew of IUCN species is also evident from an examination of the taxonomic groups included in the IUCN map set; see Fig.
-XXX).
-While both data sets indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps data set shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN data set.
-
------
-
-Each data set offers spatial distribution information for large numbers of species.
-However, the data sets vary in terms of taxonomic coverage and regional coverage.
-For spatial assessments of biodiversity, the choice of one data set over the other is likely to create significantly different results.
-For studies confined to a narrow range of taxa or to a narrow spatial scale, one data set may offer an advantage over the other in the number of species maps available.
-For global scale biodiversity studies, however, the selection of one data set over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
-___find a reference that describes what makes a "good" data set for global biodiversity, e.g.
+Each data collection offers spatial distribution information for large numbers of species.
+However, the data collections vary in terms of taxonomic coverage and regional coverage.
+For spatial assessments of biodiversity, the choice of one data collection over the other is likely to create significantly different results.
+For studies confined to a narrow range of taxa or to a narrow spatial scale, one data collection may offer an advantage over the other in the number of species maps available.
+For global scale biodiversity studies, however, the selection of one data collection over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
+___find a reference that describes what makes a "good" data collection for global biodiversity, e.g.
 OHI___ _or maybe species richness vs diversity vs "health" or whatnot_
@@
 
-# Analysis of spatial alignment by species
+## Analysis of spatial alignment by species
 
-## Defining spatial alignment between the two data sets
+### Defining spatial alignment between the two data collections
 
-AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, we would expect to see some spatial correlation.
+<!--
+AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, --> For species described in both spatial data collections, we would expect to see some spatial correlation, both in the global pattern of species distribution and the extent of species range.
+Large discrepancies in distribution and range area could indicate species that require further study to consolidate expert knowledge and improve species distribution models.
 
-To facilitate our analysis, we selected the subset of marine species that have range maps in both IUCN (filtering out ranges identified as extinct) and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`), and used genus and species binomials to identify paired maps.
-We then rasterized the IUCN shapefiles to match the Aquamaps half-cell raster.
-We flattened each map by assigning to each cell a simple "presence/absence" value:
+Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
+Assigning "presence" values using the same methodology above, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
 
-* For IUCN data, we defined "presence" within a cell as any non-zero overlap with a range map polygon.
-* For AquaMaps data, we defined presence as any cell with non-zero probability ___(or RES?)___.  
-* In both cases, these criteria for presence will maximize the number of cells included.
-
-Using these criteria, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
-
-Overlaying distribution maps from the two data sets for a given species, we calculated two dimensions of spatial alignment: distribution match, which we defined as the ratio of the smaller range that overlapped with the larger range; and area match, which we defined as the ratio of the smaller range to the larger range.
+Overlaying distribution maps from the two data sets for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the ratio of the smaller range that overlapped with the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
 
@@
 
-For a species whose distribution is well understood, and whose range is described in both data sets, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (distribution alignment) and very similar extent of range (area alignment).
+For a species whose distribution is well understood, and whose range is described in both data collections, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
 
-## Comparing spatial alignment between the two data sets
+-----
 
-For each of the `r nrow(spp_map_pairs)` species whose range is described by both data sets, we calculated the distribution match and the area ratio, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
+For each of the `r nrow(spp_map_pairs)` species whose range is described by both data collections, we calculated the distribution alignment and the area alignment, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
 
@@
 ___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
-
-___ditch the species identifiers here?  or highlight ones that are used in discussion of possible mechanistic explanations for differences?___
 
@@
 
-The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
+___caption?___
+
+*The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
 For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
 (n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
-
-The other quadrants each suggest a different story.
+*The other quadrants each suggest a different story.
 Species map pairs that fall within the upper left quadrant (quadrant 2: well-matched in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
 (n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
-
-The lower right quadrant (quadrant 3: well-matched in area but poorly matched in distribution) includes species for which the two maps generally agree in range area, but disagree in where those ranges occur.
+*The lower right quadrant (quadrant 3: well-matched in area but poorly matched in distribution) includes species for which the two maps generally agree in range area, but disagree in where those ranges occur.
 Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being more problematic than map pairs in either of the upper quadrants.
 (n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
+*The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
+(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
 
-The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
+![](figures/barchart_spp_risk_quads.png)
 
@@
 
-Figure 2 divides the `r nrow(spp_map_pairs)` species map pairs into taxonomic groups as classified by the published IUCN range maps, and examines the breakdown of each taxonomic group by quadrant.
+Breaking down the quadrants by IUCN extinction risk categories, species with higher extinction risk tend to be better matched between the two data collections; this is not surprising, as these at-risk species are likely to face closer expert scrutiny than species ranked as Least Concern.
 
-### Mechanistic explanations for some of the variation?
+When broken down by taxonomic group, we found that certain taxonomic groups were far more likely than others to be spatially well-matched; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2), .
+Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two data collections, predicted areas are not well-matched; AquaMaps accounts for seafloor depth as a limiting factor, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, and thus likely overestimate the actual distribution.
 
-We were not surprised to find that map pairs for marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2); many of these species are global-ranging pelagic organisms, with some ranges encompassing nearly the entire ocean.
-
-_Map pairs for corals, most of which fall into quadrant 2, ***hypothesis...
-check a bunch of maps to see if this is a real trend*** generally agree well in distribution, but in many cases IUCN ranges include buffers that extend much farther from coastlines than AquaMaps modeled ranges, which use seafloor depth as a limiting factor._
-
-___Other hypotheses/case studies to highlight possible mechanistic differences?___
+___Other hypotheses/case studies to highlight other possible mechanistic differences?___
 
@@
 
-# AquaMaps: Effect of changing "presence" threshold on area representation
+## AquaMaps: Effect of changing "presence" threshold on area representation
+
+Why should anyone care about AquaMaps thresholds. ___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
+
+_what would it mean if there were non-linearities or generally linear trend?_
 
@@
 For those species that are are represented within the IUCN range maps, AquaMaps distribution maps were used.
-Species with no spatial data in either data set were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
+Species with no spatial data in either data collection were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
 
@@
 doi:10.1371/journal.pone.0060284.
+
+* _IUCN reference:_ 
+    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
+    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources

### 2015-12-18 R/data_explore_am_thresh.Rmd
adding ribbon plot for threshold

@@
 
-### create dataframe that gives me all species, their total number of cells, 
-### total number of cells >= 0.4 prob, difference between the two and threat
 spp_df3 <- spp_cells %>%
@@
 
-ggsave('../figures/AMThreshold.png', width = 6, height = 4)
+ggsave('figures/AMThreshold.png', width = 6, height = 4)
 
@@
 
-ggsave('../figures/AMThreshold2.png', width = 6, height = 4)
+ggsave('figures/AMThreshold2.png', width = 6, height = 4)
 
@@
   summarize(mean_prop   = mean(proportion, na.rm = TRUE),
-            median_prop = median(proportion, na.rm = TRUE))
+            median_prop = median(proportion, na.rm = TRUE),
+            quant09 = quantile(proportion, .09),
+            quant25 = quantile(proportion, .25),
+            quant75 = quantile(proportion, .75),
+            quant91 = quantile(proportion, .91))
 
@@
 
-ggsave('../figures/scatter_AMThreshold.png', width = 6, height = 4)
+ggsave('figures/scatter_AMThreshold.png', width = 6, height = 4)
 
@@
 ```
+
+```{r boxplot_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
+
+area_mean_model <- lm(mean_prop ~ thresh, data = thresh_df_summary)
+area_median_model  <- lm(median_prop ~ thresh,  data = thresh_df_summary)
+
+ribbon_area_vs_threshold <- ggplot(thresh_df_summary, aes(x = thresh, y = median_prop)) + 
+  ggtheme_plot + 
+#  geom_point(position = position_jitter(w = .1), aes(color = scenario), alpha = .5) +
+
+  geom_ribbon(aes(ymax = quant91, ymin = quant09), alpha = .5, color = NA, fill = 'grey80') +
+  geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = 'grey70') +
+  geom_point(color = 'grey20') + 
+  geom_point(aes(y = mean_prop), color = 'darkred') + 
+  geom_abline(intercept = area_mean_model[['coefficients']][1],
+              slope     = area_mean_model[['coefficients']][2],
+              color = 'red') + 
+  geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
+  geom_vline(x = 0.50, color = 'blue', linetype = 'dashed') +
+  geom_vline(x = 0.60, color = 'blue', linetype = 'dashed') +
+  annotate('text', x = .40, y = .05, hjust = 0, vjust = -.5,
+           angle = 90, size = 2, color = 'grey20',
+           label = 'Ocean Health Index: 40%') + 
+  annotate('text', x = .50, y = .05, hjust = 0, vjust = -.5,
+           angle = 90, size = 2, color = 'grey20',
+           label = 'MPA Gap Analysis: 50%') + 
+  annotate('text', x = .60, y = .05, hjust = 0, vjust = -.5,
+           angle = 90, size = 2, color = 'grey20',
+           label = 'Marine mammals: 60%') + 
+  annotate('text', x = .75, y = .85,
+           size = 2, color = 'grey20',
+           label = 'R^2 = .995') + 
+  ylim(0, 1) +
+  labs(title = 'Relative species range after presence threshold',
+       x = 'Presence threshold cutoff', 
+       y = 'Mean (median) relative species range')
+
+print(ribbon_area_vs_threshold)
+
+ggsave('figures/ribbon_AMThreshold.png', width = 6, height = 4)
+
+```

### 2015-12-19 R/data_explore_am_thresh.Rmd
added mean/median to histograms

@@
 
+spp_df3_mean <- mean(spp_df3$prop, na.rm = TRUE)
+spp_df3_med <- median(spp_df3$prop, na.rm = TRUE)
+
 hist_area_vs_threshold <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
@@
   geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
+  geom_vline(x = spp_df3_mean, color = 'darkred', linetype = 'dashed') + 
+  annotate('text', x = spp_df3_mean, y = .16, hjust = 0, vjust = -.5,
+           angle = 90, size = 2, color = 'grey20',
+           label = 'mean') + 
+  geom_vline(x = spp_df3_med, color = 'grey30', linetype = 'dashed') + 
+  annotate('text', x = spp_df3_med, y = .16, hjust = 0, vjust = -.5,
+           angle = 90, size = 2, color = 'grey20',
+           label = 'median') + 
   scale_fill_manual(values = c('LC' = 'green4', 
@@
 
+print(hist_area_vs_threshold)
+
 ggsave('figures/AMThreshold.png', width = 6, height = 4)
+```
 
-print(hist_area_vs_threshold)
+```{r hist_lost_area_w_threshold_ne_dd_removed, warning = FALSE, message = FALSE, echo = FALSE}
 
@@
 
+spp_df3a_mean <- mean(spp_df3$prop, na.rm = TRUE)
+spp_df3a_med <- median(spp_df3$prop, na.rm = TRUE)
+
 hist_area_vs_threshold2 <- ggplot(spp_df3a, aes(prop, fill = iucn_code)) + 
@@
   geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
+  geom_vline(x = spp_df3a_mean, color = 'darkred', linetype = 'dashed') + 
+  annotate('text', x = spp_df3a_mean, y = .16, hjust = 0, vjust = -.5,
+           angle = 90, size = 2, color = 'grey20',
+           label = 'mean') + 
+  geom_vline(x = spp_df3a_med, color = 'grey30', linetype = 'dashed') + 
+  annotate('text', x = spp_df3a_med, y = .16, hjust = 0, vjust = -.5,
+           angle = 90, size = 2, color = 'grey20',
+           label = 'median') + 
   scale_fill_manual(values = c('LC' = 'green4', 
@@
   geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = 'grey70') +
-  geom_point(color = 'grey20') + 
+  geom_point(color = 'grey30') + 
   geom_point(aes(y = mean_prop), color = 'darkred') + 

### 2015-12-19 data_explore.Rmd
update manuscript and graphs

@@
 
+***
+
+Quadrant map and quad breakdown by taxa and risk
+
 ``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
@@
 
-### define 25% and 75% quartiles for both x and y axes
+### define 25%, 50%, and 75% quartiles for both x and y axes
 area_ratio_quartiles <- quantile(spp_list$area_ratio, na.rm = TRUE)[2:4]
@@
                      ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
-  geom_point() + 
+  geom_point(alpha = .6)
+
+### Manage scales for color and size 
+scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
   scale_colour_manual(values = c('LC' = 'green4', 
@@
                                  'DD' = 'grey30')) + 
-  geom_vline(x = area_ratio_quartiles[c(1, 3)],
-             color = 'grey50', linetype = 'dashed') +
-  geom_hline(y = pct_overlap_quartiles[c(1, 3)],
-             color = 'grey50', linetype = 'dashed') +
+  scale_size_continuous(limits = c(0, 100),
+                        breaks = c( 0,    25,    50,    75,    100), 
+                        labels = c('0%', '25%', '50%', '75%', '100%')) + 
+  scale_x_continuous(expand = c(0, 0), limits = c(0, 100)) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
+
+### here are median labels:
+scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+  annotate(geom = 'text',
+           x = area_ratio_quartiles[2], y = 5,
+           hjust = 0, vjust = 0,
+           color = 'grey50', size = 2,
+           fontface = 'bold.italic', angle = 90,
+           label = sprintf('Median = %s%%', round(area_ratio_quartiles[2], 1))) +
+  annotate(geom = 'text',
+           x = 5, y = pct_overlap_quartiles[2],
+           hjust = 0, vjust = 0,
+           color = 'grey50', size = 2,
+           fontface = 'bold.italic', angle = 0,
+           label = sprintf('Median = %s%%', round(pct_overlap_quartiles[2], 1)))
+
+### here are median lines:
+# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+#   geom_vline(x = area_ratio_quartiles[2],
+#              color = 'grey50', linetype = 'dashed') +
+#   geom_hline(y = pct_overlap_quartiles[2],
+#              color = 'grey50', linetype = 'dashed')
+
+### here are 25% and 75% quantile lines:
+# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+#   geom_vline(x = area_ratio_quartiles[c(1, 3)],
+#              color = 'grey50', linetype = 'dashed') +
+#   geom_hline(y = pct_overlap_quartiles[c(1, 3)],
+#              color = 'grey50', linetype = 'dashed')
+
+### here are individual species, identified and named on the plot:
+# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
 #   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
@@
 #             hjust = .25,
-#             vjust = 1.5) +
-  labs(x = 'Area match: \n(range area smaller : range area larger) * 100%', 
-       y = 'Distribution match: \n % overlap of smaller range w/ larger', 
-       title = 'Area match vs distribution match',
+#             vjust = 1.5)
+
+scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+  labs(x = bquote('Area alignment '*alpha['area']*' (%)'), 
+       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
+       title = 'Distribution alignment vs area alignment',
        color = 'Risk', size = '% ocean range')
+
+print(scatter_arearatio_vs_overlap)
 
@@
 
-print(scatter_arearatio_vs_overlap)
-
 ```
 
-```{r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+``` {r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
                                'q4 poorly matched' = 'orangered2')) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 4) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
+  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 2, color = 'grey20') +
   coord_flip() +
-  labs(x = 'Species Group', 
-       y = 'Relative number of species per quadrant', 
-       title = 'Species group breakdown by quadrant',
-       fill = 'Match quadrant')
+  labs(x = 'Taxonomic Group', 
+       y = 'Relative proportion of species per quadrant', 
+       title = 'Taxonomic group by quadrant',
+       fill = 'Alignment')
 
@@
                                'q4 poorly matched' = 'orangered2')) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 4) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
+  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 2, color = 'grey20') +
   coord_flip() +
   labs(x = 'Risk category', 
-       y = 'Relative number of species per quadrant', 
-       title = 'Risk category breakdown by quadrant',
-       fill = 'Match quadrant')
+       y = 'Relative proportion of species per quadrant', 
+       title = 'Risk category by quadrant',
+       fill = 'Alignment')
 
@@
        filename = file.path(dir_fig, 'barchart_spp_risk_quads.png'),
-       width = 6, height = 4)
+       width = 6, height = 1.5)
 
@@
   left_join(sciname_lookup %>% 
-              select(sciname, class) %>%
+              dplyr::select(sciname, class) %>%
               unique(), 
@@
 knitr::kable(taxa_df %>%
-               select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both))
+               dplyr::select(class, 
+                             AquaMaps = n_am_class_all, 
+                             IUCN = n_iucn_class_all, 
+                             Both = n_class_both))
 ```
@@
 taxa_df_long1 <- taxa_df %>%
-  select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both) %>%
+  dplyr::select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both) %>%
   gather(dataset, n_spp, AquaMaps:Both)
@@
 taxa_df_long2 <- taxa_df %>%
-  select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
+  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
   gather(dataset, n_spp, AquaMaps:Both) %>%
@@
   scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
-  geom_text(aes(label = sprintf('n = %s', n_total), y = .05), hjust = 0, size = 4, angle = 90, hjust = 1) +
+  geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
+            size = 3, angle = 90, hjust = 0, color = 'grey30') +
   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
   labs(x = 'Class', y = 'Percent of mapped species',
@@
 taxa_df_long3 <- taxa_df %>%
-  select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
+  dplyr::select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
   gather(dataset, n_spp, AquaMaps:Both) %>%
@@
   scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
-  geom_text(aes(label = sprintf('n = %s', n_total), y = .05), hjust = 0, size = 4, angle = 90, hjust = 1) +
+  geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
+            size = 3, angle = 90, hjust = 0, color = 'grey30') +
   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
   labs(x = 'Class', y = 'Percent of mapped and assessed species',

### 2015-12-19 ms_outline2.Rmd
update manuscript and graphs

@@
 ---
-title: "A comparative analysis of AquaMaps and IUCN species distribution maps"
-author: "JA, CO, BH"
+title: "Aligning species range data to better serve science and conservation"
+author: "Casey O'Hara, Jamie Afflerbach, Ben Halpern"
 output:
-  word_document: null
   pdf_document:
@@
     toc: yes
+  word_document: null
 ---
@@
 
+## Abstract
+
+_get to this later..._
+
+-----
+
 ## Significance
 
-_This is important because of some reason or another...!!!_
+_subset of the abstract_
 
@@
 
-<!--By 2050 ___some threshold year?___ our oceans will __in some way, find a study that predicts big changes?__ be very different from the oceans we enjoy today.--> Climate change, ocean temperature rise, ocean acidification, sea ice loss, human impacts from fishing and introduction of nonnative species are likely to significantly alter the global distribution of marine species - but how well do we really understand the current distribution?
-Understanding species abundances and distributions is a critical component to assessing global biodiversity and detecting patterns and trends through time.
-<!--Many efforts to map terrestrial, freshwater and marine species from local to global scales inform our understanding of taxonomic assemblages, biogeographical limits to species ranges, suitable habitat for species.-->
+Mapping and predicting species’ distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
+Knowing where individuals of a species are, and what allows them to persist there, provides foundational information for understanding species ranges, how best to protect and manage species, and how they may respond to increasing human impacts and a changing climate.
+A rich literature exists tackling the many dimensions of these questions.
 
-AquaMaps and IUCN marine species distribution maps are both global-scale spatial data collections that have been used to analyze marine species status (Halpern et al.
+One very important outcome of this body of science is the various compiled databases of species range maps.
+In the oceans there are now two such global repositories – Aquamaps (REF) and International Union for Conservation of Nature (IUCN) species distribution maps (REF) – that are used for a wide range of purposes, including assessing marine species status (Halpern et al.
 2012, Selig et al.
-2013), global biodiversity patterns and trends (other references), predicting range shifts (Molinos et al.
+2013), evaluating global biodiversity patterns and trends (REF), predicting range shifts (Molinos et al.
 2015), and setting conservation priorities (Klein et al.
 2015).
-The two data collections ostensibly describe the same information, but significant differences in methodology, intent, and taxonomic and geographic sampling could lead to dramatically different understandings of our marine ecosystems.
+The two data collections ostensibly describe the same information, but significant differences in methodology, intent, and taxonomic and geographic sampling could lead to dramatically different understandings of our marine ecosystems and resulting policy and conservation recommendations
 
-<!--_Differences in methodology (model-driven vs expert knowledge-driven); differences in intent (what does it communicate); differences in represented species and regions_-->
+_Importantly, biases in taxanomic or spatial coverage of a dataset could shift management or conservation actions towards places or species that aren’t actually the most in need. [elaborate a bit?]_
 
-To better understand the similarities and differences between the two data collections, we examined three fundamental questions:
+To understand the implications of similarities and differences between the two datasets, we examined how they compare in their representation of global spatial distribution of species, and for the relatively small number of species represented in both datasets, how closely their spatial distribution maps align.
+We then reanalyze two recent studies using these datasets based on different versions of the data  to illustrate how and why decisions about which species range data to use affect our understanding of the status of marine biodiversity.
 
-* How do the two spatial data collections compare in their representation of global spatial distribution of species diversity? 
-* How do the two spatial data collections compare in their representation of species diversity across taxonomic groups?
-* For species represented in both collections, how closely do the spatial distribution maps align?
+# Overview of the datasets
 
-The Ocean Health Index and MPA Gap Analysis study both provide case studies examined the implications of alternate data use scenarios on two large-scale studies of global biodiversity and conservation.
+___We may need to shorten this section to fit this paper into the format guidelines for PNAS.  If so, some of this can be pushed to the Suppl. Materials. - BH___
 
-# Understanding each dataset separately
-
-The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.
+The IUCN publishes species distribution maps as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.
 Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.
-These maps are typically based on species occurrence records from databases such as ___OBIS___ and ___GBIF___; IUCN provides general guidelines to facilitate consistency in how the maps are generated.
+These maps are typically based on species occurrence records from databases such as the Ocean Biogeographic Information System (OBIS) (REF) and the Global Biodiversity Information Facility (GBIF) (REF); IUCN provides _general guidelines_ ___Do we need to describe these?
+- BH___ to facilitate consistency in how the maps are generated.
 The IUCN presents these maps as 'limits of distribution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed within the polygon.
@@
 While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN had published species distribution maps for nearly 4500 marine species across 24 taxonomic groups.
-(For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.)
+As of December 2015, IUCN had published species distribution maps for `r nrow(spp_iucn_maps)` marine species across 24 taxonomic groups.
+For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.
 
@@
 For each assessed species, data from occurrence records such as OBIS and GBIF, published species databases such as FishBase, and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, sea ice concentration, and distance to land.
-The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence, with cell values ranging from 0 to 1.00.
+The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence, with cell values ranging from 0.00 to 1.00.
 These maps can then be reviewed by experts to fine-tune the parameters to better match known occurrences.
-As of December 2015, AquaMaps current native distribution maps have been produced for nearly 23,000 species, though only 5,600 of these species have been evaluated for the IUCN Red List.
-_See (Kaschner et al.
-2013) for more detailed methods._
+As of December 2015, AquaMaps current native distribution maps have been produced for `r nrow(spp_am)` species.
+Use of these data to predict species presence or absence requires decisions about which probability threshold to use.
 
@@
 
-# Methods
+_Analyses [briefly describe the analyses done]_
 
-## Analysis of taxonomic and regional inclusion
+* _simple taxanomic comparison (describe very briefly)_
+* _spatial overlap comparison (describe very briefly)_
+* _redo of OHI biodiversity goal (describe briefly what was done previously, and how that was changed for here)_
+* _redo global MPA gap analysis (describe briefly what was done previously, and how that was changed for here)_
+
+# Results and Discussion
 
@@
 
-Due to differences in both methodology and intent, there are significant differences in the coverage afforded by the species distribution maps published by AquaMaps and those published by the IUCN.
-This is immediately clear from an examination of the number of species maps offered by each source, and the small proportion of species with maps from both sources.
+``` {r calc species numbers, echo = FALSE}
+n_spp_am    <- nrow(spp_am)
+n_spp_am_rl <- nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))
+p_spp_am_rl <- round(n_spp_am_rl/n_spp_am * 100, 1)
+n_spp_iucn  <- nrow(spp_iucn_maps)
+n_spp_both  <- nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))
+p_both_am   <- round(n_spp_both/n_spp_am, 1)
+p_both_iucn <- round(n_spp_both/n_spp_iucn, 1)
+```
 
+Due to differences in both methodology and intent, significant differences exist between AquaMaps and IUCN species distribution maps (Fig.
+1).
+Only `r n_spp_both` species were covered by both datasets (`r p_both_am`% of total Aquamaps species; `p_both_iucn`% of total IUCN species; Fig.
+1, 2).
+__All IUCN-mapped species are also included in Red List species, but only `r p_spp_am_rl`% of AquaMaps species.__
+
+<!--
 | Map set   | n~total~       | n~Red_List~     |
-| :-------: | ----------------------: | :---------------------------------------------------------------: | 
-| AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` | 
-| IUCN      | `r nrow(spp_iucn_maps)` | all                                                               | 
-| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))` | all                       |
+| :-------: | -------------: | :-------------: | 
+| AquaMaps  | `r n_spp_am`   | `r n_spp_am_rl` | 
+| IUCN      | `r n_spp_iucn` | all             | 
+| Both      | `r n_spp_both` | all             |
+-->
 
-### Taxonomic distribution between data collections
+### Taxonomic distribution between datasets
 
-Studies of species richness and biodiversity depend on sampling a representative cross section of ecosystems, and thus benefit from a broad representation across taxonomic groups.
-To examine the overall taxonomic distribution across the spatial data collections, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
-Since biodiversity studies often rely on both spatial distribution and indicators of species health, we further divided the data by whether species had been evaluated for the IUCN Red List of Threatened Species.
+___Either move to brief methods section above, and maybe even shorten - BH___: To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
+We further divided the data by whether species had been evaluated for the IUCN Red List of Threatened Species.
 
-![](figures/bar_by_taxa_all.png)
-
-![](figures/bar_by_taxa_assessed.png)
-
-The IUCN only releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed.
+___Either cut or move to data description section of methods above.
+- BH___: The IUCN only releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed.
 As such, spatial data for many taxonomic classes are unavailable, and within a class, the assessed sub-groups may not represent the entire class.
-For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
+___Move to supplementary materials.
+- BH___: For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
 However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
 
-The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
+___Probably just cut, but maybe include in Suppl Materials.
+- BH___: The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
 
-Some global-scale studies of biodiversity, such as the Ocean Health Index, require both spatial distribution information and non-spatial information such as extinction risk, population trend, or native/non-native status provided by IUCN Red List assessments.
-While all species represented in IUCN distribution maps have also been evaluated for Red List inclusion, we found that only `r round(nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))/nrow(spp_am) * 100, 1)`% (`r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` of `r nrow(spp_am)`) of AquaMaps species have been similarly evaluated, severely limiting the catalog of available species maps for studies of biodiversity and ecosystem health.
-
-Broad taxonomic inclusion is critical for understanding distribution and health of ecosystems at any spatial scale, but global and large regional studies of species diversity also require representative sampling across regions.
-To compare the spatial representation of the two data collections directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
-For the AquaMaps data collection, we determined per-cell species richness by summing all species with non-zero probability of occurrence.
-We represented relative distribution of species richness for each data collection by plotting average species count against latitude and longitude.
-To highlight the differences between the two data collections, we also examined the relative difference in per-cell species richness against latitude and longitude.
-
-![](figures/all_aquamaps_map.png)
-
-![](figures/all_iucn_map.png)
+___Move to methods section above.
+- BH___: To compare the spatial representation of the two datasets directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
+For the AquaMaps dataset, we determined per-cell species richness by summing all species with non-zero probability of occurrence.
+We represented relative distribution of species richness for each dataset by plotting average species count against latitude and longitude.
+To highlight the differences between the two datasets, we also examined the relative difference in per-cell species richness against latitude and longitude.
 
@@
 XXX).
-While both data collections indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps data collection shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN data collection.
+While both datasets indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps dataset shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN dataset.
 
-Each data collection offers spatial distribution information for large numbers of species.
-However, the data collections vary in terms of taxonomic coverage and regional coverage.
-For spatial assessments of biodiversity, the choice of one data collection over the other is likely to create significantly different results.
-For studies confined to a narrow range of taxa or to a narrow spatial scale, one data collection may offer an advantage over the other in the number of species maps available.
-For global scale biodiversity studies, however, the selection of one data collection over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
-___find a reference that describes what makes a "good" data collection for global biodiversity, e.g.
+Each dataset offers spatial distribution information for large numbers of species.
+However, the datasets vary in terms of taxonomic coverage and regional coverage.
+For spatial assessments of biodiversity, the choice of one dataset over the other is likely to create significantly different results.
+For studies confined to a narrow range of taxa or to a narrow spatial scale, one dataset may offer an advantage over the other in the number of species maps available.
+For global scale biodiversity studies, however, the selection of one dataset over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
+___find a reference that describes what makes a "good" dataset for global biodiversity, e.g.
 OHI___ _or maybe species richness vs diversity vs "health" or whatnot_
@@
 
-### Defining spatial alignment between the two data collections
+### Defining spatial alignment between the two datasets
 
 <!--
-AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, --> For species described in both spatial data collections, we would expect to see some spatial correlation, both in the global pattern of species distribution and the extent of species range.
-Large discrepancies in distribution and range area could indicate species that require further study to consolidate expert knowledge and improve species distribution models.
+AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, --> For species described in both spatial datasets, we would expect to see spatial correlation both in the global pattern of species distribution and the extent of species range.
+Large discrepancies in distribution and range area could indicate species that require further study to consolidate expert knowledge for IUCN and improve the inputs to AquaMaps species distribution models.
 
 Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
-Assigning "presence" values using the same methodology above, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
+Assigning "presence" values using the same methodology above, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of paired-map species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
+This aligns with the expectations and early analysis of Kaschner et al (REF).
 
-Overlaying distribution maps from the two data sets for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the ratio of the smaller range that overlapped with the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
+Overlaying paired distribution maps for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
 
-$$Match_{dist} = \frac{A_{overlap}}{A_{larger}} * 100\%$$
+$$\alpha_{dist} = \frac{A_{small \cap large}}{A_{large}} * 100\%$$
 
-$$Match_{area} = \frac{A_{smaller}}{A_{larger}} * 100\%$$
+$$\alpha_{area} = \frac{A_{small}}{A_{large}} * 100\%$$
 
-For a species whose distribution is well understood, and whose range is described in both data collections, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
+For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
 
@@
 
-For each of the `r nrow(spp_map_pairs)` species whose range is described by both data collections, we calculated the distribution alignment and the area alignment, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
-
-![](figures/AreaRatiovsPercOverlap.png)
-
-___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
+For each of the `r nrow(spp_map_pairs)` species whose range is described by both datasets, we calculated the distribution alignment and the area alignment, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
 
@@
 
-| Description              |  Distribution match  |     Area match       |
-| :----------------------: | :------------------: | :------------------: |
-| q1: Well-aligned         | >= `r dist_median` % | >= `r area_median` % |
-| q2: Distribution-aligned |  < `r dist_median` % | >= `r area_median` % |
-| q3: Area-aligned         | >= `r dist_median` % |  < `r area_median` % |
-| q4: Poorly-aligned       |  < `r dist_median` % |  < `r area_median` % |
-
-___caption?___
-
-*The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
+* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
 For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
 (n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The other quadrants each suggest a different story.
-Species map pairs that fall within the upper left quadrant (quadrant 2: well-matched in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
+* The other quadrants each suggest a different story.
+Species map pairs that fall within the upper left quadrant (quadrant 2: well-aligned in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
 (n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The lower right quadrant (quadrant 3: well-matched in area but poorly matched in distribution) includes species for which the two maps generally agree in range area, but disagree in where those ranges occur.
-Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being more problematic than map pairs in either of the upper quadrants.
+* The lower right quadrant (quadrant 3: well-aligned in area but poorly matched in distribution) includes species for which the paired maps generally agree in range area, but disagree in locations of those ranges.
+Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being lower quality alignment than map pairs in either of the upper quadrants.
 (n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
+* The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
 (n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
 
-![](figures/barchart_spp_risk_quads.png)
+Breaking down the quadrants by IUCN extinction risk categories, species with higher extinction risk tend to be better aligned between the two datasets.
+Well-aligned range map pairs (membership in quadrants 1 and 2) may suggest increased expert scrutiny, though this could be explained as higher perceived risk leading to increased attention, or as increased attention revealing deeper risk.
+Likely both mechanisms are at play in a feedback loop.
+In any case, poorly-aligned rangemaps, regardless of extinction risk category, seem to indicate species who could benefit from further expert study.
 
-![](figures/barchart_spp_gp_quads.png)
-
-Breaking down the quadrants by IUCN extinction risk categories, species with higher extinction risk tend to be better matched between the two data collections; this is not surprising, as these at-risk species are likely to face closer expert scrutiny than species ranked as Least Concern.
-
-When broken down by taxonomic group, we found that certain taxonomic groups were far more likely than others to be spatially well-matched; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2), .
-Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two data collections, predicted areas are not well-matched; AquaMaps accounts for seafloor depth as a limiting factor, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, and thus likely overestimate the actual distribution.
+When broken down by taxonomic group, we found that certain taxonomic groups were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
+Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two datasets, predicted areas are not well-aligned.
+AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range.
+A similar mechanism may also drive poor area alignment of other demersal and reef-associated organisms.
 
@@
 
-Why should anyone care about AquaMaps thresholds. ___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
+AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value indicated by IUCN maps.
+Frequently, studies convert this probability to a simple presence value by assigning a threshold value (XXX references here).
+A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
+For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to approximate the presence criteria used by IUCN.
+To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0.05 to 1.00 and calculated the average species range relative to a zero threshold.
 
-_what would it mean if there were non-linearities or generally linear trend?_
+_Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
+Do regression - shallow slope indicates that sensitivity is less than 1:1._
 
-![](figures/AMThreshold.png) ___add a mean/median vline on this?___ do we need this figure?
-it looks cool though...
-why broken down by risk category?
-Ditch NE category since most studies probably care about the species health as well as distribution?
-
-___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
-
-![](figures/scatter_AMThreshold_mockup.png)
-Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
-Do regression - shallow slope indicates that sensitivity is less than 1:1.
+At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a (XXX to XXX %, median = XXX) decrease in represented range, though some species lose nearly their entire range.
+Examining the entire species catalog as a single set, a unit shift in presence threshold results in a generally smaller proportional shift in represented area, indicating a moderate and consistent sensitivity to threshold with no surprising tradeoffs to inform the identification of a "correct" threshold.
+This pattern may not hold true for all subsets of AquaMaps, however, whether subsetting by taxa or by georegion.
 
@@
 The biodiversity goal is made up of two parts; Species and Habitats.
-As it is currently calculated, the Species subgoal uses the IUCN range map data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
-For those species that are are represented within the IUCN range maps, AquaMaps distribution maps were used.
-Species with no spatial data in either data collection were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
+As it is currently calculated, the Species subgoal uses species spatial distribution data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
+Spatial distributions were gleaned from both IUCN and AquaMaps datasets.
+A probability threshold of 40% is used to determine presence for AquaMaps data.
+For those species that are are represented within both datasets, IUCN distribution maps were used as the preferred data source.
+Species with no spatial data in either dataset were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
 
-This decision to use IUCN data preferentially to AquaMaps data was made based on the fact that experts are developing the IUCN maps and AquaMaps data is based off a computer model, which in some cases is then consulted by an expert and appropriately altered.
-___is this really the reasoning behind the decision?
-let's get the real reason in here___
+Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
+Reducing the threshold will always increase the apparent range of a species, therefore the slight decrease in average score suggests increased spatial representation of threatened species.
 
-Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
+Scenario 2 shows the effect of preferring AquaMaps data over IUCN, while maintaining the same presence threshold.
+This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so many but not all overlapping species will see a decrease in represented range.
+The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
 
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
+Scenario 3 shows the effect of preferring AquaMaps data over IUCN, while eliminating the presence threshold.
+Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
+The large decrease seems to indicate that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.
 
-| Scenario              | Priority data source | AquaMaps presence threshold |
-| --------------------- | :------------------: | :-------------------------: |
-| Scenario 0 (current)  | IUCN                 |  >= 40%                     |
-| Scenario 1            | IUCN                 |   > 0%                      |
-| Scenario 2            | AquaMaps             |  >= 40%                     |
-| Scenario 3            | AquaMaps             |   > 0%                      |
-
-Scenario 1 shows the effect of reducing the threshold for AquaMaps presence.
-This will always increase the apparent range of any AquaMaps species in the analysis.
-The slight decrease in average score suggests increased spatial representation of species at greater risk of extinction.
-
-Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.
-This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so we could predict an average decrease in range for these overlapping species.
-_The slight bump in mean score indicates that a slight increased representation of low-risk species counterbalances a decreased representation of high-risk species._ ___can we really make any generalizations here?
-not really anything useful without further analysis___
-
-Scenario 3 shows the combination of the two previous scenarios.  The larger decrease in score indicates that the ___hmm gotta think about this a little more - come back to it___
+Since the Ocean Health Index Species subgoal depends on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted.
+When IUCN data is the preferred data source, only the non-paired AquaMaps species will be affected by threshold changes; and even when AquaMaps is the preferred source, the non-paired IUCN species will dampen the effect of a threshold change.
 
@@
 
-#References
+-----
 
@@
     * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources
+
+# Figures
+
+Figure 1 (a, b):
+
+![](figures/bar_by_taxa_all.png)
+
+![](figures/bar_by_taxa_assessed.png)
+
+Figure 2 (a, b):
+
+![](figures/all_aquamaps_map.png)
+
+![](figures/all_iucn_map.png)
+
+Figure 3:
+
+![](figures/AreaRatiovsPercOverlap.png)
+
+___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
+
+Figure 4 (a, b):
+
+![](figures/barchart_spp_risk_quads.png)
+
+![](figures/barchart_spp_gp_quads.png)
+
+Figure 5 (a, b):
+
+![](figures/AMThreshold.png)
+
+![](figures/ribbon_AMThreshold.png)
+
+Figure 6
+
+![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
+
+Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
+
+| Scenario              | Priority data source | AquaMaps presence threshold |
+| --------------------- | :------------------: | :-------------------------: |
+| Scenario 0 (current)  | IUCN                 |  >= 40%                     |
+| Scenario 1            | IUCN                 |   > 0%                      |
+| Scenario 2            | AquaMaps             |  >= 40%                     |
+| Scenario 3            | AquaMaps             |   > 0%                      |

### 2015-12-19 R/data_explore_fxns.R

@@ -1,10 +1,19 @@
 ### data_explore_fxns.R
 
 ### generic theme for all plots
-ggtheme_plot <- theme(axis.ticks = element_blank(),
-                      text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-                      plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-                      legend.position = 'right')
+ggtheme_basic <- theme(axis.ticks = element_blank(),
+        text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
+        plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
+        legend.position = 'right')
+
+ggtheme_plot <- ggtheme_basic + 
+  theme(panel.border = element_blank(),
+        panel.grid.minor = element_blank(), 
+        panel.grid.major = element_line(colour = 'grey90'),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey30"))
+
+  
 
 # ggtheme_map <- theme(axis.ticks = element_blank(), axis.text = element_blank(),
 #                     text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
@@ -12,7 +21,7 @@ ggtheme_plot <- theme(axis.ticks = element_blank(),
 #                     legend.position = 'right')
 
 ### theme for species range maps - based on generic plot theme
-ggtheme_map <- ggtheme_plot +
+ggtheme_map <- ggtheme_basic +
   theme(axis.text = element_blank())
 
 ### Function to create a species list from scratch... use when new data is available

### 2015-12-20 data_explore/compare_am_2015_files.Rmd
cleaning up directory, separating manuscript into main and children

---
title: "Comparing Aquamap 2015 csv files"
author: "Jamie Afflerbach"
date: "12/3/2015"
output: html_document
---

Since the 2015 Aquamaps dataset was sent as .sql files, we implemented two different methods of converting the .sql to .csvs.

Casey wrote an R script to convert the files found here.

Jamie used the old method of MySQL to convert the files. A solution in R is preferred so this script compares the output of the two methods to make sure the R script produces what we want.

###Setup
```{r}

library(data.table)
library(dplyr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

```

###Casey's data tables

```{r}

hcaf_native_co = read.csv(file.path(dir_N,'git-annex/globalprep/_raw_data/aquamaps/v2015/csv/hcaf_sp_native_trunc.csv'))
hcaf_co        = read.csv(file.path(dir_N,'git-annex/globalprep/_raw_data/aquamaps/v2015/csv/hcaf_truncated.csv'))
spp_co         = read.csv(file.path(dir_N,'git-annex/globalprep/_raw_data/aquamaps/v2015/csv/speciesoccursum.csv'))
```

###Jamie's data tables

```{r}

hcaf_native_ja = fread(file.path(dir_N,'git-annex/globalprep/_raw_data/aquamaps/v2015/tables/hcaf_species_native_ohi.csv'))%>%as.data.frame()
hcaf_ja        = read.csv(file.path(dir_N,'git-annex/globalprep/_raw_data/aquamaps/v2015/tables/hcaf_ohi.csv'))
spp_ja         = read.csv(file.path(dir_N,'git-annex/globalprep/_raw_data/aquamaps/v2015/tables/speciesoccursum_ohi.csv'))
```

##Comparison

```{r}

# First look at species list

nrow(spp_ja)-nrow(spp_co)

setdiff(spp_ja$SPECIESID,spp_co$speciesid)
setdiff(spp_co$speciesid,spp_ja$SPECIESID)
setdiff(spp_ja$Species,spp_co$species)

# Looks like no differences

# Look at cells files

nrow(spp_ja)-nrow(spp_co)

setdiff(hcaf_ja$CsquareCode,hcaf_co$csquarecode)
setdiff(hcaf_co$csquarecode,hcaf_ja$CsquareCode)

#looks like no differences

nrow(hcaf_native_ja) - nrow(hcaf_native_co)

setdiff(unique(hcaf_native_ja$SpeciesID),unique(hcaf_native_co$speciesid))
setdiff(unique(hcaf_native_co$speciesid),unique(hcaf_native_ja$SpeciesID))

#look at number of cells per species

h = hcaf_native_ja%>%group_by(SpeciesID)%>%summarize(count=n())%>%rename(speciesid=SpeciesID)%>%as.data.frame()
k = hcaf_native_co%>%group_by(speciesid)%>%summarize(count=n())%>%as.data.frame()%>%mutate(speciesid=as.character(speciesid))

setdiff(h,k)
setdiff(k,h)

#seems like just one species is causing the difference
f = filter(hcaf_native_co, speciesid=='Fis-32347')%>% #grab only rows with this species from Caseys data
            mutate(speciesid=as.character(speciesid))
g = filter(hcaf_native_ja, SpeciesID=='Fis-32347')%>%
     left_join(hcaf_co, by = c('CsquareCode' = 'csquarecode'))%>%
        select(speciesid = SpeciesID,probability,loiczid)

setdiff(g,f)

f = mutate(f,dup = duplicated(f))
filter(f,dup==TRUE)

#There are 6 cells that are duplicated for some reason....

filter(f,loiczid==150423)
filter(f,loiczid==150424)

filter(g,loiczid==150423) # but in the data that Casey created using R, he got rid of the duplicates

```

###Conclusion

It looks like there are 6 cells that are duplicated for Species ID **Fis-32347** (Bluntsnout lanternfish). This came from the raw Aquamaps dataset so Casey's version is just fine to work with!

### 2015-12-20 data_explore/data_explore.Rmd
cleaning up directory, separating manuscript into main and children

--- 2015-12-19 data_explore.Rmd
@@
 spp_list <- spp_list %>%
-  left_join(read.csv('sciname_lookups/spp_groupnames.csv', stringsAsFactors = FALSE),
+  left_join(read.csv('data/sciname_lookups/spp_groupnames.csv', stringsAsFactors = FALSE),
             by = 'spp_group')
@@
 
-sciname_lookup <- read.csv('sciname_lookups/spp_iucn_all.csv', stringsAsFactors = FALSE)
+sciname_lookup <- read.csv('data/sciname_lookups/spp_iucn_all.csv', stringsAsFactors = FALSE)
 

### 2015-12-20 data_explore/data_explore_global_maps.Rmd
cleaning up directory, separating manuscript into main and children

--- 2015-12-14 R/data_explore_global_maps.Rmd
@@
 
+am_ln_map <- log(am_sp_map + 1)
+iucn_ln_map <- log(iucn_sp_map + 1)
+
 ### Plot both maps
@@
           colorkey = list(at = seq(0, 1063, length = 15)))
+
+levelplot(am_ln_map, par.settings = myTheme, 
+          main = 'Aquamaps Dataset', 
+          colorkey = list( at = seq(0, maxValue(am_ln_map), length = 15)))
+
+levelplot(iucn_ln_map, par.settings = myTheme, 
+          main = 'IUCN Dataset', 
+          colorkey = list(at = seq(0, maxValue(iucn_ln_map), length = 15)))
 
@@
 levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
+
+all_am_ln_map <- log(all_am_map + 1)
+all_am_ln_map <- all_am_ln_map/maxValue(all_am_ln_map)
+levelplot(all_am_ln_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset: log-transformed relative species richness', 
+          colorkey = list(at = seq(0, 1, length = 15)))
+
+am_ln_pts <- as.data.frame(rasterToPoints(all_am_ln_map))
+spp_richness_am_plot <- ggplot(am_ln_pts, aes(x = x, y = y)) +
+  ggtheme_map + 
+  geom_raster(aes(fill = layer), alpha = .7) +
+  scale_fill_continuous(low = 'white', high = 'blue') +
+  borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
+  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
+  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
+  labs(title = 'AquaMaps Dataset: log-transformed relative species richness', x = NULL, y = NULL, fill = NULL)
+
+spp_richness_am_plot
+ggsave(file.path(dir_fig, 'all_am_ln_map.png'), height = 4, width = 8)
+
 ```
@@
 levelplot(all_iucn_map, par.settings = myTheme, main = 'Entire IUCN Dataset')
+
+all_iucn_ln_map <- log(all_iucn_map + 1)
+all_iucn_ln_map <- all_iucn_ln_map/maxValue(all_iucn_ln_map)
+levelplot(all_iucn_ln_map, 
+          par.settings = myTheme, 
+          margin = list(axis = TRUE),
+          main = 'Entire IUCN Dataset: log-transformed relative species richness', 
+          colorkey = list(at = seq(0, 1, length = 15)))
+
+iucn_ln_pts <- as.data.frame(rasterToPoints(all_iucn_ln_map))
+spp_richness_iucn_plot <- ggplot(iucn_ln_pts, aes(x = x, y = y)) +
+  ggtheme_map + 
+  geom_raster(aes(fill = layer), alpha = .7) +
+  scale_fill_continuous(low = 'white', high = 'blue') +
+  borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
+  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
+  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
+  labs(title = 'IUCN Dataset: log-transformed relative species richness', x = NULL, y = NULL, fill = NULL)
+
+spp_richness_iucn_plot
+ggsave(file.path(dir_fig, 'all_iucn_ln_map.png'), height = 4, width = 8)
+
 ```
@@
 levelplot(all_am_pct_map, par.settings = myTheme, 
-          main = 'Aquamaps Dataset percent of max', 
-          colorkey = list( at = seq(0, 1, length = 15)))
+          main = 'Aquamaps Dataset percent of max')
 levelplot(all_iucn_pct_map, par.settings = myTheme, 
-          main = 'IUCN Dataset percent of max', 
-          colorkey = list(at = seq(0, 1, length = 15)))
+          main = 'IUCN Dataset percent of max')
 ### this is the difference one. Should it be min value to max value?
 levelplot(diff_pct_map, par.settings = myTheme, 
-          main = 'Difference: AquaMaps - IUCN by percent', 
+          main = 'Difference: AquaMaps - IUCN by percent')
+
+diff_ln_map <- all_am_ln_map - all_iucn_ln_map
+levelplot(diff_ln_map, 
+          par.settings = myTheme, 
+          pretty = TRUE,
+          margin = list(axis = TRUE),
+          main = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', 
           colorkey = list(at = seq(-1, 1, length = 15)))
+maxValue(all_am_ln_map)
+minValue(all_am_ln_map)
+maxValue(all_iucn_ln_map)
+minValue(all_iucn_ln_map)
+maxValue(diff_ln_map)
+minValue(diff_ln_map)
+diff_ln_pts <- as.data.frame(rasterToPoints(diff_ln_map))
+
+spp_richness_diff_plot <- ggplot(diff_ln_pts, aes(x = x, y = y)) +
+  ggtheme_map + 
+  geom_raster(aes(fill = layer)) +
+  scale_fill_gradient2(low = 'red', mid = 'white', high = 'blue', midpoint = 0) +
+  borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
+  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
+  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
+  labs(title = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', x = NULL, y = NULL)
+
+spp_richness_diff_plot
 

### 2015-12-20 data_explore/data_explore_pt3.Rmd
cleaning up directory, separating manuscript into main and children

Plotting species from q1 (well-matched) (see data_explore_quadmaps.html for plots)
```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

x <- 20
y <- nrow(spp_list_q1)

spp_names_q1 <- spp_list_q1$sciname[seq(1, y, y/x)]

for (spp in spp_names_q1) { # i = 1
  spp_plot <- plot_rangemap(spp)
  spp_solo <- spp_list %>% filter(sciname == spp)
  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
  spp_plot <- spp_plot +
    labs(title = sprintf('%s, quadrant 1: %s', spp, text_tag))

  print(spp_plot)

  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q1_%s.png', str_replace(tolower(spp), ' ', '_'))),
  #        plot = spp_plot, 
  #        width = 8, height = 4, dpi = 200)
}

```

Plotting species from q2 (distribution-matched)
```{r plot_spp_from_q2, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

x <- 20
y <- nrow(spp_list_q2)

spp_names_q2 <- spp_list_q2$sciname[seq(1, y, y/x)]

for (spp in spp_names_q2) { # i = 1
  spp_plot <- plot_rangemap(spp)
  spp_solo <- spp_list %>% filter(sciname == spp)
  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
  spp_plot <- spp_plot +
    labs(title = sprintf('%s, quadrant 2: %s', spp, text_tag))
  print(spp_plot)

  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q2_%s.png', str_replace(tolower(spp), ' ', '_'))),
  #        plot = spp_plot, 
  #        width = 8, height = 4, dpi = 200)
}

```

Plotting species from q3 (area-matched)
```{r plot_spp_from_q3, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

x <- 20
y <- nrow(spp_list_q3)

spp_names_q3 <- spp_list_q3$sciname[seq(1, y, y/x)]

for (spp in spp_names_q3) { # i = 1
  spp_plot <- plot_rangemap(spp)
  spp_solo <- spp_list %>% filter(sciname == spp)
  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
  spp_plot <- spp_plot +
    labs(title = sprintf('%s, quadrant 3: %s', spp, text_tag))
  print(spp_plot)

  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q3_%s.png', str_replace(tolower(spp), ' ', '_'))),
  #        plot = spp_plot, 
  #        width = 8, height = 4, dpi = 200)
}

```

Plotting species from q4 (poorly matched)
```{r plot_spp_from_q4, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

x <- 20
y <- nrow(spp_list_q4)

spp_names_q4 <- spp_list_q4$sciname[seq(1, y, y/x)]

for (spp in spp_names_q4) { # i = 1
  spp_plot <- plot_rangemap(spp)
  spp_solo <- spp_list %>% filter(sciname == spp)
  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
  spp_plot <- spp_plot +
    labs(title = sprintf('%s, quadrant 4: %s', spp, text_tag))
  print(spp_plot)

  # ggsave(filename = file.path(dir_fig, sprintf('rangemap_q4_%s.png', str_replace(tolower(spp), ' ', '_'))),
  #        plot = spp_plot, 
  #        width = 8, height = 4, dpi = 200)
}

```

### 2015-12-20 data_explore/data_explore_fxns.R

@@ -1,0 +1,246 @@
+### data_explore_fxns.R
+
+### generic theme for all plots
+ggtheme_basic <- theme(axis.ticks = element_blank(),
+        text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
+        plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
+        legend.position = 'right')
+
+ggtheme_plot <- ggtheme_basic + 
+  theme(panel.border = element_blank(),
+        panel.grid.minor = element_blank(), 
+        panel.grid.major = element_line(colour = 'grey90'),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey30"))
+
+  
+
+# ggtheme_map <- theme(axis.ticks = element_blank(), axis.text = element_blank(),
+#                     text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
+#                     plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
+#                     legend.position = 'right')
+
+### theme for species range maps - based on generic plot theme
+ggtheme_map <- ggtheme_basic +
+  theme(axis.text = element_blank())
+
+### Function to create a species list from scratch... use when new data is available
+create_spp_list <- function(spp_list_base, data_file) {
+  spp_list <- spp_list_base %>%
+    mutate(total_area = NA,
+           am_area    = NA,
+           iucn_area  = NA,
+           perc       = NA,
+           am_perc    = NA,
+           iucn_perc  = NA,
+           sm_perc    = NA, #filled with % of smaller range that is within the larger range
+           sm_range   = NA) #what range is smaller
+  
+  
+  #there's likely a better way to do this than a for loop (lapply or sapply?)
+  for (i in 1:nrow(spp_list)){ # i = 1
+    spp     <- spp_list[i, ]$sciname #species name
+    spp_map <- get_spp_map(spp) #get species map
+    
+    ###get just the aquamaps cells
+    am_spp_map <- spp_map %>%
+      filter(!is.na(am_prob))%>%
+      dplyr::select(loiczid, am_sid, am_prob) %>% 
+      ### add in am_prob when we want to look at probability
+      unique()
+    
+    if(any(duplicated(am_spp_map$loiczid))){
[…]

### 2015-12-20 data_explore/spp_compare_AM_IUCN.R

@@ -1,0 +1,128 @@
+### Compare status and trend 
+
+setwd('~/github/ohiprep')
+library(ggplot2)
+source('src/R/common.R')
+
+goal     <- 'globalprep/SPP_ICO'
+scenario <- 'vAM_IUCN'
+dir_data  <- file.path('~/github/ohiprep', goal, scenario, 'data') 
+
+
+#############################################################################=
+scatterPlot <- function(csv_orig, csv_new, title_text, zero_one_new = FALSE, zero_one_old = FALSE,
+                        fig_save = file.path(dir_data, '../graphs', paste0(title_text, '_scatterPlot.png'))) {
+  
+  require(git2r)
+  require(ggplot2)
+  require(RColorBrewer)
+  
+  names <- read.csv("~/github/ohi-global/eez2013/layers/rgn_labels.csv") %>%
+    filter(type=="eez") %>%
+    select(rgn_id, label)
+  
+  data_orig <- read.csv(csv_orig)
+  names(data_orig) <- c('rgn_id', 'score')
+  data_new <- read.csv(csv_new)
+  names(data_new) <- c('rgn_id', 'score')
+  if(zero_one_old) data_orig$score = data_orig$score*100
+  if(zero_one_new) data_new$score = data_new$score*100
+  
+  data_combo <- data_orig %>%
+    dplyr::rename(scores_old = score) %>%
+    left_join(data_new %>%
+                dplyr::rename(scores_new = score), 
+              by=c('rgn_id')) %>%
+    dplyr::mutate(change = scores_new - scores_old) %>%
+    dplyr::mutate(mean = mean(change, na.rm=TRUE),
+           sd =  sd(change, na.rm=TRUE)) %>%
+    ungroup() %>%
+    dplyr::mutate(z_score = (change-mean)/sd) %>%
+    dplyr::mutate(z_greater_1 = ifelse(abs(z_score) > 1, "yes", "no")) %>%
+    left_join(names, by='rgn_id') %>%
+    filter(rgn_id != 0) %>%
+    dplyr::mutate(
+#      label     = ifelse(is.na(label), as.character(rgn_id), label),
+      plotLabel = ifelse(z_greater_1=="yes", as.character(label), NA)
+    ) 
+    
+  ggplot(data_combo, aes(x = scores_old, y = scores_new)) +
+    geom_point(shape = 19) +
+    theme_bw() + 
+    labs(title=sprintf('Score differences for %s', title_text), 
[…]

### 2015-12-20 R/data_explore_fxns.R
more cleanup

@@ -1,246 +1,0 @@
-### data_explore_fxns.R
-
-### generic theme for all plots
-ggtheme_basic <- theme(axis.ticks = element_blank(),
-        text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-        plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-        legend.position = 'right')
-
-ggtheme_plot <- ggtheme_basic + 
-  theme(panel.border = element_blank(),
-        panel.grid.minor = element_blank(), 
-        panel.grid.major = element_line(colour = 'grey90'),
-        panel.background = element_blank(),
-        axis.line = element_line(colour = "grey30"))
-
-  
-
-# ggtheme_map <- theme(axis.ticks = element_blank(), axis.text = element_blank(),
-#                     text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-#                     plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-#                     legend.position = 'right')
-
-### theme for species range maps - based on generic plot theme
-ggtheme_map <- ggtheme_basic +
-  theme(axis.text = element_blank())
-
-### Function to create a species list from scratch... use when new data is available
-create_spp_list <- function(spp_list_base, data_file) {
-  spp_list <- spp_list_base %>%
-    mutate(total_area = NA,
-           am_area    = NA,
-           iucn_area  = NA,
-           perc       = NA,
-           am_perc    = NA,
-           iucn_perc  = NA,
-           sm_perc    = NA, #filled with % of smaller range that is within the larger range
-           sm_range   = NA) #what range is smaller
-  
-  
-  #there's likely a better way to do this than a for loop (lapply or sapply?)
-  for (i in 1:nrow(spp_list)){ # i = 1
-    spp     <- spp_list[i, ]$sciname #species name
-    spp_map <- get_spp_map(spp) #get species map
-    
-    ###get just the aquamaps cells
-    am_spp_map <- spp_map %>%
-      filter(!is.na(am_prob))%>%
-      dplyr::select(loiczid, am_sid, am_prob) %>% 
-      ### add in am_prob when we want to look at probability
-      unique()
-    
-    if(any(duplicated(am_spp_map$loiczid))){
[…]

### 2015-12-20 R/map_am_v_iucn.R

@@ -1,173 +1,0 @@
-#load data for species map function, function, and plotting a species
-
-#JAfflerbach
-#9.11.2014
-
-## SETUP ##
-
-#libraries
-
-library(data.table)
-library(readr)
-library(raster)
-library(rgdal)
-library(dplyr)
-library(RColorBrewer)
-library(maps)
-
-
-#increase R memory size to deal with reading in the large files
-
-memory.limit(size=200000000)
-
-
-dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-           'Darwin'  = '/Volumes/data_edit',
-           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-dir_anx      <- file.path(dir_N, 'git-annex/globalprep/SpeciesDiversity')
-dir_aquamaps <- file.path(dir_anx, 'raw/aquamaps_2014')
-dir_iucn     <- file.path(dir_anx, 'iucn_intersections')
-dir_cells    <- file.path(dir_N, 'git-annex/Global/NCEAS-SpatialFishCatch_v2014/raw/ohi_spp/data')
-
-#read in data
-#fread comes from data.table package and is much more efficient at reading in the larger .csv files
-
-spp_list <- read.csv('spp_am_v_iucn.csv', stringsAsFactors = FALSE)
-### list of species with both IUCN and Aquamaps spatial data
-spp_list_ltd <- spp_list %>% filter(!is.na(parent_sid) | !is.na(subpop_sid))
-### truncated list
-
-# am_spp   <- fread(input = file.path(dir_aquamaps, 'tables/ohi_speciesoccursum.csv'), header = TRUE) %>%
-#               mutate(sciname = paste(Genus, Species, sep = ' '))
-
-
-### get cell ID lookup table
-# cells    <- read.csv(file.path(dir_cells, 'cells.csv')) %>%
-# cells <- cells %>%
-#   dplyr::select(cid, csq = csquarecod, loiczid, faoaream)
-# 
-# cells2 <- read.csv(file.path(dir_anx, 'raw/aquamaps_2014/tables/hcaf.csv'))
-# cells2 <- cells2 %>% 
-#   dplyr::select(csq = CsquareCode, loiczid = LOICZID, faoaream = FAOAreaM)
-# write_csv(cells,  file.path(dir_anx, 'explore_am_v_iucn/cells.csv'))
[…]

### 2015-12-20 R/mapping aquamaps.R

@@ -1,87 +1,0 @@
-#load data for species map function, function, and plotting a species
-
-#JAfflerbach
-#9.11.2014
-
-## SETUP ##
-
-#libraries
-
-library(data.table)
-library(raster)
-library(rgdal)
-library(dplyr)
-library(RColorBrewer)
-library(maps)
-
-
-#increase R memory size to deal with reading in the large files
-
-memory.limit(size=200000000)
-
-
-dir_N = c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-          'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-dir_aquamaps = file.path(dir_N,"git-annex/globalprep/SPP_ICO/raw/aquamaps_2014")
-dir_cells    = file.path(dir_N,"git-annex/Global/NCEAS-SpatialFishCatch_v2014/raw/ohi_spp/data")
-
-#read in data
-#fread comes from data.table package and is much more efficient at reading in the larger .csv files
-
-spp       = fread(input=file.path(dir_aquamaps,"tables/ohi_speciesoccursum.csv"),header=T)%>%
-              mutate(scientific = paste(Genus,Species,sep=" "))
-
-cells     = read.csv(file.path(dir_cells,'cells.csv'))%>%
-              dplyr::select(cid,csquarecod,faoaream)
-    names(cells)<-c("cid","csquare_code","faoaream")
-
-spp_cells = fread(input=file.path(dir_aquamaps,"tables/ohi_hcaf_species_native.csv"),header=T,
-                  colClasses=c('NULL','character',NA,'numeric','NULL','NULL')) #1 min 49 sec
-
-setnames(spp_cells,c('SpeciesID','CsquareCode','probability'),c("species_id","csquare_code","probability"))
-
-
-#convert spp_cells, which is both a 'data.table' and a 'data.frame' to just a 'data.frame' to allow for plotting. A weird
-# caveat of fread() is that the output is both class 'data.table' and 'data.frame'
-
-spp_cells = as.data.frame(spp_cells)
-
-
-### Species Map Function ###
-# This function takes a single species scientific name as input, then grabs all occurrence cells and associated probability per cell
-
-sp_map_fun <- function(species){
-  
-  sp_id  = filter(spp,scientific==species)$SPECIESID
-  sp_map = filter(spp_cells,species_id==sp_id)%>%
[…]

### 2015-12-20 R/spp_compare_AM_IUCN.R

@@ -1,128 +1,0 @@
-### Compare status and trend 
-
-setwd('~/github/ohiprep')
-library(ggplot2)
-source('src/R/common.R')
-
-goal     <- 'globalprep/SPP_ICO'
-scenario <- 'vAM_IUCN'
-dir_data  <- file.path('~/github/ohiprep', goal, scenario, 'data') 
-
-
-#############################################################################=
-scatterPlot <- function(csv_orig, csv_new, title_text, zero_one_new = FALSE, zero_one_old = FALSE,
-                        fig_save = file.path(dir_data, '../graphs', paste0(title_text, '_scatterPlot.png'))) {
-  
-  require(git2r)
-  require(ggplot2)
-  require(RColorBrewer)
-  
-  names <- read.csv("~/github/ohi-global/eez2013/layers/rgn_labels.csv") %>%
-    filter(type=="eez") %>%
-    select(rgn_id, label)
-  
-  data_orig <- read.csv(csv_orig)
-  names(data_orig) <- c('rgn_id', 'score')
-  data_new <- read.csv(csv_new)
-  names(data_new) <- c('rgn_id', 'score')
-  if(zero_one_old) data_orig$score = data_orig$score*100
-  if(zero_one_new) data_new$score = data_new$score*100
-  
-  data_combo <- data_orig %>%
-    dplyr::rename(scores_old = score) %>%
-    left_join(data_new %>%
-                dplyr::rename(scores_new = score), 
-              by=c('rgn_id')) %>%
-    dplyr::mutate(change = scores_new - scores_old) %>%
-    dplyr::mutate(mean = mean(change, na.rm=TRUE),
-           sd =  sd(change, na.rm=TRUE)) %>%
-    ungroup() %>%
-    dplyr::mutate(z_score = (change-mean)/sd) %>%
-    dplyr::mutate(z_greater_1 = ifelse(abs(z_score) > 1, "yes", "no")) %>%
-    left_join(names, by='rgn_id') %>%
-    filter(rgn_id != 0) %>%
-    dplyr::mutate(
-#      label     = ifelse(is.na(label), as.character(rgn_id), label),
-      plotLabel = ifelse(z_greater_1=="yes", as.character(label), NA)
-    ) 
-    
-  ggplot(data_combo, aes(x = scores_old, y = scores_new)) +
-    geom_point(shape = 19) +
-    theme_bw() + 
-    labs(title=sprintf('Score differences for %s', title_text), 
[…]

### 2015-12-21 data_explore/data_explore.Rmd
trying this again... cleaning up data_explore plots and .csvs

--- 2015-12-19 data_explore.Rmd
@@
 library(data.table)
+library(raster)
 library(tidyr)
@@
 
+dir_git <- '~/github/IUCN-AquaMaps'
+
 dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-dir_fig <- file.path(getwd(), 'figures')
+dir_fig <- file.path(dir_git, 'figures')
+dir_data <- file.path(dir_git, 'data')
 
-source('R/data_explore_fxns.R')
+source('data_explore_fxns.R')
+### When knitting this, it automatically sets WD to be this directory...
 ```
@@
 ### LOICZID cell ID numbers - used in plotting rasters
-loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
+loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
 loiczid_raster       <- raster(loiczid_raster_file)
@@
 ```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
+data_file <- file.path(dir_data, 'spp_list_w_area_all.csv')
 
@@
   ### species info list - scinames, IUCN species ID, AM species ID, category etc
-  spp_list_base <- read.csv('data/spp_am_v_iucn_2015_inclDD.csv', stringsAsFactors = FALSE)
+  spp_list_base <- read.csv(file.path(dir_data, 'spp_am_v_iucn.csv'),
+                            stringsAsFactors = FALSE)
 
@@
 
-spp_list <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(sm_perc)) 
+spp_list <- read.csv(data_file, stringsAsFactors = FALSE) 
+### At this point, spp_list contains subpops and aliases, and zero-area IUCNs
+
+spp_list <- spp_list %>%
+  filter(!is.na(sm_perc)) %>%
   ### will ditch the ones where no IUCN polygon exists - or are these species
   ### whose ranges only occur in cells beyond the LOICZID grid?
+  filter(!str_detect(spatial_source, 'subpop') & !str_detect(spatial_source, 'alias'))
 spp_list <- spp_list %>%
-  left_join(read.csv('sciname_lookups/spp_groupnames.csv', stringsAsFactors = FALSE),
+  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
+                     stringsAsFactors = FALSE),
             by = 'spp_group')
+
+write.csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
 
@@
 
-``` {r child = 'R/data_explore_pt1.Rmd', eval = FALSE}
+``` {r child = 'data_explore_pt1.Rmd', eval = FALSE}
 ```
@@
 
-``` {r, child = 'R/data_explore_pt2.Rmd', eval = FALSE}
+``` {r, child = 'data_explore_pt2.Rmd', eval = FALSE}
 ```
@@
 
-``` {r child = 'R/data_explore_pt3.Rmd', eval = FALSE}
+<!-- # Multiple species maps per quadrant -->
+``` {r child = 'data_explore_pt3.Rmd', eval = FALSE}
 ```
-<!-- # Multiple species maps per quadrant -->
 
@@
 
-data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
+data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
 spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
@@
 
-spp_all_file <- 'data/spp_all_2015.csv'
+spp_all_file <- file.path(dir_data, 'spp_all.csv')
 spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
 
-iucn_marine_file <- 'data/spp_iucn_marine.csv'
+iucn_marine_file <- file.path(dir_data, 'spp_iucn_marine.csv')
 spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
 
-iucn_maps_file <- 'data/spp_iucn_maps.csv'
+message('Getting IUCN maps list')
+iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps.csv')
 spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
@@
 
-spp_am_file <- 'data/spp_am.csv'
+spp_am_file <- file.path(dir_data, 'spp_am.csv')
 spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
@@
 
-sciname_lookup <- read.csv('sciname_lookups/spp_iucn_all.csv', stringsAsFactors = FALSE)
+sciname_lookup <- read.csv(file.path(dir_data, 'sciname_lookups/spp_iucn_scinames.csv'), stringsAsFactors = FALSE)
 
@@
   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) + 
   labs(x = 'Class', y = 'Percent of mapped species',
@@
   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
   labs(x = 'Class', y = 'Percent of mapped and assessed species',
@@
 
-``` {r, child = 'R/data_explore_global_maps.Rmd', eval = FALSE}
+``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
 ```
 
-``` {r, child = 'R/data_explore_am_thresh.Rmd', eval = FALSE}
+``` {r, child = 'data_explore_am_thresh.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
 ```
-
-## Changes in SPP for scenarios
-
-Comparing three scenarios to the current (v2015) global model for SPP:
-
-* Baseline is: preferring IUCN rangemaps over Aquamaps rangemaps; Aquamaps 'presence' threshold is set at 40% or greater.
-* Scenario 1: preferring IUCN rangemaps over Aquamaps; AM presence threshold set at >= 1% (effectively, all occurrences).
-* Scenario 2: preferring AM maps over IUCN; AM presence >= 40%.
-* Scenario 3: preferring AM over IUCN; AM presence >= 1%.
-
-Boxplots:
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
-
-![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)

### 2015-12-21 data_explore/data_explore_global_maps.Rmd
trying this again... cleaning up data_explore plots and .csvs

@@
 
-``` {r setup, message = FALSE, eval = FALSE}
-library(rasterVis)
-library(ggplot2)
-library(maps)
-library(raster)
-library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-library(readr)
-
-dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-           'Darwin'  = '/Volumes/data_edit',
-           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')
-dir_fig <- file.path(getwd(), 'figures')
-
-source('R/data_explore_fxns.R')
-
-data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
-
-spp_list <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(sm_perc)) ### will ditch the ones where no IUCN polygon exists
-
-```
-
-``` {r load data sets, echo = FALSE, message = FALSE}
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
-                           col_types = 'cdddd', progress = TRUE) %>%
-  as.data.frame() %>%
-  filter(presence != 5) ### removes extinct subpopulations
-
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), progress = TRUE)
-
-### LOICZID cell ID numbers - used in plotting rasters
-loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
-loiczid_raster       <- raster(loiczid_raster_file)
-names(loiczid_raster) <- 'loiczid'
-```
-
-```{r global map matched species, echo = FALSE, warning = FALSE, message = FALSE}
+```{r global map matched species, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 ### Number of aquamaps species per cell
-iucn_pair_rast_file <- 'rasters/iucn_pair_rast.tif'
-am_pair_rast_file   <- 'rasters/am_pair_rast.tif'
+iucn_pair_rast_file <- file.path(dir_data, 'rasters/iucn_pair_rast.tif')
+am_pair_rast_file   <- file.path(dir_data, 'rasters/am_pair_rast.tif')
 
@@
 
-all_am_rast_file   <- 'rasters/am_all_rast.tif'
+all_am_rast_file   <- file.path(dir_data, 'rasters/am_all_rast.tif')
 
@@
 
-levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
+am_levelplot1 <- levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
 
@@
 all_am_ln_map <- all_am_ln_map/maxValue(all_am_ln_map)
-levelplot(all_am_ln_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset: log-transformed relative species richness', 
+am_levelplot1 <- levelplot(all_am_ln_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset: log-transformed relative species richness', 
           colorkey = list(at = seq(0, 1, length = 15)))
 
-am_ln_pts <- as.data.frame(rasterToPoints(all_am_ln_map))
-spp_richness_am_plot <- ggplot(am_ln_pts, aes(x = x, y = y)) +
-  ggtheme_map + 
-  geom_raster(aes(fill = layer), alpha = .7) +
-  scale_fill_continuous(low = 'white', high = 'blue') +
-  borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
-  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
-  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
-  labs(title = 'AquaMaps Dataset: log-transformed relative species richness', x = NULL, y = NULL, fill = NULL)
+print(am_levelplot1)
+print(am_levelplot2)
 
-spp_richness_am_plot
-ggsave(file.path(dir_fig, 'all_am_ln_map.png'), height = 4, width = 8)
+# am_ln_pts <- as.data.frame(rasterToPoints(all_am_ln_map))
+# spp_richness_am_plot <- ggplot(am_ln_pts, aes(x = x, y = y)) +
+#   ggtheme_map + 
+#   geom_raster(aes(fill = layer), alpha = .7) +
+#   scale_fill_continuous(low = 'white', high = 'blue') +
+#   borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
+#   scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
+#   scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
+#   labs(title = 'AquaMaps Dataset: log-transformed relative species richness', x = NULL, y = NULL, fill = NULL) 
+# 
+# spp_richness_am_plot
+# ggsave(file.path(dir_fig, 'all_am_ln_map.png'), height = 4, width = 8)
 
@@
 
-```{r all_aquamaps, eval = FALSE, echo = FALSE}
+```{r all_iucn, eval = FALSE, echo = FALSE}
 
-all_iucn_rast_file   <- 'rasters/iucn_all_rast.tif'
+all_iucn_rast_file   <- file.path(dir_data, 'rasters/iucn_all_rast.tif')
 
@@
 
-levelplot(all_iucn_map, par.settings = myTheme, main = 'Entire IUCN Dataset')
+iucn_levelplot1 <- levelplot(all_iucn_map, par.settings = myTheme, main = 'Entire IUCN Dataset')
 
@@
 all_iucn_ln_map <- all_iucn_ln_map/maxValue(all_iucn_ln_map)
-levelplot(all_iucn_ln_map, 
+iucn_levelplot2 <- levelplot(all_iucn_ln_map, 
           par.settings = myTheme, 
@@
 
-iucn_ln_pts <- as.data.frame(rasterToPoints(all_iucn_ln_map))
-spp_richness_iucn_plot <- ggplot(iucn_ln_pts, aes(x = x, y = y)) +
-  ggtheme_map + 
-  geom_raster(aes(fill = layer), alpha = .7) +
-  scale_fill_continuous(low = 'white', high = 'blue') +
-  borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
-  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
-  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
-  labs(title = 'IUCN Dataset: log-transformed relative species richness', x = NULL, y = NULL, fill = NULL)
-
-spp_richness_iucn_plot
-ggsave(file.path(dir_fig, 'all_iucn_ln_map.png'), height = 4, width = 8)
+print(iucn_levelplot1)
+print(iucn_levelplot2)
+# iucn_ln_pts <- as.data.frame(rasterToPoints(all_iucn_ln_map))
+# spp_richness_iucn_plot <- ggplot(iucn_ln_pts, aes(x = x, y = y)) +
+#   ggtheme_map + 
+#   geom_raster(aes(fill = layer), alpha = .7) +
+#   scale_fill_continuous(low = 'white', high = 'blue') +
+#   borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
+#   scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
+#   scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
+#   labs(title = 'IUCN Dataset: log-transformed relative species richness', x = NULL, y = NULL, fill = NULL) 
+# 
+# spp_richness_iucn_plot
+# ggsave(file.path(dir_fig, 'all_iucn_ln_map.png'), height = 4, width = 8)
 
 ```
-
-![image](figures/all_aquamaps_map.png)
 

### 2015-12-21 ms_0_main.Rmd
trying this again... cleaning up data_explore plots and .csvs

--- 2015-12-19 ms_outline2.Rmd
@@
 ``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.
+### use data_explore.Rmd to create this file, if it does not yet exist.  This
+### filters out presence = Extinct from IUCN cells.
 
-data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
+data_file <- 'data/spp_list_w_area_all.csv'
 spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%

### 2015-12-21 data_explore/data_explore_fxns.R

@@ -26,7 +26,7 @@ ggtheme_map <- ggtheme_basic +
 
 ### Function to create a species list from scratch... use when new data is available
 create_spp_list <- function(spp_list_base, data_file) {
-  spp_list <- spp_list_base %>%
+  spp_list_build <- spp_list_base %>%
     mutate(total_area = NA,
            am_area    = NA,
            iucn_area  = NA,
@@ -38,8 +38,8 @@ create_spp_list <- function(spp_list_base, data_file) {
   
   
   #there's likely a better way to do this than a for loop (lapply or sapply?)
-  for (i in 1:nrow(spp_list)){ # i = 1
-    spp     <- spp_list[i, ]$sciname #species name
+  for (i in 1:nrow(spp_list_build)){ # i = 1
+    spp     <- spp_list_build[i, ]$sciname #species name
     spp_map <- get_spp_map(spp) #get species map
     
     ###get just the aquamaps cells
@@ -62,7 +62,7 @@ create_spp_list <- function(spp_list_base, data_file) {
                      which = 'am_sid', 
                      subsWithNA = TRUE)
     
-    spp_list[i, ]$am_area <- area(r_am_spp, na.rm = TRUE) %>%
+    spp_list_build[i, ]$am_area <- area(r_am_spp, na.rm = TRUE) %>%
       cellStats(., stat='sum')
     
     
@@ -85,7 +85,7 @@ create_spp_list <- function(spp_list_base, data_file) {
                        which = 'iucn_pres', 
                        subsWithNA = TRUE)
     
-    spp_list[i, ]$iucn_area <- area(r_iucn_spp, na.rm = TRUE) %>%
+    spp_list_build[i, ]$iucn_area <- area(r_iucn_spp, na.rm = TRUE) %>%
       cellStats(., stat = 'sum')
     
     
@@ -102,7 +102,7 @@ create_spp_list <- function(spp_list_base, data_file) {
                       which = 'value', 
                       subsWithNA = TRUE)
     
-    spp_list[i, ]$total_area <- area(fullRange, na.rm = TRUE) %>%
+    spp_list_build[i, ]$total_area <- area(fullRange, na.rm = TRUE) %>%
       cellStats(., stat = 'sum')
     
     ### get percent overlap of map
@@ -120,7 +120,7 @@ create_spp_list <- function(spp_list_base, data_file) {
     
     #percent overlap
[…]

### 2015-12-21 ms_0_main.Rmd
dividing up manuscript into sections

--- 2015-12-17 ms_outline2.Rmd
@@
 ---
-title: "A comparative analysis of AquaMaps and IUCN species distribution maps"
-author: "JA, CO, BH"
+title: "Aligning species range data to better serve science and conservation"
+author: "Casey O'Hara, Jamie Afflerbach, Ben Halpern"
 output:
-  word_document: null
-  pdf_document:
-    number_sections: yes
-    toc: no
   html_document:
@@
     toc: yes
+  pdf_document:
+    number_sections: yes
+    toc: no
+  word_document: null
 ---
@@
 
-data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
+data_file <- 'data/spp_list_w_area_trimmed.csv'
 spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
@@
 
-## Significance
+-----
 
-_This is important because of some reason or another...!!!_
+``` {r abstract, child = 'ms_1_abst_sig.Rmd'}
+```
 
@@
 
-# Introduction
-
-<!--By 2050 ___some threshold year?___ our oceans will __in some way, find a study that predicts big changes?__ be very different from the oceans we enjoy today.--> Climate change, ocean temperature rise, ocean acidification, sea ice loss, human impacts from fishing and introduction of nonnative species are likely to significantly alter the global distribution of marine species - but how well do we really understand the current distribution?
-Understanding species abundances and distributions is a critical component to assessing global biodiversity and detecting patterns and trends through time.
-<!--Many efforts to map terrestrial, freshwater and marine species from local to global scales inform our understanding of taxonomic assemblages, biogeographical limits to species ranges, suitable habitat for species.-->
-
-AquaMaps and IUCN marine species distribution maps are both global-scale spatial data collections that have been used to analyze marine species status (Halpern et al.
-2012, Selig et al.
-2013), global biodiversity patterns and trends (other references), predicting range shifts (Molinos et al.
-2015), and setting conservation priorities (Klein et al.
-2015).
-The two data collections ostensibly describe the same information, but significant differences in methodology, intent, and taxonomic and geographic sampling could lead to dramatically different understandings of our marine ecosystems.
-
-<!--_Differences in methodology (model-driven vs expert knowledge-driven); differences in intent (what does it communicate); differences in represented species and regions_-->
-
-To better understand the similarities and differences between the two data collections, we examined three fundamental questions:
-
-* How do the two spatial data collections compare in their representation of global spatial distribution of species diversity? 
-* How do the two spatial data collections compare in their representation of species diversity across taxonomic groups?
-* For species represented in both collections, how closely do the spatial distribution maps align?
-
-The Ocean Health Index and MPA Gap Analysis study both provide case studies examined the implications of alternate data use scenarios on two large-scale studies of global biodiversity and conservation.
-
-# Understanding each dataset separately
-
-The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.
-Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.
-These maps are typically based on species occurrence records from databases such as ___OBIS___ and ___GBIF___; IUCN provides general guidelines to facilitate consistency in how the maps are generated.
-The IUCN presents these maps as 'limits of distribution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed within the polygon.
-While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
-
-IUCN publishes spatial distribution map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
-in which at least 90% of the species within the taxonomic group have been evaluated.
-While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN had published species distribution maps for nearly 4500 marine species across 24 taxonomic groups.
-(For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.)
-
-AquaMaps (Kaschner et al.
-2013) offers species range maps based on modeled relative environmental suitability.
-For each assessed species, data from occurrence records such as OBIS and GBIF, published species databases such as FishBase, and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, sea ice concentration, and distance to land.
-The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence, with cell values ranging from 0 to 1.00.
-These maps can then be reviewed by experts to fine-tune the parameters to better match known occurrences.
-As of December 2015, AquaMaps current native distribution maps have been produced for nearly 23,000 species, though only 5,600 of these species have been evaluated for the IUCN Red List.
-_See (Kaschner et al.
-2013) for more detailed methods._
+``` {r abstract, child = 'ms_2_intro_res_disc.Rmd'}
+```
 
@@
 
-# Methods
-
-## Analysis of taxonomic and regional inclusion
-
-### Overlap between assessed species
-
-Due to differences in both methodology and intent, there are significant differences in the coverage afforded by the species distribution maps published by AquaMaps and those published by the IUCN.
-This is immediately clear from an examination of the number of species maps offered by each source, and the small proportion of species with maps from both sources.
-
-| Map set   | n~total~                | n~Red_List~                                                       |
-| :-------: | ----------------------: | :---------------------------------------------------------------: | 
-| AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` | 
-| IUCN      | `r nrow(spp_iucn_maps)` | all                                                               | 
-| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))` | all                       |
-
-### Taxonomic distribution between data collections
-
-Studies of species richness and biodiversity depend on sampling a representative cross section of ecosystems, and thus benefit from a broad representation across taxonomic groups.
-To examine the overall taxonomic distribution across the spatial data collections, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
-Since biodiversity studies often rely on both spatial distribution and indicators of species health, we further divided the data by whether species had been evaluated for the IUCN Red List of Threatened Species.
-
-![](figures/bar_by_taxa_all.png)
-
-![](figures/bar_by_taxa_assessed.png)
-
-The IUCN only releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed.
-As such, spatial data for many taxonomic classes are unavailable, and within a class, the assessed sub-groups may not represent the entire class.
-For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
-However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
-
-The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
-
-Some global-scale studies of biodiversity, such as the Ocean Health Index, require both spatial distribution information and non-spatial information such as extinction risk, population trend, or native/non-native status provided by IUCN Red List assessments.
-While all species represented in IUCN distribution maps have also been evaluated for Red List inclusion, we found that only `r round(nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))/nrow(spp_am) * 100, 1)`% (`r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` of `r nrow(spp_am)`) of AquaMaps species have been similarly evaluated, severely limiting the catalog of available species maps for studies of biodiversity and ecosystem health.
-
-Broad taxonomic inclusion is critical for understanding distribution and health of ecosystems at any spatial scale, but global and large regional studies of species diversity also require representative sampling across regions.
-To compare the spatial representation of the two data collections directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
-For the AquaMaps data collection, we determined per-cell species richness by summing all species with non-zero probability of occurrence.
-We represented relative distribution of species richness for each data collection by plotting average species count against latitude and longitude.
-To highlight the differences between the two data collections, we also examined the relative difference in per-cell species richness against latitude and longitude.
-
-![](figures/all_aquamaps_map.png)
-
-![](figures/all_iucn_map.png)
-
-The balance of species represented in IUCN spatial data skews toward tropical latitudes compared to the distribution represented in AquaMaps distribution maps.
-This may be driven by the fact that the taxonomic distribution of IUCN distribution maps focuses heavily on coral reef-associated species (see fig.
-XXX).
-While both data collections indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps data collection shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN data collection.
-
-Each data collection offers spatial distribution information for large numbers of species.
-However, the data collections vary in terms of taxonomic coverage and regional coverage.
-For spatial assessments of biodiversity, the choice of one data collection over the other is likely to create significantly different results.
-For studies confined to a narrow range of taxa or to a narrow spatial scale, one data collection may offer an advantage over the other in the number of species maps available.
-For global scale biodiversity studies, however, the selection of one data collection over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
-___find a reference that describes what makes a "good" data collection for global biodiversity, e.g.
-OHI___ _or maybe species richness vs diversity vs "health" or whatnot_
+``` {r abstract, child = 'ms_3_methods.Rmd'}
+```
 
@@
 
-## Analysis of spatial alignment by species
-
-### Defining spatial alignment between the two data collections
-
-<!--
-AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, --> For species described in both spatial data collections, we would expect to see some spatial correlation, both in the global pattern of species distribution and the extent of species range.
-Large discrepancies in distribution and range area could indicate species that require further study to consolidate expert knowledge and improve species distribution models.
-
-Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
-Assigning "presence" values using the same methodology above, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
-
-Overlaying distribution maps from the two data sets for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the ratio of the smaller range that overlapped with the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
-
-$$Match_{dist} = \frac{A_{overlap}}{A_{larger}} * 100\%$$
-
-$$Match_{area} = \frac{A_{smaller}}{A_{larger}} * 100\%$$
-
-For a species whose distribution is well understood, and whose range is described in both data collections, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
+``` {r abstract, child = 'ms_4_figs.Rmd'}
+```
 
@@
 
-For each of the `r nrow(spp_map_pairs)` species whose range is described by both data collections, we calculated the distribution alignment and the area alignment, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
-
-![](figures/AreaRatiovsPercOverlap.png)
-
-___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
-
-We divided the plot into quadrants based on the median values for each axis to differentiate between four different qualities of alignment:
-
-``` {r calc medians and quadrant counts, echo = FALSE}
-dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
-area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
-q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
+``` {r abstract, child = 'ms_5_refs.Rmd'}
 ```
-
-| Description              |  Distribution match  |     Area match       |
-| :----------------------: | :------------------: | :------------------: |
-| q1: Well-aligned         | >= `r dist_median` % | >= `r area_median` % |
-| q2: Distribution-aligned |  < `r dist_median` % | >= `r area_median` % |
-| q3: Area-aligned         | >= `r dist_median` % |  < `r area_median` % |
-| q4: Poorly-aligned       |  < `r dist_median` % |  < `r area_median` % |
-
-___caption?___
-
-*The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
-For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
-(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The other quadrants each suggest a different story.
-Species map pairs that fall within the upper left quadrant (quadrant 2: well-matched in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
-(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The lower right quadrant (quadrant 3: well-matched in area but poorly matched in distribution) includes species for which the two maps generally agree in range area, but disagree in where those ranges occur.
-Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being more problematic than map pairs in either of the upper quadrants.
-(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
-
-![](figures/barchart_spp_risk_quads.png)
-
-![](figures/barchart_spp_gp_quads.png)
-
-Breaking down the quadrants by IUCN extinction risk categories, species with higher extinction risk tend to be better matched between the two data collections; this is not surprising, as these at-risk species are likely to face closer expert scrutiny than species ranked as Least Concern.
-
-When broken down by taxonomic group, we found that certain taxonomic groups were far more likely than others to be spatially well-matched; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2), .
-Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two data collections, predicted areas are not well-matched; AquaMaps accounts for seafloor depth as a limiting factor, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, and thus likely overestimate the actual distribution.
-
-___Other hypotheses/case studies to highlight other possible mechanistic differences?___
-
------
-
-## AquaMaps: Effect of changing "presence" threshold on area representation
-
-Why should anyone care about AquaMaps thresholds. ___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
-
-_what would it mean if there were non-linearities or generally linear trend?_
-
-![](figures/AMThreshold.png) ___add a mean/median vline on this?___ do we need this figure?
-it looks cool though...
-why broken down by risk category?
-Ditch NE category since most studies probably care about the species health as well as distribution?
-
-___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
-
-![](figures/scatter_AMThreshold_mockup.png)
-Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
-Do regression - shallow slope indicates that sensitivity is less than 1:1.
-
------
-
-# Implications
-
-## Application to OHI
-
-The global Ocean Health Index (Halpern et al.
-2012), an index made up of 10 goals, utilizes both of these datasets to inform the Biodiversity goal.
-The biodiversity goal is made up of two parts; Species and Habitats.
-As it is currently calculated, the Species subgoal uses the IUCN range map data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
-For those species that are are represented within the IUCN range maps, AquaMaps distribution maps were used.
-Species with no spatial data in either data collection were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
-
-This decision to use IUCN data preferentially to AquaMaps data was made based on the fact that experts are developing the IUCN maps and AquaMaps data is based off a computer model, which in some cases is then consulted by an expert and appropriately altered.
-___is this really the reasoning behind the decision?
-let's get the real reason in here___
-
-Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
-
-| Scenario              | Priority data source | AquaMaps presence threshold |
-| --------------------- | :------------------: | :-------------------------: |
-| Scenario 0 (current)  | IUCN                 |  >= 40%                     |
-| Scenario 1            | IUCN                 |   > 0%                      |
-| Scenario 2            | AquaMaps             |  >= 40%                     |
-| Scenario 3            | AquaMaps             |   > 0%                      |
-
-Scenario 1 shows the effect of reducing the threshold for AquaMaps presence.
-This will always increase the apparent range of any AquaMaps species in the analysis.
-The slight decrease in average score suggests increased spatial representation of species at greater risk of extinction.
-
-Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.
-This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so we could predict an average decrease in range for these overlapping species.
-_The slight bump in mean score indicates that a slight increased representation of low-risk species counterbalances a decreased representation of high-risk species._ ___can we really make any generalizations here?
-not really anything useful without further analysis___
-
-Scenario 3 shows the combination of the two previous scenarios.  The larger decrease in score indicates that the ___hmm gotta think about this a little more - come back to it___
-
------
-
-## Application to MPA Gap Analysis
-
-___yeah still gotta do something on this one___
-
------
-
-# Conclusion
-
------
-
-#References
-
-__Key AquaMaps publications:__
-
-Kaschner, K., D.P. Tittensor, J. Ready, T Gerrodette and B. Worm (2011). Current and Future Patterns of Global Marine Mammal Biodiversity. PLoS ONE 6(5): e19653. PDF
-
-Ready, J., K.
-Kaschner, A.B.
-South, P.D Eastwood, T.
-Rees, J.
-Rius, E.
-Agbayani, S.
-Kullander and R.
-Froese (2010).
-Predicting the distributions of marine organisms at the global scale.
-Ecological Modelling 221(3): 467-478.
-PDF
-
-__Papers based on AquaMaps:__ ___Use these for section 4?___
-
-Jones, M.C., S.R.
-Dyeb, J.K.
-Pinnegar and W.W.L.
-Cheung (2012).
-Modelling commercial fish distributions: Prediction and assessment using different approaches.
-Ecological Modelling 225(2012): 133-145.
-PDF
-
-Coll, M., C. Piroddi, J. Steenbeek, K. Kaschner, F. Ben Rais Lasram et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. PDF
-
-Martin C.S., Fletcher R., Jones M.C., Kaschner K., Sullivan E., Tittensor D.P., Mcowen C., Geffert J.L., van Bochove J.W., Thomas H., Blyth S., Ravillious C., Tolley M., Stanwell-Smith D.
-(2014).
-Manual of marine nd coastal datasets of biodiversity importance.
-May 2014 release.
-Cambridge (UK): UNEP World Conservation Monitoring Centre.
-28 pp.
-(+ 4 annexes totalling 174 pp.
-and one e-supplement).
-PDF
-
-#References
-
-García Molinos, Jorge, Benjamin S.
-Halpern, David S.
-Schoeman, Christopher J.
-Brown, Wolfgang Kiessling, Pippa J.
-Moore, John M.
-Pandolfi, Elvira S.
-Poloczanska, Anthony J.
-Richardson, and Michael T.
-Burrows.
-“Climate Velocity and the Future Global Redistribution of Marine Biodiversity.” Nature Climate Change advance online publication (August 31, 2015).
-doi:10.1038/nclimate2769.
-
-Halpern, Benjamin S., Catherine Longo, Darren Hardy, Karen L.
-McLeod, Jameal F.
-Samhouri, Steven K.
-Katona, Kristin Kleisner, et al.
-“An Index to Assess the Health and Benefits of the Global Ocean.” Nature 488, no.
-7413 (August 30, 2012): 615–20.
-doi:10.1038/nature11397.
-
-<!--
-Halpin, P.N., A.J.
-Read, E.
-Fujioka, B.D.
-Best, B.
-Donnelly, L.J.
-Hazen, C.
-Kot, K.
-Urian, E.
-LaBrecque, A.
-Dimatteo, J.
-Cleary, C.
-Good, L.B.
-Crowder, and K.D.
-Hyrenbach.
-2009.
-OBIS-SEAMAP: The world data center for marine mammal, sea bird, and sea turtle distributions.
-Oceanography 22(2):104-115
--->
-
-Klein, Carissa J., Christopher J.
-Brown, Benjamin S.
-Halpern, Daniel B.
-Segan, Jennifer McGowan, Maria Beger, and James E.M.
-Watson.
-“Shortfalls in the Global Protected Area Network at Representing Marine Biodiversity.” Scientific Reports 5 (December 3, 2015): 17539.
-doi:10.1038/srep17539.
-
-Selig, Elizabeth R., Catherine Longo, Benjamin S.
-Halpern, Benjamin D.
-Best, Darren Hardy, Cristiane T.
-Elfes, Courtney Scarborough, Kristin M.
-Kleisner, and Steven K.
-Katona.
-“Assessing Global Marine Biodiversity Status within a Coupled Socio-Ecological Perspective.” PLoS ONE 8, no.
-4 (April 11, 2013): e60284.
-doi:10.1371/journal.pone.0060284.
-
-* _IUCN reference:_ 
-    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
-    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources

### 2015-12-21 ms_1_abst_sig.Rmd
dividing up manuscript into sections

--- 2015-12-17 ms_outline2.Rmd
@@
----
-title: "A comparative analysis of AquaMaps and IUCN species distribution maps"
-author: "JA, CO, BH"
-output:
-  word_document: null
-  pdf_document:
-    number_sections: yes
-    toc: no
-  html_document:
-    highlight: haddock
-    number_sections: yes
-    theme: spacelab
-    toc: yes
----
+## Abstract
 
-``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
+_get to this later..._
 
-# library(ggplot2)
-# library(maps)
-# library(readr)
-# library(raster)
-# library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-
-# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-#            'Darwin'  = '/Volumes/data_edit',
-#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-# dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-# dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
-# dir_fig <- file.path(getwd(), 'figures')
-```
-
-``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.
-
-data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
-
-### generate other lists
-
-spp_all_file <- 'data/spp_all_2015.csv'
-spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
-
-iucn_marine_file <- 'data/spp_iucn_marine.csv'
-spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
-
-iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
-  rename(sciname = binomial) %>%
-  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
-  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
-
-spp_am_file <- 'data/spp_am.csv'
-spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
-
-```
+-----
 
@@
 
-_This is important because of some reason or another...!!!_
-
------
-
-# Introduction
-
-<!--By 2050 ___some threshold year?___ our oceans will __in some way, find a study that predicts big changes?__ be very different from the oceans we enjoy today.--> Climate change, ocean temperature rise, ocean acidification, sea ice loss, human impacts from fishing and introduction of nonnative species are likely to significantly alter the global distribution of marine species - but how well do we really understand the current distribution?
-Understanding species abundances and distributions is a critical component to assessing global biodiversity and detecting patterns and trends through time.
-<!--Many efforts to map terrestrial, freshwater and marine species from local to global scales inform our understanding of taxonomic assemblages, biogeographical limits to species ranges, suitable habitat for species.-->
-
-AquaMaps and IUCN marine species distribution maps are both global-scale spatial data collections that have been used to analyze marine species status (Halpern et al.
-2012, Selig et al.
-2013), global biodiversity patterns and trends (other references), predicting range shifts (Molinos et al.
-2015), and setting conservation priorities (Klein et al.
-2015).
-The two data collections ostensibly describe the same information, but significant differences in methodology, intent, and taxonomic and geographic sampling could lead to dramatically different understandings of our marine ecosystems.
-
-<!--_Differences in methodology (model-driven vs expert knowledge-driven); differences in intent (what does it communicate); differences in represented species and regions_-->
-
-To better understand the similarities and differences between the two data collections, we examined three fundamental questions:
-
-* How do the two spatial data collections compare in their representation of global spatial distribution of species diversity? 
-* How do the two spatial data collections compare in their representation of species diversity across taxonomic groups?
-* For species represented in both collections, how closely do the spatial distribution maps align?
-
-The Ocean Health Index and MPA Gap Analysis study both provide case studies examined the implications of alternate data use scenarios on two large-scale studies of global biodiversity and conservation.
-
-# Understanding each dataset separately
-
-The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.
-Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.
-These maps are typically based on species occurrence records from databases such as ___OBIS___ and ___GBIF___; IUCN provides general guidelines to facilitate consistency in how the maps are generated.
-The IUCN presents these maps as 'limits of distribution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed within the polygon.
-While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
-
-IUCN publishes spatial distribution map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
-in which at least 90% of the species within the taxonomic group have been evaluated.
-While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN had published species distribution maps for nearly 4500 marine species across 24 taxonomic groups.
-(For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.)
-
-AquaMaps (Kaschner et al.
-2013) offers species range maps based on modeled relative environmental suitability.
-For each assessed species, data from occurrence records such as OBIS and GBIF, published species databases such as FishBase, and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, sea ice concentration, and distance to land.
-The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence, with cell values ranging from 0 to 1.00.
-These maps can then be reviewed by experts to fine-tune the parameters to better match known occurrences.
-As of December 2015, AquaMaps current native distribution maps have been produced for nearly 23,000 species, though only 5,600 of these species have been evaluated for the IUCN Red List.
-_See (Kaschner et al.
-2013) for more detailed methods._
-
------
-
-# Methods
-
-## Analysis of taxonomic and regional inclusion
-
-### Overlap between assessed species
-
-Due to differences in both methodology and intent, there are significant differences in the coverage afforded by the species distribution maps published by AquaMaps and those published by the IUCN.
-This is immediately clear from an examination of the number of species maps offered by each source, and the small proportion of species with maps from both sources.
-
-| Map set   | n~total~                | n~Red_List~                                                       |
-| :-------: | ----------------------: | :---------------------------------------------------------------: | 
-| AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` | 
-| IUCN      | `r nrow(spp_iucn_maps)` | all                                                               | 
-| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))` | all                       |
-
-### Taxonomic distribution between data collections
-
-Studies of species richness and biodiversity depend on sampling a representative cross section of ecosystems, and thus benefit from a broad representation across taxonomic groups.
-To examine the overall taxonomic distribution across the spatial data collections, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
-Since biodiversity studies often rely on both spatial distribution and indicators of species health, we further divided the data by whether species had been evaluated for the IUCN Red List of Threatened Species.
-
-![](figures/bar_by_taxa_all.png)
-
-![](figures/bar_by_taxa_assessed.png)
-
-The IUCN only releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed.
-As such, spatial data for many taxonomic classes are unavailable, and within a class, the assessed sub-groups may not represent the entire class.
-For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
-However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
-
-The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
-
-Some global-scale studies of biodiversity, such as the Ocean Health Index, require both spatial distribution information and non-spatial information such as extinction risk, population trend, or native/non-native status provided by IUCN Red List assessments.
-While all species represented in IUCN distribution maps have also been evaluated for Red List inclusion, we found that only `r round(nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))/nrow(spp_am) * 100, 1)`% (`r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` of `r nrow(spp_am)`) of AquaMaps species have been similarly evaluated, severely limiting the catalog of available species maps for studies of biodiversity and ecosystem health.
-
-Broad taxonomic inclusion is critical for understanding distribution and health of ecosystems at any spatial scale, but global and large regional studies of species diversity also require representative sampling across regions.
-To compare the spatial representation of the two data collections directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
-For the AquaMaps data collection, we determined per-cell species richness by summing all species with non-zero probability of occurrence.
-We represented relative distribution of species richness for each data collection by plotting average species count against latitude and longitude.
-To highlight the differences between the two data collections, we also examined the relative difference in per-cell species richness against latitude and longitude.
-
-![](figures/all_aquamaps_map.png)
-
-![](figures/all_iucn_map.png)
-
-The balance of species represented in IUCN spatial data skews toward tropical latitudes compared to the distribution represented in AquaMaps distribution maps.
-This may be driven by the fact that the taxonomic distribution of IUCN distribution maps focuses heavily on coral reef-associated species (see fig.
-XXX).
-While both data collections indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps data collection shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN data collection.
-
-Each data collection offers spatial distribution information for large numbers of species.
-However, the data collections vary in terms of taxonomic coverage and regional coverage.
-For spatial assessments of biodiversity, the choice of one data collection over the other is likely to create significantly different results.
-For studies confined to a narrow range of taxa or to a narrow spatial scale, one data collection may offer an advantage over the other in the number of species maps available.
-For global scale biodiversity studies, however, the selection of one data collection over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
-___find a reference that describes what makes a "good" data collection for global biodiversity, e.g.
-OHI___ _or maybe species richness vs diversity vs "health" or whatnot_
-
------
-
-## Analysis of spatial alignment by species
-
-### Defining spatial alignment between the two data collections
-
-<!--
-AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, --> For species described in both spatial data collections, we would expect to see some spatial correlation, both in the global pattern of species distribution and the extent of species range.
-Large discrepancies in distribution and range area could indicate species that require further study to consolidate expert knowledge and improve species distribution models.
-
-Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
-Assigning "presence" values using the same methodology above, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
-
-Overlaying distribution maps from the two data sets for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the ratio of the smaller range that overlapped with the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
-
-$$Match_{dist} = \frac{A_{overlap}}{A_{larger}} * 100\%$$
-
-$$Match_{area} = \frac{A_{smaller}}{A_{larger}} * 100\%$$
-
-For a species whose distribution is well understood, and whose range is described in both data collections, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
-
------
-
-For each of the `r nrow(spp_map_pairs)` species whose range is described by both data collections, we calculated the distribution alignment and the area alignment, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
-
-![](figures/AreaRatiovsPercOverlap.png)
-
-___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
-
-We divided the plot into quadrants based on the median values for each axis to differentiate between four different qualities of alignment:
-
-``` {r calc medians and quadrant counts, echo = FALSE}
-dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
-area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
-q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
-```
-
-| Description              |  Distribution match  |     Area match       |
-| :----------------------: | :------------------: | :------------------: |
-| q1: Well-aligned         | >= `r dist_median` % | >= `r area_median` % |
-| q2: Distribution-aligned |  < `r dist_median` % | >= `r area_median` % |
-| q3: Area-aligned         | >= `r dist_median` % |  < `r area_median` % |
-| q4: Poorly-aligned       |  < `r dist_median` % |  < `r area_median` % |
-
-___caption?___
-
-*The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
-For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
-(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The other quadrants each suggest a different story.
-Species map pairs that fall within the upper left quadrant (quadrant 2: well-matched in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
-(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The lower right quadrant (quadrant 3: well-matched in area but poorly matched in distribution) includes species for which the two maps generally agree in range area, but disagree in where those ranges occur.
-Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being more problematic than map pairs in either of the upper quadrants.
-(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
-
-![](figures/barchart_spp_risk_quads.png)
-
-![](figures/barchart_spp_gp_quads.png)
-
-Breaking down the quadrants by IUCN extinction risk categories, species with higher extinction risk tend to be better matched between the two data collections; this is not surprising, as these at-risk species are likely to face closer expert scrutiny than species ranked as Least Concern.
-
-When broken down by taxonomic group, we found that certain taxonomic groups were far more likely than others to be spatially well-matched; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2), .
-Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two data collections, predicted areas are not well-matched; AquaMaps accounts for seafloor depth as a limiting factor, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, and thus likely overestimate the actual distribution.
-
-___Other hypotheses/case studies to highlight other possible mechanistic differences?___
-
------
-
-## AquaMaps: Effect of changing "presence" threshold on area representation
-
-Why should anyone care about AquaMaps thresholds. ___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
-
-_what would it mean if there were non-linearities or generally linear trend?_
-
-![](figures/AMThreshold.png) ___add a mean/median vline on this?___ do we need this figure?
-it looks cool though...
-why broken down by risk category?
-Ditch NE category since most studies probably care about the species health as well as distribution?
-
-___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
-
-![](figures/scatter_AMThreshold_mockup.png)
-Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
-Do regression - shallow slope indicates that sensitivity is less than 1:1.
-
------
-
-# Implications
-
-## Application to OHI
-
-The global Ocean Health Index (Halpern et al.
-2012), an index made up of 10 goals, utilizes both of these datasets to inform the Biodiversity goal.
-The biodiversity goal is made up of two parts; Species and Habitats.
-As it is currently calculated, the Species subgoal uses the IUCN range map data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
-For those species that are are represented within the IUCN range maps, AquaMaps distribution maps were used.
-Species with no spatial data in either data collection were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
-
-This decision to use IUCN data preferentially to AquaMaps data was made based on the fact that experts are developing the IUCN maps and AquaMaps data is based off a computer model, which in some cases is then consulted by an expert and appropriately altered.
-___is this really the reasoning behind the decision?
-let's get the real reason in here___
-
-Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
-
-| Scenario              | Priority data source | AquaMaps presence threshold |
-| --------------------- | :------------------: | :-------------------------: |
-| Scenario 0 (current)  | IUCN                 |  >= 40%                     |
-| Scenario 1            | IUCN                 |   > 0%                      |
-| Scenario 2            | AquaMaps             |  >= 40%                     |
-| Scenario 3            | AquaMaps             |   > 0%                      |
-
-Scenario 1 shows the effect of reducing the threshold for AquaMaps presence.
-This will always increase the apparent range of any AquaMaps species in the analysis.
-The slight decrease in average score suggests increased spatial representation of species at greater risk of extinction.
-
-Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.
-This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so we could predict an average decrease in range for these overlapping species.
-_The slight bump in mean score indicates that a slight increased representation of low-risk species counterbalances a decreased representation of high-risk species._ ___can we really make any generalizations here?
-not really anything useful without further analysis___
-
-Scenario 3 shows the combination of the two previous scenarios.  The larger decrease in score indicates that the ___hmm gotta think about this a little more - come back to it___
-
------
-
-## Application to MPA Gap Analysis
-
-___yeah still gotta do something on this one___
-
------
-
-# Conclusion
-
------
-
-#References
-
-__Key AquaMaps publications:__
-
-Kaschner, K., D.P. Tittensor, J. Ready, T Gerrodette and B. Worm (2011). Current and Future Patterns of Global Marine Mammal Biodiversity. PLoS ONE 6(5): e19653. PDF
-
-Ready, J., K.
-Kaschner, A.B.
-South, P.D Eastwood, T.
-Rees, J.
-Rius, E.
-Agbayani, S.
-Kullander and R.
-Froese (2010).
-Predicting the distributions of marine organisms at the global scale.
-Ecological Modelling 221(3): 467-478.
-PDF
-
-__Papers based on AquaMaps:__ ___Use these for section 4?___
-
-Jones, M.C., S.R.
-Dyeb, J.K.
-Pinnegar and W.W.L.
-Cheung (2012).
-Modelling commercial fish distributions: Prediction and assessment using different approaches.
-Ecological Modelling 225(2012): 133-145.
-PDF
-
-Coll, M., C. Piroddi, J. Steenbeek, K. Kaschner, F. Ben Rais Lasram et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. PDF
-
-Martin C.S., Fletcher R., Jones M.C., Kaschner K., Sullivan E., Tittensor D.P., Mcowen C., Geffert J.L., van Bochove J.W., Thomas H., Blyth S., Ravillious C., Tolley M., Stanwell-Smith D.
-(2014).
-Manual of marine nd coastal datasets of biodiversity importance.
-May 2014 release.
-Cambridge (UK): UNEP World Conservation Monitoring Centre.
-28 pp.
-(+ 4 annexes totalling 174 pp.
-and one e-supplement).
-PDF
-
-#References
-
-García Molinos, Jorge, Benjamin S.
-Halpern, David S.
-Schoeman, Christopher J.
-Brown, Wolfgang Kiessling, Pippa J.
-Moore, John M.
-Pandolfi, Elvira S.
-Poloczanska, Anthony J.
-Richardson, and Michael T.
-Burrows.
-“Climate Velocity and the Future Global Redistribution of Marine Biodiversity.” Nature Climate Change advance online publication (August 31, 2015).
-doi:10.1038/nclimate2769.
-
-Halpern, Benjamin S., Catherine Longo, Darren Hardy, Karen L.
-McLeod, Jameal F.
-Samhouri, Steven K.
-Katona, Kristin Kleisner, et al.
-“An Index to Assess the Health and Benefits of the Global Ocean.” Nature 488, no.
-7413 (August 30, 2012): 615–20.
-doi:10.1038/nature11397.
-
-<!--
-Halpin, P.N., A.J.
-Read, E.
-Fujioka, B.D.
-Best, B.
-Donnelly, L.J.
-Hazen, C.
-Kot, K.
-Urian, E.
-LaBrecque, A.
-Dimatteo, J.
-Cleary, C.
-Good, L.B.
-Crowder, and K.D.
-Hyrenbach.
-2009.
-OBIS-SEAMAP: The world data center for marine mammal, sea bird, and sea turtle distributions.
-Oceanography 22(2):104-115
--->
-
-Klein, Carissa J., Christopher J.
-Brown, Benjamin S.
-Halpern, Daniel B.
-Segan, Jennifer McGowan, Maria Beger, and James E.M.
-Watson.
-“Shortfalls in the Global Protected Area Network at Representing Marine Biodiversity.” Scientific Reports 5 (December 3, 2015): 17539.
-doi:10.1038/srep17539.
-
-Selig, Elizabeth R., Catherine Longo, Benjamin S.
-Halpern, Benjamin D.
-Best, Darren Hardy, Cristiane T.
-Elfes, Courtney Scarborough, Kristin M.
-Kleisner, and Steven K.
-Katona.
-“Assessing Global Marine Biodiversity Status within a Coupled Socio-Ecological Perspective.” PLoS ONE 8, no.
-4 (April 11, 2013): e60284.
-doi:10.1371/journal.pone.0060284.
-
-* _IUCN reference:_ 
-    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
-    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources
+_subset of the abstract_

### 2015-12-21 ms_2_intro_res_disc.Rmd
dividing up manuscript into sections

--- 2015-12-19 ms_outline2.Rmd
@@
----
-title: "Aligning species range data to better serve science and conservation"
-author: "Casey O'Hara, Jamie Afflerbach, Ben Halpern"
-output:
-  pdf_document:
-    number_sections: yes
-    toc: no
-  html_document:
-    highlight: haddock
-    number_sections: yes
-    theme: spacelab
-    toc: yes
-  word_document: null
----
-
-``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
-
-# library(ggplot2)
-# library(maps)
-# library(readr)
-# library(raster)
-# library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-
-# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-#            'Darwin'  = '/Volumes/data_edit',
-#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-# dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-# dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
-# dir_fig <- file.path(getwd(), 'figures')
-```
-
-``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.
-
-data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
-
-### generate other lists
-
-spp_all_file <- 'data/spp_all_2015.csv'
-spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
-
-iucn_marine_file <- 'data/spp_iucn_marine.csv'
-spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
-
-iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
-  rename(sciname = binomial) %>%
-  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
-  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
-
-spp_am_file <- 'data/spp_am.csv'
-spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
-
-```
-
-## Abstract
-
-_get to this later..._
-
------
-
-## Significance
-
-_subset of the abstract_
-
------
-
 # Introduction
@@
 ___yeah still gotta do something on this one___
-
------
-
-# Conclusion
-
------
-
-#References
-
-__Key AquaMaps publications:__
-
-Kaschner, K., D.P. Tittensor, J. Ready, T Gerrodette and B. Worm (2011). Current and Future Patterns of Global Marine Mammal Biodiversity. PLoS ONE 6(5): e19653. PDF
-
-Ready, J., K.
-Kaschner, A.B.
-South, P.D Eastwood, T.
-Rees, J.
-Rius, E.
-Agbayani, S.
-Kullander and R.
-Froese (2010).
-Predicting the distributions of marine organisms at the global scale.
-Ecological Modelling 221(3): 467-478.
-PDF
-
-__Papers based on AquaMaps:__ ___Use these for section 4?___
-
-Jones, M.C., S.R.
-Dyeb, J.K.
-Pinnegar and W.W.L.
-Cheung (2012).
-Modelling commercial fish distributions: Prediction and assessment using different approaches.
-Ecological Modelling 225(2012): 133-145.
-PDF
-
-Coll, M., C. Piroddi, J. Steenbeek, K. Kaschner, F. Ben Rais Lasram et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. PDF
-
-Martin C.S., Fletcher R., Jones M.C., Kaschner K., Sullivan E., Tittensor D.P., Mcowen C., Geffert J.L., van Bochove J.W., Thomas H., Blyth S., Ravillious C., Tolley M., Stanwell-Smith D.
-(2014).
-Manual of marine nd coastal datasets of biodiversity importance.
-May 2014 release.
-Cambridge (UK): UNEP World Conservation Monitoring Centre.
-28 pp.
-(+ 4 annexes totalling 174 pp.
-and one e-supplement).
-PDF
-
------
-
-García Molinos, Jorge, Benjamin S.
-Halpern, David S.
-Schoeman, Christopher J.
-Brown, Wolfgang Kiessling, Pippa J.
-Moore, John M.
-Pandolfi, Elvira S.
-Poloczanska, Anthony J.
-Richardson, and Michael T.
-Burrows.
-“Climate Velocity and the Future Global Redistribution of Marine Biodiversity.” Nature Climate Change advance online publication (August 31, 2015).
-doi:10.1038/nclimate2769.
-
-Halpern, Benjamin S., Catherine Longo, Darren Hardy, Karen L.
-McLeod, Jameal F.
-Samhouri, Steven K.
-Katona, Kristin Kleisner, et al.
-“An Index to Assess the Health and Benefits of the Global Ocean.” Nature 488, no.
-7413 (August 30, 2012): 615–20.
-doi:10.1038/nature11397.
-
-<!--
-Halpin, P.N., A.J.
-Read, E.
-Fujioka, B.D.
-Best, B.
-Donnelly, L.J.
-Hazen, C.
-Kot, K.
-Urian, E.
-LaBrecque, A.
-Dimatteo, J.
-Cleary, C.
-Good, L.B.
-Crowder, and K.D.
-Hyrenbach.
-2009.
-OBIS-SEAMAP: The world data center for marine mammal, sea bird, and sea turtle distributions.
-Oceanography 22(2):104-115
--->
-
-Klein, Carissa J., Christopher J.
-Brown, Benjamin S.
-Halpern, Daniel B.
-Segan, Jennifer McGowan, Maria Beger, and James E.M.
-Watson.
-“Shortfalls in the Global Protected Area Network at Representing Marine Biodiversity.” Scientific Reports 5 (December 3, 2015): 17539.
-doi:10.1038/srep17539.
-
-Selig, Elizabeth R., Catherine Longo, Benjamin S.
-Halpern, Benjamin D.
-Best, Darren Hardy, Cristiane T.
-Elfes, Courtney Scarborough, Kristin M.
-Kleisner, and Steven K.
-Katona.
-“Assessing Global Marine Biodiversity Status within a Coupled Socio-Ecological Perspective.” PLoS ONE 8, no.
-4 (April 11, 2013): e60284.
-doi:10.1371/journal.pone.0060284.
-
-* _IUCN reference:_ 
-    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
-    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources
-
-# Figures
-
-Figure 1 (a, b):
-
-![](figures/bar_by_taxa_all.png)
-
-![](figures/bar_by_taxa_assessed.png)
-
-Figure 2 (a, b):
-
-![](figures/all_aquamaps_map.png)
-
-![](figures/all_iucn_map.png)
-
-Figure 3:
-
-![](figures/AreaRatiovsPercOverlap.png)
-
-___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
-
-Figure 4 (a, b):
-
-![](figures/barchart_spp_risk_quads.png)
-
-![](figures/barchart_spp_gp_quads.png)
-
-Figure 5 (a, b):
-
-![](figures/AMThreshold.png)
-
-![](figures/ribbon_AMThreshold.png)
-
-Figure 6
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
-
-Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
-
-| Scenario              | Priority data source | AquaMaps presence threshold |
-| --------------------- | :------------------: | :-------------------------: |
-| Scenario 0 (current)  | IUCN                 |  >= 40%                     |
-| Scenario 1            | IUCN                 |   > 0%                      |
-| Scenario 2            | AquaMaps             |  >= 40%                     |
-| Scenario 3            | AquaMaps             |   > 0%                      |

### 2015-12-21 ms_4_figs.Rmd
dividing up manuscript into sections

--- 2015-12-17 ms_outline2.Rmd
@@
----
-title: "A comparative analysis of AquaMaps and IUCN species distribution maps"
-author: "JA, CO, BH"
-output:
-  word_document: null
-  pdf_document:
-    number_sections: yes
-    toc: no
-  html_document:
-    highlight: haddock
-    number_sections: yes
-    theme: spacelab
-    toc: yes
----
+# Figures
 
-``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
-
-# library(ggplot2)
-# library(maps)
-# library(readr)
-# library(raster)
-# library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-
-# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-#            'Darwin'  = '/Volumes/data_edit',
-#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-# dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-# dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
-# dir_fig <- file.path(getwd(), 'figures')
-```
-
-``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.
-
-data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
-
-### generate other lists
-
-spp_all_file <- 'data/spp_all_2015.csv'
-spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
-
-iucn_marine_file <- 'data/spp_iucn_marine.csv'
-spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
-
-iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
-  rename(sciname = binomial) %>%
-  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
-  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
-
-spp_am_file <- 'data/spp_am.csv'
-spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
-
-```
-
-## Significance
-
-_This is important because of some reason or another...!!!_
-
------
-
-# Introduction
-
-<!--By 2050 ___some threshold year?___ our oceans will __in some way, find a study that predicts big changes?__ be very different from the oceans we enjoy today.--> Climate change, ocean temperature rise, ocean acidification, sea ice loss, human impacts from fishing and introduction of nonnative species are likely to significantly alter the global distribution of marine species - but how well do we really understand the current distribution?
-Understanding species abundances and distributions is a critical component to assessing global biodiversity and detecting patterns and trends through time.
-<!--Many efforts to map terrestrial, freshwater and marine species from local to global scales inform our understanding of taxonomic assemblages, biogeographical limits to species ranges, suitable habitat for species.-->
-
-AquaMaps and IUCN marine species distribution maps are both global-scale spatial data collections that have been used to analyze marine species status (Halpern et al.
-2012, Selig et al.
-2013), global biodiversity patterns and trends (other references), predicting range shifts (Molinos et al.
-2015), and setting conservation priorities (Klein et al.
-2015).
-The two data collections ostensibly describe the same information, but significant differences in methodology, intent, and taxonomic and geographic sampling could lead to dramatically different understandings of our marine ecosystems.
-
-<!--_Differences in methodology (model-driven vs expert knowledge-driven); differences in intent (what does it communicate); differences in represented species and regions_-->
-
-To better understand the similarities and differences between the two data collections, we examined three fundamental questions:
-
-* How do the two spatial data collections compare in their representation of global spatial distribution of species diversity? 
-* How do the two spatial data collections compare in their representation of species diversity across taxonomic groups?
-* For species represented in both collections, how closely do the spatial distribution maps align?
-
-The Ocean Health Index and MPA Gap Analysis study both provide case studies examined the implications of alternate data use scenarios on two large-scale studies of global biodiversity and conservation.
-
-# Understanding each dataset separately
-
-The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.
-Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.
-These maps are typically based on species occurrence records from databases such as ___OBIS___ and ___GBIF___; IUCN provides general guidelines to facilitate consistency in how the maps are generated.
-The IUCN presents these maps as 'limits of distribution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed within the polygon.
-While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
-
-IUCN publishes spatial distribution map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
-in which at least 90% of the species within the taxonomic group have been evaluated.
-While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN had published species distribution maps for nearly 4500 marine species across 24 taxonomic groups.
-(For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.)
-
-AquaMaps (Kaschner et al.
-2013) offers species range maps based on modeled relative environmental suitability.
-For each assessed species, data from occurrence records such as OBIS and GBIF, published species databases such as FishBase, and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, sea ice concentration, and distance to land.
-The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence, with cell values ranging from 0 to 1.00.
-These maps can then be reviewed by experts to fine-tune the parameters to better match known occurrences.
-As of December 2015, AquaMaps current native distribution maps have been produced for nearly 23,000 species, though only 5,600 of these species have been evaluated for the IUCN Red List.
-_See (Kaschner et al.
-2013) for more detailed methods._
-
------
-
-# Methods
-
-## Analysis of taxonomic and regional inclusion
-
-### Overlap between assessed species
-
-Due to differences in both methodology and intent, there are significant differences in the coverage afforded by the species distribution maps published by AquaMaps and those published by the IUCN.
-This is immediately clear from an examination of the number of species maps offered by each source, and the small proportion of species with maps from both sources.
-
-| Map set   | n~total~                | n~Red_List~                                                       |
-| :-------: | ----------------------: | :---------------------------------------------------------------: | 
-| AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` | 
-| IUCN      | `r nrow(spp_iucn_maps)` | all                                                               | 
-| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))` | all                       |
-
-### Taxonomic distribution between data collections
-
-Studies of species richness and biodiversity depend on sampling a representative cross section of ecosystems, and thus benefit from a broad representation across taxonomic groups.
-To examine the overall taxonomic distribution across the spatial data collections, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
-Since biodiversity studies often rely on both spatial distribution and indicators of species health, we further divided the data by whether species had been evaluated for the IUCN Red List of Threatened Species.
+Figure 1 (a, b):
 
@@
 
-The IUCN only releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed.
-As such, spatial data for many taxonomic classes are unavailable, and within a class, the assessed sub-groups may not represent the entire class.
-For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
-However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
-
-The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
-
-Some global-scale studies of biodiversity, such as the Ocean Health Index, require both spatial distribution information and non-spatial information such as extinction risk, population trend, or native/non-native status provided by IUCN Red List assessments.
-While all species represented in IUCN distribution maps have also been evaluated for Red List inclusion, we found that only `r round(nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))/nrow(spp_am) * 100, 1)`% (`r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` of `r nrow(spp_am)`) of AquaMaps species have been similarly evaluated, severely limiting the catalog of available species maps for studies of biodiversity and ecosystem health.
-
-Broad taxonomic inclusion is critical for understanding distribution and health of ecosystems at any spatial scale, but global and large regional studies of species diversity also require representative sampling across regions.
-To compare the spatial representation of the two data collections directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
-For the AquaMaps data collection, we determined per-cell species richness by summing all species with non-zero probability of occurrence.
-We represented relative distribution of species richness for each data collection by plotting average species count against latitude and longitude.
-To highlight the differences between the two data collections, we also examined the relative difference in per-cell species richness against latitude and longitude.
+Figure 2 (a, b):
 
@@
 
-The balance of species represented in IUCN spatial data skews toward tropical latitudes compared to the distribution represented in AquaMaps distribution maps.
-This may be driven by the fact that the taxonomic distribution of IUCN distribution maps focuses heavily on coral reef-associated species (see fig.
-XXX).
-While both data collections indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps data collection shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN data collection.
-
-Each data collection offers spatial distribution information for large numbers of species.
-However, the data collections vary in terms of taxonomic coverage and regional coverage.
-For spatial assessments of biodiversity, the choice of one data collection over the other is likely to create significantly different results.
-For studies confined to a narrow range of taxa or to a narrow spatial scale, one data collection may offer an advantage over the other in the number of species maps available.
-For global scale biodiversity studies, however, the selection of one data collection over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
-___find a reference that describes what makes a "good" data collection for global biodiversity, e.g.
-OHI___ _or maybe species richness vs diversity vs "health" or whatnot_
-
------
-
-## Analysis of spatial alignment by species
-
-### Defining spatial alignment between the two data collections
-
-<!--
-AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, --> For species described in both spatial data collections, we would expect to see some spatial correlation, both in the global pattern of species distribution and the extent of species range.
-Large discrepancies in distribution and range area could indicate species that require further study to consolidate expert knowledge and improve species distribution models.
-
-Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
-Assigning "presence" values using the same methodology above, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
-
-Overlaying distribution maps from the two data sets for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the ratio of the smaller range that overlapped with the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
-
-$$Match_{dist} = \frac{A_{overlap}}{A_{larger}} * 100\%$$
-
-$$Match_{area} = \frac{A_{smaller}}{A_{larger}} * 100\%$$
-
-For a species whose distribution is well understood, and whose range is described in both data collections, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
-
------
-
-For each of the `r nrow(spp_map_pairs)` species whose range is described by both data collections, we calculated the distribution alignment and the area alignment, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
+Figure 3:
 
@@
 
-We divided the plot into quadrants based on the median values for each axis to differentiate between four different qualities of alignment:
-
-``` {r calc medians and quadrant counts, echo = FALSE}
-dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
-area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
-q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
-```
-
-| Description              |  Distribution match  |     Area match       |
-| :----------------------: | :------------------: | :------------------: |
-| q1: Well-aligned         | >= `r dist_median` % | >= `r area_median` % |
-| q2: Distribution-aligned |  < `r dist_median` % | >= `r area_median` % |
-| q3: Area-aligned         | >= `r dist_median` % |  < `r area_median` % |
-| q4: Poorly-aligned       |  < `r dist_median` % |  < `r area_median` % |
-
-___caption?___
-
-*The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
-For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
-(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The other quadrants each suggest a different story.
-Species map pairs that fall within the upper left quadrant (quadrant 2: well-matched in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
-(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The lower right quadrant (quadrant 3: well-matched in area but poorly matched in distribution) includes species for which the two maps generally agree in range area, but disagree in where those ranges occur.
-Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being more problematic than map pairs in either of the upper quadrants.
-(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
+Figure 4 (a, b):
 
@@
 
-Breaking down the quadrants by IUCN extinction risk categories, species with higher extinction risk tend to be better matched between the two data collections; this is not surprising, as these at-risk species are likely to face closer expert scrutiny than species ranked as Least Concern.
+Figure 5 (a, b):
 
-When broken down by taxonomic group, we found that certain taxonomic groups were far more likely than others to be spatially well-matched; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2), .
-Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two data collections, predicted areas are not well-matched; AquaMaps accounts for seafloor depth as a limiting factor, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, and thus likely overestimate the actual distribution.
+![](figures/AMThreshold.png)
 
-___Other hypotheses/case studies to highlight other possible mechanistic differences?___
+![](figures/ribbon_AMThreshold.png)
 
------
+Figure 6
 
-## AquaMaps: Effect of changing "presence" threshold on area representation
-
-Why should anyone care about AquaMaps thresholds. ___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
-
-_what would it mean if there were non-linearities or generally linear trend?_
-
-![](figures/AMThreshold.png) ___add a mean/median vline on this?___ do we need this figure?
-it looks cool though...
-why broken down by risk category?
-Ditch NE category since most studies probably care about the species health as well as distribution?
-
-___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
-
-![](figures/scatter_AMThreshold_mockup.png)
-Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
-Do regression - shallow slope indicates that sensitivity is less than 1:1.
-
------
-
-# Implications
-
-## Application to OHI
-
-The global Ocean Health Index (Halpern et al.
-2012), an index made up of 10 goals, utilizes both of these datasets to inform the Biodiversity goal.
-The biodiversity goal is made up of two parts; Species and Habitats.
-As it is currently calculated, the Species subgoal uses the IUCN range map data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
-For those species that are are represented within the IUCN range maps, AquaMaps distribution maps were used.
-Species with no spatial data in either data collection were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
-
-This decision to use IUCN data preferentially to AquaMaps data was made based on the fact that experts are developing the IUCN maps and AquaMaps data is based off a computer model, which in some cases is then consulted by an expert and appropriately altered.
-___is this really the reasoning behind the decision?
-let's get the real reason in here___
+![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
 
 Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
 
@@
 | Scenario 3            | AquaMaps             |   > 0%                      |
-
-Scenario 1 shows the effect of reducing the threshold for AquaMaps presence.
-This will always increase the apparent range of any AquaMaps species in the analysis.
-The slight decrease in average score suggests increased spatial representation of species at greater risk of extinction.
-
-Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.
-This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so we could predict an average decrease in range for these overlapping species.
-_The slight bump in mean score indicates that a slight increased representation of low-risk species counterbalances a decreased representation of high-risk species._ ___can we really make any generalizations here?
-not really anything useful without further analysis___
-
-Scenario 3 shows the combination of the two previous scenarios.  The larger decrease in score indicates that the ___hmm gotta think about this a little more - come back to it___
-
------
-
-## Application to MPA Gap Analysis
-
-___yeah still gotta do something on this one___
-
------
-
-# Conclusion
-
------
-
-#References
-
-__Key AquaMaps publications:__
-
-Kaschner, K., D.P. Tittensor, J. Ready, T Gerrodette and B. Worm (2011). Current and Future Patterns of Global Marine Mammal Biodiversity. PLoS ONE 6(5): e19653. PDF
-
-Ready, J., K.
-Kaschner, A.B.
-South, P.D Eastwood, T.
-Rees, J.
-Rius, E.
-Agbayani, S.
-Kullander and R.
-Froese (2010).
-Predicting the distributions of marine organisms at the global scale.
-Ecological Modelling 221(3): 467-478.
-PDF
-
-__Papers based on AquaMaps:__ ___Use these for section 4?___
-
-Jones, M.C., S.R.
-Dyeb, J.K.
-Pinnegar and W.W.L.
-Cheung (2012).
-Modelling commercial fish distributions: Prediction and assessment using different approaches.
-Ecological Modelling 225(2012): 133-145.
-PDF
-
-Coll, M., C. Piroddi, J. Steenbeek, K. Kaschner, F. Ben Rais Lasram et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. PDF
-
-Martin C.S., Fletcher R., Jones M.C., Kaschner K., Sullivan E., Tittensor D.P., Mcowen C., Geffert J.L., van Bochove J.W., Thomas H., Blyth S., Ravillious C., Tolley M., Stanwell-Smith D.
-(2014).
-Manual of marine nd coastal datasets of biodiversity importance.
-May 2014 release.
-Cambridge (UK): UNEP World Conservation Monitoring Centre.
-28 pp.
-(+ 4 annexes totalling 174 pp.
-and one e-supplement).
-PDF
-
-#References
-
-García Molinos, Jorge, Benjamin S.
-Halpern, David S.
-Schoeman, Christopher J.
-Brown, Wolfgang Kiessling, Pippa J.
-Moore, John M.
-Pandolfi, Elvira S.
-Poloczanska, Anthony J.
-Richardson, and Michael T.
-Burrows.
-“Climate Velocity and the Future Global Redistribution of Marine Biodiversity.” Nature Climate Change advance online publication (August 31, 2015).
-doi:10.1038/nclimate2769.
-
-Halpern, Benjamin S., Catherine Longo, Darren Hardy, Karen L.
-McLeod, Jameal F.
-Samhouri, Steven K.
-Katona, Kristin Kleisner, et al.
-“An Index to Assess the Health and Benefits of the Global Ocean.” Nature 488, no.
-7413 (August 30, 2012): 615–20.
-doi:10.1038/nature11397.
-
-<!--
-Halpin, P.N., A.J.
-Read, E.
-Fujioka, B.D.
-Best, B.
-Donnelly, L.J.
-Hazen, C.
-Kot, K.
-Urian, E.
-LaBrecque, A.
-Dimatteo, J.
-Cleary, C.
-Good, L.B.
-Crowder, and K.D.
-Hyrenbach.
-2009.
-OBIS-SEAMAP: The world data center for marine mammal, sea bird, and sea turtle distributions.
-Oceanography 22(2):104-115
--->
-
-Klein, Carissa J., Christopher J.
-Brown, Benjamin S.
-Halpern, Daniel B.
-Segan, Jennifer McGowan, Maria Beger, and James E.M.
-Watson.
-“Shortfalls in the Global Protected Area Network at Representing Marine Biodiversity.” Scientific Reports 5 (December 3, 2015): 17539.
-doi:10.1038/srep17539.
-
-Selig, Elizabeth R., Catherine Longo, Benjamin S.
-Halpern, Benjamin D.
-Best, Darren Hardy, Cristiane T.
-Elfes, Courtney Scarborough, Kristin M.
-Kleisner, and Steven K.
-Katona.
-“Assessing Global Marine Biodiversity Status within a Coupled Socio-Ecological Perspective.” PLoS ONE 8, no.
-4 (April 11, 2013): e60284.
-doi:10.1371/journal.pone.0060284.
-
-* _IUCN reference:_ 
-    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
-    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources

### 2015-12-21 ms_5_refs.Rmd
dividing up manuscript into sections

--- 2015-12-17 ms_outline2.Rmd
@@
----
-title: "A comparative analysis of AquaMaps and IUCN species distribution maps"
-author: "JA, CO, BH"
-output:
-  word_document: null
-  pdf_document:
-    number_sections: yes
-    toc: no
-  html_document:
-    highlight: haddock
-    number_sections: yes
-    theme: spacelab
-    toc: yes
----
-
-``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
-
-# library(ggplot2)
-# library(maps)
-# library(readr)
-# library(raster)
-# library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-
-# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-#            'Darwin'  = '/Volumes/data_edit',
-#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-# dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-# dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
-# dir_fig <- file.path(getwd(), 'figures')
-```
-
-``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.
-
-data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
-
-### generate other lists
-
-spp_all_file <- 'data/spp_all_2015.csv'
-spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
-
-iucn_marine_file <- 'data/spp_iucn_marine.csv'
-spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
-
-iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
-  rename(sciname = binomial) %>%
-  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
-  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
-
-spp_am_file <- 'data/spp_am.csv'
-spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
-
-```
-
-## Significance
-
-_This is important because of some reason or another...!!!_
-
------
-
-# Introduction
-
-<!--By 2050 ___some threshold year?___ our oceans will __in some way, find a study that predicts big changes?__ be very different from the oceans we enjoy today.--> Climate change, ocean temperature rise, ocean acidification, sea ice loss, human impacts from fishing and introduction of nonnative species are likely to significantly alter the global distribution of marine species - but how well do we really understand the current distribution?
-Understanding species abundances and distributions is a critical component to assessing global biodiversity and detecting patterns and trends through time.
-<!--Many efforts to map terrestrial, freshwater and marine species from local to global scales inform our understanding of taxonomic assemblages, biogeographical limits to species ranges, suitable habitat for species.-->
-
-AquaMaps and IUCN marine species distribution maps are both global-scale spatial data collections that have been used to analyze marine species status (Halpern et al.
-2012, Selig et al.
-2013), global biodiversity patterns and trends (other references), predicting range shifts (Molinos et al.
-2015), and setting conservation priorities (Klein et al.
-2015).
-The two data collections ostensibly describe the same information, but significant differences in methodology, intent, and taxonomic and geographic sampling could lead to dramatically different understandings of our marine ecosystems.
-
-<!--_Differences in methodology (model-driven vs expert knowledge-driven); differences in intent (what does it communicate); differences in represented species and regions_-->
-
-To better understand the similarities and differences between the two data collections, we examined three fundamental questions:
-
-* How do the two spatial data collections compare in their representation of global spatial distribution of species diversity? 
-* How do the two spatial data collections compare in their representation of species diversity across taxonomic groups?
-* For species represented in both collections, how closely do the spatial distribution maps align?
-
-The Ocean Health Index and MPA Gap Analysis study both provide case studies examined the implications of alternate data use scenarios on two large-scale studies of global biodiversity and conservation.
-
-# Understanding each dataset separately
-
-The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.
-Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.
-These maps are typically based on species occurrence records from databases such as ___OBIS___ and ___GBIF___; IUCN provides general guidelines to facilitate consistency in how the maps are generated.
-The IUCN presents these maps as 'limits of distribution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed within the polygon.
-While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
-
-IUCN publishes spatial distribution map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
-in which at least 90% of the species within the taxonomic group have been evaluated.
-While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN had published species distribution maps for nearly 4500 marine species across 24 taxonomic groups.
-(For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.)
-
-AquaMaps (Kaschner et al.
-2013) offers species range maps based on modeled relative environmental suitability.
-For each assessed species, data from occurrence records such as OBIS and GBIF, published species databases such as FishBase, and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, sea ice concentration, and distance to land.
-The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence, with cell values ranging from 0 to 1.00.
-These maps can then be reviewed by experts to fine-tune the parameters to better match known occurrences.
-As of December 2015, AquaMaps current native distribution maps have been produced for nearly 23,000 species, though only 5,600 of these species have been evaluated for the IUCN Red List.
-_See (Kaschner et al.
-2013) for more detailed methods._
-
------
-
-# Methods
-
-## Analysis of taxonomic and regional inclusion
-
-### Overlap between assessed species
-
-Due to differences in both methodology and intent, there are significant differences in the coverage afforded by the species distribution maps published by AquaMaps and those published by the IUCN.
-This is immediately clear from an examination of the number of species maps offered by each source, and the small proportion of species with maps from both sources.
-
-| Map set   | n~total~                | n~Red_List~                                                       |
-| :-------: | ----------------------: | :---------------------------------------------------------------: | 
-| AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` | 
-| IUCN      | `r nrow(spp_iucn_maps)` | all                                                               | 
-| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))` | all                       |
-
-### Taxonomic distribution between data collections
-
-Studies of species richness and biodiversity depend on sampling a representative cross section of ecosystems, and thus benefit from a broad representation across taxonomic groups.
-To examine the overall taxonomic distribution across the spatial data collections, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
-Since biodiversity studies often rely on both spatial distribution and indicators of species health, we further divided the data by whether species had been evaluated for the IUCN Red List of Threatened Species.
-
-![](figures/bar_by_taxa_all.png)
-
-![](figures/bar_by_taxa_assessed.png)
-
-The IUCN only releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed.
-As such, spatial data for many taxonomic classes are unavailable, and within a class, the assessed sub-groups may not represent the entire class.
-For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
-However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
-
-The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
-
-Some global-scale studies of biodiversity, such as the Ocean Health Index, require both spatial distribution information and non-spatial information such as extinction risk, population trend, or native/non-native status provided by IUCN Red List assessments.
-While all species represented in IUCN distribution maps have also been evaluated for Red List inclusion, we found that only `r round(nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))/nrow(spp_am) * 100, 1)`% (`r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` of `r nrow(spp_am)`) of AquaMaps species have been similarly evaluated, severely limiting the catalog of available species maps for studies of biodiversity and ecosystem health.
-
-Broad taxonomic inclusion is critical for understanding distribution and health of ecosystems at any spatial scale, but global and large regional studies of species diversity also require representative sampling across regions.
-To compare the spatial representation of the two data collections directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
-For the AquaMaps data collection, we determined per-cell species richness by summing all species with non-zero probability of occurrence.
-We represented relative distribution of species richness for each data collection by plotting average species count against latitude and longitude.
-To highlight the differences between the two data collections, we also examined the relative difference in per-cell species richness against latitude and longitude.
-
-![](figures/all_aquamaps_map.png)
-
-![](figures/all_iucn_map.png)
-
-The balance of species represented in IUCN spatial data skews toward tropical latitudes compared to the distribution represented in AquaMaps distribution maps.
-This may be driven by the fact that the taxonomic distribution of IUCN distribution maps focuses heavily on coral reef-associated species (see fig.
-XXX).
-While both data collections indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps data collection shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN data collection.
-
-Each data collection offers spatial distribution information for large numbers of species.
-However, the data collections vary in terms of taxonomic coverage and regional coverage.
-For spatial assessments of biodiversity, the choice of one data collection over the other is likely to create significantly different results.
-For studies confined to a narrow range of taxa or to a narrow spatial scale, one data collection may offer an advantage over the other in the number of species maps available.
-For global scale biodiversity studies, however, the selection of one data collection over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
-___find a reference that describes what makes a "good" data collection for global biodiversity, e.g.
-OHI___ _or maybe species richness vs diversity vs "health" or whatnot_
-
------
-
-## Analysis of spatial alignment by species
-
-### Defining spatial alignment between the two data collections
-
-<!--
-AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, --> For species described in both spatial data collections, we would expect to see some spatial correlation, both in the global pattern of species distribution and the extent of species range.
-Large discrepancies in distribution and range area could indicate species that require further study to consolidate expert knowledge and improve species distribution models.
-
-Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
-Assigning "presence" values using the same methodology above, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
-
-Overlaying distribution maps from the two data sets for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the ratio of the smaller range that overlapped with the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
-
-$$Match_{dist} = \frac{A_{overlap}}{A_{larger}} * 100\%$$
-
-$$Match_{area} = \frac{A_{smaller}}{A_{larger}} * 100\%$$
-
-For a species whose distribution is well understood, and whose range is described in both data collections, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
-
------
-
-For each of the `r nrow(spp_map_pairs)` species whose range is described by both data collections, we calculated the distribution alignment and the area alignment, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
-
-![](figures/AreaRatiovsPercOverlap.png)
-
-___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
-
-We divided the plot into quadrants based on the median values for each axis to differentiate between four different qualities of alignment:
-
-``` {r calc medians and quadrant counts, echo = FALSE}
-dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
-area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
-q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
-```
-
-| Description              |  Distribution match  |     Area match       |
-| :----------------------: | :------------------: | :------------------: |
-| q1: Well-aligned         | >= `r dist_median` % | >= `r area_median` % |
-| q2: Distribution-aligned |  < `r dist_median` % | >= `r area_median` % |
-| q3: Area-aligned         | >= `r dist_median` % |  < `r area_median` % |
-| q4: Poorly-aligned       |  < `r dist_median` % |  < `r area_median` % |
-
-___caption?___
-
-*The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
-For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
-(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The other quadrants each suggest a different story.
-Species map pairs that fall within the upper left quadrant (quadrant 2: well-matched in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
-(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The lower right quadrant (quadrant 3: well-matched in area but poorly matched in distribution) includes species for which the two maps generally agree in range area, but disagree in where those ranges occur.
-Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being more problematic than map pairs in either of the upper quadrants.
-(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
-*The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
-
-![](figures/barchart_spp_risk_quads.png)
-
-![](figures/barchart_spp_gp_quads.png)
-
-Breaking down the quadrants by IUCN extinction risk categories, species with higher extinction risk tend to be better matched between the two data collections; this is not surprising, as these at-risk species are likely to face closer expert scrutiny than species ranked as Least Concern.
-
-When broken down by taxonomic group, we found that certain taxonomic groups were far more likely than others to be spatially well-matched; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2), .
-Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two data collections, predicted areas are not well-matched; AquaMaps accounts for seafloor depth as a limiting factor, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, and thus likely overestimate the actual distribution.
-
-___Other hypotheses/case studies to highlight other possible mechanistic differences?___
-
------
-
-## AquaMaps: Effect of changing "presence" threshold on area representation
-
-Why should anyone care about AquaMaps thresholds. ___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
-
-_what would it mean if there were non-linearities or generally linear trend?_
-
-![](figures/AMThreshold.png) ___add a mean/median vline on this?___ do we need this figure?
-it looks cool though...
-why broken down by risk category?
-Ditch NE category since most studies probably care about the species health as well as distribution?
-
-___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis
-
-![](figures/scatter_AMThreshold_mockup.png)
-Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
-Do regression - shallow slope indicates that sensitivity is less than 1:1.
-
------
-
-# Implications
-
-## Application to OHI
-
-The global Ocean Health Index (Halpern et al.
-2012), an index made up of 10 goals, utilizes both of these datasets to inform the Biodiversity goal.
-The biodiversity goal is made up of two parts; Species and Habitats.
-As it is currently calculated, the Species subgoal uses the IUCN range map data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
-For those species that are are represented within the IUCN range maps, AquaMaps distribution maps were used.
-Species with no spatial data in either data collection were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
-
-This decision to use IUCN data preferentially to AquaMaps data was made based on the fact that experts are developing the IUCN maps and AquaMaps data is based off a computer model, which in some cases is then consulted by an expert and appropriately altered.
-___is this really the reasoning behind the decision?
-let's get the real reason in here___
-
-Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
-
-| Scenario              | Priority data source | AquaMaps presence threshold |
-| --------------------- | :------------------: | :-------------------------: |
-| Scenario 0 (current)  | IUCN                 |  >= 40%                     |
-| Scenario 1            | IUCN                 |   > 0%                      |
-| Scenario 2            | AquaMaps             |  >= 40%                     |
-| Scenario 3            | AquaMaps             |   > 0%                      |
-
-Scenario 1 shows the effect of reducing the threshold for AquaMaps presence.
-This will always increase the apparent range of any AquaMaps species in the analysis.
-The slight decrease in average score suggests increased spatial representation of species at greater risk of extinction.
-
-Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.
-This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so we could predict an average decrease in range for these overlapping species.
-_The slight bump in mean score indicates that a slight increased representation of low-risk species counterbalances a decreased representation of high-risk species._ ___can we really make any generalizations here?
-not really anything useful without further analysis___
-
-Scenario 3 shows the combination of the two previous scenarios.  The larger decrease in score indicates that the ___hmm gotta think about this a little more - come back to it___
-
------
-
-## Application to MPA Gap Analysis
-
-___yeah still gotta do something on this one___
-
------
-
-# Conclusion
-
------
-
 #References
@@
 
-#References
+-----
+
+Hurlbert 2007
+
+Jetz 2008
+
+Pimm 2014
+
+Rondinini 2006
 

### 2015-12-21 data_explore/data_explore_am_thresh.Rmd

--- 2015-12-18 R/data_explore_am_thresh.Rmd
@@
                                'NE' = 'grey70')) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
   labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
@@
 
-ggsave('figures/AMThreshold.png', width = 6, height = 4)
+print(hist_area_vs_threshold)
 
-print(hist_area_vs_threshold)
+ggsave(file.path(dir_fig, 'AMThreshold.png'), width = 6, height = 4)
+```
+
+```{r hist_lost_area_w_threshold_ne_dd_removed, warning = FALSE, message = FALSE, echo = FALSE}
 
@@
 
-ggsave('figures/AMThreshold2.png', width = 6, height = 4)
+ggsave(file.path(dir_fig, 'AMThreshold2.png'), width = 6, height = 4)
 
@@
   geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
-  ylim(0, 1) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
   labs(title = 'Mean (median) species range after presence threshold cutoff',
@@
 
-ggsave('figures/scatter_AMThreshold.png', width = 6, height = 4)
+ggsave(file.path(dir_fig, 'scatter_AMThreshold.png'), width = 6, height = 4)
 
@@
 
-```{r boxplot_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
+```{r ribbon_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
 
@@
            label = 'R^2 = .995') + 
-  ylim(0, 1) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
   labs(title = 'Relative species range after presence threshold',
@@
 
-ggsave('figures/ribbon_AMThreshold.png', width = 6, height = 4)
+ggsave(file.path(dir_fig, 'ribbon_AMThreshold.png'), width = 6, height = 4)
 

### 2015-12-21 data_explore/data_explore_pt1.Rmd

--- 2015-12-01 R/data_explore_pt1.Rmd
@@
 
+<!--
 ####How this breaks down between AquaMaps and IUCN
@@
 Here is an example of a single species, the yellow speckled chromis (*Chromis alpha*)  
+-->
 
-```{r map chromis alpha, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+```{r map chromis alpha, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 
+<!--
 For `r spp`, `r round(pi[1], 2)`% of the total area is covered by both IUCN and AquaMaps, `r round(p1[2, ], 2)`% is covered just by AquaMaps and `r round(p1[3, ], 2)`% is covered just by IUCN.
+-->
 

### 2015-12-21 data_explore/data_explore_pt3.Rmd

@@
 Plotting species from q1 (well-matched) (see data_explore_quadmaps.html for plots)
+
 ```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

### 2015-12-21 data_explore/data_explore.Rmd
updated to ggplot2 2.0 and redid plots

@@
   ggtheme_plot + 
-  geom_bar(stat = 'bin', alpha = .5) +
+  geom_bar(stat = 'count', alpha = .5) +
   scale_fill_manual(values = c('q1 well-matched'   = 'green4', 

### 2015-12-21 data_explore/data_explore_am_thresh.Rmd
updated to ggplot2 2.0 and redid plots

--- 2015-12-18 R/data_explore_am_thresh.Rmd
@@
 
-```{r aquamaps probabilities, warning = FALSE, message = FALSE, cache.lazy = FALSE, cache = TRUE, verbose = FALSE, results = FALSE, echo = FALSE, include = FALSE}
+```{r aquamaps_load_data, warning = FALSE, message = FALSE, echo = FALSE, include = FALSE}
 
@@
                       col_types = '_ccd__', progress = TRUE)
-
-# spp_cells <- fread(input = file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), verbose = FALSE, header = TRUE, 
-#                   colClasses = c('NULL', 'character', NA, 'numeric', 'NULL', 'NULL')) #1 min 49 sec
 
@@
   geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
+  geom_vline(xintercept = spp_df3_mean, color = 'darkred', linetype = 'dashed') + 
+  annotate('text', x = spp_df3_mean, y = .16, hjust = 0, vjust = -.5,
+           angle = 90, size = 2, color = 'grey20',
+           label = 'mean') + 
+  geom_vline(xintercept = spp_df3_med, color = 'grey30', linetype = 'dashed') + 
+  annotate('text', x = spp_df3_med, y = .16, hjust = 0, vjust = -.5,
+           angle = 90, size = 2, color = 'grey20',
+           label = 'median') + 
   scale_fill_manual(values = c('LC' = 'green4', 
@@
   geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
+  geom_vline(xintercept = spp_df3a_mean, color = 'darkred', linetype = 'dashed') + 
+  annotate('text', x = spp_df3a_mean, y = .16, hjust = 0, vjust = -.5,
+           angle = 90, size = 2, color = 'grey20',
+           label = 'mean') + 
+  geom_vline(xintercept = spp_df3a_med, color = 'grey30', linetype = 'dashed') + 
+  annotate('text', x = spp_df3a_med, y = .16, hjust = 0, vjust = -.5,
+           angle = 90, size = 2, color = 'grey20',
+           label = 'median') + 
   scale_fill_manual(values = c('LC' = 'green4', 
@@
   geom_abline(slope = -1, intercept = 1, color = 'red') + 
-  geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
-  ylim(0, 1) +
+  geom_vline(xintercept = 0.40, color = 'blue', linetype = 'dashed') +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
   labs(title = 'Mean (median) species range after presence threshold cutoff',
@@
               color = 'red') + 
-  geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
-  geom_vline(x = 0.50, color = 'blue', linetype = 'dashed') +
-  geom_vline(x = 0.60, color = 'blue', linetype = 'dashed') +
+  geom_vline(xintercept = 0.40, color = 'blue', linetype = 'dashed') +
+  geom_vline(xintercept = 0.50, color = 'blue', linetype = 'dashed') +
+  geom_vline(xintercept = 0.60, color = 'blue', linetype = 'dashed') +
   annotate('text', x = .40, y = .05, hjust = 0, vjust = -.5,

### 2015-12-21 data_explore/data_explore.Rmd
rescaled graphs and graph titles

@@
 
-ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 6, height = 4)
+ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 6, height = 3.5)
 
@@
        filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
-       width = 6, height = 4)
+       width = 6, height = 3.5)
 
@@
        filename = file.path(dir_fig, 'barchart_spp_risk_quads.png'),
-       width = 6, height = 1.5)
+       width = 6, height = 1.75)
 
@@
   labs(x = 'Class', y = 'Percent of mapped species',
-       title = 'Distribution of species maps by taxonomic group (class)')
+       title = 'Species maps by taxonomic class')
 print(bar_taxa2)
-ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), height = 4, width = 8)
+ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), height = 3, width = 6)
 
@@
   scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
-  labs(x = 'Class', y = 'Percent of mapped and assessed species',
-       title = 'Distribution of species maps by taxonomic group (class)')
+  labs(x = 'Class', y = 'Percent of assessed mapped species',
+       title = 'Species maps by taxonomic class, Red List-assessed only')
 print(bar_taxa3)
-ggsave(filename = file.path(dir_fig, 'bar_by_taxa_assessed.png'), height = 4, width = 8)
+ggsave(filename = file.path(dir_fig, 'bar_by_taxa_assessed.png'), height = 3, width = 6)
 

### 2015-12-21 data_explore/data_explore_am_thresh.Rmd
rescaled graphs and graph titles

@@
 
-ggsave(file.path(dir_fig, 'AMThreshold.png'), width = 6, height = 4)
+ggsave(file.path(dir_fig, 'AMThreshold.png'), width = 6, height = 3.5)
 ```
@@
 
-ggsave(file.path(dir_fig, 'AMThreshold2.png'), width = 6, height = 4)
+ggsave(file.path(dir_fig, 'AMThreshold2.png'), width = 6, height = 3.5)
 
@@
 
-ggsave(file.path(dir_fig, 'scatter_AMThreshold.png'), width = 6, height = 4)
+ggsave(file.path(dir_fig, 'scatter_AMThreshold.png'), width = 6, height = 3.5)
 
@@
 
-ggsave(file.path(dir_fig, 'ribbon_AMThreshold.png'), width = 6, height = 4)
+ggsave(file.path(dir_fig, 'ribbon_AMThreshold.png'), width = 6, height = 3.5)
 

### 2015-12-21 data_explore/data_explore_fxns.R

@@ -3,7 +3,7 @@
 ### generic theme for all plots
 ggtheme_basic <- theme(axis.ticks = element_blank(),
         text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-        plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
+        plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
         legend.position = 'right')
 
 ggtheme_plot <- ggtheme_basic + 

### 2015-12-21 ms_2_intro_res_disc.Rmd
updating manuscript

@@
 
-Due to differences in both methodology and intent, significant differences exist between AquaMaps and IUCN species distribution maps (Fig.
+Differences in methodology and intent drive significant differences between AquaMaps and IUCN species distribution maps (Fig.
 1).
-Only `r n_spp_both` species were covered by both datasets (`r p_both_am`% of total Aquamaps species; `p_both_iucn`% of total IUCN species; Fig.
+Only `r n_spp_both` species were included in both datasets (`r p_both_am`% of total Aquamaps species; `r p_both_iucn`% of total IUCN species; Fig.
 1, 2).
@@
 
-<!--
-| Map set   | n~total~       | n~Red_List~     |
-| :-------: | -------------: | :-------------: | 
-| AquaMaps  | `r n_spp_am`   | `r n_spp_am_rl` | 
-| IUCN      | `r n_spp_iucn` | all             | 
-| Both      | `r n_spp_both` | all             |
--->
-
 ### Taxonomic distribution between datasets
-
-___Either move to brief methods section above, and maybe even shorten - BH___: To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
-We further divided the data by whether species had been evaluated for the IUCN Red List of Threatened Species.
-
-___Either cut or move to data description section of methods above.
-- BH___: The IUCN only releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed.
-As such, spatial data for many taxonomic classes are unavailable, and within a class, the assessed sub-groups may not represent the entire class.
-___Move to supplementary materials.
-- BH___: For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
-However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
-
-___Probably just cut, but maybe include in Suppl Materials.
-- BH___: The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
-
-___Move to methods section above.
-- BH___: To compare the spatial representation of the two datasets directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
-For the AquaMaps dataset, we determined per-cell species richness by summing all species with non-zero probability of occurrence.
-We represented relative distribution of species richness for each dataset by plotting average species count against latitude and longitude.
-To highlight the differences between the two datasets, we also examined the relative difference in per-cell species richness against latitude and longitude.
 
@@
 
-<!--
-AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, --> For species described in both spatial datasets, we would expect to see spatial correlation both in the global pattern of species distribution and the extent of species range.
+For species described in both spatial datasets, we would expect to see spatial correlation both in the global pattern of species distribution and the extent of species range.
 Large discrepancies in distribution and range area could indicate species that require further study to consolidate expert knowledge for IUCN and improve the inputs to AquaMaps species distribution models.
@@
 Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
-Assigning "presence" values using the same methodology above, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of paired-map species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
+We found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of paired-map species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
 This aligns with the expectations and early analysis of Kaschner et al (REF).
@@
 
-$$\alpha_{dist} = \frac{A_{small \cap large}}{A_{large}} * 100\%$$
+For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
+Dividing the map-paired species into quadrants based on median values for each axis, we can examine the implications of four different qualities of alignment.
 
-$$\alpha_{area} = \frac{A_{small}}{A_{large}} * 100\%$$
-
-For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
-
------
-
-For each of the `r nrow(spp_map_pairs)` species whose range is described by both datasets, we calculated the distribution alignment and the area alignment, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
-
-We divided the plot into quadrants based on the median values for each axis to differentiate between four different qualities of alignment:
-
-``` {r calc medians and quadrant counts, echo = FALSE}
-dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
-area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
-q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
-```
+<!--
+___moving to figure caption.  Summarize results here.___
 
@@
 For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
-(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The other quadrants each suggest a different story.
-Species map pairs that fall within the upper left quadrant (quadrant 2: well-aligned in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
-(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
+* Species map pairs that fall within the upper left quadrant (quadrant 2: well-aligned in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
 * The lower right quadrant (quadrant 3: well-aligned in area but poorly matched in distribution) includes species for which the paired maps generally agree in range area, but disagree in locations of those ranges.
 Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being lower quality alignment than map pairs in either of the upper quadrants.
-(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
 * The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution.
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
 
-Breaking down the quadrants by IUCN extinction risk categories, species with higher extinction risk tend to be better aligned between the two datasets.
-Well-aligned range map pairs (membership in quadrants 1 and 2) may suggest increased expert scrutiny, though this could be explained as higher perceived risk leading to increased attention, or as increased attention revealing deeper risk.
+-->
+
+Breaking down the quadrants by IUCN extinction risk categories, we found that species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
+Does higher perceived risk lead to increased attention, and thus better understanding of species distribution?
+Or conversely, does increased attention to species distribution reveal deeper risk?
 Likely both mechanisms are at play in a feedback loop.
-In any case, poorly-aligned rangemaps, regardless of extinction risk category, seem to indicate species who could benefit from further expert study.
 
-When broken down by taxonomic group, we found that certain taxonomic groups were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
+Examining quadrants by taxonomic group, we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
 Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two datasets, predicted areas are not well-aligned.
-AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range.
+AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range (see SI).
 A similar mechanism may also drive poor area alignment of other demersal and reef-associated organisms.
 
+Poorly-aligned rangemaps, regardless of taxon or extinction risk category, indicate species who could benefit from further expert study.
+
 ___Other hypotheses/case studies to highlight other possible mechanistic differences?___
+
+Confusion in species names also seems to correlate with spatial misalignment.
+AquaMaps identifies each species map by binomial as well as a unique species ID code.
+While the IUCN assigns a unique numeric code to every Red List-assessed species, this code is not used consistently to identify range map polygons.
+___NOTE: NEW DATA as of Nov 24?
+Need to re-extract all the IUCN species :( ___
 
@@
 AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value indicated by IUCN maps.
-Frequently, studies convert this probability to a simple presence value by assigning a threshold value (XXX references here).
+Frequently, studies convert this probability to a simple presence value by assigning a threshold value (REF references here).
 A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
-For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to approximate the presence criteria used by IUCN.
-To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0.05 to 1.00 and calculated the average species range relative to a zero threshold.
+_For the comparisons above, thresh = 0_
 
-_Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
-Do regression - shallow slope indicates that sensitivity is less than 1:1._
-
-At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a (XXX to XXX %, median = XXX) decrease in represented range, though some species lose nearly their entire range.
-Examining the entire species catalog as a single set, a unit shift in presence threshold results in a generally smaller proportional shift in represented area, indicating a moderate and consistent sensitivity to threshold with no surprising tradeoffs to inform the identification of a "correct" threshold.
+At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a _(XXX to XXX %, median = XXX)_ decrease in represented range, though some species lose nearly their entire range.
+Incrementing the presence threshold from zero to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.
 This pattern may not hold true for all subsets of AquaMaps, however, whether subsetting by taxa or by georegion.
@@
 
-The global Ocean Health Index (Halpern et al.
-2012), an index made up of 10 goals, utilizes both of these datasets to inform the Biodiversity goal.
-The biodiversity goal is made up of two parts; Species and Habitats.
+The global Ocean Health Index (OHI) (Halpern et al.
+2012), an index made up of 10 goals, utilizes both of these datasets to inform the Species subgoal of the Biodiversity goal.
 As it is currently calculated, the Species subgoal uses species spatial distribution data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
-Spatial distributions were gleaned from both IUCN and AquaMaps datasets.
-A probability threshold of 40% is used to determine presence for AquaMaps data.
-For those species that are are represented within both datasets, IUCN distribution maps were used as the preferred data source.
+Spatial distributions were gleaned from both IUCN and AquaMaps datasets, preferring IUCN data for species represented in both data sets.
+OHI uses a probability threshold of 40% to determine presence for AquaMaps data.
 Species with no spatial data in either dataset were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
 
-Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
+___caption?___
+
+* Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
 Reducing the threshold will always increase the apparent range of a species, therefore the slight decrease in average score suggests increased spatial representation of threatened species.
 
-Scenario 2 shows the effect of preferring AquaMaps data over IUCN, while maintaining the same presence threshold.
+* Scenario 2 shows the effect of preferring AquaMaps data over IUCN, while maintaining the same presence threshold.
 This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so many but not all overlapping species will see a decrease in represented range.
@@
 
-Scenario 3 shows the effect of preferring AquaMaps data over IUCN, while eliminating the presence threshold.
+* Scenario 3 shows the effect of preferring AquaMaps data over IUCN, while eliminating the presence threshold.
 Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
@@
 
-Since the Ocean Health Index Species subgoal depends on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted.
-When IUCN data is the preferred data source, only the non-paired AquaMaps species will be affected by threshold changes; and even when AquaMaps is the preferred source, the non-paired IUCN species will dampen the effect of a threshold change.
+___end caption___
+
+_Briefly summarize results_
+
+Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted.
+When IUCN data is the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change.
+But
 
@@
 
-___yeah still gotta do something on this one___
+___do the analysis...___
+
+Predictions:
+
+IUCN overestimates area esp for coastal species?
+More species will be included in MPAs so fewer apparent gap species.
+Included range area inside MPAs will increase.
+Policy implications?
+Less pressure to carefully consider MPA design; less resolution on

### 2015-12-21 ms_3_methods.Rmd
updating manuscript

--- 2015-12-19 ms_outline2.Rmd
@@
----
-title: "Aligning species range data to better serve science and conservation"
-author: "Casey O'Hara, Jamie Afflerbach, Ben Halpern"
-output:
-  pdf_document:
-    number_sections: yes
-    toc: no
-  html_document:
-    highlight: haddock
-    number_sections: yes
-    theme: spacelab
-    toc: yes
-  word_document: null
----
+# Methods
 
-``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
-
-# library(ggplot2)
-# library(maps)
-# library(readr)
-# library(raster)
-# library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-
-# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-#            'Darwin'  = '/Volumes/data_edit',
-#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-# dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-# dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
-# dir_fig <- file.path(getwd(), 'figures')
-```
-
-``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.
-
-data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
-
-### generate other lists
-
-spp_all_file <- 'data/spp_all_2015.csv'
-spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
-
-iucn_marine_file <- 'data/spp_iucn_marine.csv'
-spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
-
-iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
-  rename(sciname = binomial) %>%
-  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
-  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
-
-spp_am_file <- 'data/spp_am.csv'
-spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
-
-```
-
-## Abstract
-
-_get to this later..._
-
------
-
-## Significance
-
-_subset of the abstract_
-
------
-
-# Introduction
-
-Mapping and predicting species’ distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
-Knowing where individuals of a species are, and what allows them to persist there, provides foundational information for understanding species ranges, how best to protect and manage species, and how they may respond to increasing human impacts and a changing climate.
-A rich literature exists tackling the many dimensions of these questions.
-
-One very important outcome of this body of science is the various compiled databases of species range maps.
-In the oceans there are now two such global repositories – Aquamaps (REF) and International Union for Conservation of Nature (IUCN) species distribution maps (REF) – that are used for a wide range of purposes, including assessing marine species status (Halpern et al.
-2012, Selig et al.
-2013), evaluating global biodiversity patterns and trends (REF), predicting range shifts (Molinos et al.
-2015), and setting conservation priorities (Klein et al.
-2015).
-The two data collections ostensibly describe the same information, but significant differences in methodology, intent, and taxonomic and geographic sampling could lead to dramatically different understandings of our marine ecosystems and resulting policy and conservation recommendations
-
-_Importantly, biases in taxanomic or spatial coverage of a dataset could shift management or conservation actions towards places or species that aren’t actually the most in need. [elaborate a bit?]_
-
-To understand the implications of similarities and differences between the two datasets, we examined how they compare in their representation of global spatial distribution of species, and for the relatively small number of species represented in both datasets, how closely their spatial distribution maps align.
-We then reanalyze two recent studies using these datasets based on different versions of the data  to illustrate how and why decisions about which species range data to use affect our understanding of the status of marine biodiversity.
-
-# Overview of the datasets
-
-___We may need to shorten this section to fit this paper into the format guidelines for PNAS.  If so, some of this can be pushed to the Suppl. Materials. - BH___
-
-The IUCN publishes species distribution maps as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.
-Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.
-These maps are typically based on species occurrence records from databases such as the Ocean Biogeographic Information System (OBIS) (REF) and the Global Biodiversity Information Facility (GBIF) (REF); IUCN provides _general guidelines_ ___Do we need to describe these?
-- BH___ to facilitate consistency in how the maps are generated.
-The IUCN presents these maps as 'limits of distribution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed within the polygon.
-While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
-
-IUCN publishes spatial distribution map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
-in which at least 90% of the species within the taxonomic group have been evaluated.
-While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN had published species distribution maps for `r nrow(spp_iucn_maps)` marine species across 24 taxonomic groups.
-For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.
-
-AquaMaps (Kaschner et al.
-2013) offers species range maps based on modeled relative environmental suitability.
-For each assessed species, data from occurrence records such as OBIS and GBIF, published species databases such as FishBase, and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, sea ice concentration, and distance to land.
-The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence, with cell values ranging from 0.00 to 1.00.
-These maps can then be reviewed by experts to fine-tune the parameters to better match known occurrences.
-As of December 2015, AquaMaps current native distribution maps have been produced for `r nrow(spp_am)` species.
-Use of these data to predict species presence or absence requires decisions about which probability threshold to use.
-
------
-
-_Analyses [briefly describe the analyses done]_
-
-* _simple taxanomic comparison (describe very briefly)_
-* _spatial overlap comparison (describe very briefly)_
-* _redo of OHI biodiversity goal (describe briefly what was done previously, and how that was changed for here)_
-* _redo global MPA gap analysis (describe briefly what was done previously, and how that was changed for here)_
-
-# Results and Discussion
-
-### Overlap between assessed species
-
-``` {r calc species numbers, echo = FALSE}
-n_spp_am    <- nrow(spp_am)
-n_spp_am_rl <- nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))
-p_spp_am_rl <- round(n_spp_am_rl/n_spp_am * 100, 1)
-n_spp_iucn  <- nrow(spp_iucn_maps)
-n_spp_both  <- nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))
-p_both_am   <- round(n_spp_both/n_spp_am, 1)
-p_both_iucn <- round(n_spp_both/n_spp_iucn, 1)
-```
-
-Due to differences in both methodology and intent, significant differences exist between AquaMaps and IUCN species distribution maps (Fig.
-1).
-Only `r n_spp_both` species were covered by both datasets (`r p_both_am`% of total Aquamaps species; `p_both_iucn`% of total IUCN species; Fig.
-1, 2).
-__All IUCN-mapped species are also included in Red List species, but only `r p_spp_am_rl`% of AquaMaps species.__
-
-<!--
-| Map set   | n~total~       | n~Red_List~     |
-| :-------: | -------------: | :-------------: | 
-| AquaMaps  | `r n_spp_am`   | `r n_spp_am_rl` | 
-| IUCN      | `r n_spp_iucn` | all             | 
-| Both      | `r n_spp_both` | all             |
--->
-
-### Taxonomic distribution between datasets
+_taxonomic distribution comparison_
 
@@
 ___Either cut or move to data description section of methods above.
-- BH___: The IUCN only releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed.
-As such, spatial data for many taxonomic classes are unavailable, and within a class, the assessed sub-groups may not represent the entire class.
-___Move to supplementary materials.
-- BH___: For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
-However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
+- BH___: The IUCN releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed, to guard against sample bias (though non-comprehensive datasets are available for reptiles and marine fish).
+As such, spatial data for many taxonomic classes remain unavailable, and within a class, the assessed sub-groups may not represent the entire class.
 
-___Probably just cut, but maybe include in Suppl Materials.
-- BH___: The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
+_global distribution spatial comparison_
 
-___Move to methods section above.
-- BH___: To compare the spatial representation of the two datasets directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
+To compare the spatial representation of the two datasets directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
 For the AquaMaps dataset, we determined per-cell species richness by summing all species with non-zero probability of occurrence.
@@
 
-The balance of species represented in IUCN spatial data skews toward tropical latitudes compared to the distribution represented in AquaMaps distribution maps.
-This may be driven by the fact that the taxonomic distribution of IUCN distribution maps focuses heavily on coral reef-associated species (see fig.
-XXX).
-While both datasets indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps dataset shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN dataset.
+For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to approximate the presence criteria used by IUCN.
+To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0.05 to 1.00 and calculated the average species range relative to a zero threshold.
 
-Each dataset offers spatial distribution information for large numbers of species.
-However, the datasets vary in terms of taxonomic coverage and regional coverage.
-For spatial assessments of biodiversity, the choice of one dataset over the other is likely to create significantly different results.
-For studies confined to a narrow range of taxa or to a narrow spatial scale, one dataset may offer an advantage over the other in the number of species maps available.
-For global scale biodiversity studies, however, the selection of one dataset over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
-___find a reference that describes what makes a "good" dataset for global biodiversity, e.g.
-OHI___ _or maybe species richness vs diversity vs "health" or whatnot_
+_map pairs comparison_
 
------
-
-## Analysis of spatial alignment by species
-
-### Defining spatial alignment between the two datasets
-
-<!--
-AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, --> For species described in both spatial datasets, we would expect to see spatial correlation both in the global pattern of species distribution and the extent of species range.
-Large discrepancies in distribution and range area could indicate species that require further study to consolidate expert knowledge for IUCN and improve the inputs to AquaMaps species distribution models.
-
-Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
-Assigning "presence" values using the same methodology above, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of paired-map species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
-This aligns with the expectations and early analysis of Kaschner et al (REF).
-
-Overlaying paired distribution maps for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
+___from results/discussion:___ Overlaying paired distribution maps for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
 
@@
 $$\alpha_{area} = \frac{A_{small}}{A_{large}} * 100\%$$
-
-For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
-
------
-
-For each of the `r nrow(spp_map_pairs)` species whose range is described by both datasets, we calculated the distribution alignment and the area alignment, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
-
-We divided the plot into quadrants based on the median values for each axis to differentiate between four different qualities of alignment:
-
-``` {r calc medians and quadrant counts, echo = FALSE}
-dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
-area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
-q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
-```
-
-* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
-For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
-(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The other quadrants each suggest a different story.
-Species map pairs that fall within the upper left quadrant (quadrant 2: well-aligned in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
-(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The lower right quadrant (quadrant 3: well-aligned in area but poorly matched in distribution) includes species for which the paired maps generally agree in range area, but disagree in locations of those ranges.
-Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being lower quality alignment than map pairs in either of the upper quadrants.
-(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
-
-Breaking down the quadrants by IUCN extinction risk categories, species with higher extinction risk tend to be better aligned between the two datasets.
-Well-aligned range map pairs (membership in quadrants 1 and 2) may suggest increased expert scrutiny, though this could be explained as higher perceived risk leading to increased attention, or as increased attention revealing deeper risk.
-Likely both mechanisms are at play in a feedback loop.
-In any case, poorly-aligned rangemaps, regardless of extinction risk category, seem to indicate species who could benefit from further expert study.
-
-When broken down by taxonomic group, we found that certain taxonomic groups were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
-Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two datasets, predicted areas are not well-aligned.
-AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range.
-A similar mechanism may also drive poor area alignment of other demersal and reef-associated organisms.
-
-___Other hypotheses/case studies to highlight other possible mechanistic differences?___
-
------
-
-## AquaMaps: Effect of changing "presence" threshold on area representation
-
-AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value indicated by IUCN maps.
-Frequently, studies convert this probability to a simple presence value by assigning a threshold value (XXX references here).
-A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
-For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to approximate the presence criteria used by IUCN.
-To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0.05 to 1.00 and calculated the average species range relative to a zero threshold.
-
-_Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
-Do regression - shallow slope indicates that sensitivity is less than 1:1._
-
-At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a (XXX to XXX %, median = XXX) decrease in represented range, though some species lose nearly their entire range.
-Examining the entire species catalog as a single set, a unit shift in presence threshold results in a generally smaller proportional shift in represented area, indicating a moderate and consistent sensitivity to threshold with no surprising tradeoffs to inform the identification of a "correct" threshold.
-This pattern may not hold true for all subsets of AquaMaps, however, whether subsetting by taxa or by georegion.
-
------
-
-# Implications
-
-## Application to OHI
-
-The global Ocean Health Index (Halpern et al.
-2012), an index made up of 10 goals, utilizes both of these datasets to inform the Biodiversity goal.
-The biodiversity goal is made up of two parts; Species and Habitats.
-As it is currently calculated, the Species subgoal uses species spatial distribution data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
-Spatial distributions were gleaned from both IUCN and AquaMaps datasets.
-A probability threshold of 40% is used to determine presence for AquaMaps data.
-For those species that are are represented within both datasets, IUCN distribution maps were used as the preferred data source.
-Species with no spatial data in either dataset were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
-
-Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
-Reducing the threshold will always increase the apparent range of a species, therefore the slight decrease in average score suggests increased spatial representation of threatened species.
-
-Scenario 2 shows the effect of preferring AquaMaps data over IUCN, while maintaining the same presence threshold.
-This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so many but not all overlapping species will see a decrease in represented range.
-The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
-
-Scenario 3 shows the effect of preferring AquaMaps data over IUCN, while eliminating the presence threshold.
-Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
-The large decrease seems to indicate that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.
-
-Since the Ocean Health Index Species subgoal depends on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted.
-When IUCN data is the preferred data source, only the non-paired AquaMaps species will be affected by threshold changes; and even when AquaMaps is the preferred source, the non-paired IUCN species will dampen the effect of a threshold change.
-
------
-
-## Application to MPA Gap Analysis
-
-___yeah still gotta do something on this one___
-
------
-
-# Conclusion
-
------
-
-#References
-
-__Key AquaMaps publications:__
-
-Kaschner, K., D.P. Tittensor, J. Ready, T Gerrodette and B. Worm (2011). Current and Future Patterns of Global Marine Mammal Biodiversity. PLoS ONE 6(5): e19653. PDF
-
-Ready, J., K.
-Kaschner, A.B.
-South, P.D Eastwood, T.
-Rees, J.
-Rius, E.
-Agbayani, S.
-Kullander and R.
-Froese (2010).
-Predicting the distributions of marine organisms at the global scale.
-Ecological Modelling 221(3): 467-478.
-PDF
-
-__Papers based on AquaMaps:__ ___Use these for section 4?___
-
-Jones, M.C., S.R.
-Dyeb, J.K.
-Pinnegar and W.W.L.
-Cheung (2012).
-Modelling commercial fish distributions: Prediction and assessment using different approaches.
-Ecological Modelling 225(2012): 133-145.
-PDF
-
-Coll, M., C. Piroddi, J. Steenbeek, K. Kaschner, F. Ben Rais Lasram et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. PDF
-
-Martin C.S., Fletcher R., Jones M.C., Kaschner K., Sullivan E., Tittensor D.P., Mcowen C., Geffert J.L., van Bochove J.W., Thomas H., Blyth S., Ravillious C., Tolley M., Stanwell-Smith D.
-(2014).
-Manual of marine nd coastal datasets of biodiversity importance.
-May 2014 release.
-Cambridge (UK): UNEP World Conservation Monitoring Centre.
-28 pp.
-(+ 4 annexes totalling 174 pp.
-and one e-supplement).
-PDF
-
------
-
-García Molinos, Jorge, Benjamin S.
-Halpern, David S.
-Schoeman, Christopher J.
-Brown, Wolfgang Kiessling, Pippa J.
-Moore, John M.
-Pandolfi, Elvira S.
-Poloczanska, Anthony J.
-Richardson, and Michael T.
-Burrows.
-“Climate Velocity and the Future Global Redistribution of Marine Biodiversity.” Nature Climate Change advance online publication (August 31, 2015).
-doi:10.1038/nclimate2769.
-
-Halpern, Benjamin S., Catherine Longo, Darren Hardy, Karen L.
-McLeod, Jameal F.
-Samhouri, Steven K.
-Katona, Kristin Kleisner, et al.
-“An Index to Assess the Health and Benefits of the Global Ocean.” Nature 488, no.
-7413 (August 30, 2012): 615–20.
-doi:10.1038/nature11397.
-
-<!--
-Halpin, P.N., A.J.
-Read, E.
-Fujioka, B.D.
-Best, B.
-Donnelly, L.J.
-Hazen, C.
-Kot, K.
-Urian, E.
-LaBrecque, A.
-Dimatteo, J.
-Cleary, C.
-Good, L.B.
-Crowder, and K.D.
-Hyrenbach.
-2009.
-OBIS-SEAMAP: The world data center for marine mammal, sea bird, and sea turtle distributions.
-Oceanography 22(2):104-115
--->
-
-Klein, Carissa J., Christopher J.
-Brown, Benjamin S.
-Halpern, Daniel B.
-Segan, Jennifer McGowan, Maria Beger, and James E.M.
-Watson.
-“Shortfalls in the Global Protected Area Network at Representing Marine Biodiversity.” Scientific Reports 5 (December 3, 2015): 17539.
-doi:10.1038/srep17539.
-
-Selig, Elizabeth R., Catherine Longo, Benjamin S.
-Halpern, Benjamin D.
-Best, Darren Hardy, Cristiane T.
-Elfes, Courtney Scarborough, Kristin M.
-Kleisner, and Steven K.
-Katona.
-“Assessing Global Marine Biodiversity Status within a Coupled Socio-Ecological Perspective.” PLoS ONE 8, no.
-4 (April 11, 2013): e60284.
-doi:10.1371/journal.pone.0060284.
-
-* _IUCN reference:_ 
-    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
-    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources
-
-# Figures
-
-Figure 1 (a, b):
-
-![](figures/bar_by_taxa_all.png)
-
-![](figures/bar_by_taxa_assessed.png)
-
-Figure 2 (a, b):
-
-![](figures/all_aquamaps_map.png)
-
-![](figures/all_iucn_map.png)
-
-Figure 3:
-
-![](figures/AreaRatiovsPercOverlap.png)
-
-___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
-
-Figure 4 (a, b):
-
-![](figures/barchart_spp_risk_quads.png)
-
-![](figures/barchart_spp_gp_quads.png)
-
-Figure 5 (a, b):
-
-![](figures/AMThreshold.png)
-
-![](figures/ribbon_AMThreshold.png)
-
-Figure 6
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
-
-Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
-
-| Scenario              | Priority data source | AquaMaps presence threshold |
-| --------------------- | :------------------: | :-------------------------: |
-| Scenario 0 (current)  | IUCN                 |  >= 40%                     |
-| Scenario 1            | IUCN                 |   > 0%                      |
-| Scenario 2            | AquaMaps             |  >= 40%                     |
-| Scenario 3            | AquaMaps             |   > 0%                      |

### 2015-12-21 ms_4_figs.Rmd
updating manuscript

@@
 
-Figure 1 (a, b):
+### Figure 1 (a, b):
 
@@
 
-Figure 2 (a, b):
+### Figure 2 (a, b):
 
@@
 
-Figure 3:
+### Figure 3:
 
@@
 ___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
+``` {r tally_quad_counts, echo = FALSE}
+dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
+area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
+q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
+q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
+q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
+q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
+```
 
-Figure 4 (a, b):
+* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
+For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
+(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
+* Species map pairs that fall within the upper left quadrant (quadrant 2: well-aligned in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
+(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
+* The lower right quadrant (quadrant 3: well-aligned in area but poorly matched in distribution) includes species for which the paired maps generally agree in range area, but disagree in locations of those ranges.
+Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being lower quality alignment than map pairs in either of the upper quadrants.
+(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
+* The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
+(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
+
+### Figure 4 (a, b):
 
@@
 
-Figure 5 (a, b):
+### Figure 5 (a, b):
 
@@
 
-Figure 6
+### Figure 6
 

### 2015-12-21 ms_5_refs.Rmd
updating manuscript

@@
     * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources
+IUCN Red List - accessed 12/21/2015

### 2015-12-21 ms_6_supplemental.Rmd
updating manuscript

--- 2015-12-19 ms_outline2.Rmd
@@
----
-title: "Aligning species range data to better serve science and conservation"
-author: "Casey O'Hara, Jamie Afflerbach, Ben Halpern"
-output:
-  pdf_document:
-    number_sections: yes
-    toc: no
-  html_document:
-    highlight: haddock
-    number_sections: yes
-    theme: spacelab
-    toc: yes
-  word_document: null
----
-
-``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
-
-# library(ggplot2)
-# library(maps)
-# library(readr)
-# library(raster)
-# library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-
-# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-#            'Darwin'  = '/Volumes/data_edit',
-#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-# dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-# dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
-# dir_fig <- file.path(getwd(), 'figures')
-```
-
-``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.
-
-data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
-
-### generate other lists
-
-spp_all_file <- 'data/spp_all_2015.csv'
-spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
-
-iucn_marine_file <- 'data/spp_iucn_marine.csv'
-spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
-
-iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
-  rename(sciname = binomial) %>%
-  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
-  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
-
-spp_am_file <- 'data/spp_am.csv'
-spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
-
-```
-
-## Abstract
-
-_get to this later..._
-
------
-
-## Significance
-
-_subset of the abstract_
-
------
-
-# Introduction
-
-Mapping and predicting species’ distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
-Knowing where individuals of a species are, and what allows them to persist there, provides foundational information for understanding species ranges, how best to protect and manage species, and how they may respond to increasing human impacts and a changing climate.
-A rich literature exists tackling the many dimensions of these questions.
-
-One very important outcome of this body of science is the various compiled databases of species range maps.
-In the oceans there are now two such global repositories – Aquamaps (REF) and International Union for Conservation of Nature (IUCN) species distribution maps (REF) – that are used for a wide range of purposes, including assessing marine species status (Halpern et al.
-2012, Selig et al.
-2013), evaluating global biodiversity patterns and trends (REF), predicting range shifts (Molinos et al.
-2015), and setting conservation priorities (Klein et al.
-2015).
-The two data collections ostensibly describe the same information, but significant differences in methodology, intent, and taxonomic and geographic sampling could lead to dramatically different understandings of our marine ecosystems and resulting policy and conservation recommendations
-
-_Importantly, biases in taxanomic or spatial coverage of a dataset could shift management or conservation actions towards places or species that aren’t actually the most in need. [elaborate a bit?]_
-
-To understand the implications of similarities and differences between the two datasets, we examined how they compare in their representation of global spatial distribution of species, and for the relatively small number of species represented in both datasets, how closely their spatial distribution maps align.
-We then reanalyze two recent studies using these datasets based on different versions of the data  to illustrate how and why decisions about which species range data to use affect our understanding of the status of marine biodiversity.
-
-# Overview of the datasets
-
-___We may need to shorten this section to fit this paper into the format guidelines for PNAS.  If so, some of this can be pushed to the Suppl. Materials. - BH___
-
-The IUCN publishes species distribution maps as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.
-Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.
-These maps are typically based on species occurrence records from databases such as the Ocean Biogeographic Information System (OBIS) (REF) and the Global Biodiversity Information Facility (GBIF) (REF); IUCN provides _general guidelines_ ___Do we need to describe these?
-- BH___ to facilitate consistency in how the maps are generated.
-The IUCN presents these maps as 'limits of distribution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed within the polygon.
-While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
-
-IUCN publishes spatial distribution map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
-in which at least 90% of the species within the taxonomic group have been evaluated.
-While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN had published species distribution maps for `r nrow(spp_iucn_maps)` marine species across 24 taxonomic groups.
-For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.
-
-AquaMaps (Kaschner et al.
-2013) offers species range maps based on modeled relative environmental suitability.
-For each assessed species, data from occurrence records such as OBIS and GBIF, published species databases such as FishBase, and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, sea ice concentration, and distance to land.
-The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence, with cell values ranging from 0.00 to 1.00.
-These maps can then be reviewed by experts to fine-tune the parameters to better match known occurrences.
-As of December 2015, AquaMaps current native distribution maps have been produced for `r nrow(spp_am)` species.
-Use of these data to predict species presence or absence requires decisions about which probability threshold to use.
-
------
-
-_Analyses [briefly describe the analyses done]_
-
-* _simple taxanomic comparison (describe very briefly)_
-* _spatial overlap comparison (describe very briefly)_
-* _redo of OHI biodiversity goal (describe briefly what was done previously, and how that was changed for here)_
-* _redo global MPA gap analysis (describe briefly what was done previously, and how that was changed for here)_
-
-# Results and Discussion
-
-### Overlap between assessed species
-
-``` {r calc species numbers, echo = FALSE}
-n_spp_am    <- nrow(spp_am)
-n_spp_am_rl <- nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))
-p_spp_am_rl <- round(n_spp_am_rl/n_spp_am * 100, 1)
-n_spp_iucn  <- nrow(spp_iucn_maps)
-n_spp_both  <- nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))
-p_both_am   <- round(n_spp_both/n_spp_am, 1)
-p_both_iucn <- round(n_spp_both/n_spp_iucn, 1)
-```
-
-Due to differences in both methodology and intent, significant differences exist between AquaMaps and IUCN species distribution maps (Fig.
-1).
-Only `r n_spp_both` species were covered by both datasets (`r p_both_am`% of total Aquamaps species; `p_both_iucn`% of total IUCN species; Fig.
-1, 2).
-__All IUCN-mapped species are also included in Red List species, but only `r p_spp_am_rl`% of AquaMaps species.__
-
-<!--
-| Map set   | n~total~       | n~Red_List~     |
-| :-------: | -------------: | :-------------: | 
-| AquaMaps  | `r n_spp_am`   | `r n_spp_am_rl` | 
-| IUCN      | `r n_spp_iucn` | all             | 
-| Both      | `r n_spp_both` | all             |
--->
-
-### Taxonomic distribution between datasets
-
-___Either move to brief methods section above, and maybe even shorten - BH___: To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
-We further divided the data by whether species had been evaluated for the IUCN Red List of Threatened Species.
-
-___Either cut or move to data description section of methods above.
-- BH___: The IUCN only releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed.
-As such, spatial data for many taxonomic classes are unavailable, and within a class, the assessed sub-groups may not represent the entire class.
  ___Move to supplementary materials.
@@
 
-___Move to methods section above.
-- BH___: To compare the spatial representation of the two datasets directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
-For the AquaMaps dataset, we determined per-cell species richness by summing all species with non-zero probability of occurrence.
-We represented relative distribution of species richness for each dataset by plotting average species count against latitude and longitude.
-To highlight the differences between the two datasets, we also examined the relative difference in per-cell species richness against latitude and longitude.
-
-The balance of species represented in IUCN spatial data skews toward tropical latitudes compared to the distribution represented in AquaMaps distribution maps.
-This may be driven by the fact that the taxonomic distribution of IUCN distribution maps focuses heavily on coral reef-associated species (see fig.
-XXX).
-While both datasets indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps dataset shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN dataset.
-
-Each dataset offers spatial distribution information for large numbers of species.
-However, the datasets vary in terms of taxonomic coverage and regional coverage.
-For spatial assessments of biodiversity, the choice of one dataset over the other is likely to create significantly different results.
-For studies confined to a narrow range of taxa or to a narrow spatial scale, one dataset may offer an advantage over the other in the number of species maps available.
-For global scale biodiversity studies, however, the selection of one dataset over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
-___find a reference that describes what makes a "good" dataset for global biodiversity, e.g.
-OHI___ _or maybe species richness vs diversity vs "health" or whatnot_
-
------
-
-## Analysis of spatial alignment by species
-
-### Defining spatial alignment between the two datasets
-
-<!--
-AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to variation between range maps for any given species, --> For species described in both spatial datasets, we would expect to see spatial correlation both in the global pattern of species distribution and the extent of species range.
-Large discrepancies in distribution and range area could indicate species that require further study to consolidate expert knowledge for IUCN and improve the inputs to AquaMaps species distribution models.
-
-Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
-Assigning "presence" values using the same methodology above, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of paired-map species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
-This aligns with the expectations and early analysis of Kaschner et al (REF).
-
-Overlaying paired distribution maps for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
-
-$$\alpha_{dist} = \frac{A_{small \cap large}}{A_{large}} * 100\%$$
-
-$$\alpha_{area} = \frac{A_{small}}{A_{large}} * 100\%$$
-
-For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
-
------
-
-For each of the `r nrow(spp_map_pairs)` species whose range is described by both datasets, we calculated the distribution alignment and the area alignment, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).
-
-We divided the plot into quadrants based on the median values for each axis to differentiate between four different qualities of alignment:
-
-``` {r calc medians and quadrant counts, echo = FALSE}
-dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
-area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
-q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
-```
-
-* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
-For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
-(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The other quadrants each suggest a different story.
-Species map pairs that fall within the upper left quadrant (quadrant 2: well-aligned in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
-(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The lower right quadrant (quadrant 3: well-aligned in area but poorly matched in distribution) includes species for which the paired maps generally agree in range area, but disagree in locations of those ranges.
-Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being lower quality alignment than map pairs in either of the upper quadrants.
-(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
-
-Breaking down the quadrants by IUCN extinction risk categories, species with higher extinction risk tend to be better aligned between the two datasets.
-Well-aligned range map pairs (membership in quadrants 1 and 2) may suggest increased expert scrutiny, though this could be explained as higher perceived risk leading to increased attention, or as increased attention revealing deeper risk.
-Likely both mechanisms are at play in a feedback loop.
-In any case, poorly-aligned rangemaps, regardless of extinction risk category, seem to indicate species who could benefit from further expert study.
-
-When broken down by taxonomic group, we found that certain taxonomic groups were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
-Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two datasets, predicted areas are not well-aligned.
-AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range.
-A similar mechanism may also drive poor area alignment of other demersal and reef-associated organisms.
-
-___Other hypotheses/case studies to highlight other possible mechanistic differences?___
-
------
-
-## AquaMaps: Effect of changing "presence" threshold on area representation
-
-AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value indicated by IUCN maps.
-Frequently, studies convert this probability to a simple presence value by assigning a threshold value (XXX references here).
-A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
-For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to approximate the presence criteria used by IUCN.
-To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0.05 to 1.00 and calculated the average species range relative to a zero threshold.
-
-_Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.
-Do regression - shallow slope indicates that sensitivity is less than 1:1._
-
-At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a (XXX to XXX %, median = XXX) decrease in represented range, though some species lose nearly their entire range.
-Examining the entire species catalog as a single set, a unit shift in presence threshold results in a generally smaller proportional shift in represented area, indicating a moderate and consistent sensitivity to threshold with no surprising tradeoffs to inform the identification of a "correct" threshold.
-This pattern may not hold true for all subsets of AquaMaps, however, whether subsetting by taxa or by georegion.
-
------
-
-# Implications
-
-## Application to OHI
-
-The global Ocean Health Index (Halpern et al.
-2012), an index made up of 10 goals, utilizes both of these datasets to inform the Biodiversity goal.
-The biodiversity goal is made up of two parts; Species and Habitats.
-As it is currently calculated, the Species subgoal uses species spatial distribution data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
-Spatial distributions were gleaned from both IUCN and AquaMaps datasets.
-A probability threshold of 40% is used to determine presence for AquaMaps data.
-For those species that are are represented within both datasets, IUCN distribution maps were used as the preferred data source.
-Species with no spatial data in either dataset were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
-
-Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
-Reducing the threshold will always increase the apparent range of a species, therefore the slight decrease in average score suggests increased spatial representation of threatened species.
-
-Scenario 2 shows the effect of preferring AquaMaps data over IUCN, while maintaining the same presence threshold.
-This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so many but not all overlapping species will see a decrease in represented range.
-The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
-
-Scenario 3 shows the effect of preferring AquaMaps data over IUCN, while eliminating the presence threshold.
-Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
-The large decrease seems to indicate that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.
-
-Since the Ocean Health Index Species subgoal depends on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted.
-When IUCN data is the preferred data source, only the non-paired AquaMaps species will be affected by threshold changes; and even when AquaMaps is the preferred source, the non-paired IUCN species will dampen the effect of a threshold change.
-
------
-
-## Application to MPA Gap Analysis
-
-___yeah still gotta do something on this one___
-
------
-
-# Conclusion
-
------
-
-#References
-
-__Key AquaMaps publications:__
-
-Kaschner, K., D.P. Tittensor, J. Ready, T Gerrodette and B. Worm (2011). Current and Future Patterns of Global Marine Mammal Biodiversity. PLoS ONE 6(5): e19653. PDF
-
-Ready, J., K.
-Kaschner, A.B.
-South, P.D Eastwood, T.
-Rees, J.
-Rius, E.
-Agbayani, S.
-Kullander and R.
-Froese (2010).
-Predicting the distributions of marine organisms at the global scale.
-Ecological Modelling 221(3): 467-478.
-PDF
-
-__Papers based on AquaMaps:__ ___Use these for section 4?___
-
-Jones, M.C., S.R.
-Dyeb, J.K.
-Pinnegar and W.W.L.
-Cheung (2012).
-Modelling commercial fish distributions: Prediction and assessment using different approaches.
-Ecological Modelling 225(2012): 133-145.
-PDF
-
-Coll, M., C. Piroddi, J. Steenbeek, K. Kaschner, F. Ben Rais Lasram et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. PDF
-
-Martin C.S., Fletcher R., Jones M.C., Kaschner K., Sullivan E., Tittensor D.P., Mcowen C., Geffert J.L., van Bochove J.W., Thomas H., Blyth S., Ravillious C., Tolley M., Stanwell-Smith D.
-(2014).
-Manual of marine nd coastal datasets of biodiversity importance.
-May 2014 release.
-Cambridge (UK): UNEP World Conservation Monitoring Centre.
-28 pp.
-(+ 4 annexes totalling 174 pp.
-and one e-supplement).
-PDF
-
------
-
-García Molinos, Jorge, Benjamin S.
-Halpern, David S.
-Schoeman, Christopher J.
-Brown, Wolfgang Kiessling, Pippa J.
-Moore, John M.
-Pandolfi, Elvira S.
-Poloczanska, Anthony J.
-Richardson, and Michael T.
-Burrows.
-“Climate Velocity and the Future Global Redistribution of Marine Biodiversity.” Nature Climate Change advance online publication (August 31, 2015).
-doi:10.1038/nclimate2769.
-
-Halpern, Benjamin S., Catherine Longo, Darren Hardy, Karen L.
-McLeod, Jameal F.
-Samhouri, Steven K.
-Katona, Kristin Kleisner, et al.
-“An Index to Assess the Health and Benefits of the Global Ocean.” Nature 488, no.
-7413 (August 30, 2012): 615–20.
-doi:10.1038/nature11397.
-
-<!--
-Halpin, P.N., A.J.
-Read, E.
-Fujioka, B.D.
-Best, B.
-Donnelly, L.J.
-Hazen, C.
-Kot, K.
-Urian, E.
-LaBrecque, A.
-Dimatteo, J.
-Cleary, C.
-Good, L.B.
-Crowder, and K.D.
-Hyrenbach.
-2009.
-OBIS-SEAMAP: The world data center for marine mammal, sea bird, and sea turtle distributions.
-Oceanography 22(2):104-115
--->
-
-Klein, Carissa J., Christopher J.
-Brown, Benjamin S.
-Halpern, Daniel B.
-Segan, Jennifer McGowan, Maria Beger, and James E.M.
-Watson.
-“Shortfalls in the Global Protected Area Network at Representing Marine Biodiversity.” Scientific Reports 5 (December 3, 2015): 17539.
-doi:10.1038/srep17539.
-
-Selig, Elizabeth R., Catherine Longo, Benjamin S.
-Halpern, Benjamin D.
-Best, Darren Hardy, Cristiane T.
-Elfes, Courtney Scarborough, Kristin M.
-Kleisner, and Steven K.
-Katona.
-“Assessing Global Marine Biodiversity Status within a Coupled Socio-Ecological Perspective.” PLoS ONE 8, no.
-4 (April 11, 2013): e60284.
-doi:10.1371/journal.pone.0060284.
-
-* _IUCN reference:_ 
-    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
-    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources
-
-# Figures
-
-Figure 1 (a, b):
-
-![](figures/bar_by_taxa_all.png)
-
-![](figures/bar_by_taxa_assessed.png)
-
-Figure 2 (a, b):
-
-![](figures/all_aquamaps_map.png)
-
-![](figures/all_iucn_map.png)
-
-Figure 3:
-
-![](figures/AreaRatiovsPercOverlap.png)
-
-___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
-
-Figure 4 (a, b):
-
-![](figures/barchart_spp_risk_quads.png)
-
-![](figures/barchart_spp_gp_quads.png)
-
-Figure 5 (a, b):
-
-![](figures/AMThreshold.png)
-
-![](figures/ribbon_AMThreshold.png)
-
-Figure 6
-
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
-
-Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
-
-| Scenario              | Priority data source | AquaMaps presence threshold |
-| --------------------- | :------------------: | :-------------------------: |
-| Scenario 0 (current)  | IUCN                 |  >= 40%                     |
-| Scenario 1            | IUCN                 |   > 0%                      |
-| Scenario 2            | AquaMaps             |  >= 40%                     |
-| Scenario 3            | AquaMaps             |   > 0%                      |
+__Add'l maps to show mechanistic causes of quadrants__

### 2015-12-23 ms_0_main.Rmd
updating manuscript... adding a few annotated refs

@@
 output:
+  pdf_document:
+    number_sections: yes
+    toc: no
   html_document:
@@
     toc: yes
-  pdf_document:
-    number_sections: yes
-    toc: no
-  word_document: null
+  word_document: default
 ---
@@
 ``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.
+### use data_explore.Rmd to create this file, if it does not yet exist.  This
+### filters out presence = Extinct from IUCN cells.
 
@@
 
-spp_all_file <- 'data/spp_all_2015.csv'
+spp_all_file <- 'data/spp_all.csv'
 spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
@@
 spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
-  rename(sciname = binomial) %>%
   filter(spp_group != 'hagfishes') %>%           ### not from IUCN...

### 2015-12-23 ms_2_intro_res_disc.Rmd
updating manuscript... adding a few annotated refs

@@
 
-Mapping and predicting species’ distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
-Knowing where individuals of a species are, and what allows them to persist there, provides foundational information for understanding species ranges, how best to protect and manage species, and how they may respond to increasing human impacts and a changing climate.
-A rich literature exists tackling the many dimensions of these questions.
+Mapping and predicting species distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
+Knowing where individuals of a species exist, and what allows them to persist there, provides foundational information for understanding species ranges, how best to protect and manage species, and how they may respond to increasing human impacts and a changing climate.
+A rich literature tackles the many dimensions of these questions.
 
-One very important outcome of this body of science is the various compiled databases of species range maps.
-In the oceans there are now two such global repositories – Aquamaps (REF) and International Union for Conservation of Nature (IUCN) species distribution maps (REF) – that are used for a wide range of purposes, including assessing marine species status (Halpern et al.
+One very important outcome of this body of science is the various compiled databases of species distribution maps.
+In the oceans there are now two such global repositories _(if we define it narrowly; OBIS is global species distribution, but not quite range maps)_ – Aquamaps predicted distribution maps (Kaschner et al.
+2013) and International Union for Conservation of Nature (IUCN) range maps (REF) – that are used for a wide range of purposes, including assessing marine species status (Halpern et al.
 2012, Selig et al.
-2013), evaluating global biodiversity patterns and trends (REF), predicting range shifts (Molinos et al.
+2013), evaluating global biodiversity patterns (Coll et al.
+2010, Martin et al.
+2014), predicting range shifts (Molinos et al.
 2015), and setting conservation priorities (Klein et al.
 2015).
-The two data collections ostensibly describe the same information, but significant differences in methodology, intent, and taxonomic and geographic sampling could lead to dramatically different understandings of our marine ecosystems and resulting policy and conservation recommendations
+The two data collections ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
 
-_Importantly, biases in taxanomic or spatial coverage of a dataset could shift management or conservation actions towards places or species that aren’t actually the most in need. [elaborate a bit?]_
+Importantly, biases in taxonomic or spatial coverage of a dataset could shift management or conservation actions _towards places or species that aren’t actually the most in need (shift management or conservation actions away from places or species that are most in need?)_.(maybe Jetz 2008?) False indications of presence (commission errors) overestimate the protection afforded by a marine reserve, while false indications of absence (omission errors) reduce the adequacy of a reserve system by underestimating species distributions.
+(Rondinini et al.
+2006)
 
 To understand the implications of similarities and differences between the two datasets, we examined how they compare in their representation of global spatial distribution of species, and for the relatively small number of species represented in both datasets, how closely their spatial distribution maps align.
-We then reanalyze two recent studies using these datasets based on different versions of the data  to illustrate how and why decisions about which species range data to use affect our understanding of the status of marine biodiversity.
+_We then reanalyze two recent studies using these datasets based on different versions of the data to illustrate how and why decisions about which species range data to use affect our understanding of the status of marine biodiversity._
 
@@
 
-The IUCN publishes species distribution maps as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.
-Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.
-These maps are typically based on species occurrence records from databases such as the Ocean Biogeographic Information System (OBIS) (REF) and the Global Biodiversity Information Facility (GBIF) (REF); IUCN provides _general guidelines_ ___Do we need to describe these?
-- BH___ to facilitate consistency in how the maps are generated.
-The IUCN presents these maps as 'limits of distribution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed within the polygon.
-While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
+The IUCN publishes species distribution maps as vector shapefiles gathered into taxonomic groups _that have been comprehensively assessed (> 90% of species evaluated)_.
+Species extents of occurrence are outlined by experts based upon known occurrences of the species, typically from point locality databases such as the Ocean Biogeographic Information System (OBIS) (REF) and the Global Biodiversity Information Facility (GBIF) (REF), and refined by expert understanding of range and habitat preferences.
 
-IUCN publishes spatial distribution map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
-in which at least 90% of the species within the taxonomic group have been evaluated.
+_IUCN publishes range map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
+in which at least 90% of the species within the taxonomic group have been evaluated_.
 While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-As of December 2015, IUCN had published species distribution maps for `r nrow(spp_iucn_maps)` marine species across 24 taxonomic groups.
-For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.
 
-AquaMaps (Kaschner et al.
-2013) offers species range maps based on modeled relative environmental suitability.
-For each assessed species, data from occurrence records such as OBIS and GBIF, published species databases such as FishBase, and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, sea ice concentration, and distance to land.
-The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence, with cell values ranging from 0.00 to 1.00.
-These maps can then be reviewed by experts to fine-tune the parameters to better match known occurrences.
-As of December 2015, AquaMaps current native distribution maps have been produced for `r nrow(spp_am)` species.
-Use of these data to predict species presence or absence requires decisions about which probability threshold to use.
+AquaMaps develops species distribution maps based on modeled relative environmental suitability.
+For each modeled species, environmental preferences (e.g.
+temperature, depth, salinity) are deduced from occurrence records such as OBIS and GBIF, published species databases such as FishBase, and expert knowledge.
+These environmental preferences are compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence.
+
+* _How does AquaMaps select species?  based on data available from OBIS/GBIF and Fishbase?_
 
@@
 
+# Results and Discussion
+
+<!--
 _Analyses [briefly describe the analyses done]_
 
-* _simple taxanomic comparison (describe very briefly)_
-* _spatial overlap comparison (describe very briefly)_
-* _redo of OHI biodiversity goal (describe briefly what was done previously, and how that was changed for here)_
+* _simple taxonomic and global spatial comparison (describe very briefly)_ 
+* _spatial overlap comparison (describe very briefly)_ For the relatively small number of species with maps in both datasets, we examined how well the maps align, and examined the implications of various categories of spatial misalignment.
+* _redo of OHI biodiversity goal (describe briefly what was done previously, and how that was changed for here)_ Using the Ocean Health Index biodiversity goal as one case study, we predicted the impact on OHI Species scores based on alternate decisions regarding how to use the two datasets.
 * _redo global MPA gap analysis (describe briefly what was done previously, and how that was changed for here)_ 
-
-# Results and Discussion
+-->
 
@@
 For species described in both spatial datasets, we would expect to see spatial correlation both in the global pattern of species distribution and the extent of species range.
-Large discrepancies in distribution and range area could indicate species that require further study to consolidate expert knowledge for IUCN and improve the inputs to AquaMaps species distribution models.
+Large discrepancies in distribution and range extent could indicate species that require further study to consolidate expert knowledge for IUCN range maps and improve the accuracy of AquaMaps species distribution models.
 
@@
 We found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of paired-map species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
-This aligns with the expectations and early analysis of Kaschner et al (REF).
+This aligns with the expectations of Rondinini et al.
+(2006) that range maps are more likely to falsely predict presence while predicted distribution models are more likely to falsely predict absence.
 
-Overlaying paired distribution maps for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
+Overlaying paired distribution maps for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range; and _extent alignment_, which we defined as the ratio of the smaller range to the larger range.
 
-For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and very similar extent of range (_area alignment_).
+For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and similar extent of occurrence (_extent alignment_).
 Dividing the map-paired species into quadrants based on median values for each axis, we can examine the implications of four different qualities of alignment.
@@
 
-* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
+* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges.
 For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
-* Species map pairs that fall within the upper left quadrant (quadrant 2: well-aligned in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
-* The lower right quadrant (quadrant 3: well-aligned in area but poorly matched in distribution) includes species for which the paired maps generally agree in range area, but disagree in locations of those ranges.
-Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being lower quality alignment than map pairs in either of the upper quadrants.
-* The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution.
+* Species map pairs that fall within the upper left quadrant (quadrant 2: well-aligned in distribution but poorly-matched in extent) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
+* The lower right quadrant (quadrant 3: well-aligned in extent but poorly matched in distribution) includes species for which the paired maps generally agree in range extent, but disagree in locations of those ranges.
+Disagreement in distribution seems to indicate a greater issue than disagreement in extent, so we categorized map pairs in this quadrant as being lower quality alignment than map pairs in either of the upper quadrants.
+* The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both extent and distribution.
 
@@
 Examining quadrants by taxonomic group, we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
-Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two datasets, predicted areas are not well-aligned.
+Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two datasets, predicted extents are not well-aligned.
 AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range (see SI).
-A similar mechanism may also drive poor area alignment of other demersal and reef-associated organisms.
+A similar mechanism may also drive poor extent alignment of other demersal and reef-associated organisms.
 
-Poorly-aligned rangemaps, regardless of taxon or extinction risk category, indicate species who could benefit from further expert study.
+Poorly-aligned rangemaps, regardless of taxon or extinction risk category, indicate species that could benefit from further expert study.
 
@@
 
+___range maps less likely to commit omission errors and more likely to commit commission errors; predicted distribution moderate chance of each compared to range maps and point locality data___ (based on Rondinini et al.
+2006) this in discussion after the distribution/extent analysis
+
 -----
 
-## AquaMaps: Effect of changing "presence" threshold on area representation
+## AquaMaps: Effect of changing "presence" threshold on apparent distribution
 
-AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value indicated by IUCN maps.
-Frequently, studies convert this probability to a simple presence value by assigning a threshold value (REF references here).
+AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value as indicated by IUCN maps.
+Many studies convert this AquaMaps probability to a simple presence value by assigning a threshold value (REF references here).
 A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
@@
 
-At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a _(XXX to XXX %, median = XXX)_ decrease in represented range, though some species lose nearly their entire range.
-Incrementing the presence threshold from zero to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.
-This pattern may not hold true for all subsets of AquaMaps, however, whether subsetting by taxa or by georegion.
+At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a significant decrease in represented range, and some species lose nearly their entire range.
+Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.
+This pattern may not hold true for all subsets of AquaMaps species, however, whether subsetting by taxa or by georegion.
 
@@
 Species with no spatial data in either dataset were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
-
-___caption?___
-
-* Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
-Reducing the threshold will always increase the apparent range of a species, therefore the slight decrease in average score suggests increased spatial representation of threatened species.
-
-* Scenario 2 shows the effect of preferring AquaMaps data over IUCN, while maintaining the same presence threshold.
-This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so many but not all overlapping species will see a decrease in represented range.
-The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
-
-* Scenario 3 shows the effect of preferring AquaMaps data over IUCN, while eliminating the presence threshold.
-Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
-The large decrease seems to indicate that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.
-
-___end caption___
 
@@
 
-IUCN overestimates area esp for coastal species?
+IUCN overestimates extent esp for coastal species?
 More species will be included in MPAs so fewer apparent gap species.

### 2015-12-23 ms_4_figs.Rmd
updating manuscript... adding a few annotated refs

@@
 | Scenario 3            | AquaMaps             |   > 0%                      |
+
+* Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
+Reducing the threshold will always increase the apparent range of a species, therefore the slight decrease in average score suggests increased spatial representation of threatened species.
+
+* Scenario 2 shows the effect of preferring AquaMaps data over IUCN, while maintaining the same presence threshold.
+This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so many but not all overlapping species will see a decrease in represented range.
+The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
+
+* Scenario 3 shows the effect of preferring AquaMaps data over IUCN, while eliminating the presence threshold.
+Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
+The large decrease seems to indicate that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.

### 2015-12-23 ms_6_supplemental.Rmd
updating manuscript... adding a few annotated refs

@@
+# Supplemental Information
+
+### descriptions of IUCN and AM data sets
+
+IUCN: While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
+
+As of December 2015, IUCN had published species distribution maps for `r nrow(spp_iucn_maps)` marine species across 24 taxonomic groups.
+For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.
+
+ As of December 2015, AquaMaps current native distribution maps have been produced for `r nrow(spp_am)` species.
+
  ___Move to supplementary materials.

### 2015-12-24 data_explore/data_explore.Rmd
updating manuscript

@@
 
+setwd('data_explore')
 source('data_explore_fxns.R')
 ### When knitting this, it automatically sets WD to be this directory...
+### the 'setwd()' is there for running by chunk
 ```
@@
 scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-  labs(x = bquote('Area alignment '*alpha['area']*' (%)'), 
+  labs(x = bquote('Extent alignment '*alpha['ext']*' (%)'), 
        y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
-       title = 'Distribution alignment vs area alignment',
+       title = 'Distribution alignment vs Extent alignment',
        color = 'Risk', size = '% ocean range')
@@
   filter(area_ratio < median(area_ratio) & sm_perc < median(sm_perc))
+
+spp_list_quads <- spp_list1 %>%
+  mutate(quad = NA,
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_q1$sciname, 'q1', quad),
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_q2$sciname, 'q2', quad),
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_q3$sciname, 'q3', quad),
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_q4$sciname, 'q4', quad))
+write.csv(spp_list_quads, file.path(dir_data, 'spp_list_quads.csv'), row.names = FALSE)
 
@@
 
+write.csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads.csv'), row.names = FALSE)
+
 ```
@@
 ``` {r examine_risk_quartiles, echo = FALSE, eval = TRUE}
-
-spp_list1 <- spp_list %>%
-  group_by(popn_category) %>%
-  mutate(n_spp = n()) %>%
-  ungroup()
-
-### Create species lists by quadrant
-spp_list_q1 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio) & sm_perc >= median(sm_perc))
-spp_list_q2 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio) & sm_perc >= median(sm_perc))
-spp_list_q3 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio) & sm_perc < median(sm_perc))
-spp_list_q4 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio) & sm_perc < median(sm_perc))
 

### 2015-12-24 ms_0_main.Rmd
updating manuscript

@@
 output:
-  pdf_document:
-    number_sections: yes
-    toc: no
   html_document:
@@
     theme: spacelab
-    toc: yes
+    toc: no
+  pdf_document:
+    number_sections: yes
+    toc: no
   word_document: default
@@
 
-``` {r abstract, child = 'ms_2_intro_res_disc.Rmd'}
+``` {r intro_results_discussion, child = 'ms_2_intro_res_disc.Rmd'}
 ```
@@
 
-``` {r abstract, child = 'ms_3_methods.Rmd'}
+``` {r methods, child = 'ms_3_methods.Rmd'}
 ```
@@
 
-``` {r abstract, child = 'ms_4_figs.Rmd'}
+``` {r figures, child = 'ms_4_figs.Rmd'}
 ```
@@
 
-``` {r abstract, child = 'ms_5_refs.Rmd'}
+``` {r references, child = 'ms_5_refs.Rmd'}
 ```

### 2015-12-24 ms_2_intro_res_disc.Rmd
updating manuscript

@@
 
-Importantly, biases in taxonomic or spatial coverage of a dataset could shift management or conservation actions _towards places or species that aren’t actually the most in need (shift management or conservation actions away from places or species that are most in need?)_.(maybe Jetz 2008?) False indications of presence (commission errors) overestimate the protection afforded by a marine reserve, while false indications of absence (omission errors) reduce the adequacy of a reserve system by underestimating species distributions.
+Importantly, biases in taxonomic or spatial coverage of a dataset could shift management or conservation actions _towards places or species that aren’t actually the most in need (shift management or conservation actions away from places or species that are most in need?)_.(maybe Jetz 2008?) Errors of commission (false indications of presence) or omission (false indications of absence) could lead to inefficient or ineffective marine reserve systems and management plans.
 (Rondinini et al.
-2006)
+2006) (Jetz 2008?)
 
-To understand the implications of similarities and differences between the two datasets, we examined how they compare in their representation of global spatial distribution of species, and for the relatively small number of species represented in both datasets, how closely their spatial distribution maps align.
-_We then reanalyze two recent studies using these datasets based on different versions of the data to illustrate how and why decisions about which species range data to use affect our understanding of the status of marine biodiversity._
+To understand the implications of differences between the two datasets, we compared how each data set represents the global spatial distribution of species.
+For the relatively small number of species mapped in both datasets, we examined how well their spatial distribution maps align.
+We then subjected two recent marine biodiversity studies to a sensitivity analysis, highlighting the consequences of different data use decisions _on our understanding of the status of marine biodiversity._
+
+-----
 
@@
 
-The IUCN publishes species distribution maps as vector shapefiles gathered into taxonomic groups _that have been comprehensively assessed (> 90% of species evaluated)_.
-Species extents of occurrence are outlined by experts based upon known occurrences of the species, typically from point locality databases such as the Ocean Biogeographic Information System (OBIS) (REF) and the Global Biodiversity Information Facility (GBIF) (REF), and refined by expert understanding of range and habitat preferences.
-
-_IUCN publishes range map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
-in which at least 90% of the species within the taxonomic group have been evaluated_.
+The IUCN publishes species distribution maps as vector shapefiles, collected by taxonomic groups.
+Species experts outline polygons to represent a species' extent of occurrence, based on observation records <!--point locality databases such as the Ocean Biogeographic Information System (OBIS) (REF?
+not actually using this data) and the Global Biodiversity Information Facility (GBIF) (REF?),--> and refined by the expert's understanding of the species' range and habitat preferences.
+IUCN publishes range map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
+in which at least 90% of the species within the taxonomic group have been evaluated.
 While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
@@
 For each modeled species, environmental preferences (e.g.
-temperature, depth, salinity) are deduced from occurrence records such as OBIS and GBIF, published species databases such as FishBase, and expert knowledge.
-These environmental preferences are compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence.
+temperature, depth, salinity) are deduced from occurrence records _such as OBIS and GBIF_ ___(should we mention these, or leave out?
+I left them out when describing the IUCN data sets)___, published species databases such as FishBase, and expert knowledge.
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence.
 
@@
 
-Differences in methodology and intent drive significant differences between AquaMaps and IUCN species distribution maps (Fig.
+IUCN maps are best categorized as geographic range data, while AquaMaps maps can be categorized as predicted distribution data (Rondinini et al., 2006) _(the categories come from here, but not my categorizations of these particular datasets...)_; the two types of spatial data differ in methodology and intent, and this drives significant differences between the resulting species distribution maps (Fig.
 1).
@@
 1, 2).
-__All IUCN-mapped species are also included in Red List species, but only `r p_spp_am_rl`% of AquaMaps species.__
+_All IUCN-mapped species are also included in Red List species, but only `r p_spp_am_rl`% of AquaMaps species._
 
@@
 
-The balance of species represented in IUCN spatial data skews toward tropical latitudes compared to the distribution represented in AquaMaps distribution maps.
-This may be driven by the fact that the taxonomic distribution of IUCN distribution maps focuses heavily on coral reef-associated species (see fig.
+The distribution of IUCN-mapped species skews toward tropical latitudes and away from the Atlantic and Eastern Pacific compared to the distribution of AquaMaps-mapped species.
+This skew likely reflects the fact that the IUCN dataset focuses more heavily on coral reef-associated taxa than does the AquaMaps dataset (see fig.
 XXX).
-While both datasets indicate highest species richness in the Coral Triangle and western Indian Ocean, the AquaMaps dataset shows a relatively larger representation of species in the Atlantic, Caribbean, and eastern Pacific than does the IUCN dataset.
 
@@
 For studies confined to a narrow range of taxa or to a narrow spatial scale, one dataset may offer an advantage over the other in the number of species maps available.
-For global scale biodiversity studies, however, the selection of one dataset over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth.
-___find a reference that describes what makes a "good" dataset for global biodiversity, e.g.
-OHI___ _or maybe species richness vs diversity vs "health" or whatnot_
+For global scale biodiversity studies, however, the selection of one dataset over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth._use something from Rondinini here?_
+
+_find a reference that describes what makes a "good" dataset for global biodiversity, e.g. OHI, or maybe species richness vs diversity vs "health" or whatnot:  (Tittensor et al., 2010)_
 
@@
 
-For species described in both spatial datasets, we would expect to see spatial correlation both in the global pattern of species distribution and the extent of species range.
+The IUCN and AquaMaps spatial datasets share `r nrow(spp_map_pairs)` species in common.
+If we were to examine each species' pair of maps side by side, we would hope to see spatial correlation both in the global pattern of species distribution (where on the map) and the extent of species range (how much of the map).
 Large discrepancies in distribution and range extent could indicate species that require further study to consolidate expert knowledge for IUCN range maps and improve the accuracy of AquaMaps species distribution models.
 
-Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
-We found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of paired-map species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.
-This aligns with the expectations of Rondinini et al.
-(2006) that range maps are more likely to falsely predict presence while predicted distribution models are more likely to falsely predict absence.
+For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range to the larger range (how much of the map).
+For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.
 
-Overlaying paired distribution maps for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range; and _extent alignment_, which we defined as the ratio of the smaller range to the larger range.
+We found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of paired-map species, the IUCN distribution map indicated a larger species range than the AquaMaps map.
+This concurs with the general expectation that geographic range maps (e.g.
+IUCN) are more likely to over-predict presence than predicted distribution models (e.g.
+AquaMaps), while predicted distribution models are more likely to over-predict absence.(Rondinini et al., 2006) _errors of commission vs errors of omission, essentially type I and type II errors_
 
-For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (_distribution alignment_) and similar extent of occurrence (_extent alignment_).
-Dividing the map-paired species into quadrants based on median values for each axis, we can examine the implications of four different qualities of alignment.
+Dividing the map-paired species into quadrants based on median values for each dimension, we can examine the _implications_ of four different qualities of alignment. (FIG 3a)
 
-<!--
-___moving to figure caption.  Summarize results here.___
+``` {r load_spp_list_by_quad, echo = FALSE}
+spp_list_quads <- read.csv('data/spp_list_quads.csv', stringsAsFactors = FALSE)
+n_q1 <- nrow(spp_list_quads %>% filter(quad == 'q1'))
+n_q2 <- nrow(spp_list_quads %>% filter(quad == 'q2'))
+n_q2_am_sm  <- nrow(spp_list_quads %>% filter(sm_range == 'AM' & quad == 'q2'))
+n_q2_corals <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral') & quad == 'q2'))
+n_corals    <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral')))
+n_q3 <- nrow(spp_list_quads %>% filter(quad == 'q3'))
+n_q4 <- nrow(spp_list_quads %>% filter(quad == 'q4'))
+```
 
-* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges.
-For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution.
-* Species map pairs that fall within the upper left quadrant (quadrant 2: well-aligned in distribution but poorly-matched in extent) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map.
-* The lower right quadrant (quadrant 3: well-aligned in extent but poorly matched in distribution) includes species for which the paired maps generally agree in range extent, but disagree in locations of those ranges.
-Disagreement in distribution seems to indicate a greater issue than disagreement in extent, so we categorized map pairs in this quadrant as being lower quality alignment than map pairs in either of the upper quadrants.
-* The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both extent and distribution.
+___could use some help on organizing this part...
+too much info?
+should some go in the caption, and some in the body?
+thinking this way: quickie description of quadrants in caption of figure; then use the bar charts in Fig 4 to help explain some mechanisms of quadrants___
 
--->
+* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges, as we hoped for well-understood species.
+Excellent, but not particularly interesting.
+* Species map pairs that fall within the upper left quadrant (quadrant 2) agree well in distribution, but disagree in extent.
+* The lower right quadrant (quadrant 3 ___note, "math" quadrant 3 would be lower left, not lower right, will this quadrant ID scheme be confusing?___) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more worrisome mismatch than that indicated by quadrant 2.
+* The lower left quadrant (quadrant 4) indicates species for which the map pairs fail to agree in both extent and distribution.
 
-Breaking down the quadrants by IUCN extinction risk categories, we found that species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
+Examining spatial alignment by taxonomic group (FIG 3a), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
+_explain reasoning?_
+
+Coral species are predominantly found in quadrant 2 (`r n_q2_corals` coral species - `r round(n_q2_corals/n_corals * 100, 1)`% of all corals and `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in quadrant 2).
+<!--AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range (see SI).
+A similar mechanism may also drive poor extent alignment of other demersal and reef-associated organisms.-->
+
+___how to transition from above paragraph to this idea:___ Looking more closely at the species included in quadrant 2, we can identify several mechanisms that could cause extent misalignment.
+For `r n_q2_am_sm` of `r n_q2` species in this quadrant (`r round((n_q2_am_sm/n_q2) * 100, 1)`%), the IUCN extent is larger than the AquaMaps extent.
+By itself, this is not surprising; but for many of these species, a quick look at the maps (see SI for examples) shows that the IUCN range hews closely to the AquaMaps range, while including a wide buffer zone.
+We suspect that many of these extent-misaligned map pairs can be easily explained: most corals and reef-associated organisms prefer shallower waters; seafloor depth is explicitly modeled in AquaMaps, but not explicitly included in IUCN range considerations.
+
+* _Occasionally, the AquaMaps map conformed to a small, local subset of the IUCN map (see SI for examples), in some cases due to some point locality observations being rejected by AquaMaps experts (EXAMPLES) and in other cases due to differences in species identification._ ___bleh - this needs help.___
+
+Breaking down the quadrants by IUCN extinction risk categories (FIG 3c), we found that species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
 Does higher perceived risk lead to increased attention, and thus better understanding of species distribution?
-Or conversely, does increased attention to species distribution reveal deeper risk?
-Likely both mechanisms are at play in a feedback loop.
-
-Examining quadrants by taxonomic group, we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
-Coral species are predominantly found in quadrant 2, indicating that while general global distribution trends seem to be consistent between the two datasets, predicted extents are not well-aligned.
-AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range (see SI).
-A similar mechanism may also drive poor extent alignment of other demersal and reef-associated organisms.
+Or conversely, does increased attention to species distribution reveal more species at risk?
+Likely both mechanisms are at play on a case-by-case basis, depending on the species' taxon and region.
+_does this argument bear up to closer scrutiny?
+CR isn't dominated by Q1 any more_
 
 Poorly-aligned rangemaps, regardless of taxon or extinction risk category, indicate species that could benefit from further expert study.
+___is this just 'duh'?
+is there something more important I can say?___
 
-___Other hypotheses/case studies to highlight other possible mechanistic differences?___
-
-Confusion in species names also seems to correlate with spatial misalignment.
-AquaMaps identifies each species map by binomial as well as a unique species ID code.
-While the IUCN assigns a unique numeric code to every Red List-assessed species, this code is not used consistently to identify range map polygons.
-___NOTE: NEW DATA as of Nov 24?
-Need to re-extract all the IUCN species :( ___
-
-___range maps less likely to commit omission errors and more likely to commit commission errors; predicted distribution moderate chance of each compared to range maps and point locality data___ (based on Rondinini et al.
-2006) this in discussion after the distribution/extent analysis
+_Other hypotheses/case studies to highlight other possible mechanistic differences?_
 
@@
 
-___do the analysis...___
+_do the analysis..._
 
-Predictions:
-
-IUCN overestimates extent esp for coastal species?
+_Predictions:  IUCN overestimates extent esp for coastal species?
 More species will be included in MPAs so fewer apparent gap species.
@@
 Policy implications?
-Less pressure to carefully consider MPA design; less resolution on
+Less pressure to carefully consider MPA design; less resolution on_

### 2015-12-24 ms_3_methods.Rmd
updating manuscript... adding a few annotated refs
updating manuscript

@@
 ___Either move to brief methods section above, and maybe even shorten - BH___: To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
-We further divided the data by whether species had been evaluated for the IUCN Red List of Threatened Species.
+We then filtered the species list to those that have been evaluated for the IUCN Red List of Threatened Species.
 
@@
 
-For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to approximate the presence criteria used by IUCN.
+For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to best approximate the "extent of occurrence" as generally indicated by IUCN maps.
 To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0.05 to 1.00 and calculated the average species range relative to a zero threshold.
@@
 _map pairs comparison_
+
+Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
 

### 2015-12-24 ms_4_figs.Rmd
updating manuscript

--- 2015-12-21 ms_4_figs.Rmd
@@
 
-Figure 2 (a, b):
+Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both. (a) All species with distribution maps; (b) IUCN Red List-assessed species with distribution maps.
+
+### Figure 2 (a, b):
 
@@
 
-Figure 3:
+Global marine species richness according to (a) AquaMaps dataset and (b) IUCN dataset.
+
+### Figure 3 (a):
 
@@
 
-___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___
+Distribution alignment vs Extent alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN datasets.
 
-Figure 4 (a, b):
+``` {r tally_quad_counts, echo = FALSE}
+dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
+area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
+q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
+q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
+q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
+q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
+```
+
+* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
+(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
+* The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent.
+(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
+* The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more problematic mismatch than that indicated by quadrant 2.
+(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
+* The lower left quadrant (quadrant 4) indicates species for which the map pairs fail to agree in both area and distribution. 
+(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
+
+### Figure 3 (b, c):
+
+![](figures/barchart_spp_gp_quads.png)
 
@@
 
-![](figures/barchart_spp_gp_quads.png)
+Spatial alignment of paired-map species by (a) taxonomic group and (b) extinction risk category.
 
-Figure 5 (a, b):
+### Figure 5 (a, b):
 
@@
 
-Figure 6
+AquaMaps distribution map extent remaining after applying a presence threshold.
+(a) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of XXX, with a wide distribution in which some species lose nearly all of their apparent range.
+(b) Mean (median) remaining extent at increments of presence threshold.
+Dark grey ribbon includes 25% to 75% quantiles, while light grey ribbon includes 9% to 91% quantiles.
+
+### Figure 6
 

### 2015-12-24 data_explore/data_explore_fxns.R

@@ -184,7 +184,7 @@ create_spp_list <- function(spp_list_base, data_file) {
 # occurrence cells and associated Aquamaps probability and/or IUCN proportional area
 # per cell
 get_spp_map <- function(species){ # species = 'Acanthopagrus latus' # species = 'Abudefduf concolor'
-  spp_id <- spp_list_build %>%
+  spp_id <- spp_list %>%
     filter(sciname == species & is.na(parent_sid)) %>%
     dplyr::select(am_sid, iucn_sid, sciname) %>%
     unique()

### 2015-12-28 data_explore/data_explore.Rmd
updates to manuscript; also data_explore still needs to be updated with new species lists and the like.

@@
                      stringsAsFactors = FALSE),
-            by = 'spp_group')
+            by = 'spp_group') %>%
+  mutate(spp_group = ifelse(str_detect(spp_group, 'corals'), 'corals', spp_group))
+    ### lump all corals together just for ease of handling
 

### 2015-12-28 ms_4_figs.Rmd
updates to manuscript; also data_explore still needs to be updated with new species lists and the like.

@@
 
-### Figure 1 (a, b):
+### Figure 1(a)?:
 
@@
 
-![](figures/bar_by_taxa_assessed.png)
+Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.
 
-Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both. (a) All species with distribution maps; (b) IUCN Red List-assessed species with distribution maps.
-
-### Figure 2 (a, b):
+### Figure 1 (b, c):
 
@@
 
-Distribution alignment vs Extent alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN datasets.
+For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range to the larger range (how much of the map).
+For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.
 
@@
 
-* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.
+* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges. 
 (n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
@@
 (n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more problematic mismatch than that indicated by quadrant 2.
+* The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more worrisome mismatch than that indicated by quadrant 2.
 (n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The lower left quadrant (quadrant 4) indicates species for which the map pairs fail to agree in both area and distribution. 
+* The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution. 
 (n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
 
-### Figure 3 (b, c):
+### Figure 3 (b):
 
@@
 
-![](figures/barchart_spp_risk_quads.png)
+Spatial alignment of paired-map species by taxonomic group.
 
-Spatial alignment of paired-map species by (a) taxonomic group and (b) extinction risk category.
-
-### Figure 5 (a, b):
-
-![](figures/AMThreshold.png)
-
-![](figures/ribbon_AMThreshold.png)
-
-AquaMaps distribution map extent remaining after applying a presence threshold.
-(a) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of XXX, with a wide distribution in which some species lose nearly all of their apparent range.
-(b) Mean (median) remaining extent at increments of presence threshold.
-Dark grey ribbon includes 25% to 75% quantiles, while light grey ribbon includes 9% to 91% quantiles.
-
-### Figure 6
+### Figure 4
 

### 2015-12-28 data_explore/data_explore.Rmd
updating for 2015 IUCN data

--- 2015-12-24 data_explore/data_explore.Rmd
@@
 ```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-data_file <- file.path(dir_data, 'spp_list_w_area_all.csv')
-
-if (!file.exists(data_file)) {
-  ### species info list - scinames, IUCN species ID, AM species ID, category etc
-  spp_list_base <- read.csv(file.path(dir_data, 'spp_am_v_iucn.csv'),
-                            stringsAsFactors = FALSE)
-
-  create_spp_list(spp_list_base, data_file)
-}
-
-spp_list <- read.csv(data_file, stringsAsFactors = FALSE) 
-### At this point, spp_list contains subpops and aliases, and zero-area IUCNs
-
-spp_list <- spp_list %>%
-  filter(!is.na(sm_perc)) %>%
-  ### will ditch the ones where no IUCN polygon exists - or are these species
-  ### whose ranges only occur in cells beyond the LOICZID grid?
-  filter(!str_detect(spatial_source, 'subpop') & !str_detect(spatial_source, 'alias'))
-spp_list <- spp_list %>%
-  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
-                     stringsAsFactors = FALSE),
-            by = 'spp_group')
-
-write.csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
+### this file created in data_prep
+read.csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE)
 

### 2015-12-28 data_explore/data_prep.Rmd
updating for 2015 IUCN data

---
title: 'Data Prep: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(readr)
library(data.table)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')
dir_data_remote <- file.path(dir_N, 'git-annex/globalprep/_raw_data')

setwd('data_explore')
source('data_explore_fxns.R')
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

Need to:

gather spp_all, spp_am, spp_am_v_iucn, spp_list_w_area_all and _trimmed into data folder.

```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
### get spp_all_complete from git-annex, bring to local repo
spp_all_remote <- read.csv(file.path(dir_anx, 'v2016/int/spp_all_complete.csv'), stringsAsFactors = FALSE)
write.csv(spp_all_remote, file.path(dir_data, 'spp_all.csv'), row.names = FALSE)

### create spp_am_v_iucn
spp_am_v_iucn <- spp_all_remote %>%
  filter(str_detect(spatial_source, 'iucn') & !is.na(am_sid)) %>%
  unique()
write.csv(spp_am_v_iucn, file.path(dir_data, 'spp_am_v_iucn.csv'), row.names = FALSE)

```

```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                           col_types = 'cdddd', progress = TRUE) %>%
  as.data.frame() %>%
  filter(presence != 5) ### removes extinct subpopulations

am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), progress = TRUE)

```

```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
data_file <- file.path(dir_data, 'spp_list_w_area_all.csv')
reload = TRUE
if (!file.exists(data_file) | reload) {
  ### species info list - scinames, IUCN species ID, AM species ID, category etc
  spp_list_base <- read.csv(file.path(dir_data, 'spp_am_v_iucn.csv'),
                            stringsAsFactors = FALSE)

  create_spp_list(spp_list_base, data_file)
}

spp_list <- read.csv(data_file, stringsAsFactors = FALSE) 
### At this point, spp_list contains subpops and aliases, and zero-area IUCNs

spp_list <- spp_list %>%
  filter(!is.na(sm_perc)) %>%
  ### will ditch the ones where no IUCN polygon exists - or are these species
  ### whose ranges only occur in cells beyond the LOICZID grid?
  filter(!str_detect(spatial_source, 'subpop') & !str_detect(spatial_source, 'alias'))
spp_list <- spp_list %>%
  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
                     stringsAsFactors = FALSE),
            by = 'spp_group')

write.csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'))

```

### 2015-12-28 data_explore/data_explore.Rmd

@@
 ### this file created in data_prep
-read.csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE)
+spp_list <- read.csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE)
 

### 2015-12-28 ms_2_intro_res_disc.Rmd

@@
+---
+output: pdf_document
+---
+
 # Introduction
 
-Mapping and predicting species distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
-Knowing where individuals of a species exist, and what allows them to persist there, provides foundational information for understanding species ranges, how best to protect and manage species, and how they may respond to increasing human impacts and a changing climate.
+___Peer?
+Dive?
+Cast a scientific eye?___ _beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean.
+Which species would you expect to find, and which species would be woefully out of place?_ Mapping and predicting species distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
+Knowing where individuals of a species exist, and where they thrive, provides foundational information for
+ understanding species ranges and diversity,
+ predicting species responses to human impacts and climate change,
+ and managing and protecting species effectively.
 A rich literature tackles the many dimensions of these questions.
 
-One very important outcome of this body of science is the various compiled databases of species distribution maps.
-In the oceans there are now two such global repositories _(if we define it narrowly; OBIS is global species distribution, but not quite range maps)_ – Aquamaps predicted distribution maps (Kaschner et al.
-2013) and International Union for Conservation of Nature (IUCN) range maps (REF) – that are used for a wide range of purposes, including assessing marine species status (Halpern et al.
+One major outcome of this body of science is the various compiled databases of species distribution maps.
+<!--In the oceans there are now two such global repositories--> Two important repositories _predict_ marine species ranges throughout the world's oceans  <!--_(if we define it narrowly; OBIS is global species occurrences, but not quite range or distribution maps)_--> – AquaMaps modeled species distribution maps (Kaschner et al.
+2013) and International Union for Conservation of Nature (IUCN) species range maps (REF).
+These spatial datasets are used for a wide range of purposes, including assessing marine species status (Halpern et al.
 2012, Selig et al.
@@
 2015).
-The two data collections ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
+The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
 
-Importantly, biases in taxonomic or spatial coverage of a dataset could shift management or conservation actions _towards places or species that aren’t actually the most in need (shift management or conservation actions away from places or species that are most in need?)_.(maybe Jetz 2008?) Errors of commission (false indications of presence) or omission (false indications of absence) could lead to inefficient or ineffective marine reserve systems and management plans.
+Importantly, biases in taxonomic or spatial coverage within a dataset could shift management or conservation actions away from places or species that are most in need.(maybe Jetz 2008?)  Errors of commission (false indications of presence) or omission (false indications of absence) could lead to inefficient or ineffective marine reserve systems and management plans.
 (Rondinini et al.
@@
 
-To understand the implications of differences between the two datasets, we compared how each data set represents the global spatial distribution of species.
-For the relatively small number of species mapped in both datasets, we examined how well their spatial distribution maps align.
-We then subjected two recent marine biodiversity studies to a sensitivity analysis, highlighting the consequences of different data use decisions _on our understanding of the status of marine biodiversity._
+To understand the implications of differences between these two datasets, we compared how each data set represents the global spatial and taxonomic distribution of species.
+For the relatively small number of species mapped in both datasets, we examined how well the species maps align.
+We then subjected two recent marine biodiversity studies - the Ocean Health Index biodiversity goal (Halpern et al.
+2012) and a global MPA gap analysis (Klein et al.
+2015) - to a sensitivity analysis, substituting one dataset over the other, to highlight the possible consequences of different data use decisions on our understanding of _the status of_ marine biodiversity.
 
@@
 
-# Overview of the datasets
+# Overview of AquaMaps and IUCN datasets
 
+<!--
 ___We may need to shorten this section to fit this paper into the format guidelines for PNAS.  If so, some of this can be pushed to the Suppl. Materials. - BH___
+-->
 
-The IUCN publishes species distribution maps as vector shapefiles, collected by taxonomic groups.
-Species experts outline polygons to represent a species' extent of occurrence, based on observation records <!--point locality databases such as the Ocean Biogeographic Information System (OBIS) (REF?
-not actually using this data) and the Global Biodiversity Information Facility (GBIF) (REF?),--> and refined by the expert's understanding of the species' range and habitat preferences.
-IUCN publishes range map sets only for taxonomic groups that have been "comprehensively assessed," i.e.
-in which at least 90% of the species within the taxonomic group have been evaluated.
-While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
+The IUCN publishes species range maps as spatial vector polygon shapefiles, bundled by taxonomic groups.
+Using GIS, experts outline spatial polygons to represent a given species' extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
+_IUCN releases a range map bundle for a taxonomic group once it has been "comprehensively assessed," i.e.
+at least 90% of the species within the taxonomic group have been evaluated (REF)._  While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
 
 AquaMaps develops species distribution maps based on modeled relative environmental suitability.
-For each modeled species, environmental preferences (e.g.
-temperature, depth, salinity) are deduced from occurrence records _such as OBIS and GBIF_ ___(should we mention these, or leave out?
-I left them out when describing the IUCN data sets)___, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree global grid, creating a global map of probability of occurrence.
+For each mapped species, environmental preferences (e.g.
+temperature, depth, salinity) are deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species.
 
+IUCN range maps are intended to include all possible regions in which a species may be present, with the caveat that this does not imply the species is evenly distributed everywhere within the boundaries of the map(REF).
+AquaMaps modeled distribution maps provide a more nuanced prediction of species presence within the extent of occurrence, but the model may not capture all the complexities that drive species distribution (REF).
+Due to these differences in methodology and intent, IUCN range maps are more likely to overpredict species presence while AquaMaps modeled distribution maps are more likely to underpredict species presence.(REF and REF)
+
+<!--
 * _How does AquaMaps select species?  based on data available from OBIS/GBIF and Fishbase?_
+-->
 
@@
 
-### Overlap between assessed species
+### Taxonomic distribution between datasets
 
@@
 n_spp_both  <- nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))
-p_both_am   <- round(n_spp_both/n_spp_am, 1)
-p_both_iucn <- round(n_spp_both/n_spp_iucn, 1)
+p_both_am   <- round(n_spp_both/n_spp_am * 100, 1)
+p_both_iucn <- round(n_spp_both/n_spp_iucn * 100, 1)
 ```
 
-IUCN maps are best categorized as geographic range data, while AquaMaps maps can be categorized as predicted distribution data (Rondinini et al., 2006) _(the categories come from here, but not my categorizations of these particular datasets...)_; the two types of spatial data differ in methodology and intent, and this drives significant differences between the resulting species distribution maps (Fig.
-1).
-Only `r n_spp_both` species were included in both datasets (`r p_both_am`% of total Aquamaps species; `r p_both_iucn`% of total IUCN species; Fig.
+Each dataset offers spatial distribution information for large numbers of species (`r n_spp_am` AquaMaps-mapped species; `r n_spp_iucn` IUCN-mapped marine species).
+However, the datasets vary in terms of taxonomic coverage and regional coverage, with only `r n_spp_both` species included in both datasets (`r p_both_am`% of AquaMaps species; `r p_both_iucn`% of IUCN marine species; Fig.
 1, 2).
-_All IUCN-mapped species are also included in Red List species, but only `r p_spp_am_rl`% of AquaMaps species._
 
-### Taxonomic distribution between datasets
+The distribution of IUCN-mapped species skews toward tropical latitudes and away from the Atlantic and Eastern Pacific compared to the distribution of AquaMaps-mapped species.(FIG)  This skew likely reflects the fact that the IUCN dataset focuses more heavily on coral reef-associated taxa than does the AquaMaps dataset (see FIG).
 
-The distribution of IUCN-mapped species skews toward tropical latitudes and away from the Atlantic and Eastern Pacific compared to the distribution of AquaMaps-mapped species.
-This skew likely reflects the fact that the IUCN dataset focuses more heavily on coral reef-associated taxa than does the AquaMaps dataset (see fig.
-XXX).
+For spatial assessments of biodiversity, the choice of one dataset over the other is likely to create significantly different results.
+For studies confined to a narrow range of taxa or to a narrow spatial scale, one dataset may offer an advantage over the other in the number of mapped species available.
+For global scale biodiversity studies, however, the selection of one dataset over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth._use something from Rondinini here?_
 
-Each dataset offers spatial distribution information for large numbers of species.
-However, the datasets vary in terms of taxonomic coverage and regional coverage.
-For spatial assessments of biodiversity, the choice of one dataset over the other is likely to create significantly different results.
-For studies confined to a narrow range of taxa or to a narrow spatial scale, one dataset may offer an advantage over the other in the number of species maps available.
-For global scale biodiversity studies, however, the selection of one dataset over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth._use something from Rondinini here?_
+_Small overlap means that using both datasets in conjunction will add a huge number of species without duplicating efforts, if the differences between the two methods can be reconciled appropriately_
 
@@
 
-## Analysis of spatial alignment by species
-
 ### Defining spatial alignment between the two datasets
@@
 The IUCN and AquaMaps spatial datasets share `r nrow(spp_map_pairs)` species in common.
-If we were to examine each species' pair of maps side by side, we would hope to see spatial correlation both in the global pattern of species distribution (where on the map) and the extent of species range (how much of the map).
-Large discrepancies in distribution and range extent could indicate species that require further study to consolidate expert knowledge for IUCN range maps and improve the accuracy of AquaMaps species distribution models.
+If we were to examine each of these species' pair of maps side by side, we would hope to see spatial correlation both in the global pattern of species distribution (where on the map) and the extent of species range (how much of the map).
 
+<!-- moved to caption
 For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range to the larger range (how much of the map).
 For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.
+-->
 
-We found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of paired-map species, the IUCN distribution map indicated a larger species range than the AquaMaps map.
-This concurs with the general expectation that geographic range maps (e.g.
+In `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of cases, the IUCN range map indicated a larger species range than the AquaMaps map.
+This finding concurs with the general expectation that geographic range maps (e.g.
 IUCN) are more likely to over-predict presence than predicted distribution models (e.g.
@@
 
-___could use some help on organizing this part...
-too much info?
-should some go in the caption, and some in the body?
-thinking this way: quickie description of quadrants in caption of figure; then use the bar charts in Fig 4 to help explain some mechanisms of quadrants___
-
-* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges, as we hoped for well-understood species.
-Excellent, but not particularly interesting.
-* Species map pairs that fall within the upper left quadrant (quadrant 2) agree well in distribution, but disagree in extent.
-* The lower right quadrant (quadrant 3 ___note, "math" quadrant 3 would be lower left, not lower right, will this quadrant ID scheme be confusing?___) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more worrisome mismatch than that indicated by quadrant 2.
-* The lower left quadrant (quadrant 4) indicates species for which the map pairs fail to agree in both extent and distribution.
-
-Examining spatial alignment by taxonomic group (FIG 3a), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
+Examining spatial alignment by taxonomic group (FIG 3b), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
 _explain reasoning?_
 
-Coral species are predominantly found in quadrant 2 (`r n_q2_corals` coral species - `r round(n_q2_corals/n_corals * 100, 1)`% of all corals and `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in quadrant 2).
-<!--AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range (see SI).
+<!--AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range (see SOM).
 A similar mechanism may also drive poor extent alignment of other demersal and reef-associated organisms.-->
 
-___how to transition from above paragraph to this idea:___ Looking more closely at the species included in quadrant 2, we can identify several mechanisms that could cause extent misalignment.
+Quadrant 1 contains species whose map pairs agree in both spatial distribution and the extent of described ranges, as we would expect for well-understood species.
+Excellent and valuable, but not particularly interesting.
+The extent-misaligned maps contained in quadrant 2, however, provide a richer set of examples to understand the fundamental differences between the datasets; examining these, we can identify several mechanisms that could cause such extent misalignment.
+
 For `r n_q2_am_sm` of `r n_q2` species in this quadrant (`r round((n_q2_am_sm/n_q2) * 100, 1)`%), the IUCN extent is larger than the AquaMaps extent.
-By itself, this is not surprising; but for many of these species, a quick look at the maps (see SI for examples) shows that the IUCN range hews closely to the AquaMaps range, while including a wide buffer zone.
-We suspect that many of these extent-misaligned map pairs can be easily explained: most corals and reef-associated organisms prefer shallower waters; seafloor depth is explicitly modeled in AquaMaps, but not explicitly included in IUCN range considerations.
+By itself, this is not surprising; but for many of these species, a quick look at the maps (see SOM for examples) shows that the IUCN range hews closely to the AquaMaps range, while adding a significant buffer zone.
+Coral species dominate this quadrant (`r n_q2_corals` coral species - `r round(n_q2_corals/n_corals * 100, 1)`% of all corals and `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in quadrant 2).
+We suspect that many of these extent-misaligned map pairs can be easily explained: most corals and reef-associated organisms prefer shallower waters; seafloor depth is explicitly modeled in AquaMaps, but not explicitly considered in IUCN range considerations.
+___include recommendation for IUCN range maps to be clipped to bathymetry where appropriate?___
 
-* _Occasionally, the AquaMaps map conformed to a small, local subset of the IUCN map (see SI for examples), in some cases due to some point locality observations being rejected by AquaMaps experts (EXAMPLES) and in other cases due to differences in species identification._ ___bleh - this needs help.___
-
-Breaking down the quadrants by IUCN extinction risk categories (FIG 3c), we found that species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
-Does higher perceived risk lead to increased attention, and thus better understanding of species distribution?
-Or conversely, does increased attention to species distribution reveal more species at risk?
-Likely both mechanisms are at play on a case-by-case basis, depending on the species' taxon and region.
-_does this argument bear up to closer scrutiny?
-CR isn't dominated by Q1 any more_
+___wrap up something with quadrants 3 and 4:___ _In several instances, the AquaMaps map conformed to a small, local subset of the IUCN map (see SOM for examples - what was the Q4 graph?), in some cases due to some point locality observations being rejected by AquaMaps experts (EXAMPLES) and in other cases due to differences in species identification._ ___bleh - this needs help.
+Evidence?
+examples?___
 
@@
 _Other hypotheses/case studies to highlight other possible mechanistic differences?_
-
------
-
-## AquaMaps: Effect of changing "presence" threshold on apparent distribution
-
-AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value as indicated by IUCN maps.
-Many studies convert this AquaMaps probability to a simple presence value by assigning a threshold value (REF references here).
-A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
-_For the comparisons above, thresh = 0_
-
-At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a significant decrease in represented range, and some species lose nearly their entire range.
-Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.
-This pattern may not hold true for all subsets of AquaMaps species, however, whether subsetting by taxa or by georegion.
 
@@
 Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted.
-When IUCN data is the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change.
+When IUCN range maps are the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change.
 But
@@
 
-_Predictions:  IUCN overestimates extent esp for coastal species?
+_Predictions:  What happens when using a 0% threshold instead of a 50%?
+What happens when using IUCN rather than AM?
+IUCN overestimates extent esp for coastal species?
 More species will be included in MPAs so fewer apparent gap species.
 Included range area inside MPAs will increase.
-Policy implications?
-Less pressure to carefully consider MPA design; less resolution on_
+Policy implications?_

### 2015-12-28 ms_3_methods.Rmd

@@
 
-_taxonomic distribution comparison_
+### taxonomic distribution comparison
 
-___Either move to brief methods section above, and maybe even shorten - BH___: To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
-We then filtered the species list to those that have been evaluated for the IUCN Red List of Threatened Species.
+To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
+We then filtered the species list to those that have been evaluated for the IUCN Red List of Threatened Species (in SOM).
 
-___Either cut or move to data description section of methods above.
-- BH___: The IUCN releases spatial data sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed, to guard against sample bias (though non-comprehensive datasets are available for reptiles and marine fish).
+___Currently this is also included in body - take out and leave in here?
+Per BH: "Either cut or move to data description section of methods above."___: The IUCN releases spatial data sets when a taxonomic group (typically on the scale of order or family) has been comprehensively assessed, to guard against sample bias (though non-comprehensive datasets are available for reptiles and marine fish).
 As such, spatial data for many taxonomic classes remain unavailable, and within a class, the assessed sub-groups may not represent the entire class.
 
-_global distribution spatial comparison_
+### global spatial distribution comparison
 
-To compare the spatial representation of the two datasets directly, we first rasterized the IUCN species polygons to the same 0.5° cells as the AquaMaps species maps; species presence within a cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the sum of the species present.
-For the AquaMaps dataset, we determined per-cell species richness by summing all species with non-zero probability of occurrence.
+To compare the spatial representation of the two datasets directly, we first rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; species presence within a grid cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the count of the species present.
+For the AquaMaps dataset, we determined per-cell species richness by counting all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.
 We represented relative distribution of species richness for each dataset by plotting average species count against latitude and longitude.
-To highlight the differences between the two datasets, we also examined the relative difference in per-cell species richness against latitude and longitude.
-
-For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to best approximate the "extent of occurrence" as generally indicated by IUCN maps.
-To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0.05 to 1.00 and calculated the average species range relative to a zero threshold.
 
@@
 Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
+We used the same criteria as outlined above to determine species presence within a cell.
 

### 2015-12-29 data_explore/data_explore.Rmd
pushing updates to data_explore to see quadmaps.  Updates include changes to incorporate new IUCN data

@@
 
-setwd('data_explore')
+if(basename(getwd()) != 'data_explore') setwd('data_explore')
 source('data_explore_fxns.R')
@@
 <!-- # Multiple species maps per quadrant -->
-``` {r child = 'data_explore_pt3.Rmd', eval = FALSE}
+``` {r child = 'data_explore_pt3.Rmd', eval = TRUE}
 ```

### 2015-12-29 data_explore/data_explore_pt3.Rmd
pushing updates to data_explore to see quadmaps.  Updates include changes to incorporate new IUCN data

@@
 
-```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
-x <- 20
+x <- 10
 y <- nrow(spp_list_q1)
 
-spp_names_q1 <- spp_list_q1$sciname[seq(1, y, y/x)]
+spp_names_q1 <- spp_list_q1$sciname[seq(1, y, round(y/x))]
 
@@
   spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
+  text_tag <- sprintf('Distribution align: %s%%, Extent align: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
   spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 1: %s', spp, text_tag))
+    labs(title = sprintf('%s (%s), quadrant 1: %s', spp, spp_solo$spp_group_text, text_tag))
 
@@
 Plotting species from q2 (distribution-matched)
-```{r plot_spp_from_q2, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+```{r plot_spp_from_q2, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
-x <- 20
+x <- 10
 y <- nrow(spp_list_q2)
 
-spp_names_q2 <- spp_list_q2$sciname[seq(1, y, y/x)]
+spp_names_q2 <- spp_list_q2$sciname[seq(1, y, round(y/x))]
 
@@
   spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
+  text_tag <- sprintf('Distribution align: %s%%, Extent align: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
   spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 2: %s', spp, text_tag))
+    labs(title = sprintf('%s (%s), quadrant 2: %s', spp, spp_solo$spp_group_text, text_tag))
   print(spp_plot)
@@
 Plotting species from q3 (area-matched)
-```{r plot_spp_from_q3, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+```{r plot_spp_from_q3, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
@@
 
-spp_names_q3 <- spp_list_q3$sciname[seq(1, y, y/x)]
+spp_names_q3 <- spp_list_q3$sciname[seq(1, y, round(y/x))]
 
@@
   spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
+  text_tag <- sprintf('Distribution align: %s%%, Extent align: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
   spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 3: %s', spp, text_tag))
+    labs(title = sprintf('%s (%s), quadrant 3: %s', spp, spp_solo$spp_group_text, text_tag))
   print(spp_plot)
@@
 Plotting species from q4 (poorly matched)
-```{r plot_spp_from_q4, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+```{r plot_spp_from_q4, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
@@
 
-spp_names_q4 <- spp_list_q4$sciname[seq(1, y, y/x)]
+spp_names_q4 <- spp_list_q4$sciname[seq(1, y, round(y/x))]
 
@@
   spp_solo <- spp_list %>% filter(sciname == spp)
-  text_tag <- sprintf('Overlap: %s%%, area ratio: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
+  text_tag <- sprintf('Distribution align: %s%%, Extent align: %s%%', round(spp_solo$sm_perc, 1), round(spp_solo$area_ratio, 1))
   spp_plot <- spp_plot +
-    labs(title = sprintf('%s, quadrant 4: %s', spp, text_tag))
+    labs(title = sprintf('%s (%s), quadrant 4: %s', spp, spp_solo$spp_group_text, text_tag))
   print(spp_plot)

### 2015-12-29 ms_4_figs.Rmd
pushing updates to data_explore to see quadmaps.  Updates include changes to incorporate new IUCN data

@@
 
-### Figure 1(a)?:
-
-![](figures/bar_by_taxa_all.png)
-
-Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.
-
-### Figure 1 (b, c):
-
-![](figures/all_aquamaps_map.png)
-
-![](figures/all_iucn_map.png)
-
-Global marine species richness according to (a) AquaMaps dataset and (b) IUCN dataset.
-
-### Figure 3 (a):
-
-![](figures/AreaRatiovsPercOverlap.png)
-
-For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range to the larger range (how much of the map).
-For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.
-
-``` {r tally_quad_counts, echo = FALSE}
-dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
-area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
-q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
-```
-
-* The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges. 
-(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent.
-(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more worrisome mismatch than that indicated by quadrant 2.
-(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)
-* The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution. 
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)
-
-### Figure 3 (b):
-
-![](figures/barchart_spp_gp_quads.png)
-
-Spatial alignment of paired-map species by taxonomic group.
+Keep figures in the text while working, for easier editing and captioning.  
+Put figures here once the manuscript is done
 

### 2015-12-29 data_explore/data_explore.Rmd
data_explore updates

--- 2015-12-28 data_explore/data_explore.Rmd
@@
 setwd('data_explore') 
+  ### When knitting, it automatically sets WD to be this directory...
+  ### this command is just here to facilitate running when not in knit mode.
 source('data_explore_fxns.R')
-### When knitting this, it automatically sets WD to be this directory...
-### the 'setwd()' is there for running by chunk
 ```
@@
 ```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-### this file created in data_prep
-read.csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE)
+data_file <- file.path(dir_data, 'spp_list_w_area_all.csv')
+
+if (!file.exists(data_file)) {
+  ### species info list - scinames, IUCN species ID, AM species ID, category etc
+  spp_list_base <- read.csv(file.path(dir_data, 'spp_am_v_iucn.csv'),
+                            stringsAsFactors = FALSE)
+
+  create_spp_list(spp_list_base, data_file)
+}
+
+spp_list <- read.csv(data_file, stringsAsFactors = FALSE) 
+### At this point, spp_list contains subpops and aliases, and zero-area IUCNs
+
+spp_list <- spp_list %>%
+  filter(!is.na(sm_perc)) %>%
+  ### will ditch the ones where no IUCN polygon exists - or are these species
+  ### whose ranges only occur in cells beyond the LOICZID grid?
+  filter(!str_detect(spatial_source, 'subpop') & !str_detect(spatial_source, 'alias'))
+spp_list <- spp_list %>%
+  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
+                     stringsAsFactors = FALSE),
+            by = 'spp_group')
+
+write.csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
 
@@
 scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-  labs(x = bquote('Extent alignment '*alpha['ext']*' (%)'), 
+  labs(x = bquote('Area alignment '*alpha['area']*' (%)'), 
        y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
-       title = 'Distribution alignment vs Extent alignment',
+       title = 'Distribution alignment vs area alignment',
        color = 'Risk', size = '% ocean range')
@@
   filter(area_ratio < median(area_ratio) & sm_perc < median(sm_perc))
-
-spp_list_quads <- spp_list1 %>%
-  mutate(quad = NA,
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q1$sciname, 'q1', quad),
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q2$sciname, 'q2', quad),
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q3$sciname, 'q3', quad),
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q4$sciname, 'q4', quad))
-write.csv(spp_list_quads, file.path(dir_data, 'spp_list_quads.csv'), row.names = FALSE)
 
@@
 
-write.csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads.csv'), row.names = FALSE)
-
 ```
@@
 ``` {r examine_risk_quartiles, echo = FALSE, eval = TRUE}
+
+spp_list1 <- spp_list %>%
+  group_by(popn_category) %>%
+  mutate(n_spp = n()) %>%
+  ungroup()
+
+### Create species lists by quadrant
+spp_list_q1 <- spp_list1 %>%
+  filter(area_ratio >= median(area_ratio) & sm_perc >= median(sm_perc))
+spp_list_q2 <- spp_list1 %>%
+  filter(area_ratio < median(area_ratio) & sm_perc >= median(sm_perc))
+spp_list_q3 <- spp_list1 %>%
+  filter(area_ratio >= median(area_ratio) & sm_perc < median(sm_perc))
+spp_list_q4 <- spp_list1 %>%
+  filter(area_ratio < median(area_ratio) & sm_perc < median(sm_perc))
 

### 2015-12-29 ms_0_main.Rmd
data_explore updates

--- 2015-12-23 ms_0_main.Rmd
@@
     toc: yes
-  word_document: default
+  pdf_document:
+    number_sections: yes
+    toc: no
+  word_document: null
 ---

### 2015-12-29 data_explore/data_explore_quadmaps.Rmd

--- 2015-12-29 data_explore/data_explore.Rmd
@@
 ---
-title: 'Data Exploration: IUCN and AquaMaps'
+title: 'Data Exploration quad maps: IUCN and AquaMaps'
 output: html_document
@@
 
-``` {r child = 'data_explore_pt1.Rmd', eval = FALSE}
-```
-<!-- # bar chart: number of spp per group
-# scatter plot: area compare IUCN vs AM, including log plot and plot by species group
-# rangemap: Chromis alpha -->
-
-``` {r, child = 'data_explore_pt2.Rmd', eval = FALSE}
-```
-
-***
-
 Quadrant map and quad breakdown by taxa and risk
@@
 pct_overlap_quartiles <- quantile(spp_list$sm_perc, na.rm = TRUE)[2:4]
-
-### define representative species from each quadrant, for over-plotting
-# spp_quads <- c('q1' = 'Kajikia albida', 
-#                'q2' = 'Galaxea cryptoramosa', 
-#                'q4' = 'Acanthurus nigroris', 
-#                'q4a' = 'Metanephrops australiensis', 
-#                'q3' = 'Conus magnificus')
 
@@
 
-ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 6, height = 3.5)
-
 ```
 
-``` {r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+``` {r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+### define representative species from each quadrant, for over-plotting
+spp_quads <- c('q1' = 'Kajikia albida', 
+               'q2' = 'Conus episcopatus',
+               'q3' = 'Conus magnificus', 
+               'q4' = 'Acanthurus nigroris')
 
@@
 
+  cat(sprintf('%s: %s', spp_quad, spp))
   spp_plot <- plot_rangemap(spp)
   print(spp_plot)
-
-  ggsave(filename = file.path(dir_fig, sprintf('rangemap_%s_%s.png', spp_quad,
-                                               str_replace(tolower(spp), ' ', '_'))),
-         plot = spp_plot, 
-         width = 8, height = 4, dpi = 200)
   }
@@
 
-### Plot the bar chart
-barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
-                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
-                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'count', alpha = .5) +
-  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
-                               'q2 dist-matched'   = 'greenyellow', 
-                               'q3 area-matched'   = 'orange', 
-                               'q4 poorly matched' = 'orangered2')) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 2, color = 'grey20') +
-  coord_flip() +
-  labs(x = 'Taxonomic Group', 
-       y = 'Relative proportion of species per quadrant', 
-       title = 'Taxonomic group by quadrant',
-       fill = 'Alignment')
-
-print(barchart_spp_gp_quads)
-ggsave(plot = barchart_spp_gp_quads,
-       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
-       width = 6, height = 3.5)
-
-write.csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads.csv'), row.names = FALSE)
-
-```
-
-``` {r examine_risk_quartiles, echo = FALSE, eval = TRUE}
-
-### Summarize quadrant species list by species group (e.g. CORALS1)
-spp_risk_q1 <- spp_list_q1 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp = first(n_spp), n_spp_q1 = n())
-spp_risk_q2 <- spp_list_q2 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp = first(n_spp), n_spp_q2 = n())
-spp_risk_q3 <- spp_list_q3 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp = first(n_spp), n_spp_q3 = n())
-spp_risk_q4 <- spp_list_q4 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp = first(n_spp), n_spp_q4 = n())
-
-### Join species group by quadrant into a single dataframe
-spp_risk_quadrants <- spp_risk_q1 %>%
-  full_join(spp_risk_q2, by = c('popn_category', 'n_spp')) %>%
-  full_join(spp_risk_q3, by = c('popn_category', 'n_spp')) %>%
-  full_join(spp_risk_q4, by = c('popn_category', 'n_spp')) %>%
-  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
-  mutate(quad = str_replace(quad, 'n_spp_', ''),
-         pct_quad = n_quad/n_spp,
-         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
-         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))
-
-spp_risk_quadrants <- spp_risk_quadrants %>%
-  left_join(spp_risk_quadrants %>% 
-              filter(quad == 'q1') %>%
-              dplyr::select(popn_category, pct_q1 = pct_quad),
-            by = 'popn_category') %>%
-  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
-         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
-         quad = str_replace(quad, 'q3', 'q3 area-matched'),
-         quad = str_replace(quad, 'q4', 'q4 poorly matched')) %>%
-  left_join(spp_list %>%
-              dplyr::select(popn_category, cat_rank = category_score) %>%
-              unique() %>%
-              mutate(cat_rank = ifelse(popn_category == 'DD', -1, cat_rank)),
-            by = 'popn_category')
-
-### Plot the bar chart
-barchart_spp_risk_quads <- ggplot(spp_risk_quadrants %>% 
-                                  transform(popn_category = reorder(popn_category, cat_rank)), 
-                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'count', alpha = .5) +
-  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
-                               'q2 dist-matched'   = 'greenyellow', 
-                               'q3 area-matched'   = 'orange', 
-                               'q4 poorly matched' = 'orangered2')) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 2, color = 'grey20') +
-  coord_flip() +
-  labs(x = 'Risk category', 
-       y = 'Relative proportion of species per quadrant', 
-       title = 'Risk category by quadrant',
-       fill = 'Alignment')
-
-print(barchart_spp_risk_quads)
-ggsave(plot = barchart_spp_risk_quads,
-       filename = file.path(dir_fig, 'barchart_spp_risk_quads.png'),
-       width = 6, height = 1.75)
-
 ```
@@
 ```
-
-***
-
-Breakdown maps by taxa
-
-``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.
-
-data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
-
-### generate other lists
-
-spp_all_file <- file.path(dir_data, 'spp_all.csv')
-spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
-
-iucn_marine_file <- file.path(dir_data, 'spp_iucn_marine.csv')
-spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
-
-message('Getting IUCN maps list')
-iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps.csv')
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
-  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
-  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
-
-spp_am_file <- file.path(dir_data, 'spp_am.csv')
-spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
-
-```
-
-``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE}
-
-sciname_lookup <- read.csv(file.path(dir_data, 'sciname_lookups/spp_iucn_scinames.csv'), stringsAsFactors = FALSE)
-
-iucn_maps_class <- spp_iucn_maps %>% 
-  left_join(sciname_lookup %>% 
-              dplyr::select(sciname, class) %>%
-              unique(), 
-            by = 'sciname')
-
-taxa_iucn_all <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
-  filter(!is.na(class)) %>% ### don't count the unknowns
-  mutate(class = str_to_title(class),
-         n_iucn_tot = n()) %>%        ### length of whole maps list
-  group_by(class) %>%
-  summarize(n_iucn_class_all  = n(),
-            p_iucn_class_all  = round((n_iucn_class_all/first(n_iucn_tot) * 100), 2))
-
-taxa_iucn_solo <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
-  filter(!is.na(class)) %>% ### don't count the unknowns
-  filter(!sciname %in% spp_am$sciname) %>%  ### filter list to IUCN only
-  mutate(class = str_to_title(class),
-         n_iucn_solo = n()) %>%       ### length of IUCN only list
-  group_by(class) %>%
-  summarize(n_iucn_class_solo = n(),
-            p_iucn_class_solo = round((n_iucn_class_solo/first(n_iucn_solo) * 100), 2))
-
-taxa_am_all <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  mutate(class = str_to_title(class),
-         n_am_tot = n()) %>%        ### length of whole AM maps list
-  group_by(class) %>%
-  summarize(n_am_class_all  = n(),
-            p_am_class_all  = round((n_am_class_all/first(n_am_tot) * 100), 2))
-
-taxa_am_solo <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
-  mutate(class = str_to_title(class),
-         n_am_solo = n()) %>%       ### length of AM only list
-  group_by(class) %>%
-  summarize(n_am_class_solo = n(),
-            p_am_class_solo = round((n_am_class_solo/first(n_am_solo) * 100), 2))
-
-taxa_both <- spp_am %>%
-  filter(sciname %in% spp_iucn_maps$sciname) %>%
-  filter(!class %in% c('Not assigned', 'null')) %>%
-  mutate(class = str_to_title(class),
-         n_both_tot = n()) %>%
-  group_by(class) %>%
-  summarize(n_class_both = n(),
-            p_class_both = round(n_class_both/first(n_both_tot) * 100, 2))
-
-taxa_am_all_assessed <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(sciname %in% spp_iucn_marine$sciname) %>%
-  mutate(class = str_to_title(class),
-         n_am_tot = n()) %>%        ### length of whole AM maps list
-  group_by(class) %>%
-  summarize(n_am_class_ass  = n(),
-            p_am_class_ass  = round((n_am_class_ass/first(n_am_tot) * 100), 2))
-
-taxa_am_solo_assessed <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(sciname %in% spp_iucn_marine$sciname) %>%
-  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
-  mutate(class = str_to_title(class),
-         n_am_solo = n()) %>%       ### length of AM only list
-  group_by(class) %>%
-  summarize(n_am_class_solo_ass = n(),
-            p_am_class_solo_ass = round((n_am_class_solo_ass/first(n_am_solo) * 100), 2))
-
-taxa_df <- taxa_am_all %>%
-  full_join(taxa_am_solo,   by = 'class') %>%
-  full_join(taxa_am_all_assessed,  by = 'class') %>%
-  full_join(taxa_am_solo_assessed, by = 'class') %>%
-  full_join(taxa_iucn_all,  by = 'class') %>%
-  full_join(taxa_iucn_solo, by = 'class') %>%
-  full_join(taxa_both,      by = 'class') %>%
-  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))
-
-knitr::kable(taxa_df %>%
-               dplyr::select(class, 
-                             AquaMaps = n_am_class_all, 
-                             IUCN = n_iucn_class_all, 
-                             Both = n_class_both))
-```
-
-``` {r plot bar charts by taxa, echo = FALSE, message = FALSE, warning = FALSE}
-### bar charts comparing number of species in each class for each data set
-
-taxa_df_long1 <- taxa_df %>%
-  dplyr::select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both) %>%
-  gather(dataset, n_spp, AquaMaps:Both)
-bar_taxa1 <- ggplot(taxa_df_long1, aes(x = class, y = n_spp)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'identity') +
-  facet_grid(dataset ~ ., scales = 'free_y') +
-  theme(axis.text.x = element_text(angle = 75, hjust = 1))
-print(bar_taxa1)
-
-taxa_df_long2 <- taxa_df %>%
-  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
-  gather(dataset, n_spp, AquaMaps:Both) %>%
-  group_by(class) %>%
-  mutate(n_total = sum(n_spp, na.rm = TRUE),
-         percent = n_spp/n_total * 100)
-bar_taxa2 <- ggplot(taxa_df_long2, aes(x = class, y = percent, fill = dataset)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'identity', alpha = .5) +
-  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
-  geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
-            size = 3, angle = 90, hjust = 0, color = 'grey30') +
-  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) + 
-  labs(x = 'Class', y = 'Percent of mapped species',
-       title = 'Species maps by taxonomic class')
-print(bar_taxa2)
-ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), height = 3, width = 6)
-
-taxa_df_long3 <- taxa_df %>%
-  dplyr::select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
-  gather(dataset, n_spp, AquaMaps:Both) %>%
-  group_by(class) %>%
-  mutate(n_total = sum(n_spp, na.rm = TRUE),
-         percent = n_spp/n_total * 100)
-bar_taxa3 <- ggplot(taxa_df_long3, aes(x = class, y = percent, fill = dataset)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'identity', alpha = .5) +
-  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
-  geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
-            size = 3, angle = 90, hjust = 0, color = 'grey30') +
-  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
-  labs(x = 'Class', y = 'Percent of assessed mapped species',
-       title = 'Species maps by taxonomic class, Red List-assessed only')
-print(bar_taxa3)
-ggsave(filename = file.path(dir_fig, 'bar_by_taxa_assessed.png'), height = 3, width = 6)
-
-```
-
-***
-
-``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
-```
-
-``` {r, child = 'data_explore_am_thresh.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
-```

### 2015-12-29 data_explore/data_prep.Rmd

@@
 library(data.table)
+library(ggplot2)
 library(tidyr)
@@
 dir_data_remote <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
+scenario <- 'v2016'
 
-setwd('data_explore')
+if(basename(getwd()) != 'data_explore') setwd('data_explore')
 source('data_explore_fxns.R')
@@
 ```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-### get spp_all_complete from git-annex, bring to local repo
-spp_all_remote <- read.csv(file.path(dir_anx, 'v2016/int/spp_all_complete.csv'), stringsAsFactors = FALSE)
+### get spp_all_complete from git-annex, bring to local repo.  The spp_all_complete
+### file was created in ohiprep/globalprep/SPP_ICO/data_prep_SPP.R, running the
+### function create_spp_master_lookup(source_pref = 'iucn', fn_tag = '_complete', reload = TRUE)
+spp_all_remote <- read.csv(file.path(dir_anx, scenario, 'int/spp_all_complete.csv'), 
+                           stringsAsFactors = FALSE)
 write.csv(spp_all_remote, file.path(dir_data, 'spp_all.csv'), row.names = FALSE)
 
-### create spp_am_v_iucn
+### create spp_am_v_iucn: the whole maplist narrowed down to just the overlapping species.
+### Overlaps detected from spatial_source containing 'iucn' (b/c source_pref = 'iucn' above),
+### that also have a non-NA AquaMaps ID code.
 spp_am_v_iucn <- spp_all_remote %>%
@@
   unique()
-write.csv(spp_am_v_iucn, file.path(dir_data, 'spp_am_v_iucn.csv'), row.names = FALSE)
+write.csv(spp_am_v_iucn, file.path(dir_data, 'spp_am_v_iucn.csv'), 
+          row.names = FALSE)
+
+### pull spp_iucn_marine.csv from git-annex and bring to local repo
+spp_iucn_marine <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_marine.csv'), 
+                            stringsAsFactors = FALSE)
+spp_iucn_marine <- spp_iucn_marine %>%
+  select(sciname, iucn_sid, iucn_category = category, popn_trend, parent_sid, subpop_sid)
+write.csv(spp_iucn_marine, file.path(dir_data, 'spp_iucn_marine.csv'), 
+          row.names = FALSE)
+
+### pull spp_iucn_maps_all.csv from git-annex and bring to local repo
+spp_iucn_maps <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv'), 
+                            stringsAsFactors = FALSE)
+spp_iucn_maps <- spp_iucn_maps %>%
+  filter(!str_detect(spp_group, 'MARINEFISH')) %>%
+  filter(id_no %in% spp_iucn_marine$iucn_sid)   ### ditch terrestrial and freshwater spp
+write.csv(spp_iucn_maps, file.path(dir_data, 'spp_iucn_maps.csv'), 
+          row.names = FALSE)
+
+### pull speciesoccursum.csv from git-annex and bring to local repo
+spp_am <- read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
+                   stringsAsFactors = FALSE)
+spp_am <- spp_am %>%
+  mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' '))
+write.csv(spp_am, file.path(dir_data, 'spp_am.csv'), 
+          row.names = FALSE)
 
@@
 ```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+### Create files for species list with area info, both a complete list and a trimmed list
+
 data_file <- file.path(dir_data, 'spp_list_w_area_all.csv')

### 2015-12-29 ms_0_main.Rmd

--- 2015-12-24 ms_0_main.Rmd
@@
     theme: spacelab
-    toc: no
-  pdf_document:
-    number_sections: yes
     toc: no

### 2015-12-29 ms_2_intro_res_disc.Rmd

@@
 
+![](figures/bar_by_taxa_all.png)
+
+> __Figure 1(a):__ Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.
+
+![](figures/all_aquamaps_map.png)
+
+![](figures/all_iucn_map.png)
+
+> __Figure 1 (b, c):__ Global marine species richness according to (a) AquaMaps dataset and (b) IUCN dataset.
+
 ``` {r calc species numbers, echo = FALSE}
@@
 
-<!-- moved to caption
-For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range to the larger range (how much of the map).
-For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.
--->
-
 In `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of cases, the IUCN range map indicated a larger species range than the AquaMaps map.
@@
 AquaMaps), while predicted distribution models are more likely to over-predict absence.(Rondinini et al., 2006) _errors of commission vs errors of omission, essentially type I and type II errors_
+
+![](figures/AreaRatiovsPercOverlap.png)
+
+> __Figure 3 (a):__ For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range to the larger range (how much of the map).
+For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.
+
+``` {r tally_quad_counts, echo = FALSE}
+dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
+area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
+q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
+q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
+q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
+q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
+```
+
+> The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
+(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
+> The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent
+(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
+> The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more worrisome mismatch than that indicated by quadrant 2
+(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
+> The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
+(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).
 
@@
 
-Examining spatial alignment by taxonomic group (FIG 3b), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
-_explain reasoning?_
+Examining spatial alignment by taxonomic group (FIG 3b), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms (50% or more of ocean area) such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
+_propose an explanation here?_
+
+![](figures/barchart_spp_gp_quads.png)
+
+> __Figure 3 (b):__ Spatial alignment of paired-map species by taxonomic group.
 
@@
 We suspect that many of these extent-misaligned map pairs can be easily explained: most corals and reef-associated organisms prefer shallower waters; seafloor depth is explicitly modeled in AquaMaps, but not explicitly considered in IUCN range considerations.
-___include recommendation for IUCN range maps to be clipped to bathymetry where appropriate?___
+___include recommendation for IUCN range maps to be clipped to bathymetry where appropriate?
+or save til end?___
 
@@
 Policy implications?_
+
+-----
+
+## Wrap it up
+
+___Recommendations on data use?___
+
+* IUCN mapmakers
+    * at least clip polygons to bathymetry for depth-sensitive species, come on now people
+    * think carefully about quality of data in -- e.g. are all OBIS points equally valuable?  AquaMaps seems to disagree - but are they justified in doing so?
+
+* AquaMaps mapmakers
+    * get expert eyeballs on all the maps
+    * Explain rationale for discarding some of the data points - can there be a defined decision tool for this?
+
+* Users
+    * understand that range maps tend to overestimate presence; AquaMaps are less likely to overestimate presence, but trade off for overestimating absence.
+Both better than simple point locality data.
+    * to use both together:
+        * consider setting AM presence threshold to 0% to create an extent of occurrence map, more like IUCN map.
+        * consider de-weighting IUCN for area-weighted calcs, for taxa in which IUCN is consistently overestimating presence

### 2015-12-29 ms_5_refs.Rmd

--- 2015-12-21 ms_5_refs.Rmd
@@
-#References
+# References
 
@@
 
-Kaschner, K., D.P. Tittensor, J. Ready, T Gerrodette and B. Worm (2011). Current and Future Patterns of Global Marine Mammal Biodiversity. PLoS ONE 6(5): e19653. PDF
+Kaschner, K., R.
+Watson, A.
+W.
+Trites, D.
+Pauly (2006).
+Mapping world-wide distributions of marine mammal species using a relative environmental suitability (RES) model.
+Marine Ecology Progress Series 316: 285–310.
+_check this citation journal name..._ _This outlines the basic RES methodology - AM development_
+
+Kaschner, K., D.P.
+Tittensor, J.
+Ready, T Gerrodette and B.
+Worm (2011).
+Current and Future Patterns of Global Marine Mammal Biodiversity.
+PLoS ONE 6(5): e19653.
+PDF _just what the title says - AM development, presence threshold 60%, also analyzes richness as a function of threshold_
 
@@
 Ecological Modelling 221(3): 467-478.
-PDF
+PDF _Presents AM; assessing AquaMaps against other presence-only species models_
 
-__Papers based on AquaMaps:__ ___Use these for section 4?___
+__Papers based on AquaMaps:__
 
@@
 Ecological Modelling 225(2012): 133-145.
-PDF
+PDF _comparison of species distribution models including AquaMaps, Maxent and the Sea Around Us Project_
 
-Coll, M., C. Piroddi, J. Steenbeek, K. Kaschner, F. Ben Rais Lasram et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. PDF
+Coll, M., C.
+Piroddi, J.
+Steenbeek, K.
+Kaschner, F.
+Ben Rais Lasram et al.
+(2010).
+The biodiversity of the Mediterranean Sea: estimates, patterns, and threats.
+PLoS ONE 5(8): e11842.
+PDF _used AquaMaps to predict Med biodiversity.
+Also: Threshold = 0._
 
@@
 (2014).
-Manual of marine nd coastal datasets of biodiversity importance.
+Manual of marine and coastal datasets of biodiversity importance.
 May 2014 release.
@@
 and one e-supplement).
-PDF
+PDF _report on marine data sets and data gaps etc, incl both IUCN and AM_
 
@@
 
-Hurlbert 2007
+Hurlbert 2007 Species richness, hotspots, and the scale dependence of range maps in ecology and conservation.
+_mostly rasters of range maps?
+"The scale dependence of range-map accuracy poses clear limitations on braod-scale ecological analyses and conservation assessments.
+...
+we provide guidance about the approriate scale of their use_
 
-Jetz 2008
+Jetz 2008 Ecological Correlates and Conservation Implications of Overestimating Species Geographic Ranges _EOO maps are usually highly interpolated and overestimate small-scale occurrence, which may bias research outcomes_
 
-Pimm 2014
+Pimm 2014 The biodiversity of species and their rates of extinction, distribtuion, and protection.
+_uses range maps to show biodiversity areas; may use IUCN range maps.
+Also discusses gaps and possible things that can be done about them._
 
-Rondinini 2006
+Rondinini 2006 Tradeoffs of different types of species occurrence data for use in systematic conservation planning _compares point locality, range maps, and distribution models in terms of omission and commission errors; also outlines Extent of Occurrence and Area of Occupancy distinctions._
 
@@
 
-<!--
-Halpin, P.N., A.J.
-Read, E.
-Fujioka, B.D.
-Best, B.
-Donnelly, L.J.
-Hazen, C.
-Kot, K.
-Urian, E.
-LaBrecque, A.
-Dimatteo, J.
-Cleary, C.
-Good, L.B.
-Crowder, and K.D.
-Hyrenbach.
-2009.
-OBIS-SEAMAP: The world data center for marine mammal, sea bird, and sea turtle distributions.
-Oceanography 22(2):104-115
--->
-
 Klein, Carissa J., Christopher J.
@@
 
-* _IUCN reference:_ 
+* IUCN reference:
     * https://www.conservationtraining.org/course/view.php?id=217&lang=en
     * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources
+    * IUCN Red List - accessed 12/21/2015
+
+* AquaMaps reference, and accessed date?

### 2015-12-29 ms_6_supplemental.Rmd

@@
 # Supplemental Information
+
+### illustrative maps for different quadrants and different mechanistic problems
 
@@
 
+___Red List inclusion:___
+
+![](figures/bar_by_taxa_assessed.png)
+
+* _All IUCN-mapped species are also included in Red List species, but only `r p_spp_am_rl`% of AquaMaps species._
+
+![](figures/barchart_spp_risk_quads.png)
+
+* Breaking down the quadrants by IUCN extinction risk categories (FIG 3c), we found that species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
+Does higher perceived risk lead to increased attention, and thus better understanding of species distribution?
+Or conversely, does increased attention to species distribution reveal more species at risk?
+Likely both mechanisms are at play on a case-by-case basis, depending on the species' taxon and region.
+_does this argument bear up to closer scrutiny?
+CR isn't dominated by Q1 any more_
+
 __Add'l maps to show mechanistic causes of quadrants__
+
+## AquaMaps: Effect of changing "presence" threshold on apparent distribution
+
+### AquaMaps presence threshold analysis - move to SOM
+
+For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to best approximate the "extent of occurrence" as generally indicated by IUCN maps.
+To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0.05 to 1.00 and calculated the average species range relative to a zero threshold.
+
+### Figure 5 (a, b):
+
+![](figures/AMThreshold.png)
+
+![](figures/ribbon_AMThreshold.png)
+
+>AquaMaps distribution map extent remaining after applying a presence threshold.
+(a) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of XXX, with a wide distribution in which some species lose nearly all of their apparent range.
+(b) Mean (median) remaining extent at increments of presence threshold.
+Dark grey ribbon includes 25% to 75% quantiles, while light grey ribbon includes 9% to 91% quantiles.
+
+AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value as indicated by IUCN maps.
+Many studies convert this AquaMaps probability to a simple presence value by assigning a threshold value (REF references here).
+A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
+
+At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a significant decrease in represented range, and some species lose nearly their entire range.
+Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.
+This pattern may not hold true for all subsets of AquaMaps species, however, whether subsetting by taxa or by georegion.
+
+-----

### 2016-01-12 data_explore/data_explore_quadmaps.Rmd
need to update the Rmd header on this - output info

@@
 title: 'Data Exploration quad maps: IUCN and AquaMaps'
-output: html_document
+output: pdf_document
 ---
@@
 
-  cat(sprintf('%s: %s', spp_quad, spp))
   spp_plot <- plot_rangemap(spp)
+  spp_plot <- spp_plot + 
+    labs(title = sprintf('%s: %s', spp_quad, spp))
+
   print(spp_plot)

### 2016-01-13 data_explore/data_explore.Rmd
updating manuscript, and incorporating info on AM occurrence cells

--- 2015-12-29 data_explore/data_explore.Rmd
@@
 ### this file created in data_prep
-spp_list <- read.csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE)
+spp_list <- read.csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'), 
+                     stringsAsFactors = FALSE)
 
@@
 
+spp_list <- spp_list %>% mutate(data_poor = occurcells < 10)
 ### mongo plot time
@@
                                        aes(x = area_ratio, y = sm_perc, 
-                                           color = popn_category,
+                                           color = data_poor,
                                            size = lg_area_pct)) + 

### 2016-01-13 data_explore/data_prep.Rmd
updating manuscript, and incorporating info on AM occurrence cells

@@
                      stringsAsFactors = FALSE),
-            by = 'spp_group')
+            by = 'spp_group') %>%
+  left_join(spp_am %>% select(am_sid = speciesid, occurcells),
+            by = 'am_sid')
 
-write.csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
+write.csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'), 
+          row.names = FALSE)
 

### 2016-01-13 ms_1_abst_sig.Rmd
updating manuscript, and incorporating info on AM occurrence cells

@@
 ## Abstract
+<!--
+Mapping and predicting species distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
+Knowing where individuals of a species exist, and where they thrive, provides foundational information for
+ understanding species ranges and diversity,
+ predicting species responses to human impacts and climate change,
+ and managing and protecting species effectively.
+A rich literature tackles the many dimensions of these questions.
+-->
+- Why are species distribution maps important
 
-_get to this later..._
+<!--
+Two important repositories _predict_ marine species ranges throughout the world's oceans – AquaMaps modeled species distribution maps (Kaschner et al.
+2013) and International Union for Conservation of Nature (IUCN) species range maps (REF).
+These spatial datasets are used for a wide range of purposes, including assessing marine species status (Halpern et al.
+2012, Selig et al.
+2013), evaluating global biodiversity patterns (Coll et al.
+2010, Martin et al.
+2014), predicting range shifts (Molinos et al.
+2015), and setting conservation priorities (Klein et al.
+2015).
+-->
+Species maps from AquaMaps and the International Union for Conservation of Nature describe marine species ranges throughout the world's oceans, and are widely used in a wide range of biodiversity and conservation purposes.
+___get across the point that these two datasets are the big ones___
+
+<!--
+The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
+
+Importantly, biases in taxonomic or spatial coverage within a dataset could shift management or conservation actions away from places or species that are most in need.(maybe Jetz 2008?)  Inaccurate indications of presence or absence could lead to inefficient or ineffective marine reserve systems and management plans.
+(Rondinini et al.
+2006) (Jetz 2008?)
+-->
+
+- These data sets are different, due to intent and methodology.
+- Spatial and taxa coverages are important to understand impacts on different analyses 
+    - taxa (probs) drives spatial
+    - small overlap provides opportunity to increase species without much redundancy
+- Comparisons of maps for overlapping species informs possible improvements
+    - improve method for outlining IUCN maps
+    - identify common problems and point toward solutions
+- How differences can affect outcomes of research and conservation studies
+
+<!--
+To understand the implications of differences between these two datasets, we compared how each data set represents the global spatial and taxonomic distribution of species.
+For the relatively small number of species mapped in both datasets, we examined how well the species maps align.
+We then subjected two recent marine biodiversity studies - the Ocean Health Index biodiversity goal (Halpern et al.
+2012) and a global MPA gap analysis (Klein et al.
+2015) - to a sensitivity analysis, substituting one dataset over the other, to highlight the possible consequences of different data use decisions on our understanding of _the status of_ marine biodiversity.
+-->
 

### 2016-01-13 ms_2_intro_res_disc.Rmd
updating manuscript, and incorporating info on AM occurrence cells

@@
 
-___Peer?
-Dive?
-Cast a scientific eye?___ _beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean.
-Which species would you expect to find, and which species would be woefully out of place?_ Mapping and predicting species distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
+_Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean.
+Which species would you expect to encounter, and which species would be woefully out of place?_ Mapping and predicting species distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
 Knowing where individuals of a species exist, and where they thrive, provides foundational information for
@@
 One major outcome of this body of science is the various compiled databases of species distribution maps.
-<!--In the oceans there are now two such global repositories--> Two important repositories _predict_ marine species ranges throughout the world's oceans  <!--_(if we define it narrowly; OBIS is global species occurrences, but not quite range or distribution maps)_--> – AquaMaps modeled species distribution maps (Kaschner et al.
+<!--In the oceans there are now two such global repositories--> Two important repositories _predict_ marine species ranges throughout the world's oceans – AquaMaps modeled species distribution maps (Kaschner et al.
 2013) and International Union for Conservation of Nature (IUCN) species range maps (REF).
@@
 2015).
+
 The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
 
-Importantly, biases in taxonomic or spatial coverage within a dataset could shift management or conservation actions away from places or species that are most in need.(maybe Jetz 2008?)  Errors of commission (false indications of presence) or omission (false indications of absence) could lead to inefficient or ineffective marine reserve systems and management plans.
+Importantly, biases in taxonomic or spatial coverage within a dataset could shift management or conservation actions away from places or species that are most in need.(maybe Jetz 2008?)  Inaccurate indications of presence or absence could lead to inefficient or ineffective marine reserve systems and management plans.
 (Rondinini et al.
@@
 AquaMaps modeled distribution maps provide a more nuanced prediction of species presence within the extent of occurrence, but the model may not capture all the complexities that drive species distribution (REF).
-Due to these differences in methodology and intent, IUCN range maps are more likely to overpredict species presence while AquaMaps modeled distribution maps are more likely to underpredict species presence.(REF and REF)
+Due to these differences in methodology and intent, IUCN range maps are more likely to overpredict species presence while AquaMaps modeled distribution maps are relatively more likely to underpredict species presence.(REF and REF)
 
@@
 Excellent and valuable, but not particularly interesting.
-The extent-misaligned maps contained in quadrant 2, however, provide a richer set of examples to understand the fundamental differences between the datasets; examining these, we can identify several mechanisms that could cause such extent misalignment.
+
+The extent-misaligned maps contained in quadrant 2 provide a richer set of examples to understand the fundamental differences between the datasets; examining these, we can identify several mechanisms that could cause such extent misalignment.
 
@@
 We suspect that many of these extent-misaligned map pairs can be easily explained: most corals and reef-associated organisms prefer shallower waters; seafloor depth is explicitly modeled in AquaMaps, but not explicitly considered in IUCN range considerations.
-___include recommendation for IUCN range maps to be clipped to bathymetry where appropriate?
+(For examples and details, see SOM) ___include recommendation for IUCN range maps to be clipped to bathymetry where appropriate?
 or save til end?___
 
+The extent-alignment of species maps found in quadrant 3 can in many cases be attributed to "two wrongs make a right."  For these species, IUCN ranges are frequently overextended into unsuitable habitat, as in the case of many quadrant 2 species, but at the same time, AquaMaps aggressively extrapolates distribution of presence into locations where IUCN predicts absence.
+This results in an area ratio close to 1, giving the impression of extent alignment.
 ___wrap up something with quadrants 3 and 4:___ _In several instances, the AquaMaps map conformed to a small, local subset of the IUCN map (see SOM for examples - what was the Q4 graph?), in some cases due to some point locality observations being rejected by AquaMaps experts (EXAMPLES) and in other cases due to differences in species identification._ ___bleh - this needs help.
@@
 
-Poorly-aligned rangemaps, regardless of taxon or extinction risk category, indicate species that could benefit from further expert study.
-___is this just 'duh'?
-is there something more important I can say?___
+Quadrant 4 species maps seem to suffer from the same problem as those in quadrant 3, but here the AquaMaps extrapolation extends even further, pushing the extent alignment ratio lower.
+In this quadrant, AquaMaps predicts a larger extent than IUCN for 70% of species represented in this quadrant, compared with just 30% of species in the entire set of overlapping maps.
+
+_how many species are "reviewed" in each quadrant? how many are data poor?_
+overall: 21% of overlapping species are data poor (model based on fewer than 10 cell-occurrences)
+q1: 8.2% (33/401); q2: 22.1% (151/682); q3: 16.5% (114/692); q4: 39.7% (159/401)
+Overall: 13.4% of species reviewed...
+q1: 24.9%; q2: 14.7%; q3: 9.5%; q4: 6.2%
 
@@
         * consider setting AM presence threshold to 0% to create an extent of occurrence map, more like IUCN map.
-        * consider de-weighting IUCN for area-weighted calcs, for taxa in which IUCN is consistently overestimating presence
+        * consider additional filters and criteria, e.g. "reviewed" or "occurcells" 
+        * consider clipping IUCN ranges to bathymetry for depth-limited species or taxa
+        * or consider de-weighting IUCN for area-weighted calcs, for taxa in which IUCN is consistently overestimating presence

### 2016-01-13 ms_6_supplemental.Rmd
updating manuscript, and incorporating info on AM occurrence cells

@@
 -----
+
+species to look more closely at:
+a few codes to indicate likely problems?
+- DC = depth clipping; IUCN extents go beyond reasonable depth for the species
+- DX = data excluded; AM excludes some observations that appear to be used for IUCN
+- DP = data poor (AquaMaps acknowledges data poor status)
+- NX = needs evaluation by expert; AM model likely predicts areas where species is unlikely to be found based on observed occurrences - does AM model reflect same as AM suitable habitat?
+
+Q2:
+Conus episcopatus (cones) DP, DX - d-match around Coral Triangle, but IUCN shows much more around Indian Ocean and S Pacific; also IUCN extents
+Ctenochaetus binotatus - IUCN extents beyond AM
+Naso vlamingii (acanthuridae) - IUCN extents beyond AM
+Thalassoma purpureum (wrasses) - IUCN extents beyond AM
+Porites nigrescens (corals 2) - DP - AM limited to CT; IUCN shows far greater dist, and extents
+Conus ammiralis (cones) - DP - AM limited to CT, IUCN shows in Madagascar; also extents
+Conus tessulatus (cones) - extents
+Acropora sarmentosa (corals 1) - extents; also AM shows HI but IUCN does not?
+Holothuria fuscogilva (sea cucumbers) - extents; also AM shows Mex and Central Am west coast, but IUCN does not (not available online?)
+Oculina Varicosa (corals 2) - lim to Caribbean; extents - this might be a good close-up map
+Acanthocybium solandri (tunas/billfish) - mostly tropical by IUCN; AM shows much farther N and S
+
+Q3:
+Conus magnificus (q3 - IUCN distribution is more limited than AM distribution)
+Abudefduf concolor (damselfish) - Baja for AM, southern Central Am for IUCN; maybe a good closeup map?
+AM points from Baja to Peru incl Galapagos; some rejected pts from Caribbean and eastern Atlantic
+Centropyge aurantonotus (angelfish) - AM shows Caribbean; IUCN shows northern South America, also maybe a good closeup map? AM points from FL to south Brazil, but no Gulf of Mexico.
+Chlorurus perspicillatus (parrotfish) - IUCN shows only HI; AM shows dots all over S Pacific and near Marianas Islands
+Sarda orientalis (tunas/billfish) - limited dists by IUCN
+Montastraea franksi (corals 2) - DP - IUCN shows just in Caribbean; AM shows all up & down east coast of N America and S America
+
+Q4:
+Acanthurus nigroris (q4 - IUCN dist limited to HI, while AM shows all over S Pacific and CT) - AM observations all over S pacific...
+Praealticus tanegasimae (blenny) - IUCN shows limited dist around Japan/Marianas; AM shows much broader through CT and W Pacific - good local map?
+Stethojulis marquesensis (wrasse) - IUCN shows only small cluster of islands (Marquesas?) while AM shows much broader across S Pacific
+Canthigaster leoparda (puffer) - IUCN limited spots around a couple of islands; AM shows all throughout CT
+pentacheles snyderi (lobster) - DP - AM shows all over - sim to depth profiles?
+while IUCN shows only a couple patches - HI?
+S Pacific, and S Indian oceans?
+AM limits to specific FAO regions, which results in odd cutoffs; also shows some data points in N Atlantic which has an odd result.
+Nephropsis sulcata (lobster) - IUCN shows odd ranges, most of which don't overlap with AM

### 2016-01-13 ms_1_abst_sig.Rmd
Getting the abstract started

@@
 ## Abstract
+
+_questions for abstract style: voice (passive vs active, 'we' vs 'this paper'), tense (present vs past)..._
+
 <!--
@@
 -->
-- Why are species distribution maps important
 
@@
 -->
-Species maps from AquaMaps and the International Union for Conservation of Nature describe marine species ranges throughout the world's oceans, and are widely used in a wide range of biodiversity and conservation purposes.
-___get across the point that these two datasets are the big ones___
+Two global datasets of species distribution maps - AquaMaps and the International Union for Conservation of Nature - dominate our understanding of marine species ranges throughout the world's oceans, and inform a wide range of biodiversity and conservation purposes.
+Differences in methodology and intent between these two datasets drive differences in predicted species ranges, with significant implications for management and conservation decisions.
 
@@
 -->
+Together, the two datasets provide range information for XXX species, with only `r nrow(spp_map_pairs)` species common to both datasets.
+___what is the key takeaway for the taxonomic and spatial coverage analyses?
+can we say something about benefits of combining the two for better taxo coverage?___  For the subset of species included in both datasets, we examine the alignment of range maps by comparing range distribution and extent.
+Categorizing the results based on quality of alignment, we identify several _problems?_ that commonly lead to disagreement in species range predictions.
+Notably, IUCN maps often predict species presence at unsuitable depths, and AquaMaps maps often extrapolate species presence far afield from known occurrences.
+___something about recommendations here?
+how this understanding can lead to better use of these datasets?___
 
-- These data sets are different, due to intent and methodology.
-- Spatial and taxa coverages are important to understand impacts on different analyses 
-    - taxa (probs) drives spatial
-    - small overlap provides opportunity to increase species without much redundancy
-- Comparisons of maps for overlapping species informs possible improvements
-    - improve method for outlining IUCN maps
-    - identify common problems and point toward solutions
-- How differences can affect outcomes of research and conservation studies
-
+To understand the implications of these differences, we reexamined two recent global analyses that depend on these datasets: the Ocean Health Index and a global analysis of gaps in coverage of marine protected areas.
+___how far to describe results here?___
 <!--

### 2016-01-14 data_explore/data_explore_global_maps.Rmd
making rasters from aquamaps csvs

@@
 ```{r global map matched species, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+
+library(rasterVis)
 
@@
 
+dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')
+
 if(!file.exists(all_am_rast_file)) {
@@
 
-am_levelplot1 <- levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
+am_levelplot1 <- levelplot(all_am_map, main = 'Entire Aquamaps Dataset') # add in par.settings = myTheme
 
-all_am_ln_map <- log(all_am_map + 1)
-all_am_ln_map <- all_am_ln_map/maxValue(all_am_ln_map)
-am_levelplot1 <- levelplot(all_am_ln_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset: log-transformed relative species richness', 
-          colorkey = list(at = seq(0, 1, length = 15)))
+# all_am_ln_map <- log(all_am_map + 1)
+# all_am_ln_map <- all_am_ln_map/maxValue(all_am_ln_map)
+# am_levelplot1 <- levelplot(all_am_ln_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset: log-transformed relative species richness', 
+#           colorkey = list(at = seq(0, 1, length = 15)))
 
 print(am_levelplot1)
-print(am_levelplot2)
+# print(am_levelplot2)
 
@@
 all_iucn_rast_file   <- file.path(dir_data, 'rasters/iucn_all_rast.tif')
+dir_data_iucn <- file.path(dir_N, 'git-annex/globalprep/_raw_data/iucn_spp')
 
@@
 
-iucn_levelplot1 <- levelplot(all_iucn_map, par.settings = myTheme, main = 'Entire IUCN Dataset')
+iucn_levelplot1 <- levelplot(all_iucn_map, main = 'Entire IUCN Dataset') # add in par.settings = myTheme
 
-all_iucn_ln_map <- log(all_iucn_map + 1)
-all_iucn_ln_map <- all_iucn_ln_map/maxValue(all_iucn_ln_map)
-iucn_levelplot2 <- levelplot(all_iucn_ln_map, 
-          par.settings = myTheme, 
-          margin = list(axis = TRUE),
-          main = 'Entire IUCN Dataset: log-transformed relative species richness', 
-          colorkey = list(at = seq(0, 1, length = 15)))
+# all_iucn_ln_map <- log(all_iucn_map + 1)
+# all_iucn_ln_map <- all_iucn_ln_map/maxValue(all_iucn_ln_map)
+# iucn_levelplot2 <- levelplot(all_iucn_ln_map, 
+#           par.settings = myTheme, 
+#           margin = list(axis = TRUE),
+#           main = 'Entire IUCN Dataset: log-transformed relative species richness', 
+#           colorkey = list(at = seq(0, 1, length = 15)))
 
 print(iucn_levelplot1)
-print(iucn_levelplot2)
+# print(iucn_levelplot2)
 # iucn_ln_pts <- as.data.frame(rasterToPoints(all_iucn_ln_map))

### 2016-01-14 data_explore/data_explore_fxns.R

@@ -253,4 +253,20 @@ plot_rangemap <- function(spp) {
     labs(title = spp, x = NULL, y = NULL) 
 }
 
+create_am_raster <- function(spp) {
+  map <- get_spp_map(spp)
+  
+  am_map <- map %>%
+    filter(!is.na(am_prob))
+  
+  r_am_spp  <-  subs(loiczid_raster, 
+                     am_map[ , c('loiczid', 'am_prob')], 
+                     by = 'loiczid', 
+                     which = 'am_prob', 
+                     subsWithNA = TRUE)
+  
+  write.raster(r_am_spp, file.path(dir_git, 'rasters', paste(str_replace(spp, ' ', '_'), '.tif', sep = '')))
+}
 
+# spp <- 'Oculina varicosa'
+# create_am_raster(spp)

### 2016-01-14 ms_6_supplemental.Rmd
updating SOM

--- 2015-12-29 ms_6_supplemental.Rmd
@@
 
-### illustrative maps for different quadrants and different mechanistic problems
+## Info on data prep
 
-### descriptions of IUCN and AM data sets
+### processing AquaMaps
+
+* start with .sql files - three of 'em - how to get 'em?
+* turn into .csvs, which columns are critical for this analysis? simplify
+
+### downloading and processing IUCN maps
+
+* which data sets are included?
+* raster::extract() to convert polys to csvs
+* which columns are included?
+
+### creating master species list for co-listed species
+
+* using AquaMaps and IUCN, and IUCN master list, create the big list
+* which columns are included?
+* adjust this master list to use IUCN SID only for parent - eliminate the whole question of parent/subpop? for the purpose of this analysis
+
+## descriptions of IUCN and AM data sets
 
@@
 
-__Add'l maps to show mechanistic causes of quadrants__
+## illustrative maps for different quadrants and different mechanistic problems
+
+a few codes to indicate likely problems?
+
+- DC = depth clipping; IUCN extents go beyond reasonable depth for the species
+- DX = data excluded; AM excludes some observations that appear to be used for IUCN
+- DP = data poor (AquaMaps acknowledges data poor status? or use less than 10 data points as criteria)
+- NX = needs evaluation by expert; AM model likely predicts areas where species is unlikely to be found based on observed occurrences - does AM model reflect same as AM suitable habitat?
+
+Q2:
+
+| species                | quad | group           | codes   | notes |
+| -------                | ---- | -----           | -----   | ----- |
+| Conus episcopatus      | q2   | cones           |  DP, DX | d-match around Coral Triangle, but IUCN shows much more around Indian Ocean and S Pacific; also IUCN extents
+| Ctenochaetus binotatus | q2   |                 |         | IUCN extents beyond AM
+| Naso vlamingii         | q2   | acanthuridae    |         | IUCN extents beyond AM
+| Thalassoma purpureum   | q2   | wrasses         |         | IUCN extents beyond AM
+| Porites nigrescens     | q2   | corals 2        | DP      | AM limited to CT; IUCN shows far greater dist, and extents
+| Conus ammiralis        | q2   | cones           | DP      | AM  | limited to CT, IUCN shows in Madagascar; also extents
+| Conus tessulatus       | q2   | cones           |         | extents
+| Acropora sarmentosa    | q2   | corals 1        |         | extents; also AM shows HI but IUCN does not?
+| Holothuria fuscogilva  | q2   | sea cucumbers   |         | extents; also AM shows Mex and Central Am west coast, but IUCN does not (not available online?)
+| Oculina Varicosa       | q2   | corals 2        |         | lim to Caribbean; extents - this might be a good close-up map
+| Acanthocybium solandri | q2   | tunas/billfish  |         | mostly tropical by IUCN; AM shows much farther N and S
+
+Q3
+
+| species                | quad | group           | codes   | notes |
+| -------                | ---- | -----           | -----   | ----- |
+| Conus magnificus       | q3   |  cones          |         | IUCN distribution is more limited than AM distribution)
+| Abudefduf concolor     | q3   |  damselfish     |         | Baja for AM, southern Central Am for IUCN; maybe a good closeup map?
+AM points from Baja to Peru incl Galapagos; some rejected pts from Caribbean and eastern Atlantic
+| Centropyge aurantonotus | q3  |  angelfish      |         | AM shows Caribbean; IUCN shows northern South America, also maybe a good closeup map?
+AM points from FL to south Brazil, but no Gulf of Mexico.
+| Chlorurus perspicillatus | q3 |  parrotfish     |         | IUCN shows only HI; AM shows dots all over S Pacific and near Marianas Islands
+| Sarda orientalis        | q3  |  tunas/billfish |         | limited dists by IUCN
+| Montastraea franksi     | q3  |  corals 2       | DP      | IUCN shows just in Caribbean; AM shows all up & down east coast of N America and S America
+
+Q4
+
+| species                  | quad | group   | codes   | notes |
+| -------                  | ---- | -----   | -----   | ----- |
+| Acanthurus nigroris      | q4   |         |         | IUCN dist limited to HI, while AM shows all over S Pacific and CT - AM observations all over S pacific...
+| Praealticus tanegasimae  | q4   | blenny  |         | IUCN shows limited dist around Japan/Marianas; AM shows much broader through CT and W Pacific - good local map?
+| Stethojulis marquesensis | q4   | wrasse  |         | IUCN shows only small cluster of islands (Marquesas?) while AM shows much broader across S Pacific
+| Canthigaster leoparda    | q4   | puffer  |         | IUCN limited spots around a couple of islands; AM shows all throughout CT
+| pentacheles snyderi      | q4   | lobster | DP      | AM shows all over - sim to depth profiles?
+while IUCN shows only a couple patches - HI?
+S Pacific, and S Indian oceans?
+AM limits to specific FAO regions, which results in odd cutoffs; also shows some data points in N Atlantic which has an odd result.
+| Nephropsis sulcata       | q4   | lobster |         | IUCN shows odd ranges, most of which don't overlap with AM
 

### 2016-01-14 ms_0_main.Rmd
updating SOM
various updates to manuscript

--- 2015-12-24 ms_0_main.Rmd
@@
 # dir_fig  <- file.path(getwd(), 'figures')
+
 ```

### 2016-01-14 ms_2_intro_res_disc.Rmd
updating SOM
various updates to manuscript

@@
 ---
-output: pdf_document
+output: html_document
 ---
@@
 We then subjected two recent marine biodiversity studies - the Ocean Health Index biodiversity goal (Halpern et al.
-2012) and a global MPA gap analysis (Klein et al.
+2012) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (Klein et al.
 2015) - to a sensitivity analysis, substituting one dataset over the other, to highlight the possible consequences of different data use decisions on our understanding of _the status of_ marine biodiversity.
@@
 
-![](figures/bar_by_taxa_all.png)
-
-> __Figure 1(a):__ Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.
-
-![](figures/all_aquamaps_map.png)
-
-![](figures/all_iucn_map.png)
-
-> __Figure 1 (b, c):__ Global marine species richness according to (a) AquaMaps dataset and (b) IUCN dataset.
-
 ``` {r calc species numbers, echo = FALSE}
@@
 n_spp_iucn  <- nrow(spp_iucn_maps)
-n_spp_both  <- nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))
+n_spp_both  <- nrow(spp_map_pairs)
 p_both_am   <- round(n_spp_both/n_spp_am * 100, 1)
@@
 Each dataset offers spatial distribution information for large numbers of species (`r n_spp_am` AquaMaps-mapped species; `r n_spp_iucn` IUCN-mapped marine species).
-However, the datasets vary in terms of taxonomic coverage and regional coverage, with only `r n_spp_both` species included in both datasets (`r p_both_am`% of AquaMaps species; `r p_both_iucn`% of IUCN marine species; Fig.
-1, 2).
+However, the datasets vary in terms of taxonomic coverage and regional coverage, with only `r n_spp_both` species included in both datasets (`r p_both_am`% of AquaMaps species; `r p_both_iucn`% of IUCN marine species; Figure 1a).
 
-The distribution of IUCN-mapped species skews toward tropical latitudes and away from the Atlantic and Eastern Pacific compared to the distribution of AquaMaps-mapped species.(FIG)  This skew likely reflects the fact that the IUCN dataset focuses more heavily on coral reef-associated taxa than does the AquaMaps dataset (see FIG).
+The distribution of IUCN-mapped species skews toward tropical latitudes and away from the Atlantic and Eastern Pacific compared to the distribution of AquaMaps-mapped species.(Figure 1b and 1c)  This skew likely reflects the fact that the IUCN dataset focuses more heavily on coral reef-associated taxa than does the AquaMaps dataset (see Figure 1a).
 
@@
 In `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of cases, the IUCN range map indicated a larger species range than the AquaMaps map.
-This finding concurs with the general expectation that geographic range maps (e.g.
-IUCN) are more likely to over-predict presence than predicted distribution models (e.g.
-AquaMaps), while predicted distribution models are more likely to over-predict absence.(Rondinini et al., 2006) _errors of commission vs errors of omission, essentially type I and type II errors_
+This finding concurs with the general expectation that geographic range maps such as IUCN are more likely to over-predict presence than predicted distribution models such as AquaMaps, while predicted distribution models are more likely to over-predict absence.(Rondinini et al., 2006) _errors of commission vs errors of omission, essentially type I and type II errors_
 
-![](figures/AreaRatiovsPercOverlap.png)
-
-> __Figure 3 (a):__ For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range to the larger range (how much of the map).
-For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.
-
-``` {r tally_quad_counts, echo = FALSE}
-dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
-area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
-q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
-```
-
-> The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
-(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
-> The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent
-(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
-> The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more worrisome mismatch than that indicated by quadrant 2
-(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
-> The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).
-
-Dividing the map-paired species into quadrants based on median values for each dimension, we can examine the _implications_ of four different qualities of alignment. (FIG 3a)
+Dividing the map-paired species into quadrants based on median values for each dimension, we can examine the _implications_ of four different qualities of alignment. (Figure 2a)
 
@@
 
-Examining spatial alignment by taxonomic group (FIG 3b), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms (50% or more of ocean area) such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
+Examining spatial alignment by taxonomic group (Figure 2b), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms (50% or more of ocean area) such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
 _propose an explanation here?_
-
-![](figures/barchart_spp_gp_quads.png)
-
-> __Figure 3 (b):__ Spatial alignment of paired-map species by taxonomic group.
-
-<!--AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range (see SOM).
-A similar mechanism may also drive poor extent alignment of other demersal and reef-associated organisms.-->
 
@@
 We suspect that many of these extent-misaligned map pairs can be easily explained: most corals and reef-associated organisms prefer shallower waters; seafloor depth is explicitly modeled in AquaMaps, but not explicitly considered in IUCN range considerations.
-(For examples and details, see SOM) ___include recommendation for IUCN range maps to be clipped to bathymetry where appropriate?
-or save til end?___
+(For examples and details, see SOM)
+
+* ___verify some of the coral maps by saving raster, showing raster and polygons in QGIS against a bathymetry layer___
+* ___include recommendation for IUCN range maps to be clipped to bathymetry where appropriate? or save til end?___
 
 The extent-alignment of species maps found in quadrant 3 can in many cases be attributed to "two wrongs make a right."  For these species, IUCN ranges are frequently overextended into unsuitable habitat, as in the case of many quadrant 2 species, but at the same time, AquaMaps aggressively extrapolates distribution of presence into locations where IUCN predicts absence.
-This results in an area ratio close to 1, giving the impression of extent alignment.
-___wrap up something with quadrants 3 and 4:___ _In several instances, the AquaMaps map conformed to a small, local subset of the IUCN map (see SOM for examples - what was the Q4 graph?), in some cases due to some point locality observations being rejected by AquaMaps experts (EXAMPLES) and in other cases due to differences in species identification._ ___bleh - this needs help.
-Evidence?
-examples?___
+This results in an area ratio close to 100%, giving the impression of extent alignment.
+___this quadrant seems to be a false alignment then?___
 
 Quadrant 4 species maps seem to suffer from the same problem as those in quadrant 3, but here the AquaMaps extrapolation extends even further, pushing the extent alignment ratio lower.
-In this quadrant, AquaMaps predicts a larger extent than IUCN for 70% of species represented in this quadrant, compared with just 30% of species in the entire set of overlapping maps.
+In this quadrant, AquaMaps predicts a larger extent than IUCN for 70% of species represented in this quadrant, compared with just 30% of species in the entire set of overlapping maps _(this is supposed to show that much of this quadrant is based on overextrapolation of AquaMaps)_.
 
-_how many species are "reviewed" in each quadrant? how many are data poor?_
-overall: 21% of overlapping species are data poor (model based on fewer than 10 cell-occurrences)
-q1: 8.2% (33/401); q2: 22.1% (151/682); q3: 16.5% (114/692); q4: 39.7% (159/401)
-Overall: 13.4% of species reviewed...
-q1: 24.9%; q2: 14.7%; q3: 9.5%; q4: 6.2%
+* ___Q4: more data-poor models and fewer reviewed species maps; data-rich models and expert review improve AquaMaps duh...___
+* ___recommend filtering for occurcells >= 10, and where possible, giving extra weight to reviewed maps___
+
+___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot___
+
+| quadrant | n species | n data poor species | n reviewed species |
+| :------: | :-------: | :-----------------: | :----------------: |
+|  total   |   22889   |        8749 (38.2%) |       1296  (5.7%) |
+|  paired  |    2166   |         457 (21.1%) |        290 (13.4%) |
+|   q1     |     401   |          33  (8.2%) |        100 (24.9%) |
+|   q2     |     682   |         151 (22.1%) |        100 (14.7%) |
+|   q3     |     682   |         114 (16.5%) |         65  (9.5%) |
+|   q4     |     410   |         159 (39.7%) |         25  (6.2%) |
+
+_data poor = fewer than 10 points used to generate the environmental envelopes; reviewed = map checked or formally expert-reviewed_
 
@@
 
-_do the analysis..._
+do the analysis...:
+
+* rerun based on old AquaMaps at 50% (to verify methodology gets a close result to paper), then new AquaMaps for new baseline
+* run with threshold at 0%
+* run with filtering out data-poor species?
+* run with IUCN maps instead of AquaMaps
 
@@
 What happens when using IUCN rather than AM?
-IUCN overestimates extent esp for coastal species?
+IUCN tends to overestimate extent esp for coastal species?
 More species will be included in MPAs so fewer apparent gap species.
@@
 * IUCN mapmakers
-    * at least clip polygons to bathymetry for depth-sensitive species, come on now people
-    * think carefully about quality of data in -- e.g. are all OBIS points equally valuable?  AquaMaps seems to disagree - but are they justified in doing so?
+    * at least clip polygons to bathymetry for depth-sensitive species, come on now people, it's really pretty obvious.  
+    * Note something about how photosynthetic marine species, and the species that depend on them, are far more sensitive to bathymetry than terrestrial species are to altitude... 
+    * Note that this is explicitly mentioned in the [training for IUCN red list marine maps](https://www.conservationtraining.org/mod/scorm/player.php?a=1412&currentorg=M5L1_ORG&scoid=6492) - "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species"
 
@@
     * get expert eyeballs on all the maps
-    * Explain rationale for discarding some of the data points - can there be a defined decision tool for this?
 
@@
         * consider setting AM presence threshold to 0% to create an extent of occurrence map, more like IUCN map.
-        * consider additional filters and criteria, e.g. "reviewed" or "occurcells" 
+        * consider additional filters and criteria, e.g. "reviewed" and/or "occurcells" to avoid over-extrapolation
         * consider clipping IUCN ranges to bathymetry for depth-limited species or taxa

### 2016-01-14 ms_4_figs.Rmd
various updates to manuscript

@@
 
-Keep figures in the text while working, for easier editing and captioning.  
-Put figures here once the manuscript is done
+### Figure 1
 
-### Figure 4
+![](figures/bar_by_taxa_all.png)
+``` {r barchart_by_taxa, echo = FALSE}
+n_spp <- spp_map_pairs %>%
+  mutate(spp_group = ifelse(str_detect(spp_group, 'CORAL'), 'CORAL', spp_group)) %>%
+  group_by(spp_group) %>%
+  summarize(n_spp = n()) %>%
+  arrange(desc(n_spp))
+
+n_corals   <- nrow(spp_map_pairs %>% 
+                     filter(str_detect(spp_group, 'CORAL')))
+n_bonyfish <- nrow(spp_map_pairs %>% 
+                     filter(spp_group %in% c('WRASSES', 'COMBTOOTHBLENNIES', 'BUTTERFLYFISH', 
+                                             'PUFFERFISH', 'GROUPERS', 'SEABREAMS_PORGIES', 'PARROTFISH', 
+                                             'SURGEONFISH_TANGS_UNICORNFISH', 'ANGELFISH', 'TUNAS_BILLFISHES', 
+                                             'DAMSELFISH', 'BONEFISH_TARPONS')))
+```
+> __Figure 1(a):__ Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.
+AquaMaps encompasses a broader range of taxa than IUCN, while IUCN focuses on comprehensively assessing select taxonomic groups, typically at the level of order or family.
+Overlapping species are dominated by bony fishes (`r n_bonyfish` species, primarily tropical taxa) and corals (`r n_corals` species).
+
+![](figures/all_aquamaps_map.png)
+
+![](figures/all_iucn_map.png)
+
+> __Figure 1 (b, c):__ Global marine species richness according to (a) AquaMaps dataset and (b) IUCN dataset.
+The frequency plot to the right of each map shows relative species count per cell at each latitude; while both datasets peak in tropical latitudes near the equator, the frequency for IUCN maps drops quickly beyond 30°N and 30°S, while the frequency for AquaMaps remains robust well into temperate latitudes.
+The frequency plot above each map shows relative species count at each longitude, showing a slight bias in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
+
+### Figure 2
+
+![](figures/AreaRatiovsPercOverlap.png)
+
+> __Figure 2 (a):__ For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range to the larger range (how much of the map).
+For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.
+
+``` {r tally_quad_counts, echo = FALSE}
+dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
+area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
+q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
+q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
+q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
+q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
+```
+
+> The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
+(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
+> The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent
+(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
+> The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more worrisome mismatch than that indicated by quadrant 2
+(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
+> The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
+(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).
+
+![](figures/barchart_spp_gp_quads.png)
+
+> __Figure 2 (b):__ Spatial alignment of paired-map species by taxonomic group.
+
+<!--AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range (see SOM).
+A similar mechanism may also drive poor extent alignment of other demersal and reef-associated organisms.-->
+
+### Figure 3
 

### 2016-01-14 ms_6_supplemental.Rmd
various updates to manuscript

--- 2016-01-13 ms_6_supplemental.Rmd
@@
 Holothuria fuscogilva (sea cucumbers) - extents; also AM shows Mex and Central Am west coast, but IUCN does not (not available online?)
-Oculina Varicosa (corals 2) - lim to Caribbean; extents - this might be a good close-up map
+Oculina varicosa (URM-Cn-5921) (corals part 1) - lim to Caribbean; extents - this might be a good close-up map
 Acanthocybium solandri (tunas/billfish) - mostly tropical by IUCN; AM shows much farther N and S

### 2016-01-14 ms_2_intro_res_disc.Rmd

@@
 
-For `r n_q2_am_sm` of `r n_q2` species in this quadrant (`r round((n_q2_am_sm/n_q2) * 100, 1)`%), the IUCN extent is larger than the AquaMaps extent.
+For `r n_q2_am_sm` of `r n_q2` species in quadrant 2 (`r round((n_q2_am_sm/n_q2) * 100, 1)`%), the IUCN extent is larger than the AquaMaps extent.
 By itself, this is not surprising; but for many of these species, a quick look at the maps (see SOM for examples) shows that the IUCN range hews closely to the AquaMaps range, while adding a significant buffer zone.
@@
 In this quadrant, AquaMaps predicts a larger extent than IUCN for 70% of species represented in this quadrant, compared with just 30% of species in the entire set of overlapping maps _(this is supposed to show that much of this quadrant is based on overextrapolation of AquaMaps)_.
-
-* ___Q4: more data-poor models and fewer reviewed species maps; data-rich models and expert review improve AquaMaps duh...___
-* ___recommend filtering for occurcells >= 10, and where possible, giving extra weight to reviewed maps___
 

### 2016-01-14 data_explore/data_explore.Rmd
updated species lists...?

--- 2015-12-29 data_explore/data_explore.Rmd
@@
                                          transform(popn_category = reorder(popn_category, category_score)),
-                                       aes(x = area_ratio, y = sm_perc, 
+                                       aes(x = area_ratio, 
+                                           y = sm_perc, 
                                            color = popn_category,
@@
            hjust = 0, vjust = 0,
-           color = 'grey50', size = 2,
+           color = 'grey40', size = 2,
            fontface = 'bold.italic', angle = 90,
@@
            hjust = 0, vjust = 0,
-           color = 'grey50', size = 2,
+           color = 'grey40', size = 2,
            fontface = 'bold.italic', angle = 0,
@@
 ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 6, height = 3.5)
+
+```
+
+``` {r pct_overlap_vs_ratio_of_area_w_data_poor, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+spp_list <- spp_list %>% mutate(data_poor = ifelse(occurcells < 10, '< 10', '>= 10')
+### mongo plot time
+scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
+                                         transform(popn_category = reorder(popn_category, category_score)),
+                                       aes(x = area_ratio, y = sm_perc, 
+                                           color = data_poor,
+                                           size = lg_area_pct)) + 
+  ggtheme_plot + 
+  ### color the quadrant backgrounds:
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
+                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
+                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
+                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
+    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
+                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
+  geom_point(alpha = .6)
+
+### Manage scales for color and size 
+scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+  scale_colour_manual(values = c('>= 10' = 'green4', 
+                                 '< 10'  = 'orangered2')) + 
+  scale_size_continuous(limits = c(0, 100),
+                        breaks = c( 0,    25,    50,    75,    100), 
+                        labels = c('0%', '25%', '50%', '75%', '100%')) + 
+  scale_x_continuous(expand = c(0, 0), limits = c(0, 100)) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
+
+### here are median labels:
+scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+  annotate(geom = 'text',
+           x = area_ratio_quartiles[2], y = 5,
+           hjust = 0, vjust = 0,
+           color = 'grey40', size = 2,
+           fontface = 'bold.italic', angle = 90,
+           label = sprintf('Median = %s%%', round(area_ratio_quartiles[2], 1))) +
+  annotate(geom = 'text',
+           x = 5, y = pct_overlap_quartiles[2],
+           hjust = 0, vjust = 0,
+           color = 'grey40', size = 2,
+           fontface = 'bold.italic', angle = 0,
+           label = sprintf('Median = %s%%', round(pct_overlap_quartiles[2], 1)))
+
+### here are median lines:
+# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+#   geom_vline(x = area_ratio_quartiles[2],
+#              color = 'grey50', linetype = 'dashed') +
+#   geom_hline(y = pct_overlap_quartiles[2],
+#              color = 'grey50', linetype = 'dashed')
+
+### here are 25% and 75% quantile lines:
+# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+#   geom_vline(x = area_ratio_quartiles[c(1, 3)],
+#              color = 'grey50', linetype = 'dashed') +
+#   geom_hline(y = pct_overlap_quartiles[c(1, 3)],
+#              color = 'grey50', linetype = 'dashed')
+
+### here are individual species, identified and named on the plot:
+# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+#   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
+#   geom_text(data = spp_list_label, 
+#             aes(label = sciname_label), 
+#             color = 'grey30',
+#             fontface = 'bold.italic',
+#             size = 2,
+#             hjust = .25,
+#             vjust = 1.5)
+
+scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+  labs(x = bquote('Extent alignment '*alpha['ext']*' (%)'), 
+       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
+       title = 'Distribution alignment vs Extent alignment',
+       color = 'AM model cells', size = '% ocean range')
+
+print(scatter_arearatio_vs_overlap)
+
+ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap-data_poor.png'), width = 6, height = 3.5)
 

### 2016-01-14 data_explore/data_explore_fxns.R
fixed write raster function call

@@ -265,7 +265,7 @@ create_am_raster <- function(spp) {
                      which = 'am_prob', 
                      subsWithNA = TRUE)
   
-  write.raster(r_am_spp, file.path(dir_git, 'rasters', paste(str_replace(spp, ' ', '_'), '.tif', sep = '')))
+  writeRaster(r_am_spp, file.path(dir_git, 'rasters', paste(str_replace(spp, ' ', '_'), '.tif', sep = '')))
 }
 
 # spp <- 'Oculina varicosa'

### 2016-01-14 data_explore/data_prep.Rmd

@@
 
-```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 ### get spp_all_complete from git-annex, bring to local repo.  The spp_all_complete
@@
 ### function create_spp_master_lookup(source_pref = 'iucn', fn_tag = '_complete', reload = TRUE)
-spp_all_remote <- read.csv(file.path(dir_anx, scenario, 'int/spp_all_complete.csv'), 
-                           stringsAsFactors = FALSE)
+spp_all_remote <- read.csv(file.path(dir_anx, scenario, 'int/spp_all.csv'), 
+                           stringsAsFactors = FALSE) %>%
+  filter(!str_detect(spatial_source, 'subpop')) %>%
+  filter(!str_detect(spatial_source,  'alias')) %>%
+  select(am_sid, sciname, reviewed, iucn_sid, popn_trend,
+         popn_category, spp_group, id_no, spatial_source)
+
 write.csv(spp_all_remote, file.path(dir_data, 'spp_all.csv'), row.names = FALSE)
@@
 data_file <- file.path(dir_data, 'spp_list_w_area_all.csv')
-reload = TRUE
+reload <- TRUE
 if (!file.exists(data_file) | reload) {

### 2016-01-14 ms_1_abst_sig.Rmd

@@
 ___what is the key takeaway for the taxonomic and spatial coverage analyses?
-can we say something about benefits of combining the two for better taxo coverage?___  For the subset of species included in both datasets, we examine the alignment of range maps by comparing range distribution and extent.
-Categorizing the results based on quality of alignment, we identify several _problems?_ that commonly lead to disagreement in species range predictions.
-Notably, IUCN maps often predict species presence at unsuitable depths, and AquaMaps maps often extrapolate species presence far afield from known occurrences.
+can we say something about benefits of combining the two for better taxo coverage?___
+
+For the subset of species included in both datasets, we examine the alignment of range maps by comparing distribution and extent of predicted ranges.
+Categorizing the results based on quality of alignment, we identify several issues that commonly lead to disagreement in species range predictions.
+Notably, IUCN maps often neglect considerations of bathymetry, leading to predictions of species presence at unsuitable depths, and AquaMaps maps for data-poor species often extrapolate presence far afield from known occurrences.
 ___something about recommendations here?
@@
 To understand the implications of these differences, we reexamined two recent global analyses that depend on these datasets: the Ocean Health Index and a global analysis of gaps in coverage of marine protected areas.
-___how far to describe results here?___
+Variations in which dataset is used, and how, lead to significantly different estimates of species diversity and conservation effectiveness.
+
 <!--

### 2016-01-14 ms_2_intro_res_disc.Rmd

@@
 Coral species dominate this quadrant (`r n_q2_corals` coral species - `r round(n_q2_corals/n_corals * 100, 1)`% of all corals and `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in quadrant 2).
-We suspect that many of these extent-misaligned map pairs can be easily explained: most corals and reef-associated organisms prefer shallower waters; seafloor depth is explicitly modeled in AquaMaps, but not explicitly considered in IUCN range considerations.
+We suspect that many of these extent-misaligned map pairs can be easily explained: most corals and reef-associated organisms prefer shallower waters; seafloor depth is explicitly modeled in AquaMaps, while depth is frequently ignored as a factor in IUCN range maps.
 (For examples and details, see SOM)
@@
 
-The extent-alignment of species maps found in quadrant 3 can in many cases be attributed to "two wrongs make a right."  For these species, IUCN ranges are frequently overextended into unsuitable habitat, as in the case of many quadrant 2 species, but at the same time, AquaMaps aggressively extrapolates distribution of presence into locations where IUCN predicts absence.
+The extent-alignment of species maps found in quadrant 3 can in many cases be attributed to "two wrongs make a right."  For these species, IUCN ranges are frequently overextended into unsuitable habitat, as in the case of many quadrant 2 species, while at the same time, AquaMaps aggressively extrapolates distribution of presence into locations where IUCN predicts absence.
 This results in an area ratio close to 100%, giving the impression of extent alignment.
@@
 In this quadrant, AquaMaps predicts a larger extent than IUCN for 70% of species represented in this quadrant, compared with just 30% of species in the entire set of overlapping maps _(this is supposed to show that much of this quadrant is based on overextrapolation of AquaMaps)_.
+Quadrant 4 in particular includes a high proportion of data-poor species in which the environmental envelope model is based on fewer than 10 known cell occurrences, and at the same time includes a smaller number of reviewed species maps than the other quadrants.
 
-___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot___
+___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot - should the quadrant plot show data-poor species instead of, say, red-list category?___
 
@@
 | :------: | :-------: | :-----------------: | :----------------: |
-|  total   |   22889   |        8749 (38.2%) |       1296  (5.7%) |
-|  paired  |    2166   |         457 (21.1%) |        290 (13.4%) |
+| overall  |   22889   |        8749 (38.2%) |       1296  (5.7%) |
+| AM&IUCN  |    2166   |         457 (21.1%) |        290 (13.4%) |
 |   q1     |     401   |          33  (8.2%) |        100 (24.9%) |

### 2016-01-15 data_explore/data_prep.Rmd
updated species area table

@@
 
+library(raster)
 library(readr)
@@
 
-if(basename(getwd()) != 'data_explore') setwd('data_explore')
+if(basename(getwd()) != 'data_explore') setwd('~/github/IUCN-AquaMaps/data_explore')
 source('data_explore_fxns.R')
@@
 
+loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
+loiczid_raster       <- raster(loiczid_raster_file)
+names(loiczid_raster) <- 'loiczid'
+
 ```
@@
 
-  create_spp_list(spp_list_base, data_file)
+  create_spp_list(spp_list_base, data_file, am_spp_cells, iucn_spp_cells, loiczid_raster)
 }
@@
 ### At this point, spp_list contains subpops and aliases, and zero-area IUCNs
+
+spp_am <- read.csv(file.path(dir_data, 'spp_am.csv'), stringsAsFactors = FALSE)
 

### 2016-01-15 data_explore/data_explore_fxns.R

@@ -25,7 +25,7 @@ ggtheme_map <- ggtheme_basic +
   theme(axis.text = element_blank())
 
 ### Function to create a species list from scratch... use when new data is available
-create_spp_list <- function(spp_list_base, data_file) {
+create_spp_list <- function(spp_list_base, data_file, am_spp_cells, iucn_spp_cells, loiczid_raster) {
   spp_list_build <- spp_list_base %>%
     mutate(total_area = NA,
            am_area    = NA,
@@ -38,9 +38,10 @@ create_spp_list <- function(spp_list_base, data_file) {
   
   
   #there's likely a better way to do this than a for loop (lapply or sapply?)
-  for (i in 1:nrow(spp_list_build)){ # i = 1
+  for (i in 1:nrow(spp_list_build)){ # i = 3
     spp     <- spp_list_build[i, ]$sciname #species name
-    spp_map <- get_spp_map(spp) #get species map
+    message(sprintf('Species: %s | am_sid: %s | iucn_sid: %s', spp, spp_list_build[i, ]$am_sid, spp_list_build[i, ]$iucn_sid))
+    spp_map <- get_spp_map(spp, spp_list_build, am_spp_cells, iucn_spp_cells) # get species map
     
     ###get just the aquamaps cells
     am_spp_map <- spp_map %>%
@@ -72,10 +73,6 @@ create_spp_list <- function(spp_list_base, data_file) {
       dplyr::select(loiczid, iucn_area, am_sid, am_prob, iucn_pres) %>%
       unique()
     
-    #TODO: incorporate the iucn_area (area of cell that species overlaps with...I think thats what it is)
-    
-    ###this is incorporated because there are some duplications for loiczids with IUCN cells. I'm guessing this is because of the parent/subpopulations?
-    
     if(TRUE %in% duplicated(iucn_map$loiczid)) next()
     
     ### rasterize IUCN map
@@ -92,7 +89,7 @@ create_spp_list <- function(spp_list_base, data_file) {
     ### Get total area (IUCN and AM combined)
     allCells <- spp_map %>%
       dplyr::select(loiczid) %>%
-      mutate(value=1) %>%
+      mutate(value = 1) %>%
       unique()
     
     
@@ -183,20 +180,22 @@ create_spp_list <- function(spp_list_base, data_file) {
 # This function takes a single species scientific name as input, then grabs all 
[…]

### 2016-01-15 data_explore/mpa_analysis.Rmd
trying to get MPA analysis started

---
title: 'MPA Gap Analysis: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(rgdal)
library(maptools)
library(raster)
library(readr)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

if(basename(getwd()) != 'data_explore') setwd('data_explore')

### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

# MPA gap analysis

Several steps to this:

* set up WDPA database in a raster format for easy use
* perform gap analysis for WDPA-MPA data base (which year?) vs AquaMaps 2014, with 50% threshold - replicate the original analysis
* reperform the analysis with latest/greatest WDPA-MPA database, and AquaMaps 2015, with 50% threshold and 0% threshold
* perform the analysis with latest/greatest WDPA-MPA database, and IUCN shapefiles

## set up WDPA raster

* direct extract method: extract WDPA polygons against half-degree cells, with proportion; value = year (or combo code)
    * all protected areas
    * 'marine' = 1
    * 'marine' = 1 and IUCN I - IV
    * clip WDPA to ocean regions, then IUCN I - IV.
    * NOTE: using extract() could result in double-counting - at least make sure total protected area per cell is not greater than 1.
        * maybe check any cells with protection > 50% and protected areas n > 1?
    * this method would be excruciatingly slow.
* to avoid double-counting for overlapping polygons - rasterize WDPA to fine scale first
    * less than 1km^2^ was mentioned in paper as resolution of MPA data.  Rasterize to 0.01 degree x 0.01 degree?  
        * this results in max cell size of 1.11 km x 1.11 km at the equator.
        * -180 to +180 longitude; min/max latitude use min/max from AquaMaps
    * WDPA_MPA layer: value = IUCN category (1-6, with NA = 7?) combined with marine flag and marine overlap flag?
        * first digit IUCN category; second digit marine flag; third digit marine overlap?
            * rasterize by IUCN category (unprotected, call NA?)
            * then rasterize again by marine field (0 or 1)
            * then rasterize land (or ocean) (0 or 1)
            * then add 'em up to construct full raster at 0.01 degree resolution
        * for year-to-year analysis, could recreate this layer for older years by filtering out newer protected areas
            * this allows for a cell's protection class to change over time?
            * could do this every, say, 5 years - to more quickly capture the trend across a span of, say, 25 years.
    * Cell area layer: use raster::area() to create a single layer of cell areas at this new resolution.
    * aggregate cells up to 0.5 degree cells, by protected area within each cell for each type of analysis
        * summarize to data.frame of loiczid, with area of protection (and total area?)?
        * create different protected-area layer for each protection scenario
* AquaMaps layer: use as is
* To analyze for each species:
    * Summarize total area and total protected area (for a given protection scenario) for the species
    * Include species extinction risk in there too?

``` {r set_up_base_raster}
### import AquaMaps half-degree loiczid raster - for extents and such
rast_loiczid_file   <- file.path(dir_git, 'data/rasters/loiczid_raster.grd')
rast_loiczid        <- raster(rast_loiczid_file)
names(rast_loiczid) <- 'loiczid'

rast_base      <- raster(ext = extent(rast_loiczid), resolution = 0.01)
rast_area_base <- raster::area(rast_base) ### takes a while...

### as long as we're here, let's set up some objects for WGS 84 projection
p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
crs_wgs84 <- CRS(p4s_wgs84)

```

``` {r set_up_spatially_marine_raster}
### import OHI regions polygons, to determine spatially-marine areas
dir_rgn   <- file.path(dir_N, 'git-annex/globalprep/spatial/v2015/data')
lyr_rgn <- 'regions_gcs'
poly_rgn <- readShapePoly(file.path(dir_rgn, lyr_rgn), 
                           proj4string = crs_wgs84)

### filter out all polygons that are land or landlocked, and flag the rest as spatially-marine = 1
poly_marine <- poly_rgn[!str_detect(poly_rgn@data$rgn_typ, 'land'), ]
poly_marine@data <- poly_marine@data %>%
  mutate(sp_mar_flag = 1)

# writePolyShape(poly_marine, file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_sp_mar'))

# rast_sp_mar <- rasterize(poly_marine, rast_base, value = sp_mar_flag) ### pretty slow... consider gdal_rasterize?

# library(gdalUtils)
# rast_sp_mar <- gdal_rasterize(
#     src_datasource = file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_sp_mar.shp')),
#     dst_filename = file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_sp_mar.tif'), 
#       # destination for output
#     a = 'sp_mar_flag', 
#       # the attribute in the shapefile to be assigned to the cell values
#     te = c(-180, -90, 180, 90), 
#       # extents for output raster
#     tr = c(.01, .01),  
#       # resolution for output raster
#     tap = TRUE, 
#       # target aligned pixels - align coords of extent of output to values of -tr
#     a_nodata = NA, 
#       # nodata value for raster; otherwise they will be filled in with zeroes
#     output_Raster = TRUE, 
#       # return output as a RasterBrick? 
#     verbose = TRUE)

```

``` {r set_up_wdpa_raster}

### import WDPA database polygons from original .gdb
# gdb_wdpa   <- file.path(dir_N, 'git-annex/globalprep/_raw_data/wdpa_mpa/WDPA_Jan2015_Public/WDPA_Jan2015_Public.gdb')
# lyr_wdpa <- 'WDPA_poly_Jan2015'
# 
# poly_wdpa <- readOGR(dsn = gdb_wdpa, layer = lyr_wdpa, stringsAsFactors = FALSE)

### import WDPA database polygons from saved .shp
dir_wdpa   <- file.path(dir_N, 'git-annex/globalprep/_raw_data/wdpa_mpa/WDPA_Jan2015_Public')
lyr_wdpa <- 'WDPA_poly_Jan2015'
poly_wdpa <- readShapePoly(file.path(dir_wdpa, lyr_wdpa), 
                           proj4string = crs_wgs84)

prot_code_df <- data.frame('IUCN_CLASS' = c('I', 'IIa', 'IIb', 'III', 'IV', 'V', 'V', NA),
                           'prot_code'  = c( 1,    2,     2,     3,     4,   5,   6,   7))
poly_wdpa@data <- poly_wdpa@data %>%
  left_join(prot_code_df, by = 'IUCN_CLASS') %>%
  mutate(prot_code = prot_code + MARINE)

```

### 2016-01-15 ms_1_abst_sig.Rmd
trying to get MPA analysis started

@@
 
-<!--
-Mapping and predicting species distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
-Knowing where individuals of a species exist, and where they thrive, provides foundational information for
- understanding species ranges and diversity,
- predicting species responses to human impacts and climate change,
- and managing and protecting species effectively.
-A rich literature tackles the many dimensions of these questions.
--->
-
-<!--
-Two important repositories _predict_ marine species ranges throughout the world's oceans – AquaMaps modeled species distribution maps (Kaschner et al.
-2013) and International Union for Conservation of Nature (IUCN) species range maps (REF).
-These spatial datasets are used for a wide range of purposes, including assessing marine species status (Halpern et al.
-2012, Selig et al.
-2013), evaluating global biodiversity patterns (Coll et al.
-2010, Martin et al.
-2014), predicting range shifts (Molinos et al.
-2015), and setting conservation priorities (Klein et al.
-2015).
--->
 Two global datasets of species distribution maps - AquaMaps and the International Union for Conservation of Nature - dominate our understanding of marine species ranges throughout the world's oceans, and inform a wide range of biodiversity and conservation purposes.
@@
 Categorizing the results based on quality of alignment, we identify several issues that commonly lead to disagreement in species range predictions.
-Notably, IUCN maps often neglect considerations of bathymetry, leading to predictions of species presence at unsuitable depths, and AquaMaps maps for data-poor species often extrapolate presence far afield from known occurrences.
+Notably, IUCN maps often disregard bathymetry for depth-limited species, leading to predictions of species presence at unsuitable depths, and AquaMaps maps for data-poor species often extrapolate presence far afield from known occurrences.
 ___something about recommendations here?
@@
 
-<!--
-To understand the implications of differences between these two datasets, we compared how each data set represents the global spatial and taxonomic distribution of species.
-For the relatively small number of species mapped in both datasets, we examined how well the species maps align.
-We then subjected two recent marine biodiversity studies - the Ocean Health Index biodiversity goal (Halpern et al.
-2012) and a global MPA gap analysis (Klein et al.
-2015) - to a sensitivity analysis, substituting one dataset over the other, to highlight the possible consequences of different data use decisions on our understanding of _the status of_ marine biodiversity.
--->
-
 -----

### 2016-01-15 ms_2_intro_res_disc.Rmd
trying to get MPA analysis started

@@
 AquaMaps modeled distribution maps provide a more nuanced prediction of species presence within the extent of occurrence, but the model may not capture all the complexities that drive species distribution (REF).
-Due to these differences in methodology and intent, IUCN range maps are more likely to overpredict species presence while AquaMaps modeled distribution maps are relatively more likely to underpredict species presence.(REF and REF)
+Due to these differences in methodology and intent, IUCN range maps are more likely to overpredict species presence (and underpredict absence) while AquaMaps modeled distribution maps are relatively more likely to underpredict species presence.(REF and REF)
 
@@
 Each dataset offers spatial distribution information for large numbers of species (`r n_spp_am` AquaMaps-mapped species; `r n_spp_iucn` IUCN-mapped marine species).
-However, the datasets vary in terms of taxonomic coverage and regional coverage, with only `r n_spp_both` species included in both datasets (`r p_both_am`% of AquaMaps species; `r p_both_iucn`% of IUCN marine species; Figure 1a).
+However, the datasets vary in terms of taxonomic coverage and regional coverage, with only `r n_spp_both` species included in both datasets (`r p_both_am`% of AquaMaps species; `r p_both_iucn`% of IUCN marine species; <a href = 'fig1'>Figure 1a</a>).
 
-The distribution of IUCN-mapped species skews toward tropical latitudes and away from the Atlantic and Eastern Pacific compared to the distribution of AquaMaps-mapped species.(Figure 1b and 1c)  This skew likely reflects the fact that the IUCN dataset focuses more heavily on coral reef-associated taxa than does the AquaMaps dataset (see Figure 1a).
+The distribution of IUCN-mapped species skews toward tropical latitudes and away from the Atlantic and Eastern Pacific compared to the distribution of AquaMaps-mapped species.(<a href = 'fig1'>Figure 1b and 1c</a>)  This skew likely reflects the fact that the IUCN dataset focuses more heavily on coral reef-associated taxa than does the AquaMaps dataset (see <a href = 'fig1'>Figure 1a</a>).
 
@@
 
-Dividing the map-paired species into quadrants based on median values for each dimension, we can examine the _implications_ of four different qualities of alignment. (Figure 2a)
+Dividing the map-paired species into quadrants based on median values for each dimension, we can examine the _implications_ of four different qualities of alignment. (<a href = 'fig2'>Figure 2a</a>)
 
@@
 
-Examining spatial alignment by taxonomic group (Figure 2b), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms (50% or more of ocean area) such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
+Examining spatial alignment by taxonomic group (<a href = 'fig2'>Figure 2b</a>), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms (50% or more of ocean area) such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
 _propose an explanation here?_
@@
 For `r n_q2_am_sm` of `r n_q2` species in quadrant 2 (`r round((n_q2_am_sm/n_q2) * 100, 1)`%), the IUCN extent is larger than the AquaMaps extent.
-By itself, this is not surprising; but for many of these species, a quick look at the maps (see SOM for examples) shows that the IUCN range hews closely to the AquaMaps range, while adding a significant buffer zone.
+For many of these species, a quick look at the maps (see SOM for examples) shows that the IUCN range hews closely to the AquaMaps range, while adding a significant buffer zone.
 Coral species dominate this quadrant (`r n_q2_corals` coral species - `r round(n_q2_corals/n_corals * 100, 1)`% of all corals and `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in quadrant 2).
@@
 
-Quadrant 4 species maps seem to suffer from the same problem as those in quadrant 3, but here the AquaMaps extrapolation extends even further, pushing the extent alignment ratio lower.
+Quadrant 4 species maps seem to suffer from the same issues as those in quadrant 3, but here the AquaMaps extrapolation extends even further, pushing the extent alignment ratio lower.
 In this quadrant, AquaMaps predicts a larger extent than IUCN for 70% of species represented in this quadrant, compared with just 30% of species in the entire set of overlapping maps _(this is supposed to show that much of this quadrant is based on overextrapolation of AquaMaps)_.
@@
 Spatial distributions were gleaned from both IUCN and AquaMaps datasets, preferring IUCN data for species represented in both data sets.
-OHI uses a probability threshold of 40% to determine presence for AquaMaps data.
+OHI uses a probability threshold of 40% to determine species presence for AquaMaps data.
 Species with no spatial data in either dataset were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
@@
 
-Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted.
+Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted (<a href = 'fig3'>Figure 3</a>).
 When IUCN range maps are the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change.
-But
 

### 2016-01-15 ms_4_figs.Rmd
trying to get MPA analysis started

@@
 ### Figure 1
+
+<a id = 'fig1'></a>
 
@@
 ```
-> __Figure 1(a):__ Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.
-AquaMaps encompasses a broader range of taxa than IUCN, while IUCN focuses on comprehensively assessing select taxonomic groups, typically at the level of order or family.
-Overlapping species are dominated by bony fishes (`r n_bonyfish` species, primarily tropical taxa) and corals (`r n_corals` species).
 
@@
 ![](figures/all_iucn_map.png)
+
+>  __Figure 1(a):__ Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.
+AquaMaps encompasses a broader range of taxa than IUCN, while IUCN focuses on comprehensively assessing select taxonomic groups, typically at the level of order or family.
+Overlapping species are dominated by bony fishes (`r n_bonyfish` species, primarily tropical taxa) and corals (`r n_corals` species).
 
@@
 ### Figure 2
+
+<a id = 'fig2'></a>
 
@@
 
+<a id = 'fig3'></a>
+
 ![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)

### 2016-01-16 ms_1_abst_sig.Rmd
Updating overall manuscript to get a draft to Ben

@@
 Two global datasets of species distribution maps - AquaMaps and the International Union for Conservation of Nature - dominate our understanding of marine species ranges throughout the world's oceans, and inform a wide range of biodiversity and conservation purposes.
-Differences in methodology and intent between these two datasets drive differences in predicted species ranges, with significant implications for management and conservation decisions.
+This paper examines differences in predicted species ranges between the two datasets, proposes mechanistic causes and potential solutions for such differences, and explores the implications of these differences for management and conservation decisions.
 
@@
 -->
-Together, the two datasets provide range information for XXX species, with only `r nrow(spp_map_pairs)` species common to both datasets.
+Together, the two datasets provide range information for `r nrow(spp_all)` unique species, with only `r nrow(spp_map_pairs)` species common to both datasets.
+When selecting one dataset over the other, spatial and taxonomic coverage should be considered against the scope of the project.
+<!--
 ___what is the key takeaway for the taxonomic and spatial coverage analyses?
-can we say something about benefits of combining the two for better taxo coverage?___
+can we say something about benefits of combining the two for better taxo coverage?___  Is this just tangential to the real storyline?
+I think the taxonomic differences are important considerations when exploring the implications for OHI and MPA gap analysis
+-->
 
-For the subset of species included in both datasets, we examine the alignment of range maps by comparing distribution and extent of predicted ranges.
-Categorizing the results based on quality of alignment, we identify several issues that commonly lead to disagreement in species range predictions.
+To understand mechanistic reasons for differences in species range predictions, we compare the distribution and extent of the pair of predicted ranges for the subset of species included in both datasets.
+Categorizing the results based on how well each pair of species maps aligns, we identify several issues that commonly lead to misalignment in species range predictions.
 Notably, IUCN maps often disregard bathymetry for depth-limited species, leading to predictions of species presence at unsuitable depths, and AquaMaps maps for data-poor species often extrapolate presence far afield from known occurrences.
-___something about recommendations here?
-how this understanding can lead to better use of these datasets?___
+Understanding these issues points toward solutions for providers and users of these datasets.
 
-To understand the implications of these differences, we reexamined two recent global analyses that depend on these datasets: the Ocean Health Index and a global analysis of gaps in coverage of marine protected areas.
+We reexamine two recent global analyses that depend on these datasets: the Ocean Health Index biodiversity goal and a global analysis of gaps in coverage of marine protected areas.
 Variations in which dataset is used, and how, lead to significantly different estimates of species diversity and conservation effectiveness.

### 2016-01-16 ms_2_intro_res_disc.Rmd
Updating overall manuscript to get a draft to Ben

@@
 
-_Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean.
-Which species would you expect to encounter, and which species would be woefully out of place?_ Mapping and predicting species distributions is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
-Knowing where individuals of a species exist, and where they thrive, provides foundational information for
+_Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean - and you will find a unique ecosystem defined as much by which species are absent as by which species are present._ Knowing where individuals of a species exist, and where they thrive, is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
+This knowledge provides foundational information for
  understanding species ranges and diversity,
@@
 One major outcome of this body of science is the various compiled databases of species distribution maps.
-<!--In the oceans there are now two such global repositories--> Two important repositories _predict_ marine species ranges throughout the world's oceans – AquaMaps modeled species distribution maps (Kaschner et al.
+Two global-scale repositories _predict_ marine species ranges throughout the world's oceans – AquaMaps modeled species distribution maps (Kaschner et al.
 2013) and International Union for Conservation of Nature (IUCN) species range maps (REF).
-These spatial datasets are used for a wide range of purposes, including assessing marine species status (Halpern et al.
+These two spatial datasets are used for a wide range of purposes, including assessing marine species status (Halpern et al.
 2012, Selig et al.
@@
 The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
-
-Importantly, biases in taxonomic or spatial coverage within a dataset could shift management or conservation actions away from places or species that are most in need.(maybe Jetz 2008?)  Inaccurate indications of presence or absence could lead to inefficient or ineffective marine reserve systems and management plans.
-(Rondinini et al.
+Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
+Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans.(Rondinini et al.
 2006) (Jetz 2008?)
@@
 To understand the implications of differences between these two datasets, we compared how each data set represents the global spatial and taxonomic distribution of species.
-For the relatively small number of species mapped in both datasets, we examined how well the species maps align.
+For the relatively small number of species mapped in both datasets, we examined how well the species maps align and determined several issues that lead to misalignment between predicted species distributions.
 We then subjected two recent marine biodiversity studies - the Ocean Health Index biodiversity goal (Halpern et al.
@@
 
-# Overview of AquaMaps and IUCN datasets
+## Overview of AquaMaps and IUCN datasets
 
-<!--
-___We may need to shorten this section to fit this paper into the format guidelines for PNAS.  If so, some of this can be pushed to the Suppl. Materials. - BH___
--->
+``` {r calc species numbers, echo = FALSE}
+n_spp_am    <- nrow(spp_am)
+n_spp_am_reviewed <- nrow(spp_am %>% filter(reviewed == 1))
+p_spp_am_reviewed <- round(n_spp_am_reviewed/n_spp_am * 100, 1)
+n_spp_am_rl <- nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))
+p_spp_am_rl <- round(n_spp_am_rl/n_spp_am * 100, 1)
+n_spp_iucn  <- nrow(spp_iucn_maps)
+n_spp_both  <- nrow(spp_map_pairs)
+p_both_am   <- round(n_spp_both/n_spp_am * 100, 1)
+p_both_iucn <- round(n_spp_both/n_spp_iucn * 100, 1)
+```
 
@@
 The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species.
+A subset of the resulting distribution maps (`r n_spp_am_reviewed` of `r n_spp_am` species maps; `r p_spp_am_reviewed`%) have been checked or formally reviewed by experts, to validate and improve the parameters and underlying model.
 
@@
 
-<!--
-* _How does AquaMaps select species?  based on data available from OBIS/GBIF and Fishbase?_
--->
-
 -----
@@
 
-<!--
-_Analyses [briefly describe the analyses done]_
-
-* _simple taxonomic and global spatial comparison (describe very briefly)_ 
-* _spatial overlap comparison (describe very briefly)_ For the relatively small number of species with maps in both datasets, we examined how well the maps align, and examined the implications of various categories of spatial misalignment.
-* _redo of OHI biodiversity goal (describe briefly what was done previously, and how that was changed for here)_ Using the Ocean Health Index biodiversity goal as one case study, we predicted the impact on OHI Species scores based on alternate decisions regarding how to use the two datasets.
-* _redo global MPA gap analysis (describe briefly what was done previously, and how that was changed for here)_ 
--->
+## Comparing the two datasets
 
@@
 
-``` {r calc species numbers, echo = FALSE}
-n_spp_am    <- nrow(spp_am)
-n_spp_am_rl <- nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))
-p_spp_am_rl <- round(n_spp_am_rl/n_spp_am * 100, 1)
-n_spp_iucn  <- nrow(spp_iucn_maps)
-n_spp_both  <- nrow(spp_map_pairs)
-p_both_am   <- round(n_spp_both/n_spp_am * 100, 1)
-p_both_iucn <- round(n_spp_both/n_spp_iucn * 100, 1)
-```
+___I wonder if this section is becoming less a part of the "story", maybe this part should go in SOM?___ _- though I think it's pretty important to understand that there is a huge difference in which species and taxonomic groups are covered in each set.
+That also helps to explain why there is such a small overlap, and helps show that the overlapping species are not necessarily representative?
+It also provides a logical lead-in from description of the overall datasets to the analysis of just the overlapping portion of the datasets._
 
 Each dataset offers spatial distribution information for large numbers of species (`r n_spp_am` AquaMaps-mapped species; `r n_spp_iucn` IUCN-mapped marine species).
-However, the datasets vary in terms of taxonomic coverage and regional coverage, with only `r n_spp_both` species included in both datasets (`r p_both_am`% of AquaMaps species; `r p_both_iucn`% of IUCN marine species; <a href = 'fig1'>Figure 1a</a>).
+However, the datasets vary in terms of taxonomic coverage and regional coverage, with only `r n_spp_both` species included in both datasets (`r p_both_am`% of AquaMaps species; `r p_both_iucn`% of IUCN marine species; <a href = '#fig1'>Figure 1a</a>).
 
-The distribution of IUCN-mapped species skews toward tropical latitudes and away from the Atlantic and Eastern Pacific compared to the distribution of AquaMaps-mapped species.(<a href = 'fig1'>Figure 1b and 1c</a>)  This skew likely reflects the fact that the IUCN dataset focuses more heavily on coral reef-associated taxa than does the AquaMaps dataset (see <a href = 'fig1'>Figure 1a</a>).
+The distribution of IUCN-mapped species skews toward tropical latitudes and away from the Atlantic and Eastern Pacific compared to the distribution of AquaMaps-mapped species.(<a href = '#fig1'>Figure 1b and 1c</a>)  This skew likely reflects the fact that the IUCN dataset focuses more heavily on coral reef-associated taxa than does the AquaMaps dataset (see <a href = '#fig1'>Figure 1a</a>).
 
-For spatial assessments of biodiversity, the choice of one dataset over the other is likely to create significantly different results.
+For spatial assessments of biodiversity, the choice of one dataset over the other is likely to result in significantly different conclusions.
 For studies confined to a narrow range of taxa or to a narrow spatial scale, one dataset may offer an advantage over the other in the number of mapped species available.
-For global scale biodiversity studies, however, the selection of one dataset over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth._use something from Rondinini here?_
+For global scale biodiversity studies, however, the selection of one dataset over the other will entail tradeoffs in spatial coverage and taxonomic breadth versus depth._use something from Rondinini here?_
 
@@
 The IUCN and AquaMaps spatial datasets share `r nrow(spp_map_pairs)` species in common.
-If we were to examine each of these species' pair of maps side by side, we would hope to see spatial correlation both in the global pattern of species distribution (where on the map) and the extent of species range (how much of the map).
+Placing each of these species' pair of maps side by side, we would hope to see spatial correlation both in the global pattern of species distribution (where on the map) and the extent of species range (how much of the map).
 
+<!-- NOT REALLY PART OF THE STORY? INCLUDE IN SOM?
 In `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of cases, the IUCN range map indicated a larger species range than the AquaMaps map.
-This finding concurs with the general expectation that geographic range maps such as IUCN are more likely to over-predict presence than predicted distribution models such as AquaMaps, while predicted distribution models are more likely to over-predict absence.(Rondinini et al., 2006) _errors of commission vs errors of omission, essentially type I and type II errors_
+This finding concurs with the general expectation that geographic range maps such as IUCN are more likely to over-predict presence and under-predict absence than distribution models such as AquaMaps.(Rondinini et al., 2006) _errors of commission vs errors of omission, essentially type I and type II errors - should I use one of these terms to clarify?_
+-->
 
-Dividing the map-paired species into quadrants based on median values for each dimension, we can examine the _implications_ of four different qualities of alignment. (<a href = 'fig2'>Figure 2a</a>)
+Dividing the map-paired species into quadrants based on median values for each dimension allows us to examine the _implications_ of four different qualities of alignment.
+(<a href = '#fig2'>Figure 2a</a>)
 
@@
 
-Examining spatial alignment by taxonomic group (<a href = 'fig2'>Figure 2b</a>), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms (50% or more of ocean area) such as marine mammals, tunas, and billfishes were more consistently well-aligned (quadrants 1 and 2) than demersal and reef organisms.
-_propose an explanation here?_
-
-Quadrant 1 contains species whose map pairs agree in both spatial distribution and the extent of described ranges, as we would expect for well-understood species.
-Excellent and valuable, but not particularly interesting.
+Quadrant 1 highlights species whose map pairs agree in both spatial distribution and the extent of described ranges, as we would expect for well-understood species.
+Examining spatial alignment by taxonomic group (<a href = '#fig2'>Figure 2b</a>), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms (50% or more of ocean area) such as marine mammals, tunas, and billfishes were more consistently well-aligned than demersal and reef organisms.
+_propose an explanation here?
+basically - these species go pretty much anywhere, so both maps predict most of the ocean, you'll get a good overlap - not really that revolutionary_
 
@@
 Coral species dominate this quadrant (`r n_q2_corals` coral species - `r round(n_q2_corals/n_corals * 100, 1)`% of all corals and `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in quadrant 2).
-We suspect that many of these extent-misaligned map pairs can be easily explained: most corals and reef-associated organisms prefer shallower waters; seafloor depth is explicitly modeled in AquaMaps, while depth is frequently ignored as a factor in IUCN range maps.
+Most corals and reef-associated organisms prefer shallower waters, and we suspect that many of these extent-misaligned map pairs can be easily explained: seafloor depth is explicitly included in AquaMaps models, while depth is frequently ignored as a factor in IUCN range maps, leading to overprediction of species range into areas too deep to support the species.
 (For examples and details, see SOM)
 
-* ___verify some of the coral maps by saving raster, showing raster and polygons in QGIS against a bathymetry layer___
+* ___verify some of the coral maps by saving raster, showing raster and polygons in QGIS against a bathymetry layer___ - put a map for Oculina varicosa (and others?) in the SOM
 * ___include recommendation for IUCN range maps to be clipped to bathymetry where appropriate? or save til end?___
 
-The extent-alignment of species maps found in quadrant 3 can in many cases be attributed to "two wrongs make a right."  For these species, IUCN ranges are frequently overextended into unsuitable habitat, as in the case of many quadrant 2 species, while at the same time, AquaMaps aggressively extrapolates distribution of presence into locations where IUCN predicts absence.
+The extent-alignment of species maps found in quadrant 3 can in many cases be attributed to "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable habitat, as in the case of many quadrant 2 species, while at the same time, AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence.
 This results in an area ratio close to 100%, giving the impression of extent alignment.
-___this quadrant seems to be a false alignment then?___
+___this quadrant seems to be a false alignment then?
+should there be a stronger warning here___
 
-Quadrant 4 species maps seem to suffer from the same issues as those in quadrant 3, but here the AquaMaps extrapolation extends even further, pushing the extent alignment ratio lower.
-In this quadrant, AquaMaps predicts a larger extent than IUCN for 70% of species represented in this quadrant, compared with just 30% of species in the entire set of overlapping maps _(this is supposed to show that much of this quadrant is based on overextrapolation of AquaMaps)_.
-Quadrant 4 in particular includes a high proportion of data-poor species in which the environmental envelope model is based on fewer than 10 known cell occurrences, and at the same time includes a smaller number of reviewed species maps than the other quadrants.
+Quadrant 4 species maps suffer from the same issues as those in quadrant 3, but here the AquaMaps extrapolation extends even further, pushing the extent alignment ratio lower.
+In this quadrant, AquaMaps predicts a larger extent than IUCN for 70% of species represented in this quadrant, compared with just 30% of species in the entire set of overlapping maps _(this is supposed to show that much of this quadrant is based on overextrapolation of AquaMaps...
+not sure that's clear yet)_.
+AquaMaps range maps in quadrant 4 include a high proportion of data-poor species in which the environmental envelope model is based on fewer than 10 known cell occurrences, and at the same time include fewer reviewed species maps than the other quadrants.
 
-___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot - should the quadrant plot show data-poor species instead of, say, red-list category?___
+___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot___ _- perhaps update the quadrant plot to reveal data-poor species (and reviewed species?
+little overlap of reviewed & data-poor) instead of, say, red-list category which doesn't get discussed in the body of the paper?
+then this table can go in SOM if we like it_
 
@@
 
-_Other hypotheses/case studies to highlight other possible mechanistic differences?_
+_Other hypotheses/case studies to highlight other possible mechanistic differences?_  I think we found a couple of the big ones; good enough?
 
 -----
-
-# Implications
 
@@
 The global Ocean Health Index (OHI) (Halpern et al.
-2012), an index made up of 10 goals, utilizes both of these datasets to inform the Species subgoal of the Biodiversity goal.
+2012), a composite index made up of 10 goals, utilizes both of these datasets to inform the Species subgoal of the Biodiversity goal.
 As it is currently calculated, the Species subgoal uses species spatial distribution data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
@@
 
-Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted (<a href = 'fig3'>Figure 3</a>).
+Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted (<a href = '#fig3'>Figure 3</a>).
 When IUCN range maps are the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change.
@@
 
-do the analysis...:
+Klein et al.
+(2015) compared the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect key species, and to identify which species fall through the gaps in protection.
+The MPA gap analysis relied on the AquaMaps database available in early 2015, comprising distribution maps for XXX species.
+Species presence was defined using a probability threshold of 50% or greater.
+The study defined four subsets of protected areas using the World Database of Protected Areas, based upon combinations of IUCN protected area classification, marine designation, and spatial marine coverage.
+The study determined that currently defined marine protected areas (IUCN categories I through IV overlapping with marine areas) are woefully inadequate in protecting marine biodiversity; 90.5% of species have less than 5% of their overall range represented within marine protected areas.
 
-* rerun based on old AquaMaps at 50% (to verify methodology gets a close result to paper), then new AquaMaps for new baseline
-* run with threshold at 0%
+working on the analysis...
+
+* rerun based on old AquaMaps at 50% (to verify methodology gets a close result to paper), then new AquaMaps for updated baseline
+    * consider also: filter on STATUS == 'Designated', and filter out those pesky fishery management areas and manatee speed zones...
+* run with threshold at 0% (not all the same as original paper)
 * run with filtering out data-poor species?
 * run with IUCN maps instead of AquaMaps
+* Which aspects of the analysis should we focus on?:
+    * overall protection, for 4th subset of protected areas? or overall for all subsets? (see paper, table 1)
+    * overall vs broken down by taxonomic groups? (see paper, figure 1)
+    * spatial distribution of gap species (see paper, figure 2)
 
@@
     * Note something about how photosynthetic marine species, and the species that depend on them, are far more sensitive to bathymetry than terrestrial species are to altitude... 
-    * Note that this is explicitly mentioned in the [training for IUCN red list marine maps](https://www.conservationtraining.org/mod/scorm/player.php?a=1412&currentorg=M5L1_ORG&scoid=6492) - "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species"
+    * Note that this is explicitly mentioned in the [training for IUCN red list marine maps](https://www.conservationtraining.org/mod/scorm/player.php?a=1412&currentorg=M5L1_ORG&scoid=6492) 
+        - "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species"
 
@@
     * get expert eyeballs on all the maps
+    * revisit the decision to use data-poor species... or at least make it more up-front about the ramifications of including the data-poor species
 
@@
         * consider setting AM presence threshold to 0% to create an extent of occurrence map, more like IUCN map.
-        * consider additional filters and criteria, e.g. "reviewed" and/or "occurcells" to avoid over-extrapolation
+        * consider additional filters and criteria, e.g. `reviewed` and/or `occurcells` to avoid over-extrapolation problem
         * consider clipping IUCN ranges to bathymetry for depth-limited species or taxa

### 2016-01-16 ms_3_methods.Rmd
Updating overall manuscript to get a draft to Ben

@@
 
-___from results/discussion:___ Overlaying paired distribution maps for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
+___from results/discussion: paraphrase or expand upon this___ Overlaying paired distribution maps for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
 
@@
 $$\alpha_{area} = \frac{A_{small}}{A_{large}} * 100\%$$
+
+### Methods for OHI comparison
+
+### Methods for MPA Gap Analysis comparison

### 2016-01-16 ms_1_abst_sig.Rmd
a few little tweaks to wording, and incorporated a few more of Julie's comments

@@
 
-To understand mechanistic reasons for differences in species range predictions, we compare the distribution and extent of the pair of predicted ranges for the subset of species included in both datasets.
-Categorizing the results based on how well each pair of species maps aligns, we identify several issues that commonly lead to misalignment in species range predictions.
+For the subset of species included in both datasets, we compare the distribution and extent of the pair of predicted ranges and identify several issues that commonly lead to misalignment between the two datasets.
 Notably, IUCN maps often disregard bathymetry for depth-limited species, leading to predictions of species presence at unsuitable depths, and AquaMaps maps for data-poor species often extrapolate presence far afield from known occurrences.

### 2016-01-16 ms_2_intro_res_disc.Rmd
a few little tweaks to wording, and incorporated a few more of Julie's comments

@@
 
+``` {r calc species numbers, echo = FALSE}
+n_spp_am    <- nrow(spp_am)
+n_spp_am_reviewed <- nrow(spp_am %>% filter(reviewed == 1))
+p_spp_am_reviewed <- round(n_spp_am_reviewed/n_spp_am * 100, 1)
+n_spp_am_rl <- nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))
+p_spp_am_rl <- round(n_spp_am_rl/n_spp_am * 100, 1)
+n_spp_iucn  <- nrow(spp_iucn_maps)
+n_spp_both  <- nrow(spp_map_pairs)
+p_both_am   <- round(n_spp_both/n_spp_am * 100, 1)
+p_both_iucn <- round(n_spp_both/n_spp_iucn * 100, 1)
+p_both_all  <- round(n_spp_both/nrow(spp_all) * 100, 1)
+```
+
 # Introduction
 
-_Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean - and you will find a unique ecosystem defined as much by which species are absent as by which species are present._ Knowing where individuals of a species exist, and where they thrive, is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
+_Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean - and you will find a unique ecosystem defined as much by which species are absent as by which species are abundant._ Knowing where individuals of a species thrive is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
 This knowledge provides foundational information for
@@
 
-To understand the implications of differences between these two datasets, we compared how each data set represents the global spatial and taxonomic distribution of species.
-For the relatively small number of species mapped in both datasets, we examined how well the species maps align and determined several issues that lead to misalignment between predicted species distributions.
+To understand the implications of differences between the AquaMaps and IUCN datasets, we compared how each data source represents the global spatial and taxonomic distribution of species.
+For the `r p_both_all`% of species mapped in both datasets, we examined how well the species maps align and determined several issues that lead to misalignment between predicted species distributions, and outlined possible improvements.
+
 We then subjected two recent marine biodiversity studies - the Ocean Health Index biodiversity goal (Halpern et al.
@@
 2015) - to a sensitivity analysis, substituting one dataset over the other, to highlight the possible consequences of different data use decisions on our understanding of _the status of_ marine biodiversity.
+___We found XXX...
+very briefly outline results?___
 
@@
 
-``` {r calc species numbers, echo = FALSE}
-n_spp_am    <- nrow(spp_am)
-n_spp_am_reviewed <- nrow(spp_am %>% filter(reviewed == 1))
-p_spp_am_reviewed <- round(n_spp_am_reviewed/n_spp_am * 100, 1)
-n_spp_am_rl <- nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))
-p_spp_am_rl <- round(n_spp_am_rl/n_spp_am * 100, 1)
-n_spp_iucn  <- nrow(spp_iucn_maps)
-n_spp_both  <- nrow(spp_map_pairs)
-p_both_am   <- round(n_spp_both/n_spp_am * 100, 1)
-p_both_iucn <- round(n_spp_both/n_spp_iucn * 100, 1)
-```
-
-The IUCN publishes species range maps as spatial vector polygon shapefiles, bundled by taxonomic groups.
-Using GIS, experts outline spatial polygons to represent a given species' extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
+The IUCN publishes species range maps as polygon shapefiles, bundled by taxonomic groups (REF).
+Using a geographic information system (GIS), experts outline spatial polygons to represent a given species' extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
 _IUCN releases a range map bundle for a taxonomic group once it has been "comprehensively assessed," i.e.
@@
 -----
+# Implications
+
+___From JSL:___ _I think you need a broader discussion of implications.
+How many times has IUCN/Aquamaps been cited?
+How long have they been around?
+What scales are the studies?
+Then you can muse about how results/interpretations would be different depending on which dataset they used, and reference Fig 3’s._
+
+_Then, say for example, the OHI uses these datasets together (don’t need to get too much into OHI background I think, maybe they don’t even need to know about the biodiversity goal?), but say that to try to combine them they use a threshold at 40% since others use this threshold (but check out the SOM for analyses about thresholds)_
 

### 2016-01-16 data_explore/mpa_analysis.Rmd
pushing to Neptune to work - too little space on my laptop! damn R can hog some memory

@@
 ``` {r set_up_base_raster}
+### base raster will contain 0.01 degree cells from -180 to +180,
+### and -90 to +90, with the associated area of each cell.
+### Save it as a .tif in GitHub - about 50 
+rast_base_file <- file.path(dir_git, 'data/mpa_analysis/rast_base.tif')
+
+if(!file.exists(rast_base_file)) {
   ### import AquaMaps half-degree loiczid raster - for extents and such
@@
 
-rast_base      <- raster(ext = extent(rast_loiczid), resolution = 0.01)
-rast_area_base <- raster::area(rast_base) ### takes a while...
+  rast_empty <- raster(ext = extent(rast_loiczid), resolution = 0.01)
+  rast_base  <- raster::area(rast_empty) ### takes a while...
+
+  writeRaster(rast_base, rast_base_file)
+} else {
+  rast_base <- raster(rast_base_file)
+}
 

### 2016-01-16 data_explore/mpa_analysis.Rmd
trying to get shapefiles saved and rasters burned... saving shapefile so we can use gdal_rasterize effectively

@@
 
-  writeRaster(rast_base, rast_base_file)
+  writeRaster(rast_base, rast_base_file, overwrite = TRUE)
 } else {
@@
 ``` {r set_up_spatially_marine_raster}
+### This raster (at 0.01 degree resolution) contains flags on whether a cell
+### is "spatially marine" - i.e. overlaps (center?) with an ocean area.
+### This is separate from the WDPA "marine" flag.
+
+rast_marine_file <- file.path(dir_git, 'data/mpa_analysis/rast_marine.tif')
+reload <- FALSE
+if(reload == TRUE) {
 ### import OHI regions polygons, to determine spatially-marine areas
+  poly_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_marine.shp')
+  if(!file.exists(poly_marine_file)) {
     dir_rgn   <- file.path(dir_N, 'git-annex/globalprep/spatial/v2015/data')
@@
 
-### filter out all polygons that are land or landlocked, and flag the rest as spatially-marine = 1
+    ### filter out all polygons that are land or landlocked, and flag 
+    ### the rest as spatially-marine = 1
     poly_marine <- poly_rgn[!str_detect(poly_rgn@data$rgn_typ, 'land'), ]
     poly_marine@data <- poly_marine@data %>%
-  mutate(sp_mar_flag = 1)
+      mutate(sp_mar = 1)
 
-# writePolyShape(poly_marine, file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_sp_mar'))
+    # writePolyShape(poly_marine, poly_marine_file)
+    writeOGR(obj    = poly_marine,
+             dsn    = path.expand(dirname(poly_marine_file)),
+             layer  = str_replace(basename(poly_marine_file), '.shp', ''),
+             driver = 'ESRI Shapefile',
+             overwrite_layer = TRUE,
+             verbose = TRUE)
 
-# rast_sp_mar <- rasterize(poly_marine, rast_base, value = sp_mar_flag) ### pretty slow... consider gdal_rasterize?
+    } else {
+    poly_marine <- readShapePoly(poly_marine_file, 
+                               proj4string = crs_wgs84)
+  }
 
-# library(gdalUtils)
-# rast_sp_mar <- gdal_rasterize(
-#     src_datasource = file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_sp_mar.shp')),
-#     dst_filename = file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_sp_mar.tif'), 
-#       # destination for output
-#     a = 'sp_mar_flag', 
-#       # the attribute in the shapefile to be assigned to the cell values
-#     te = c(-180, -90, 180, 90), 
-#       # extents for output raster
-#     tr = c(.01, .01),  
-#       # resolution for output raster
-#     tap = TRUE, 
-#       # target aligned pixels - align coords of extent of output to values of -tr
-#     a_nodata = NA, 
-#       # nodata value for raster; otherwise they will be filled in with zeroes
-#     output_Raster = TRUE, 
-#       # return output as a RasterBrick? 
-#     verbose = TRUE)
+  # rast_marine <- rasterize(poly_marine, rast_base, value = sp_mar_flag) 
+  ### excruciatingly slow (interminable?)... consider gdal_rasterize?
+
+  library(gdalUtils)
+  rast_marine <- gdal_rasterize(
+      src_datasource = poly_marine_file),
+      dst_filename   = rast_marine_file), 
+        # destination for output
+      a = 'sp_mar_flag', 
+        # the attribute in the shapefile to be assigned to the cell values
+      te = c(-180, -90, 180, 90), 
+        # extents for output raster
+      tr = c(.01, .01),  
+        # resolution for output raster
+      tap = TRUE, 
+        # target aligned pixels - align coords of extent of output to values of -tr
+      a_nodata = NA, 
+        # nodata value for raster; otherwise they will be filled in with zeroes
+      output_Raster = TRUE, 
+        # return output as a RasterBrick? 
+      verbose = TRUE)
+}
 

### 2016-01-17 data_explore/mpa_analysis.Rmd
updating so marine raster goes to git-annex instead of github - file too large...

@@
 
-rast_marine_file <- file.path(dir_git, 'data/mpa_analysis/rast_marine.tif')
+rast_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters/rast_marine.tif')
 reload <- FALSE
@@
 ### import OHI regions polygons, to determine spatially-marine areas
-  poly_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/tmp_rasters/poly_marine.shp')
+  poly_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters/poly_marine.shp')
   if(!file.exists(poly_marine_file)) {
@@
         # destination for output
-      a = 'sp_mar_flag', 
+      a = 'sp_mar', 
         # the attribute in the shapefile to be assigned to the cell values

### 2016-01-17 data_explore/mpa_analysis.Rmd
still trying to get these damn rasters to work... argh!

@@
 
+  plot(poly_marine, col = 'blue')
+
   # rast_marine <- rasterize(poly_marine, rast_base, value = sp_mar_flag) 
@@
   rast_marine <- gdal_rasterize(
-      src_datasource = poly_marine_file),
-      dst_filename   = rast_marine_file), 
+      src_datasource = poly_marine_file,
+      dst_filename   = rast_marine_file, 
         # destination for output
-      a = 'sp_mar', 
+      a = 'rgn_id', 
+#      a = 'sp_mar', 
         # the attribute in the shapefile to be assigned to the cell values

### 2016-01-19 ms_1_abst_sig.Rmd
incorporating comments from BSH

--- 2016-01-15 ms_1_abst_sig.Rmd
@@
 
-_questions for abstract style: voice (passive vs active, 'we' vs 'this paper'), tense (present vs past)..._
-
-Two global datasets of species distribution maps - AquaMaps and the International Union for Conservation of Nature - dominate our understanding of marine species ranges throughout the world's oceans, and inform a wide range of biodiversity and conservation purposes.
-Differences in methodology and intent between these two datasets drive differences in predicted species ranges, with significant implications for management and conservation decisions.
-
-<!--
-The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
-
-Importantly, biases in taxonomic or spatial coverage within a dataset could shift management or conservation actions away from places or species that are most in need.(maybe Jetz 2008?)  Inaccurate indications of presence or absence could lead to inefficient or ineffective marine reserve systems and management plans.
-(Rondinini et al.
-2006) (Jetz 2008?)
--->
-Together, the two datasets provide range information for XXX species, with only `r nrow(spp_map_pairs)` species common to both datasets.
-___what is the key takeaway for the taxonomic and spatial coverage analyses?
-can we say something about benefits of combining the two for better taxo coverage?___
-
-For the subset of species included in both datasets, we examine the alignment of range maps by comparing distribution and extent of predicted ranges.
-Categorizing the results based on quality of alignment, we identify several issues that commonly lead to disagreement in species range predictions.
-Notably, IUCN maps often disregard bathymetry for depth-limited species, leading to predictions of species presence at unsuitable depths, and AquaMaps maps for data-poor species often extrapolate presence far afield from known occurrences.
-___something about recommendations here?
-how this understanding can lead to better use of these datasets?___
-
-To understand the implications of these differences, we reexamined two recent global analyses that depend on these datasets: the Ocean Health Index and a global analysis of gaps in coverage of marine protected areas.
-Variations in which dataset is used, and how, lead to significantly different estimates of species diversity and conservation effectiveness.
+Species distribution data provide the foundation for a wide range of ecological research and conservation science and management, yet most species ranges remain unknown, and existing range maps have little overlap.
+In the ocean, two global species distribution datasets - AquaMaps and the International Union for Conservation of Nature (IUCN) - dominate our understanding of marine species ranges throughout the world's oceans, representing `r nrow(spp_all)` species overall, mostly from AquaMaps, with only `r nrow(spp_map_pairs)` species overlapping.
+Here we examine differences in predicted species ranges between the two datasets, propose mechanistic causes and potential solutions for such differences, and explore the implications of these differences for management and conservation decisions.
+We find that IUCN maps often disregard bathymetry for depth-limited species, leading to predictions of species presence at unsuitable depths, and AquaMaps ranges for data-poor species often extrapolate presence far afield from known occurrences.
+We illustrate the implications of these differences by repeating two recent applications - the Ocean Health Index biodiversity goal and a global analysis of gaps in coverage of marine protected areas - and find significantly different estimates of the status of species diversity and effectiveness of conservation depending on which dataset was used.
+Understanding these issues points toward solutions for providers and users of these datasets.
 

### 2016-01-22 data_explore/data_explore.Rmd
update manuscript and figs based on Ben's comments

@@
        y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
-       title = 'Distribution alignment vs Extent alignment',
+#       title = 'Distribution alignment vs Extent alignment',
        color = 'Risk', size = '% ocean range')
@@
 
-spp_list <- spp_list %>% mutate(data_poor = ifelse(occurcells < 10, '< 10', '>= 10')
+spp_list <- spp_list %>% mutate(data_poor = ifelse(occurcells < 10, '< 10', '>= 10'))
 ### mongo plot time
@@
        y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
-       title = 'Distribution alignment vs Extent alignment',
+#       title = 'Distribution alignment vs Extent alignment',
        color = 'AM model cells', size = '% ocean range')
@@
        y = 'Relative proportion of species per quadrant', 
-       title = 'Taxonomic group by quadrant',
+#       title = 'Taxonomic group by quadrant',
        fill = 'Alignment')
@@
        y = 'Relative proportion of species per quadrant', 
-       title = 'Risk category by quadrant',
+#       title = 'Risk category by quadrant',
        fill = 'Alignment')
@@
   scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) + 
-  labs(x = 'Class', y = 'Percent of mapped species',
-       title = 'Species maps by taxonomic class')
+  labs(# title = 'Species maps by taxonomic class',
+       x = 'Class', 
+       y = 'Percent of mapped species')
 print(bar_taxa2)
@@
   scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
-  labs(x = 'Class', y = 'Percent of assessed mapped species',
-       title = 'Species maps by taxonomic class, Red List-assessed only')
+  labs(# title = 'Species maps by taxonomic class, Red List-assessed only',
+       x = 'Class', 
+       y = 'Percent of assessed mapped species')
 print(bar_taxa3)

### 2016-01-22 manuscript-main.Rmd
update manuscript and figs based on Ben's comments

--- 2016-01-14 ms_0_main.Rmd
@@
 title: "Aligning species range data to better serve science and conservation"
-author: "Casey O'Hara, Jamie Afflerbach, Ben Halpern"
+author: "Casey O'Hara, Jamie Afflerbach, Benjamin S. Halpern"
 output:
@@
 # library(readr)
+
 # library(raster)
@@
 ```
+
+-----
+
+``` {r SOM, child = 'ms_6_supplemental.Rmd'}
+```

### 2016-01-22 ms_1_abst_sig.Rmd
update manuscript and figs based on Ben's comments

@@
 Species distribution data provide the foundation for a wide range of ecological research and conservation science and management, yet most species ranges remain unknown, and existing range maps have little overlap.
-In the ocean, two global species distribution datasets - AquaMaps and the International Union for Conservation of Nature (IUCN) - dominate our understanding of marine species ranges throughout the world's oceans, representing `r nrow(spp_all)` species overall, mostly from AquaMaps, with only `r nrow(spp_map_pairs)` species overlapping.
+In the ocean, two global species distribution datasets - AquaMaps and the International Union for Conservation of Nature (IUCN) - dominate our understanding of marine species ranges throughout the world's oceans, representing `r format(nrow(spp_all), big.mark = ',')` species overall, mostly from AquaMaps, with only `r format(nrow(spp_map_pairs), big.mark = ',')` species overlapping.
 Here we examine differences in predicted species ranges between the two datasets, propose mechanistic causes and potential solutions for such differences, and explore the implications of these differences for management and conservation decisions.

### 2016-01-22 ms_2_intro_res_disc.Rmd
update manuscript and figs based on Ben's comments

@@
 ``` {r calc species numbers, echo = FALSE}
+n_spp_all    <- nrow(spp_all)
 n_spp_am     <- nrow(spp_am)
-n_spp_am_reviewed <- nrow(spp_am %>% filter(reviewed == 1))
-p_spp_am_reviewed <- round(n_spp_am_reviewed/n_spp_am * 100, 1)
-n_spp_am_rl <- nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))
-p_spp_am_rl <- round(n_spp_am_rl/n_spp_am * 100, 1)
+n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
 n_spp_iucn   <- nrow(spp_iucn_maps)
 n_spp_both   <- nrow(spp_map_pairs)
+
+p_spp_am     <- round(nrow(spp_am)/n_spp_all * 100, 1)
+p_spp_am_rev <- round(n_spp_am_rev/n_spp_am * 100, 1)
 p_both_am    <- round(n_spp_both/n_spp_am   * 100, 1)
 p_both_iucn  <- round(n_spp_both/n_spp_iucn * 100, 1)
-p_both_all  <- round(n_spp_both/nrow(spp_all) * 100, 1)
+p_both_all   <- round(n_spp_both/n_spp_all  * 100, 1)
 ```
@@
 One major outcome of this body of science is the various compiled databases of species distribution maps.
-Two global-scale repositories _predict_ marine species ranges throughout the world's oceans – AquaMaps modeled species distribution maps (Kaschner et al.
+Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps modeled species distribution maps (Kaschner et al.
 2013) and International Union for Conservation of Nature (IUCN) species range maps (REF).
@@
 Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
-Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans.(Rondinini et al.
-2006) (Jetz 2008?)
+Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans (Rondinini et al.
+2006) (Jetz 2008?).
 
 To understand the implications of differences between the AquaMaps and IUCN datasets, we compared how each data source represents the global spatial and taxonomic distribution of species.
-For the `r p_both_all`% of species mapped in both datasets, we examined how well the species maps align and determined several issues that lead to misalignment between predicted species distributions, and outlined possible improvements.
+Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')`; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` species).
+For the `r format(n_spp_both, big.mark = ',')` (`r p_both_all`% of total) species mapped in both datasets, we examined how well the maps align and determined several issues that lead to misalignment between predicted species distributions, and outline possible improvements.
 
-We then subjected two recent marine biodiversity studies - the Ocean Health Index biodiversity goal (Halpern et al.
-2012) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (Klein et al.
-2015) - to a sensitivity analysis, substituting one dataset over the other, to highlight the possible consequences of different data use decisions on our understanding of _the status of_ marine biodiversity.
-___We found XXX...
-very briefly outline results?___
+We then subjected two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (Halpern et al.
+2012, 2015) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (Klein et al.
+2015) - to a sensitivity analysis.
+...
+___substituting___ one dataset over the other, to highlight the possible consequences of different data use decisions on our understanding of _the status of_ marine biodiversity.
+___BSH suggestion: We used all unique species range data from both sources but substituted one dataset over the other for shared species to highlight the possible consequences of different data use decisions on our understanding of the status of and protection of marine biodiversity.___
 
 -----
+<!--
+## Overview of AquaMaps and IUCN datasets
 
-## Overview of AquaMaps and IUCN datasets
+___BSH comment: Fold this section into the methods and/or Supplementary Methods, or maybe into the results section just below on differences between the datasets – but make the description VERY brief.___
 
@@
 The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species.
-A subset of the resulting distribution maps (`r n_spp_am_reviewed` of `r n_spp_am` species maps; `r p_spp_am_reviewed`%) have been checked or formally reviewed by experts, to validate and improve the parameters and underlying model.
+A subset of the resulting distribution maps (`r format(n_spp_am_rev, big.mark = ',')` of `r format(n_spp_am, big.mark = ',')` species maps; `r p_spp_am_rev`%) have been checked or formally reviewed by experts, to validate and improve the parameters and underlying model.
 
@@
 Due to these differences in methodology and intent, IUCN range maps are more likely to overpredict species presence (and underpredict absence) while AquaMaps modeled distribution maps are relatively more likely to underpredict species presence.(REF and REF)
-
------
+-->
 
@@
 
-## Comparing the two datasets
+## How and why the datasets differ
 
-### Taxonomic distribution between datasets
+The IUCN publishes species range maps as polygon shapefiles, bundled by taxonomic groups (REF).
+Using a geographic information system (GIS), experts outline spatial polygons to represent a given species' "limits of distribution" (IUCN 2014) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
+AquaMaps models species distribution maps based environmental preferences (e.g.
+temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species.
 
-___I wonder if this section is becoming less a part of the "story", maybe this part should go in SOM?___ _- though I think it's pretty important to understand that there is a huge difference in which species and taxonomic groups are covered in each set.
-That also helps to explain why there is such a small overlap, and helps show that the overlapping species are not necessarily representative?
-It also provides a logical lead-in from description of the overall datasets to the analysis of just the overlapping portion of the datasets._
-
-Each dataset offers spatial distribution information for large numbers of species (`r n_spp_am` AquaMaps-mapped species; `r n_spp_iucn` IUCN-mapped marine species).
-However, the datasets vary in terms of taxonomic coverage and regional coverage, with only `r n_spp_both` species included in both datasets (`r p_both_am`% of AquaMaps species; `r p_both_iucn`% of IUCN marine species; <a href = '#fig1'>Figure 1a</a>).
-
-The distribution of IUCN-mapped species skews toward tropical latitudes and away from the Atlantic and Eastern Pacific compared to the distribution of AquaMaps-mapped species.(<a href = '#fig1'>Figure 1b and 1c</a>)  This skew likely reflects the fact that the IUCN dataset focuses more heavily on coral reef-associated taxa than does the AquaMaps dataset (see <a href = '#fig1'>Figure 1a</a>).
-
-For spatial assessments of biodiversity, the choice of one dataset over the other is likely to result in significantly different conclusions.
-For studies confined to a narrow range of taxa or to a narrow spatial scale, one dataset may offer an advantage over the other in the number of mapped species available.
-For global scale biodiversity studies, however, the selection of one dataset over the other will entail tradeoffs in spatial coverage and taxonomic breadth versus depth._use something from Rondinini here?_
-
-_Small overlap means that using both datasets in conjunction will add a huge number of species without duplicating efforts, if the differences between the two methods can be reconciled appropriately_
-
-_find a reference that describes what makes a "good" dataset for global biodiversity, e.g. OHI, or maybe species richness vs diversity vs "health" or whatnot:  (Tittensor et al., 2010)_
+The two datasets have notably different taxonomic (Fig.
+1A) and regional (Fig.
+1B,C) coverage.
+IUCN-mapped species focus on tropical latitudes and away from the Atlantic and Eastern Pacific compared to AquaMaps-mapped species.
+These differences can be mitigated by using both datasets, but the underlying methodological differences complicate such direct comparisons, and furthermore ___requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species.___
 
 -----
-
-### Defining spatial alignment between the two datasets
-
-The IUCN and AquaMaps spatial datasets share `r nrow(spp_map_pairs)` species in common.
-Placing each of these species' pair of maps side by side, we would hope to see spatial correlation both in the global pattern of species distribution (where on the map) and the extent of species range (how much of the map).
-
-<!-- NOT REALLY PART OF THE STORY? INCLUDE IN SOM?
-In `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of cases, the IUCN range map indicated a larger species range than the AquaMaps map.
-This finding concurs with the general expectation that geographic range maps such as IUCN are more likely to over-predict presence and under-predict absence than distribution models such as AquaMaps.(Rondinini et al., 2006) _errors of commission vs errors of omission, essentially type I and type II errors - should I use one of these terms to clarify?_
--->
-
-Dividing the map-paired species into quadrants based on median values for each dimension allows us to examine the _implications_ of four different qualities of alignment.
-(<a href = '#fig2'>Figure 2a</a>)
 
@@
 
-Quadrant 1 highlights species whose map pairs agree in both spatial distribution and the extent of described ranges, as we would expect for well-understood species.
-Examining spatial alignment by taxonomic group (<a href = '#fig2'>Figure 2b</a>), we found that certain taxa were far more likely than others to be spatially well-aligned; in particular, wide-ranging pelagic organisms (50% or more of ocean area) such as marine mammals, tunas, and billfishes were more consistently well-aligned than demersal and reef organisms.
-_propose an explanation here?
-basically - these species go pretty much anywhere, so both maps predict most of the ocean, you'll get a good overlap - not really that revolutionary_
+___BSH suggestion___ To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (where on the map) against the ratio of extent (how much of the map) for each shared species (Fig.
+2a).
+__Overall we see a negative linear relationship, where increasing similarity in extent (extent ratio approaches 100%) correlates with a decrease in alignment.
+[explain/discuss].__
 
-The extent-misaligned maps contained in quadrant 2 provide a richer set of examples to understand the fundamental differences between the datasets; examining these, we can identify several mechanisms that could cause such extent misalignment.
+Dividing the map-paired species into quadrants highlights different categories of relationships.
+The upper right quadrant includes species (n = `r n_q1`) whose map pairs agree in both spatial distribution and extent of described ranges.
+These species tend to be well-studied and predominantly fall within wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig.
+2B).
+This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because they exist nearly everywhere.
 
-For `r n_q2_am_sm` of `r n_q2` species in quadrant 2 (`r round((n_q2_am_sm/n_q2) * 100, 1)`%), the IUCN extent is larger than the AquaMaps extent.
-For many of these species, a quick look at the maps (see SOM for examples) shows that the IUCN range hews closely to the AquaMaps range, while adding a significant buffer zone.
-Coral species dominate this quadrant (`r n_q2_corals` coral species - `r round(n_q2_corals/n_corals * 100, 1)`% of all corals and `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in quadrant 2).
-Most corals and reef-associated organisms prefer shallower waters, and we suspect that many of these extent-misaligned map pairs can be easily explained: seafloor depth is explicitly included in AquaMaps models, while depth is frequently ignored as a factor in IUCN range maps, leading to overprediction of species range into areas too deep to support the species.
+The extent-mismatched maps contained in the upper left quadrant (n = `r n_q2`) provide a richer set of examples to understand the fundamental differences between the datasets.
+For many of these species, while the spatial distribution is similar, the IUCN range predicts a significant buffer zone beyond the AquaMaps range.
+Corals dominate this quadrant (`r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant).
+Most corals and reef-associated organisms prefer shallower waters, and we suspect that many of these disagreements in extent can be explained by bathymetry: seafloor depth is explicitly included in AquaMaps models, while depth is frequently ignored as a factor in IUCN range maps, leading to overprediction of species range into areas too deep to support the species.
 (For examples and details, see SOM)
 
-* ___verify some of the coral maps by saving raster, showing raster and polygons in QGIS against a bathymetry layer___ - put a map for Oculina varicosa (and others?) in the SOM
-* ___include recommendation for IUCN range maps to be clipped to bathymetry where appropriate? or save til end?___
+The differences in extent of species maps found in the lower right quadrant (n = `r n_q3`) can in many cases be attributed to "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence.
+This results in an area ratio close to 100%, but similar extents are meaningless when the distributions are poorly aligned.
 
-The extent-alignment of species maps found in quadrant 3 can in many cases be attributed to "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable habitat, as in the case of many quadrant 2 species, while at the same time, AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence.
-This results in an area ratio close to 100%, giving the impression of extent alignment.
-___this quadrant seems to be a false alignment then?
-should there be a stronger warning here___
+Moving from the lower right to lower left quadrant, AquaMaps more frequently extrapolates species presence well beyond known occurrences, resulting in poor matches in both distribution and extent.
+Of the `r n_q4` species in this quadrant, 159 (39.7%) are data-poor (in which the AquaMaps environmental envelope model is based on fewer than 10 known cell occurrences), compared with 21.1% for the full set of map-paired species and only 8.2% of species in the upper left quadrant.
+A mere 6.2% of AquaMaps maps in this quadrant have undergone expert review, compared with 13.4% of all map-paired species and 24.9% of species in the upper left quadrant.
 
-Quadrant 4 species maps suffer from the same issues as those in quadrant 3, but here the AquaMaps extrapolation extends even further, pushing the extent alignment ratio lower.
-In this quadrant, AquaMaps predicts a larger extent than IUCN for 70% of species represented in this quadrant, compared with just 30% of species in the entire set of overlapping maps _(this is supposed to show that much of this quadrant is based on overextrapolation of AquaMaps...
-not sure that's clear yet)_.
-AquaMaps range maps in quadrant 4 include a high proportion of data-poor species in which the environmental envelope model is based on fewer than 10 known cell occurrences, and at the same time include fewer reviewed species maps than the other quadrants.
-
-___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot___ _- perhaps update the quadrant plot to reveal data-poor species (and reviewed species?
-little overlap of reviewed & data-poor) instead of, say, red-list category which doesn't get discussed in the body of the paper?
-then this table can go in SOM if we like it_
-
-| quadrant | n species | n data poor species | n reviewed species |
-| :------: | :-------: | :-----------------: | :----------------: |
-| overall  |   22889   |        8749 (38.2%) |       1296  (5.7%) |
-| AM&IUCN  |    2166   |         457 (21.1%) |        290 (13.4%) |
-|   q1     |     401   |          33  (8.2%) |        100 (24.9%) |
-|   q2     |     682   |         151 (22.1%) |        100 (14.7%) |
-|   q3     |     682   |         114 (16.5%) |         65  (9.5%) |
-|   q4     |     410   |         159 (39.7%) |         25  (6.2%) |
-
-_data poor = fewer than 10 points used to generate the environmental envelopes; reviewed = map checked or formally expert-reviewed_
-
-_Other hypotheses/case studies to highlight other possible mechanistic differences?_  I think we found a couple of the big ones; good enough?
+This breakdown by quadrant helps explain the negative linear relationship seen in the plot, in which increasing similarity in extent correlates with decreasing distribution alignment.
+Expanding species range beyond known occurrences, the marginal range predicted by the AquaMaps model will fall in different locations than the marginal range predicted using IUCN methodology.
+For species with dissimilar extents, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
+As we examine species with increasingly similar extents, the differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.
 
 -----
-# Implications
+
+# Implications of the differences
 
@@
 The global Ocean Health Index (OHI) (Halpern et al.
-2012), a composite index made up of 10 goals, utilizes both of these datasets to inform the Species subgoal of the Biodiversity goal.
-As it is currently calculated, the Species subgoal uses species spatial distribution data and IUCN Red List conservation status to calculate an area-weighted mean species status in each of 221 exclusive economic zones.
-Spatial distributions were gleaned from both IUCN and AquaMaps datasets, preferring IUCN data for species represented in both data sets.
+2012, 2015), a composite index made up of 10 goals, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status for each of 221 exclusive economic zones.
+To maximize the number of represented species, spatial distributions were gleaned from both IUCN and AquaMaps datasets, preferring IUCN data for species in both sources.
 OHI uses a probability threshold of 40% to determine species presence for AquaMaps data.
@@
 
-## Wrap it up
+# Conclusions: What can be done to address the differences
 
-___Recommendations on data use?___
+_While the differences between species distributions described by AquaMaps and IUCN can drive different conclusions for conservation research, some of the mechanistic causes can be addressed._ ___oh god this section is so passive voice...
+arrrrgghhhh___
 
-* IUCN mapmakers
-    * at least clip polygons to bathymetry for depth-sensitive species, come on now people, it's really pretty obvious.  
-    * Note something about how photosynthetic marine species, and the species that depend on them, are far more sensitive to bathymetry than terrestrial species are to altitude... 
-    * Note that this is explicitly mentioned in the [training for IUCN red list marine maps](https://www.conservationtraining.org/mod/scorm/player.php?a=1412&currentorg=M5L1_ORG&scoid=6492) 
-        - "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species"
+__For data providers:__
 
-* AquaMaps mapmakers
-    * get expert eyeballs on all the maps
-    * revisit the decision to use data-poor species... or at least make it more up-front about the ramifications of including the data-poor species
+* The training manual for experts creating IUCN Red List maps for marine species suggests that "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species" (REF https://www.conservationtraining.org/mod/scorm/player.php?a=1412&currentorg=M5L1_ORG&scoid=6492).
+Unfortunately, this recommendation appears to be applied haphazardly.
+Especially for demersal communities dependent on photosynthesis, bathymetry should be a primary consideration; even a low resolution cutoff of 200 meters would improve the utility of species range maps.
 
-* Users
-    * understand that range maps tend to overestimate presence; AquaMaps are less likely to overestimate presence, but trade off for overestimating absence.
+* The AquaMaps model is a powerful tool for quickly estimating species presence based on limited information; however, the model's output is only as good as the input data, which for marine species is often sorely limited.
+AquaMaps encourages experts to review and comment on the predicted species distribution maps; lessons from the small sample of reviewed maps can be generalized, for example to other species in the same taxonomic group, to further refine the predictions, especially for data-poor species.
+
+* In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.
+Currently the area restrictions are denoted by FAO Major Fishing Areas (REF find a reference in AquaMaps somewhere); however, allowing area restrictions based on biogeographical criteria, a la MEOW, would likely have better resolution and predictive power, especially for data-poor species.
+
+__For data users:__
+
+* understand that range maps a la IUCN tend to overestimate presence; AquaMaps are less likely to overestimate presence, but trade off for overestimating absence.
 Both better than simple point locality data.
-    * to use both together:
-        * consider setting AM presence threshold to 0% to create an extent of occurrence map, more like IUCN map.
-        * consider additional filters and criteria, e.g. `reviewed` and/or `occurcells` to avoid over-extrapolation problem
-        * consider clipping IUCN ranges to bathymetry for depth-limited species or taxa
-        * or consider de-weighting IUCN for area-weighted calcs, for taxa in which IUCN is consistently overestimating presence
+(Rondinini etc)
+* Using both datasets together can increase the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized.  Consider:
+    * An AquaMaps presence threshold of 0% more closely approximates the "limits of distribution" outlined by IUCN range maps.
+    * Additional filters and criteria on AquaMaps distribution maps can help avoid overextrapolation for data poor species.
+The AquaMaps native range dataset includes information on expert review and the number of occurrence cells used to generate the environmental envelope.
+    * For depth-limited species or taxa, clipping IUCN range polygons to a reasonable bathymetry contour can limit overprediction of species presence into unsuitable habitats.

### 2016-01-22 ms_4_figs.Rmd
Updating overall manuscript to get a draft to Ben
update manuscript and figs based on Ben's comments

@@
 
-> __Figure 2 (a):__ For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range to the larger range (how much of the map).
+> __Figure 2 (a):__ For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller predicted range intersecting the larger predicted range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range area to the larger range area (how much of the map).
 For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.
@@
 (n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
-> The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges lie - a more worrisome mismatch than that indicated by quadrant 2
+> The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges occur
 (n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
@@
 > __Figure 2 (b):__ Spatial alignment of paired-map species by taxonomic group.
+___REDO with all corals grouped together!___  _Also - is it distracting to use order names for some groups and not others?
+Some of the groups had pretty long names ("acanthuridae" is way shorter than "surgeonfish, tangs, and unicornfish")_
 

### 2016-01-22 ms_6_supplemental.Rmd
Updating overall manuscript to get a draft to Ben
update manuscript and figs based on Ben's comments

--- 2016-01-14 ms_6_supplemental.Rmd
@@
 
-* _All IUCN-mapped species are also included in Red List species, but only `r p_spp_am_rl`% of AquaMaps species._
-
 ![](figures/barchart_spp_risk_quads.png)
@@
 
+___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot___ _- perhaps update the quadrant plot to reveal data-poor species (and reviewed species?
+little overlap of reviewed & data-poor) instead of, say, red-list category which doesn't get discussed in the body of the paper?
+then this table can go in SOM if we like it_
+
+| quadrant | n species | n data poor species | mean (median) data points | n reviewed species |
+| :------: | :-------: | :-----------------: | :-----------------------: | :----------------: |
+| all AM   |   22889   |        8749 (38.2%) |          57.1 (16)        |       1296  (5.7%) |
+| AM&IUCN  |    2166   |         457 (21.1%) |          89.9 (33)        |        290 (13.4%) |
+|   q1     |     401   |          33  (8.2%) |         233.0 (78)        |        100 (24.9%) |
+|   q2     |     682   |         151 (22.1%) |          77.4 (39)        |        100 (14.7%) |
+|   q3     |     682   |         114 (16.5%) |          52.4 (29)        |         65  (9.5%) |
+|   q4     |     410   |         159 (39.7%) |          32.1 (13)        |         25  (6.2%) |
+
 ## illustrative maps for different quadrants and different mechanistic problems
+
+Oculina varicosa - Ivory tree coral ![oculina_pic.jpg](figures/oculina/oculina_pic.jpg)
+
+* from IUCN: Colonies are found to depths of 152 m depth on limestone rubble, low-relief limestone outcrops, high-relief, steeply sloping prominences, and soft-bottom sloping habitats.
+Colonies are semi-isolated, patchy and low-growing in shallow water, or they form larger, massive coalescing aggregates (thickets or coppices) with substantial topographic relief in 50-100 m depth.
+In shallow waters (2-30m) the form is zooxanthellate, inhabiting limestone ledges.
+In deeper waters, an azooxanthellate form is known from the shelf edge off eastern Florida, USA from Ft.
+Pierce to Daytona (Reed 1980, 1983, 2002; Brooke and Young 2003).
+
+plotted against a 200 m bathymetry line: ![oculina_map.png](figures/oculina/oculina_map.png)
+
+<!--
 
@@
 
+-->
+
 ## AquaMaps: Effect of changing "presence" threshold on apparent distribution
@@
 This pattern may not hold true for all subsets of AquaMaps species, however, whether subsetting by taxa or by georegion.
-
------

### 2016-01-22 data_explore/data_explore_am_thresh.Rmd
updated figures without titles; boxplot = violin plot

@@
   scale_y_continuous(expand = c(0, 0)) +
-  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
+  labs(x = 'Proportion of extent included when selecting a threshold of 0.4', 
        y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3)))
@@
   scale_y_continuous(expand = c(0, 0)) +
-  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
+  labs(x = 'Proportion of extent included when selecting a threshold of 0.4', 
        y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3a)))
@@
   scale_y_continuous(expand = c(0, 0)) +
-  labs(title = 'Mean (median) species range after presence threshold cutoff',
+  labs(# title = 'Mean (median) species range after presence threshold cutoff',
        x = 'Presence threshold cutoff', 
@@
   scale_y_continuous(expand = c(0, 0)) +
-  labs(title = 'Relative species range after presence threshold',
+  labs(# title = 'Relative species range after presence threshold',
        x = 'Presence threshold cutoff', 

### 2016-01-22 data_explore/data_explore_global_maps.Rmd
updated figures without titles; boxplot = violin plot

@@
 library(rasterVis)
+
+cat('First off: map-paired species.  Map 1: AM dataset; map 2: IUCN dataset; maps 3 & 4: log-transformed AM & IUCN\n')
 
@@
 levelplot(am_sp_map, par.settings = myTheme, 
-          main = 'Aquamaps Dataset', 
+#          main = 'Aquamaps Dataset', 
           colorkey = list( at = seq(0, 1063, length = 15)))
@@
 levelplot(iucn_sp_map, par.settings = myTheme, 
-          main = 'IUCN Dataset', 
+#          main = 'IUCN Dataset', 
           colorkey = list(at = seq(0, 1063, length = 15)))
@@
 levelplot(am_ln_map, par.settings = myTheme, 
-          main = 'Aquamaps Dataset', 
+#          main = 'Aquamaps Dataset', 
           colorkey = list( at = seq(0, maxValue(am_ln_map), length = 15)))
@@
 levelplot(iucn_ln_map, par.settings = myTheme, 
-          main = 'IUCN Dataset', 
+#          main = 'IUCN Dataset', 
           colorkey = list(at = seq(0, maxValue(iucn_ln_map), length = 15)))
@@
 ```{r all_aquamaps, eval = FALSE, echo = FALSE}
+
+cat('Full AM dataset, species richness\n')
 
@@
 
-am_levelplot1 <- levelplot(all_am_map, main = 'Entire Aquamaps Dataset') # add in par.settings = myTheme
+am_levelplot1 <- levelplot(all_am_map, 
+#                           main = 'Entire Aquamaps Dataset',
+                           par.settings = myTheme)
 
@@
 ```{r all_iucn, eval = FALSE, echo = FALSE}
+
+cat('Full IUCN dataset, species richness\n')
 
@@
 
-iucn_levelplot1 <- levelplot(all_iucn_map, main = 'Entire IUCN Dataset') # add in par.settings = myTheme
+iucn_levelplot1 <- levelplot(all_iucn_map, 
+                             #main = 'Entire IUCN Dataset',
+                             par.settings = myTheme)
 
@@
           margin = list(axis = TRUE),
-          main = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', 
+#          main = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', 
           colorkey = list(at = seq(-1, 1, length = 15)))

### 2016-01-22 data_explore/data_explore_pt3.Rmd
updated figures without titles; boxplot = violin plot

@@
 
-```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+```{r plot_spp_from_q1, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 Plotting species from q3 (area-matched)
-```{r plot_spp_from_q3, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+```{r plot_spp_from_q3, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 Plotting species from q4 (poorly matched)
-```{r plot_spp_from_q4, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+```{r plot_spp_from_q4, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 

### 2016-01-22 data_explore/mpa_analysis.Rmd
updated figures without titles; boxplot = violin plot

@@
 
+Dealing with rasters:
+
+* how to ensure 1 km or better resolution?  WGS84 at 0.01 degree or Mollweide at 1 km
+    * WGS84 --> 36000 * 18000 cells = 648,000,000 cells...
+    * Mollweide --> 
+* options for creating rasters
+    * use arcpy as in LSP... global Mollweide raster of WDPA is about 50 MB; includes .tif and supporting files.  This is .5 km raster.  How is it so small?
+    * use gdal_rasterize seems to create multi-GB rasters; .tif only
+* why rasters?  avoid double-counting for overlapping parks.  WDPA is most important, but since that is in raster and AM is raster, easiest to keep it all in raster
+    * WDPA - capture park classification and year and marine flag
+    * spatial marine - capture whether cell includes ocean or not - clip to max N and S of WDPA?
+        * xMin,yMin -180.007,-72.95 : xMax,yMax 180,83.73
+* Use this R script to source arcPy script modified from LSP - need to run it on a PC with appropriate GDAL installation, most likely
+
 ``` {r set_up_spatially_marine_raster}

### 2016-01-22 ms_4_figs.Rmd
updated figures without titles; boxplot = violin plot

@@
 # Figures
+
+___consider using HCL or HSV values for colors on plots?___ or ask Mel re: good color choices.
+Make sure that colors have good contrast to show up on a black-and-white print...
+Should the quad plot background be colored?
+How else to quickly code the four quadrants relative to the quadrant-breakdown-by-taxa plot?
+
+**`hcl {grDevices}`**: Create a vector of colors from vectors specifying hue, chroma and luminance.
+
+*Usage* `hcl(h = 0, c = 35, l = 85, alpha, fixup = TRUE)`
+Arguments
+
+*arguments*
+- `h`: The hue of the color specified as an angle in the range [0,360]. 0 yields red, 120 yields green 240 yields blue, etc.
+- `c`: The chroma of the color. The upper bound for chroma depends on hue and luminance.
+- `l`: A value in the range [0,100] giving the luminance of the colour. For a given combination of hue and chroma, only a subset of this range is possible.
+- `alpha`: numeric vector of values in the range [0,1] for alpha transparency channel (0 means transparent and 1 means opaque).
+- `fixup`: a logical value which indicates whether the resulting RGB values should be corrected to ensure that a real color results.
+if fixup is FALSE RGB components lying outside the range [0,1] will result in an NA value.
+
+**`hsv {grDevices}`**: Create a vector of colors from vectors specifying hue, saturation and value.
+
+*usage* `hsv(h = 1, s = 1, v = 1, alpha)`
+
+*arguments*
+
+- `h`, `s`,`v`: numeric vectors of values in the range [0, 1] for ‘hue’, ‘saturation’ and ‘value’ to be combined to form a vector of colors. Values in shorter arguments are recycled.
+- `alpha`: numeric vector of values in the range [0, 1] for alpha transparency channel (0 means transparent and 1 means opaque).
+
+-----
 
@@
 
-> __Figure 1 (b, c):__ Global marine species richness according to (a) AquaMaps dataset and (b) IUCN dataset.
+> __Figure 1 (b, c):__ Global marine species richness according to (b) AquaMaps dataset and (c) IUCN dataset.
 The frequency plot to the right of each map shows relative species count per cell at each latitude; while both datasets peak in tropical latitudes near the equator, the frequency for IUCN maps drops quickly beyond 30°N and 30°S, while the frequency for AquaMaps remains robust well into temperate latitudes.
@@
 
-![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)
+![Absolute change in SPP score](figures/boxplot_abs_diff.png)
 
-Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
+___NOTE: These figures still need to be recreated using the latest and greatest data!___ _currently they use data generated in November or so, in ohiprep._
+
+> __Figure 3:__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
 
@@
 
-* Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
+> * Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
 Reducing the threshold will always increase the apparent range of a species, therefore the slight decrease in average score suggests increased spatial representation of threatened species.
-
-* Scenario 2 shows the effect of preferring AquaMaps data over IUCN, while maintaining the same presence threshold.
+> * Scenario 2 shows the effect of preferring AquaMaps data over IUCN, while maintaining the same presence threshold.
 This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so many but not all overlapping species will see a decrease in represented range.
 The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
-
-* Scenario 3 shows the effect of preferring AquaMaps data over IUCN, while eliminating the presence threshold.
+> * Scenario 3 shows the effect of preferring AquaMaps data over IUCN, while eliminating the presence threshold.
 Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.

### 2016-01-22 data_explore/boxplots.R

@@ -13,11 +13,27 @@ library(tidyr)
 library(stringr)
 library(ggplot2)
 
-dir_git  <- '~/github/ohiprep/globalprep/SPP_ICO/vAM_IUCN'
-st_base  <- read.csv(file.path(dir_git, 'data/spp_status_global_IUCNpref_prob0.4.csv'),  stringsAsFactors = FALSE)
-st_scen1 <- read.csv(file.path(dir_git, 'data/spp_status_global_IUCNpref_prob0.01.csv'), stringsAsFactors = FALSE)
-st_scen2 <- read.csv(file.path(dir_git, 'data/spp_status_global_AMpref_prob0.4.csv'),    stringsAsFactors = FALSE)
-st_scen3 <- read.csv(file.path(dir_git, 'data/spp_status_global_AMpref_prob0.01.csv'),   stringsAsFactors = FALSE)
+dir_data  <- '~/github/ohiprep/globalprep/SPP_ICO/vAM_IUCN'
+dir_fig <- '~/github/IUCN-AquaMaps/figures'
+
+### generic theme for all plots
+ggtheme_basic <- theme(axis.ticks = element_blank(),
+                       text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
+                       plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
+                       legend.position = 'none')
+
+ggtheme_plot <- ggtheme_basic + 
+  theme(panel.border = element_blank(),
+        panel.grid.minor = element_blank(), 
+        panel.grid.major = element_line(colour = 'grey90'),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey30"))
+
+
+st_base  <- read.csv(file.path(dir_data, 'data/spp_status_global_IUCNpref_prob0.4.csv'),  stringsAsFactors = FALSE)
+st_scen1 <- read.csv(file.path(dir_data, 'data/spp_status_global_IUCNpref_prob0.01.csv'), stringsAsFactors = FALSE)
+st_scen2 <- read.csv(file.path(dir_data, 'data/spp_status_global_AMpref_prob0.4.csv'),    stringsAsFactors = FALSE)
+st_scen3 <- read.csv(file.path(dir_data, 'data/spp_status_global_AMpref_prob0.01.csv'),   stringsAsFactors = FALSE)
 
 ### scenario 1: prefer IUCN data; drop AM probability threshold to 0.01
 ### scenario 2: prefer AM data; keep AM prob thresh at 0.40
[…]

### 2016-01-22 data_explore/data_explore_fxns.R

@@ -25,7 +25,12 @@ ggtheme_map <- ggtheme_basic +
   theme(axis.text = element_blank())
 
 ### Function to create a species list from scratch... use when new data is available
-create_spp_list <- function(spp_list_base, data_file, am_spp_cells, iucn_spp_cells, loiczid_raster) {
+create_spp_list <- function(spp_list_base, 
+                            data_file, 
+                            am_spp_cells, 
+                            iucn_spp_cells, 
+                            loiczid_raster) {
+
   spp_list_build <- spp_list_base %>%
     mutate(total_area = NA,
            am_area    = NA,

### 2016-01-23 manuscript-main.Rmd
updating the manuscript...

@@
     toc: no
-  pdf_document:
-    number_sections: yes
-    toc: no
   word_document: default
 ---
+
+<!--
+
+from PNAS (http://www.pnas.org/site/authors/preparation.xhtml):
+
+PNAS generally uses a two-column format averaging 67 characters, including spaces, per line.
+The maximum length of a regular research article is 6 pages (or 10 pages for PNAS Plus articles), including all text, spaces, and the number of characters displaced by figures, tables, and equations.
+
+An online submission tool provides authors with an estimation of whether their manuscript fits within the PNAS length requirements (see Length Estimate FAQ).
+Use the following guidelines to determine whether your manuscript exceeds the page-length requirements before submission:
+
+Text: When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces).
+Figures: Calculated at 180 characters per cm in height for one column and 360 characters per cm in height for two columns; a figure wider than 8.7 cm is two columns wide.
+Tables and equations: Calculated at 60 characters per line for one column and 120 per line for two columns; a table with more than 60 characters per line is two columns wide.
+If the text in a figure or table is deemed illegible for production, the figure or table may be enlarged to improve legibility.
+Add the word count to the character count to obtain a character count that includes spaces.
+
+See also: http://www.pnascentral.org/html/pnas_length_estimate_instructions.html
+
+-->
 

### 2016-01-23 ms_2_intro_res_disc.Rmd
updated figures without titles; boxplot = violin plot
updating the manuscript...

@@
 To understand the implications of differences between the AquaMaps and IUCN datasets, we compared how each data source represents the global spatial and taxonomic distribution of species.
-Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')`; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` species).
-For the `r format(n_spp_both, big.mark = ',')` (`r p_both_all`% of total) species mapped in both datasets, we examined how well the maps align and determined several issues that lead to misalignment between predicted species distributions, and outline possible improvements.
+Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` species).
+For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examined how well the maps align and determined several issues that lead to misalignment between predicted species distributions, and outline possible improvements.
 
@@
 -----
+
 <!--
@@
 The IUCN publishes species range maps as polygon shapefiles, bundled by taxonomic groups (REF).
-Using a geographic information system (GIS), experts outline spatial polygons to represent a given species' "limits of distribution" (IUCN 2014) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
-AquaMaps models species distribution maps based environmental preferences (e.g.
+Using a geographic information system (GIS), experts outline spatial polygons to represent a given species' "limits of distribution" (REF IUCN 2014) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
+AquaMaps models species distribution based environmental preferences (e.g.
 temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species.
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species (AquaMaps 2013)
+
+The methodologies behind these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).
+Geographic range data such as IUCN range maps generally include large commission errors, while predicted distribution models such as AquaMaps will likely include fewer commission errors but more omission errors.
+Each type of error bears different implications for conservation goals (Rondinini, 2006).
+
+<!-- Geographic ranges generate large commission errors because they assume homogenous species distributions.
+Predicted distribution data make explicit inferences on species occurrence and their commission and omission errors depend on model structure, on the omission of variables that determine species distribution and on data resolution.
+Omission errors lead to identifying networks of areas for conservation action that are smaller than required and centred on known species occurrences, thus affecting the comprehensiveness, representativeness and efficiency of selected areas.
+Commission errors lead to selecting areas not relevant to conservation, thus affecting the representativeness and adequacy of reserve networks.
+-->
 
@@
 IUCN-mapped species focus on tropical latitudes and away from the Atlantic and Eastern Pacific compared to AquaMaps-mapped species.
-These differences can be mitigated by using both datasets, but the underlying methodological differences complicate such direct comparisons, and furthermore ___requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species.___
+These differences can be mitigated by using both datasets, but the underlying methodological differences complicate such direct comparisons.
+___and furthermore requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species.___
 
@@
 
-___BSH suggestion___ To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (where on the map) against the ratio of extent (how much of the map) for each shared species (Fig.
-2a).
-__Overall we see a negative linear relationship, where increasing similarity in extent (extent ratio approaches 100%) correlates with a decrease in alignment.
-[explain/discuss].__
+To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (where on the map) against the ratio of extent (how much of the map) for each shared species (Fig.
+2A).
 
@@
 For many of these species, while the spatial distribution is similar, the IUCN range predicts a significant buffer zone beyond the AquaMaps range.
-Corals dominate this quadrant (`r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant).
+Corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant).
 Most corals and reef-associated organisms prefer shallower waters, and we suspect that many of these disagreements in extent can be explained by bathymetry: seafloor depth is explicitly included in AquaMaps models, while depth is frequently ignored as a factor in IUCN range maps, leading to overprediction of species range into areas too deep to support the species.
@@
 The global Ocean Health Index (OHI) (Halpern et al.
-2012, 2015), a composite index made up of 10 goals, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status for each of 221 exclusive economic zones.
-To maximize the number of represented species, spatial distributions were gleaned from both IUCN and AquaMaps datasets, preferring IUCN data for species in both sources.
+2012, 2015), a composite index made up of 10 sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
+To maximize the number of represented species, OHI gleans spatial distributions from both IUCN and AquaMaps datasets, prioritizing IUCN data for the subset of species in both sources.
 OHI uses a probability threshold of 40% to determine species presence for AquaMaps data.
-Species with no spatial data in either dataset were excluded, as were species with insufficient information to determine conservation status (including species listed as not evaluated or data deficient).
 
-_Briefly summarize results_
+We recalculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig.
+3</a>).
+Most country scores varied by fewer than five points across the three scenarios; a few country scores decreased by as many as 12 points, indicating a more dire assessment of species conservation status.
 
-Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted (<a href = '#fig3'>Figure 3</a>).
-When IUCN range maps are the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change.
+_Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted.
+When IUCN range maps are the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change._
 
@@
 
-working on the analysis...
-
-* rerun based on old AquaMaps at 50% (to verify methodology gets a close result to paper), then new AquaMaps for updated baseline
-    * consider also: filter on STATUS == 'Designated', and filter out those pesky fishery management areas and manatee speed zones...
-* run with threshold at 0% (not all the same as original paper)
-* run with filtering out data-poor species?
-* run with IUCN maps instead of AquaMaps
-* Which aspects of the analysis should we focus on?:
-    * overall protection, for 4th subset of protected areas? or overall for all subsets? (see paper, table 1)
-    * overall vs broken down by taxonomic groups? (see paper, figure 1)
-    * spatial distribution of gap species (see paper, figure 2)
+___working on the analysis...___
 
@@
 
-_While the differences between species distributions described by AquaMaps and IUCN can drive different conclusions for conservation research, some of the mechanistic causes can be addressed._ ___oh god this section is so passive voice...
-arrrrgghhhh___
+Understanding the causes and implications of the differences between these two important datasets can help us better align the predictions of species distributions, increasing their utility for conservation goals.
 
-__For data providers:__
+__Suggestions for data providers:__
 
@@
 Unfortunately, this recommendation appears to be applied haphazardly.
-Especially for demersal communities dependent on photosynthesis, bathymetry should be a primary consideration; even a low resolution cutoff of 200 meters would improve the utility of species range maps.
+Especially for demersal communities dependent on photosynthesis, bathymetry should be a primary consideration; even a simple cutoff at the photosynthetic limit of 200 meters would improve the utility of species range maps.
+* The AquaMaps model is a powerful tool for quickly estimating species distribution based on limited information; however, the model's output is only as good as the input data, which for marine species is often sorely limited.
+AquaMaps encourages experts to review and comment on the predicted species distribution maps; lessons from the small sample of reviewed maps can be generalized, for example to other species in the same taxonomic group, to further refine the predictions, especially for data-poor species.
+* In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.
+Currently the area restrictions are denoted by FAO Major Fishing Areas (REF find a reference in AquaMaps somewhere); however, allowing area restrictions based on biogeographical criteria, a la Marine Ecoregions of the World (), would likely have better resolution and predictive power, especially for data-poor species.
 
-* The AquaMaps model is a powerful tool for quickly estimating species presence based on limited information; however, the model's output is only as good as the input data, which for marine species is often sorely limited.
-AquaMaps encourages experts to review and comment on the predicted species distribution maps; lessons from the small sample of reviewed maps can be generalized, for example to other species in the same taxonomic group, to further refine the predictions, especially for data-poor species.
-
-* In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.
-Currently the area restrictions are denoted by FAO Major Fishing Areas (REF find a reference in AquaMaps somewhere); however, allowing area restrictions based on biogeographical criteria, a la MEOW, would likely have better resolution and predictive power, especially for data-poor species.
-
-__For data users:__
+__Suggestions for data users:__
 

### 2016-01-23 ms_3_methods.Rmd
updating the manuscript...

@@
 
-### taxonomic distribution comparison
+### comparison of taxonomic and regional distribution
 
-To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
-We then filtered the species list to those that have been evaluated for the IUCN Red List of Threatened Species (in SOM).
+To examine the overall taxonomic distribution across the spatial datasets (Fig.
+1a), we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
 
-___Currently this is also included in body - take out and leave in here?
-Per BH: "Either cut or move to data description section of methods above."___: The IUCN releases spatial data sets when a taxonomic group (typically on the scale of order or family) has been comprehensively assessed, to guard against sample bias (though non-comprehensive datasets are available for reptiles and marine fish).
-As such, spatial data for many taxonomic classes remain unavailable, and within a class, the assessed sub-groups may not represent the entire class.
+To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; species presence within a grid cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the count of the species present.
+For the AquaMaps dataset, we determined per-cell species richness by counting all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.
+We represented relative distribution of species richness for each dataset by plotting average species count against latitude and longitude (Fig.
+1b, 1c).
 
-### global spatial distribution comparison
+### comparison of paired maps
 
-To compare the spatial representation of the two datasets directly, we first rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; species presence within a grid cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the count of the species present.
-For the AquaMaps dataset, we determined per-cell species richness by counting all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.
-We represented relative distribution of species richness for each dataset by plotting average species count against latitude and longitude.
+Using genus and species binomials as a matching key, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.
+To avoid double-counting, we removed subpopulations and species aliases.
+We determined species presence within each spatial cell using the same criteria as outlined above.
 
-_map pairs comparison_
-
-Using genus and species binomials to identify paired maps, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`).
-We used the same criteria as outlined above to determine species presence within a cell.
-
-___from results/discussion: paraphrase or expand upon this___ Overlaying paired distribution maps for a given species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller range intersecting the larger range; and _area alignment_, which we defined as the ratio of the smaller range to the larger range.
+Overlaying paired distribution maps for a given species, we defined and calculated _distribution alignment_ and _extent alignment_ and plotted these in Fig. 2:
 
@@
 
+We visually inspected a random selection of paired distribution maps from each quadrant to identify possible mechanistic causes of misalignment.
+To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS (REF) to overlay a selection of IUCN and AquaMaps maps with a 200 meter bathymetry contour.
+
 ### Methods for OHI comparison
 
+Using methods and supplemental materials from OHI (Halpern et al.
+2012, 2015), we modified the original code for OHI 2015 Species status (SPP) (REF), allowing for flexibility in prioritized data source and AquaMaps presence threshold.
+We ran the SPP code three times, prioritizing IUCN over AquaMaps for a 0% threshold, and prioritizing AquaMaps over IUCN for both a 40% and 0% threshold.
+We compared each of these to the output of the published OHI 2015 SPP model (prioritizing IUCN over AquaMaps at a 40% threshold).
+
 ### Methods for MPA Gap Analysis comparison
+
+Based upon the methods described in Klein et al.
+2015, we reconstructed the analysis using the subset of protected areas spatially covering a marine area and classified as IUCN I-IV ___(Klein paper, table 1, subset 4 - seems to be the one they focus on the most)___, converting the World Database of Protected Areas (WDPA, REF) dataset to a raster of 0.5° cells indicating proportion of protected area within the cell ___(consider also: filter on STATUS == 'Designated', and filter out those pesky fishery management areas and manatee speed zones...)___.
+To verify our reconstruction we ran the analysis using the AquaMaps dataset available in April 2015 (REF) at a presence threshold of 50%.
+We then updated the analysis using the most recent AquaMaps dataset (REF?) at 50% threshold
+
+* Possible analyses:
+    * run AquaMaps (new data) with threshold at 0%
+        * already done in original paper, but redo it to compare to AM + IUCN scenario below
+    * run with AquaMaps only, but filtering out data-poor species?
+    * run with IUCN maps only in place of AquaMaps?
+    * run with AquaMaps and IUCN both, using 0% threshold; which preferred? 
+        * Should we clip corals (and coral-associated taxa) to 200 meter bathymetry, just for fun?
+* Which aspects of the analysis should we focus on?:
+    * overall vs broken down by taxonomic groups? (see paper, figure 1)
+    * spatial distribution of gap species (see paper, figure 2)

### 2016-01-23 ms_4_figs.Rmd
updating the manuscript...

@@
 # Figures
+
+<!--
+
+From PNAS http://www.pnas.org/site/authors/preparation.xhtml:
+
+Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text.
+Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
+(See SI below for supplementary material.) Color images must be in RGB (red, green, blue) mode.
+Include the font files for any text.
+Images must be final size, preferably one column width (8.7 cm).
+Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
+Numbers, letters, and symbols should be no smaller than 6 points (2 mm) and no larger than 12 points (6 mm) after reduction and must be consistent.
+Composite figures must be preassembled.
+Figures must be submitted as separate files, not embedded in manuscript text.
+See the Digital Art Guidelines.
+
+See also: http://www.pnas.org/site/misc/digitalart.pdf
+
+Label like so: 
+Fig. 2.
+Diffraction patterns at three different angles and time-dependent structural changes of JDF-L1.
+(A–C) Diffraction images observed at different zone axes before time 0; n is the electron direction.
+(D) Femtosecond intensity changes of (200) and (201) spots with time.
+Two different types of dynamics were observed: a decay with a femtosecond time constant (see main text) and another without a decay.
 
@@
 
+-->
+
 -----
 
-### Figure 1
+### Fig. 1
 
@@
 
->  __Figure 1(a):__ Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.
+>  __Fig.
+1.__ (A) Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.
 AquaMaps encompasses a broader range of taxa than IUCN, while IUCN focuses on comprehensively assessing select taxonomic groups, typically at the level of order or family.
 Overlapping species are dominated by bony fishes (`r n_bonyfish` species, primarily tropical taxa) and corals (`r n_corals` species).
-
-> __Figure 1 (b, c):__ Global marine species richness according to (b) AquaMaps dataset and (c) IUCN dataset.
+(B, C) Global marine species richness according to (B) AquaMaps dataset and (C) IUCN dataset.
 The frequency plot to the right of each map shows relative species count per cell at each latitude; while both datasets peak in tropical latitudes near the equator, the frequency for IUCN maps drops quickly beyond 30°N and 30°S, while the frequency for AquaMaps remains robust well into temperate latitudes.
@@
 ![](figures/AreaRatiovsPercOverlap.png)
-
-> __Figure 2 (a):__ For each paired-map species, we calculated two dimensions of spatial alignment: _distribution alignment_, which we defined as the proportion of the smaller predicted range intersecting the larger predicted range (where on the map); and _extent alignment_, which we defined as the ratio of the smaller range area to the larger range area (how much of the map).
-For a species whose distribution is well understood and described in both datasets, we would expect to see a value near 100% for each dimension of alignment.
+![](figures/barchart_spp_gp_quads.png)
 
@@
 
-> The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
+> __Fig.
+2.__ (A) DESCRIPTION, then explanation...
+The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
 (n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
@@
 (n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).  
-
-![](figures/barchart_spp_gp_quads.png)
-
-> __Figure 2 (b):__ Spatial alignment of paired-map species by taxonomic group.
-___REDO with all corals grouped together!___  _Also - is it distracting to use order names for some groups and not others?
+(B) Spatial alignment of paired-map species by taxonomic group.
+_is it distracting to use order names for some groups and not others?
 Some of the groups had pretty long names ("acanthuridae" is way shorter than "surgeonfish, tangs, and unicornfish")_
-
-<!--AquaMaps accounts for seafloor depth as a limiting factor, highly constraining the predicted habitat for photosynthesizing corals, while IUCN maps for coral species often include wide buffers that would extend far beyond appropriate depth for coral habitat, thus likely overestimating the actual species range (see SOM).
-A similar mechanism may also drive poor extent alignment of other demersal and reef-associated organisms.-->
 
@@
 
-> __Figure 3:__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
+> __Fig. 3.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
 
@@
 > * Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
-Reducing the threshold will always increase the apparent range of a species, therefore the slight decrease in average score suggests increased spatial representation of threatened species.
-> * Scenario 2 shows the effect of preferring AquaMaps data over IUCN, while maintaining the same presence threshold.
-This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so many but not all overlapping species will see a decrease in represented range.
+Reducing the threshold increases the apparent range of a species; the slight decrease in average score suggests increased spatial representation of threatened species.
+> * Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.
+This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so most overlapping species will see a decrease in represented range.
 The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
-> * Scenario 3 shows the effect of preferring AquaMaps data over IUCN, while eliminating the presence threshold.
+> * Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, while eliminating the presence threshold.
 Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
-The large decrease seems to indicate that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.
+The large decrease indicates that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.

### 2016-01-23 ms_6_supplemental.Rmd
updating the manuscript...

@@
 # Supplemental Information
+
+<!--
+
+from PNAS http://www.pnas.org/site/authors/preparation.xhtml:
+
+The main text of the paper must stand on its own without the SI.
+Refer to SI in the manuscript at an appropriate point in the text.
+Number supporting figures and tables starting with S1, S2, etc.
+Authors are limited to no more than 10 SI files, not including movie files.
+
+Authors who place detailed materials and methods in SI must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the online methods.
+If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
+
+Because PNAS edits SI and composes it into a single PDF, authors must provide the following file formats only:
+
+Text: Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).
+Figures: Provide a brief legend for each supporting figure after the supporting text.
+Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text.
+When saving TIFF files, use only LZW compression; do not use JPEG compression.
+Do not save figure numbers, legends, or author names as part of the image.
+Composite figures must be preassembled.
+3D figures: Supply a composable U3D or PRC file so that it may be edited and composed.
+Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.
+Tables: Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file.
+Do not use tabs or spaces to separate columns in Word tables.
+Datasets: Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
+Movies: Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file.
+All movies should be submitted at the desired reproduction size and length.
+Movies should be no more than 10 MB in size.
+Still images: Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.
+Appendices: PNAS prefers that authors submit individual source files to ensure readability.
+If this is not possible, supply a single PDF file that contains all of the SI associated with the paper.
+This file type will be published in raw format and will not be edited or composed.
+
+-->
 

### 2016-01-25 manuscript-main.Rmd
updating manuscript to match PNAS format...

@@
 title: "Aligning species range data to better serve science and conservation"
-author: "Casey O'Hara, Jamie Afflerbach, Benjamin S. Halpern"
+author: "Casey O'Hara^1^, Jamie Afflerbach^1^, Benjamin S. Halpern^1, 2, 3^"
 output:
+  word_document: default
   pdf_document:
@@
     toc: no
-  word_document: default
 ---
+
+<!--Title: Titles should be no more than three typeset lines (generally 135 characters including spaces) and should be comprehensible to a broad scientific audience.
+Please provide a short title of fewer than 50 characters for mobile devices and RSS feeds.-->
+
+_Classification:_
+
+* BIOLOGICAL SCIENCES: 
+    * Ecology? Environmental Sciences? Sustainability Science?
+
+_Author affiliation:_ <!--Include department, institution, and complete address, with the ZIP/postal code, for each author.
+Use superscripts to match authors with institutions.
+Authors are encouraged to supply their ORCID ID at submission-->
+
+1. National Center for Ecological Analysis and Synthesis, 735 State Street, Suite 300, Santa Barbara CA 93101
+2. Bren School of Environmental Science and Management, University of California, Santa Barbara CA 93106
+3. Imperial College London, Silwood Park Campus, Ascot, United Kingdom (need address?)
+
+_Corresponding author:_ <!--Corresponding author: The name, complete address, telephone number, and email address of the author to whom correspondence and proofs should be sent.
+Email addresses will appear in the article footnotes.-->
+
+* Casey O'Hara, National Center for Ecological Analysis and Synthesis, 735 State Street, Suite 300, Santa Barbara CA 93101; _NCEAS phone #?_; ohara@nceas.ucsb.edu
+
+_Keywords:_ <!--Keywords are listed below the abstract of the published article. At least three keywords are required at submission.--> TBD
 

### 2016-01-25 ms_1_abst_sig.Rmd
updating manuscript to match PNAS format...

@@
 ## Abstract
+
+_Abstract.
+Provide an abstract of no more than 250 words on page 2 of the manuscript.
+Abstracts should explain to the general reader the major contributions of the article.
+References in the abstract must be cited in full within the abstract itself and cited in the text._
 
@@
 
-_subset of the abstract_
+_Significance Statement.
+Authors must submit a 120-word-maximum statement about the significance of their research paper written at a level understandable to an undergraduate-educated scientist outside their field of specialty.
+The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
+The Significance Statement appears in the paper itself and is required for all research papers._

### 2016-01-25 ms_2_intro_res_disc.Rmd
updating manuscript to match PNAS format...

@@
 
-_Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean - and you will find a unique ecosystem defined as much by which species are absent as by which species are abundant._ Knowing where individuals of a species thrive is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
+<!--
+research questions: How do the two datasets differ in prediction of species distribution?
+Why do they differ?
+What are the implications of these differences?
+and how can we improve our confidence in their predictions?_ _this should also go in the abstract_
+-->
+
+Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean - and you will find a unique ecosystem defined as much by which species are absent as by which species are abundant.
+Knowing where individuals of a species thrive is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
 This knowledge provides foundational information for
@@
 
-<!--
-## Overview of AquaMaps and IUCN datasets
-
-___BSH comment: Fold this section into the methods and/or Supplementary Methods, or maybe into the results section just below on differences between the datasets – but make the description VERY brief.___
-
-The IUCN publishes species range maps as polygon shapefiles, bundled by taxonomic groups (REF).
-Using a geographic information system (GIS), experts outline spatial polygons to represent a given species' extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
-_IUCN releases a range map bundle for a taxonomic group once it has been "comprehensively assessed," i.e.
-at least 90% of the species within the taxonomic group have been evaluated (REF)._  While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.
-
-AquaMaps develops species distribution maps based on modeled relative environmental suitability.
-For each mapped species, environmental preferences (e.g.
-temperature, depth, salinity) are deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species.
-A subset of the resulting distribution maps (`r format(n_spp_am_rev, big.mark = ',')` of `r format(n_spp_am, big.mark = ',')` species maps; `r p_spp_am_rev`%) have been checked or formally reviewed by experts, to validate and improve the parameters and underlying model.
-
-IUCN range maps are intended to include all possible regions in which a species may be present, with the caveat that this does not imply the species is evenly distributed everywhere within the boundaries of the map(REF).
-AquaMaps modeled distribution maps provide a more nuanced prediction of species presence within the extent of occurrence, but the model may not capture all the complexities that drive species distribution (REF).
-Due to these differences in methodology and intent, IUCN range maps are more likely to overpredict species presence (and underpredict absence) while AquaMaps modeled distribution maps are relatively more likely to underpredict species presence.(REF and REF)
--->
-
 # Results and Discussion
@@
 
-<!-- Geographic ranges generate large commission errors because they assume homogenous species distributions.
-Predicted distribution data make explicit inferences on species occurrence and their commission and omission errors depend on model structure, on the omission of variables that determine species distribution and on data resolution.
-Omission errors lead to identifying networks of areas for conservation action that are smaller than required and centred on known species occurrences, thus affecting the comprehensiveness, representativeness and efficiency of selected areas.
-Commission errors lead to selecting areas not relevant to conservation, thus affecting the representativeness and adequacy of reserve networks.
--->
-
 The two datasets have notably different taxonomic (Fig.
@@
 These differences can be mitigated by using both datasets, but the underlying methodological differences complicate such direct comparisons.
-___and furthermore requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species.___
+_and furthermore requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species._
 
@@
 The extent-mismatched maps contained in the upper left quadrant (n = `r n_q2`) provide a richer set of examples to understand the fundamental differences between the datasets.
-For many of these species, while the spatial distribution is similar, the IUCN range predicts a significant buffer zone beyond the AquaMaps range.
+For many of these species, while the spatial distribution is similar, the IUCN range includes a significant buffer zone beyond the AquaMaps range.
 Corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant).
-Most corals and reef-associated organisms prefer shallower waters, and we suspect that many of these disagreements in extent can be explained by bathymetry: seafloor depth is explicitly included in AquaMaps models, while depth is frequently ignored as a factor in IUCN range maps, leading to overprediction of species range into areas too deep to support the species.
-(For examples and details, see SOM)
+Most corals and reef-associated organisms prefer shallower waters, and we suspect that many of these disagreements in extent can be explained by bathymetry: seafloor depth is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps, leading to overprediction of species range into areas too deep to support the species (For examples and details, see SOM).
+_commission error terminology?_
 
@@
 This results in an area ratio close to 100%, but similar extents are meaningless when the distributions are poorly aligned.
+_commission error terminology?_
 
-Moving from the lower right to lower left quadrant, AquaMaps more frequently extrapolates species presence well beyond known occurrences, resulting in poor matches in both distribution and extent.
+Moving from the lower right to lower left quadrant, AquaMaps more frequently relies on limited data to predict species presence well beyond known occurrences, resulting in poor matches in both distribution and extent.
 Of the `r n_q4` species in this quadrant, 159 (39.7%) are data-poor (in which the AquaMaps environmental envelope model is based on fewer than 10 known cell occurrences), compared with 21.1% for the full set of map-paired species and only 8.2% of species in the upper left quadrant.
@@
 This breakdown by quadrant helps explain the negative linear relationship seen in the plot, in which increasing similarity in extent correlates with decreasing distribution alignment.
-Expanding species range beyond known occurrences, the marginal range predicted by the AquaMaps model will fall in different locations than the marginal range predicted using IUCN methodology.
+When extrapolating species range beyond known occurrences, the marginal range predicted by the AquaMaps model will fall in different locations than the marginal range predicted using IUCN methodology.
 For species with dissimilar extents, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
@@
 
-# Implications of the differences
+# Implications
 
-___From JSL:___ _I think you need a broader discussion of implications.
+_brief intro to implications - where/how long have IUCN/AquaMaps been used for various purposes; then introduce two case studies_
+
+* ___From JSL:___ _I think you need a broader discussion of implications.
 How many times has IUCN/Aquamaps been cited?
@@
 Then you can muse about how results/interpretations would be different depending on which dataset they used, and reference Fig 3’s._
+* _Then, say for example, the OHI uses these datasets together (don’t need to get too much into OHI background I think, maybe they don’t even need to know about the biodiversity goal?), but say that to try to combine them they use a threshold at 40% since others use this threshold (but check out the SOM for analyses about thresholds)_
 
-_Then, say for example, the OHI uses these datasets together (don’t need to get too much into OHI background I think, maybe they don’t even need to know about the biodiversity goal?), but say that to try to combine them they use a threshold at 40% since others use this threshold (but check out the SOM for analyses about thresholds)_
-
-## Application to OHI
+**Case Study: The Ocean Health Index**
 
 The global Ocean Health Index (OHI) (Halpern et al.
-2012, 2015), a composite index made up of 10 sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
+2012, 2015), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
 To maximize the number of represented species, OHI gleans spatial distributions from both IUCN and AquaMaps datasets, prioritizing IUCN data for the subset of species in both sources.
@@
 
-## Application to MPA Gap Analysis
+**Case Study: MPA Gap Analysis**
 
 Klein et al.
-(2015) compared the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect key species, and to identify which species fall through the gaps in protection.
-The MPA gap analysis relied on the AquaMaps database available in early 2015, comprising distribution maps for XXX species.
-Species presence was defined using a probability threshold of 50% or greater.
-The study defined four subsets of protected areas using the World Database of Protected Areas, based upon combinations of IUCN protected area classification, marine designation, and spatial marine coverage.
+(2015) compared the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect key species and to identify which species fall through gaps in protection.
+The study relied on the AquaMaps database, using a probability threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
 The study determined that currently defined marine protected areas (IUCN categories I through IV overlapping with marine areas) are woefully inadequate in protecting marine biodiversity; 90.5% of species have less than 5% of their overall range represented within marine protected areas.
@@
 
-# Conclusions: What can be done to address the differences
+# Conclusions
 
-Understanding the causes and implications of the differences between these two important datasets can help us better align the predictions of species distributions, increasing their utility for conservation goals.
+What can be done to address the differences between these two important marine species distribution datasets?
+We have identified several potential causes of commission errors among these datasets.
+Understanding these causes points toward a few improvements that may help us better align the predictions of species distributions, increasing their utility for ecology research and conservation actions.
 
@@
 
-* The training manual for experts creating IUCN Red List maps for marine species suggests that "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species" (REF https://www.conservationtraining.org/mod/scorm/player.php?a=1412&currentorg=M5L1_ORG&scoid=6492).
-Unfortunately, this recommendation appears to be applied haphazardly.
-Especially for demersal communities dependent on photosynthesis, bathymetry should be a primary consideration; even a simple cutoff at the photosynthetic limit of 200 meters would improve the utility of species range maps.
+* The IUCN Red List instructional materials for mapping marine species suggest that "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species" (REF https://www.conservationtraining.org/mod/scorm/player.php?a=1412&currentorg=M5L1_ORG&scoid=6492).
+Unfortunately, this recommendation appears to be applied inconsistently.
+Especially for demersal communities dependent on photosynthesis, bathymetry should be a primary consideration; even a simple cutoff at the photosynthetic limit of 200 meters would minimize a likely source of commission errors.
 * The AquaMaps model is a powerful tool for quickly estimating species distribution based on limited information; however, the model's output is only as good as the input data, which for marine species is often sorely limited.
-AquaMaps encourages experts to review and comment on the predicted species distribution maps; lessons from the small sample of reviewed maps can be generalized, for example to other species in the same taxonomic group, to further refine the predictions, especially for data-poor species.
+AquaMaps encourages experts to review and comment on the predicted species distribution maps; lessons from the small sample of reviewed maps can be generalized, for example to other species in the same taxonomic group, to further refine the predictions.
+This is especially important for data-poor species.
 * In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.
-Currently the area restrictions are denoted by FAO Major Fishing Areas (REF find a reference in AquaMaps somewhere); however, allowing area restrictions based on biogeographical criteria, a la Marine Ecoregions of the World (), would likely have better resolution and predictive power, especially for data-poor species.
+Currently the area restrictions are denoted by FAO Major Fishing Areas (REF find a reference in AquaMaps somewhere); however, allowing area restrictions based on biogeographical criteria, e.g.
+Marine Ecoregions of the World (REF), would likely have better resolution and predictive power, especially for data-poor species.
 
@@
 
-* understand that range maps a la IUCN tend to overestimate presence; AquaMaps are less likely to overestimate presence, but trade off for overestimating absence.
-Both better than simple point locality data.
-(Rondinini etc)
+* For depth-limited species and taxa, clipping IUCN range polygons to a reasonable bathymetry contour can reduce commission errors due to overprediction of species presence into unsuitable habitats.
 * Using both datasets together can increase the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized.  Consider:
-    * An AquaMaps presence threshold of 0% more closely approximates the "limits of distribution" outlined by IUCN range maps.
-    * Additional filters and criteria on AquaMaps distribution maps can help avoid overextrapolation for data poor species.
-The AquaMaps native range dataset includes information on expert review and the number of occurrence cells used to generate the environmental envelope.
-    * For depth-limited species or taxa, clipping IUCN range polygons to a reasonable bathymetry contour can limit overprediction of species presence into unsuitable habitats.
+    * An AquaMaps presence threshold of 0% most closely approximates the "limits of distribution" defined by IUCN range maps.
+    * Additional filters on AquaMaps distribution criteria, such as expert review status or the number of occurrence cells used to generate the species model, can help avoid overextrapolation for data poor species.

### 2016-01-25 ms_3_methods.Rmd
updating manuscript to match PNAS format...

@@
 
-### comparison of taxonomic and regional distribution
-
+**Comparison of taxonomic and regional distribution:** 
 To examine the overall taxonomic distribution across the spatial datasets (Fig.
-1a), we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
+1A), we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
 
@@
 We represented relative distribution of species richness for each dataset by plotting average species count against latitude and longitude (Fig.
-1b, 1c).
+1B, 1C).
 
-### comparison of paired maps
-
+**Comparison of paired maps:** 
 Using genus and species binomials as a matching key, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.
@@
 
-### Methods for OHI comparison
-
+**Methods for OHI case study: **
 Using methods and supplemental materials from OHI (Halpern et al.
@@
 
-### Methods for MPA Gap Analysis comparison
-
+**Methods for MPA Gap Analysis case study:** 
 Based upon the methods described in Klein et al.
@@
     * run AquaMaps (new data) with threshold at 0%
-        * already done in original paper, but redo it to compare to AM + IUCN scenario below
+        * already done in original paper for old data, but redo it to compare to AM + IUCN scenario below
     * run with AquaMaps only, but filtering out data-poor species?

### 2016-01-25 ms_4_figs.Rmd
updating manuscript to match PNAS format...

@@
-# Figures
+# Figures and captions
+
+___still need work on captions; then separate captions from images___
 
@@
 n_bonyfish <- nrow(spp_map_pairs %>% 
-                     filter(spp_group %in% c('WRASSES', 'COMBTOOTHBLENNIES', 'BUTTERFLYFISH', 
-                                             'PUFFERFISH', 'GROUPERS', 'SEABREAMS_PORGIES', 'PARROTFISH', 
-                                             'SURGEONFISH_TANGS_UNICORNFISH', 'ANGELFISH', 'TUNAS_BILLFISHES', 
+                     filter(spp_group %in% c('WRASSES', 'BUTTERFLYFISH', 'PARROTFISH',
+                                             'COMBTOOTHBLENNIES', 'SEABREAMS_PORGIES', 
+                                             'PUFFERFISH', 'GROUPERS', 'TUNAS_BILLFISHES', 
+                                             'SURGEONFISH_TANGS_UNICORNFISH', 'ANGELFISH', 
                                              'DAMSELFISH', 'BONEFISH_TARPONS')))
@@
 >  __Fig.
-1.__ (A) Number and proportion of species, listed by taxa, included in each dataset: IUCN, AquaMaps, or both.
+1.__ (A) Number and proportion of species by taxa included in each dataset.
 AquaMaps encompasses a broader range of taxa than IUCN, while IUCN focuses on comprehensively assessing select taxonomic groups, typically at the level of order or family.
@@
 (B, C) Global marine species richness according to (B) AquaMaps dataset and (C) IUCN dataset.
-The frequency plot to the right of each map shows relative species count per cell at each latitude; while both datasets peak in tropical latitudes near the equator, the frequency for IUCN maps drops quickly beyond 30°N and 30°S, while the frequency for AquaMaps remains robust well into temperate latitudes.
-The frequency plot above each map shows relative species count at each longitude, showing a slight bias in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
+The margin frequency plots show relative species count per cell at each latitude and longitude; while both datasets peak in tropical latitudes near the equator, the frequency for IUCN maps drops quickly beyond 30°N and 30°S, while the frequency for AquaMaps remains robust well into temperate latitudes.
+The longitude frequency plots show a slight bias in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
 
@@
 > __Fig.
-2.__ (A) DESCRIPTION, then explanation...
+2.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r n_spp_both` species included in both IUCN and AquaMaps datasets.
 The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
@@
 (n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).  
-(B) Spatial alignment of paired-map species by taxonomic group.
+(B) Alignment breakdown of paired-map species by taxonomic group.
 _is it distracting to use order names for some groups and not others?
-Some of the groups had pretty long names ("acanthuridae" is way shorter than "surgeonfish, tangs, and unicornfish")_
+Some of the IUCN taxonomic groups had pretty long names ("acanthuridae" is way shorter than "surgeonfish, tangs, and unicornfish")_
 

### 2016-01-25 ms_5_refs.Rmd
updating manuscript to match PNAS format...

@@
 # References
+
+___still need to check citations, format properly, put into order___
+
+<!--
+Acknowledgments. List acknowledgments and funding sources. PNAS collaborates with FundRef to use a standard taxonomy of funder names. Dedications are rarely allowed.
+
+References.
+References must be in PNAS style.
+Only published or in-press papers and books may be cited in the reference list.
+For references to in-press articles, please confirm with the cited journal that the article is in press and include a DOI number and online publication date.
+Unpublished abstracts of papers presented at meetings or references to "data not shown" are not permitted.
+References should be cited in numerical order as they appear in text, and all references, including for the SI, should be included in the main manuscript file.
+Because tables and figures will be inserted in the text where first cited, references in these sections should be numbered accordingly.
+Include the full title for each cited article.
+All authors (unless there are more than five) should be named in the citation.
+If there are more than five, list the first author's name followed by et al.
+Provide inclusive page ranges for journal articles and book chapters.
+Cite databases in the text or as footnotes.
+
+Journal articles are cited as follows:
+
+10.
+Neuhaus J-M, Sitcher L, Meins F, Jr, Boller T (1991) A short C-terminal sequence is necessary and sufficient for the targeting of chitinases to the plant vacuole.
+Proc Natl Acad Sci USA 88(22):10362–10366.
+-->
 

### 2016-01-25 data_explore/data_explore.Rmd
updated figures without titles; boxplot = violin plot
updated figures; added word count; etc

@@
            hjust = 0, vjust = 0,
-           color = 'grey40', size = 2,
+           color = 'grey30', 
+           size = 2,
            fontface = 'bold.italic', angle = 90,
@@
            hjust = 0, vjust = 0,
-           color = 'grey40', size = 2,
+           color = 'grey30', 
+           size = 2,
            fontface = 'bold.italic', angle = 0,
@@
 
-ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 6, height = 3.5)
+ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 17.8, height = 9, units = 'cm')
 
@@
 
-ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap-data_poor.png'), width = 6, height = 3.5)
+ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap-data_poor.png'), width = 17.8, height = 9, units = 'cm')
 
@@
   scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 2, color = 'grey20') +
+  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, 
+            #size = 2, 
+            color = 'grey20') +
   coord_flip() +
@@
        filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
-       width = 6, height = 3.5)
+       width = 17.8, height = 7, units = 'cm')
 
@@
        filename = file.path(dir_fig, 'barchart_spp_risk_quads.png'),
-       width = 6, height = 1.75)
+       width = 17.8, height = 7, units = 'cm')
 
@@
 
-taxa_df_long1 <- taxa_df %>%
-  dplyr::select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both) %>%
-  gather(dataset, n_spp, AquaMaps:Both)
-bar_taxa1 <- ggplot(taxa_df_long1, aes(x = class, y = n_spp)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'identity') +
-  facet_grid(dataset ~ ., scales = 'free_y') +
-  theme(axis.text.x = element_text(angle = 75, hjust = 1))
-print(bar_taxa1)
+# taxa_df_long1 <- taxa_df %>%
+#   dplyr::select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both) %>%
+#   gather(dataset, n_spp, AquaMaps:Both)
+# bar_taxa1 <- ggplot(taxa_df_long1, aes(x = class, y = n_spp)) +
+#   ggtheme_plot + 
+#   geom_bar(stat = 'identity') +
+#   facet_grid(dataset ~ ., scales = 'free_y') +
+#   theme(axis.text.x = element_text(angle = 75, hjust = 1))
+# print(bar_taxa1)
 
@@
   geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
-            size = 3, angle = 90, hjust = 0, color = 'grey30') +
+            size = 3, 
+            angle = 90, hjust = 0, color = 'grey30') +
   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
@@
 print(bar_taxa2)
-ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), height = 3, width = 6)
+ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
+       width = 17.8, height = 7, units = 'cm')
 
@@
   geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
-            size = 3, angle = 90, hjust = 0, color = 'grey30') +
+            #size = 3, 
+            angle = 90, hjust = 0, color = 'grey30') +
   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
@@
 print(bar_taxa3)
-ggsave(filename = file.path(dir_fig, 'bar_by_taxa_assessed.png'), height = 3, width = 6)
+ggsave(filename = file.path(dir_fig, 'bar_by_taxa_assessed.png'), 
+       width = 17.8, height = 6, units = 'cm')
 
@@
 
-``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
+``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = FALSE}
 ```
+
+All AquaMaps species richness map:
+
+![figures/all_aquamaps_map.png](../figures/all_aquamaps_map.png)
+
+All IUCN species richness map:
+
+![figures/all_iucn_map.png](../figures/all_iucn_map.png)
 

### 2016-01-25 data_explore/data_explore_am_thresh.Rmd
updated figures; added word count; etc

@@
 
-This looks at the full Aquamaps dataset... should it be limited to species with valid categories?
+This looks at the full 2014 Aquamaps dataset... should it be limited to species with valid categories?
 
@@
   annotate('text', x = spp_df3_mean, y = .16, hjust = 0, vjust = -.5,
-           angle = 90, size = 2, color = 'grey20',
+           angle = 90, 
+           #size = 2, 
+           color = 'grey20',
            label = 'mean') + 
@@
   annotate('text', x = spp_df3_med, y = .16, hjust = 0, vjust = -.5,
-           angle = 90, size = 2, color = 'grey20',
+           angle = 90, 
+           #size = 2, 
+           color = 'grey20',
            label = 'median') + 
@@
 
-ggsave(file.path(dir_fig, 'AMThreshold.png'), width = 6, height = 3.5)
+ggsave(file.path(dir_fig, 'AMThreshold.png'), 
+       width = 17.8, height = 9, units = 'cm')
 ```
@@
   annotate('text', x = spp_df3a_mean, y = .16, hjust = 0, vjust = -.5,
-           angle = 90, size = 2, color = 'grey20',
+           angle = 90, 
+           #size = 2, 
+           color = 'grey20',
            label = 'mean') + 
@@
   annotate('text', x = spp_df3a_med, y = .16, hjust = 0, vjust = -.5,
-           angle = 90, size = 2, color = 'grey20',
+           angle = 90, 
+           #size = 2, 
+           color = 'grey20',
            label = 'median') + 
@@
 
-ggsave(file.path(dir_fig, 'AMThreshold2.png'), width = 6, height = 3.5)
+ggsave(file.path(dir_fig, 'AMThreshold2.png'), 
+       width = 17.8, height = 9, units = 'cm')
 
@@
 
-ggsave(file.path(dir_fig, 'scatter_AMThreshold.png'), width = 6, height = 3.5)
+ggsave(file.path(dir_fig, 'scatter_AMThreshold.png'), 
+       width = 17.8, height = 9, units = 'cm')
 
@@
   annotate('text', x = .40, y = .05, hjust = 0, vjust = -.5,
-           angle = 90, size = 2, color = 'grey20',
+           angle = 90, 
+           #size = 2, 
+           color = 'grey20',
            label = 'Ocean Health Index: 40%') + 
   annotate('text', x = .50, y = .05, hjust = 0, vjust = -.5,
-           angle = 90, size = 2, color = 'grey20',
+           angle = 90, 
+           #size = 2, 
+           color = 'grey20',
            label = 'MPA Gap Analysis: 50%') + 
   annotate('text', x = .60, y = .05, hjust = 0, vjust = -.5,
-           angle = 90, size = 2, color = 'grey20',
+           angle = 90, 
+           #size = 2, 
+           color = 'grey20',
            label = 'Marine mammals: 60%') + 
   annotate('text', x = .75, y = .85,
-           size = 2, color = 'grey20',
-           label = 'R^2 = .995') + 
+           #size = 2, 
+           color = 'grey20',
+           label = 'R^2^ = .995') + 
   scale_x_continuous(expand = c(0, 0)) +
@@
 
-ggsave(file.path(dir_fig, 'ribbon_AMThreshold.png'), width = 6, height = 3.5)
+ggsave(file.path(dir_fig, 'ribbon_AMThreshold.png'), 
+       width = 17.8, height = 6, units = 'cm')
 

### 2016-01-25 manuscript-main.Rmd
updated figures; added word count; etc

@@
 title: "Aligning species range data to better serve science and conservation"
-author: "Casey O'Hara^1^, Jamie Afflerbach^1^, Benjamin S. Halpern^1, 2, 3^"
+author: "Casey O'Hara^1^, Jamie Afflerbach^1^, Benjamin S. Halpern<sup>1, 2, 3</sup>"
 output:
-  word_document: default
-  pdf_document:
-    number_sections: yes
-    toc: no
   html_document:
@@
     toc: no
+  pdf_document:
+    number_sections: yes
+    toc: no
+  word_document: default
 ---
@@
 
-_Classification:_
-
-* BIOLOGICAL SCIENCES: 
-    * Ecology? Environmental Sciences? Sustainability Science?
-
 _Author affiliation:_ <!--Include department, institution, and complete address, with the ZIP/postal code, for each author.
@@
 
-1. National Center for Ecological Analysis and Synthesis, 735 State Street, Suite 300, Santa Barbara CA 93101
+1. National Center for Ecological Analysis and Synthesis, 735 State Street Suite 300, Santa Barbara CA 93101
 2. Bren School of Environmental Science and Management, University of California, Santa Barbara CA 93106
-3. Imperial College London, Silwood Park Campus, Ascot, United Kingdom (need address?)
+3. Department of Life Sciences, Imperial College London, Silwood Park Campus, Buckhurst Rd, Ascot, West Berkshire SL5 7PY, United Kingdom (ZIP/postal code?)
 
@@
 Email addresses will appear in the article footnotes.-->
+Casey O'Hara, National Center for Ecological Analysis and Synthesis, 735 State Street, Suite 300, Santa Barbara CA 93101; _NCEAS phone #?
+or one that's actually likely to reach me?_; ohara@nceas.ucsb.edu
 
-* Casey O'Hara, National Center for Ecological Analysis and Synthesis, 735 State Street, Suite 300, Santa Barbara CA 93101; _NCEAS phone #?_; ohara@nceas.ucsb.edu
+_Classification:_ BIOLOGICAL SCIENCES: Ecology? Environmental Sciences? Sustainability Science?
 
-_Keywords:_ <!--Keywords are listed below the abstract of the published article. At least three keywords are required at submission.--> TBD
+_Keywords:_ <!--Keywords are listed below the abstract of the published article.
+At least three keywords are required at submission.
+maybe species distribution maps, AquaMaps, IUCN, marine protected area planning, biodiversity status--> TBD
 

### 2016-01-25 ms_2_intro_res_disc.Rmd
updated figures; added word count; etc

@@
 ```
+
+| Section                   |   Word #   | Char # (w/spc) |
+| :------                   |    ------: |    ------: |
+| Introduction              |  512 words |  3411 char |
+| Results/Discussion        |  802 words |  5410 char |
+| Implications (incomplete) |  532 words |  3334 char |
+| Conclusions               |  390 words |  2813 char |
+| Methods (incomplete)      |  612 words |  3889 char |
+| total body                | 2848 words | 18857 char |
 

### 2016-01-25 data_explore/data_explore_fxns.R

@@ -7,7 +7,7 @@ ggtheme_basic <- theme(axis.ticks = element_blank(),
         legend.position = 'right')
 
 ggtheme_plot <- ggtheme_basic + 
-  theme(panel.border = element_blank(),
+  theme(panel.border     = element_blank(),
         panel.grid.minor = element_blank(), 
         panel.grid.major = element_line(colour = 'grey90'),
         panel.background = element_blank(),

### 2016-01-25 ms_2_intro_res_disc.Rmd
commented up draft for Ben, organized the drafts folder a bit

@@
 These differences can be mitigated by using both datasets, but the underlying methodological differences complicate such direct comparisons.
-_and furthermore requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species._
+<!--_and furthermore requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species._-->
 
@@
 
-_Predictions:  What happens when using a 0% threshold instead of a 50%?
-What happens when using IUCN rather than AM?
-IUCN tends to overestimate extent esp for coastal species?
-More species will be included in MPAs so fewer apparent gap species.
-Included range area inside MPAs will increase.
-Policy implications?_
+_Predictions: What happens when using a 0% threshold instead of a 50%? What happens when using IUCN rather than AM?_
+* _IUCN tends to overestimate extent esp for coastal species?  How much of this area will fall within MPAs and how much will fall outside, proportionally? this affects protection ratio._
+* _More area means more species (relatively) will be at least partially included in MPAs so fewer apparent gap species?_
+* _Policy implications?  Can also tie this back to the predictions of Rondinini et al re: effects of commission/omission errors_
 

### 2016-01-31 manuscript-main.Rmd
updating manuscript based on Ben's comments

@@
 ---
-title: "Aligning species range data to better serve science and conservation"
+title: "Aligning marine species range data to better serve science and conservation"
 author: "Casey O'Hara^1^, Jamie Afflerbach^1^, Benjamin S. Halpern<sup>1, 2, 3</sup>"
@@
 2. Bren School of Environmental Science and Management, University of California, Santa Barbara CA 93106
-3. Department of Life Sciences, Imperial College London, Silwood Park Campus, Buckhurst Rd, Ascot, West Berkshire SL5 7PY, United Kingdom (ZIP/postal code?)
+3. Department of Life Sciences, Imperial College London, Silwood Park Campus, Buckhurst Rd, Ascot, West Berkshire SL5 7PY, United Kingdom
 
@@
 Email addresses will appear in the article footnotes.-->
-Casey O'Hara, National Center for Ecological Analysis and Synthesis, 735 State Street, Suite 300, Santa Barbara CA 93101; _NCEAS phone #?
-or one that's actually likely to reach me?_; ohara@nceas.ucsb.edu
+Casey O'Hara, National Center for Ecological Analysis and Synthesis, 735 State Street, Suite 300, Santa Barbara CA 93101; (805) 892-2500; ohara@nceas.ucsb.edu
 

### 2016-01-31 ms_1_abst_sig.Rmd
updated figures; added word count; etc
updating manuscript based on Ben's comments

@@
 Species distribution data provide the foundation for a wide range of ecological research and conservation science and management, yet most species ranges remain unknown, and existing range maps have little overlap.
-In the ocean, two global species distribution datasets - AquaMaps and the International Union for Conservation of Nature (IUCN) - dominate our understanding of marine species ranges throughout the world's oceans, representing `r format(nrow(spp_all), big.mark = ',')` species overall, mostly from AquaMaps, with only `r format(nrow(spp_map_pairs), big.mark = ',')` species overlapping.
+In the ocean, two global species distribution datasets, produced by AquaMaps and the International Union for Conservation of Nature (IUCN), dominate our understanding of marine species ranges throughout the world's oceans.
+Together they represent `r format(nrow(spp_all), big.mark = ',')` species, mostly from AquaMaps, with only `r format(nrow(spp_map_pairs), big.mark = ',')` species overlapping.
 Here we examine differences in predicted species ranges between the two datasets, propose mechanistic causes and potential solutions for such differences, and explore the implications of these differences for management and conservation decisions.
 We find that IUCN maps often disregard bathymetry for depth-limited species, leading to predictions of species presence at unsuitable depths, and AquaMaps ranges for data-poor species often extrapolate presence far afield from known occurrences.
-We illustrate the implications of these differences by repeating two recent applications - the Ocean Health Index biodiversity goal and a global analysis of gaps in coverage of marine protected areas - and find significantly different estimates of the status of species diversity and effectiveness of conservation depending on which dataset was used.
-Understanding these issues points toward solutions for providers and users of these datasets.
+We illustrate the implications of these differences by repeating two recent applications - the Ocean Health Index biodiversity goal and a global analysis of gaps in coverage of marine protected areas - and find significantly different estimates of the status of global biodiversity and effectiveness of conservation depending on which dataset was used.
+___Creating a single, highly accurate dataset of global marine species ranges will be difficult.
+Understanding the implications of dataset differences for conservation planning and decision-making is essential.___
+<!--Understanding these issues points toward solutions for providers and users of these datasets.-->
+
+(223 words; 1593 char with spaces)
 

### 2016-01-31 ms_2_intro_res_disc.Rmd
updating manuscript based on Ben's comments

@@
 
-Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean - and you will find a unique ecosystem defined as much by which species are absent as by which species are abundant.
-Knowing where individuals of a species thrive is fundamental to the sciences of ecology, biogeography, and conservation, among many others.
-This knowledge provides foundational information for
+<!--Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean - and you will find a unique ecosystem defined as much by which species are absent as by which species are abundant.-->
+Knowing where individuals of a species live is fundamental to the sciences of ecology, biogeography, and conservation, among many others.  This knowledge provides foundational information for 
  understanding species ranges and diversity,
@@
 One major outcome of this body of science is the various compiled databases of species distribution maps.
-Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps modeled species distribution maps (Kaschner et al.
-2013) and International Union for Conservation of Nature (IUCN) species range maps (REF).
-These two spatial datasets are used for a wide range of purposes, including assessing marine species status (Halpern et al.
-2012, Selig et al.
-2013), evaluating global biodiversity patterns (Coll et al.
-2010, Martin et al.
-2014), predicting range shifts (Molinos et al.
-2015), and setting conservation priorities (Klein et al.
-2015).
+Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps (Kaschner et al.
+2013) and International Union for Conservation of Nature (IUCN) (REF).
+These two spatial datasets have been used for a wide range of purposes, including assessing marine species status (Halpern et al.
+2012, 2015 _both_; Selig et al.
+2013 _both_), evaluating global biodiversity patterns (Coll et al.
+2010 _AM_; Martin et al.
+2014 _both and others_, Pimm 2014 _IUCN, mostly focusing on terrestrial but some marine_), predicting range shifts (Molinos et al.
+2015 _AquaMaps_, and setting conservation priorities (Klein et al.
+2015 _AquaMaps_).
+<!-- BSH: Any of these use IUCN?
+Would be good to make sure at least 2-3 examples were based on IUCN.
+___none of these exclusively use IUCN - other suggestions?___-->
 
@@
 
-To understand the implications of differences between the AquaMaps and IUCN datasets, we compared how each data source represents the global spatial and taxonomic distribution of species.
+To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of species.
 Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` species).
-For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examined how well the maps align and determined several issues that lead to misalignment between predicted species distributions, and outline possible improvements.
+For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and outline possible improvements.
+<!--BSH: I think Jamie makes this comment below, but we may need to be able to substantiate this somehow.
+Is there a way to quickly figure out how many studies have used AM vs.
+IUCN?-->
 
-We then subjected two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (Halpern et al.
+We then subject two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (Halpern et al.
 2012, 2015) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (Klein et al.
@@
 
-The IUCN publishes species range maps as polygon shapefiles, bundled by taxonomic groups (REF).
-Using a geographic information system (GIS), experts outline spatial polygons to represent a given species' "limits of distribution" (REF IUCN 2014) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
-AquaMaps models species distribution based environmental preferences (e.g.
+The IUCN publishes species range maps, bundled by taxonomic groups (REF), based on expert input on spatial boundaries of a given species' "limits of distribution" (REF IUCN 2014) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
+In contrast, AquaMaps models species distribution based environmental preferences (e.g.
 temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
@@
 Each type of error bears different implications for conservation goals (Rondinini, 2006).
+<!-- This study basically predicted some of the problems we’re seeing here and predicts implications; I’d like to meaningfully acknowledge it, but not sure how deeply to get into it.
+Later comments ask about whether to include the omission/commission terminology.
+
+BSH: I haven’t read the paper, but it sounds like we need to acknowledge what was already discussed/done there.
+This may actually simplify things, such that we condense the first three paragraphs of this section into a single paragraph that restates what that paper found (and cite it) and add anything else we want to point out.
+But we will definitely need a succinct statement about how our paper moves well beyond what that paper did.
+Not sure best approach without reading the paper.-->
 
@@
 <!--_and furthermore requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species._-->
-
------
 
@@
 n_q4 <- nrow(spp_list_quads %>% filter(quad == 'q4'))
+
+n_q4_datapoor <- nrow(spp_list_quads %>% filter(quad == 'q4' & occurcells < 10))
+p_q4_datapoor <- round(n_q4_datapoor/n_q4 * 100, 1)
+n_q123_datapoor <- nrow(spp_list_quads %>% filter(quad != 'q4' & occurcells < 10))
+p_q123_datapoor <- round(n_q123_datapoor/n_spp_both * 100, 1)
+
+n_q4_rev <- nrow(spp_list_quads %>% filter(quad == 'q4' & !is.na(reviewed)))
+p_q4_rev <- round(n_q4_rev/n_q4 * 100, 1)
+n_q123_rev <- nrow(spp_list_quads %>% filter(quad != 'q4' &  !is.na(reviewed)))
+p_q123_rev <- round(n_q123_rev/n_spp_both * 100, 1)
 ```
@@
 
-Dividing the map-paired species into quadrants highlights different categories of relationships.
-The upper right quadrant includes species (n = `r n_q1`) whose map pairs agree in both spatial distribution and extent of described ranges.
-These species tend to be well-studied and predominantly fall within wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig.
+<!--BSH: I think it’s stronger to start with the negative relationship overall, then talk about the 4 quadrants.  So I rearranged, in turn requiring a bit of editing-->
+The negative linear relationship <!--BSH: Any stats for this?
+If so, insert just after this phrase.
+Otherwise maybe phrase it as 'The general pattern that emerges of a negative linear relationship, where increasing similarity...'--> indicates that increasing similarity in extent correlates with decreasing distribution alignment.
+<!--BSH: Could also be said as increasing similarity in alignment leads to decreasing similarity in extent.
+The phrasing implies causality, even though you are careful to use ‘correlates’--> AquaMaps tends to extrapolate species ranges beyond known occurrences, such that the marginal range predicted by AquaMaps will fall in different locations than the marginal range predicted using IUCN methodology.
+For species with dissimilar extents, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
+For species with increasingly similar extents, differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.
+
+Dividing the map-paired species into quadrants highlights different categories of relationships that in turn help further explain the general pattern.
+The upper right quadrant includes species (n = `r n_q1`) whose described ranges agree in both spatial distribution and extent.
+These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig.
 2B).
@@
 
-The extent-mismatched maps contained in the upper left quadrant (n = `r n_q2`) provide a richer set of examples to understand the fundamental differences between the datasets.
-For many of these species, while the spatial distribution is similar, the IUCN range includes a significant buffer zone beyond the AquaMaps range.
-Corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant).
-Most corals and reef-associated organisms prefer shallower waters, and we suspect that many of these disagreements in extent can be explained by bathymetry: seafloor depth is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps, leading to overprediction of species range into areas too deep to support the species (For examples and details, see SOM).
+The extent-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where the IUCN range is notably larger, often extending into deeper water.
+For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their depth tolerance.
+<!-- BSH: Worth a few spot checks to confirm this is more or less true.
+And then maybe include 2-3 examples as figures in the SOM and refer to them here.-->
+Ocean depth is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps.
+_BSH: Simply clipping IUCN range maps to known depth preferences would resolve many of these mismatches._ <!--BSH: This is kinda redundant, just emphasizing a solution to the problem.
+Not sure it's needed, or perhaps could be woven in to previous sentence.--> _commission error terminology?_
+
+Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence.
+Consequently, area ratios are close to 100%, but similar extents are meaningless when the distributions are poorly aligned.
 _commission error terminology?_
 
-The differences in extent of species maps found in the lower right quadrant (n = `r n_q3`) can in many cases be attributed to "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence.
-This results in an area ratio close to 100%, but similar extents are meaningless when the distributions are poorly aligned.
-_commission error terminology?_
+The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match.
+Often these mismatches arise when AquaMaps relies on limited data to predict species presence well beyond known occurrences.
+Indeed, `r p_q4_datapoor`% (n = `r n_q4_datapoor`) of these cases are data-poor (in which the AquaMaps environmental envelope model is based on fewer than 10 known cell occurrences), compared with `r p_q123_datapoor`% for species in the other three quadrants.
+Only `r p_q4_rev`% (n = `r n_q4_rev`) of AquaMaps maps in this quadrant have undergone expert review, compared with `r p_q123_rev`% of species in the other three quadrants.
+<!--BSH: This implies the problem is solely with AM.
+Are there no issues with IUCN maps that cause problems in this quadrant?--> ___The data-poor condition likely applies to IUCN as well - AM using OBIS points to generate models, and IUCN is probably doing the same.
+Check, and rework.
+This means AM is creating errors of commission, while IUCN may be committing errors of omission for poorly-studied species and/or regions - compounding the total problem.___
 
-Moving from the lower right to lower left quadrant, AquaMaps more frequently relies on limited data to predict species presence well beyond known occurrences, resulting in poor matches in both distribution and extent.
-Of the `r n_q4` species in this quadrant, 159 (39.7%) are data-poor (in which the AquaMaps environmental envelope model is based on fewer than 10 known cell occurrences), compared with 21.1% for the full set of map-paired species and only 8.2% of species in the upper left quadrant.
-A mere 6.2% of AquaMaps maps in this quadrant have undergone expert review, compared with 13.4% of all map-paired species and 24.9% of species in the upper left quadrant.
-
-This breakdown by quadrant helps explain the negative linear relationship seen in the plot, in which increasing similarity in extent correlates with decreasing distribution alignment.
+<!-- OLD: This breakdown by quadrant helps explain the negative linear relationship seen in the plot, in which increasing similarity in extent correlates with decreasing distribution alignment.
 When extrapolating species range beyond known occurrences, the marginal range predicted by the AquaMaps model will fall in different locations than the marginal range predicted using IUCN methodology.
 For species with dissimilar extents, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
-As we examine species with increasingly similar extents, the differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.
+As we examine species with increasingly similar extents, the differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.-->
 
@@
 
-**Case Study: The Ocean Health Index**
+<!--BSH: I think the first bullet point can/should go in the Intro (very briefly!) - I've noted where in my comments to that section.
+I do agree there needs to be a tiny bit of explanation of how OHI uses the two datasets.-->
+
+**Case Study: The Ocean Health Index** <!--BSH: I stopped editing and worksmithing here and just focused on addressing your comments/questions from here on-->
 
@@
 _Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted.
-When IUCN range maps are the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change._
+When IUCN range maps are the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change._ <!--CCO: This seems overly specific and I think could probably be cut, but the score changes weren’t earth-shattering – and that’s because we are buffering it by using both datasets
+BSH: I would maybe reframe this as 'given that only 2166 species overlap, score changes are not likely to be large for OHI and yet we still see important differences .
+.
+.
+" (remember that at the global scale, 2-3 points is a big deal)-->
 

### 2016-01-31 data_explore/data_prep.Rmd
tracking down the discrepancies in species numbers

@@
 
-```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+```{r get_spp_all, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 ### get spp_all_complete from git-annex, bring to local repo.  The spp_all_complete
@@
 spp_all_remote <- read.csv(file.path(dir_anx, scenario, 'int/spp_all.csv'), 
-                           stringsAsFactors = FALSE) %>%
+                           stringsAsFactors = FALSE) 
+spp_all_remote_clean <- spp_all_remote %>%
   filter(!str_detect(spatial_source, 'subpop')) %>%
-  filter(!str_detect(spatial_source,  'alias')) %>%
-  select(am_sid, sciname, reviewed, iucn_sid, popn_trend,
-         popn_category, spp_group, id_no, spatial_source)
+  filter(!str_detect(spatial_source,  'alias'))
+### At this point, plenty of duplicates:
+### * This code returns 518 duplicated names:
+###     length(spp_all_remote_clean$sciname) - length(unique(spp_all_remote_clean$sciname))
+### * some are AM subpops which were ignored in processing the remote spp_all file.
+###   Note: all subpops have an identified parent, so can filter them out without
+###   losing any species.
+###   code to verify this:
+###     z_parent <- spp_all_remote_clean %>% 
+###       filter(!is.na(subpop_sid) & spatial_source == 'am')
+###     z_subpop <- spp_all_remote_clean %>% 
+###       filter(!is.na(parent_sid) & spatial_source == 'am')
+###     nrow(z_subpop %>% filter(!sciname %in% z_parent$sciname))
+### * some are duplicated IUCN parents (e.g. Dermochelys coriacea) that will be 
+###   eliminated when select(-iucn_subpop) column and then unique()
 
-write.csv(spp_all_remote, file.path(dir_data, 'spp_all.csv'), row.names = FALSE)
+### filter out spatial_source == 'am' that also have parent_sid (i.e.
+### these are AM-mapped subpopulations); then: 
+### select away the iucn_subpop, parent_sid, and subpop_sid (among others),
+### these are creating false unique rows.
+spp_all <- spp_all_remote_clean %>%
+  filter(!(spatial_source == 'am' & !is.na(parent_sid))) %>%
+  dplyr::select(am_sid, sciname, reviewed, iucn_sid, popn_trend,
+         popn_category, spp_group, id_no, spatial_source) %>%
+  unique()
+### At this point, still some duped scinames - these are ones with multiple IUCN
+### SIDs that are not designated as parents or subpops.  We will brute-force
+### this by just eliminating the duplicated rows, keeping only the first instance.
+### By arranging by iucn_sid, we will keep the lowest value, which seems to
+### generally correspond to the parent...
+spp_all <- spp_all %>%
+  arrange(iucn_sid)
+spp_all <- spp_all[!duplicated(spp_all$sciname), ]
 
+length(spp_all$sciname) - length(unique(spp_all_remote_clean$sciname))
+### no duped scinames - same length as original (cleaned) list
+
+write.csv(spp_all, file.path(dir_data, 'spp_all.csv'), row.names = FALSE)
+
+```
+
+```{r get_other_vars, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 ### create spp_am_v_iucn: the whole maplist narrowed down to just the overlapping species.
@@
 ### that also have a non-NA AquaMaps ID code.
-spp_am_v_iucn <- spp_all_remote %>%
+spp_all <- read.csv(file.path(dir_data, 'spp_all.csv'), stringsAsFactors = FALSE)
+
+spp_am_v_iucn <- spp_all %>%
   filter(str_detect(spatial_source, 'iucn') & !is.na(am_sid)) %>%
@@
 spp_iucn_marine <- spp_iucn_marine %>%
-  select(sciname, iucn_sid, iucn_category = category, popn_trend, parent_sid, subpop_sid)
+  dplyr::select(sciname, iucn_sid, iucn_category = category, 
+                popn_trend, parent_sid, subpop_sid) %>%
+  unique()
 write.csv(spp_iucn_marine, file.path(dir_data, 'spp_iucn_marine.csv'), 
@@
 spp_iucn_maps <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv'), 
-                            stringsAsFactors = FALSE)
+                            stringsAsFactors = FALSE) %>%
+  rename(iucn_sid = id_no)
+
 spp_iucn_maps <- spp_iucn_maps %>%
   filter(!str_detect(spp_group, 'MARINEFISH')) %>%
-  filter(id_no %in% spp_iucn_marine$iucn_sid)   ### ditch terrestrial and freshwater spp
+  filter(iucn_sid %in% spp_iucn_marine$iucn_sid) %>% ### ditch terrestrial and freshwater spp
+  unique() %>%
+  arrange(iucn_sid)
+spp_iucn_maps <- spp_iucn_maps[!duplicated(spp_iucn_maps$sciname), ]
 write.csv(spp_iucn_maps, file.path(dir_data, 'spp_iucn_maps.csv'), 

### 2016-01-31 manuscript-main.Rmd
tracking down the discrepancies in species numbers

@@
 
+if(basename(getwd()) != 'IUCN-AquaMaps') setwd('~/github/IUCN-AquaMaps')
+
 ```
@@
 spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(popn_category)) %>% ### filter out unassessed species
+#  filter(!is.na(popn_category)) %>% ### filter out unassessed species
   filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)

### 2016-02-01 data_explore/data_prep.Rmd
small changes

@@
 spp_list <- read.csv(data_file, stringsAsFactors = FALSE) 
-### At this point, spp_list contains subpops and aliases, and zero-area IUCNs
+### At this point, spp_list zero-area IUCNs due to missing (?) polygons (maybe points?)
 
@@
 spp_list <- spp_list %>%
-  filter(!is.na(sm_perc)) %>%
+  mutate(sm_perc = ifelse(!is.na(sm_perc), 0, sm_perc))
+  # filter(!is.na(sm_perc)) %>%
   ### will ditch the ones where no IUCN polygon exists - or are these species
   ### whose ranges only occur in cells beyond the LOICZID grid?
+spp_list <- spp_list %>%
   filter(!str_detect(spatial_source, 'subpop') & !str_detect(spatial_source, 'alias')) 

### 2016-02-01 data_explore/data_prep.Rmd
updating species list (spp_all.csv) with missed or dropped species

--- 2016-01-31 data_explore/data_prep.Rmd
@@
 gather spp_all, spp_am, spp_am_v_iucn, spp_list_w_area_all and _trimmed into data folder.
+
+``` {r pull_files_to_local, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+### pull spp_iucn_marine.csv from git-annex and bring to local repo
+spp_iucn_marine <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_marine.csv'), 
+                            stringsAsFactors = FALSE)
+spp_iucn_marine <- spp_iucn_marine %>%
+  dplyr::select(sciname, iucn_sid, iucn_category = category, 
+                popn_trend, parent_sid, subpop_sid) %>%
+  unique()
+write.csv(spp_iucn_marine, file.path(dir_data, 'spp_iucn_marine.csv'), 
+          row.names = FALSE)
+
+### pull spp_iucn_maps_all.csv from git-annex and bring to local repo
+spp_iucn_maps <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv'), 
+                            stringsAsFactors = FALSE) %>%
+  rename(iucn_sid = id_no)
+
+spp_iucn_maps <- spp_iucn_maps %>%
+  filter(!str_detect(spp_group, 'MARINEFISH')) %>%
+  filter(iucn_sid %in% spp_iucn_marine$iucn_sid) %>% ### ditch terrestrial and freshwater spp
+  unique() %>%
+  arrange(iucn_sid)
+spp_iucn_maps <- spp_iucn_maps[!duplicated(spp_iucn_maps$sciname), ]
+write.csv(spp_iucn_maps, file.path(dir_data, 'spp_iucn_maps.csv'), 
+          row.names = FALSE)
+
+### pull speciesoccursum.csv from git-annex and bring to local repo
+spp_am <- read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
+                   stringsAsFactors = FALSE)
+spp_am <- spp_am %>%
+  mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' '))
+write.csv(spp_am, file.path(dir_data, 'spp_am.csv'), 
+          row.names = FALSE)
+
+```
 
@@
 
+### But, some of the extracted IUCN maps don't appear on this list! add 'em in.
+spp_iucn_marine   <- read.csv(file.path(dir_data, 'spp_iucn_marine.csv'), 
+                              stringsAsFactors = FALSE)
+spp_iucn_maps_new <- read.csv(file.path(dir_data, 'spp_iucn_maps.csv'), 
+                              stringsAsFactors = FALSE) %>%
+  filter(!sciname %in% spp_all$sciname) %>%
+  filter(!str_detect(sciname, 'subpop')) %>% ### a Hawaiian subpop... ditch it!
+  select(-iucn_subpop)
+spp_all1 <- spp_all %>%
+  full_join(spp_iucn_maps_new)
+
+### Also, some of the AM listings have disappeared due to name conflicts
+### - aliases, etc.
+spp_am_new <- read.csv(file.path(dir_data, 'spp_am.csv'), 
+                       stringsAsFactors = FALSE) %>%
+  rename(am_sid = speciesid) %>%
+  filter(!am_sid %in% spp_all$am_sid) %>%
+  dplyr::select(am_sid, sciname, reviewed, iucn_sid = iucn_id,
+                popn_category = iucn_code) %>%
+  mutate(spatial_source = 'am',
+         iucn_sid      = as.integer(iucn_sid),
+         popn_category = ifelse(popn_category == 'LR/lc', 'LC', popn_category),
+         popn_category = ifelse(popn_category == 'LR/nt', 'NT', popn_category),
+         popn_category = ifelse(popn_category == 'N.E.',  'NE', popn_category))
+
+### 600 new spp_am not in spp_all from remote source, by am_sid.
+### 12 of these, however, list iucn_sids that match existing iucn_sid.
+### * for species with iucn_sid that match, and no am_sid, add am_sid to 
+###   existing record (use existing values for other fields; drop new values).
+### * for species with iucn_sid that match, and with a different am_sid, no action.
+### * for species with non-matching or NA iucn_sid, just add to list.
+spp_am_new_iucnmatch <- spp_am_new %>% 
+  filter(!is.na(iucn_sid) & iucn_sid %in% spp_all$iucn_sid) %>%
+  dplyr::select(iucn_sid, am_sid_new = am_sid, reviewed_new = reviewed)
+spp_all2 <- spp_all1 %>%
+  left_join(spp_am_new_iucnmatch, by = 'iucn_sid') %>%
+  mutate(am_sid   = ifelse(is.na(am_sid) & !is.na(am_sid_new), am_sid_new, am_sid),
+         reviewed = ifelse(is.na(am_sid) & !is.na(am_sid_new), reviewed_new, reviewed)) %>%
+  select(-am_sid_new, -reviewed_new)
+spp_all <- spp_all2 %>%
+  full_join(spp_am_new %>%
+              filter(!iucn_sid %in% spp_am_new_iucnmatch$iucn_sid),
+            by = c("am_sid", "sciname", "reviewed", "iucn_sid", "popn_category", "spatial_source"))
+
 write.csv(spp_all, file.path(dir_data, 'spp_all.csv'), row.names = FALSE)
@@
           row.names = FALSE)
+```
 
+``` {r pull_files_to_local, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 ### pull spp_iucn_marine.csv from git-annex and bring to local repo

### 2016-02-01 data_explore/data_prep.Rmd
updating species area calculations with the latest and greatest species list

@@
 
-``` {r pull_files_to_local, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
-### pull spp_iucn_marine.csv from git-annex and bring to local repo
-spp_iucn_marine <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_marine.csv'), 
-                            stringsAsFactors = FALSE)
-spp_iucn_marine <- spp_iucn_marine %>%
-  dplyr::select(sciname, iucn_sid, iucn_category = category, 
-                popn_trend, parent_sid, subpop_sid) %>%
-  unique()
-write.csv(spp_iucn_marine, file.path(dir_data, 'spp_iucn_marine.csv'), 
-          row.names = FALSE)
-
-### pull spp_iucn_maps_all.csv from git-annex and bring to local repo
-spp_iucn_maps <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv'), 
-                            stringsAsFactors = FALSE) %>%
-  rename(iucn_sid = id_no)
-
-spp_iucn_maps <- spp_iucn_maps %>%
-  filter(!str_detect(spp_group, 'MARINEFISH')) %>%
-  filter(iucn_sid %in% spp_iucn_marine$iucn_sid) %>% ### ditch terrestrial and freshwater spp
-  unique() %>%
-  arrange(iucn_sid)
-spp_iucn_maps <- spp_iucn_maps[!duplicated(spp_iucn_maps$sciname), ]
-write.csv(spp_iucn_maps, file.path(dir_data, 'spp_iucn_maps.csv'), 
-          row.names = FALSE)
-
-### pull speciesoccursum.csv from git-annex and bring to local repo
-spp_am <- read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
-                   stringsAsFactors = FALSE)
-spp_am <- spp_am %>%
-  mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' '))
-write.csv(spp_am, file.path(dir_data, 'spp_am.csv'), 
-          row.names = FALSE)
-
-```
-
 ```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
@@
 spp_list <- spp_list %>%
-  filter(!str_detect(spatial_source, 'subpop') & !str_detect(spatial_source, 'alias')) 
-spp_list <- spp_list %>%
   left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 

### 2016-02-01 manuscript-main.Rmd
updating species area calculations with the latest and greatest species list

@@
 data_file <- 'data/spp_list_w_area_trimmed.csv'
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-#  filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
+spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) 
+### NOTE: as of Jan 31, includes all map pairs, even ones where IUCN area == 0
 
 ### generate other lists
-
 spp_all_file <- 'data/spp_all.csv'
@@
 spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
-  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
-  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
+  filter(spp_group != 'hagfishes')           ### not from IUCN...
 

### 2016-02-01 manuscript-main.Rmd
notes re: missing map pairs...

--- 2016-01-31 manuscript-main.Rmd
@@
 data_file <- 'data/spp_list_w_area_trimmed.csv'
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-#  filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
+spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE)
+### NOTE: as of 1/31, there are supposedly 2246 paired maps; but
+### only 2003 unique IUCN species names in the maps list - presumably 
+### the remainder are NA area (no polygons in LOICZID area)
 

### 2016-02-01 manuscript-main.Rmd

--- 2016-02-01 manuscript-main.Rmd
@@
 spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE)
+spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) 
 ### NOTE: as of Jan 31, includes all map pairs, even ones where IUCN area == 0

### 2016-02-05 ms_3_methods.Rmd

@@
 
-* Possible analyses:
+* Possible analyses: <!--BSH: priorities are 1 (AM only), 3 (IUCN only) and 4 (AM + IUCN).-->
     * run AquaMaps (new data) with threshold at 0%
         * already done in original paper for old data, but redo it to compare to AM + IUCN scenario below
-    * run with AquaMaps only, but filtering out data-poor species?
-    * run with IUCN maps only in place of AquaMaps?
+    * run with IUCN maps only in place of AquaMaps? _<!--BSH: this seems a bit silly, since AM has so many more species, but worth doing to make the point, I think.-->_
     * run with AquaMaps and IUCN both, using 0% threshold; which preferred?
-        * Should we clip corals (and coral-associated taxa) to 200 meter bathymetry, just for fun?
-* Which aspects of the analysis should we focus on?:
+_<!--BSH: I think this would be good to do, with AM as priority.
+We can then compare to AM only, as you note above.
+This tells us exactly what is gained by including the additional 1500 species (or whatever) that are only in IUCN.-->_
+* Which aspects of the analysis should we focus on?: _<!--BSH: I’m guessing it will be hard to see differences in the maps (except for the IUCN only example), so probably best to focus on figure 1 presentation of results.-->_
     * overall vs broken down by taxonomic groups? (see paper, figure 1)
-    * spatial distribution of gap species (see paper, figure 2)
+    * spatial distribution of gap species (see paper, figure 2) <!--

### 2016-02-05 ms_4_figs.Rmd
updated figures; added word count; etc

@@
 <!--
+
+Figures: Calculated at 180 characters per cm in height for one column and 360 characters per cm in height for two columns; a figure wider than 8.7 cm is two columns wide.
 
@@
 
-### Fig. 1
-
+### Fig. 1 <!--can you put N = XXX above each bar? Would make it a little easier to read compared to within the bar.-->
 <a id = 'fig1'></a>
@@
 
-### Figure 2
+### Figure 2 <!--Ditto of Fig. 2B – can you move the N = XX values to be outside the bars (to the right) instead of within them?-->
 
@@
 
+_CCO: a lot of this is duplicated in body… tell the story more briefly here; BSH: Yes, make very brief, probably cutting most of this._
 > __Fig.
@@
 _is it distracting to use order names for some groups and not others?
-Some of the IUCN taxonomic groups had pretty long names ("acanthuridae" is way shorter than "surgeonfish, tangs, and unicornfish")_
+Some of the IUCN taxonomic groups had pretty long names ("acanthuridae" is way shorter than "surgeonfish, tangs, and unicornfish")_ _BSH: No, but just be sure to note this difference in the legend._ ___NOTE: including the order names might imply the whole order is covered...
+not necessarily so?
+Use IUCN taxo groups!___
 
@@
 
-___NOTE: These figures still need to be recreated using the latest and greatest data!___ _currently they use data generated in November or so, in ohiprep._
+___NOTE: These figures still need to be recreated using the latest and greatest data!___ _currently they use data generated in November or so, in ohiprep._ | _BSH: It will be worth noting in the main text of the paper the range of values, i.e., that individual countries can see differences from -12 to +6 (depending on scenario).
+Those are pretty big differences within a country, even if the overall global number doesn’t change more than a few points._
 
@@
 
-> * Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
+> * Scenario 1 shows the effect of reducing the probability threshold for AquaMaps presence.
 Reducing the threshold increases the apparent range of a species; the slight decrease in average score suggests increased spatial representation of threatened species.
+_CCO: Reference how this more accurately tracks the definition of IUCN “limits of distribution”?
+BSH: I don’t think you need to directly reference this comparison, but these scenarios are helpful to understand what’s going on, so they’ll need some explanation/implications._
 > * Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.
@@
 The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
+_CCO: Too much detail here?
+Just get the main point – probably a mixture of more/less good/bad – can’t draw much of a conclusion other than that scores change all over.
+BSH: You’ll need a sentence or two explaining this result, but keep it brief.
+Extra text, if deemed useful, can be put into the SOM._
 > * Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, while eliminating the presence threshold.
@@
 The large decrease indicates that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.
+_similar note as scenario 2_

### 2016-02-06 ms_2_intro_res_disc.Rmd
blarg

@@
 Is there a way to quickly figure out how many studies have used AM vs.
-IUCN?-->
+IUCN?
+CCO: What about age - AquaMaps is pretty recent; before that, would IUCN have been the dominant dataset for marine?
+Maybe we can track citations of AquaMaps, but citations of IUCN I think are generic - just cite the Red List - not referencing spatial information specifically, much less marine.-->
 
@@
 These differences can be mitigated by using both datasets, but the underlying methodological differences complicate such direct comparisons.
-<!--_and furthermore requires resolving differences (or choosing one source over the other) for the `r format(n_spp_both, big.mark = ',')` shared species._-->
 
@@
 
-<!--BSH: I think it’s stronger to start with the negative relationship overall, then talk about the 4 quadrants.  So I rearranged, in turn requiring a bit of editing-->
 The negative linear relationship <!--BSH: Any stats for this?
@@
 The extent-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where the IUCN range is notably larger, often extending into deeper water.
-For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their depth tolerance.
+For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their preferred depths, likely introducing errors of commission.
 <!-- BSH: Worth a few spot checks to confirm this is more or less true.
@@
 _BSH: Simply clipping IUCN range maps to known depth preferences would resolve many of these mismatches._ <!--BSH: This is kinda redundant, just emphasizing a solution to the problem.
-Not sure it's needed, or perhaps could be woven in to previous sentence.--> _commission error terminology?_
+Not sure it's needed, or perhaps could be woven in to previous sentence.-->
 
-Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence.
-Consequently, area ratios are close to 100%, but similar extents are meaningless when the distributions are poorly aligned.
-_commission error terminology?_
+Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, introducing additional commission errors.
+Consequently, area ratios are close to 100%, though similar extents are meaningless when the distributions are poorly aligned.
 
-The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match.
+<!--The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match.
 Often these mismatches arise when AquaMaps relies on limited data to predict species presence well beyond known occurrences.
@@
 Check, and rework.
+This means AM is creating errors of commission, while IUCN may be committing errors of omission for poorly-studied species and/or regions - compounding the total problem.___-->
+
+The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match.
+When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences; at the same time, IUCN range maps .
+Indeed, `r p_q4_datapoor`% (n = `r n_q4_datapoor`) of these cases are data-poor (in which the AquaMaps environmental envelope model is based on fewer than 10 known cell occurrences ___is there a good way to get the number of OBIS observations for each of these species, as an external validation of data-poor?)___, compared with `r p_q123_datapoor`% for species in the other three quadrants.
+Only `r p_q4_rev`% (n = `r n_q4_rev`) of AquaMaps maps in this quadrant have undergone expert review, compared with `r p_q123_rev`% of species in the other three quadrants.
+<!--BSH: This implies the problem is solely with AM.
+Are there no issues with IUCN maps that cause problems in this quadrant?--> ___The data-poor condition likely applies to IUCN as well - AM using OBIS points to generate models, and IUCN is probably doing the same.
+Check, and rework.
 This means AM is creating errors of commission, while IUCN may be committing errors of omission for poorly-studied species and/or regions - compounding the total problem.___
-
-<!-- OLD: This breakdown by quadrant helps explain the negative linear relationship seen in the plot, in which increasing similarity in extent correlates with decreasing distribution alignment.
-When extrapolating species range beyond known occurrences, the marginal range predicted by the AquaMaps model will fall in different locations than the marginal range predicted using IUCN methodology.
-For species with dissimilar extents, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
-As we examine species with increasingly similar extents, the differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.-->
 

### 2016-02-11 data_explore/data_prep.Rmd
committing - so close to getting rid of all the friggin' duped species names

@@
 ### pull spp_iucn_marine.csv from git-annex and bring to local repo
-spp_iucn_marine <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_marine.csv'), 
+spp_iucn <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_cleaned.csv'), 
                             stringsAsFactors = FALSE)
-spp_iucn_marine <- spp_iucn_marine %>%
-  dplyr::select(sciname, iucn_sid, iucn_category = category, 
+spp_iucn <- spp_iucn %>%
+  dplyr::select(sciname, iucn_sid, iucn_category, 
                 popn_trend, parent_sid, subpop_sid) %>%
   unique()
-write.csv(spp_iucn_marine, file.path(dir_data, 'spp_iucn_marine.csv'), 
-          row.names = FALSE)
+write_csv(spp_iucn, file.path(dir_data, 'spp_iucn.csv'))
 
@@
   filter(!str_detect(spp_group, 'MARINEFISH')) %>%
-  filter(iucn_sid %in% spp_iucn_marine$iucn_sid) %>% ### ditch terrestrial and freshwater spp
   unique() %>%
   arrange(iucn_sid)
-spp_iucn_maps <- spp_iucn_maps[!duplicated(spp_iucn_maps$sciname), ]
-write.csv(spp_iucn_maps, file.path(dir_data, 'spp_iucn_maps.csv'), 
-          row.names = FALSE)
 
-### pull speciesoccursum.csv from git-annex and bring to local repo
-spp_am <- read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
+### At this point, still some duplicated iucn_sids, due to several
+### names matching a single ID.  Not sure which names might be in
+### AquaMaps though, so leave 'em in.
+
+write_csv(spp_iucn_maps, file.path(dir_data, 'spp_iucn_maps.csv'))
+
+### pull cleaned spp_am data from git-annex and bring to local repo
+spp_am <- read.csv(file.path(dir_anx, scenario, 'int/spp_am_cleaned.csv'), 
                    stringsAsFactors = FALSE)
-spp_am <- spp_am %>%
-  mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' '))
-write.csv(spp_am, file.path(dir_data, 'spp_am.csv'), 
-          row.names = FALSE)
+write_csv(spp_am, file.path(dir_data, 'spp_am.csv'))
+
+### pull raw aquamaps species list from git-annex and bring to local repo
+spp_am_raw <- read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
+                   stringsAsFactors = FALSE)
+write_csv(spp_am_raw, file.path(dir_data, 'speciesoccursum.csv'))
 
@@
 ### At this point, plenty of duplicates:
-### * This code returns 518 duplicated names:
+### * This code returns 984 duplicated names:
 ###     length(spp_all_remote_clean$sciname) - length(unique(spp_all_remote_clean$sciname))
 ### * some are AM subpops which were ignored in processing the remote spp_all file.
-###   Note: all subpops have an identified parent, so can filter them out without
+###   Note: All subpops have an identified parent, so can filter them out without
 ###   losing any species.
@@
   dplyr::select(am_sid, sciname, reviewed, iucn_sid, popn_trend,
-         popn_category, spp_group, id_no, spatial_source) %>%
+         popn_category, spp_group, id_no, spatial_source, name_verified) %>%
   unique()
@@
 ### But, some of the extracted IUCN maps don't appear on this list! add 'em in.
-spp_iucn_marine   <- read.csv(file.path(dir_data, 'spp_iucn_marine.csv'), 
-                              stringsAsFactors = FALSE)
 spp_iucn_maps_new <- read.csv(file.path(dir_data, 'spp_iucn_maps.csv'), 
@@
   select(-iucn_subpop)
+
 spp_all1 <- spp_all %>%
@@
 ### - aliases, etc.
-spp_am_new <- read.csv(file.path(dir_data, 'spp_am.csv'), 
+spp_am_new <- read.csv(file.path(dir_data, 'speciesoccursum.csv'), 
                        stringsAsFactors = FALSE) %>%
-  rename(am_sid = speciesid) %>%
+  mutate(sciname = paste(str_trim(genus), str_trim(species)),
+         iucn_sid = as.integer(iucn_id)) %>%
+  dplyr::select(am_sid = speciesid, sciname, reviewed, iucn_sid,
+                popn_category = iucn_code) %>%
   filter(!am_sid %in% spp_all$am_sid) %>%
-  dplyr::select(am_sid, sciname, reviewed, iucn_sid = iucn_id,
-                popn_category = iucn_code) %>%
   mutate(spatial_source = 'am',
-         iucn_sid      = as.integer(iucn_sid),
          popn_category = ifelse(popn_category == 'LR/lc', 'LC', popn_category),
@@
 
-### 600 new spp_am not in spp_all from remote source, by am_sid.
-### 12 of these, however, list iucn_sids that match existing iucn_sid.
+### 545 new spp_am not in spp_all from remote source, by am_sid.
+### 24 of these, however, list iucn_sids that match existing iucn_sid.
 ### * for species with iucn_sid that match, and no am_sid, add am_sid to 
@@
   select(-am_sid_new, -reviewed_new)
-spp_all <- spp_all2 %>%
+spp_all3 <- spp_all2 %>%
   full_join(spp_am_new %>%
@@
 
-write.csv(spp_all, file.path(dir_data, 'spp_all.csv'), row.names = FALSE)
+write_csv(spp_all3, file.path(dir_data, 'spp_all.csv'))
 
@@
   filter(str_detect(spatial_source, 'iucn') & !is.na(am_sid)) %>%
+  filter(!is.na(id_no)) %>%
   unique()
-write.csv(spp_am_v_iucn, file.path(dir_data, 'spp_am_v_iucn.csv'), 
-          row.names = FALSE)
+# length(unique(spp_am_v_iucn$am_sid)); length(unique(spp_am_v_iucn$iucn_sid))
+### NOTE: 2377 unique scientific names; but 2315 unique am_sid and 2280 unique iucn_sid.
+x <- spp_am_v_iucn[duplicated(spp_am_v_iucn$am_sid)   | duplicated(spp_am_v_iucn$am_sid,   fromLast = TRUE), ] %>%
+  select(sciname, iucn_sid, am_sid, id_no, spatial_source, name_verified)
+y <- spp_am_v_iucn[duplicated(spp_am_v_iucn$iucn_sid) | duplicated(spp_am_v_iucn$iucn_sid, fromLast = TRUE), ] %>%
+  select(sciname, iucn_sid, am_sid, id_no, spatial_source, name_verified)
+
+write_csv(spp_am_v_iucn, file.path(dir_data, 'spp_am_v_iucn.csv'))
 ```
@@
 ```
+
+``` {r get_counts, echo = FALSE}
+x <- occ_search(scientificName = spp_iucn_marine$sciname[1:100], fields = 'minimal', limit = 25)
+y <- sapply(x, function(x) x$meta$count)
+```

### 2016-02-11 data_explore/data_prep.Rmd
I think I have all the name matching issues worked out?

@@
   filter(!is.na(id_no)) %>%
+  filter(!str_detect(spp_group, 'MARINEFISH')) %>%
   unique()
 # length(unique(spp_am_v_iucn$am_sid)); length(unique(spp_am_v_iucn$iucn_sid))
-### NOTE: 2377 unique scientific names; but 2315 unique am_sid and 2280 unique iucn_sid.
-x <- spp_am_v_iucn[duplicated(spp_am_v_iucn$am_sid)   | duplicated(spp_am_v_iucn$am_sid,   fromLast = TRUE), ] %>%
-  select(sciname, iucn_sid, am_sid, id_no, spatial_source, name_verified)
-y <- spp_am_v_iucn[duplicated(spp_am_v_iucn$iucn_sid) | duplicated(spp_am_v_iucn$iucn_sid, fromLast = TRUE), ] %>%
-  select(sciname, iucn_sid, am_sid, id_no, spatial_source, name_verified)
+### NOTE: 2345 unique scientific names; but 2283 unique am_sid and 2249 unique iucn_sid.
+### Some unique species (as identified by am_sid and iucn_sid) have multiple
+### scinames.  At this point, not really important what the sciname is?
+spp_am_v_iucn1 <- spp_am_v_iucn %>%
+  group_by(am_sid, iucn_sid) %>%
+  mutate(dupes = n() - 1,
+         first_instance = sciname == first(sciname)) %>%
+  filter(!(dupes > 0 & !first_instance)) %>%
+  select(-dupes, -first_instance) %>%
+  ungroup()
+# length(unique(spp_am_v_iucn1$am_sid)); length(unique(spp_am_v_iucn1$iucn_sid))
 
-write_csv(spp_am_v_iucn, file.path(dir_data, 'spp_am_v_iucn.csv'))
+### What species are represented by the duplicated IUCN ID numbers?  These
+### are species that IUCN experts lumped together (via synonyms etc) 
+### but are treated as separate by AquaMaps (the reverse does not seem
+### to happen)...  There are 64 AM species IDs and only 30 IUCN species IDs.  
+### How to treat in the mapping?
+### * lump AM species together and plot against a single IUCN species map?
+###   That would occur in the area determining function - group by iucn_sid
+###   and map all am_sids against that iucn_sid
+x <- spp_am_v_iucn1 %>%
+  group_by(iucn_sid) %>%
+  filter(n() > 1)
+# length(unique(x$am_sid)); length(unique(x$iucn_sid))
+
+write_csv(spp_am_v_iucn1, file.path(dir_data, 'spp_am_v_iucn.csv'))
 ```
@@
   ### species info list - scinames, IUCN species ID, AM species ID, category etc
-  spp_list_base <- read.csv(file.path(dir_data, 'spp_am_v_iucn.csv'),
-                            stringsAsFactors = FALSE)
+  spp_list_base <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))
 

### 2016-02-11 manuscript-main.Rmd
committing so I can re-run area calcs

@@
 
-iucn_marine_file <- 'data/spp_iucn_marine.csv'
-spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
+# iucn_marine_file <- 'data/spp_iucn_marine.csv'
+# spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
 
 iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
-  filter(spp_group != 'hagfishes')           ### not from IUCN...
+spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE)
 

### 2016-02-12 data_explore/data_prep.Rmd
got the name crap figured out.  next steps: identify data-poor species etc

@@
 
-```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
+```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE}
+
+spp_am_v_iucn <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))
 
@@
 #iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
-                           col_types = 'cdddd', progress = TRUE) %>%
-  as.data.frame() %>%
-  filter(presence != 5) ### removes extinct subpopulations
+message('Building IUCN species to cell table.  This might take a few minutes.')
+iucn_map_files      <- file.path(dir_anx, 'iucn_intersections', list.files(file.path(dir_anx, 'iucn_intersections')))
+iucn_spp_cells <- lapply(iucn_map_files, data.table::fread) %>%
+  bind_rows()
+iucn_spp_cells <- iucn_spp_cells %>%
+  rename(iucn_sid = id_no, loiczid = LOICZID) %>%
+  mutate(iucn_sid = as.integer(iucn_sid),
+         presence = as.integer(presence),
+         loiczid  = as.integer(loiczid)) %>%
+  filter(iucn_sid %in% spp_am_v_iucn$iucn_sid) %>%
+  filter(presence != 5)
+### length(unique(iucn_spp_cells$iucn_sid)) results in 2013 species found...
+### some species with no cells reported.  2249 listed in the am v iucn file.
 
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), progress = TRUE)
+message('Loading AquaMaps species to cell table.  This might take a few minutes.')
+am_spp_cells <- fread(file.path(dir_data_remote, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv'), 
+                         showProgress = TRUE)
+am_spp_cells <- am_spp_cells %>%
+  rename(am_sid = speciesid, prob = probability) %>%
+  filter(am_sid %in% spp_am_v_iucn$am_sid)
+### length(unique(am_spp_cells$am_sid)) results in 2283 species found - great!
 
@@
 
-spp_am <- read.csv(file.path(dir_data, 'spp_am.csv'), stringsAsFactors = FALSE)
-
 spp_list <- spp_list %>%
@@
             by = 'spp_group') %>%
-  left_join(spp_am %>% select(am_sid = speciesid, occurcells),
+  left_join(read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
+                   stringsAsFactors = FALSE) %>% 
+              select(am_sid = speciesid, occurcells),
             by = 'am_sid')
 
-write.csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'), 
-          row.names = FALSE)
+write_csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
 

### 2016-02-12 data_explore/data_explore_fxns.R

@@ -24,218 +24,180 @@ ggtheme_plot <- ggtheme_basic +
 ggtheme_map <- ggtheme_basic +
   theme(axis.text = element_blank())
 
+#################################################################=
 ### Function to create a species list from scratch... use when new data is available
-create_spp_list <- function(spp_list_base, 
-                            data_file, 
-                            am_spp_cells, 
-                            iucn_spp_cells, 
-                            loiczid_raster) {
-
+create_spp_list <- function(spp_list_base, data_file, am_spp_cells, 
+                            iucn_spp_cells, loiczid_raster) {
   spp_list_build <- spp_list_base %>%
-    mutate(total_area = NA,
-           am_area    = NA,
-           iucn_area  = NA,
-           perc       = NA,
-           am_perc    = NA,
-           iucn_perc  = NA,
+    mutate(area_total = NA,  area_am = NA,  area_iucn = NA,
+           area_overlap = NA,
            sm_perc    = NA, #filled with % of smaller range that is within the larger range
            sm_range   = NA) #what range is smaller
   
+  i         <- 0 ### initialize the counting index
+  am_thresh <- 0 ### set the aquamaps presence threshold
+  sid_vector <- unique(spp_list_build$iucn_sid)
+  message(sprintf('There are a total of %s distinct species, by IUCN SID', 
+                  length(sid_vector)))
   
-  #there's likely a better way to do this than a for loop (lapply or sapply?)
-  for (i in 1:nrow(spp_list_build)){ # i = 3
-    spp     <- spp_list_build[i, ]$sciname #species name
-    message(sprintf('Species: %s | am_sid: %s | iucn_sid: %s', spp, spp_list_build[i, ]$am_sid, spp_list_build[i, ]$iucn_sid))
-    spp_map <- get_spp_map(spp, spp_list_build, am_spp_cells, iucn_spp_cells) # get species map
+  for (sid in sid_vector) { 
+    # sid <- spp_list_build$iucn_sid[1]
+    # sid <- 2058  sid <- 41711  sid <- 64412 ### these are doubled-up IDs
+    ### use iucn_sid as index, since some include multiple am_sid
+    i <- i + 1
+    sids <- spp_list_build %>% 
+      select(am_sid, iucn_sid, sciname) %>%
[…]

### 2016-02-12 data_explore/data_explore.Rmd
oops, write the processed cell files

@@
 iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
-                           col_types = 'cdddd', progress = TRUE) %>%
-  as.data.frame() %>%
-  filter(presence != 5) ### removes extinct subpopulations
+                           col_types = 'cdddd', progress = TRUE)
 

### 2016-02-12 data_explore/data_explore.Rmd
updated plots with newly cleaned data

@@
 iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
-                           col_types = 'cdddd', progress = TRUE)
+                           progress = TRUE)
 
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), progress = TRUE)
+am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), 
+                           progress = TRUE)
 
@@
 
-spp_list <- spp_list %>%
-  mutate(area_ratio = am_area / iucn_area,
-         area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
-         area_ratio = area_ratio * 100)
-
 ### define 25%, 50%, and 75% quartiles for both x and y axes
@@
 #                'q3' = 'Conus magnificus')
-
-# spp_rando <- c("Lagocephalus lagocephalus",
-#                "Makaira nigricans",
-#                "Pentacheles laevis",
-#                "Stereomastis suhmi",
-#                "Stereomastis surda") 
 # 
@@
 
-spp_list <- spp_list %>% mutate(data_poor = ifelse(occurcells < 10, '< 10', '>= 10'))
+spp_list <- spp_list %>% 
+  mutate(data_poor = ifelse(occurcells < 10, '< 10', '>= 10'))
 ### mongo plot time
-scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
+scatter_arearatio_vs_overlap2 <- ggplot(spp_list %>% 
                                          transform(popn_category = reorder(popn_category, category_score)),
@@
 ### Manage scales for color and size 
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 +
   scale_colour_manual(values = c('>= 10' = 'green4', 
@@
 ### here are median labels:
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 +
   annotate(geom = 'text',
@@
 
-### here are median lines:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_vline(x = area_ratio_quartiles[2],
-#              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = pct_overlap_quartiles[2],
-#              color = 'grey50', linetype = 'dashed')
-
-### here are 25% and 75% quantile lines:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_vline(x = area_ratio_quartiles[c(1, 3)],
-#              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = pct_overlap_quartiles[c(1, 3)],
-#              color = 'grey50', linetype = 'dashed')
-
-### here are individual species, identified and named on the plot:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
-#   geom_text(data = spp_list_label, 
-#             aes(label = sciname_label), 
-#             color = 'grey30',
-#             fontface = 'bold.italic',
-#             size = 2,
-#             hjust = .25,
-#             vjust = 1.5)
-
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 +
   labs(x = bquote('Extent alignment '*alpha['ext']*' (%)'), 
@@
 
-print(scatter_arearatio_vs_overlap)
+print(scatter_arearatio_vs_overlap2)
 
@@
          quad = ifelse(is.na(quad) & sciname %in% spp_list_q4$sciname, 'q4', quad))
-write.csv(spp_list_quads, file.path(dir_data, 'spp_list_quads.csv'), row.names = FALSE)
+write_csv(spp_list_quads, file.path(dir_data, 'spp_list_quads.csv'))
 
@@
   ggtheme_plot + 
+  theme(panel.grid.major.y = element_blank()) +
   geom_bar(stat = 'count', alpha = .5) +
@@
                                'q4 poorly matched' = 'orangered2')) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, 
-            #size = 2, 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.1)) +
+  geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
+            size = 2, 
             color = 'grey20') +
@@
 
-write.csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads.csv'), row.names = FALSE)
+write_csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads.csv'))
 
@@
   group_by(popn_category) %>%
-  summarize(n_spp = first(n_spp), n_spp_q1 = n())
+  summarize(n_spp_q1 = n())
 spp_risk_q2 <- spp_list_q2 %>%
   group_by(popn_category) %>%
-  summarize(n_spp = first(n_spp), n_spp_q2 = n())
+  summarize(n_spp_q2 = n())
 spp_risk_q3 <- spp_list_q3 %>%
   group_by(popn_category) %>%
-  summarize(n_spp = first(n_spp), n_spp_q3 = n())
+  summarize(n_spp_q3 = n())
 spp_risk_q4 <- spp_list_q4 %>%
   group_by(popn_category) %>%
-  summarize(n_spp = first(n_spp), n_spp_q4 = n())
+  summarize(n_spp_q4 = n())
 
@@
 spp_risk_quadrants <- spp_risk_q1 %>%
-  full_join(spp_risk_q2, by = c('popn_category', 'n_spp')) %>%
-  full_join(spp_risk_q3, by = c('popn_category', 'n_spp')) %>%
-  full_join(spp_risk_q4, by = c('popn_category', 'n_spp')) %>%
-  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
+  full_join(spp_risk_q2, by = c('popn_category')) %>%
+  full_join(spp_risk_q3, by = c('popn_category')) %>%
+  full_join(spp_risk_q4, by = c('popn_category')) %>%
+  gather(quad, n_spp_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
+  group_by(popn_category) %>%
+  mutate(n_tot_cat = sum(n_spp_quad)) %>%
+  ungroup() %>%
   mutate(quad = str_replace(quad, 'n_spp_', ''),
-         pct_quad = n_quad/n_spp,
-         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
+         pct_quad   = n_spp_quad/n_tot_cat,
          pct_quad   = ifelse(is.na(pct_quad),   0, pct_quad))
 
-spp_risk_quadrants <- spp_risk_quadrants %>%
+spp_risk_quadrants1 <- spp_risk_quadrants %>%
   left_join(spp_risk_quadrants %>% 
@@
               unique() %>%
-              mutate(cat_rank = ifelse(popn_category == 'DD', -1, cat_rank)),
+              mutate(cat_rank = ifelse(popn_category %in% c('DD', 'NE'), -1, cat_rank)),
             by = 'popn_category')
@@
 ### Plot the bar chart
-barchart_spp_risk_quads <- ggplot(spp_risk_quadrants %>% 
+barchart_spp_risk_quads <- ggplot(spp_risk_quadrants1 %>% 
                                   transform(popn_category = reorder(popn_category, cat_rank)), 
@@
                                'q4 poorly matched' = 'orangered2')) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, 
-            #size = 2, 
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.1)) +
+  geom_text(aes(label = sprintf('n = %s', n_tot_cat), y = 1.01), hjust = 0, 
+            size = 2, 
             color = 'grey20') +
@@
 
-iucn_marine_file <- file.path(dir_data, 'spp_iucn_marine.csv')
-spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
+iucn_file <- file.path(dir_data, 'spp_iucn.csv')
+spp_iucn <- read.csv(iucn_file, stringsAsFactors = FALSE)
 
@@
   filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
-  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp
+  filter(sciname %in% spp_iucn$sciname)   ### ditch freshwater spp
 
@@
 
-``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE}
+``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE}
 
@@
   filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(sciname %in% spp_iucn_marine$sciname) %>%
+  filter(sciname %in% spp_iucn$sciname) %>%
   mutate(class = str_to_title(class),
@@
   filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(sciname %in% spp_iucn_marine$sciname) %>%
+  filter(sciname %in% spp_iucn$sciname) %>%
   filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
@@
 
-``` {r plot bar charts by taxa, echo = FALSE, message = FALSE, warning = FALSE}
+``` {r plot bar charts by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE}
 ### bar charts comparing number of species in each class for each data set

### 2016-02-12 data_explore/data_explore_am_thresh.Rmd
updated plots with newly cleaned data

@@
             median_prop = median(proportion, na.rm = TRUE),
-            quant09 = quantile(proportion, .09),
+            quant09 = quantile(proportion, .05),
             quant25 = quantile(proportion, .25),
             quant75 = quantile(proportion, .75),
-            quant91 = quantile(proportion, .91))
+            quant91 = quantile(proportion, .95))
 

### 2016-02-12 data_explore/data_prep.Rmd
oops, write the processed cell files
updated plots with newly cleaned data

@@
 
-gather spp_all, spp_am, spp_am_v_iucn, spp_list_w_area_all and _trimmed into data folder.
+gather spp_all, spp_am, spp_iucn, spp_iucn_maps, and spp_am_raw into local data folder.
 
@@
   dplyr::select(am_sid, sciname, reviewed, iucn_sid, popn_trend,
-         popn_category, spp_group, id_no, spatial_source, name_verified) %>%
+                popn_category, category_score, spp_group, 
+                id_no, spatial_source, name_verified) %>%
   unique()
@@
 ### some species with no cells reported.  2249 listed in the am v iucn file.
+write_csv(iucn_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'))
 
@@
 ### length(unique(am_spp_cells$am_sid)) results in 2283 species found - great!
+write_csv(am_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'))
 
@@
 spp_list <- spp_list %>%
-  mutate(sm_perc = ifelse(!is.na(sm_perc), 0, sm_perc))
-  # filter(!is.na(sm_perc)) %>%
+  mutate(sm_perc = ifelse(is.na(sm_perc), 0, sm_perc)) %>%
+  filter(!area_iucn == 0)
   ### will ditch the ones where no IUCN polygon exists - or are these species
   ### whose ranges only occur in cells beyond the LOICZID grid?
+
+### attach group name text field, and occurcells field
 spp_list <- spp_list %>%
@@
 
+cat_score_df <- data.frame(popn_category  = c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE'),
+                           category_score = c( 0.0,  0.2,  0.4,  0.6,  0.8,  1.0,   NA,  NA))
+spp_list <- spp_list %>%
+  left_join(cat_score_df, by = 'popn_category')
+
 write_csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'))

### 2016-02-12 data_explore/data_prep.Rmd
got some code for finding occurrence info... let it run on Neptune!

@@
 
-``` {r get_counts, echo = FALSE}
-x <- occ_search(scientificName = spp_iucn_marine$sciname[1:100], fields = 'minimal', limit = 25)
+``` {r spp_all with observation counts}
+spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))
+
+library(rgbif)
+# library(robis)
+
+x <- occ_search(scientificName = spp_all$sciname[1:100], fields = 'minimal', limit = 5)
 y <- sapply(x, function(x) x$meta$count)
+z <- data.frame(sciname = names(y), obs = y)
+
+z <- z %>% 
+  left_join(spp_all %>%
+              select(iucn_sid, am_sid, sciname),
+            by = 'sciname') %>%
+  left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
+                       select(am_sid = speciesid, occurcells),
+                     by = 'am_sid') %>%
+  mutate(ln_obs        = log(obs + 1),
+         ln_occurcells = log(occurcells + 1))
+
+plot(ln_obs ~ ln_occurcells, data = z)
+
 ```

### 2016-02-12 data_explore/data_prep.Rmd
updating so it will pull counts for all the species in spp_all...

@@
 
-``` {r spp_all with observation counts}
+``` {r counts_spp_all_observations}
 spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))
@@
 
-x <- occ_search(scientificName = spp_all$sciname[1:100], fields = 'minimal', limit = 5)
+system.time({
+  x <- occ_search(scientificName = spp_all$sciname, fields = 'minimal', limit = 1)
+  ### seems linear - 10 names ~ 2.5 sec, 100 names ~ 25 sec, 1000 names, ~250 sec
+  ### so 30000 names ~ 7500 sec or 2 hours
+})
 y <- sapply(x, function(x) x$meta$count)
@@
 plot(ln_obs ~ ln_occurcells, data = z)
+plot(obs ~ occurcells, data = z)
+print(summary(lm(obs ~ occurcells, data = z)))
+
+write_csv(z, file.path(dir_anx, 'explore_am_v_iucn/occurrence_df.csv'))
 

### 2016-02-16 data_explore/data_prep.Rmd
pushing new version of acquiring observation information

--- 2016-02-12 data_explore/data_prep.Rmd
@@
 
-``` {r get_counts, echo = FALSE}
-x <- occ_search(scientificName = spp_iucn_marine$sciname[1:100], fields = 'minimal', limit = 25)
+``` {r counts_spp_all_observations}
+spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))
+
+library(rgbif)
+# library(robis)
+
+system.time({
+  x <- occ_search(scientificName = spp_all$sciname[1:100], fields = 'minimal', limit = 1)
+  ### seems linear - 10 names ~ 2.5 sec, 100 names ~ 25 sec, 1000 names, ~250 sec
+  ### so 30000 names ~ 7500 sec or 2 hours
+})
 y <- sapply(x, function(x) x$meta$count)
+gbif_occur <- data.frame(sciname = names(y), gbif_occ = y)
+
+gbif_occur <- gbif_occur %>% 
+  left_join(spp_all %>%
+              select(iucn_sid, am_sid, sciname),
+            by = 'sciname') %>%
+  left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
+                       select(am_sid = speciesid, occurcells),
+                     by = 'am_sid')
+
+# plot(ln_obs ~ ln_occurcells, data = gbif_occur)
+# plot(obs ~ occurcells, data = gbif_occur)
+# print(summary(lm(obs ~ occurcells, data = gbif_occur)))
+
+write_csv(gbif_occur, file.path(dir_data, 'gbif_occurrence_df.csv'))
+
+source('obis_count.R')
+scinames <- spp_all$sciname[1:100]
+
+obis_occur <- obis_count(scinames)
+
+obis_occur <- obis_occur %>% 
+  left_join(spp_all %>%
+              select(iucn_sid, am_sid, sciname),
+            by = 'sciname') %>%
+  left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
+                       select(am_sid = speciesid, occurcells),
+                     by = 'am_sid')
+
+write_csv(obis_occur, file.path(dir_data, 'obis_occurrence_df.csv'))
+
+totes_occur <- obis_occur %>%
+  left_join(gbif_occur)
+
 ```

### 2016-02-16 data_explore/obis_count.R

@@ -1,0 +1,87 @@
+### support functions for obis_count()
+
+obis_url <- function() {
+  getOption("robis_url", "http://api.iobis.org/")
+}
+
+max_characters <- function() {
+  getOption("robis_max_characters", 50000)
+}
+
+handle_date <- function(date) {
+  if(!is.null(date) && class(date) == "Date") {
+    as.character(date)
+  } else {
+    date
+  }
+}
+
+handle_vector <- function(x) {
+  if(!is.null(x)) {
+    paste0(x, collapse = ",")
+  } else {
+    x
+  }
+}
+
+log_request <- function(result) {
+  cat("\n", paste(result$request$method, result$request$url, result$status_code, result$headers$age, result$times[["total"]]), "\n", sep="")
+}
+
+log_progress <- function(total, count) {
+  cat("\rRetrieved ", total, " records of ", count, " (", floor(total/count*100),"%)", sep="")
+}
+
+http_request <- function(method, path, query) {
+  if (method == "GET") {
+    httr::GET(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
+        path = path, query = query)
+  } else if (method == "POST") {
+    httr::POST(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
+         path = path, body = query)
+  }
+}
+
+##########################################################################
+
+obis_count <- function(
+  scinames,
+  verbose = FALSE) {
+
+  offset <- 0
+  i <- 0
+  occ_list <- list() ### initialize occurrence count list
+  
+  for (sciname in scinames) {
+    i <- i + 1
+    lastpage <- FALSE
+    ptm <- proc.time()
+    
+    query <- list(scientificname = sciname,
+                  offset = format(offset, scientific = FALSE))
+    
+    result <- http_request("GET", "occurrence", query)
+    
+    stop_for_status(result)
+    res <- content(result)
+    
+    if(!is.null(res$message)) {
+      lastpage = TRUE
+      warning(res$message)
+      obis_occ <- NA
+    } else {
+      obis_occ <- res$count
+    }
+    tmp_df <- data.frame('sciname' = sciname, 'occ_count' = obis_occ, stringsAsFactors = FALSE)
+    
+    ptm <- round((proc.time() - ptm)[["elapsed"]], 2)
[…]

### 2016-02-16 data_explore/data_prep.Rmd
updates

@@
 library(rgbif)
-# library(robis)
 
-system.time({
-  x <- occ_search(scientificName = spp_all$sciname[1:100], fields = 'minimal', limit = 1)
-  ### seems linear - 10 names ~ 2.5 sec, 100 names ~ 25 sec, 1000 names, ~250 sec
-  ### so 30000 names ~ 7500 sec or 2 hours
-})
-y <- sapply(x, function(x) x$meta$count)
-gbif_occur <- data.frame(sciname = names(y), gbif_occ = y)
-
-gbif_occur <- gbif_occur %>% 
-  left_join(spp_all %>%
-              select(iucn_sid, am_sid, sciname),
-            by = 'sciname') %>%
-  left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
-                       select(am_sid = speciesid, occurcells),
-                     by = 'am_sid')
+# system.time({
+#   x <- occ_search(scientificName = spp_all$sciname[1:100], fields = 'minimal', limit = 1)
+#   ### seems linear - 10 names ~ 2.5 sec, 100 names ~ 25 sec, 1000 names, ~250 sec
+#   ### so 30000 names ~ 7500 sec or 2 hours
+# })
+# y <- sapply(x, function(x) x$meta$count)
+# gbif_occur <- data.frame(sciname = names(y), gbif_occ = y)
+# 
+# gbif_occur <- gbif_occur %>% 
+#   left_join(spp_all %>%
+#               select(iucn_sid, am_sid, sciname),
+#             by = 'sciname') %>%
+#   left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
+#                        select(am_sid = speciesid, occurcells),
+#                      by = 'am_sid')
 
@@
 
-write_csv(gbif_occur, file.path(dir_data, 'gbif_occurrence_df.csv'))
+# write_csv(gbif_occur, file.path(dir_data, 'gbif_occurrence_df.csv'))
+gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence_df.csv'))
 
 source('obis_count.R')
-scinames <- spp_all$sciname[1:100]
+scinames <- spp_all$sciname
 
@@
 
-write_csv(obis_occur, file.path(dir_data, 'obis_occurrence_df.csv'))
+write_csv(obis_occur, file.path(dir_data, 'obis_occurrence.csv'))
 

### 2016-02-16 data_explore/obis_count.R

@@ -1,4 +1,5 @@
 ### support functions for obis_count()
+library(httr)
 
 obis_url <- function() {
   getOption("robis_url", "http://api.iobis.org/")
@@ -51,6 +52,7 @@ obis_count <- function(
   offset <- 0
   i <- 0
   occ_list <- list() ### initialize occurrence count list
+  ptm_total <- proc.time()
   
   for (sciname in scinames) {
     i <- i + 1
@@ -68,18 +70,27 @@ obis_count <- function(
     if(!is.null(res$message)) {
       lastpage = TRUE
       warning(res$message)
-      obis_occ <- NA
+      occ_count <- NA
     } else {
-      obis_occ <- res$count
+      occ_count <- res$count
     }
-    tmp_df <- data.frame('sciname' = sciname, 'occ_count' = obis_occ, stringsAsFactors = FALSE)
+    tmp_df <- data.frame('sciname' = sciname, 'obis_occ' = occ_count, stringsAsFactors = FALSE)
     
-    ptm <- round((proc.time() - ptm)[["elapsed"]], 2)
-    if(verbose) message(i, '. Species ', sciname, '; Occurrence count: ', obis_occ, '; Total time: ', ptm, ' seconds')
-    else if(i == round(i, -2)) message('# species processed: ', i)
+    if(verbose) {
+      ptm <- round((proc.time() - ptm)[["elapsed"]], 2)
+      message(i, '. Species ', sciname, '; Occurrence count: ', occ_count, 
+              '; Elapsed time: ', ptm, ' seconds')
+    } else if(i == round(i, -2)) {
+      ptm <- round((proc.time() - ptm_total)[["elapsed"]], 2)
+      message(i, ' species processed of ', length(scinames), 
+              '; Total time: ', ptm, ' seconds')
+    } 
     
     occ_list[[i]] <- tmp_df
   }
+  
+  ptm <- round((proc.time() - ptm_total)[["elapsed"]], 2)
+  message('Total: ', i, ' species processed; Total time: ', ptm, ' seconds')
   
   occ_df <- rbind_all(occ_list)
 

### 2016-02-17 data_explore/data_prep.Rmd
Updating numbers etc in manuscript... done with "data poor" section?

@@
 
-library(raster)
 library(readr)
 library(data.table)
-library(ggplot2)
 library(tidyr)
@@
 if(basename(getwd()) != 'data_explore') setwd('~/github/IUCN-AquaMaps/data_explore')
-source('data_explore_fxns.R')
+# source('data_explore_fxns.R')
 ### When knitting this, it automatically sets WD to be this directory...
@@
 ### pull spp_iucn_marine.csv from git-annex and bring to local repo
-spp_iucn <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_cleaned.csv'), 
-                            stringsAsFactors = FALSE)
+spp_iucn <- read_csv(file.path(dir_anx, scenario, 'int/spp_iucn_cleaned.csv'))
 spp_iucn <- spp_iucn %>%
@@
 ### pull spp_iucn_maps_all.csv from git-annex and bring to local repo
-spp_iucn_maps <- read.csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv'), 
-                            stringsAsFactors = FALSE) %>%
+spp_iucn_maps <- read_csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv')) %>%
   rename(iucn_sid = id_no)
@@
 ### pull cleaned spp_am data from git-annex and bring to local repo
-spp_am <- read.csv(file.path(dir_anx, scenario, 'int/spp_am_cleaned.csv'), 
-                   stringsAsFactors = FALSE)
+spp_am <- read_csv(file.path(dir_anx, scenario, 'int/spp_am_cleaned.csv'))
 write_csv(spp_am, file.path(dir_data, 'spp_am.csv'))
@@
 ### pull raw aquamaps species list from git-annex and bring to local repo
-spp_am_raw <- read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
-                   stringsAsFactors = FALSE)
+spp_am_raw <- read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'))
 write_csv(spp_am_raw, file.path(dir_data, 'speciesoccursum.csv'))
@@
 ### function create_spp_master_lookup(source_pref = 'iucn', fn_tag = '_complete', reload = TRUE)
-spp_all_remote <- read.csv(file.path(dir_anx, scenario, 'int/spp_all.csv'), 
-                           stringsAsFactors = FALSE) 
+spp_all_remote <- read_csv(file.path(dir_anx, scenario, 'int/spp_all.csv'))
 spp_all_remote_clean <- spp_all_remote %>%
@@
 ### But, some of the extracted IUCN maps don't appear on this list! add 'em in.
-spp_iucn_maps_new <- read.csv(file.path(dir_data, 'spp_iucn_maps.csv'), 
-                              stringsAsFactors = FALSE) %>%
+spp_iucn_maps_new <- read_csv(file.path(dir_data, 'spp_iucn_maps.csv')) %>%
   filter(!sciname %in% spp_all$sciname) %>%
@@
 ### - aliases, etc.
-spp_am_new <- read.csv(file.path(dir_data, 'speciesoccursum.csv'), 
-                       stringsAsFactors = FALSE) %>%
+spp_am_new <- read_csv(file.path(dir_data, 'speciesoccursum.csv')) %>%
   mutate(sciname = paste(str_trim(genus), str_trim(species)),
@@
 ### that also have a non-NA AquaMaps ID code.
-spp_all <- read.csv(file.path(dir_data, 'spp_all.csv'), stringsAsFactors = FALSE)
+spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))
 
@@
 
-loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
-loiczid_raster       <- raster(loiczid_raster_file)
-names(loiczid_raster) <- 'loiczid'
-
 ```
@@
 
+loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
+loiczid_raster       <- raster::raster(loiczid_raster_file)
+names(loiczid_raster) <- 'loiczid'
+
 data_file <- file.path(dir_data, 'spp_list_w_area_all.csv')
-reload <- TRUE
+reload <- FALSE
 if (!file.exists(data_file) | reload) {
@@
 
-spp_list <- read.csv(data_file, stringsAsFactors = FALSE) 
+spp_list <- read_csv(data_file) 
 ### At this point, spp_list zero-area IUCNs due to missing (?) polygons (maybe points?)
@@
 spp_list <- spp_list %>%
-  left_join(read.csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv'), 
-                     stringsAsFactors = FALSE),
+  left_join(x <- read_csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv')),
             by = 'spp_group') %>%
-  left_join(read.csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'), 
-                   stringsAsFactors = FALSE) %>% 
+  left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>% 
               select(am_sid = speciesid, occurcells),
@@
 
-``` {r counts_spp_all_observations}
+``` {r counts_spp_all_observations, echo = FALSE}
 spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))
 
-library(rgbif)
+# library(rgbif)
 
@@
 
-# write_csv(gbif_occur, file.path(dir_data, 'gbif_occurrence_df.csv'))
-gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence_df.csv'))
+# write_csv(gbif_occur, file.path(dir_data, 'gbif_occurrence.csv'))
+gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence.csv'))
 
-source('obis_count.R')
-scinames <- spp_all$sciname
-
-obis_occur <- obis_count(scinames)
-
-obis_occur <- obis_occur %>% 
-  left_join(spp_all %>%
-              select(iucn_sid, am_sid, sciname),
-            by = 'sciname') %>%
-  left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
-                       select(am_sid = speciesid, occurcells),
-                     by = 'am_sid')
-
-write_csv(obis_occur, file.path(dir_data, 'obis_occurrence.csv'))
+# source('obis_count.R')
+# scinames <- spp_all$sciname
+# 
+# obis_occur <- obis_count(scinames)
+# obis_occur <- obis_count(scinames[26500:26600], verbose = TRUE)
+# 
+# obis_occur <- obis_occur %>% 
+#   left_join(spp_all %>%
+#               select(iucn_sid, am_sid, sciname),
+#             by = 'sciname') %>%
+#   left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
+#                        select(am_sid = speciesid, occurcells),
+#                      by = 'am_sid')
+# 
+# write_csv(obis_occur, file.path(dir_data, 'obis_occurrence.csv'))
+obis_occur <- read_csv(file.path(dir_data, 'obis_occurrence.csv'))
 
 totes_occur <- obis_occur %>%
-  left_join(gbif_occur)
+  left_join(gbif_occur) %>%
+  rowwise() %>%
+  mutate(mean_occ = mean(c(gbif_occ, obis_occ), na.rm = TRUE))
 
 ```
+
+__Possible "data poor" indicators:__
+
+* (of matched species) how does bottom decile get distributed across quadrants
+* use mean of OBIS and GBIF, and just report average (or median?) occurrences across all species in each quadrant
+* also use similar method for AM occurcells
+
+``` {r data_poor_analysis, echo = FALSE, eval = FALSE}
+spp_quads <- read_csv(file.path(dir_data, 'spp_list_quads.csv')) %>% 
+  select(am_sid, iucn_sid, occurcells, quad) %>%
+  unique()
+### This contains all mapped species, discarding those with no actual IUCN maps
+
+spp_occ_counts <- spp_quads %>%
+  left_join(totes_occur) %>%
+  group_by(iucn_sid, am_sid, quad) %>%
+  summarize(occurcells = mean(occurcells, na.rm = TRUE), occ = mean(mean_occ, na.rm = TRUE)) %>%
+  group_by(iucn_sid, quad) %>%
+  summarize(occurcells = sum(occurcells, na.rm = TRUE), occ = sum(occ, na.rm = TRUE)) %>%
+  ungroup()
+
+btm_dec_occurcells <- spp_occ_counts$occurcells %>% quantile(c(.1, .25))
+btm_dec_occ <- spp_occ_counts$occ %>% quantile(c(.1, .25))
+
+quad_count_means <- spp_occ_counts %>%
+  mutate(quad = ifelse(quad == 'q4', 'q4', 'q123')) %>%
+  group_by(quad) %>%
+  summarize(mean_occurcells = mean(occurcells),
+            mean_occ        = mean(occ),
+            med_occurcells  = median(occurcells),
+            med_occ         = median(occ),
+            dp_am_10 = sum(occurcells <= btm_dec_occurcells[1])/n(),
+            dp_am_25 = sum(occurcells <= btm_dec_occurcells[2])/n(),
+            dp_gb_10 = sum(occ <= btm_dec_occ[1])/n(),
+            dp_gb_25 = sum(occ <= btm_dec_occ[2])/n())
+
+write_csv(quad_count_means, file.path(dir_data, 'quads_data_poor.csv'))
+
+```

### 2016-02-17 manuscript-main.Rmd
Updating numbers etc in manuscript... done with "data poor" section?

@@
 # library(maps)
-# library(readr)
 
@@
 # dir_anx  <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-# dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
-# dir_fig  <- file.path(getwd(), 'figures')
 
 if(basename(getwd()) != 'IUCN-AquaMaps') setwd('~/github/IUCN-AquaMaps')
+
+dir_data <- 'data'
+dir_fig  <- 'figures'
 
@@
 data_file <- 'data/spp_list_w_area_trimmed.csv'
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE)
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) 
+spp_map_pairs <- read_csv(data_file)
 ### NOTE: as of Jan 31, includes all map pairs, even ones where IUCN area == 0
@@
 spp_all_file <- 'data/spp_all.csv'
-spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
+spp_all <- read_csv(spp_all_file)
 
 # iucn_marine_file <- 'data/spp_iucn_marine.csv'
-# spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)
+# spp_iucn_marine <- read_csv(iucn_marine_file)
 
 iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE)
+spp_iucn_maps <- read_csv(iucn_maps_file)
 
 spp_am_file <- 'data/spp_am.csv'
-spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
+spp_am <- read_csv(spp_am_file)
 

### 2016-02-17 ms_2_intro_res_disc.Rmd
committing so I can re-run area calcs
Updating numbers etc in manuscript... done with "data poor" section?

@@
 n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
-n_spp_iucn   <- nrow(spp_iucn_maps)
-n_spp_both   <- nrow(spp_map_pairs)
+n_spp_iucn   <- length(spp_iucn_maps$iucn_sid %>% unique())
+n_spp_both   <- length(spp_map_pairs$iucn_sid %>% unique())
 
@@
 To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of species.
-Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` species).
+Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique species).
 For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and outline possible improvements.
@@
 p_q123_rev <- round(n_q123_rev/n_spp_both * 100, 1)
+
+quad_count_means <- read_csv(file.path(dir_data, 'quads_data_poor.csv')) %>%
+  arrange(quad)
+dp_med_am_q4    <- quad_count_means$med_occurcells[4]
+dp_med_am_q123  <- quad_count_means$med_occurcells_total[1]
+dp_med_occ_q4   <- quad_count_means$med_occ[4]
+dp_med_occ_q123 <- quad_count_means$med_occ_total[1]
+
 ```
@@
 <!--BSH: This implies the problem is solely with AM.
-Are there no issues with IUCN maps that cause problems in this quadrant?--> ___The data-poor condition likely applies to IUCN as well - AM using OBIS points to generate models, and IUCN is probably doing the same.
+Are there no issues with IUCN maps that cause problems in this quadrant?--> <!--___The data-poor condition likely applies to IUCN as well - AM using OBIS points to generate models, and IUCN is probably doing the same.
 Check, and rework.
@@
 
-The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match.
-When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences; at the same time, IUCN range maps .
-Indeed, `r p_q4_datapoor`% (n = `r n_q4_datapoor`) of these cases are data-poor (in which the AquaMaps environmental envelope model is based on fewer than 10 known cell occurrences ___is there a good way to get the number of OBIS observations for each of these species, as an external validation of data-poor?)___, compared with `r p_q123_datapoor`% for species in the other three quadrants.
-Only `r p_q4_rev`% (n = `r n_q4_rev`) of AquaMaps maps in this quadrant have undergone expert review, compared with `r p_q123_rev`% of species in the other three quadrants.
-<!--BSH: This implies the problem is solely with AM.
-Are there no issues with IUCN maps that cause problems in this quadrant?--> ___The data-poor condition likely applies to IUCN as well - AM using OBIS points to generate models, and IUCN is probably doing the same.
-Check, and rework.
-This means AM is creating errors of commission, while IUCN may be committing errors of omission for poorly-studied species and/or regions - compounding the total problem.___
+The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match well.
+Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (summing (averaging?) OBIS and GBIF occurrences) for this quadrant is `r dp_med_am_q4`, compared to a median of `r dp_med_am_q123` for species in the other three quadrants.
+When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species.
 

### 2016-02-17 data_explore/obis_count.R

@@ -45,15 +45,13 @@ http_request <- function(method, path, query) {
 
 ##########################################################################
 
-obis_count <- function(
-  scinames,
-  verbose = FALSE) {
+obis_count <- function(scinames, verbose = FALSE) {
 
   offset <- 0
   i <- 0
-  occ_list <- list() ### initialize occurrence count list
+  occ_list <- vector('list', length = length(scinames)) ### initialize occurrence count list
   ptm_total <- proc.time()
-  
+ 
   for (sciname in scinames) {
     i <- i + 1
     lastpage <- FALSE

### 2016-02-18 mpa_analysis/mpa_analysis.Rmd
setting up to do raster crap on Neptune

--- 2016-01-22 data_explore/mpa_analysis.Rmd
@@
 
-if(basename(getwd()) != 'data_explore') setwd('data_explore')
+if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)
 
@@
         * -180 to +180 longitude; min/max latitude use min/max from AquaMaps
-    * WDPA_MPA layer: value = IUCN category (1-6, with NA = 7?) combined with marine flag and marine overlap flag?
+    * WDPA_MPA layer: value = IUCN category (1-4, with NA = 7?) combined with marine flag and marine overlap flag?
         * first digit IUCN category; second digit marine flag; third digit marine overlap?
@@
 ### Save it as a .tif in GitHub - about 50 
-rast_base_file <- file.path(dir_git, 'data/mpa_analysis/rast_base.tif')
+rast_base_file <- file.path(dir_git, 'mpa_analysis/rast_base.tif')
 
@@
 ``` {r set_up_spatially_marine_raster}
+
 ### This raster (at 0.01 degree resolution) contains flags on whether a cell
@@
 
-rast_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters/rast_marine.tif')
+rast_marine_file <- file.path(dir_git, 'mpa_analysis/rast_marine.tif')
 reload <- FALSE
@@
 
+    poly_mar_union <- raster::union(poly_marine)
+    poly_mar_union <- rgeos::gUnaryUnion(poly_marine)
+
     # writePolyShape(poly_marine, poly_marine_file)
-    writeOGR(obj    = poly_marine,
+    writeOGR(obj    = poly_mar_union,
              dsn    = path.expand(dirname(poly_marine_file)),

### 2016-02-19 mpa_analysis/mpa_analysis.Rmd
kjhgf

@@
 
-### as long as we're here, let's set up some objects for WGS 84 projection
-p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
-crs_wgs84 <- CRS(p4s_wgs84)
-
 ```
@@
 
+### as long as we're here, let's set up some objects for WGS 84 projection
+p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
+crs_wgs84 <- CRS(p4s_wgs84)
+
 rast_marine_file <- file.path(dir_git, 'mpa_analysis/rast_marine.tif')
@@
 ### import OHI regions polygons, to determine spatially-marine areas
-  poly_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters/poly_marine.shp')
+  poly_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters/poly_marine1.shp')
   if(!file.exists(poly_marine_file)) {
@@
 
-    poly_mar_union <- raster::union(poly_marine)
-    poly_mar_union <- rgeos::gUnaryUnion(poly_marine)
+    poly_list <- list()
+    poly_rows <- nrow(poly_marine)
+    n_entities <- 20
+    for (i in 1:ceiling(poly_rows/n_entities)) { # i = 1
+      min_row <- (i - 1) * n_entities + 1
+      max_row <- min(i * n_entities, poly_rows)
+      poly_tmp <- poly_marine[min_row:max_row, ]
+      message(sprintf('%s. Entities [%s:%s]', i, min_row, max_row))
+      print(system.time({
+        poly_tmp2 <- rgeos::gUnaryUnion(poly_tmp)
+      })[3])
+      poly_list[i] <- poly_tmp2
+    }
+
+    poly_mar_frankenstein <- spRbind(poly_list)
+
+    # poly_mar_union <- rgeos::gUnaryUnion(poly_marine)
 
     # writePolyShape(poly_marine, poly_marine_file)
-    writeOGR(obj    = poly_mar_union,
+    writeOGR(obj    = poly_mar_frankenstein,
              dsn    = path.expand(dirname(poly_marine_file)),

### 2016-02-19 mpa_analysis/wdpa_prep.Rmd
added wdpa_prep to access and filter WDPA database

--- 2016-02-18 mpa_analysis/mpa_analysis.Rmd
@@
 * Use this R script to source arcPy script modified from LSP - need to run it on a PC with appropriate GDAL installation, most likely
-
-``` {r set_up_spatially_marine_raster}
-
-### This raster (at 0.01 degree resolution) contains flags on whether a cell
-### is "spatially marine" - i.e. overlaps (center?) with an ocean area.
-### This is separate from the WDPA "marine" flag.
-
-rast_marine_file <- file.path(dir_git, 'mpa_analysis/rast_marine.tif')
-reload <- FALSE
-if(reload == TRUE) {
-### import OHI regions polygons, to determine spatially-marine areas
-  poly_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters/poly_marine.shp')
-  if(!file.exists(poly_marine_file)) {
-    dir_rgn   <- file.path(dir_N, 'git-annex/globalprep/spatial/v2015/data')
-    lyr_rgn <- 'regions_gcs'
-    poly_rgn <- readShapePoly(file.path(dir_rgn, lyr_rgn), 
-                               proj4string = crs_wgs84)
-
-    ### filter out all polygons that are land or landlocked, and flag 
-    ### the rest as spatially-marine = 1
-    poly_marine <- poly_rgn[!str_detect(poly_rgn@data$rgn_typ, 'land'), ]
-    poly_marine@data <- poly_marine@data %>%
-      mutate(sp_mar = 1)
-
-    poly_mar_union <- raster::union(poly_marine)
-    poly_mar_union <- rgeos::gUnaryUnion(poly_marine)
-
-    # writePolyShape(poly_marine, poly_marine_file)
-    writeOGR(obj    = poly_mar_union,
-             dsn    = path.expand(dirname(poly_marine_file)),
-             layer  = str_replace(basename(poly_marine_file), '.shp', ''),
-             driver = 'ESRI Shapefile',
-             overwrite_layer = TRUE,
-             verbose = TRUE)
-
-    } else {
-    poly_marine <- readShapePoly(poly_marine_file, 
-                               proj4string = crs_wgs84)
-    }
-
-  plot(poly_marine, col = 'blue')
-
-  # rast_marine <- rasterize(poly_marine, rast_base, value = sp_mar_flag) 
-  ### excruciatingly slow (interminable?)... consider gdal_rasterize?
-
-  library(gdalUtils)
-  rast_marine <- gdal_rasterize(
-      src_datasource = poly_marine_file,
-      dst_filename   = rast_marine_file, 
-        # destination for output
-      a = 'rgn_id', 
-#      a = 'sp_mar', 
-        # the attribute in the shapefile to be assigned to the cell values
-      te = c(-180, -90, 180, 90), 
-        # extents for output raster
-      tr = c(.01, .01),  
-        # resolution for output raster
-      tap = TRUE, 
-        # target aligned pixels - align coords of extent of output to values of -tr
-      a_nodata = NA, 
-        # nodata value for raster; otherwise they will be filled in with zeroes
-      output_Raster = TRUE, 
-        # return output as a RasterBrick? 
-      verbose = TRUE)
-}
-
-```
 
@@
 poly_wdpa@data <- poly_wdpa@data %>%
-  left_join(prot_code_df, by = 'IUCN_CLASS') %>%
-  mutate(prot_code = prot_code + MARINE)
+  left_join(prot_code_df, by = 'IUCN_CLASS')
+
+poly_mpa <- poly_wdpa[poly_wdpa@data$MARINE == 1, ]
+poly_mpa <- poly_wdpa[poly_wdpa@data$prot_code <= 4, ]
 

### 2016-02-22 mpa_analysis/mpa_gap_analysis.Rmd
oh god did I finally get past this too-large-file bullshit?

---
title: 'MPA Gap Analysis: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(raster)
library(readr)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)

### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

# MPA gap analysis

Several steps to this:

* set up WDPA database in a raster format for easy use
* perform gap analysis for WDPA-MPA data base (which year?) vs AquaMaps 2014, with 50% threshold - replicate the original analysis
* reperform the analysis with latest/greatest WDPA-MPA database, and AquaMaps 2015, with 50% threshold and 0% threshold
* perform the analysis with latest/greatest WDPA-MPA database, and IUCN shapefiles

## set up WDPA raster

* Create polygon subset of WDPA database
    * STATUS = 'Designated'
    * IUCN_CAT = I - IV
    * Simplify polygons to make it easier for rasterization
        * cut any main polygons less than 10% of a raster cell.  Raster
          cells are .01 x .01 degree, so area = .0001; cut any less than 1e-5.
          Do the same for subpolygons within a larger polygon.
    * Inverse order by `STATUS_YR`, so oldest is at the end (for `gdal_rasterize()`)
* To avoid double-counting for overlapping polygons - rasterize WDPA to fine scale first
    * less than 1km^2^ was mentioned in paper as resolution of MPA data.  
      Rasterize to 0.01 degree x 0.01 degree  
        * this results in max cell size of 1.11 km x 1.11 km at the equator.
    * Use STATUS_YR as field; see note above about ordering polygons
      for use in `gdal_rasterize()`

## Set up marine raster

* Create polygons subset of OHI regions, for only marine regions
* Rasterize to same resolution as above.

## Create raster of WDPA IUCN I-IV and spatially marine

* Intersect the two rasters so only protected areas within spatially marine
  areas are included.

## Aggregate protected area layer to LOICZID scale (0.5 deg)

* Use raster::extract? each LOICZID gets 2500 subcells noting NA (not
  protected) or a value (protected); turn this into proportion of
  LOICZID cell that is protected.
* Also need to do this to find the proportion of LOICZID cell that is
  spatially marine - if not fully marine, the protected proportion will
  be larger; but the area will be smaller (from AquaMaps, or recalc)
    * e.g. a single LOICZID is 60% marine (1500 marine cells), and 48%
      protected (1200 protected cells, which are by def marine).  So
      protected proportion = 48/60 = 80% protection of the marine area
      within this cell.

* AquaMaps layer: use as is
* To analyze for each species:
    * Summarize total area and total protected area (for a given protection scenario) for the species
    * Include species extinction risk in there too?

``` {r set_up_base_raster}
### DO WE NEED THIS? Use area at 0.5 degree level; most cells this will be pretty close.
### base raster will contain 0.01 degree cells from -180 to +180,
### and -90 to +90, with the associated area of each cell.
### Save it as a .tif in GitHub - about 50 
rast_base_file <- file.path(dir_git, 'mpa_analysis/rast_base.tif')

if(!file.exists(rast_base_file)) {
  ### import AquaMaps half-degree loiczid raster - for extents and such
  rast_loiczid_file   <- file.path(dir_git, 'data/rasters/loiczid_raster.grd')
  rast_loiczid        <- raster(rast_loiczid_file)
  names(rast_loiczid) <- 'loiczid'

  rast_empty <- raster(ext = extent(rast_loiczid), resolution = 0.01)
  rast_base  <- raster::area(rast_empty) ### takes a while...

  writeRaster(rast_base, rast_base_file, overwrite = TRUE)
} else {
  rast_base <- raster(rast_base_file)
}

```

``` {r load_rasters}

dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
rast_mpa_mar_file <- file.path(dir_rast, 'rast_mpa_mar.tif')
if(!file.exists(rast_mpa_mar_file)) {
  rast_mar_file <- file.path(dir_rast, 'rast_marine.tif')
  rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa1.tif')

  rast_mar <- raster(rast_mar_file)
  rast_mpa <- raster(rast_mpa_file)

  rast_mpa_mar <- mask(rast_mpa, rast_mar)
  writeRaster(rast_mpa_mar, rast_mpa_mar_file)
} else {
  rast_mar            <- raster(rast_mar_file)
  names(rast_mar)     <- 'rgn_id'
  rast_mpa_mar        <- raster(rast_mpa_mar_file)
  names(rast_mpa_mar) <- 'status_yr'
}

```

``` {r aggregate_to_loiczid}
rast_loiczid_file   <- file.path(dir_git, 'data/rasters/loiczid_raster.grd')
rast_loiczid        <- raster(rast_loiczid_file)
names(rast_loiczid) <- 'loiczid'

rast_loiczid_01_file <- file.path(dir_rast, 'rast_loiczid_01.tif')
rast_loiczid_01 <- resample(rast_loiczid, rast_mpa_mar, method = 'ngb')
writeRaster(rast_loiczid_01, rast_loiczid_01_file)

### Turn all marine cells to value of 1 (so can uze zonal sum)
rast_is_mar <- rast_mar
values(rast_is_mar)[!is.na(values(rast_mar))] <- 1
names(rast_is_mar) <- 'is_marine'

### Turn all MPA cells to value of 1 (so can use zonal sum)
### NOTE: loses status_yr values - so filter by year first
rast_is_mpa <- rast_mpa_mar
values(rast_is_mpa)[!is.na(values(rast_mpa_mar))] <- 1
names(rast_is_mpa) <- 'is_mpa'

rast_test <- rast_test2 <- raster(ext = extent(rast_loiczid), resolution = c(10, 10))
values(rast_test) <- 1:648
values(rast_test2) <- rep(c(1:10), each = 64, length.out = length(rast_test))
x <- zonal(rast_test, rast_test2, fun = sum)

### 2016-02-22 mpa_analysis/wdpa_prep.Rmd
oh god did I finally get past this too-large-file bullshit?

@@
 
-# MPA gap analysis
-
-Several steps to this:
-
-* set up WDPA database in a raster format for easy use
-* perform gap analysis for WDPA-MPA data base (which year?) vs AquaMaps 2014, with 50% threshold - replicate the original analysis
-* reperform the analysis with latest/greatest WDPA-MPA database, and AquaMaps 2015, with 50% threshold and 0% threshold
-* perform the analysis with latest/greatest WDPA-MPA database, and IUCN shapefiles
-
-## set up WDPA raster
-
-* direct extract method: extract WDPA polygons against half-degree cells, with proportion; value = year (or combo code)
-    * all protected areas
-    * 'marine' = 1
-    * 'marine' = 1 and IUCN I - IV
-    * clip WDPA to ocean regions, then IUCN I - IV.
-    * NOTE: using extract() could result in double-counting - at least make sure total protected area per cell is not greater than 1.
-        * maybe check any cells with protection > 50% and protected areas n > 1?
-    * this method would be excruciatingly slow.
-* to avoid double-counting for overlapping polygons - rasterize WDPA to fine scale first
-    * less than 1km^2^ was mentioned in paper as resolution of MPA data.  Rasterize to 0.01 degree x 0.01 degree?  
-        * this results in max cell size of 1.11 km x 1.11 km at the equator.
-        * -180 to +180 longitude; min/max latitude use min/max from AquaMaps
-    * WDPA_MPA layer: value = IUCN category (1-4, with NA = 7?) combined with marine flag and marine overlap flag?
-        * first digit IUCN category; second digit marine flag; third digit marine overlap?
-            * rasterize by IUCN category (unprotected, call NA?)
-            * then rasterize again by marine field (0 or 1)
-            * then rasterize land (or ocean) (0 or 1)
-            * then add 'em up to construct full raster at 0.01 degree resolution
-        * for year-to-year analysis, could recreate this layer for older years by filtering out newer protected areas
-            * this allows for a cell's protection class to change over time?
-            * could do this every, say, 5 years - to more quickly capture the trend across a span of, say, 25 years.
-    * Cell area layer: use raster::area() to create a single layer of cell areas at this new resolution.
-    * aggregate cells up to 0.5 degree cells, by protected area within each cell for each type of analysis
-        * summarize to data.frame of loiczid, with area of protection (and total area?)?
-        * create different protected-area layer for each protection scenario
-* AquaMaps layer: use as is
-* To analyze for each species:
-    * Summarize total area and total protected area (for a given protection scenario) for the species
-    * Include species extinction risk in there too?
-
-``` {r set_up_base_raster}
-### base raster will contain 0.01 degree cells from -180 to +180,
-### and -90 to +90, with the associated area of each cell.
-### Save it as a .tif in GitHub - about 50 
-rast_base_file <- file.path(dir_git, 'mpa_analysis/rast_base.tif')
-
-if(!file.exists(rast_base_file)) {
-  ### import AquaMaps half-degree loiczid raster - for extents and such
-  rast_loiczid_file   <- file.path(dir_git, 'data/rasters/loiczid_raster.grd')
-  rast_loiczid        <- raster(rast_loiczid_file)
-  names(rast_loiczid) <- 'loiczid'
-
-  rast_empty <- raster(ext = extent(rast_loiczid), resolution = 0.01)
-  rast_base  <- raster::area(rast_empty) ### takes a while...
-
-  writeRaster(rast_base, rast_base_file, overwrite = TRUE)
-} else {
-  rast_base <- raster(rast_base_file)
-}
-
-```
-
-Dealing with rasters:
-
-* how to ensure 1 km or better resolution?  WGS84 at 0.01 degree or Mollweide at 1 km
-    * WGS84 --> 36000 * 18000 cells = 648,000,000 cells...
-    * Mollweide --> 
-* options for creating rasters
-    * use arcpy as in LSP... global Mollweide raster of WDPA is about 50 MB; includes .tif and supporting files.  This is .5 km raster.  How is it so small?
-    * use gdal_rasterize seems to create multi-GB rasters; .tif only
-* why rasters?  avoid double-counting for overlapping parks.  WDPA is most important, but since that is in raster and AM is raster, easiest to keep it all in raster
-    * WDPA - capture park classification and year and marine flag
-    * spatial marine - capture whether cell includes ocean or not - clip to max N and S of WDPA?
-        * xMin,yMin -180.007,-72.95 : xMax,yMax 180,83.73
-* Use this R script to source arcPy script modified from LSP - need to run it on a PC with appropriate GDAL installation, most likely
-
-``` {r set_up_wdpa_raster}
+``` {r set_up_wdpa_mpa_polys}
 
@@
 lyr_wdpa    <- 'WDPA_poly_Jan2015'
-poly_wdpa <- readShapePoly(file.path(dir_wdpa, lyr_wdpa), 
-                           proj4string = crs_wgs84)
+lyr_mpa     <- 'WDPA_poly_Jan2015_mpa'
+lyr_mpa_mar <- 'WDPA_poly_Jan2015_mpa_mar'
 
-prot_code_df <- data.frame('IUCN_CLASS' = c('I', 'IIa', 'IIb', 'III', 'IV', 'V', 'V', NA),
-                           'prot_code'  = c( 1,    2,     2,     3,     4,   5,   6,   7))
+if(!file.exists(file.path(dir_wdpa, paste(lyr_mpa, '.shp', sep = '')))) {
+  poly_wdpa <- readOGR(dsn = dir_wdpa, layer = lyr_wdpa, stringsAsFactors = FALSE)
+
+#   nrow(poly_wdpa@data)
+#   poly_wdpa@data$IUCN_CAT %>% unique()
+
+  prot_code_df <- data.frame('IUCN_CAT'  = c('I', 'Ia', 'Ib', 'II', 'III', 'IV', 'V', 'VI'),
+                             'prot_code' = c(  1,   1,    1,    2,     3,    4,   5,    6))
+# poly_wdpa_save <- poly_wdpa
   poly_wdpa@data <- poly_wdpa@data %>%
-  left_join(prot_code_df, by = 'IUCN_CLASS')
+    left_join(prot_code_df, by = 'IUCN_CAT') %>%
+    mutate(prot_code = ifelse(is.na(prot_code), 7, prot_code))
 
-poly_mpa <- poly_wdpa[poly_wdpa@data$MARINE == 1, ]
-poly_mpa <- poly_wdpa[poly_wdpa@data$prot_code <= 4, ]
+  poly_mpa <- poly_wdpa[poly_wdpa@data$STATUS == 'Designated', ]
+  poly_mpa <- poly_mpa[poly_mpa@data$prot_code <= 4, ]
+
+  poly_mpa_marine <- poly_mpa[poly_mpa@data$MARINE == 1, ]
+
+  writeOGR(poly_mpa, dsn = dir_wdpa, layer = lyr_mpa,
+           driver = 'ESRI Shapefile')
+  writeOGR(poly_mpa_marine, dsn = dir_wdpa, layer = lyr_mpa_mar,
+           driver = 'ESRI Shapefile')
+}
+
+poly_mpa <- readOGR(dir_wdpa, lyr_mpa, stringsAsFactors = FALSE)
+### create simplified polygon
+polys_mpa_area <- lapply(poly_mpa@polygons, function(x) x@area)
+poly_mpa1 <- poly_mpa[polys_mpa_area > 1e-5, ]
+poly_mpa1 <- poly_simplify(poly_mpa1, thresh = 1e-5, verbose = TRUE)
+
+### sort by STATUS_YR, descending, so gdal_rasterize will assign newer
+### values then overwrite (where applicable) with older values
+poly_mpa2 <- poly_mpa1[order(z <- poly_mpa1@data$STATUS_YR, decreasing = TRUE), ]
+
+writeOGR(poly_mpa1, dsn = dir_wdpa,
+         layer = paste(lyr_mpa, '1', sep = ''),
+         driver = 'ESRI Shapefile')
+
+}
+```
+
+unique(poly_mpa@data$STATUS_YR)
+table(poly_mpa@data$STATUS_YR)
+hist(poly_mpa@data$STATUS_YR[poly_mpa@data$STATUS_YR > 0])
+summary(poly_mpa@data$STATUS_YR[poly_mpa@data$STATUS_YR > 0])
+
+``` {r rasterize_that_shit}
+dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
+
+### Get source from WDPA_MPA data location
+src <- path.expand(file.path(dir_wdpa, paste(lyr_mpa, '1.shp', sep = '')))
+### gdal_rasterize to temp directory
+tmp_dst <- file.path(dir_rast, 'tmp', paste(lyr_mpa, '1.tif', sep = '')) %>%
+  path.expand
+### then writeRaster to final directory, to save smaller file
+fnl_dst <- file.path(dir_rast, paste(lyr_mpa, '1.tif', sep = '')) %>%
+  path.expand
+
+### Set extents and resolution
+ext  <- c(-180, -90, 180, 90)
+reso <- c(.01, .01)
+
+ptm <- proc.time()
+tmp_rast <- gdalUtils::gdal_rasterize(
+  src_datasource = src,
+  dst_filename   = tmp_dst, 
+    # destination for output
+  a = 'STATUS_YR', 
+    # the attribute in the shapefile to be assigned to the cell values
+  te = ext, 
+    # extents for output raster
+  tr = reso,  
+    # resolution for output raster
+  a_nodata = NA, 
+    # nodata value for raster; otherwise they will be filled in with zeroes
+  output_Raster = TRUE,
+    # return output as a RasterBrick? 
+  verbose = TRUE)
+message('Rasterized ', src, '; \n  ', round((proc.time() - ptm)[3], 2), ' seconds')
+
+ptm <- proc.time()
+raster::writeRaster(tmp_rast, fnl_dst, overwrite = TRUE)
+message('Write ', src, '\n  to ', fnl_dst, '; \n  ', round((proc.time() - ptm)[3], 2), ' seconds')
+remove(tmp_rast)
+unlink(tmp_dst)
 
 ```
+
+``` {r simplify_polys_function}
+poly_simplify <- function(poly_set, thresh, verbose = FALSE) {
+  if(class(poly_set) == 'list') poly_set <- poly_set[[1]]
+#   if(verbose) {
+#     message('Counting subpolys in poly set:')
+#     poly_counts <- sapply(poly_set@polygons, function(x) length(x@Polygons))
+#     print(poly_counts)
+#   }
+
+  ### filter out tiny sub-polygons -----
+  ### create area list for filtering.  Area will be in square degrees... 
+  polys_area <- lapply(poly_set@polygons, function(x) sapply(x@Polygons, function(y) y@area))
+
+  if(verbose) {
+    message('Polygon area quantiles:')
+    print(quantile(unlist(polys_area)))
+  }
+
+  ### check to make sure threshold 
+#   if(verbose) {
+#     message('Smallest "main" subpolygon, i.e. smallest EEZ, in this set:')
+#     poly_small <- min(unlist(lapply(polys_area, function(x) (max(x)))))
+#     if(poly_small < thresh) message('Smallest poly smaller than threshold value of ', thresh)
+#     for (j in 1:length(polys_area)) {
+#       message(sprintf('rgn %s: largest polygon %.5f (%s)', poly_set@data$rgn_id[j], 
+#                       max(polys_area[[j]]), poly_set@data$rgn_nam[j]))
+#     }
+#   }
+
+  mainPolys <- lapply(polys_area, function(x) which(x > thresh))
+
+  for(i in 1:length(mainPolys)) {
+    if(length(mainPolys[[i]]) >= 1 && mainPolys[[i]][1] >= 1) {
+      poly_set@polygons[[i]]@Polygons  <- poly_set@polygons[[i]]@Polygons[mainPolys[[i]]]
+      poly_set@polygons[[i]]@plotOrder <- 1:length(poly_set@polygons[[i]]@Polygons)
+    }
+  }
+
+  plot_order_max <- max(poly_set@plotOrder)
+  poly_list_length <- length(poly_set@polygons)
+  message('Plot order max: ', plot_order_max, '; length poly list: ', poly_list_length)
+
+  if(plot_order_max != poly_list_length) {
+    message('NOT SAME LENGTH?!')
+    stop()
+    # poly_set@plotOrder <- 1:length(poly_set@polygons)
+    # is plotOrder necessary to be continuous?  that would explain this line.
+  }
+
+  if(verbose) {
+#     message('Counting subpolys in modified poly set:')
+#     poly_counts <- sapply(poly_set@polygons, function(x) length(x@Polygons))
+#     print(poly_counts)
+
+    polys_area_new <- lapply(poly_set@polygons, function(x) sapply(x@Polygons, function(y) y@area))
+
+    message('Polygon area quantiles after adjustment:')
+    print(quantile(unlist(polys_area_new)))
+  }
+
+  return(poly_set)
+}
+
+```

### 2016-02-25 mpa_analysis/wdpa_prep.Rmd
got the mpa analysis going well.

@@
 library(rgdal)
-library(maptools)
 library(raster)
@@
 ### the 'setwd()' is there for running by chunk
-```
 
-``` {r set_up_wdpa_mpa_polys}
-
-### import WDPA database polygons from original .gdb
-# gdb_wdpa   <- file.path(dir_N, 'git-annex/globalprep/_raw_data/wdpa_mpa/WDPA_Jan2015_Public/WDPA_Jan2015_Public.gdb')
-# lyr_wdpa <- 'WDPA_poly_Jan2015'
-# 
-# poly_wdpa <- readOGR(dsn = gdb_wdpa, layer = lyr_wdpa, stringsAsFactors = FALSE)
-
-### import WDPA database polygons from saved .shp
-dir_wdpa   <- file.path(dir_N, 'git-annex/globalprep/_raw_data/wdpa_mpa/WDPA_Jan2015_Public')
-lyr_wdpa    <- 'WDPA_poly_Jan2015'
-lyr_mpa     <- 'WDPA_poly_Jan2015_mpa'
-lyr_mpa_mar <- 'WDPA_poly_Jan2015_mpa_mar'
-
-if(!file.exists(file.path(dir_wdpa, paste(lyr_mpa, '.shp', sep = '')))) {
-  poly_wdpa <- readOGR(dsn = dir_wdpa, layer = lyr_wdpa, stringsAsFactors = FALSE)
-
-#   nrow(poly_wdpa@data)
-#   poly_wdpa@data$IUCN_CAT %>% unique()
-
-  prot_code_df <- data.frame('IUCN_CAT'  = c('I', 'Ia', 'Ib', 'II', 'III', 'IV', 'V', 'VI'),
-                             'prot_code' = c(  1,   1,    1,    2,     3,    4,   5,    6))
-# poly_wdpa_save <- poly_wdpa
-  poly_wdpa@data <- poly_wdpa@data %>%
-    left_join(prot_code_df, by = 'IUCN_CAT') %>%
-    mutate(prot_code = ifelse(is.na(prot_code), 7, prot_code))
-
-  poly_mpa <- poly_wdpa[poly_wdpa@data$STATUS == 'Designated', ]
-  poly_mpa <- poly_mpa[poly_mpa@data$prot_code <= 4, ]
-
-  poly_mpa_marine <- poly_mpa[poly_mpa@data$MARINE == 1, ]
-
-  writeOGR(poly_mpa, dsn = dir_wdpa, layer = lyr_mpa,
-           driver = 'ESRI Shapefile')
-  writeOGR(poly_mpa_marine, dsn = dir_wdpa, layer = lyr_mpa_mar,
-           driver = 'ESRI Shapefile')
-}
-
-poly_mpa <- readOGR(dir_wdpa, lyr_mpa, stringsAsFactors = FALSE)
-### create simplified polygon
-polys_mpa_area <- lapply(poly_mpa@polygons, function(x) x@area)
-poly_mpa1 <- poly_mpa[polys_mpa_area > 1e-5, ]
-poly_mpa1 <- poly_simplify(poly_mpa1, thresh = 1e-5, verbose = TRUE)
-
-### sort by STATUS_YR, descending, so gdal_rasterize will assign newer
-### values then overwrite (where applicable) with older values
-poly_mpa2 <- poly_mpa1[order(z <- poly_mpa1@data$STATUS_YR, decreasing = TRUE), ]
-
-writeOGR(poly_mpa1, dsn = dir_wdpa,
-         layer = paste(lyr_mpa, '1', sep = ''),
-         driver = 'ESRI Shapefile')
-
-}
-```
-
-unique(poly_mpa@data$STATUS_YR)
-table(poly_mpa@data$STATUS_YR)
-hist(poly_mpa@data$STATUS_YR[poly_mpa@data$STATUS_YR > 0])
-summary(poly_mpa@data$STATUS_YR[poly_mpa@data$STATUS_YR > 0])
-
-``` {r rasterize_that_shit}
-dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
-
-### Get source from WDPA_MPA data location
-src <- path.expand(file.path(dir_wdpa, paste(lyr_mpa, '1.shp', sep = '')))
-### gdal_rasterize to temp directory
-tmp_dst <- file.path(dir_rast, 'tmp', paste(lyr_mpa, '1.tif', sep = '')) %>%
-  path.expand
-### then writeRaster to final directory, to save smaller file
-fnl_dst <- file.path(dir_rast, paste(lyr_mpa, '1.tif', sep = '')) %>%
-  path.expand
-
-### Set extents and resolution
-ext  <- c(-180, -90, 180, 90)
-reso <- c(.01, .01)
-
-ptm <- proc.time()
-tmp_rast <- gdalUtils::gdal_rasterize(
-  src_datasource = src,
-  dst_filename   = tmp_dst, 
-    # destination for output
-  a = 'STATUS_YR', 
-    # the attribute in the shapefile to be assigned to the cell values
-  te = ext, 
-    # extents for output raster
-  tr = reso,  
-    # resolution for output raster
-  a_nodata = NA, 
-    # nodata value for raster; otherwise they will be filled in with zeroes
-  output_Raster = TRUE,
-    # return output as a RasterBrick? 
-  verbose = TRUE)
-message('Rasterized ', src, '; \n  ', round((proc.time() - ptm)[3], 2), ' seconds')
-
-ptm <- proc.time()
-raster::writeRaster(tmp_rast, fnl_dst, overwrite = TRUE)
-message('Write ', src, '\n  to ', fnl_dst, '; \n  ', round((proc.time() - ptm)[3], 2), ' seconds')
-remove(tmp_rast)
-unlink(tmp_dst)
+p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
+crs_wgs84 <- CRS(p4s_wgs84)
 
@@
 ```
+
+``` {r set_up_wdpa_mpa_polys}
+
+### import WDPA database polygons from original .gdb
+# gdb_wdpa   <- file.path(dir_N, 'git-annex/globalprep/_raw_data/wdpa_mpa/WDPA_Jan2015_Public/WDPA_Jan2015_Public.gdb')
+# lyr_wdpa <- 'WDPA_poly_Jan2015'
+# 
+# poly_wdpa <- readOGR(dsn = gdb_wdpa, layer = lyr_wdpa, stringsAsFactors = FALSE)
+
+### import WDPA database polygons from saved .shp
+dir_wdpa   <- file.path(dir_N, 'git-annex/globalprep/_raw_data/wdpa_mpa/WDPA_Jan2015_Public')
+lyr_wdpa    <- 'WDPA_poly_Jan2015'
+lyr_mpa     <- 'WDPA_poly_Jan2015_mpa'
+lyr_mpa_mar <- 'WDPA_poly_Jan2015_mpa_mar'
+
+if(!file.exists(file.path(dir_wdpa, paste(lyr_mpa, '.shp', sep = '')))) {
+  poly_wdpa <- readOGR(dsn = dir_wdpa, layer = lyr_wdpa, stringsAsFactors = FALSE)
+
+#   nrow(poly_wdpa@data)
+#   poly_wdpa@data$IUCN_CAT %>% unique()
+
+  prot_code_df <- data.frame('IUCN_CAT'  = c('Ia', 'Ib', 'II', 'III', 'IV', 'V', 'VI'),
+                             'prot_code' = c(  1,    1,    2,     3,    4,   5,    6))
+# poly_wdpa_save <- poly_wdpa
+  poly_wdpa@data <- poly_wdpa@data %>%
+    left_join(prot_code_df, by = 'IUCN_CAT') %>%
+    mutate(prot_code = ifelse(is.na(prot_code), 7, prot_code))
+
+  poly_mpa <- poly_wdpa[poly_wdpa@data$STATUS == 'Designated', ]
+  poly_mpa <- poly_mpa[poly_mpa@data$prot_code <= 4, ]
+
+  poly_mpa_marine <- poly_mpa[poly_mpa@data$MARINE == 1, ]
+
+  writeOGR(poly_mpa, dsn = dir_wdpa, layer = lyr_mpa,
+           driver = 'ESRI Shapefile', overwrite_layer = TRUE)
+  writeOGR(poly_mpa_marine, dsn = dir_wdpa, layer = lyr_mpa_mar,
+           driver = 'ESRI Shapefile', overwrite_layer = TRUE)
+}
+
+poly_mpa <- readOGR(dir_wdpa, lyr_mpa, stringsAsFactors = FALSE)
+# poly_mpa <- readShapePoly(fn = file.path(dir_wdpa, lyr_mpa),
+#                           proj4string = crs_wgs84,
+#                           stringsAsFactors = FALSE)
+
+### create simplified polygon
+polys_mpa_area <- lapply(poly_mpa@polygons, function(x) x@area)
+poly_mpa1 <- poly_mpa[polys_mpa_area > 1e-5, ]
+poly_mpa1 <- poly_simplify(poly_mpa1, thresh = 1e-5, verbose = TRUE)
+
+### sort by STATUS_YR, descending, so gdal_rasterize will assign newer
+### values then overwrite (where applicable) with older values
+poly_mpa1 <- poly_mpa1[order(poly_mpa1@data$STATUS_YR, decreasing = TRUE), ]
+
+# writeOGR(poly_mpa1, dsn = dir_wdpa,
+#          layer = paste(lyr_mpa, '1', sep = ''),
+#          driver = 'ESRI Shapefile')
+maptools::writePolyShape(poly_mpa1, 
+                         fn = file.path(dir_wdpa, paste(lyr_mpa, '1', sep = '')))
+### copy .prj:
+file.copy(from = file.path(dir_wdpa, paste(lyr_mpa, '.prj', sep = '')), 
+          to = file.path(dir_wdpa, paste(lyr_mpa, '1.prj', sep = '')))
+
+```
+
+unique(poly_mpa@data$STATUS_YR)
+table(poly_mpa@data$STATUS_YR)
+hist(poly_mpa@data$STATUS_YR[poly_mpa@data$STATUS_YR > 0])
+summary(poly_mpa@data$STATUS_YR[poly_mpa@data$STATUS_YR > 0])
+
+``` {r rasterize_that_shit}
+dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
+
+### Get source from WDPA_MPA data location
+src <- path.expand(file.path(dir_wdpa, paste(lyr_mpa, '1.shp', sep = '')))
+### gdal_rasterize to temp directory
+tmp_dst <- file.path(dir_rast, 'tmp', paste(lyr_mpa, '1.tif', sep = '')) %>%
+  path.expand
+### then writeRaster to final directory, to save smaller file
+fnl_dst <- file.path(dir_rast, paste(lyr_mpa, '1.tif', sep = '')) %>%
+  path.expand
+
+### Set extents and resolution
+ext  <- c(-180, -90, 180, 90)
+reso <- c(.01, .01)
+
+ptm <- proc.time()
+tmp_rast <- gdalUtils::gdal_rasterize(
+  src_datasource = src,
+  dst_filename   = tmp_dst, 
+    # destination for output
+  a = 'STATUS_YR', 
+    # the attribute in the shapefile to be assigned to the cell values
+  te = ext, 
+    # extents for output raster
+  tr = reso,  
+    # resolution for output raster
+  a_nodata = NA, 
+    # nodata value for raster; otherwise they will be filled in with zeroes
+  output_Raster = TRUE,
+    # return output as a RasterBrick? 
+  verbose = TRUE)
+message('Rasterized ', src, '; \n  ', round((proc.time() - ptm)[3], 2), ' seconds')
+
+ptm <- proc.time()
+raster::writeRaster(tmp_rast, fnl_dst, 
+                    progress = 'text', overwrite = TRUE)
+message('Write ', src, '\n  to ', fnl_dst, '; \n  ', round((proc.time() - ptm)[3], 2), ' seconds')
+remove(tmp_rast)
+unlink(tmp_dst)
+
+```

### 2016-02-25 ms_2_intro_res_disc.Rmd
meh

@@
     * Additional filters on AquaMaps distribution criteria, such as expert review status or the number of occurrence cells used to generate the species model, can help avoid overextrapolation for data poor species.
+
+Conclusions:
+combine datasets for increased number of species; identify data-poor species to improve confidence in range estimates
+differences in methodology create significant differences in results; errors of both commission and omission
+importance to conservation management - different types of errors have different implications

### 2016-02-25 mpa_analysis/marine_prep.Rmd

---
title: 'MPA Gap Analysis: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(rgdal)
library(maptools)
library(raster)
library(readr)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)

### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

``` {r set_up_spatially_marine_raster}

### This raster (at 0.01 degree resolution) contains flags on whether a cell
### is "spatially marine" - i.e. overlaps (center?) with an ocean area.
### This is separate from the WDPA "marine" flag.

### as long as we're here, let's set up some objects for WGS 84 projection
p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
crs_wgs84 <- CRS(p4s_wgs84)

rast_marine_file <- file.path(dir_git, 'mpa_analysis/rast_marine.tif')
reload <- FALSE
if(reload == TRUE) {
### import OHI regions polygons, to determine spatially-marine areas
  poly_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters/poly_marine1.shp')
  if(!file.exists(poly_marine_file)) {
    dir_rgn   <- file.path(dir_N, 'git-annex/globalprep/spatial/v2015/data')
    lyr_rgn <- 'regions_gcs'
    poly_rgn <- readShapePoly(file.path(dir_rgn, lyr_rgn), 
                               proj4string = crs_wgs84)

    ### filter out all polygons that are land or landlocked, and flag 
    ### the rest as spatially-marine = 1
    poly_marine <- poly_rgn[!str_detect(poly_rgn@data$rgn_typ, 'land'), ]

    # writePolyShape(poly_marine, poly_marine_file)
    writeOGR(obj    = poly_marine,
             dsn    = path.expand(dirname(poly_marine_file)),
             layer  = str_replace(basename(poly_marine_file), '.shp', ''),
             driver = 'ESRI Shapefile',
             overwrite_layer = TRUE,
             verbose = TRUE)
  }

  library(gdalUtils)
  rast_marine <- gdal_rasterize(
      src_datasource = poly_marine_file,
      dst_filename   = rast_marine_file, 
        # destination for output
      a = 'rgn_id', 
        # the attribute in the shapefile to be assigned to the cell values
      te = c(-180, -90, 180, 90), 
        # extents for output raster
      tr = c(.01, .01),  
        # resolution for output raster
      tap = TRUE, 
        # target aligned pixels - align coords of extent of output to values of -tr
      a_nodata = NA, 
        # nodata value for raster; otherwise they will be filled in with zeroes
      output_Raster = TRUE, 
        # return output as a RasterBrick? 
      verbose = TRUE)
}

```

``` {r simplify_polys_function}
poly_simplify <- function(poly_set, thresh, verbose = FALSE) {
  if(class(poly_set) == 'list') poly_set <- poly_set[[1]]
#   if(verbose) {
#     message('Counting subpolys in poly set:')
#     poly_counts <- sapply(poly_set@polygons, function(x) length(x@Polygons))
#     print(poly_counts)
#   }

  ### filter out tiny sub-polygons -----
  ### create area list for filtering.  Area will be in square degrees... 
  polys_area <- lapply(poly_set@polygons, function(x) sapply(x@Polygons, function(y) y@area))

  if(verbose) {
    message('Polygon area quantiles:')
    print(quantile(unlist(polys_area)))
  }

  ### check to make sure threshold 
#   if(verbose) {
#     message('Smallest "main" subpolygon, i.e. smallest EEZ, in this set:')
#     poly_small <- min(unlist(lapply(polys_area, function(x) (max(x)))))
#     if(poly_small < thresh) message('Smallest poly smaller than threshold value of ', thresh)
#     for (j in 1:length(polys_area)) {
#       message(sprintf('rgn %s: largest polygon %.5f (%s)', poly_set@data$rgn_id[j], 
#                       max(polys_area[[j]]), poly_set@data$rgn_nam[j]))
#     }
#   }

  mainPolys <- lapply(polys_area, function(x) which(x > thresh))

  for(i in 1:length(mainPolys)) {
    if(length(mainPolys[[i]]) >= 1 && mainPolys[[i]][1] >= 1) {
      poly_set@polygons[[i]]@Polygons  <- poly_set@polygons[[i]]@Polygons[mainPolys[[i]]]
      poly_set@polygons[[i]]@plotOrder <- 1:length(poly_set@polygons[[i]]@Polygons)
    }
  }

  plot_order_max <- max(poly_set@plotOrder)
  poly_list_length <- length(poly_set@polygons)
  message('Plot order max: ', plot_order_max, '; length poly list: ', poly_list_length)

  if(plot_order_max != poly_list_length) {
    message('NOT SAME LENGTH?!')
    stop()
    # poly_set@plotOrder <- 1:length(poly_set@polygons)
    # is plotOrder necessary to be continuous?  that would explain this line.
  }

  if(verbose) {
#     message('Counting subpolys in modified poly set:')
#     poly_counts <- sapply(poly_set@polygons, function(x) length(x@Polygons))
#     print(poly_counts)

    polys_area_new <- lapply(poly_set@polygons, function(x) sapply(x@Polygons, function(y) y@area))

    message('Polygon area quantiles after adjustment:')
    print(quantile(unlist(polys_area_new)))
  }

  return(poly_set)
}

```

### 2016-02-25 mpa_analysis/mpa_analysis2.Rmd

--- 2016-02-18 mpa_analysis/mpa_analysis.Rmd
@@
 library(stringr)
+library(rgeos)
 
@@
 
-  rast_empty <- raster(ext = extent(rast_loiczid), resolution = 0.01)
+  rast_empty <- raster(ext = extent(rast_loiczid), resolution = 0.02)
   rast_base  <- raster::area(rast_empty) ### takes a while...
@@
 }
-
-### as long as we're here, let's set up some objects for WGS 84 projection
-p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
-crs_wgs84 <- CRS(p4s_wgs84)
 
@@
 
-rast_marine_file <- file.path(dir_git, 'mpa_analysis/rast_marine.tif')
-reload <- FALSE
-if(reload == TRUE) {
-### import OHI regions polygons, to determine spatially-marine areas
-  poly_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters/poly_marine.shp')
-  if(!file.exists(poly_marine_file)) {
+### set up some objects for WGS 84 projection
+p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
+crs_wgs84 <- CRS(p4s_wgs84)
+
+poly_dir <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
+rast_dir <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
+
 dir_rgn   <- file.path(dir_N, 'git-annex/globalprep/spatial/v2015/data')
 lyr_rgn <- 'regions_gcs'
-    poly_rgn <- readShapePoly(file.path(dir_rgn, lyr_rgn), 
+poly_marine <- readShapePoly(file.path(dir_rgn, lyr_rgn), 
                              proj4string = crs_wgs84)
@@
 ### the rest as spatially-marine = 1
-    poly_marine <- poly_rgn[!str_detect(poly_rgn@data$rgn_typ, 'land'), ]
-    poly_marine@data <- poly_marine@data %>%
-      mutate(sp_mar = 1)
 
-    poly_mar_union <- raster::union(poly_marine)
-    poly_mar_union <- rgeos::gUnaryUnion(poly_marine)
+poly_marine <- poly_marine[!str_detect(poly_marine@data$rgn_typ, 'land'), ]
+poly_list <- list()
+poly_list[['poly_fao']]     <- poly_marine[str_detect(poly_marine@data$ant_typ, 'fao'), ]
+poly_list[['poly_ccamlr']]  <- poly_marine[str_detect(poly_marine@data$ant_typ, 'ccamlr'), ]
+poly_list[['poly_disp']]    <- poly_marine[str_detect(poly_marine@data$ant_typ, 'disputed'), ]
+poly_eez    <- poly_marine[!(str_detect(poly_marine@data$ant_typ, 'ccamlr')) & 
+                           !(str_detect(poly_marine@data$ant_typ, 'fao'))   &
+                           !(str_detect(poly_marine@data$ant_typ, 'disputed')), ]
+n_eez <- 5
+for (i in 1:ceiling(nrow(poly_eez)/n_eez)) {
+  min_row <- n_eez * (i - 1) + 1
+  max_row <- n_eez * i
+  poly_list[[paste('poly_eez', i, sep = '')]] <- poly_eez[min_row:max_row, ]
+}
 
-    # writePolyShape(poly_marine, poly_marine_file)
-    writeOGR(obj    = poly_mar_union,
-             dsn    = path.expand(dirname(poly_marine_file)),
-             layer  = str_replace(basename(poly_marine_file), '.shp', ''),
+for (fn in names(poly_list)) { # fn = names(poly_list)[1] 
+  message('Simplifying polygon ', fn)
+  poly_list[[fn]] <- poly_simplify(poly_list[[fn]], thresh = 1e-5, verbose = TRUE)
+
+#   if(!gIsValid(poly_list[[fn]], byid = TRUE, reason = TRUE)) {
+#     message('INVALID POLYGONS FOUND in polygon ', fn)
+#   }
+  message('Writing polygon ', fn)
+  writeOGR(obj    = poly_list[[fn]],
+           dsn    = path.expand(poly_dir),
+           layer  = fn,
            driver = 'ESRI Shapefile',
@@
            verbose = TRUE)
-
-    } else {
-    poly_marine <- readShapePoly(poly_marine_file, 
-                               proj4string = crs_wgs84)
 }
 
-  plot(poly_marine, col = 'blue')
+```
 
-  # rast_marine <- rasterize(poly_marine, rast_base, value = sp_mar_flag) 
-  ### excruciatingly slow (interminable?)... consider gdal_rasterize?
+``` {r turn saved files into rasters}
+poly_names <- names(poly_list)
+remove(poly_list, poly_eez, poly_marine)
 
-  library(gdalUtils)
-  rast_marine <- gdal_rasterize(
-      src_datasource = poly_marine_file,
-      dst_filename   = rast_marine_file, 
+ext  <- c(-180, -90, 180, 90), 
+          # extents for output raster
+reso <- c(.05, .05),
+
+# tmpdir <- '~/home_big/ohara/R_raster_tmp'
+# dir.create(tmpdir, showWarnings = FALSE)
+# rasterOptions(tmpdir = tmpdir)
+
+### gdalUtils::gdal_rasterize() method:
+for (fn in poly_names) { # fn = poly_names[1]
+  message('Rasterizing polygon ', fn)
+
+  src <- path.expand(file.path(poly_dir, paste(fn, '.shp', sep = '')))
+  rfn <- str_replace(fn, 'poly', 'rast') 
+  tmp_dst <- file.path(rast_dir, 'tmp', paste(rfn, '.tif', sep = '')) %>%
+    path.expand
+  fnl_dst <- file.path(rast_dir, paste(rfn, '.tif', sep = '')) %>%
+    path.expand
+
+# if(!file.exists(fnl_dst)) {
+#     if(file.size(src) > 7e6) {
+#       message(sprintf('%s: Too damn big... %.2f', src, file.size(src)/1e6))
+#     } else {
+      ptm <- proc.time()
+      tmp_rast <- gdalUtils::gdal_rasterize(
+        src_datasource = src,
+        dst_filename   = tmp_dst, 
           # destination for output
         a = 'rgn_id', 
-#      a = 'sp_mar', 
           # the attribute in the shapefile to be assigned to the cell values
-      te = c(-180, -90, 180, 90), 
+        te = ext, 
           # extents for output raster
-      tr = c(.01, .01),  
+        tr = reso,  
           # resolution for output raster
-      tap = TRUE, 
-        # target aligned pixels - align coords of extent of output to values of -tr
         a_nodata = NA, 
@@
         verbose = TRUE)
+      message('Rasterized ', rfn, '; ', (proc.time() - ptm)[3], ' seconds')
+
+      ptm <- proc.time()
+      raster::writeRaster(tmp_rast, fnl_dst, overwrite = TRUE)
+      message('Write ', rfn, '; ', (proc.time() - ptm)[3], ' seconds')
+      remove(tmp_rast)
+      unlink(tmp_dst)
+    # }
+#   } else {
+#     message('File ', fnl_dst, ' already exists.')
+#   }
 }
@@
 ```
-
-``` {r set_up_wdpa_raster}
-
-### import WDPA database polygons from original .gdb
-# gdb_wdpa   <- file.path(dir_N, 'git-annex/globalprep/_raw_data/wdpa_mpa/WDPA_Jan2015_Public/WDPA_Jan2015_Public.gdb')
-# lyr_wdpa <- 'WDPA_poly_Jan2015'
-# 
-# poly_wdpa <- readOGR(dsn = gdb_wdpa, layer = lyr_wdpa, stringsAsFactors = FALSE)
-
-### import WDPA database polygons from saved .shp
-dir_wdpa   <- file.path(dir_N, 'git-annex/globalprep/_raw_data/wdpa_mpa/WDPA_Jan2015_Public')
-lyr_wdpa <- 'WDPA_poly_Jan2015'
-poly_wdpa <- readShapePoly(file.path(dir_wdpa, lyr_wdpa), 
-                           proj4string = crs_wgs84)
-
-prot_code_df <- data.frame('IUCN_CLASS' = c('I', 'IIa', 'IIb', 'III', 'IV', 'V', 'V', NA),
-                           'prot_code'  = c( 1,    2,     2,     3,     4,   5,   6,   7))
-poly_wdpa@data <- poly_wdpa@data %>%
-  left_join(prot_code_df, by = 'IUCN_CLASS') %>%
-  mutate(prot_code = prot_code + MARINE)
-
-```

### 2016-02-25 mpa_analysis/mpa_analysis3.Rmd

--- 2016-02-19 mpa_analysis/mpa_analysis.Rmd
@@
     poly_marine <- poly_rgn[!str_detect(poly_rgn@data$rgn_typ, 'land'), ]
-    poly_marine@data <- poly_marine@data %>%
-      mutate(sp_mar = 1)
-
-    poly_list <- list()
-    poly_rows <- nrow(poly_marine)
-    n_entities <- 20
-    for (i in 1:ceiling(poly_rows/n_entities)) { # i = 1
-      min_row <- (i - 1) * n_entities + 1
-      max_row <- min(i * n_entities, poly_rows)
-      poly_tmp <- poly_marine[min_row:max_row, ]
-      message(sprintf('%s. Entities [%s:%s]', i, min_row, max_row))
-      print(system.time({
-        poly_tmp2 <- rgeos::gUnaryUnion(poly_tmp)
-      })[3])
-      poly_list[i] <- poly_tmp2
-    }
-
-    poly_mar_frankenstein <- spRbind(poly_list)
-
-    # poly_mar_union <- rgeos::gUnaryUnion(poly_marine)
 
     # writePolyShape(poly_marine, poly_marine_file)
-    writeOGR(obj    = poly_mar_frankenstein,
+    writeOGR(obj    = poly_marine,
              dsn    = path.expand(dirname(poly_marine_file)),
@@
              verbose = TRUE)
-
-    } else {
-    poly_marine <- readShapePoly(poly_marine_file, 
-                               proj4string = crs_wgs84)
   }
-
-  plot(poly_marine, col = 'blue')
-
-  # rast_marine <- rasterize(poly_marine, rast_base, value = sp_mar_flag) 
-  ### excruciatingly slow (interminable?)... consider gdal_rasterize?
 
@@
       a = 'rgn_id', 
-#      a = 'sp_mar', 
         # the attribute in the shapefile to be assigned to the cell values

### 2016-02-25 mpa_analysis/mpa_gap_analysis.Rmd

@@
 ### Save it as a .tif in GitHub - about 50 
-rast_base_file <- file.path(dir_git, 'mpa_analysis/rast_base.tif')
-
-if(!file.exists(rast_base_file)) {
-  ### import AquaMaps half-degree loiczid raster - for extents and such
-  rast_loiczid_file   <- file.path(dir_git, 'data/rasters/loiczid_raster.grd')
-  rast_loiczid        <- raster(rast_loiczid_file)
-  names(rast_loiczid) <- 'loiczid'
-
-  rast_empty <- raster(ext = extent(rast_loiczid), resolution = 0.01)
-  rast_base  <- raster::area(rast_empty) ### takes a while...
-
-  writeRaster(rast_base, rast_base_file, overwrite = TRUE)
-} else {
-  rast_base <- raster(rast_base_file)
-}
+# rast_base_file <- file.path(dir_git, 'mpa_analysis/rast_base.tif')
+# 
+# if(!file.exists(rast_base_file)) {
+#   ### import AquaMaps half-degree loiczid raster - for extents and such
+#   rast_loiczid_file   <- file.path(dir_git, 'data/rasters/loiczid_raster.grd')
+#   rast_loiczid        <- raster(rast_loiczid_file)
+#   names(rast_loiczid) <- 'loiczid'
+# 
+#   rast_empty <- raster(ext = extent(rast_loiczid), resolution = 0.01)
+#   rast_base  <- raster::area(rast_empty) ### takes a while...
+#   
+#   writeRaster(rast_base, rast_base_file,
+#               progress = 'text', overwrite = TRUE)
+# } else {
+#   rast_base <- raster(rast_base_file)
+# }
 
@@
 rast_mpa_mar_file <- file.path(dir_rast, 'rast_mpa_mar.tif')
+rast_mar_file <- file.path(dir_rast, 'rast_marine.tif')
+
 if(!file.exists(rast_mpa_mar_file)) {
-  rast_mar_file <- file.path(dir_rast, 'rast_marine.tif')
   rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa1.tif')
@@
 
-  rast_mpa_mar <- mask(rast_mpa, rast_mar)
-  writeRaster(rast_mpa_mar, rast_mpa_mar_file)
+  rast_mpa_mar <- mask(rast_mpa, rast_mar, 
+                       filename = rast_mpa_mar_file, 
+                       overwrite = TRUE,
+                       progress = 'text')
 } else {
@@
 rast_loiczid_01_file <- file.path(dir_rast, 'rast_loiczid_01.tif')
-rast_loiczid_01 <- resample(rast_loiczid, rast_mpa_mar, method = 'ngb')
-writeRaster(rast_loiczid_01, rast_loiczid_01_file)
+if(!file.exists(rast_loiczid_01_file)) {
+  rast_loiczid_01 <- resample(rast_loiczid, rast_mpa_mar, 
+                            method = 'ngb',
+                            progress = 'text',
+                            filename = rast_loiczid_01_file)
+# writeRaster(rast_loiczid_01, rast_loiczid_01_file, progress = 'text')
+} else {
+  rast_loiczid_01 <- raster(rast_loiczid_01_file)
+}
+```
 
-### Turn all marine cells to value of 1 (so can uze zonal sum)
-rast_is_mar <- rast_mar
-values(rast_is_mar)[!is.na(values(rast_mar))] <- 1
-names(rast_is_mar) <- 'is_marine'
+``` {r summarize counts of marine and MPAs by cell}
 
-### Turn all MPA cells to value of 1 (so can use zonal sum)
-### NOTE: loses status_yr values - so filter by year first
+### Count marine cells in each LOICZID cell
+mar_df_file <- file.path(dir_git, 'mpa_analysis/loiczid_marine.csv')
+if(!file.exists(mar_df_file)) {
+  is_marine_df <- zonal(rast_mar, 
+                        z = rast_loiczid_01, 
+                        fun = 'count',
+                        progress = 'text') %>%
+    as.data.frame() %>%
+    rename(loiczid = zone)
+  write_csv(is_marine_df, mar_df_file)
+}
+
+### Count all protected cells in each LOICZID cell
+### NOTE: loses status_yr values - so filter by year for each iteration
+yr_list <- c('y1990', 'y1995', 'y2000', 'y2005', 'y2010', 'y2014', 'y2015')
+mpa_csvs_missing <- !file.exists(file.path(dir_git, sprintf('mpa_analysis/loiczid_mpa_%s.csv', yr_list)))
+
+if(any(mpa_csvs_missing)) {
+  yr_list1 <- yr_list[mpa_csvs_missing]
+  message('Creating marine protected area rasters for years:')
+  message('  ', paste(yr_list, collapse = ', '))
+  for (yr in yr_list1) { # yr <- yr_list1[1]
+    message('filtering MPAs to those designated prior to ', str_replace(yr, 'y', ''))
     rast_is_mpa <- rast_mpa_mar
-values(rast_is_mpa)[!is.na(values(rast_mpa_mar))] <- 1
-names(rast_is_mpa) <- 'is_mpa'
+    data_year <- as.integer(str_replace(yr, 'y', ''))
+    values(rast_is_mpa)[values(rast_mpa_mar) >= data_year] <- NA
+      ### NOTE: 2015 MPA dataset includes up to 2014; 2014 dataset goes
+      ###   through 2013; etc.
+    message('counting protected cells per LOICZID for ', yr)
+    is_mpa_df <- zonal(rast_is_mpa, 
+                       z = rast_loiczid_01, 
+                       fun = 'count',
+                       progress = 'text') %>%
+      as.data.frame() %>%
+      rename(loiczid = zone)
+    write_csv(is_mpa_df, 
+              file.path(dir_git, sprintf('mpa_analysis/loiczid_mpa_%s.csv', yr)))
+  }
+}
+```
 
-rast_test <- rast_test2 <- raster(ext = extent(rast_loiczid), resolution = c(10, 10))
-values(rast_test) <- 1:648
-values(rast_test2) <- rep(c(1:10), each = 64, length.out = length(rast_test))
-x <- zonal(rast_test, rast_test2, fun = sum)
+First step:
+
+* Create data frame of LOICZID vs MPA coverage and marine area
+    * note: using raster-generated marine area, not AquaMaps reported marine
+      area - need to look into how different these are.
+        * 7% of cells differ in calculated vs reported area
+        * most of these differences cluster very near zero
+
+For each species:
+
+* for each "present" LOICZID cell
+    * determine total area: overall cell area (in km2) * marine area (cells/2500)
+    * determine protected area: overall cell area (in km2) * protected area (cells/2500)
+* sum total area and protected area across all "present" LOICZID cells
+* report: 
+    * ratio of (protected area:total area) (dimensionless, percentage)
+    * total marine area and total protected area (km2)
+    * species name, ID, taxonomic group, extinction risk
+
+``` {r combine_mpa_areas}
+
+dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv/')
+
+hcaf <- read_csv(file.path(dir_am_data, 'hcaf_truncated.csv')) %>%
+  select(loiczid, cellarea, oceanarea)
+
+cell_df <- read_csv(file.path(dir_git, 'mpa_analysis/loiczid_marine.csv'),
+  col_types = 'dd') %>%
+  rename(marine_ct = count) %>%
+  left_join(hcaf, by = 'loiczid') %>%
+  mutate(marine_pct  = marine_ct/2500,
+         marine_area = marine_pct * cellarea) %>%
+  select(-marine_ct)
+
+for (yr in yr_list) { # yr <- yr_list[1]
+  fn <- file.path(dir_git, sprintf('mpa_analysis/loiczid_mpa_%s.csv', yr))
+  df <- read_csv(fn, col_types = 'dd')
+  names(df)[2] <- yr
+  cell_df <- cell_df %>%
+    left_join(df, by = 'loiczid')
+  cell_df[ , paste(yr, '_pct', sep = '')]  <- cell_df[ , yr]/2500
+  cell_df[ , paste(yr, '_area', sep = '')] <- cell_df[ , paste(yr, '_pct', sep = '')] * cell_df$cellarea
+
+  ### remove intermediate columns
+  cell_df[ , yr] <- NULL
+  cell_df[ , paste(yr, '_pct', sep = '')] <- NULL
+}
+
+# ### check diffs in oceanarea vs marine_area
+# x <- cell_df %>% 
+#   filter(oceanarea != marine_area) %>%
+#   select(loiczid, oceanarea, marine_area, cellarea)
+# ### 18158 cells out of 259200 = 7%
+# x <- x %>% mutate(pct_diff = (marine_area - oceanarea)/cellarea)
+# hist(x$pct_diff)
+# ### most differences cluster right around zero.  Not going to worry about it!
+
+```
+
+``` {r apply loiczid area df to species list}
+
+### Assemble the AM species-cell df with info on marine area and 
+### protected area per cell
+am_spp_cells <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv'))
+am_spp_cells <- am_spp_cells %>%
+  left_join(cell_df, by = 'loiczid')
+
+### filter to the appropriate threshold...
+thresh <- 0.50
+
+am_spp_cells_filt <- am_spp_cells %>%
+  filter(probability >= thresh)
+
+### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
+am_spp_mpas <- am_spp_cells %>%
+  group_by(speciesid) %>%
+  summarize(cellarea_tot = sum(cellarea), 
+            ocean_range  = sum(oceanarea), 
+            marine_range = sum(marine_area),
+            mpa1990 = sum(y1990_area),
+            mpa1995 = sum(y1995_area),
+            mpa2000 = sum(y2000_area),
+            mpa2005 = sum(y2005_area),
+            mpa2010 = sum(y2010_area),
+            mpa2014 = sum(y2014_area),
+            mpa2015 = sum(y2015_area),
+            pct1990 = mpa1990/marine_range,
+            pct1995 = mpa1995/marine_range,
+            pct2000 = mpa2000/marine_range,
+            pct2005 = mpa2005/marine_range,
+            pct2010 = mpa2010/marine_range,
+            pct2014 = mpa2014/marine_range,
+            pct2015 = mpa2015/marine_range)
+```
+
+``` {r recreate Klein results}
+
+area_mpa <- cell_df %>%
+  summarize(marine_area  = sum(marine_area), 
+            mpa_tot_1990 = sum(y1990_area), 
+            mpa_tot_1995 = sum(y1995_area), 
+            mpa_tot_2000 = sum(y2000_area), 
+            mpa_tot_2005 = sum(y2005_area), 
+            mpa_tot_2010 = sum(y2010_area), 
+            mpa_tot_2014 = sum(y2014_area), 
+            mpa_tot_2015 = sum(y2015_area))
+#   marine_area  mpa_tot_1990  mpa_tot_1995  mpa_tot_2000  mpa_tot_2005   mpa_tot_2010   mpa_tot_2014
+#   362,657,307       351,592       532,098       748,172       932,817      1,137,142      2,885,465
+
+### Recreate Klein results, using 2014 AquaMaps
+mpa_klein_am2014 <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc_2014.csv'))
+mpa_klein_am2014 <- mpa_klein_am2014 %>%
+  left_join(cell_df, by = 'loiczid')
+
+### filter to the appropriate threshold...
+thresh <- 0.50
+
+mpa_klein_am2014_filt <- mpa_klein_am2014 %>%
+  filter(probability >= thresh)
+
+### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
+mpa_klein_am2014 <- mpa_klein_am2014 %>%
+  group_by(speciesid) %>%
+  summarize(marine_range = sum(marine_area),
+            mpa2014 = sum(y2014_area),
+            pct2014 = mpa2014/marine_range)
+
+mpa_klein_am2014 <- mpa_klein_am2014 %>%
+  mutate(cat2014 = ifelse(pct2014 == 0,                       'gap', NA),
+         cat2014 = ifelse(pct2014 > 0    & pct2014 <= 0.02,  '0-2%', cat2014),
+         cat2014 = ifelse(pct2014 > 0.02 & pct2014 <= 0.05,  '2-5%', cat2014),
+         cat2014 = ifelse(pct2014 > 0.05 & pct2014 <= 0.10, '5-10%', cat2014),
+         cat2014 = ifelse(pct2014 > 0.10,                    '>10%', cat2014))
+
+mpa_klein_orig <- mpa_klein_am2014 %>%
+  group_by(cat2014) %>%
+  summarize(n_tot = nrow(mpa_klein_am2014), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
+
+#   cat2014 n_tot n_cat pct_cat
+#       gap 17348   378     2.2
+#      0-2% 17348  8859    51.2
+#      2-5% 17348  6728    38.7
+#     5-10% 17348  1096     6.3
+#      >10% 17348   287     1.7
+
+### Redo analysis using 50% threshold, but new AquaMaps database
+### * select just 2014 and 2015, at 50% threshold, and group by
+###   same categories that the original paper did.
+mpa_klein_am2015 <- am_spp_mpas %>%
+  select(speciesid, marine_range, mpa2014, mpa2015, pct2014, pct2015)
+
+mpa_klein_am2015 <- mpa_klein_am2015 %>%
+  mutate(cat2014 = ifelse(pct2014 == 0,                       'gap', NA),
+         cat2014 = ifelse(pct2014 > 0    & pct2014 <= 0.02,  '0-2%', cat2014),
+         cat2014 = ifelse(pct2014 > 0.02 & pct2014 <= 0.05,  '2-5%', cat2014),
+         cat2014 = ifelse(pct2014 > 0.05 & pct2014 <= 0.10, '5-10%', cat2014),
+         cat2014 = ifelse(pct2014 > 0.10,                    '>10%', cat2014),
+         cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
+         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
+         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
+         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
+         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))
+
+mpa_klein_sum2014 <- mpa_klein_am2015 %>%
+  group_by(cat2014) %>%
+  summarize(n_tot = nrow(mpa_klein_am2015), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
+
+mpa_klein_sum2015 <- mpa_klein_am2015 %>%
+  group_by(cat2015) %>%
+  summarize(n_tot = nrow(mpa_klein_am2015), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
+
+#   cat2014 n_tot n_cat pct_cat
+#       gap 22889   552    2.4
+#      0-2% 22889 11546   50.4
+#      2-5% 22889  8947   39.1
+#     5-10% 22889  1456    6.4
+#      >10% 22889   388    1.7
+
+#   cat2015 n_tot n_cat pct_cat
+#       gap 22889   552     2.4
+#      0-2% 22889 11520    50.3
+#      2-5% 22889  8960    39.1
+#     5-10% 22889  1466     6.4
+#      >10% 22889   391     1.7
+
+```
+
+### Original study (Klein 2015) vs new analyses
+
+| study           | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
+| :-------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
+| Klein 2015      | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
+| am2014, mpa2014 | IUCN I-IV, Sp. Mar |   2,885,465  | 2.2% | 51.1% | 38.8% |  6.3% | 1.7% |
+| am2015, mpa2015 | IUCN I-IV, Sp. Mar |   2,889,758  | 2.4% | 50.3% | 39.1% |  6.4% | 1.7% |
+| am2015, mpa2014 | IUCN I-IV, Sp. Mar |   2,885,465  | 2.4% | 50.4% | 39.1% |  6.4% | 1.7% |
+| am2015, mpa2000 | IUCN I-IV, Sp. Mar |     748,172  | 3.5% | 91.6% |  4.4% |  0.4% | 0.1% |
+| am2015, mpa1990 | IUCN I-IV, Sp. Mar |     351,592  | 4.0% | 94.2% |  1.6% |  0.1% | 0.0% |
+
+``` {r check results for 1990 and 2000}
+mpa_spp <- am_spp_mpas %>%
+  select(speciesid, marine_range, mpa1990, mpa2000, pct1990, pct2000)
+
+mpa_spp <- mpa_spp %>%
+  mutate(cat1990 = ifelse(pct1990 == 0,                       'gap', NA),
+         cat1990 = ifelse(pct1990 > 0    & pct1990 <= 0.02,  '0-2%', cat1990),
+         cat1990 = ifelse(pct1990 > 0.02 & pct1990 <= 0.05,  '2-5%', cat1990),
+         cat1990 = ifelse(pct1990 > 0.05 & pct1990 <= 0.10, '5-10%', cat1990),
+         cat1990 = ifelse(pct1990 > 0.10,                    '>10%', cat1990),
+         cat1995 = ifelse(pct1995 == 0,                       'gap', NA),
+         cat1995 = ifelse(pct1995 > 0    & pct1995 <= 0.02,  '0-2%', cat1995),
+         cat1995 = ifelse(pct1995 > 0.02 & pct1995 <= 0.05,  '2-5%', cat1995),
+         cat1995 = ifelse(pct1995 > 0.05 & pct1995 <= 0.10, '5-10%', cat1995),
+         cat1995 = ifelse(pct1995 > 0.10,                    '>10%', cat1995),
+         cat2000 = ifelse(pct2000 == 0,                       'gap', NA),
+         cat2000 = ifelse(pct2000 > 0    & pct2000 <= 0.02,  '0-2%', cat2000),
+         cat2000 = ifelse(pct2000 > 0.02 & pct2000 <= 0.05,  '2-5%', cat2000),
+         cat2000 = ifelse(pct2000 > 0.05 & pct2000 <= 0.10, '5-10%', cat2000),
+         cat2000 = ifelse(pct2000 > 0.10,                    '>10%', cat2000),
+         cat2005 = ifelse(pct2005 == 0,                       'gap', NA),
+         cat2005 = ifelse(pct2005 > 0    & pct2005 <= 0.02,  '0-2%', cat2005),
+         cat2005 = ifelse(pct2005 > 0.02 & pct2005 <= 0.05,  '2-5%', cat2005),
+         cat2005 = ifelse(pct2005 > 0.05 & pct2005 <= 0.10, '5-10%', cat2005),
+         cat2005 = ifelse(pct2005 > 0.10,                    '>10%', cat2005),
+         cat2010 = ifelse(pct2010 == 0,                       'gap', NA),
+         cat2010 = ifelse(pct2010 > 0    & pct2010 <= 0.02,  '0-2%', cat2010),
+         cat2010 = ifelse(pct2010 > 0.02 & pct2010 <= 0.05,  '2-5%', cat2010),
+         cat2010 = ifelse(pct2010 > 0.05 & pct2010 <= 0.10, '5-10%', cat2010),
+         cat2010 = ifelse(pct2010 > 0.10,                    '>10%', cat2010))
+
+mpa_spp_1990 <- mpa_spp %>%
+  group_by(cat1990) %>%
+  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
+
+mpa_spp_2000 <- mpa_spp %>%
+  group_by(cat2000) %>%
+  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
+
+#   cat1990 n_tot n_cat pct_cat
+#       gap 22889   925     4.0
+#      0-2% 22889 21569    94.2
+#      2-5% 22889   376     1.6
+#     5-10% 22889    18     0.1
+#      >10% 22889     1     0.0
+# 
+#   cat2000 n_tot n_cat pct_cat
+#       gap 22889   801     3.5
+#      0-2% 22889 20973    91.6
+#      2-5% 22889  1004     4.4
+#     5-10% 22889    92     0.4
+#      >10% 22889    19     0.1
+
+```

### 2016-03-01 mpa_analysis/mpa_gap_analysis.Rmd
let's get some of this up on line...

@@
 
-``` {r set_up_base_raster}
-### DO WE NEED THIS? Use area at 0.5 degree level; most cells this will be pretty close.
-### base raster will contain 0.01 degree cells from -180 to +180,
-### and -90 to +90, with the associated area of each cell.
-### Save it as a .tif in GitHub - about 50 
-# rast_base_file <- file.path(dir_git, 'mpa_analysis/rast_base.tif')
-# 
-# if(!file.exists(rast_base_file)) {
-#   ### import AquaMaps half-degree loiczid raster - for extents and such
-#   rast_loiczid_file   <- file.path(dir_git, 'data/rasters/loiczid_raster.grd')
-#   rast_loiczid        <- raster(rast_loiczid_file)
-#   names(rast_loiczid) <- 'loiczid'
-# 
-#   rast_empty <- raster(ext = extent(rast_loiczid), resolution = 0.01)
-#   rast_base  <- raster::area(rast_empty) ### takes a while...
-#   
-#   writeRaster(rast_base, rast_base_file,
-#               progress = 'text', overwrite = TRUE)
-# } else {
-#   rast_base <- raster(rast_base_file)
-# }
-
-```
-
 ``` {r load_rasters}
@@
 dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
-rast_mpa_mar_file <- file.path(dir_rast, 'rast_mpa_mar.tif')
+rast_mpa_mar_file <- file.path(dir_rast, 'rast_mpa_mar_i-iv.tif')
+# rast_mpa_mar_file <- file.path(dir_rast, 'wdpa_all_mar_klein_i-iv.tif')
 rast_mar_file <- file.path(dir_rast, 'rast_marine.tif')
@@
 if(!file.exists(rast_mpa_mar_file)) {
-  rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa1.tif')
+#   rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa_i-iv-des.tif')
+   rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa_i-iv-all.tif')
 
@@
                             filename = rast_loiczid_01_file)
-# writeRaster(rast_loiczid_01, rast_loiczid_01_file, progress = 'text')
 } else {
@@
 ### Count marine cells in each LOICZID cell
-mar_df_file <- file.path(dir_git, 'mpa_analysis/loiczid_marine.csv')
+mar_df_file <- file.path(dir_git, 'mpa_analysis/celldata/loiczid_marine.csv')
 if(!file.exists(mar_df_file)) {
@@
 yr_list <- c('y1990', 'y1995', 'y2000', 'y2005', 'y2010', 'y2014', 'y2015')
-mpa_csvs_missing <- !file.exists(file.path(dir_git, sprintf('mpa_analysis/loiczid_mpa_%s.csv', yr_list)))
+mpa_cvs_files <- c(file.path(dir_git, 
+                             sprintf('mpa_analysis/celldata/loiczid_mpa_i-iv_%s.csv', 
+                                     yr_list)))
+# mpa_cvs_files <- c(file.path(dir_git, 
+#                              sprintf('mpa_analysis/celldata/loiczid_mpa_klein_i-iv_%s.csv', 
+#                                      yr_list)))
+yr_list1 <- yr_list[!file.exists(mpa_cvs_files)]
 
-if(any(mpa_csvs_missing)) {
-  yr_list1 <- yr_list[mpa_csvs_missing]
+if(length(yr_list1) > 0) {
   message('Creating marine protected area rasters for years:')
-  message('  ', paste(yr_list, collapse = ', '))
+  message('  ', paste(yr_list1, collapse = ', '))
   for (yr in yr_list1) { # yr <- yr_list1[1]
@@
     write_csv(is_mpa_df, 
-              file.path(dir_git, sprintf('mpa_analysis/loiczid_mpa_%s.csv', yr)))
+              mpa_cvs_files[yr_list == yr])
   }
 }
+### These files get loaded later; no need to hog memory by loading them here.
+
 ```
@@
 dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv/')
+yr_list <- c('y1990', 'y1995', 'y2000', 'y2005', 'y2010', 'y2014', 'y2015')
 
+### get AquaMaps file of cells to translate cell IDs to area and ocean area
 hcaf <- read_csv(file.path(dir_am_data, 'hcaf_truncated.csv')) %>%
@@
 
-cell_df <- read_csv(file.path(dir_git, 'mpa_analysis/loiczid_marine.csv'),
+### attach the AquaMaps areas to the calculated percentages of marine area
+### note: AquaMaps calculated areas differ from my calculated areas - see below
+cell_df <- read_csv(file.path(dir_git, 'mpa_analysis/celldata/loiczid_marine.csv'),
   col_types = 'dd') %>%
@@
 
+### load each analysis year (in yr_list) and join them all together
+### to note protection status year of each cell.
+
+### Use this to point toward the 'i-iv-all' files (for all STATUS) or
+### 'i-iv-des' (for STATUS == 'Designated')
+mpa_file_base <- 'mpa_analysis/celldata/loiczid_mpa_i-iv_%s.csv'
+# mpa_file_base <- 'mpa_analysis/celldata/loiczid_mpa_klein_i-iv_%s.csv'
+# mpa_file_base <- 'mpa_analysis/celldata/loiczid_mpa_i-iv-des_%s.csv'
+
 for (yr in yr_list) { # yr <- yr_list[1]
-  fn <- file.path(dir_git, sprintf('mpa_analysis/loiczid_mpa_%s.csv', yr))
+  fn <- file.path(dir_git, sprintf(mpa_file_base, yr))
   df <- read_csv(fn, col_types = 'dd')
@@
 
+# recreate Klein results using aquamaps 2014 data
+
+``` {r recreate Klein results}
+
+area_mpa <- cell_df %>%
+  summarize(marine_area  = sum(marine_area), 
+            mpa_tot_1990 = sum(y1990_area), 
+            mpa_tot_1995 = sum(y1995_area), 
+            mpa_tot_2000 = sum(y2000_area), 
+            mpa_tot_2005 = sum(y2005_area), 
+            mpa_tot_2010 = sum(y2010_area), 
+            mpa_tot_2014 = sum(y2014_area), 
+            mpa_tot_2015 = sum(y2015_area))
+
+### Recreate Klein results, using 2014 AquaMaps
+mpa_klein_am2014 <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc_2014.csv'))
+mpa_klein_am2014 <- mpa_klein_am2014 %>%
+  left_join(cell_df, by = 'loiczid')
+
+### filter to the appropriate threshold...
+thresh <- 0.50
+
+mpa_klein_am2014 <- mpa_klein_am2014 %>%
+  filter(probability >= thresh)
+
+### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
+mpa_klein_am2014 <- mpa_klein_am2014 %>%
+  group_by(speciesid) %>%
+  summarize(marine_range = sum(marine_area),
+            mpa2014 = sum(y2014_area),
+            pct2014 = mpa2014/marine_range)
+
+mpa_klein_am2014 <- mpa_klein_am2014 %>%
+  mutate(cat2014 = ifelse(pct2014 == 0,                       'gap', NA),
+         cat2014 = ifelse(pct2014 > 0    & pct2014 <= 0.02,  '0-2%', cat2014),
+         cat2014 = ifelse(pct2014 > 0.02 & pct2014 <= 0.05,  '2-5%', cat2014),
+         cat2014 = ifelse(pct2014 > 0.05 & pct2014 <= 0.10, '5-10%', cat2014),
+         cat2014 = ifelse(pct2014 > 0.10,                    '>10%', cat2014))
+
+mpa_klein_orig <- mpa_klein_am2014 %>%
+  group_by(cat2014) %>%
+  summarize(n_tot = nrow(mpa_klein_am2014), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
+
+### Not filtering on status, but all in shapefile are status == 'Designated'
+#   cat2014 n_tot n_cat pct_cat
+#       gap 17344   282     1.6
+#      0-2% 17344  7728    44.6
+#      2-5% 17344  7738    44.6
+#     5-10% 17344  1188     6.8
+#      >10% 17344   408     2.4
+
+```
+
+# run analysis using 2015 aquamaps
+
 ``` {r apply loiczid area df to species list}
@@
 
-### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
-am_spp_mpas <- am_spp_cells %>%
+### group by speciesid, then summarize sum(marine area), 
+### sum(protected area), ratio
+am_spp_mpas_filt <- am_spp_cells_filt %>%
   group_by(speciesid) %>%
@@
 
-``` {r recreate Klein results}
+``` {r redo klein for new AquaMaps with 2015 mpas}
 
@@
             mpa_tot_2015 = sum(y2015_area))
-#   marine_area  mpa_tot_1990  mpa_tot_1995  mpa_tot_2000  mpa_tot_2005   mpa_tot_2010   mpa_tot_2014
-#   362,657,307       351,592       532,098       748,172       932,817      1,137,142      2,885,465
 
-### Recreate Klein results, using 2014 AquaMaps
-mpa_klein_am2014 <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc_2014.csv'))
-mpa_klein_am2014 <- mpa_klein_am2014 %>%
-  left_join(cell_df, by = 'loiczid')
-
-### filter to the appropriate threshold...
-thresh <- 0.50
-
-mpa_klein_am2014_filt <- mpa_klein_am2014 %>%
-  filter(probability >= thresh)
-
-### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
-mpa_klein_am2014 <- mpa_klein_am2014 %>%
-  group_by(speciesid) %>%
-  summarize(marine_range = sum(marine_area),
-            mpa2014 = sum(y2014_area),
-            pct2014 = mpa2014/marine_range)
-
-mpa_klein_am2014 <- mpa_klein_am2014 %>%
-  mutate(cat2014 = ifelse(pct2014 == 0,                       'gap', NA),
-         cat2014 = ifelse(pct2014 > 0    & pct2014 <= 0.02,  '0-2%', cat2014),
-         cat2014 = ifelse(pct2014 > 0.02 & pct2014 <= 0.05,  '2-5%', cat2014),
-         cat2014 = ifelse(pct2014 > 0.05 & pct2014 <= 0.10, '5-10%', cat2014),
-         cat2014 = ifelse(pct2014 > 0.10,                    '>10%', cat2014))
-
-mpa_klein_orig <- mpa_klein_am2014 %>%
-  group_by(cat2014) %>%
-  summarize(n_tot = nrow(mpa_klein_am2014), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
-
-#   cat2014 n_tot n_cat pct_cat
-#       gap 17348   378     2.2
-#      0-2% 17348  8859    51.2
-#      2-5% 17348  6728    38.7
-#     5-10% 17348  1096     6.3
-#      >10% 17348   287     1.7
+#   marine_area mpa_tot_1990 mpa_tot_1995 mpa_tot_2000 mpa_tot_2005 mpa_tot_2010 mpa_tot_2014 mpa_tot_2015
+#         (dbl)        (dbl)        (dbl)        (dbl)        (dbl)        (dbl)        (dbl)        (dbl)
+# 1   362657307     353355.3     534137.2     750217.9     935137.1      1140272      2888600      2896875
 
@@
 ###   same categories that the original paper did.
-mpa_klein_am2015 <- am_spp_mpas %>%
+mpa_klein_am2015 <- am_spp_mpas_filt %>%
   select(speciesid, marine_range, mpa2014, mpa2015, pct2014, pct2015)
@@
 
+# > mpa_klein_sum2014
 #   cat2014 n_tot n_cat pct_cat
-#       gap 22889   552    2.4
-#      0-2% 22889 11546   50.4
-#      2-5% 22889  8947   39.1
-#     5-10% 22889  1456    6.4
-#      >10% 22889   388    1.7
-
+#       gap 22885   797     3.5
+#      0-2% 22885 11038    48.2
+#      2-5% 22885  8713    38.1
+#     5-10% 22885  1650     7.2
+#      >10% 22885   687     3.0
+# 
+# > mpa_klein_sum2015
 #   cat2015 n_tot n_cat pct_cat
-#       gap 22889   552     2.4
-#      0-2% 22889 11520    50.3
-#      2-5% 22889  8960    39.1
-#     5-10% 22889  1466     6.4
-#      >10% 22889   391     1.7
+#       gap 22885   797     3.5
+#      0-2% 22885 10990    48.0
+#      2-5% 22885  8747    38.2
+#     5-10% 22885  1661     7.3
+#      >10% 22885   690     3.0
 
@@
 
-### Original study (Klein 2015) vs new analyses
-
-| study           | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
-| :-------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
-| Klein 2015      | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
-| am2014, mpa2014 | IUCN I-IV, Sp. Mar |   2,885,465  | 2.2% | 51.1% | 38.8% |  6.3% | 1.7% |
-| am2015, mpa2015 | IUCN I-IV, Sp. Mar |   2,889,758  | 2.4% | 50.3% | 39.1% |  6.4% | 1.7% |
-| am2015, mpa2014 | IUCN I-IV, Sp. Mar |   2,885,465  | 2.4% | 50.4% | 39.1% |  6.4% | 1.7% |
-| am2015, mpa2000 | IUCN I-IV, Sp. Mar |     748,172  | 3.5% | 91.6% |  4.4% |  0.4% | 0.1% |
-| am2015, mpa1990 | IUCN I-IV, Sp. Mar |     351,592  | 4.0% | 94.2% |  1.6% |  0.1% | 0.0% |
-
-``` {r check results for 1990 and 2000}
-mpa_spp <- am_spp_mpas %>%
-  select(speciesid, marine_range, mpa1990, mpa2000, pct1990, pct2000)
+``` {r check results for 1990 through 2010 at 50%}
+mpa_spp <- am_spp_mpas_filt %>%
+  select(speciesid, marine_range, mpa1990:mpa2010, pct1990:pct2010)
 
@@
 
+mpa_spp_1995 <- mpa_spp %>%
+  group_by(cat1995) %>%
+  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
+
 mpa_spp_2000 <- mpa_spp %>%
   group_by(cat2000) %>%
+  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
+
+mpa_spp_2005 <- mpa_spp %>%
+  group_by(cat2005) %>%
+  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
+
+mpa_spp_2010 <- mpa_spp %>%
+  group_by(cat2010) %>%
   summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
@@
 # 
+#   cat1995 n_tot n_cat pct_cat
+#       gap 22889   878     3.8
+#      0-2% 22889 21111    92.2
+#      2-5% 22889   817     3.6
+#     5-10% 22889    70     0.3
+#      >10% 22889    13     0.1
+#
 #   cat2000 n_tot n_cat pct_cat
@@
 #      >10% 22889    19     0.1
+#
+#   cat2005 n_tot n_cat pct_cat
+#       gap 22889   608     2.7
+#      0-2% 22889 19828    86.6
+#      2-5% 22889  2174     9.5
+#     5-10% 22889   205     0.9
+#      >10% 22889    74     0.3
+# 
+#   cat2010 n_tot n_cat pct_cat
+#       gap 22889   578     2.5
+#      0-2% 22889 18670    81.6
+#      2-5% 22889  3278    14.3
+#     5-10% 22889   288     1.3
+#      >10% 22889    75     0.3
+
+argh <- am_spp_mpas_filt %>%
+  select(-(mpa1990:mpa2015)) %>%
+  gather(key = year, value = pct, pct1990:pct2015) %>%
+  mutate(year = str_replace(year, 'pct', '')) %>%
+  group_by(year)
+
+hist_plot <- ggplot(data = argh, aes(x = pct, color = year)) + 
+  geom_density() +
+  scale_x_continuous(limits = c(0, .05))
+#  scale_x_log10()
+
+hist_plot
+ggsave(file.path(dir_git, 'figures/mpa_analysis_dist.png'))
+```
+
+### Original study (Klein 2015) vs new analyses
+
+| study           | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
+| :-------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
+| Klein et al 2015| IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
+| am2014, mpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
+| .               | .                  | .            | .    | .     | .     | .     | .    |
+| am2015, mpa2015 | IUCN I-IV, Sp. Mar |   2,889,758  | 2.4% | 50.3% | 39.1% |  6.4% | 1.7% |
+| am2015, mpa2014 | IUCN I-IV, Sp. Mar |   2,885,465  | 2.4% | 50.4% | 39.1% |  6.4% | 1.7% |
+| am2015, mpa2010 | IUCN I-IV, Sp. Mar |   1,137,142  | 2.5% | 81.6% | 14.3% |  1.3% | 0.3% |
+| am2015, mpa2005 | IUCN I-IV, Sp. Mar |     932,817  | 2.7% | 86.6% |  9.5% |  0.9% | 0.3% |
+| am2015, mpa2000 | IUCN I-IV, Sp. Mar |     748,172  | 3.5% | 91.6% |  4.4% |  0.4% | 0.1% |
+| am2015, mpa1995 | IUCN I-IV, Sp. Mar |     532,098  | 3.8% | 92.2% |  3.6% |  0.3% | 0.1% |
+| am2015, mpa1990 | IUCN I-IV, Sp. Mar |     351,592  | 4.0% | 94.2% |  1.6% |  0.1% | 0.0% |
+
+max protected species = Fis-160882
+
+``` {r iucn_only}
+### compare results using only IUCN species maps
+dir_iucn_data <- '/var/data/ohi/git-annex/globalprep/_raw_data/iucn_spp'
+iucn_cells_file <- file.path(dir_iucn_data, 'iucn_cells_2015.csv')
+iucn_spp_cells <- read_csv(iucn_cells_file, col_types = 'cddddc') %>%
+  as.data.frame(stringsAsFactors = FALSE)
+
+# iucn_spp_cells <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv'))
+iucn_spp_cells <- iucn_spp_cells %>%
+  left_join(cell_df, by = 'loiczid')
+
+### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
+iucn_spp_sum <- iucn_spp_cells %>%
+  group_by(iucn_sid) %>%
+  summarize(cellarea_tot = sum(cellarea), 
+            ocean_range  = sum(oceanarea), 
+            marine_range = sum(marine_area),
+            mpa2015 = sum(y2015_area),
+            pct2015 = mpa2015/marine_range)
+
+iucn_spp_sum <- iucn_spp_sum %>%
+  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
+         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
+         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
+         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
+         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))
+
+iucn_spp_sum2 <- iucn_spp_sum %>%
+  group_by(cat2015) %>%
+  summarize(n_tot = nrow(iucn_spp_sum), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))
 
 ```
+
+``` {r combine IUCN and AM with 0% thresh, AM on top}
+
+### use iucn_spp_sum from above, and create am_spp_sum for 0% thresh.
+iucn_spp_sum
+
+am_spp_mpas_0 <- am_spp_cells %>%
+  group_by(speciesid) %>%
+  summarize(cellarea_tot = sum(cellarea), 
+            ocean_range  = sum(oceanarea), 
+            marine_range = sum(marine_area),
+            mpa2015 = sum(y2015_area),
+            pct2015 = mpa2015/marine_range)
+
+am_spp_sum <- am_spp_sum %>%
+  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
+         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
+         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
+         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
+         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))
+
+# combine by bringing in species list with IUCN SID mapped to AM SID;
+# use all with AM SID, and then add in IUCN spp that are not mapped to an AM SID
+
+```
+
+``` {r get_iucn_cells}
+reload = TRUE
+  dir_iucn_data <- '/var/data/ohi/git-annex/globalprep/_raw_data/iucn_spp'
+  iucn_cells_file <- file.path(dir_iucn_data, 'iucn_cells_2015.csv')
+  if(!file.exists(iucn_cells_file) | reload) {
+    message(sprintf('Building IUCN species to cell table.  This might take a few minutes.\n'))
+    iucn_map_files      <- file.path(dir_anx, 'iucn_intersections', list.files(file.path(dir_anx, 'iucn_intersections')))
+    iucn_cells_spp_list <- lapply(iucn_map_files, read.csv) # read each into dataframe within a list
+    iucn_cells_spp      <- bind_rows(iucn_cells_spp_list)   # combine list of dataframes to single dataframe
+    # This creates a full data frame of all IUCN species, across all species groups, for all cells.
+    # Probably big...
+    names(iucn_cells_spp) <- tolower(names(iucn_cells_spp))
+    write_csv(iucn_cells_spp, iucn_cells_file)
+  } else {
+    iucn_cells_spp <- read_csv(iucn_cells_file, col_types = 'cddddc') %>%
+      as.data.frame(stringsAsFactors = FALSE)
+  }
+
+}
+```

### 2016-03-01 mpa_analysis/wdpa_prep.Rmd
meh
let's get some of this up on line...

@@
 lyr_wdpa    <- 'WDPA_poly_Jan2015'
-lyr_mpa     <- 'WDPA_poly_Jan2015_mpa'
-lyr_mpa_mar <- 'WDPA_poly_Jan2015_mpa_mar'
+lyr_mpa     <- 'WDPA_poly_Jan2015_mpa_i-iv-all'
 
@@
 
-  poly_mpa <- poly_wdpa[poly_wdpa@data$STATUS == 'Designated', ]
-  poly_mpa <- poly_mpa[poly_mpa@data$prot_code <= 4, ]
-
-  poly_mpa_marine <- poly_mpa[poly_mpa@data$MARINE == 1, ]
-
-  writeOGR(poly_mpa, dsn = dir_wdpa, layer = lyr_mpa,
-           driver = 'ESRI Shapefile', overwrite_layer = TRUE)
-  writeOGR(poly_mpa_marine, dsn = dir_wdpa, layer = lyr_mpa_mar,
-           driver = 'ESRI Shapefile', overwrite_layer = TRUE)
-}
-
-poly_mpa <- readOGR(dir_wdpa, lyr_mpa, stringsAsFactors = FALSE)
-# poly_mpa <- readShapePoly(fn = file.path(dir_wdpa, lyr_mpa),
-#                           proj4string = crs_wgs84,
-#                           stringsAsFactors = FALSE)
-
-### create simplified polygon
-polys_mpa_area <- lapply(poly_mpa@polygons, function(x) x@area)
-poly_mpa1 <- poly_mpa[polys_mpa_area > 1e-5, ]
-poly_mpa1 <- poly_simplify(poly_mpa1, thresh = 1e-5, verbose = TRUE)
+  poly_mpa <- poly_wdpa[poly_wdpa@data$prot_code <= 4, ]
+  #  poly_mpa <- poly_mpa[poly_mpa@data$STATUS == 'Designated', ]
 
@@
   ### values then overwrite (where applicable) with older values
-poly_mpa1 <- poly_mpa1[order(poly_mpa1@data$STATUS_YR, decreasing = TRUE), ]
+  poly_mpa <- poly_mpa[order(poly_mpa@data$STATUS_YR, decreasing = TRUE), ]
 
-# writeOGR(poly_mpa1, dsn = dir_wdpa,
-#          layer = paste(lyr_mpa, '1', sep = ''),
-#          driver = 'ESRI Shapefile')
-maptools::writePolyShape(poly_mpa1, 
-                         fn = file.path(dir_wdpa, paste(lyr_mpa, '1', sep = '')))
+  maptools::writePolyShape(poly_mpa, 
+                           fn = file.path(dir_wdpa, lyr_mpa))
   ### copy .prj:
-file.copy(from = file.path(dir_wdpa, paste(lyr_mpa, '.prj', sep = '')), 
-          to = file.path(dir_wdpa, paste(lyr_mpa, '1.prj', sep = '')))
+  file.copy(from = file.path(dir_wdpa, paste(lyr_wdpa, '.prj', sep = '')), 
+            to = file.path(dir_wdpa, paste(lyr_mpa, '.prj', sep = '')))
+
+# writeOGR(poly_mpa, dsn = dir_wdpa, layer = lyr_mpa,
+#            driver = 'ESRI Shapefile', overwrite_layer = TRUE)
+#   writeOGR(poly_mpa_marine, dsn = dir_wdpa, layer = lyr_mpa_mar,
+#            driver = 'ESRI Shapefile', overwrite_layer = TRUE)
+}
 
@@
 ### Get source from WDPA_MPA data location
-src <- path.expand(file.path(dir_wdpa, paste(lyr_mpa, '1.shp', sep = '')))
+src <- path.expand(file.path(dir_wdpa, paste(lyr_mpa, '.shp', sep = '')))
 ### gdal_rasterize to temp directory
-tmp_dst <- file.path(dir_rast, 'tmp', paste(lyr_mpa, '1.tif', sep = '')) %>%
+tmp_dst <- file.path(dir_rast, 'tmp', paste(lyr_mpa, '.tif', sep = '')) %>%
   path.expand
 ### then writeRaster to final directory, to save smaller file
-fnl_dst <- file.path(dir_rast, paste(lyr_mpa, '1.tif', sep = '')) %>%
+fnl_dst <- file.path(dir_rast, paste(lyr_mpa, '.tif', sep = '')) %>%
   path.expand
@@
 ```
+
+``` {r set_up_wdpa_mpa_polys_klein}
+
+### import WDPA database polygons from saved .shp
+dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
+dir_wdpa   <- file.path(dir_rast, 'WDPA2014_mar_klein')
+lyr_wdpa    <- 'wdpa_all_mar_klein'
+lyr_mpa     <- 'wdpa_all_mar_klein_i-iv'
+
+if(!file.exists(file.path(dir_wdpa, paste(lyr_mpa, '.shp', sep = '')))) {
+  poly_wdpa <- readOGR(dsn = dir_wdpa, layer = lyr_wdpa, stringsAsFactors = FALSE)
+
+  prot_code_df <- data.frame('IUCN_CAT'  = c('Ia', 'Ib', 'II', 'III', 'IV', 'V', 'VI'),
+                             'prot_code' = c(  1,    1,    2,     3,    4,   5,    6))
+
+  poly_wdpa@data <- poly_wdpa@data %>%
+    left_join(prot_code_df, by = 'IUCN_CAT') %>%
+    mutate(prot_code = ifelse(is.na(prot_code), 7, prot_code))
+
+  poly_mpa <- poly_wdpa[poly_wdpa@data$prot_code <= 4, ]
+
+  # poly_mpa <- poly_mpa[poly_mpa@data$STATUS == 'Designated', ]
+
+  if(poly_mpa@proj4string@projargs != p4s_wgs84) {
+    message('Polygons not in WGS84 projection.  \n  Polygon projection: ', 
+            poly_mpa@proj4string@projargs)
+    poly_mpa <- spTransform(poly_mpa, crs_wgs84)
+  }
+
+  poly_mpa <- poly_mpa[order(poly_mpa1@data$STATUS_YR, decreasing = TRUE), ]
+
+  writeOGR(poly_mpa, dsn = dir_wdpa, layer = lyr_mpa,
+             driver = 'ESRI Shapefile', overwrite_layer = TRUE)
+  #   writeOGR(poly_mpa_marine, dsn = dir_wdpa, layer = lyr_mpa_mar,
+  #            driver = 'ESRI Shapefile', overwrite_layer = TRUE)
+}
+
+```

### 2016-03-01 ms_6_supplemental.Rmd

@@
 
-## Info on data prep
+#Outline
 
-### processing AquaMaps
+## Data processing
 
-* start with .sql files - three of 'em - how to get 'em?
-* turn into .csvs, which columns are critical for this analysis? simplify
+All code is publicly available and freely downloadable [here on GitHub](link to github).
 
-### downloading and processing IUCN maps
+### AquaMaps
 
-* which data sets are included?
-* raster::extract() to convert polys to csvs
-* which columns are included?
+AquaMaps data was provided by AquaMaps directly via ftp.
+Three SQL files were sent that contained (1) list of all `r nrow(spp_am)` species and their taxonomic information, (2) information for all (292,000?) half degree cells used for range mapping and (3) the probability of occurrence per cell for all `r nrow(spp_am)` species.
+These were extracted into .csv form using R Statistical Software.
 
-### creating master species list for co-listed species
+___Probably just cut, but maybe include in Suppl Materials.
+- BH___: The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
 
-* using AquaMaps and IUCN, and IUCN master list, create the big list
-* which columns are included?
-* adjust this master list to use IUCN SID only for parent - eliminate the whole question of parent/subpop? for the purpose of this analysis
+#### Possible figures
+- Map of all cells
+- Map of # of species per cell
+- Example aquamaps (one species range)
 
-## descriptions of IUCN and AM data sets
+### IUCN
 
@@
 
- As of December 2015, AquaMaps current native distribution maps have been produced for `r nrow(spp_am)` species.
+* which data sets are included?
+* raster::extract() to convert polys to csvs
+* which columns are included?
 
- ___Move to supplementary materials.
-- BH___: For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
+JA: I think here we only need a few sentences describing how we downloaded the data and on what date. Link to the IUCN spatial data website.
+
+IUCN data is provided as polygon shapefiles. These vectors were rasterized to the same cell resolution as AquaMaps data (0.5 degrees).
+
+#### Possible figures
+
+#### Caveats
+For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
 However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
-
-___Probably just cut, but maybe include in Suppl Materials.
-- BH___: The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
 

### 2016-03-03 mpa_analysis/2_wdpa_rast_to_csv.Rmd
finished MPA analysis(?)

--- 2016-02-22 mpa_analysis/mpa_gap_analysis.Rmd
@@
 
-``` {r set_up_base_raster}
-### DO WE NEED THIS? Use area at 0.5 degree level; most cells this will be pretty close.
-### base raster will contain 0.01 degree cells from -180 to +180,
-### and -90 to +90, with the associated area of each cell.
-### Save it as a .tif in GitHub - about 50 
-rast_base_file <- file.path(dir_git, 'mpa_analysis/rast_base.tif')
-
-if(!file.exists(rast_base_file)) {
-  ### import AquaMaps half-degree loiczid raster - for extents and such
-  rast_loiczid_file   <- file.path(dir_git, 'data/rasters/loiczid_raster.grd')
-  rast_loiczid        <- raster(rast_loiczid_file)
-  names(rast_loiczid) <- 'loiczid'
-
-  rast_empty <- raster(ext = extent(rast_loiczid), resolution = 0.01)
-  rast_base  <- raster::area(rast_empty) ### takes a while...
-
-  writeRaster(rast_base, rast_base_file, overwrite = TRUE)
-} else {
-  rast_base <- raster(rast_base_file)
-}
-
-```
-
 ``` {r load_rasters}
@@
 dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
-rast_mpa_mar_file <- file.path(dir_rast, 'rast_mpa_mar.tif')
+rast_mpa_mar_file <- file.path(dir_rast, 'rast_mpa_mar_i-iv.tif')
+# rast_mpa_mar_file <- file.path(dir_rast, 'wdpa_all_mar_klein_i-iv.tif')
+rast_mar_file <- file.path(dir_rast, 'rast_marine.tif')
+
 if(!file.exists(rast_mpa_mar_file)) {
-  rast_mar_file <- file.path(dir_rast, 'rast_marine.tif')
-  rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa1.tif')
+#   rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa_i-iv-des.tif')
+   rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa_i-iv-all.tif')
 
@@
 rast_loiczid_01_file <- file.path(dir_rast, 'rast_loiczid_01.tif')
-rast_loiczid_01 <- resample(rast_loiczid, rast_mpa_mar, method = 'ngb')
-writeRaster(rast_loiczid_01, rast_loiczid_01_file)
+if(!file.exists(rast_loiczid_01_file)) {
+  rast_loiczid_01 <- resample(rast_loiczid, rast_mpa_mar, 
+                            method = 'ngb',
+                            progress = 'text',
+                            filename = rast_loiczid_01_file)
+} else {
+  rast_loiczid_01 <- raster(rast_loiczid_01_file)
+}
+```
 
-### Turn all marine cells to value of 1 (so can uze zonal sum)
-rast_is_mar <- rast_mar
-values(rast_is_mar)[!is.na(values(rast_mar))] <- 1
-names(rast_is_mar) <- 'is_marine'
+``` {r summarize counts of marine and MPAs by cell}
 
-### Turn all MPA cells to value of 1 (so can use zonal sum)
-### NOTE: loses status_yr values - so filter by year first
+### Count marine cells in each LOICZID cell
+mar_df_file <- file.path(dir_git, 'mpa_analysis/celldata/loiczid_marine.csv')
+if(!file.exists(mar_df_file)) {
+  is_marine_df <- zonal(rast_mar, 
+                        z = rast_loiczid_01, 
+                        fun = 'count',
+                        progress = 'text') %>%
+    as.data.frame() %>%
+    rename(loiczid = zone)
+  write_csv(is_marine_df, mar_df_file)
+}
+
+### Count all protected cells in each LOICZID cell
+### NOTE: loses status_yr values - so filter by year for each iteration
+yr_list <- c('y1990', 'y1995', 'y2000', 'y2005', 'y2010', 'y2014', 'y2015')
+mpa_cvs_files <- c(file.path(dir_git, 
+                             sprintf('mpa_analysis/celldata/loiczid_mpa_i-iv_%s.csv', 
+                                     yr_list)))
+# mpa_cvs_files <- c(file.path(dir_git, 
+#                              sprintf('mpa_analysis/celldata/loiczid_mpa_klein_i-iv_%s.csv', 
+#                                      yr_list)))
+yr_list1 <- yr_list[!file.exists(mpa_cvs_files)]
+
+if(length(yr_list1) > 0) {
+  message('Creating marine protected area rasters for years:')
+  message('  ', paste(yr_list1, collapse = ', '))
+  for (yr in yr_list1) { # yr <- yr_list1[1]
+    message('filtering MPAs to those designated prior to ', str_replace(yr, 'y', ''))
     rast_is_mpa <- rast_mpa_mar
-values(rast_is_mpa)[!is.na(values(rast_mpa_mar))] <- 1
-names(rast_is_mpa) <- 'is_mpa'
+    data_year <- as.integer(str_replace(yr, 'y', ''))
+    values(rast_is_mpa)[values(rast_mpa_mar) >= data_year] <- NA
+      ### NOTE: 2015 MPA dataset includes up to 2014; 2014 dataset goes
+      ###   through 2013; etc.
+    message('counting protected cells per LOICZID for ', yr)
+    is_mpa_df <- zonal(rast_is_mpa, 
+                       z = rast_loiczid_01, 
+                       fun = 'count',
+                       progress = 'text') %>%
+      as.data.frame() %>%
+      rename(loiczid = zone)
+    write_csv(is_mpa_df, 
+              mpa_cvs_files[yr_list == yr])
+  }
+}
+### These files get loaded later; no need to hog memory by loading them here.
 
-rast_test <- rast_test2 <- raster(ext = extent(rast_loiczid), resolution = c(10, 10))
-values(rast_test) <- 1:648
-values(rast_test2) <- rep(c(1:10), each = 64, length.out = length(rast_test))
-x <- zonal(rast_test, rast_test2, fun = sum)
+```

### 2016-03-03 mpa_analysis/3_mpa_spp_analysis.Rmd
finished MPA analysis(?)

---
title: 'MPA Gap Analysis: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(raster)
library(readr)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')
dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')

if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)

### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

# MPA gap analysis

For each species:

* for each "present" LOICZID cell
    * determine total area: overall cell area (in km2) * marine area (cells/2500)
    * determine protected area: overall cell area (in km2) * protected area (cells/2500)
* sum total area and protected area across all "present" LOICZID cells
* report: 
    * ratio of (protected area:total area) (dimensionless, percentage)
    * total marine area and total protected area (km2)
    * species name, ID, taxonomic group, extinction risk

``` {r combine_mpa_areas_all}
yr_list <- c('y1990', 'y1995', 'y2000', 'y2005', 'y2010', 'y2014', 'y2015')

combine_mpa_areas <- function(mpa_file_base, yr_list) {

  hcaf_fn <- file.path(dir_am_data, 'hcaf_truncated.csv')
  message('Reading hcaf cells to get area per cell: ', hcaf_fn)
  ### get AquaMaps file of cells to translate cell IDs to area and ocean area
  hcaf <- read_csv(hcaf_fn) %>%
    select(loiczid, cellarea, oceanarea)

  ### attach the AquaMaps areas to the calculated percentages of marine area
  ### note: AquaMaps calculated areas differ from my calculated areas - see below
  cell_df <- read_csv(file.path(dir_git, 'mpa_analysis/celldata/loiczid_marine.csv'),
    col_types = 'dd') %>%
    rename(marine_ct = count) %>%
    left_join(hcaf, by = 'loiczid') %>%
    mutate(marine_pct  = marine_ct/2500,
           marine_area = marine_pct * cellarea) %>%
    select(-marine_ct)

  ### load each analysis year (in yr_list) and join them all together
  ### to note protection status year of each cell.

  for (yr in yr_list) { # yr <- yr_list[1]
    fn <- file.path(dir_git, sprintf(mpa_file_base, yr))
    message('Reading in mpa cell file: ', fn)
    df <- read_csv(fn, col_types = 'dd')
    names(df)[2] <- yr
    cell_df <- cell_df %>%
      left_join(df, by = 'loiczid')
    cell_df[ , paste(yr, '_pct', sep = '')]  <- cell_df[ , yr]/2500
    cell_df[ , paste(yr, '_area', sep = '')] <- cell_df[ , paste(yr, '_pct', sep = '')] * cell_df$cellarea

    ### remove intermediate columns
    cell_df[ , yr] <- NULL
    cell_df[ , paste(yr, '_pct', sep = '')] <- NULL
  }
  return(cell_df)
}

cell_df_klein <- combine_mpa_areas('mpa_analysis/celldata/loiczid_mpa_klein_i-iv_%s.csv', yr_list)
cell_df_all   <- combine_mpa_areas('mpa_analysis/celldata/loiczid_mpa_i-iv_%s.csv',       yr_list)
cell_df_des   <- combine_mpa_areas('mpa_analysis/celldata/loiczid_mpa_i-iv-des_%s.csv',   yr_list)

# ### check diffs in oceanarea vs marine_area
# x <- cell_df %>% 
#   filter(oceanarea != marine_area) %>%
#   select(loiczid, oceanarea, marine_area, cellarea)
# ### 18158 cells out of 259200 = 7%
# x <- x %>% mutate(pct_diff = (marine_area - oceanarea)/cellarea)
# hist(x$pct_diff)
# ### most differences cluster right around zero.  Not going to worry about it!

```

# recreate Klein results using aquamaps 2014 data

``` {r recreate Klein results 2014 AM data, echo = FALSE, eval = FALSE}

area_mpa <- cell_df_klein %>%
  summarize(marine_area  = sum(marine_area),
            mpa_tot_2014 = sum(y2014_area))

### Recreate Klein results, using 2014 AquaMaps
am_spp_cells2014 <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc_2014.csv'))
mpa_klein_am2014 <- am_spp_cells2014 %>%
  left_join(cell_df_klein, by = 'loiczid')

### filter to the appropriate threshold...
thresh <- 0.50

mpa_klein_am2014 <- mpa_klein_am2014 %>%
  filter(probability >= thresh)

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
mpa_klein_am2014 <- mpa_klein_am2014 %>%
  group_by(speciesid) %>%
  summarize(marine_range = sum(marine_area),
            mpa2014 = sum(y2014_area),
            pct2014 = mpa2014/marine_range)

mpa_klein_am2014 <- mpa_klein_am2014 %>%
  mutate(cat2014 = ifelse(pct2014 == 0,                       'gap', NA),
         cat2014 = ifelse(pct2014 > 0    & pct2014 <= 0.02,  '0-2%', cat2014),
         cat2014 = ifelse(pct2014 > 0.02 & pct2014 <= 0.05,  '2-5%', cat2014),
         cat2014 = ifelse(pct2014 > 0.05 & pct2014 <= 0.10, '5-10%', cat2014),
         cat2014 = ifelse(pct2014 > 0.10,                    '>10%', cat2014))

mpa_klein_orig <- mpa_klein_am2014 %>%
  group_by(cat2014) %>%
  summarize(n_tot   = nrow(mpa_klein_am2014), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2014 = factor(cat2014, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2014)

knitr::kable(mpa_klein_orig)
# |cat2014 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     | 17344|   282|     1.6|
# |0-2%    | 17344|  7728|    44.6|
# |2-5%    | 17344|  7738|    44.6|
# |5-10%   | 17344|  1188|     6.8|
# |>10%    | 17344|   408|     2.4|

```

| study            | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :--------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015 | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, wdpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |

# recreate Klein analysis using 2015 aquamaps

``` {r load AM 2015 data, echo = FALSE}
am_spp_cells <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) ### AM 2015

```

``` {r recreate Klein results 2015 AM data, echo = FALSE, eval = FALSE}

area_mpa <- cell_df_klein %>%
  summarize(marine_area  = sum(marine_area),
            mpa_tot_2015 = sum(y2015_area))
### Note: should be same as using 2014 data, since cell_df_klein is based
### on 2014 WDPA database

### Recreate Klein results, using 2015 AquaMaps
mpa_klein_am2015 <- am_spp_cells
mpa_klein_am2015 <- mpa_klein_am2015 %>%
  left_join(cell_df_klein, by = 'loiczid')

### filter to the appropriate threshold...
thresh <- 0.50

mpa_klein_am2015 <- mpa_klein_am2015 %>%
  filter(probability >= thresh)

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
mpa_klein_am2015 <- mpa_klein_am2015 %>%
  group_by(speciesid) %>%
  summarize(marine_range = sum(marine_area),
            mpa2015 = sum(y2015_area),
            pct2015 = mpa2015/marine_range)

mpa_klein_am2015 <- mpa_klein_am2015 %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))

mpa_klein_update <- mpa_klein_am2015 %>%
  group_by(cat2015) %>%
  summarize(n_tot   = nrow(mpa_klein_am2015), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2015 = factor(cat2015, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2015)

knitr::kable(mpa_klein_update)
# |cat2015 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     | 22885|   379|     1.7|
# |0-2%    | 22885| 10454|    45.7|
# |2-5%    | 22885|  9777|    42.7|
# |5-10%   | 22885|  1654|     7.2|
# |>10%    | 22885|   621|     2.7|

```

| study            | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :--------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015 | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, wdpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| am2015, wdpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 1.7% | 45.7% | 42.7% |  7.2% | 2.7% |

# Switch to WDPA 2015 data set?

No.  Save that for time-series analysis.

``` {r iucn_only}

### Load IUCN cells data - 
### load spp_all from SPP_ICO v2016;
###

### compare results using only IUCN species maps
dir_iucn_data <- '/var/data/ohi/git-annex/globalprep/_raw_data/iucn_spp'
iucn_cells_file <- file.path(dir_iucn_data, 'iucn_cells_v2016.csv')
iucn_spp_cells <- read_csv(iucn_cells_file, col_types = 'cddddc') %>%
  as.data.frame(stringsAsFactors = FALSE)

iucn_spp_cells <- iucn_spp_cells %>%
  left_join(cell_df_klein, by = 'loiczid')

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
iucn_spp_sum <- iucn_spp_cells %>%
  rename(iucn_sid = id_no) %>%
  group_by(iucn_sid) %>%
  summarize(cellarea_tot = sum(cellarea), 
            ocean_range  = sum(oceanarea), 
            marine_range = sum(marine_area),
            mpa2015 = sum(y2015_area),
            pct2015 = mpa2015/marine_range)

iucn_spp_sum <- iucn_spp_sum %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015)) %>%
  mutate(speciesid = paste('iucn', iucn_sid, sep = ''))

iucn_spp_sum2 <- iucn_spp_sum %>%
  group_by(cat2015) %>%
  summarize(n_tot   = nrow(iucn_spp_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2015 = factor(cat2015, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2015)

knitr::kable(iucn_spp_sum2)

# |cat2015 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     |  4024|   257|     6.4|
# |0-2%    |  4024|  2687|    66.8|
# |2-5%    |  4024|   810|    20.1|
# |5-10%   |  4024|   152|     3.8|
# |>10%    |  4024|   118|     2.9|

```

| study              | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :----------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015   | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, wdpa2014   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| am2015, wdpa2014   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.7% | 45.7% | 42.7% |  7.2% | 2.7% |
| iucn2015, wdpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 6.4% | 66.8% | 20.1% |  3.8% | 2.9% |

``` {r 2015 AM data w 0 threshold, echo = FALSE, eval = FALSE}

### Set up AquaMaps species with zero threshold
mpa_am2015_0 <- am_spp_cells
mpa_am2015_0 <- mpa_am2015_0 %>%
  left_join(cell_df_klein, by = 'loiczid')

### skip filtering step - threshold = 0

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
mpa_am2015_0 <- mpa_am2015_0 %>%
  group_by(speciesid) %>%
  summarize(marine_range = sum(marine_area),
            mpa2015 = sum(y2015_area),
            pct2015 = mpa2015/marine_range)

mpa_am2015_0 <- mpa_am2015_0 %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))

mpa_am2015_0_sum <- mpa_am2015_0 %>%
  group_by(cat2015) %>%
  summarize(n_tot   = nrow(mpa_am2015_0), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2015 = factor(cat2015, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2015)

knitr::kable(mpa_am2015_0_sum)
# |cat2015 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     | 22889|   268|     1.2|
# |0-2%    | 22889| 10643|    46.5|
# |2-5%    | 22889| 10187|    44.5|
# |5-10%   | 22889|  1455|     6.4|
# |>10%    | 22889|   336|     1.5|

```

``` {r combine IUCN and AM with 0% thresh AM on top}

head(iucn_spp_sum)
head(mpa_am2015_0)

### get list of overlap species and remove from IUCN species (they are
### included in the AM species list)
spp_overlap <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))

mpa_am2015_iucn <- mpa_am2015_0 %>%
  bind_rows(iucn_spp_sum %>%
              filter(!iucn_sid %in% spp_overlap$iucn_sid) %>%
              select(-iucn_sid, cellarea_tot, ocean_range))

mpa_am2015_iucn_sum <- mpa_am2015_iucn %>%
  group_by(cat2015) %>%
  summarize(n_tot   = nrow(mpa_am2015_iucn), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2015 = factor(cat2015, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2015)

knitr::kable(mpa_am2015_iucn_sum)

# |cat2015 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     | 24634|   494|     2.0|
# |0-2%    | 24634| 11823|    48.0|
# |2-5%    | 24634| 10401|    42.2|
# |5-10%   | 24634|  1521|     6.2|
# |>10%    | 24634|   395|     1.6|

```

| study              | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :----------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015   | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014 50%   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| am2015 50%   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.7% | 45.7% | 42.7% |  7.2% | 2.7% |
| am2015 0%   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.2% | 46.5% | 44.5% |  6.4% | 1.5% |
| iucn2015 | IUCN I-IV, Sp. Mar |   2,915,745  | 6.4% | 66.8% | 20.1% |  3.8% | 2.9% |
| am2015 0% + iucn2015 | IUCN I-IV, Sp. Mar |   2,915,745  | 2.0% | 48.0% | 42.2% |  6.2% | 1.6% |

``` {r combine all runs into one df}
mpa_all_runs <- mpa_klein_am2014 %>%
  mutate(trial = 'am2014 (50%)') %>%
  rename(cat = cat2014, pct = pct2014) %>%
  bind_rows(mpa_klein_am2015 %>%
    mutate(trial = 'am2015 (50%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_0 %>%
    mutate(trial = 'am2015 (0%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_iucn %>%
    mutate(trial = 'am2015 (0%) + iucn2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(iucn_spp_sum %>%
    mutate(trial = 'iucn2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  select(speciesid, cat, pct, trial)

# names(mpa_all_runs)[3:7] <- paste('mpa', names(mpa_all_runs)[3:7], sep = '')

# mpa_all_runs <- mpa_all_runs %>%
#   select(trial, n_tot, 3:7) %>%
#   gather(mpa_cat, pct, 3:7) %>%
#   mutate(mpa_cat = factor(mpa_cat, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
#   arrange(mpa_cat)

write_csv(mpa_all_runs, file.path(dir_data, 'mpa_all_runs.csv'))
```

``` {r plot this beast}
library(ggplot2)
source('data_explore/data_explore_fxns.R')

if(!exists('mpa_all_runs'))
  mpa_all_runs <- read_csv(file.path(dir_data, 'mpa_all_runs.csv'))

mpa_all_runs1 <- mpa_all_runs %>%
  group_by(trial) %>%
  mutate(n_trial = n()) %>%
  group_by(trial, cat, n_trial) %>%
  summarize(p_cat = n()/first(n_trial)) %>%
  ungroup() %>%
  mutate(cat = factor(cat, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  mutate(trial = factor(trial, levels = c('iucn2015', 'am2015 (0%) + iucn2015', 'am2015 (0%)', 'am2015 (50%)', 'am2014 (50%)'))) %>%
  mutate(p_cat = p_cat * 100) %>%
  arrange(cat)

### Plot the bar chart
barchart_mpa_gap <- ggplot(mpa_all_runs1, 
                           aes(x = trial, fill = cat, weight = p_cat)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = .5) +
  scale_fill_manual(values = c('>10%'  = 'green4', 
                               '5-10%' = 'greenyellow', 
                               '2-5%'  = 'orange', 
                               '0-2%'  = 'orangered2',
                               'gap'   = 'red')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 115), breaks = c(seq(0, 100, 10))) +
  geom_text(aes(label = sprintf('n = %s', n_trial), y = 101), hjust = 0, 
            size = 2, 
            color = 'grey30') +
  coord_flip() +
  labs(x = 'scenario', 
       y = 'Percent of species by MPA coverage', 
       fill = 'MPA coverage')

print(barchart_mpa_gap)
ggsave(plot = barchart_mpa_gap,
       filename = file.path(dir_fig, 'barchart_mpa_gap.png'),
       width = 17.8, height = 5, units = 'cm')

mpa_all_runs_wide <- mpa_all_runs1 %>%
  mutate(p_cat = paste(round(p_cat, 1), '%', sep = '')) %>%
  spread(cat, p_cat) %>%
  rename(Scenario = trial, n = n_trial)
names(mpa_all_runs_wide)[4:7] <- paste('covered', names(mpa_all_runs_wide)[4:7], sep = ' ')
mpa_all_runs_wide <- mpa_all_runs_wide[rev(mpa_all_runs_wide$Scenario), ]

write_csv(mpa_all_runs_wide, file.path(dir_data, 'mpa_all_runs_summary.csv'))
```

### 2016-03-03 data_explore/data_explore.Rmd
updated manuscript

@@
 
+# Years for IUCN maps?
+
+``` {r get_iucn_citation_info, echo = FALSE}
+
+  ### Queries shapefiles and extracts column headers.  Use to help identify which
+  ### species groups have no id_no field, and possibly other fields to help 
+  ### differentiate.
+
+dir_iucn_shp <- file.path(dir_N, 'git-annex/globalprep/_raw_data/iucn_spp/iucn_shp')
+groups_list <- as.data.frame(list.files(dir_iucn_shp)) %>%
+  rename(shp_fn = `list.files(dir_iucn_shp)`) %>%
+  filter(tools::file_ext(shp_fn) == 'shp') %>%
+  mutate(shp_fn = str_replace(shp_fn, '.shp', ''))
+
+iucn_shp_info <- data.frame()
+
+for (spp_gp in groups_list$shp_fn) { # spp_gp <- groups_list$shp_fn[3]
+  spp_dbf <- foreign::read.dbf(file.path(dir_iucn_shp, sprintf('%s.dbf', spp_gp)))
+  message('file read successfully... ')
+  cites <- spp_dbf %>%
+    select(id_no, compiler, year, citation) %>%
+    unique() %>%
+    mutate(spp_group = spp_gp)
+
+  iucn_shp_info <- bind_rows(iucn_shp_info, cites)
+  message('binding to list...\n')
+}
+
+### oldest records go back to late 70s for terrestrial mammals; then skip to 2004 for a marine turtle map.
+
+```
+
+***
+
 ``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = FALSE}

### 2016-03-03 manuscript-main.Rmd
updated manuscript

@@
 output:
+  word_document: default
   pdf_document:
@@
     toc: no
-  word_document: default
 ---

### 2016-03-03 mpa_analysis/3_mpa_spp_analysis.Rmd
updated manuscript

@@
 
+knitr::kable(mpa_all_runs_wide)
+
 write_csv(mpa_all_runs_wide, file.path(dir_data, 'mpa_all_runs_summary.csv'))
 ```
+
+|Scenario               |     n|gap  |covered 0-2% |covered 2-5% |covered 5-10% |covered >10% |
+|:----------------------|-----:|:----|:------------|:------------|:-------------|:------------|
+|am2014 (50%)           | 17344|1.6% |44.6%        |44.6%        |6.8%          |2.4%         |
+|am2015 (50%)           | 22885|1.7% |45.7%        |42.7%        |7.2%          |2.7%         |
+|am2015 (0%)            | 22889|1.2% |46.5%        |44.5%        |6.4%          |1.5%         |
+|am2015 (0%) + iucn2015 | 24634|2%   |48%          |42.2%        |6.2%          |1.6%         |
+|iucn2015               |  4024|6.4% |66.8%        |20.1%        |3.8%          |2.9%         |

### 2016-03-03 ms_2_intro_res_disc.Rmd
finished MPA analysis(?)
updated manuscript

--- 2016-02-17 ms_2_intro_res_disc.Rmd
@@
 
+___BSH:___ _I think Julie makes this comment below, but we may need to be able to substantiate this somehow.  Is there a way to quickly figure out how many studies have used AM vs. IUCN?_
+
+___CCO:___ _What about age - AquaMaps is pretty recent (2006 at earliest); before that, would IUCN have been the dominant dataset for marine?
+Maybe we can track citations of AquaMaps, but citations of IUCN are generic - just cite the Red List - not referencing spatial information specifically, much less marine.
+Going by citation year field on the maps themselves, the oldest marine record I can find is 2004._
+
+_Four foundational papers relating to AquaMaps (Kaschner et al. 2006, 2011; 2013; Ready et al. 2010) have been cited a total of 379 times._
+
 The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
@@
 For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and outline possible improvements.
-<!--BSH: I think Jamie makes this comment below, but we may need to be able to substantiate this somehow.
-Is there a way to quickly figure out how many studies have used AM vs.
-IUCN?
-CCO: What about age - AquaMaps is pretty recent; before that, would IUCN have been the dominant dataset for marine?
-Maybe we can track citations of AquaMaps, but citations of IUCN I think are generic - just cite the Red List - not referencing spatial information specifically, much less marine.-->
 
-We then subject two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (Halpern et al.
+We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (Halpern et al.
 2012, 2015) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (Klein et al.
-2015) - to a sensitivity analysis.
-...
-___substituting___ one dataset over the other, to highlight the possible consequences of different data use decisions on our understanding of _the status of_ marine biodiversity.
-___BSH suggestion: We used all unique species range data from both sources but substituted one dataset over the other for shared species to highlight the possible consequences of different data use decisions on our understanding of the status of and protection of marine biodiversity.___
+2015) - as case studies to explore the implications of prioritizing one data set over the other.
+The results highlight the possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
 
@@
 
-The methodologies behind these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).
+The methodologies behind these datasets imply differences in prediction of species distribution due to errors of commission (falsely _(erroneously?
+less accusatory, but an annoying word)_ indicating species presence) and omission (falsely indicating species absence).
 Geographic range data such as IUCN range maps generally include large commission errors, while predicted distribution models such as AquaMaps will likely include fewer commission errors but more omission errors.
-Each type of error bears different implications for conservation goals (Rondinini, 2006).
+Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range <!-- from Rondinini: Omission errors lead to identifying networks of areas for conservation action that are smaller than required and centred on known species occurrences, thus affecting the comprehensiveness, representativeness and efficiency of selected areas.
+Commission errors lead to selecting areas not relevant to conservation, thus affecting the representativeness and adequacy of reserve network --> (Rondinini, 2006).
 <!-- This study basically predicted some of the problems we’re seeing here and predicts implications; I’d like to meaningfully acknowledge it, but not sure how deeply to get into it.
@@
 
-The negative linear relationship <!--BSH: Any stats for this?
+A general negative linear pattern emerges, <!--BSH: Any stats for this?
 If so, insert just after this phrase.
-Otherwise maybe phrase it as 'The general pattern that emerges of a negative linear relationship, where increasing similarity...'--> indicates that increasing similarity in extent correlates with decreasing distribution alignment.
+Otherwise maybe phrase it as 'The general pattern that emerges of a negative linear relationship, where increasing similarity...'--> suggesting that increasing similarity in extent correlates with decreasing distribution alignment.
 <!--BSH: Could also be said as increasing similarity in alignment leads to decreasing similarity in extent.
@@
 
-Dividing the map-paired species into quadrants highlights different categories of relationships that in turn help further explain the general pattern.
+By dividing the map-paired species into quadrants, we highlight different categories of relationships that in turn help further explain the general pattern.
 The upper right quadrant includes species (n = `r n_q1`) whose described ranges agree in both spatial distribution and extent.
@@
 For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their preferred depths, likely introducing errors of commission.
-<!-- BSH: Worth a few spot checks to confirm this is more or less true.
-And then maybe include 2-3 examples as figures in the SOM and refer to them here.-->
+<!-- BSH: maybe include 2-3 examples as figures in the SOM and refer to them here.-->
 Ocean depth is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps.
-_BSH: Simply clipping IUCN range maps to known depth preferences would resolve many of these mismatches._ <!--BSH: This is kinda redundant, just emphasizing a solution to the problem.
-Not sure it's needed, or perhaps could be woven in to previous sentence.-->
+Simply clipping IUCN range maps to known depth preferences would resolve many of these mismatches.
 
@@
 
-<!--The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match.
-Often these mismatches arise when AquaMaps relies on limited data to predict species presence well beyond known occurrences.
-Indeed, `r p_q4_datapoor`% (n = `r n_q4_datapoor`) of these cases are data-poor (in which the AquaMaps environmental envelope model is based on fewer than 10 known cell occurrences), compared with `r p_q123_datapoor`% for species in the other three quadrants.
-Only `r p_q4_rev`% (n = `r n_q4_rev`) of AquaMaps maps in this quadrant have undergone expert review, compared with `r p_q123_rev`% of species in the other three quadrants.
-<!--BSH: This implies the problem is solely with AM.
-Are there no issues with IUCN maps that cause problems in this quadrant?--> <!--___The data-poor condition likely applies to IUCN as well - AM using OBIS points to generate models, and IUCN is probably doing the same.
-Check, and rework.
-This means AM is creating errors of commission, while IUCN may be committing errors of omission for poorly-studied species and/or regions - compounding the total problem.___-->
-
 The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match well.
-Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (summing (averaging?) OBIS and GBIF occurrences) for this quadrant is `r dp_med_am_q4`, compared to a median of `r dp_med_am_q123` for species in the other three quadrants.
+Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (summing (averaging?) OBIS and GBIF occurrences) for this quadrant is `r dp_med_am_q4`, compared to a median of `r dp_med_am_q123` for species across the other three quadrants.
 When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species.
@@
 
-_brief intro to implications - where/how long have IUCN/AquaMaps been used for various purposes; then introduce two case studies_
-
-* ___From JSL:___ _I think you need a broader discussion of implications.
-How many times has IUCN/Aquamaps been cited?
-How long have they been around?
-What scales are the studies?
-Then you can muse about how results/interpretations would be different depending on which dataset they used, and reference Fig 3’s._
-* _Then, say for example, the OHI uses these datasets together (don’t need to get too much into OHI background I think, maybe they don’t even need to know about the biodiversity goal?), but say that to try to combine them they use a threshold at 40% since others use this threshold (but check out the SOM for analyses about thresholds)_
-
-<!--BSH: I think the first bullet point can/should go in the Intro (very briefly!) - I've noted where in my comments to that section.
-I do agree there needs to be a tiny bit of explanation of how OHI uses the two datasets.-->
+Clear and significant differences in species range descriptions between AquaMaps and IUCN range maps, due to method-driven differences in errors of commission and omission.
+To examine the implications of these differences, we replicated two recent studies, varying only the prioritization of one data set over the other.
 
@@
 2012, 2015), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
-To maximize the number of represented species, OHI gleans spatial distributions from both IUCN and AquaMaps datasets, prioritizing IUCN data for the subset of species in both sources.
-OHI uses a probability threshold of 40% to determine species presence for AquaMaps data.
+To maximize the number of represented species, OHI gleans spatial distributions from both IUCN and AquaMaps datasets, prioritizing IUCN data for the `r n_spp_both` species included in both sources.
+OHI currently uses a probability threshold of 40% to determine species presence for AquaMaps data.
 
@@
 3</a>).
-Most country scores varied by fewer than five points across the three scenarios; a few country scores decreased by as many as 12 points, indicating a more dire assessment of species conservation status.
-
-_Since the Ocean Health Index Species subgoal relies on spatial data from both datasets, the impacts of these threshold and preference changes will be somewhat muted.
-When IUCN range maps are the preferred data source, only the subset of AquaMaps-only species will be affected by threshold changes; and when AquaMaps is the preferred source, the IUCN-only species will dampen the effect of a threshold change._ <!--CCO: This seems overly specific and I think could probably be cut, but the score changes weren’t earth-shattering – and that’s because we are buffering it by using both datasets
-BSH: I would maybe reframe this as 'given that only 2166 species overlap, score changes are not likely to be large for OHI and yet we still see important differences .
-.
-.
-" (remember that at the global scale, 2-3 points is a big deal)-->
+Given that only `r n_spp_both` species overlap out of `r n_spp_all`, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in such large score shifts as seen in scenarios 2 and 3.
+While many country scores showed only small variation across all three scenarios, select countries gained as many as 5 points while others dropped up to 12 points.
+Significantly, the global score decreased by two full points in scenario 3, indicating that an arbitrary change in how the two datasets are combined results in a more dire assessment of species conservation status.
 
@@
 The study relied on the AquaMaps database, using a probability threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
-The study determined that currently defined marine protected areas (IUCN categories I through IV overlapping with marine areas) are woefully inadequate in protecting marine biodiversity; 90.5% of species have less than 5% of their overall range represented within marine protected areas.
+The study determined that currently established MPAs (IUCN categories I through IV overlapping with marine areas) are woefully inadequate in protecting marine biodiversity; 90.5% of species have less than 5% of their overall range represented within MPAs and 1.4% of species have no protection at all ("gap" species).
 
-___working on the analysis...___
-
-_Predictions: What happens when using a 0% threshold instead of a 50%? What happens when using IUCN rather than AM?_
-* _IUCN tends to overestimate extent esp for coastal species?  How much of this area will fall within MPAs and how much will fall outside, proportionally? this affects protection ratio._
-* _More area means more species (relatively) will be at least partially included in MPAs so fewer apparent gap species?_
-* _Policy implications?  Can also tie this back to the predictions of Rondinini et al re: effects of commission/omission errors_
+ (<a href = '#fig4'>Figure 4</a>) displays the results of applying this methodology under several different scenarios.
+When comparing the conclusions based upon an IUCN-only scenario to an AquaMaps-only scenario (using 2015 data and a 0% threshold sets the most meaningful comparison) we found a five-fold increase in the proportion of gap species (6.4% vs 1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% vs.
+47.7%).
+However, this comparison also indicates a larger proportion of well-protected species with greater than 10% of range protected (2.9% vs.
+1.5%).
+Conclusions drawn from each of these datasets would paint dramatically different pictures of the protection afforded by our current global MPA network.
 
@@
 
-What can be done to address the differences between these two important marine species distribution datasets?
-We have identified several potential causes of commission errors among these datasets.
-Understanding these causes points toward a few improvements that may help us better align the predictions of species distributions, increasing their utility for ecology research and conservation actions.
+AquaMaps and IUCN range maps provide important spatial information on marine species ranges, and show reasonable agreement for many well-studied species, but illustrate uncertainty in our understanding of spatial distribution for many others.
+Identifying and addressing differences in these datasets will increase their utility for ecology research and conservation actions.
+We have identified several likely drivers of commission errors and omission errors between these datasets, pointing toward improvements to better align the predictions of species distributions.
 
@@
 * In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.
-Currently the area restrictions are denoted by FAO Major Fishing Areas (REF find a reference in AquaMaps somewhere); however, allowing area restrictions based on biogeographical criteria, e.g.
-Marine Ecoregions of the World (REF), would likely have better resolution and predictive power, especially for data-poor species.
+Currently the area restrictions are denoted by FAO Major Fishing Areas (REF find a reference in AquaMaps somewhere); however, area restrictions based on biogeographical criteria, e.g.
+Marine Ecoregions of the World (REF), would likely provide better resolution and predictive power, especially for data-poor species.
 
@@
     * Additional filters on AquaMaps distribution criteria, such as expert review status or the number of occurrence cells used to generate the species model, can help avoid overextrapolation for data poor species.
+
+_(do we need a wrap-it-up paragraph here?  I put that more at the top of the conclusions section... could shift it down here, but probably not needed in both places)_

### 2016-03-03 ms_4_figs.Rmd
finished MPA analysis(?)
updated manuscript

--- 2016-01-25 ms_4_figs.Rmd
@@
 
-### Figure 2
-
 <a id = 'fig2'></a>
@@
 (B) Alignment breakdown of paired-map species by taxonomic group.
-_is it distracting to use order names for some groups and not others?
-Some of the IUCN taxonomic groups had pretty long names ("acanthuridae" is way shorter than "surgeonfish, tangs, and unicornfish")_
 
@@
 
-___NOTE: These figures still need to be recreated using the latest and greatest data!___ _currently they use data generated in November or so, in ohiprep._
+> __Fig. 3.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
+> Scenario 1 shows the effect of reducing the probability threshold for AquaMaps presence to more accurately track the definition of IUCN “limits of distribution."  Reducing the threshold increases the apparent range of a species; the slight decrease in average score suggests increased spatial representation of threatened species.
+> Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the 40% presence threshold.
+In general, AquaMaps ranges are smaller than IUCN ranges, so most overlapping species will see a decrease in represented range.
+The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or a combination of both.
+> Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, and simultaneously eliminating the presence threshold.
+The zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
+The large decrease indicates that a zero threshold greatly increases the spatial representation of high-risk species.
 
-> __Fig. 3.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
+___NOTE: This figure still needs to be recreated using the latest and greatest data!___ _currently they use data generated in November or so, in ohiprep._
 
-| Scenario              | Priority data source | AquaMaps presence threshold |
-| --------------------- | :------------------: | :-------------------------: |
-| Scenario 0 (current)  | IUCN                 |  >= 40%                     |
-| Scenario 1            | IUCN                 |   > 0%                      |
-| Scenario 2            | AquaMaps             |  >= 40%                     |
-| Scenario 3            | AquaMaps             |   > 0%                      |
+### Figure 4
 
-> * Scenario 1 shows the effect of reducing the presence threshold for AquaMaps presence.
-Reducing the threshold increases the apparent range of a species; the slight decrease in average score suggests increased spatial representation of threatened species.
-> * Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the same presence threshold.
-This will have different effects depending on the species; in general, AquaMaps ranges are smaller than IUCN ranges, so most overlapping species will see a decrease in represented range.
-The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or more likely a combination of both.
-> * Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, while eliminating the presence threshold.
-Just as a presence threshold of zero in scenario 1 drives a decrease in average score relative to the baseline, the zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
-The large decrease indicates that within the set of paired-map species, a zero threshold greatly increases the spatial representation of high-risk species relative to low-risk species.
+<a id = 'fig4'></a>
+
+![Change in MPA coverage](figures/barchart_mpa_gap.png)
+
+> __Fig.
+4.__  Percent of species range covered by MPAs based upon methods in Klein et al.
+(2015).
+Scenario 1 replicates the original results, measuring protected range of species in AquaMaps 2014 dataset, with a 50% presence threshold, against the 2014 World Database of Protected Areas, filtered for IUCN categories I-IV that overlap marine areas.
+Scenario 2 updates the results using the 2015 AquaMaps dataset; the small changes despite the inclusion of an additional 5545 species demonstrates the robustness of the conclusions of Klein et al.
+(2015).
+Scenario 3, using 2015 AquaMaps data, drops the presence threshold to zero, showing an expected decrease in gap species, but also a decrease in species with 5% or greater protected range.
+Scenario 4 adds an additional 1745 species unique to the IUCN dataset, resulting in increases in gap species and species with less than 2% coverage.
+Scenario 5 examines species MPA coverage using only the IUCN dataset.
+
+`r knitr::kable(read_csv(file.path(dir_data, 'mpa_all_runs_summary.csv')))`

### 2016-03-03 ms_5_refs.Rmd
updated manuscript

@@
 and one e-supplement).
-PDF _report on marine data sets and data gaps etc, incl both IUCN and AM_
+PDF _review of marine data sets and data gaps etc, incl both IUCN and AM as well as many others_
 
@@
 * AquaMaps reference, and accessed date?
+
+Predicting the distributions of marine organisms at the global scale
+Ready, J., Kaschner, K., South, A.
+B., Eastwood, P.
+D., Rees, T., Rius, J., ...
+& Froese, R.
+(2010).
+Predicting the distributions of marine organisms at the global scale.
+Ecological Modelling, 221(3), 467-478.
+Cited by 75
+
+[PDF] Mapping world-wide distributions of marine mammal species using a relative environmental suitability (RES) model
+K Kaschner, R Watson, AW Trites, D Pauly - Marine Ecology Progress …, 2006 - vliz.be
+ABSTRACT: The lack of comprehensive sighting data sets precludes the application of 
+standard habitat suitability modeling approaches to predict distributions of the majority of 
+marine mammal species on very large scales. As an alternative, we developed an ...
+Cited by 167 Related articles All 12 versions Web of Science: 86 Cite Save
+
+[DOC] from aquamaps.org
+[DOC] AquaMaps
+K Kaschner, J Rius-Barile, K Kesner-Reyes… - Predicted range maps …, 2013 - aquamaps.org
+AquaMaps is an approach to generating model-based, large-scale predictions of currently 
+known natural occurrence of marine species. Models are constructed from estimates of the 
+environmental tolerance of a given species with respect to depth, salinity, temperature, ...
+Cited by 64 Related articles All 3 versions Cite Save More
+
+[HTML] Current and future patterns of global marine mammal biodiversity
+K Kaschner, DP Tittensor, J Ready, T Gerrodette… - PLoS …, 2011 - journals.plos.org
+Abstract Quantifying the spatial distribution of taxa is an important prerequisite for the 
+preservation of biodiversity, and can provide a baseline against which to measure the 
+impacts of climate change. Here we analyse patterns of marine mammal species richness ...
+Cited by 73 Related articles All 20 versions Web of Science: 34 Cite Save More

### 2016-03-04 manuscript-main.Rmd
updating refs

@@
 output:
-  word_document: default
-  pdf_document:
-    number_sections: yes
-    toc: no
   html_document:

### 2016-03-04 ms_1_abst_sig.Rmd
updating refs

@@
 Species distribution data provide the foundation for a wide range of ecological research and conservation science and management, yet most species ranges remain unknown, and existing range maps have little overlap.
-In the ocean, two global species distribution datasets, produced by AquaMaps and the International Union for Conservation of Nature (IUCN), dominate our understanding of marine species ranges throughout the world's oceans.
+In the ocean, two global species distribution datasets, produced by AquaMaps (REF) and the International Union for Conservation of Nature (IUCN) (REF), dominate our understanding of marine species ranges throughout the world's oceans.
 Together they represent `r format(nrow(spp_all), big.mark = ',')` species, mostly from AquaMaps, with only `r format(nrow(spp_map_pairs), big.mark = ',')` species overlapping.
@@
 We find that IUCN maps often disregard bathymetry for depth-limited species, leading to predictions of species presence at unsuitable depths, and AquaMaps ranges for data-poor species often extrapolate presence far afield from known occurrences.
-We illustrate the implications of these differences by repeating two recent applications - the Ocean Health Index biodiversity goal and a global analysis of gaps in coverage of marine protected areas - and find significantly different estimates of the status of global biodiversity and effectiveness of conservation depending on which dataset was used.
+We illustrate the implications of these differences by repeating two recent applications - the Ocean Health Index (Halpern et al.
+2012, 2015) biodiversity goal and a global analysis of gaps in coverage of marine protected areas (Klein et al.
+2015) - and find significantly different estimates of the status of global biodiversity and effectiveness of conservation depending on how the two datasets were prioritized.
 ___Creating a single, highly accurate dataset of global marine species ranges will be difficult.

### 2016-03-04 ms_2_intro_res_disc.Rmd
updating refs

@@
 2010 _AM_; Martin et al.
-2014 _both and others_, Pimm 2014 _IUCN, mostly focusing on terrestrial but some marine_), predicting range shifts (Molinos et al.
+2014 _both and others_, Pimm et al.
+2014 _IUCN, mostly focusing on terrestrial but some marine_, Kaschner et al.
+2011), predicting range shifts (Molinos et al.
 2015 _AquaMaps_, and setting conservation priorities (Klein et al.
 2015 _AquaMaps_).
-<!-- BSH: Any of these use IUCN?
-Would be good to make sure at least 2-3 examples were based on IUCN.
-___none of these exclusively use IUCN - other suggestions?___-->
 
@@
 
-_Four foundational papers relating to AquaMaps (Kaschner et al. 2006, 2011; 2013; Ready et al. 2010) have been cited a total of 379 times._
+_Three foundational papers relating to AquaMaps (Kaschner et al. 2006, 2011; Ready et al. 2010) have been cited a total of 315 times._
 
@@
 
-The IUCN publishes species range maps, bundled by taxonomic groups (REF), based on expert input on spatial boundaries of a given species' "limits of distribution" (REF IUCN 2014) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
+The IUCN publishes species range maps, bundled by taxonomic groups, based on expert input on spatial boundaries of a given species' "limits of distribution" (IUCN 2015) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
 In contrast, AquaMaps models species distribution based environmental preferences (e.g.
 temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species (AquaMaps 2013)
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species (Kaschner et al.
+2006, Ready et al.
+2010)
 
@@
 Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range <!-- from Rondinini: Omission errors lead to identifying networks of areas for conservation action that are smaller than required and centred on known species occurrences, thus affecting the comprehensiveness, representativeness and efficiency of selected areas.
-Commission errors lead to selecting areas not relevant to conservation, thus affecting the representativeness and adequacy of reserve network --> (Rondinini, 2006).
+Commission errors lead to selecting areas not relevant to conservation, thus affecting the representativeness and adequacy of reserve network --> (Rondinini et al.
+2006).
 <!-- This study basically predicted some of the problems we’re seeing here and predicts implications; I’d like to meaningfully acknowledge it, but not sure how deeply to get into it.
@@
 2A).
-
 A general negative linear pattern emerges, <!--BSH: Any stats for this?
@@
 
-* The IUCN Red List instructional materials for mapping marine species suggest that "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species" (REF https://www.conservationtraining.org/mod/scorm/player.php?a=1412&currentorg=M5L1_ORG&scoid=6492).
+* The IUCN Red List instructional materials for mapping marine species suggest that "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species" (IUCN 2015).
 Unfortunately, this recommendation appears to be applied inconsistently.
@@
 * The AquaMaps model is a powerful tool for quickly estimating species distribution based on limited information; however, the model's output is only as good as the input data, which for marine species is often sorely limited.
-AquaMaps encourages experts to review and comment on the predicted species distribution maps; lessons from the small sample of reviewed maps can be generalized, for example to other species in the same taxonomic group, to further refine the predictions.
+AquaMaps encourages experts to review and comment on the predicted species distribution maps; lessons from the small sample of reviewed maps could be generalized, for example to other species in the same taxonomic group, to further refine the predictions.
 This is especially important for data-poor species.
 * In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.
-Currently the area restrictions are denoted by FAO Major Fishing Areas (REF find a reference in AquaMaps somewhere); however, area restrictions based on biogeographical criteria, e.g.
-Marine Ecoregions of the World (REF), would likely provide better resolution and predictive power, especially for data-poor species.
+Currently area restrictions are denoted by UN Food and Agriculture Organisation (FAO) fisheries reporting areas (Ready et al.
+2010); however, area restrictions based on biogeographical criteria, e.g.
+Marine Ecoregions of the World (Spalding et al.
+2007), would likely provide better resolution and predictive power, especially for data-poor species.
 

### 2016-03-04 ms_5_refs.Rmd
updating refs

--- 2016-01-25 ms_5_refs.Rmd
@@
 
-__Key AquaMaps publications:__
+Kaschner K et al. (2015) AquaMaps: Predicted range maps for aquatic species. World wide web electronic publication, www.aquamaps.org, Version 08/2015.
 
-Kaschner, K., R.
-Watson, A.
-W.
-Trites, D.
-Pauly (2006).
-Mapping world-wide distributions of marine mammal species using a relative environmental suitability (RES) model.
-Marine Ecology Progress Series 316: 285–310.
-_check this citation journal name..._ _This outlines the basic RES methodology - AM development_
+IUCN (2015). The IUCN Red List of Threatened Species. Version 2015-4. <http://www.iucnredlist.org>. Downloaded on 21 December 2015.
 
-Kaschner, K., D.P.
-Tittensor, J.
-Ready, T Gerrodette and B.
-Worm (2011).
-Current and Future Patterns of Global Marine Mammal Biodiversity.
-PLoS ONE 6(5): e19653.
-PDF _just what the title says - AM development, presence threshold 60%, also analyzes richness as a function of threshold_
+Halpern BS et al. "An index to assess the health and benefits of the global ocean." Nature 488.7413 (2012): 615-620. _doi:10.1038/nature11397._
 
-Ready, J., K.
-Kaschner, A.B.
-South, P.D Eastwood, T.
-Rees, J.
-Rius, E.
-Agbayani, S.
-Kullander and R.
-Froese (2010).
-Predicting the distributions of marine organisms at the global scale.
-Ecological Modelling 221(3): 467-478.
-PDF _Presents AM; assessing AquaMaps against other presence-only species models_
+Halpern BS et al. (2015). Patterns and emerging trends in global ocean health. PloS one, 10(3), e0117863.
 
-__Papers based on AquaMaps:__
+Klein CJ et al. (2015). Shortfalls in the global protected area network at representing marine biodiversity. Scientific Reports 5: 17539. _doi:10.1038/srep17539._
 
-Jones, M.C., S.R.
-Dyeb, J.K.
-Pinnegar and W.W.L.
-Cheung (2012).
-Modelling commercial fish distributions: Prediction and assessment using different approaches.
-Ecological Modelling 225(2012): 133-145.
-PDF _comparison of species distribution models including AquaMaps, Maxent and the Sea Around Us Project_
+Selig ER et al. (2013). Assessing global marine biodiversity status within a coupled socio-ecological perspective. PloS one, 8(4), e60284. _doi:10.1371/journal.pone.0060284._
 
-Coll, M., C.
-Piroddi, J.
-Steenbeek, K.
-Kaschner, F.
-Ben Rais Lasram et al.
-(2010).
-The biodiversity of the Mediterranean Sea: estimates, patterns, and threats.
-PLoS ONE 5(8): e11842.
-PDF _used AquaMaps to predict Med biodiversity.
-Also: Threshold = 0._
+Coll M et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. _used AquaMaps to predict Med biodiversity.  Also: Threshold = 0._
 
-Martin C.S., Fletcher R., Jones M.C., Kaschner K., Sullivan E., Tittensor D.P., Mcowen C., Geffert J.L., van Bochove J.W., Thomas H., Blyth S., Ravillious C., Tolley M., Stanwell-Smith D.
+*Martin CS et al.
 (2014).
@@
 and one e-supplement).
-PDF _report on marine data sets and data gaps etc, incl both IUCN and AM_
+_review of marine data sets and data gaps etc, incl both IUCN and AM as well as many others_
 
------
+Pimm SL et al.
+(2014).
+The biodiversity of species and their rates of extinction, distribution, and protection.
+Science, 344(6187), 1246752.
+_uses range maps to show biodiversity areas; may use IUCN range maps.
+Also discusses gaps and possible things that can be done about them._
 
-Hurlbert 2007 Species richness, hotspots, and the scale dependence of range maps in ecology and conservation.
+Kaschner K, Tittensor DP, Ready J, Gerrodette T, Worm B (2011).
+Current and Future Patterns of Global Marine Mammal Biodiversity.
+PLoS ONE 6(5): e19653.
+_just what the title says - AM development, presence threshold 60%, also analyzes richness as a function of threshold_
+
+Molinos JG et al. (2015). Climate velocity and the future global redistribution of marine biodiversity. Nature Climate Change. _doi:10.1038/nclimate2769._
+
+Rondinini C, Wilson KA, Boitani L, Grantham H, & Possingham HP (2006).
+Tradeoffs of different types of species occurrence data for use in systematic conservation planning.
+Ecology letters, 9(10), 1136-1145.
+_compares point locality, range maps, and distribution models in terms of omission and commission errors; also outlines Extent of Occurrence and Area of Occupancy distinctions._
+
+Jetz W, Sekercioglu CH, and Watson JE (2008).
+Ecological correlates and conservation implications of overestimating species geographic ranges.
+Conservation Biology, 22(1), 110-119.
+_EOO maps are usually highly interpolated and overestimate small-scale occurrence, which may bias research outcomes_
+
+Kaschner K, Watson R, Trites AW, Pauly D (2006).
+Mapping world-wide distributions of marine mammal species using a relative environmental suitability (RES) model.
+Marine Ecology Progress Series 316: 285–310.
+_check this citation journal name..._ _This outlines the basic RES methodology - AM development_
+
+Ready J et al.
+(2010).
+Predicting the distributions of marine organisms at the global scale.
+Ecological Modelling 221(3): 467-478.
+_Presents AM; assessing AquaMaps against other presence-only species models_
+
+Spalding MD et al. (2007). Marine ecoregions of the world: a bioregionalization of coastal and shelf areas. BioScience, 57(7), 573-583.
+
+### NOT CURRENTLY USED:
+
+Jones MC, Dyeb SR, Pinnegar JK, Cheung WWL (2012).
+Modelling commercial fish distributions: Prediction and assessment using different approaches.
+Ecological Modelling 225(2012): 133-145.
+PDF _comparison of species distribution models including AquaMaps, Maxent and the Sea Around Us Project_
+
+Hurlbert AH, and Jetz W (2007).
+Species richness, hotspots, and the scale dependence of range maps in ecology and conservation.
+Proceedings of the National Academy of Sciences, 104(33), 13384-13389.
 _mostly rasters of range maps?
@@
 we provide guidance about the approriate scale of their use_
-
-Jetz 2008 Ecological Correlates and Conservation Implications of Overestimating Species Geographic Ranges _EOO maps are usually highly interpolated and overestimate small-scale occurrence, which may bias research outcomes_
-
-Pimm 2014 The biodiversity of species and their rates of extinction, distribtuion, and protection.
-_uses range maps to show biodiversity areas; may use IUCN range maps.
-Also discusses gaps and possible things that can be done about them._
-
-Rondinini 2006 Tradeoffs of different types of species occurrence data for use in systematic conservation planning _compares point locality, range maps, and distribution models in terms of omission and commission errors; also outlines Extent of Occurrence and Area of Occupancy distinctions._
-
-García Molinos, Jorge, Benjamin S.
-Halpern, David S.
-Schoeman, Christopher J.
-Brown, Wolfgang Kiessling, Pippa J.
-Moore, John M.
-Pandolfi, Elvira S.
-Poloczanska, Anthony J.
-Richardson, and Michael T.
-Burrows.
-“Climate Velocity and the Future Global Redistribution of Marine Biodiversity.” Nature Climate Change advance online publication (August 31, 2015).
-doi:10.1038/nclimate2769.
-
-Halpern, Benjamin S., Catherine Longo, Darren Hardy, Karen L.
-McLeod, Jameal F.
-Samhouri, Steven K.
-Katona, Kristin Kleisner, et al.
-“An Index to Assess the Health and Benefits of the Global Ocean.” Nature 488, no.
-7413 (August 30, 2012): 615–20.
-doi:10.1038/nature11397.
-
-Klein, Carissa J., Christopher J.
-Brown, Benjamin S.
-Halpern, Daniel B.
-Segan, Jennifer McGowan, Maria Beger, and James E.M.
-Watson.
-“Shortfalls in the Global Protected Area Network at Representing Marine Biodiversity.” Scientific Reports 5 (December 3, 2015): 17539.
-doi:10.1038/srep17539.
-
-Selig, Elizabeth R., Catherine Longo, Benjamin S.
-Halpern, Benjamin D.
-Best, Darren Hardy, Cristiane T.
-Elfes, Courtney Scarborough, Kristin M.
-Kleisner, and Steven K.
-Katona.
-“Assessing Global Marine Biodiversity Status within a Coupled Socio-Ecological Perspective.” PLoS ONE 8, no.
-4 (April 11, 2013): e60284.
-doi:10.1371/journal.pone.0060284.
-
-* IUCN reference:
-    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
-    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources
-    * IUCN Red List - accessed 12/21/2015
-
-* AquaMaps reference, and accessed date?

### 2016-03-04 data_explore/data_explore.Rmd
updating refs, numbers etc in manuscript

@@
 iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
-                           progress = TRUE)
+                           progress = TRUE,
+                           col_types = 'cddddc')
 
@@
 ### this file created in data_prep
-spp_list <- read.csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'), 
-                     stringsAsFactors = FALSE)
+spp_list <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
 

### 2016-03-04 data_explore/data_prep.Rmd
updating refs, numbers etc in manuscript

@@
 if(basename(getwd()) != 'data_explore') setwd('~/github/IUCN-AquaMaps/data_explore')
-# source('data_explore_fxns.R')
+source('data_explore_fxns.R')
 ### When knitting this, it automatically sets WD to be this directory...
@@
 # Data files
-#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-message('Building IUCN species to cell table.  This might take a few minutes.')
-iucn_map_files <- file.path(dir_anx, 'iucn_intersections', list.files(file.path(dir_anx, 'iucn_intersections')))
-iucn_spp_cells <- lapply(iucn_map_files, data.table::fread) %>%
-  bind_rows()
-iucn_spp_cells <- iucn_spp_cells %>%
-  rename(iucn_sid = id_no, loiczid = LOICZID) %>%
-  mutate(iucn_sid = as.integer(iucn_sid),
-         presence = as.integer(presence),
-         loiczid  = as.integer(loiczid)) %>%
+
+### This file is generated in globalprep/SPP_ICO:
+iucn_spp_cells_all <- read_csv(file.path(dir_data_remote, 'iucn_spp/iucn_cells_v2016.csv'),
+                            col_types = 'cddddc')
+
+iucn_spp_cells <- iucn_spp_cells_all %>%
+  rename(iucn_sid = id_no) %>%
   filter(iucn_sid %in% spp_am_v_iucn$iucn_sid) %>%
   filter(presence != 5)
-### length(unique(iucn_spp_cells$iucn_sid)) results in 2013 species found...
-### some species with no cells reported.  2249 listed in the am v iucn file.
+### length(unique(iucn_spp_cells$iucn_sid)) results in 2279 species found...
+### some species with no cells reported.  2279 listed in the am v iucn file.
 write_csv(iucn_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'))
@@
 message('Loading AquaMaps species to cell table.  This might take a few minutes.')
-am_spp_cells <- fread(file.path(dir_data_remote, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv'), 
+am_spp_cells_all <- fread(file.path(dir_data_remote, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv'), 
                          showProgress = TRUE)
-am_spp_cells <- am_spp_cells %>%
+am_spp_cells <- am_spp_cells_all %>%
   rename(am_sid = speciesid, prob = probability) %>%
   filter(am_sid %in% spp_am_v_iucn$am_sid)
-### length(unique(am_spp_cells$am_sid)) results in 2283 species found - great!
+### length(unique(am_spp_cells$am_sid)) results in 2314 species found - great!
 write_csv(am_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'))
@@
 if (!file.exists(data_file) | reload) {
+
+  if(!exists('am_spp_cells'))
+    am_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'))
+  if(!exists('iucn_spp_cells'))
+    iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'),
+                               col_types = 'cddddc')
   ### species info list - scinames, IUCN species ID, AM species ID, category etc
@@
   filter(!area_iucn == 0)
-  ### will ditch the ones where no IUCN polygon exists - or are these species
-  ### whose ranges only occur in cells beyond the LOICZID grid?
+  ### will ditch the ones with no IUCN area - formerly caused by DD species
+  ### being filtered out.  Now, none suffer this consequence.
 

### 2016-03-04 ms_1_abst_sig.Rmd
updating refs, numbers etc in manuscript

--- 2016-01-31 ms_1_abst_sig.Rmd
@@
 In the ocean, two global species distribution datasets, produced by AquaMaps and the International Union for Conservation of Nature (IUCN), dominate our understanding of marine species ranges throughout the world's oceans.
-Together they represent `r format(nrow(spp_all), big.mark = ',')` species, mostly from AquaMaps, with only `r format(nrow(spp_map_pairs), big.mark = ',')` species overlapping.
+Together they represent `r format(n_spp_all, big.mark = ',')` species, mostly from AquaMaps, with only `r format(n_spp_both, big.mark = ',')` species overlapping.
 Here we examine differences in predicted species ranges between the two datasets, propose mechanistic causes and potential solutions for such differences, and explore the implications of these differences for management and conservation decisions.

### 2016-03-04 ms_2_intro_res_disc.Rmd
updating refs, numbers etc in manuscript

--- 2016-03-03 ms_2_intro_res_disc.Rmd
@@
 ---
-
-``` {r calc species numbers, echo = FALSE}
-n_spp_all    <- nrow(spp_all)
-n_spp_am     <- nrow(spp_am)
-n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
-n_spp_iucn   <- length(spp_iucn_maps$iucn_sid %>% unique())
-n_spp_both   <- length(spp_map_pairs$iucn_sid %>% unique())
-
-p_spp_am     <- round(nrow(spp_am)/n_spp_all * 100, 1)
-p_spp_am_rev <- round(n_spp_am_rev/n_spp_am * 100, 1)
-p_both_am    <- round(n_spp_both/n_spp_am   * 100, 1)
-p_both_iucn  <- round(n_spp_both/n_spp_iucn * 100, 1)
-p_both_all   <- round(n_spp_both/n_spp_all  * 100, 1)
-```
 
@@
 Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps (Kaschner et al.
-2013) and International Union for Conservation of Nature (IUCN) (REF).
+2013) and International Union for Conservation of Nature (IUCN) (IUCN 2015).
 These two spatial datasets have been used for a wide range of purposes, including assessing marine species status (Halpern et al.

### 2016-03-04 data_explore/data_explore_fxns.R

@@ -1,5 +1,7 @@
 ### data_explore_fxns.R
 
+library(ggplot2)
+
 ### generic theme for all plots
 ggtheme_basic <- theme(axis.ticks = element_blank(),
         text = element_text(family = 'Helvetica', color = 'gray30', size = 8),

### 2016-03-04 ms_1_abst_sig.Rmd

--- 2016-03-04 ms_1_abst_sig.Rmd
@@
 In the ocean, two global species distribution datasets, produced by AquaMaps (REF) and the International Union for Conservation of Nature (IUCN) (REF), dominate our understanding of marine species ranges throughout the world's oceans.
-Together they represent `r format(nrow(spp_all), big.mark = ',')` species, mostly from AquaMaps, with only `r format(nrow(spp_map_pairs), big.mark = ',')` species overlapping.
+Together they represent `r format(n_spp_all, big.mark = ',')` species, mostly from AquaMaps, with only `r format(n_spp_both, big.mark = ',')` species overlapping.
 Here we examine differences in predicted species ranges between the two datasets, propose mechanistic causes and potential solutions for such differences, and explore the implications of these differences for management and conservation decisions.
@@
 Understanding the implications of dataset differences for conservation planning and decision-making is essential.___
+
 <!--Understanding these issues points toward solutions for providers and users of these datasets.-->

### 2016-03-04 ms_3_methods.Rmd

--- 2016-01-25 ms_3_methods.Rmd
@@
 Based upon the methods described in Klein et al.
-2015, we reconstructed the analysis using the subset of protected areas spatially covering a marine area and classified as IUCN I-IV ___(Klein paper, table 1, subset 4 - seems to be the one they focus on the most)___, converting the World Database of Protected Areas (WDPA, REF) dataset to a raster of 0.5° cells indicating proportion of protected area within the cell ___(consider also: filter on STATUS == 'Designated', and filter out those pesky fishery management areas and manatee speed zones...)___.
-To verify our reconstruction we ran the analysis using the AquaMaps dataset available in April 2015 (REF) at a presence threshold of 50%.
-We then updated the analysis using the most recent AquaMaps dataset (REF?) at 50% threshold
+2015, we reconstructed the analysis using the subset of protected areas (WDPA 2014) spatially covering a marine area and classified as IUCN I-IV.
+The WDPA polygons and marine polygons were rasterized to 0.01° and then aggregated to AquaMaps native 0.5° cells, to calculate proportion of protected area and marine within each cell.
+After verifying our results using the 2014 AquaMaps dataset, we updated the analysis using the 2015 AquaMaps dataset at a presence threshold of 50% (to compare to Klein et al.
+directly) and 0% (to better compare with IUCN spatial data).
+To analyze MPA coverage against IUCN spatial data, we extracted IUCN polygon weights per 0.5° cell for each species and compared against the protected area raster.
+Finally, we combined AquaMaps data (at 0% threshold) and IUCN data, using AquaMaps for the `r n_spp_both` overlapping species and again compared against the protected area raster.
 
-* Possible analyses:
-    * run AquaMaps (new data) with threshold at 0%
-        * already done in original paper for old data, but redo it to compare to AM + IUCN scenario below
-    * run with AquaMaps only, but filtering out data-poor species?
-    * run with IUCN maps only in place of AquaMaps?
-    * run with AquaMaps and IUCN both, using 0% threshold; which preferred? 
-        * Should we clip corals (and coral-associated taxa) to 200 meter bathymetry, just for fun?
-* Which aspects of the analysis should we focus on?:
-    * overall vs broken down by taxonomic groups? (see paper, figure 1)
-    * spatial distribution of gap species (see paper, figure 2)
+All processing was completed in R (REF), and all code and intermediate data are available on GitHub (REF?) at https://github.com/OHI-Science/IUCN-AquaMaps.

### 2016-03-05 manuscript-main.Rmd
tweaks to manuscript - Abstract bringing in conservation importance earlier; refs adding note on how to include DOIs; discussion - clarifying percents in MPA gap analysis, noting questions about AquaMaps citations, and question about the top right cluster - should it be noted more directly

--- 2016-03-03 manuscript-main.Rmd
@@
 spp_all_file <- 'data/spp_all.csv'
-spp_all <- read_csv(spp_all_file)
+spp_all <- read_csv(spp_all_file) ### NOTE: contains duped scinames
 
@@
 iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read_csv(iucn_maps_file)
+spp_iucn_maps <- read_csv(iucn_maps_file) %>%
+  filter(!duplicated(iucn_sid))
 
@@
 
+-----
+
+``` {r calc species numbers, echo = FALSE}
+n_spp_all    <- nrow(spp_am) + nrow(spp_iucn_maps) - nrow(spp_map_pairs)
+n_spp_am     <- nrow(spp_am)
+n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
+n_spp_iucn   <- nrow(spp_iucn_maps)
+n_spp_both   <- nrow(spp_map_pairs)
+
+p_spp_am     <- round(n_spp_am/n_spp_all * 100, 1)
+p_spp_am_rev <- round(n_spp_am_rev/n_spp_am * 100, 1)
+p_both_am    <- round(n_spp_both/n_spp_am   * 100, 1)
+p_both_iucn  <- round(n_spp_both/n_spp_iucn * 100, 1)
+p_both_all   <- round(n_spp_both/n_spp_all  * 100, 1)
+```
 -----

### 2016-03-05 ms_1_abst_sig.Rmd
tweaks to manuscript - Abstract bringing in conservation importance earlier; refs adding note on how to include DOIs; discussion - clarifying percents in MPA gap analysis, noting questions about AquaMaps citations, and question about the top right cluster - should it be noted more directly

@@
 Species distribution data provide the foundation for a wide range of ecological research and conservation science and management, yet most species ranges remain unknown, and existing range maps have little overlap.
-In the ocean, two global species distribution datasets, produced by AquaMaps (REF) and the International Union for Conservation of Nature (IUCN) (REF), dominate our understanding of marine species ranges throughout the world's oceans.
+In the ocean, two global species distribution datasets, produced by AquaMaps (Kaschner et al.
+___2015???___) and the International Union for Conservation of Nature (IUCN) (IUCN 2015), dominate our understanding of marine species ranges throughout the world's oceans ___and drive conservation policy and management decisions___.
 Together they represent `r format(n_spp_all, big.mark = ',')` species, mostly from AquaMaps, with only `r format(n_spp_both, big.mark = ',')` species overlapping.

### 2016-03-05 ms_2_intro_res_disc.Rmd
tweaks to manuscript - Abstract bringing in conservation importance earlier; refs adding note on how to include DOIs; discussion - clarifying percents in MPA gap analysis, noting questions about AquaMaps citations, and question about the top right cluster - should it be noted more directly

@@
 Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps (Kaschner et al.
-2013) and International Union for Conservation of Nature (IUCN) (IUCN 2015).
+___2013???___) and International Union for Conservation of Nature (IUCN) (IUCN 2015).
 These two spatial datasets have been used for a wide range of purposes, including assessing marine species status (Halpern et al.
@@
 
+___directly address the cluster along the top right of the plot?___
+
 By dividing the map-paired species into quadrants, we highlight different categories of relationships that in turn help further explain the general pattern.
@@
  (<a href = '#fig4'>Figure 4</a>) displays the results of applying this methodology under several different scenarios.
-When comparing the conclusions based upon an IUCN-only scenario to an AquaMaps-only scenario (using 2015 data and a 0% threshold sets the most meaningful comparison) we found a five-fold increase in the proportion of gap species (6.4% vs 1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% vs.
+When comparing the conclusions based upon an IUCN-only scenario to an AquaMaps-only scenario (using 2015 data and a 0% threshold sets the most meaningful comparison) we found a five-fold increase in the proportion of gap species (6.4% of species vs 1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% of species vs.
 47.7%).
-However, this comparison also indicates a larger proportion of well-protected species with greater than 10% of range protected (2.9% vs.
+However, this comparison also indicates a larger proportion of well-protected species with greater than 10% of range protected (2.9% of species vs.
 1.5%).

### 2016-03-05 ms_5_refs.Rmd
tweaks to manuscript - Abstract bringing in conservation importance earlier; refs adding note on how to include DOIs; discussion - clarifying percents in MPA gap analysis, noting questions about AquaMaps citations, and question about the top right cluster - should it be noted more directly

@@
+# Achnowledgments
+<!-- Acknowledgments. List acknowledgments and funding sources. PNAS collaborates with FundRef to use a standard taxonomy of funder names. Dedications are rarely allowed. -->
+
 # References
 
-___still need to check citations, format properly, put into order___
+___double check formats; also - include DOIs for references? which if any?___
 
 <!--
-Acknowledgments. List acknowledgments and funding sources. PNAS collaborates with FundRef to use a standard taxonomy of funder names. Dedications are rarely allowed.
-
 References.
@@
 
-Kaschner K et al. (2015) AquaMaps: Predicted range maps for aquatic species. World wide web electronic publication, www.aquamaps.org, Version 08/2015.
+Kaschner K et al.
+(2015) AquaMaps: Predicted range maps for aquatic species.
+World wide web electronic publication, www.aquamaps.org, Version 08/2015.
+___since this is not a peer-reviewed publication, can we include this for PNAS reference list?___
 
-IUCN (2015). The IUCN Red List of Threatened Species. Version 2015-4. <http://www.iucnredlist.org>. Downloaded on 21 December 2015.
+IUCN (2015).
+The IUCN Red List of Threatened Species.
+Version 2015-4.
+<http://www.iucnredlist.org>.
+Downloaded on 21 December 2015.
+___since this is not a peer-reviewed publication, can we include this for PNAS reference list?___
 
-Halpern BS et al. "An index to assess the health and benefits of the global ocean." Nature 488.7413 (2012): 615-620. _doi:10.1038/nature11397._
+Halpern BS et al. (2012). An index to assess the health and benefits of the global ocean. Nature, 488(7413), 615-620. _doi:10.1038/nature11397._
 
-Halpern BS et al. (2015). Patterns and emerging trends in global ocean health. PloS one, 10(3), e0117863.
+Halpern BS et al. (2015). Patterns and emerging trends in global ocean health. PloS one, 10(3), e0117863. ___doi?___
 

### 2016-03-08 data_explore/data_explore.Rmd
updating fig captions; changed colors on figs to color-blind friendly.

@@
 
-```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
-
-# Data files
-#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
-                           progress = TRUE,
-                           col_types = 'cddddc')
-
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), 
-                           progress = TRUE)
-
-### LOICZID cell ID numbers - used in plotting rasters
-loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
-loiczid_raster       <- raster(loiczid_raster_file)
-names(loiczid_raster) <- 'loiczid'
-```
-
 ```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
@@
 ```
+<!-- percent overlap vs total area exploration -->
 
@@
 ### define 25%, 50%, and 75% quartiles for both x and y axes
-area_ratio_quartiles <- quantile(spp_list$area_ratio, na.rm = TRUE)[2:4]
-pct_overlap_quartiles <- quantile(spp_list$sm_perc, na.rm = TRUE)[2:4]
+area_align_med <- median(spp_list$area_ratio, na.rm = TRUE)
+dist_align_med <- median(spp_list$sm_perc, na.rm = TRUE)
 
@@
   ### color the quadrant backgrounds:
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
+    annotate("rect", xmin = area_align_med, xmax = 100, 
+                     ymin = dist_align_med, ymax = 100, 
+             alpha = .3, 
+#             fill= "green4")  + 
+             fill= "#4dac26")  + 
+    annotate("rect", xmax = area_align_med, xmin =   0, 
+                     ymin = dist_align_med, ymax = 100, 
+             alpha = .3, 
+#             fill= "greenyellow") + 
+             fill= "#b8e186") + 
+    annotate("rect", xmin = area_align_med, xmax = 100, 
+                     ymax = dist_align_med, ymin =   0, 
+             alpha = .3, 
+#             fill= "orange") + 
+             fill= "#f1b6da") + 
+    annotate("rect", xmax = area_align_med, xmin =   0, 
+                     ymax = dist_align_med, ymin =   0, 
+             alpha = .3, 
+#             fill= "orangered2") + 
+             fill= "#d01c8b") + 
   geom_point(alpha = .6)
@@
                         labels = c('0%', '25%', '50%', '75%', '100%')) + 
-  scale_x_continuous(expand = c(0, 0), limits = c(0, 100)) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
+  scale_x_continuous(expand = c(0, 0), 
+                     limits = c(0, 102),
+                     breaks = c(seq(0, 100, 25))) +
+  scale_y_continuous(expand = c(0, 0),
+                     limits = c(0, 102),
+                     breaks = c(seq(0, 100, 25)))
 
@@
   annotate(geom = 'text',
-           x = area_ratio_quartiles[2], y = 5,
+           x = area_align_med, y = 5,
            hjust = 0, vjust = 0,
@@
            fontface = 'bold.italic', angle = 90,
-           label = sprintf('Median = %s%%', round(area_ratio_quartiles[2], 1))) +
+           label = sprintf('Median = %s%%', round(area_align_med, 1))) +
   annotate(geom = 'text',
-           x = 5, y = pct_overlap_quartiles[2],
+           x = 5, y = dist_align_med,
            hjust = 0, vjust = 0,
@@
            fontface = 'bold.italic', angle = 0,
-           label = sprintf('Median = %s%%', round(pct_overlap_quartiles[2], 1)))
+           label = sprintf('Median = %s%%', round(dist_align_med, 1)))
 
@@
 # scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_vline(x = area_ratio_quartiles[2],
+#   geom_vline(x = area_align_med,
 #              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = pct_overlap_quartiles[2],
-#              color = 'grey50', linetype = 'dashed')
-
-### here are 25% and 75% quantile lines:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_vline(x = area_ratio_quartiles[c(1, 3)],
-#              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = pct_overlap_quartiles[c(1, 3)],
+#   geom_hline(y = dist_align_med,
 #              color = 'grey50', linetype = 'dashed')
@@
 ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 17.8, height = 9, units = 'cm')
-
-```
-
-``` {r pct_overlap_vs_ratio_of_area_w_data_poor, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-
-spp_list <- spp_list %>% 
-  mutate(data_poor = ifelse(occurcells < 10, '< 10', '>= 10'))
-### mongo plot time
-scatter_arearatio_vs_overlap2 <- ggplot(spp_list %>% 
-                                         transform(popn_category = reorder(popn_category, category_score)),
-                                       aes(x = area_ratio, y = sm_perc, 
-                                           color = data_poor,
-                                           size = lg_area_pct)) + 
-  ggtheme_plot + 
-  ### color the quadrant backgrounds:
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
-  geom_point(alpha = .6)
-
-### Manage scales for color and size 
-scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 +
-  scale_colour_manual(values = c('>= 10' = 'green4', 
-                                 '< 10'  = 'orangered2')) + 
-  scale_size_continuous(limits = c(0, 100),
-                        breaks = c( 0,    25,    50,    75,    100), 
-                        labels = c('0%', '25%', '50%', '75%', '100%')) + 
-  scale_x_continuous(expand = c(0, 0), limits = c(0, 100)) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
-
-### here are median labels:
-scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 +
-  annotate(geom = 'text',
-           x = area_ratio_quartiles[2], y = 5,
-           hjust = 0, vjust = 0,
-           color = 'grey30', 
-           size = 2,
-           fontface = 'bold.italic', angle = 90,
-           label = sprintf('Median = %s%%', round(area_ratio_quartiles[2], 1))) +
-  annotate(geom = 'text',
-           x = 5, y = pct_overlap_quartiles[2],
-           hjust = 0, vjust = 0,
-           color = 'grey30', 
-           size = 2,
-           fontface = 'bold.italic', angle = 0,
-           label = sprintf('Median = %s%%', round(pct_overlap_quartiles[2], 1)))
-
-scatter_arearatio_vs_overlap2 <- scatter_arearatio_vs_overlap2 +
-  labs(x = bquote('Extent alignment '*alpha['ext']*' (%)'), 
-       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
-#       title = 'Distribution alignment vs Extent alignment',
-       color = 'AM model cells', size = '% ocean range')
-
-print(scatter_arearatio_vs_overlap2)
-
-ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap-data_poor.png'), width = 17.8, height = 9, units = 'cm')
-
-```
-
-``` {r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
-
-for (i in 1:length(spp_quads)) { # i = 1
-  spp <- spp_quads[i]
-  spp_quad <- names(spp_quads)[i]
-
-  spp_plot <- plot_rangemap(spp)
-  print(spp_plot)
-
-  ggsave(filename = file.path(dir_fig, sprintf('rangemap_%s_%s.png', spp_quad,
-                                               str_replace(tolower(spp), ' ', '_'))),
-         plot = spp_plot, 
-         width = 8, height = 4, dpi = 200)
-}
 
@@
   theme(panel.grid.major.y = element_blank()) +
-  geom_bar(stat = 'count', alpha = .5) +
-  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
-                               'q2 dist-matched'   = 'greenyellow', 
-                               'q3 area-matched'   = 'orange', 
-                               'q4 poorly matched' = 'orangered2')) +
+  geom_bar(stat = 'count', alpha = .8) +
+  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
+                               'q2 dist-matched'   = '#b8e186', 
+                               'q3 area-matched'   = '#f1b6da', 
+                               'q4 poorly matched' = '#d01c8b')) +
+#   scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
+#                                'q2 dist-matched'   = 'greenyellow', 
+#                                'q3 area-matched'   = 'orange', 
+#                                'q4 poorly matched' = 'orangered2')) +
   scale_y_continuous(expand = c(0, 0), limits = c(0, 1.1)) +
@@
             size = 2, 
-            color = 'grey20') +
+            color = 'grey30') +
   coord_flip() +
@@
 
-``` {r examine_risk_quartiles, echo = FALSE, eval = TRUE}
+``` {r examine_risk_quartiles, echo = FALSE, eval = FALSE}
 
@@
   theme(panel.grid.major.y = element_blank()) +
-  geom_bar(stat = 'count', alpha = .5) +
-  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
-                               'q2 dist-matched'   = 'greenyellow', 
-                               'q3 area-matched'   = 'orange', 
-                               'q4 poorly matched' = 'orangered2')) +
+  geom_bar(stat = 'count', alpha = .8) +
+  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
+                               'q2 dist-matched'   = '#b8e186', 
+                               'q3 area-matched'   = '#f1b6da', 
+                               'q4 poorly matched' = '#d01c8b')) +
   scale_y_continuous(expand = c(0, 0), limits = c(0, 1.1)) +
@@
 <!-- # Multiple species maps per quadrant -->
-``` {r child = 'data_explore_pt3.Rmd', eval = TRUE}
+``` {r child = 'data_explore_pt3.Rmd', eval = FALSE}
 ```
@@
 ``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.
+### use data_prep.Rmd to create this file, if it does not yet exist.
 
@@
 spp_all_file <- file.path(dir_data, 'spp_all.csv')
-spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)
+spp_all <- read_csv(spp_all_file)
 
@@
 spp_am_file <- file.path(dir_data, 'spp_am.csv')
-spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)
+spp_am <- read_csv(spp_am_file)
 
@@
 
-``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE}
+``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
 
 sciname_lookup <- read.csv(file.path(dir_data, 'sciname_lookups/spp_iucn_scinames.csv'), stringsAsFactors = FALSE)
+
+spp_am <- spp_am %>%
+  left_join(read_csv(file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv/speciesoccursum.csv')) %>% 
+              dplyr::select(am_sid = speciesid, class) %>%
+              unique(), 
+            by = 'am_sid')
 
@@
   mutate(n_total = sum(n_spp, na.rm = TRUE),
-         percent = n_spp/n_total * 100)
+         percent = n_spp/n_total * 100,
+         dataset = factor(dataset, levels = c('AquaMaps', 'Both', 'IUCN')))
+
 bar_taxa2 <- ggplot(taxa_df_long2, aes(x = class, y = percent, fill = dataset)) +
   ggtheme_plot + 
-  geom_bar(stat = 'identity', alpha = .5) +
-  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
-  geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
-            size = 3, 
+  theme(panel.grid.major.x = element_blank()) +
+  geom_bar(stat = 'identity', alpha = .8) +
+#  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
+  scale_fill_manual(values = c('AquaMaps' = '#1b9e77', 'Both' = '#d95f02', 'IUCN' = '#7570b3')) +
+  geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
+            size = 2, 
             angle = 90, hjust = 0, color = 'grey30') +
   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) + 
+  scale_y_continuous(expand = c(0, 0), 
+                     limits = c(0, 130),
+                     breaks = c(0, 25, 50, 75, 100)) + 
   labs(# title = 'Species maps by taxonomic class',
        x = 'Class', 
-       y = 'Percent of mapped species')
+       y = 'Percent of mapped species',
+       fill = 'Dataset')
 print(bar_taxa2)
@@
 
-taxa_df_long3 <- taxa_df %>%
-  dplyr::select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
-  gather(dataset, n_spp, AquaMaps:Both) %>%
-  group_by(class) %>%
-  mutate(n_total = sum(n_spp, na.rm = TRUE),
-         percent = n_spp/n_total * 100)
-bar_taxa3 <- ggplot(taxa_df_long3, aes(x = class, y = percent, fill = dataset)) +
-  ggtheme_plot + 
-  geom_bar(stat = 'identity', alpha = .5) +
-  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
-  geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
-            #size = 3, 
-            angle = 90, hjust = 0, color = 'grey30') +
-  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
-  labs(# title = 'Species maps by taxonomic class, Red List-assessed only',
-       x = 'Class', 
-       y = 'Percent of assessed mapped species')
-print(bar_taxa3)
-ggsave(filename = file.path(dir_fig, 'bar_by_taxa_assessed.png'), 
-       width = 17.8, height = 6, units = 'cm')
+# taxa_df_long3 <- taxa_df %>%
+#   dplyr::select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
+#   gather(dataset, n_spp, AquaMaps:Both) %>%
+#   group_by(class) %>%
+#   mutate(n_total = sum(n_spp, na.rm = TRUE),
+#          percent = n_spp/n_total * 100)
+#
+# bar_taxa3 <- ggplot(taxa_df_long3, aes(x = class, y = percent, fill = dataset)) +
+#   ggtheme_plot + 
+#   geom_bar(stat = 'identity', alpha = .5) +
+#   scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
+#   geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
+#             #size = 3, 
+#             angle = 90, hjust = 0, color = 'grey30') +
+#   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
+#   scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
+#   labs(# title = 'Species maps by taxonomic class, Red List-assessed only',
+#        x = 'Class', 
+#        y = 'Percent of assessed mapped species')
+# print(bar_taxa3)
+# ggsave(filename = file.path(dir_fig, 'bar_by_taxa_assessed.png'), 
+#        width = 17.8, height = 6, units = 'cm')
 
@@
 
-``` {r, child = 'data_explore_am_thresh.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
+``` {r, child = 'data_explore_am_thresh.Rmd', echo = FALSE, message = FALSE, eval = FALSE}
 ```

### 2016-03-08 data_explore/data_explore_global_maps.Rmd
updating fig captions; changed colors on figs to color-blind friendly.

@@
 By counting the number of unique species are found in each of the half degree cells we can then create a single global raster of number of species.
+
+```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+
+# Data files
+#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
+iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
+                           progress = TRUE,
+                           col_types = 'cddddc')
+
+am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), 
+                           progress = TRUE)
+
+### LOICZID cell ID numbers - used in plotting rasters
+loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
+loiczid_raster       <- raster(loiczid_raster_file)
+names(loiczid_raster) <- 'loiczid'
+```
 

### 2016-03-08 manuscript-main.Rmd
updating fig captions; changed colors on figs to color-blind friendly.

@@
 output:
-  word_document: default
   pdf_document:
@@
     toc: no
+  word_document: default
 ---

### 2016-03-08 mpa_analysis/3_mpa_spp_analysis.Rmd
updating fig captions; changed colors on figs to color-blind friendly.

@@
          cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))
+
+write_csv(mpa_klein_am2015, 'mpa_analysis/celldata/mpa_by_spp_am2015.csv')
 
@@
   theme(panel.grid.major.y = element_blank()) +
-  geom_bar(stat = 'count', alpha = .5) +
-  scale_fill_manual(values = c('>10%'  = 'green4', 
-                               '5-10%' = 'greenyellow', 
-                               '2-5%'  = 'orange', 
-                               '0-2%'  = 'orangered2',
-                               'gap'   = 'red')) +
+  geom_bar(stat = 'count', alpha = .8) +
+  scale_fill_manual(values = c('>10%'  = '#4dac26', 
+                               '5-10%' = '#b8e186', 
+                               '2-5%'  = '#f7f7f7', 
+                               '0-2%'  = '#f1b6da',
+                               'gap'   = '#d01c8b')) +
   scale_y_continuous(expand = c(0, 0), limits = c(0, 115), breaks = c(seq(0, 100, 10))) +

### 2016-03-08 ms_4_figs.Rmd
updating fig captions; changed colors on figs to color-blind friendly.

@@
 1.__ (A) Number and proportion of species by taxa included in each dataset.
-AquaMaps encompasses a broader range of taxa than IUCN, while IUCN focuses on comprehensively assessing select taxonomic groups, typically at the level of order or family.
+AquaMaps encompasses a broader range of taxa than IUCN, while IUCN focuses on comprehensively assessing select taxonomic groups.
 Overlapping species are dominated by bony fishes (`r n_bonyfish` species, primarily tropical taxa) and corals (`r n_corals` species).
-(B, C) Global marine species richness according to (B) AquaMaps dataset and (C) IUCN dataset.
-The margin frequency plots show relative species count per cell at each latitude and longitude; while both datasets peak in tropical latitudes near the equator, the frequency for IUCN maps drops quickly beyond 30°N and 30°S, while the frequency for AquaMaps remains robust well into temperate latitudes.
+(B, C) Global marine species richness according to (B) AquaMaps and (C) IUCN.
+The margin frequency plots show relative species count per cell at each latitude and longitude; while both datasets peak in tropical latitudes, the frequency for IUCN maps drops quickly beyond 30°N and 30°S, while the frequency for AquaMaps remains robust well into temperate latitudes.
 The longitude frequency plots show a slight bias in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
@@
 Scenario 1 replicates the original results, measuring protected range of species in AquaMaps 2014 dataset, with a 50% presence threshold, against the 2014 World Database of Protected Areas, filtered for IUCN categories I-IV that overlap marine areas.
-Scenario 2 updates the results using the 2015 AquaMaps dataset; the small changes despite the inclusion of an additional 5545 species demonstrates the robustness of the conclusions of Klein et al.
-(2015).
+Scenario 2 updates the results using AquaMaps 2015, showing very small changes despite the inclusion of an additional 5545 species.
 Scenario 3, using 2015 AquaMaps data, drops the presence threshold to zero, showing an expected decrease in gap species, but also a decrease in species with 5% or greater protected range.
-Scenario 4 adds an additional 1745 species unique to the IUCN dataset, resulting in increases in gap species and species with less than 2% coverage.
+Scenario 4 adds an additional 1745 species unique to IUCN, resulting in increases in gap species and species with less than 2% coverage.
 Scenario 5 examines species MPA coverage using only the IUCN dataset.

### 2016-03-10 data_explore/data_explore.Rmd
updates to manuscript based on CS and JA comments

@@
 ### define 25%, 50%, and 75% quartiles for both x and y axes
-area_align_med <- median(spp_list$area_ratio, na.rm = TRUE)
-dist_align_med <- median(spp_list$sm_perc, na.rm = TRUE)
+area_align_mean <- mean(spp_list$area_ratio, na.rm = TRUE)
+dist_align_mean <- mean(spp_list$sm_perc, na.rm = TRUE)
 
@@
   ### color the quadrant backgrounds:
-    annotate("rect", xmin = area_align_med, xmax = 100, 
-                     ymin = dist_align_med, ymax = 100, 
+    annotate("rect", xmin = area_align_mean, xmax = 100, 
+                     ymin = dist_align_mean, ymax = 100, 
              alpha = .3, 
@@
              fill= "#4dac26")  + 
-    annotate("rect", xmax = area_align_med, xmin =   0, 
-                     ymin = dist_align_med, ymax = 100, 
+    annotate("rect", xmax = area_align_mean, xmin =   0, 
+                     ymin = dist_align_mean, ymax = 100, 
              alpha = .3, 
@@
              fill= "#b8e186") + 
-    annotate("rect", xmin = area_align_med, xmax = 100, 
-                     ymax = dist_align_med, ymin =   0, 
+    annotate("rect", xmin = area_align_mean, xmax = 100, 
+                     ymax = dist_align_mean, ymin =   0, 
              alpha = .3, 
@@
              fill= "#f1b6da") + 
-    annotate("rect", xmax = area_align_med, xmin =   0, 
-                     ymax = dist_align_med, ymin =   0, 
+    annotate("rect", xmax = area_align_mean, xmin =   0, 
+                     ymax = dist_align_mean, ymin =   0, 
              alpha = .3, 
@@
                      limits = c(0, 102),
-                     breaks = c(seq(0, 100, 25))) +
+                     breaks = c(seq(0, 100, 25)),
+                     labels = c('0%', '25%', '50%', '75%', '100%')) +
   scale_y_continuous(expand = c(0, 0),
                      limits = c(0, 102),
-                     breaks = c(seq(0, 100, 25)))
+                     breaks = c(seq(0, 100, 25)),
+                     labels = c('0%', '25%', '50%', '75%', '100%'))
 
-### here are median labels:
+### here are mean labels:
 scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
   annotate(geom = 'text',
-           x = area_align_med, y = 5,
+           x = area_align_mean, y = 5,
            hjust = 0, vjust = 0,
@@
            fontface = 'bold.italic', angle = 90,
-           label = sprintf('Median = %s%%', round(area_align_med, 1))) +
+           label = sprintf('Mean = %s%%', round(area_align_mean, 1))) +
   annotate(geom = 'text',
-           x = 5, y = dist_align_med,
+           x = 5, y = dist_align_mean,
            hjust = 0, vjust = 0,
@@
            fontface = 'bold.italic', angle = 0,
-           label = sprintf('Median = %s%%', round(dist_align_med, 1)))
+           label = sprintf('Mean = %s%%', round(dist_align_mean, 1)))
 
-### here are median lines:
+### here are mean lines:
 # scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_vline(x = area_align_med,
+#   geom_vline(x = area_align_mean,
 #              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = dist_align_med,
+#   geom_hline(y = dist_align_mean,
 #              color = 'grey50', linetype = 'dashed')
 
-### here are individual species, identified and named on the plot:
+### here are individual species, identified and namean on the plot:
 # scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
@@
 scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-  labs(x = bquote('Extent alignment '*alpha['ext']*' (%)'), 
+  labs(x = bquote('Area alignment '*alpha['ext']*' (%)'), 
        y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
@@
 spp_list_q1 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio) & sm_perc >= median(sm_perc))
+  filter(area_ratio >= mean(area_ratio) & sm_perc >= mean(sm_perc))
 spp_list_q2 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio) & sm_perc >= median(sm_perc))
+  filter(area_ratio < mean(area_ratio) & sm_perc >= mean(sm_perc))
 spp_list_q3 <- spp_list1 %>%
-  filter(area_ratio >= median(area_ratio) & sm_perc < median(sm_perc))
+  filter(area_ratio >= mean(area_ratio) & sm_perc < mean(sm_perc))
 spp_list_q4 <- spp_list1 %>%
-  filter(area_ratio < median(area_ratio) & sm_perc < median(sm_perc))
+  filter(area_ratio < mean(area_ratio) & sm_perc < mean(sm_perc))
 
@@
 #                                'q4 poorly matched' = 'orangered2')) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.1)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     limits = c(0, 1.1),
+                     breaks = c(0, .25, .5, .75, 1.0),
+                     labels = c('0%', '25%', '50%', '75%', '100%')) +
   geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
@@
   labs(x = 'Taxonomic Group', 
-       y = 'Relative proportion of species per quadrant', 
+       y = 'Percent of species by quadrant', 
 #       title = 'Taxonomic group by quadrant',
@@
 
-``` {r examine_risk_quartiles, echo = FALSE, eval = FALSE}
+``` {r count_occurrences_by_quad, echo = FALSE, warning = FALSE}
+
+gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence.csv'))
+obis_occur <- read_csv(file.path(dir_data, 'obis_occurrence.csv'))
+
+totes_occur <- obis_occur %>%
+  left_join(gbif_occur) %>%
+  rowwise() %>%
+  mutate(mean_occ = mean(c(gbif_occ, obis_occ), na.rm = TRUE)) %>%
+  group_by(iucn_sid) %>%
+  summarize(mean_occ = mean(mean_occ, na.rm = TRUE),
+            tots_occ = sum(mean_occ, na.rm = TRUE))
+
+spp_quads_occ <- spp_list_quads %>%
+  left_join(totes_occur %>%
+              dplyr::select(iucn_sid, mean_occ),
+            by = 'iucn_sid')
+
+write_csv(spp_quads_occ, file.path(dir_data, 'spp_quads_occ.csv'))
+
+```
+
+``` {r examine_risk_quartiles, echo = FALSE, eval = TRUE}
 
@@
                                'q4 poorly matched' = '#d01c8b')) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.1)) +
+  scale_y_continuous(expand = c(0, 0), 
+                     limits = c(0, 1.1),
+                     breaks = c(0, .25, .5, .75, 1.0),
+                     labels = c('0%', '25%', '50%', '75%', '100%')) +
   geom_text(aes(label = sprintf('n = %s', n_tot_cat), y = 1.01), hjust = 0, 
@@
   labs(x = 'Risk category', 
-       y = 'Relative proportion of species per quadrant', 
+       y = 'Percent of species by quadrant', 
 #       title = 'Risk category by quadrant',
@@
 
-``` {r plot bar charts by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE}
+``` {r plot bar charts by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
 ### bar charts comparing number of species in each class for each data set
@@
                      limits = c(0, 130),
-                     breaks = c(0, 25, 50, 75, 100)) + 
+                     breaks = c(0, 25, 50, 75, 100),
+                     labels = c('0%', '25%', '50%', '75%', '100%')) + 
   labs(# title = 'Species maps by taxonomic class',
@@
 
-``` {r get_iucn_citation_info, echo = FALSE}
+``` {r get_iucn_citation_info, echo = FALSE, warning = FALSE, eval = FALSE}
 

### 2016-03-10 manuscript-main.Rmd
updates to manuscript based on CS and JA comments

--- 2016-03-04 manuscript-main.Rmd
@@
 
+-----
+
+``` {r calc species numbers, echo = FALSE}
+n_spp_am     <- nrow(spp_am)
+n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
+n_spp_iucn   <- length(unique(spp_iucn_maps$iucn_sid))
+n_spp_both   <- length(unique(spp_map_pairs$iucn_sid))
+n_spp_all    <- n_spp_am + n_spp_iucn - n_spp_both
+
+p_spp_am     <- round(n_spp_am/n_spp_all * 100, 1)
+p_spp_iucn   <- round(n_spp_iucn/n_spp_all * 100, 1)
+p_both_am    <- round(n_spp_both/n_spp_am   * 100, 1)
+p_both_iucn  <- round(n_spp_both/n_spp_iucn * 100, 1)
+p_both_all   <- round(n_spp_both/n_spp_all  * 100, 1)
+
+spp_list <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
+area_align_mean <- round(mean(spp_list$area_ratio, na.rm = TRUE), 1)
+dist_align_mean <- round(mean(spp_list$sm_perc, na.rm = TRUE), 1)
+
+```
 -----

### 2016-03-10 ms_1_abst_sig.Rmd
updates to manuscript based on CS and JA comments

--- 2016-03-04 ms_1_abst_sig.Rmd
@@
 
-Species distribution data provide the foundation for a wide range of ecological research and conservation science and management, yet most species ranges remain unknown, and existing range maps have little overlap.
-In the ocean, two global species distribution datasets, produced by AquaMaps (REF) and the International Union for Conservation of Nature (IUCN) (REF), dominate our understanding of marine species ranges throughout the world's oceans.
-Together they represent `r format(n_spp_all, big.mark = ',')` species, mostly from AquaMaps, with only `r format(n_spp_both, big.mark = ',')` species overlapping.
-Here we examine differences in predicted species ranges between the two datasets, propose mechanistic causes and potential solutions for such differences, and explore the implications of these differences for management and conservation decisions.
-We find that IUCN maps often disregard bathymetry for depth-limited species, leading to predictions of species presence at unsuitable depths, and AquaMaps ranges for data-poor species often extrapolate presence far afield from known occurrences.
-We illustrate the implications of these differences by repeating two recent applications - the Ocean Health Index (Halpern et al.
-2012, 2015) biodiversity goal and a global analysis of gaps in coverage of marine protected areas (Klein et al.
+Species distribution data provide the foundation for a wide range of ecological research studies and conservation management decisions, yet most species ranges remain unknown, and existing range maps often suffer from data limitations and show inconsistencies.
+AquaMaps (Kaschner et al.
+___2015???___) and the International Union for Conservation of Nature (IUCN) (IUCN 2015) are two distinct efforts to map marine species distributions at a global scale.
+Together these two databases represent `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` species included in both.
+Here we examine differences in predicted species ranges between the two datasets, and find that these misalignments mainly result from divergent methodologies that introduce differing frequencies of commission and omission errors.
+We illustrate the scientific and management implications of these differences by repeating two recent applications - an assessment of global biodiversity within the Ocean Health Index (Halpern et al.
+2012, 2015) and a global analysis of gaps in coverage of marine protected areas (Klein et al.
 2015) - and find significantly different estimates of the status of global biodiversity and effectiveness of conservation depending on how the two datasets were prioritized.
-___Creating a single, highly accurate dataset of global marine species ranges will be difficult.
-Understanding the implications of dataset differences for conservation planning and decision-making is essential.___
+Until a single, highly accurate dataset of global marine species ranges becomes available, understanding the implications of dataset differences for conservation planning and decision-making will be essential.
 

### 2016-03-10 ms_2_intro_res_disc.Rmd
updates to manuscript based on CS and JA comments

@@
 
-<!--
-research questions: How do the two datasets differ in prediction of species distribution?
-Why do they differ?
-What are the implications of these differences?
-and how can we improve our confidence in their predictions?_ _this should also go in the abstract_
--->
-
-<!--Peer beneath the waves anywhere in the world - a moody kelp forest, a bustling coral reefscape, the frigid depths beneath a polar ice sheet, the endless blue of the open ocean - and you will find a unique ecosystem defined as much by which species are absent as by which species are abundant.-->
-Knowing where individuals of a species live is fundamental to the sciences of ecology, biogeography, and conservation, among many others.  This knowledge provides foundational information for 
+Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation, among many others.  This knowledge provides foundational information for 
  understanding species ranges and diversity,
@@
 2011), predicting range shifts (Molinos et al.
-2015 _AquaMaps_, and setting conservation priorities (Klein et al.
+2015 _AquaMaps_), and setting conservation priorities (Klein et al.
 2015 _AquaMaps_).
@@
 Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans (Rondinini et al.
-2006) (Jetz 2008?).
+2006, Jetz et al.
+2008).
 
 To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of species.
-Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique species).
+Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).
 For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and outline possible improvements.
@@
 
-The IUCN publishes species range maps, bundled by taxonomic groups, based on expert input on spatial boundaries of a given species' "limits of distribution" (IUCN 2015) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
+The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" (IUCN 2015) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
 In contrast, AquaMaps models species distribution based environmental preferences (e.g.
 temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of probability of occurrence for each species (Kaschner et al.
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of "probability of occurrence" for each species (Kaschner et al.
 2006, Ready et al.
-2010)
+2010).
+Studies frequently define "presence" for AquaMaps by applying a probability threshold, e.g.
+40% or greater probability of occurrence (Halpern et al.
+2012, 2015).
 
-The methodologies behind these datasets imply differences in prediction of species distribution due to errors of commission (falsely _(erroneously?
-less accusatory, but an annoying word)_ indicating species presence) and omission (falsely indicating species absence).
-Geographic range data such as IUCN range maps generally include large commission errors, while predicted distribution models such as AquaMaps will likely include fewer commission errors but more omission errors.
-Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range <!-- from Rondinini: Omission errors lead to identifying networks of areas for conservation action that are smaller than required and centred on known species occurrences, thus affecting the comprehensiveness, representativeness and efficiency of selected areas.
-Commission errors lead to selecting areas not relevant to conservation, thus affecting the representativeness and adequacy of reserve network --> (Rondinini et al.
+The methodologies behind the creation of these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).
+Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.
+Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range (Rondinini et al.
 2006).
-<!-- This study basically predicted some of the problems we’re seeing here and predicts implications; I’d like to meaningfully acknowledge it, but not sure how deeply to get into it.
-Later comments ask about whether to include the omission/commission terminology.
+___By comparing maps resulting from each methodology, we can identify and possibly address mechanistic causes for each type of error.___
 
-BSH: I haven’t read the paper, but it sounds like we need to acknowledge what was already discussed/done there.
+<!--BSH: I haven’t read the paper, but it sounds like we need to acknowledge what was already discussed/done there.
 This may actually simplify things, such that we condense the first three paragraphs of this section into a single paragraph that restates what that paper found (and cite it) and add anything else we want to point out.
@@
 The two datasets have notably different taxonomic (Fig.
-1A) and regional (Fig.
-1B,C) coverage.
-IUCN-mapped species focus on tropical latitudes and away from the Atlantic and Eastern Pacific compared to AquaMaps-mapped species.
-These differences can be mitigated by using both datasets, but the underlying methodological differences complicate such direct comparisons.
+1) and regional (Fig.
+2) coverage.
+AquaMaps encompasses a broader range of taxa than IUCN, as IUCN spatial data files are only available for select taxonomic groups that have been comprehensively assessed.
+While species numbers in both datasets peak in tropical latitudes near the equator, species counts for IUCN maps drops quickly beyond 30°N and 30°S, while species counts for AquaMaps remains robust well into temperate latitudes.
+The longitude frequency plots show a slight bias in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
+To achieve more comprehensive global coverage of species ranges these two datasets can be used together, but the underlying methodological differences complicate such direct comparisons.
 
@@
 n_q2 <- nrow(spp_list_quads %>% filter(quad == 'q2'))
-n_q2_am_sm  <- nrow(spp_list_quads %>% filter(sm_range == 'AM' & quad == 'q2'))
+# n_q2_am_sm  <- nrow(spp_list_quads %>% filter(sm_range == 'AM' & quad == 'q2'))
 n_q2_corals <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral') & quad == 'q2'))
@@
 
-n_q4_datapoor <- nrow(spp_list_quads %>% filter(quad == 'q4' & occurcells < 10))
-p_q4_datapoor <- round(n_q4_datapoor/n_q4 * 100, 1)
-n_q123_datapoor <- nrow(spp_list_quads %>% filter(quad != 'q4' & occurcells < 10))
-p_q123_datapoor <- round(n_q123_datapoor/n_spp_both * 100, 1)
+# n_q4_datapoor <- nrow(spp_list_quads %>% filter(quad == 'q4' & occurcells < 10))
+# p_q4_datapoor <- round(n_q4_datapoor/n_q4 * 100, 1)
+# n_q123_datapoor <- nrow(spp_list_quads %>% filter(quad != 'q4' & occurcells < 10))
+# p_q123_datapoor <- round(n_q123_datapoor/n_spp_both * 100, 1)
 
-n_q4_rev <- nrow(spp_list_quads %>% filter(quad == 'q4' & !is.na(reviewed)))
-p_q4_rev <- round(n_q4_rev/n_q4 * 100, 1)
-n_q123_rev <- nrow(spp_list_quads %>% filter(quad != 'q4' &  !is.na(reviewed)))
-p_q123_rev <- round(n_q123_rev/n_spp_both * 100, 1)
+# n_q4_rev <- nrow(spp_list_quads %>% filter(quad == 'q4' & !is.na(reviewed)))
+# p_q4_rev <- round(n_q4_rev/n_q4 * 100, 1)
+# n_q123_rev <- nrow(spp_list_quads %>% filter(quad != 'q4' &  !is.na(reviewed)))
+# p_q123_rev <- round(n_q123_rev/n_spp_both * 100, 1)
 
-quad_count_means <- read_csv(file.path(dir_data, 'quads_data_poor.csv')) %>%
-  arrange(quad)
-dp_med_am_q4    <- quad_count_means$med_occurcells[4]
-dp_med_am_q123  <- quad_count_means$med_occurcells_total[1]
-dp_med_occ_q4   <- quad_count_means$med_occ[4]
-dp_med_occ_q123 <- quad_count_means$med_occ_total[1]
+# quad_count_means <- read_csv(file.path(dir_data, 'quads_data_poor.csv')) %>%
+#   arrange(quad)
+# dp_med_am_q4    <- quad_count_means$med_occurcells[4]
+# dp_med_am_q123  <- quad_count_means$med_occurcells_total[1]
+# dp_med_occ_q4   <- quad_count_means$med_occ[4]
+# dp_med_occ_q123 <- quad_count_means$med_occ_total[1]
+
+spp_quads_occ <- read_csv(file.path(dir_data, 'spp_quads_occ.csv'))
+med_occ_q4   <- round(median(spp_quads_occ$mean_occ[spp_quads_occ$quad == 'q4']))
+med_occ_q123 <- round(median(spp_quads_occ$mean_occ[spp_quads_occ$quad %in% c('q1', 'q2', 'q3')]))
 
@@
 
-To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (where on the map) against the ratio of extent (how much of the map) for each shared species (Fig.
-2A).
-A general negative linear pattern emerges, <!--BSH: Any stats for this?
-If so, insert just after this phrase.
-Otherwise maybe phrase it as 'The general pattern that emerges of a negative linear relationship, where increasing similarity...'--> suggesting that increasing similarity in extent correlates with decreasing distribution alignment.
-<!--BSH: Could also be said as increasing similarity in alignment leads to decreasing similarity in extent.
-The phrasing implies causality, even though you are careful to use ‘correlates’--> AquaMaps tends to extrapolate species ranges beyond known occurrences, such that the marginal range predicted by AquaMaps will fall in different locations than the marginal range predicted using IUCN methodology.
-For species with dissimilar extents, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
-For species with increasingly similar extents, differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.
+To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e.
+where on the map) against the area alignment (ratio of smaller range area to larger range area, i.e.
+how much of the map) for each shared species (Fig.
+3A).
+ This analysis revealed a general negative linear pattern, suggesting that increasing similarity in range area correlates with decreasing distribution alignment.
+ AquaMaps tends to extrapolate species ranges beyond known occurrences, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
+For species with dissimilar range areas, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
+For species with increasingly similar range areas, differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.
 
-___directly address the cluster along the top right of the plot?___
+The mean distribution alignment for species included in both datasets was `r dist_align_mean`%; the mean area alignment was `r area_align_mean`%.
+By dividing the map-paired species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
+The upper right quadrant includes species (n = `r n_q1`) whose described ranges are above average in both spatial distribution and extent.
+These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig.
+3B).
+This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their ranges span nearly the entire map.
 
-By dividing the map-paired species into quadrants, we highlight different categories of relationships that in turn help further explain the general pattern.
-The upper right quadrant includes species (n = `r n_q1`) whose described ranges agree in both spatial distribution and extent.
-These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig.
-2B).
-This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because they exist nearly everywhere.
-
-The extent-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where the IUCN range is notably larger, often extending into deeper water.
+The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where the IUCN range is notably larger, often extending into deeper water.
 For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their preferred depths, likely introducing errors of commission.
-<!-- BSH: maybe include 2-3 examples as figures in the SOM and refer to them here.-->
+___See supplemental information for maps that highlight this issue.___ _should there just be a blanket "see SOM for maps that typify each quadrant"?_
 Ocean depth is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps.
@@
 Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, introducing additional commission errors.
-Consequently, area ratios are close to 100%, though similar extents are meaningless when the distributions are poorly aligned.
+Consequently, area ratios are close to 100%, though similar areas are meaningless when the distributions are poorly aligned.
 
-The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor extent match well.
-Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (summing (averaging?) OBIS and GBIF occurrences) for this quadrant is `r dp_med_am_q4`, compared to a median of `r dp_med_am_q123` for species across the other three quadrants.
+The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor area match well.
+Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging OBIS and GBIF occurrences) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
 When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species.
+
+___While we examined IUCN Red List extinction risk categories by quadrant, we found no compelling patterns.  See SOM for details.___
 
@@
 
-Clear and significant differences in species range descriptions between AquaMaps and IUCN range maps, due to method-driven differences in errors of commission and omission.
+Clear and significant differences in species range descriptions between AquaMaps and IUCN range maps due to method-driven differences in commission and omission errors.
 To examine the implications of these differences, we replicated two recent studies, varying only the prioritization of one data set over the other.
 
-**Case Study: The Ocean Health Index** <!--BSH: I stopped editing and worksmithing here and just focused on addressing your comments/questions from here on-->
+**Case Study: The Ocean Health Index**
 
@@
 2012, 2015), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
-To maximize the number of represented species, OHI gleans spatial distributions from both IUCN and AquaMaps datasets, prioritizing IUCN data for the `r n_spp_both` species included in both sources.
-OHI currently uses a probability threshold of 40% to determine species presence for AquaMaps data.
+To maximize the number of represented species, OHI gleans spatial distributions from both IUCN and AquaMaps datasets, prioritizing IUCN data for the `r format(n_spp_both, big.mark = ',')` species included in both sources.
+OHI currently uses a probability of occurrence threshold of 40% to determine species presence for AquaMaps data.
 
-We recalculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig.
-3</a>).
-Given that only `r n_spp_both` species overlap out of `r n_spp_all`, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in such large score shifts as seen in scenarios 2 and 3.
-While many country scores showed only small variation across all three scenarios, select countries gained as many as 5 points while others dropped up to 12 points.
-Significantly, the global score decreased by two full points in scenario 3, indicating that an arbitrary change in how the two datasets are combined results in a more dire assessment of species conservation status.
+We recalculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig4'>Fig.
+4</a>).
+ Reducing the threshold increases the apparent range of a species; the slight decrease in average score for scenario 1 suggests increased spatial representation of threatened species.  
+ The slight bump in mean score for scenario 2 may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or a combination of both.
+ The large decrease shown in scenario 3 indicates that a zero threshold greatly increases the spatial representation of high-risk species.
+
+Given that only `r format(p_both_all, big.mark = ',')` percent of species overlap, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in such large country-level score shifts as seen in scenarios 2 and 3.
+While many country scores showed only small variation across all three scenarios, select countries gained up to 5 points while others dropped as many as 12 points.
+Significantly, the global score decreased by two full points in scenario 3, indicating that an arbitrary change in how the two datasets are combined can result in a more dire assessment of species conservation status.
 
@@
 Klein et al.
-(2015) compared the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect key species and to identify which species fall through gaps in protection.
-The study relied on the AquaMaps database, using a probability threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
-The study determined that currently established MPAs (IUCN categories I through IV overlapping with marine areas) are woefully inadequate in protecting marine biodiversity; 90.5% of species have less than 5% of their overall range represented within MPAs and 1.4% of species have no protection at all ("gap" species).
+(2015) compare the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect key species and identify which species fall through gaps in protection.
+The study relied on the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
+The study determined that the global MPA network is woefully inadequate in protecting marine biodiversity: 90.5% of species have less than 5% of their overall range represented within MPAs and 1.4% of species have no protection at all ("gap" species).
 
- (<a href = '#fig4'>Figure 4</a>) displays the results of applying this methodology under several different scenarios.
-When comparing the conclusions based upon an IUCN-only scenario to an AquaMaps-only scenario (using 2015 data and a 0% threshold sets the most meaningful comparison) we found a five-fold increase in the proportion of gap species (6.4% of species vs 1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% of species vs.
+ (<a href = '#fig5'>Figure 5</a>) displays the results of applying this methodology under new scenarios.
+When comparing the conclusions based upon an IUCN-only scenario to an AquaMaps-only scenario (using 2015 AquaMaps data and a 0% threshold sets the most meaningful comparison) we found a five-fold increase in the proportion of gap species (6.4% of species vs.
+1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% of species vs.
 47.7%).
@@
 Conclusions drawn from each of these datasets would paint dramatically different pictures of the protection afforded by our current global MPA network.
+
+___Should end with a more solid statement of how choosing one or the other datasets/thresholds would have different results – using our calculated differences.  JA___
+
+___If you have room, I think you could add a bit more interpretation of these results here.
+Could discuss the commission and omission stuff here too to describe these somewhat non-intuitive results???
+CS___
 

### 2016-03-10 ms_4_figs.Rmd
updates to manuscript based on CS and JA comments

@@
 
-### Fig. 1 <!--can you put N = XXX above each bar? Would make it a little easier to read compared to within the bar.-->
+### Fig. 1
 <a id = 'fig1'></a>
@@
 
+>  __Fig.
+1.__ (A) Number and proportion of species by taxa included in each dataset.
+Overlapping species are dominated by bony fishes (`r n_bonyfish` species, primarily tropical taxa) and corals (`r n_corals` species).
+
+### Fig. 2
+<a id = 'fig2'></a>
+
 ![](figures/all_aquamaps_map.png)
@@
 
->  __Fig.
-1.__ (A) Number and proportion of species by taxa included in each dataset.
-AquaMaps encompasses a broader range of taxa than IUCN, while IUCN focuses on comprehensively assessing select taxonomic groups.
-Overlapping species are dominated by bony fishes (`r n_bonyfish` species, primarily tropical taxa) and corals (`r n_corals` species).
-(B, C) Global marine species richness according to (B) AquaMaps and (C) IUCN.
-The margin frequency plots show relative species count per cell at each latitude and longitude; while both datasets peak in tropical latitudes, the frequency for IUCN maps drops quickly beyond 30°N and 30°S, while the frequency for AquaMaps remains robust well into temperate latitudes.
-The longitude frequency plots show a slight bias in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
+>  __Fig. 2.__(A, B) Number of global marine species according to (A) AquaMaps and (B) IUCN.  The margin frequency plots show relative species count per cell at each latitude and longitude.
 
-<a id = 'fig2'></a>
+### Fig. 3
+<a id = 'fig3'></a>
 
@@
 > __Fig.
-2.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r n_spp_both` species included in both IUCN and AquaMaps datasets.
+3.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r n_spp_both` species included in both IUCN and AquaMaps datasets.
 The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
@@
 (n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).  
-(B) Alignment breakdown of paired-map species by taxonomic group.
-
-### Figure 3
-
-<a id = 'fig3'></a>
-
-![Absolute change in SPP score](figures/boxplot_abs_diff.png)
-
-> __Fig. 3.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
-> Scenario 1 shows the effect of reducing the probability threshold for AquaMaps presence to more accurately track the definition of IUCN “limits of distribution."  Reducing the threshold increases the apparent range of a species; the slight decrease in average score suggests increased spatial representation of threatened species.
-> Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the 40% presence threshold.
-In general, AquaMaps ranges are smaller than IUCN ranges, so most overlapping species will see a decrease in represented range.
-The slight bump in mean score may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or a combination of both.
-> Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, and simultaneously eliminating the presence threshold.
-The zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
-The large decrease indicates that a zero threshold greatly increases the spatial representation of high-risk species.
-
-___NOTE: This figure still needs to be recreated using the latest and greatest data!___ _currently they use data generated in November or so, in ohiprep._
+(B) Alignment quadrant breakdown of paired-map species by taxonomic group.
 
@@
 
+![Absolute change in SPP score](figures/boxplot_abs_diff.png)
+
+> __Fig. 4.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
+> Scenario 1 shows the effect of reducing the probability threshold to 0% for AquaMaps presence to more accurately track the definition of IUCN “limits of distribution."  
+> Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the 40% presence threshold.
+In general, AquaMaps ranges are smaller than IUCN ranges, so most overlapping species will see a decrease in represented range.
+> Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, and simultaneously eliminating the presence threshold.
+The zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
+
+___NOTE: This figure still needs to be recreated using the latest and greatest data!___ _currently they use data generated in November or so, in ohiprep._
+
+### Figure 5
+
+<a id = 'fig5'></a>
+
 ![Change in MPA coverage](figures/barchart_mpa_gap.png)
@@
 > __Fig.
-4.__  Percent of species range covered by MPAs based upon methods in Klein et al.
+5.__  Percent of species range covered by MPAs based upon methods in Klein et al.
 (2015).
@@
 
-`r knitr::kable(read_csv(file.path(dir_data, 'mpa_all_runs_summary.csv')))`
+<!-- `r knitr::kable(read_csv(file.path(dir_data, 'mpa_all_runs_summary.csv')))` -->

### 2016-03-10 manuscript-main.Rmd
updating manuscript

@@
 title: "Aligning marine species range data to better serve science and conservation"
-author: "Casey O'Hara^1^, Jamie Afflerbach^1^, Benjamin S. Halpern<sup>1, 2, 3</sup>"
+author: "Casey O'Hara^1^, Jamie Afflerbach^1^, Benjamin S. Halpern<sup>1, 2, 3</sup>, Courtney Scarborough^1^"
 output:

### 2016-03-10 mpa_analysis/3_mpa_spp_analysis.Rmd
created IUCN spp cell summary to identify gap spp and 10%+ spp
updating manuscript

@@
 ### compare results using only IUCN species maps
-dir_iucn_data <- '/var/data/ohi/git-annex/globalprep/_raw_data/iucn_spp'
+dir_iucn_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/iucn_spp')
 iucn_cells_file <- file.path(dir_iucn_data, 'iucn_cells_v2016.csv')

### 2016-03-10 ms_2_intro_res_disc.Rmd
updating manuscript

@@
 2006).
-___By comparing maps resulting from each methodology, we can identify and possibly address mechanistic causes for each type of error.___
-
-<!--BSH: I haven’t read the paper, but it sounds like we need to acknowledge what was already discussed/done there.
-This may actually simplify things, such that we condense the first three paragraphs of this section into a single paragraph that restates what that paper found (and cite it) and add anything else we want to point out.
-But we will definitely need a succinct statement about how our paper moves well beyond what that paper did.
-Not sure best approach without reading the paper.-->
+By comparing maps resulting from IUCN and AquaMaps methodolog, we can identify and possibly address mechanistic causes for each type of error.
 
@@
  This analysis revealed a general negative linear pattern, suggesting that increasing similarity in range area correlates with decreasing distribution alignment.
- AquaMaps tends to extrapolate species ranges beyond known occurrences, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
+ AquaMaps tends to extrapolate species ranges into suitable areas beyond known occurrences, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
 For species with dissimilar range areas, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
@@
 By dividing the map-paired species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
+For representative maps from each category, please refer to the supplemental materials (SI Fig.
+XXX).
+
 The upper right quadrant includes species (n = `r n_q1`) whose described ranges are above average in both spatial distribution and extent.
@@
 The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where the IUCN range is notably larger, often extending into deeper water.
-For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their preferred depths, likely introducing errors of commission.
-___See supplemental information for maps that highlight this issue.___ _should there just be a blanket "see SOM for maps that typify each quadrant"?_
+For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their preferred depths, likely introducing errors of commission  (SI Figs.
+XXX, XXX).
  Ocean depth is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps.
@@
 
-___While we examined IUCN Red List extinction risk categories by quadrant, we found no compelling patterns.  See SOM for details.___
+While we examined IUCN Red List extinction risk categories by quadrant, expecting to see more concurrence between maps for species of concern; however, we found no compelling patterns (SI Fig. XXX).
 
@@
 
-Clear and significant differences in species range descriptions between AquaMaps and IUCN range maps due to method-driven differences in commission and omission errors.
+Method-driven differences in commission and omission errors drive clear and significant differences in species range descriptions between AquaMaps and IUCN datasets.
 To examine the implications of these differences, we replicated two recent studies, varying only the prioritization of one data set over the other.
@@
 
-___Should end with a more solid statement of how choosing one or the other datasets/thresholds would have different results – using our calculated differences.  JA___
-
 ___If you have room, I think you could add a bit more interpretation of these results here.
@@
 CS___
+
+- check which species are gap species and look at some maps
+- check which species are 10%+ species and look at some maps
+- where do these species fall georegionally? is there anything odd about their ranges?
 
@@
 
-__Suggestions for data providers:__
+The IUCN Red List instructional materials for mapping marine species suggest that "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species" (IUCN 2015).
+Especially for demersal communities dependent on photosynthesis, bathymetry should be a primary consideration; data users might consider a simple cutoff at the photosynthetic limit of 200 meters to minimize a likely source of commission errors.
 
-* The IUCN Red List instructional materials for mapping marine species suggest that "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species" (IUCN 2015).
-Unfortunately, this recommendation appears to be applied inconsistently.
-Especially for demersal communities dependent on photosynthesis, bathymetry should be a primary consideration; even a simple cutoff at the photosynthetic limit of 200 meters would minimize a likely source of commission errors.
-* The AquaMaps model is a powerful tool for quickly estimating species distribution based on limited information; however, the model's output is only as good as the input data, which for marine species is often sorely limited.
-AquaMaps encourages experts to review and comment on the predicted species distribution maps; lessons from the small sample of reviewed maps could be generalized, for example to other species in the same taxonomic group, to further refine the predictions.
-This is especially important for data-poor species.
-* In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.
-Currently area restrictions are denoted by UN Food and Agriculture Organisation (FAO) fisheries reporting areas (Ready et al.
+The AquaMaps methodology is a powerful means of quickly estimating species distribution based on limited information; however, the model's output is only as good as the input data, which for marine species is often sorely limited.
+In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.
+Currently, these area restrictions are denoted by UN Food and Agriculture Organisation (FAO) fisheries reporting areas (Ready et al.
 2010); however, area restrictions based on biogeographical criteria, e.g.
 Marine Ecoregions of the World (Spalding et al.
-2007), would likely provide better resolution and predictive power, especially for data-poor species.
+2007), would likely decrease commission errors and improve predictive power, especially for data-poor species.
 
-__Suggestions for data users:__
+When using AquaMaps data alone, changing the probability of occurrence threshold to determine presence becomes a tradeoff between errors of commission (low threshold) and omission (high threshold).
+Using AquaMaps in conjunction with IUCN can increase the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized.
+In this case, we recommend a presence threshold of 0% as it most closely approximates the "limits of distribution" criterion defined by IUCN data providers.
 
-* For depth-limited species and taxa, clipping IUCN range polygons to a reasonable bathymetry contour can reduce commission errors due to overprediction of species presence into unsuitable habitats.
-* Using both datasets together can increase the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized.  Consider:
-    * An AquaMaps presence threshold of 0% most closely approximates the "limits of distribution" defined by IUCN range maps.
-    * Additional filters on AquaMaps distribution criteria, such as expert review status or the number of occurrence cells used to generate the species model, can help avoid overextrapolation for data poor species.
+<!-- * Additional filters on AquaMaps distribution criteria, such as expert review status or the number of occurrence cells used to generate the species model, can help avoid overextrapolation for data poor species.
+_using other attributes and variables to filter?
+this seems outside the scope of this paper_ -->
 

### 2016-03-10 ms_3_methods.Rmd
updating manuscript

@@
 To examine the overall taxonomic distribution across the spatial datasets (Fig.
-1A), we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
+1), we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
 
-To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; species presence within a grid cell was determined by any non-zero overlap of a species polygon with the cell, and species richness per cell was simply the count of the species present.
-For the AquaMaps dataset, we determined per-cell species richness by counting all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.
-We represented relative distribution of species richness for each dataset by plotting average species count against latitude and longitude (Fig.
-1B, 1C).
+To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; species presence within a grid cell was determined by any non-zero overlap of a species polygon with the cell, and number of species per cell was simply the count of the species present.
+For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.
+We represented relative distribution of species count for each dataset by plotting average species count against latitude and longitude (Fig.
+A, 2B).
 
@@
 
-Overlaying paired distribution maps for a given species, we defined and calculated _distribution alignment_ and _extent alignment_ and plotted these in Fig. 2:
+Overlaying paired distribution maps for a given species, we defined and calculated _distribution alignment_ and _extent alignment_ and plotted these in Fig. 3A:
 
@@
 
+For each species with ranges described in both IUCN and AquaMaps, $A_{small}$ and $A_{large}$ indicate the smaller and larger range representation (regardless of which dataset).
+$A_{small \cap large}$ represents the amount of overlapping area between the two datasets.
 We visually inspected a random selection of paired distribution maps from each quadrant to identify possible mechanistic causes of misalignment.

### 2016-03-11 data_explore/data_explore.Rmd
created mapping chunk for iucn gap and 10+ species cells

@@
 
+``` {r plot_spp_iucn_gaps_tenplus, eval = TRUE, echo = FALSE}
+
+iucn_gap10     <- read_csv(file.path(dir_data, 'iucn_gaps_tenplus.csv'))
+iucn_gap10_spp <- unique(iucn_gap10[iucn_gap10$sciname %in% spp_list$sciname, ])
+
+iucn_gapmaps <- sample_n(iucn_gap10_spp %>% filter(cat == 'gap'), 10)
+iucn_10maps  <- sample_n(iucn_gap10_spp %>% filter(cat != 'gap'), 10)
+
+if(!exists('iucn_spp_cells')) 
+  iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'))
+
+for (spp in iucn_gapmaps$sciname) { # i = 1
+  spp_plot <- plot_rangemap(spp)
+  spp_solo <- spp_list %>% filter(sciname == spp)
+  spp_plot <- spp_plot +
+    labs(title = sprintf('%s: %s', spp, spp_solo$spp_group_text))
+
+  print(spp_plot)
+}
+
+```
 ***

### 2016-03-11 data_explore/data_explore.Rmd
updating manuscript, and creating maps for gaps and 10+% MPAs

--- 2016-03-10 data_explore/data_explore.Rmd
@@
 
+``` {r plot_spp_iucn_gaps_tenplus, eval = TRUE, echo = FALSE}
+
+iucn_gap10     <- read_csv(file.path(dir_data, 'iucn_gaps_tenplus.csv'))
+iucn_gap10_spp <- unique(iucn_gap10[iucn_gap10$iucn_sid %in% spp_list$iucn_sid, ])
+
+iucn_gapmaps <- sample_n(iucn_gap10_spp %>% filter(cat == 'gap'), 10)
+iucn_10maps  <- sample_n(iucn_gap10_spp %>% filter(cat != 'gap'), 10)
+
+if(!exists('iucn_spp_cells')) {
+  iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'),
+                             col_types = 'cddddc')
+  am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'))
+}
+
+am_thresh = 0
+
+message('Printing gap maps')
+for (i in 1:nrow(iucn_gapmaps)) { ### i <- 1
+  spp <- iucn_gapmaps[i, ]
+
+  spp_plot <- plot_rangemap(spp)
+  spp_plot <- spp_plot +
+    labs(title = sprintf('%s: %s', spp$sciname, spp$spp_group_text))
+
+  print(spp_plot)
+}
+
+message('Printing >10% maps')
+
+for (i in 1:nrow(iucn_10maps)) { ### i <- 1
+  spp <- iucn_10maps[i, ]
+  spp_plot <- plot_rangemap(spp)
+  spp_plot <- spp_plot +
+    labs(title = sprintf('%s: %s', spp$sciname, spp$spp_group_text))
+
+  print(spp_plot)
+}
+
+```
 ***

### 2016-03-11 manuscript-main.Rmd
updating manuscript, and creating maps for gaps and 10+% MPAs

--- 2016-03-10 manuscript-main.Rmd
@@
 title: "Aligning marine species range data to better serve science and conservation"
-author: "Casey O'Hara^1^, Jamie Afflerbach^1^, Benjamin S. Halpern<sup>1, 2, 3</sup>"
+author: Casey O'Hara^1^, Jamie Afflerbach^1^, Benjamin S. Halpern<sup>1, 2, 3</sup>,
+  Courtney Scarborough^1^
 output:
+  word_document: default
+  pdf_document:
+    number_sections: yes
+    toc: no
   html_document:
@@
     toc: no
-  pdf_document:
-    number_sections: yes
-    toc: no
-  word_document: default
 ---

### 2016-03-11 ms_2_intro_res_disc.Rmd
updating manuscript, and creating maps for gaps and 10+% MPAs

@@
 
-___BSH:___ _I think Julie makes this comment below, but we may need to be able to substantiate this somehow.  Is there a way to quickly figure out how many studies have used AM vs. IUCN?_
+___BSH comment:___ _I think Julie makes this comment below, but we may need to be able to substantiate this somehow.
+Is there a way to quickly figure out how many studies have used AM vs.
+IUCN?_  ___CCO: I can give # of citations for AquaMaps foundational papers (~315); IUCN is tougher since IUCN citations aren't necessarily spatial or marine.
+How important is this - i.e.
+can I get away without making a big deal here?
+or making a sweeping statement of some sort??___
 
+<!--
 ___CCO:___ _What about age - AquaMaps is pretty recent (2006 at earliest); before that, would IUCN have been the dominant dataset for marine?
@@
 _Three foundational papers relating to AquaMaps (Kaschner et al. 2006, 2011; Ready et al. 2010) have been cited a total of 315 times._
+-->
 
@@
  (<a href = '#fig5'>Figure 5</a>) displays the results of applying this methodology under new scenarios.
-When comparing the conclusions based upon an IUCN-only scenario to an AquaMaps-only scenario (using 2015 AquaMaps data and a 0% threshold sets the most meaningful comparison) we found a five-fold increase in the proportion of gap species (6.4% of species vs.
+When comparing the conclusions based upon an IUCN-only scenario to an AquaMaps-only scenario (using 2015 AquaMaps data and a 0% threshold sets the most meaningful comparison to IUCN's "limits of distribution") we found a five-fold increase in the proportion of gap species (6.4% of species vs.
 1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% of species vs.
@@
 
-___If you have room, I think you could add a bit more interpretation of these results here.
-Could discuss the commission and omission stuff here too to describe these somewhat non-intuitive results???
-CS___
+___working on checking why so many gaps and 10+ species show up with IUCN.  I will write up a paragraph once I've had a little more time to examine maps___
 
-- check which species are gap species and look at some maps
-- check which species are 10%+ species and look at some maps
-- where do these species fall georegionally? is there anything odd about their ranges?
+- check which species are gap species; check maps
+- check which species are 10%+ species and check maps
+- where do these species fall georegionally? is there anything odd about their ranges? any reason why IUCN would be 
+- hypothesis to explain more gap species: Aquamaps more diffuse than IUCN, catching individual cells that could be suitable; a single cell (even low probability) moves a species from "gap" to non-gap status.
+IUCN more clustered by polygon boundaries; so no scattered cells to accidentally fall into an MPA.
+    - introducing errors: AM = likely commission (overestimates species with tiny bits of range in MPAs) vs IUCN = possible omission (misses possible real habitat that falls within MPAs)
+- hypothesis to explain more 10%+ species - ten random samples most have IUCN polygons just east of Australia where a large MPA is indicated - Great Barrier Reef I assume - a single large poly falling in GBR gets a high score, while for the same species, many of the AM ranges (also in same area) scatter suitable area over larger extent, so lower percent of species range is protected
+    - introducing errors: IUCN = likely omission (skipping possible real habitat outside of MPAs) and likely commission (possibly overcounting real habitat in MPAs e.g.
+GBR); AquaMaps = probably a little commission (possibly underestimating, if it overpredicts lots of range outside MPAs, thus reducing the proportion of protected range)
 
@@
 
-<!-- * Additional filters on AquaMaps distribution criteria, such as expert review status or the number of occurrence cells used to generate the species model, can help avoid overextrapolation for data poor species.
-_using other attributes and variables to filter?
-this seems outside the scope of this paper_ -->
-
-_(do we need a wrap-it-up paragraph here?  I put that more at the top of the conclusions section... could shift it down here, but probably not needed in both places)_
+AquaMaps and IUCN are two of the most important databases providing species distribution information across a wide range of marine species, taxa, and regions.
+The different methodologies behind the creation of each database is likely to introduce different rates of commission and omission errors, which can have a profound impact on our understanding of species ranges.
+By comparing overlapping species maps from these two databases, we have identified likely sources of error, allowing data users to address or account for them.
+For users of AquaMaps, IUCN, and other datasets of species distribution, reducing the likelihood of these errors is critical to effective biodiversity research and conservation action.

### 2016-03-11 data_explore/data_explore_fxns.R

@@ -181,12 +181,12 @@ get_spp_map <- function(sid, spp_cells) {
 ### Plot Species Map function
 ### uses ggplot to create a formatted map of species ranges, including
 ### Aquamaps, IUCN, and overlapped ranges; returns the plot object.
-plot_rangemap <- function(sid) {
-  map_iucn <- get_spp_map(sid, iucn_spp_cells) %>%
+plot_rangemap <- function(spp) {
+  map_iucn <- get_spp_map(spp$iucn_sid, iucn_spp_cells) %>%
     mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0)) %>%
     select(-area_iucn, -iucn_sid) %>%
     unique()
-  map_am   <- get_spp_map(sids$am_sid, am_spp_cells) %>%
+  map_am   <- get_spp_map(spp$am_sid, am_spp_cells) %>%
     mutate(am_pres = ifelse(am_prob >= am_thresh, 1, 0)) %>%
     select(-am_sid, -am_prob) %>%
     unique()
@@ -218,7 +218,7 @@ plot_rangemap <- function(sid) {
     borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
     scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
     scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
-    labs(title = sprintf('IUCN species ID: %s', sid), x = NULL, y = NULL) 
+    labs(title = sprintf('IUCN species ID: %s', spp$iucn_sid), x = NULL, y = NULL) 
 }
 
 #################################################################=

### 2016-03-18 ms_2_intro_res_disc.Rmd
Ben's comments, some fixes

@@
 
-While we examined IUCN Red List extinction risk categories by quadrant, expecting to see more concurrence between maps for species of concern; however, we found no compelling patterns (SI Fig. XXX).
+We examined IUCN Red List extinction risk categories by quadrant, expecting to see more concurrence between maps for species of concern; however, we found no compelling patterns (SI Fig. XXX).
 
@@
 
-Given that only `r format(p_both_all, big.mark = ',')` percent of species overlap, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in such large country-level score shifts as seen in scenarios 2 and 3.
+Given that only `r format(p_both_all, big.mark = ',')`% of species overlap, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in such large country-level score shifts as seen in scenarios 2 and 3.
 While many country scores showed only small variation across all three scenarios, select countries gained up to 5 points while others dropped as many as 12 points.
@@
 
- (<a href = '#fig5'>Figure 5</a>) displays the results of applying this methodology under new scenarios.
+<a href = '#fig5'>Figure 5</a> displays the results of applying this methodology under new scenarios.
 When comparing the conclusions based upon an IUCN-only scenario to an AquaMaps-only scenario (using 2015 AquaMaps data and a 0% threshold sets the most meaningful comparison to IUCN's "limits of distribution") we found a five-fold increase in the proportion of gap species (6.4% of species vs.

### 2016-03-21 data_explore/data_explore_global_maps.Rmd
incorporated Ben's comments; reframing a conclusion; preparing figures.  Moved the MPA analysis explanation of shifts into the SI

@@
 ###Global representation of species
+
+```{r setup, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+
+# Libraries and Paths
+
+library(ggplot2)
+library(maps)
+library(readr)
+library(data.table)
+library(raster)
+library(tidyr)
+library(dplyr)
+library(stringr)
+
+dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
+           'Darwin'  = '/Volumes/data_edit',
+           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
+
+dir_git <- '~/github/IUCN-AquaMaps'
+
+dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
+dir_fig <- file.path(dir_git, 'figures')
+dir_data <- file.path(dir_git, 'data')
+
+if(basename(getwd()) != 'data_explore') setwd('data_explore')
+source('data_explore_fxns.R')
+### When knitting this, it automatically sets WD to be this directory...
+### the 'setwd()' is there for running by chunk
+```
 
@@
 
-```{r global map matched species, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+```{r global_maps, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 
-cat('First off: map-paired species.  Map 1: AM dataset; map 2: IUCN dataset; maps 3 & 4: log-transformed AM & IUCN\n')
-
-### find total species in each dataset for later percent calcs
-# am_n_tot   <- am_spp_cells %>% 
-#   filter(am_sid %in% unique(spp_list$am_sid)) %>%
-#   dplyr::select(am_sid) %>% 
-#   unique() %>% nrow()
-# iucn_n_tot <- iucn_spp_cells %>% 
-#   filter(sciname %in% unique(spp_list$sciname)) %>%
-#   dplyr::select(sciname) %>% 
-#   unique() %>% nrow()
-### good - the same number
-
-### Number of aquamaps species per cell
-iucn_pair_rast_file <- file.path(dir_data, 'rasters/iucn_pair_rast.tif')
-am_pair_rast_file   <- file.path(dir_data, 'rasters/am_pair_rast.tif')
-
-if(!file.exists(iucn_pair_rast_file) | !file.exists(am_pair_rast_file)) {
-  am_n_per_cell <- am_spp_cells %>%
-  filter(am_sid %in% unique(spp_list$am_sid)) %>%
-  group_by(loiczid) %>%
-  summarize(count_am = n()) %>%
-#   summarize(count_am = n(),
-#             pct_am = count_am/am_n_tot) %>%
-  as.data.frame()
-
-  am_sp_map <- subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'count_am')], 
-                    by = 'loiczid', 
-                    which = 'count_am', 
-                    subsWithNA = TRUE)
-
-  ### Number of IUCN species per cell
-  iucn_n_per_cell <- iucn_spp_cells %>%
-    group_by(loiczid) %>%
-    summarize(count_iucn = n()) %>%
-  #   summarize(count_iucn = n(),
-  #             pct_iucn = count_iucn/iucn_n_tot) %>%
-    as.data.frame()
-
-  iucn_sp_map <- subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'count_iucn')], 
-                      by = 'loiczid', 
-                      which = 'count_iucn', 
-                      subsWithNA = TRUE)
-
-  writeRaster(iucn_sp_map, filename = iucn_pair_rast_file, overwrite = TRUE)
-  writeRaster(am_sp_map, filename = am_pair_rast_file, overwrite = TRUE)
-} else {
-  am_sp_map   <- raster(am_pair_rast_file)
-  iucn_sp_map <- raster(iucn_pair_rast_file)
-}
-
-am_ln_map <- log(am_sp_map + 1)
-iucn_ln_map <- log(iucn_sp_map + 1)
-
-### Plot both maps
 myTheme <- rasterTheme(region = brewer.pal('Blues', n = 9))
 
-levelplot(am_sp_map, par.settings = myTheme, 
-#          main = 'Aquamaps Dataset', 
-          colorkey = list( at = seq(0, 1063, length = 15)))
-
-levelplot(iucn_sp_map, par.settings = myTheme, 
-#          main = 'IUCN Dataset', 
-          colorkey = list(at = seq(0, 1063, length = 15)))
-
-levelplot(am_ln_map, par.settings = myTheme, 
-#          main = 'Aquamaps Dataset', 
-          colorkey = list( at = seq(0, maxValue(am_ln_map), length = 15)))
-
-levelplot(iucn_ln_map, par.settings = myTheme, 
-#          main = 'IUCN Dataset', 
-          colorkey = list(at = seq(0, maxValue(iucn_ln_map), length = 15)))
-
-```
-
-Now look at full aquamaps species
-
-```{r all_aquamaps, eval = FALSE, echo = FALSE}
-
-cat('Full AM dataset, species richness\n')
+message('Full AM dataset, species count\n')
 
@@
 
-  ### find total spp to allow for percent calcs later
-  # am_all_n_tot <- am_spp_cells_big %>% 
-  #   dplyr::select(SpeciesID) %>% 
-  #   unique() %>% nrow()
-
   all_am <- am_spp_cells_big %>%
@@
     summarize(all_count = n()) %>%
-  #   summarize(all_count = n(),
-  #             all_pct   = all_count/am_all_n_tot) %>%
     as.data.frame()
@@
 
+### set up boundaries - from http://oscarperpinan.github.io/rastervis/FAQ.html#sec-7
+ext <- as.vector(extent(all_am_map))
+
+boundaries <- maps::map('world', fill=TRUE,
+    xlim=ext[1:2], ylim=ext[3:4],
+    plot=FALSE)
+
+## read the map2SpatialPolygons help page for details
+IDs <- sapply(strsplit(boundaries$names, ":"), function(x) x[1])
+bPols <- maptools::map2SpatialPolygons(boundaries, IDs = IDs,
+                              proj4string = CRS(projection(all_am_map)))
+
+### create plot, add boundary polygons
 am_levelplot1 <- levelplot(all_am_map, 
-#                           main = 'Entire Aquamaps Dataset',
-                           par.settings = myTheme)
+                           scales = list(at = seq(-180, 180, 45)),
+                           xlab = NULL, ylab = NULL,
+                           par.settings = myTheme) + 
+  layer(sp.polygons(bPols, fill = 'grey90', col = 'grey30', lwd = .25))
 
-# all_am_ln_map <- log(all_am_map + 1)
-# all_am_ln_map <- all_am_ln_map/maxValue(all_am_ln_map)
-# am_levelplot1 <- levelplot(all_am_ln_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset: log-transformed relative species richness', 
-#           colorkey = list(at = seq(0, 1, length = 15)))
-
+png(file.path(dir_fig, 'all_aquamaps_map1.png'),
+    width = 8.7, height = 6, units = 'in', res = 300)
 print(am_levelplot1)
-# print(am_levelplot2)
-
-# am_ln_pts <- as.data.frame(rasterToPoints(all_am_ln_map))
-# spp_richness_am_plot <- ggplot(am_ln_pts, aes(x = x, y = y)) +
-#   ggtheme_map + 
-#   geom_raster(aes(fill = layer), alpha = .7) +
-#   scale_fill_continuous(low = 'white', high = 'blue') +
-#   borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
-#   scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
-#   scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
-#   labs(title = 'AquaMaps Dataset: log-transformed relative species richness', x = NULL, y = NULL, fill = NULL) 
-# 
-# spp_richness_am_plot
-# ggsave(file.path(dir_fig, 'all_am_ln_map.png'), height = 4, width = 8)
+dev.off()
 
@@
 
-cat('Full IUCN dataset, species richness\n')
+message('Full IUCN dataset, species count\n')
 
@@
 
-  ### find total spp to allow for percent calcs later
-  # am_all_n_tot <- am_spp_cells_big %>% 
-  #   dplyr::select(SpeciesID) %>% 
-  #   unique() %>% nrow()
-
   all_iucn <- iucn_spp_cells_all %>%
@@
   all_iucn_map   <- raster(all_iucn_rast_file)
+  values(all_iucn_map)[is.na(values(all_am_map))] <- NA
 }
@@
 iucn_levelplot1 <- levelplot(all_iucn_map, 
-                             #main = 'Entire IUCN Dataset',
-                             par.settings = myTheme)
+                             scales = list(at = seq(-180, 180, 45)),
+                             xlab = NULL, ylab = NULL,
+                             par.settings = myTheme) + 
+  layer(sp.polygons(bPols, fill = 'grey90', col = 'grey30', lwd = .25))
 
-# all_iucn_ln_map <- log(all_iucn_map + 1)
-# all_iucn_ln_map <- all_iucn_ln_map/maxValue(all_iucn_ln_map)
-# iucn_levelplot2 <- levelplot(all_iucn_ln_map, 
-#           par.settings = myTheme, 
-#           margin = list(axis = TRUE),
-#           main = 'Entire IUCN Dataset: log-transformed relative species richness', 
-#           colorkey = list(at = seq(0, 1, length = 15)))
-
+png(file.path(dir_fig, 'all_iucn_map1.png'),
+    width = 8.7, height = 6, units = 'in', res = 300)
 print(iucn_levelplot1)
-# print(iucn_levelplot2)
-# iucn_ln_pts <- as.data.frame(rasterToPoints(all_iucn_ln_map))
-# spp_richness_iucn_plot <- ggplot(iucn_ln_pts, aes(x = x, y = y)) +
-#   ggtheme_map + 
-#   geom_raster(aes(fill = layer), alpha = .7) +
-#   scale_fill_continuous(low = 'white', high = 'blue') +
-#   borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
-#   scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
-#   scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
-#   labs(title = 'IUCN Dataset: log-transformed relative species richness', x = NULL, y = NULL, fill = NULL) 
-# 
-# spp_richness_iucn_plot
-# ggsave(file.path(dir_fig, 'all_iucn_ln_map.png'), height = 4, width = 8)
+dev.off()
 

### 2016-03-21 ms_1_abst_sig.Rmd
incorporated Ben's comments; reframing a conclusion; preparing figures.  Moved the MPA analysis explanation of shifts into the SI

@@
 Species distribution data provide the foundation for a wide range of ecological research studies and conservation management decisions, yet most species ranges remain unknown, and existing range maps often suffer from data limitations and show inconsistencies.
-AquaMaps (Kaschner et al.
-___2015???___) and the International Union for Conservation of Nature (IUCN) (IUCN 2015) are two distinct efforts to map marine species distributions at a global scale.
-Together these two databases represent `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` species included in both.
-Here we examine differences in predicted species ranges between the two datasets, and find that these misalignments mainly result from divergent methodologies that introduce differing frequencies of commission and omission errors.
-We illustrate the scientific and management implications of these differences by repeating two recent applications - an assessment of global biodiversity within the Ocean Health Index (Halpern et al.
-2012, 2015) and a global analysis of gaps in coverage of marine protected areas (Klein et al.
-2015) - and find significantly different estimates of the status of global biodiversity and effectiveness of conservation depending on how the two datasets were prioritized.
-Until a single, highly accurate dataset of global marine species ranges becomes available, understanding the implications of dataset differences for conservation planning and decision-making will be essential.
+AquaMaps and the International Union for Conservation of Nature (IUCN) are two distinct efforts to map marine species distributions at a global scale.
+Together these databases represent `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` shared species.
+Here we examine differences in predicted species ranges between the two datasets and find that these misalignments mainly result from divergent methodologies that introduce differing frequencies of commission and omission errors.
+We illustrate the scientific and management implications of these differences by repeating two recent applications - an assessment of global biodiversity within the Ocean Health Index and a global analysis of gaps in coverage of marine protected areas - and find significantly different results depending on how the two datasets were used.
+Until a single, highly accurate dataset of global marine species ranges becomes available, understanding the implications of dataset differences for conservation planning and decision-making is essential.
 
@@
 
-(223 words; 1593 char with spaces)
+(184 words; 1307 char with spaces)
 

### 2016-03-21 ms_2_intro_res_disc.Rmd
incorporated Ben's comments; reframing a conclusion; preparing figures.  Moved the MPA analysis explanation of shifts into the SI

--- 2016-03-10 ms_2_intro_res_disc.Rmd
@@
 ___2013???___) and International Union for Conservation of Nature (IUCN) (IUCN 2015).
-These two spatial datasets have been used for a wide range of purposes, including assessing marine species status (Halpern et al.
-2012, 2015 _both_; Selig et al.
-2013 _both_), evaluating global biodiversity patterns (Coll et al.
-2010 _AM_; Martin et al.
-2014 _both and others_, Pimm et al.
-2014 _IUCN, mostly focusing on terrestrial but some marine_, Kaschner et al.
+These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status (Halpern et al.
+2012, 2015; Selig et al.
+2013), evaluating global biodiversity patterns (Coll et al.
+2010; Martin et al.
+2014, Pimm et al.
+2014, Kaschner et al.
 2011), predicting range shifts (Molinos et al.
-2015 _AquaMaps_), and setting conservation priorities (Klein et al.
-2015 _AquaMaps_).
+2015), and setting conservation priorities (Klein et al.
+2015).
 
-___BSH:___ _I think Julie makes this comment below, but we may need to be able to substantiate this somehow.  Is there a way to quickly figure out how many studies have used AM vs. IUCN?_
-
-___CCO:___ _What about age - AquaMaps is pretty recent (2006 at earliest); before that, would IUCN have been the dominant dataset for marine?
-Maybe we can track citations of AquaMaps, but citations of IUCN are generic - just cite the Red List - not referencing spatial information specifically, much less marine.
-Going by citation year field on the maps themselves, the oldest marine record I can find is 2004._
-
-_Three foundational papers relating to AquaMaps (Kaschner et al. 2006, 2011; Ready et al. 2010) have been cited a total of 315 times._
+<!--
+Three foundational papers relating to AquaMaps (Kaschner et al. 2006, 2011; Ready et al. 2010) have been cited a total of 315 times.
+-->
 
@@
 2015) - as case studies to explore the implications of prioritizing one data set over the other.
-The results highlight the possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
+The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
 
@@
 
-The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" (IUCN 2015) - essentially a refined extent of occurrence, based on observation records and refined by expert understanding of the species' range and habitat preferences.
-In contrast, AquaMaps models species distribution based environmental preferences (e.g.
-temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
+The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" (IUCN 2015) - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.
+In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
 The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of "probability of occurrence" for each species (Kaschner et al.
@@
 2010).
-Studies frequently define "presence" for AquaMaps by applying a probability threshold, e.g.
-40% or greater probability of occurrence (Halpern et al.
+Studies frequently define "presence" for AquaMaps by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence (Halpern et al.
 2012, 2015).
@@
 AquaMaps encompasses a broader range of taxa than IUCN, as IUCN spatial data files are only available for select taxonomic groups that have been comprehensively assessed.
-While species numbers in both datasets peak in tropical latitudes near the equator, species counts for IUCN maps drops quickly beyond 30°N and 30°S, while species counts for AquaMaps remains robust well into temperate latitudes.
+While species numbers in both datasets peak in tropical latitudes near the equator, species counts for IUCN maps drop quickly beyond 30°N and 30°S, while species counts for AquaMaps remain robust well into temperate latitudes.
 The longitude frequency plots show a slight bias in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
@@
 
-To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e.
-where on the map) against the area alignment (ratio of smaller range area to larger range area, i.e.
-how much of the map) for each shared species (Fig.
+To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area alignment (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig.
 3A).
@@
 This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their ranges span nearly the entire map.
+We expected to see significant concurrence between maps for IUCN Red List species of concern, assuming that increased attention would lead to improved understanding of species ranges; however, we found no compelling patterns to indicate that this was the case (SI Fig.
+XXX).
 
@@
 When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species.
-
-While we examined IUCN Red List extinction risk categories by quadrant, expecting to see more concurrence between maps for species of concern; however, we found no compelling patterns (SI Fig. XXX).
 
@@
 Reducing the threshold increases the apparent range of a species; the slight decrease in average score for scenario 1 suggests increased spatial representation of threatened species.
- The slight bump in mean score for scenario 2 may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or a combination of both.
+The slight increase in mean score for scenario 2 may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or a combination of both.
 The large decrease shown in scenario 3 indicates that a zero threshold greatly increases the spatial representation of high-risk species.
 
-Given that only `r format(p_both_all, big.mark = ',')` percent of species overlap, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in such large country-level score shifts as seen in scenarios 2 and 3.
+Given that only `r format(p_both_all, big.mark = ',')`% of species overlap, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in such large country-level score shifts as seen in scenarios 2 and 3.
 While many country scores showed only small variation across all three scenarios, select countries gained up to 5 points while others dropped as many as 12 points.
@@
 The study relied on the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
-The study determined that the global MPA network is woefully inadequate in protecting marine biodiversity: 90.5% of species have less than 5% of their overall range represented within MPAs and 1.4% of species have no protection at all ("gap" species).
+They found that the global MPA network leaves 90.5% of marine species with less than 5% of their overall range represented within MPAs, and 1.4% of species have no protection at all ("gap" species).
 
- (<a href = '#fig5'>Figure 5</a>) displays the results of applying this methodology under new scenarios.
-When comparing the conclusions based upon an IUCN-only scenario to an AquaMaps-only scenario (using 2015 AquaMaps data and a 0% threshold sets the most meaningful comparison) we found a five-fold increase in the proportion of gap species (6.4% of species vs.
+We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using 2015 AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig5'>Fig.
+5</a>).
+We found a five-fold increase in the proportion of gap species (6.4% of species vs.
 1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% of species vs.
@@
 
-___If you have room, I think you could add a bit more interpretation of these results here.
-Could discuss the commission and omission stuff here too to describe these somewhat non-intuitive results???
-CS___
-
-- check which species are gap species and look at some maps
-- check which species are 10%+ species and look at some maps
-- where do these species fall georegionally? is there anything odd about their ranges?
-
 -----
@@
 
-AquaMaps and IUCN range maps provide important spatial information on marine species ranges, and show reasonable agreement for many well-studied species, but illustrate uncertainty in our understanding of spatial distribution for many others.
-Identifying and addressing differences in these datasets will increase their utility for ecology research and conservation actions.
-We have identified several likely drivers of commission errors and omission errors between these datasets, pointing toward improvements to better align the predictions of species distributions.
+AquaMaps and IUCN range maps show reasonable agreement for many well-studied species, but substantial differences illustrate uncertainty in our understanding of spatial distribution for many others.
+Although many other appriaches exist for species distribution modeling, these two are the only ones applied globally to marine species.
+Identifying and addressing differences in these datasets will increase their utility for research and conservation actions.
+Several likely drivers of commission errors and omission errors between these datasets point to a few important ways to improve range alignment.
 
-The IUCN Red List instructional materials for mapping marine species suggest that "bathymetry can be used to delineate species' ranges limited by depth in the same way as elevation is used for terrestrial species" (IUCN 2015).
-Especially for demersal communities dependent on photosynthesis, bathymetry should be a primary consideration; data users might consider a simple cutoff at the photosynthetic limit of 200 meters to minimize a likely source of commission errors.
+For IUCN range data, clipping ranges to known depth limits improves output for many species, most notably corals and other demersal species dependent on photosynthesis.
+If species' depth limits are not known, simple rules of thumb will likely reduce commission errors without introducing substantial omission errors.
+For example, for most corals, researchers could clip range maps to the photosynthetic limit of 200 meters.
+For AquaMaps range data, dependent primarily on environmental and physical preferences and conditions, implementing area restrictions based on biogeographical criteria such as Marine Ecoregions of the World (Spalding et al.
+2007) would likely decrease commission errors and improve predictive power, especially for data-poor species.
 
-The AquaMaps methodology is a powerful means of quickly estimating species distribution based on limited information; however, the model's output is only as good as the input data, which for marine species is often sorely limited.
-In addition to environmental preferences and conditions, AquaMaps mapping parameters include area restrictions to help limit over-extrapolation of AquaMaps models.
-Currently, these area restrictions are denoted by UN Food and Agriculture Organisation (FAO) fisheries reporting areas (Ready et al.
-2010); however, area restrictions based on biogeographical criteria, e.g.
-Marine Ecoregions of the World (Spalding et al.
-2007), would likely decrease commission errors and improve predictive power, especially for data-poor species.
-
-When using AquaMaps data alone, changing the probability of occurrence threshold to determine presence becomes a tradeoff between errors of commission (low threshold) and omission (high threshold).
-Using AquaMaps in conjunction with IUCN can increase the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized.
+Many studies using AquaMaps data test the sensitivity of results by varying the probability of occurrence threshold to determine presence.
+This decision ultimately represents a tradeoff between errors of commission (low threshold) and omission (high threshold).
+Using AquaMaps in conjunction with IUCN can mitigate potential errors, while also increasing the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized.
 In this case, we recommend a presence threshold of 0% as it most closely approximates the "limits of distribution" criterion defined by IUCN data providers.
 
-<!-- * Additional filters on AquaMaps distribution criteria, such as expert review status or the number of occurrence cells used to generate the species model, can help avoid overextrapolation for data poor species.
-_using other attributes and variables to filter?
-this seems outside the scope of this paper_ -->
+<!-- summarize back to big picture: CBD, MPAs, climate velocity, etc... understanding species distributions, and therefore minimizing errors, is important to all these things -->
+_AquaMaps and IUCN are two of the most important databases providing species distribution information across a wide range of marine species, taxa, and regions.
+The different methodologies behind the creation of each database is likely to introduce different rates of commission and omission errors, which can have a profound impact on our understanding of species ranges.
+By comparing overlapping species maps from these two databases, we have identified likely sources of error, allowing data users to address or account for them.
+For users of AquaMaps, IUCN, and other datasets of species distribution, reducing the likelihood of these errors is critical to effective biodiversity research and conservation action._
 
-_(do we need a wrap-it-up paragraph here?  I put that more at the top of the conclusions section... could shift it down here, but probably not needed in both places)_
+Human impacts on oceans (pollution, ocean acidification, overfishing, extractive use) are increasingly pressuring the marine ecosystems that provide food, economic opportunity, cultural value, and life support.
+Climate change and direct human impacts are driving shifts in species richness and abundance.
+Managing and protecting the oceans to deliver these benefits sustainably requires much deeper understanding of the flora and fauna.
+Managing and protecting marine species improves the resilience of critical ocean ecosystems to human-induced pressures.
+But effective management depends critically on accurate and robust understanding of the distribution of species we want to protect.

### 2016-03-21 ms_3_methods.Rmd
Ben's comments, some fixes
incorporated Ben's comments; reframing a conclusion; preparing figures.  Moved the MPA analysis explanation of shifts into the SI

@@
 We represented relative distribution of species count for each dataset by plotting average species count against latitude and longitude (Fig.
-A, 2B).
+2).
 
@@
 We ran the SPP code three times, prioritizing IUCN over AquaMaps for a 0% threshold, and prioritizing AquaMaps over IUCN for both a 40% and 0% threshold.
-We compared each of these to the output of the published OHI 2015 SPP model (prioritizing IUCN over AquaMaps at a 40% threshold).
+We compared each of these to the output of the published OHI 2015 SPP model (which prioritizes IUCN over AquaMaps at a 40% threshold).
 
@@
 To analyze MPA coverage against IUCN spatial data, we extracted IUCN polygon weights per 0.5° cell for each species and compared against the protected area raster.
-Finally, we combined AquaMaps data (at 0% threshold) and IUCN data, using AquaMaps for the `r n_spp_both` overlapping species and again compared against the protected area raster.
+Finally, we combined AquaMaps data (at 0% threshold) and IUCN data, using AquaMaps for the `r format(n_spp_both, big.mark = ',')` overlapping species and again compared against the protected area raster.
 
-All processing was completed in R (REF), and all code and intermediate data are available on GitHub (REF?) at https://github.com/OHI-Science/IUCN-AquaMaps.
+All processing was completed using R statistical software (REF), and all code and intermediate data are available on GitHub (REF?) at https://github.com/OHI-Science/IUCN-AquaMaps.

### 2016-03-21 ms_6_supplemental.Rmd
incorporated Ben's comments; reframing a conclusion; preparing figures.  Moved the MPA analysis explanation of shifts into the SI

@@
 This pattern may not hold true for all subsets of AquaMaps species, however, whether subsetting by taxa or by georegion.
+
+### MPA gap analysis - explanation
+
+___why so many gaps and 10+ species show up with IUCN.  I will write up a paragraph once I've had a little more time to examine maps___
+
+- check which species are gap species; check maps
+- check which species are 10%+ species and check maps
+- where do these species fall georegionally? is there anything odd about their ranges? any reason why IUCN would be 
+- hypothesis to explain more gap species: Aquamaps more diffuse than IUCN, catching individual cells that could be suitable; a single cell (even low probability) moves a species from "gap" to non-gap status.
+IUCN more clustered by polygon boundaries; so no scattered cells to accidentally fall into an MPA.
+    - introducing errors: AM = likely commission (overestimates species with tiny bits of range in MPAs) vs IUCN = possible omission (misses possible real habitat that falls within MPAs)
+- hypothesis to explain more 10%+ species - ten random samples most have IUCN polygons just east of Australia where a large MPA is indicated - Great Barrier Reef I assume - a single large poly falling in GBR gets a high score, while for the same species, many of the AM ranges (also in same area) scatter suitable area over larger extent, so lower percent of species range is protected
+    - introducing errors: IUCN = likely omission (skipping possible real habitat outside of MPAs) and likely commission (possibly overcounting real habitat in MPAs e.g.
+GBR); AquaMaps = probably a little commission (possibly underestimating, if it overpredicts lots of range outside MPAs, thus reducing the proportion of protected range)

### 2016-03-21 data_explore/data_explore.Rmd
cleaning up data explore

--- 2016-03-10 data_explore/data_explore.Rmd
@@
 
-``` {r child = 'data_explore_pt1.Rmd', eval = FALSE}
+# Figures 1, 2
+
+``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
 ```
-<!-- # bar chart: number of spp per group
-# scatter plot: area compare IUCN vs AM, including log plot and plot by species group
-# rangemap: Chromis alpha -->
-
-``` {r, child = 'data_explore_pt2.Rmd', eval = FALSE}
-```
-<!-- percent overlap vs total area exploration -->
-
-***
-
-Quadrant map and quad breakdown by taxa and risk
-
-``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-
-### define 25%, 50%, and 75% quartiles for both x and y axes
-area_align_mean <- mean(spp_list$area_ratio, na.rm = TRUE)
-dist_align_mean <- mean(spp_list$sm_perc, na.rm = TRUE)
-
-### define representative species from each quadrant, for over-plotting
-# spp_quads <- c('q1' = 'Kajikia albida', 
-#                'q2' = 'Galaxea cryptoramosa', 
-#                'q4' = 'Acanthurus nigroris', 
-#                'q4a' = 'Metanephrops australiensis', 
-#                'q3' = 'Conus magnificus')
-# 
-### truncate the species names for the representative species
-# spp_list_label <- spp_list %>%
-#   filter(sciname %in% spp_rando) %>%
-#   separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
-#   mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
-#   dplyr::select(-g, -s)
-
-### mongo plot time
-scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
-                                         transform(popn_category = reorder(popn_category, category_score)),
-                                       aes(x = area_ratio, 
-                                           y = sm_perc, 
-                                           color = popn_category,
-                                           size = lg_area_pct)) + 
-  ggtheme_plot + 
-  ### color the quadrant backgrounds:
-    annotate("rect", xmin = area_align_mean, xmax = 100, 
-                     ymin = dist_align_mean, ymax = 100, 
-             alpha = .3, 
-#             fill= "green4")  + 
-             fill= "#4dac26")  + 
-    annotate("rect", xmax = area_align_mean, xmin =   0, 
-                     ymin = dist_align_mean, ymax = 100, 
-             alpha = .3, 
-#             fill= "greenyellow") + 
-             fill= "#b8e186") + 
-    annotate("rect", xmin = area_align_mean, xmax = 100, 
-                     ymax = dist_align_mean, ymin =   0, 
-             alpha = .3, 
-#             fill= "orange") + 
-             fill= "#f1b6da") + 
-    annotate("rect", xmax = area_align_mean, xmin =   0, 
-                     ymax = dist_align_mean, ymin =   0, 
-             alpha = .3, 
-#             fill= "orangered2") + 
-             fill= "#d01c8b") + 
-  geom_point(alpha = .6)
-
-### Manage scales for color and size 
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-  scale_colour_manual(values = c('LC' = 'green4', 
-                                 'NT' = 'greenyellow', 
-                                 'VU' = 'yellow', 
-                                 'EN' = 'orange', 
-                                 'CR' = 'orangered2',
-                                 'DD' = 'grey30')) + 
-  scale_size_continuous(limits = c(0, 100),
-                        breaks = c( 0,    25,    50,    75,    100), 
-                        labels = c('0%', '25%', '50%', '75%', '100%')) + 
-  scale_x_continuous(expand = c(0, 0), 
-                     limits = c(0, 102),
-                     breaks = c(seq(0, 100, 25)),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) +
-  scale_y_continuous(expand = c(0, 0),
-                     limits = c(0, 102),
-                     breaks = c(seq(0, 100, 25)),
-                     labels = c('0%', '25%', '50%', '75%', '100%'))
-
-### here are mean labels:
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-  annotate(geom = 'text',
-           x = area_align_mean, y = 5,
-           hjust = 0, vjust = 0,
-           color = 'grey30', 
-           size = 2,
-           fontface = 'bold.italic', angle = 90,
-           label = sprintf('Mean = %s%%', round(area_align_mean, 1))) +
-  annotate(geom = 'text',
-           x = 5, y = dist_align_mean,
-           hjust = 0, vjust = 0,
-           color = 'grey30', 
-           size = 2,
-           fontface = 'bold.italic', angle = 0,
-           label = sprintf('Mean = %s%%', round(dist_align_mean, 1)))
-
-### here are mean lines:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_vline(x = area_align_mean,
-#              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = dist_align_mean,
-#              color = 'grey50', linetype = 'dashed')
-
-### here are individual species, identified and namean on the plot:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
-#   geom_text(data = spp_list_label, 
-#             aes(label = sciname_label), 
-#             color = 'grey30',
-#             fontface = 'bold.italic',
-#             size = 2,
-#             hjust = .25,
-#             vjust = 1.5)
-
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-  labs(x = bquote('Area alignment '*alpha['ext']*' (%)'), 
-       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
-#       title = 'Distribution alignment vs Extent alignment',
-       color = 'Risk', size = '% ocean range')
-
-print(scatter_arearatio_vs_overlap)
-
-ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 17.8, height = 9, units = 'cm')
-
-```
-
-``` {r examine_spp_gp_quartiles, echo = FALSE, eval = TRUE}
-
-spp_list1 <- spp_list %>%
-  group_by(spp_group_text) %>%
-  mutate(n_spp = n()) %>%
-  ungroup()
-
-### Create species lists by quadrant
-spp_list_q1 <- spp_list1 %>%
-  filter(area_ratio >= mean(area_ratio) & sm_perc >= mean(sm_perc))
-spp_list_q2 <- spp_list1 %>%
-  filter(area_ratio < mean(area_ratio) & sm_perc >= mean(sm_perc))
-spp_list_q3 <- spp_list1 %>%
-  filter(area_ratio >= mean(area_ratio) & sm_perc < mean(sm_perc))
-spp_list_q4 <- spp_list1 %>%
-  filter(area_ratio < mean(area_ratio) & sm_perc < mean(sm_perc))
-
-spp_list_quads <- spp_list1 %>%
-  mutate(quad = NA,
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q1$sciname, 'q1', quad),
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q2$sciname, 'q2', quad),
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q3$sciname, 'q3', quad),
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q4$sciname, 'q4', quad))
-write_csv(spp_list_quads, file.path(dir_data, 'spp_list_quads.csv'))
-
-### Summarize quadrant species list by species group (e.g. CORALS1)
-spp_gp_q1 <- spp_list_q1 %>%
-  group_by(spp_group_text) %>%
-  summarize(n_spp = first(n_spp), n_spp_q1 = n())
-spp_gp_q2 <- spp_list_q2 %>%
-  group_by(spp_group_text) %>%
-  summarize(n_spp = first(n_spp), n_spp_q2 = n())
-spp_gp_q3 <- spp_list_q3 %>%
-  group_by(spp_group_text) %>%
-  summarize(n_spp = first(n_spp), n_spp_q3 = n())
-spp_gp_q4 <- spp_list_q4 %>%
-  group_by(spp_group_text) %>%
-  summarize(n_spp = first(n_spp), n_spp_q4 = n())
-
-### Join species group by quadrant into a single dataframe
-spp_gp_quadrants <- spp_gp_q1 %>%
-  full_join(spp_gp_q2, by = c('spp_group_text', 'n_spp')) %>%
-  full_join(spp_gp_q3, by = c('spp_group_text', 'n_spp')) %>%
-  full_join(spp_gp_q4, by = c('spp_group_text', 'n_spp')) %>%
-  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
-  mutate(quad = str_replace(quad, 'n_spp_', ''),
-         pct_quad = n_quad/n_spp,
-         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
-         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))
-
-spp_gp_quadrants <- spp_gp_quadrants %>%
-  left_join(spp_gp_quadrants %>% 
-              filter(quad == 'q1') %>%
-              dplyr::select(spp_group_text, pct_q1 = pct_quad),
-            by = 'spp_group_text') %>%
-  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
-         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
-         quad = str_replace(quad, 'q3', 'q3 area-matched'),
-         quad = str_replace(quad, 'q4', 'q4 poorly matched'))
-
-### Plot the bar chart
-barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
-                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
-                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
-  theme(panel.grid.major.y = element_blank()) +
-  geom_bar(stat = 'count', alpha = .8) +
-  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
-                               'q2 dist-matched'   = '#b8e186', 
-                               'q3 area-matched'   = '#f1b6da', 
-                               'q4 poorly matched' = '#d01c8b')) +
-#   scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
-#                                'q2 dist-matched'   = 'greenyellow', 
-#                                'q3 area-matched'   = 'orange', 
-#                                'q4 poorly matched' = 'orangered2')) +
-  scale_y_continuous(expand = c(0, 0), 
-                     limits = c(0, 1.1),
-                     breaks = c(0, .25, .5, .75, 1.0),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
-            size = 2, 
-            color = 'grey30') +
-  coord_flip() +
-  labs(x = 'Taxonomic Group', 
-       y = 'Percent of species by quadrant', 
-#       title = 'Taxonomic group by quadrant',
-       fill = 'Alignment')
-
-print(barchart_spp_gp_quads)
-ggsave(plot = barchart_spp_gp_quads,
-       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
-       width = 17.8, height = 7, units = 'cm')
-
-write_csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads.csv'))
-
-```
-
-``` {r count_occurrences_by_quad, echo = FALSE, warning = FALSE}
-
-gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence.csv'))
-obis_occur <- read_csv(file.path(dir_data, 'obis_occurrence.csv'))
-
-totes_occur <- obis_occur %>%
-  left_join(gbif_occur) %>%
-  rowwise() %>%
-  mutate(mean_occ = mean(c(gbif_occ, obis_occ), na.rm = TRUE)) %>%
-  group_by(iucn_sid) %>%
-  summarize(mean_occ = mean(mean_occ, na.rm = TRUE),
-            tots_occ = sum(mean_occ, na.rm = TRUE))
-
-spp_quads_occ <- spp_list_quads %>%
-  left_join(totes_occur %>%
-              dplyr::select(iucn_sid, mean_occ),
-            by = 'iucn_sid')
-
-write_csv(spp_quads_occ, file.path(dir_data, 'spp_quads_occ.csv'))
-
-```
-
-``` {r examine_risk_quartiles, echo = FALSE, eval = TRUE}
-
-### Summarize quadrant species list by species group (e.g. CORALS1)
-spp_risk_q1 <- spp_list_q1 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp_q1 = n())
-spp_risk_q2 <- spp_list_q2 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp_q2 = n())
-spp_risk_q3 <- spp_list_q3 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp_q3 = n())
-spp_risk_q4 <- spp_list_q4 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp_q4 = n())
-
-### Join species group by quadrant into a single dataframe
-spp_risk_quadrants <- spp_risk_q1 %>%
-  full_join(spp_risk_q2, by = c('popn_category')) %>%
-  full_join(spp_risk_q3, by = c('popn_category')) %>%
-  full_join(spp_risk_q4, by = c('popn_category')) %>%
-  gather(quad, n_spp_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
-  group_by(popn_category) %>%
-  mutate(n_tot_cat = sum(n_spp_quad)) %>%
-  ungroup() %>%
-  mutate(quad = str_replace(quad, 'n_spp_', ''),
-         pct_quad   = n_spp_quad/n_tot_cat,
-         pct_quad   = ifelse(is.na(pct_quad),   0, pct_quad))
-
-spp_risk_quadrants1 <- spp_risk_quadrants %>%
-  left_join(spp_risk_quadrants %>% 
-              filter(quad == 'q1') %>%
-              dplyr::select(popn_category, pct_q1 = pct_quad),
-            by = 'popn_category') %>%
-  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
-         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
-         quad = str_replace(quad, 'q3', 'q3 area-matched'),
-         quad = str_replace(quad, 'q4', 'q4 poorly matched')) %>%
-  left_join(spp_list %>%
-              dplyr::select(popn_category, cat_rank = category_score) %>%
-              unique() %>%
-              mutate(cat_rank = ifelse(popn_category %in% c('DD', 'NE'), -1, cat_rank)),
-            by = 'popn_category')
-
-### Plot the bar chart
-barchart_spp_risk_quads <- ggplot(spp_risk_quadrants1 %>% 
-                                  transform(popn_category = reorder(popn_category, cat_rank)), 
-                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
-  theme(panel.grid.major.y = element_blank()) +
-  geom_bar(stat = 'count', alpha = .8) +
-  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
-                               'q2 dist-matched'   = '#b8e186', 
-                               'q3 area-matched'   = '#f1b6da', 
-                               'q4 poorly matched' = '#d01c8b')) +
-  scale_y_continuous(expand = c(0, 0), 
-                     limits = c(0, 1.1),
-                     breaks = c(0, .25, .5, .75, 1.0),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) +
-  geom_text(aes(label = sprintf('n = %s', n_tot_cat), y = 1.01), hjust = 0, 
-            size = 2, 
-            color = 'grey30') +
-  coord_flip() +
-  labs(x = 'Risk category', 
-       y = 'Percent of species by quadrant', 
-#       title = 'Risk category by quadrant',
-       fill = 'Alignment')
-
-print(barchart_spp_risk_quads)
-ggsave(plot = barchart_spp_risk_quads,
-       filename = file.path(dir_fig, 'barchart_spp_risk_quads.png'),
-       width = 17.8, height = 7, units = 'cm')
-
-```
-
-<!-- # Multiple species maps per quadrant -->
-``` {r child = 'data_explore_pt3.Rmd', eval = FALSE}
-```
-
-***
-
-Breakdown maps by taxa
 
@@
 # print(bar_taxa3)
-# ggsave(filename = file.path(dir_fig, 'bar_by_taxa_assessed.png'), 
+# ggsave(filename = file.path(dir_fig, 'si_bar_by_taxa_assessed.png'), 
 #        width = 17.8, height = 6, units = 'cm')
@@
 
-***
+# Figure 3
 
-# Years for IUCN maps?
+## Quadrant map
 
-``` {r get_iucn_citation_info, echo = FALSE, warning = FALSE, eval = FALSE}
+``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
-  ### Queries shapefiles and extracts column headers.  Use to help identify which
-  ### species groups have no id_no field, and possibly other fields to help 
-  ### differentiate.
+### define 25%, 50%, and 75% quartiles for both x and y axes
+area_align_mean <- mean(spp_list$area_ratio, na.rm = TRUE)
+dist_align_mean <- mean(spp_list$sm_perc, na.rm = TRUE)
 
-dir_iucn_shp <- file.path(dir_N, 'git-annex/globalprep/_raw_data/iucn_spp/iucn_shp')
-groups_list <- as.data.frame(list.files(dir_iucn_shp)) %>%
-  rename(shp_fn = `list.files(dir_iucn_shp)`) %>%
-  filter(tools::file_ext(shp_fn) == 'shp') %>%
-  mutate(shp_fn = str_replace(shp_fn, '.shp', ''))
+### define representative species from each quadrant, for over-plotting
+# spp_quads <- c('q1' = 'Kajikia albida', 
+#                'q2' = 'Galaxea cryptoramosa', 
+#                'q4' = 'Acanthurus nigroris', 
+#                'q4a' = 'Metanephrops australiensis', 
+#                'q3' = 'Conus magnificus')
+# 
+### truncate the species names for the representative species
+# spp_list_label <- spp_list %>%
+#   filter(sciname %in% spp_rando) %>%
+#   separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
+#   mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
+#   dplyr::select(-g, -s)
 
-iucn_shp_info <- data.frame()
+### mongo plot time
+scatter_quadplot <- ggplot(spp_list %>% 
+                             transform(popn_category = reorder(popn_category, category_score)),
+                           aes(x = area_ratio, 
+                               y = sm_perc, 
+                               color = popn_category,
+                               size = lg_area_pct)) + 
+  ggtheme_plot + 
+  ### color the quadrant backgrounds:
+    annotate("rect", xmin = area_align_mean, xmax = 100, 
+                     ymin = dist_align_mean, ymax = 100, 
+             alpha = .3, 
+#             fill= "green4")  + 
+             fill= "#4dac26")  + 
+    annotate("rect", xmax = area_align_mean, xmin =   0, 
+                     ymin = dist_align_mean, ymax = 100, 
+             alpha = .3, 
+#             fill= "greenyellow") + 
+             fill= "#b8e186") + 
+    annotate("rect", xmin = area_align_mean, xmax = 100, 
+                     ymax = dist_align_mean, ymin =   0, 
+             alpha = .3, 
+#             fill= "orange") + 
+             fill= "#f1b6da") + 
+    annotate("rect", xmax = area_align_mean, xmin =   0, 
+                     ymax = dist_align_mean, ymin =   0, 
+             alpha = .3, 
+#             fill= "orangered2") + 
+             fill= "#d01c8b") + 
+  geom_point(alpha = .6)
 
-for (spp_gp in groups_list$shp_fn) { # spp_gp <- groups_list$shp_fn[3]
-  spp_dbf <- foreign::read.dbf(file.path(dir_iucn_shp, sprintf('%s.dbf', spp_gp)))
-  message('file read successfully... ')
-  cites <- spp_dbf %>%
-    select(id_no, compiler, year, citation) %>%
-    unique() %>%
-    mutate(spp_group = spp_gp)
+### Manage scales for color and size 
+scatter_quadplot <- scatter_quadplot +
+  scale_colour_manual(values = c('LC' = 'green4', 
+                                 'NT' = 'greenyellow', 
+                                 'VU' = 'yellow', 
+                                 'EN' = 'orange', 
+                                 'CR' = 'orangered2',
+                                 'DD' = 'grey30')) + 
+  scale_size_continuous(limits = c(0, 100),
+                        breaks = c( 0,    25,    50,    75,    100), 
+                        labels = c('0%', '25%', '50%', '75%', '100%')) + 
+  scale_x_continuous(expand = c(0, 0), 
+                     limits = c(0, 102),
+                     breaks = c(seq(0, 100, 25)),
+                     labels = c('0%', '25%', '50%', '75%', '100%')) +
+  scale_y_continuous(expand = c(0, 0),
+                     limits = c(0, 102),
+                     breaks = c(seq(0, 100, 25)),
+                     labels = c('0%', '25%', '50%', '75%', '100%'))
 
-  iucn_shp_info <- bind_rows(iucn_shp_info, cites)
-  message('binding to list...\n')
-}
+### here are mean labels:
+scatter_quadplot <- scatter_quadplot +
+  annotate(geom = 'text',
+           x = area_align_mean, y = 5,
+           hjust = 0, vjust = 0,
+           color = 'grey30', 
+           size = 2,
+           fontface = 'bold.italic', angle = 90,
+           label = sprintf('Mean = %s%%', round(area_align_mean, 1))) +
+  annotate(geom = 'text',
+           x = 5, y = dist_align_mean,
+           hjust = 0, vjust = 0,
+           color = 'grey30', 
+           size = 2,
+           fontface = 'bold.italic', angle = 0,
+           label = sprintf('Mean = %s%%', round(dist_align_mean, 1)))
 
-### oldest records go back to late 70s for terrestrial mammals; then skip to 2004 for a marine turtle map.
+### here are mean lines:
+# scatter_quadplot <- scatter_quadplot +
+#   geom_vline(x = area_align_mean,
+#              color = 'grey50', linetype = 'dashed') +
+#   geom_hline(y = dist_align_mean,
+#              color = 'grey50', linetype = 'dashed')
+
+### here are individual species, identified and namean on the plot:
+# scatter_quadplot <- scatter_quadplot +
+#   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
+#   geom_text(data = spp_list_label, 
+#             aes(label = sciname_label), 
+#             color = 'grey30',
+#             fontface = 'bold.italic',
+#             size = 2,
+#             hjust = .25,
+#             vjust = 1.5)
+
+scatter_quadplot <- scatter_quadplot +
+  labs(x = bquote('Area alignment '*alpha['ext']*' (%)'), 
+       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
+#       title = 'Distribution alignment vs Extent alignment',
+       color = 'Risk', size = '% ocean range')
+
+# print(scatter_quadplot)
+
+ggsave(file.path(dir_fig, 'scatter_quadmap.png'), width = 17.8, height = 9, units = 'cm')
 
@@
 
-***
+![](../figures/scatter_quadmap.png)
 
-``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = FALSE}
+## Break down quads by species group
+
+``` {r examine_spp_gp_quartiles, echo = FALSE, eval = TRUE}
+
+spp_list1 <- spp_list %>%
+  group_by(spp_group_text) %>%
+  mutate(n_spp = n()) %>%
+  ungroup()
+
+### Create species lists by quadrant
+spp_list_q1 <- spp_list1 %>%
+  filter(area_ratio >= mean(area_ratio) & sm_perc >= mean(sm_perc))
+spp_list_q2 <- spp_list1 %>%
+  filter(area_ratio < mean(area_ratio) & sm_perc >= mean(sm_perc))
+spp_list_q3 <- spp_list1 %>%
+  filter(area_ratio >= mean(area_ratio) & sm_perc < mean(sm_perc))
+spp_list_q4 <- spp_list1 %>%
+  filter(area_ratio < mean(area_ratio) & sm_perc < mean(sm_perc))
+
+spp_list_quads <- spp_list1 %>%
+  mutate(quad = NA,
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_q1$sciname, 'q1', quad),
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_q2$sciname, 'q2', quad),
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_q3$sciname, 'q3', quad),
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_q4$sciname, 'q4', quad))
+write_csv(spp_list_quads, file.path(dir_data, 'spp_list_quads.csv'))
+
+### Summarize quadrant species list by species group (e.g. CORALS1)
+spp_gp_q1 <- spp_list_q1 %>%
+  group_by(spp_group_text) %>%
+  summarize(n_spp = first(n_spp), n_spp_q1 = n())
+spp_gp_q2 <- spp_list_q2 %>%
+  group_by(spp_group_text) %>%
+  summarize(n_spp = first(n_spp), n_spp_q2 = n())
+spp_gp_q3 <- spp_list_q3 %>%
+  group_by(spp_group_text) %>%
+  summarize(n_spp = first(n_spp), n_spp_q3 = n())
+spp_gp_q4 <- spp_list_q4 %>%
+  group_by(spp_group_text) %>%
+  summarize(n_spp = first(n_spp), n_spp_q4 = n())
+
+### Join species group by quadrant into a single dataframe
+spp_gp_quadrants <- spp_gp_q1 %>%
+  full_join(spp_gp_q2, by = c('spp_group_text', 'n_spp')) %>%
+  full_join(spp_gp_q3, by = c('spp_group_text', 'n_spp')) %>%
+  full_join(spp_gp_q4, by = c('spp_group_text', 'n_spp')) %>%
+  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
+  mutate(quad = str_replace(quad, 'n_spp_', ''),
+         pct_quad = n_quad/n_spp,
+         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
+         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))
+
+spp_gp_quadrants <- spp_gp_quadrants %>%
+  left_join(spp_gp_quadrants %>% 
+              filter(quad == 'q1') %>%
+              dplyr::select(spp_group_text, pct_q1 = pct_quad),
+            by = 'spp_group_text') %>%
+  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
+         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
+         quad = str_replace(quad, 'q3', 'q3 area-matched'),
+         quad = str_replace(quad, 'q4', 'q4 poorly matched'))
+
+### Plot the bar chart
+barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
+                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
+                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
+  ggtheme_plot + 
+  theme(panel.grid.major.y = element_blank()) +
+  geom_bar(stat = 'count', alpha = .8) +
+  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
+                               'q2 dist-matched'   = '#b8e186', 
+                               'q3 area-matched'   = '#f1b6da', 
+                               'q4 poorly matched' = '#d01c8b')) +
+#   scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
+#                                'q2 dist-matched'   = 'greenyellow', 
+#                                'q3 area-matched'   = 'orange', 
+#                                'q4 poorly matched' = 'orangered2')) +
+  scale_y_continuous(expand = c(0, 0), 
+                     limits = c(0, 1.1),
+                     breaks = c(0, .25, .5, .75, 1.0),
+                     labels = c('0%', '25%', '50%', '75%', '100%')) +
+  geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
+            size = 2, 
+            color = 'grey30') +
+  coord_flip() +
+  labs(x = 'Taxonomic Group', 
+       y = 'Percent of species by quadrant', 
+#       title = 'Taxonomic group by quadrant',
+       fill = 'Alignment')
+
+# print(barchart_spp_gp_quads)
+ggsave(plot = barchart_spp_gp_quads,
+       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
+       width = 17.8, height = 7, units = 'cm')
+
+write_csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads.csv'))
+
 ```
 
-All AquaMaps species richness map:
+![](../figures/barchart_spp_gp_quads.png)
 
-![figures/all_aquamaps_map.png](../figures/all_aquamaps_map.png)
+# Plots for SI
 
-All IUCN species richness map:
+## Break down quadrants by risk category
 
-![figures/all_iucn_map.png](../figures/all_iucn_map.png)
+``` {r examine_risk_quartiles, echo = FALSE, eval = TRUE}
 
-``` {r, child = 'data_explore_am_thresh.Rmd', echo = FALSE, message = FALSE, eval = FALSE}
+### Summarize quadrant species list by species group (e.g. CORALS1)
+spp_risk_q1 <- spp_list_q1 %>%
+  group_by(popn_category) %>%
+  summarize(n_spp_q1 = n())
+spp_risk_q2 <- spp_list_q2 %>%
+  group_by(popn_category) %>%
+  summarize(n_spp_q2 = n())
+spp_risk_q3 <- spp_list_q3 %>%
+  group_by(popn_category) %>%
+  summarize(n_spp_q3 = n())
+spp_risk_q4 <- spp_list_q4 %>%
+  group_by(popn_category) %>%
+  summarize(n_spp_q4 = n())
+
+### Join species group by quadrant into a single dataframe
+spp_risk_quadrants <- spp_risk_q1 %>%
+  full_join(spp_risk_q2, by = c('popn_category')) %>%
+  full_join(spp_risk_q3, by = c('popn_category')) %>%
+  full_join(spp_risk_q4, by = c('popn_category')) %>%
+  gather(quad, n_spp_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
+  group_by(popn_category) %>%
+  mutate(n_tot_cat = sum(n_spp_quad)) %>%
+  ungroup() %>%
+  mutate(quad = str_replace(quad, 'n_spp_', ''),
+         pct_quad   = n_spp_quad/n_tot_cat,
+         pct_quad   = ifelse(is.na(pct_quad),   0, pct_quad))
+
+spp_risk_quadrants1 <- spp_risk_quadrants %>%
+  left_join(spp_risk_quadrants %>% 
+              filter(quad == 'q1') %>%
+              dplyr::select(popn_category, pct_q1 = pct_quad),
+            by = 'popn_category') %>%
+  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
+         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
+         quad = str_replace(quad, 'q3', 'q3 area-matched'),
+         quad = str_replace(quad, 'q4', 'q4 poorly matched')) %>%
+  left_join(spp_list %>%
+              dplyr::select(popn_category, cat_rank = category_score) %>%
+              unique() %>%
+              mutate(cat_rank = ifelse(popn_category %in% c('DD', 'NE'), -1, cat_rank)),
+            by = 'popn_category')
+
+### Plot the bar chart
+barchart_spp_risk_quads <- ggplot(spp_risk_quadrants1 %>% 
+                                  transform(popn_category = reorder(popn_category, cat_rank)), 
+                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
+  ggtheme_plot + 
+  theme(panel.grid.major.y = element_blank()) +
+  geom_bar(stat = 'count', alpha = .8) +
+  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
+                               'q2 dist-matched'   = '#b8e186', 
+                               'q3 area-matched'   = '#f1b6da', 
+                               'q4 poorly matched' = '#d01c8b')) +
+  scale_y_continuous(expand = c(0, 0), 
+                     limits = c(0, 1.1),
+                     breaks = c(0, .25, .5, .75, 1.0),
+                     labels = c('0%', '25%', '50%', '75%', '100%')) +
+  geom_text(aes(label = sprintf('n = %s', n_tot_cat), y = 1.01), hjust = 0, 
+            size = 2, 
+            color = 'grey30') +
+  coord_flip() +
+  labs(x = 'Risk category', 
+       y = 'Percent of species by quadrant', 
+#       title = 'Risk category by quadrant',
+       fill = 'Alignment')
+
+# print(barchart_spp_risk_quads)
+ggsave(plot = barchart_spp_risk_quads,
+       filename = file.path(dir_fig, 'si_barchart_spp_risk_quads.png'),
+       width = 17.8, height = 5, units = 'cm')
+
 ```
+![](../figures/si_barchart_spp_risk_quads.png)
+
+-----
+
+## AquaMaps threshold exploration
+
+``` {r, child = 'data_explore_am_thresh.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
+```
+
+-----
+
+## Count species occurrences as proxy for data poverty
+
+``` {r count_occurrences_by_quad, echo = FALSE, warning = FALSE}
+
+gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence.csv'))
+obis_occur <- read_csv(file.path(dir_data, 'obis_occurrence.csv'))
+
+totes_occur <- obis_occur %>%
+  left_join(gbif_occur) %>%
+  rowwise() %>%
+  mutate(mean_occ = mean(c(gbif_occ, obis_occ), na.rm = TRUE)) %>%
+  group_by(iucn_sid) %>%
+  summarize(mean_occ = mean(mean_occ, na.rm = TRUE),
+            tots_occ = sum(mean_occ, na.rm = TRUE))
+
+spp_quads_occ <- spp_list_quads %>%
+  left_join(totes_occur %>%
+              dplyr::select(iucn_sid, mean_occ),
+            by = 'iucn_sid')
+
+write_csv(spp_quads_occ, file.path(dir_data, 'spp_quads_occ.csv'))
+
+```
+
+-----
+
+# Obsolete plots; not evaluated
+
+* bar chart: number of spp per group
+* scatter plot: area compare IUCN vs AM, including log plot and plot by species group
+* rangemap: Chromis alpha
+
+``` {r child = 'data_explore_pt1.Rmd', eval = FALSE}
+```
+
+-----
+
+* percent overlap vs total area exploration
+
+``` {r, child = 'data_explore_pt2.Rmd', eval = FALSE}
+```
+
+-----
+
+* Multiple species maps per quadrant
+
+``` {r child = 'data_explore_pt3.Rmd', eval = FALSE}
+```
+
+-----

### 2016-03-21 data_explore/data_explore_am_thresh.Rmd
cleaning up data explore

@@
 
-```{r aquamaps_load_data, warning = FALSE, message = FALSE, echo = FALSE, include = FALSE}
+```{r aquamaps_load_data, warning = FALSE, message = FALSE, echo = FALSE, include = FALSE, eval = TRUE}
 
@@
 
-```{r hist_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
+```{r hist_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE, eval = TRUE}
 
@@
   geom_vline(xintercept = spp_df3_mean, color = 'darkred', linetype = 'dashed') + 
-  annotate('text', x = spp_df3_mean, y = .16, hjust = 0, vjust = -.5,
+  annotate('text', x = spp_df3_mean, y = .13, hjust = 0, vjust = -.5,
            angle = 90, 
-           #size = 2, 
+           size = 3, 
            color = 'grey20',
@@
   geom_vline(xintercept = spp_df3_med, color = 'grey30', linetype = 'dashed') + 
-  annotate('text', x = spp_df3_med, y = .16, hjust = 0, vjust = -.5,
+  annotate('text', x = spp_df3_med, y = .13, hjust = 0, vjust = -.5,
            angle = 90, 
-           #size = 2, 
+           size = 3, 
            color = 'grey20',
@@
   scale_y_continuous(expand = c(0, 0)) +
-  labs(x = 'Proportion of extent included when selecting a threshold of 0.4', 
-       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3)))
+  labs(x = 'Proportion of range included when selecting a threshold of 0.4', 
+       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3)),
+       fill = 'IUCN code')
 
-print(hist_area_vs_threshold)
+# print(hist_area_vs_threshold)
 
-ggsave(file.path(dir_fig, 'AMThreshold.png'), 
+ggsave(file.path(dir_fig, 'am_threshold.png'), 
        width = 17.8, height = 9, units = 'cm')
@@
 
-```{r hist_lost_area_w_threshold_ne_dd_removed, warning = FALSE, message = FALSE, echo = FALSE}
+![](../figures/am_threshold.png)
+
+```{r hist_lost_area_w_threshold_ne_dd_removed, warning = FALSE, message = FALSE, echo = FALSE, eval = TRUE}
 
@@
            angle = 90, 
-           #size = 2, 
+           size = 3, 
            color = 'grey20',
@@
   scale_y_continuous(expand = c(0, 0)) +
-  labs(x = 'Proportion of extent included when selecting a threshold of 0.4', 
-       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3a)))
+  labs(x = 'Proportion of range included when selecting a threshold of 0.4', 
+       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3a)),
+       fill = 'IUCN code')
 
-ggsave(file.path(dir_fig, 'AMThreshold2.png'), 
+ggsave(file.path(dir_fig, 'am_threshold2.png'), 
        width = 17.8, height = 9, units = 'cm')
-
-print(hist_area_vs_threshold2)
 
@@
 
-```{r scatter_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
+![](../figures/am_threshold2.png)
 
+```{r df_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE, eval = TRUE}
 ### create dataframe that gives me all species, their total number of cells, 
@@
             median_prop = median(proportion, na.rm = TRUE),
-            quant09 = quantile(proportion, .05),
+            quant05 = quantile(proportion, .05),
             quant25 = quantile(proportion, .25),
             quant75 = quantile(proportion, .75),
-            quant91 = quantile(proportion, .95))
-
-scatter_area_vs_threshold <- ggplot(thresh_df_summary, aes(x = thresh, y = mean_prop)) + 
-  ggtheme_plot + 
-  geom_point(aes(y = median_prop), color = 'grey60') +
-  geom_point() + 
-  geom_abline(slope = -1, intercept = 1, color = 'red') + 
-  geom_vline(xintercept = 0.40, color = 'blue', linetype = 'dashed') +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  labs(# title = 'Mean (median) species range after presence threshold cutoff',
-       x = 'Presence threshold cutoff', 
-       y = 'Mean (median) relative species range')
-
-ggsave(file.path(dir_fig, 'scatter_AMThreshold.png'), 
-       width = 17.8, height = 9, units = 'cm')
-
-print(scatter_area_vs_threshold)
+            quant95 = quantile(proportion, .95))
 
@@
 
-```{r ribbon_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}
-
-area_mean_model <- lm(mean_prop ~ thresh, data = thresh_df_summary)
-area_median_model  <- lm(median_prop ~ thresh,  data = thresh_df_summary)
+```{r ribbon_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE, eval = TRUE}
 
@@
 
-  geom_ribbon(aes(ymax = quant91, ymin = quant09), alpha = .5, color = NA, fill = 'grey80') +
+  geom_ribbon(aes(ymax = quant95, ymin = quant05), alpha = .5, color = NA, fill = 'grey80') +
   geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = 'grey70') +
-  geom_point(color = 'grey30') + 
-  geom_point(aes(y = mean_prop), color = 'darkred') + 
-  geom_abline(intercept = area_mean_model[['coefficients']][1],
-              slope     = area_mean_model[['coefficients']][2],
-              color = 'red') + 
+  geom_line(color = 'grey30') + 
+  # geom_line(aes(y = mean_prop), color = 'darkgreen') + 
+  # geom_abline(intercept = area_mean_model[['coefficients']][1],
+  #             slope     = area_mean_model[['coefficients']][2],
+  #             color = 'red') + 
   geom_vline(xintercept = 0.40, color = 'blue', linetype = 'dashed') +
@@
            angle = 90, 
-           #size = 2, 
+           size = 3, 
            color = 'grey20',
-           label = 'Ocean Health Index: 40%') + 
+           label = 'Halpern et al. 2012') + 
   annotate('text', x = .50, y = .05, hjust = 0, vjust = -.5,
            angle = 90, 
-           #size = 2, 
+           size = 3, 
            color = 'grey20',
-           label = 'MPA Gap Analysis: 50%') + 
+           label = 'Klein et al. 2015') + 
   annotate('text', x = .60, y = .05, hjust = 0, vjust = -.5,
            angle = 90, 
-           #size = 2, 
+           size = 3, 
            color = 'grey20',
-           label = 'Marine mammals: 60%') + 
-  annotate('text', x = .75, y = .85,
-           #size = 2, 
-           color = 'grey20',
-           label = 'R^2^ = .995') + 
-  scale_x_continuous(expand = c(0, 0)) +
+           label = 'Kaschner et al. 2006') + 
+  # annotate('text', x = .75, y = .85,
+  #          #size = 3, 
+  #          color = 'grey20',
+  #          label = 'R^2^ = .995') + 
+  scale_x_continuous(expand = c(0, 0), 
+                     breaks = c(  .2,    .4,    .5,    .6,    .8,   1.0),
+                     labels = c('20%', '40%', '50%', '60%', '80%', '100%')
+                     limits = c(0, 1.02)) +
   scale_y_continuous(expand = c(0, 0)) +
   labs(# title = 'Relative species range after presence threshold',
-       x = 'Presence threshold cutoff', 
-       y = 'Mean (median) relative species range')
+       x = 'AquaMaps presence threshold cutoff', 
+       y = 'Relative species range')
 
@@
 
-ggsave(file.path(dir_fig, 'ribbon_AMThreshold.png'), 
+ggsave(file.path(dir_fig, 'ribbon_am_threshold.png'), 
        width = 17.8, height = 6, units = 'cm')
@@
 ```
+
+![](../figures/ribbon_am_threshold.png)

### 2016-03-21 data_explore/data_explore_global_maps.Rmd
cleaning up data explore

@@
 
-```{r setup, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+```{r setup_from_data_explore, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
@@
 
-```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+### All AquaMaps species count by cell
 
-# Data files
-#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
-                           progress = TRUE,
-                           col_types = 'cddddc')
+```{r all_am_map, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), 
-                           progress = TRUE)
+library(rasterVis)
+myTheme <- rasterTheme(region = brewer.pal('Blues', n = 9))
 
-### LOICZID cell ID numbers - used in plotting rasters
+all_am_rast_file   <- file.path(dir_data, 'rasters/am_all_rast.tif')
+dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')
+
+if(!file.exists(all_am_rast_file)) {
   loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
@@
   names(loiczid_raster) <- 'loiczid'
-```
 
-```{r global_maps, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
-
-library(rasterVis)
-
-myTheme <- rasterTheme(region = brewer.pal('Blues', n = 9))
-
-message('Full AM dataset, species count\n')
-
-all_am_rast_file   <- file.path(dir_data, 'rasters/am_all_rast.tif')
-
-dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')
-
-if(!file.exists(all_am_rast_file)) {
   am_spp_cells_big <- fread(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) %>%
@@
 
-png(file.path(dir_fig, 'all_aquamaps_map1.png'),
+png(file.path(dir_fig, 'all_aquamaps_map.png'),
     width = 8.7, height = 6, units = 'in', res = 300)
@@
 
-Now look at full IUCN species
+![](../figures/all_aquamaps_map.png)
 
-```{r all_iucn, eval = FALSE, echo = FALSE}
+### All IUCN species count by cell
 
-message('Full IUCN dataset, species count\n')
+```{r all_iucn, eval = TRUE, echo = FALSE}
 
@@
 if(!file.exists(all_iucn_rast_file)) {
+  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
+  loiczid_raster       <- raster(loiczid_raster_file)
+  names(loiczid_raster) <- 'loiczid'
+
   iucn_cells_file    <- file.path(dir_data_iucn, 'iucn_cells_2015.csv')
@@
   all_iucn_map   <- raster(all_iucn_rast_file)
-  values(all_iucn_map)[is.na(values(all_am_map))] <- NA
 }
@@
 
-png(file.path(dir_fig, 'all_iucn_map1.png'),
+png(file.path(dir_fig, 'all_iucn_map.png'),
     width = 8.7, height = 6, units = 'in', res = 300)
@@
 
-``` {r variations_on_gl_dist_maps, echo = FALSE, eval = FALSE}
-### rescale values relative to maximum cell value
-
-am_map_max <- maxValue(all_am_map)
-all_am_pct_map <- all_am_map/am_map_max
-
-iucn_map_max <- maxValue(all_iucn_map)
-all_iucn_pct_map <- all_iucn_map/iucn_map_max
-
-diff_pct_map    <- all_am_pct_map - all_iucn_pct_map
-
-### plot maps with distributions as percentages.  Should it be 0 - 1, or 
-### 0 - max value?
-levelplot(all_am_pct_map, par.settings = myTheme, 
-          main = 'Aquamaps Dataset percent of max')
-levelplot(all_iucn_pct_map, par.settings = myTheme, 
-          main = 'IUCN Dataset percent of max')
-### this is the difference one. Should it be min value to max value?
-levelplot(diff_pct_map, par.settings = myTheme, 
-          main = 'Difference: AquaMaps - IUCN by percent')
-
-diff_ln_map <- all_am_ln_map - all_iucn_ln_map
-levelplot(diff_ln_map, 
-          par.settings = myTheme, 
-          pretty = TRUE,
-          margin = list(axis = TRUE),
-#          main = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', 
-          colorkey = list(at = seq(-1, 1, length = 15)))
-maxValue(all_am_ln_map)
-minValue(all_am_ln_map)
-maxValue(all_iucn_ln_map)
-minValue(all_iucn_ln_map)
-maxValue(diff_ln_map)
-minValue(diff_ln_map)
-diff_ln_pts <- as.data.frame(rasterToPoints(diff_ln_map))
-
-spp_richness_diff_plot <- ggplot(diff_ln_pts, aes(x = x, y = y)) +
-  ggtheme_map + 
-  geom_raster(aes(fill = layer)) +
-  scale_fill_gradient2(low = 'red', mid = 'white', high = 'blue', midpoint = 0) +
-  borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
-  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
-  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
-  labs(title = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', x = NULL, y = NULL)
-
-spp_richness_diff_plot
-
-```
+![](../figures/all_iucn_map.png)

### 2016-03-21 ms_1_abst_sig.Rmd
cleaning up data explore

@@
 The Significance Statement appears in the paper itself and is required for all research papers._
+
+- ocean ecosystems provide broad benefits to humans, but are under pressure from human activity and climate change; ensuring sustainability of benefits depends on effective management and protection of resources.
+- Effective management and protection of our ocean ecosystems depends on understanding how species are distributed globally.  
+- Two key datasets of marine species ranges communicate an understanding of the global distribution of marine species - and are commonly used to inform management and protection research and policy.
+- substantial differences highlight important gaps in understanding (or communicating) species ranges, leading to ineffective management and protection.
+- identifying these differences points to solutions to address them and improve understanding.

### 2016-03-21 ms_2_intro_res_disc.Rmd
cleaning up data explore

@@
 
-Human impacts on oceans (pollution, ocean acidification, overfishing, extractive use) are increasingly pressuring the marine ecosystems that provide food, economic opportunity, cultural value, and life support.
-Climate change and direct human impacts are driving shifts in species richness and abundance.
-Managing and protecting the oceans to deliver these benefits sustainably requires much deeper understanding of the flora and fauna.
-Managing and protecting marine species improves the resilience of critical ocean ecosystems to human-induced pressures.
-But effective management depends critically on accurate and robust understanding of the distribution of species we want to protect.
+Human impacts on oceans (pollution, ocean acidification, overfishing, extractive use, climate change) are increasingly pressuring the marine species and ecosystems we depend on for food, economic opportunity, cultural value, and life support.
+Managing and protecting marine species improves the resilience of critical ocean ecosystems oceans to deliver these benefits sustainably, but effective management depends critically on accurate and robust understanding of the distribution of species we want to protect.
+Highlighting the differences between the two currently most important datasets of marine species distributions can help us
+
+- ocean ecosystems provide broad benefits to humans, but are under pressure from human activity and climate change; ensuring sustainability of benefits depends on effective management and protection of resources.
+- Effective management and protection of our ocean ecosystems depends on understanding how species are distributed globally.  
+- Two key datasets of marine species ranges communicate an understanding of the global distribution of marine species - and are commonly used to inform management and protection research and policy.
+- substantial differences highlight important gaps in understanding (or communicating) species ranges, leading to ineffective management and protection.
+- identifying these differences points to solutions to address them and improve understanding.

### 2016-03-22 data_explore/data_explore.Rmd
All figures done? slap 'em together in Photoshop now.

@@
 
-knitr::kable(taxa_df %>%
-               dplyr::select(class, 
-                             AquaMaps = n_am_class_all, 
-                             IUCN = n_iucn_class_all, 
-                             Both = n_class_both))
+# knitr::kable(taxa_df %>%
+#                dplyr::select(class, 
+#                              AquaMaps = n_am_class_all, 
+#                              IUCN = n_iucn_class_all, 
+#                              Both = n_class_both))
 ```
@@
        fill = 'Dataset')
-print(bar_taxa2)
+
+# print(bar_taxa2)
 ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
@@
 ```
+
+![](../figures/bar_by_taxa_all.png)
 
@@
   scale_y_continuous(expand = c(0, 0),
-                     limits = c(0, 102),
+                     limits = c(0, 103),
                      breaks = c(seq(0, 100, 25)),
@@
        filename = file.path(dir_fig, 'si_barchart_spp_risk_quads.png'),
-       width = 17.8, height = 5, units = 'cm')
+       width = 17.8, height = 4.5, units = 'cm')
 
 ```
+
 ![](../figures/si_barchart_spp_risk_quads.png)

### 2016-03-22 data_explore/data_explore_am_thresh.Rmd
All figures done? slap 'em together in Photoshop now.

@@
                                'NE' = 'grey70')) +
-  scale_x_continuous(expand = c(0, 0)) +
+  scale_x_continuous(expand = c(0, 0), limits = c(0, 1.02)) +
   scale_y_continuous(expand = c(0, 0)) +
-  labs(x = 'Proportion of range included when selecting a threshold of 0.4', 
+  labs(x = 'Proportion of range included when selecting a threshold of 40%', 
        y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3)),
@@
                                'NE' = 'grey70')) +
-  scale_x_continuous(expand = c(0, 0)) +
+  scale_x_continuous(expand = c(0, 0), limits = c(0, 1.02)) +
   scale_y_continuous(expand = c(0, 0)) +
-  labs(x = 'Proportion of range included when selecting a threshold of 0.4', 
+  labs(x = 'Proportion of range included when selecting a threshold of 40%', 
        y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3a)),
@@
   geom_vline(xintercept = 0.60, color = 'blue', linetype = 'dashed') +
-  annotate('text', x = .40, y = .05, hjust = 0, vjust = -.5,
+  annotate('text', x = .39, y = .03, hjust = 0, vjust = 0,
            angle = 90, 
@@
            label = 'Halpern et al. 2012') + 
-  annotate('text', x = .50, y = .05, hjust = 0, vjust = -.5,
+  annotate('text', x = .49, y = .03, hjust = 0, vjust = 0,
            angle = 90, 
@@
            label = 'Klein et al. 2015') + 
-  annotate('text', x = .60, y = .05, hjust = 0, vjust = -.5,
+  annotate('text', x = .59, y = .03, hjust = 0, vjust = 0,
            angle = 90, 
@@
                      breaks = c(  .2,    .4,    .5,    .6,    .8,   1.0),
-                     labels = c('20%', '40%', '50%', '60%', '80%', '100%')
+                     labels = c('20%', '40%', '50%', '60%', '80%', '100%'),
                      limits = c(0, 1.02)) +
-  scale_y_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0),
+                     limits = c(0, 1.01)) +
   labs(# title = 'Relative species range after presence threshold',
@@
 ggsave(file.path(dir_fig, 'ribbon_am_threshold.png'), 
-       width = 17.8, height = 6, units = 'cm')
+       width = 17.8, height = 7, units = 'cm')
 

### 2016-03-23 data_explore/data_explore.Rmd
updating figs

@@
 #  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
-  scale_fill_manual(values = c('AquaMaps' = '#1b9e77', 'Both' = '#d95f02', 'IUCN' = '#7570b3')) +
+  scale_fill_manual(values = c('AquaMaps' = '#1b9e77', 'Both' = '#d95f02', 'IUCN' = '#6f6ba0')) +
   geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
@@
 # print(bar_taxa3)
-# ggsave(filename = file.path(dir_fig, 'si_bar_by_taxa_assessed.png'), 
+# ggsave(filename = file.path(dir_fig, 'si_bar_by_taxa_assessed.tif'), device = 'tiff', 
 #        width = 17.8, height = 6, units = 'cm')
@@
 
-ggsave(file.path(dir_fig, 'scatter_quadmap.png'), width = 17.8, height = 9, units = 'cm')
+ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
+       width = 17.8, height = 9, units = 'cm')
 
@@
 ggsave(plot = barchart_spp_risk_quads,
+       filename = file.path(dir_fig, 'si_barchart_spp_risk_quads.tif'), device = 'tiff',
+       width = 17.8, height = 4.5, units = 'cm')
+ggsave(plot = barchart_spp_risk_quads,
        filename = file.path(dir_fig, 'si_barchart_spp_risk_quads.png'),

### 2016-03-23 data_explore/data_explore_am_thresh.Rmd
updating figs

@@
 
-ggsave(file.path(dir_fig, 'am_threshold.png'), 
+ggsave(file.path(dir_fig, 'si_am_threshold.tif'), device = 'tiff', 
        width = 17.8, height = 9, units = 'cm')
+ggsave(file.path(dir_fig, 'si_am_threshold.png'), 
+       width = 17.8, height = 9, units = 'cm')
+
 ```
 
-![](../figures/am_threshold.png)
+![](../figures/si_am_threshold.png)
 
@@
 
-ggsave(file.path(dir_fig, 'am_threshold2.png'), 
+ggsave(file.path(dir_fig, 'si_am_threshold2.tif'), device = 'tiff', 
+       width = 17.8, height = 9, units = 'cm')
+ggsave(file.path(dir_fig, 'si_am_threshold2.png'), 
        width = 17.8, height = 9, units = 'cm')
@@
 
-![](../figures/am_threshold2.png)
+![](../figures/si_am_threshold2.png)
 
@@
 
-  geom_ribbon(aes(ymax = quant95, ymin = quant05), alpha = .5, color = NA, fill = 'grey80') +
-  geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = 'grey70') +
+  geom_ribbon(aes(ymax = quant95, ymin = quant05), alpha = .5, color = NA, fill = '#b8e186') +
+  geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = '#4dac26') +
   geom_line(color = 'grey30') + 
@@
 
-print(ribbon_area_vs_threshold)
-
-ggsave(file.path(dir_fig, 'ribbon_am_threshold.png'), 
+ggsave(file.path(dir_fig, 'si_ribbon_am_threshold.tif'), device = 'tiff', 
+       width = 17.8, height = 7, units = 'cm')
+ggsave(file.path(dir_fig, 'si_ribbon_am_threshold.png'), 
        width = 17.8, height = 7, units = 'cm')
@@
 
-![](../figures/ribbon_am_threshold.png)
+![](../figures/si_ribbon_am_threshold.png)

### 2016-03-23 manuscript-main.Rmd
incorporated Ben's comments; reframing a conclusion; preparing figures.  Moved the MPA analysis explanation of shifts into the SI
updating figs

--- 2016-03-10 manuscript-main.Rmd
@@
 title: "Aligning marine species range data to better serve science and conservation"
-author: "Casey O'Hara^1^, Jamie Afflerbach^1^, Benjamin S. Halpern<sup>1, 2, 3</sup>"
+author: Casey O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin S. Halpern<sup>1, 2, 3</sup>
+
 output:
+  word_document: default
+  pdf_document:
+    number_sections: yes
+    toc: no
   html_document:
@@
 
-_Classification:_ BIOLOGICAL SCIENCES: Ecology? Environmental Sciences? Sustainability Science?
+_Classification:_ BIOLOGICAL SCIENCES: Sustainability Science
 

### 2016-03-23 mpa_analysis/3_mpa_spp_analysis.Rmd
created mapping chunk for iucn gap and 10+ species cells
updating figs

@@
 write_csv(mpa_all_runs, file.path(dir_data, 'mpa_all_runs.csv'))
+
+# mpa_runs <- read_csv(file.path(dir_data, 'mpa_all_runs.csv'))
 ```
@@
                                '5-10%' = '#b8e186', 
-                               '2-5%'  = '#f7f7f7', 
+                               '2-5%'  = '#fbe5c5', 
                                '0-2%'  = '#f1b6da',
@@
        filename = file.path(dir_fig, 'barchart_mpa_gap.png'),
+       width = 17.8, height = 5, units = 'cm')
+ggsave(plot = barchart_mpa_gap,
+       filename = file.path(dir_fig, 'barchart_mpa_gap.tif'), device = 'tiff',
        width = 17.8, height = 5, units = 'cm')

### 2016-03-23 ms_1_abst_sig.Rmd
updating figs

--- 2016-03-21 ms_1_abst_sig.Rmd
@@
 
-Species distribution data provide the foundation for a wide range of ecological research studies and conservation management decisions, yet most species ranges remain unknown, and existing range maps often suffer from data limitations and show inconsistencies.
+Species distribution data provide the foundation for a wide range of ecological research studies and conservation management decisions, yet most species ranges remain unknown, and existing range maps often suffer from data limitations and inconsistencies.
 AquaMaps and the International Union for Conservation of Nature (IUCN) are two distinct efforts to map marine species distributions at a global scale.
@@
 We illustrate the scientific and management implications of these differences by repeating two recent applications - an assessment of global biodiversity within the Ocean Health Index and a global analysis of gaps in coverage of marine protected areas - and find significantly different results depending on how the two datasets were used.
-Until a single, highly accurate dataset of global marine species ranges becomes available, understanding the implications of dataset differences for conservation planning and decision-making is essential.
-
-<!--Understanding these issues points toward solutions for providers and users of these datasets.-->
-
-(184 words; 1307 char with spaces)
-
------
+Until a single, highly accurate dataset of global marine species ranges becomes available, understanding the implications of dataset differences for conservation planning and decision-making remains essential.
 
@@
 The Significance Statement appears in the paper itself and is required for all research papers._
+
+Successful conservation of marine species critically depends on detailed and accurate understanding of where species exist and where they do not.
+By systematically comparing two widely used global datasets of marine species range maps, we find substantial differences that highlight inconsistencies in our understanding and communication of species ranges.
+Knowing where and why these inconsistencies occur will substantially improve our ability to develop effective and efficient conservation strategies that support resilient ocean ecosystems.

### 2016-03-23 ms_2_intro_res_disc.Rmd
updating figs

--- 2016-03-21 ms_2_intro_res_disc.Rmd
@@
 
+<!--
 | Section                   |   Word #   | Char # (w/spc) |
@@
 | total body                | 2848 words | 18857 char |
+-->
 
@@
 Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps (Kaschner et al.
-___2013???___) and International Union for Conservation of Nature (IUCN) (IUCN 2015).
+2013) and International Union for Conservation of Nature (IUCN) (IUCN 2015).
 These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status (Halpern et al.
@@
 The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
-
------
 
@@
 
------
-
 # Implications
@@
 
------
-
 **Case Study: MPA Gap Analysis**
@@
 Conclusions drawn from each of these datasets would paint dramatically different pictures of the protection afforded by our current global MPA network.
-
------
 
@@
 
-<!-- summarize back to big picture: CBD, MPAs, climate velocity, etc... understanding species distributions, and therefore minimizing errors, is important to all these things -->
-_AquaMaps and IUCN are two of the most important databases providing species distribution information across a wide range of marine species, taxa, and regions.
-The different methodologies behind the creation of each database is likely to introduce different rates of commission and omission errors, which can have a profound impact on our understanding of species ranges.
-By comparing overlapping species maps from these two databases, we have identified likely sources of error, allowing data users to address or account for them.
-For users of AquaMaps, IUCN, and other datasets of species distribution, reducing the likelihood of these errors is critical to effective biodiversity research and conservation action._
-
-Human impacts on oceans (pollution, ocean acidification, overfishing, extractive use) are increasingly pressuring the marine ecosystems that provide food, economic opportunity, cultural value, and life support.
-Climate change and direct human impacts are driving shifts in species richness and abundance.
-Managing and protecting the oceans to deliver these benefits sustainably requires much deeper understanding of the flora and fauna.
-Managing and protecting marine species improves the resilience of critical ocean ecosystems to human-induced pressures.
-But effective management depends critically on accurate and robust understanding of the distribution of species we want to protect.
+Effective management and protection of marine species depends on a robust understanding of where species exist and where they do not; without this knowledge we risk wasting resources protecting low-value regions while missing opportunities to protect critical ones.
+By identifying the differences between these two fundamental marine species range datasets and understanding the likely mechanisms causing these discrepancies, we improve our ability to develop strategic and effective conservation policy that supports a resilient ocean ecosystem.

### 2016-03-23 ms_4_figs.Rmd
Ben's comments, some fixes
updating figs

@@
 
-![](figures/bar_by_taxa_all.png)
+![](figures/fig1abc_am_iucn_maps.tif)
+
 ``` {r barchart_by_taxa, echo = FALSE}
@@
 Overlapping species are dominated by bony fishes (`r n_bonyfish` species, primarily tropical taxa) and corals (`r n_corals` species).
+(B, C) Number of global marine species according to (A) AquaMaps and (B) IUCN.
+The margin frequency plots show relative species count per cell at each latitude and longitude.
 
@@
 <a id = 'fig2'></a>
-
-![](figures/all_aquamaps_map.png)
-
-![](figures/all_iucn_map.png)
-
->  __Fig. 2.__(A, B) Number of global marine species according to (A) AquaMaps and (B) IUCN.  The margin frequency plots show relative species count per cell at each latitude and longitude.
-
-### Fig. 3
-<a id = 'fig3'></a>
 
@@
 > __Fig.
-3.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r n_spp_both` species included in both IUCN and AquaMaps datasets.
+2.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r n_spp_both` species included in both IUCN and AquaMaps datasets.
 The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
@@
 
-### Figure 4
+### Figure 3
 
@@
 
-> __Fig. 4.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
+> __Fig. 3.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
 > Scenario 1 shows the effect of reducing the probability threshold to 0% for AquaMaps presence to more accurately track the definition of IUCN “limits of distribution."  
@@
 
-### Figure 5
+### Figure 4
 
-<a id = 'fig5'></a>
+<a id = 'fig4'></a>
 
@@
 > __Fig.
-5.__  Percent of species range covered by MPAs based upon methods in Klein et al.
+4.__  Percent of species range covered by MPAs based upon methods in Klein et al.
 (2015).

### 2016-03-24 case_study_mpa/3_mpa_spp_analysis.Rmd
updated figures; trying to redo violin plots - need to verify results

--- 2016-03-23 mpa_analysis/3_mpa_spp_analysis.Rmd
@@
 
-if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)
+# if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)
 
@@
   ### note: AquaMaps calculated areas differ from my calculated areas - see below
-  cell_df <- read_csv(file.path(dir_git, 'mpa_analysis/celldata/loiczid_marine.csv'),
+  cell_df <- read_csv(file.path(dir_git, 'case_study_mpa/celldata/loiczid_marine.csv'),
     col_types = 'dd') %>%
@@
 
-cell_df_klein <- combine_mpa_areas('mpa_analysis/celldata/loiczid_mpa_klein_i-iv_%s.csv', yr_list)
-cell_df_all   <- combine_mpa_areas('mpa_analysis/celldata/loiczid_mpa_i-iv_%s.csv',       yr_list)
-cell_df_des   <- combine_mpa_areas('mpa_analysis/celldata/loiczid_mpa_i-iv-des_%s.csv',   yr_list)
+cell_df_klein <- combine_mpa_areas('case_study_mpa/celldata/loiczid_mpa_klein_i-iv_%s.csv', yr_list)
+cell_df_all   <- combine_mpa_areas('case_study_mpa/celldata/loiczid_mpa_i-iv_%s.csv',       yr_list)
+cell_df_des   <- combine_mpa_areas('case_study_mpa/celldata/loiczid_mpa_i-iv-des_%s.csv',   yr_list)
 
@@
 
-``` {r recreate Klein results 2014 AM data, echo = FALSE, eval = FALSE}
+``` {r recreate Klein results 2014 AM data, echo = FALSE, eval = TRUE}
 
@@
 
-``` {r recreate Klein results 2015 AM data, echo = FALSE, eval = FALSE}
+``` {r recreate Klein results 2015 AM data, echo = FALSE, eval = TRUE}
 
@@
 
-write_csv(mpa_klein_am2015, 'mpa_analysis/celldata/mpa_by_spp_am2015.csv')
+write_csv(mpa_klein_am2015, file.path(dir_git, 'case_study_mpa/celldata/mpa_by_spp_am2015.csv'))
 
@@
 
-``` {r 2015 AM data w 0 threshold, echo = FALSE, eval = FALSE}
+``` {r 2015 AM data w 0 threshold, echo = FALSE, eval = TRUE}
 
@@
 
-head(iucn_spp_sum)
-head(mpa_am2015_0)
+# head(iucn_spp_sum)
+# head(mpa_am2015_0)
 
@@
 mpa_all_runs <- mpa_klein_am2014 %>%
-  mutate(trial = 'am2014 (50%)') %>%
+  mutate(trial = 'AM 2014 (50%)') %>%
   rename(cat = cat2014, pct = pct2014) %>%
   bind_rows(mpa_klein_am2015 %>%
-    mutate(trial = 'am2015 (50%)') %>%
+    mutate(trial = 'AM 2015 (50%)') %>%
     rename(cat = cat2015, pct = pct2015)) %>%
   bind_rows(mpa_am2015_0 %>%
-    mutate(trial = 'am2015 (0%)') %>%
+    mutate(trial = 'AM 2015 (0%)') %>%
     rename(cat = cat2015, pct = pct2015)) %>%
   bind_rows(mpa_am2015_iucn %>%
-    mutate(trial = 'am2015 (0%) + iucn2015') %>%
+    mutate(trial = 'AM 2015 (0%) +\nIUCN 2015') %>%
     rename(cat = cat2015, pct = pct2015)) %>%
   bind_rows(iucn_spp_sum %>%
-    mutate(trial = 'iucn2015') %>%
+    mutate(trial = 'IUCN 2015') %>%
     rename(cat = cat2015, pct = pct2015)) %>%
@@
 library(ggplot2)
-source('data_explore/data_explore_fxns.R')
+source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
 
@@
   mutate(cat = factor(cat, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
-  mutate(trial = factor(trial, levels = c('iucn2015', 'am2015 (0%) + iucn2015', 'am2015 (0%)', 'am2015 (50%)', 'am2014 (50%)'))) %>%
+  mutate(trial = factor(trial, levels = c('IUCN 2015', 'AM 2015 (0%) +\nIUCN 2015', 'AM 2015 (0%)', 'AM 2015 (50%)', 'AM 2014 (50%)'))) %>%
   mutate(p_cat = p_cat * 100) %>%

### 2016-03-24 data_explore/data_explore.Rmd
updated figures; trying to redo violin plots - need to verify results

@@
 
-# Figures 1, 2
-
-``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
-```
+# Figures 1 a b c
 
@@
          percent = n_spp/n_total * 100,
-         dataset = factor(dataset, levels = c('AquaMaps', 'Both', 'IUCN')))
+         dataset = factor(dataset, levels = c('Both', 'IUCN', 'AquaMaps')))
 
@@
 #  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
-  scale_fill_manual(values = c('AquaMaps' = '#1b9e77', 'Both' = '#d95f02', 'IUCN' = '#6f6ba0')) +
+ scale_fill_manual(values = c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')) +
+#  scale_fill_manual(values = c('AquaMaps' = '#4dac26', 'Both' = '#fbe5c5', 'IUCN' = '#d01c8b')) +
   geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
@@
 
-# print(bar_taxa2)
+print(bar_taxa2)
 ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
@@
 
-# Figure 3
+``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
+```
+
+# Figure 2a b
 

### 2016-03-24 manuscript-main.Rmd
updated figures; trying to redo violin plots - need to verify results

@@
 title: "Aligning marine species range data to better serve science and conservation"
-author: Casey O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin S. Halpern<sup>1, 2, 3</sup>
-
+author: Casey O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin
+  S. Halpern<sup>1, 2, 3</sup>
 output:
-  word_document: default
-  pdf_document:
-    number_sections: yes
-    toc: no
   html_document:
@@
 
-``` {r SOM, child = 'ms_6_supplemental.Rmd'}
+``` {r SOM, child = 'ms_6_supplemental.Rmd', eval = FALSE}
 ```

### 2016-03-24 ms_4_figs.Rmd
updated figures; trying to redo violin plots - need to verify results

@@
 
-![](figures/fig1abc_am_iucn_maps.tif)
+![](figures/fig1abc_am_iucn_maps.png)
 
@@
 
-![](figures/AreaRatiovsPercOverlap.png)
-![](figures/barchart_spp_gp_quads.png)
+![](figures/fig2ab_quadmap_sppbars.png)
 
@@
 
-![Absolute change in SPP score](figures/boxplot_abs_diff.png)
+![Absolute change in SPP score](figures/fig3_boxplot_abs_diff.png)
 
@@
 
-![Change in MPA coverage](figures/barchart_mpa_gap.png)
+![Change in MPA coverage](figures/fig4_barchart_mpa_gap.png)
 
@@
 Scenario 1 replicates the original results, measuring protected range of species in AquaMaps 2014 dataset, with a 50% presence threshold, against the 2014 World Database of Protected Areas, filtered for IUCN categories I-IV that overlap marine areas.
-Scenario 2 updates the results using AquaMaps 2015, showing very small changes despite the inclusion of an additional 5545 species.
+Scenario 2 updates the results using AquaMaps 2015, showing very small changes despite the inclusion of an additional 5,545 species.
 Scenario 3, using 2015 AquaMaps data, drops the presence threshold to zero, showing an expected decrease in gap species, but also a decrease in species with 5% or greater protected range.

### 2016-03-24 case_study_mpa/model_wdpa2014_i-iv.py

@@ -1,0 +1,100 @@
+# File: ohiprep/globalprep/WDPA_MPA/v2015/model.py
+#
+# Modified March 2015 by Casey O'Hara
+#
+# Filters by 
+# * "STATUS" == 'Designated':  includes only officially designated areas, not
+#   areas still in planning or proposal phases.
+# * "MANG_PLAN" NOT LIKE 'Non-MPA%': eliminates all areas specifically designated
+#   as Non-MPA Programmatic Management Plans (code only used for U.S.).  Eliminates
+#   groundfish management areas, '25 mph manatee' zones, and the like.
+
+import arcpy, os, subprocess, csv, sys, socket
+from arcpy.sa import *
+
+# configuration based on machine name
+dirs =  {'git'    :'C:/Users/ohara/Documents/github/IUCN-AquaMaps', 
+         'neptune':'N:',
+         'tmp'    :'C:/Users/ohara/tmp',
+         }
+print 'Setting pathnames...'
+print dirs
+
+# paths
+prod     = 'mpa_analysis/WDPA_MPA'          # name of product
+dir_git  = '%s/%s' % (dirs['git'], prod)    # local github directory inside IUCN-AquaMaps
+dir_tmp  = '%s/%s' % (dirs['tmp'], prod)    # local temp directory
+dir_anx  = '%s/git-annex/Global/WDPA-MPA_v2014' % (dirs['neptune']) # git annex directory on neptune
+gdb      = '%s/geodb.gdb' % dir_tmp         # file geodatabase
+print 'dir_git = ' + dir_git
+print 'dir_tmp = ' + dir_tmp
+print 'dir_anx = ' + dir_anx
+
+# inputs
+poly_wdpa  = '%s/raw/WDPA_Apr2014_Public1/WDPA_Apr2014.gdb/WDPA_poly_Apr2014' % dir_anx
+dir_rgn    = '%s/git-annex/globalprep/spatial/v2015/data' % dirs['neptune']
+
+msk = '%s/rgn_marine.tif' % dir_git   # .tif of 500 m marine raster? will be huge - put on git-annex
+
+print 'input poly_wdpa = ' + poly_wdpa
+print 'input dir_rgn = ' + dir_rgn
+print 'input rgn_off_3nm = ' + rgn_off_3nm
+print 'input rgn_inl_1km = ' + rgn_inl_1km
+print 'input msk = ' + msk
+
+
+# outputs
[…]

### 2016-03-24 case_study_ohi/boxplots.R

@@ -1,0 +1,93 @@
+### Create box plot for each scenario
+
+# get list of files (for all scenarios) - status (then eventually trend)
+# define base status as IUCN pref, 0.4 aquamaps thresh
+
+# smash together into a data frame of:
+# rgn_id | base status | scenario 1 status | scen 2 st | scen 3 st |
+
+# then box plot those bastards
+
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(ggplot2)
+
+dir_data  <- '~/github/ohiprep/globalprep/SPP_ICO/vAM_IUCN'
+dir_fig <- '~/github/IUCN-AquaMaps/figures'
+
+### generic theme for all plots
+ggtheme_basic <- theme(axis.ticks = element_blank(),
+                       text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
+                       plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
+                       legend.position = 'none')
+
+ggtheme_plot <- ggtheme_basic + 
+  theme(panel.border = element_blank(),
+        panel.grid.minor = element_blank(), 
+        panel.grid.major = element_line(colour = 'grey90'),
+        panel.background = element_blank(),
+        axis.line = element_line(colour = "grey30"))
+
+
+st_base  <- read_csv(file.path(dir_data, 'data/spp_status_global_IUCNpref_prob0.4.csv'))
+st_scen1 <- read_csv(file.path(dir_data, 'data/spp_status_global_IUCNpref_prob0.csv'))
+st_scen2 <- read_csv(file.path(dir_data, 'data/spp_status_global_AMpref_prob0.4.csv'))
+st_scen3 <- read_csv(file.path(dir_data, 'data/spp_status_global_AMpref_prob0.csv'))
+
+### scenario 1: prefer IUCN data; drop AM probability threshold to 0.01
+### scenario 2: prefer AM data; keep AM prob thresh at 0.40
+### scenario 3: prefer AM data; drop AM prob thresh to 0.01
+
+st_all <- st_base %>% rename(st_b = score) %>%
+  left_join(st_scen1 %>% rename(st_1 = score), by = 'rgn_id') %>%
+  left_join(st_scen2 %>% rename(st_2 = score), by = 'rgn_id') %>%
+  left_join(st_scen3 %>% rename(st_3 = score), by = 'rgn_id')
+  
+st_all <- st_all %>% mutate(absdiff1 = (st_1 - st_b)*100,
[…]

### 2016-03-24 ohi_analysis/boxplots.R

@@ -1,93 +1,0 @@
-### Create box plot for each scenario
-
-# get list of files (for all scenarios) - status (then eventually trend)
-# define base status as IUCN pref, 0.4 aquamaps thresh
-
-# smash together into a data frame of:
-# rgn_id | base status | scenario 1 status | scen 2 st | scen 3 st |
-
-# then box plot those bastards
-
-library(dplyr)
-library(tidyr)
-library(stringr)
-library(ggplot2)
-
-dir_data  <- '~/github/ohiprep/globalprep/SPP_ICO/vAM_IUCN'
-dir_fig <- '~/github/IUCN-AquaMaps/figures'
-
-### generic theme for all plots
-ggtheme_basic <- theme(axis.ticks = element_blank(),
-                       text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-                       plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
-                       legend.position = 'none')
-
-ggtheme_plot <- ggtheme_basic + 
-  theme(panel.border = element_blank(),
-        panel.grid.minor = element_blank(), 
-        panel.grid.major = element_line(colour = 'grey90'),
-        panel.background = element_blank(),
-        axis.line = element_line(colour = "grey30"))
-
-
-st_base  <- read.csv(file.path(dir_data, 'data/spp_status_global_IUCNpref_prob0.4.csv'),  stringsAsFactors = FALSE)
-st_scen1 <- read.csv(file.path(dir_data, 'data/spp_status_global_IUCNpref_prob0.01.csv'), stringsAsFactors = FALSE)
-st_scen2 <- read.csv(file.path(dir_data, 'data/spp_status_global_AMpref_prob0.4.csv'),    stringsAsFactors = FALSE)
-st_scen3 <- read.csv(file.path(dir_data, 'data/spp_status_global_AMpref_prob0.01.csv'),   stringsAsFactors = FALSE)
-
-### scenario 1: prefer IUCN data; drop AM probability threshold to 0.01
-### scenario 2: prefer AM data; keep AM prob thresh at 0.40
-### scenario 3: prefer AM data; drop AM prob thresh to 0.01
-
-st_all <- st_base %>% rename(st_b = score) %>%
-  left_join(st_scen1 %>% rename(st_1 = score), by = 'rgn_id') %>%
-  left_join(st_scen2 %>% rename(st_2 = score), by = 'rgn_id') %>%
[…]

### 2016-03-26 case_study_mpa/3_mpa_spp_analysis.Rmd
updating figures, text, SI

@@
                                'gap'   = '#d01c8b')) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 115), breaks = c(seq(0, 100, 10))) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 125), breaks = c(seq(0, 100, 20))) +
   geom_text(aes(label = sprintf('n = %s', n_trial), y = 101), hjust = 0, 
@@
 ggsave(plot = barchart_mpa_gap,
-       filename = file.path(dir_fig, 'barchart_mpa_gap.png'),
-       width = 17.8, height = 5, units = 'cm')
+       filename = file.path(dir_fig, 'fig4_barchart_mpa_gap.png'),
+       width = 11.4, height = 5, units = 'cm')
 ggsave(plot = barchart_mpa_gap,
-       filename = file.path(dir_fig, 'barchart_mpa_gap.tif'), device = 'tiff',
-       width = 17.8, height = 5, units = 'cm')
+       filename = file.path(dir_fig, 'fig4_barchart_mpa_gap.tif'), device = 'tiff',
+       width = 11.4, height = 5, units = 'cm')
 

### 2016-03-26 ms_2_intro_res_disc.Rmd
updating figures, text, SI

@@
 ---
-output: html_document
+output: pdf_document
 ---
@@
 By dividing the map-paired species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
-For representative maps from each category, please refer to the supplemental materials (SI Fig.
-XXX).
+For representative maps from each category, please refer to the supplemental materials (Fig.
+S1).
 
@@
 This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their ranges span nearly the entire map.
-We expected to see significant concurrence between maps for IUCN Red List species of concern, assuming that increased attention would lead to improved understanding of species ranges; however, we found no compelling patterns to indicate that this was the case (SI Fig.
-XXX).
+We expected to see significant concurrence between maps for IUCN Red List species of concern, assuming that increased attention would lead to improved understanding of species ranges; however, we found no compelling patterns to indicate that this was the case (Fig.
+S2).
 
 The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where the IUCN range is notably larger, often extending into deeper water.
-For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their preferred depths, likely introducing errors of commission  (SI Figs.
-XXX, XXX).
+For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their preferred depths, likely introducing errors of commission  (Fig.
+S3).
  Ocean depth is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps.
@@
 
+``` {r calc_ohi_numbers, echo = FALSE, message = FALSE, warning = FALSE}
+spp_ohi <- spp_all %>%
+  filter(!is.na(spatial_source) & !is.na(category_score)) %>%
+  select(am_sid, iucn_sid, spatial_source) %>%
+  unique()
+
+n_spp_am_ohi <- spp_ohi %>% 
+                    filter(str_detect(spatial_source, 'am')) %>%
+                    select(am_sid) %>%
+                    unique() %>%
+                    nrow()
+n_spp_iucn_ohi <- spp_ohi %>%
+                    filter(str_detect(spp_ohi$spatial_source, 'iucn')) %>%
+                    filter(is.na(am_sid)) %>%
+                    unique() %>%
+                    nrow()
+n_spp_both_ohi <- spp_ohi %>% 
+                    filter(!is.na(am_sid)) %>%
+                    filter(str_detect(spatial_source, 'iucn')) %>%
+                    select(iucn_sid) %>%
+                    unique() %>%
+                    nrow()
+
+n_spp_ohi <- n_spp_iucn_ohi + n_spp_am_ohi + n_spp_both_ohi
+
+p_spp_both_ohi <- round(n_spp_both_ohi/n_spp_ohi * 100, 1)
+
+```
+
 The global Ocean Health Index (OHI) (Halpern et al.
 2012, 2015), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
-To maximize the number of represented species, OHI gleans spatial distributions from both IUCN and AquaMaps datasets, prioritizing IUCN data for the `r format(n_spp_both, big.mark = ',')` species included in both sources.
+To maximize the number of represented species, OHI gleans spatial distributions from Red List-assessed species in both IUCN and AquaMaps datasets (n = `r format(n_spp_ohi, big.mark = ',')`), prioritizing IUCN data for the `r format(n_spp_both_ohi, big.mark = ',')` species included in both sources.
 OHI currently uses a probability of occurrence threshold of 40% to determine species presence for AquaMaps data.
 
-We recalculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig4'>Fig.
-4</a>).
+We calculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig.
+3</a>).
 Reducing the threshold increases the apparent range of a species; the slight decrease in average score for scenario 1 suggests increased spatial representation of threatened species.
@@
 
-Given that only `r format(p_both_all, big.mark = ',')`% of species overlap, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in such large country-level score shifts as seen in scenarios 2 and 3.
-While many country scores showed only small variation across all three scenarios, select countries gained up to 5 points while others dropped as many as 12 points.
-Significantly, the global score decreased by two full points in scenario 3, indicating that an arbitrary change in how the two datasets are combined can result in a more dire assessment of species conservation status.
+Given that only `r p_spp_both_ohi`% of species overlap, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in such large country-level score shifts as seen in scenarios 2 and 3.
+While the mean global score did not vary significantly from scenario to scenario, select countries gained up to 7.5 points while others dropped as many as 5 points.
+This result indicates that especially on a national or regional scale, an arbitrary change in how the two datasets are combined can result in a very different assessment of species conservation status.
 
@@
 
-We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using 2015 AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig5'>Fig.
-5</a>).
+We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using 2015 AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig4'>Fig.
+4</a>).
 We found a five-fold increase in the proportion of gap species (6.4% of species vs.
@@
 
-For IUCN range data, clipping ranges to known depth limits improves output for many species, most notably corals and other demersal species dependent on photosynthesis.
+For IUCN range data, clipping ranges to known depth limits improves output for many species, most notably corals and reef-associated fishes.
 If species' depth limits are not known, simple rules of thumb will likely reduce commission errors without introducing substantial omission errors.

### 2016-03-26 ms_4_figs.Rmd
updating figures, text, SI

@@
 # Figures and captions
-
-___still need work on captions; then separate captions from images___
 
@@
 1.__ (A) Number and proportion of species by taxa included in each dataset.
-Overlapping species are dominated by bony fishes (`r n_bonyfish` species, primarily tropical taxa) and corals (`r n_corals` species).
+Overlapping species are dominated by bony fishes (`r format(n_bonyfish, big.mark = ',')` species, primarily tropical taxa) and corals (`r format(n_corals, big.mark = ',')` species).
 (B, C) Number of global marine species according to (A) AquaMaps and (B) IUCN.
@@
 > __Fig.
-2.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r n_spp_both` species included in both IUCN and AquaMaps datasets.
+2.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r format(n_spp_both, big.mark = ',')` species included in both IUCN and AquaMaps datasets.
 The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
@@
 
-___NOTE: This figure still needs to be recreated using the latest and greatest data!___ _currently they use data generated in November or so, in ohiprep._
-
 ### Figure 4

### 2016-03-26 ms_6_supplemental.Rmd
updating figures, text, SI

@@
 
-___Probably just cut, but maybe include in Suppl Materials.
-- BH___: The release of AquaMaps distribution maps is not limited to comprehensively-assessed taxa, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.
-
-#### Possible figures
-- Map of all cells
-- Map of # of species per cell
-- Example aquamaps (one species range)
-
 ### IUCN
-
-IUCN: While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
 
@@
 For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.
+
+Each shapefile contains polygon limits of distribution indicating regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
 
@@
 
-JA: I think here we only need a few sentences describing how we downloaded the data and on what date. Link to the IUCN spatial data website.
-
-IUCN data is provided as polygon shapefiles. These vectors were rasterized to the same cell resolution as AquaMaps data (0.5 degrees).
-
 #### Possible figures
 
+S1: sample maps from each quadrant
+
+S2: risk by quadrant
+
+S3: coral depth map
+
+<!--
 #### Caveats
@@
 However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
+-->
 
@@
 
-![](figures/bar_by_taxa_assessed.png)
+<!--![](figures/bar_by_taxa_assessed.png)-->
 
@@
 
-* Breaking down the quadrants by IUCN extinction risk categories (FIG 3c), we found that species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
+* Breaking down the quadrants by IUCN extinction risk categories (FIG 3c), we found little  with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
 Does higher perceived risk lead to increased attention, and thus better understanding of species distribution?
@@
 
+<!--
 ___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot___ _- perhaps update the quadrant plot to reveal data-poor species (and reviewed species?
@@
 |   q4     |     410   |         159 (39.7%) |          32.1 (13)        |         25  (6.2%) |
+
+-->
 
@@
 
-### AquaMaps presence threshold analysis - move to SOM
+### AquaMaps presence threshold analysis
 
-For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to best approximate the "extent of occurrence" as generally indicated by IUCN maps.
-To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0.05 to 1.00 and calculated the average species range relative to a zero threshold.
+For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to best approximate the "limits of distribution" as indicated by IUCN maps.
+To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0% to 100% and calculated the average species range relative to a zero threshold.
 
@@
 
-![](figures/AMThreshold.png)
+![](figures/si_am_threshold.png)
 
-![](figures/ribbon_AMThreshold.png)
+![](figures/si_ribbon_am_threshold.png)
 
@@
 (b) Mean (median) remaining extent at increments of presence threshold.
-Dark grey ribbon includes 25% to 75% quantiles, while light grey ribbon includes 9% to 91% quantiles.
+Dark grey ribbon includes 25% to 75% quantiles, while light grey ribbon includes 5% to 95% quantiles.
 

### 2016-03-26 case_study_ohi/boxplots.R

@@ -12,6 +12,7 @@ library(dplyr)
 library(tidyr)
 library(stringr)
 library(ggplot2)
+library(readr)
 
 dir_data  <- '~/github/ohiprep/globalprep/SPP_ICO/vAM_IUCN'
 dir_fig <- '~/github/IUCN-AquaMaps/figures'
@@ -51,6 +52,21 @@ st_all <- st_all %>% mutate(absdiff1 = (st_1 - st_b)*100,
                             reldiff2 = absdiff2/st_b,
                             reldiff3 = absdiff3/st_b) %>%
   gather(scenario, difference, absdiff1:reldiff3)
+
+# georgns <- read_csv('~/github/ohi-global/eez2013/layers/rgn_georegions.csv') %>%
+#      spread(level, georgn_id) 
+# 
+# georgn_lbls <- read.csv('~/github/ohi-global/eez2013/layers/rgn_georegion_labels.csv') %>%
+#   mutate(level_label = sprintf('%s_label', level)) %>%
+#   select(-level) %>%
+#   spread(level_label, label) %>%
+#   arrange(r0_label, r1_label, r2_label, rgn_id)
+
+
+# st_all <- st_all %>%
+#   left_join(georgns, by = 'rgn_id') %>%
+#   left_join(georgn_lbls, by = 'rgn_id') %>%
+#   mutate(r1 = factor(r1, levels = unique(r1)))
   
 bplot_abs <- ggplot(data = st_all %>% filter(str_detect(scenario, 'abs')),
                     aes(x = scenario, y = difference)) +
@@ -58,19 +74,38 @@ bplot_abs <- ggplot(data = st_all %>% filter(str_detect(scenario, 'abs')),
 #   theme(text = element_text(family = 'Helvetica', color = 'gray30', size = 12),
 #         plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
 #         legend.position = 'none') +
-  geom_violin(fill = 'grey90', draw_quantiles = c(.25, .5, .75)) +
-  geom_point(position = position_jitter(w = .3), color = '#4dac26', alpha = .5) +
+  geom_hline(yintercept = 0, color = 'gray50', size = .5) +
+  geom_violin(fill = 'grey80', draw_quantiles = c(.25, .5, .75), 
+              alpha = .5) +
+  geom_point(position = position_jitter(w = .3), 
+             aes(color = scenario), 
+             alpha = .5,
+             show.legend = TRUE) +
[…]

### 2016-03-27 case_study_mpa/3_mpa_spp_analysis.Rmd
updated figures

@@
 |Scenario               |     n|gap  |covered 0-2% |covered 2-5% |covered 5-10% |covered >10% |
-|:----------------------|-----:|:----|:------------|:------------|:-------------|:------------|
+|:----------------------|-----:|----:|------------:|------------:|-------------:|------------:|
 |am2014 (50%)           | 17344|1.6% |44.6%        |44.6%        |6.8%          |2.4%         |
@@
 |am2015 (0%)            | 22889|1.2% |46.5%        |44.5%        |6.4%          |1.5%         |
-|am2015 (0%) + iucn2015 | 24634|2%   |48%          |42.2%        |6.2%          |1.6%         |
+|am2015 (0%) & iucn2015 | 24634|2%   |48%          |42.2%        |6.2%          |1.6%         |
 |iucn2015               |  4024|6.4% |66.8%        |20.1%        |3.8%          |2.9%         |

### 2016-03-27 data_explore/data_explore_am_thresh.Rmd
updated figures

@@
 # get spp_cells raw data for aquamaps
-dir_aquamaps <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO/raw/aquamaps_2014')
+dir_aquamaps <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015')
 
-spp_cells <- read_csv(file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), 
-                      col_types = '_ccd__', progress = TRUE)
+spp_cells <- read_csv(file.path(dir_aquamaps, 'csv/hcaf_sp_native_trunc.csv'), 
+                      col_types = 'ccd', progress = TRUE)
 
-setnames(spp_cells, c('SpeciesID', 'CsquareCode', 'probability'), c('species_id', 'csquare_code', 'probability'))
-
-spp_cells <- as.data.frame(spp_cells)
-
-spp <- fread(input = file.path(dir_aquamaps, 'tables/ohi_speciesoccursum.csv'), 
+spp <- fread(input = file.path(dir_aquamaps, 'csv/speciesoccursum.csv'), 
              verbose = FALSE, header = TRUE) %>%
-  mutate(sciname = paste(Genus, Species, sep = ' '),
+  mutate(sciname = paste(genus, species, sep = ' '),
          iucn_code = str_replace(toupper(iucn_code), 'LR/', ''),
@@
 spp_df3 <- spp_cells %>%
-  group_by(species_id) %>%
+  group_by(speciesid) %>%
   summarize(cells_0.4 = sum(probability >= 0.4),
             cells_t = sum(probability >= 0)) %>%
-  left_join(spp, by = c('species_id' = 'SPECIESID')) %>%
+  left_join(spp, by = c('speciesid')) %>%
   mutate(prop = cells_0.4 / cells_t,
@@
   # filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')) %>%
-  select(sciname, species_id, cells_0.4, cells_t, prop, iucn_code)
+  select(sciname, speciesid, cells_0.4, cells_t, prop, iucn_code)
 
@@
                                'CR' = 'red4',
+                               'EX' = 'black',
                                'DD' = 'grey60',
@@
   labs(x = 'Proportion of range included when selecting a threshold of 40%', 
-       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3)),
+       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp)),
        fill = 'IUCN code')
@@
 
-```{r hist_lost_area_w_threshold_ne_dd_removed, warning = FALSE, message = FALSE, echo = FALSE, eval = TRUE}
+```{r hist_lost_area_w_threshold_ne_dd_removed, warning = FALSE, message = FALSE, echo = FALSE, eval = FALSE}
 
@@
 
-![](../figures/si_am_threshold2.png)
+<!--![](../figures/si_am_threshold2.png)-->
 
@@
   temp_df <- spp_cells %>%
-    group_by(species_id) %>%
+    group_by(speciesid) %>%
     summarize(thresh       = i,
@@
               filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
-            by = c('species_id' = 'SPECIESID')) %>%
-    select(sciname, species_id, thresh, cells_thresh, cells_total, proportion, iucn_code)
+            by = c('speciesid')) %>%
+    select(sciname, speciesid, thresh, cells_thresh, cells_total, proportion, iucn_code)
 

### 2016-03-27 data_explore/data_explore_quadmaps.Rmd
updated figures

--- 2015-12-29 data_explore/data_explore_quadmaps.Rmd
@@
 iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
-                           col_types = 'cdddd', progress = TRUE) %>%
+                           col_types = 'cddddc', progress = TRUE) %>%
   as.data.frame() %>%
@@
                'q4' = 'Acanthurus nigroris')
+am_thresh <- 0
 
 for (i in 1:length(spp_quads)) { # i = 1
-  spp <- spp_quads[i]
+  spp <- spp_list %>% filter(sciname == spp_quads[i])
   spp_quad <- names(spp_quads)[i]
 
-  cat(sprintf('%s: %s', spp_quad, spp))
   spp_plot <- plot_rangemap(spp)
+  spp_plot <- spp_plot + 
+    labs(title = element_blank())
+
   print(spp_plot)
+  plotfile <- file.path(dir_fig, 'spp_maps', 
+                        sprintf('%s_%s.png', 
+                                spp_quad, 
+                                tolower(str_replace(spp_quads[i], ' ', '_'))))
+  ggsave(plotfile, height = 4.5, width = 8.7, units = 'cm')
   }
@@
 
-``` {r examine_spp_gp_quartiles, echo = FALSE, eval = TRUE}
+``` {r examine_spp_gp_quartiles, echo = FALSE, eval = FALSE}
 

### 2016-03-27 data_explore/data_explore_fxns.R

@@ -207,14 +207,14 @@ plot_rangemap <- function(spp) {
   spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
     full_join(as.data.frame(rasterToPoints(r_am_spp)),
               by = c('x', 'y')) %>%
-    mutate(presence = ifelse(am_pres & is.na(iucn_pres), 'am',
-                             ifelse(iucn_pres & is.na(am_pres), 'iucn',
-                                    ifelse(iucn_pres & am_pres, 'both', NA))))
+    mutate(presence = ifelse(am_pres & is.na(iucn_pres), 'AquaMaps',
+                             ifelse(iucn_pres & is.na(am_pres), 'IUCN',
+                                    ifelse(iucn_pres & am_pres, 'Both', NA))))
   
   spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
     ggtheme_map + 
-    geom_raster(aes(fill = presence), alpha = .5) +
-    scale_fill_manual(values = c('iucn' = 'blue', 'am' = 'red', 'both' = 'mediumorchid')) +
+    geom_raster(aes(fill = presence), alpha = .8) +
+    scale_fill_manual(values = c('AquaMaps' = '#1b9e77', 'Both' = '#cc50dc', 'IUCN' = '#d95f02')) +
     borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
     scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
     scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +

### 2016-03-27 case_study_mpa/3_mpa_spp_analysis.Rmd
pulling figures to update the supplemental info

--- 2016-03-26 case_study_mpa/3_mpa_spp_analysis.Rmd
@@
     rename(cat = cat2015, pct = pct2015)) %>%
-  select(speciesid, cat, pct, trial)
+  select(speciesid, cat, pct, trial) %>%
+  unique()
 

### 2016-03-27 ms_6_supplemental.Rmd
pulling figures to update the supplemental info

@@
 
-## Data processing
+## Data processing - should this go in methods???
 
@@
 AquaMaps data was provided by AquaMaps directly via ftp.
-Three SQL files were sent that contained (1) list of all `r nrow(spp_am)` species and their taxonomic information, (2) information for all (292,000?) half degree cells used for range mapping and (3) the probability of occurrence per cell for all `r nrow(spp_am)` species.
+Three SQL files were sent that contained (1) list of all `r nrow(spp_am)` species and their taxonomic information, (2) information for all (259,200?) half degree cells used for range mapping and (3) the probability of occurrence per cell for all `r nrow(spp_am)` species.
 These were extracted into .csv form using R Statistical Software.
@@
 
-#### Possible figures
+## S1: sample maps from each quadrant
 
-S1: sample maps from each quadrant
+## S2: risk by quadrant
 
-S2: risk by quadrant
+![](figures/si_barchart_spp_risk_quads.png)
 
-S3: coral depth map
-
-<!--
-#### Caveats
-For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
-However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
--->
-
-___Red List inclusion:___
-
-<!--![](figures/bar_by_taxa_assessed.png)-->
-
-![](figures/barchart_spp_risk_quads.png)
-
-* Breaking down the quadrants by IUCN extinction risk categories (FIG 3c), we found little  with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
+* Breaking down the quadrants by IUCN extinction risk categories (Fig.
+S2), we found little support for our hypothesis that maps for species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
 Does higher perceived risk lead to increased attention, and thus better understanding of species distribution?
@@
 
-<!--
-___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot___ _- perhaps update the quadrant plot to reveal data-poor species (and reviewed species?
-little overlap of reviewed & data-poor) instead of, say, red-list category which doesn't get discussed in the body of the paper?
-then this table can go in SOM if we like it_
-
-| quadrant | n species | n data poor species | mean (median) data points | n reviewed species |
-| :------: | :-------: | :-----------------: | :-----------------------: | :----------------: |
-| all AM   |   22889   |        8749 (38.2%) |          57.1 (16)        |       1296  (5.7%) |
-| AM&IUCN  |    2166   |         457 (21.1%) |          89.9 (33)        |        290 (13.4%) |
-|   q1     |     401   |          33  (8.2%) |         233.0 (78)        |        100 (24.9%) |
-|   q2     |     682   |         151 (22.1%) |          77.4 (39)        |        100 (14.7%) |
-|   q3     |     682   |         114 (16.5%) |          52.4 (29)        |         65  (9.5%) |
-|   q4     |     410   |         159 (39.7%) |          32.1 (13)        |         25  (6.2%) |
-
--->
-
-## illustrative maps for different quadrants and different mechanistic problems
+## S3: coral depth map
 
@@
 plotted against a 200 m bathymetry line: ![oculina_map.png](figures/oculina/oculina_map.png)
+
+## S4: AquaMaps threshold examination
+
+For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to best approximate the "limits of distribution" as indicated by IUCN maps.
+To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0% to 100% and calculated the average species range relative to a zero threshold.
+
+S4a:
+
+![](figures/si_am_threshold.png)
+
+S4b:
+
+![](figures/si_ribbon_am_threshold.png)
+
+>AquaMaps distribution map extent remaining after applying a presence threshold.
+(a) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of XXX, with a wide distribution in which some species lose nearly all of their apparent range.
+(b) Mean (median) remaining extent at increments of presence threshold.
+Dark grey ribbon includes 25% to 75% quantiles, while light grey ribbon includes 5% to 95% quantiles.
+
+AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value as indicated by IUCN maps.
+Many studies convert this AquaMaps probability to a simple presence value by assigning a threshold value (REF references here).
+A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
+
+At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a significant decrease in represented range, and some species lose nearly their entire range.
+Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.
+
+## MPA gap analysis - explanation
+
+___why so many gaps and 10+ species show up with IUCN.  I will write up a paragraph once I've had a little more time to examine maps___
+
+- check which species are gap species; check maps
+- check which species are 10%+ species and check maps
+- where do these species fall georegionally? is there anything odd about their ranges? any reason why IUCN would be 
+- hypothesis to explain more gap species: Aquamaps more diffuse than IUCN, catching individual cells that could be suitable; a single cell (even low probability) moves a species from "gap" to non-gap status.
+IUCN more clustered by polygon boundaries; so no scattered cells to accidentally fall into an MPA.
+    - introducing errors: AM = likely commission (overestimates species with tiny bits of range in MPAs) vs IUCN = possible omission (misses possible real habitat that falls within MPAs)
+- hypothesis to explain more 10%+ species - ten random samples most have IUCN polygons just east of Australia where a large MPA is indicated - Great Barrier Reef I assume - a single large poly falling in GBR gets a high score, while for the same species, many of the AM ranges (also in same area) scatter suitable area over larger extent, so lower percent of species range is protected
+    - introducing errors: IUCN = likely omission (skipping possible real habitat outside of MPAs) and likely commission (possibly overcounting real habitat in MPAs e.g.
+GBR); AquaMaps = probably a little commission (possibly underestimating, if it overpredicts lots of range outside MPAs, thus reducing the proportion of protected range)
+
+<!--
+#### Caveats
+For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
+However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
+-->
+
+<!--
+___Red List inclusion:___
+
+![](figures/bar_by_taxa_assessed.png)-->
 
@@
 
-## AquaMaps: Effect of changing "presence" threshold on apparent distribution
+<!--
+___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot___ _- perhaps update the quadrant plot to reveal data-poor species (and reviewed species?
+little overlap of reviewed & data-poor) instead of, say, red-list category which doesn't get discussed in the body of the paper?
+then this table can go in SOM if we like it_
 
-### AquaMaps presence threshold analysis
+| quadrant | n species | n data poor species | mean (median) data points | n reviewed species |
+| :------: | :-------: | :-----------------: | :-----------------------: | :----------------: |
+| all AM   |   22889   |        8749 (38.2%) |          57.1 (16)        |       1296  (5.7%) |
+| AM&IUCN  |    2166   |         457 (21.1%) |          89.9 (33)        |        290 (13.4%) |
+|   q1     |     401   |          33  (8.2%) |         233.0 (78)        |        100 (24.9%) |
+|   q2     |     682   |         151 (22.1%) |          77.4 (39)        |        100 (14.7%) |
+|   q3     |     682   |         114 (16.5%) |          52.4 (29)        |         65  (9.5%) |
+|   q4     |     410   |         159 (39.7%) |          32.1 (13)        |         25  (6.2%) |
 
-For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to best approximate the "limits of distribution" as indicated by IUCN maps.
-To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0% to 100% and calculated the average species range relative to a zero threshold.
-
-### Figure 5 (a, b):
-
-![](figures/si_am_threshold.png)
-
-![](figures/si_ribbon_am_threshold.png)
-
->AquaMaps distribution map extent remaining after applying a presence threshold.
-(a) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of XXX, with a wide distribution in which some species lose nearly all of their apparent range.
-(b) Mean (median) remaining extent at increments of presence threshold.
-Dark grey ribbon includes 25% to 75% quantiles, while light grey ribbon includes 5% to 95% quantiles.
-
-AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value as indicated by IUCN maps.
-Many studies convert this AquaMaps probability to a simple presence value by assigning a threshold value (REF references here).
-A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
-
-At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a significant decrease in represented range, and some species lose nearly their entire range.
-Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.
-This pattern may not hold true for all subsets of AquaMaps species, however, whether subsetting by taxa or by georegion.
-
-### MPA gap analysis - explanation
-
-___why so many gaps and 10+ species show up with IUCN.  I will write up a paragraph once I've had a little more time to examine maps___
-
-- check which species are gap species; check maps
-- check which species are 10%+ species and check maps
-- where do these species fall georegionally? is there anything odd about their ranges? any reason why IUCN would be 
-- hypothesis to explain more gap species: Aquamaps more diffuse than IUCN, catching individual cells that could be suitable; a single cell (even low probability) moves a species from "gap" to non-gap status.
-IUCN more clustered by polygon boundaries; so no scattered cells to accidentally fall into an MPA.
-    - introducing errors: AM = likely commission (overestimates species with tiny bits of range in MPAs) vs IUCN = possible omission (misses possible real habitat that falls within MPAs)
-- hypothesis to explain more 10%+ species - ten random samples most have IUCN polygons just east of Australia where a large MPA is indicated - Great Barrier Reef I assume - a single large poly falling in GBR gets a high score, while for the same species, many of the AM ranges (also in same area) scatter suitable area over larger extent, so lower percent of species range is protected
-    - introducing errors: IUCN = likely omission (skipping possible real habitat outside of MPAs) and likely commission (possibly overcounting real habitat in MPAs e.g.
-GBR); AquaMaps = probably a little commission (possibly underestimating, if it overpredicts lots of range outside MPAs, thus reducing the proportion of protected range)
+-->

### 2016-03-27 data_explore/data_explore_am_thresh.Rmd
updating boxplot figs and am threshold figs

@@
   geom_vline(xintercept = spp_df3_med, color = 'grey30', linetype = 'dashed') + 
-  annotate('text', x = spp_df3_med, y = .13, hjust = 0, vjust = -.5,
+  annotate('text', x = spp_df3_med + .02, y = .13, hjust = 0, vjust = 1,
            angle = 90, 
@@
   scale_y_continuous(expand = c(0, 0)) +
-  labs(x = 'Proportion of range included when selecting a threshold of 40%', 
-       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp)),
+  labs(x = 'Proportion of range included for presence threshold = 40%', 
+       y = sprintf('Proportion of all \nAquaMaps species (n = %s)', nrow(spp)),
        fill = 'IUCN code')
@@
 ggsave(file.path(dir_fig, 'si_am_threshold.tif'), device = 'tiff', 
-       width = 17.8, height = 9, units = 'cm')
+       width = 11.4, height = 6, units = 'cm')
 ggsave(file.path(dir_fig, 'si_am_threshold.png'), 
-       width = 17.8, height = 9, units = 'cm')
+       width = 11.4, height = 6, units = 'cm')
 
@@
            angle = 90, 
-           size = 3, 
+           size = 2, 
            color = 'grey20',
@@
 ggsave(file.path(dir_fig, 'si_ribbon_am_threshold.tif'), device = 'tiff', 
-       width = 17.8, height = 7, units = 'cm')
+       width = 11.4, height = 6, units = 'cm')
 ggsave(file.path(dir_fig, 'si_ribbon_am_threshold.png'), 
-       width = 17.8, height = 7, units = 'cm')
+       width = 11.4, height = 6, units = 'cm')
 

### 2016-03-27 data_explore/data_explore.Rmd

@@
   filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)
+  filter(!is.na(sm_perc))           ### filter out NA maps
 

### 2016-03-27 data_explore/data_explore_coralmaps.Rmd
trying to get a coral map going

--- 2016-03-27 data_explore/data_explore_quadmaps.Rmd
@@
 ---
-title: 'Data Exploration quad maps: IUCN and AquaMaps'
+title: 'Data Exploration coral maps: IUCN and AquaMaps'
 output: html_document
@@
 
+``` {r define_fxns, echo = FALSE}
+#################################################################=
+### Species Map Function ###
+# This function takes a dataframe of loiczid per species ID, and a target species ID 
+### (as single or vector).  Return a df of cells ready to be rasterized.  
+### If AquaMaps, returned df includes aquamaps probability; if IUCN, returned
+### df includes proportional area.
+get_iucn_spp_map <- function(sid, spp_cells) {
+
+  spp_map <- spp_cells %>%
+    filter(iucn_sid %in% sid)
+  if(nrow(spp_map) == 0) {
+    message('no IUCN cells found')
+    spp_map <- spp_map %>%
+      as.data.frame() %>%
+      select(iucn_sid, loiczid) %>%
+      mutate(area_iucn = NA)
+  } else {
+    spp_map <- spp_map %>%
+      group_by(iucn_sid, loiczid) %>%        
+      summarize(area_iucn = max(prop_area))
+    ### The group_by() and summarize() are to collapse cells with 
+    ### overlapped polygons, fixing the duplicate 'by' issue.
+    ### Note: this drops presence, sciname, and subpop columns
+  }
+
+  return(spp_map %>% unique() %>% as.data.frame())
+}
+
+#################################################################=
+### Map vs Depth function
+### uses ggplot to create a formatted map of species ranges, including
+### Aquamaps, IUCN, and overlapped ranges; returns the plot object.
+plot_depthmap <- function(spp) {
+  map_iucn <- get_spp_map(spp$iucn_sid, iucn_spp_cells) %>%
+    mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0)) %>%
+    select(-area_iucn, -iucn_sid) %>%
+    unique()
+
+  r_iucn_spp <- subs(loiczid_raster, 
+                     map_iucn[ , c('loiczid', 'iucn_pres')], 
+                     by = 'loiczid', 
+                     which = 'iucn_pres', 
+                     subsWithNA = TRUE)
+
+  spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
+    mutate(presence = 'IUCN')
+
+  depth_poly_df <- rgdal::readOGR(dsn = 'coralmaps', layer = 'bath_200m') %>%
+    fortify() %>%
+    mutate(x = long, y = lat)
+
+  spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
+    ggtheme_map + 
+    geom_raster(aes(fill = presence), alpha = .8) +
+    scale_fill_manual(values = c('IUCN' = '#d95f02')) +
+    geom_polygon(data = depth_poly_df, color = 'black', fill = 'yellow', alpha = .8) +
+    borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
+    scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
+    scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2))
+}
+
+```
+
 ```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
+
+spp_corals <- 'Oculina varicosa'
 
@@
 #iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
+iucn_coral_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                            col_types = 'cddddc', progress = TRUE) %>%
   as.data.frame() %>%
-  filter(presence != 5) ### removes extinct subpopulations
-
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), progress = TRUE)
+  filter(presence != 5) %>% ### removes extinct subpopulations
+  filter(sciname %in% spp_corals)
 
@@
 names(loiczid_raster) <- 'loiczid'
-```
 
-```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 ### this file created in data_prep
-spp_list <- read.csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE)
+spp_list_corals <- read.csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE) %>%
+  filter(sciname %in% spp_corals)
 
@@
 
-Quadrant map and quad breakdown by taxa and risk
+``` {r plot_coral_spp_at_depth, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+### define representative species from each quadrant, for over-plotting
 
-``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+  spp <- spp_list_corals
 
-spp_list <- spp_list %>%
-  mutate(area_ratio = am_area / iucn_area,
-         area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
-         area_ratio = area_ratio * 100)
-
-### define 25%, 50%, and 75% quartiles for both x and y axes
-area_ratio_quartiles <- quantile(spp_list$area_ratio, na.rm = TRUE)[2:4]
-pct_overlap_quartiles <- quantile(spp_list$sm_perc, na.rm = TRUE)[2:4]
-
-# spp_rando <- c("Lagocephalus lagocephalus",
-#                "Makaira nigricans",
-#                "Pentacheles laevis",
-#                "Stereomastis suhmi",
-#                "Stereomastis surda") 
-# 
-### truncate the species names for the representative species
-# spp_list_label <- spp_list %>%
-#   filter(sciname %in% spp_rando) %>%
-#   separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
-#   mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
-#   dplyr::select(-g, -s)
-
-### mongo plot time
-scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
-                                         transform(popn_category = reorder(popn_category, category_score)),
-                                       aes(x = area_ratio, y = sm_perc, 
-                                           color = popn_category,
-                                           size = lg_area_pct)) + 
-  ggtheme_plot + 
-  ### color the quadrant backgrounds:
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
-  geom_point(alpha = .6)
-
-### Manage scales for color and size 
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-  scale_colour_manual(values = c('LC' = 'green4', 
-                                 'NT' = 'greenyellow', 
-                                 'VU' = 'yellow', 
-                                 'EN' = 'orange', 
-                                 'CR' = 'orangered2',
-                                 'DD' = 'grey30')) + 
-  scale_size_continuous(limits = c(0, 100),
-                        breaks = c( 0,    25,    50,    75,    100), 
-                        labels = c('0%', '25%', '50%', '75%', '100%')) + 
-  scale_x_continuous(expand = c(0, 0), limits = c(0, 100)) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
-
-### here are median labels:
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-  annotate(geom = 'text',
-           x = area_ratio_quartiles[2], y = 5,
-           hjust = 0, vjust = 0,
-           color = 'grey50', size = 2,
-           fontface = 'bold.italic', angle = 90,
-           label = sprintf('Median = %s%%', round(area_ratio_quartiles[2], 1))) +
-  annotate(geom = 'text',
-           x = 5, y = pct_overlap_quartiles[2],
-           hjust = 0, vjust = 0,
-           color = 'grey50', size = 2,
-           fontface = 'bold.italic', angle = 0,
-           label = sprintf('Median = %s%%', round(pct_overlap_quartiles[2], 1)))
-
-### here are median lines:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_vline(x = area_ratio_quartiles[2],
-#              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = pct_overlap_quartiles[2],
-#              color = 'grey50', linetype = 'dashed')
-
-### here are 25% and 75% quantile lines:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_vline(x = area_ratio_quartiles[c(1, 3)],
-#              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = pct_overlap_quartiles[c(1, 3)],
-#              color = 'grey50', linetype = 'dashed')
-
-### here are individual species, identified and named on the plot:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
-#   geom_text(data = spp_list_label, 
-#             aes(label = sciname_label), 
-#             color = 'grey30',
-#             fontface = 'bold.italic',
-#             size = 2,
-#             hjust = .25,
-#             vjust = 1.5)
-
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-  labs(x = bquote('Extent alignment '*alpha['ext']*' (%)'), 
-       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
-       title = 'Distribution alignment vs Extent alignment',
-       color = 'Risk', size = '% ocean range')
-
-print(scatter_arearatio_vs_overlap)
-
-```
-
-``` {r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-### define representative species from each quadrant, for over-plotting
-spp_quads <- c('q1' = 'Kajikia albida', 
-               'q2' = 'Conus episcopatus',
-               'q3' = 'Conus magnificus', 
-               'q4' = 'Acanthurus nigroris')
-am_thresh <- 0
-
-for (i in 1:length(spp_quads)) { # i = 1
-  spp <- spp_list %>% filter(sciname == spp_quads[i])
-  spp_quad <- names(spp_quads)[i]
-
-  spp_plot <- plot_rangemap(spp)
-  spp_plot <- spp_plot + 
-    labs(title = element_blank())
+  spp_plot <- plot_depthmap(spp)
 

### 2016-03-27 data_explore/data_explore_coralmaps.Rmd
writing the raster and just gonna make the map in QGIS

@@
 
-``` {r define_fxns, echo = FALSE}
-#################################################################=
-### Species Map Function ###
-# This function takes a dataframe of loiczid per species ID, and a target species ID 
-### (as single or vector).  Return a df of cells ready to be rasterized.  
-### If AquaMaps, returned df includes aquamaps probability; if IUCN, returned
-### df includes proportional area.
-get_iucn_spp_map <- function(sid, spp_cells) {
-
-  spp_map <- spp_cells %>%
-    filter(iucn_sid %in% sid)
-  if(nrow(spp_map) == 0) {
-    message('no IUCN cells found')
-    spp_map <- spp_map %>%
-      as.data.frame() %>%
-      select(iucn_sid, loiczid) %>%
-      mutate(area_iucn = NA)
-  } else {
-    spp_map <- spp_map %>%
-      group_by(iucn_sid, loiczid) %>%        
-      summarize(area_iucn = max(prop_area))
-    ### The group_by() and summarize() are to collapse cells with 
-    ### overlapped polygons, fixing the duplicate 'by' issue.
-    ### Note: this drops presence, sciname, and subpop columns
-  }
-
-  return(spp_map %>% unique() %>% as.data.frame())
-}
-
-#################################################################=
-### Map vs Depth function
-### uses ggplot to create a formatted map of species ranges, including
-### Aquamaps, IUCN, and overlapped ranges; returns the plot object.
-plot_depthmap <- function(spp) {
-  map_iucn <- get_spp_map(spp$iucn_sid, iucn_spp_cells) %>%
-    mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0)) %>%
-    select(-area_iucn, -iucn_sid) %>%
-    unique()
-
-  r_iucn_spp <- subs(loiczid_raster, 
-                     map_iucn[ , c('loiczid', 'iucn_pres')], 
-                     by = 'loiczid', 
-                     which = 'iucn_pres', 
-                     subsWithNA = TRUE)
-
-  spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
-    mutate(presence = 'IUCN')
-
-  depth_poly_df <- rgdal::readOGR(dsn = 'coralmaps', layer = 'bath_200m') %>%
-    fortify() %>%
-    mutate(x = long, y = lat)
-
-  spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
-    ggtheme_map + 
-    geom_raster(aes(fill = presence), alpha = .8) +
-    scale_fill_manual(values = c('IUCN' = '#d95f02')) +
-    geom_polygon(data = depth_poly_df, color = 'black', fill = 'yellow', alpha = .8) +
-    borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
-    scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
-    scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2))
-}
-
-```
-
-```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
+```{r read_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}
 
@@
 
-  spp_plot <- plot_depthmap(spp)
+  map_iucn <- get_spp_map(spp$iucn_sid, iucn_spp_cells) %>%
+    mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0)) %>%
+    select(-area_iucn, -iucn_sid) %>%
+    unique()
 
-  print(spp_plot)
-  plotfile <- file.path(dir_fig, 'spp_maps', 
-                        sprintf('%s_%s.png', 
-                                spp_quad, 
-                                tolower(str_replace(spp_quads[i], ' ', '_'))))
-  ggsave(plotfile, height = 4.5, width = 8.7, units = 'cm')
+  r_iucn_spp <- subs(loiczid_raster, 
+                     map_iucn[ , c('loiczid', 'iucn_pres')], 
+                     by = 'loiczid', 
+                     which = 'iucn_pres', 
+                     subsWithNA = TRUE)
+  writeRaster(r_iucn_spp, 'coralmaps/oculina_varicosa.tif')
+
+#   spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
+#     mutate(presence = 'IUCN')
+#   
+#   rast_fine <- loiczid_raster %>%
+#     crop(extent(x = c(-135, -45), y = c(0, 45)))
+#   res(rast_fine) <- 0.05
+
+#   depth_poly <- rgdal::readOGR(dsn = 'coralmaps', layer = 'bath_200m')
+#   depth_pts <- rasterize(depth_poly, rast_fine) %>%
+#     rasterToPoints() %>%
+#     as.data.frame() %>%
+#     rename(bath200m = layer)
+#   
+#   spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
+#     ggtheme_map + 
+#     theme(panel.background = element_rect(fill = 'lightblue')) +
+# #    geom_raster(aes(fill = presence), alpha = 1) +
+#     geom_raster(data = depth_pts, aes(x = x, y = y, fill = bath200m),
+#                 alpha = .5) +
+# #    scale_fill_manual(values = c('IUCN' = '#d95f02')) +
+#     borders('world', color='gray40', fill='gray45', size = .1) +
+#     coord_cartesian(xlim = c(-105, -60), ylim = c(7.5, 37.5)) +
+#     scale_x_continuous(breaks = seq(-105, -60, by = 15)) +
+#     scale_y_continuous(breaks = seq( 0,  45, by = 15)) +
+#     theme(legend.position = 'none',
+#           axis.text = element_blank())
+#   
+#   print(spp_plot)
+#   plotfile <- file.path(dir_fig, 'spp_maps', 
+#                         sprintf('%s_%s.png', 
+#                                 spp_quad, 
+#                                 tolower(str_replace(spp_quads[i], ' ', '_'))))
+#   ggsave(plotfile, height = 4.5, width = 8.7, units = 'cm')
   }

### 2016-03-27 data_explore/data_explore_fxns.R

@@ -16,12 +16,6 @@ ggtheme_plot <- ggtheme_basic +
         axis.line = element_line(colour = "grey30"))
 
   
-
-# ggtheme_map <- theme(axis.ticks = element_blank(), axis.text = element_blank(),
-#                     text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-#                     plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-#                     legend.position = 'right')
-
 ### theme for species range maps - based on generic plot theme
 ggtheme_map <- ggtheme_basic +
   theme(axis.text = element_blank())

### 2016-03-27 manuscript-main.Rmd
created four maps image
still working on supplemental

--- 2016-03-23 manuscript-main.Rmd
@@
 title: "Aligning marine species range data to better serve science and conservation"
-author: Casey O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin S. Halpern<sup>1, 2, 3</sup>
-
+author: Casey O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin
+  S. Halpern<sup>1, 2, 3</sup>
 output:
@@
 
-``` {r SOM, child = 'ms_6_supplemental.Rmd'}
+``` {r SOM, child = 'ms_6_supplemental.Rmd', eval = TRUE}
 ```

### 2016-03-27 ms_6_supplemental.Rmd
created four maps image
still working on supplemental

@@
 
+<!--
+![](figures/si_four_maps.png)
+-->
+
+a) quad 1: Kajikia albida
+b) quad 2: Conus episcopatus
+c) quad 3: Conus magnificus
+d) quad 4: Acanthurus nigroris
+
 ## S2: risk by quadrant
@@
 
+<!--
 Oculina varicosa - Ivory tree coral ![oculina_pic.jpg](figures/oculina/oculina_pic.jpg)
+-->
 
@@
 
-plotted against a 200 m bathymetry line: ![oculina_map.png](figures/oculina/oculina_map.png)
+plotted against a 200 m bathymetry line:
+
+![oculina varicosa](figures/si_coralmap.png)
 

### 2016-03-27 manuscript-main.Rmd
done(ish)?

@@
   pdf_document:
-    number_sections: yes
+    number_sections: no
     toc: no
@@
     highlight: haddock
-    number_sections: yes
+    number_sections: no
     theme: spacelab
@@
 
------
-
 ``` {r calc species numbers, echo = FALSE}

### 2016-03-27 ms_1_abst_sig.Rmd
done(ish)?

@@
 
-_Abstract.
+<!--_Abstract.
 Provide an abstract of no more than 250 words on page 2 of the manuscript.
 Abstracts should explain to the general reader the major contributions of the article.
-References in the abstract must be cited in full within the abstract itself and cited in the text._
+References in the abstract must be cited in full within the abstract itself and cited in the text._-->
 
@@
 
-_Significance Statement.
+<!--_Significance Statement.
 Authors must submit a 120-word-maximum statement about the significance of their research paper written at a level understandable to an undergraduate-educated scientist outside their field of specialty.
 The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
-The Significance Statement appears in the paper itself and is required for all research papers._
+The Significance Statement appears in the paper itself and is required for all research papers._-->
 

### 2016-03-27 ms_3_methods.Rmd
done(ish)?

@@
 To examine the overall taxonomic distribution across the spatial datasets (Fig.
-1), we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
+1A), we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
 
@@
 We visually inspected a random selection of paired distribution maps from each quadrant to identify possible mechanistic causes of misalignment.
-To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS (REF) to overlay a selection of IUCN and AquaMaps maps with a 200 meter bathymetry contour.
+To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS (QGIS Development Team 2015) to overlay a selection of IUCN and AquaMaps maps with a 200 meter bathymetry contour.
 
@@
 Using methods and supplemental materials from OHI (Halpern et al.
-2012, 2015), we modified the original code for OHI 2015 Species status (SPP) (REF), allowing for flexibility in prioritized data source and AquaMaps presence threshold.
+2012, 2015), we modified the original code for OHI 2015 Species status (SPP), allowing for flexibility in prioritized data source and AquaMaps presence threshold.
 We ran the SPP code three times, prioritizing IUCN over AquaMaps for a 0% threshold, and prioritizing AquaMaps over IUCN for both a 40% and 0% threshold.
@@
 Based upon the methods described in Klein et al.
-2015, we reconstructed the analysis using the subset of protected areas (WDPA 2014) spatially covering a marine area and classified as IUCN I-IV.
+2015, we reconstructed the analysis using the subset of protected areas (IUCN and UNEP-WCMC 2014) spatially covering a marine area and classified as IUCN I-IV.
 The WDPA polygons and marine polygons were rasterized to 0.01° and then aggregated to AquaMaps native 0.5° cells, to calculate proportion of protected area and marine within each cell.
@@
 
-All processing was completed using R statistical software (REF), and all code and intermediate data are available on GitHub (REF?) at https://github.com/OHI-Science/IUCN-AquaMaps.
+All processing was completed using R statistical software (R Core Team 2015), and all code and intermediate data are available on GitHub at https://github.com/OHI-Science/IUCN-AquaMaps.

### 2016-03-27 ms_5_refs.Rmd
updating manuscript
updating manuscript, and creating maps for gaps and 10+% MPAs
done(ish)?

@@
-# Achnowledgments
+# Acknowledgments
 <!-- Acknowledgments. List acknowledgments and funding sources. PNAS collaborates with FundRef to use a standard taxonomy of funder names. Dedications are rarely allowed. -->
@@
 # References
-
-___double check formats; also - include DOIs for references? which if any?___
 
@@
 World wide web electronic publication, www.aquamaps.org, Version 08/2015.
-___since this is not a peer-reviewed publication, can we include this for PNAS reference list?___
+<!--___since this is not a peer-reviewed publication, can we include this for PNAS reference list?___-->
 
@@
 Downloaded on 21 December 2015.
-___since this is not a peer-reviewed publication, can we include this for PNAS reference list?___
+<!--___since this is not a peer-reviewed publication, can we include this for PNAS reference list?___-->
 
-Halpern BS et al. (2012). An index to assess the health and benefits of the global ocean. Nature, 488(7413), 615-620. _doi:10.1038/nature11397._
+Halpern BS et al. (2012). An index to assess the health and benefits of the global ocean. Nature, 488(7413), 615-620. <!--_doi:10.1038/nature11397._-->
 
-Halpern BS et al. (2015). Patterns and emerging trends in global ocean health. PloS one, 10(3), e0117863. ___doi?___
+Halpern BS et al. (2015). Patterns and emerging trends in global ocean health. PloS one, 10(3), e0117863. <!--___doi?___-->
 
-Klein CJ et al. (2015). Shortfalls in the global protected area network at representing marine biodiversity. Scientific Reports 5: 17539. _doi:10.1038/srep17539._
+Klein CJ et al. (2015). Shortfalls in the global protected area network at representing marine biodiversity. Scientific Reports 5: 17539. <!--_doi:10.1038/srep17539._-->
 
-Selig ER et al. (2013). Assessing global marine biodiversity status within a coupled socio-ecological perspective. PloS one, 8(4), e60284. _doi:10.1371/journal.pone.0060284._
+Selig ER et al. (2013). Assessing global marine biodiversity status within a coupled socio-ecological perspective. PloS one, 8(4), e60284. <!--_doi:10.1371/journal.pone.0060284._-->
 
-Coll M et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. _used AquaMaps to predict Med biodiversity.  Also: Threshold = 0._
+Coll M et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. <!--_used AquaMaps to predict Med biodiversity.  Also: Threshold = 0._-->
 
@@
 28 pp.
-(+ 4 annexes totalling 174 pp.
+<!--(+ 4 annexes totalling 174 pp.
 and one e-supplement).
-_review of marine data sets and data gaps etc, incl both IUCN and AM as well as many others_
+_review of marine data sets and data gaps etc, incl both IUCN and AM as well as many others_-->
 
@@
 Science, 344(6187), 1246752.
-_uses range maps to show biodiversity areas; may use IUCN range maps.
-Also discusses gaps and possible things that can be done about them._
+<!--_uses range maps to show biodiversity areas; may use IUCN range maps.
+Also discusses gaps and possible things that can be done about them._-->
 
@@
 PLoS ONE 6(5): e19653.
-_just what the title says - AM development, presence threshold 60%, also analyzes richness as a function of threshold_
+<!--_just what the title says - AM development, presence threshold 60%, also analyzes richness as a function of threshold_-->
 
-Molinos JG et al. (2015). Climate velocity and the future global redistribution of marine biodiversity. Nature Climate Change. _doi:10.1038/nclimate2769._
+Molinos JG et al. (2015). Climate velocity and the future global redistribution of marine biodiversity. Nature Climate Change. <!--_doi:10.1038/nclimate2769._-->
 
@@
 Ecology letters, 9(10), 1136-1145.
-_compares point locality, range maps, and distribution models in terms of omission and commission errors; also outlines Extent of Occurrence and Area of Occupancy distinctions._
+<!--_compares point locality, range maps, and distribution models in terms of omission and commission errors; also outlines Extent of Occurrence and Area of Occupancy distinctions._-->
 
@@
 Conservation Biology, 22(1), 110-119.
-_EOO maps are usually highly interpolated and overestimate small-scale occurrence, which may bias research outcomes_
+<!--_EOO maps are usually highly interpolated and overestimate small-scale occurrence, which may bias research outcomes_-->
 
@@
 Marine Ecology Progress Series 316: 285–310.
-_check this citation journal name..._ _This outlines the basic RES methodology - AM development_
+<!--_check this citation journal name..._ _This outlines the basic RES methodology - AM development_-->
 
@@
 Ecological Modelling 221(3): 467-478.
-_Presents AM; assessing AquaMaps against other presence-only species models_
+<!--_Presents AM; assessing AquaMaps against other presence-only species models_-->
 
@@
 
-### NOT CURRENTLY USED:
+QGIS Development Team (2015). QGIS Geographic Information System. Open Source Geospatial Foundation Project. http://qgis.osgeo.org
 
-Jones MC, Dyeb SR, Pinnegar JK, Cheung WWL (2012).
-Modelling commercial fish distributions: Prediction and assessment using different approaches.
-Ecological Modelling 225(2012): 133-145.
-PDF _comparison of species distribution models including AquaMaps, Maxent and the Sea Around Us Project_
+IUCN and UNEP-WCMC (2014). The World Database on Protected Areas (WDPA) [On-line]. Cambridge, UK: UNEP- WCMC. Available at: www.protectedplanet.net [Accessed December 1, 2014].
 
-Hurlbert AH, and Jetz W (2007).
-Species richness, hotspots, and the scale dependence of range maps in ecology and conservation.
-Proceedings of the National Academy of Sciences, 104(33), 13384-13389.
-_mostly rasters of range maps?
-"The scale dependence of range-map accuracy poses clear limitations on braod-scale ecological analyses and conservation assessments.
-...
-we provide guidance about the approriate scale of their use_
+R Core Team (2015). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

### 2016-03-27 ms_6_supplemental.Rmd
done(ish)?

@@
 
-#Outline
-
-## Data processing - should this go in methods???
+<!--## Data processing - should this go in methods???
 
@@
 AquaMaps data was provided by AquaMaps directly via ftp.
-Three SQL files were sent that contained (1) list of all `r nrow(spp_am)` species and their taxonomic information, (2) information for all (259,200?) half degree cells used for range mapping and (3) the probability of occurrence per cell for all `r nrow(spp_am)` species.
+Three SQL files were sent that contained (1) list of all `r nrow(spp_am)` species and their taxonomic information, (2) information for all (259,200) half degree cells used for range mapping and (3) the probability of occurrence per cell for all `r nrow(spp_am)` species.
 These were extracted into .csv form using R Statistical Software.
@@
 
-Each shapefile contains polygon limits of distribution indicating regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
-
-* which data sets are included?
-* raster::extract() to convert polys to csvs
-* which columns are included?
+Each shapefile contains polygon limits of distribution indicating regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.-->
 
@@
 
-<!--
 ![](figures/si_four_maps.png)
--->
 
-a) quad 1: Kajikia albida
-b) quad 2: Conus episcopatus
-c) quad 3: Conus magnificus
-d) quad 4: Acanthurus nigroris
+>  __Fig. S1.__ Representative species maps to illustrate each quadrant from Fig. 2A.  Note that A-D are arranged to match quadrants in Fig. 2A.
+(A: well-matched) Kajikia albida
+(B: distribution-matched) Conus episcopatus
+(C: area-matched) Conus magnificus
+(D: poorly-matched) Acanthurus nigroris
 
-## S2: risk by quadrant
+## S2: Risk by quadrant
+
+<a id = 'S2'></a>
 
@@
 
-* Breaking down the quadrants by IUCN extinction risk categories (Fig.
+>  __Fig.
+S2.__ Breaking down the quadrants by IUCN extinction risk categories (Fig.
 S2), we found little support for our hypothesis that maps for species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
@@
 
-## S3: coral depth map
+### S3: Coral depth map
+
+<a id = 'S3'></a>
+
+![oculina varicosa](figures/si_coralmap.png)
 
@@
 
-* from IUCN: Colonies are found to depths of 152 m depth on limestone rubble, low-relief limestone outcrops, high-relief, steeply sloping prominences, and soft-bottom sloping habitats.
-Colonies are semi-isolated, patchy and low-growing in shallow water, or they form larger, massive coalescing aggregates (thickets or coppices) with substantial topographic relief in 50-100 m depth.
-In shallow waters (2-30m) the form is zooxanthellate, inhabiting limestone ledges.
-In deeper waters, an azooxanthellate form is known from the shelf edge off eastern Florida, USA from Ft.
-Pierce to Daytona (Reed 1980, 1983, 2002; Brooke and Young 2003).
+>  __Fig.
+S3.__ IUCN-defined range of _Oculina varicosa_, a typical photosynthetic coral, plotted against a 200 m bathymetry contour.
+According to IUCN, _O.
+varicosa_ can be found to depths of 152 m.
+In waters shallower than 30 m, _O.
+varicosa_ is zooxanthellate, while an azooxanthellate form can be found in deeper waters.
 
-plotted against a 200 m bathymetry line:
+### S4: AquaMaps threshold examination
 
-![oculina varicosa](figures/si_coralmap.png)
+<a id = 'S4'></a>
 
-## S4: AquaMaps threshold examination
+AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value as indicated by IUCN maps.
+Many studies convert this AquaMaps probability to a simple presence value by assigning a threshold value.
+A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
 
@@
 
-S4a:
-
+A. 
 ![](figures/si_am_threshold.png)
 
-S4b:
-
+B. 
 ![](figures/si_ribbon_am_threshold.png)
 
->AquaMaps distribution map extent remaining after applying a presence threshold.
-(a) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of XXX, with a wide distribution in which some species lose nearly all of their apparent range.
-(b) Mean (median) remaining extent at increments of presence threshold.
-Dark grey ribbon includes 25% to 75% quantiles, while light grey ribbon includes 5% to 95% quantiles.
+>  __Fig.
+S4.__ AquaMaps distribution map extent remaining after applying a presence threshold.
+(A) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of 28% of total range, with a wide distribution in which some species lose nearly all of their apparent range.
+(B) Median remaining extent at increments of presence threshold.
+Dark green ribbon includes 25% to 75% quantiles of remaining species range, while lighter ribbon includes 5% to 95% quantiles.
+Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity of apparent species range to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.
 
-AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value as indicated by IUCN maps.
-Many studies convert this AquaMaps probability to a simple presence value by assigning a threshold value (REF references here).
-A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
-
-At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a significant decrease in represented range, and some species lose nearly their entire range.
-Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.
-
+<!--
 ## MPA gap analysis - explanation
@@
 GBR); AquaMaps = probably a little commission (possibly underestimating, if it overpredicts lots of range outside MPAs, thus reducing the proportion of protected range)
+-->
 

### 2016-03-27 ms_2_intro_res_disc.Rmd
ugh

@@
 3</a>).
-Reducing the threshold increases the apparent range of a species; the slight decrease in average score for scenario 1 suggests increased spatial representation of threatened species.
-The slight increase in mean score for scenario 2 may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or a combination of both.
-The large decrease shown in scenario 3 indicates that a zero threshold greatly increases the spatial representation of high-risk species.
+Reducing the threshold increases the apparent range of a species, so a decrease in a region’s score under scenario 1 would indicate increased spatial representation of threatened species for that region.
+Shifting priority from IUCN to AquaMaps may increase or decrease a species’ apparent range, so a decrease in region score for scenario 2 may indicate a decrease in spatial representation of low-risk species, an increase in spatial representation of high-risk species, or a combination of both.
+Scenario three combines effects of scenarios 1 and 2.
 
-Given that only `r p_spp_both_ohi`% of species overlap, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in such large country-level score shifts as seen in scenarios 2 and 3.
+Given that only `r p_spp_both_ohi`% of species overlap, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in substantial country-level score shifts as seen in scenarios 2 and 3.
 While the mean global score did not vary significantly from scenario to scenario, select countries gained up to 7.5 points while others dropped as many as 5 points.

### 2016-03-28 manuscript-main.Rmd
added Ben's edits

@@
 output:
-  word_document: default
-  pdf_document:
-    number_sections: no
-    toc: no
   html_document:
@@
     toc: no
+  pdf_document:
+    number_sections: no
+    toc: no
+  word_document: default
 ---
@@
 At least three keywords are required at submission.
-maybe species distribution maps, AquaMaps, IUCN, marine protected area planning, biodiversity status--> TBD
+maybe species distribution maps, AquaMaps, IUCN, marine protected area planning, biodiversity status--> TBD: These should be words NOT in the title or abstract (BSH comment)
 

### 2016-03-28 ms_2_intro_res_disc.Rmd
added Ben's edits

@@
 The two datasets have notably different taxonomic (Fig.
-1) and regional (Fig.
-2) coverage.
+1A) and regional (Figs.
+1B, 1C) coverage.
 AquaMaps encompasses a broader range of taxa than IUCN, as IUCN spatial data files are only available for select taxonomic groups that have been comprehensively assessed.
@@
 To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area alignment (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig.
-3A).
+2A).
  This analysis revealed a general negative linear pattern, suggesting that increasing similarity in range area correlates with decreasing distribution alignment.
@@
 By dividing the map-paired species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
-For representative maps from each category, please refer to the supplemental materials (Fig.
+Representative maps from each category are provided in the supplemental materials (Fig.
 S1).
@@
 These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig.
-3B).
+2B).
 This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their ranges span nearly the entire map.
@@
 Method-driven differences in commission and omission errors drive clear and significant differences in species range descriptions between AquaMaps and IUCN datasets.
-To examine the implications of these differences, we replicated two recent studies, varying only the prioritization of one data set over the other.
+To examine the implications of these differences, we repeated two recent studies, varying only the prioritization of one data set over the other.
 
@@
 Klein et al.
-(2015) compare the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect key species and identify which species fall through gaps in protection.
+(2015) compare the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect species and identify which species fall through gaps in protection.
 The study relied on the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
-They found that the global MPA network leaves 90.5% of marine species with less than 5% of their overall range represented within MPAs, and 1.4% of species have no protection at all ("gap" species).
+They found that the global MPA network leaves 90.5% of marine species with less than 5% of their overall range represented within MPAs, and 1.4% of species have no protection at all (i.e., "gap" species).
 
@@
 AquaMaps and IUCN range maps show reasonable agreement for many well-studied species, but substantial differences illustrate uncertainty in our understanding of spatial distribution for many others.
-Although many other appriaches exist for species distribution modeling, these two are the only ones applied globally to marine species.
+Although many other approaches exist for species distribution modeling, these two are the only ones applied globally to marine species.
 Identifying and addressing differences in these datasets will increase their utility for research and conservation actions.
-Several likely drivers of commission errors and omission errors between these datasets point to a few important ways to improve range alignment.
+Several likely drivers of commission and omission errors between these datasets point to a few important ways to improve range alignment.
 

### 2016-03-28 ms_3_methods.Rmd
ugh
added Ben's edits

@@
 We represented relative distribution of species count for each dataset by plotting average species count against latitude and longitude (Fig.
-2).
+1B, 1C).
 
@@
 
-Overlaying paired distribution maps for a given species, we defined and calculated _distribution alignment_ and _extent alignment_ and plotted these in Fig. 3A:
+Overlaying paired distribution maps for a given species, we defined and calculated _distribution alignment_ and _area alignment_ and plotted these in Fig. 2A:
 

### 2016-03-29 manuscript-main.Rmd
updated formats incl. references, figures, etc; split SI and figures into separate docs.  Also added bit about OBIS/GBIF

@@
 author: Casey O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin
-  S. Halpern<sup>1, 2, 3</sup>
+  S. Halpern^1,2,3^
 output:
+  word_document: default
+  pdf_document:
+    number_sections: no
+    toc: no
   html_document:
@@
     toc: no
-  pdf_document:
-    number_sections: no
-    toc: no
-  word_document: default
 ---
@@
 
+<!--
+| Section        |   Word #   | Char/spc # |
+| :------        |    ------: |    ------: |
+| Abstract       |  184 words |  1312 char |
+| Significance   |   73 words |   545 char |
+-->
+
 -----
+
+\\body
 
@@
 ```
+
+<!--
+| Section                   |   Word #   | Char/spc # |
+| :------                   |    ------: |    ------: |
+| Introduction              |  429 words |  2905 char |
+| Results/Discussion        |  999 words |  6697 char |
+| Implications (incomplete) |  597 words |  3794 char |
+| Conclusions               |  354 words |  2505 char |
+| Methods (incomplete)      |  601 words |  3919 char |
+| total body                | 2987 words | 19885 char |
+-->
 
@@
 
-``` {r figures, child = 'ms_4_figs.Rmd'}
+``` {r references, child = 'ms_4_refs.Rmd'}
 ```
@@
 
-``` {r references, child = 'ms_5_refs.Rmd'}
+``` {r figure_legends, child = 'ms_5_fig_legends.Rmd'}
 ```
 
------
-
-``` {r SOM, child = 'ms_6_supplemental.Rmd', eval = TRUE}
-```
+<!--
+| Figure  |  w x h (px) | w x h (cm)  | char # |
+| :------ |     ------: |     ------: |------: |
+| 1 A - C | 2102 x 1536 | 17.8 x 13.0 |   4680 |
+| caption |             |             |    367 | 
+| 2 A - B | 2102 x 1940 | 17.8 x 16.5 |   5940 |
+| caption |             |             |    934 | 
+|       3 | 1027 x  531 |  8.7 x  4.5 |    810 |
+| caption |             |             |    744 | 
+|       4 | 1346 x  590 | 11.4 x  5.0 |   1800 |
+| caption |             |             |    870 | 
+|  total  |             |             |  16145 |
+-->

### 2016-03-29 ms_2_intro_res_disc.Rmd
updated formats incl. references, figures, etc; split SI and figures into separate docs.  Also added bit about OBIS/GBIF

@@
 ---
-
-<!--
-| Section                   |   Word #   | Char # (w/spc) |
-| :------                   |    ------: |    ------: |
-| Introduction              |  512 words |  3411 char |
-| Results/Discussion        |  802 words |  5410 char |
-| Implications (incomplete) |  532 words |  3334 char |
-| Conclusions               |  390 words |  2813 char |
-| Methods (incomplete)      |  612 words |  3889 char |
-| total body                | 2848 words | 18857 char |
--->
 
@@
 One major outcome of this body of science is the various compiled databases of species distribution maps.
-Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps (Kaschner et al.
-2013) and International Union for Conservation of Nature (IUCN) (IUCN 2015).
-These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status (Halpern et al.
-2012, 2015; Selig et al.
-2013), evaluating global biodiversity patterns (Coll et al.
-2010; Martin et al.
-2014, Pimm et al.
-2014, Kaschner et al.
-2011), predicting range shifts (Molinos et al.
-2015), and setting conservation priorities (Klein et al.
-2015).
+Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps (1) and International Union for Conservation of Nature (IUCN) (2).
+These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status (3, 4, 5), evaluating global biodiversity patterns (6, 7, 8, 9), predicting range shifts (10), and setting conservation priorities (11).
 
@@
 Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
-Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans (Rondinini et al.
-2006, Jetz et al.
-2008).
+Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans (12, 13).
 
@@
 
-We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (Halpern et al.
-2012, 2015) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (Klein et al.
-2015) - as case studies to explore the implications of prioritizing one data set over the other.
+We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (3, 4) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (11) - as case studies to explore the implications of prioritizing one data set over the other.
 The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
@@
 
-The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" (IUCN 2015) - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.
+The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" (2) - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.
 In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of "probability of occurrence" for each species (Kaschner et al.
-2006, Ready et al.
-2010).
-Studies frequently define "presence" for AquaMaps by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence (Halpern et al.
-2012, 2015).
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of "probability of occurrence" for each species (14, 15).
+Studies frequently define "presence" for AquaMaps by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence (3, 4).
 
@@
 Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.
-Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range (Rondinini et al.
-2006).
+Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range (12).
 By comparing maps resulting from IUCN and AquaMaps methodolog, we can identify and possibly address mechanistic causes for each type of error.
@@
 The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor area match well.
-Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging OBIS and GBIF occurrences) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
+Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) (16) and the Global Biodiversity Information Facility (GBIF) (17)) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
 When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species.
@@
 
-The global Ocean Health Index (OHI) (Halpern et al.
-2012, 2015), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
+The global Ocean Health Index (OHI) (3, 4), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
 To maximize the number of represented species, OHI gleans spatial distributions from Red List-assessed species in both IUCN and AquaMaps datasets (n = `r format(n_spp_ohi, big.mark = ',')`), prioritizing IUCN data for the `r format(n_spp_both_ohi, big.mark = ',')` species included in both sources.
@@
 Klein et al.
-(2015) compare the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect species and identify which species fall through gaps in protection.
+(11) compare the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect species and identify which species fall through gaps in protection.
 The study relied on the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
@@
 For example, for most corals, researchers could clip range maps to the photosynthetic limit of 200 meters.
-For AquaMaps range data, dependent primarily on environmental and physical preferences and conditions, implementing area restrictions based on biogeographical criteria such as Marine Ecoregions of the World (Spalding et al.
-2007) would likely decrease commission errors and improve predictive power, especially for data-poor species.
+ For AquaMaps range data, dependent primarily on environmental and physical preferences and conditions, implementing area restrictions based on biogeographical criteria such as Marine Ecoregions of the World (18) would likely decrease commission errors and improve predictive power, especially for data-poor species.
+ For either data set, maps based on few occurrences are more likely to include errors of commission and omission.
+Occurrence counts from external sources such as OBIS or GBIF can help identify relatively data-poor species; additionally, the "occurcells" attribute in the AquaMaps data set, which counts the number of half-degree cells used to generate the environmental envelope for each species, can be used in a similar manner.
 

### 2016-03-29 ms_3_methods.Rmd
updated formats incl. references, figures, etc; split SI and figures into separate docs.  Also added bit about OBIS/GBIF

@@
 
-**Comparison of taxonomic and regional distribution:** 
+__Comparison of taxonomic and regional distribution__:
 To examine the overall taxonomic distribution across the spatial datasets (Fig.
@@
 
-**Comparison of paired maps:** 
+__Comparison of paired maps__:
 Using genus and species binomials as a matching key, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.
@@
 We visually inspected a random selection of paired distribution maps from each quadrant to identify possible mechanistic causes of misalignment.
-To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS (QGIS Development Team 2015) to overlay a selection of IUCN and AquaMaps maps with a 200 meter bathymetry contour.
+To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS (19) to overlay a selection of IUCN and AquaMaps maps with a 200 meter bathymetry contour.
 
 **Methods for OHI case study: **
-Using methods and supplemental materials from OHI (Halpern et al.
-2012, 2015), we modified the original code for OHI 2015 Species status (SPP), allowing for flexibility in prioritized data source and AquaMaps presence threshold.
+Using methods and supplemental materials from OHI (3, 4), we modified the original code for OHI 2015 Species status (SPP), allowing for flexibility in prioritized data source and AquaMaps presence threshold.
 We ran the SPP code three times, prioritizing IUCN over AquaMaps for a 0% threshold, and prioritizing AquaMaps over IUCN for both a 40% and 0% threshold.
@@
 Based upon the methods described in Klein et al.
-2015, we reconstructed the analysis using the subset of protected areas (IUCN and UNEP-WCMC 2014) spatially covering a marine area and classified as IUCN I-IV.
+(11), we reconstructed the analysis using the subset of protected areas from the World Database on Protected Areas (WDPA) (20) spatially covering a marine area and classified as IUCN I-IV.
 The WDPA polygons and marine polygons were rasterized to 0.01° and then aggregated to AquaMaps native 0.5° cells, to calculate proportion of protected area and marine within each cell.
@@
 
-All processing was completed using R statistical software (R Core Team 2015), and all code and intermediate data are available on GitHub at https://github.com/OHI-Science/IUCN-AquaMaps.
+All processing was completed using R statistical software (21), and all code and intermediate data are available on GitHub at https://github.com/OHI-Science/IUCN-AquaMaps.

### 2016-03-29 ms_5_fig_legends.Rmd
updated formats incl. references, figures, etc; split SI and figures into separate docs.  Also added bit about OBIS/GBIF

--- 2016-03-26 ms_4_figs.Rmd
@@
-# Figures and captions
-
-<!--
-
-Figures: Calculated at 180 characters per cm in height for one column and 360 characters per cm in height for two columns; a figure wider than 8.7 cm is two columns wide.
-
-From PNAS http://www.pnas.org/site/authors/preparation.xhtml:
-
-Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text.
-Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
-(See SI below for supplementary material.) Color images must be in RGB (red, green, blue) mode.
-Include the font files for any text.
-Images must be final size, preferably one column width (8.7 cm).
-Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Numbers, letters, and symbols should be no smaller than 6 points (2 mm) and no larger than 12 points (6 mm) after reduction and must be consistent.
-Composite figures must be preassembled.
-Figures must be submitted as separate files, not embedded in manuscript text.
-See the Digital Art Guidelines.
-
-See also: http://www.pnas.org/site/misc/digitalart.pdf
-
-Label like so: 
-Fig. 2.
-Diffraction patterns at three different angles and time-dependent structural changes of JDF-L1.
-(A–C) Diffraction images observed at different zone axes before time 0; n is the electron direction.
-(D) Femtosecond intensity changes of (200) and (201) spots with time.
-Two different types of dynamics were observed: a decay with a femtosecond time constant (see main text) and another without a decay.
-
-___consider using HCL or HSV values for colors on plots?___ or ask Mel re: good color choices.
-Make sure that colors have good contrast to show up on a black-and-white print...
-Should the quad plot background be colored?
-How else to quickly code the four quadrants relative to the quadrant-breakdown-by-taxa plot?
-
-**`hcl {grDevices}`**: Create a vector of colors from vectors specifying hue, chroma and luminance.
-
-*Usage* `hcl(h = 0, c = 35, l = 85, alpha, fixup = TRUE)`
-Arguments
-
-*arguments*
-- `h`: The hue of the color specified as an angle in the range [0,360]. 0 yields red, 120 yields green 240 yields blue, etc.
-- `c`: The chroma of the color. The upper bound for chroma depends on hue and luminance.
-- `l`: A value in the range [0,100] giving the luminance of the colour. For a given combination of hue and chroma, only a subset of this range is possible.
-- `alpha`: numeric vector of values in the range [0,1] for alpha transparency channel (0 means transparent and 1 means opaque).
-- `fixup`: a logical value which indicates whether the resulting RGB values should be corrected to ensure that a real color results.
-if fixup is FALSE RGB components lying outside the range [0,1] will result in an NA value.
-
-**`hsv {grDevices}`**: Create a vector of colors from vectors specifying hue, saturation and value.
-
-*usage* `hsv(h = 1, s = 1, v = 1, alpha)`
-
-*arguments*
-
-- `h`, `s`,`v`: numeric vectors of values in the range [0, 1] for ‘hue’, ‘saturation’ and ‘value’ to be combined to form a vector of colors. Values in shorter arguments are recycled.
-- `alpha`: numeric vector of values in the range [0, 1] for alpha transparency channel (0 means transparent and 1 means opaque).
-
--->
-
------
-
-### Fig. 1
-<a id = 'fig1'></a>
-
-![](figures/fig1abc_am_iucn_maps.png)
+# Figure Legends
 
@@
 
->  __Fig.
+__Fig.
 1.__ (A) Number and proportion of species by taxa included in each dataset.
 Overlapping species are dominated by bony fishes (`r format(n_bonyfish, big.mark = ',')` species, primarily tropical taxa) and corals (`r format(n_corals, big.mark = ',')` species).
-(B, C) Number of global marine species according to (A) AquaMaps and (B) IUCN.
+(B, C) Global marine species count per cell according to (B) AquaMaps and (C) IUCN.
 The margin frequency plots show relative species count per cell at each latitude and longitude.
-
-### Fig. 2
-<a id = 'fig2'></a>
-
-![](figures/fig2ab_quadmap_sppbars.png)
 
@@
 
-> __Fig.
+__Fig.
 2.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r format(n_spp_both, big.mark = ',')` species included in both IUCN and AquaMaps datasets.
-The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
-(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
-> The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent
+The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges (n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
+The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent
 (n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
-> The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges occur
+The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges occur
 (n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
-> The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
+The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
 (n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).
@@
 
-### Figure 3
-
-<a id = 'fig4'></a>
-
-![Absolute change in SPP score](figures/fig3_boxplot_abs_diff.png)
-
-> __Fig. 3.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
-> Scenario 1 shows the effect of reducing the probability threshold to 0% for AquaMaps presence to more accurately track the definition of IUCN “limits of distribution."  
-> Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the 40% presence threshold.
+__Fig. 3.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
+Scenario 1 shows the effect of reducing the probability threshold to 0% for AquaMaps presence to more accurately track the definition of IUCN “limits of distribution."
+Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the 40% presence threshold.
 In general, AquaMaps ranges are smaller than IUCN ranges, so most overlapping species will see a decrease in represented range.
-> Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, and simultaneously eliminating the presence threshold.
+Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, and simultaneously eliminating the presence threshold.
 The zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
 
-### Figure 4
-
-<a id = 'fig4'></a>
-
-![Change in MPA coverage](figures/fig4_barchart_mpa_gap.png)
-
-> __Fig.
+__Fig.
 4.__  Percent of species range covered by MPAs based upon methods in Klein et al.
@@
 Scenario 5 examines species MPA coverage using only the IUCN dataset.
-
-<!-- `r knitr::kable(read_csv(file.path(dir_data, 'mpa_all_runs_summary.csv')))` -->

### 2016-03-29 ms_figures.Rmd
updated formats incl. references, figures, etc; split SI and figures into separate docs.  Also added bit about OBIS/GBIF

--- 2016-03-26 ms_4_figs.Rmd
@@
-# Figures and captions
+---
+title: "Figures"
+output:
+  word_document: default
+  pdf_document:
+    number_sections: no
+    toc: no
+  html_document:
+    highlight: haddock
+    number_sections: no
+    theme: spacelab
+    toc: no
+---
 
@@
 
------
-
 ### Fig. 1
@@
 
-``` {r barchart_by_taxa, echo = FALSE}
-n_spp <- spp_map_pairs %>%
-  mutate(spp_group = ifelse(str_detect(spp_group, 'CORAL'), 'CORAL', spp_group)) %>%
-  group_by(spp_group) %>%
-  summarize(n_spp = n()) %>%
-  arrange(desc(n_spp))
-
-n_corals   <- nrow(spp_map_pairs %>% 
-                     filter(str_detect(spp_group, 'CORAL')))
-n_bonyfish <- nrow(spp_map_pairs %>% 
-                     filter(spp_group %in% c('WRASSES', 'BUTTERFLYFISH', 'PARROTFISH',
-                                             'COMBTOOTHBLENNIES', 'SEABREAMS_PORGIES', 
-                                             'PUFFERFISH', 'GROUPERS', 'TUNAS_BILLFISHES', 
-                                             'SURGEONFISH_TANGS_UNICORNFISH', 'ANGELFISH', 
-                                             'DAMSELFISH', 'BONEFISH_TARPONS')))
-```
-
->  __Fig.
-1.__ (A) Number and proportion of species by taxa included in each dataset.
-Overlapping species are dominated by bony fishes (`r format(n_bonyfish, big.mark = ',')` species, primarily tropical taxa) and corals (`r format(n_corals, big.mark = ',')` species).
-(B, C) Number of global marine species according to (A) AquaMaps and (B) IUCN.
-The margin frequency plots show relative species count per cell at each latitude and longitude.
-
 ### Fig. 2
@@
 ![](figures/fig2ab_quadmap_sppbars.png)
-
-``` {r tally_quad_counts, echo = FALSE}
-dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
-area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
-q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
-```
-
-> __Fig.
-2.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r format(n_spp_both, big.mark = ',')` species included in both IUCN and AquaMaps datasets.
-The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges
-(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
-> The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent
-(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
-> The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges occur
-(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
-> The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).  
-(B) Alignment quadrant breakdown of paired-map species by taxonomic group.
 
@@
 
-> __Fig. 3.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
-> Scenario 1 shows the effect of reducing the probability threshold to 0% for AquaMaps presence to more accurately track the definition of IUCN “limits of distribution."  
-> Scenario 2 shows the effect of prioritizing AquaMaps data over IUCN, while maintaining the 40% presence threshold.
-In general, AquaMaps ranges are smaller than IUCN ranges, so most overlapping species will see a decrease in represented range.
-> Scenario 3 shows the effect of prioritizing AquaMaps data over IUCN, and simultaneously eliminating the presence threshold.
-The zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
-
 ### Figure 4
@@
 ![Change in MPA coverage](figures/fig4_barchart_mpa_gap.png)
-
-> __Fig.
-4.__  Percent of species range covered by MPAs based upon methods in Klein et al.
-(2015).
-Scenario 1 replicates the original results, measuring protected range of species in AquaMaps 2014 dataset, with a 50% presence threshold, against the 2014 World Database of Protected Areas, filtered for IUCN categories I-IV that overlap marine areas.
-Scenario 2 updates the results using AquaMaps 2015, showing very small changes despite the inclusion of an additional 5,545 species.
-Scenario 3, using 2015 AquaMaps data, drops the presence threshold to zero, showing an expected decrease in gap species, but also a decrease in species with 5% or greater protected range.
-Scenario 4 adds an additional 1745 species unique to IUCN, resulting in increases in gap species and species with less than 2% coverage.
-Scenario 5 examines species MPA coverage using only the IUCN dataset.
-
-<!-- `r knitr::kable(read_csv(file.path(dir_data, 'mpa_all_runs_summary.csv')))` -->

### 2016-03-29 ms_supplemental.Rmd
updated formats incl. references, figures, etc; split SI and figures into separate docs.  Also added bit about OBIS/GBIF

--- 2016-03-27 ms_6_supplemental.Rmd
@@
-# Supplemental Information
+---
+title: "Supplemental Information"
+output:
+  word_document: default
+  pdf_document:
+    number_sections: no
+    toc: no
+  html_document:
+    highlight: haddock
+    number_sections: no
+    theme: spacelab
+    toc: no
+---
 
@@
 This file type will be published in raw format and will not be edited or composed.
-
 -->
 
-<!--## Data processing - should this go in methods???
+<!--## Data processing - should this go in methods??? -->
 
+<!--
 All code is publicly available and freely downloadable [here on GitHub](link to github).
 
-### AquaMaps
-
 AquaMaps data was provided by AquaMaps directly via ftp.
-Three SQL files were sent that contained (1) list of all `r nrow(spp_am)` species and their taxonomic information, (2) information for all (259,200) half degree cells used for range mapping and (3) the probability of occurrence per cell for all `r nrow(spp_am)` species.
+Three SQL files were sent that contained (1) list of all 22,889 modeled species and their taxonomic information, (2) information for all (259,200) half degree cells used for range mapping and (3) the probability of occurrence per cell for all 22,889 species.
 These were extracted into .csv form using R Statistical Software.
 
-### IUCN
-
-As of December 2015, IUCN had published species distribution maps for `r nrow(spp_iucn_maps)` marine species across 24 taxonomic groups.
+As of December 2015, IUCN had published species distribution maps for XXX marine species across 24 taxonomic groups.
 For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.

### 2016-03-30 data_explore/data_explore.Rmd
fixing up figs

@@
 
+#library(colorspace)
+#hclvec <- heat_hcl(3); colvec <- c('AquaMaps' = hclvec[1], 'Both' = hclvec[2], 'IUCN' = hclvec[3])
+#colvec <- c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')
+colvec <- c('AquaMaps' = '#76A94C', 'Both' = '#e18258', 'IUCN' = '#9168a9')
+
 taxa_df_long2 <- taxa_df %>%
@@
          percent = n_spp/n_total * 100,
-         dataset = factor(dataset, levels = c('Both', 'IUCN', 'AquaMaps')))
+         dataset = factor(dataset, levels = c('AquaMaps', 'Both', 'IUCN'), ordered = TRUE)) %>%
+  arrange(desc(dataset))
 
@@
   geom_bar(stat = 'identity', alpha = .8) +
-#  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
- scale_fill_manual(values = c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')) +
-#  scale_fill_manual(values = c('AquaMaps' = '#4dac26', 'Both' = '#fbe5c5', 'IUCN' = '#d01c8b')) +
+  scale_fill_manual(values = colvec) +
   geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
@@
 
-### define representative species from each quadrant, for over-plotting
-# spp_quads <- c('q1' = 'Kajikia albida', 
-#                'q2' = 'Galaxea cryptoramosa', 
-#                'q4' = 'Acanthurus nigroris', 
-#                'q4a' = 'Metanephrops australiensis', 
-#                'q3' = 'Conus magnificus')
-# 
-### truncate the species names for the representative species
-# spp_list_label <- spp_list %>%
-#   filter(sciname %in% spp_rando) %>%
-#   separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
-#   mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
-#   dplyr::select(-g, -s)
-
 ### mongo plot time
@@
                                y = sm_perc, 
-                               color = popn_category,
-                               size = lg_area_pct)) + 
+ #                               size = lg_area_pct,
+                               color = popn_category)) + 
   ggtheme_plot + 
@@
              alpha = .3, 
-#             fill= "green4")  + 
              fill= "#4dac26")  + 
@@
              alpha = .3, 
-#             fill= "greenyellow") + 
              fill= "#b8e186") + 
@@
              alpha = .3, 
-#             fill= "orange") + 
              fill= "#f1b6da") + 
@@
              alpha = .3, 
-#             fill= "orangered2") + 
              fill= "#d01c8b") + 
@@
                                  'DD' = 'grey30')) + 
-  scale_size_continuous(limits = c(0, 100),
-                        breaks = c( 0,    25,    50,    75,    100), 
-                        labels = c('0%', '25%', '50%', '75%', '100%')) + 
+#   scale_size_continuous(limits = c(0, 100),
+#                         breaks = c( 0,    25,    50,    75,    100), 
+#                         labels = c('0%', '25%', '50%', '75%', '100%')) + 
   scale_x_continuous(expand = c(0, 0), 
@@
 
-### here are mean lines:
-# scatter_quadplot <- scatter_quadplot +
-#   geom_vline(x = area_align_mean,
-#              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = dist_align_mean,
-#              color = 'grey50', linetype = 'dashed')
-
-### here are individual species, identified and namean on the plot:
-# scatter_quadplot <- scatter_quadplot +
-#   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
-#   geom_text(data = spp_list_label, 
-#             aes(label = sciname_label), 
-#             color = 'grey30',
-#             fontface = 'bold.italic',
-#             size = 2,
-#             hjust = .25,
-#             vjust = 1.5)
-
 scatter_quadplot <- scatter_quadplot +
-  labs(x = bquote('Area alignment '*alpha['ext']*' (%)'), 
+  labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
        y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
 #       title = 'Distribution alignment vs Extent alignment',
-       color = 'Risk', size = '% ocean range')
+#       size = '% ocean range',
+       color = 'Extinction\nRisk')
 

### 2016-03-30 data_explore/data_explore_quadmaps.Rmd
fixing up figs

@@
 
-spp_list <- spp_list %>%
-  mutate(area_ratio = am_area / iucn_area,
-         area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
-         area_ratio = area_ratio * 100)
+### define 25%, 50%, and 75% quartiles for both x and y axes
+area_align_mean <- mean(spp_list$area_ratio, na.rm = TRUE)
+dist_align_mean <- mean(spp_list$sm_perc, na.rm = TRUE)
 
-### define 25%, 50%, and 75% quartiles for both x and y axes
-area_ratio_quartiles <- quantile(spp_list$area_ratio, na.rm = TRUE)[2:4]
-pct_overlap_quartiles <- quantile(spp_list$sm_perc, na.rm = TRUE)[2:4]
+### define representative species from each quadrant, for over-plotting
+spp_quads <- c('q1' = 'Kajikia albida', 
+               'q2' = 'Conus episcopatus', 
+               'q4' = 'Acanthurus nigroris', 
+               'q3' = 'Conus magnificus')
 
-# spp_rando <- c("Lagocephalus lagocephalus",
-#                "Makaira nigricans",
-#                "Pentacheles laevis",
-#                "Stereomastis suhmi",
-#                "Stereomastis surda") 
-# 
 ### truncate the species names for the representative species
-# spp_list_label <- spp_list %>%
-#   filter(sciname %in% spp_rando) %>%
-#   separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
-#   mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
-#   dplyr::select(-g, -s)
+spp_list_label <- spp_list %>%
+  filter(sciname %in% spp_quads) %>%
+  separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
+  mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
+  dplyr::select(-g, -s)
 
 ### mongo plot time
-scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
+scatter_quadplot <- ggplot(spp_list %>% 
                              transform(popn_category = reorder(popn_category, category_score)),
-                                       aes(x = area_ratio, y = sm_perc, 
-                                           color = popn_category,
-                                           size = lg_area_pct)) + 
+                           aes(x = area_ratio, 
+                               y = sm_perc, 
+                               color = popn_category)) + 
   ggtheme_plot + 
   ### color the quadrant backgrounds:
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
-    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
-                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
+    annotate("rect", xmin = area_align_mean, xmax = 100, 
+                     ymin = dist_align_mean, ymax = 100, 
+             alpha = .3, 
+             fill= "#4dac26")  + 
+    annotate("rect", xmax = area_align_mean, xmin =   0, 
+                     ymin = dist_align_mean, ymax = 100, 
+             alpha = .3, 
+             fill= "#b8e186") + 
+    annotate("rect", xmin = area_align_mean, xmax = 100, 
+                     ymax = dist_align_mean, ymin =   0, 
+             alpha = .3, 
+             fill= "#f1b6da") + 
+    annotate("rect", xmax = area_align_mean, xmin =   0, 
+                     ymax = dist_align_mean, ymin =   0, 
+             alpha = .3, 
+             fill= "#d01c8b") + 
   geom_point(alpha = .6)
@@
 ### Manage scales for color and size 
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+scatter_quadplot <- scatter_quadplot +
   scale_colour_manual(values = c('LC' = 'green4', 
@@
                                  'DD' = 'grey30')) + 
-  scale_size_continuous(limits = c(0, 100),
-                        breaks = c( 0,    25,    50,    75,    100), 
+  scale_x_continuous(expand = c(0, 0), 
+                     limits = c(0, 102),
+                     breaks = c(seq(0, 100, 25)),
                      labels = c('0%', '25%', '50%', '75%', '100%')) +
-  scale_x_continuous(expand = c(0, 0), limits = c(0, 100)) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))
+  scale_y_continuous(expand = c(0, 0),
+                     limits = c(0, 103),
+                     breaks = c(seq(0, 100, 25)),
+                     labels = c('0%', '25%', '50%', '75%', '100%'))
 
-### here are median labels:
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
+### here are mean labels:
+scatter_quadplot <- scatter_quadplot +
   annotate(geom = 'text',
-           x = area_ratio_quartiles[2], y = 5,
+           x = area_align_mean, y = 5,
            hjust = 0, vjust = 0,
-           color = 'grey50', size = 2,
+           color = 'grey30', 
+           size = 2,
            fontface = 'bold.italic', angle = 90,
-           label = sprintf('Median = %s%%', round(area_ratio_quartiles[2], 1))) +
+           label = sprintf('Mean = %s%%', round(area_align_mean, 1))) +
   annotate(geom = 'text',
-           x = 5, y = pct_overlap_quartiles[2],
+           x = 5, y = dist_align_mean,
            hjust = 0, vjust = 0,
-           color = 'grey50', size = 2,
+           color = 'grey30', 
+           size = 2,
            fontface = 'bold.italic', angle = 0,
-           label = sprintf('Median = %s%%', round(pct_overlap_quartiles[2], 1)))
+           label = sprintf('Mean = %s%%', round(dist_align_mean, 1)))
 
-### here are median lines:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_vline(x = area_ratio_quartiles[2],
-#              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = pct_overlap_quartiles[2],
-#              color = 'grey50', linetype = 'dashed')
+### here are individual species, identified and namean on the plot:
+scatter_quadplot <- scatter_quadplot +
+  geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
+  geom_text(data = spp_list_label, 
+            aes(label = sciname_label), 
+            color = 'grey30',
+            fontface = 'bold.italic',
+            size = 3,
+            hjust = .75,
+            vjust = 1.5)
 
-### here are 25% and 75% quantile lines:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_vline(x = area_ratio_quartiles[c(1, 3)],
-#              color = 'grey50', linetype = 'dashed') +
-#   geom_hline(y = pct_overlap_quartiles[c(1, 3)],
-#              color = 'grey50', linetype = 'dashed')
+scatter_quadplot <- scatter_quadplot +
+  labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
+       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
+       color = 'Extinction\nRisk')
 
-### here are individual species, identified and named on the plot:
-# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-#   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
-#   geom_text(data = spp_list_label, 
-#             aes(label = sciname_label), 
-#             color = 'grey30',
-#             fontface = 'bold.italic',
-#             size = 2,
-#             hjust = .25,
-#             vjust = 1.5)
+# print(scatter_quadplot)
 
-scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
-  labs(x = bquote('Extent alignment '*alpha['ext']*' (%)'), 
-       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
-       title = 'Distribution alignment vs Extent alignment',
-       color = 'Risk', size = '% ocean range')
-
-print(scatter_arearatio_vs_overlap)
+ggsave(file.path(dir_fig, 'si_scatter_quadmap_spp.png'),
+       width = 17.8, height = 9, units = 'cm')
 
 ```
+
+![](../figures/si_scatter_quadmap_spp.png)
 
@@
   spp_plot <- spp_plot + 
-    labs(title = element_blank())
+    labs(title = element_blank()) +
+    theme(legend.position = 'none')
 

### 2016-03-30 ms_5_fig_legends.Rmd
fixing up figs

@@
 __Fig.
-2.__ (A) Distribution alignment (overlap of smaller range within larger) versus extent ratio (the ratio of smaller range area to the larger range area) for `r format(n_spp_both, big.mark = ',')` species included in both IUCN and AquaMaps datasets.
+2.__ (A) Distribution alignment (overlap of smaller range within larger) versus area ratio (the ratio of smaller range area to the larger range area) for `r format(n_spp_both, big.mark = ',')` species included in both IUCN and AquaMaps datasets.
 The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges (n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
-The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in extent
+The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in area
 (n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
-The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range extent, but disagree on where those ranges occur
+The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range area, but disagree on where those ranges occur
 (n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
@@
 (n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).
-(B) Alignment quadrant breakdown of paired-map species by taxonomic group.
+(B) Alignment quadrant breakdown of species by taxonomic group.
 

### 2016-03-30 ms_supplemental.Rmd
fixing up figs

@@
 S2), we found little support for our hypothesis that maps for species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
-Does higher perceived risk lead to increased attention, and thus better understanding of species distribution?
-Or conversely, does increased attention to species distribution reveal more species at risk?
-Likely both mechanisms are at play on a case-by-case basis, depending on the species' taxon and region.
-_does this argument bear up to closer scrutiny?
-CR isn't dominated by Q1 any more_
 
@@
 
->  __Fig.
-S3.__ IUCN-defined range of _Oculina varicosa_, a typical photosynthetic coral, plotted against a 200 m bathymetry contour.
-According to IUCN, _O.
-varicosa_ can be found to depths of 152 m.
-In waters shallower than 30 m, _O.
-varicosa_ is zooxanthellate, while an azooxanthellate form can be found in deeper waters.
+>  __Fig. S3.__ IUCN-defined range of _Oculina varicosa_, a typical photosynthetic coral, plotted against a 200 m bathymetry contour. According to IUCN, _O. varicosa_ can be found to depths of 152 m.
 

### 2016-03-30 case_study_ohi/boxplots.R

@@ -91,9 +91,9 @@ bplot_abs <- ggplot(data = st_all %>% filter(str_detect(scenario, 'abs')),
   labs(# title = 'SPP status difference by scenario',
        x = 'Scenario',
        y = 'Change in SPP status') +
-  scale_x_discrete(labels = c('IUCN/AquaMaps\nno threshold', 
-                              'AquaMaps/IUCN\n40% threshold', 
-                              'AquaMaps/IUCN\nno threshold'))
+  scale_x_discrete(labels = c('1. IUCN/AquaMaps\nno threshold', 
+                              '2. AquaMaps/IUCN\n40% threshold', 
+                              '3. AquaMaps/IUCN\nno threshold'))
 
 x <- st_all %>% filter(str_detect(scenario, 'abs'))
 y <- x %>% 

### 2016-03-30 data_explore/data_explore_fxns.R

@@ -6,19 +6,21 @@ library(ggplot2)
 ggtheme_basic <- theme(axis.ticks = element_blank(),
         text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
         plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
+        panel.background = element_blank(),
         legend.position = 'right')
 
 ggtheme_plot <- ggtheme_basic + 
   theme(panel.border     = element_blank(),
         panel.grid.minor = element_blank(), 
         panel.grid.major = element_line(colour = 'grey90'),
-        panel.background = element_blank(),
         axis.line = element_line(colour = "grey30"))
 
   
 ### theme for species range maps - based on generic plot theme
 ggtheme_map <- ggtheme_basic +
-  theme(axis.text = element_blank())
+  theme(axis.text = element_blank(),
+        axis.title = element_blank())
+
 
 #################################################################=
 ### Function to create a species list from scratch... use when new data is available
@@ -205,14 +207,20 @@ plot_rangemap <- function(spp) {
                              ifelse(iucn_pres & is.na(am_pres), 'IUCN',
                                     ifelse(iucn_pres & am_pres, 'Both', NA))))
   
+  #library(colorspace)
+  #hclvec <- heat_hcl(3); colvec <- c('AquaMaps' = hclvec[1], 'Both' = hclvec[2], 'IUCN' = hclvec[3])
+  #colvec <- c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')
+  colvec <- c('AquaMaps' = '#76A94C', 'Both' = '#e18258', 'IUCN' = '#9168a9')
+  
   spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
     ggtheme_map + 
     geom_raster(aes(fill = presence), alpha = .8) +
-    scale_fill_manual(values = c('AquaMaps' = '#1b9e77', 'Both' = '#cc50dc', 'IUCN' = '#d95f02')) +
-    borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
-    scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
-    scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
-    labs(title = sprintf('IUCN species ID: %s', spp$iucn_sid), x = NULL, y = NULL) 
[…]

### 2016-03-30 case_study_mpa/3_mpa_spp_analysis.Rmd

@@
             color = 'grey30') +
+  scale_x_discrete(labels = c('5. IUCN 2015',
+                              '4. AM 2015 (0%) +\nIUCN 2015', 
+                              '3. AM 2015 (0%)', 
+                              '2. AM 2015 (50%)', 
+                              '1. AM 2014 (50%)')) +
   coord_flip() +
-  labs(x = 'scenario', 
+  labs(x = 'Scenario', 
        y = 'Percent of species by MPA coverage', 

### 2016-03-30 ms_2_intro_res_disc.Rmd

@@
 Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps (1) and International Union for Conservation of Nature (IUCN) (2).
-These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status (3, 4, 5), evaluating global biodiversity patterns (6, 7, 8, 9), predicting range shifts (10), and setting conservation priorities (11).
+These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status (3, 4, 5), evaluating global biodiversity patterns (6, 7, 8, 9), predicting species range shifts (10), and setting conservation priorities (11).
 
@@
 
-To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of species.
-Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).
+To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of marine species.
+Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).
 For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and outline possible improvements.
@@
 The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of "probability of occurrence" for each species (14, 15).
-Studies frequently define "presence" for AquaMaps by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence (3, 4).
+Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence (3, 4).
 
@@
 While species numbers in both datasets peak in tropical latitudes near the equator, species counts for IUCN maps drop quickly beyond 30°N and 30°S, while species counts for AquaMaps remain robust well into temperate latitudes.
-The longitude frequency plots show a slight bias in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
+The longitude frequency plots show a slight shift in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
 To achieve more comprehensive global coverage of species ranges these two datasets can be used together, but the underlying methodological differences complicate such direct comparisons.
@@
 
-The upper right quadrant includes species (n = `r n_q1`) whose described ranges are above average in both spatial distribution and extent.
+The upper right quadrant includes species (n = `r n_q1`) whose described ranges are above average in alignment of both spatial distribution and area.
 These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig.
@@
 S3).
- Ocean depth is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps.
+ Ocean depth preference is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps.
 Simply clipping IUCN range maps to known depth preferences would resolve many of these mismatches.
@@
 
-Method-driven differences in commission and omission errors drive clear and significant differences in species range descriptions between AquaMaps and IUCN datasets.
+Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.
 To examine the implications of these differences, we repeated two recent studies, varying only the prioritization of one data set over the other.
@@
 
-Given that only `r p_spp_both_ohi`% of species overlap, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in substantial country-level score shifts as seen in scenarios 2 and 3.
+Given that only `r p_spp_both_ohi`% of species are included in both datasets, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in substantial country-level score shifts as seen in scenarios 2 and 3.
 While the mean global score did not vary significantly from scenario to scenario, select countries gained up to 7.5 points while others dropped as many as 5 points.
-This result indicates that especially on a national or regional scale, an arbitrary change in how the two datasets are combined can result in a very different assessment of species conservation status.
+This result indicates that especially on a national or regional scale, an arbitrary change in how the two datasets are combined can result in a different assessment of species conservation status.
 
@@
 AquaMaps and IUCN range maps show reasonable agreement for many well-studied species, but substantial differences illustrate uncertainty in our understanding of spatial distribution for many others.
-Although many other approaches exist for species distribution modeling, these two are the only ones applied globally to marine species.
+Although many other approaches exist for species distribution modeling, these two are the only ones applied globally across different marine taxa.
 Identifying and addressing differences in these datasets will increase their utility for research and conservation actions.

### 2016-03-30 ms_3_methods.Rmd

@@
 
-Overlaying paired distribution maps for a given species, we defined and calculated _distribution alignment_ and _area alignment_ and plotted these in Fig. 2A:
+Overlaying paired distribution maps for a given species, we defined and calculated _distribution alignment_ and _area ratio_ and plotted these in Fig. 2A:
 

### 2016-03-31 data_explore/data_explore.Rmd
added code to ditch IUCN risk colors

@@
                            aes(x = area_ratio, 
-                               y = sm_perc, 
+                               y = sm_perc)) + #, 
  #                               size = lg_area_pct,
-                               color = popn_category)) + 
+ #                              color = popn_category)) + 
   ggtheme_plot + 
@@
              fill= "#d01c8b") + 
-  geom_point(alpha = .6)
+  geom_point(color = 'grey40', alpha = .6)
 
@@
 scatter_quadplot <- scatter_quadplot +
-  scale_colour_manual(values = c('LC' = 'green4', 
-                                 'NT' = 'greenyellow', 
-                                 'VU' = 'yellow', 
-                                 'EN' = 'orange', 
-                                 'CR' = 'orangered2',
-                                 'DD' = 'grey30')) + 
+#   scale_colour_manual(values = c('LC' = 'green4', 
+#                                  'NT' = 'greenyellow', 
+#                                  'VU' = 'yellow', 
+#                                  'EN' = 'orange', 
+#                                  'CR' = 'orangered2',
+#                                  'DD' = 'grey30')) + 
 #   scale_size_continuous(limits = c(0, 100),
@@
   labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
-       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
+       y = bquote('Distribution alignment '*alpha['dist']*' (%)')) #, 
 #       title = 'Distribution alignment vs Extent alignment',
 #       size = '% ocean range',
-       color = 'Extinction\nRisk')
+ #      color = 'Extinction\nRisk')
 
@@
 
-ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
+ggsave(file.path(dir_fig, 'scatter_quadmap_norisk.png'),
        width = 17.8, height = 9, units = 'cm')
@@
 
+``` {r risk quads linear model}
+
+spp_risk_df <- spp_list_quads %>%
+  select(iucn_sid, spp_group = spp_group_text, popn_category, category_score, quad) %>%
+  mutate(quad_good = ifelse(quad == 'q1', 1, 0))
+
+risk_mdl1 <- glm(as.factor(quad_good) ~ category_score + spp_group, 
+                 family = binomial,
+                 data = spp_risk_df)
+
+risk_mdl2 <- glm(as.factor(quad_good) ~ category_score, 
+                 family = binomial,
+                 data = spp_risk_df)
+
+risk_mdl3 <- glm(as.factor(quad_good) ~ spp_group, 
+                 family = binomial,
+                 data = spp_risk_df)
+
+BIC(risk_mdl1, risk_mdl2, risk_mdl3)
+summary(risk_mdl1)
+summary(risk_mdl2)
+summary(risk_mdl1)
+
+binomial logistic regression model
+
+# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q1'])
+# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q2'])
+# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q3'])
+# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q4'])
+
+```
+
 -----

### 2016-03-31 data_explore/data_explore_quadmaps.Rmd
added code to ditch IUCN risk colors

@@
 
-```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-### this file created in data_prep
-spp_list <- read.csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE)
-
-```
-
 Quadrant map and quad breakdown by taxa and risk
@@
 ``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+spp_list <- read.csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE)
+
+spp_list <- spp_list %>% 
+  mutate(popn_category = ifelse(popn_category == 'LC', 'Least concern', popn_category),
+         popn_category = ifelse(popn_category == 'NT', 'Near threatened', popn_category),
+         popn_category = ifelse(popn_category == 'VU', 'Vulnerable', popn_category),
+         popn_category = ifelse(popn_category == 'EN', 'Endangered', popn_category),
+         popn_category = ifelse(popn_category == 'CR', 'Critically endangered', popn_category),
+         popn_category = ifelse(popn_category == 'DD', 'Data deficient', popn_category)) %>%
+  transform(popn_category = reorder(popn_category, category_score))
 
@@
 ### mongo plot time
-scatter_quadplot <- ggplot(spp_list %>% 
-                             transform(popn_category = reorder(popn_category, category_score)),
+scatter_quadplot <- ggplot(spp_list,
                            aes(x = area_ratio, 
@@
              fill= "#d01c8b") + 
-  geom_point(alpha = .6)
+  geom_point(alpha = .5)
 
@@
 scatter_quadplot <- scatter_quadplot +
-  scale_colour_manual(values = c('LC' = 'green4', 
-                                 'NT' = 'greenyellow', 
-                                 'VU' = 'yellow', 
-                                 'EN' = 'orange', 
-                                 'CR' = 'orangered2',
-                                 'DD' = 'grey30')) + 
+  scale_colour_manual(values = c('Least concern' = 'green4', 
+                                 'Near threatened' = 'greenyellow', 
+                                 'Vulnerable' = 'yellow', 
+                                 'Endangered' = 'orange', 
+                                 'Critically endangered' = 'orangered2',
+                                 'Data deficient' = 'grey30')) + 
   scale_x_continuous(expand = c(0, 0), 
@@
             fontface = 'bold.italic',
-            size = 3,
+            size = 2.5,
             hjust = .75,
@@
        y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
-       color = 'Extinction\nRisk')
+       color = 'Extinction Risk')
 

### 2016-04-02 manuscript-main.Rmd
submit this bastard

@@
 At least three keywords are required at submission.
-maybe species distribution maps, AquaMaps, IUCN, marine protected area planning, biodiversity status--> TBD: These should be words NOT in the title or abstract (BSH comment)
+maybe species distribution maps, AquaMaps, IUCN, marine protected area planning, biodiversity status--> Marine spatial planning, species distribution model, extent of occurrence, spatial ecology
 

### 2016-04-02 ms_2_intro_res_disc.Rmd
submit this bastard

@@
 
-## How and why the datasets differ
-
 The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" (2) - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.
 In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of "probability of occurrence" for each species (14, 15).
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of "probability of occurrence" for each species (14, 15).
 Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence (3, 4).
@@
 Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range (12).
-By comparing maps resulting from IUCN and AquaMaps methodolog, we can identify and possibly address mechanistic causes for each type of error.
+By comparing maps resulting from IUCN and AquaMaps methodologies, we can identify and possibly address mechanistic causes for each type of error.
 
@@
 This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their ranges span nearly the entire map.
-We expected to see significant concurrence between maps for IUCN Red List species of concern, assuming that increased attention would lead to improved understanding of species ranges; however, we found no compelling patterns to indicate that this was the case (Fig.
-S2).
 
@@
 Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, introducing additional commission errors.
-Consequently, area ratios are close to 100%, though similar areas are meaningless when the distributions are poorly aligned.
+Consequently, area ratios are close to 100%, though similar areas are unhelpful when the distributions are poorly aligned.
 
@@
 The global Ocean Health Index (OHI) (3, 4), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
-To maximize the number of represented species, OHI gleans spatial distributions from Red List-assessed species in both IUCN and AquaMaps datasets (n = `r format(n_spp_ohi, big.mark = ',')`), prioritizing IUCN data for the `r format(n_spp_both_ohi, big.mark = ',')` species included in both sources.
+To maximize the number of represented species, OHI gleans spatial distribution data from Red List-assessed species in both IUCN and AquaMaps datasets (n = `r format(n_spp_ohi, big.mark = ',')`), prioritizing IUCN data for the `r format(n_spp_both_ohi, big.mark = ',')` species included in both sources.
 OHI currently uses a probability of occurrence threshold of 40% to determine species presence for AquaMaps data.
@@
 Klein et al.
-(11) compare the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect species and identify which species fall through gaps in protection.
+(11) compare the global distribution of species to the global distribution of marine protected areas to assess how well current MPAs overlap with species ranges and identify which species fall through gaps in protection.
 The study relied on the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
@@
 
-We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using 2015 AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig4'>Fig.
+We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using the most recent AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig4'>Fig.
 4</a>).

### 2016-04-02 ms_3_methods.Rmd
submit this bastard

@@
 To avoid double-counting, we removed subpopulations and species aliases.
-We determined species presence within each spatial cell using the same criteria as outlined above.
+We determined species presence within each spatial cell using the same criteria outlined above.
 
@@
 We visually inspected a random selection of paired distribution maps from each quadrant to identify possible mechanistic causes of misalignment.
-To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS (19) to overlay a selection of IUCN and AquaMaps maps with a 200 meter bathymetry contour.
+To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS (19) to overlay a selection of IUCN and AquaMaps maps with a 200-meter bathymetry contour.
 
-**Methods for OHI case study: **
+**Methods for OHI case study:**
 Using methods and supplemental materials from OHI (3, 4), we modified the original code for OHI 2015 Species status (SPP), allowing for flexibility in prioritized data source and AquaMaps presence threshold.

### 2016-04-02 ms_supplemental.Rmd
submit this bastard

@@
 ---
-title: "Supplemental Information"
+title: "Supplemental Information: Aligning marine species range data to better serve science and conservation"
+author: "Casey O'Hara, Jamie C. Afflerbach, Courtney Scarborough, Benjamin S. Halpern"
 output:
@@
 -->
+<!--
+# Data sources and methods
 
-<!--## Data processing - should this go in methods??? -->
+All code is publicly available and freely downloadable [here on GitHub](link to github).
 
-<!--
-All code is publicly available and freely downloadable [here on GitHub](link to github).
+## AquaMaps
 
@@
 
+## IUCN
+
 As of December 2015, IUCN had published species distribution maps for XXX marine species across 24 taxonomic groups.
@@
 
-Each shapefile contains polygon limits of distribution indicating regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.-->
+Each shapefile contains spatial polygons outlining limits of distribution for each species, indicating regions of presence/absence; additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
+-->
 
-## S1: sample maps from each quadrant
+## Representative maps from each quadrant
 
@@
 
->  __Fig. S1.__ Representative species maps to illustrate each quadrant from Fig. 2A.  Note that A-D are arranged to match quadrants in Fig. 2A.
-(A: well-matched) Kajikia albida
-(B: distribution-matched) Conus episcopatus
-(C: area-matched) Conus magnificus
-(D: poorly-matched) Acanthurus nigroris
+>  __Fig.
+S1.__ Representative species maps to illustrate each quadrant from Fig.
+2A.
+Note that B-E are arranged to match quadrants in Fig.
+S1A.
+(A) Alternative version of Figure 2A that includes extinction risk categories for each species.
+Four representative species, one from each quadrant, are mapped in (B-E).
+(B) Well-matched: Kajikia albida, the Atlantic white marlin.  Ranges from each data set show nearly complete overlap, and very similar area of coverage.
+(C) Distribution-matched: Conus episcopatus, the dignified cone snail.  Ranges shows excellent overlap in the western Pacific, though IUCN range extends beyond the bounds of the AquaMaps range.
+(D) Area-matched: Conus magnificus, the magnificent cone snail.  Ranges overlap in the southern Pacific, but align poorly elsewhere.  The total area for each range is similar.
+(E) Poorly-matched: Acanthurus nigroris, the blue-lined surgeonfish.
+IUCN predicts species range only near the Hawaiian islands; AquaMaps predicts extensive range throughout the central and western Pacific Ocean.
+The datasets align in neither distribution nor area of coverage.
 
-## S2: Risk by quadrant
+## Risk by quadrant
 
@@
 >  __Fig.
-S2.__ Breaking down the quadrants by IUCN extinction risk categories (Fig.
-S2), we found little support for our hypothesis that maps for species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
+S2.__ Breakdown of quadrants (Fig.
+S1A) by IUCN extinction risk categories.
+IUCN and Aquamaps ranges were in better agreement (i.e., had a higher probability of being in the upper-right, "well-aligned" quadrant, Figure 2A/Figure S1A) for species at greater risk according to IUCN risk categories (p < 0.001, generalized linear model, family = binomial, Table S1).
+This suggests at-risk species may receive more attention in terms of data collection or model refinement.
 
-### S3: Coral depth map
+> __Table S1.__  Generalized linear model (family = binomial) describing how the agreement between IUCN and Aquamap range maps (i.e., probability of occurring in quadrant 1, Figure 2A/S1A) increases for a species as its IUCN risk status increases.
+For this analysis, we excluded all species that had "data deficient" IUCN status (n = 255).
+We analyzed the IUCN risk categories as a continuous variable (LC = 0.0, NT = 0.2, VU = 0.4, EN = 0.6, CR = 0.8, EX = 1.0).
+
+_Model: glm(quadrant 1 membership ~ IUCN category, family = binomial) n = 2,059_
+
+| Coefficients:  | Estimate | Std. Error | z value | p value      |
+| :-----         |   -----: |     -----: |  -----: |       -----: |
+| Intercept      | -1.32933 |   0.05804  | -22.902 | < 2e-16 ***  |
+| IUCN category  |  1.18672 |   0.33469  |   3.546 | 0.000391 *** |
+
+### Coral depth map
 
@@
 
-<!--
-Oculina varicosa - Ivory tree coral ![oculina_pic.jpg](figures/oculina/oculina_pic.jpg)
--->
+>  __Fig.
+S3.__ IUCN-defined range of _Oculina varicosa_, a typical photosynthetic coral, plotted against a 200 m bathymetry contour.
+According to IUCN, _O.
+varicosa_ can be found to depths of 152 m.
+Maps of many other coral species and other reef-associated organisms followed similar patterns.
 
->  __Fig. S3.__ IUCN-defined range of _Oculina varicosa_, a typical photosynthetic coral, plotted against a 200 m bathymetry contour. According to IUCN, _O. varicosa_ can be found to depths of 152 m.
-
-### S4: AquaMaps threshold examination
+### AquaMaps threshold examination
 
@@
 Many studies convert this AquaMaps probability to a simple presence value by assigning a threshold value.
-A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
 
-For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to best approximate the "limits of distribution" as indicated by IUCN maps.
 To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0% to 100% and calculated the average species range relative to a zero threshold.
@@
 S4.__ AquaMaps distribution map extent remaining after applying a presence threshold.
-(A) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of 28% of total range, with a wide distribution in which some species lose nearly all of their apparent range.
+A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
+(A) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of 28% of total range compared to a 0% threshold, with a wide distribution in which some species lose nearly all of their apparent range.
 (B) Median remaining extent at increments of presence threshold.
@@
 -->
-
-<!--
-___Red List inclusion:___
-
-![](figures/bar_by_taxa_assessed.png)-->
-
-<!--
-
-a few codes to indicate likely problems?
-
-- DC = depth clipping; IUCN extents go beyond reasonable depth for the species
-- DX = data excluded; AM excludes some observations that appear to be used for IUCN
-- DP = data poor (AquaMaps acknowledges data poor status? or use less than 10 data points as criteria)
-- NX = needs evaluation by expert; AM model likely predicts areas where species is unlikely to be found based on observed occurrences - does AM model reflect same as AM suitable habitat?
-
-Q2:
-
-| species                | quad | group           | codes   | notes |
-| -------                | ---- | -----           | -----   | ----- |
-| Conus episcopatus      | q2   | cones           |  DP, DX | d-match around Coral Triangle, but IUCN shows much more around Indian Ocean and S Pacific; also IUCN extents
-| Ctenochaetus binotatus | q2   |                 |         | IUCN extents beyond AM
-| Naso vlamingii         | q2   | acanthuridae    |         | IUCN extents beyond AM
-| Thalassoma purpureum   | q2   | wrasses         |         | IUCN extents beyond AM
-| Porites nigrescens     | q2   | corals 2        | DP      | AM limited to CT; IUCN shows far greater dist, and extents
-| Conus ammiralis        | q2   | cones           | DP      | AM  | limited to CT, IUCN shows in Madagascar; also extents
-| Conus tessulatus       | q2   | cones           |         | extents
-| Acropora sarmentosa    | q2   | corals 1        |         | extents; also AM shows HI but IUCN does not?
-| Holothuria fuscogilva  | q2   | sea cucumbers   |         | extents; also AM shows Mex and Central Am west coast, but IUCN does not (not available online?)
-| Oculina Varicosa       | q2   | corals 2        |         | lim to Caribbean; extents - this might be a good close-up map
-| Acanthocybium solandri | q2   | tunas/billfish  |         | mostly tropical by IUCN; AM shows much farther N and S
-
-Q3
-
-| species                | quad | group           | codes   | notes |
-| -------                | ---- | -----           | -----   | ----- |
-| Conus magnificus       | q3   |  cones          |         | IUCN distribution is more limited than AM distribution)
-| Abudefduf concolor     | q3   |  damselfish     |         | Baja for AM, southern Central Am for IUCN; maybe a good closeup map?
-AM points from Baja to Peru incl Galapagos; some rejected pts from Caribbean and eastern Atlantic
-| Centropyge aurantonotus | q3  |  angelfish      |         | AM shows Caribbean; IUCN shows northern South America, also maybe a good closeup map?
-AM points from FL to south Brazil, but no Gulf of Mexico.
-| Chlorurus perspicillatus | q3 |  parrotfish     |         | IUCN shows only HI; AM shows dots all over S Pacific and near Marianas Islands
-| Sarda orientalis        | q3  |  tunas/billfish |         | limited dists by IUCN
-| Montastraea franksi     | q3  |  corals 2       | DP      | IUCN shows just in Caribbean; AM shows all up & down east coast of N America and S America
-
-Q4
-
-| species                  | quad | group   | codes   | notes |
-| -------                  | ---- | -----   | -----   | ----- |
-| Acanthurus nigroris      | q4   |         |         | IUCN dist limited to HI, while AM shows all over S Pacific and CT - AM observations all over S pacific...
-| Praealticus tanegasimae  | q4   | blenny  |         | IUCN shows limited dist around Japan/Marianas; AM shows much broader through CT and W Pacific - good local map?
-| Stethojulis marquesensis | q4   | wrasse  |         | IUCN shows only small cluster of islands (Marquesas?) while AM shows much broader across S Pacific
-| Canthigaster leoparda    | q4   | puffer  |         | IUCN limited spots around a couple of islands; AM shows all throughout CT
-| pentacheles snyderi      | q4   | lobster | DP      | AM shows all over - sim to depth profiles?
-while IUCN shows only a couple patches - HI?
-S Pacific, and S Indian oceans?
-AM limits to specific FAO regions, which results in odd cutoffs; also shows some data points in N Atlantic which has an odd result.
-| Nephropsis sulcata       | q4   | lobster |         | IUCN shows odd ranges, most of which don't overlap with AM
-
--->
-
-<!--
-___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot___ _- perhaps update the quadrant plot to reveal data-poor species (and reviewed species?
-little overlap of reviewed & data-poor) instead of, say, red-list category which doesn't get discussed in the body of the paper?
-then this table can go in SOM if we like it_
-
-| quadrant | n species | n data poor species | mean (median) data points | n reviewed species |
-| :------: | :-------: | :-----------------: | :-----------------------: | :----------------: |
-| all AM   |   22889   |        8749 (38.2%) |          57.1 (16)        |       1296  (5.7%) |
-| AM&IUCN  |    2166   |         457 (21.1%) |          89.9 (33)        |        290 (13.4%) |
-|   q1     |     401   |          33  (8.2%) |         233.0 (78)        |        100 (24.9%) |
-|   q2     |     682   |         151 (22.1%) |          77.4 (39)        |        100 (14.7%) |
-|   q3     |     682   |         114 (16.5%) |          52.4 (29)        |         65  (9.5%) |
-|   q4     |     410   |         159 (39.7%) |          32.1 (13)        |         25  (6.2%) |
-
--->

### 2016-05-13 manuscript-main-v2.Rmd
added zotero functionality; preparing for submission to ecology letters

--- 2016-04-02 manuscript-main.Rmd
@@
     toc: no
+  pdf_document:
+    number_sections: no
+    toc: no
+  word_document: default
+bibliography: refs/zotero_lib.bib
+csl: refs/ecology-letters.csl
 ---
@@
 
-``` {r intro_results_discussion, child = 'ms_2_intro_res_disc.Rmd'}
+``` {r intro, child = 'ms_2_intro.Rmd'}
+```
+
+``` {r methods, child = 'ms_3_methods.Rmd'}
+```
+
+``` {r results_discussion, child = 'ms_2a_res_disc.Rmd'}
 ```
@@
 
-``` {r methods, child = 'ms_3_methods.Rmd'}
-```
-
------
-
 ``` {r references, child = 'ms_4_refs.Rmd'}
-```
-
------
-
-``` {r figure_legends, child = 'ms_5_fig_legends.Rmd'}
 ```

### 2016-05-13 ms_2a_res_disc.Rmd
added zotero functionality; preparing for submission to ecology letters

--- 2016-04-02 ms_2_intro_res_disc.Rmd
@@
 
-# Introduction
-
-Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation, among many others.  This knowledge provides foundational information for 
- understanding species ranges and diversity,
- predicting species responses to human impacts and climate change,
- and managing and protecting species effectively.
-A rich literature tackles the many dimensions of these questions.
-
-One major outcome of this body of science is the various compiled databases of species distribution maps.
-Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps (1) and International Union for Conservation of Nature (IUCN) (2).
-These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status (3, 4, 5), evaluating global biodiversity patterns (6, 7, 8, 9), predicting species range shifts (10), and setting conservation priorities (11).
-
-<!--
-Three foundational papers relating to AquaMaps (Kaschner et al. 2006, 2011; Ready et al. 2010) have been cited a total of 315 times.
--->
-
-The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
-Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
-Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans (12, 13).
-
-To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of marine species.
-Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).
-For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and outline possible improvements.
-
-We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (3, 4) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (11) - as case studies to explore the implications of prioritizing one data set over the other.
-The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
-
 # Results and Discussion
 
-The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" (2) - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.
+The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" [@iucn_iucn_2015] - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.
 In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of "probability of occurrence" for each species (14, 15).
-Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence (3, 4).
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of "probability of occurrence" for each species [@kaschner_mapping_2006; @ready_predicting_2010].
+Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
 
@@
 Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.
-Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range (12).
+Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006].
 By comparing maps resulting from IUCN and AquaMaps methodologies, we can identify and possibly address mechanistic causes for each type of error.
@@
 The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor area match well.
-Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) (16) and the Global Biodiversity Information Facility (GBIF) (17)) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
+Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
 When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species.
@@
 
-The global Ocean Health Index (OHI) (3, 4), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
+The global Ocean Health Index (OHI) [@halpern_index_2012; @halpern_patterns_2015], a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
 To maximize the number of represented species, OHI gleans spatial distribution data from Red List-assessed species in both IUCN and AquaMaps datasets (n = `r format(n_spp_ohi, big.mark = ',')`), prioritizing IUCN data for the `r format(n_spp_both_ohi, big.mark = ',')` species included in both sources.
@@
 Klein et al.
-(11) compare the global distribution of species to the global distribution of marine protected areas to assess how well current MPAs overlap with species ranges and identify which species fall through gaps in protection.
+[-@klein_shortfalls_2015] compare the global distribution of species to the global distribution of marine protected areas to assess how well current MPAs overlap with species ranges and identify which species fall through gaps in protection.
 The study relied on the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
@@
 For example, for most corals, researchers could clip range maps to the photosynthetic limit of 200 meters.
- For AquaMaps range data, dependent primarily on environmental and physical preferences and conditions, implementing area restrictions based on biogeographical criteria such as Marine Ecoregions of the World (18) would likely decrease commission errors and improve predictive power, especially for data-poor species.
+ For AquaMaps range data, dependent primarily on environmental and physical preferences and conditions, implementing area restrictions based on biogeographical criteria such as Marine Ecoregions of the World [@spalding_marine_2007] would likely decrease commission errors and improve predictive power, especially for data-poor species.
  For either data set, maps based on few occurrences are more likely to include errors of commission and omission.

### 2016-05-13 ms_3_methods.Rmd
added zotero functionality; preparing for submission to ecology letters

@@
 We visually inspected a random selection of paired distribution maps from each quadrant to identify possible mechanistic causes of misalignment.
-To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS (19) to overlay a selection of IUCN and AquaMaps maps with a 200-meter bathymetry contour.
+To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS [@qgis_development_team_qgis_2015] to overlay a selection of IUCN and AquaMaps maps with a 200-meter bathymetry contour.
 
 **Methods for OHI case study:**
-Using methods and supplemental materials from OHI (3, 4), we modified the original code for OHI 2015 Species status (SPP), allowing for flexibility in prioritized data source and AquaMaps presence threshold.
+Using methods and supplemental materials from OHI [@halpern_index_2012; @halpern_patterns_2015], we modified the original code for OHI 2015 Species status (SPP), allowing for flexibility in prioritized data source and AquaMaps presence threshold.
 We ran the SPP code three times, prioritizing IUCN over AquaMaps for a 0% threshold, and prioritizing AquaMaps over IUCN for both a 40% and 0% threshold.
@@
 Based upon the methods described in Klein et al.
-(11), we reconstructed the analysis using the subset of protected areas from the World Database on Protected Areas (WDPA) (20) spatially covering a marine area and classified as IUCN I-IV.
+[-@klein_shortfalls_2015], we reconstructed the analysis using the subset of protected areas from the World Database on Protected Areas (WDPA) [@iucn_world_2014] spatially covering a marine area and classified as IUCN I-IV.
 The WDPA polygons and marine polygons were rasterized to 0.01° and then aggregated to AquaMaps native 0.5° cells, to calculate proportion of protected area and marine within each cell.
@@
 
-All processing was completed using R statistical software (21), and all code and intermediate data are available on GitHub at https://github.com/OHI-Science/IUCN-AquaMaps.
+All processing was completed using R statistical software [@r_core_team_r:_2016], and all code and intermediate data are available on GitHub at https://github.com/OHI-Science/IUCN-AquaMaps.

### 2016-05-13 ms_4_refs_manual.Rmd
added zotero functionality; preparing for submission to ecology letters

--- 2016-03-27 ms_5_refs.Rmd
@@
 # Acknowledgments
+
 <!-- Acknowledgments. List acknowledgments and funding sources. PNAS collaborates with FundRef to use a standard taxonomy of funder names. Dedications are rarely allowed. -->
+
+We are thankful to Melanie Frazier and Julia Stewart Lowndes for their insightful comments on earlier versions of this manuscript, and to the National Center for Ecological Analysis and Synthesis for computation support.
+We acknowledge financial support from the Gordon and Betty Moore Foundation.
+
+-----
 
@@
 
+1.
 Kaschner K et al.
@@
 
+2.
 IUCN (2015).
@@
 
-Halpern BS et al. (2012). An index to assess the health and benefits of the global ocean. Nature, 488(7413), 615-620. <!--_doi:10.1038/nature11397._-->
+3. Halpern BS et al. (2012). An index to assess the health and benefits of the global ocean. Nature, 488(7413), 615-620. <!--_doi:10.1038/nature11397._-->
 
-Halpern BS et al. (2015). Patterns and emerging trends in global ocean health. PloS one, 10(3), e0117863. <!--___doi?___-->
+4. Halpern BS et al. (2015). Patterns and emerging trends in global ocean health. PloS one, 10(3), e0117863. <!--___doi?___-->
 
-Klein CJ et al. (2015). Shortfalls in the global protected area network at representing marine biodiversity. Scientific Reports 5: 17539. <!--_doi:10.1038/srep17539._-->
+5. Selig ER et al. (2013). Assessing global marine biodiversity status within a coupled socio-ecological perspective. PloS one, 8(4), e60284. <!--_doi:10.1371/journal.pone.0060284._-->
 
-Selig ER et al. (2013). Assessing global marine biodiversity status within a coupled socio-ecological perspective. PloS one, 8(4), e60284. <!--_doi:10.1371/journal.pone.0060284._-->
+6. Coll M et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. <!--_used AquaMaps to predict Med biodiversity.  Also: Threshold = 0._-->
 
-Coll M et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. <!--_used AquaMaps to predict Med biodiversity.  Also: Threshold = 0._-->
-
-*Martin CS et al.
+7.
+Martin CS et al.
 (2014).
@@
 
+8.
 Pimm SL et al.
@@
 
+9.
 Kaschner K, Tittensor DP, Ready J, Gerrodette T, Worm B (2011).
@@
 
-Molinos JG et al. (2015). Climate velocity and the future global redistribution of marine biodiversity. Nature Climate Change. <!--_doi:10.1038/nclimate2769._-->
+10. Molinos JG et al. (2015). Climate velocity and the future global redistribution of marine biodiversity. Nature Climate Change. <!--_doi:10.1038/nclimate2769._-->
 
+11. Klein CJ et al. (2015). Shortfalls in the global protected area network at representing marine biodiversity. Scientific Reports 5: 17539. <!--_doi:10.1038/srep17539._-->
+
+12.
 Rondinini C, Wilson KA, Boitani L, Grantham H, & Possingham HP (2006).
@@
 
+13.
 Jetz W, Sekercioglu CH, and Watson JE (2008).
@@
 
+14.
 Kaschner K, Watson R, Trites AW, Pauly D (2006).
@@
 
+15.
 Ready J et al.
@@
 
-Spalding MD et al. (2007). Marine ecoregions of the world: a bioregionalization of coastal and shelf areas. BioScience, 57(7), 573-583.
+16. OBIS (2016). Data from the Ocean Biogeographic Information System. Intergovernmental Oceanographic Commission of UNESCO. Web. http://www.iobis.org (consulted on 2016/02/29).
 
-QGIS Development Team (2015). QGIS Geographic Information System. Open Source Geospatial Foundation Project. http://qgis.osgeo.org
+17. GBIF (2010). GBIF Memorandum of Understanding. Copenhagen: Global Biodiversity Information Facility, 20 pp. Accessible online at http://www.gbif.org/resource/80661.
 
-IUCN and UNEP-WCMC (2014). The World Database on Protected Areas (WDPA) [On-line]. Cambridge, UK: UNEP- WCMC. Available at: www.protectedplanet.net [Accessed December 1, 2014].
+18. Spalding MD et al. (2007). Marine ecoregions of the world: a bioregionalization of coastal and shelf areas. BioScience, 57(7), 573-583.
 
-R Core Team (2015). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
+19. QGIS Development Team (2015). QGIS Geographic Information System. Open Source Geospatial Foundation Project. http://qgis.osgeo.org
+
+20. IUCN and UNEP-WCMC (2014). The World Database on Protected Areas (WDPA) [On-line]. Cambridge, UK: UNEP- WCMC. Available at: www.protectedplanet.net [Accessed December 1, 2014].
+
+21. R Core Team (2015). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

### 2016-05-13 testbib.Rmd
added zotero functionality; preparing for submission to ecology letters

--- 2016-05-13 manuscript-main-v2.Rmd
@@
 
-<!--Title: Titles should be no more than three typeset lines (generally 135 characters including spaces) and should be comprehensible to a broad scientific audience.
-Please provide a short title of fewer than 50 characters for mobile devices and RSS feeds.-->
+All processing was completed using R statistical software [@r_core_team_r:_2016], and all code and intermediate data are available on GitHub at https://github.com/OHI-Science/IUCN-AquaMaps.
 
-_Author affiliation:_ <!--Include department, institution, and complete address, with the ZIP/postal code, for each author.
-Use superscripts to match authors with institutions.
-Authors are encouraged to supply their ORCID ID at submission-->
+evaluating global biodiversity patterns [@kaschner_current_2011].
 
-1. National Center for Ecological Analysis and Synthesis, 735 State Street Suite 300, Santa Barbara CA 93101
-2. Bren School of Environmental Science and Management, University of California, Santa Barbara CA 93106
-3. Department of Life Sciences, Imperial College London, Silwood Park Campus, Buckhurst Rd, Ascot, West Berkshire SL5 7PY, United Kingdom
+[@halpern_index_2012]
 
-_Corresponding author:_ <!--Corresponding author: The name, complete address, telephone number, and email address of the author to whom correspondence and proofs should be sent.
-Email addresses will appear in the article footnotes.-->
-Casey O'Hara, National Center for Ecological Analysis and Synthesis, 735 State Street, Suite 300, Santa Barbara CA 93101; (805) 892-2500; ohara@nceas.ucsb.edu
+[@jones_modelling_2012]
 
-_Classification:_ BIOLOGICAL SCIENCES: Sustainability Science
+[@selig_assessing_2013]
 
-_Keywords:_ <!--Keywords are listed below the abstract of the published article.
-At least three keywords are required at submission.
-maybe species distribution maps, AquaMaps, IUCN, marine protected area planning, biodiversity status--> Marine spatial planning, species distribution model, extent of occurrence, spatial ecology
-
-<!--
-
-from PNAS (http://www.pnas.org/site/authors/preparation.xhtml):
-
-PNAS generally uses a two-column format averaging 67 characters, including spaces, per line.
-The maximum length of a regular research article is 6 pages (or 10 pages for PNAS Plus articles), including all text, spaces, and the number of characters displaced by figures, tables, and equations.
-
-An online submission tool provides authors with an estimation of whether their manuscript fits within the PNAS length requirements (see Length Estimate FAQ).
-Use the following guidelines to determine whether your manuscript exceeds the page-length requirements before submission:
-
-Text: When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces).
-Figures: Calculated at 180 characters per cm in height for one column and 360 characters per cm in height for two columns; a figure wider than 8.7 cm is two columns wide.
-Tables and equations: Calculated at 60 characters per line for one column and 120 per line for two columns; a table with more than 60 characters per line is two columns wide.
-If the text in a figure or table is deemed illegible for production, the figure or table may be enlarged to improve legibility.
-Add the word count to the character count to obtain a character count that includes spaces.
-
-See also: http://www.pnascentral.org/html/pnas_length_estimate_instructions.html
-
--->
-
-``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
-
-# library(ggplot2)
-# library(maps)
-
-# library(raster)
-# library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-library(readr)
-
-# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-#            'Darwin'  = '/Volumes/data_edit',
-#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-# dir_anx  <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
-
-if(basename(getwd()) != 'IUCN-AquaMaps') setwd('~/github/IUCN-AquaMaps')
-
-dir_data <- 'data'
-dir_fig  <- 'figures'
-
-```
-
-``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.  This
-### filters out presence = Extinct from IUCN cells.
-
-data_file <- 'data/spp_list_w_area_trimmed.csv'
-spp_map_pairs <- read_csv(data_file)
-### NOTE: as of Jan 31, includes all map pairs, even ones where IUCN area == 0
-
-### generate other lists
-spp_all_file <- 'data/spp_all.csv'
-spp_all <- read_csv(spp_all_file) ### NOTE: contains duped scinames
-
-# iucn_marine_file <- 'data/spp_iucn_marine.csv'
-# spp_iucn_marine <- read_csv(iucn_marine_file)
-
-iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read_csv(iucn_maps_file) %>%
-  filter(!duplicated(iucn_sid))
-
-spp_am_file <- 'data/spp_am.csv'
-spp_am <- read_csv(spp_am_file)
-
-```
-
-``` {r calc species numbers, echo = FALSE}
-n_spp_am     <- nrow(spp_am)
-n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
-n_spp_iucn   <- length(unique(spp_iucn_maps$iucn_sid))
-n_spp_both   <- length(unique(spp_map_pairs$iucn_sid))
-n_spp_all    <- n_spp_am + n_spp_iucn - n_spp_both
-
-p_spp_am     <- round(n_spp_am/n_spp_all * 100, 1)
-p_spp_iucn   <- round(n_spp_iucn/n_spp_all * 100, 1)
-p_both_am    <- round(n_spp_both/n_spp_am   * 100, 1)
-p_both_iucn  <- round(n_spp_both/n_spp_iucn * 100, 1)
-p_both_all   <- round(n_spp_both/n_spp_all  * 100, 1)
-
-spp_list <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
-area_align_mean <- round(mean(spp_list$area_ratio, na.rm = TRUE), 1)
-dist_align_mean <- round(mean(spp_list$sm_perc, na.rm = TRUE), 1)
-
-```
------
-
-``` {r abstract, child = 'ms_1_abst_sig.Rmd'}
-```
-
-<!--
-| Section        |   Word #   | Char/spc # |
-| :------        |    ------: |    ------: |
-| Abstract       |  184 words |  1312 char |
-| Significance   |   73 words |   545 char |
--->
-
------
-
-\\body
-
-``` {r intro, child = 'ms_2_intro.Rmd'}
-```
-
-``` {r methods, child = 'ms_3_methods.Rmd'}
-```
-
-``` {r results_discussion, child = 'ms_2a_res_disc.Rmd'}
-```
-
-<!--
-| Section                   |   Word #   | Char/spc # |
-| :------                   |    ------: |    ------: |
-| Introduction              |  429 words |  2905 char |
-| Results/Discussion        |  999 words |  6697 char |
-| Implications (incomplete) |  597 words |  3794 char |
-| Conclusions               |  354 words |  2505 char |
-| Methods (incomplete)      |  601 words |  3919 char |
-| total body                | 2987 words | 19885 char |
--->
-
------
-
-``` {r references, child = 'ms_4_refs.Rmd'}
-```
-
-<!--
-| Figure  |  w x h (px) | w x h (cm)  | char # |
-| :------ |     ------: |     ------: |------: |
-| 1 A - C | 2102 x 1536 | 17.8 x 13.0 |   4680 |
-| caption |             |             |    367 | 
-| 2 A - B | 2102 x 1940 | 17.8 x 16.5 |   5940 |
-| caption |             |             |    934 | 
-|       3 | 1027 x  531 |  8.7 x  4.5 |    810 |
-| caption |             |             |    744 | 
-|       4 | 1346 x  590 | 11.4 x  5.0 |   1800 |
-| caption |             |             |    870 | 
-|  total  |             |             |  16145 |
--->
+# REFERENCES

### 2016-05-17 data_explore/data_explore_am_thresh.Rmd
updating manuscript to ecology letters standards

@@
 ggsave(file.path(dir_fig, 'si_am_threshold.tif'), device = 'tiff', 
-       width = 11.4, height = 6, units = 'cm')
+       width = 11.4, height = 6, units = 'cm', dpi = 600)
 ggsave(file.path(dir_fig, 'si_am_threshold.png'), 
-       width = 11.4, height = 6, units = 'cm')
+       width = 11.4, height = 6, units = 'cm', dpi = 600)
 
@@
 ggsave(file.path(dir_fig, 'si_am_threshold2.tif'), device = 'tiff', 
-       width = 17.8, height = 9, units = 'cm')
+       width = 17.8, height = 9, units = 'cm', dpi = 600)
 ggsave(file.path(dir_fig, 'si_am_threshold2.png'), 
-       width = 17.8, height = 9, units = 'cm')
+       width = 17.8, height = 9, units = 'cm', dpi = 600)
 
@@
 ggsave(file.path(dir_fig, 'si_ribbon_am_threshold.tif'), device = 'tiff', 
-       width = 11.4, height = 6, units = 'cm')
+       width = 11.4, height = 6, units = 'cm', dpi = 600)
 ggsave(file.path(dir_fig, 'si_ribbon_am_threshold.png'), 
-       width = 11.4, height = 6, units = 'cm')
+       width = 11.4, height = 6, units = 'cm', dpi = 600)
 

### 2016-05-17 manuscript-main-v2.Rmd
updating manuscript to ecology letters standards

@@
   S. Halpern^1,2,3^
+csl: refs/ecology-letters.csl
 output:
+  word_document: default
+  pdf_document:
+    number_sections: no
+    toc: no
   html_document:
@@
     toc: no
-  pdf_document:
-    number_sections: no
-    toc: no
-  word_document: default
 bibliography: refs/zotero_lib.bib
-csl: refs/ecology-letters.csl
 ---
@@
 
-``` {r abstract, child = 'ms_1_abst_sig.Rmd'}
+``` {r abstract, child = 'ms_1_abst_150wd.Rmd'}
 ```

### 2016-05-17 manuscript-main.Rmd
updating manuscript to ecology letters standards

@@
   S. Halpern^1,2,3^
+csl: refs/ecology-letters.csl
 output:
@@
     toc: no
+bibliography: refs/zotero_lib.bib
 ---
@@
 
-``` {r intro_results_discussion, child = 'ms_2_intro_res_disc.Rmd'}
+``` {r intro, child = 'ms_2_intro.Rmd'}
+```
+
+``` {r results_discussion, child = 'ms_2a_res_disc.Rmd'}
 ```
@@
 
------
-
-``` {r figure_legends, child = 'ms_5_fig_legends.Rmd'}
-```
-
 <!--

### 2016-05-17 data_explore/data_explore.Rmd

@@
 
+``` {r pct_overlap_vs_ratio_of_area2, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+### define 25%, 50%, and 75% quartiles for both x and y axes
+area_align_mean <- mean(spp_list$area_ratio, na.rm = TRUE)
+dist_align_mean <- mean(spp_list$sm_perc, na.rm = TRUE)
+
+spp_list_quads <- read_csv(file.path(dir_data, 'spp_list_quads.csv'))
+
+ggtheme_plot2 <- ggtheme_basic + 
+  theme(panel.border     = element_blank(),
+        panel.grid.minor = element_blank(), 
+#        panel.grid.major = element_line(colour = 'grey90'),
+        panel.grid.major = element_blank(),
+        axis.title.x = element_text(hjust = 0),
+        axis.title.y = element_text(hjust = 1),
+        axis.line = element_blank())
+
+### mongo plot time
+scatter_quadplot <- ggplot(spp_list_quads %>% 
+                             transform(popn_category = reorder(popn_category, category_score)) %>%
+                             mutate(quad = factor(quad, levels = c('q1', 'q2', 'q3', 'q4'))),
+                           aes(x = area_ratio, 
+                               y = sm_perc, 
+                               color = quad)) +
+  ggtheme_plot2
+
+### add mean lines (quadrant dividers) under the points
+scatter_quadplot <- scatter_quadplot +
+  geom_vline(xintercept = area_align_mean,
+             color      = 'grey80') +
+  geom_hline(yintercept = dist_align_mean,
+             color      = 'grey80')
+
+### add the points!
+scatter_quadplot <- scatter_quadplot +
+  geom_point(alpha = .8) +
+  scale_colour_manual(values = c('q1' = '#4dac26',
+                                 'q2' = '#b8e186',
+                                 'q3' = '#f1b6da',
+                                 'q4' = '#d01c8b')) +
+  scale_x_continuous(expand = c(0, 0), 
+                     limits = c(0, 102),
+                     breaks = c(seq(0, 100, 25)),
+                     labels = c('0%', '25%', '50%', '75%', '100%')) +
+  scale_y_continuous(expand = c(0, 0),
+                     limits = c(0, 103),
+                     breaks = c(seq(0, 100, 25)),
+                     labels = c('0%', '25%', '50%', '75%', '100%'))
+
+### add axis lines and mean labels:
+scatter_quadplot <- scatter_quadplot  +
+  geom_vline(xintercept = 0,
+             color      = 'grey80') +
+  geom_hline(yintercept = 0,
+             color      = 'grey80') +
+  annotate(geom = 'text',
+           x = (area_align_mean - .5), y = 5,
+           hjust = 0, vjust = 0,
+           color = 'grey30', size = 2,
+           fontface = 'bold.italic', angle = 90,
+           label = sprintf('Mean = %s%%', round(area_align_mean, 1))) +
+  annotate(geom = 'text',
+           x = 5, y = (dist_align_mean + .5),
+           hjust = 0, vjust = 0,
+           color = 'grey30', size = 2,
+           fontface = 'bold.italic', angle = 0,
+           label = sprintf('Mean = %s%%', round(dist_align_mean, 1)))
+
+### quadrant labels
+scatter_quadplot <- scatter_quadplot +
+  annotate(geom = 'text', color = 'grey30', size = 3, fontface = 'bold.italic',
+           x = 2, y = 2,
+           hjust = 0, vjust = 0,
+           label = 'Poorly aligned') +
+  annotate(geom = 'text', color = 'grey30', size = 3, fontface = 'bold.italic',
+           x = 98, y = 98,
+           hjust = 1, vjust = 1,
+           label = 'Well aligned') +
+  annotate(geom = 'text', color = 'grey30', size = 3, fontface = 'bold.italic',
+           x = 2, y = 98,
+           hjust = 0, vjust = 1,
+           label = 'Distribution aligned') +
+  annotate(geom = 'text', color = 'grey30', size = 3, fontface = 'bold.italic',
+           x = 98, y = 2,
+           hjust = 1, vjust = 0,
+           label = 'Area aligned')
+
+scatter_quadplot <- scatter_quadplot +
+  labs(x = bquote('Area ratio '*alpha['ext']*''), 
+       y = bquote('Distribution alignment '*alpha['dist']*'')) +
+  theme(legend.position = 'none')
+
+# print(scatter_quadplot)
+
+ggsave(file.path(dir_fig, 'scatter_quadmap_colbyquad.png'),
+       width = 17.8, height = 9, units = 'cm')
+
+```
+
 ![](../figures/scatter_quadmap.png)

### 2016-05-17 data_explore/data_explore.Rmd
including dpi = 600 for plotting

--- 2016-03-31 data_explore/data_explore.Rmd
@@
 ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
-       width = 17.8, height = 7, units = 'cm')
+       width = 17.8, height = 7, units = 'cm', dpi = 600)
 
@@
 # ggsave(filename = file.path(dir_fig, 'si_bar_by_taxa_assessed.tif'), device = 'tiff', 
-#        width = 17.8, height = 6, units = 'cm')
+#        width = 17.8, height = 6, units = 'cm', dpi = 600)
 
@@
 ggsave(file.path(dir_fig, 'scatter_quadmap_norisk.png'),
-       width = 17.8, height = 9, units = 'cm')
+       width = 17.8, height = 9, units = 'cm', dpi = 600)
 
@@
        filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
-       width = 17.8, height = 7, units = 'cm')
+       width = 17.8, height = 7, units = 'cm', dpi = 600)
 
@@
        filename = file.path(dir_fig, 'si_barchart_spp_risk_quads.tif'), device = 'tiff',
-       width = 17.8, height = 4.5, units = 'cm')
+       width = 17.8, height = 4.5, units = 'cm', dpi = 600)
 ggsave(plot = barchart_spp_risk_quads,
        filename = file.path(dir_fig, 'si_barchart_spp_risk_quads.png'),
-       width = 17.8, height = 4.5, units = 'cm')
+       width = 17.8, height = 4.5, units = 'cm', dpi = 600)
 

### 2016-05-18 case_study_mpa/3_mpa_spp_analysis.Rmd
updated to Mazu to rerun the figures.

@@
 
-dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-           'Darwin'  = '/Volumes/data_edit',
-           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
+dir_M             <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
+                       'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
+                       'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]
 
@@
 dir_data <- file.path(dir_git, 'data')
-dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')
+dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/d2015/csv')
 
@@
        filename = file.path(dir_fig, 'fig4_barchart_mpa_gap.png'),
-       width = 11.4, height = 5, units = 'cm')
+       width = 11.4, height = 5, units = 'cm', dpi = 600)
 ggsave(plot = barchart_mpa_gap,
        filename = file.path(dir_fig, 'fig4_barchart_mpa_gap.tif'), device = 'tiff',
-       width = 11.4, height = 5, units = 'cm')
+       width = 11.4, height = 5, units = 'cm', dpi = 600)
 

### 2016-05-18 case_study_ohi/boxplots.R

@@ -75,16 +75,18 @@ bplot_abs <- ggplot(data = st_all %>% filter(str_detect(scenario, 'abs')),
 #         plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
 #         legend.position = 'none') +
   geom_hline(yintercept = 0, color = 'gray50', size = .5) +
-  geom_violin(fill = 'grey80', draw_quantiles = c(.25, .5, .75), 
+  geom_violin(fill = 'grey80', 
+              draw_quantiles = c(.25, .5, .75), 
               alpha = .5) +
   geom_point(position = position_jitter(w = .3), 
-             aes(color = scenario), 
+             # aes(color = scenario), 
+             color = 'grey40',
              alpha = .5,
              show.legend = TRUE) +
 #  stat_summary(fun.y=mean, geom="point", color = 'red', size = 3) +
-  scale_color_manual(values = c('absdiff1' = '#1b9e77', 
-                                'absdiff2' = '#565080', 
-                                'absdiff3' = '#d95f02')) +
+  # scale_color_manual(values = c('absdiff1' = '#1b9e77', 
+  #                               'absdiff2' = '#565080', 
+  #                               'absdiff3' = '#d95f02')) +
   
   geom_violin(fill = NA, draw_quantiles = c(.25, .5, .75)) +
   #  geom_boxplot(size = .75, fill = NA) +
@@ -103,9 +105,9 @@ y <- x %>%
 
 bplot_abs
 ggsave(file.path(dir_fig, 'fig3_boxplot_abs_diff.png'), 
-       height = 4.5, width = 8.7, units = 'cm')
+       height = 4.5, width = 8.7, units = 'cm', dpi = 600)
 ggsave(file.path(dir_fig, 'fig3_boxplot_abs_diff.tif'), device = 'tiff',
-       height = 4.5, width = 8.7, units = 'cm')
+       height = 4.5, width = 8.7, units = 'cm', dpi = 600)
 
 # bplot_rel <- ggplot(data = st_all %>% filter(str_detect(scenario, 'rel')),
 #                     aes(x = scenario, y = difference)) +

### 2016-05-18 case_study_mpa/3_mpa_spp_analysis.Rmd
updating figs to 600 dpi for ecology letters

@@
 
-dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
+dir_anx <- file.path(dir_M, 'git-annex/globalprep/SPP_ICO')
 dir_fig <- file.path(dir_git, 'figures')
 dir_data <- file.path(dir_git, 'data')
-dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/d2015/csv')
+dir_am_data <- file.path(dir_M, 'git-annex/globalprep/_raw_data/aquamaps/d2015/csv')
 
@@
 ### compare results using only IUCN species maps
-dir_iucn_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/iucn_spp')
-iucn_cells_file <- file.path(dir_iucn_data, 'iucn_cells_v2016.csv')
-iucn_spp_cells <- read_csv(iucn_cells_file, col_types = 'cddddc') %>%
-  as.data.frame(stringsAsFactors = FALSE)
+dir_iucn_data <- file.path(dir_M, 'git-annex/globalprep/_raw_data/iucn_spp')
+dir_iucn_data <- file.path(dir_M, 'git-annex/globalprep/spp_ico')
+iucn_cells_file <- file.path(dir_iucn_data, 'vAM_IUCN/int/iucn_cells_spp1.csv')
+iucn_spp_cells <- read_csv(iucn_cells_file, col_types = 'cddddc')
+write_csv(iucn_spp_cells, iucn_cells_file)
+x <- read_csv(file.path(dir_iucn_data, 'v2016/int/spp_all_cleaned.csv')) %>%
+  filter(spp_group != 'BOTW' & !is.na(spp_group))
+
+iucn_spp_cells1 <- iucn_spp_cells %>% filter(iucn_sid %in% x$iucn_sid)
 

### 2016-05-18 data_explore/data_explore.Rmd
updating figs to 600 dpi for ecology letters

@@
 
-dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
+dir_anx <- file.path(dir_M, 'git-annex/globalprep/SPP_ICO')
 dir_fig <- file.path(dir_git, 'figures')
@@
 spp_am <- spp_am %>%
-  left_join(read_csv(file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv/speciesoccursum.csv')) %>% 
+  left_join(read_csv(file.path(dir_M, 'git-annex/globalprep/_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>% 
               dplyr::select(am_sid = speciesid, class) %>%
@@
              fill= "#d01c8b") + 
-  geom_point(color = 'grey40', alpha = .6)
+  geom_point(color = '#4d4dac', alpha = .6)
 

### 2016-05-18 data_explore/data_explore_am_thresh.Rmd
updating figs to 600 dpi for ecology letters

@@
 # get spp_cells raw data for aquamaps
-dir_aquamaps <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015')
+dir_aquamaps <- file.path(dir_M, 'git-annex/globalprep/_raw_data/aquamaps/d2015')
 
@@
 ```{r df_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE, eval = TRUE}
-### create dataframe that gives me all species, their total number of cells, 
-### total number of cells >= 0.4 prob, difference between the two and threat
 
@@
 for (i in seq(0, 1, .05)) { # i = .05
+  message('Calculating species cell losses for threshold = ', i, '...')
   temp_df <- spp_cells %>%
@@
     left_join(spp %>%
-#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')), 
-              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
+             filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')),
+#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
             by = c('speciesid')) %>%
@@
 
+message('Summarizing all down to quantiles, medians, and means...')
 thresh_df_summary <- threshold_df %>%
@@
 
-  geom_ribbon(aes(ymax = quant95, ymin = quant05), alpha = .5, color = NA, fill = '#b8e186') +
-  geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = '#4dac26') +
+  geom_ribbon(aes(ymax = quant95, ymin = quant05), alpha = .5, color = NA, fill = '#b8b8e1') +
+  geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = '#4d4dac') +
   geom_line(color = 'grey30') + 

### 2016-05-18 case_study_ohi/boxplots.R

@@ -14,7 +14,7 @@ library(stringr)
 library(ggplot2)
 library(readr)
 
-dir_data  <- '~/github/ohiprep/globalprep/SPP_ICO/vAM_IUCN'
+dir_data  <- '~/github/ohiprep/globalprep/spp_ico/vAM_IUCN'
 dir_fig <- '~/github/IUCN-AquaMaps/figures'
 
 ### generic theme for all plots
@@ -80,7 +80,7 @@ bplot_abs <- ggplot(data = st_all %>% filter(str_detect(scenario, 'abs')),
               alpha = .5) +
   geom_point(position = position_jitter(w = .3), 
              # aes(color = scenario), 
-             color = 'grey40',
+             color = '#4d4dac',
              alpha = .5,
              show.legend = TRUE) +
 #  stat_summary(fun.y=mean, geom="point", color = 'red', size = 3) +
@@ -97,11 +97,11 @@ bplot_abs <- ggplot(data = st_all %>% filter(str_detect(scenario, 'abs')),
                               '2. AquaMaps/IUCN\n40% threshold', 
                               '3. AquaMaps/IUCN\nno threshold'))
 
-x <- st_all %>% filter(str_detect(scenario, 'abs'))
-y <- x %>% 
-  group_by(scenario, r1) %>% 
-  summarize(mean_diff = mean(difference, na.rm = TRUE),
-            n_rgn = n())
+# x <- st_all %>% filter(str_detect(scenario, 'abs'))
+# y <- x %>% 
+#   group_by(scenario, r1) %>% 
+#   summarize(mean_diff = mean(difference, na.rm = TRUE),
+#             n_rgn = n())
 
 bplot_abs
 ggsave(file.path(dir_fig, 'fig3_boxplot_abs_diff.png'), 

### 2016-06-15 manuscript-main-j_app_ecol.Rmd
tweakin a few things to clean up

--- 2016-05-13 manuscript-main-v2.Rmd
@@
   S. Halpern^1,2,3^
+csl: refs/journal-of-applied-ecology.csl
 output:
+  word_document: default
+  pdf_document:
+    number_sections: no
+    toc: no
   html_document:
@@
 
-``` {r abstract, child = 'ms_1_abst_sig.Rmd'}
+``` {r abstract, child = 'ms_1_abst_j_appl_ecol.Rmd'}
 ```

### 2016-07-30 coralmaps/coral_maps.Rmd
setting up for in-depth coral analysis

--- 2016-05-18 data_explore/data_explore.Rmd
@@
 
-```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-### this file created in data_prep
-spp_list <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
+```{r get_bathy_raster, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+### get shapefile
+### rasterize to half-degree grid; or raster::extract
+
+###
 
 ```
-
-# Figures 1 a b c
-
-``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_prep.Rmd to create this file, if it does not yet exist.
-
-data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps
-
-### generate other lists
-
-spp_all_file <- file.path(dir_data, 'spp_all.csv')
-spp_all <- read_csv(spp_all_file)
-
-iucn_file <- file.path(dir_data, 'spp_iucn.csv')
-spp_iucn <- read.csv(iucn_file, stringsAsFactors = FALSE)
-
-message('Getting IUCN maps list')
-iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps.csv')
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
-  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
-  filter(sciname %in% spp_iucn$sciname)   ### ditch freshwater spp
-
-spp_am_file <- file.path(dir_data, 'spp_am.csv')
-spp_am <- read_csv(spp_am_file)
-
-```
-
-``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
-
-sciname_lookup <- read.csv(file.path(dir_data, 'sciname_lookups/spp_iucn_scinames.csv'), stringsAsFactors = FALSE)
-
-spp_am <- spp_am %>%
-  left_join(read_csv(file.path(dir_M, 'git-annex/globalprep/_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>% 
-              dplyr::select(am_sid = speciesid, class) %>%
-              unique(), 
-            by = 'am_sid')
-
-iucn_maps_class <- spp_iucn_maps %>% 
-  left_join(sciname_lookup %>% 
-              dplyr::select(sciname, class) %>%
-              unique(), 
-            by = 'sciname')
-
-taxa_iucn_all <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
-  filter(!is.na(class)) %>% ### don't count the unknowns
-  mutate(class = str_to_title(class),
-         n_iucn_tot = n()) %>%        ### length of whole maps list
-  group_by(class) %>%
-  summarize(n_iucn_class_all  = n(),
-            p_iucn_class_all  = round((n_iucn_class_all/first(n_iucn_tot) * 100), 2))
-
-taxa_iucn_solo <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
-  filter(!is.na(class)) %>% ### don't count the unknowns
-  filter(!sciname %in% spp_am$sciname) %>%  ### filter list to IUCN only
-  mutate(class = str_to_title(class),
-         n_iucn_solo = n()) %>%       ### length of IUCN only list
-  group_by(class) %>%
-  summarize(n_iucn_class_solo = n(),
-            p_iucn_class_solo = round((n_iucn_class_solo/first(n_iucn_solo) * 100), 2))
-
-taxa_am_all <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  mutate(class = str_to_title(class),
-         n_am_tot = n()) %>%        ### length of whole AM maps list
-  group_by(class) %>%
-  summarize(n_am_class_all  = n(),
-            p_am_class_all  = round((n_am_class_all/first(n_am_tot) * 100), 2))
-
-taxa_am_solo <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
-  mutate(class = str_to_title(class),
-         n_am_solo = n()) %>%       ### length of AM only list
-  group_by(class) %>%
-  summarize(n_am_class_solo = n(),
-            p_am_class_solo = round((n_am_class_solo/first(n_am_solo) * 100), 2))
-
-taxa_both <- spp_am %>%
-  filter(sciname %in% spp_iucn_maps$sciname) %>%
-  filter(!class %in% c('Not assigned', 'null')) %>%
-  mutate(class = str_to_title(class),
-         n_both_tot = n()) %>%
-  group_by(class) %>%
-  summarize(n_class_both = n(),
-            p_class_both = round(n_class_both/first(n_both_tot) * 100, 2))
-
-taxa_am_all_assessed <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(sciname %in% spp_iucn$sciname) %>%
-  mutate(class = str_to_title(class),
-         n_am_tot = n()) %>%        ### length of whole AM maps list
-  group_by(class) %>%
-  summarize(n_am_class_ass  = n(),
-            p_am_class_ass  = round((n_am_class_ass/first(n_am_tot) * 100), 2))
-
-taxa_am_solo_assessed <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(sciname %in% spp_iucn$sciname) %>%
-  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
-  mutate(class = str_to_title(class),
-         n_am_solo = n()) %>%       ### length of AM only list
-  group_by(class) %>%
-  summarize(n_am_class_solo_ass = n(),
-            p_am_class_solo_ass = round((n_am_class_solo_ass/first(n_am_solo) * 100), 2))
-
-taxa_df <- taxa_am_all %>%
-  full_join(taxa_am_solo,   by = 'class') %>%
-  full_join(taxa_am_all_assessed,  by = 'class') %>%
-  full_join(taxa_am_solo_assessed, by = 'class') %>%
-  full_join(taxa_iucn_all,  by = 'class') %>%
-  full_join(taxa_iucn_solo, by = 'class') %>%
-  full_join(taxa_both,      by = 'class') %>%
-  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))
-
-# knitr::kable(taxa_df %>%
-#                dplyr::select(class, 
-#                              AquaMaps = n_am_class_all, 
-#                              IUCN = n_iucn_class_all, 
-#                              Both = n_class_both))
-```
-
-``` {r plot bar charts by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
-### bar charts comparing number of species in each class for each data set
-
-# taxa_df_long1 <- taxa_df %>%
-#   dplyr::select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both) %>%
-#   gather(dataset, n_spp, AquaMaps:Both)
-# bar_taxa1 <- ggplot(taxa_df_long1, aes(x = class, y = n_spp)) +
-#   ggtheme_plot + 
-#   geom_bar(stat = 'identity') +
-#   facet_grid(dataset ~ ., scales = 'free_y') +
-#   theme(axis.text.x = element_text(angle = 75, hjust = 1))
-# print(bar_taxa1)
-
-#library(colorspace)
-#hclvec <- heat_hcl(3); colvec <- c('AquaMaps' = hclvec[1], 'Both' = hclvec[2], 'IUCN' = hclvec[3])
-#colvec <- c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')
-colvec <- c('AquaMaps' = '#76A94C', 'Both' = '#e18258', 'IUCN' = '#9168a9')
-
-taxa_df_long2 <- taxa_df %>%
-  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
-  gather(dataset, n_spp, AquaMaps:Both) %>%
-  group_by(class) %>%
-  mutate(n_total = sum(n_spp, na.rm = TRUE),
-         percent = n_spp/n_total * 100,
-         dataset = factor(dataset, levels = c('AquaMaps', 'Both', 'IUCN'), ordered = TRUE)) %>%
-  arrange(desc(dataset))
-
-bar_taxa2 <- ggplot(taxa_df_long2, aes(x = class, y = percent, fill = dataset)) +
-  ggtheme_plot + 
-  theme(panel.grid.major.x = element_blank()) +
-  geom_bar(stat = 'identity', alpha = .8) +
-  scale_fill_manual(values = colvec) +
-  geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
-            size = 2, 
-            angle = 90, hjust = 0, color = 'grey30') +
-  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
-  scale_y_continuous(expand = c(0, 0), 
-                     limits = c(0, 130),
-                     breaks = c(0, 25, 50, 75, 100),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) + 
-  labs(# title = 'Species maps by taxonomic class',
-       x = 'Class', 
-       y = 'Percent of mapped species',
-       fill = 'Dataset')
-
-print(bar_taxa2)
-ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
-       width = 17.8, height = 7, units = 'cm', dpi = 600)
-
-# taxa_df_long3 <- taxa_df %>%
-#   dplyr::select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
-#   gather(dataset, n_spp, AquaMaps:Both) %>%
-#   group_by(class) %>%
-#   mutate(n_total = sum(n_spp, na.rm = TRUE),
-#          percent = n_spp/n_total * 100)
-#
-# bar_taxa3 <- ggplot(taxa_df_long3, aes(x = class, y = percent, fill = dataset)) +
-#   ggtheme_plot + 
-#   geom_bar(stat = 'identity', alpha = .5) +
-#   scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
-#   geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
-#             #size = 3, 
-#             angle = 90, hjust = 0, color = 'grey30') +
-#   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
-#   scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
-#   labs(# title = 'Species maps by taxonomic class, Red List-assessed only',
-#        x = 'Class', 
-#        y = 'Percent of assessed mapped species')
-# print(bar_taxa3)
-# ggsave(filename = file.path(dir_fig, 'si_bar_by_taxa_assessed.tif'), device = 'tiff', 
-#        width = 17.8, height = 6, units = 'cm', dpi = 600)
-
-```
-
-![](../figures/bar_by_taxa_all.png)
-
-``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
-```
-
-# Figure 2a b
-
-## Quadrant map
-
-``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-
-### define 25%, 50%, and 75% quartiles for both x and y axes
-area_align_mean <- mean(spp_list$area_ratio, na.rm = TRUE)
-dist_align_mean <- mean(spp_list$sm_perc, na.rm = TRUE)
-
-### mongo plot time
-scatter_quadplot <- ggplot(spp_list %>% 
-                             transform(popn_category = reorder(popn_category, category_score)),
-                           aes(x = area_ratio, 
-                               y = sm_perc)) + #, 
- #                               size = lg_area_pct,
- #                              color = popn_category)) + 
-  ggtheme_plot + 
-  ### color the quadrant backgrounds:
-    annotate("rect", xmin = area_align_mean, xmax = 100, 
-                     ymin = dist_align_mean, ymax = 100, 
-             alpha = .3, 
-             fill= "#4dac26")  + 
-    annotate("rect", xmax = area_align_mean, xmin =   0, 
-                     ymin = dist_align_mean, ymax = 100, 
-             alpha = .3, 
-             fill= "#b8e186") + 
-    annotate("rect", xmin = area_align_mean, xmax = 100, 
-                     ymax = dist_align_mean, ymin =   0, 
-             alpha = .3, 
-             fill= "#f1b6da") + 
-    annotate("rect", xmax = area_align_mean, xmin =   0, 
-                     ymax = dist_align_mean, ymin =   0, 
-             alpha = .3, 
-             fill= "#d01c8b") + 
-  geom_point(color = '#4d4dac', alpha = .6)
-
-### Manage scales for color and size 
-scatter_quadplot <- scatter_quadplot +
-#   scale_colour_manual(values = c('LC' = 'green4', 
-#                                  'NT' = 'greenyellow', 
-#                                  'VU' = 'yellow', 
-#                                  'EN' = 'orange', 
-#                                  'CR' = 'orangered2',
-#                                  'DD' = 'grey30')) + 
-#   scale_size_continuous(limits = c(0, 100),
-#                         breaks = c( 0,    25,    50,    75,    100), 
-#                         labels = c('0%', '25%', '50%', '75%', '100%')) + 
-  scale_x_continuous(expand = c(0, 0), 
-                     limits = c(0, 102),
-                     breaks = c(seq(0, 100, 25)),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) +
-  scale_y_continuous(expand = c(0, 0),
-                     limits = c(0, 103),
-                     breaks = c(seq(0, 100, 25)),
-                     labels = c('0%', '25%', '50%', '75%', '100%'))
-
-### here are mean labels:
-scatter_quadplot <- scatter_quadplot +
-  annotate(geom = 'text',
-           x = area_align_mean, y = 5,
-           hjust = 0, vjust = 0,
-           color = 'grey30', 
-           size = 2,
-           fontface = 'bold.italic', angle = 90,
-           label = sprintf('Mean = %s%%', round(area_align_mean, 1))) +
-  annotate(geom = 'text',
-           x = 5, y = dist_align_mean,
-           hjust = 0, vjust = 0,
-           color = 'grey30', 
-           size = 2,
-           fontface = 'bold.italic', angle = 0,
-           label = sprintf('Mean = %s%%', round(dist_align_mean, 1)))
-
-scatter_quadplot <- scatter_quadplot +
-  labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
-       y = bquote('Distribution alignment '*alpha['dist']*' (%)')) #, 
-#       title = 'Distribution alignment vs Extent alignment',
-#       size = '% ocean range',
- #      color = 'Extinction\nRisk')
-
-# print(scatter_quadplot)
-
-ggsave(file.path(dir_fig, 'scatter_quadmap_norisk.png'),
-       width = 17.8, height = 9, units = 'cm', dpi = 600)
-
-```
-
-![](../figures/scatter_quadmap.png)
-
-## Break down quads by species group
-
-``` {r examine_spp_gp_quartiles, echo = FALSE, eval = TRUE}
-
-spp_list1 <- spp_list %>%
-  group_by(spp_group_text) %>%
-  mutate(n_spp = n()) %>%
-  ungroup()
-
-### Create species lists by quadrant
-spp_list_q1 <- spp_list1 %>%
-  filter(area_ratio >= mean(area_ratio) & sm_perc >= mean(sm_perc))
-spp_list_q2 <- spp_list1 %>%
-  filter(area_ratio < mean(area_ratio) & sm_perc >= mean(sm_perc))
-spp_list_q3 <- spp_list1 %>%
-  filter(area_ratio >= mean(area_ratio) & sm_perc < mean(sm_perc))
-spp_list_q4 <- spp_list1 %>%
-  filter(area_ratio < mean(area_ratio) & sm_perc < mean(sm_perc))
-
-spp_list_quads <- spp_list1 %>%
-  mutate(quad = NA,
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q1$sciname, 'q1', quad),
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q2$sciname, 'q2', quad),
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q3$sciname, 'q3', quad),
-         quad = ifelse(is.na(quad) & sciname %in% spp_list_q4$sciname, 'q4', quad))
-write_csv(spp_list_quads, file.path(dir_data, 'spp_list_quads.csv'))
-
-### Summarize quadrant species list by species group (e.g. CORALS1)
-spp_gp_q1 <- spp_list_q1 %>%
-  group_by(spp_group_text) %>%
-  summarize(n_spp = first(n_spp), n_spp_q1 = n())
-spp_gp_q2 <- spp_list_q2 %>%
-  group_by(spp_group_text) %>%
-  summarize(n_spp = first(n_spp), n_spp_q2 = n())
-spp_gp_q3 <- spp_list_q3 %>%
-  group_by(spp_group_text) %>%
-  summarize(n_spp = first(n_spp), n_spp_q3 = n())
-spp_gp_q4 <- spp_list_q4 %>%
-  group_by(spp_group_text) %>%
-  summarize(n_spp = first(n_spp), n_spp_q4 = n())
-
-### Join species group by quadrant into a single dataframe
-spp_gp_quadrants <- spp_gp_q1 %>%
-  full_join(spp_gp_q2, by = c('spp_group_text', 'n_spp')) %>%
-  full_join(spp_gp_q3, by = c('spp_group_text', 'n_spp')) %>%
-  full_join(spp_gp_q4, by = c('spp_group_text', 'n_spp')) %>%
-  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
-  mutate(quad = str_replace(quad, 'n_spp_', ''),
-         pct_quad = n_quad/n_spp,
-         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
-         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))
-
-spp_gp_quadrants <- spp_gp_quadrants %>%
-  left_join(spp_gp_quadrants %>% 
-              filter(quad == 'q1') %>%
-              dplyr::select(spp_group_text, pct_q1 = pct_quad),
-            by = 'spp_group_text') %>%
-  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
-         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
-         quad = str_replace(quad, 'q3', 'q3 area-matched'),
-         quad = str_replace(quad, 'q4', 'q4 poorly matched'))
-
-### Plot the bar chart
-barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
-                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
-                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
-  theme(panel.grid.major.y = element_blank()) +
-  geom_bar(stat = 'count', alpha = .8) +
-  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
-                               'q2 dist-matched'   = '#b8e186', 
-                               'q3 area-matched'   = '#f1b6da', 
-                               'q4 poorly matched' = '#d01c8b')) +
-#   scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
-#                                'q2 dist-matched'   = 'greenyellow', 
-#                                'q3 area-matched'   = 'orange', 
-#                                'q4 poorly matched' = 'orangered2')) +
-  scale_y_continuous(expand = c(0, 0), 
-                     limits = c(0, 1.1),
-                     breaks = c(0, .25, .5, .75, 1.0),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
-            size = 2, 
-            color = 'grey30') +
-  coord_flip() +
-  labs(x = 'Taxonomic Group', 
-       y = 'Percent of species by quadrant', 
-#       title = 'Taxonomic group by quadrant',
-       fill = 'Alignment')
-
-# print(barchart_spp_gp_quads)
-ggsave(plot = barchart_spp_gp_quads,
-       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
-       width = 17.8, height = 7, units = 'cm', dpi = 600)
-
-write_csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads.csv'))
-
-```
-
-![](../figures/barchart_spp_gp_quads.png)
-
-# Plots for SI
-
-## Break down quadrants by risk category
-
-``` {r examine_risk_quartiles, echo = FALSE, eval = TRUE}
-
-### Summarize quadrant species list by species group (e.g. CORALS1)
-spp_risk_q1 <- spp_list_q1 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp_q1 = n())
-spp_risk_q2 <- spp_list_q2 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp_q2 = n())
-spp_risk_q3 <- spp_list_q3 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp_q3 = n())
-spp_risk_q4 <- spp_list_q4 %>%
-  group_by(popn_category) %>%
-  summarize(n_spp_q4 = n())
-
-### Join species group by quadrant into a single dataframe
-spp_risk_quadrants <- spp_risk_q1 %>%
-  full_join(spp_risk_q2, by = c('popn_category')) %>%
-  full_join(spp_risk_q3, by = c('popn_category')) %>%
-  full_join(spp_risk_q4, by = c('popn_category')) %>%
-  gather(quad, n_spp_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
-  group_by(popn_category) %>%
-  mutate(n_tot_cat = sum(n_spp_quad)) %>%
-  ungroup() %>%
-  mutate(quad = str_replace(quad, 'n_spp_', ''),
-         pct_quad   = n_spp_quad/n_tot_cat,
-         pct_quad   = ifelse(is.na(pct_quad),   0, pct_quad))
-
-spp_risk_quadrants1 <- spp_risk_quadrants %>%
-  left_join(spp_risk_quadrants %>% 
-              filter(quad == 'q1') %>%
-              dplyr::select(popn_category, pct_q1 = pct_quad),
-            by = 'popn_category') %>%
-  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
-         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
-         quad = str_replace(quad, 'q3', 'q3 area-matched'),
-         quad = str_replace(quad, 'q4', 'q4 poorly matched')) %>%
-  left_join(spp_list %>%
-              dplyr::select(popn_category, cat_rank = category_score) %>%
-              unique() %>%
-              mutate(cat_rank = ifelse(popn_category %in% c('DD', 'NE'), -1, cat_rank)),
-            by = 'popn_category')
-
-### Plot the bar chart
-barchart_spp_risk_quads <- ggplot(spp_risk_quadrants1 %>% 
-                                  transform(popn_category = reorder(popn_category, cat_rank)), 
-                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
-  theme(panel.grid.major.y = element_blank()) +
-  geom_bar(stat = 'count', alpha = .8) +
-  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
-                               'q2 dist-matched'   = '#b8e186', 
-                               'q3 area-matched'   = '#f1b6da', 
-                               'q4 poorly matched' = '#d01c8b')) +
-  scale_y_continuous(expand = c(0, 0), 
-                     limits = c(0, 1.1),
-                     breaks = c(0, .25, .5, .75, 1.0),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) +
-  geom_text(aes(label = sprintf('n = %s', n_tot_cat), y = 1.01), hjust = 0, 
-            size = 2, 
-            color = 'grey30') +
-  coord_flip() +
-  labs(x = 'Risk category', 
-       y = 'Percent of species by quadrant', 
-#       title = 'Risk category by quadrant',
-       fill = 'Alignment')
-
-# print(barchart_spp_risk_quads)
-ggsave(plot = barchart_spp_risk_quads,
-       filename = file.path(dir_fig, 'si_barchart_spp_risk_quads.tif'), device = 'tiff',
-       width = 17.8, height = 4.5, units = 'cm', dpi = 600)
-ggsave(plot = barchart_spp_risk_quads,
-       filename = file.path(dir_fig, 'si_barchart_spp_risk_quads.png'),
-       width = 17.8, height = 4.5, units = 'cm', dpi = 600)
-
-```
-
-![](../figures/si_barchart_spp_risk_quads.png)
-
-``` {r risk quads linear model}
-
-spp_risk_df <- spp_list_quads %>%
-  select(iucn_sid, spp_group = spp_group_text, popn_category, category_score, quad) %>%
-  mutate(quad_good = ifelse(quad == 'q1', 1, 0))
-
-risk_mdl1 <- glm(as.factor(quad_good) ~ category_score + spp_group, 
-                 family = binomial,
-                 data = spp_risk_df)
-
-risk_mdl2 <- glm(as.factor(quad_good) ~ category_score, 
-                 family = binomial,
-                 data = spp_risk_df)
-
-risk_mdl3 <- glm(as.factor(quad_good) ~ spp_group, 
-                 family = binomial,
-                 data = spp_risk_df)
-
-BIC(risk_mdl1, risk_mdl2, risk_mdl3)
-summary(risk_mdl1)
-summary(risk_mdl2)
-summary(risk_mdl1)
-
-binomial logistic regression model
-
-# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q1'])
-# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q2'])
-# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q3'])
-# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q4'])
-
-```
-
------
-
-## AquaMaps threshold exploration
-
-``` {r, child = 'data_explore_am_thresh.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
-```
-
------
-
-## Count species occurrences as proxy for data poverty
-
-``` {r count_occurrences_by_quad, echo = FALSE, warning = FALSE}
-
-gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence.csv'))
-obis_occur <- read_csv(file.path(dir_data, 'obis_occurrence.csv'))
-
-totes_occur <- obis_occur %>%
-  left_join(gbif_occur) %>%
-  rowwise() %>%
-  mutate(mean_occ = mean(c(gbif_occ, obis_occ), na.rm = TRUE)) %>%
-  group_by(iucn_sid) %>%
-  summarize(mean_occ = mean(mean_occ, na.rm = TRUE),
-            tots_occ = sum(mean_occ, na.rm = TRUE))
-
-spp_quads_occ <- spp_list_quads %>%
-  left_join(totes_occur %>%
-              dplyr::select(iucn_sid, mean_occ),
-            by = 'iucn_sid')
-
-write_csv(spp_quads_occ, file.path(dir_data, 'spp_quads_occ.csv'))
-
-```
-
------
-
-# Obsolete plots; not evaluated
-
-* bar chart: number of spp per group
-* scatter plot: area compare IUCN vs AM, including log plot and plot by species group
-* rangemap: Chromis alpha
-
-``` {r child = 'data_explore_pt1.Rmd', eval = FALSE}
-```
-
------
-
-* percent overlap vs total area exploration
-
-``` {r, child = 'data_explore_pt2.Rmd', eval = FALSE}
-```
-
------
-
-* Multiple species maps per quadrant
-
-``` {r child = 'data_explore_pt3.Rmd', eval = FALSE}
-```
-
------

### 2016-07-30 coralmaps/coral_maps.Rmd
first pass analysis - assuming that removing deep cells does not affect AquaMaps (since AM is theoretically filtered to shallows as well)

@@
 library(readr)
-library(data.table)
 library(raster)
+library(rgdal)
 library(tidyr)
@@
 
-dir_anx <- file.path(dir_M, 'git-annex/globalprep/SPP_ICO')
+dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico')
 dir_fig <- file.path(dir_git, 'figures')
@@
 
-###
+if(!file.exists(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))) {
+
+  poly_bath <- readOGR(dsn = file.path(path.expand(dir_git), 'coralmaps/ne_10m_bathymetry_K_200'), 
+                       layer = 'ne_10m_bathymetry_K_200', 
+                       stringsAsFactors = TRUE)
+
+  ### LOICZID cell ID numbers - used in plotting rasters
+  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
+  loiczid_raster       <- raster(loiczid_raster_file)
+  names(loiczid_raster) <- 'loiczid'
+
+  bath_list <- raster::extract(loiczid_raster, poly_bath, 
+                             weights = TRUE, normalizeWeights = FALSE,
+                             progress = 'text')
+  bath_df <- lapply(bath_list, function(x) {as.data.frame(x)}) %>%
+    bind_rows() %>%
+    rename(loiczid = value) %>%
+    group_by(loiczid) %>%
+    summarize(weight = sum(weight))
+
+  ### generate raster?
+  bath_raster <- subs(loiczid_raster, bath_df, by = 'loiczid', which = 'weight')
+  values(bath_raster)[values(bath_raster) > 5] <- NA
+
+  write_csv(bath_df, file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))
+  ### probably want to filter to just 100% 200m+
+}
 
 ```
+
+``` {r}
+### find spp depth limits
+### focus just on corals?
+
+if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_depth_raw.csv'))) {
+
+  library(jsonlite)
+  api_key <- 'fb71ae836f415f04f41176f6d30c4a9e4cea620d46b9e5021bf2fb142ea51bf5'
+
+  get_from_api <- function(url, param) {
+      api_info <- fromJSON(sprintf(url, param, api_key)) %>%
+        data.frame(stringsAsFactors = FALSE)
+  }
+
+  mc_get_from_api <- function(url, param_vec) {
+    numcores = ifelse(Sys.info()[['nodename']] == 'mazu', 12, 1)
+    out_df <- parallel::mclapply(param_vec, 
+                            function(x) get_from_api(url, x),
+                            mc.cores = numcores) %>% 
+      bind_rows() 
+    out_df <- out_df %>%
+      setNames(names(.) %>%
+                 str_replace('result.', ''))
+  }
+
+  ### get species list
+  data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
+  spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
+    filter(!is.na(popn_category)) %>% ### filter out unassessed species
+    filter(!is.na(sm_perc))           ### filter out NA maps
+
+  coral_spp_sid <- spp_map_pairs %>%
+    filter(str_detect(spp_group, 'CORAL')) %>%
+    .$iucn_sid
+
+  spp_narr_url <- 'http://apiv3.iucnredlist.org/api/v3/species/narrative/id/%s?token=%s'
+  spp_narr     <- mc_get_from_api(spp_narr_url, coral_spp_sid)
+
+  write_csv(spp_narr, file.path(dir_git, 'coralmaps', 'coral_spp_narratives.csv'))
+
+  spp_depth <- spp_narr %>%
+    select(iucn_sid = species_id, habitat)
+  write_csv(spp_depth, file.path(dir_git, 'coralmaps', 'coral_spp_depth_raw.csv'))
+  ### manually find max depths from this
+}
+
+```
+
+``` {r}
+
+### load iucn spp_cells
+coral_files <- list.files(file.path(dir_anx, 'v2016', 'iucn_intersections'), 
+                          pattern = 'CORAL*.*',
+                          full.names = TRUE)
+iucn_coral_cells <- lapply(coral_files, read_csv)
+
+```

### 2016-07-31 coralmaps/coral_maps.Rmd
basics of coral depth-clipping analysis done

@@
 
+spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))
+bath_df <-  read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd') %>%
+  filter(weight >= 1) %>%
+  mutate(deep = 1) %>%
+  select(-weight)
+
 ### load iucn spp_cells
@@
                           full.names = TRUE)
-iucn_coral_cells <- lapply(coral_files, read_csv)
+iucn_coral_cells <- lapply(coral_files, read_csv) %>%
+  bind_rows() %>%
+  select(iucn_sid, loiczid = LOICZID) %>%
+  filter(iucn_sid %in% spp_depth$iucn_sid) %>%
+  left_join(bath_df, by = 'loiczid') %>%
+  mutate(deep = ifelse(is.na(deep), 0, deep))
+
+write_csv(iucn_coral_cells, file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv'))
 
 ```
+
+``` {r}
+### calc areas
+### read hcaf to get areas; left join to iucn_coral_cells
+### group_by iucn_sid and depth, summarize area (deep and shallow)
+### group_by iucn_sid, summarize total area
+iucn_coral_cells <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv'))
+
+hcaf <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
+  select(loiczid, cellarea, oceanarea)
+
+iucn_coral_areas <- iucn_coral_cells %>%
+  left_join(hcaf, by = 'loiczid') %>%
+  group_by(iucn_sid, deep) %>%
+  summarize(deeparea = sum(oceanarea)) %>%
+  group_by(iucn_sid) %>%
+  mutate(totarea = sum(deeparea),
+         shallow_pct = (totarea - deeparea)/totarea) %>%
+  filter(deep == 1) %>%
+  select(-deep)
+
+  ### get species list
+  data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
+  spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
+    filter(!is.na(popn_category)) %>%   ### filter out unassessed species
+    filter(!is.na(sm_perc))             ### filter out NA maps
+
+iucn_coral_areas <- iucn_coral_areas %>%
+  left_join(spp_depth,     by = 'iucn_sid') %>%
+  left_join(spp_map_pairs, by = 'iucn_sid')
+
+write_csv(iucn_coral_areas, file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv'))
+
+```
+
+``` {r}
+if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv'))) {
+
+  ### get species list
+  iucn_coral_cells_clipped <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv')) %>%
+    filter(deep == 0) %>%
+    mutate(prop_area = 1)
+
+  spp_coralmaps <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
+    filter(iucn_sid %in% iucn_coral_cells_clipped$iucn_sid) %>%
+    select(am_sid, sciname, iucn_sid)
+
+  ### copy aquamaps cells to local file, keeping only corals in the paired maps
+  if(!file.exists(file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))) {
+    am_coral_cells <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv', 'hcaf_sp_native_trunc.csv')) %>%
+      rename(am_sid = speciesid, prob = probability) %>%
+      filter(am_sid %in% spp_coralmaps$am_sid) 
+    write_csv(am_coral_cells, file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))
+  }
+
+  am_coral_cells <- read_csv(file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))
+
+  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
+  loiczid_raster       <- raster(loiczid_raster_file)
+  names(loiczid_raster) <- 'loiczid'
+
+  spp_coralmaps <- spp_coralmaps %>%
+    mutate(area_total = NA,  area_am = NA,  area_iucn = NA,
+           area_overlap = NA,
+           sm_perc    = NA, #filled with % of smaller range that is within the larger range
+           sm_range   = NA) #what range is smaller
+
+  i         <- 0 ### initialize the counting index
+  am_thresh <- 0 ### set the aquamaps presence threshold
+  sid_vector <- unique(spp_coralmaps$iucn_sid)
+  message(sprintf('There are a total of %s distinct species, by IUCN SID', 
+                  length(sid_vector)))
+
+  for (sid in sid_vector) { 
+    # sid <- spp_coralmaps$iucn_sid[1]
+    i <- i + 1
+    sids <- spp_coralmaps %>% 
+      select(am_sid, iucn_sid, sciname) %>%
+      filter(iucn_sid == sid)
+    rowindex <- (spp_coralmaps$iucn_sid == sid)
+
+    message(sprintf('%s. Processing species iucn_sid: %s \n   am_sid(s):  %s \n   sciname(s): %s', 
+                    i, sid, paste(sids$am_sid, collapse = ', '), paste(sids$sciname, collapse = ' ')))
+    map_iucn <- get_spp_map(sid, iucn_coral_cells_clipped) %>%
+      mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0))
+    map_am   <- get_spp_map(sids$am_sid, am_coral_cells) %>%
+      mutate(am_pres = ifelse(am_prob >= am_thresh, 1, 0))
+
+    ### rasterize aquamaps map
+    if(any(duplicated(map_am$loiczid))) message('Duplicated cells in am spp map')
+    map_am <- map_am %>%
+      dplyr::select(loiczid, am_pres) %>% 
+      unique()
+
+    r_am_spp <- subs(x = loiczid_raster, y = map_am %>% select(loiczid, am_pres), 
+                     by = 'loiczid', which = 'am_pres', 
+                     subsWithNA = TRUE)
+
+    ### rasterize IUCN map
+    if(TRUE %in% duplicated(map_iucn$loiczid)) stop('Duplicated cells in IUCN map')
+    r_iucn_spp <- subs(loiczid_raster, map_iucn %>% select(loiczid, iucn_pres), 
+                       by = 'loiczid', which = 'iucn_pres', 
+                       subsWithNA = TRUE)
+
+    ### Start calculating areas etc.
+    area_iucn <- area(r_iucn_spp, na.rm = TRUE) %>%
+      cellStats(., stat = 'sum')
+    area_am   <- area(r_am_spp, na.rm = TRUE) %>%
+      cellStats(., stat = 'sum')
+    spp_coralmaps[rowindex, ]$area_iucn <- area_iucn
+    spp_coralmaps[rowindex, ]$area_am   <- area_am
+
+    ### Get total area (IUCN and AM combined)
+    all_cells <- data.frame(loiczid = unique(c(map_iucn$loiczid, map_am$loiczid)),
+                           spp_pres = 1)
+    r_all_cells <- subs(loiczid_raster, all_cells, 
+                       by = 'loiczid', which = 'spp_pres', 
+                       subsWithNA = TRUE)
+    area_total  <- area(r_all_cells, na.rm = TRUE) %>% cellStats(., stat = 'sum')
+    spp_coralmaps[rowindex, ]$area_total <- area_total
+
+    ### Get total overlap (IUCN and AM together). If area_iucn = 0, just assign 0
+    if(area_iucn != 0) {
+      overlap_cells <- unique(map_iucn$loiczid[map_iucn$loiczid %in% map_am$loiczid])
+      if(length(overlap_cells) > 0) {
+        overlap_cells <- data.frame(loiczid = overlap_cells,
+                                  spp_pres = 1)
+        r_overlap_cells <- subs(loiczid_raster, overlap_cells, 
+                                by = 'loiczid', which = 'spp_pres', 
+                                subsWithNA = TRUE)
+        area_overlap <- area(r_overlap_cells, na.rm = TRUE) %>% cellStats(., stat = 'sum')
+      } else area_overlap <- 0 ### if no overlapping cells, assign zero
+    } else area_overlap <- 0 ### if no IUCN map cells, assign zero
+    spp_coralmaps[rowindex, ]$area_overlap <- area_overlap
+
+    message(sprintf('   Species iucn_sid %s: smaller range is %s. Smaller-in-larger is %.2f%%, A_sm:A_lg is %.2f%%.',
+                    sid, ifelse(area_am < area_iucn, 'AquaMaps', 'IUCN'), 
+                    ifelse(area_am < area_iucn, area_overlap/area_am*100, area_overlap/area_iucn*100),
+                    ifelse(area_am < area_iucn, area_am/area_iucn*100, area_iucn/area_am*100)))
+  }
+
+  spp_coralmaps <- spp_coralmaps %>%
+    mutate(sm_range   = ifelse(area_am < area_iucn, 'AM', 'IUCN'),
+           sm_perc    = ifelse(area_am < area_iucn, (area_overlap / area_am) * 100, (area_overlap / area_iucn) * 100),
+           area_ratio = area_am / area_iucn,
+           area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
+           area_ratio = area_ratio * 100)
+
+  spp_coralmaps <- spp_coralmaps %>%
+    mutate(lg_area = ifelse(area_am > area_iucn, area_am, area_iucn),
+           sm_area = ifelse(area_am < area_iucn, area_am, area_iucn),
+           lg_area_pct = 100 * lg_area / max(area_total, na.rm = TRUE)) %>% 
+             ### use largest data set as approx for total ocean range
+    select(-lg_area, -sm_area)
+
+  spp_coralmaps <- spp_coralmaps %>%
+    mutate(area_total  = round(area_total, 1),
+           area_am     = round(area_am, 1),
+           area_iucn   = round(area_iucn, 1),
+           sm_perc     = round(sm_perc, 3),
+           area_ratio  = round(area_ratio, 3),
+           lg_area_pct = round(lg_area_pct, 3))
+
+  print(head(spp_coralmaps))
+  write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv'))
+}
+```
+
+``` {r}
+### combine old and new area dataframes
+spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv')) %>%
+  select(iucn_sid, am_sid, sciname, area_am, 
+         area_iucn_clipped = area_iucn, area_overlap_clipped = area_overlap, 
+         sm_perc_clipped = sm_perc, sm_range_clipped = sm_range, area_ratio_clipped = area_ratio) %>%
+  left_join(read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
+              select(am_sid, iucn_sid, 
+                     area_iucn_raw = area_iucn, area_overlap_raw = area_overlap, 
+                     sm_perc_raw = sm_perc, sm_range_raw = sm_range, area_ratio_raw = area_ratio),
+            by = c('iucn_sid', 'am_sid'))
+
+write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
+```
+
+``` {r}
+
+spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
+
+### check raw iucn_area vs clipped - looks proportional; scaling factor due to difference in area calcs?
+scatterplot1 <- ggplot(spp_coralmaps, aes(x = area_iucn_raw, y = area_iucn_clipped)) +
+  geom_point() +
+  geom_abline(color = 'red')
+scatterplot1
+
+### rough recalcs
+### assume that by removing the deep cells, we will not remove cells that are
+### included in AquaMaps, since AquaMaps already filters by depth.  This means
+### the overlap area will not be affected.
+iucn_coral_fixed <- spp_coralmaps %>%
+  gather(dist, perc, c(sm_perc_raw, sm_perc_clipped)) %>%
+  gather(area, ratio, c(area_ratio_raw, area_ratio_clipped)) %>%
+  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
+  filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
+  mutate(method = ifelse(str_detect(dist, 'clip'), '200m clip', 'all depth'))
+
+clipped_quads <- ggplot(iucn_coral_fixed %>%
+                      filter(!iucn_sid %in% (iucn_coral_fixed %>% 
+                                               filter(perc > 100) %>% 
+                                               .$iucn_sid)),
+                    aes(x = ratio, y = perc, group = iucn_sid)) +
+  geom_line(color = 'grey50', alpha = .5) +
+  geom_point(aes(color = method)) +
+  labs(x = 'Area ratio (smaller/larger)',
+       y = 'Distribtion (%smaller in larger)',
+       title = 'Coral species alignment, IUCN full range vs depth-clipped')
+
+clipped_quads
+```

### 2016-08-02 coralmaps/coral_maps.Rmd
ok the deployment finally took.

@@
                     aes(x = ratio, y = perc, group = iucn_sid)) +
-  geom_line(color = 'grey50', alpha = .5) +
-  geom_point(aes(color = method)) +
+  geom_line(color = 'grey50', size = .5, alpha = .3) +
+  geom_point(aes(color = method), size = 2) +
   labs(x = 'Area ratio (smaller/larger)',

### 2016-08-02 shiny_am_iucn/server.R

@@ -30,6 +30,7 @@ server <- shinyServer(function(input, output, session) {
     {
       message('observed change in input$species or input$am_cutoff; getting spp_map dataframe')
       spp_map$df <- get_spp_map_df(input$species, input$am_cutoff)
+      print(head(spp_map$df))
     }
   )
   

### 2016-08-02 shiny_am_iucn/server_fxns.R

@@ -52,7 +52,7 @@ get_spp_map_df <- function(species, am_cutoff = 0){ ### species <- spp_list$scin
   return(spp_map_df)
 }
 
-get_rast <- function(spp_map_df, type = 'both') {
+get_rast <- function(spp_map_df, type) {
   message('in get_rast()')
   rast_obj  <-  subs(loiczid_raster, spp_map_df, 
                      by    = 'loiczid', 
@@ -208,6 +208,18 @@ create_quadplot <- function() {
 
 create_coralplot <- function() {
   spp_coralmaps <- read_csv('data/coral_spp_areas.csv')
+  
+  ggtheme_basic <- theme(axis.ticks = element_blank(),
+                         text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
+                         plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
+                         panel.background = element_blank(),
+                         legend.position = 'right')
+  
+  ggtheme_plot <- ggtheme_basic + 
+    theme(panel.border     = element_blank(),
+          panel.grid.minor = element_blank(), 
+          panel.grid.major = element_line(colour = 'grey90'),
+          axis.line = element_line(colour = "grey30"))
   
   iucn_coral_fixed <- spp_coralmaps %>%
     rename(dist_align_raw = sm_perc_raw, dist_align_clipped = sm_perc_clipped) %>%

### 2016-08-02 shiny_am_iucn/ui.R

@@ -1,15 +1,6 @@
 #ui.r
 
-library(readr)
-library(dplyr)
-library(tidyr)
-library(stringr)
-library(plotly)
-
-library(shinythemes)
-
-spp_list       <- read_csv('data/spp_list.csv')
-
+source('ui_setup.R')
 
 ### Tab 1: intro/explanation/link to published paper
 ### Tab 2: sample maps showing AM/IUCN/both; to select a species, can it be 

### 2016-08-02 shiny_am_iucn/ui_setup.R

@@ -1,0 +1,10 @@
+library(readr)
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(plotly)
+
+library(shinythemes)
+
+message('working directory ', getwd())
+spp_list <- read_csv('data/spp_list.csv')

### 2016-08-07 coralmaps/coral_maps.Rmd
cleaning up the old drafts; cleaning up the data_explores

--- 2016-07-31 coralmaps/coral_maps.Rmd
@@
 
-  ### get species list
-  data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
-  spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
+  ### get paired species list
+  paired_maps_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
+  spp_map_pairs <- read_csv(paired_maps_file) %>%
     filter(!is.na(popn_category)) %>% ### filter out unassessed species
@@
 
-  coral_spp_sid <- spp_map_pairs %>%
-    filter(str_detect(spp_group, 'CORAL')) %>%
-    .$iucn_sid
+  ### get full species list
+  spp_all_file <- file.path(dir_data, 'spp_all.csv')
+  spp_all <- read_csv(spp_all_file)
+
+  coral_spp_sid <- spp_all %>%
+    filter(str_detect(tolower(spp_group), 'coral')) %>%
+    .$iucn_sid %>%
+    unique()
 
@@
   spp_depth <- spp_narr %>%
-    select(iucn_sid = species_id, habitat)
-  write_csv(spp_depth, file.path(dir_git, 'coralmaps', 'coral_spp_depth_raw.csv'))
-  ### manually find max depths from this
-}
+    select(iucn_sid = species_id, habitat) %>%
+    mutate(paired_map = (iucn_sid %in% spp_map_pairs$iucn_sid))
 
+spp_depth1 <- spp_depth %>%
+  mutate(depth = str_extract_all(habitat, '\\d++(?:\\.\\d++)?(?=\\s*+(?:m))')) %>%
+  unnest(depth) %>%
+  group_by(iucn_sid) %>%
+  summarize(max_depth = max(as.integer(depth)))
+
+spp_depth <- spp_depth %>%
+  left_join(spp_depth1, by = 'iucn_sid') %>%
+  mutate(shallow_str = str_detect(tolower(habitat), 'shallow'),
+         shallow = max_depth <= 200 | shallow_str)
+
+write_csv(spp_depth, file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))
+
+summary(spp_depth)
+ #    iucn_sid        habitat          paired_map        max_depth      shallow_str     shallow       
+ # Min.   :132859   Length:562         Mode :logical   Min.   :  1.00   Mode :logical   Mode:logical  
+ # 1st Qu.:133081   Class :character   FALSE:174       1st Qu.: 20.00   FALSE:194       TRUE:544      
+ # Median :133284   Mode  :character   TRUE :388       Median : 20.00   TRUE :367       NA's:18       
+ # Mean   :133283                      NA's :0         Mean   : 27.55   NA's :1                       
+ # 3rd Qu.:133486                                      3rd Qu.: 30.00                                 
+ # Max.   :133703                                      Max.   :200.00                                 
+ #                                                     NA's   :46
+
+nrow(spp_depth %>% filter(max_depth > 100))
+# [1] 5
+### of 562 species, only 5 are listed as occurring below 100 m; none are below 200 m
+
+spp_depth_na <- spp_depth %>%
+  filter(is.na(shallow))
+
+### 17 species (of 396) have no depth information; one is listed as being dredged from 13 fathoms (~ 25 m)
 ```
 
-``` {r}
+``` {r extract_IUCN_cells_for_coral_pairs}
 
-spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))
 bath_df <-  read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd') %>%
@@
                           full.names = TRUE)
-iucn_coral_cells <- lapply(coral_files, read_csv) %>%
+coral_cells_iucn <- lapply(coral_files, read_csv) %>%
   bind_rows() %>%
   select(iucn_sid, loiczid = LOICZID) %>%
-  filter(iucn_sid %in% spp_depth$iucn_sid) %>%
+  filter(iucn_sid %in% spp_depth$iucn_sid[spp_depth$paired_map]) %>%
   left_join(bath_df, by = 'loiczid') %>%
@@
 
-write_csv(iucn_coral_cells, file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv'))
+write_csv(coral_cells_iucn, file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))
 
@@
 
-``` {r}
+``` {r calc_areas_for_iucn_corals}
 ### calc areas
-### read hcaf to get areas; left join to iucn_coral_cells
+### read hcaf to get areas; left join to coral_cells_iucn
 ### group_by iucn_sid and depth, summarize area (deep and shallow)
 ### group_by iucn_sid, summarize total area
-iucn_coral_cells <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv'))
+coral_cells_iucn <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))
 
@@
 
-iucn_coral_areas <- iucn_coral_cells %>%
+iucn_coral_areas <- coral_cells_iucn %>%
   left_join(hcaf, by = 'loiczid') %>%
@@
          shallow_pct = (totarea - deeparea)/totarea) %>%
-  filter(deep == 1) %>%
-  select(-deep)
+  arrange(desc(deep)) %>%
+  .[!duplicated(.$iucn_sid), ]
 
@@
 
-``` {r}
-if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv'))) {
+``` {r calc_overlaps}
+if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))) {
 
   ### get species list
-  iucn_coral_cells_clipped <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv')) %>%
+  coral_cells_iucn_clipped <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
     filter(deep == 0) %>%
@@
   spp_coralmaps <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
-    filter(iucn_sid %in% iucn_coral_cells_clipped$iucn_sid) %>%
+    filter(iucn_sid %in% coral_cells_iucn_clipped$iucn_sid) %>%
     select(am_sid, sciname, iucn_sid)
@@
   ### copy aquamaps cells to local file, keeping only corals in the paired maps
-  if(!file.exists(file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))) {
-    am_coral_cells <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv', 'hcaf_sp_native_trunc.csv')) %>%
+  if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))) {
+    coral_cells_am <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv', 'hcaf_sp_native_trunc.csv')) %>%
       rename(am_sid = speciesid, prob = probability) %>%
       filter(am_sid %in% spp_coralmaps$am_sid) 
-    write_csv(am_coral_cells, file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))
+    write_csv(coral_cells_am, file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))
   }
 
-  am_coral_cells <- read_csv(file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))
+  coral_cells_am <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))
 
@@
                     i, sid, paste(sids$am_sid, collapse = ', '), paste(sids$sciname, collapse = ' ')))
-    map_iucn <- get_spp_map(sid, iucn_coral_cells_clipped) %>%
+    map_iucn <- get_spp_map(sid, coral_cells_iucn_clipped) %>%
       mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0))
-    map_am   <- get_spp_map(sids$am_sid, am_coral_cells) %>%
+    map_am   <- get_spp_map(sids$am_sid, coral_cells_am) %>%
       mutate(am_pres = ifelse(am_prob >= am_thresh, 1, 0))
@@
   print(head(spp_coralmaps))
-  write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv'))
-}
-```
 
-``` {r}
   ### combine old and new area dataframes
-spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv')) %>%
+  spp_coralmaps <- spp_coralmaps %>%
     select(iucn_sid, am_sid, sciname, area_am, 
@@
   write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
+}
 ```
@@
 
+source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
 clipped_quads <- ggplot(iucn_coral_fixed %>%
@@
                     aes(x = ratio, y = perc, group = iucn_sid)) +
-  geom_line(color = 'grey50', alpha = .5) +
-  geom_point(aes(color = method)) +
+  ggtheme_plot + 
+  ### color the quadrant backgrounds:
+  # annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100, 
+  #          alpha = .3, fill= "#4dac26")  + 
+  # annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100, 
+  #          alpha = .3, fill= "#b8e186") + 
+  # annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0, 
+  #          alpha = .3, fill= "#f1b6da") + 
+  # annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0, 
+  #          alpha = .3, fill= "#d01c8b") +
+  scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
+                     breaks = c(seq(0, 100, 25)),
+                     labels = c('0%', '25%', '50%', '75%', '100%')) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 103),
+                     breaks = c(seq(0, 100, 25)),
+                     labels = c('0%', '25%', '50%', '75%', '100%')) +
+
+  geom_line(color = 'grey50', size = .5, alpha = .3) +
+  geom_point(aes(color = method), size = 2, alpha = .8) +
   labs(x = 'Area ratio (smaller/larger)',
-       y = 'Distribtion (%smaller in larger)',
-       title = 'Coral species alignment, IUCN full range vs depth-clipped')
+       y = 'Distribtion (%smaller in larger)')
 
-clipped_quads
+print(clipped_quads)
+
+ggsave(plot = clipped_quads,
+       filename = file.path(dir_fig, 'scatter_quadmap_corals.png'), device = 'png',
+       width = 17.8, height = 9, units = 'cm', dpi = 600)
+
 ```

### 2016-08-07 data_explore/data_explore.Rmd
cleaning up the old drafts; cleaning up the data_explores

@@
 iucn_file <- file.path(dir_data, 'spp_iucn.csv')
-spp_iucn <- read.csv(iucn_file, stringsAsFactors = FALSE)
+spp_iucn <- read_csv(iucn_file)
 
@@
 iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps.csv')
-spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
+spp_iucn_maps <- read_csv(iucn_maps_file) %>%
   filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
   filter(sciname %in% spp_iucn$sciname)   ### ditch freshwater spp
-
-spp_am_file <- file.path(dir_data, 'spp_am.csv')
-spp_am <- read_csv(spp_am_file)
 
@@
 
-sciname_lookup <- read.csv(file.path(dir_data, 'sciname_lookups/spp_iucn_scinames.csv'), stringsAsFactors = FALSE)
+sciname_lookup <- read_csv(file.path(dir_data, 'sciname_lookups/spp_iucn_scinames.csv'))
+
+spp_am_file <- file.path(dir_data, 'spp_am.csv')
+spp_am <- read_csv(spp_am_file)
 
@@
 
-taxa_am_all_assessed <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(sciname %in% spp_iucn$sciname) %>%
-  mutate(class = str_to_title(class),
-         n_am_tot = n()) %>%        ### length of whole AM maps list
-  group_by(class) %>%
-  summarize(n_am_class_ass  = n(),
-            p_am_class_ass  = round((n_am_class_ass/first(n_am_tot) * 100), 2))
-
-taxa_am_solo_assessed <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(sciname %in% spp_iucn$sciname) %>%
-  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
-  mutate(class = str_to_title(class),
-         n_am_solo = n()) %>%       ### length of AM only list
-  group_by(class) %>%
-  summarize(n_am_class_solo_ass = n(),
-            p_am_class_solo_ass = round((n_am_class_solo_ass/first(n_am_solo) * 100), 2))
-
 taxa_df <- taxa_am_all %>%
   full_join(taxa_am_solo,   by = 'class') %>%
-  full_join(taxa_am_all_assessed,  by = 'class') %>%
-  full_join(taxa_am_solo_assessed, by = 'class') %>%
   full_join(taxa_iucn_all,  by = 'class') %>%
@@
 
-# taxa_df_long1 <- taxa_df %>%
-#   dplyr::select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both) %>%
-#   gather(dataset, n_spp, AquaMaps:Both)
-# bar_taxa1 <- ggplot(taxa_df_long1, aes(x = class, y = n_spp)) +
-#   ggtheme_plot + 
-#   geom_bar(stat = 'identity') +
-#   facet_grid(dataset ~ ., scales = 'free_y') +
-#   theme(axis.text.x = element_text(angle = 75, hjust = 1))
-# print(bar_taxa1)
-
 #library(colorspace)
@@
 
-taxa_df_long2 <- taxa_df %>%
+taxa_df_long <- taxa_df %>%
   dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
@@
 
-bar_taxa2 <- ggplot(taxa_df_long2, aes(x = class, y = percent, fill = dataset)) +
+bar_taxa <- ggplot(taxa_df_long, aes(x = class, y = percent, fill = dataset)) +
   ggtheme_plot + 
   theme(panel.grid.major.x = element_blank()) +
-  geom_bar(stat = 'identity', alpha = .8) +
+  geom_bar(stat = 'identity', alpha = .9) +
   scale_fill_manual(values = colvec) +
@@
 
-print(bar_taxa2)
+print(bar_taxa)
 ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
        width = 17.8, height = 7, units = 'cm', dpi = 600)
-
-# taxa_df_long3 <- taxa_df %>%
-#   dplyr::select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
-#   gather(dataset, n_spp, AquaMaps:Both) %>%
-#   group_by(class) %>%
-#   mutate(n_total = sum(n_spp, na.rm = TRUE),
-#          percent = n_spp/n_total * 100)
-#
-# bar_taxa3 <- ggplot(taxa_df_long3, aes(x = class, y = percent, fill = dataset)) +
-#   ggtheme_plot + 
-#   geom_bar(stat = 'identity', alpha = .5) +
-#   scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
-#   geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
-#             #size = 3, 
-#             angle = 90, hjust = 0, color = 'grey30') +
-#   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
-#   scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
-#   labs(# title = 'Species maps by taxonomic class, Red List-assessed only',
-#        x = 'Class', 
-#        y = 'Percent of assessed mapped species')
-# print(bar_taxa3)
-# ggsave(filename = file.path(dir_fig, 'si_bar_by_taxa_assessed.tif'), device = 'tiff', 
-#        width = 17.8, height = 6, units = 'cm', dpi = 600)
 
@@
                                y = sm_perc)) + #, 
- #                               size = lg_area_pct,
- #                              color = popn_category)) + 
   ggtheme_plot + 
   ### color the quadrant backgrounds:
-    annotate("rect", xmin = area_align_mean, xmax = 100, 
-                     ymin = dist_align_mean, ymax = 100, 
-             alpha = .3, 
-             fill= "#4dac26")  + 
-    annotate("rect", xmax = area_align_mean, xmin =   0, 
-                     ymin = dist_align_mean, ymax = 100, 
-             alpha = .3, 
-             fill= "#b8e186") + 
-    annotate("rect", xmin = area_align_mean, xmax = 100, 
-                     ymax = dist_align_mean, ymin =   0, 
-             alpha = .3, 
-             fill= "#f1b6da") + 
-    annotate("rect", xmax = area_align_mean, xmin =   0, 
-                     ymax = dist_align_mean, ymin =   0, 
-             alpha = .3, 
-             fill= "#d01c8b") + 
+  annotate("rect", xmin = area_align_mean, xmax = 100, ymin = dist_align_mean, ymax = 100, 
+           alpha = .3, fill= "#4dac26")  + 
+  annotate("rect", xmax = area_align_mean, xmin =   0, ymin = dist_align_mean, ymax = 100, 
+           alpha = .3, fill= "#b8e186") + 
+  annotate("rect", xmin = area_align_mean, xmax = 100, ymax = dist_align_mean, ymin =   0, 
+           alpha = .3, fill= "#f1b6da") + 
+  annotate("rect", xmax = area_align_mean, xmin =   0, ymax = dist_align_mean, ymin =   0, 
+           alpha = .3, fill= "#d01c8b")
+
+  ### plot the data points!
+scatter_quadplot <- scatter_quadplot +
   geom_point(color = '#4d4dac', alpha = .6)
 
-### Manage scales for color and size 
+  ### label the quadrants:
 scatter_quadplot <- scatter_quadplot +
-#   scale_colour_manual(values = c('LC' = 'green4', 
-#                                  'NT' = 'greenyellow', 
-#                                  'VU' = 'yellow', 
-#                                  'EN' = 'orange', 
-#                                  'CR' = 'orangered2',
-#                                  'DD' = 'grey30')) + 
-#   scale_size_continuous(limits = c(0, 100),
-#                         breaks = c( 0,    25,    50,    75,    100), 
-#                         labels = c('0%', '25%', '50%', '75%', '100%')) + 
-  scale_x_continuous(expand = c(0, 0), 
-                     limits = c(0, 102),
+  annotate("rect", xmin = 82, xmax = 98, ymin = 85, ymax = 91, 
+           alpha = .7, fill= 'white')  + 
+  annotate("rect", xmax = 24, xmin =  2, ymin = 85, ymax = 91, 
+           alpha = .7, fill= "white") + 
+  annotate("rect", xmin = 82, xmax = 98, ymax =  9, ymin = 15, 
+           alpha = .7, fill= "white") + 
+  annotate("rect", xmax = 20, xmin =  2, ymax =  9, ymin = 15, 
+           alpha = .7, fill= "white") +
+  annotate("text", x = 96, y = 88, hjust = 1, vjust = .5, size = 3, color = 'grey20', 
+           fontface = 'bold.italic', label = "Well-aligned") + 
+  annotate("text", x =  4, y = 88, hjust = 0, vjust = .5, size = 3, color = 'grey20',
+           fontface = 'bold.italic', label = "Distribution-aligned") + 
+  annotate("text", x = 96, y = 12, hjust = 1, vjust = .5, size = 3, color = 'grey20',
+           fontface = 'bold.italic', label = "Area-aligned") + 
+  annotate("text", x =  4, y = 12, hjust = 0, vjust = .5, size = 3, color = 'grey20',
+           fontface = 'bold.italic', label = "Poorly aligned")
+
+### Manage axis scales 
+scatter_quadplot <- scatter_quadplot +
+  scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
                      breaks = c(seq(0, 100, 25)),
                      labels = c('0%', '25%', '50%', '75%', '100%')) +
-  scale_y_continuous(expand = c(0, 0),
-                     limits = c(0, 103),
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 103),
                      breaks = c(seq(0, 100, 25)),
@@
 scatter_quadplot <- scatter_quadplot +
-  annotate(geom = 'text',
-           x = area_align_mean, y = 5,
-           hjust = 0, vjust = 0,
-           color = 'grey30', 
-           size = 2,
+  annotate(geom = 'text', x = area_align_mean, y = 5,
+           hjust = 0, vjust = 0, color = 'grey30', size = 2,
            fontface = 'bold.italic', angle = 90,
            label = sprintf('Mean = %s%%', round(area_align_mean, 1))) +
-  annotate(geom = 'text',
-           x = 5, y = dist_align_mean,
-           hjust = 0, vjust = 0,
-           color = 'grey30', 
-           size = 2,
+  annotate(geom = 'text', x = 5, y = dist_align_mean,
+           hjust = 0, vjust = 0, color = 'grey30', size = 2,
            fontface = 'bold.italic', angle = 0,
@@
   labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
-       y = bquote('Distribution alignment '*alpha['dist']*' (%)')) #, 
-#       title = 'Distribution alignment vs Extent alignment',
-#       size = '% ocean range',
- #      color = 'Extinction\nRisk')
+       y = bquote('Distribution alignment '*alpha['dist']*' (%)'))
 
@@
 
-ggsave(file.path(dir_fig, 'scatter_quadmap_norisk.png'),
+ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
        width = 17.8, height = 9, units = 'cm', dpi = 600)
@@
 
+quad_names <- data.frame('quad' = c('q1', 'q2', 'q3', 'q4'),
+                         'quad_name' = factor(c('well-matched', 'dist-matched', 
+                                               'area-matched', 'poorly matched'),
+                                             ordered = TRUE))
+
 spp_gp_quadrants <- spp_gp_quadrants %>%
@@
             by = 'spp_group_text') %>%
-  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
-         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
-         quad = str_replace(quad, 'q3', 'q3 area-matched'),
-         quad = str_replace(quad, 'q4', 'q4 poorly matched'))
+  left_join(quad_names, by = 'quad')
 
@@
   geom_bar(stat = 'count', alpha = .8) +
-  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
-                               'q2 dist-matched'   = '#b8e186', 
-                               'q3 area-matched'   = '#f1b6da', 
-                               'q4 poorly matched' = '#d01c8b')) +
-#   scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
-#                                'q2 dist-matched'   = 'greenyellow', 
-#                                'q3 area-matched'   = 'orange', 
-#                                'q4 poorly matched' = 'orangered2')) +
+  scale_fill_manual(values = c('q1'   = '#4dac26', 
+                               'q2'   = '#b8e186', 
+                               'q3'   = '#f1b6da', 
+                               'q4' = '#d01c8b'),
+                    labels = levels(quad_names)) +
   scale_y_continuous(expand = c(0, 0), 
@@
        y = 'Percent of species by quadrant', 
-#       title = 'Taxonomic group by quadrant',
        fill = 'Alignment')
@@
 -----
-
-# Obsolete plots; not evaluated
-
-* bar chart: number of spp per group
-* scatter plot: area compare IUCN vs AM, including log plot and plot by species group
-* rangemap: Chromis alpha
-
-``` {r child = 'data_explore_pt1.Rmd', eval = FALSE}
-```
-
------
-
-* percent overlap vs total area exploration
-
-``` {r, child = 'data_explore_pt2.Rmd', eval = FALSE}
-```
-
------
-
-* Multiple species maps per quadrant
-
-``` {r child = 'data_explore_pt3.Rmd', eval = FALSE}
-```
-
------

### 2016-08-07 drafts/initial_version/ms_2_intro.Rmd
cleaning up the old drafts; cleaning up the data_explores

--- 2016-04-02 ms_2_intro_res_disc.Rmd
@@
 One major outcome of this body of science is the various compiled databases of species distribution maps.
-Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps (1) and International Union for Conservation of Nature (IUCN) (2).
-These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status (3, 4, 5), evaluating global biodiversity patterns (6, 7, 8, 9), predicting species range shifts (10), and setting conservation priorities (11).
+Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps [@kaschner_aquamaps:_2015] and International Union for Conservation of Nature (IUCN) [@iucn_iucn_2015].
+These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
 
@@
 Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
-Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans (12, 13).
+Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans [@rondinini_tradeoffs_2006; @jetz_ecological_2008].
 
@@
 
-We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (3, 4) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (11) - as case studies to explore the implications of prioritizing one data set over the other.
+We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index [@halpern_index_2012; @halpern_patterns_2015] and a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as case studies to explore the implications of prioritizing one data set over the other.
 The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
-
-# Results and Discussion
-
-The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" (2) - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.
-In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of "probability of occurrence" for each species (14, 15).
-Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence (3, 4).
-
-The methodologies behind the creation of these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).
-Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.
-Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range (12).
-By comparing maps resulting from IUCN and AquaMaps methodologies, we can identify and possibly address mechanistic causes for each type of error.
-
-The two datasets have notably different taxonomic (Fig.
-1A) and regional (Figs.
-1B, 1C) coverage.
-AquaMaps encompasses a broader range of taxa than IUCN, as IUCN spatial data files are only available for select taxonomic groups that have been comprehensively assessed.
-While species numbers in both datasets peak in tropical latitudes near the equator, species counts for IUCN maps drop quickly beyond 30°N and 30°S, while species counts for AquaMaps remain robust well into temperate latitudes.
-The longitude frequency plots show a slight shift in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
-To achieve more comprehensive global coverage of species ranges these two datasets can be used together, but the underlying methodological differences complicate such direct comparisons.
-
-``` {r load_spp_list_by_quad, echo = FALSE}
-spp_list_quads <- read.csv('data/spp_list_quads.csv', stringsAsFactors = FALSE)
-n_q1 <- nrow(spp_list_quads %>% filter(quad == 'q1'))
-n_q2 <- nrow(spp_list_quads %>% filter(quad == 'q2'))
-# n_q2_am_sm  <- nrow(spp_list_quads %>% filter(sm_range == 'AM' & quad == 'q2'))
-n_q2_corals <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral') & quad == 'q2'))
-n_corals    <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral')))
-n_q3 <- nrow(spp_list_quads %>% filter(quad == 'q3'))
-n_q4 <- nrow(spp_list_quads %>% filter(quad == 'q4'))
-
-# n_q4_datapoor <- nrow(spp_list_quads %>% filter(quad == 'q4' & occurcells < 10))
-# p_q4_datapoor <- round(n_q4_datapoor/n_q4 * 100, 1)
-# n_q123_datapoor <- nrow(spp_list_quads %>% filter(quad != 'q4' & occurcells < 10))
-# p_q123_datapoor <- round(n_q123_datapoor/n_spp_both * 100, 1)
-
-# n_q4_rev <- nrow(spp_list_quads %>% filter(quad == 'q4' & !is.na(reviewed)))
-# p_q4_rev <- round(n_q4_rev/n_q4 * 100, 1)
-# n_q123_rev <- nrow(spp_list_quads %>% filter(quad != 'q4' &  !is.na(reviewed)))
-# p_q123_rev <- round(n_q123_rev/n_spp_both * 100, 1)
-
-# quad_count_means <- read_csv(file.path(dir_data, 'quads_data_poor.csv')) %>%
-#   arrange(quad)
-# dp_med_am_q4    <- quad_count_means$med_occurcells[4]
-# dp_med_am_q123  <- quad_count_means$med_occurcells_total[1]
-# dp_med_occ_q4   <- quad_count_means$med_occ[4]
-# dp_med_occ_q123 <- quad_count_means$med_occ_total[1]
-
-spp_quads_occ <- read_csv(file.path(dir_data, 'spp_quads_occ.csv'))
-med_occ_q4   <- round(median(spp_quads_occ$mean_occ[spp_quads_occ$quad == 'q4']))
-med_occ_q123 <- round(median(spp_quads_occ$mean_occ[spp_quads_occ$quad %in% c('q1', 'q2', 'q3')]))
-
-```
-
-To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area alignment (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig.
-2A).
- This analysis revealed a general negative linear pattern, suggesting that increasing similarity in range area correlates with decreasing distribution alignment.
- AquaMaps tends to extrapolate species ranges into suitable areas beyond known occurrences, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
-For species with dissimilar range areas, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
-For species with increasingly similar range areas, differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.
-
-The mean distribution alignment for species included in both datasets was `r dist_align_mean`%; the mean area alignment was `r area_align_mean`%.
-By dividing the map-paired species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
-Representative maps from each category are provided in the supplemental materials (Fig.
-S1).
-
-The upper right quadrant includes species (n = `r n_q1`) whose described ranges are above average in alignment of both spatial distribution and area.
-These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig.
-2B).
-This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their ranges span nearly the entire map.
-
-The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where the IUCN range is notably larger, often extending into deeper water.
-For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their preferred depths, likely introducing errors of commission  (Fig.
-S3).
- Ocean depth preference is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps.
-Simply clipping IUCN range maps to known depth preferences would resolve many of these mismatches.
-
-Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, introducing additional commission errors.
-Consequently, area ratios are close to 100%, though similar areas are unhelpful when the distributions are poorly aligned.
-
-The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor area match well.
-Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) (16) and the Global Biodiversity Information Facility (GBIF) (17)) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
-When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species.
-
-# Implications
-
-Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.
-To examine the implications of these differences, we repeated two recent studies, varying only the prioritization of one data set over the other.
-
-**Case Study: The Ocean Health Index**
-
-``` {r calc_ohi_numbers, echo = FALSE, message = FALSE, warning = FALSE}
-spp_ohi <- spp_all %>%
-  filter(!is.na(spatial_source) & !is.na(category_score)) %>%
-  select(am_sid, iucn_sid, spatial_source) %>%
-  unique()
-
-n_spp_am_ohi <- spp_ohi %>% 
-                    filter(str_detect(spatial_source, 'am')) %>%
-                    select(am_sid) %>%
-                    unique() %>%
-                    nrow()
-n_spp_iucn_ohi <- spp_ohi %>%
-                    filter(str_detect(spp_ohi$spatial_source, 'iucn')) %>%
-                    filter(is.na(am_sid)) %>%
-                    unique() %>%
-                    nrow()
-n_spp_both_ohi <- spp_ohi %>% 
-                    filter(!is.na(am_sid)) %>%
-                    filter(str_detect(spatial_source, 'iucn')) %>%
-                    select(iucn_sid) %>%
-                    unique() %>%
-                    nrow()
-
-n_spp_ohi <- n_spp_iucn_ohi + n_spp_am_ohi + n_spp_both_ohi
-
-p_spp_both_ohi <- round(n_spp_both_ohi/n_spp_ohi * 100, 1)
-
-```
-
-The global Ocean Health Index (OHI) (3, 4), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
-To maximize the number of represented species, OHI gleans spatial distribution data from Red List-assessed species in both IUCN and AquaMaps datasets (n = `r format(n_spp_ohi, big.mark = ',')`), prioritizing IUCN data for the `r format(n_spp_both_ohi, big.mark = ',')` species included in both sources.
-OHI currently uses a probability of occurrence threshold of 40% to determine species presence for AquaMaps data.
-
-We calculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig.
-3</a>).
-Reducing the threshold increases the apparent range of a species, so a decrease in a region’s score under scenario 1 would indicate increased spatial representation of threatened species for that region.
-Shifting priority from IUCN to AquaMaps may increase or decrease a species’ apparent range, so a decrease in region score for scenario 2 may indicate a decrease in spatial representation of low-risk species, an increase in spatial representation of high-risk species, or a combination of both.
-Scenario three combines effects of scenarios 1 and 2.
-
-Given that only `r p_spp_both_ohi`% of species are included in both datasets, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in substantial country-level score shifts as seen in scenarios 2 and 3.
-While the mean global score did not vary significantly from scenario to scenario, select countries gained up to 7.5 points while others dropped as many as 5 points.
-This result indicates that especially on a national or regional scale, an arbitrary change in how the two datasets are combined can result in a different assessment of species conservation status.
-
-**Case Study: MPA Gap Analysis**
-
-Klein et al.
-(11) compare the global distribution of species to the global distribution of marine protected areas to assess how well current MPAs overlap with species ranges and identify which species fall through gaps in protection.
-The study relied on the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
-They found that the global MPA network leaves 90.5% of marine species with less than 5% of their overall range represented within MPAs, and 1.4% of species have no protection at all (i.e., "gap" species).
-
-We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using the most recent AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig4'>Fig.
-4</a>).
-We found a five-fold increase in the proportion of gap species (6.4% of species vs.
-1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% of species vs.
-47.7%).
-However, this comparison also indicates a larger proportion of well-protected species with greater than 10% of range protected (2.9% of species vs.
-1.5%).
-Conclusions drawn from each of these datasets would paint dramatically different pictures of the protection afforded by our current global MPA network.
-
-# Conclusions
-
-AquaMaps and IUCN range maps show reasonable agreement for many well-studied species, but substantial differences illustrate uncertainty in our understanding of spatial distribution for many others.
-Although many other approaches exist for species distribution modeling, these two are the only ones applied globally across different marine taxa.
-Identifying and addressing differences in these datasets will increase their utility for research and conservation actions.
-Several likely drivers of commission and omission errors between these datasets point to a few important ways to improve range alignment.
-
-For IUCN range data, clipping ranges to known depth limits improves output for many species, most notably corals and reef-associated fishes.
-If species' depth limits are not known, simple rules of thumb will likely reduce commission errors without introducing substantial omission errors.
-For example, for most corals, researchers could clip range maps to the photosynthetic limit of 200 meters.
- For AquaMaps range data, dependent primarily on environmental and physical preferences and conditions, implementing area restrictions based on biogeographical criteria such as Marine Ecoregions of the World (18) would likely decrease commission errors and improve predictive power, especially for data-poor species.
- For either data set, maps based on few occurrences are more likely to include errors of commission and omission.
-Occurrence counts from external sources such as OBIS or GBIF can help identify relatively data-poor species; additionally, the "occurcells" attribute in the AquaMaps data set, which counts the number of half-degree cells used to generate the environmental envelope for each species, can be used in a similar manner.
-
-Many studies using AquaMaps data test the sensitivity of results by varying the probability of occurrence threshold to determine presence.
-This decision ultimately represents a tradeoff between errors of commission (low threshold) and omission (high threshold).
-Using AquaMaps in conjunction with IUCN can mitigate potential errors, while also increasing the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized.
-In this case, we recommend a presence threshold of 0% as it most closely approximates the "limits of distribution" criterion defined by IUCN data providers.
-
-Effective management and protection of marine species depends on a robust understanding of where species exist and where they do not; without this knowledge we risk wasting resources protecting low-value regions while missing opportunities to protect critical ones.
-By identifying the differences between these two fundamental marine species range datasets and understanding the likely mechanisms causing these discrepancies, we improve our ability to develop strategic and effective conservation policy that supports a resilient ocean ecosystem.

### 2016-08-07 drafts/initial_version/ms_4_refs.Rmd
cleaning up the old drafts; cleaning up the data_explores

--- 2016-03-27 ms_5_refs.Rmd
@@
 -->
-
-Kaschner K et al.
-(2015) AquaMaps: Predicted range maps for aquatic species.
-World wide web electronic publication, www.aquamaps.org, Version 08/2015.
-<!--___since this is not a peer-reviewed publication, can we include this for PNAS reference list?___-->
-
-IUCN (2015).
-The IUCN Red List of Threatened Species.
-Version 2015-4.
-<http://www.iucnredlist.org>.
-Downloaded on 21 December 2015.
-<!--___since this is not a peer-reviewed publication, can we include this for PNAS reference list?___-->
-
-Halpern BS et al. (2012). An index to assess the health and benefits of the global ocean. Nature, 488(7413), 615-620. <!--_doi:10.1038/nature11397._-->
-
-Halpern BS et al. (2015). Patterns and emerging trends in global ocean health. PloS one, 10(3), e0117863. <!--___doi?___-->
-
-Klein CJ et al. (2015). Shortfalls in the global protected area network at representing marine biodiversity. Scientific Reports 5: 17539. <!--_doi:10.1038/srep17539._-->
-
-Selig ER et al. (2013). Assessing global marine biodiversity status within a coupled socio-ecological perspective. PloS one, 8(4), e60284. <!--_doi:10.1371/journal.pone.0060284._-->
-
-Coll M et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. <!--_used AquaMaps to predict Med biodiversity.  Also: Threshold = 0._-->
-
-*Martin CS et al.
-(2014).
-Manual of marine and coastal datasets of biodiversity importance.
-May 2014 release.
-Cambridge (UK): UNEP World Conservation Monitoring Centre.
-28 pp.
-<!--(+ 4 annexes totalling 174 pp.
-and one e-supplement).
-_review of marine data sets and data gaps etc, incl both IUCN and AM as well as many others_-->
-
-Pimm SL et al.
-(2014).
-The biodiversity of species and their rates of extinction, distribution, and protection.
-Science, 344(6187), 1246752.
-<!--_uses range maps to show biodiversity areas; may use IUCN range maps.
-Also discusses gaps and possible things that can be done about them._-->
-
-Kaschner K, Tittensor DP, Ready J, Gerrodette T, Worm B (2011).
-Current and Future Patterns of Global Marine Mammal Biodiversity.
-PLoS ONE 6(5): e19653.
-<!--_just what the title says - AM development, presence threshold 60%, also analyzes richness as a function of threshold_-->
-
-Molinos JG et al. (2015). Climate velocity and the future global redistribution of marine biodiversity. Nature Climate Change. <!--_doi:10.1038/nclimate2769._-->
-
-Rondinini C, Wilson KA, Boitani L, Grantham H, & Possingham HP (2006).
-Tradeoffs of different types of species occurrence data for use in systematic conservation planning.
-Ecology letters, 9(10), 1136-1145.
-<!--_compares point locality, range maps, and distribution models in terms of omission and commission errors; also outlines Extent of Occurrence and Area of Occupancy distinctions._-->
-
-Jetz W, Sekercioglu CH, and Watson JE (2008).
-Ecological correlates and conservation implications of overestimating species geographic ranges.
-Conservation Biology, 22(1), 110-119.
-<!--_EOO maps are usually highly interpolated and overestimate small-scale occurrence, which may bias research outcomes_-->
-
-Kaschner K, Watson R, Trites AW, Pauly D (2006).
-Mapping world-wide distributions of marine mammal species using a relative environmental suitability (RES) model.
-Marine Ecology Progress Series 316: 285–310.
-<!--_check this citation journal name..._ _This outlines the basic RES methodology - AM development_-->
-
-Ready J et al.
-(2010).
-Predicting the distributions of marine organisms at the global scale.
-Ecological Modelling 221(3): 467-478.
-<!--_Presents AM; assessing AquaMaps against other presence-only species models_-->
-
-Spalding MD et al. (2007). Marine ecoregions of the world: a bioregionalization of coastal and shelf areas. BioScience, 57(7), 573-583.
-
-QGIS Development Team (2015). QGIS Geographic Information System. Open Source Geospatial Foundation Project. http://qgis.osgeo.org
-
-IUCN and UNEP-WCMC (2014). The World Database on Protected Areas (WDPA) [On-line]. Cambridge, UK: UNEP- WCMC. Available at: www.protectedplanet.net [Accessed December 1, 2014].
-
-R Core Team (2015). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

### 2016-08-07 drafts/initial_version/ms_figures.Rmd
cleaning up the old drafts; cleaning up the data_explores

--- 2016-03-29 ms_figures.Rmd
@@
 ---
+
+``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
+
+# library(ggplot2)
+# library(maps)
+
+# library(raster)
+# library(data.table)
+library(tidyr)
+library(dplyr)
+library(stringr)
+library(readr)
+
+# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
+#            'Darwin'  = '/Volumes/data_edit',
+#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
+
+# dir_anx  <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
+
+if(basename(getwd()) != 'IUCN-AquaMaps') setwd('~/github/IUCN-AquaMaps')
+
+dir_data <- 'data'
+dir_fig  <- 'figures'
+
+```
+
+``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
+### use data_explore.Rmd to create this file, if it does not yet exist.  This
+### filters out presence = Extinct from IUCN cells.
+
+data_file <- 'data/spp_list_w_area_trimmed.csv'
+spp_map_pairs <- read_csv(data_file)
+### NOTE: as of Jan 31, includes all map pairs, even ones where IUCN area == 0
+
+### generate other lists
+spp_all_file <- 'data/spp_all.csv'
+spp_all <- read_csv(spp_all_file) ### NOTE: contains duped scinames
+
+# iucn_marine_file <- 'data/spp_iucn_marine.csv'
+# spp_iucn_marine <- read_csv(iucn_marine_file)
+
+iucn_maps_file <- 'data/spp_iucn_maps.csv'
+spp_iucn_maps <- read_csv(iucn_maps_file) %>%
+  filter(!duplicated(iucn_sid))
+
+spp_am_file <- 'data/spp_am.csv'
+spp_am <- read_csv(spp_am_file)
+
+```
+
+``` {r calc species numbers, echo = FALSE}
+n_spp_am     <- nrow(spp_am)
+n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
+n_spp_iucn   <- length(unique(spp_iucn_maps$iucn_sid))
+n_spp_both   <- length(unique(spp_map_pairs$iucn_sid))
+n_spp_all    <- n_spp_am + n_spp_iucn - n_spp_both
+
+p_spp_am     <- round(n_spp_am/n_spp_all * 100, 1)
+p_spp_iucn   <- round(n_spp_iucn/n_spp_all * 100, 1)
+p_both_am    <- round(n_spp_both/n_spp_am   * 100, 1)
+p_both_iucn  <- round(n_spp_both/n_spp_iucn * 100, 1)
+p_both_all   <- round(n_spp_both/n_spp_all  * 100, 1)
+
+spp_list <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
+area_align_mean <- round(mean(spp_list$area_ratio, na.rm = TRUE), 1)
+dist_align_mean <- round(mean(spp_list$sm_perc, na.rm = TRUE), 1)
+
+```
 
@@
 
+# Figures
+
 ### Fig. 1
@@
 ![Change in MPA coverage](figures/fig4_barchart_mpa_gap.png)
+
+-----
+
+``` {r figure_legends, child = 'ms_5_fig_legends.Rmd'}
+```

### 2016-08-07 ms_1_intro.Rmd
cleaning up the old drafts; cleaning up the data_explores

--- 2016-08-07 drafts/initial_version/ms_2_intro.Rmd
@@
  and managing and protecting species effectively.
-A rich literature tackles the many dimensions of these questions.
+<!--___A rich literature tackles the many dimensions of these questions.___-->
 
@@
 <!--
-Three foundational papers relating to AquaMaps (Kaschner et al. 2006, 2011; Ready et al. 2010) have been cited a total of 315 times.
+The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
+Discuss "truth" and uncertainty in species distributions.
+There is an objective "truth"; 
 -->
+Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each presents one understanding of the "truth" - one database relying primarily on expert opinion and the other on model predictions.
+Uncertainties inherent in the method and intent inevitably drive differences in the range predictions made by each dataset; significant differences could ___lead to, communicate, imply___ dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
 
-The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
 Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
@@
 Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).
-For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and outline possible improvements.
+
+It seems fair to assume that range predictions from each dataset would be more closely aligned for well-understood species.
+For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and suggest methods to improve confidence in species range predictions.
 

### 2016-08-07 ms_2_methods.Rmd
cleaning up the old drafts; cleaning up the data_explores

# Methods

_About the datasets__:

The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" [@iucn_iucn_2015] - essentially a refined extent of occurrence (EOO), based on observation records and informed by expert understanding of species' range and habitat preferences.  In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.  The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of "probability of occurrence" for each species [@kaschner_mapping_2006; @ready_predicting_2010].  Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].

___should this go in the discussion?___
_The methodologies behind the creation of these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).  Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.  Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006].  By comparing maps resulting from IUCN and AquaMaps methodologies, we can identify and possibly address mechanistic causes for each type of error._

__Comparison of taxonomic and regional distribution__:
To examine the overall taxonomic distribution across the spatial datasets (Fig. 1A), we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.

To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; we determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell.  For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.  We represented relative distribution of species count for each dataset by plotting average species count against latitude and longitude (Fig. 1B, 1C).

__Comparison of paired maps__:
Using genus and species binomials as a matching key, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.  ___To avoid double-counting, we removed subpopulations and species aliases. Used the taxize package? which version?___  We determined species presence within each spatial cell using the same criteria outlined above.

Overlaying paired distribution maps for a given species, we defined and calculated _distribution alignment_ and _area ratio_ and plotted these in Fig. 2A:

$$\alpha_{dist} = \frac{A_{small \cap large}}{A_{large}} * 100\%$$

$$\alpha_{area} = \frac{A_{small}}{A_{large}} * 100\%$$

For each species with ranges described in both IUCN and AquaMaps, $A_{small}$ and $A_{large}$ indicate the smaller and larger range representation (regardless of which dataset).  $A_{small \cap large}$ represents the amount of overlapping area between the two datasets.

_Distribution alignment - overlap as proxy for common understanding; use Bayesian argument to show that these areas are high-confidence areas; area ratio as proxy for errors of commission/omission.  While we cannot know which dataset is closer to the "truth" we can make inferences based on expectations of EOO and AOO maps - is AquaMaps AOO?_

Dividing the plot into quadrants based on mean values along each axis, we roughly categorized four qualities of alignment, and for each taxonomic group, based on IUCN taxonomic groupings, determined the proportion of species that fell into each quality category (2B).

Noting that a large proportion of corals were identified in the upper-left quadrant, we examined a handful of IUCN coral maps and found that each predicted coral presence in far deeper waters than would be expected based on the IUCN's habitat description.  To examine this issue further, we created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth???] and masked our IUCN coral rasters to constrain the maps to shallow waters.  None of the paired-map corals is described as occurring deeper than 200 m, ___and most are found in far shallower waters.___  The resulting maps were again compared to the AquaMaps rasters; ___figure X___ plots shifts in coral species alignment due to this depth clipping.

**Methods for Ocean Health Index (OHI) case study:**
The Species component of the OHI Biodiversity goal (SPP) uses IUCN Red List extinction risk as a proxy for a species' health.  Each species is assigned a risk score based on IUCN Red List extinction risk category, from Least Concern (0.0) to Extinct (1.0) (Data Deficient and unassessed species are excluded).  Species range is determined from its IUCN range map, where available, or its AquaMaps range map ("presence" defined as a 40% or greater probability of occurrence) if an IUCN map is unavailable.  For each assessed OHI region, an area-weighted mean extinction risk across all species is calculated, and rescaled from 0 (mean extinction risk = 0.75) to 100 (mean extinction risk = 0) [@halpern_index_2012; @halpern_patterns_2015].

___For our analysis, we modified the original code for OHI 2015 Species status, allowing for flexibility in which data source was used for the overlapping species, and flexibility in modifying the AquaMaps presence threshold.  Using the output of the published OHI 2015 SPP status as our baseline, we ran the modified SPP code three times: once with IUCN  prioritizing IUCN over AquaMaps for a 0% threshold, and prioritizing AquaMaps over IUCN for both a 40% and 0% threshold.  We compared each of these to the output of the published OHI 2015 SPP status.___

_To highlight the differences in the datasets, we selected the subset of species whose ranges are represented in both datasets.  In the original OHI SPP analysis, IUCN data was used to determine ranges for these overlapping species; we calculated the SPP status for these species as our baseline.  We then recalculated using AquaMaps ranges at a 40% presence threshold (as used in the original OHI analysis) and a 0% threshold (to better simulate the "limits of distribution" defined by IUCN maps) and compared them to our IUCN baseline._

_We then modified the original OHI source code to recalculate the 2015 SPP status, using AquaMaps rather than IUCN for the overlapping species, and setting a 0% presence threshold on all AquaMaps species ranges.  This we compared to the original SPP status scores for 2015._

* ___overlapping species only? IUCN as baseline; AM 40%; AM 0%___

**Methods for MPA Gap Analysis case study:** 
To assess the effectiveness of MPAs in protecting biodiversity, Klein et al. [-@klein_shortfalls_2015] compared the coverage of the global MPA network presented by the World Database on Protected Areas (WDPA) [@iucn_world_2014] to the species ranges described in the 2014 AquaMaps dataset.  For this study, the researchers defined "presence" as 50% or greater probability of occurrence.

To reconstruct the primary analysis, we selected the subset of protected areas from the 2014 WDPA dataset classified as IUCN protected area management categories I-IV and spatially covering a marine area. The WDPA polygons and marine polygons were rasterized to 0.01° and then aggregated to AquaMaps native 0.5° cells, to calculate proportion of marine protected area within each cell.  After verifying our results using the 2014 AquaMaps dataset, we updated the analysis to use the 2015 AquaMaps dataset, at a presence threshold of 50% (to compare to Klein et al. directly) and 0% (to better compare with IUCN spatial data).  To analyze MPA coverage against IUCN spatial data, we extracted IUCN polygon weights per 0.5° cell for each species and compared against the protected area raster.  Finally, we combined AquaMaps data (at 0% threshold) and IUCN data, using AquaMaps for the `r format(n_spp_both, big.mark = ',')` overlapping species and again compared against the protected area raster.

All processing was completed using R statistical software [@r_core_team_r:_2016], and all code and intermediate data are available on GitHub at https://github.com/OHI-Science/IUCN-AquaMaps.

### 2016-08-07 ms_3_res_disc.Rmd
cleaning up the old drafts; cleaning up the data_explores

--- 2016-05-13 ms_2a_res_disc.Rmd
@@
 # Results and Discussion
-
-The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" [@iucn_iucn_2015] - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.
-In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of "probability of occurrence" for each species [@kaschner_mapping_2006; @ready_predicting_2010].
-Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
-
-The methodologies behind the creation of these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).
-Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.
-Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006].
-By comparing maps resulting from IUCN and AquaMaps methodologies, we can identify and possibly address mechanistic causes for each type of error.
 
@@
 By dividing the map-paired species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
-Representative maps from each category are provided in the supplemental materials (Fig.
+Representative maps from each category are provided in the supporting materials (Fig.
 S1).

### 2016-08-07 case_study_ohi/ohi_paired_maps.Rmd
push to Mazu to work remotely

--- 2016-08-07 coralmaps/coral_maps.Rmd
@@
 library(ggplot2)
-library(maps)
 library(readr)
@@
 
-```{r get_bathy_raster, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+```{r get_spp_cells, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
-### get shapefile
-### rasterize to half-degree grid; or raster::extract
-
-if(!file.exists(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))) {
-
-  poly_bath <- readOGR(dsn = file.path(path.expand(dir_git), 'coralmaps/ne_10m_bathymetry_K_200'), 
-                       layer = 'ne_10m_bathymetry_K_200', 
-                       stringsAsFactors = TRUE)
-
-  ### LOICZID cell ID numbers - used in plotting rasters
 loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
@@
 
-  bath_list <- raster::extract(loiczid_raster, poly_bath, 
-                             weights = TRUE, normalizeWeights = FALSE,
-                             progress = 'text')
-  bath_df <- lapply(bath_list, function(x) {as.data.frame(x)}) %>%
-    bind_rows() %>%
-    rename(loiczid = value) %>%
-    group_by(loiczid) %>%
-    summarize(weight = sum(weight))
+paired_maps_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
+spp_map_pairs <- read_csv(paired_maps_file) %>%
+  filter(!is.na(popn_category)) %>% ### filter out unassessed species
+  filter(!is.na(sm_perc))           ### filter out NA maps
 
-  ### generate raster?
-  bath_raster <- subs(loiczid_raster, bath_df, by = 'loiczid', which = 'weight')
-  values(bath_raster)[values(bath_raster) > 5] <- NA
+# load IUCN cells, filter to just paired maps
+spp_cells_iucn
 
-  write_csv(bath_df, file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))
-  ### probably want to filter to just 100% 200m+
-}
+# load AM cells, filter to just paired maps; include column of prob of occur
 
@@
   filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
-  mutate(method = ifelse(str_detect(dist, 'clip'), '200m clip', 'all depth'))
+  mutate(method = ifelse(str_detect(dist, 'clip'), '200 m clip', 'original'))
 
@@
   ### color the quadrant backgrounds:
-  # annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100, 
-  #          alpha = .3, fill= "#4dac26")  + 
-  # annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100, 
-  #          alpha = .3, fill= "#b8e186") + 
-  # annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0, 
-  #          alpha = .3, fill= "#f1b6da") + 
-  # annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0, 
-  #          alpha = .3, fill= "#d01c8b") +
+  annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100,
+           alpha = .3, fill= "#4dac26")  +
+  annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100,
+           alpha = .3, fill= "#b8e186") +
+  annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0,
+           alpha = .3, fill= "#f1b6da") +
+  annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0,
+           alpha = .3, fill= "#d01c8b") +
   scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
@@
 
-  geom_line(color = 'grey50', size = .5, alpha = .3) +
-  geom_point(aes(color = method), size = 2, alpha = .8) +
+  geom_line(color = 'grey50', size = .25, alpha = .3) +
+  # geom_point(color = 'grey80', size = 2, alpha = 1) + ### blank behind points
+  geom_point(aes(color = method), size = 2, alpha = .7) +
+  scale_color_manual(values = c('original' = '#4d4dac', '200 m clip' = '#00cc00')) +
   labs(x = 'Area ratio (smaller/larger)',

### 2016-08-07 data_explore/data_explore.Rmd
push to Mazu to work remotely

@@
 
+### define windows for labels
+q1 <- c('x1' = 84, 'x2' = 98, 'y1' = 92.5, 'y2' = 97.5)
+q2 <- c('x1' =  2, 'x2' = 22, 'y1' = 92.5, 'y2' = 97.5)
+q3 <- c('x1' = 84, 'x2' = 98, 'y1' =  2.5, 'y2' = 7.5)
+q4 <- c('x1' =  2, 'x2' = 18, 'y1' =  2.5, 'y2' = 7.5)
+
+spp_list_labs <- spp_list %>%
+  rename(x = area_ratio, y = sm_perc) %>%
+  mutate(fade = FALSE,
+         fade = ifelse((x > q1[1] & x < q1[2] & y > q1[3] & y < q1[4]), TRUE, fade),
+         fade = ifelse((x > q2[1] & x < q2[2] & y > q2[3] & y < q2[4]), TRUE, fade),
+         fade = ifelse((x > q3[1] & x < q3[2] & y > q3[3] & y < q3[4]), TRUE, fade),
+         fade = ifelse((x > q4[1] & x < q4[2] & y > q4[3] & y < q4[4]), TRUE, fade))
+
 ### mongo plot time
-scatter_quadplot <- ggplot(spp_list %>% 
-                             transform(popn_category = reorder(popn_category, category_score)),
-                           aes(x = area_ratio, 
-                               y = sm_perc)) + #, 
+scatter_quadplot <- ggplot(spp_list_labs, aes(x = x, y = y)) +
   ggtheme_plot + 
@@
 scatter_quadplot <- scatter_quadplot +
-  geom_point(color = '#4d4dac', alpha = .6)
+  # geom_point(color = 'white',   alpha = 1) + ### blank behind the points
+  geom_point(color = '#4d4dac', aes(alpha = fade), show.legend = FALSE) +
+  scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = .7))
 
@@
 scatter_quadplot <- scatter_quadplot +
-  annotate("rect", xmin = 82, xmax = 98, ymin = 85, ymax = 91, 
-           alpha = .7, fill= 'white')  + 
-  annotate("rect", xmax = 24, xmin =  2, ymin = 85, ymax = 91, 
-           alpha = .7, fill= "white") + 
-  annotate("rect", xmin = 82, xmax = 98, ymax =  9, ymin = 15, 
-           alpha = .7, fill= "white") + 
-  annotate("rect", xmax = 20, xmin =  2, ymax =  9, ymin = 15, 
-           alpha = .7, fill= "white") +
-  annotate("text", x = 96, y = 88, hjust = 1, vjust = .5, size = 3, color = 'grey20', 
+  annotate("text", x = 97, y = 95, hjust = 1, vjust = .5, size = 3, color = 'grey20', 
            fontface = 'bold.italic', label = "Well-aligned") + 
-  annotate("text", x =  4, y = 88, hjust = 0, vjust = .5, size = 3, color = 'grey20',
+  annotate("text", x =  3, y = 95, hjust = 0, vjust = .5, size = 3, color = 'grey20',
            fontface = 'bold.italic', label = "Distribution-aligned") + 
-  annotate("text", x = 96, y = 12, hjust = 1, vjust = .5, size = 3, color = 'grey20',
+  annotate("text", x = 97, y =  5, hjust = 1, vjust = .5, size = 3, color = 'grey20',
            fontface = 'bold.italic', label = "Area-aligned") + 
-  annotate("text", x =  4, y = 12, hjust = 0, vjust = .5, size = 3, color = 'grey20',
+  annotate("text", x =  3, y =  5, hjust = 0, vjust = .5, size = 3, color = 'grey20',
            fontface = 'bold.italic', label = "Poorly aligned")

### 2016-08-09 ms_3_res_disc.Rmd
working on results/discussion - notes and outlines

@@
 The longitude frequency plots show a slight shift in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
-To achieve more comprehensive global coverage of species ranges these two datasets can be used together, but the underlying methodological differences complicate such direct comparisons.
+___To achieve more comprehensive global coverage of species ranges these two datasets can be used together, but the underlying methodological differences complicate such direct comparisons.___
 
@@
 
-To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area alignment (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig.
+___Distribution alignment - overlap as proxy for common understanding; use Bayesian argument to show that these areas are high-confidence areas; area ratio as proxy for errors of commission/omission.
+While we cannot know which dataset is closer to the "truth" we can make inferences based on expectations of EOO and AOO maps - is AquaMaps AOO?___
+
+_Distribution alignment - Bayesian description of higher confidence_
+_Area ratio is defined as the ratio of the smaller range area to the larger range area.
+This can provide insights into commission and omission errors: a large difference in areas implies a greater number of errors in one or both datasets.
+While we cannot know which dataset is closer to representing the "truth" of each species' range, we can expect that IUCN (representing EOO) will introduce more commission errors, while AquaMaps (representing AOO) is likely to reduce these at the expense of introducing more omission errors._
+
+___is this sufficiently described in methods?
+or revisit here?___ To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area alignment (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig.
 2A).
@@
 
-The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where the IUCN range is notably larger, often extending into deeper water.
+The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where the IUCN range is notably larger, often extending into deeper water than the AquaMaps ranges suggest.
 For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their preferred depths, likely introducing errors of commission (Fig.
 S3).
- Ocean depth preference is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps.
-Simply clipping IUCN range maps to known depth preferences would resolve many of these mismatches.
+ _Ocean depth preference is explicitly included in AquaMaps models; while depth is recommended by the IUCN as a criterion for range maps, it seems to be overlooked in some cases._  Simply clipping IUCN range maps to known depth preferences would resolve many of these commission errors.
 
-Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, introducing additional commission errors.
+___from IUCN:  The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range.
+http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources___
+
+___area ratio suggests similar commission/omission errors, but dist suggests otherwise___ Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, likely introducing additional commission errors.
 Consequently, area ratios are close to 100%, though similar areas are unhelpful when the distributions are poorly aligned.
 
-The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor area match well.
-Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
+*The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor area match well.* ___present a solution here?___ Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
 When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species.
@@
 Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.
-To examine the implications of these differences, we repeated two recent studies, varying only the prioritization of one data set over the other.
+___come back to the idea that neither represents the "truth", and different studies might choose one over the other without recognizing the implications___ To examine the implications of these differences, we repeated two recent studies, varying only the prioritization of one data set over the other.
 

### 2016-08-10 coralmaps/coral_maps.Rmd
fixing up the plots: new quad plot, new coral quad plot, new dataset by taxa plot, new OHI plot

@@
 
-if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_depth_raw.csv'))) {
+if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))) {
 
@@
   write_csv(spp_depth, file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))
+
+} else 
+  spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))
 
@@
 
-spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
+spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv')) %>% 
+                      mutate(loss_ratio = 100 * area_iucn_clipped/area_iucn_raw)
 
+spp_small_effect <- spp_coralmaps %>%
+  filter(loss_ratio > 80) 
+### only five with such small effects; one of those did not change at all.
+
+summary(spp_coralmaps$loss_ratio)
+  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+  # 23.80   52.89   55.48   56.33   60.59  100.00
+
+hist(spp_coralmaps$loss_ratio)
 ### check raw iucn_area vs clipped - looks proportional; scaling factor due to difference in area calcs?
+
 scatterplot1 <- ggplot(spp_coralmaps, aes(x = area_iucn_raw, y = area_iucn_clipped)) +
@@
 
-### rough recalcs
-### assume that by removing the deep cells, we will not remove cells that are
-### included in AquaMaps, since AquaMaps already filters by depth.  This means
-### the overlap area will not be affected.
+violplot1 <- ggplot(spp_coralmaps, aes(x = 'sm_range_raw', y = loss_ratio)) +
+  ggtheme_plot +
+  geom_violin(fill = 'grey80', alpha = .5,
+              draw_quantiles = c(.25, .5, .75)) +
+  geom_point(position = position_jitter(w = .3), 
+             color = '#4d4dac', alpha = .5) +
+  geom_violin(fill = NA, draw_quantiles = c(.25, .5, .75)) +
+  geom_hline(color = 'grey30', yintercept = 100) +
+  scale_y_continuous(limits = c(0, 100)) +
+  scale_x_discrete(labels = element_blank()) +
+  theme(axis.title.x = element_blank()) +
+  labs(y = 'IUCN coral range: (<= 200 m) / (total)')
+print(scatterplot1)
+
+### recalc areas and overlaps
 iucn_coral_fixed <- spp_coralmaps %>%
@@
   filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
-  mutate(method = ifelse(str_detect(dist, 'clip'), '200m clip', 'all depth'))
+  mutate(method = ifelse(str_detect(dist, 'clip'), '200 m clip', 'original'))
 
 source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
+
+break_nums <- seq(0, 100, 20)
+
 clipped_quads <- ggplot(iucn_coral_fixed %>%
@@
   ### color the quadrant backgrounds:
-  # annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100, 
-  #          alpha = .3, fill= "#4dac26")  + 
-  # annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100, 
-  #          alpha = .3, fill= "#b8e186") + 
-  # annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0, 
-  #          alpha = .3, fill= "#f1b6da") + 
-  # annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0, 
-  #          alpha = .3, fill= "#d01c8b") +
-  scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
-                     breaks = c(seq(0, 100, 25)),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 103),
-                     breaks = c(seq(0, 100, 25)),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) +
+  annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100,
+           alpha = .3, fill= "#4dac26")  +
+  annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100,
+           alpha = .3, fill= "#b8e186") +
+  annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0,
+           alpha = .3, fill= "#f1b6da") +
+  annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0,
+           alpha = .3, fill= "#d01c8b") +
+  ### add grid lines
+  geom_vline(xintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
+  geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
 
-  geom_line(color = 'grey50', size = .5, alpha = .3) +
-  geom_point(aes(color = method), size = 2, alpha = .8) +
-  labs(x = 'Area ratio (smaller/larger)',
-       y = 'Distribtion (%smaller in larger)')
+  scale_x_continuous(expand = c(0, 0), limits = c(-1, 101),
+                     breaks = break_nums,
+                     labels = sprintf('%s%%', break_nums)) +
+  scale_y_continuous(expand = c(0, 0), limits = c(-1, 101),
+                     breaks = break_nums,
+                     labels = sprintf('%s%%', break_nums)) +
+
+  geom_line(color = 'grey50', size = .25, alpha = .3) +
+  geom_point(aes(color = method, size = method), alpha = 1) +
+  scale_color_manual(values = c('original' = 'grey60', '200 m clip' = 'black')) +
+  scale_size_manual(values = c('original' = .25, '200 m clip' = .25)) +
+    labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
+         y = bquote('Distribution alignment '*alpha['dist']*' (%)'))
 

### 2016-08-10 data_explore/data_explore.Rmd
fixing up the plots: new quad plot, new coral quad plot, new dataset by taxa plot, new OHI plot

@@
 
+break_nums <- seq(0, 100, 20)
+
 bar_taxa <- ggplot(taxa_df_long, aes(x = class, y = percent, fill = dataset)) +
@@
   theme(panel.grid.major.x = element_blank()) +
-  geom_bar(stat = 'identity', alpha = .9) +
+  geom_bar(stat = 'identity', alpha = 1) +
+  geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
   scale_fill_manual(values = colvec) +
@@
                      limits = c(0, 130),
-                     breaks = c(0, 25, 50, 75, 100),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) + 
+                     breaks = break_nums,
+                     labels = sprintf('%s%%', break_nums)) + 
   labs(# title = 'Species maps by taxonomic class',
@@
 ![](../figures/bar_by_taxa_all.png)
+
+``` {r plot bar-line charts by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
+### bar charts comparing number of species in each class for each data set
+
+#library(colorspace)
+#hclvec <- heat_hcl(3); colvec <- c('AquaMaps' = hclvec[1], 'Both' = hclvec[2], 'IUCN' = hclvec[3])
+#colvec <- c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')
+colvec <- c('AquaMaps' = '#76A94C', 'Both' = '#e18258', 'IUCN' = '#9168a9')
+
+taxa_df1 <- taxa_df %>%
+  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
+  mutate(AquaMaps = ifelse(is.na(AquaMaps), 0, AquaMaps),
+         IUCN     = ifelse(is.na(IUCN),     0, IUCN),
+         Both     = ifelse(is.na(Both),     0, Both),
+         n_total  = AquaMaps + IUCN + Both,
+         IUCN_end = 100 * (IUCN + Both)/n_total,
+         IUCN_start = 0,
+         AquaMaps_end   = 100,
+         AquaMaps_start = 100 * (n_total - AquaMaps - Both)/n_total,
+         Both_end = AquaMaps_start,
+         Both_start = IUCN_end) %>%
+  gather(key = dataset,  value = start, ends_with('start')) %>%
+  gather(key = dataset2, value = end, ends_with('end')) %>%
+  mutate(dataset = str_replace(dataset, '_start', ''),
+         dataset2 = str_replace(dataset2, '_end', '')) %>%
+  filter(dataset == dataset2) %>%
+  select(-dataset2)
+
+taxa_all_df <- taxa_df1 %>%
+  group_by(dataset) %>%
+  summarize(class = 'Total',
+            AquaMaps_total = sum(AquaMaps) + sum(Both),
+            IUCN_total = sum(IUCN) + sum(Both),
+            n_total  = sum(n_total),
+            IUCN_end = 100 * IUCN_total/n_total,
+            IUCN_start = 0,
+            AquaMaps_end   = 100,
+            AquaMaps_start = 100 * (n_total - AquaMaps_total)/n_total,
+            Both_end = AquaMaps_start,
+            Both_start = IUCN_end) %>%
+  select(-dataset) %>% distinct() %>%
+  gather(key = dataset,  value = start, ends_with('start')) %>%
+  gather(key = dataset2, value = end, ends_with('end')) %>%
+  mutate(dataset = str_replace(dataset, '_start', ''),
+         dataset2 = str_replace(dataset2, '_end', '')) %>%
+  filter(dataset == dataset2) %>%
+  select(-dataset2, -AquaMaps_total, -IUCN_total)
+
+taxa_df2 <- taxa_all_df %>%
+  bind_rows(taxa_all_df %>%
+              mutate(class = ' ',
+                     start = 0,
+                     end   = 0,
+                     n_total = NA)) %>%
+  bind_rows(taxa_df1 %>%
+              select(-IUCN, -AquaMaps, -Both)) %>%
+  mutate(class = factor(class, levels = unique(.$class), ordered = TRUE),
+         n_label = ifelse(!is.na(n_total), paste0('n = ', n_total), ''))
+
+break_nums <- seq(0, 100, 20)
+
+dodge <- position_dodge(width = 0.1)
+
+bar_taxa <- ggplot(taxa_df2, aes(x = class, color = dataset)) +
+  ggtheme_plot + 
+  theme(panel.grid.major.x = element_blank()) +
+  geom_linerange(aes(ymin = start, ymax = end, size = dataset),
+                 position = position_dodge(width = 0.5)) +
+  scale_size_manual(values = c('IUCN' = .25, 'AquaMaps' = .25, 'Both' = 2.5)) +
+  scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey60', 'Both' = 'black')) +
+  geom_text(aes(label = n_label, y = 102),
+            size = 2,
+            angle = 90, hjust = 0, color = 'grey30') +
+  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
+  scale_y_continuous(expand = c(0, 0),
+                     limits = c(0, 130),
+                     breaks = break_nums,
+                     labels = sprintf('%s%%', break_nums)) +
+  labs(# title = 'Species maps by taxonomic class',
+       x = 'Class',
+       y = 'Percent of mapped species',
+       fill = 'Dataset')
+
+print(bar_taxa)
+ggsave(filename = file.path(dir_fig, 'bar2_by_taxa_all.png'), 
+       width = 17.8, height = 7, units = 'cm', dpi = 600)
+
+```
+
+![](../figures/bar2_by_taxa_all.png)
 
@@
 
+break_nums <- seq(0, 100, 20)
+
 ### mongo plot time
@@
   annotate("rect", xmax = area_align_mean, xmin =   0, ymax = dist_align_mean, ymin =   0, 
-           alpha = .3, fill= "#d01c8b")
+           alpha = .3, fill= "#d01c8b") +
+  ### add grid lines
+  geom_vline(xintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
+  geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5)
 
@@
 scatter_quadplot <- scatter_quadplot +
-  # geom_point(color = 'white',   alpha = 1) + ### blank behind the points
-  geom_point(color = '#4d4dac', aes(alpha = fade), show.legend = FALSE) +
-  scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = .7))
+  # geom_point(color = '#4d4dac', aes(alpha = fade), show.legend = FALSE) +
+  # geom_point(data = spp_list_labs %>%
+  #              filter(str_detect(spp_group_text, 'Corals')),
+  #            color = 'red', aes(alpha = fade), 
+  #            shape = 4, stroke = .5, 
+  #            size = 1, show.legend = FALSE) +
+  # scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = 1)) +
+  geom_point(color = 'black', aes(alpha = fade), size = .1, show.legend = FALSE) +
+  scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = 1))
 
@@
 scatter_quadplot <- scatter_quadplot +
-  scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
-                     breaks = c(seq(0, 100, 25)),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 103),
-                     breaks = c(seq(0, 100, 25)),
-                     labels = c('0%', '25%', '50%', '75%', '100%'))
+  scale_x_continuous(expand = c(0, 0), limits = c(-1, 101),
+                     breaks = break_nums,
+                     labels = sprintf('%s%%', break_nums)) + 
+  scale_y_continuous(expand = c(0, 0), limits = c(-1, 101),
+                     breaks = break_nums,
+                     labels = sprintf('%s%%', break_nums))
 
@@
   theme(panel.grid.major.y = element_blank()) +
-  geom_bar(stat = 'count', alpha = .8) +
-  scale_fill_manual(values = c('q1'   = '#4dac26', 
+  geom_bar(stat = 'count', alpha = 1) +
+  scale_fill_manual(values = c('q4' = '#d01c8b', 
+                               'q3' = '#f1b6da', 
                                'q2' = '#b8e186',
-                               'q3'   = '#f1b6da', 
-                               'q4' = '#d01c8b'),
-                    labels = levels(quad_names)) +
+                               'q1' = '#4dac26'),
+                    labels = quad_names$quad_name) +
   scale_y_continuous(expand = c(0, 0), 
                      limits = c(0, 1.1),
-                     breaks = c(0, .25, .5, .75, 1.0),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) +
-  geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
+                     breaks = break_nums/100,
+                     labels = sprintf('%s%%', break_nums)) + 
+  ### add grid lines; horizontal but then get flipped
+  geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
+  ### add text
+  geom_text(# data = spp_gp_quadrants %>%
+              # select(spp_group_text, n_spp) %>%
+              # distinct(),
+            aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
             size = 2, 
@@
   theme(panel.grid.major.y = element_blank()) +
-  geom_bar(stat = 'count', alpha = .8) +
+  geom_bar(stat = 'count', alpha = 1) +
   scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
@@
                      limits = c(0, 1.1),
-                     breaks = c(0, .25, .5, .75, 1.0),
-                     labels = c('0%', '25%', '50%', '75%', '100%')) +
+                     breaks = break_nums/100,
+                     labels = sprintf('%s%%', break_nums)) + 
+  ### add grid lines
+  geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
+  ### add text
   geom_text(aes(label = sprintf('n = %s', n_tot_cat), y = 1.01), hjust = 0, 

### 2016-08-10 ms_1_intro.Rmd
fixing up the plots: new quad plot, new coral quad plot, new dataset by taxa plot, new OHI plot

@@
 
-We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index [@halpern_index_2012; @halpern_patterns_2015] and a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as case studies to explore the implications of prioritizing one data set over the other.
+We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (OHI) [@halpern_index_2012; @halpern_patterns_2015] and a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as case studies to explore the implications of prioritizing one data set over the other.
 The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
+
+___we are not recommending one data set as better than the other!___
+
+___acknowledge or debunk other global data sets - e.g. SAUP?___

### 2016-08-10 ms_2_methods.Rmd
fixing up the plots: new quad plot, new coral quad plot, new dataset by taxa plot, new OHI plot

@@
 Noting that a large proportion of corals were identified in the upper-left quadrant, we examined a handful of IUCN coral maps and found that each predicted coral presence in far deeper waters than would be expected based on the IUCN's habitat description.
-To examine this issue further, we created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth???] and masked our IUCN coral rasters to constrain the maps to shallow waters.
-None of the paired-map corals is described as occurring deeper than 200 m, ___and most are found in far shallower waters.___  The resulting maps were again compared to the AquaMaps rasters; ___figure X___ plots shifts in coral species alignment due to this depth clipping.
+Using the IUCN's API [REF] to find life history information for all IUCN-mapped corals, we found that none of the paired-map corals is described as occurring deeper than 200 m, ___and most are found in far shallower waters.___  We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] and masked our IUCN coral rasters to exclude presence below 200 m.
+The resulting maps were again compared to the AquaMaps rasters; ___figure XXXX___ plots shifts in coral species alignment due to this depth clipping.
+___does some of this belong in the results/discussion?___
 
@@
 Species range is determined from its IUCN range map, where available, or its AquaMaps range map ("presence" defined as a 40% or greater probability of occurrence) if an IUCN map is unavailable.
-For each assessed OHI region, an area-weighted mean extinction risk across all species is calculated, and rescaled from 0 (mean extinction risk = 0.75) to 100 (mean extinction risk = 0) [@halpern_index_2012; @halpern_patterns_2015].
+For each assessed OHI region, an area-weighted mean extinction risk across all species is calculated, and rescaled to calculate a status score from 0 (mean extinction risk = 0.75) to 100 (mean extinction risk = 0) [@halpern_index_2012; @halpern_patterns_2015].
 
-___For our analysis, we modified the original code for OHI 2015 Species status, allowing for flexibility in which data source was used for the overlapping species, and flexibility in modifying the AquaMaps presence threshold.
-Using the output of the published OHI 2015 SPP status as our baseline, we ran the modified SPP code three times: once with IUCN  prioritizing IUCN over AquaMaps for a 0% threshold, and prioritizing AquaMaps over IUCN for both a 40% and 0% threshold.
-We compared each of these to the output of the published OHI 2015 SPP status.___
+_To highlight the differences in the datasets, we selected the subset of species whose ranges are represented in both datasets, eliminating species found in only one or the other dataset.
+In the original OHI SPP analysis, IUCN data was used to determine ranges for these overlapping species; we calculated all regional SPP status scores for just these species as our baseline.
+We then recalculated the scores using AquaMaps ranges at a 40% presence threshold (as used in the original OHI analysis) and a 0% threshold (to better simulate the "limits of distribution" defined by IUCN maps) and compared them to our IUCN baseline._
 
-_To highlight the differences in the datasets, we selected the subset of species whose ranges are represented in both datasets.
-In the original OHI SPP analysis, IUCN data was used to determine ranges for these overlapping species; we calculated the SPP status for these species as our baseline.
-We then recalculated using AquaMaps ranges at a 40% presence threshold (as used in the original OHI analysis) and a 0% threshold (to better simulate the "limits of distribution" defined by IUCN maps) and compared them to our IUCN baseline._
-
-_We then modified the original OHI source code to recalculate the 2015 SPP status, using AquaMaps rather than IUCN for the overlapping species, and setting a 0% presence threshold on all AquaMaps species ranges.
-This we compared to the original SPP status scores for 2015._
-
-* ___overlapping species only? IUCN as baseline; AM 40%; AM 0%___
+_We then modified the original OHI source code to recalculate the 2015 SPP status scores for all regions and all species to examine the effects of changing the AquaMaps presence threshold from 40% to 0% (all range included), and the effects of changing the data source for the overlapping species from IUCN to AquaMaps.
+We calculated the differences between these results and the published SPP status scores for 2015._
 
@@
 To assess the effectiveness of MPAs in protecting biodiversity, Klein et al.
-[-@klein_shortfalls_2015] compared the coverage of the global MPA network presented by the World Database on Protected Areas (WDPA) [@iucn_world_2014] to the species ranges described in the 2014 AquaMaps dataset.
-For this study, the researchers defined "presence" as 50% or greater probability of occurrence.
+[-@klein_shortfalls_2015] compared the coverage of the global MPA network presented by the World Database on Protected Areas (WDPA) [@iucn_world_2014] to the species ranges described in the 2014 AquaMaps dataset [@aquamaps_XXX].
+For the primary analysis, the researchers defined species "presence" as 50% or greater probability of occurrence.
 

### 2016-08-10 ms_figures.Rmd
added zotero functionality; preparing for submission to ecology letters
fixing up the plots: new quad plot, new coral quad plot, new dataset by taxa plot, new OHI plot

@@
 
-<a id = 'fig4'></a>
+<a id = 'fig3'></a>
 
 ![Absolute change in SPP score](figures/fig3_boxplot_abs_diff.png)
+![Absolute change in SPP score](figures/fig3_boxplot2_abs_diff.png)
 

### 2016-08-12 coralmaps/coral_maps.Rmd
tracking down coral bathymetry questions

@@
   ### LOICZID cell ID numbers - used in plotting rasters
-  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
+  loiczid_raster_file  <- file.path(dir_git, 'shiny_am_iucn/data/loiczid_raster.tif')
   loiczid_raster       <- raster(loiczid_raster_file)
@@
 
+  writeRaster(bath_raster, file.path(dir_git, 'coralmaps', 'bathy_cells.tif'))
+
   write_csv(bath_df, file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))
+  # bath_df <- read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd')
   ### probably want to filter to just 100% 200m+
@@
 
+spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))
+
 ### load iucn spp_cells
@@
   select(iucn_sid, loiczid = LOICZID) %>%
-  filter(iucn_sid %in% spp_depth$iucn_sid[spp_depth$paired_map]) %>%
+  # filter(iucn_sid %in% spp_depth$iucn_sid[spp_depth$paired_map]) %>%
   left_join(bath_df, by = 'loiczid') %>%
@@
 data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
+spp_map_pairs <- read_csv(data_file) %>%
   filter(!is.na(popn_category)) %>%   ### filter out unassessed species
@@
   ### get species list
-  coral_cells_iucn_clipped <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
+  coral_cells_iucn_shallow <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
     filter(deep == 0) %>%
@@
   spp_coralmaps <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
-    filter(iucn_sid %in% coral_cells_iucn_clipped$iucn_sid) %>%
+    filter(iucn_sid %in% coral_cells_iucn_shallow$iucn_sid) %>%
     select(am_sid, sciname, iucn_sid)
@@
 
-  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
+  loiczid_raster_file  <- file.path(dir_git, 'shiny_am_iucn/data/loiczid_raster.tif')
   loiczid_raster       <- raster(loiczid_raster_file)
@@
                     i, sid, paste(sids$am_sid, collapse = ', '), paste(sids$sciname, collapse = ' ')))
-    map_iucn <- get_spp_map(sid, coral_cells_iucn_clipped) %>%
+    map_iucn <- get_spp_map(sid, coral_cells_iucn_shallow) %>%
       mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0))
@@
     select(iucn_sid, am_sid, sciname, area_am, 
-           area_iucn_clipped = area_iucn, area_overlap_clipped = area_overlap, 
-           sm_perc_clipped = sm_perc, sm_range_clipped = sm_range, area_ratio_clipped = area_ratio) %>%
+           area_iucn_shallow = area_iucn, area_overlap_shallow = area_overlap, 
+           sm_perc_shallow = sm_perc, sm_range_shallow = sm_range, area_ratio_shallow = area_ratio) %>%
     left_join(read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
@@
 
-``` {r}
+``` {r generate_area_plots}
 
-spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv')) %>% 
-                      mutate(loss_ratio = 100 * area_iucn_clipped/area_iucn_raw)
+spp_coral_clip <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv')) %>%
+  select(iucn_sid, shallow_pct, totarea, deeparea) %>%
+  mutate(shallow_pct = 100 * shallow_pct)
 
-spp_small_effect <- spp_coralmaps %>%
-  filter(loss_ratio > 80) 
+spp_small_effect <- spp_coral_clip %>%
+  filter(shallow_pct > 80) 
 ### only five with such small effects; one of those did not change at all.
 
-summary(spp_coralmaps$loss_ratio)
+summary(spp_coral_clip$shallow_pct)
   #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
@@
 
-hist(spp_coralmaps$loss_ratio)
-### check raw iucn_area vs clipped - looks proportional; scaling factor due to difference in area calcs?
-
-scatterplot1 <- ggplot(spp_coralmaps, aes(x = area_iucn_raw, y = area_iucn_clipped)) +
-  geom_point() +
+scatterplot1 <- ggplot(spp_coral_clip, aes(x = totalarea, y = deeparea)) +
+  geom_point(size = 0.25, color = 'black') +
   geom_abline(color = 'red')
@@
 
-violplot1 <- ggplot(spp_coralmaps, aes(x = 'sm_range_raw', y = loss_ratio)) +
+violplot1 <- ggplot(spp_coral_clip, aes(x = 'sm_range_raw', y = shallow_pct)) +
   ggtheme_plot +
-  geom_violin(fill = 'grey80', alpha = .5,
-              draw_quantiles = c(.25, .5, .75)) +
-  geom_point(position = position_jitter(w = .3), 
-             color = '#4d4dac', alpha = .5) +
-  geom_violin(fill = NA, draw_quantiles = c(.25, .5, .75)) +
+  # geom_violin(fill = 'grey80', alpha = .5,
+  #             draw_quantiles = c(.25, .5, .75)) +
+  # geom_point(position = position_jitter(w = .3), 
+  #            size = 0.25,
+  #            color = 'black', alpha = 1) +
+  geom_violin(fill = NA, size = 0.1, weight = 0.5,
+              draw_quantiles = c(seq(.005, .05, .005), seq(.05, .95, .02), seq(.95, 1.0, .005)),
+              color = 'black') +
+  geom_violin(fill = NA, size = 1,
+              draw_quantiles = NULL,
+              color = 'white') +
   geom_hline(color = 'grey30', yintercept = 100) +
@@
   labs(y = 'IUCN coral range: (<= 200 m) / (total)')
-print(scatterplot1)
+print(violplot1)
+
+spp_coral_clip2 <- spp_coral_clip %>%
+  select(iucn_sid, shallow = shallow_pct) %>%
+  arrange(desc(shallow))
+
+spp_coral_clip2$l_to_h <- c(1:nrow(spp_coral_clip))
+
+spp_coral_clip2 <- spp_coral_clip2 %>%
+  mutate(deep = 100 - shallow) %>%
+  gather(depth, pct_value, c(shallow, deep)) %>%
+  mutate(depth = ifelse(depth == 'shallow', '0 - 200 m', '200 m +'))
+
+break_nums <- seq(0, 100, 20)
+coral_area_plot <- ggplot(spp_coral_clip2, aes(x = l_to_h, y = pct_value, fill = depth)) +
+  ggtheme_plot +
+  # geom_linerange(aes(ymin = kept_ratio, ymax = 100),
+  #                position = position_dodge(width = 0.5)) +
+  geom_area() +
+  scale_fill_manual(values = c('0 - 200 m' = 'grey90', '200 m +' = 'grey10')) +
+  geom_hline(yintercept = break_nums, color = 'white', size = .25) +
+  scale_y_continuous(limits = c(0, 100), breaks = break_nums,
+                     labels = paste0(break_nums, '%')) +
+  # scale_x_discrete(labels = element_blank()) +
+  # theme(axis.title.x = element_blank()) +
+  labs(y = 'IUCN % of coral range',
+       x = 'Count of coral species')
+print(coral_area_plot)
+
+ggsave(plot = coral_area_plot,
+       filename = file.path(dir_fig, 'coral_area_plot.png'), device = 'png',
+       width = 9, height = 5, units = 'cm', dpi = 600)
+
+```
+
+``` {r generate_coral_quadmap}
+
+spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
 
@@
 iucn_coral_fixed <- spp_coralmaps %>%
-  gather(dist, perc, c(sm_perc_raw, sm_perc_clipped)) %>%
-  gather(area, ratio, c(area_ratio_raw, area_ratio_clipped)) %>%
+  gather(dist, perc, c(sm_perc_raw, sm_perc_shallow)) %>%
+  gather(area, ratio, c(area_ratio_raw, area_ratio_shallow)) %>%
   filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%

### 2016-08-12 case_study_mpa/marine_prep.Rmd

--- 2016-02-25 mpa_analysis/marine_prep.Rmd
@@
 ### import OHI regions polygons, to determine spatially-marine areas
-  poly_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters/poly_marine1.shp')
+  poly_marine_file <- file.path(dir_anx, 'vAM_IUCN/mpa_rasters/poly_marine1.shp')
   if(!file.exists(poly_marine_file)) {

### 2016-08-12 case_study_ohi/ohi_full_analysis.Rmd

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

if(basename(getwd()) != 'data_explore') setwd('data_explore')
source('data_explore_fxns.R')
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

```{r get_spp_cells, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

paired_maps_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')

rgn_cells_file <- file.path(dir_git, 'case_study_ohi/rgn_cells.csv')
if(!file.exists(rgn_cells_file)) 
  file.copy(file.path(dir_anx, 'rgns', 'cellID_rgn_gcs.csv'), rgn_cells_file)

ohi_rgn_cells <- read_csv(rgn_cells_file, col_types = 'icdd_d')

spp_map_pairs <- read_csv(paired_maps_file) %>%
  filter(!is.na(popn_category)) %>% ### filter out unassessed species
  filter(!is.na(sm_perc))           ### filter out NA maps

### Load species cells; filter to those in paired maps and those in OHI regions
am_paired_cells <- read_csv(file.path(dir_anx, 'vAM_IUCN/am_spp_cells_2015.csv')) %>%
  filter(am_sid %in% spp_map_pairs$am_sid) %>%
  filter(loiczid %in% ohi_rgn_cells$loiczid)

iucn_paired_cells <- read_csv(file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'),
                           col_types = 'cddddc') %>%
  filter(presence != 5) %>%
  filter(iucn_sid %in% spp_map_pairs$iucn_sid) %>%
  filter(loiczid %in% ohi_rgn_cells$loiczid)

```

``` {r calc_ohi_iucn_paired_spp}

### bind IUCN loiczid dataframe to species risk category dataframe
iucn_paired_cells <- iucn_paired_cells %>%
  left_join(spp_map_pairs %>% select(iucn_sid, cat_score = category_score),
            by = 'iucn_sid')

### group by loiczid, then summarize to find means (and counts)
iucn_cell_sum <- iucn_paired_cells %>%
  filter(!is.na(cat_score)) %>%
  group_by(loiczid) %>%
  summarize(mean_cat = mean(cat_score),
            n_spp    = n())

write_csv(iucn_cell_sum, file.path(dir_git, 'case_study_ohi/iucn_paired_cell_sum.csv'))

### bind cells to regions dataframe and summarize
iucn_rgn_sums <- iucn_cell_sum %>%
  inner_join(ohi_rgn_cells,
             by = 'loiczid') %>%
  # mutate(rgn_area = cell_area * proportionArea,
  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
  mutate(cell_area_weight = cell_area * n_spp * proportionArea,
         area_weighted_mean_cat = mean_cat * cell_area_weight) %>%
  arrange(loiczid)

iucn_rgn_sums <- iucn_rgn_sums %>%
  group_by(rgn_id) %>%
  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))

iucn_rgn_sums <- iucn_rgn_sums %>%
  mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)

write_csv(iucn_rgn_sums, file.path(dir_git, 'case_study_ohi/iucn_paired_region_sums.csv'))

```

``` {r calc_ohi_am_paired_spp_0_thresh}

### bind IUCN loiczid dataframe to species risk category dataframe
am_paired_cells_0 <- am_paired_cells %>%
  left_join(spp_map_pairs %>% select(am_sid, cat_score = category_score),
            by = 'am_sid')

### group by loiczid, then summarize to find means (and counts)
am_cell_sum_0 <- am_paired_cells_0 %>%
  filter(!is.na(cat_score)) %>%
  group_by(loiczid) %>%
  summarize(mean_cat = mean(cat_score),
            n_spp    = n())

write_csv(am_cell_sum_0, file.path(dir_git, 'case_study_ohi/am_paired_cell_sum_0.csv'))

### bind cells to regions dataframe and summarize
am_rgn_sums_0 <- am_cell_sum_0 %>%
  inner_join(ohi_rgn_cells,
             by = 'loiczid') %>%
  # mutate(rgn_area = cell_area * proportionArea,
  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
  mutate(cell_area_weight = cell_area * n_spp * proportionArea,
         area_weighted_mean_cat = mean_cat   * cell_area_weight) %>%
  arrange(loiczid)

am_rgn_sums_0 <- am_rgn_sums_0 %>%
  group_by(rgn_id) %>%
  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))

am_rgn_sums_0 <- am_rgn_sums_0 %>%
  mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)

write_csv(am_rgn_sums_0, file.path(dir_git, 'case_study_ohi/am_paired_region_sums_0.csv'))

```

``` {r calc_ohi_am_paired_spp40_thresh}

### bind IUCN loiczid dataframe to species risk category dataframe
am_paired_cells40 <- am_paired_cells %>%
  filter(prob >= 0.40) %>%
  left_join(spp_map_pairs %>% select(am_sid, cat_score = category_score),
            by = 'am_sid')

### group by loiczid, then summarize to find means (and counts)
am_cell_sum40 <- am_paired_cells40 %>%
  filter(!is.na(cat_score)) %>%
  group_by(loiczid) %>%
  summarize(mean_cat = mean(cat_score),
            n_spp    = n())

write_csv(am_cell_sum40, file.path(dir_git, 'case_study_ohi/am_paired_cell_sum40.csv'))

### bind cells to regions dataframe and summarize
am_rgn_sums40 <- am_cell_sum40 %>%
  inner_join(ohi_rgn_cells,
             by = 'loiczid') %>%
  # mutate(rgn_area = cell_area * proportionArea,
  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
  mutate(cell_area_weight = cell_area * n_spp * proportionArea,
         area_weighted_mean_cat = mean_cat   * cell_area_weight) %>%
  arrange(loiczid)

am_rgn_sums40 <- am_rgn_sums40 %>%
  group_by(rgn_id) %>%
  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))

am_rgn_sums40 <- am_rgn_sums40 %>%
  mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)

write_csv(am_rgn_sums40, file.path(dir_git, 'case_study_ohi/am_paired_region_sums40.csv'))

```

``` {r create_box_plots}
dir_fig <- '~/github/IUCN-AquaMaps/figures'
source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))

iucn_rgn_sums  <- read_csv(file.path(dir_git, 'case_study_ohi/iucn_paired_region_sums.csv'))
am_rgn_sums_0  <- read_csv(file.path(dir_git, 'case_study_ohi/am_paired_region_sums_0.csv'))
am_rgn_sums40 <- read_csv(file.path(dir_git, 'case_study_ohi/am_paired_region_sums40.csv'))

status_df <- iucn_rgn_sums %>%
  select(rgn_id, iucn_st = status) %>%
  full_join(am_rgn_sums_0 %>% select(rgn_id, am_st_0 = status), by = 'rgn_id') %>%
  full_join(am_rgn_sums40 %>% select(rgn_id, am_st40 = status), by = 'rgn_id') %>%
  mutate(diff_am_0 = am_st_0 - iucn_st,
         diff_am40 = am_st40 - iucn_st,
         diff_am40_am0 = am_st_0 - am_st40) %>%
  gather(scenario, diff, diff_am_0:diff_am40_am0) %>%
  mutate(scenario = factor(scenario, levels = c('diff_am40', 'diff_am_0', 'diff_am40_am0')))

bplot_abs <- ggplot(data = status_df, aes(x = scenario, y = diff)) +
  ggtheme_plot + 
#   theme(text = element_text(family = 'Helvetica', color = 'gray30', size = 12),
#         plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
#         legend.position = 'none') +
  geom_hline(yintercept = 0, color = 'gray50', size = .5) +
  geom_violin(fill = 'grey80', alpha = .5,
              draw_quantiles = c(.25, .5, .75)) +
  geom_point(position = position_jitter(w = .3), 
             color = '#4d4dac', alpha = .5) +
  geom_violin(fill = NA, draw_quantiles = c(.25, .5, .75)) +
  labs(x = 'Scenario',
       y = 'Change in SPP status') +
  scale_x_discrete(labels = c('1. AquaMaps \n(40% threshold)', 
                              '2. AquaMaps \n(no threshold)', 
                              '3. AquaMaps \n(40% -> no threshold)'))

print(bplot_abs)
# ggsave(file.path(dir_fig, 'fig3_boxplot_abs_diff.png'), 
#        height = 4.5, width = 8.7, units = 'cm', dpi = 600)
# ggsave(file.path(dir_fig, 'fig3_boxplot_abs_diff.tif'), device = 'tiff',
#        height = 4.5, width = 8.7, units = 'cm', dpi = 600)

```

### 2016-08-12 case_study_ohi/ohi_paired_maps.Rmd

--- 2016-08-02 coralmaps/coral_maps.Rmd
@@
 
-```{r get_bathy_raster, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+```{r get_spp_cells, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
-### get shapefile
-### rasterize to half-degree grid; or raster::extract
+paired_maps_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
 
-if(!file.exists(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))) {
+rgn_cells_file <- file.path(dir_git, 'case_study_ohi/rgn_cells.csv')
+if(!file.exists(rgn_cells_file)) 
+  file.copy(file.path(dir_anx, 'rgns', 'cellID_rgn_gcs.csv'), rgn_cells_file)
 
-  poly_bath <- readOGR(dsn = file.path(path.expand(dir_git), 'coralmaps/ne_10m_bathymetry_K_200'), 
-                       layer = 'ne_10m_bathymetry_K_200', 
-                       stringsAsFactors = TRUE)
+ohi_rgn_cells <- read_csv(rgn_cells_file, col_types = 'icdd_d')
 
-  ### LOICZID cell ID numbers - used in plotting rasters
-  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
-  loiczid_raster       <- raster(loiczid_raster_file)
-  names(loiczid_raster) <- 'loiczid'
+spp_map_pairs <- read_csv(paired_maps_file) %>%
+  filter(!is.na(popn_category)) %>% ### filter out unassessed species
+  filter(!is.na(sm_perc))           ### filter out NA maps
 
-  bath_list <- raster::extract(loiczid_raster, poly_bath, 
-                             weights = TRUE, normalizeWeights = FALSE,
-                             progress = 'text')
-  bath_df <- lapply(bath_list, function(x) {as.data.frame(x)}) %>%
-    bind_rows() %>%
-    rename(loiczid = value) %>%
-    group_by(loiczid) %>%
-    summarize(weight = sum(weight))
+### Load species cells; filter to those in paired maps and those in OHI regions
+am_paired_cells <- read_csv(file.path(dir_anx, 'vAM_IUCN/am_spp_cells_2015.csv')) %>%
+  filter(am_sid %in% spp_map_pairs$am_sid) %>%
+  filter(loiczid %in% ohi_rgn_cells$loiczid)
 
-  ### generate raster?
-  bath_raster <- subs(loiczid_raster, bath_df, by = 'loiczid', which = 'weight')
-  values(bath_raster)[values(bath_raster) > 5] <- NA
-
-  write_csv(bath_df, file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))
-  ### probably want to filter to just 100% 200m+
-}
+iucn_paired_cells <- read_csv(file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'),
+                           col_types = 'cddddc') %>%
+  filter(presence != 5) %>%
+  filter(iucn_sid %in% spp_map_pairs$iucn_sid) %>%
+  filter(loiczid %in% ohi_rgn_cells$loiczid)
 
@@
 
-``` {r}
-### find spp depth limits
-### focus just on corals?
+``` {r calc_ohi_iucn_paired_spp}
 
-if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_depth_raw.csv'))) {
+### bind IUCN loiczid dataframe to species risk category dataframe
+iucn_paired_cells <- iucn_paired_cells %>%
+  left_join(spp_map_pairs %>% select(iucn_sid, cat_score = category_score),
+            by = 'iucn_sid')
 
-  library(jsonlite)
-  api_key <- 'fb71ae836f415f04f41176f6d30c4a9e4cea620d46b9e5021bf2fb142ea51bf5'
+### group by loiczid, then summarize to find means (and counts)
+iucn_cell_sum <- iucn_paired_cells %>%
+  filter(!is.na(cat_score)) %>%
+  group_by(loiczid) %>%
+  summarize(mean_cat = mean(cat_score),
+            n_spp    = n())
 
-  get_from_api <- function(url, param) {
-      api_info <- fromJSON(sprintf(url, param, api_key)) %>%
-        data.frame(stringsAsFactors = FALSE)
-  }
+write_csv(iucn_cell_sum, file.path(dir_git, 'case_study_ohi/iucn_paired_cell_sum.csv'))
 
-  mc_get_from_api <- function(url, param_vec) {
-    numcores = ifelse(Sys.info()[['nodename']] == 'mazu', 12, 1)
-    out_df <- parallel::mclapply(param_vec, 
-                            function(x) get_from_api(url, x),
-                            mc.cores = numcores) %>% 
-      bind_rows() 
-    out_df <- out_df %>%
-      setNames(names(.) %>%
-                 str_replace('result.', ''))
-  }
+### bind cells to regions dataframe and summarize
+iucn_rgn_sums <- iucn_cell_sum %>%
+  inner_join(ohi_rgn_cells,
+             by = 'loiczid') %>%
+  # mutate(rgn_area = cell_area * proportionArea,
+  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
+  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
+  mutate(cell_area_weight = cell_area * n_spp * proportionArea,
+         area_weighted_mean_cat = mean_cat * cell_area_weight) %>%
+  arrange(loiczid)
 
-  ### get species list
-  data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
-  spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-    filter(!is.na(popn_category)) %>% ### filter out unassessed species
-    filter(!is.na(sm_perc))           ### filter out NA maps
+iucn_rgn_sums <- iucn_rgn_sums %>%
+  group_by(rgn_id) %>%
+  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))
 
-  coral_spp_sid <- spp_map_pairs %>%
-    filter(str_detect(spp_group, 'CORAL')) %>%
-    .$iucn_sid
+iucn_rgn_sums <- iucn_rgn_sums %>%
+  mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)
 
-  spp_narr_url <- 'http://apiv3.iucnredlist.org/api/v3/species/narrative/id/%s?token=%s'
-  spp_narr     <- mc_get_from_api(spp_narr_url, coral_spp_sid)
-
-  write_csv(spp_narr, file.path(dir_git, 'coralmaps', 'coral_spp_narratives.csv'))
-
-  spp_depth <- spp_narr %>%
-    select(iucn_sid = species_id, habitat)
-  write_csv(spp_depth, file.path(dir_git, 'coralmaps', 'coral_spp_depth_raw.csv'))
-  ### manually find max depths from this
-}
+write_csv(iucn_rgn_sums, file.path(dir_git, 'case_study_ohi/iucn_paired_region_sums.csv'))
 
@@
 
-``` {r}
+``` {r calc_ohi_am_paired_spp_0_thresh}
 
-spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))
-bath_df <-  read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd') %>%
-  filter(weight >= 1) %>%
-  mutate(deep = 1) %>%
-  select(-weight)
+### bind IUCN loiczid dataframe to species risk category dataframe
+am_paired_cells_0 <- am_paired_cells %>%
+  left_join(spp_map_pairs %>% select(am_sid, cat_score = category_score),
+            by = 'am_sid')
 
-### load iucn spp_cells
-coral_files <- list.files(file.path(dir_anx, 'v2016', 'iucn_intersections'), 
-                          pattern = 'CORAL*.*',
-                          full.names = TRUE)
-iucn_coral_cells <- lapply(coral_files, read_csv) %>%
-  bind_rows() %>%
-  select(iucn_sid, loiczid = LOICZID) %>%
-  filter(iucn_sid %in% spp_depth$iucn_sid) %>%
-  left_join(bath_df, by = 'loiczid') %>%
-  mutate(deep = ifelse(is.na(deep), 0, deep))
+### group by loiczid, then summarize to find means (and counts)
+am_cell_sum_0 <- am_paired_cells_0 %>%
+  filter(!is.na(cat_score)) %>%
+  group_by(loiczid) %>%
+  summarize(mean_cat = mean(cat_score),
+            n_spp    = n())
 
-write_csv(iucn_coral_cells, file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv'))
+write_csv(am_cell_sum_0, file.path(dir_git, 'case_study_ohi/am_paired_cell_sum_0.csv'))
+
+### bind cells to regions dataframe and summarize
+am_rgn_sums_0 <- am_cell_sum_0 %>%
+  inner_join(ohi_rgn_cells,
+             by = 'loiczid') %>%
+  # mutate(rgn_area = cell_area * proportionArea,
+  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
+  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
+  mutate(cell_area_weight = cell_area * n_spp * proportionArea,
+         area_weighted_mean_cat = mean_cat   * cell_area_weight) %>%
+  arrange(loiczid)
+
+am_rgn_sums_0 <- am_rgn_sums_0 %>%
+  group_by(rgn_id) %>%
+  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))
+
+am_rgn_sums_0 <- am_rgn_sums_0 %>%
+  mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)
+
+write_csv(am_rgn_sums_0, file.path(dir_git, 'case_study_ohi/am_paired_region_sums_0.csv'))
 
@@
 
-``` {r}
-### calc areas
-### read hcaf to get areas; left join to iucn_coral_cells
-### group_by iucn_sid and depth, summarize area (deep and shallow)
-### group_by iucn_sid, summarize total area
-iucn_coral_cells <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv'))
+``` {r calc_ohi_am_paired_spp40_thresh}
 
-hcaf <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
-  select(loiczid, cellarea, oceanarea)
+### bind IUCN loiczid dataframe to species risk category dataframe
+am_paired_cells40 <- am_paired_cells %>%
+  filter(prob >= 0.40) %>%
+  left_join(spp_map_pairs %>% select(am_sid, cat_score = category_score),
+            by = 'am_sid')
 
-iucn_coral_areas <- iucn_coral_cells %>%
-  left_join(hcaf, by = 'loiczid') %>%
-  group_by(iucn_sid, deep) %>%
-  summarize(deeparea = sum(oceanarea)) %>%
-  group_by(iucn_sid) %>%
-  mutate(totarea = sum(deeparea),
-         shallow_pct = (totarea - deeparea)/totarea) %>%
-  filter(deep == 1) %>%
-  select(-deep)
+### group by loiczid, then summarize to find means (and counts)
+am_cell_sum40 <- am_paired_cells40 %>%
+  filter(!is.na(cat_score)) %>%
+  group_by(loiczid) %>%
+  summarize(mean_cat = mean(cat_score),
+            n_spp    = n())
 
-  ### get species list
-  data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
-  spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-    filter(!is.na(popn_category)) %>%   ### filter out unassessed species
-    filter(!is.na(sm_perc))             ### filter out NA maps
+write_csv(am_cell_sum40, file.path(dir_git, 'case_study_ohi/am_paired_cell_sum40.csv'))
 
-iucn_coral_areas <- iucn_coral_areas %>%
-  left_join(spp_depth,     by = 'iucn_sid') %>%
-  left_join(spp_map_pairs, by = 'iucn_sid')
+### bind cells to regions dataframe and summarize
+am_rgn_sums40 <- am_cell_sum40 %>%
+  inner_join(ohi_rgn_cells,
+             by = 'loiczid') %>%
+  # mutate(rgn_area = cell_area * proportionArea,
+  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
+  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
+  mutate(cell_area_weight = cell_area * n_spp * proportionArea,
+         area_weighted_mean_cat = mean_cat   * cell_area_weight) %>%
+  arrange(loiczid)
 
-write_csv(iucn_coral_areas, file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv'))
+am_rgn_sums40 <- am_rgn_sums40 %>%
+  group_by(rgn_id) %>%
+  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))
+
+am_rgn_sums40 <- am_rgn_sums40 %>%
+  mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)
+
+write_csv(am_rgn_sums40, file.path(dir_git, 'case_study_ohi/am_paired_region_sums40.csv'))
 
@@
 
-``` {r}
-if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv'))) {
+``` {r create_box_plots}
+dir_fig <- '~/github/IUCN-AquaMaps/figures'
+source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
 
-  ### get species list
-  iucn_coral_cells_clipped <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_cells.csv')) %>%
-    filter(deep == 0) %>%
-    mutate(prop_area = 1)
+iucn_rgn_sums  <- read_csv(file.path(dir_git, 'case_study_ohi/iucn_paired_region_sums.csv'))
+am_rgn_sums_0  <- read_csv(file.path(dir_git, 'case_study_ohi/am_paired_region_sums_0.csv'))
+am_rgn_sums40 <- read_csv(file.path(dir_git, 'case_study_ohi/am_paired_region_sums40.csv'))
 
-  spp_coralmaps <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
-    filter(iucn_sid %in% iucn_coral_cells_clipped$iucn_sid) %>%
-    select(am_sid, sciname, iucn_sid)
+status_df <- iucn_rgn_sums %>%
+  select(rgn_id, iucn_st = status) %>%
+  full_join(am_rgn_sums_0 %>% select(rgn_id, am_st_0 = status), by = 'rgn_id') %>%
+  full_join(am_rgn_sums40 %>% select(rgn_id, am_st40 = status), by = 'rgn_id') %>%
+  mutate(diff_am_0 = am_st_0 - iucn_st,
+         diff_am40 = am_st40 - iucn_st,
+         diff_am40_am0 = am_st_0 - am_st40) %>%
+  gather(scenario, diff, diff_am_0:diff_am40_am0) %>%
+  mutate(scenario = factor(scenario, levels = c('diff_am40', 'diff_am_0', 'diff_am40_am0')))
 
-  ### copy aquamaps cells to local file, keeping only corals in the paired maps
-  if(!file.exists(file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))) {
-    am_coral_cells <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv', 'hcaf_sp_native_trunc.csv')) %>%
-      rename(am_sid = speciesid, prob = probability) %>%
-      filter(am_sid %in% spp_coralmaps$am_sid) 
-    write_csv(am_coral_cells, file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))
-  }
+bplot_abs <- ggplot(data = status_df, aes(x = scenario, y = diff)) +
+  ggtheme_plot + 
+  theme(panel.grid.major = element_line(colour = 'grey90', size = .25),
+        panel.grid.major.x = element_blank()) +
 
-  am_coral_cells <- read_csv(file.path(dir_git, 'coralmaps', 'am_coral_cells.csv'))
+#   theme(text = element_text(family = 'Helvetica', color = 'gray30', size = 12),
+#         plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
+#         legend.position = 'none') +
+  geom_hline(yintercept = 0, color = 'gray50', size = .5) +
+  geom_violin(fill = 'grey80', alpha = .5,
+              draw_quantiles = c(.25, .5, .75)) +
+  geom_point(position = position_jitter(w = .3), 
+             color = '#4d4dac', alpha = .5) +
+  geom_violin(fill = NA, draw_quantiles = c(.25, .5, .75)) +
+  labs(x = 'Scenario',
+       y = 'Change in SPP status') +
+  scale_x_discrete(labels = c('1. AquaMaps \n(40% threshold)', 
+                              '2. AquaMaps \n(no threshold)', 
+                              '3. AquaMaps \n(40% -> no threshold)'))
 
-  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
-  loiczid_raster       <- raster(loiczid_raster_file)
-  names(loiczid_raster) <- 'loiczid'
+print(bplot_abs)
+ggsave(file.path(dir_fig, 'fig3_boxplot2_abs_diff.png'),
+       height = 4.5, width = 8.7, units = 'cm', dpi = 600)
+ggsave(file.path(dir_fig, 'fig3_boxplot2_abs_diff.tif'), device = 'tiff',
+       height = 4.5, width = 8.7, units = 'cm', dpi = 600)
 
-  spp_coralmaps <- spp_coralmaps %>%
-    mutate(area_total = NA,  area_am = NA,  area_iucn = NA,
-           area_overlap = NA,
-           sm_perc    = NA, #filled with % of smaller range that is within the larger range
-           sm_range   = NA) #what range is smaller
-
-  i         <- 0 ### initialize the counting index
-  am_thresh <- 0 ### set the aquamaps presence threshold
-  sid_vector <- unique(spp_coralmaps$iucn_sid)
-  message(sprintf('There are a total of %s distinct species, by IUCN SID', 
-                  length(sid_vector)))
-
-  for (sid in sid_vector) { 
-    # sid <- spp_coralmaps$iucn_sid[1]
-    i <- i + 1
-    sids <- spp_coralmaps %>% 
-      select(am_sid, iucn_sid, sciname) %>%
-      filter(iucn_sid == sid)
-    rowindex <- (spp_coralmaps$iucn_sid == sid)
-
-    message(sprintf('%s. Processing species iucn_sid: %s \n   am_sid(s):  %s \n   sciname(s): %s', 
-                    i, sid, paste(sids$am_sid, collapse = ', '), paste(sids$sciname, collapse = ' ')))
-    map_iucn <- get_spp_map(sid, iucn_coral_cells_clipped) %>%
-      mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0))
-    map_am   <- get_spp_map(sids$am_sid, am_coral_cells) %>%
-      mutate(am_pres = ifelse(am_prob >= am_thresh, 1, 0))
-
-    ### rasterize aquamaps map
-    if(any(duplicated(map_am$loiczid))) message('Duplicated cells in am spp map')
-    map_am <- map_am %>%
-      dplyr::select(loiczid, am_pres) %>% 
-      unique()
-
-    r_am_spp <- subs(x = loiczid_raster, y = map_am %>% select(loiczid, am_pres), 
-                     by = 'loiczid', which = 'am_pres', 
-                     subsWithNA = TRUE)
-
-    ### rasterize IUCN map
-    if(TRUE %in% duplicated(map_iucn$loiczid)) stop('Duplicated cells in IUCN map')
-    r_iucn_spp <- subs(loiczid_raster, map_iucn %>% select(loiczid, iucn_pres), 
-                       by = 'loiczid', which = 'iucn_pres', 
-                       subsWithNA = TRUE)
-
-    ### Start calculating areas etc.
-    area_iucn <- area(r_iucn_spp, na.rm = TRUE) %>%
-      cellStats(., stat = 'sum')
-    area_am   <- area(r_am_spp, na.rm = TRUE) %>%
-      cellStats(., stat = 'sum')
-    spp_coralmaps[rowindex, ]$area_iucn <- area_iucn
-    spp_coralmaps[rowindex, ]$area_am   <- area_am
-
-    ### Get total area (IUCN and AM combined)
-    all_cells <- data.frame(loiczid = unique(c(map_iucn$loiczid, map_am$loiczid)),
-                           spp_pres = 1)
-    r_all_cells <- subs(loiczid_raster, all_cells, 
-                       by = 'loiczid', which = 'spp_pres', 
-                       subsWithNA = TRUE)
-    area_total  <- area(r_all_cells, na.rm = TRUE) %>% cellStats(., stat = 'sum')
-    spp_coralmaps[rowindex, ]$area_total <- area_total
-
-    ### Get total overlap (IUCN and AM together). If area_iucn = 0, just assign 0
-    if(area_iucn != 0) {
-      overlap_cells <- unique(map_iucn$loiczid[map_iucn$loiczid %in% map_am$loiczid])
-      if(length(overlap_cells) > 0) {
-        overlap_cells <- data.frame(loiczid = overlap_cells,
-                                  spp_pres = 1)
-        r_overlap_cells <- subs(loiczid_raster, overlap_cells, 
-                                by = 'loiczid', which = 'spp_pres', 
-                                subsWithNA = TRUE)
-        area_overlap <- area(r_overlap_cells, na.rm = TRUE) %>% cellStats(., stat = 'sum')
-      } else area_overlap <- 0 ### if no overlapping cells, assign zero
-    } else area_overlap <- 0 ### if no IUCN map cells, assign zero
-    spp_coralmaps[rowindex, ]$area_overlap <- area_overlap
-
-    message(sprintf('   Species iucn_sid %s: smaller range is %s. Smaller-in-larger is %.2f%%, A_sm:A_lg is %.2f%%.',
-                    sid, ifelse(area_am < area_iucn, 'AquaMaps', 'IUCN'), 
-                    ifelse(area_am < area_iucn, area_overlap/area_am*100, area_overlap/area_iucn*100),
-                    ifelse(area_am < area_iucn, area_am/area_iucn*100, area_iucn/area_am*100)))
-  }
-
-  spp_coralmaps <- spp_coralmaps %>%
-    mutate(sm_range   = ifelse(area_am < area_iucn, 'AM', 'IUCN'),
-           sm_perc    = ifelse(area_am < area_iucn, (area_overlap / area_am) * 100, (area_overlap / area_iucn) * 100),
-           area_ratio = area_am / area_iucn,
-           area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
-           area_ratio = area_ratio * 100)
-
-  spp_coralmaps <- spp_coralmaps %>%
-    mutate(lg_area = ifelse(area_am > area_iucn, area_am, area_iucn),
-           sm_area = ifelse(area_am < area_iucn, area_am, area_iucn),
-           lg_area_pct = 100 * lg_area / max(area_total, na.rm = TRUE)) %>% 
-             ### use largest data set as approx for total ocean range
-    select(-lg_area, -sm_area)
-
-  spp_coralmaps <- spp_coralmaps %>%
-    mutate(area_total  = round(area_total, 1),
-           area_am     = round(area_am, 1),
-           area_iucn   = round(area_iucn, 1),
-           sm_perc     = round(sm_perc, 3),
-           area_ratio  = round(area_ratio, 3),
-           lg_area_pct = round(lg_area_pct, 3))
-
-  print(head(spp_coralmaps))
-  write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv'))
-}
 ```
-
-``` {r}
-### combine old and new area dataframes
-spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_with_clipped_area.csv')) %>%
-  select(iucn_sid, am_sid, sciname, area_am, 
-         area_iucn_clipped = area_iucn, area_overlap_clipped = area_overlap, 
-         sm_perc_clipped = sm_perc, sm_range_clipped = sm_range, area_ratio_clipped = area_ratio) %>%
-  left_join(read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
-              select(am_sid, iucn_sid, 
-                     area_iucn_raw = area_iucn, area_overlap_raw = area_overlap, 
-                     sm_perc_raw = sm_perc, sm_range_raw = sm_range, area_ratio_raw = area_ratio),
-            by = c('iucn_sid', 'am_sid'))
-
-write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
-```
-
-``` {r}
-
-spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
-
-### check raw iucn_area vs clipped - looks proportional; scaling factor due to difference in area calcs?
-scatterplot1 <- ggplot(spp_coralmaps, aes(x = area_iucn_raw, y = area_iucn_clipped)) +
-  geom_point() +
-  geom_abline(color = 'red')
-scatterplot1
-
-### rough recalcs
-### assume that by removing the deep cells, we will not remove cells that are
-### included in AquaMaps, since AquaMaps already filters by depth.  This means
-### the overlap area will not be affected.
-iucn_coral_fixed <- spp_coralmaps %>%
-  gather(dist, perc, c(sm_perc_raw, sm_perc_clipped)) %>%
-  gather(area, ratio, c(area_ratio_raw, area_ratio_clipped)) %>%
-  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
-  filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
-  mutate(method = ifelse(str_detect(dist, 'clip'), '200m clip', 'all depth'))
-
-clipped_quads <- ggplot(iucn_coral_fixed %>%
-                      filter(!iucn_sid %in% (iucn_coral_fixed %>% 
-                                               filter(perc > 100) %>% 
-                                               .$iucn_sid)),
-                    aes(x = ratio, y = perc, group = iucn_sid)) +
-  geom_line(color = 'grey50', size = .5, alpha = .3) +
-  geom_point(aes(color = method), size = 2) +
-  labs(x = 'Area ratio (smaller/larger)',
-       y = 'Distribtion (%smaller in larger)',
-       title = 'Coral species alignment, IUCN full range vs depth-clipped')
-
-clipped_quads
-```

### 2016-08-12 coralmaps/coral_maps.Rmd

@@
 
-if(basename(getwd()) != 'data_explore') setwd('data_explore')
-source('data_explore_fxns.R')
+source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
 ### When knitting this, it automatically sets WD to be this directory...
@@
 # [1] 5
-### of 562 species, only 5 are listed as occurring below 100 m; none are below 200 m
+### of 562 species, only 5 are listed as occurring below 100 m; none are below 200 m.
+### NOTE: Many of these 562 species have no maps in IUCN; they are mapped in AquaMaps though.
 
@@
 
-### 17 species (of 396) have no depth information; one is listed as being dredged from 13 fathoms (~ 25 m)
+### 18 species (of 562) have no depth information; one is listed as being dredged from 13 fathoms (~ 25 m)
 ```
@@
 bath_df <-  read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd') %>%
-  filter(weight >= 1) %>%
+  filter(weight >= 1) %>% ### cell must be 100% within the deep polygon to count as deep
   mutate(deep = 1) %>%
@@
 write_csv(coral_cells_iucn, file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))
-
+length(unique(coral_cells_iucn$iucn_sid))
+### NOTE: only 473 coral species with IUCN maps
 ```
@@
 ### group_by iucn_sid, summarize total area
-coral_cells_iucn <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))
+coral_cells_iucn <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
+  mutate(deep = ifelse(deep == 1, 'deep', 'shallow'))
 
@@
   group_by(iucn_sid, deep) %>%
-  summarize(deeparea = sum(oceanarea)) %>%
+  summarize(area = sum(oceanarea)) %>%
   group_by(iucn_sid) %>%
-  mutate(totarea = sum(deeparea),
-         shallow_pct = (totarea - deeparea)/totarea) %>%
-  arrange(desc(deep)) %>%
-  .[!duplicated(.$iucn_sid), ]
+  mutate(totarea = sum(area),
+         pct_area = area/totarea) %>%
+  filter(deep == 'shallow') %>%
+  rename(shallow_pct = pct_area) %>%
+  mutate(deep_pct = 1 - shallow_pct)
 
@@
   coral_cells_iucn_shallow <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
-    filter(deep == 0) %>%
-    mutate(prop_area = 1)
+    filter(deep == 0)
 
@@
 }
+
 ```
@@
 spp_coral_clip <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv')) %>%
-  select(iucn_sid, shallow_pct, totarea, deeparea) %>%
-  mutate(shallow_pct = 100 * shallow_pct)
+  select(iucn_sid, shallow_pct, deep_pct, totarea, shallow_area = area) %>%
+  mutate(shallow_pct = 100 * shallow_pct,
+         deep_pct    = 100 * deep_pct)
 
@@
   filter(shallow_pct > 80) 
-### only five with such small effects; one of those did not change at all.
+### only four out of 481 with such small effects; one of those did not change at all.
 
@@
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-  # 23.80   52.89   55.48   56.33   60.59  100.00
+  # 22.40   48.20   51.71   52.37   56.68  100.00
 
-scatterplot1 <- ggplot(spp_coral_clip, aes(x = totalarea, y = deeparea)) +
+scatterplot1 <- ggplot(spp_coral_clip, aes(x = totarea, y = shallow_area)) +
   geom_point(size = 0.25, color = 'black') +
@@
 
+### for area plot, arrange from highest shallow pct to lowest (best to worst)
+spp_coral_clip <- spp_coral_clip %>%
+  arrange(desc(shallow_pct))
+
+spp_coral_clip$l_to_h <- c(1:nrow(spp_coral_clip))
+
 spp_coral_clip2 <- spp_coral_clip %>%
-  select(iucn_sid, shallow = shallow_pct) %>%
-  arrange(desc(shallow))
-
-spp_coral_clip2$l_to_h <- c(1:nrow(spp_coral_clip))
-
-spp_coral_clip2 <- spp_coral_clip2 %>%
-  mutate(deep = 100 - shallow) %>%
-  gather(depth, pct_value, c(shallow, deep)) %>%
-  mutate(depth = ifelse(depth == 'shallow', '0 - 200 m', '200 m +'))
+  gather(depth, pct_value, c(shallow_pct, deep_pct)) %>%
+  mutate(depth = ifelse(depth == 'shallow_pct', '0 - 200 m', '200 m +'))
 
@@
   geom_hline(yintercept = break_nums, color = 'white', size = .25) +
-  scale_y_continuous(limits = c(0, 100), breaks = break_nums,
+  scale_y_continuous(# limits = c(0, 100), ### why is this causing a vertical line?
+                     breaks = break_nums,
                      labels = paste0(break_nums, '%')) +
@@
   filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
-  filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
-  mutate(method = ifelse(str_detect(dist, 'clip'), '200 m clip', 'original'))
-
-source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
+  filter(!(str_detect(dist, 'shallow') & !(str_detect(area, 'shallow')))) %>%
+  mutate(method = ifelse(str_detect(dist, 'shallow'), '200 m clip', 'original'))
 

### 2016-08-12 data_explore/data_prep.Rmd

@@
 ### some species with no cells reported.  2279 listed in the am v iucn file.
-write_csv(iucn_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'))
+write_csv(iucn_spp_cells, file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'))
 
@@
 ### length(unique(am_spp_cells$am_sid)) results in 2314 species found - great!
-write_csv(am_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'))
+write_csv(am_spp_cells, file.path(dir_anx, 'vAM_IUCN/am_spp_cells_2015.csv'))
 
@@
   if(!exists('am_spp_cells'))
-    am_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'))
+    am_spp_cells <- read_csv(file.path(dir_anx, 'vAM_IUCN/am_spp_cells_2015.csv'))
   if(!exists('iucn_spp_cells'))
-    iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'),
+    iucn_spp_cells <- read_csv(file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'),
                                col_types = 'cddddc')

### 2016-08-12 ms_3_res_disc.Rmd

@@
 
-___Distribution alignment - overlap as proxy for common understanding; use Bayesian argument to show that these areas are high-confidence areas; area ratio as proxy for errors of commission/omission.
+___Distribution alignment - overlap as proxy for common understanding; these areas are high-confidence areas.
+Area ratio as proxy for errors of commission/omission.
 While we cannot know which dataset is closer to the "truth" we can make inferences based on expectations of EOO and AOO maps - is AquaMaps AOO?___
 
-_Distribution alignment - Bayesian description of higher confidence_
+_Distribution alignment - higher confidence when both datasets predict presence (or absence) in the same area_
 _Area ratio is defined as the ratio of the smaller range area to the larger range area.

### 2016-08-12 case_study_mpa/3_mpa_spp_analysis.Rmd
cleaning up plots

@@
 
+break_nums <- seq(0, 100, 20)
+
 ### Plot the bar chart
@@
   theme(panel.grid.major.y = element_blank()) +
-  geom_bar(stat = 'count', alpha = .8) +
+  geom_bar(stat = 'count', alpha = 1) +
   scale_fill_manual(values = c('>10%'  = '#4dac26', 
@@
                                'gap'   = '#d01c8b')) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 125), breaks = c(seq(0, 100, 20))) +
+  ### add grid lines; horizontal but then get flipped
+  geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 125), breaks = break_nums,
+                     labels = sprintf('%s%%', break_nums)) + 
   geom_text(aes(label = sprintf('n = %s', n_trial), y = 101), hjust = 0, 

### 2016-08-12 case_study_ohi/ohi_full_datasets.Rmd
cleaning up plots

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico')
dir_fig <- file.path(dir_git, 'figures')
dir_ohi_data <- file.path(dir_git, 'case_study_ohi/ohi_global_data')

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))

```

### NOTE:

This analysis depends on data generated by the full Ocean Health Index Species goal scripts and data.  The data loaded here was output from multiple runs of that original script.

The original OHI Species goal script(s) can be found in the `github/ohi-science/ohiprep` repository; browse to the `globalprep/spp_ico/v2016` folder for the `data_prep_spp.Rmd` and supporting files, scripts, and intermediate data.

``` {r load_data_from_global_analyses}

st_base  <- read_csv(file.path(dir_ohi_data, 'spp_status_global_IUCNpref_prob0.4.csv'))
st_scen1 <- read_csv(file.path(dir_ohi_data, 'spp_status_global_IUCNpref_prob0.csv'))
st_scen2 <- read_csv(file.path(dir_ohi_data, 'spp_status_global_AMpref_prob0.4.csv'))
st_scen3 <- read_csv(file.path(dir_ohi_data, 'spp_status_global_AMpref_prob0.csv'))

### scenario 1: prefer IUCN data; drop AM probability threshold to 0.01
### scenario 2: prefer AM data; keep AM prob thresh at 0.40
### scenario 3: prefer AM data; drop AM prob thresh to 0.01

st_all <- st_base %>% rename(st_b = score) %>%
  left_join(st_scen1 %>% rename(st_1 = score), by = 'rgn_id') %>%
  left_join(st_scen2 %>% rename(st_2 = score), by = 'rgn_id') %>%
  left_join(st_scen3 %>% rename(st_3 = score), by = 'rgn_id')

st_all <- st_all %>% mutate(absdiff1 = (st_1 - st_b)*100,
                            absdiff2 = (st_2 - st_b)*100,
                            absdiff3 = (st_3 - st_b)*100) %>%
  gather(scenario, difference, absdiff1:absdiff3)

```

``` {r build_violin_plots}
viol_plot_global <- ggplot(data = st_all %>% filter(str_detect(scenario, 'abs')),
                    aes(x = scenario, y = difference)) +
  ggtheme_plot + 
  geom_hline(yintercept = 0, color = 'gray50', size = .5) +
  geom_violin(fill = 'grey80', alpha = .5,
              scale = 'width',
              draw_quantiles = c(.25, .5, .75)) +
  geom_point(position = position_jitter(w = .3), 
             color = 'black', alpha = 1, size = 0.25) +
  geom_violin(fill = NA, 
              draw_quantiles = c(.25, .5, .75),
              scale = 'width') +
  labs(x = 'Scenario',
       y = 'Change in SPP status') +
  scale_x_discrete(labels = c('4. IUCN/AquaMaps\nno threshold', 
                              '5. AquaMaps/IUCN\n40% threshold', 
                              '6. AquaMaps/IUCN\nno threshold'))

print(viol_plot_global)

ggsave(file.path(dir_fig, 'fig3b_boxplot_ohi_global.png'), 
       height = 4.5, width = 8.7, units = 'cm', dpi = 600)
ggsave(file.path(dir_fig, 'fig3b_boxplot_ohi_global.tif'), device = 'tiff',
       height = 4.5, width = 8.7, units = 'cm', dpi = 600)

### 2016-08-12 case_study_ohi/ohi_paired_maps.Rmd
cleaning up plots

@@
 
-if(basename(getwd()) != 'data_explore') setwd('data_explore')
-source('data_explore_fxns.R')
-### When knitting this, it automatically sets WD to be this directory...
-### the 'setwd()' is there for running by chunk
+source('~/github/IUCN-AquaMaps/data_explore/data_explore_fxns.R')
+
 ```
@@
 
-#   theme(text = element_text(family = 'Helvetica', color = 'gray30', size = 12),
-#         plot.title = element_text(size = rel(1.5), hjust = 0, face = 'bold'),
-#         legend.position = 'none') +
   geom_hline(yintercept = 0, color = 'gray50', size = .5) +
   geom_violin(fill = 'grey80', alpha = .5,
+              scale = 'width',
               draw_quantiles = c(.25, .5, .75)) +
   geom_point(position = position_jitter(w = .3), 
-             color = '#4d4dac', alpha = .5) +
-  geom_violin(fill = NA, draw_quantiles = c(.25, .5, .75)) +
+             color = 'black', alpha = 1, size = 0.25) +
+  geom_violin(fill = NA, 
+              draw_quantiles = c(.25, .5, .75),
+              scale = 'width') +
   labs(x = 'Scenario',
@@
 print(bplot_abs)
-ggsave(file.path(dir_fig, 'fig3_boxplot2_abs_diff.png'),
+ggsave(file.path(dir_fig, 'fig3a_boxplot_paired_maps.png'),
        height = 4.5, width = 8.7, units = 'cm', dpi = 600)
-ggsave(file.path(dir_fig, 'fig3_boxplot2_abs_diff.tif'), device = 'tiff',
+ggsave(file.path(dir_fig, 'fig3a_boxplot_paired_maps.tif'), device = 'tiff',
        height = 4.5, width = 8.7, units = 'cm', dpi = 600)

### 2016-08-12 coralmaps/coral_maps.Rmd
cleaning up plots

@@
 break_nums <- seq(0, 100, 20)
+num_spp <- max(spp_coral_clip2$l_to_h)
 coral_area_plot <- ggplot(spp_coral_clip2, aes(x = l_to_h, y = pct_value, fill = depth)) +
   ggtheme_plot +
-  # geom_linerange(aes(ymin = kept_ratio, ymax = 100),
-  #                position = position_dodge(width = 0.5)) +
-  geom_area() +
+  geom_area(show.legend = FALSE) +
   scale_fill_manual(values = c('0 - 200 m' = 'grey90', '200 m +' = 'grey10')) +
   geom_hline(yintercept = break_nums, color = 'white', size = .25) +
+  annotate("text", x = (num_spp - 10), y = 95, hjust = 1, vjust = .5, size = 3, color = 'white', 
+           fontface = 'bold.italic', label = "200 m and deeper") + 
+  annotate("text", x = 10, y = 5, hjust = 0, vjust = .5, size = 3, color = 'grey10', 
+           fontface = 'bold.italic', label = "0 - 200 m") + 
   scale_y_continuous(# limits = c(0, 100), ### why is this causing a vertical line?
@@
                      labels = paste0(break_nums, '%')) +
-  # scale_x_discrete(labels = element_blank()) +
-  # theme(axis.title.x = element_blank()) +
+  scale_x_continuous(limits = c(0, num_spp), ### why is this causing a vertical line?
+                     breaks = c(seq(0, 400, 100), num_spp)) +
   labs(y = 'IUCN % of coral range',
-       x = 'Count of coral species')
+       x = 'Count of IUCN coral species maps')
 print(coral_area_plot)

### 2016-08-12 data_explore/data_explore.Rmd
cleaning up plots

@@
 
-``` {r plot bar charts by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
-### bar charts comparing number of species in each class for each data set
-
-#library(colorspace)
-#hclvec <- heat_hcl(3); colvec <- c('AquaMaps' = hclvec[1], 'Both' = hclvec[2], 'IUCN' = hclvec[3])
-#colvec <- c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')
-colvec <- c('AquaMaps' = '#76A94C', 'Both' = '#e18258', 'IUCN' = '#9168a9')
-
-taxa_df_long <- taxa_df %>%
-  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
-  gather(dataset, n_spp, AquaMaps:Both) %>%
-  group_by(class) %>%
-  mutate(n_total = sum(n_spp, na.rm = TRUE),
-         percent = n_spp/n_total * 100,
-         dataset = factor(dataset, levels = c('AquaMaps', 'Both', 'IUCN'), ordered = TRUE)) %>%
-  arrange(desc(dataset))
-
-break_nums <- seq(0, 100, 20)
-
-bar_taxa <- ggplot(taxa_df_long, aes(x = class, y = percent, fill = dataset)) +
-  ggtheme_plot + 
-  theme(panel.grid.major.x = element_blank()) +
-  geom_bar(stat = 'identity', alpha = 1) +
-  geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
-  scale_fill_manual(values = colvec) +
-  geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
-            size = 2, 
-            angle = 90, hjust = 0, color = 'grey30') +
-  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
-  scale_y_continuous(expand = c(0, 0), 
-                     limits = c(0, 130),
-                     breaks = break_nums,
-                     labels = sprintf('%s%%', break_nums)) + 
-  labs(# title = 'Species maps by taxonomic class',
-       x = 'Class', 
-       y = 'Percent of mapped species',
-       fill = 'Dataset')
-
-print(bar_taxa)
-ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
-       width = 17.8, height = 7, units = 'cm', dpi = 600)
+``` {r plot_bar_charts_by_taxa_old, echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE}
+# ### bar charts comparing number of species in each class for each data set
+# 
+# #library(colorspace)
+# #hclvec <- heat_hcl(3); colvec <- c('AquaMaps' = hclvec[1], 'Both' = hclvec[2], 'IUCN' = hclvec[3])
+# #colvec <- c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')
+# colvec <- c('AquaMaps' = '#76A94C', 'Both' = '#e18258', 'IUCN' = '#9168a9')
+# 
+# taxa_df_long <- taxa_df %>%
+#   dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
+#   gather(dataset, n_spp, AquaMaps:Both) %>%
+#   group_by(class) %>%
+#   mutate(n_total = sum(n_spp, na.rm = TRUE),
+#          percent = n_spp/n_total * 100,
+#          dataset = factor(dataset, levels = c('AquaMaps', 'Both', 'IUCN'), ordered = TRUE)) %>%
+#   arrange(desc(dataset))
+# 
+# break_nums <- seq(0, 100, 20)
+# 
+# bar_taxa <- ggplot(taxa_df_long, aes(x = class, y = percent, fill = dataset)) +
+#   ggtheme_plot + 
+#   theme(panel.grid.major.x = element_blank()) +
+#   geom_bar(stat = 'identity', alpha = 1) +
+#   geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
+#   scale_fill_manual(values = colvec) +
+#   geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
+#             size = 2, 
+#             angle = 90, hjust = 0, color = 'grey30') +
+#   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
+#   scale_y_continuous(expand = c(0, 0), 
+#                      limits = c(0, 130),
+#                      breaks = break_nums,
+#                      labels = sprintf('%s%%', break_nums)) + 
+#   labs(# title = 'Species maps by taxonomic class',
+#        x = 'Class', 
+#        y = 'Percent of mapped species',
+#        fill = 'Dataset')
+# 
+# print(bar_taxa)
+# ggsave(filename = file.path(dir_fig, 'fig1a_bar_by_taxa_all.png'), 
+#        width = 17.8, height = 7, units = 'cm', dpi = 600)
 
@@
 
-![](../figures/bar_by_taxa_all.png)
-
-``` {r plot bar-line charts by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
+``` {r plot_barline_charts_by_taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
 ### bar charts comparing number of species in each class for each data set
-
-#library(colorspace)
-#hclvec <- heat_hcl(3); colvec <- c('AquaMaps' = hclvec[1], 'Both' = hclvec[2], 'IUCN' = hclvec[3])
-#colvec <- c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')
-colvec <- c('AquaMaps' = '#76A94C', 'Both' = '#e18258', 'IUCN' = '#9168a9')
 
@@
                      labels = sprintf('%s%%', break_nums)) +
-  labs(# title = 'Species maps by taxonomic class',
-       x = 'Class',
+  labs(x = 'Class',
        y = 'Percent of mapped species',
@@
 print(bar_taxa)
-ggsave(filename = file.path(dir_fig, 'bar2_by_taxa_all.png'), 
+ggsave(filename = file.path(dir_fig, 'fig1a_bar_by_taxa_all.png'), 
        width = 17.8, height = 7, units = 'cm', dpi = 600)
@@
 
-ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
+ggsave(file.path(dir_fig, 'fig2a_scatter_quadmap.png'),
        width = 17.8, height = 9, units = 'cm', dpi = 600)
@@
 
-# print(barchart_spp_gp_quads)
+print(barchart_spp_gp_quads)
+
 ggsave(plot = barchart_spp_gp_quads,
-       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
+       filename = file.path(dir_fig, 'fig2b_barchart_spp_gp_quads.png'),
        width = 17.8, height = 7, units = 'cm', dpi = 600)

### 2016-08-12 ms_figures.Rmd
cleaning up plots

@@
 
+![Change in MPA coverage](figures/bar2_by_taxa_all.png)
+
 ### Fig. 2
@@
 
+### New figures - corals
+
+![Change in MPA coverage](figures/coral_area_plot.png)
+
+![Change in MPA coverage](figures/scatter_quadmap_corals.png)
+
 -----

### 2016-08-15 manuscript-main.Rmd
pushing edits so I can edit manuscript locally

--- 2016-04-02 manuscript-main.Rmd
@@
   S. Halpern^1,2,3^
+csl: refs/journal-of-applied-ecology.csl
 output:
@@
 ``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
+knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
+                      echo = FALSE, message = FALSE, warning = FALSE)
 
@@
 
-``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
+``` {r load data}
 ### use data_explore.Rmd to create this file, if it does not yet exist.  This
@@
 
-``` {r calc species numbers, echo = FALSE}
+``` {r calc species numbers}
 n_spp_am     <- nrow(spp_am)
@@
 
-``` {r abstract, child = 'ms_1_abst_sig.Rmd'}
+``` {r abstract, child = 'ms_0_abst_j_appl_ecol.Rmd'}
 ```
@@
 
-\\body
+``` {r intro, child = 'ms_1_intro.Rmd'}
+```
 
-``` {r intro_results_discussion, child = 'ms_2_intro_res_disc.Rmd'}
+``` {r methods, child = 'ms_2_methods.Rmd'}
+```
+
+``` {r results_discussion, child = 'ms_3_res_disc.Rmd'}
 ```

### 2016-08-15 ms_1_intro.Rmd
pushing edits so I can edit manuscript locally

@@
 
+___sentence about SAUP - how can we rephrase the "two global-scale repositories" sentence to exclude it?
+"broad taxonomic coverage"?
+something about not focusing just on economically important species?
+(is that a valid assumption about SAUP)___
+
+___or: "Several global-scale repositories predict ... In this paper we focus on two datasets that cover a broad taxonomic range: ..."___
+
 <!--
@@
 -->
-Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each presents one understanding of the "truth" - one database relying primarily on expert opinion and the other on model predictions.
-Uncertainties inherent in the method and intent inevitably drive differences in the range predictions made by each dataset; significant differences could ___lead to, communicate, imply___ dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
-
+Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each presents one understanding of the truth - one database relying primarily on expert opinion and the other on modeling.
+Uncertainties inherent in the methods and intents inevitably drive differences in the range predictions made by each dataset.
+Significant differences could ___lead to, communicate, imply___ dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
 Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
@@
 
-It seems fair to assume that range predictions from each dataset would be more closely aligned for well-understood species.
-For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and suggest methods to improve confidence in species range predictions.
+We can reasonably assume that range predictions from these two datasets would be more tightly aligned for well-understood species.
+For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the predicted ranges align in distribution and area, and determine several issues that lead to poor alignment between the two datasets.
+___
 
@@
 
-___we are not recommending one data set as better than the other!___
+___While we cannot recommend one data set as more accurate than the other, we do suggest methods to improve confidence in species range predictions and conservation outcomes based on these two datasets.___
 

### 2016-08-15 ms_2_methods.Rmd
pushing edits so I can edit manuscript locally

@@
 The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of "probability of occurrence" for each species [@kaschner_mapping_2006; @ready_predicting_2010].
-Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
+Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., previously, the Ocean Health Index defined presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
+In the 2016 OHI assessment, the presence threshold was eliminated, to better approximate the "extent of occurrence" generally indicated by IUCN maps, since the two datasets are used together.
+[@REF???]
 
@@
 To examine the overall taxonomic distribution across the spatial datasets (Fig.
-1A), we grouped species by taxonomic class and data source (IUCN, AquaMaps, or both), and examined the proportion of each class represented in each data source category.
+1A), we grouped species by taxonomic class and data source, and determined the proportion of each class represented in each dataset.
+We defined "paired map" species as those with spatial data in both datasets.
 
-To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; we determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell.
-For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.
+To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; we determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell  ___(Fig.
+SXXX)___.
+For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps ___(Fig.
+SXXX)___.
 We represented relative distribution of species count for each dataset by plotting average species count against latitude and longitude (Fig.
@@
 __Comparison of paired maps__:
-Using genus and species binomials as a matching key, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.
-___To avoid double-counting, we removed subpopulations and species aliases.
+Using genus and species binomials as a matching key, we selected the "paired map" species - the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.
+_We used the taxize package [@REF] in R [@REF] to standardize species names and synonyms; for species with separate subpopulation maps in IUCN, we combined all subpopulations to create a single global population._  ___To avoid double-counting, we removed subpopulations and species aliases.
 Used the taxize package?
@@
 
-For each species with ranges described in both IUCN and AquaMaps, $A_{small}$ and $A_{large}$ indicate the smaller and larger range representation (regardless of which dataset).
+For each paired map species, $A_{small}$ and $A_{large}$ indicate the smaller and larger range representation (regardless of which dataset).
 $A_{small \cap large}$ represents the amount of overlapping area between the two datasets.
@@
 
-Dividing the plot into quadrants based on mean values along each axis, we roughly categorized four qualities of alignment, and for each taxonomic group, based on IUCN taxonomic groupings, determined the proportion of species that fell into each quality category (2B).
+Dividing the plot into quadrants based on mean values along each axis, we roughly categorized four qualities of alignment, and for each taxonomic group, based on IUCN taxonomic groupings, determined the proportion of paired map species that fell into each quality category (2B).
 
 Noting that a large proportion of corals were identified in the upper-left quadrant, we examined a handful of IUCN coral maps and found that each predicted coral presence in far deeper waters than would be expected based on the IUCN's habitat description.
-Using the IUCN's API [REF] to find life history information for all IUCN-mapped corals, we found that none of the paired-map corals is described as occurring deeper than 200 m, ___and most are found in far shallower waters.___  We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] and masked our IUCN coral rasters to exclude presence below 200 m.
-The resulting maps were again compared to the AquaMaps rasters; ___figure XXXX___ plots shifts in coral species alignment due to this depth clipping.
+Using the IUCN's API [REF] to find life history information for all IUCN-mapped corals, we found that none of the paired-map corals is described as occurring deeper than 200 m; 94% are noted as being limited to 50 m or shallower.
+We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] ___(Fig.
+SXXX)___ and masked our IUCN coral rasters to exclude presence below 200 m.
+The resulting maps were again compared to the AquaMaps rasters; ___Fig.
+XXXX___ plots shifts in coral species alignment due to this depth clipping.
 ___does some of this belong in the results/discussion?___
+
+___To identify data-poor species, we used the robis???
+and rgbif???
+packages in R to identify known occurrences of each species; occurrences were averaged between the two bases as a proxy for data richness.___
 
@@
 Each species is assigned a risk score based on IUCN Red List extinction risk category, from Least Concern (0.0) to Extinct (1.0) (Data Deficient and unassessed species are excluded).
-Species range is determined from its IUCN range map, where available, or its AquaMaps range map ("presence" defined as a 40% or greater probability of occurrence) if an IUCN map is unavailable.
+Species range is determined from its IUCN range map, where available, or its AquaMaps range map if an IUCN map is unavailable.
 For each assessed OHI region, an area-weighted mean extinction risk across all species is calculated, and rescaled to calculate a status score from 0 (mean extinction risk = 0.75) to 100 (mean extinction risk = 0) [@halpern_index_2012; @halpern_patterns_2015].
+
+_To highlight the differences in the datasets, we recalculated the OHI SPP status values using only one complete dataset at a time.
+For the AquaMaps dataset, we used a presence threshold of 0% to align with the OHI 2016 assessment._  ___Fig.
+3 compares these single-dataset scores to the 2016 OHI SPP status values.___
 

### 2016-08-15 ms_5_figures.Rmd
pushing edits so I can edit manuscript locally

--- 2016-08-07 drafts/initial_version/ms_figures.Rmd
@@
 
-![](figures/fig1abc_am_iucn_maps.png)
+![](figures/fig1a_bar_by_taxa_all.png)
+![](figures/fig1b_all_aquamaps_map.png)
+![](figures/fig1c_all_iucn_map.png)
+
+-----
 
@@
 
-![](figures/fig2ab_quadmap_sppbars.png)
+![](figures/fig2a_scatter_quadmap.png)
+![](figures/fig2b_barchart_spp_gp_quads.png)
+
+-----
 
@@
 
-<a id = 'fig4'></a>
+<a id = 'fig3'></a>
 
-![Absolute change in SPP score](figures/fig3_boxplot_abs_diff.png)
+![paired map comparison](figures/fig3a_violinplot_paired_maps.png)
+
+why is the global one not showing up?:
+![global full species list comparison](figures/fig3a_violinplot_ohi_global.png)
+
+-----
 
@@
 
+### New figures - corals
+
+![coral area change due to depth clipping](figures/fig5a_coral_area_plot.png)
+
+![coral quadmap change due to depth clipping](figures/fig5b_scatter_quadmap_corals.png)
+
 -----

### 2016-08-15 ms_1_intro.Rmd
pushing to merge with remote edits

--- 2016-08-10 ms_1_intro.Rmd
@@
 Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each presents one understanding of the "truth" - one database relying primarily on expert opinion and the other on model predictions.
-Uncertainties inherent in the method and intent inevitably drive differences in the range predictions made by each dataset; significant differences could ___lead to, communicate, imply___ dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
+Uncertainties inherent in the method and intent inevitably drive differences in the range predictions made by each dataset; significant differences could indicate dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
 
+Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans [@rondinini_tradeoffs_2006; @jetz_ecological_2008].
 Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
-Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans [@rondinini_tradeoffs_2006; @jetz_ecological_2008].
 
@@
 
-It seems fair to assume that range predictions from each dataset would be more closely aligned for well-understood species.
+___It seems fair to assume that/we can reasonably assume that___ range predictions from each dataset would be more closely aligned for well-understood species.
 For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and suggest methods to improve confidence in species range predictions.
@@
 
-___we are not recommending one data set as better than the other!___
+While we cannot recommmend one data set as more accurate than the other, we suggest methods to improve confidence in species range predictions to better inform conservation management and policy decisions.
 

### 2016-08-15 ms_2_methods.Rmd
pushing to merge with remote edits

--- 2016-08-10 ms_2_methods.Rmd
@@
 In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of "probability of occurrence" for each species [@kaschner_mapping_2006; @ready_predicting_2010].
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of suitable habitat.
+The model masks habitat to certain FAO Major Fishing Areas as a rough cut of geographic range, resulting in a "probability of occurrence" for each species.
+Of the resulting maps, 1296 (5.7%) have been further refined under expert review [@kaschner_mapping_2006; @ready_predicting_2010].
 Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
+
+___short paragraph about other datasets we didn't include, i.e. SAUP?___
 

### 2016-08-15 case_study_ohi/ohi_paired_maps.Rmd

@@
 
-``` {r create_box_plots}
+``` {r create_violin_plots}
 dir_fig <- '~/github/IUCN-AquaMaps/figures'
@@
 print(bplot_abs)
-ggsave(file.path(dir_fig, 'fig3a_boxplot_paired_maps.png'),
+ggsave(file.path(dir_fig, 'fig3a_violinplot_paired_maps.png'),
        height = 4.5, width = 8.7, units = 'cm', dpi = 600)
-ggsave(file.path(dir_fig, 'fig3a_boxplot_paired_maps.tif'), device = 'tiff',
+ggsave(file.path(dir_fig, 'fig3a_violinplot_paired_maps.tif'), device = 'tiff',
        height = 4.5, width = 8.7, units = 'cm', dpi = 600)

### 2016-08-15 case_study_ohi/ohi_single_datasets.Rmd

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

source('~/github/IUCN-AquaMaps/data_explore/data_explore_fxns.R')

```

```{r get_spp_cells, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

### get rgn_cells - local copy
ohi_rgn_cells <- read_csv(file.path(dir_git, 'case_study_ohi/rgn_cells.csv'), 
                          col_types = 'icdd_d')

### get spp_info - git-annex original
ohi_spp_info <- read_csv(file.path(dir_anx, 'v2016/int/spp_all_cleaned.csv')) %>%
  filter(!is.na(cat_score))  %>% ### filter out unassessed and DD species
  filter(!is.na(spatial_source)) ### filter out unmapped species

### Load species cells; filter to species IDs in assessed species list and cell IDs in OHI regions
am_spp_cells <- read_csv(file.path(dir_anx, 'v2016/int/am_cells_spp_prob0.csv')) %>%
  filter(am_sid %in% ohi_spp_info$am_sid) %>%
  filter(loiczid %in% ohi_rgn_cells$loiczid)

iucn_spp_cells <- read_csv(file.path(dir_anx, 'v2016/int/iucn_cells_spp.csv'),
                           col_types = 'cddddc') %>%
  filter(presence != 5) %>%
  filter(iucn_sid %in% ohi_spp_info$iucn_sid) %>%
  filter(loiczid %in% ohi_rgn_cells$loiczid)

### NOTE: by using the 2016 cell files we're also getting the BOTW species...
### either mention in methods, or filter out!!!

```

``` {r calc_ohi_iucn_spp}

### bind IUCN loiczid dataframe to species risk category dataframe
iucn_spp_cells <- iucn_spp_cells %>%
  left_join(ohi_spp_info %>% select(iucn_sid, cat_score),
            by = 'iucn_sid')

### group by loiczid, then summarize to find means (and counts)
iucn_cell_sums <- iucn_spp_cells %>%
  filter(!is.na(cat_score)) %>%
  group_by(loiczid) %>%
  summarize(mean_cat = mean(cat_score),
            n_spp    = n())

write_csv(iucn_cell_sums, file.path(dir_git, 'case_study_ohi/iucn_spp_cell_sums.csv'))

### bind cells to regions dataframe and summarize
iucn_rgn_sums <- iucn_cell_sums %>%
  inner_join(ohi_rgn_cells,
             by = 'loiczid') %>%
  # mutate(rgn_area = cell_area * proportionArea,
  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
  mutate(cell_area_weight = cell_area * n_spp * proportionArea,
         area_weighted_mean_cat = mean_cat * cell_area_weight) %>%
  arrange(loiczid)

iucn_rgn_sums <- iucn_rgn_sums %>%
  group_by(rgn_id) %>%
  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))

iucn_rgn_sums <- iucn_rgn_sums %>%
  mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)

write_csv(iucn_rgn_sums, file.path(dir_git, 'case_study_ohi/iucn_spp_region_sums.csv'))

```

``` {r calc_ohi_am_spp}

### bind IUCN loiczid dataframe to species risk category dataframe
am_spp_cells <- am_spp_cells %>%
  left_join(ohi_spp_info %>% select(am_sid, cat_score),
            by = 'am_sid')

### group by loiczid, then summarize to find means (and counts)
am_cell_sums <- am_spp_cells %>%
  filter(!is.na(cat_score)) %>%
  group_by(loiczid) %>%
  summarize(mean_cat = mean(cat_score),
            n_spp    = n())

write_csv(am_cell_sums, file.path(dir_git, 'case_study_ohi/am_spp_cell_sums.csv'))

### bind cells to regions dataframe and summarize
am_rgn_sums <- am_cell_sums %>%
  inner_join(ohi_rgn_cells,
             by = 'loiczid') %>%
  # mutate(rgn_area = cell_area * proportionArea,
  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
  mutate(cell_area_weight = cell_area * n_spp * proportionArea,
         area_weighted_mean_cat = mean_cat   * cell_area_weight) %>%
  arrange(loiczid)

am_rgn_sums <- am_rgn_sums %>%
  group_by(rgn_id) %>%
  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))

am_rgn_sums <- am_rgn_sums %>%
  mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)

write_csv(am_rgn_sums, file.path(dir_git, 'case_study_ohi/am_spp_region_sums.csv'))

```

``` {r calc_ohi_both_spp}

### DO THIS TO SANITY CHECK/SIMPLIFY IN SAME MANNER AS OTHERS???  REMOVE BOTW SPECIES???

### bind IUCN loiczid dataframe to species risk category dataframe
### Already done in above code chunk!
# am_spp_cells <- am_spp_cells %>%
#   left_join(ohi_spp_info %>% select(am_sid, cat_score = category_score),
#             by = 'am_sid')

ohi_am_maps <- ohi_spp_info %>%
  filter(str_detect(spatial_source, 'am'))

### group by loiczid, then summarize to find means (and counts)
am_cell_sum <- am_spp_cells %>%
  filter(!is.na(cat_score)) %>%
  group_by(loiczid) %>%
  summarize(mean_cat = mean(cat_score),
            n_spp    = n())

write_csv(am_cell_sum, file.path(dir_git, 'case_study_ohi/am_spp_cell_sum.csv'))

### bind cells to regions dataframe and summarize
am_rgn_sums <- am_cell_sum %>%
  inner_join(ohi_rgn_cells,
             by = 'loiczid') %>%
  # mutate(rgn_area = cell_area * proportionArea,
  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
  mutate(cell_area_weight = cell_area * n_spp * proportionArea,
         area_weighted_mean_cat = mean_cat   * cell_area_weight) %>%
  arrange(loiczid)

am_rgn_sums <- am_rgn_sums %>%
  group_by(rgn_id) %>%
  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))

am_rgn_sums <- am_rgn_sums %>%
  mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)

write_csv(am_rgn_sums, file.path(dir_git, 'case_study_ohi/am_spp_region_sums.csv'))

```

``` {r create_violin_plots}
dir_fig <- '~/github/IUCN-AquaMaps/figures'

iucn_rgn_sums  <- read_csv(file.path(dir_git, 'case_study_ohi/iucn_spp_region_sums.csv'))
am_rgn_sums    <- read_csv(file.path(dir_git, 'case_study_ohi/am_spp_region_sums.csv'))
ohi_rgn_sums   <- read_csv(file.path('~/github/ohiprep/globalprep', 
                                     'spp_ico/v2016/output/spp_status_global.csv'))

status_df <- ohi_rgn_sums %>%
  select(rgn_id, ohi_st = score) %>%
  mutate(ohi_st = 100 * ohi_st) %>%
  full_join(am_rgn_sums %>% select(rgn_id, am_st = status), by = 'rgn_id') %>%
  full_join(iucn_rgn_sums %>% select(rgn_id, iucn_st = status), by = 'rgn_id') %>%
  mutate(diff_am   = am_st - ohi_st ,
         diff_iucn = iucn_st - ohi_st) %>%
  gather(scenario, diff, diff_am:diff_iucn)

viol_plot <- ggplot(data = status_df, aes(x = scenario, y = diff)) +
  ggtheme_plot + 
  theme(panel.grid.major = element_line(colour = 'grey90', size = .25),
        panel.grid.major.x = element_blank()) +

  geom_hline(yintercept = 0, color = 'gray50', size = .5) +
  geom_violin(fill = 'grey80', alpha = .5,
              scale = 'width',
              draw_quantiles = c(.25, .5, .75)) +
  geom_point(position = position_jitter(w = .3), 
             color = 'black', alpha = 1, size = 0.25) +
  geom_violin(fill = NA, 
              draw_quantiles = c(.25, .5, .75),
              scale = 'width') +
  labs(x = 'Scenario',
       y = 'Change in SPP status') +
  scale_x_discrete(labels = c('1. AquaMaps only', 
                              '2. IUCN only'))

print(viol_plot)

ggsave(file.path(dir_fig, 'fig3a_violinplot_ohi_maps.png'),
       height = 4.5, width = 8.7, units = 'cm', dpi = 600)
# ggsave(file.path(dir_fig, 'fig3a_violinplot_ohi_maps.png'), device = 'tiff',
#        height = 4.5, width = 8.7, units = 'cm', dpi = 600)

```

### 2016-08-15 ms_0_abst_sig.Rmd

--- 2016-03-27 ms_1_abst_sig.Rmd
@@
 Here we examine differences in predicted species ranges between the two datasets and find that these misalignments mainly result from divergent methodologies that introduce differing frequencies of commission and omission errors.
-We illustrate the scientific and management implications of these differences by repeating two recent applications - an assessment of global biodiversity within the Ocean Health Index and a global analysis of gaps in coverage of marine protected areas - and find significantly different results depending on how the two datasets were used.
+We illustrate the scientific and management implications of these differences by repeating two recent applications - an assessment of global biodiversity within the Ocean Health Index and a global analysis of gaps in coverage of marine protected areas - and find significantly different results depending on how the two datasets are used.
 Until a single, highly accurate dataset of global marine species ranges becomes available, understanding the implications of dataset differences for conservation planning and decision-making remains essential.

### 2016-08-15 ms_1_intro.Rmd

@@
 -->
+
 Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each presents one understanding of the "truth" - one database relying primarily on expert opinion and the other on model predictions.
 Uncertainties inherent in the method and intent inevitably drive differences in the range predictions made by each dataset; significant differences could indicate dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
-
 Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans [@rondinini_tradeoffs_2006; @jetz_ecological_2008].

### 2016-08-15 ms_2_methods.Rmd

--- 2016-08-15 ms_2_methods.Rmd
@@
 In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of "probability of occurrence" for each species [@kaschner_mapping_2006; @ready_predicting_2010].
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of suitable habitat.
+The model masks habitat to certain FAO Major Fishing Areas as a rough cut of geographic range, resulting in a "probability of occurrence" for each species.
+Of the resulting maps, 1296 (5.7%) have been further refined under expert review [@kaschner_mapping_2006; @ready_predicting_2010].
 Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., previously, the Ocean Health Index defined presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
 In the 2016 OHI assessment, the presence threshold was eliminated, to better approximate the "extent of occurrence" generally indicated by IUCN maps, since the two datasets are used together.
-[@REF???]
+
+___short paragraph about other datasets we didn't include, i.e. SAUP?___
 

### 2016-08-15 ms_6_supplemental.Rmd

@@
-# Supplemental Information
+---
+title: "Supplemental Information: Aligning marine species range data to better serve science and conservation"
+author: "Casey O'Hara, Jamie C. Afflerbach, Courtney Scarborough, Benjamin S. Halpern"
+output:
+  word_document: default
+  pdf_document:
+    number_sections: no
+    toc: no
+  html_document:
+    highlight: haddock
+    number_sections: no
+    theme: spacelab
+    toc: no
+---
 
@@
 This file type will be published in raw format and will not be edited or composed.
-
 -->
-
-<!--## Data processing - should this go in methods???
+<!--
+# Data sources and methods
 
@@
 
-### AquaMaps
+## AquaMaps
 
 AquaMaps data was provided by AquaMaps directly via ftp.
-Three SQL files were sent that contained (1) list of all `r nrow(spp_am)` species and their taxonomic information, (2) information for all (259,200) half degree cells used for range mapping and (3) the probability of occurrence per cell for all `r nrow(spp_am)` species.
+Three SQL files were sent that contained (1) list of all 22,889 modeled species and their taxonomic information, (2) information for all (259,200) half degree cells used for range mapping and (3) the probability of occurrence per cell for all 22,889 species.
 These were extracted into .csv form using R Statistical Software.
 
-### IUCN
+## IUCN
 
-As of December 2015, IUCN had published species distribution maps for `r nrow(spp_iucn_maps)` marine species across 24 taxonomic groups.
+As of December 2015, IUCN had published species distribution maps for XXX marine species across 24 taxonomic groups.
 For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.
 
-Each shapefile contains polygon limits of distribution indicating regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.-->
+Each shapefile contains spatial polygons outlining limits of distribution for each species, indicating regions of presence/absence; additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
+-->
 
-## S1: sample maps from each quadrant
+## Representative maps from each quadrant
 
@@
 >  __Fig.
-S2.__ Breaking down the quadrants by IUCN extinction risk categories (Fig.
-S2), we found little support for our hypothesis that maps for species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.
-Does higher perceived risk lead to increased attention, and thus better understanding of species distribution?
-Or conversely, does increased attention to species distribution reveal more species at risk?
-Likely both mechanisms are at play on a case-by-case basis, depending on the species' taxon and region.
-_does this argument bear up to closer scrutiny?
-CR isn't dominated by Q1 any more_
+S2.__ Breakdown of quadrants (Fig.
+S1A) by IUCN extinction risk categories.
+IUCN and Aquamaps ranges were in better agreement (i.e., had a higher probability of being in the upper-right, "well-aligned" quadrant, Figure 2A/Figure S1A) for species at greater risk according to IUCN risk categories (p < 0.001, generalized linear model, family = binomial, Table S1).
+This suggests at-risk species may receive more attention in terms of data collection or model refinement.
 
-### S3: Coral depth map
+> __Table S1.__  Generalized linear model (family = binomial) describing how the agreement between IUCN and Aquamap range maps (i.e., probability of occurring in quadrant 1, Figure 2A/S1A) increases for a species as its IUCN risk status increases.
+For this analysis, we excluded all species that had "data deficient" IUCN status (n = 255).
+We analyzed the IUCN risk categories as a continuous variable (LC = 0.0, NT = 0.2, VU = 0.4, EN = 0.6, CR = 0.8, EX = 1.0).
+
+_Model: glm(quadrant 1 membership ~ IUCN category, family = binomial) n = 2,059_
+
+| Coefficients:  | Estimate | Std. Error | z value | p value      |
+| :-----         |   -----: |     -----: |  -----: |       -----: |
+| Intercept      | -1.32933 |   0.05804  | -22.902 | < 2e-16 ***  |
+| IUCN category  |  1.18672 |   0.33469  |   3.546 | 0.000391 *** |
+
+### Coral range vs. depth
 
@@
 
-<!--
-Oculina varicosa - Ivory tree coral ![oculina_pic.jpg](figures/oculina/oculina_pic.jpg)
--->
-
 >  __Fig.
@@
 varicosa_ can be found to depths of 152 m.
-In waters shallower than 30 m, _O.
-varicosa_ is zooxanthellate, while an azooxanthellate form can be found in deeper waters.
+Maps of many other coral species and other reef-associated organisms followed similar patterns.
 
-### S4: AquaMaps threshold examination
+### AquaMaps threshold examination
 

### 2016-08-16 case_study_ohi/ohi_single_datasets.Rmd
cleaning out old plots; updating figures; editing manuscript

@@
 
-### get spp_info - git-annex original
+### get spp_info - git-annex original; modify spp info to exclude BOTW
 ohi_spp_info <- read_csv(file.path(dir_anx, 'v2016/int/spp_all_cleaned.csv')) %>%
   filter(!is.na(cat_score))      %>% ### filter out unassessed and DD species
-  filter(!is.na(spatial_source)) ### filter out unmapped species
+  filter(!is.na(spatial_source)) %>% ### filter out unmapped species
+  mutate(spatial_source = ifelse(!is.na(am_sid) & spatial_source == 'iucn-bli', 'am', spatial_source)) %>%
+  filter(spatial_source != 'iucn-bli')
 
@@
 
+### bind IUCN loiczid dataframe to species risk category dataframe
+am_spp_cells <- am_spp_cells %>%
+  left_join(ohi_spp_info %>% select(am_sid, cat_score),
+            by = 'am_sid')
+
 iucn_spp_cells <- read_csv(file.path(dir_anx, 'v2016/int/iucn_cells_spp.csv'),
@@
   filter(loiczid %in% ohi_rgn_cells$loiczid)
-
-### NOTE: by using the 2016 cell files we're also getting the BOTW species...
-### either mention in methods, or filter out!!!
+### bind IUCN loiczid dataframe to species risk category dataframe
+iucn_spp_cells <- iucn_spp_cells %>%
+  left_join(ohi_spp_info %>% select(iucn_sid, cat_score),
+            by = 'iucn_sid')
 
@@
 
-### bind IUCN loiczid dataframe to species risk category dataframe
-iucn_spp_cells <- iucn_spp_cells %>%
-  left_join(ohi_spp_info %>% select(iucn_sid, cat_score),
-            by = 'iucn_sid')
-
 ### group by loiczid, then summarize to find means (and counts)
-iucn_cell_sums <- iucn_spp_cells %>%
+iucn_cell_sum <- iucn_spp_cells %>%
   filter(!is.na(cat_score)) %>%
@@
 
-write_csv(iucn_cell_sums, file.path(dir_git, 'case_study_ohi/iucn_spp_cell_sums.csv'))
+write_csv(iucn_cell_sum, file.path(dir_git, 'case_study_ohi/iucn_spp_cell_sum.csv'))
 
 ### bind cells to regions dataframe and summarize
-iucn_rgn_sums <- iucn_cell_sums %>%
+iucn_rgn_sum <- iucn_cell_sum %>%
   inner_join(ohi_rgn_cells,
@@
 
-iucn_rgn_sums <- iucn_rgn_sums %>%
+iucn_rgn_sum <- iucn_rgn_sum %>%
   group_by(rgn_id) %>%
-  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))
+  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight),
+            mean_n_spp = mean(n_spp))
 
-iucn_rgn_sums <- iucn_rgn_sums %>%
+iucn_rgn_sum <- iucn_rgn_sum %>%
   mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)
 
-write_csv(iucn_rgn_sums, file.path(dir_git, 'case_study_ohi/iucn_spp_region_sums.csv'))
+write_csv(iucn_rgn_sum, file.path(dir_git, 'case_study_ohi/iucn_spp_region_sum.csv'))
 
@@
 ``` {r calc_ohi_am_spp}
-
-### bind IUCN loiczid dataframe to species risk category dataframe
-am_spp_cells <- am_spp_cells %>%
-  left_join(ohi_spp_info %>% select(am_sid, cat_score),
-            by = 'am_sid')
-
-### group by loiczid, then summarize to find means (and counts)
-am_cell_sums <- am_spp_cells %>%
-  filter(!is.na(cat_score)) %>%
-  group_by(loiczid) %>%
-  summarize(mean_cat = mean(cat_score),
-            n_spp    = n())
-
-write_csv(am_cell_sums, file.path(dir_git, 'case_study_ohi/am_spp_cell_sums.csv'))
-
-### bind cells to regions dataframe and summarize
-am_rgn_sums <- am_cell_sums %>%
-  inner_join(ohi_rgn_cells,
-             by = 'loiczid') %>%
-  # mutate(rgn_area = cell_area * proportionArea,
-  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
-  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
-  mutate(cell_area_weight = cell_area * n_spp * proportionArea,
-         area_weighted_mean_cat = mean_cat   * cell_area_weight) %>%
-  arrange(loiczid)
-
-am_rgn_sums <- am_rgn_sums %>%
-  group_by(rgn_id) %>%
-  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))
-
-am_rgn_sums <- am_rgn_sums %>%
-  mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)
-
-write_csv(am_rgn_sums, file.path(dir_git, 'case_study_ohi/am_spp_region_sums.csv'))
-
-```
-
-``` {r calc_ohi_both_spp}
-
-### DO THIS TO SANITY CHECK/SIMPLIFY IN SAME MANNER AS OTHERS???  REMOVE BOTW SPECIES???
-
-### bind IUCN loiczid dataframe to species risk category dataframe
-### Already done in above code chunk!
-# am_spp_cells <- am_spp_cells %>%
-#   left_join(ohi_spp_info %>% select(am_sid, cat_score = category_score),
-#             by = 'am_sid')
-
-ohi_am_maps <- ohi_spp_info %>%
-  filter(str_detect(spatial_source, 'am'))
 
@@
 ### bind cells to regions dataframe and summarize
-am_rgn_sums <- am_cell_sum %>%
+am_rgn_sum <- am_cell_sum %>%
   inner_join(ohi_rgn_cells,
@@
 
-am_rgn_sums <- am_rgn_sums %>%
+am_rgn_sum <- am_rgn_sum %>%
   group_by(rgn_id) %>%
-  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight))
+  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight),
+            mean_n_spp = mean(n_spp))
 
-am_rgn_sums <- am_rgn_sums %>%
+am_rgn_sum <- am_rgn_sum %>%
   mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)
 
-write_csv(am_rgn_sums, file.path(dir_git, 'case_study_ohi/am_spp_region_sums.csv'))
+write_csv(am_rgn_sum, file.path(dir_git, 'case_study_ohi/am_spp_region_sum.csv'))
+
+```
+
+``` {r calc_ohi_both_spp}
+
+### get AM spp cells for combined scores
+spp_ohi_am <- ohi_spp_info %>% 
+  filter(str_detect(spatial_source, 'am'))
+
+am_cell_sum_both <- am_spp_cells %>%
+  filter(am_sid %in% spp_ohi_am$am_sid) %>%
+  filter(!is.na(cat_score)) %>%
+  group_by(loiczid) %>%
+  summarize(mean_cat = mean(cat_score),
+            n_spp    = n(),
+            data     = 'am')
+
+### get IUCN spp cells for combined scores
+spp_ohi_iucn <- ohi_spp_info %>% 
+  filter(str_detect(spatial_source, 'iucn'))
+
+iucn_cell_sum_both <- iucn_spp_cells %>%
+  filter(iucn_sid %in% spp_ohi_iucn$iucn_sid) %>%
+  filter(!is.na(cat_score)) %>%
+  group_by(loiczid) %>%
+  summarize(mean_cat = mean(cat_score),
+            n_spp    = n(),
+            data     = 'iucn')
+
+### Stick IUCN and AM together and find spp-weighted mean category
+ohi_cell_sum <- iucn_cell_sum_both %>%
+  bind_rows(am_cell_sum_both) %>%
+  mutate(mean_weight_cat = mean_cat * n_spp) %>%
+  group_by(loiczid) %>%
+  summarize(mean_cat = sum(mean_weight_cat, na.rm = TRUE) / sum(n_spp, na.rm = TRUE),
+            n_spp = sum(n_spp, na.rm = TRUE))
+
+write_csv(ohi_cell_sum, file.path(dir_git, 'case_study_ohi/ohi_cell_sum.csv'))
+
+### bind cells to regions dataframe and summarize by region
+ohi_rgn_sum <- ohi_cell_sum %>%
+  inner_join(ohi_rgn_cells,
+             by = 'loiczid') %>%
+  # mutate(rgn_area = cell_area * proportionArea,
+  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
+  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
+  mutate(cell_area_weight = cell_area * n_spp * proportionArea,
+         area_weighted_mean_cat = mean_cat   * cell_area_weight) %>%
+  arrange(loiczid)
+
+ohi_rgn_sum <- ohi_rgn_sum %>%
+  group_by(rgn_id) %>%
+  summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight),
+            mean_n_spp = mean(n_spp))
+
+ohi_rgn_sum <- ohi_rgn_sum %>%
+  mutate(status = 100 * ((1 - rgn_mean_cat) - 0.25) / 0.75)
+
+write_csv(ohi_rgn_sum, file.path(dir_git, 'case_study_ohi/ohi_spp_region_sum.csv'))
 
@@
 
-iucn_rgn_sums  <- read_csv(file.path(dir_git, 'case_study_ohi/iucn_spp_region_sums.csv'))
-am_rgn_sums    <- read_csv(file.path(dir_git, 'case_study_ohi/am_spp_region_sums.csv'))
-ohi_rgn_sums   <- read_csv(file.path('~/github/ohiprep/globalprep', 
-                                     'spp_ico/v2016/output/spp_status_global.csv'))
+iucn_rgn_sum  <- read_csv(file.path(dir_git, 'case_study_ohi/iucn_spp_region_sum.csv'))
+am_rgn_sum    <- read_csv(file.path(dir_git, 'case_study_ohi/am_spp_region_sum.csv'))
+ohi_rgn_sum   <- read_csv(file.path(dir_git, 'case_study_ohi/ohi_spp_region_sum.csv'))
 
-status_df <- ohi_rgn_sums %>%
-  select(rgn_id, ohi_st = score) %>%
-  mutate(ohi_st = 100 * ohi_st) %>%
-  full_join(am_rgn_sums %>% select(rgn_id, am_st = status), by = 'rgn_id') %>%
-  full_join(iucn_rgn_sums %>% select(rgn_id, iucn_st = status), by = 'rgn_id') %>%
+status_df <- ohi_rgn_sum %>%
+  select(rgn_id, ohi_st = status, mean_n_spp) %>%
+  mutate(ohi_st = ohi_st) %>%
+  full_join(am_rgn_sum   %>% select(rgn_id,    am_st = status), by = 'rgn_id') %>%
+  full_join(iucn_rgn_sum %>% select(rgn_id, iucn_st = status), by = 'rgn_id') %>%
   mutate(diff_am   = am_st - ohi_st ,
@@
   geom_point(position = position_jitter(w = .3), 
-             color = 'black', alpha = 1, size = 0.25) +
+             aes(color = mean_n_spp), alpha = 1, size = 0.25) +
   geom_violin(fill = NA, 
@@
               scale = 'width') +
-  labs(x = 'Scenario',
+  labs(x = 'Scenario, compared to OHI 2016',
        y = 'Change in SPP status') +

### 2016-08-16 coralmaps/coral_maps.Rmd
cleaning out old plots; updating figures; editing manuscript

@@
 ### 18 species (of 562) have no depth information; one is listed as being dredged from 13 fathoms (~ 25 m)
+quantile(spp_depth$max_depth, seq(.9, 1, .01), na.rm = TRUE)
+ #   90%    91%    92%    93%    94%    95%    96%    97%    98%    99%   100% 
+ # 40.00  44.65  50.00  50.00  50.00  61.25  70.80  78.20  84.10 100.00 200.00
+
+spp_d1 <- spp_depth %>% filter(iucn_sid %in% coral_cells_iucn$iucn_sid) ### mapped spp only
+quantile(spp_d1$max_depth, seq(.9, 1, .01), na.rm = TRUE)
+ #   90%    91%    92%    93%    94%    95%    96%    97%    98%    99%   100% 
+ # 40.00  45.00  50.00  50.00  50.00  54.60  70.00  75.56  82.12 100.00 152.00
+
 ```
@@
 ggsave(plot = coral_area_plot,
-       filename = file.path(dir_fig, 'coral_area_plot.png'), device = 'png',
+       filename = file.path(dir_fig, 'fig3a_coral_area_plot.png'), device = 'png',
        width = 9, height = 5, units = 'cm', dpi = 600)
@@
 iucn_coral_fixed <- spp_coralmaps %>%
-  gather(dist, perc, c(sm_perc_raw, sm_perc_shallow)) %>%
-  gather(area, ratio, c(area_ratio_raw, area_ratio_shallow)) %>%
+  gather(dist, d_perc, c(sm_perc_raw, sm_perc_shallow)) %>%
+  gather(area, a_ratio, c(area_ratio_raw, area_ratio_shallow)) %>%
   filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
@@
 
+write_csv(iucn_coral_fixed, file.path(dir_git, 'coralmaps', 'iucn_coral_fixed.csv'))
+
 break_nums <- seq(0, 100, 20)
 
+### Plot the new quad graph
 clipped_quads <- ggplot(iucn_coral_fixed %>%
@@
                                                .$iucn_sid)),
-                    aes(x = ratio, y = perc, group = iucn_sid)) +
+                    aes(x = a_ratio, y = d_perc, group = iucn_sid)) +
   ggtheme_plot + 
@@
 ggsave(plot = clipped_quads,
-       filename = file.path(dir_fig, 'scatter_quadmap_corals.png'), device = 'png',
+       filename = file.path(dir_fig, 'fig3b_scatter_quadmap_corals.png'), device = 'png',
        width = 17.8, height = 9, units = 'cm', dpi = 600)
@@
 ```
+
+``` {r examine_coral_quartiles, echo = FALSE, eval = TRUE}
+
+iucn_coral_fixed <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_fixed.csv'))
+
+### identify quads based on original breaks from data_explore.Rmd
+
+area_align_mean <- 54.53334
+dist_align_mean <- 62.98212
+
+# iucn_sid  am_sid  sciname  area_am  area_iucn_shallow  area_overlap_shallow  sm_range_shallow  area_iucn_raw
+# area_overlap_raw  sm_range_raw  dist  perc  area  ratio  method
+
+iucn_coral_quads <- iucn_coral_fixed %>%
+  select(iucn_sid, sciname, d_perc, a_ratio, method) %>% ### ditch am_sid, b/c duplicates
+  distinct() %>%
+  mutate(quad = 'q1',
+         quad = ifelse(a_ratio < area_align_mean  & d_perc >= dist_align_mean, 'q2', quad),
+         quad = ifelse(a_ratio >= area_align_mean & d_perc < dist_align_mean,  'q3', quad),
+         quad = ifelse(a_ratio < area_align_mean  & d_perc < dist_align_mean,  'q4', quad)) %>%
+  group_by(method) %>%
+  mutate(n_total = n()) %>%
+  group_by(method, quad) %>%
+  mutate(n_quad = n(),
+         pct_quad = n_quad/n_total) %>%
+  ungroup()
+
+quad_names <- data.frame('quad' = c('q1', 'q2', 'q3', 'q4'),
+                         'quad_name' = factor(c('well-matched', 'dist-matched', 
+                                                'area-matched', 'poorly matched'),
+                                             ordered = TRUE))
+
+iucn_coral_quads <- iucn_coral_quads %>%
+  left_join(quad_names, by = 'quad')
+
+write_csv(iucn_coral_quads, file.path(dir_git, 'coralmaps', 'iucn_coral_quads.csv'))
+
+break_nums <- seq(0, 100, 20)
+
+iucn_coral_quad_sum <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_quads.csv')) %>%
+  select(method, quad, quad_name, pct_quad) %>%
+  distinct()
+### Plot the bar chart
+barchart_coral_quads <- ggplot(iucn_coral_quad_sum, 
+                                aes(x = method, fill = quad, weight = pct_quad)) +
+  ggtheme_plot + 
+  theme(panel.grid.major.y = element_blank()) +
+  geom_bar(stat = 'count', alpha = 1) +
+  scale_fill_manual(values = c('q4' = '#d01c8b', 
+                               'q3' = '#f1b6da', 
+                               'q2' = '#b8e186',
+                               'q1' = '#4dac26'),
+                    labels = quad_names$quad_name) +
+  scale_y_continuous(expand = c(0, 0), 
+                     limits = c(0, 1.1),
+                     breaks = break_nums/100,
+                     labels = sprintf('%s%%', break_nums)) + 
+  ### add grid lines; horizontal but then get flipped
+  geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
+  coord_flip() +
+  labs(x = 'Taxonomic Group', 
+       y = 'Percent of corals by quadrant', 
+       fill = 'Alignment')
+
+print(barchart_coral_quads)
+
+ggsave(plot = barchart_coral_quads,
+       filename = file.path(dir_fig, 'fig3c_barchart_coral_quads.png'),
+       width = 17.8, height = 4, units = 'cm', dpi = 600)
+
+write_csv(iucn_coral_quad_sum, file.path(dir_git, 'coralmaps', 'iucn_coral_quad_sum.csv'))
+
+```

### 2016-08-16 data_explore/data_explore_coralmaps.Rmd
cleaning out old plots; updating figures; editing manuscript

@@
 ### LOICZID cell ID numbers - used in plotting rasters
-loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
+loiczid_raster_file  <- file.path(dir_git, 'shiny_am_iucn/data/loiczid_raster.tif')
 loiczid_raster       <- raster(loiczid_raster_file)

### 2016-08-16 ms_2_methods.Rmd
cleaning out old plots; updating figures; editing manuscript

@@
 
-___short paragraph about other datasets we didn't include, i.e. SAUP?___
-
-___should this go in the discussion?___
-_The methodologies behind the creation of these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).
-Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.
-Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006].
-By comparing maps resulting from IUCN and AquaMaps methodologies, we can identify and possibly address mechanistic causes for each type of error._
+___short paragraph about other datasets we didn't include, i.e. SAUP? BOTW?___
 
@@
 
-Overlaying paired distribution maps for a given species, we defined and calculated _distribution alignment_ and _area ratio_ and plotted these in Fig. 2A:
+Overlaying paired distribution maps for each species, we defined and calculated _distribution alignment_ and _area ratio_ and plotted these in Fig. 2A:
 
@@
 
-_Distribution alignment - overlap as proxy for common understanding; use Bayesian argument to show that these areas are high-confidence areas; area ratio as proxy for errors of commission/omission.
-While we cannot know which dataset is closer to the "truth" we can make inferences based on expectations of EOO and AOO maps - is AquaMaps AOO?_
+_Distribution alignment uses overlapping predictions of presence as means of identifying areas of high confidence.  Area ratio provides a proxy for frequency of commission and/or omission errors._
 
@@
 
-Noting that a large proportion of corals were identified in the upper-left quadrant, we examined a handful of IUCN coral maps and found that each predicted coral presence in far deeper waters than would be expected based on the IUCN's habitat description.
-Using the IUCN's API [REF] to find life history information for all IUCN-mapped corals, we found that none of the paired-map corals is described as occurring deeper than 200 m; 94% are noted as being limited to 50 m or shallower.
+_Based on the results of the map alignment analysis, we chose to examine the effect of explicitly restricting IUCN ranges to depths based on species' life histories.
+We selected corals as a case study, due to their foundational role in biodiverse habitats, as well as their lack of mobility and general reliance on photosynthesis._  Using the IUCN's API [REF] to find life history information for all IUCN-mapped corals, we found that none of the paired-map corals is described as occurring deeper than 200 m; 94% are noted as being limited to 50 m or shallower.
 We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] ___(Fig.
@@
 XXXX___ plots shifts in coral species alignment due to this depth clipping.
-___does some of this belong in the results/discussion?___
 
@@
 _To highlight the differences in the datasets, we recalculated the OHI SPP status values using only one complete dataset at a time.
-For the AquaMaps dataset, we used a presence threshold of 0% to align with the OHI 2016 assessment._  ___Fig.
-3 compares these single-dataset scores to the 2016 OHI SPP status values.___
+For the AquaMaps dataset, we used a presence threshold of 0% to align with the OHI 2016 assessment.
+Fig.
+3 compares these single-dataset scores to the 2016 OHI SPP status values.
+While the full OHI 2016 assessment includes the Birds of the World dataset [@REF???] in addition to IUCN and AquaMaps, we omitted it from our calculations for simplicity._
 
-_To highlight the differences in the datasets, we selected the subset of species whose ranges are represented in both datasets, eliminating species found in only one or the other dataset.
+~To highlight the differences in the datasets, we selected the subset of species whose ranges are represented in both datasets, eliminating species found in only one or the other dataset.
 In the original OHI SPP analysis, IUCN data was used to determine ranges for these overlapping species; we calculated all regional SPP status scores for just these species as our baseline.
-We then recalculated the scores using AquaMaps ranges at a 40% presence threshold (as used in the original OHI analysis) and a 0% threshold (to better simulate the "limits of distribution" defined by IUCN maps) and compared them to our IUCN baseline._
+We then recalculated the scores using AquaMaps ranges at a 40% presence threshold (as used in the original OHI analysis) and a 0% threshold (to better simulate the "limits of distribution" defined by IUCN maps) and compared them to our IUCN baseline.~
 
-_We then modified the original OHI source code to recalculate the 2015 SPP status scores for all regions and all species to examine the effects of changing the AquaMaps presence threshold from 40% to 0% (all range included), and the effects of changing the data source for the overlapping species from IUCN to AquaMaps.
-We calculated the differences between these results and the published SPP status scores for 2015._
+~We then modified the original OHI source code to recalculate the 2015 SPP status scores for all regions and all species to examine the effects of changing the AquaMaps presence threshold from 40% to 0% (all range included), and the effects of changing the data source for the overlapping species from IUCN to AquaMaps.
+We calculated the differences between these results and the published SPP status scores for 2015.~
 

### 2016-08-16 ms_3_res_disc_concl.Rmd
cleaning out old plots; updating figures; editing manuscript

--- 2016-05-13 ms_2a_res_disc.Rmd
@@
 
-The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" [@iucn_iucn_2015] - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.
-In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of "probability of occurrence" for each species [@kaschner_mapping_2006; @ready_predicting_2010].
-Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
-
-The methodologies behind the creation of these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).
-Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.
-Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006].
-By comparing maps resulting from IUCN and AquaMaps methodologies, we can identify and possibly address mechanistic causes for each type of error.
-
-The two datasets have notably different taxonomic (Fig.
+(R) The two datasets have notably different taxonomic (Fig.
 1A) and regional (Figs.
@@
 The longitude frequency plots show a slight shift in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
-To achieve more comprehensive global coverage of species ranges these two datasets can be used together, but the underlying methodological differences complicate such direct comparisons.
+
+(D) _To achieve more comprehensive global coverage of species ranges these two datasets can be used together, understanding that the underlying differences complicate such direct comparisons.
+Major considerations include aligning the criteria used to determine "presence" (e.g.
+selecting an appropriate presence threshold for AquaMaps) and reducing mechanisms that introduce errors of commission and omission._
 
@@
 n_q2 <- nrow(spp_list_quads %>% filter(quad == 'q2'))
-# n_q2_am_sm  <- nrow(spp_list_quads %>% filter(sm_range == 'AM' & quad == 'q2'))
+n_q2_iucn_larger <- nrow(spp_list_quads %>% filter(sm_range == 'AM' & quad == 'q2'))
 n_q2_corals <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral') & quad == 'q2'))
@@
 
-To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area alignment (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig.
+___from Methods___
+(D) _The methodologies behind the creation of these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).
+Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.
+Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006].
+By comparing maps resulting from IUCN and AquaMaps methodologies, we can identify and possibly address mechanistic causes for each type of error._
+
+(D/M) ___is this sufficiently described in methods?
+or revisit here?___ ___Both datasets strive for accuracy in different ways; the truth likely lies somewhere in between.___ To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area alignment (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig.
 2A).
- This analysis revealed a general negative linear pattern, suggesting that increasing similarity in range area correlates with decreasing distribution alignment.
+Where two independent methods predict presence (or absence) of a species in the same location, we can assume a higher confidence in the accuracy of that prediction.
+Where one method predicts presence while the other predicts absence, we can investigate potential sources of commission and omission errors to refine our understanding of the species' range.
+Understanding these sources of error can provide guidance to improve predictions of species range even when only one source of spatial data is available.
+
+(R) This analysis revealed a general negative linear pattern, suggesting that increasing similarity in range area correlates with decreasing distribution alignment.
  AquaMaps tends to extrapolate species ranges into suitable areas beyond known occurrences, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
@@
 
-The mean distribution alignment for species included in both datasets was `r dist_align_mean`%; the mean area alignment was `r area_align_mean`%.
-By dividing the map-paired species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
-Representative maps from each category are provided in the supplemental materials (Fig.
+(R) The mean distribution alignment for species included in both datasets was `r dist_align_mean`%; the mean area alignment was `r area_align_mean`%.
+By dividing the paired map species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
+Representative maps from each category are provided in the supporting materials (Fig.
 S1).
 
-The upper right quadrant includes species (n = `r n_q1`) whose described ranges are above average in alignment of both spatial distribution and area.
+(R) The upper right quadrant includes the species (n = `r n_q1`) whose described ranges are above average in alignment of both spatial distribution and area.
 These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig.
@@
 
-The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where the IUCN range is notably larger, often extending into deeper water.
-For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their preferred depths, likely introducing errors of commission  (Fig.
-S3).
- Ocean depth preference is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps.
-Simply clipping IUCN range maps to known depth preferences would resolve many of these mismatches.
+(R) The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where one range is notably larger than the other (For `r round(n_q2_iucn_larger/n_q2 * 100, 1)`% of the species in this quadrant, the IUCN range is an average of 2.57 times larger than the AquaMaps range).
+This suggests a high rate of commission and/or omission errors by one or both datasets; further analysis is required to disentangle the source or type of error contributing to poor area alignment.
+The results of the coral analysis (described below) provide some insight.
 
-Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, introducing additional commission errors.
-Consequently, area ratios are close to 100%, though similar areas are unhelpful when the distributions are poorly aligned.
+(R) ___area ratio suggests similar commission/omission errors, but dist suggests otherwise___ Species found in the lower right quadrant (n = `r n_q3`) seem to represent cases of "two wrongs make a right."  For these species, IUCN and AquaMaps both predict ranges extending far beyond the overlapping region, but the methodological differences result in very different extrapolations.
+___OMIT:___ _For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, likely introducing additional commission errors._  Consequently, area ratios are close to 100%, though the poor distribution alignment indicates that one or both datasets are introducing errors.
+As above, further analysis is required to disentangle the causes of error.
 
-The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor area match well.
-Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
-When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species.
+(R) The lower left quadrant includes species (n = `r n_q4`) where alignment is poor in both dimensions.
+___present a solution here?___ Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
+It is not surprising that range maps based upon fewer observations bear greater uncertainty.
+___OMIT:___ _When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species._
+
+## New analysis: Coral maps (Distribution-aligned quads)
+
+(D/M) Noting from Fig.
+2B that corals dominate the upper-left "distribution-aligned" quadrant of Fig.
+2A (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), we clipped the IUCN coral range maps to a 200 m bathymetry raster to examine the impact on distribution and area alignment.
+Ocean depth preference is explicitly included in AquaMaps models.
+While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." [@REF http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources]), it is clear that this cannot be taken for granted.
+
+- Coral area plot
+
+(R) Fig.
+3A shows the 463 coral species mapped in the IUCN dataset, with their ranges broken into proportional area deeper and shallower than 200 m.
+None of these species is indicated to occur deeper than 200 m, and 94% are confined to waters shallower than 50 m; seven of the mapped species had no reported depth information.
+Clipping coral ranges to shallower than 200 m eliminated an average of 47.6% of the total predicted area while still allowing for a conservative estimate of suitable habitat.
+
+- Coral new quads
+
+(R) Masking coral ranges to shallow waters reduces a clear source of commission errors.
+In so doing, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps.
+Using the original quadrant definitions from Fig.
+2B, we see in Fig.
+3B and 3C a massive shift in 354 paired map species due to constraining coral depths to 200 m.
+Membership in the "well-aligned" quadrant jumped from 22.4% to 76.2%, with a corresponding decrease in all other quadrants.
+
+(D) We cannot know for certain the true distribution of each of these corals; much of the apparent over-prediction of coral ranges is likely due to very conservative and precautionary approach.
+But a small and sensible shift in method drastically decreases the introduction of likely commission errors, greatly improving our confidence in the remaining reported ranges.
+Note that this change applies just as well to the IUCN coral maps that are not included in the paired map analysis, and likely to other reef-associated flora and fauna.
+While species depth preferences are an easy and consistent means of constraining range predictions, other conditions such as salinity and temperature could be cautiously used to refine the results of expert opinions.
+
+***What is missing here is something to show that AquaMaps ranges, clipped to FAO Major Fishing Areas, are probably introducing errors of commission as well*** - _the fish don't see that FAO boundary and suddenly decide not to cross; most obvious in the boundaries defined by longitudes.
+Can we show this with a couple of example maps?
+That wasn't received well for the corals, but I can't think of a clear and ambiguous way to analyze this problem._
 
@@
 
+Neither IUCN nor AquaMaps claims a lock on the "truth", and the small overlap in mapped species means each provides value in taxonomic and geographic coverage.
 Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.
-To examine the implications of these differences, we repeated two recent studies, varying only the prioritization of one data set over the other.
+To examine the implications of these differences, we repeated two recent studies, varying only the selection of one data set over the other.
 
@@
 To maximize the number of represented species, OHI gleans spatial distribution data from Red List-assessed species in both IUCN and AquaMaps datasets (n = `r format(n_spp_ohi, big.mark = ',')`), prioritizing IUCN data for the `r format(n_spp_both_ohi, big.mark = ',')` species included in both sources.
-OHI currently uses a probability of occurrence threshold of 40% to determine species presence for AquaMaps data.
 
-We calculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig.
+___OMIT:___ _We calculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig.
 3</a>).
@@
 Shifting priority from IUCN to AquaMaps may increase or decrease a species’ apparent range, so a decrease in region score for scenario 2 may indicate a decrease in spatial representation of low-risk species, an increase in spatial representation of high-risk species, or a combination of both.
-Scenario three combines effects of scenarios 1 and 2.
+Scenario three combines effects of scenarios 1 and 2._
 
-Given that only `r p_spp_both_ohi`% of species are included in both datasets, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in substantial country-level score shifts as seen in scenarios 2 and 3.
+___OMIT:___ _Given that only `r p_spp_both_ohi`% of Red List-assessed species are included in both datasets, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in substantial country-level score shifts as seen in scenarios 2 and 3.
 While the mean global score did not vary significantly from scenario to scenario, select countries gained up to 7.5 points while others dropped as many as 5 points.
-This result indicates that especially on a national or regional scale, an arbitrary change in how the two datasets are combined can result in a different assessment of species conservation status.
+This result indicates that especially on a national or regional scale, an arbitrary change in how the two datasets are combined can result in a different assessment of species conservation status._
 
@@
 AquaMaps and IUCN range maps show reasonable agreement for many well-studied species, but substantial differences illustrate uncertainty in our understanding of spatial distribution for many others.
-Although many other approaches exist for species distribution modeling, these two are the only ones applied globally across different marine taxa.
+Although many other approaches exist for species distribution modeling, these two are the largest approaches applied globally across a broad range of marine taxa.
 Identifying and addressing differences in these datasets will increase their utility for research and conservation actions.
-Several likely drivers of commission and omission errors between these datasets point to a few important ways to improve range alignment.
 
-For IUCN range data, clipping ranges to known depth limits improves output for many species, most notably corals and reef-associated fishes.
-If species' depth limits are not known, simple rules of thumb will likely reduce commission errors without introducing substantial omission errors.
-For example, for most corals, researchers could clip range maps to the photosynthetic limit of 200 meters.
- For AquaMaps range data, dependent primarily on environmental and physical preferences and conditions, implementing area restrictions based on biogeographical criteria such as Marine Ecoregions of the World [@spalding_marine_2007] would likely decrease commission errors and improve predictive power, especially for data-poor species.
- For either data set, maps based on few occurrences are more likely to include errors of commission and omission.
+For IUCN range data, clipping ranges to known depth limits improves confidence in predicted ranges for coral species; similarly, depth may be a useful characteristic to consider for other reef-associated organisms.
+Other parameters such as salinity and temperature may provide useful constraints on other taxa.
+Simple but conservative rules of thumb will likely reduce commission errors without introducing substantial omission errors.
+ For AquaMaps range data, dependent primarily on environmental and physical preferences and conditions, implementing area restrictions based on biogeographical criteria such as Marine Ecoregions of the World [@spalding_marine_2007], rather than political and economic criteria such as FAO Major Fishing Areas, would likely decrease commission errors and improve predictive power, especially for data-poor species.
+ For either data set, maps based on few occurrences are more likely to bear high uncertainty in range predictions.
 Occurrence counts from external sources such as OBIS or GBIF can help identify relatively data-poor species; additionally, the "occurcells" attribute in the AquaMaps data set, which counts the number of half-degree cells used to generate the environmental envelope for each species, can be used in a similar manner.
 
-Many studies using AquaMaps data test the sensitivity of results by varying the probability of occurrence threshold to determine presence.
+___OMIT:___ _Many studies using AquaMaps data test the sensitivity of results by varying the probability of occurrence threshold to determine presence.
 This decision ultimately represents a tradeoff between errors of commission (low threshold) and omission (high threshold).
 Using AquaMaps in conjunction with IUCN can mitigate potential errors, while also increasing the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized.
-In this case, we recommend a presence threshold of 0% as it most closely approximates the "limits of distribution" criterion defined by IUCN data providers.
+In this case, a presence threshold of 0% most closely approximates the "limits of distribution" criterion defined by IUCN data providers._
 

### 2016-08-16 ms_5_figures.Rmd
cleaning out old plots; updating figures; editing manuscript

@@
 
-### Figure 3
-
+### Fig. 3
 <a id = 'fig3'></a>
 
-![paired map comparison](figures/fig3a_violinplot_paired_maps.png)
+![coral area change due to depth clipping](figures/fig3a_coral_area_plot.png)
 
-why is the global one not showing up?:
-![global full species list comparison](figures/fig3a_violinplot_ohi_global.png)
+![coral quadmap change due to depth clipping](figures/fig3b_scatter_quadmap_corals.png)
+
+![coral quadmap breakdown](figures/fig3c_barchart_coral_quads.png)
 
@@
 
-![Change in MPA coverage](figures/fig4_barchart_mpa_gap.png)
+![paired map comparison](figures/fig4a_violinplot_paired_maps.png)
 
-### New figures - corals
+why is the global one not showing up?:
+![global full species list comparison](figures/fig4a_violinplot_ohi_global.png)
 
-![coral area change due to depth clipping](figures/fig5a_coral_area_plot.png)
+-----
 
-![coral quadmap change due to depth clipping](figures/fig5b_scatter_quadmap_corals.png)
+### Figure 5
+
+<a id = 'fig5'></a>
+
+![Change in MPA coverage](figures/fig5_barchart_mpa_gap.png)
 

### 2016-08-17 data_explore/data_explore.Rmd
incorporating Jamie's comments

@@
 
-``` {r plot_bar_charts_by_taxa_old, echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE}
-# ### bar charts comparing number of species in each class for each data set
-# 
-# #library(colorspace)
-# #hclvec <- heat_hcl(3); colvec <- c('AquaMaps' = hclvec[1], 'Both' = hclvec[2], 'IUCN' = hclvec[3])
-# #colvec <- c('AquaMaps' = '#1b9e77', 'Both' = '#565080', 'IUCN' = '#d95f02')
-# colvec <- c('AquaMaps' = '#76A94C', 'Both' = '#e18258', 'IUCN' = '#9168a9')
-# 
-# taxa_df_long <- taxa_df %>%
-#   dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
-#   gather(dataset, n_spp, AquaMaps:Both) %>%
-#   group_by(class) %>%
-#   mutate(n_total = sum(n_spp, na.rm = TRUE),
-#          percent = n_spp/n_total * 100,
-#          dataset = factor(dataset, levels = c('AquaMaps', 'Both', 'IUCN'), ordered = TRUE)) %>%
-#   arrange(desc(dataset))
-# 
-# break_nums <- seq(0, 100, 20)
-# 
-# bar_taxa <- ggplot(taxa_df_long, aes(x = class, y = percent, fill = dataset)) +
-#   ggtheme_plot + 
-#   theme(panel.grid.major.x = element_blank()) +
-#   geom_bar(stat = 'identity', alpha = 1) +
-#   geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
-#   scale_fill_manual(values = colvec) +
-#   geom_text(aes(label = sprintf('n = %s', n_total), y = 102), 
-#             size = 2, 
-#             angle = 90, hjust = 0, color = 'grey30') +
-#   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
-#   scale_y_continuous(expand = c(0, 0), 
-#                      limits = c(0, 130),
-#                      breaks = break_nums,
-#                      labels = sprintf('%s%%', break_nums)) + 
-#   labs(# title = 'Species maps by taxonomic class',
-#        x = 'Class', 
-#        y = 'Percent of mapped species',
-#        fill = 'Dataset')
-# 
-# print(bar_taxa)
-# ggsave(filename = file.path(dir_fig, 'fig1a_bar_by_taxa_all.png'), 
-#        width = 17.8, height = 7, units = 'cm', dpi = 600)
-
-```
-
 ``` {r plot_barline_charts_by_taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
@@
 taxa_df1 <- taxa_df %>%
-  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
+  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Paired = n_class_both) %>%
   mutate(AquaMaps = ifelse(is.na(AquaMaps), 0, AquaMaps),
          IUCN     = ifelse(is.na(IUCN),     0, IUCN),
-         Both     = ifelse(is.na(Both),     0, Both),
-         n_total  = AquaMaps + IUCN + Both,
-         IUCN_end = 100 * (IUCN + Both)/n_total,
+         Paired     = ifelse(is.na(Paired),     0, Paired),
+         n_total  = AquaMaps + IUCN + Paired,
+         IUCN_end = 100 * (IUCN + Paired)/n_total,
          IUCN_start = 0,
          AquaMaps_end   = 100,
-         AquaMaps_start = 100 * (n_total - AquaMaps - Both)/n_total,
-         Both_end = AquaMaps_start,
-         Both_start = IUCN_end) %>%
+         AquaMaps_start = 100 * (n_total - AquaMaps - Paired)/n_total,
+         Paired_end = AquaMaps_start,
+         Paired_start = IUCN_end) %>%
   gather(key = dataset,  value = start, ends_with('start')) %>%
@@
   summarize(class = 'Total',
-            AquaMaps_total = sum(AquaMaps) + sum(Both),
-            IUCN_total = sum(IUCN) + sum(Both),
+            AquaMaps_total = sum(AquaMaps) + sum(Paired),
+            IUCN_total = sum(IUCN) + sum(Paired),
             n_total  = sum(n_total),
@@
             AquaMaps_start = 100 * (n_total - AquaMaps_total)/n_total,
-            Both_end = AquaMaps_start,
-            Both_start = IUCN_end) %>%
+            Paired_end = AquaMaps_start,
+            Paired_start = IUCN_end) %>%
   select(-dataset) %>% distinct() %>%
@@
   bind_rows(taxa_df1 %>%
-              select(-IUCN, -AquaMaps, -Both)) %>%
+              select(-IUCN, -AquaMaps, -Paired)) %>%
   mutate(class = factor(class, levels = unique(.$class), ordered = TRUE),
-         n_label = ifelse(!is.na(n_total), paste0('n = ', n_total), ''))
+         n_label = ifelse(!is.na(n_total), paste0('n = ', n_total), ''),
+         dataset = factor(dataset, levels = c('AquaMaps', 'Paired', 'IUCN'), ordered = TRUE))
 
@@
                  position = position_dodge(width = 0.5)) +
-  scale_size_manual(values = c('IUCN' = .25, 'AquaMaps' = .25, 'Both' = 2.5)) +
-  scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey60', 'Both' = 'black')) +
+  scale_size_manual(values = c('IUCN' = .25, 'AquaMaps' = .25, 'Paired' = 2.5)) +
+  scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey60', 'Paired' = 'black')) +
   geom_text(aes(label = n_label, y = 102),

### 2016-08-17 ms_3_res_disc_concl.Rmd
incorporating Jamie's comments

@@
 
-Neither IUCN nor AquaMaps claims a lock on the "truth", and the small overlap in mapped species means each provides value in taxonomic and geographic coverage.
-Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.
-To examine the implications of these differences, we repeated two recent studies, varying only the selection of one data set over the other.
+__Neither IUCN nor AquaMaps claims a lock on the "truth", and the small overlap in mapped species means each provides value in taxonomic and geographic coverage.
+Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.__  To examine the implications of these differences, we repeated two recent studies, varying only the selection of one data set over the other.
 

### 2016-08-17 ms_2_methods.Rmd
commit some changes. fuck. gonna make a conflict.

--- 2016-08-15 ms_2_methods.Rmd
@@
 
-___short paragraph about other datasets we didn't include, i.e. SAUP?___
+___short paragraph about other datasets we didn't include, i.e. SAUP?  BOTW?___
 
@@
 1A), we grouped species by taxonomic class and data source, and determined the proportion of each class represented in each dataset.
-We defined "paired map" species as those with spatial data in both datasets.
 
@@
 Using genus and species binomials as a matching key, we selected the "paired map" species - the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.
-_We used the taxize package [@REF] in R [@REF] to standardize species names and synonyms; for species with separate subpopulation maps in IUCN, we combined all subpopulations to create a single global population._  ___To avoid double-counting, we removed subpopulations and species aliases.
-Used the taxize package?
-which version?___  We determined species presence within each spatial cell using the same criteria outlined above.
+_We used the taxize package [@REF] in R [@REF] to standardize species names and synonyms; for species with separate subpopulation maps in IUCN, we combined all subpopulations to create a single global population._   For each of these paired map species, we determined species presence within each spatial cell for each dataset, using the same criteria outlined above.
 
@@
 _To highlight the differences in the datasets, we recalculated the OHI SPP status values using only one complete dataset at a time.
-For the AquaMaps dataset, we used a presence threshold of 0% to align with the OHI 2016 assessment._  ___Fig.
+For the AquaMaps dataset, we used a presence threshold of 0% to align with the OHI 2016 assessment.
+Note that the OHI 2016 analysis includes the Birds of the World dataset in addition to IUCN and AquaMaps; we excluded it for our analysis._ ___Fig.
 3 compares these single-dataset scores to the 2016 OHI SPP status values.___
 
-_To highlight the differences in the datasets, we selected the subset of species whose ranges are represented in both datasets, eliminating species found in only one or the other dataset.
+~To highlight the differences in the datasets, we selected the subset of species whose ranges are represented in both datasets, eliminating species found in only one or the other dataset.
 In the original OHI SPP analysis, IUCN data was used to determine ranges for these overlapping species; we calculated all regional SPP status scores for just these species as our baseline.
-We then recalculated the scores using AquaMaps ranges at a 40% presence threshold (as used in the original OHI analysis) and a 0% threshold (to better simulate the "limits of distribution" defined by IUCN maps) and compared them to our IUCN baseline._
+We then recalculated the scores using AquaMaps ranges at a 40% presence threshold (as used in the original OHI analysis) and a 0% threshold (to better simulate the "limits of distribution" defined by IUCN maps) and compared them to our IUCN baseline.~
 
-_We then modified the original OHI source code to recalculate the 2015 SPP status scores for all regions and all species to examine the effects of changing the AquaMaps presence threshold from 40% to 0% (all range included), and the effects of changing the data source for the overlapping species from IUCN to AquaMaps.
-We calculated the differences between these results and the published SPP status scores for 2015._
+~We then modified the original OHI source code to recalculate the 2015 SPP status scores for all regions and all species to examine the effects of changing the AquaMaps presence threshold from 40% to 0% (all range included), and the effects of changing the data source for the overlapping species from IUCN to AquaMaps.
+We calculated the differences between these results and the published SPP status scores for 2015.~
 

### 2016-08-17 case_study_ohi/ohi_single_datasets.Rmd

@@
   filter(am_sid %in% ohi_spp_info$am_sid) %>% 
-  filter(loiczid %in% ohi_rgn_cells$loiczid)
+  filter(loiczid %in% ohi_rgn_cells$loiczid) %>%
+  distinct()
 
-### bind IUCN loiczid dataframe to species risk category dataframe
+### bind AM loiczid dataframe to species risk category dataframe
 am_spp_cells <- am_spp_cells %>%
@@
             by = 'am_sid')
+
+### force exclusion of bird IUCN SIDs for IUCN analysis only
+iucn_nonbirds <- ohi_spp_info %>%
+  select(iucn_sid, spp_group) %>%
+  filter(spp_group != 'BOTW')
 
@@
   filter(iucn_sid %in% ohi_spp_info$iucn_sid) %>%
-  filter(loiczid %in% ohi_rgn_cells$loiczid)
+  filter(iucn_sid %in% iucn_nonbirds$iucn_sid) %>% ### exclude BOTW
+  filter(loiczid %in% ohi_rgn_cells$loiczid) %>%
+  distinct()
 ### bind IUCN loiczid dataframe to species risk category dataframe
@@
   filter(!is.na(cat_score)) %>%
+  distinct() %>%
   group_by(loiczid) %>%
@@
              by = 'loiczid') %>%
-  # mutate(rgn_area = cell_area * proportionArea,
-  #        area_weighted_mean_cat   = weighted_mean_cat   * rgn_area,
-  #        area_weighted_mean_trend = weighted_mean_trend * rgn_area) %>%
   mutate(cell_area_weight = cell_area * n_spp * proportionArea,
@@
 iucn_rgn_sum <- iucn_rgn_sum %>%
-  group_by(rgn_id) %>%
+  group_by(rgn_id, rgn_name) %>%
   summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight),
@@
 am_rgn_sum <- am_rgn_sum %>%
-  group_by(rgn_id) %>%
+  group_by(rgn_id, rgn_name) %>%
   summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight),
@@
 ohi_rgn_sum <- ohi_rgn_sum %>%
-  group_by(rgn_id) %>%
+  group_by(rgn_id, rgn_name) %>%
   summarize(rgn_mean_cat = sum(area_weighted_mean_cat)/sum(cell_area_weight),
@@
 status_df <- ohi_rgn_sum %>%
-  select(rgn_id, ohi_st = status, mean_n_spp) %>%
+  select(rgn_id, rgn_name, ohi_st = status, mean_n_spp) %>%
+  distinct() %>%
   mutate(ohi_st = ohi_st) %>%
-  full_join(am_rgn_sum   %>% select(rgn_id,    am_st = status), by = 'rgn_id') %>%
-  full_join(iucn_rgn_sum %>% select(rgn_id, iucn_st = status), by = 'rgn_id') %>%
+  full_join(am_rgn_sum   %>% 
+              select(rgn_id, am_st = status) %>%
+              distinct(), 
+            by = 'rgn_id') %>%
+  full_join(iucn_rgn_sum %>% 
+              select(rgn_id, iucn_st = status) %>%
+              distinct(), 
+            by = 'rgn_id') %>%
   mutate(diff_am   = am_st - ohi_st ,
@@
 ```
+
+``` {r check_weirdness}
+### noting that the mean and median by AM only and by IUCN only are both less than OHI...
+
+status_df$am_st %>% mean()     ### 91.352
+status_df$am_st %>% median()   ### 92.164
+status_df$iucn_st %>% mean()   ### 90.943
+status_df$iucn_st %>% median() ### 92.565
+status_df$ohi_st %>% mean()    ### 93.628
+status_df$ohi_st %>% median()  ### 93.803
+
+am_cell_sum_all <- read_csv(file.path(dir_git, 'case_study_ohi/am_spp_cell_sum.csv'))
+iucn_cell_sum_all <- read_csv(file.path(dir_git, 'case_study_ohi/iucn_spp_cell_sum.csv'))
+ohi_cell_sum_test <- iucn_cell_sum_both %>%
+  rename(mean_cat_i = mean_cat, n_spp_i = n_spp) %>%
+  select(-data) %>%
+  full_join(am_cell_sum_both %>%
+              rename(mean_cat_a = mean_cat, n_spp_a = n_spp) %>%
+              select(-data), 
+            by = 'loiczid')
+
+spp_info_trunc <- ohi_spp_info %>%
+  select(am_sid, iucn_sid, sciname, cat_score, spp_group, spatial_source)
+
+spp_ohi_am <- spp_info_trunc %>% 
+  filter(str_detect(spatial_source, 'am'))
+spp_ohi_iucn <- spp_info_trunc %>% 
+  filter(str_detect(spatial_source, 'iucn'))
+
+### examine one cell with lots of species to see what's happening
+a_all <- am_spp_cells %>%
+  filter(loiczid == 161563) %>%
+  inner_join(spp_info_trunc) %>%
+  distinct()
+a_ohi <- am_spp_cells %>%
+  filter(loiczid == 161563) %>%
+  inner_join(spp_ohi_am) %>%
+  mutate(type = 'ohi') %>%
+  distinct()
+a <- a_all %>%
+  full_join(a_ohi)
+i_all <- iucn_spp_cells %>%
+  filter(loiczid == 161563) %>%
+  inner_join(spp_info_trunc) %>%
+  distinct()
+i_ohi <- iucn_spp_cells %>%
+  filter(loiczid == 161563) %>%
+  inner_join(spp_ohi_iucn) %>%
+  mutate(type = 'ohi') %>%
+  distinct()
+i <- i_all %>%
+  full_join(i_ohi)
+
+# > mean(a$cat_score)
+# [1] 0.04671916
+# > mean(a$cat_score[is.na(a$type)])
+# [1] 0.2727273
+### for OHI, 44 (out of 381) relatively unhealthy species are removed from AM presence for this cell,
+### because the presence is driven by IUCN maps.
+### If they aren't replaced by IUCN presence, then this cell's score will seem artificially
+### healthier.
+
+### OK check a region with large IUCN changes and small spp numbers:
+### rgn 4 Macquarie Island
+
+macq_cells <- ohi_rgn_cells %>% filter(rgn_id == 4)
+### examine one rgn with lots of species to see what's happening
+a_rgn_all <- am_spp_cells %>%
+  filter(loiczid %in% macq_cells$loiczid) %>%
+  inner_join(spp_info_trunc) %>%
+  distinct() %>%
+  group_by(am_sid, cat_score, sciname) %>%
+  summarize(n_cells_spp = n())
+a_rgn_ohi <- am_spp_cells %>%
+  filter(loiczid %in% macq_cells$loiczid) %>%
+  inner_join(spp_ohi_am) %>%
+  distinct() %>%
+  group_by(am_sid, cat_score, sciname) %>%
+  summarize(n_cells_spp = n()) %>%
+  mutate(type = 'ohi')
+a_rgn <- a_rgn_all %>%
+  full_join(a_rgn_ohi)
+i_rgn_all <- iucn_spp_cells %>%
+  filter(loiczid %in% macq_cells$loiczid) %>%
+  inner_join(spp_info_trunc) %>%
+  distinct() %>%
+  group_by(iucn_sid, cat_score, sciname) %>%
+  summarize(n_cells_spp = n())
+i_rgn_ohi <- iucn_spp_cells %>%
+  filter(loiczid %in% macq_cells$loiczid) %>%
+  inner_join(spp_ohi_iucn) %>%
+  distinct() %>%
+  group_by(iucn_sid, cat_score, sciname) %>%
+  summarize(n_cells_spp = n()) %>%
+  mutate(type = 'ohi')
+i_rgn <- i_rgn_all %>%
+  full_join(i_rgn_ohi)
+
+a_rgn_overlap <- a_rgn %>% filter(is.na(type)) %>%
+  ungroup() %>%
+  mutate(weighted_cat = cat_score * n_cells_spp,
+         weighted_cat_mean = sum(weighted_cat) / sum(n_cells_spp))
+a_rgn_overlap$cat_score %>% mean()
+a_rgn_overlap$n_cells_spp %>% mean()
+a_rgn_overlap$weighted_cat_mean[1]
+### Removed these species and replaced with IUCN:
+### 0.1155556 mean cat score; 149.1556 mean cells; 45 species; mean weighted cat .2553
+
+i_rgn_replace <- i_rgn %>% filter(sciname %in% a_rgn_overlap$sciname) %>%
+  ungroup() %>%
+  mutate(weighted_cat = cat_score * n_cells_spp,
+         weighted_cat_mean = sum(weighted_cat) / sum(n_cells_spp))
+i_rgn_replace$cat_score %>% mean()
+i_rgn_replace$n_cells_spp %>% mean()
+i_rgn_replace$weighted_cat_mean[1]
+### These species overwrote AM species for combined score:
+### 0.225 mean cat score; 260.2 mean cells; 16 spp; .191 mean weighted cat
+
+look at weighted means globally

### 2016-08-17 ms_1_intro.Rmd

@@
 One major outcome of this body of science is the various compiled databases of species distribution maps.
-Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps [@kaschner_aquamaps:_2015] and International Union for Conservation of Nature (IUCN) [@iucn_iucn_2015].
+Here we focus on global-scale repositories that predict marine species ranges throughout the world's oceans – AquaMaps [@kaschner_aquamaps:_2015] and International Union for Conservation of Nature (IUCN) [@iucn_iucn_2015].
 These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
@@
 
-<!--
-The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
-Discuss "truth" and uncertainty in species distributions.
-There is an objective "truth"; 
--->
+__Neither IUCN nor AquaMaps claims a lock on the "truth", and the small overlap in mapped species means each provides value in taxonomic and geographic coverage.
+Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.__
 
-Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each presents one understanding of the "truth" - one database relying primarily on expert opinion and the other on model predictions.
-Uncertainties inherent in the method and intent inevitably drive differences in the range predictions made by each dataset; significant differences could indicate dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
+Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each presents an understanding of the "truth" - one relying primarily on expert opinion and the other on model predictions, each method with its own challenges and advantages.
+Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset; significant differences could indicate dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
 Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans [@rondinini_tradeoffs_2006; @jetz_ecological_2008].
@@
 
-To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of marine species.
+To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of the `r format(n_spp_all, big.mark = ',')` marine species mapped by one or both datasets.
 Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).
 
-___It seems fair to assume that/we can reasonably assume that___ range predictions from each dataset would be more closely aligned for well-understood species.
-For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and suggest methods to improve confidence in species range predictions.
+As understanding of a given species improves, we expect that range maps, regardless of method, would become more accurate in predicting both species presence and absence.
+Close alignment between two independent range maps can therefore indicate higher confidence in species presence or absence than a single map.
+For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align in both distribution and overall area; from the results we determine several issues that can introduce commission errors, and suggest methods to improve confidence in species range predictions.
 

### 2016-08-17 ms_1_intro.Rmd
sent to BH, JA, CS

@@
 something about not focusing just on economically important species?
-(is that a valid assumption about SAUP)___
+(is that a valid assumption about SAUP) - consider other datasets as well: NatureServe, BOTW, etc___
 
@@
 
-__Neither IUCN nor AquaMaps claims a lock on the "truth", and the small overlap in mapped species means each provides value in taxonomic and geographic coverage.
-Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.__
+Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each offers an understanding of the "truth" - one relying primarily on expert opinion and the other on model predictions.
+Each method presents its own challenges and advantages, and the small overlap in mapped species means each dataset provides value in taxonomic and geographic coverage.
+Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
 
-Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each presents an understanding of the "truth" - one relying primarily on expert opinion and the other on model predictions, each method with its own challenges and advantages.
-Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset; significant differences could indicate dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.
-Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans [@rondinini_tradeoffs_2006; @jetz_ecological_2008].
-Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
+Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: geographic range data such as IUCN range maps frequently introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
+Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006; @jetz_ecological_2008].
 
@@
 
-As understanding of a given species improves, we expect that range maps, regardless of method, would become more accurate in predicting both species presence and absence.
+As understanding of a given species improves, we expect that range maps, regardless of method, would become more accurate in predicting species presence and absence.
 Close alignment between two independent range maps can therefore indicate higher confidence in species presence or absence than a single map.
-For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align in both distribution and overall area; from the results we determine several issues that can introduce commission errors, and suggest methods to improve confidence in species range predictions.
+For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align in both distribution and overall area; from the results we determine methodological issues that can introduce commission errors, and suggest methods to improve confidence in species range predictions.
 
@@
 
-While we cannot recommmend one data set as more accurate than the other, we suggest methods to improve confidence in species range predictions to better inform conservation management and policy decisions.
-
-___acknowledge or debunk other global data sets - e.g. SAUP?___
+Each of these datasets provides unique value in taxonomic and geographic coverage; while we have identified some potential issues in each, we cannot simply recommmend one data set over the other.
+Instead, we recommend simple methods to reduce incidence of commission errors in each dataset, improving confidence in species range predictions to better inform conservation management and policy decisions.

### 2016-08-17 ms_2_methods.Rmd
sent to BH, JA, CS

@@
-# Methods
+# Methods and Analysis
 
-_About the datasets__:
+## About the datasets
 
-The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" [@iucn_iucn_2015] - essentially a refined extent of occurrence (EOO), based on observation records and informed by expert understanding of species' range and habitat preferences.
-In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of suitable habitat.
-The model masks habitat to certain FAO Major Fishing Areas as a rough cut of geographic range, resulting in a "probability of occurrence" for each species.
-Of the resulting maps, 1296 (5.7%) have been further refined under expert review [@kaschner_mapping_2006; @ready_predicting_2010].
-Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., previously, the Ocean Health Index defined presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
-In the 2016 OHI assessment, the presence threshold was eliminated, to better approximate the "extent of occurrence" generally indicated by IUCN maps, since the two datasets are used together.
+The IUCN publishes species range maps developed by species experts.
+These experts outline spatial boundaries of a given species' "limits of distribution" [@iucn_iucn_2015], based on observation records and informed by expert understanding of species' range and habitat preferences.
+___Fig.
+SXXX___
 
-___short paragraph about other datasets we didn't include, i.e. SAUP? BOTW?___
+In contrast, AquaMaps models species distribution based on environmental preferences, such as temperature, depth, and salinity, deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a global 0.5° grid to determine suitable habitat, resulting in a "probability of occurrence" for each species.
+Of these, 1296 (5.7%) have been further refined through an expert review process [@kaschner_mapping_2006; @ready_predicting_2010].
+___Fig.
+SXXX___
+
+___OMIT: Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., previously, the Ocean Health Index defined presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
+In the 2016 OHI assessment, the presence threshold was eliminated, to better approximate the "extent of occurrence" generally indicated by IUCN maps, since the two datasets are used together.___
 
@@
 __Comparison of paired maps__:
-Using genus and species binomials as a matching key, we selected the "paired map" species - the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.
-_We used the taxize package [@REF] in R [@REF] to standardize species names and synonyms; for species with separate subpopulation maps in IUCN, we combined all subpopulations to create a single global population._   For each of these paired map species, we determined species presence within each spatial cell for each dataset, using the same criteria outlined above.
+Using genus and species binomials as a matching key, we identified "paired map" species - the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.
+We used the taxize package [@REF] in R [@REF] to standardize species names and synonyms; for species with separate subpopulation maps in IUCN, we combined all subpopulations to create a single global population.
+For each of these paired map species, we determined species presence within each spatial cell for each dataset using the same criteria outlined above.
 
-Overlaying paired distribution maps for each species, we defined and calculated _distribution alignment_ and _area ratio_ and plotted these in Fig. 2A:
+Overlaying paired distribution maps for each species, we defined and calculated _distribution alignment_ $\alpha_{dist}$ and _area ratio_ $\alpha_{area}$:
 
@@
 
-_Distribution alignment uses overlapping predictions of presence as means of identifying areas of high confidence.  Area ratio provides a proxy for frequency of commission and/or omission errors._
+Distribution alignment uses overlapping predictions of presence as means of identifying areas of higher confidence.  Area ratio provides a proxy for frequency of commission and/or omission errors.
 
-Dividing the plot into quadrants based on mean values along each axis, we roughly categorized four qualities of alignment, and for each taxonomic group, based on IUCN taxonomic groupings, determined the proportion of paired map species that fell into each quality category (2B).
+To examine errors of commission related to depth, we selected corals as a case study due to their importance in supporting biodiversity as well as their dependence on photosynthesis.
+We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] ___(Fig.
+SXXX)___ and masked our IUCN coral rasters to identify mapped coral presence below 200 m.
+The resulting maps were again compared to the AquaMaps rasters to examine distribution alignment and area ratio.
 
-_Based on the results of the map alignment analysis, we chose to examine the effect of explicitly restricting IUCN ranges to depths based on species' life histories.
-We selected corals as a case study, due to their foundational role in biodiverse habitats, as well as their lack of mobility and general reliance on photosynthesis._  Using the IUCN's API [REF] to find life history information for all IUCN-mapped corals, we found that none of the paired-map corals is described as occurring deeper than 200 m; 94% are noted as being limited to 50 m or shallower.
-We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] ___(Fig.
-SXXX)___ and masked our IUCN coral rasters to exclude presence below 200 m.
-The resulting maps were again compared to the AquaMaps rasters; ___Fig.
-XXXX___ plots shifts in coral species alignment due to this depth clipping.
+To identify data-poor species, we used the ___robis???___ and ___rgbif???___ packages [@REFS] in R to identify known occurrences of each species; occurrences were averaged between the two occurrence databases as a proxy for data richness.
 
-___To identify data-poor species, we used the robis???
-and rgbif???
-packages in R to identify known occurrences of each species; occurrences were averaged between the two bases as a proxy for data richness.___
+## Methods for Ocean Health Index (OHI) case study
 
-**Methods for Ocean Health Index (OHI) case study:**
-The Species component of the OHI Biodiversity goal (SPP) uses IUCN Red List extinction risk as a proxy for a species' health.
+The global Ocean Health Index (OHI) [@halpern_index_2012; @halpern_patterns_2015], a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data to calculate biodiversity status for each of the world's 221 exclusive economic zones.
+
+The Species component of the OHI Biodiversity goal (abbreviated as SPP) uses IUCN Red List extinction risk as a proxy for a species' health.
 Each species is assigned a risk score based on IUCN Red List extinction risk category, from Least Concern (0.0) to Extinct (1.0) (Data Deficient and unassessed species are excluded).
@@
 
-_To highlight the differences in the datasets, we recalculated the OHI SPP status values using only one complete dataset at a time.
-For the AquaMaps dataset, we used a presence threshold of 0% to align with the OHI 2016 assessment.
-Fig.
-3 compares these single-dataset scores to the 2016 OHI SPP status values.
-While the full OHI 2016 assessment includes the Birds of the World dataset [@REF???] in addition to IUCN and AquaMaps, we omitted it from our calculations for simplicity._
+_To highlight the differences in the datasets, we recalculated the OHI SPP status values using only one complete dataset at a time, and compared to the values resulting from the combined datasets.
+For the AquaMaps dataset, we used a "probability of occurrence" threshold of 0% to determine presence, to align with the OHI 2016 assessment._
 
-~To highlight the differences in the datasets, we selected the subset of species whose ranges are represented in both datasets, eliminating species found in only one or the other dataset.
-In the original OHI SPP analysis, IUCN data was used to determine ranges for these overlapping species; we calculated all regional SPP status scores for just these species as our baseline.
-We then recalculated the scores using AquaMaps ranges at a 40% presence threshold (as used in the original OHI analysis) and a 0% threshold (to better simulate the "limits of distribution" defined by IUCN maps) and compared them to our IUCN baseline.~
-
-~We then modified the original OHI source code to recalculate the 2015 SPP status scores for all regions and all species to examine the effects of changing the AquaMaps presence threshold from 40% to 0% (all range included), and the effects of changing the data source for the overlapping species from IUCN to AquaMaps.
-We calculated the differences between these results and the published SPP status scores for 2015.~
-
-**Methods for MPA Gap Analysis case study:** 
+## Methods for MPA Gap Analysis case study
 To assess the effectiveness of MPAs in protecting biodiversity, Klein et al.
 [-@klein_shortfalls_2015] compared the coverage of the global MPA network presented by the World Database on Protected Areas (WDPA) [@iucn_world_2014] to the species ranges described in the 2014 AquaMaps dataset [@aquamaps_XXX].
-For the primary analysis, the researchers defined species "presence" as 50% or greater probability of occurrence.
+For the primary analysis, the researchers defined species presence as 50% or greater probability of occurrence.
 
-To reconstruct the primary analysis, we selected the subset of protected areas from the 2014 WDPA dataset classified as IUCN protected area management categories I-IV and spatially covering a marine area.
+To reconstruct the primary analysis, we selected the subset of protected areas from the 2014 WDPA dataset classified as IUCN protected area management categories I-IV and spatially overlapping a marine area.
 The WDPA polygons and marine polygons were rasterized to 0.01° and then aggregated to AquaMaps native 0.5° cells, to calculate proportion of marine protected area within each cell.

### 2016-08-17 ms_3_res_disc_concl.Rmd
sent to BH, JA, CS

--- 2016-08-16 ms_3_res_disc_concl.Rmd
@@
 
-(R) The two datasets have notably different taxonomic (Fig.
+## Taxonomic and geographic coverage
+
+[results] The two datasets have notably different taxonomic (Fig.
 1A) and regional (Figs.
@@
 
-(D) _To achieve more comprehensive global coverage of species ranges these two datasets can be used together, understanding that the underlying differences complicate such direct comparisons.
-Major considerations include aligning the criteria used to determine "presence" (e.g.
-selecting an appropriate presence threshold for AquaMaps) and reducing mechanisms that introduce errors of commission and omission._
+<!-- need discussion here? [discussion] Each dataset offers spatial information on a -->
 
@@
 
-___from Methods___
-(D) _The methodologies behind the creation of these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).
-Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.
-Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006].
-By comparing maps resulting from IUCN and AquaMaps methodologies, we can identify and possibly address mechanistic causes for each type of error._
-
-(D/M) ___is this sufficiently described in methods?
-or revisit here?___ ___Both datasets strive for accuracy in different ways; the truth likely lies somewhere in between.___ To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area alignment (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig.
+[discussion] To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area ratio (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig.
 2A).
@@
 
-(R) This analysis revealed a general negative linear pattern, suggesting that increasing similarity in range area correlates with decreasing distribution alignment.
+[discussion] This analysis revealed a weak negative linear pattern, suggesting that increasing similarity in range area correlates with decreasing distribution alignment.
  AquaMaps tends to extrapolate species ranges into suitable areas beyond known occurrences, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
@@
 
-(R) The mean distribution alignment for species included in both datasets was `r dist_align_mean`%; the mean area alignment was `r area_align_mean`%.
+[results] The mean distribution alignment for species included in both datasets was `r dist_align_mean`%; the mean area alignment was `r area_align_mean`%.
 By dividing the paired map species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
@@
 
-(R) The upper right quadrant includes the species (n = `r n_q1`) whose described ranges are above average in alignment of both spatial distribution and area.
+[results/discussion] The upper right quadrant includes the species (n = `r n_q1`) whose described ranges are above average in alignment of both spatial distribution and area.
 These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig.
@@
 
-(R) The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where one range is notably larger than the other (For `r round(n_q2_iucn_larger/n_q2 * 100, 1)`% of the species in this quadrant, the IUCN range is an average of 2.57 times larger than the AquaMaps range).
-This suggests a high rate of commission and/or omission errors by one or both datasets; further analysis is required to disentangle the source or type of error contributing to poor area alignment.
+[results/discussion] The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where one range is notably larger than the other (For `r round(n_q2_iucn_larger/n_q2 * 100, 1)`% of the species in this quadrant, the IUCN range is an average of 2.57 times larger than the AquaMaps range).
+This suggests a high rate of commission and/or omission errors by one or both datasets; further analysis is required to disentangle the source and type of error contributing to poor area alignment.
 The results of the coral analysis (described below) provide some insight.
 
-(R) ___area ratio suggests similar commission/omission errors, but dist suggests otherwise___ Species found in the lower right quadrant (n = `r n_q3`) seem to represent cases of "two wrongs make a right."  For these species, IUCN and AquaMaps both predict ranges extending far beyond the overlapping region, but the methodological differences result in very different extrapolations.
-___OMIT:___ _For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, likely introducing additional commission errors._  Consequently, area ratios are close to 100%, though the poor distribution alignment indicates that one or both datasets are introducing errors.
+[results/discussion] ___area ratio suggests similar commission/omission errors, but dist suggests otherwise___ Species found in the lower right quadrant (n = `r n_q3`) seem to represent cases of "two wrongs make a right."  For these species, IUCN and AquaMaps both predict ranges extending far beyond the overlapping region, but the methodological differences result in very different extrapolations.
+___OMIT:___ _For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, likely introducing additional commission errors._  Consequently, area ratios are close to 100%, though the poor distribution alignment indicates that one or both datasets are introducing significant errors.
 As above, further analysis is required to disentangle the causes of error.
 
-(R) The lower left quadrant includes species (n = `r n_q4`) where alignment is poor in both dimensions.
+[results] The lower left quadrant includes species (n = `r n_q4`) where alignment is poor in both dimensions.
 ___present a solution here?___ Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
 It is not surprising that range maps based upon fewer observations bear greater uncertainty.
-___OMIT:___ _When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species._
 
-## New analysis: Coral maps (Distribution-aligned quads)
+## Case study: Coral depth analysis
 
-(D/M) Noting from Fig.
+[discussion] Noting from Fig.
 2B that corals dominate the upper-left "distribution-aligned" quadrant of Fig.
-2A (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), we clipped the IUCN coral range maps to a 200 m bathymetry raster to examine the impact on distribution and area alignment.
+2A (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), we chose to examine the effect of explicitly restricting IUCN ranges to depths based on species' life histories.
+Corals offer an excellent case study, due to their foundational role in biodiverse habitats, as well as their lack of mobility and reliance on photosynthesis.
 Ocean depth preference is explicitly included in AquaMaps models.
-While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." [@REF http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources]), it is clear that this cannot be taken for granted.
+While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." [@REF http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources]), it is not presented as a requirement, so we cannot take its inclusion for granted.
 
@@
 
-(R) Fig.
+[results] Fig.
 3A shows the 463 coral species mapped in the IUCN dataset, with their ranges broken into proportional area deeper and shallower than 200 m.
@@
 
-(R) Masking coral ranges to shallow waters reduces a clear source of commission errors.
-In so doing, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps.
+[results] In constraining coral ranges to shallow waters, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps.
 Using the original quadrant definitions from Fig.
@@
 
-(D) We cannot know for certain the true distribution of each of these corals; much of the apparent over-prediction of coral ranges is likely due to very conservative and precautionary approach.
-But a small and sensible shift in method drastically decreases the introduction of likely commission errors, greatly improving our confidence in the remaining reported ranges.
+[discussion] We cannot know for certain the true distribution of each of these corals; much of the apparent over-prediction of coral ranges is likely due to experts taking a conservative and precautionary approach.
+But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported ranges.
 Note that this change applies just as well to the IUCN coral maps that are not included in the paired map analysis, and likely to other reef-associated flora and fauna.
@@
 
-***What is missing here is something to show that AquaMaps ranges, clipped to FAO Major Fishing Areas, are probably introducing errors of commission as well*** - _the fish don't see that FAO boundary and suddenly decide not to cross; most obvious in the boundaries defined by longitudes.
+***What is still missing here is something to show that AquaMaps ranges, clipped to FAO Major Fishing Areas, are probably introducing errors of commission as well*** - _the fish don't see that FAO boundary and suddenly decide not to cross; most obvious in the boundaries defined by longitudes - sudden vertical edges in range.
 Can we show this with a couple of example maps?
-That wasn't received well for the corals, but I can't think of a clear and ambiguous way to analyze this problem._
+That wasn't received well for the corals, but I can't think of a clear and unambiguous way to analyze this problem._
 
-# Implications
+## Case Studies
 
-Neither IUCN nor AquaMaps claims a lock on the "truth", and the small overlap in mapped species means each provides value in taxonomic and geographic coverage.
-Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.
-To examine the implications of these differences, we repeated two recent studies, varying only the selection of one data set over the other.
+To examine the implications of differences between the AquaMaps and IUCN predicted species ranges, we repeated two recent studies, varying only the selection of one data set over the other.
 
-**Case Study: The Ocean Health Index**
+### Case Study: The Ocean Health Index
 
@@
 
-The global Ocean Health Index (OHI) [@halpern_index_2012; @halpern_patterns_2015], a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones.
-To maximize the number of represented species, OHI gleans spatial distribution data from Red List-assessed species in both IUCN and AquaMaps datasets (n = `r format(n_spp_ohi, big.mark = ',')`), prioritizing IUCN data for the `r format(n_spp_both_ohi, big.mark = ',')` species included in both sources.
-
-___OMIT:___ _We calculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig.
+___THIS SECTION NEEDS TO BE REWRITTEN if we choose not to omit it___
+_We calculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig.
 3</a>).
@@
 
-___OMIT:___ _Given that only `r p_spp_both_ohi`% of Red List-assessed species are included in both datasets, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in substantial country-level score shifts as seen in scenarios 2 and 3.
+_Given that only `r p_spp_both_ohi`% of Red List-assessed species are included in both datasets, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in substantial country-level score shifts as seen in scenarios 2 and 3.
 While the mean global score did not vary significantly from scenario to scenario, select countries gained up to 7.5 points while others dropped as many as 5 points.
@@
 
-**Case Study: MPA Gap Analysis**
+### Case Study: MPA Gap Analysis
 
@@
 1.5%).
-Conclusions drawn from each of these datasets would paint dramatically different pictures of the protection afforded by our current global MPA network.
 
@@
 
-AquaMaps and IUCN range maps show reasonable agreement for many well-studied species, but substantial differences illustrate uncertainty in our understanding of spatial distribution for many others.
+No spatial dataset can ever claim to know the "truth" of the whole of marine biodiversity.
+AquaMaps and IUCN range maps show strong agreement for many well-studied species, but substantial differences illustrate uncertainty in our understanding of spatial distribution for many others.
 Although many other approaches exist for species distribution modeling, these two are the largest approaches applied globally across a broad range of marine taxa.
+Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.
+Conclusions drawn from each of these datasets would paint dramatically different pictures of global marine biodiversity or the effectiveness of conservation management decisions.
 Identifying and addressing differences in these datasets will increase their utility for research and conservation actions.
@@
 For IUCN range data, clipping ranges to known depth limits improves confidence in predicted ranges for coral species; similarly, depth may be a useful characteristic to consider for other reef-associated organisms.
-Other parameters such as salinity and temperature may provide useful constraints on other taxa.
+Other parameters such as salinity and temperature may provide useful constraints on other taxa to bolster and refine expert opinion.
 Simple but conservative rules of thumb will likely reduce commission errors without introducing substantial omission errors.
+
 For AquaMaps range data, dependent primarily on environmental and physical preferences and conditions, implementing area restrictions based on biogeographical criteria such as Marine Ecoregions of the World [@spalding_marine_2007], rather than political and economic criteria such as FAO Major Fishing Areas, would likely decrease commission errors and improve predictive power, especially for data-poor species.
+
 For either data set, maps based on few occurrences are more likely to bear high uncertainty in range predictions.
 Occurrence counts from external sources such as OBIS or GBIF can help identify relatively data-poor species; additionally, the "occurcells" attribute in the AquaMaps data set, which counts the number of half-degree cells used to generate the environmental envelope for each species, can be used in a similar manner.
+The only certain remedy for data-poor species, of course, is further research.
 
-___OMIT:___ _Many studies using AquaMaps data test the sensitivity of results by varying the probability of occurrence threshold to determine presence.
+<!-- ___OMIT:___ _Many studies using AquaMaps data test the sensitivity of results by varying the probability of occurrence threshold to determine presence.
 This decision ultimately represents a tradeoff between errors of commission (low threshold) and omission (high threshold).
 Using AquaMaps in conjunction with IUCN can mitigate potential errors, while also increasing the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized.
-In this case, a presence threshold of 0% most closely approximates the "limits of distribution" criterion defined by IUCN data providers._
+In this case, a presence threshold of 0% most closely approximates the "limits of distribution" criterion defined by IUCN data providers._ -->
+
+To achieve more comprehensive global coverage of species ranges these two datasets can be used together, understanding that the underlying differences complicate such direct comparisons.
+Major considerations include aligning the criteria used to determine "presence" (e.g.
+selecting an appropriate presence threshold for AquaMaps; _a presence threshold of 0% most closely approximates the "limits of distribution" criterion described by IUCN range maps_ ___- not really discussed or backed up here___) and addressing mechanisms that introduce errors of commission and omission.
+___as it stands, this paper really doesn't go into this very much; omit this paragraph?___
 

### 2016-08-17 data_explore/data_explore.Rmd
updated spp inclusion breakdown to be ordered by number of species

@@
   bind_rows(taxa_df1 %>%
-              select(-IUCN, -AquaMaps, -Paired)) %>%
+              select(-IUCN, -AquaMaps, -Paired) %>%
+              arrange(desc(n_total))) %>%
   mutate(class = factor(class, levels = unique(.$class), ordered = TRUE),
@@
                      labels = sprintf('%s%%', break_nums)) +
-  labs(x = 'Class',
+  labs(x = 'Class (in order of decreasing number of included species)',
        y = 'Percent of mapped species',

### 2016-08-23 clip_depth/coral_maps.Rmd
messing around with the AquaMaps FAO clipping issue.  Plots and maps for quadrant 3 and 4 species...

--- 2016-08-16 coralmaps/coral_maps.Rmd
@@
 
-if(!file.exists(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))) {
+if(!file.exists(file.path(dir_git, 'clip_depth', 'bathy_cells.csv'))) {
 
-  poly_bath <- readOGR(dsn = file.path(path.expand(dir_git), 'coralmaps/ne_10m_bathymetry_K_200'), 
+  poly_bath <- readOGR(dsn = file.path(path.expand(dir_git), 'clip_depth/ne_10m_bathymetry_K_200'), 
                        layer = 'ne_10m_bathymetry_K_200', 
@@
 
-  writeRaster(bath_raster, file.path(dir_git, 'coralmaps', 'bathy_cells.tif'))
+  writeRaster(bath_raster, file.path(dir_git, 'clip_depth', 'bathy_cells.tif'))
 
-  write_csv(bath_df, file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))
-  # bath_df <- read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd')
+  write_csv(bath_df, file.path(dir_git, 'clip_depth', 'bathy_cells.csv'))
+  # bath_df <- read_csv(file.path(dir_git, 'clip_depth', 'bathy_cells.csv'), col_types = 'dd')
   ### probably want to filter to just 100% 200m+
@@
 
-if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))) {
+if(!file.exists(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))) {
 
@@
 
-  write_csv(spp_narr, file.path(dir_git, 'coralmaps', 'coral_spp_narratives.csv'))
+  write_csv(spp_narr, file.path(dir_git, 'clip_depth', 'coral_spp_narratives.csv'))
 
@@
 
-  write_csv(spp_depth, file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))
+  write_csv(spp_depth, file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))
 
 } else 
-  spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))
+  spp_depth <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))
 
@@
 
-bath_df <-  read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd') %>%
+bath_df <-  read_csv(file.path(dir_git, 'clip_depth', 'bathy_cells.csv'), col_types = 'dd') %>%
   filter(weight >= 1) %>% ### cell must be 100% within the deep polygon to count as deep
@@
 
-spp_depth <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))
+spp_depth <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))
 
@@
 
-write_csv(coral_cells_iucn, file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))
+write_csv(coral_cells_iucn, file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv'))
 length(unique(coral_cells_iucn$iucn_sid))
@@
 ### group_by iucn_sid, summarize total area
-coral_cells_iucn <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
+coral_cells_iucn <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv')) %>%
   mutate(deep = ifelse(deep == 1, 'deep', 'shallow'))
@@
 
-write_csv(iucn_coral_areas, file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv'))
+write_csv(iucn_coral_areas, file.path(dir_git, 'clip_depth', 'iucn_coral_areas.csv'))
 
@@
 ``` {r calc_overlaps}
-if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))) {
+if(!file.exists(file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))) {
 
   ### get species list
-  coral_cells_iucn_shallow <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
+  coral_cells_iucn_shallow <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv')) %>%
     filter(deep == 0)
@@
   ### copy aquamaps cells to local file, keeping only corals in the paired maps
-  if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))) {
+  if(!file.exists(file.path(dir_git, 'clip_depth', 'coral_cells_am.csv'))) {
     coral_cells_am <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv', 'hcaf_sp_native_trunc.csv')) %>%
@@
       filter(am_sid %in% spp_coralmaps$am_sid) 
-    write_csv(coral_cells_am, file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))
+    write_csv(coral_cells_am, file.path(dir_git, 'clip_depth', 'coral_cells_am.csv'))
   }
 
-  coral_cells_am <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))
+  coral_cells_am <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_am.csv'))
 
@@
 
-  write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
+  write_csv(spp_coralmaps, file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))
 }
@@
 
-spp_coral_clip <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv')) %>%
+spp_coral_clip <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_areas.csv')) %>%
   select(iucn_sid, shallow_pct, deep_pct, totarea, shallow_area = area) %>%
@@
 
-spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
+spp_coralmaps <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))
 
@@
 
-write_csv(iucn_coral_fixed, file.path(dir_git, 'coralmaps', 'iucn_coral_fixed.csv'))
+write_csv(iucn_coral_fixed, file.path(dir_git, 'clip_depth', 'iucn_coral_fixed.csv'))
 
@@
 
-iucn_coral_fixed <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_fixed.csv'))
+iucn_coral_fixed <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_fixed.csv'))
 
@@
 
-write_csv(iucn_coral_quads, file.path(dir_git, 'coralmaps', 'iucn_coral_quads.csv'))
+write_csv(iucn_coral_quads, file.path(dir_git, 'clip_depth', 'iucn_coral_quads.csv'))
 
@@
 
-iucn_coral_quad_sum <- read_csv(file.path(dir_git, 'coralmaps', 'iucn_coral_quads.csv')) %>%
+iucn_coral_quad_sum <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_quads.csv')) %>%
   select(method, quad, quad_name, pct_quad) %>%
@@
 
-write_csv(iucn_coral_quad_sum, file.path(dir_git, 'coralmaps', 'iucn_coral_quad_sum.csv'))
+write_csv(iucn_coral_quad_sum, file.path(dir_git, 'clip_depth', 'iucn_coral_quad_sum.csv'))
 

### 2016-08-23 clip_fao/fao_clip.Rmd
messing around with the AquaMaps FAO clipping issue.  Plots and maps for quadrant 3 and 4 species...

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

- get aquamaps cells for all quadrant 3 species (and quadrant 4?)
- focus on any particular group? filter as appropriate
    - maybe by `occurcells` field? for low-data spp
- add lat/long from the hcaf file (or at least long)
- for each species:
    - identify east/west limits of range (between -180 and +180)
        - what about species who cross over in pacific? maybe a flag for pacific species?
        - maybe just select Pacific species and avoid Atlantic?
    - rescale 0-1?
        - here we'd be looking for what happens at the edges - is there a sudden jump from no presence to a bunch just at a single longitude?
        - with this, we could stack all the species and see the cumulative effects of the ends
    - plot end points, east/west bounds on a horizontal line/histogram/density plot, and put verticals on the longitudes of the FAO region boundaries
        - not even worrying about N/S? or maybe include just a big swath across the equator or whatever?

``` {r get_am_spp}

### get quadrant species list; filter to q3 and q4
spp_quad_list <- read_csv(file.path(dir_git, 'data', 'spp_list_quads.csv')) %>%
  select(sciname, am_sid, iucn_sid, reviewed, occurcells, quad) %>%
  filter(quad %in% c('q3', 'q4'))

hist(spp_quad_list$occurcells)
nrow(spp_quad_list %>% filter(occurcells > 100))
### 110 out of 1078 (~10%) have more than 100 occurrence cells

### load full AM list and filter to those in q3 and q4
am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
                                   'v2016/int/am_cells_spp_prob0.csv')) %>%
  filter(am_sid %in% spp_quad_list$am_sid) %>% 
  distinct()

### join with hcaf to get longitudes (and latitudes) for each cell
hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015',
                           'csv/hcaf_truncated.csv')) %>%
  select(-csquarecode)

am_spp_latlong <- am_spp_cells %>%
  left_join(hcaf, by = 'loiczid')

am_spp_latlong <- am_spp_latlong %>%
  mutate(clongadj  = ifelse(centerlong < 0, centerlong + 360, centerlong),
         wlimitadj = ifelse(wlimit < 0, wlimit + 360, wlimit),
         elimitadj = ifelse(elimit < 0, elimit + 360, elimit))

### filter out species from Atlantic... rough pass:
### - filter out everything west of 30 E and east of 70 W

am_spp_latlong <- am_spp_latlong %>%
  filter(clongadj >= 30 & clongadj <= 290)

```

Major fishing area boundaries
- 175W north from 25S
- 120W from 5N to 60 S
- 150E south from 25S (south of Australia)
- 80E from 0N to 55S
- 30 E from Africa on down

``` {r northmost_band}
### 5N north     - look for spike at 175 W (185E)
am_spp_north <- am_spp_latlong %>%
  filter(centerlat >= 5) %>%
  filter(clongadj <= 260) ### west of North America

am_spp_north_lims <- am_spp_north %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(30, 260), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_north_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(77, 185), color = 'red')

```

``` {r central_band}
### 25 S to 5 N  - look for spikes at 120W (240E), 175W (185E), 80E (and 77E)
am_spp_central <- am_spp_latlong %>%
  filter(centerlat <= 5 & centerlat >= -25)

am_spp_central_lims <- am_spp_central %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(30, 260), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_central_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(77, 185, 240), color = 'red')

```

``` {r southmost_band}
### 60 S to 25 S - look for spikes at 120W (240E), 150E, 80E
am_spp_south <- am_spp_latlong %>%
  filter(centerlat >= -60 & centerlat <= -25)

am_spp_south_lims <- am_spp_south %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(30, 260), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_south_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(80, 150, 240), color = 'red')

```

``` {r all_bands}

am_spp_all_lims <- am_spp_latlong %>%
  filter(clongadj <= 260) %>% ### west of North America
  filter(clongadj >= 50) %>% ### east of West Africa
  filter(centerlat >= -25) %>% ### north of Australia
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(50, 260), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_all_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(80, 150, 240), color = 'red') +
  geom_vline(xintercept = c(77, 185), color = 'red')

```

``` {r all_bands_low_data}

datapoor <- spp_quad_list %>%
  filter(is.na(reviewed)) %>%
  filter(occurcells < 50)

am_spp_dp_lims <- am_spp_latlong %>%
  filter(am_sid %in% datapoor$am_sid) %>%
  filter(clongadj <= 260) %>% ### west of North America
  filter(clongadj >= 50) %>% ### east of West Africa
  filter(centerlat >= -25) %>% ### north of Australia
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(50, 260), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_dp_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(80, 240), color = 'red') +
  geom_vline(xintercept = c(77, 185), color = 'red')

```

``` {r what_species_are_these}

### this first batch is Coral Triangle-area species
clipped_spp1 <- am_spp_all_lims %>%
  # filter(long %in% c(77, 185)) %>%
  filter((long == 77  & lim == 'w_lim') | 
         (long == 185 & lim == 'e_lim')) %>%
  select(am_sid) %>%
  distinct %>%
  left_join(spp_quad_list, by = 'am_sid')

### This next batch is open Pacific or Indian Ocean species
clipped_spp2 <- am_spp_all_lims %>%
  # filter(long %in% c(77, 185)) %>%
  filter((long == 77  & lim == 'e_lim') | 
         (long == 185 & lim == 'w_lim')) %>%
  select(am_sid) %>%
  distinct %>%
  left_join(spp_quad_list, by = 'am_sid')

write_csv(clipped_spp1, file.path(dir_git, 'clip_fao', 'clipped_se_asia.csv'))
write_csv(clipped_spp2, file.path(dir_git, 'clip_fao', 'clipped_india_pacific.csv'))

```

``` {r plot_spp, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

clipped_spp1 <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_se_asia.csv'))
clipped_spp2 <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_india_pacific.csv'))

loiczid_raster <- raster('data/rasters/loiczid_raster.tif')

spp_examples <- sample_n(clipped_spp2, 10)

for (i in 1:nrow(spp_examples)) {

  spp <- spp_examples[i, ]
  map_cells <- am_spp_cells %>%
    filter(am_sid == spp$am_sid) %>%
    mutate(am_pres = 1)

  r_am_spp  <-  subs(loiczid_raster, 
                     map_cells[ , c('loiczid', 'am_pres')], 
                     by = 'loiczid', 
                     which = 'am_pres', 
                     subsWithNA = TRUE)

  spp_pts <- as.data.frame(rasterToPoints(r_am_spp))

  spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
    ggtheme_map + 
    geom_raster(aes(fill = am_pres), alpha = .8, show.legend = FALSE) +
    borders('world', color='gray30', fill='gray40', size = .1) +  # create a layer of borders
    scale_x_continuous(breaks = seq(-180, 180, by = 30)) +
    scale_y_continuous(breaks = seq( -90,  90, by = 30))

  plot(spp_plot)
}

```

### 2016-08-25 clip_fao/fao_clip_pac.Rmd
updated FAO clipping code for mapping and whatnot

--- 2016-08-23 clip_fao/fao_clip.Rmd
@@
 library(ggplot2)
+library(tmap)
+data(World)
+
 library(readr)
@@
 am_spp_latlong <- am_spp_latlong %>%
-  filter(clongadj >= 30 & clongadj <= 290)
+  filter(clongadj >= 60 & clongadj <= 240)
 
@@
 am_spp_north <- am_spp_latlong %>%
-  filter(centerlat >= 5) %>%
-  filter(clongadj <= 260) ### west of North America
+  filter(centerlat >= 5)
 
@@
   select(-lim_n) %>%
-  mutate(long = ifelse(long %in% c(30, 260), NA, long)) %>%
+  mutate(long = ifelse(long <= 70 | long >= 200, NA, long)) %>%
   filter(!is.na(long))
@@
   geom_point() +
-  geom_vline(xintercept = c(77, 185, 240), color = 'red')
-
-```
-
-``` {r southmost_band}
-### 60 S to 25 S - look for spikes at 120W (240E), 150E, 80E
-am_spp_south <- am_spp_latlong %>%
-  filter(centerlat >= -60 & centerlat <= -25)
-
-am_spp_south_lims <- am_spp_south %>%
-  group_by(am_sid) %>%
-  summarize(w_lim = min(wlimitadj),
-            e_lim = max(elimitadj),
-            n_w_lim = sum(wlimitadj == w_lim),
-            n_e_lim = sum(elimitadj == e_lim),
-            n_tot = n()) %>%
-  gather(lim, long, w_lim:e_lim) %>%
-  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
-  filter(str_detect(lim_n, lim)) %>%
-  select(-lim_n) %>%
-  mutate(long = ifelse(long %in% c(30, 260), NA, long)) %>%
-  filter(!is.na(long))
-    ### delete artificially imposed limits
-
-tmp <- am_spp_south_lims %>%
-  group_by(long) %>%
-  summarize(n_cells = sum(ncells),
-            n_spp   = n())
-
-ggplot(tmp, aes(x = long, y = n_spp)) +
-  geom_point() +
-  geom_vline(xintercept = c(80, 150, 240), color = 'red')
-
-```
-
-``` {r all_bands}
-
-am_spp_all_lims <- am_spp_latlong %>%
-  filter(clongadj <= 260) %>% ### west of North America
-  filter(clongadj >= 50) %>% ### east of West Africa
-  filter(centerlat >= -25) %>% ### north of Australia
-  group_by(am_sid) %>%
-  summarize(w_lim = min(wlimitadj),
-            e_lim = max(elimitadj),
-            n_w_lim = sum(wlimitadj == w_lim),
-            n_e_lim = sum(elimitadj == e_lim),
-            n_tot = n()) %>%
-  gather(lim, long, w_lim:e_lim) %>%
-  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
-  filter(str_detect(lim_n, lim)) %>%
-  select(-lim_n) %>%
-  mutate(long = ifelse(long %in% c(50, 260), NA, long)) %>%
-  filter(!is.na(long))
-    ### delete artificially imposed limits
-
-tmp <- am_spp_all_lims %>%
-  group_by(long) %>%
-  summarize(n_cells = sum(ncells),
-            n_spp   = n())
-
-ggplot(tmp, aes(x = long, y = n_spp)) +
-  geom_point() +
-  geom_vline(xintercept = c(80, 150, 240), color = 'red') +
-  geom_vline(xintercept = c(77, 185), color = 'red')
-
-```
-
-``` {r all_bands_low_data}
-
-datapoor <- spp_quad_list %>%
-  filter(is.na(reviewed)) %>%
-  filter(occurcells < 50)
-
-am_spp_dp_lims <- am_spp_latlong %>%
-  filter(am_sid %in% datapoor$am_sid) %>%
-  filter(clongadj <= 260) %>% ### west of North America
-  filter(clongadj >= 50) %>% ### east of West Africa
-  filter(centerlat >= -25) %>% ### north of Australia
-  group_by(am_sid) %>%
-  summarize(w_lim = min(wlimitadj),
-            e_lim = max(elimitadj),
-            n_w_lim = sum(wlimitadj == w_lim),
-            n_e_lim = sum(elimitadj == e_lim),
-            n_tot = n()) %>%
-  gather(lim, long, w_lim:e_lim) %>%
-  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
-  filter(str_detect(lim_n, lim)) %>%
-  select(-lim_n) %>%
-  mutate(long = ifelse(long %in% c(50, 260), NA, long)) %>%
-  filter(!is.na(long))
-    ### delete artificially imposed limits
-
-tmp <- am_spp_dp_lims %>%
-  group_by(long) %>%
-  summarize(n_cells = sum(ncells),
-            n_spp   = n())
-
-ggplot(tmp, aes(x = long, y = n_spp)) +
-  geom_point() +
-  geom_vline(xintercept = c(80, 240), color = 'red') +
   geom_vline(xintercept = c(77, 185), color = 'red')
@@
 
-### this first batch is Coral Triangle-area species
-clipped_spp1 <- am_spp_all_lims %>%
+clipped_spp <- am_spp_central_lims %>%
+  # bind_rows(am_spp_north_lims) %>%
   # filter(long %in% c(77, 185)) %>%
-  filter((long == 77  & lim == 'w_lim') | 
-         (long == 185 & lim == 'e_lim')) %>%
-  select(am_sid) %>%
-  distinct %>%
+  filter(long == 185) %>%
+  select(am_sid, lim) %>%
+  distinct() %>%
   left_join(spp_quad_list, by = 'am_sid')
 
-### This next batch is open Pacific or Indian Ocean species
-clipped_spp2 <- am_spp_all_lims %>%
-  # filter(long %in% c(77, 185)) %>%
-  filter((long == 77  & lim == 'e_lim') | 
-         (long == 185 & lim == 'w_lim')) %>%
-  select(am_sid) %>%
-  distinct %>%
-  left_join(spp_quad_list, by = 'am_sid')
-
-write_csv(clipped_spp1, file.path(dir_git, 'clip_fao', 'clipped_se_asia.csv'))
-write_csv(clipped_spp2, file.path(dir_git, 'clip_fao', 'clipped_india_pacific.csv'))
+write_csv(clipped_spp, file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))
 
@@
 
-``` {r plot_spp, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-
-clipped_spp1 <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_se_asia.csv'))
-clipped_spp2 <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_india_pacific.csv'))
+``` {r plot_setup, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
 loiczid_raster <- raster('data/rasters/loiczid_raster.tif')
+fao_polys <- rgdal::readOGR(dsn = file.path(path.expand(dir_git), 'clip_fao/fao_rgns'),
+                            layer = 'fao_rgns2')
 
-spp_examples <- sample_n(clipped_spp2, 10)
+assemble_map <- function(map_rast, spp) {
+  message('...mapping ', spp)
+  map_obj <- tm_shape(map_rast) +
+    tm_raster(palette = 'Reds',
+              colorNA = NULL,
+              title = spp,
+              alpha = 1) +
+    tm_shape(World) +
+      tm_polygons() + 
+    tm_layout(basemaps = "Esri.WorldTopoMap", 
+              # title.position = 'TOP', 
+              legend.outside = TRUE, attr.outside = TRUE)
 
-for (i in 1:nrow(spp_examples)) {
+  return(map_obj)
+}
+```
 
-  spp <- spp_examples[i, ]
+``` {r plot_spp_e, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+spp_examples_e <- clipped_spp %>%
+  filter(lim == 'w_lim')
+
+rast_list <- vector('list', length = nrow(spp_examples_e))
+
+for (i in 1:nrow(spp_examples_e)) {  # i <- 1
+
+  spp <- spp_examples_e[i, ]
   map_cells <- am_spp_cells %>%
@@
 
-  spp_pts <- as.data.frame(rasterToPoints(r_am_spp))
+  r1 <- r_am_spp %>%
+    crop(extent(c(-180, 0, -50, 50)))
+  extent(r1) <- c(180, 360, -50, 50)
 
-  spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
-    ggtheme_map + 
-    geom_raster(aes(fill = am_pres), alpha = .8, show.legend = FALSE) +
-    borders('world', color='gray30', fill='gray40', size = .1) +  # create a layer of borders
-    scale_x_continuous(breaks = seq(-180, 180, by = 30)) +
-    scale_y_continuous(breaks = seq( -90,  90, by = 30))
+  r2 <- r_am_spp %>%
+    crop(extent(c(0, 180, -50, 50)))
 
-  plot(spp_plot)
+  r3 <- merge(r2, r1)
+
+  new_ext <- extent(c(50, 240, -50, 50))
+  r_am_crop <- crop(r3, new_ext)
+
+  rast_list[i] <- r_am_crop
+  names(rast_list)[i] <- spp
+
+  spp_plot <- assemble_map(r_am_crop, spp$sciname)
 }
 
+mapstack_e <- stack(rast_list) %>%
+  calc(sum, na.rm = TRUE)
+plot(mapstack_e)
+values(mapstack_e)[values(mapstack_e) == 0] <- NA
+mapstack_e <- crop(mapstack_e, extent(c(85, 220, -50, 50)))
+
+stack_plot_e <- assemble_map(mapstack_e, 'n_spp') +
+  tm_shape(fao_polys) +
+  tm_borders()
+print(stack_plot_e)
+
 ```
+
+``` {r plot_spp_w, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+spp_examples_w <- clipped_spp %>%
+  filter(lim == 'e_lim') %>%
+  sample_n(30)
+
+rast_list <- vector('list', length = nrow(spp_examples_w))
+
+for (i in 1:nrow(spp_examples_w)) {  # i <- 1
+
+  spp <- spp_examples_w[i, ]
+  map_cells <- am_spp_cells %>%
+    filter(am_sid == spp$am_sid) %>%
+    mutate(am_pres = 1)
+
+  r_am_spp  <-  subs(loiczid_raster, 
+                     map_cells[ , c('loiczid', 'am_pres')], 
+                     by = 'loiczid', 
+                     which = 'am_pres', 
+                     subsWithNA = TRUE)
+
+  r1 <- r_am_spp %>%
+    crop(extent(c(-180, 0, -50, 50)))
+  extent(r1) <- c(180, 360, -50, 50)
+
+  r2 <- r_am_spp %>%
+    crop(extent(c(0, 180, -50, 50)))
+
+  r3 <- merge(r2, r1)
+
+  new_ext <- extent(c(70, 220, -50, 50))
+  r_am_crop <- crop(r3, new_ext)
+
+  rast_list[i] <- r_am_crop
+  names(rast_list)[i] <- spp
+
+  spp_plot <- assemble_map(r_am_crop, spp$sciname)
+}
+
+mapstack_w <- stack(rast_list) %>%
+  calc(sum, na.rm = TRUE)
+plot(mapstack_w)
+values(mapstack_w)[values(mapstack_w) == 0] <- NA
+mapstack_w <- crop(mapstack_w, extent(c(85, 220, -50, 50)))
+
+stack_plot_w <- assemble_map(mapstack_w, 'n_spp') +
+  tm_shape(fao_polys) +
+  tm_borders()
+
+print(stack_plot_w)
+
+```
+
+``` {r combine_rasts}
+
+stack_plot_both <- 
+  tm_shape(mapstack_w) +
+    tm_raster(palette = 'Reds',
+              colorNA = NULL,
+              title = 'n_spp',
+              alpha = 1) +
+  tm_shape(mapstack_e) +
+    tm_raster(palette = 'Blues',
+              colorNA = NULL,
+              title = 'n_spp') +
+  tm_shape(World) +
+    tm_polygons() + 
+  tm_shape(fao_polys) +
+    tm_borders() +
+  tm_layout(basemaps = "Esri.WorldTopoMap", 
+            # title.position = 'TOP', 
+            legend.outside = TRUE, attr.outside = TRUE)
+
+print(stack_plot_both)
+```

### 2016-08-27 clip_depth/coral_maps.Rmd
pushing aquamaps border issues and manuscript updates

@@
 
+```
+
+``` {r extract_IUCN_cells_for_coral_pairs}
+
+bath_df <-  read_csv(file.path(dir_git, 'clip_depth', 'bathy_cells.csv'), col_types = 'dd') %>%
+  filter(weight >= 1) %>% ### cell must be 100% within the deep polygon to count as deep
+  mutate(deep = 1) %>%
+  select(-weight)
+
+spp_depth <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))
+
+### load iucn spp_cells
+coral_files <- list.files(file.path(dir_anx, 'v2016', 'iucn_intersections'), 
+                          pattern = 'CORAL*.*',
+                          full.names = TRUE)
+coral_cells_iucn <- lapply(coral_files, read_csv) %>%
+  bind_rows() %>%
+  select(iucn_sid, loiczid = LOICZID) %>%
+  # filter(iucn_sid %in% spp_depth$iucn_sid[spp_depth$paired_map]) %>%
+  left_join(bath_df, by = 'loiczid') %>%
+  mutate(deep = ifelse(is.na(deep), 0, deep))
+
+write_csv(coral_cells_iucn, file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv'))
+length(unique(coral_cells_iucn$iucn_sid))
+### NOTE: only 473 coral species with IUCN maps
 ```
@@
 
-``` {r extract_IUCN_cells_for_coral_pairs}
+``` {r plot_coral_composite_map}
 
-bath_df <-  read_csv(file.path(dir_git, 'clip_depth', 'bathy_cells.csv'), col_types = 'dd') %>%
-  filter(weight >= 1) %>% ### cell must be 100% within the deep polygon to count as deep
-  mutate(deep = 1) %>%
-  select(-weight)
+library(tmap)
+data(World)
 
-spp_depth <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))
+coral_comp <- coral_cells_iucn %>%
+  select(-iucn_sid) %>%
+  mutate(deep = deep * 200) %>%
+  distinct()
 
-### load iucn spp_cells
-coral_files <- list.files(file.path(dir_anx, 'v2016', 'iucn_intersections'), 
-                          pattern = 'CORAL*.*',
-                          full.names = TRUE)
-coral_cells_iucn <- lapply(coral_files, read_csv) %>%
-  bind_rows() %>%
-  select(iucn_sid, loiczid = LOICZID) %>%
-  # filter(iucn_sid %in% spp_depth$iucn_sid[spp_depth$paired_map]) %>%
-  left_join(bath_df, by = 'loiczid') %>%
-  mutate(deep = ifelse(is.na(deep), 0, deep))
+rast_coral_comp <- subs(x = loiczid_raster, y = coral_comp %>% select(loiczid, deep), 
+                     by = 'loiczid', which = 'deep', 
+                     subsWithNA = TRUE) %>%
+  crop(extent(c(-180, 180, -63, 85)))
 
-write_csv(coral_cells_iucn, file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv'))
-length(unique(coral_cells_iucn$iucn_sid))
-### NOTE: only 473 coral species with IUCN maps
+map_coral_comp <- 
+  tm_shape(rast_coral_comp) +
+    tm_raster(breaks = c(0, 200, Inf),
+              palette = c('lightblue', 'royalblue4'),
+              colorNA = NULL,
+              title = 'depth (meters)',
+              alpha = 1) +
+  tm_shape(World) +
+    tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
+  tm_layout(basemaps = "Esri.WorldTopoMap", 
+            legend.text.size = .35,
+            legend.title.size = .45,
+            # title.position = 'TOP', 
+            legend.outside = FALSE, 
+            legend.position = c('left', 'bottom'),
+            legend.bg.color = 'white',
+            legend.bg.alpha = .5,
+            attr.outside = TRUE)
+
+print(map_coral_comp)
+
+save_tmap(map_coral_comp,
+          filename = file.path(dir_fig, 'fig3a_coral_map.png'), 
+          width = 9, height = 4.2, units = 'cm', dpi = 600)
+
 ```

### 2016-08-27 clip_fao/fao_clip_pac.Rmd
pushing aquamaps border issues and manuscript updates

@@
 
-hist(spp_quad_list$occurcells)
 nrow(spp_quad_list %>% filter(occurcells > 100))
@@
 am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
-                                   'v2016/int/am_cells_spp_prob0.csv')) %>%
-  filter(am_sid %in% spp_quad_list$am_sid) %>% 
-  distinct()
+                                   'v2016/int/am_cells_spp_prob0.csv')) # %>%
+  # filter(am_sid %in% spp_quad_list$am_sid) %>% 
+  # distinct()
 
+message('loading hcaf and adjusting to se asia')
 ### join with hcaf to get longitudes (and latitudes) for each cell
@@
                            'csv/hcaf_truncated.csv')) %>%
-  select(-csquarecode)
-
-am_spp_latlong <- am_spp_cells %>%
-  left_join(hcaf, by = 'loiczid')
-
-am_spp_latlong <- am_spp_latlong %>%
+  select(-csquarecode, -cellarea, -oceanarea) %>%
   mutate(clongadj  = ifelse(centerlong < 0, centerlong + 360, centerlong),
          wlimitadj = ifelse(wlimit < 0, wlimit + 360, wlimit),
-         elimitadj = ifelse(elimit < 0, elimit + 360, elimit))
+         elimitadj = ifelse(elimit < 0, elimit + 360, elimit)) %>%
+  filter(clongadj >= 60 & clongadj <= 240)
 
@@
 ### - filter out everything west of 30 E and east of 70 W
+message('binding hcaf to am_spp_latlong')
 
-am_spp_latlong <- am_spp_latlong %>%
-  filter(clongadj >= 60 & clongadj <= 240)
+am_spp_latlong <- hcaf %>%
+  left_join(am_spp_cells, by = 'loiczid')
 
@@
 
-``` {r northmost_band}
+``` {r northmost_band, eval = FALSE}
 ### 5N north     - look for spike at 175 W (185E)
@@
   select(-lim_n) %>%
-  mutate(long = ifelse(long <= 70 | long >= 200, NA, long)) %>%
-  filter(!is.na(long))
+  filter(!(long <= 70 | long >= 200))
     ### delete artificially imposed limits
@@
 
+message('summarizing...')
 am_spp_central_lims <- am_spp_central %>%
@@
             n_e_lim = sum(elimitadj == e_lim),
-            n_tot = n()) %>%
+            n_tot = n())
+
+message('gathering...')
+am_spp_central_lims <- am_spp_central_lims %>%
   gather(lim, long, w_lim:e_lim) %>%
@@
   select(-lim_n) %>%
-  mutate(long = ifelse(long <= 70 | long >= 200, NA, long)) %>%
-  filter(!is.na(long))
+  filter(!(long <= 70 | long >= 200))
     ### delete artificially imposed limits
+
+write_csv(am_spp_central_lims, file.path(dir_git, 'clip_fao/am_spp_central_lims.csv'))
 
@@
 ``` {r plot_spp_e, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))
+
+spp_examples_e <- clipped_spp %>%
+  filter(lim == 'w_lim')
+
+message('grouping affected species and summing to loiczid')
+### group affected spp by loiczid and sum
+spp_cells_clipped <- am_spp_cells %>%
+  filter(am_sid %in% spp_examples_e$am_sid) %>%
+  group_by(loiczid) %>%
+  summarize(n_spp = n())
+
+### create raster, then shift raster
+
+rast_spp  <-  subs(loiczid_raster, 
+                   spp_cells_clipped[ , c('loiczid', 'n_spp')], 
+                   by = 'loiczid', 
+                   which = 'n_spp', 
+                   subsWithNA = TRUE)
+
+rast_spp_shift <- rast_spp %>%
+  raster::shift(360)
+
+rast_spp_e <- merge(rast_spp, rast_spp_shift)
+
+values(rast_spp_e)[values(rast_spp_e) == 0] <- NA
+values(rast_spp_e)[is.na(values(rast_spp_e))] <- 0
+rast_spp_e <- crop(rast_spp_e, extent(c(85, 220, -60, 60)))
+
+rast_spp_map_e <- 
+  tm_shape(rast_spp_e) +
+    tm_raster(palette = 'Reds',
+              breaks = c(1, 15, 100, 200, 300, Inf),
+              colorNA = 'lightblue1',
+              textNA = 'Zero',
+              title = 'Species count',
+              alpha = 1) +
+  tm_shape(World) +
+    tm_polygons(border.col = 'grey50', col = 'grey55') + 
+  tm_shape(fao_polys) +
+    tm_borders(col = 'grey40') +
+  tm_layout(basemaps = "Esri.WorldTopoMap", 
+            # title.position = 'TOP', 
+            legend.outside = TRUE, attr.outside = TRUE)
+
+print(rast_spp_map_e)
+
+```
+
+``` {r plot_spp_e, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))
 
@@
   tm_shape(mapstack_e) +
-    tm_raster(palette = 'Blues',
+    tm_raster(palette = 'Reds',
               colorNA = NULL,

### 2016-08-27 ms_1_intro.Rmd
pushing aquamaps border issues and manuscript updates

@@
 One major outcome of this body of science is the various compiled databases of species distribution maps.
-Here we focus on global-scale repositories that predict marine species ranges throughout the world's oceans – AquaMaps [@kaschner_aquamaps:_2015] and International Union for Conservation of Nature (IUCN) [@iucn_iucn_2015].
-These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
-
-___sentence about SAUP - how can we rephrase the "two global-scale repositories" sentence to exclude it?
-"broad taxonomic coverage"?
-something about not focusing just on economically important species?
-(is that a valid assumption about SAUP) - consider other datasets as well: NatureServe, BOTW, etc___
-
-___or: "Several global-scale repositories predict ... In this paper we focus on two datasets that cover a broad taxonomic range: ..."___
-
-Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each offers an understanding of the "truth" - one relying primarily on expert opinion and the other on model predictions.
-Each method presents its own challenges and advantages, and the small overlap in mapped species means each dataset provides value in taxonomic and geographic coverage.
-Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
-
+The two most comprehensive, widely-used global-scale repositories that predict marine species ranges throughout the world's oceans are AquaMaps [@kaschner_aquamaps:_2015] and range data from the International Union for Conservation of Nature (IUCN) [@iucn_iucn_2015].
+Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each offers an understanding of the "truth", with IUCN ranges relying primarily on expert opinion and AquaMaps on model predictions.
 Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: geographic range data such as IUCN range maps frequently introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
@@
 
+These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
+In most of these cases, the implications of choosing one versus the other of these datasets is not evaluated or discussed, yet strong conservation and management conclusions are drawn from the results of the studies.
+For example, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
+
 To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of the `r format(n_spp_all, big.mark = ',')` marine species mapped by one or both datasets.
 Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).
+Yet the small overlap in mapped species between the two datasets means each provides unique value in taxonomic and geographic coverage.
 
+Although relatively small in number, these overlap species present a unique opportunity to evaluate the two datasets overall.
 As understanding of a given species improves, we expect that range maps, regardless of method, would become more accurate in predicting species presence and absence.
@@
 For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align in both distribution and overall area; from the results we determine methodological issues that can introduce commission errors, and suggest methods to improve confidence in species range predictions.
+In particular, we show how solving a simple problem in coral ranges in the IUCN range data dramatically ___improves agreement with Aquamaps data.___ - _does this make it sound like agreement is the goal?
+make it clear that agreement implies better confidence_
 
@@
 
-Each of these datasets provides unique value in taxonomic and geographic coverage; while we have identified some potential issues in each, we cannot simply recommmend one data set over the other.
+While we have identified some potential issues in each dataset, we cannot simply recommend one data set over the other.
 Instead, we recommend simple methods to reduce incidence of commission errors in each dataset, improving confidence in species range predictions to better inform conservation management and policy decisions.

### 2016-08-27 ms_3_res_disc_concl.Rmd
pushing aquamaps border issues and manuscript updates

@@
 ## Taxonomic and geographic coverage
+
+- figure 1 and caption
 
@@
 
+## Distribution and range area alignment
+
 [discussion] To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area ratio (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig.
@@
 Understanding these sources of error can provide guidance to improve predictions of species range even when only one source of spatial data is available.
+
+- quad plot and bar chart and captions
 
@@
 [results/discussion] ___area ratio suggests similar commission/omission errors, but dist suggests otherwise___ Species found in the lower right quadrant (n = `r n_q3`) seem to represent cases of "two wrongs make a right."  For these species, IUCN and AquaMaps both predict ranges extending far beyond the overlapping region, but the methodological differences result in very different extrapolations.
-___OMIT:___ _For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, likely introducing additional commission errors._  Consequently, area ratios are close to 100%, though the poor distribution alignment indicates that one or both datasets are introducing significant errors.
+Consequently, area ratios are close to 100%, though the poor distribution alignment indicates that one or both datasets are introducing significant errors.
 As above, further analysis is required to disentangle the causes of error.
@@
 
-## Case study: Coral depth analysis
+## Coral depth exploration
 
@@
 3A shows the 463 coral species mapped in the IUCN dataset, with their ranges broken into proportional area deeper and shallower than 200 m.
-None of these species is indicated to occur deeper than 200 m, and 94% are confined to waters shallower than 50 m; seven of the mapped species had no reported depth information.
-Clipping coral ranges to shallower than 200 m eliminated an average of 47.6% of the total predicted area while still allowing for a conservative estimate of suitable habitat.
+___show a map of all coral species aggregated, with deeper than 200 m cells shown___  None of these species is indicated to occur deeper than 200 m, and 94% are confined to waters shallower than 50 m; seven of the mapped species had no reported depth information.
+Clipping coral ranges to shallower than 200 m eliminated an average of 47.6% of the total predicted area while still allowing for a generous estimate of suitable habitat.
 

### 2016-08-28 ms_2_methods.Rmd
some updates to methods and res/disc re: corals and aquamaps clipping

@@
 To examine errors of commission related to depth, we selected corals as a case study due to their importance in supporting biodiversity as well as their dependence on photosynthesis.
+Extracting data from the IUCN API [@REF?] we identified the depth limitations of each of the coral species mapped in the IUCN dataset.
 We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] ___(Fig.
@@
 The resulting maps were again compared to the AquaMaps rasters to examine distribution alignment and area ratio.
+
+AquaMaps uses Food and Agriculture Organization of the United Nations (FAO) Major Fishing Area [@REF] boundaries as a rough method of constraining species ranges to appropriate georegions.
+To explore the impact of these artificial constraints on species range predictions, we chose to focus on boundaries defined by longitude, as they seem less likely to relate to ecological conditions than boundaries defined by latitude.
+As an example, we identified all AquaMaps species whose eastern or western range limit between between 5° S and 25° N latitudes coincided with the large vertical boundary between FAO regions 71 and 77 at 175° W longitude.
 

### 2016-08-28 ms_3_res_disc_concl.Rmd
some updates to methods and res/disc re: corals and aquamaps clipping

@@
 3A shows the 463 coral species mapped in the IUCN dataset, with their ranges broken into proportional area deeper and shallower than 200 m.
-___show a map of all coral species aggregated, with deeper than 200 m cells shown___  None of these species is indicated to occur deeper than 200 m, and 94% are confined to waters shallower than 50 m; seven of the mapped species had no reported depth information.
+According to IUCN descriptions, none of these species is indicated to occur deeper than 200 m, and 94% are confined to waters shallower than 50 m; seven of the mapped species had no reported depth information.
 Clipping coral ranges to shallower than 200 m eliminated an average of 47.6% of the total predicted area while still allowing for a generous estimate of suitable habitat.
@@
 2B, we see in Fig.
-3B and 3C a massive shift in 354 paired map species due to constraining coral depths to 200 m.
+3B a massive shift in 354 paired-map coral species due to constraining coral depths to 200 m.
 Membership in the "well-aligned" quadrant jumped from 22.4% to 76.2%, with a corresponding decrease in all other quadrants.
+See SXXX to examine the shifts of individual species among the quadrants.
 
-[discussion] We cannot know for certain the true distribution of each of these corals; much of the apparent over-prediction of coral ranges is likely due to experts taking a conservative and precautionary approach.
+[discussion] We cannot know for certain the true distribution of each of these corals; much of the apparent over-expansion of coral ranges is likely due to experts taking a conservative and precautionary approach.
 But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported ranges.
@@
 
-***What is still missing here is something to show that AquaMaps ranges, clipped to FAO Major Fishing Areas, are probably introducing errors of commission as well*** - _the fish don't see that FAO boundary and suddenly decide not to cross; most obvious in the boundaries defined by longitudes - sudden vertical edges in range.
-Can we show this with a couple of example maps?
-That wasn't received well for the corals, but I can't think of a clear and unambiguous way to analyze this problem._
+## Georegion constraint exploration
+
+- AM clipped map
+
+[results]
 

### 2016-08-28 clip_fao/fao_clip_pac.Rmd
updated aquamaps edge map

@@
 values(rast_spp_e)[values(rast_spp_e) == 0] <- NA
-values(rast_spp_e)[is.na(values(rast_spp_e))] <- 0
-rast_spp_e <- crop(rast_spp_e, extent(c(85, 220, -60, 60)))
+# values(rast_spp_e)[is.na(values(rast_spp_e))] <- 0
+rast_spp_e <- crop(rast_spp_e, extent(c(85, 220, -59, 50)))
 
@@
               textNA = 'Zero',
-              title = 'Species count',
+              title = 'Count',
               alpha = 1) +
   tm_shape(World) +
-    tm_polygons(border.col = 'grey50', col = 'grey55') + 
+    tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
   tm_shape(fao_polys) +
-    tm_borders(col = 'grey40') +
+    tm_borders(col = 'grey40', lwd = 0.25) +
+    tm_text('F_CODE_1',
+            col = 'orangered4',
+            fontface = 'bold.italic',
+            fontfamily = 'Helvetica',
+            size = .35) +
   tm_layout(basemaps = "Esri.WorldTopoMap", 
+            legend.text.size = .35,
+            legend.title.size = .45,
             # title.position = 'TOP', 
-            legend.outside = TRUE, attr.outside = TRUE)
+            legend.outside = TRUE,
+            # legend.position = c('right', 'bottom'),
+            # legend.bg.color = 'white',
+            # legend.bg.alpha = .5,
+            attr.outside = TRUE)
 
 print(rast_spp_map_e)
+
+save_tmap(rast_spp_map_e, file.path(dir_fig, 'fig4_am_edge.png'), 
+          width = 9, height = 6, units = 'cm', dpi = 600)
 

### 2016-08-28 clip_fao/fao_clip_pac.Rmd
Omitted OHI analysis from intro, methods, and res/disc/concl; added results/discussion for AM FAO clipping

@@
   # distinct()
+
+am_spp_list <- read_csv(file.path(dir_git, 'data/spp_all.csv')) %>%
+  filter(!is.na(am_sid)) %>%
+  arrange(reviewed) %>%
+  filter(!duplicated(am_sid)) %>%
+  select(am_sid, sciname, reviewed) %>%
+  distinct()
 
@@
   distinct() %>%
-  left_join(spp_quad_list, by = 'am_sid')
+  left_join(am_spp_list, by = 'am_sid')
 
@@
 
-``` {r plot_spp_e, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+``` {r plot_spp_w, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
@@
 
-spp_examples_e <- clipped_spp %>%
-  filter(lim == 'w_lim')
+spp_examples_w <- clipped_spp %>%
+  filter(lim == 'e_lim')
 
-rast_list <- vector('list', length = nrow(spp_examples_e))
+message('grouping affected species and summing to loiczid')
+### group affected spp by loiczid and sum
+spp_cells_clipped <- am_spp_cells %>%
+  filter(am_sid %in% spp_examples_w$am_sid) %>%
+  group_by(loiczid) %>%
+  summarize(n_spp = n())
 
-for (i in 1:nrow(spp_examples_e)) {  # i <- 1
+### create raster, then shift raster
 
-  spp <- spp_examples_e[i, ]
-  map_cells <- am_spp_cells %>%
-    filter(am_sid == spp$am_sid) %>%
-    mutate(am_pres = 1)
-
-  r_am_spp  <-  subs(loiczid_raster, 
-                     map_cells[ , c('loiczid', 'am_pres')], 
+rast_spp  <-  subs(loiczid_raster, 
+                   spp_cells_clipped[ , c('loiczid', 'n_spp')], 
                    by = 'loiczid', 
-                     which = 'am_pres', 
+                   which = 'n_spp', 
                    subsWithNA = TRUE)
 
-  r1 <- r_am_spp %>%
-    crop(extent(c(-180, 0, -50, 50)))
-  extent(r1) <- c(180, 360, -50, 50)
+rast_spp_shift <- rast_spp %>%
+  raster::shift(360)
 
-  r2 <- r_am_spp %>%
-    crop(extent(c(0, 180, -50, 50)))
+rast_spp_w <- merge(rast_spp, rast_spp_shift)
 
-  r3 <- merge(r2, r1)
+values(rast_spp_w)[values(rast_spp_w) == 0] <- NA
+# values(rast_spp_w)[is.na(values(rast_spp_w))] <- 0
+rast_spp_w <- crop(rast_spp_w, extent(c(85, 220, -59, 50)))
 
-  new_ext <- extent(c(50, 240, -50, 50))
-  r_am_crop <- crop(r3, new_ext)
+rast_spp_map_w <- 
+  tm_shape(rast_spp_w) +
+    tm_raster(palette = 'Reds',
+              breaks = c(1, 15, 100, 200, 300, Inf),
+              colorNA = 'lightblue1',
+              textNA = 'Zero',
+              title = 'Count',
+              alpha = 1) +
+  tm_shape(World) +
+    tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
+  tm_shape(fao_polys) +
+    tm_borders(col = 'grey40', lwd = 0.25) +
+    tm_text('F_CODE_1',
+            col = 'orangered4',
+            fontface = 'bold.italic',
+            fontfamily = 'Helvetica',
+            size = .35) +
+  tm_layout(basemaps = "Esri.WorldTopoMap", 
+            legend.text.size = .35,
+            legend.title.size = .45,
+            # title.position = 'TOP', 
+            legend.outside = TRUE,
+            # legend.position = c('right', 'bottom'),
+            # legend.bg.color = 'white',
+            # legend.bg.alpha = .5,
+            attr.outside = TRUE)
 
-  rast_list[i] <- r_am_crop
-  names(rast_list)[i] <- spp
+print(rast_spp_map_w)
 
-  spp_plot <- assemble_map(r_am_crop, spp$sciname)
-}
-
-mapstack_e <- stack(rast_list) %>%
-  calc(sum, na.rm = TRUE)
-plot(mapstack_e)
-values(mapstack_e)[values(mapstack_e) == 0] <- NA
-mapstack_e <- crop(mapstack_e, extent(c(85, 220, -50, 50)))
-
-stack_plot_e <- assemble_map(mapstack_e, 'n_spp') +
-  tm_shape(fao_polys) +
-  tm_borders()
-print(stack_plot_e)
-
-```
-
-``` {r plot_spp_w, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-
-spp_examples_w <- clipped_spp %>%
-  filter(lim == 'e_lim') %>%
-  sample_n(30)
-
-rast_list <- vector('list', length = nrow(spp_examples_w))
-
-for (i in 1:nrow(spp_examples_w)) {  # i <- 1
-
-  spp <- spp_examples_w[i, ]
-  map_cells <- am_spp_cells %>%
-    filter(am_sid == spp$am_sid) %>%
-    mutate(am_pres = 1)
-
-  r_am_spp  <-  subs(loiczid_raster, 
-                     map_cells[ , c('loiczid', 'am_pres')], 
-                     by = 'loiczid', 
-                     which = 'am_pres', 
-                     subsWithNA = TRUE)
-
-  r1 <- r_am_spp %>%
-    crop(extent(c(-180, 0, -50, 50)))
-  extent(r1) <- c(180, 360, -50, 50)
-
-  r2 <- r_am_spp %>%
-    crop(extent(c(0, 180, -50, 50)))
-
-  r3 <- merge(r2, r1)
-
-  new_ext <- extent(c(70, 220, -50, 50))
-  r_am_crop <- crop(r3, new_ext)
-
-  rast_list[i] <- r_am_crop
-  names(rast_list)[i] <- spp
-
-  spp_plot <- assemble_map(r_am_crop, spp$sciname)
-}
-
-mapstack_w <- stack(rast_list) %>%
-  calc(sum, na.rm = TRUE)
-plot(mapstack_w)
-values(mapstack_w)[values(mapstack_w) == 0] <- NA
-mapstack_w <- crop(mapstack_w, extent(c(85, 220, -50, 50)))
-
-stack_plot_w <- assemble_map(mapstack_w, 'n_spp') +
-  tm_shape(fao_polys) +
-  tm_borders()
-
-print(stack_plot_w)
+save_tmap(rast_spp_map_w, file.path(dir_fig, 'fig4_am_edge_w.png'), 
+          width = 9, height = 6, units = 'cm', dpi = 600)
 

### 2016-08-28 ms_3_res_disc_concl.Rmd
updated aquamaps edge map
Omitted OHI analysis from intro, methods, and res/disc/concl; added results/discussion for AM FAO clipping

@@
 
-[discussion] We cannot know for certain the true distribution of each of these corals; much of the apparent over-expansion of coral ranges is likely due to experts taking a conservative and precautionary approach.
+[discussion] We cannot know for certain the true distribution of each of these corals; much of the apparent over-expansion of coral ranges may be due to experts taking a precautionary approach.
 But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported ranges.
@@
 
-## Georegion constraint exploration
+## Georegional constraint exploration
 
@@
 
-[results]
+[results] We identified 4,168 species whose equatorial ranges encounter a range limit at 175° W.
+This longitude created a western limit for 512 species, and eastern limit for 3,656.
+Plotting the entire ranges of the 512 west-limited species results in a very clear exclusion zone that matches perfectly with FAO region 71; while we did not focus on the boundary for region 57, a similar effect is readily apparent.
 
-## Case Studies
+[discussion] FAO Major Fishing Area boundaries provide a readily available method to roughly contrain AquaMaps predictions to appropriate ocean basins, enabling rapid modeling of thousands of species ranges.
+However, these boundaries are defined for statistical purposes based on economic and political considerations rather than ecological considerations, and can result in odd species range predictions where otherwise suitable habitat encounters border between regions.
+While such an odd boundary would likely be obvious when inspecting ranges of individual species, the distinction is likely to be obscured when many species ranges are aggregated as is typical for biodiversity or conservation studies.
+Applying an additional spatial constraint based upon ecological or environmental boundaries, such as Marine Ecoregions of the World [@REF???], may help boost confidence in these predicted ranges; expert review, though time-consuming, is certain to improve confidence in these predicted ranges.
+
+<!-- ## Case Studies
 
@@
 
+-->
+
+<!--
 ### Case Study: The Ocean Health Index
@@
 
-### Case Study: MPA Gap Analysis
+-->
+
+## Case Study: MPA Gap Analysis
 

### 2016-08-29 case_study_ohi/ohi_single_datasets.Rmd

@@
   geom_point(position = position_jitter(w = .3), 
-             aes(color = mean_n_spp), alpha = 1, size = 0.25) +
+             color = 'black', alpha = 1, size = 0.25) +
   geom_violin(fill = NA, 
@@
 
-ggsave(file.path(dir_fig, 'fig3a_violinplot_ohi_maps.png'),
+ggsave(file.path(dir_fig, 'fig4a_violinplot_ohi_maps.png'),
        height = 4.5, width = 8.7, units = 'cm', dpi = 600)
-# ggsave(file.path(dir_fig, 'fig3a_violinplot_ohi_maps.png'), device = 'tiff',
+# ggsave(file.path(dir_fig, 'fig4a_violinplot_ohi_maps.png'), device = 'tiff',
 #        height = 4.5, width = 8.7, units = 'cm', dpi = 600)

### 2016-08-29 clip_fao/fao_clip_atl.Rmd

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(tmap)
data(World)

library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

- get aquamaps cells for all quadrant 3 species (and quadrant 4?)
- focus on any particular group? filter as appropriate
    - maybe by `occurcells` field? for low-data spp
- add lat/long from the hcaf file (or at least long)
- for each species:
    - identify east/west limits of range (between -180 and +180)
        - what about species who cross over in pacific? maybe a flag for pacific species?
        - maybe just select Pacific species and avoid Atlantic?
    - rescale 0-1?
        - here we'd be looking for what happens at the edges - is there a sudden jump from no presence to a bunch just at a single longitude?
        - with this, we could stack all the species and see the cumulative effects of the ends
    - plot end points, east/west bounds on a horizontal line/histogram/density plot, and put verticals on the longitudes of the FAO region boundaries
        - not even worrying about N/S? or maybe include just a big swath across the equator or whatever?

``` {r get_am_spp}

### get quadrant species list; filter to q3 and q4
spp_quad_list <- read_csv(file.path(dir_git, 'data', 'spp_list_quads.csv')) %>%
  select(sciname, am_sid, iucn_sid, reviewed, occurcells, quad) %>%
  filter(quad %in% c('q3', 'q4'))

hist(spp_quad_list$occurcells)
nrow(spp_quad_list %>% filter(occurcells > 100))
### 110 out of 1078 (~10%) have more than 100 occurrence cells

### load full AM list and filter to those in q3 and q4
am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
                                   'v2016/int/am_cells_spp_prob0.csv')) %>%
  filter(am_sid %in% spp_quad_list$am_sid) %>% 
  distinct()

### join with hcaf to get longitudes (and latitudes) for each cell
hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015',
                           'csv/hcaf_truncated.csv')) %>%
  select(-csquarecode)

am_spp_latlong <- am_spp_cells %>%
  left_join(hcaf, by = 'loiczid')

### filter out species from Pacific... rough pass:
### - filter out everything west of 100 W and east of 70 E
### - this leaves southeast Pacific - coast of Chile etc

am_spp_latlong <- am_spp_latlong %>%
  filter(centerlong <= 50 & centerlong >= -65)

```

Major fishing area boundaries
- Band 1: 42 W from 35 N to 59 N (North Atlantic)
- Band 2: 40 W from 5 N to 36 N (temperate Atlantic)
- Band 3: 20 W from equator to 50 S (tropical and southern Atlantic); 30 E from 25 S on down (below Africa)

``` {r band_1}
### 35 north - look for spike at 42 W.  Not big! dammit.
am_spp_band1 <- am_spp_latlong %>%
  filter(centerlat >= 35)

am_spp_band1_lims <- am_spp_band1 %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimit),
            e_lim = max(elimit),
            n_w_lim = sum(wlimit == w_lim),
            n_e_lim = sum(elimit == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(-65, 50), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_band1_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(-42), color = 'red')

```

``` {r band_2}
### 5-36 north - look for spike at 40 W.  NONE! dammit
am_spp_band2 <- am_spp_latlong %>%
  filter(centerlat <= 36 & centerlat >= 5)

am_spp_band2_lims <- am_spp_band2 %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimit),
            e_lim = max(elimit),
            n_w_lim = sum(wlimit == w_lim),
            n_e_lim = sum(elimit == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(-65, 50), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_band2_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(-40), color = 'red')

```

``` {r band_3}
### 50 S to equator - look for spike at 20 W and 30 E
am_spp_band3 <- am_spp_latlong %>%
  filter(centerlat <= 0 & centerlat >= -70)

am_spp_band3_lims <- am_spp_band3 %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimit),
            e_lim = max(elimit),
            n_w_lim = sum(wlimit == w_lim),
            n_e_lim = sum(elimit == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(-65, 50), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_band3_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(-20, 30), color = 'red')

```

``` {r what_species_are_these}

### which species are clipped at bottom of Africa
clipped_spp <- am_spp_band3_lims %>%
  filter(long == 30) %>%
  select(am_sid, lim) %>%
  distinct() %>%
  left_join(spp_quad_list, by = 'am_sid')

write_csv(clipped_spp, file.path(dir_git, 'clip_fao', 'clipped_s_afr.csv'))
```

``` {r plot_w_lim_spp, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

loiczid_raster <- raster('data/rasters/loiczid_raster.tif')

spp_examples <- sample_n(clipped_spp %>% filter(lim == 'w_lim'), 20)

assemble_map <- function(map_rast, spp) {
  message('...mapping ', spp)
  map_obj <- tm_shape(map_rast) +
    tm_raster(palette = 'Reds',
              colorNA = NULL,
              title = spp,
              alpha = 1) +
    tm_shape(World) +
      tm_polygons() + 
    tm_layout(basemaps = "Esri.WorldTopoMap", 
              # title.position = 'TOP', 
              legend.outside = TRUE, attr.outside = TRUE)

  return(map_obj)
}

rast_list <- vector('list', length = 20)

for (i in 1:nrow(spp_examples)) {  # i <- 1

  spp <- spp_examples[i, ]
  map_cells <- am_spp_cells %>%
    filter(am_sid == spp$am_sid) %>%
    mutate(am_pres = 1)

  r_am_spp  <-  subs(loiczid_raster, 
                     map_cells[ , c('loiczid', 'am_pres')], 
                     by = 'loiczid', 
                     which = 'am_pres', 
                     subsWithNA = TRUE)

  new_ext <- extent(c(0, 70, -70, 0))
  r_am_spp <- crop(r_am_spp, new_ext)

  rast_list[i] <- r_am_spp
  names(rast_list)[i] <- spp

  spp_plot <- assemble_map(r_am_spp, spp$sciname)

  print(spp_plot)

}

mapstack <- stack(rast_list) %>%
  calc(sum, na.rm = TRUE)

values(mapstack)[values(mapstack) == 0] <- NA

fao_polys <- rgdal::readOGR(dsn = file.path(dir_anx, '_raw_data/FAO/FAO_AREAS'),
                            layer = 'FAO_AREAS')

stack_plot <- assemble_map(mapstack, 'multi') +
  tm_shape(fao_polys) +
  tm_borders()
print(stack_plot)

```

``` {r plot_e_lim_spp, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

loiczid_raster <- raster('data/rasters/loiczid_raster.tif')

spp_examples_e <- clipped_spp %>% 
  filter(lim == 'e_lim')

assemble_map <- function(map_rast, spp) {
  message('...mapping ', spp)
  map_obj <- tm_shape(map_rast) +
    tm_raster(palette = 'Reds',
              colorNA = NULL,
              title = spp,
              alpha = 1) +
    tm_shape(World) +
      tm_polygons() + 
    tm_layout(basemaps = "Esri.WorldTopoMap", 
              # title.position = 'TOP', 
              legend.outside = TRUE, attr.outside = TRUE)

  return(map_obj)
}

rast_list <- vector('list', length = length(spp_examples_e))

for (i in 1:nrow(spp_examples_e)) {  # i <- 1

  spp <- spp_examples_e[i, ]
  map_cells <- am_spp_cells %>%
    filter(am_sid == spp$am_sid) %>%
    mutate(am_pres = 1)

  r_am_spp  <-  subs(loiczid_raster, 
                     map_cells[ , c('loiczid', 'am_pres')], 
                     by = 'loiczid', 
                     which = 'am_pres', 
                     subsWithNA = TRUE)

  new_ext <- extent(c(0, 70, -70, 0))
  r_am_spp <- crop(r_am_spp, new_ext)

  rast_list[i] <- r_am_spp
  names(rast_list)[i] <- spp

  spp_plot <- assemble_map(r_am_spp, spp$sciname)

  print(spp_plot)

}

mapstack_e <- stack(rast_list) %>%
  calc(sum, na.rm = TRUE)

values(mapstack_e)[values(mapstack_e) == 0] <- NA

fao_polys <- rgdal::readOGR(dsn = file.path(dir_anx, '_raw_data/FAO/FAO_AREAS'),
                            layer = 'FAO_AREAS')

stack_plot_e <- assemble_map(mapstack_e, 'multi') +
  tm_shape(fao_polys) +
  tm_borders()
print(stack_plot_e)

```

### 2016-08-29 manuscript-main.Rmd
sent to BH, JA, CS

@@
   S. Halpern^1,2,3^
-csl: refs/journal-of-applied-ecology.csl
+csl: refs/plos-one.csl
 output:
@@
 
-1. National Center for Ecological Analysis and Synthesis, 735 State Street Suite 300, Santa Barbara CA 93101
+1. National Center for Ecological Analysis and Synthesis, University of California, 735 State Street Suite 300, Santa Barbara CA 93101
 2. Bren School of Environmental Science and Management, University of California, Santa Barbara CA 93106
@@
 Casey O'Hara, National Center for Ecological Analysis and Synthesis, 735 State Street, Suite 300, Santa Barbara CA 93101; (805) 892-2500; ohara@nceas.ucsb.edu
-
-_Classification:_ BIOLOGICAL SCIENCES: Sustainability Science
 
@@
 
-``` {r abstract, child = 'ms_0_abst_j_appl_ecol.Rmd'}
+``` {r abstract, child = 'ms_0_abst.Rmd'}
 ```
@@
 
-``` {r results_discussion, child = 'ms_3_res_disc.Rmd'}
+``` {r results_discussion, child = 'ms_3_res_disc_concl.Rmd'}
 ```

### 2016-08-29 ms_0_abst.Rmd

--- 2016-08-15 ms_0_abst_sig.Rmd
@@
 Until a single, highly accurate dataset of global marine species ranges becomes available, understanding the implications of dataset differences for conservation planning and decision-making remains essential.
-
-## Significance
-
-<!--_Significance Statement.
-Authors must submit a 120-word-maximum statement about the significance of their research paper written at a level understandable to an undergraduate-educated scientist outside their field of specialty.
-The primary goal of the Significance Statement is to explain the relevance of the work in broad context to a broad readership.
-The Significance Statement appears in the paper itself and is required for all research papers._-->
-
-Successful conservation of marine species critically depends on detailed and accurate understanding of where species exist and where they do not.
-By systematically comparing two widely used global datasets of marine species range maps, we find substantial differences that highlight inconsistencies in our understanding and communication of species ranges.
-Knowing where and why these inconsistencies occur will substantially improve our ability to develop effective and efficient conservation strategies that support resilient ocean ecosystems.

### 2016-08-29 ms_0_abst_150wd.Rmd

## Abstract

<!--_Abstract. Provide an abstract of no more than 250 words on page 2 of the manuscript. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text._-->

Species distribution data provide the foundation for a wide range of ecological research studies and conservation management decisions, yet most species ranges remain unknown, and existing range maps often suffer from data limitations and inconsistencies.  AquaMaps and the International Union for Conservation of Nature (IUCN) are two distinct efforts to map marine species distributions at a global scale, representing `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` shared species.  Here we examine differences in predicted species ranges between the two datasets and find that the divergent methodologies introduce differing frequencies of commission and omission errors.  We illustrate scientific and management implications of these differences by repeating two recent applications - an assessment of global biodiversity within the Ocean Health Index and a global analysis of marine protected area coverage gaps - and find significantly different results depending on how the two datasets are used.

### 2016-08-29 ms_0_abst_j_appl_ecol.Rmd

--- 2016-08-29 ms_0_abst.Rmd
@@
-## Abstract
+## Summary
 
@@
 
+1.
 Species distribution data provide the foundation for a wide range of ecological research studies and conservation management decisions, yet most species ranges remain unknown, and existing range maps often suffer from data limitations and inconsistencies.
@@
 Together these databases represent `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` shared species.
-Here we examine differences in predicted species ranges between the two datasets and find that these misalignments mainly result from divergent methodologies that introduce differing frequencies of commission and omission errors.
-We illustrate the scientific and management implications of these differences by repeating two recent applications - an assessment of global biodiversity within the Ocean Health Index and a global analysis of gaps in coverage of marine protected areas - and find significantly different results depending on how the two datasets are used.
+2.
+Here we examine differences in predicted species ranges between the two datasets by comparing alignment of distribution (predicted species range common to both datasets) and area (extent of range predicted by each dataset) to identify potential sources of misalignment.
+3.
+We find that these misalignments mainly result from divergent methodologies that introduce differing frequencies of commission and omission errors.
+For example, species ranges predicted by IUCN range maps frequently overextend into unsuitably deep territory, as in the case of many coral and reef-associated species; similarly, the AquaMaps species distribution model often predicts species presence far afield from known occurrences when extrapolating based on limited observations.
+4.
+__Policy implications:__ We illustrate the scientific and management implications of these differences by repeating two recent applications - an assessment of global biodiversity within the Ocean Health Index and a global analysis of gaps in coverage of marine protected areas - and find significantly different results depending on how the two datasets are used.
 Until a single, highly accurate dataset of global marine species ranges becomes available, understanding the implications of dataset differences for conservation planning and decision-making remains essential.

### 2016-08-29 ms_1_intro.Rmd

@@
 The two most comprehensive, widely-used global-scale repositories that predict marine species ranges throughout the world's oceans are AquaMaps [@kaschner_aquamaps:_2015] and range data from the International Union for Conservation of Nature (IUCN) [@iucn_iucn_2015].
-Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each offers an understanding of the "truth", with IUCN ranges relying primarily on expert opinion and AquaMaps on model predictions.
-Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: geographic range data such as IUCN range maps frequently introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
+Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each offers an understanding of the truth, with IUCN ranges relying primarily on expert opinion and AquaMaps on model predictions.
+Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: geographic range data such as IUCN range maps frequently introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps will likely introduce fewer commission errors at the expense of more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
 Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006; @jetz_ecological_2008].
@@
 In most of these cases, the implications of choosing one versus the other of these datasets is not evaluated or discussed, yet strong conservation and management conclusions are drawn from the results of the studies.
-For example, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.
+<!-- For example, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.-->
 
@@
 
+<!--
 We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (OHI) [@halpern_index_2012; @halpern_patterns_2015] and a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as case studies to explore the implications of prioritizing one data set over the other.
+The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
+-->
+
+We then reexamine a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as a case study to explore the implications of prioritizing one data set over the other.
 The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.

### 2016-08-29 ms_2_methods.Rmd

@@
 
+<!--
 ___OMIT: Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., previously, the Ocean Health Index defined presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
 In the 2016 OHI assessment, the presence threshold was eliminated, to better approximate the "extent of occurrence" generally indicated by IUCN maps, since the two datasets are used together.___
+-->
 
@@
 
+<!--
+
 ## Methods for Ocean Health Index (OHI) case study
@@
 For the AquaMaps dataset, we used a "probability of occurrence" threshold of 0% to determine presence, to align with the OHI 2016 assessment._
+
+-->
 

### 2016-08-29 clip_fao/fao_clip_pac.Rmd
new draft sent off to Courtney (and Halley and Claire)

@@
 data(World)
+library(maptools)
+data(wrld_simpl)
 
@@
                             layer = 'fao_rgns2')
+rast_ext  <- extent(c(85, 220, -59, 50))
+fao_polys <- crop(fao_polys, rast_ext)
+
+World1 <- crop(wrld_simpl, rast_ext)
 
@@
               alpha = 1) +
-    tm_shape(World) +
+    tm_shape(World1) +
       tm_polygons() + 
@@
               textNA = 'Zero',
-              title = 'Count',
+              title = 'Species count',
               alpha = 1) +
-  tm_shape(World) +
+  tm_shape(World1) +
     tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
@@
   tm_layout(basemaps = "Esri.WorldTopoMap", 
+            frame = FALSE,
             legend.text.size = .35,
@@
 
-save_tmap(rast_spp_map_e, file.path(dir_fig, 'fig4_am_edge.png'), 
+save_tmap(rast_spp_map_e, file.path(dir_fig, 'map_am_edge_map.png'), 
           width = 9, height = 6, units = 'cm', dpi = 600)
@@
     tm_raster(palette = 'Reds',
-              breaks = c(1, 15, 100, 200, 300, Inf),
+              breaks = c(1, 50, 100, 500, 1000, 2000, Inf),
               colorNA = 'lightblue1',
               textNA = 'Zero',
-              title = 'Count',
+              title = 'Species count',
               alpha = 1) +
-  tm_shape(World) +
+  tm_shape(World1) +
     tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
@@
 
-save_tmap(rast_spp_map_w, file.path(dir_fig, 'fig4_am_edge_w.png'), 
+save_tmap(rast_spp_map_w, file.path(dir_fig, 'map_am_edge_w.png'), 
           width = 9, height = 6, units = 'cm', dpi = 600)
@@
 stack_plot_both <- 
-  tm_shape(mapstack_w) +
-    tm_raster(palette = 'Reds',
+  tm_shape(rast_spp_w) +
+    tm_raster(palette = 'Greens',
               colorNA = NULL,
@@
               alpha = 1) +
-  tm_shape(mapstack_e) +
+  tm_shape(rast_spp_e) +
     tm_raster(palette = 'Reds',
@@
               title = 'n_spp') +
-  tm_shape(World) +
+  tm_shape(World1) +
     tm_polygons() + 
-  tm_shape(fao_polys) +
+  tm_shape(fao_polys_e) +
     tm_borders() +
   tm_layout(basemaps = "Esri.WorldTopoMap", 
+            frame = FALSE,
             # title.position = 'TOP', 

### 2016-08-29 manuscript-main.Rmd
new draft sent off to Courtney (and Halley and Claire)

@@
 title: "Aligning marine species range data to better serve science and conservation"
-author: Casey O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin
+author: Casey C. O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin
   S. Halpern^1,2,3^
@@
 
-<!--Title: Titles should be no more than three typeset lines (generally 135 characters including spaces) and should be comprehensible to a broad scientific audience.
-Please provide a short title of fewer than 50 characters for mobile devices and RSS feeds.-->
+<!-- Full title	250 characters	Specific, descriptive, concise, and comprehensible to readers outside the field	 -->
+<!--   Impact of Cigarette Smoke Exposure on Innate Immunity: A Caenorhabditis elegans Model -->
+<!--   Solar Drinking Water Disinfection (SODIS) to Reduce Childhood Diarrhoea in Rural Bolivia: A Cluster-Randomized, Controlled Trial -->
+<!-- Short title	70 characters	State the topic of the study	 -->
+<!--   Cigarette Smoke Exposure and Innate Immunity -->
+<!--   SODIS and Childhood Diarrhoea -->
 
@@
 Email addresses will appear in the article footnotes.-->
-Casey O'Hara, National Center for Ecological Analysis and Synthesis, 735 State Street, Suite 300, Santa Barbara CA 93101; (805) 892-2500; ohara@nceas.ucsb.edu
-
-_Keywords:_ <!--Keywords are listed below the abstract of the published article.
-At least three keywords are required at submission.
-maybe species distribution maps, AquaMaps, IUCN, marine protected area planning, biodiversity status--> Marine spatial planning, species distribution model, extent of occurrence, spatial ecology
+Casey O'Hara, National Center for Ecological Analysis and Synthesis, University of California, 735 State Street, Suite 300, Santa Barbara CA 93101; (805) 892-2500; ohara@nceas.ucsb.edu
 
 <!--
-
-from PNAS (http://www.pnas.org/site/authors/preparation.xhtml):
-
-PNAS generally uses a two-column format averaging 67 characters, including spaces, per line.
-The maximum length of a regular research article is 6 pages (or 10 pages for PNAS Plus articles), including all text, spaces, and the number of characters displaced by figures, tables, and equations.
-
-An online submission tool provides authors with an estimation of whether their manuscript fits within the PNAS length requirements (see Length Estimate FAQ).
-Use the following guidelines to determine whether your manuscript exceeds the page-length requirements before submission:
-
-Text: When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces).
-Figures: Calculated at 180 characters per cm in height for one column and 360 characters per cm in height for two columns; a figure wider than 8.7 cm is two columns wide.
-Tables and equations: Calculated at 60 characters per line for one column and 120 per line for two columns; a table with more than 60 characters per line is two columns wide.
-If the text in a figure or table is deemed illegible for production, the figure or table may be enlarged to improve legibility.
-Add the word count to the character count to obtain a character count that includes spaces.
-
-See also: http://www.pnascentral.org/html/pnas_length_estimate_instructions.html
-
+_Keywords:_  Marine spatial planning, species distribution model, extent of occurrence, spatial ecology
 -->
@@
 ```
+
+``` {r captions_for_barchart_by_taxa, echo = FALSE}
+n_spp <- spp_map_pairs %>%
+  mutate(spp_group = ifelse(str_detect(spp_group, 'CORAL'), 'CORAL', spp_group)) %>%
+  group_by(spp_group) %>%
+  summarize(n_spp = n()) %>%
+  arrange(desc(n_spp))
+
+n_corals   <- nrow(spp_map_pairs %>% 
+                     filter(str_detect(spp_group, 'CORAL')))
+n_bonyfish <- nrow(spp_map_pairs %>% 
+                     filter(spp_group %in% c('WRASSES', 'BUTTERFLYFISH', 'PARROTFISH',
+                                             'COMBTOOTHBLENNIES', 'SEABREAMS_PORGIES', 
+                                             'PUFFERFISH', 'GROUPERS', 'TUNAS_BILLFISHES', 
+                                             'SURGEONFISH_TANGS_UNICORNFISH', 'ANGELFISH', 
+                                             'DAMSELFISH', 'BONEFISH_TARPONS')))
+```
+
+``` {r captions_for_tally_quad_counts, echo = FALSE}
+dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
+area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
+q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
+q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
+q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
+q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
+```
+
 -----
@@
 | :------        |    ------: |    ------: |
-| Abstract       |  184 words |  1312 char |
-| Significance   |   73 words |   545 char |
+| Abstract       |  169 words |     - char |
 -->
@@
 | :------                   |    ------: |    ------: |
-| Introduction              |  429 words |  2905 char |
-| Results/Discussion        |  999 words |  6697 char |
-| Implications (incomplete) |  597 words |  3794 char |
-| Conclusions               |  354 words |  2505 char |
-| Methods (incomplete)      |  601 words |  3919 char |
-| total body                | 2987 words | 19885 char |
+| Introduction              |  653 words |     - char |
+| Methods                   |  878 words |     - char |
+| Results/Discussion        | 2150 words |     - char |
+| Conclusions               |  499 words |     - char |
+| total body                | 4187 words |     - char |
 -->

### 2016-08-29 ms_2_methods.Rmd
new draft sent off to Courtney (and Halley and Claire)

@@
 These experts outline spatial boundaries of a given species' "limits of distribution" [@iucn_iucn_2015], based on observation records and informed by expert understanding of species' range and habitat preferences.
-___Fig.
-SXXX___
+___Fig SXXX___
 
@@
 Of these, 1296 (5.7%) have been further refined through an expert review process [@kaschner_mapping_2006; @ready_predicting_2010].
-___Fig.
-SXXX___
+___Fig SXXX___
 
@@
 __Comparison of taxonomic and regional distribution__:
-To examine the overall taxonomic distribution across the spatial datasets (Fig.
-1A), we grouped species by taxonomic class and data source, and determined the proportion of each class represented in each dataset.
-
-To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; we determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell  ___(Fig.
-SXXX)___.
-For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps ___(Fig.
-SXXX)___.
-We represented relative distribution of species count for each dataset by plotting average species count against latitude and longitude (Fig.
-1B, 1C).
+To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source, and determined the proportion of each class represented in each dataset.
+To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; we determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell  ___(Fig SXXX)___.
+For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps ___(Fig SXXX)___.
 
@@
 Extracting data from the IUCN API [@REF?] we identified the depth limitations of each of the coral species mapped in the IUCN dataset.
-We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] ___(Fig.
-SXXX)___ and masked our IUCN coral rasters to identify mapped coral presence below 200 m.
+We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] ___(Fig SXXX)___ and masked our IUCN coral rasters to identify mapped coral presence below 200 m.
 The resulting maps were again compared to the AquaMaps rasters to examine distribution alignment and area ratio.

### 2016-08-29 ms_5_fig_legends.Rmd
new draft sent off to Courtney (and Halley and Claire)

@@
 # Figure Legends
+
+``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
+knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
+                      echo = FALSE, message = FALSE, warning = FALSE)
+
+# library(ggplot2)
+# library(maps)
+
+# library(raster)
+# library(data.table)
+library(tidyr)
+library(dplyr)
+library(stringr)
+library(readr)
+
+if(basename(getwd()) != 'IUCN-AquaMaps') setwd('~/github/IUCN-AquaMaps')
+
+dir_data <- 'data'
+dir_fig  <- 'figures'
+
+```
+
+``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
+### use data_explore.Rmd to create this file, if it does not yet exist.  This
+### filters out presence = Extinct from IUCN cells.
+
+data_file <- 'data/spp_list_w_area_trimmed.csv'
+spp_map_pairs <- read_csv(data_file)
+### NOTE: as of Jan 31, includes all map pairs, even ones where IUCN area == 0
+
+### generate other lists
+spp_all_file <- 'data/spp_all.csv'
+spp_all <- read_csv(spp_all_file) ### NOTE: contains duped scinames
+
+# iucn_marine_file <- 'data/spp_iucn_marine.csv'
+# spp_iucn_marine <- read_csv(iucn_marine_file)
+
+iucn_maps_file <- 'data/spp_iucn_maps.csv'
+spp_iucn_maps <- read_csv(iucn_maps_file) %>%
+  filter(!duplicated(iucn_sid))
+
+spp_am_file <- 'data/spp_am.csv'
+spp_am <- read_csv(spp_am_file)
+
+```
+
+``` {r calc species numbers, echo = FALSE}
+n_spp_am     <- nrow(spp_am)
+n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
+n_spp_iucn   <- length(unique(spp_iucn_maps$iucn_sid))
+n_spp_both   <- length(unique(spp_map_pairs$iucn_sid))
+n_spp_all    <- n_spp_am + n_spp_iucn - n_spp_both
+
+p_spp_am     <- round(n_spp_am/n_spp_all * 100, 1)
+p_spp_iucn   <- round(n_spp_iucn/n_spp_all * 100, 1)
+p_both_am    <- round(n_spp_both/n_spp_am   * 100, 1)
+p_both_iucn  <- round(n_spp_both/n_spp_iucn * 100, 1)
+p_both_all   <- round(n_spp_both/n_spp_all  * 100, 1)
+
+spp_list <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
+area_align_mean <- round(mean(spp_list$area_ratio, na.rm = TRUE), 1)
+dist_align_mean <- round(mean(spp_list$sm_perc, na.rm = TRUE), 1)
+
+```
 
@@
 Overlapping species are dominated by bony fishes (`r format(n_bonyfish, big.mark = ',')` species, primarily tropical taxa) and corals (`r format(n_corals, big.mark = ',')` species).
-(B, C) Global marine species count per cell according to (B) AquaMaps and (C) IUCN.
+(B, C) Global marine species count per 0.5° cell according to (B) AquaMaps and (C) IUCN.
 The margin frequency plots show relative species count per cell at each latitude and longitude.
@@
 
+<!--
 __Fig. 3.__ Change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.
@@
 The zero threshold in scenario 3 drives a decrease in scores relative to scenario 2.
+-->
 
 __Fig.
-4.__  Percent of species range covered by MPAs based upon methods in Klein et al.
+3.__ (A) Aggregate map combining ranges of the 463 coral species mapped in the IUCN dataset, showing raw ranges and ranges clipped to 200 m depth.
+(B) Alignment quadrant breakdown of paired map coral species using original data from IUCN and AquaMaps (as in Fig.
+2B) and the same species with IUCN ranges clipped to 200 m depth.
+
+__Fig.
+4.__ Composite map showing AquaMaps predicted ranges for 512 species whose mid-latitude range (between 5° S and 25° N) encounters a western limit exactly at 175° W, the boundary between FAO Major Fishing Areas 71 and 77.
+The range maps for these species conform exactly to the bounds of Area 71, as a result of AquaMaps using FAO areas as a rough georegional constraint on species ranges.
+The boundary of Area 57 creates an additional clear exclusion zone for most of these species, despite the fact that this was not used as a selection criteria.
+
+__Fig.
+5.__  Percent of species range covered by MPAs based upon methods in Klein et al.
 (2015).

### 2016-08-29 ms_5_figures.Rmd
new draft sent off to Courtney (and Halley and Claire)

--- 2016-08-07 drafts/initial_version/ms_figures.Rmd
@@
 
-``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
+<!--
+Figure File Requirements
 
-# library(ggplot2)
-# library(maps)
+The list below is an abbreviated summary of the figure specifications. Read the full details of the requirements in the corresponding sections on this page.
+  File Format:  TIFF or EPS
+  Dimensions:   Width: 789 – 2250 pixels (at 300 dpi). Height maximum: 2625 pixels (at 300 dpi).
+  Resolution:   300 – 600 dpi
+  File Size:    <10 MB
+  Text Within Figures: Arial, Times, or Symbol font only in 8-12 point
+  Figure Files: Fig1.tif, Fig2.eps, and so on. Match file name to caption label and citation.
+  Captions:     In the manuscript, not in the figure file.
 
-# library(raster)
-# library(data.table)
-library(tidyr)
-library(dplyr)
-library(stringr)
-library(readr)
+Dimensions:    Centimeters	    Inches	Pixels at 300 dpi
+Minimum width         6.68        2.63	              789
+Maximum width        19.05         7.5               2250
+Height maximum       22.23        8.75               2625
+  At the height maximum, the figure occupies the whole page and excludes the caption
+  “Dimensions” refers to the dimensions of the entire figure, excluding any white space. The closer figures match these dimensions, the closer they will meet expectations on publication.
 
-# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-#            'Darwin'  = '/Volumes/data_edit',
-#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
+TIPS
+  To align your figure with the text column of the PDF version of the article, make it no wider than 5.2 inches (13.2 cm). 
+  Ensure that individual images meet the pixel dimensions indicated above when preparing and saving figures with slideshow software (like Powerpoint, OpenOffice, or Keynote).
+  Make sure source images are 2,250 pixels wide if you want figures at full page width of 19.05cm.
 
-# dir_anx  <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
+When adjusting the figure dimensions, be sure the Chain/Lock symbol is closed. This will maintain the width and height ratio of your figure.
+Go to Image → Properties (or Image → Information) to view image dimensions.
+Resolution
+  Submit figures at the desired dimensions with a resolution no greater than 300-600 dpi. 
+  The quality of a low-resolution figure cannot be improved by simply increasing the resolution in graphics software.
+To improve the resolution of your figure, you must re-create the figure from the beginning.
+  Resolution below 300 results in blurred, jagged or pixelated published figures. Resolution above 600 dpi may lead to resizing of the published figure.
+  The quality of your figures is only as good as the lowest-resolution element placed in them.
+If you created a 72 dpi line graph and placed it in a 300 dpi TIFF, the graph will look blurred, jagged, or pixilated.
 
-if(basename(getwd()) != 'IUCN-AquaMaps') setwd('~/github/IUCN-AquaMaps')
+File size
+  Submit files at a size of 10 MB or less.
+  Reducing TIFF file size
+    Save with LZW compression.
+    Set resolution between 300-600 dpi.
+    Flatten. A flattened TIFF has a single layer called “background” and has a smaller file size than a TIFF with “Layer 1”.
 
-dir_data <- 'data'
-dir_fig  <- 'figures'
+Text within figures
+  Use only Arial, Times, or Symbol font in 8-12 point. 
+  Do not include author names, article title, or figure number/title/caption within figure files.
+That information will go into your figure caption in the manuscript.
+Read more about submitting captions.
 
-```
+Multi-panel figures
+  Place all panels from a multipart figure into a single page and file. If you have a multipart figure spanning multiple files:
+    Combine multiple panels into one page, or break them apart into separate figures.
+    Re-number all figures and in-text citations accordingly.
+  To create a multipanel figure from individual files, use a presentation program such as OpenOffice Impress, Microsoft PowerPoint, or Keynote for Mac. Then convert to TIFF.
+    To set up the page, use the values listed in Dimensions.
+    Use an Insert tool to place figures. Do not drag/drop or copy/paste images into the file, because this results in a 72 dpi image.
+    If your figures have numerous pictures, charts, or small text, they will render best at a resolution of 600 dpi.
 
-``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
-### use data_explore.Rmd to create this file, if it does not yet exist.  This
-### filters out presence = Extinct from IUCN cells.
+Color Mode
+  RGB (8 bit/channel) or grayscale only.
 
-data_file <- 'data/spp_list_w_area_trimmed.csv'
-spp_map_pairs <- read_csv(data_file)
-### NOTE: as of Jan 31, includes all map pairs, even ones where IUCN area == 0
+White space
+  A 2-point white space border around each figure is recommended to prevent inadvertent cropping of content at layout. Crop out excess white space from around image content.
 
-### generate other lists
-spp_all_file <- 'data/spp_all.csv'
-spp_all <- read_csv(spp_all_file) ### NOTE: contains duped scinames
+Orientation
+  Rotate and submit the figure in the orientation that you wish it to publish.
+  Figures will be inserted into the typeset article in the orientation in which they are supplied.
+For example, if a vertical image is submitted in a horizontal orientation, it will be set horizontally in the article.
 
-# iucn_marine_file <- 'data/spp_iucn_marine.csv'
-# spp_iucn_marine <- read_csv(iucn_marine_file)
+Additional Requirements for TIFF 
+  Layers
+    Flattened, with no layers. Figures with a single layer named “layer 1” or “layer 0” are in fact layered.
 
-iucn_maps_file <- 'data/spp_iucn_maps.csv'
-spp_iucn_maps <- read_csv(iucn_maps_file) %>%
-  filter(!duplicated(iucn_sid))
+  Compression
+    LZW compression is required. To apply: 
+      In GIMP, use “Export” instead of “Save As”. Select TIFF as the format, and then select LZW compression.
+      In Photoshop, select “LZW compression” and “Discard Layers, and Save a Copy.”
 
-spp_am_file <- 'data/spp_am.csv'
-spp_am <- read_csv(spp_am_file)
+How to Submit Figures and Captions
+  Your figures contain three elements: figure files, captions, and in-text citations.
+When naming your figures, match the figure file name with corresponding in-text citations and the caption label in the manuscript.
+Example: a figure file named “Fig1.tif” should match the citation “Fig 1” and the figure label “Fig 1.” in the caption.
 
-```
+Figure Files
+  Name files in numeric order with the format: Fig1.tif, Fig2.eps, and so on.
+  Upload each figure as an individual file that is separate from the manuscript.
+  If at any point you change the numbering order of your figures, be sure to update all figure citations, captions, and file names accordingly.
 
-``` {r calc species numbers, echo = FALSE}
-n_spp_am     <- nrow(spp_am)
-n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
-n_spp_iucn   <- length(unique(spp_iucn_maps$iucn_sid))
-n_spp_both   <- length(unique(spp_map_pairs$iucn_sid))
-n_spp_all    <- n_spp_am + n_spp_iucn - n_spp_both
+Captions
+  Organizing
+  Place figure captions in the manuscript text in read order, immediately following the paragraph in which the figure is first cited.
+Do not include captions as part of the figure files or submit them in a separate document.
 
-p_spp_am     <- round(n_spp_am/n_spp_all * 100, 1)
-p_spp_iucn   <- round(n_spp_iucn/n_spp_all * 100, 1)
-p_both_am    <- round(n_spp_both/n_spp_am   * 100, 1)
-p_both_iucn  <- round(n_spp_both/n_spp_iucn * 100, 1)
-p_both_all   <- round(n_spp_both/n_spp_all  * 100, 1)
+Formatting
+  Figure captions have two required elements: figure label and figure title. Legends are optional.
+    Label. Name figure labels using Arabic numerals, and abbreviate the word “Figure” to “Fig” (e.g., Fig 1, Fig 2, Fig 3, etc.).
+    Title. The title should be concise and descriptive. Restrict it to 15 words or less. 
+    Legend. Place the legend directly after the title of the figure to which it belongs. Place any figure credits in the last sentence of the legend.
 
-spp_list <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
-area_align_mean <- round(mean(spp_list$area_ratio, na.rm = TRUE), 1)
-dist_align_mean <- round(mean(spp_list$sm_perc, na.rm = TRUE), 1)
+  Figure legends tips
+    Be succinct: Avoid lengthy descriptions of methods. 
+    Describe the key messages of a figure: provide a description of the figure that will allow readers to understand it without referring to the text.
+    Describe each part of a multipart figure with with a lettered panel label: for example, (A) or (a).
+    Define all non-standard symbols and abbreviations.
 
-```
-
-<!--
-
-Figures: Calculated at 180 characters per cm in height for one column and 360 characters per cm in height for two columns; a figure wider than 8.7 cm is two columns wide.
-
-From PNAS http://www.pnas.org/site/authors/preparation.xhtml:
-
-Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text.
-Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
-(See SI below for supplementary material.) Color images must be in RGB (red, green, blue) mode.
-Include the font files for any text.
-Images must be final size, preferably one column width (8.7 cm).
-Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
-Numbers, letters, and symbols should be no smaller than 6 points (2 mm) and no larger than 12 points (6 mm) after reduction and must be consistent.
-Composite figures must be preassembled.
-Figures must be submitted as separate files, not embedded in manuscript text.
-See the Digital Art Guidelines.
-
-See also: http://www.pnas.org/site/misc/digitalart.pdf
-
-Label like so: 
-Fig. 2.
-Diffraction patterns at three different angles and time-dependent structural changes of JDF-L1.
-(A–C) Diffraction images observed at different zone axes before time 0; n is the electron direction.
-(D) Femtosecond intensity changes of (200) and (201) spots with time.
-Two different types of dynamics were observed: a decay with a femtosecond time constant (see main text) and another without a decay.
-
-___consider using HCL or HSV values for colors on plots?___ or ask Mel re: good color choices.
-Make sure that colors have good contrast to show up on a black-and-white print...
-Should the quad plot background be colored?
-How else to quickly code the four quadrants relative to the quadrant-breakdown-by-taxa plot?
-
-**`hcl {grDevices}`**: Create a vector of colors from vectors specifying hue, chroma and luminance.
-
-*Usage* `hcl(h = 0, c = 35, l = 85, alpha, fixup = TRUE)`
-Arguments
-
-*arguments*
-- `h`: The hue of the color specified as an angle in the range [0,360]. 0 yields red, 120 yields green 240 yields blue, etc.
-- `c`: The chroma of the color. The upper bound for chroma depends on hue and luminance.
-- `l`: A value in the range [0,100] giving the luminance of the colour. For a given combination of hue and chroma, only a subset of this range is possible.
-- `alpha`: numeric vector of values in the range [0,1] for alpha transparency channel (0 means transparent and 1 means opaque).
-- `fixup`: a logical value which indicates whether the resulting RGB values should be corrected to ensure that a real color results.
-if fixup is FALSE RGB components lying outside the range [0,1] will result in an NA value.
-
-**`hsv {grDevices}`**: Create a vector of colors from vectors specifying hue, saturation and value.
-
-*usage* `hsv(h = 1, s = 1, v = 1, alpha)`
-
-*arguments*
-
-- `h`, `s`,`v`: numeric vectors of values in the range [0, 1] for ‘hue’, ‘saturation’ and ‘value’ to be combined to form a vector of colors. Values in shorter arguments are recycled.
-- `alpha`: numeric vector of values in the range [0, 1] for alpha transparency channel (0 means transparent and 1 means opaque).
-
+  In-text citations
+    Cite figures with the format: Fig 1A, Fig 1B, Fig 2, Fig 3, etc.
+    Cite figures in ascending numeric order upon first appearance in the manuscript file.
+This includes citations to text boxes and tables.
+In the published article, figures are inserted according to the placement of their first citation and caption in the article.
+    Lettered subparts of whole figures may be cited in any order in the text, provided the first mention of each whole figure is in numerical order.
+For example, you may cite any subpart of Fig 3 in any order (e.g., Fig 3C before Fig 3A), as long as Figs 1 and 2 have already been cited.
 -->
@@
 
-![](figures/fig1abc_am_iucn_maps.png)
+![](figures/bar_by_taxa_all.png)
+![](figures/all_aquamaps_map.png)
+![](figures/all_iucn_map.png)
+
+-----
 
@@
 
-![](figures/fig2ab_quadmap_sppbars.png)
+![](figures/scatter_quadmap.png)
+![](figures/barchart_spp_gp_quads.png)
 
-### Figure 3
+-----
 
-<a id = 'fig4'></a>
+### Fig. 3
+<a id = 'fig3'></a>
 
-![Absolute change in SPP score](figures/fig3_boxplot_abs_diff.png)
+![coral area change due to depth clipping](figures/coral_map.png)
+
+![coral quadmap breakdown](figures/barchart_coral_quads.png)
+
+-----
 
@@
 
-![Change in MPA coverage](figures/fig4_barchart_mpa_gap.png)
+![aquamaps edge comparison](figures/am_edge_map.png)
 
@@
 
-``` {r figure_legends, child = 'ms_5_fig_legends.Rmd'}
+### Figure 5
+
+<a id = 'fig5'></a>
+
+![Change in MPA coverage](figures/barchart_mpa_gap.png)
+
+-----
+
+``` {r figure_legends, child = 'ms_5_fig_legends.Rmd', eval = FALSE}
 ```

### 2016-08-29 supp_figs_JA.R
init commit

@@ -1,0 +1,106 @@
+# plot for supplement
+
+library(raster)
+library(rgdal)
+library(maps)
+library(readr)
+library(maptools)
+library(tmap)
+
+data("wrld_simpl", package="maptools")
+
+source('~/github/ohiprep/src/R/common.R')
+
+sp <- "Thunnus alalunga"
+setwd('shiny_am_iucn')
+source('server_fxns.R')
+setwd('../')
+
+spp_list <- read.csv('shiny_am_iucn/data/spp_list.csv')
+
+# get rasters for tuna from IUCN and AM
+
+df     <- get_spp_map_df(sp)
+r_am   <- get_rast(df,type = 'am')
+r_iucn <- get_rast(df,type = 'iucn')
+
+plot(r_iucn)
+#playing with grids
+grid(nx = 720, ny = 360, col = 'lightgray')
+
+#crop so we can zoom in and show grid cells
+
+r_iucn_crop <- crop(r_iucn, extent(c(100,150,-25,25)))
+r_am_crop   <- crop(r_am, extent(c(100,150,-25,25)))
+
+basemap <- crop(wrld_simpl, extent(c(100,150,-25,25)))
+
+#pretty up with tmap
+tuna_iucn_raster_grid <- tm_shape(r_iucn_crop) +
+                              tm_raster(palette = 'aquamarine3',
+                                        colorNA = NULL,
+                                        alpha = .8,
+                                        legend.show = FALSE) +
+                              tm_shape(basemap) +
+                              tm_fill(col = 'gray') + 
+                              tm_layout(basemaps = "Esri.WorldTopoMap", 
+                                        # title.position = 'TOP', 
+                                        legend.outside = TRUE, attr.outside = TRUE)+
+                              tm_grid(n.x = 100, n.y = 100, projection = 'longlat', col = 'lightgray', labels.inside.frame = FALSE,labels.size=0)
+
+save_tmap(tuna_iucn_raster_grid, filename = 'figures/si_figs_JA/tuna_iucn_grid_raster.png')
+
+tuna_iucn_raster <- tm_shape(r_iucn_crop) +
+  tm_raster(palette = 'aquamarine3',
+            colorNA = NULL,
+            alpha = .8,
+            legend.show = FALSE) +
[…]

### 2016-08-29 case_study_mpa/3_mpa_spp_analysis.Rmd

@@
 ggsave(plot = barchart_mpa_gap,
-       filename = file.path(dir_fig, 'fig4_barchart_mpa_gap.png'),
+       filename = file.path(dir_fig, 'barchart_mpa_gap.png'),
        width = 11.4, height = 5, units = 'cm', dpi = 600)
-ggsave(plot = barchart_mpa_gap,
-       filename = file.path(dir_fig, 'fig4_barchart_mpa_gap.tif'), device = 'tiff',
-       width = 11.4, height = 5, units = 'cm', dpi = 600)
+# ggsave(plot = barchart_mpa_gap,
+#        filename = file.path(dir_fig, 'barchart_mpa_gap.tif'), device = 'tiff',
+#        width = 11.4, height = 5, units = 'cm', dpi = 600)
 

### 2016-08-29 data_explore/data_explore.Rmd

@@
 print(bar_taxa)
-ggsave(filename = file.path(dir_fig, 'fig1a_bar_by_taxa_all.png'), 
+ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
        width = 17.8, height = 7, units = 'cm', dpi = 600)
@@
 
-ggsave(file.path(dir_fig, 'fig2a_scatter_quadmap.png'),
+ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
        width = 17.8, height = 9, units = 'cm', dpi = 600)
@@
 ggsave(plot = barchart_spp_gp_quads,
-       filename = file.path(dir_fig, 'fig2b_barchart_spp_gp_quads.png'),
+       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
        width = 17.8, height = 7, units = 'cm', dpi = 600)

### 2016-08-29 ms_0_abst.Rmd

@@
 
-<!--_Abstract.
-Provide an abstract of no more than 250 words on page 2 of the manuscript.
-Abstracts should explain to the general reader the major contributions of the article.
-References in the abstract must be cited in full within the abstract itself and cited in the text._-->
+<!-- 
+The Abstract comes after the title page in the manuscript file. The abstract text is also entered in a separate field in the submission system.
+
+The Abstract should:
+  Describe the main objective(s) of the study
+  Explain how the study was done, including any model organisms used, without methodological detail
+  Summarize the most important results and their significance
+  Not exceed 300 words
+Abstracts should not include:
+  Citations
+  Abbreviations, if possible
+-->
 
@@
 Here we examine differences in predicted species ranges between the two datasets and find that these misalignments mainly result from divergent methodologies that introduce differing frequencies of commission and omission errors.
-We illustrate the scientific and management implications of these differences by repeating two recent applications - an assessment of global biodiversity within the Ocean Health Index and a global analysis of gaps in coverage of marine protected areas - and find significantly different results depending on how the two datasets are used.
-Until a single, highly accurate dataset of global marine species ranges becomes available, understanding the implications of dataset differences for conservation planning and decision-making remains essential.
+We illustrate the scientific and management implications of these differences by repeating a global analysis of gaps in coverage of marine protected areas, and find significantly different results depending on how the two datasets are used.
+While we suggest methods to address specific errors in each spatial dataset, it remains essential to understand the implications of dataset differences for conservation planning and decision-making.

### 2016-08-29 ms_1_intro.Rmd

@@
 ---
+
+<!--
+The introduction should:
+  Provide background that puts the manuscript into context and allows readers outside the field to understand the purpose and significance of the study
+  Define the problem addressed and why it is important
+  Include a brief review of the key literature
+  Note any relevant controversies or disagreements in the field
+  Conclude with a brief statement of the overall aim of the work and a comment about whether that aim was achieved
+-->
 
@@
 While we have identified some potential issues in each dataset, we cannot simply recommend one data set over the other.
-Instead, we recommend simple methods to reduce incidence of commission errors in each dataset, improving confidence in species range predictions to better inform conservation management and policy decisions.
+Instead, we recommend simple methods to reduce incidence of errors in each dataset, improving confidence in species range predictions to better inform conservation management and policy decisions.

### 2016-08-29 ms_3_res_disc_concl.Rmd

@@
 ---
+
+<!--
+Results, Discussion, Conclusions
+
+These sections may all be separate, or may be combined to create a mixed Results/Discussion section (commonly labeled “Results and Discussion”) or a mixed Discussion/Conclusions section (commonly labeled “Discussion”).
+These sections may be further divided into subsections, each with a concise subheading, as appropriate.
+These sections have no word limit, but the language should be clear and concise.
+
+Together, these sections should describe the results of the experiments, the interpretation of these results, and the conclusions that can be drawn.
+
+Authors should explain how the results relate to the hypothesis presented as the basis of the study and provide a succinct explanation of the implications of the findings, particularly in relation to previous related studies and potential future directions for research.
+
+PLOS ONE editorial decisions do not rely on perceived significance or impact, so authors should avoid overstating their conclusions. See the PLOS ONE Criteria for Publication for more information.
+-->
 
@@
 
-- figure 1 and caption
-
-[results] The two datasets have notably different taxonomic (Fig.
-1A) and regional (Figs.
-1B, 1C) coverage.
+[results] The two datasets have notably different taxonomic (Fig 1A) and regional (Figs 1B, 1C) coverage.
 AquaMaps encompasses a broader range of taxa than IUCN, as IUCN spatial data files are only available for select taxonomic groups that have been comprehensively assessed.
@@
 The longitude frequency plots show a slight shift in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
+
+> __Fig 1.__ (A) Number and proportion of species by taxa included in each dataset.
+Overlapping species are dominated by bony fishes (`r format(n_bonyfish, big.mark = ',')` species, primarily tropical taxa) and corals (`r format(n_corals, big.mark = ',')` species).
+(B, C) Global marine species count per 0.5° cell according to (B) AquaMaps and (C) IUCN.
+The margin frequency plots show relative species count per cell at each latitude and longitude.
 
@@
 
-[discussion] To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area ratio (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig.
-2A).
+[discussion] To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area ratio (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig 2A).
 Where two independent methods predict presence (or absence) of a species in the same location, we can assume a higher confidence in the accuracy of that prediction.
@@
 
-- quad plot and bar chart and captions
+> __Fig 2.__ (A) Distribution alignment (overlap of smaller range within larger) versus area ratio (the ratio of smaller range area to the larger range area) for `r format(n_spp_both, big.mark = ',')` species included in both IUCN and AquaMaps datasets.
+The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges (n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
+The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in area
+(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
+The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range area, but disagree on where those ranges occur
+(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
+The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
+(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).
+(B) Alignment quadrant breakdown of species by taxonomic group.
 
@@
 By dividing the paired map species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
-Representative maps from each category are provided in the supporting materials (Fig.
-S1).
+Representative maps from each category are provided in the supporting materials (Fig S1).
 
 [results/discussion] The upper right quadrant includes the species (n = `r n_q1`) whose described ranges are above average in alignment of both spatial distribution and area.
-These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig.
-2B).
+These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig 2B).
 This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their ranges span nearly the entire map.
@@
 
-[results/discussion] ___area ratio suggests similar commission/omission errors, but dist suggests otherwise___ Species found in the lower right quadrant (n = `r n_q3`) seem to represent cases of "two wrongs make a right."  For these species, IUCN and AquaMaps both predict ranges extending far beyond the overlapping region, but the methodological differences result in very different extrapolations.
+[results/discussion] Species found in the lower right quadrant (n = `r n_q3`) seem to represent cases of "two wrongs make a right."  For these species, IUCN and AquaMaps both predict ranges extending far beyond the overlapping region, but the methodological differences result in very different extrapolations.
 Consequently, area ratios are close to 100%, though the poor distribution alignment indicates that one or both datasets are introducing significant errors.
@@
 [results] The lower left quadrant includes species (n = `r n_q4`) where alignment is poor in both dimensions.
-___present a solution here?___ Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
+Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
 It is not surprising that range maps based upon fewer observations bear greater uncertainty.
@@
 
-[discussion] Noting from Fig.
-2B that corals dominate the upper-left "distribution-aligned" quadrant of Fig.
-2A (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), we chose to examine the effect of explicitly restricting IUCN ranges to depths based on species' life histories.
+[discussion] Noting from Fig 2B that corals dominate the upper-left "distribution-aligned" quadrant of Fig 2A (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), we chose to examine the effect of explicitly restricting IUCN ranges to depths based on species' life histories.
 Corals offer an excellent case study, due to their foundational role in biodiverse habitats, as well as their lack of mobility and reliance on photosynthesis.
@@
 
-- Coral area plot
-
-[results] Fig.
-3A shows the 463 coral species mapped in the IUCN dataset, with their ranges broken into proportional area deeper and shallower than 200 m.
+[results] Fig 3A shows aggregated ranges of the 463 coral species mapped in the IUCN dataset, with their ranges broken into proportional area deeper and shallower than 200 m.
 According to IUCN descriptions, none of these species is indicated to occur deeper than 200 m, and 94% are confined to waters shallower than 50 m; seven of the mapped species had no reported depth information.
@@
 
-- Coral new quads
+> __Fig 3.__ (A) Aggregate map combining ranges of the 463 coral species mapped in the IUCN dataset, showing raw ranges and ranges clipped to 200 m depth.
+(B) Alignment quadrant breakdown of paired map coral species using original data from IUCN and AquaMaps (as in Fig 2B) and the same species with IUCN ranges clipped to 200 m depth.
 
 [results] In constraining coral ranges to shallow waters, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps.
-Using the original quadrant definitions from Fig.
-2B, we see in Fig.
-3B a massive shift in 354 paired-map coral species due to constraining coral depths to 200 m.
+Using the original quadrant definitions from Fig 2B, we see in Fig 3B a massive shift in 354 paired-map coral species due to constraining coral depths to 200 m.
 Membership in the "well-aligned" quadrant jumped from 22.4% to 76.2%, with a corresponding decrease in all other quadrants.
@@
 
-- AM clipped map
+[results] We identified 4,168 species whose equatorial ranges (between 5° S and 25° N) encounter a range limit at 175° W.
+This longitudinal boundary created a western limit for 512 species, and eastern limit for 3,656 species.
+Plotting the entire ranges of the 512 west-limited species (Fig 4) results in a very clear exclusion zone that matches perfectly with FAO region 71; while we did not focus on the boundary for region 57, a similar effect is readily apparent.
 
-[results] We identified 4,168 species whose equatorial ranges encounter a range limit at 175° W.
-This longitude created a western limit for 512 species, and eastern limit for 3,656.
-Plotting the entire ranges of the 512 west-limited species results in a very clear exclusion zone that matches perfectly with FAO region 71; while we did not focus on the boundary for region 57, a similar effect is readily apparent.
+> __Fig 4.__ Composite map showing AquaMaps predicted ranges for 512 species whose equatorial range (between 5° S and 25° N) encounters a western limit exactly at 175° W, the boundary between FAO Major Fishing Areas 71 and 77.
+The range maps for these species conform exactly to the bounds of Area 71, as a result of AquaMaps using FAO areas as a rough georegional constraint on species ranges.
+The boundary of Area 57 creates an additional clear exclusion zone for most of these species, despite the fact that this was not used as a selection criteria.
 
@@
 ___THIS SECTION NEEDS TO BE REWRITTEN if we choose not to omit it___
-_We calculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig.
-3</a>).
+_We calculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig 3</a>).
 Reducing the threshold increases the apparent range of a species, so a decrease in a region’s score under scenario 1 would indicate increased spatial representation of threatened species for that region.
@@
 
-We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using the most recent AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig4'>Fig.
-4</a>).
+We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using the most recent AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig4'>Fig 4</a>).
 We found a five-fold increase in the proportion of gap species (6.4% of species vs.
@@
 1.5%).
+
+> __Fig 5.__  Percent of species range covered by MPAs based upon methods in Klein et al.
+(2015).
+Scenario 1 replicates the original results, measuring protected range of species in AquaMaps 2014 dataset, with a 50% presence threshold, against the 2014 World Database of Protected Areas, filtered for IUCN categories I-IV that overlap marine areas.
+Scenario 2 updates the results using AquaMaps 2015, showing very small changes despite the inclusion of an additional 5,545 species.
+Scenario 3, using 2015 AquaMaps data, drops the presence threshold to zero, showing an expected decrease in gap species, but also a decrease in species with 5% or greater protected range.
+Scenario 4 adds an additional 1745 species unique to IUCN, resulting in increases in gap species and species with less than 2% coverage.
+Scenario 5 examines species MPA coverage using only the IUCN dataset.
+
+[discussion?]
 
@@
 selecting an appropriate presence threshold for AquaMaps; _a presence threshold of 0% most closely approximates the "limits of distribution" criterion described by IUCN range maps_ ___- not really discussed or backed up here___) and addressing mechanisms that introduce errors of commission and omission.
-___as it stands, this paper really doesn't go into this very much; omit this paragraph?___
+___as it stands, this paper really doesn't go into combining datasets and how to do it; omit this paragraph?___
 

### 2016-09-01 clip_depth/coral_maps.Rmd
some updates to methods and res/disc re: corals and aquamaps clipping
incorporating Halley's suggestions; updated figs to match PLOS-ONE requirements; bringing in Jamie's supplemental figures

@@
 
+coral_cells_iucn <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv'))
+
 coral_comp <- coral_cells_iucn %>%
@@
 save_tmap(map_coral_comp,
-          filename = file.path(dir_fig, 'fig3a_coral_map.png'), 
-          width = 9, height = 4.2, units = 'cm', dpi = 600)
+          filename = file.path(dir_fig, 'coral_map.png'), 
+          width = 19, height = 8, units = 'cm', dpi = 600)
 
@@
 
-violplot1 <- ggplot(spp_coral_clip, aes(x = 'sm_range_raw', y = shallow_pct)) +
-  ggtheme_plot +
-  # geom_violin(fill = 'grey80', alpha = .5,
-  #             draw_quantiles = c(.25, .5, .75)) +
-  # geom_point(position = position_jitter(w = .3), 
-  #            size = 0.25,
-  #            color = 'black', alpha = 1) +
-  geom_violin(fill = NA, size = 0.1, weight = 0.5,
-              draw_quantiles = c(seq(.005, .05, .005), seq(.05, .95, .02), seq(.95, 1.0, .005)),
-              color = 'black') +
-  geom_violin(fill = NA, size = 1,
-              draw_quantiles = NULL,
-              color = 'white') +
-  geom_hline(color = 'grey30', yintercept = 100) +
-  scale_y_continuous(limits = c(0, 100)) +
-  scale_x_discrete(labels = element_blank()) +
-  theme(axis.title.x = element_blank()) +
-  labs(y = 'IUCN coral range: (<= 200 m) / (total)')
-print(violplot1)
-
 ### for area plot, arrange from highest shallow pct to lowest (best to worst)
@@
                       filter(!iucn_sid %in% (iucn_coral_fixed %>% 
-                                               filter(perc > 100) %>% 
+                                               filter(d_perc > 100) %>% 
                                                .$iucn_sid)),
@@
 ggsave(plot = clipped_quads,
-       filename = file.path(dir_fig, 'fig3b_scatter_quadmap_corals.png'), device = 'png',
-       width = 17.8, height = 9, units = 'cm', dpi = 600)
+       filename = file.path(dir_fig, 'si_scatter_quadmap_corals.png'), device = 'png',
+       width = 19, height = 9, units = 'cm', dpi = 600)
 
@@
 ggsave(plot = barchart_coral_quads,
-       filename = file.path(dir_fig, 'fig3c_barchart_coral_quads.png'),
-       width = 17.8, height = 4, units = 'cm', dpi = 600)
+       filename = file.path(dir_fig, 'barchart_coral_quads.png'),
+       width = 19, height = 4, units = 'cm', dpi = 600)
 

### 2016-09-01 data_explore/data_explore.Rmd
incorporating Halley's suggestions; updated figs to match PLOS-ONE requirements; bringing in Jamie's supplemental figures

@@
 ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
-       width = 17.8, height = 7, units = 'cm', dpi = 600)
+       width = 19.05, height = 7, units = 'cm', dpi = 600)
 
@@
 ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
-       width = 17.8, height = 9, units = 'cm', dpi = 600)
+       width = 19.05, height = 9, units = 'cm', dpi = 600)
 
@@
        filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
-       width = 17.8, height = 7, units = 'cm', dpi = 600)
+       width = 19.05, height = 7, units = 'cm', dpi = 600)
 

### 2016-09-01 data_explore/data_explore_quadmaps.Rmd
incorporating Halley's suggestions; updated figs to match PLOS-ONE requirements; bringing in Jamie's supplemental figures

@@
 library(ggplot2)
+library(tmap)
+library(maptools)
+data(wrld_simpl)
 library(maps)
@@
 
-dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-           'Darwin'  = '/Volumes/data_edit',
-           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
+dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
+           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
+           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]
 
@@
 
-dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
+dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico')
 dir_fig <- file.path(dir_git, 'figures')
@@
 #iucn spp cells WITHOUT extinct locations of IUCN subpopulations
-iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
+iucn_spp_cells <- read_csv(file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'), 
                            col_types = 'cddddc', progress = TRUE) %>%
-  as.data.frame() %>%
   filter(presence != 5) ### removes extinct subpopulations
 
-am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), progress = TRUE)
+am_spp_cells   <- read_csv(file.path(dir_anx, 'vAM_IUCN/am_spp_cells_2015.csv'), progress = TRUE)
 
 ### LOICZID cell ID numbers - used in plotting rasters
-loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
+loiczid_raster_file  <- file.path(dir_git, 'data/rasters/loiczid_raster.tif')
 loiczid_raster       <- raster(loiczid_raster_file)
@@
              fill= "#d01c8b") + 
-  geom_point(alpha = .5)
+  geom_point(alpha = .5, size = 1)
 
@@
 
+``` {r tmap_species_maps_fxns}
+
+get_spp_map_df <- function(species, am_cutoff = 0){ ### species <- spp_list$sciname[1]
+  message('in get_spp_map_df()')
+  ### am_cutoff is currently ignored
+
+  spp_id <- spp_list %>%
+    filter(sciname == species) %>%
+    dplyr::select(am_sid, iucn_sid, sciname) %>%
+    distinct()
+
+  iucn_spp_map <- iucn_spp_cells %>%
+    filter(iucn_sid %in% spp_id$iucn_sid) %>%
+    group_by(loiczid) %>%        
+    summarize(iucn_sid = first(iucn_sid))
+      ### The group_by() and summarize() are to eliminate duped cells (e.g. 
+      ### one am_sid matching two iucn_sids)
+
+  am_spp_map   <- am_spp_cells %>%
+    filter(am_sid == spp_id$am_sid)
+
+  spp_map_df <- full_join(iucn_spp_map, am_spp_map, by = 'loiczid') %>%
+    mutate(am_pres = ifelse(is.na(am_sid), NA, 1),
+           iucn_pres = ifelse(is.na(iucn_sid), NA, 1)) %>%
+    dplyr::select(loiczid, am_pres, iucn_pres) %>%
+    mutate(both_pres = ifelse(!is.na(iucn_pres), ### IUCN presence, so...
+                              ifelse(is.na(am_pres), 2, 3), ### if no AM, assign 2; if AM, assign 3. 
+                              am_pres))          ### no IUCN, so assign 1 for AM and NA for none
+
+  return(spp_map_df)
+}
+
+get_rast <- function(spp_map_df, type) {
+  message('in get_rast()')
+  rast_obj  <-  subs(loiczid_raster, spp_map_df, 
+                     by    = 'loiczid', 
+                     which = paste0(type, '_pres'), 
+                     subsWithNA = TRUE)
+
+  rast_obj_shift <- rast_obj %>%
+    raster::shift(360)
+
+  rast_obj <- raster::merge(rast_obj, rast_obj_shift)
+  return(rast_obj)
+}
+
+assemble_map <- function(map_rast, spp) {
+  message('in assemble_map()')
+  map_obj <- tm_shape(map_rast) +
+    tm_raster(palette = 'Spectral',
+              colorNA = NULL,
+              title = spp,
+              alpha = 1) +
+    tm_shape(World) +
+      tm_polygons() + 
+    tm_layout(basemaps = "Esri.WorldTopoMap", title.position = 'TOP', legend.outside = TRUE, attr.outside = TRUE)
+
+  return(map_obj)
+}
+
+wrld_simpl2 <- wrld_simpl %>%
+  rgeos::gUnion(wrld_simpl %>% shift(360))
+
+fao_polys <- rgdal::readOGR(dsn = file.path(path.expand(dir_git), 'clip_fao/fao_rgns'),
+                            layer = 'fao_rgns2')
+
+```
+
 ``` {r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
-### define representative species from each quadrant, for over-plotting
+
 spp_quads <- c('q1' = 'Kajikia albida',       
@@
                'q4' = 'Acanthurus nigroris')
-am_thresh <- 0
+
+spp_extents <- list(c(-150,  60, -75, 75), # 210 deg x 150 deg
+                    c(  30, 240, -75, 75), # 210 deg
+                    c(  90, 300, -75, 75), # 210 deg
+                    c(  30, 240, -75, 75)) # 210 deg
+
+spp_map_list <- vector('list', length = 4)
 
 for (i in 1:length(spp_quads)) { # i = 1
+  species = spp_quads[i]
+
   spp <- spp_list %>% filter(sciname == spp_quads[i])
@@
 
-  spp_plot <- plot_rangemap(spp)
-  spp_plot <- spp_plot + 
-    labs(title = element_blank()) +
-    theme(legend.position = 'none')
+  spp_map_df <- get_spp_map_df(species = spp_quads[i])
 
-  print(spp_plot)
-  plotfile <- file.path(dir_fig, 'spp_maps', 
-                        sprintf('%s_%s.png', 
+  spp_rast <- get_rast(spp_map_df, type = 'both') %>%
+    crop(extent(spp_extents[[i]]))
+
+  fao_polys_crop <- crop(fao_polys,   extent(spp_extents[[i]]))
+  world_crop     <- crop(wrld_simpl2, extent(spp_extents[[i]]))
+
+  # colvec <- c('AquaMaps' = '#76A94C', 'Both' = '#e18258', 'IUCN' = '#9168a9')
+
+  if (i == 1) {
+    spp_map <- tm_shape(spp_rast) +
+      tm_raster(palette = c('#76A94C', '#9168a9', '#e18258'),
+                title  = 'Dataset',
+                labels = c('AquaMaps', 'IUCN', 'both'),
+                colorNA = '#f0f8ff',
+                showNA = FALSE,
+                alpha = 1)
+  } else {
+    spp_map <- tm_shape(spp_rast) +
+      tm_raster(palette = c('#76A94C', '#9168a9', '#e18258'),
+                legend.show = FALSE,
+                colorNA = '#f0f8ff',
+                alpha = 1)
+  }
+
+  spp_map <- spp_map +
+    tm_shape(fao_polys_crop) +
+      tm_borders(col = 'grey70', lwd = 0.25, alpha = .3) +
+    tm_shape(world_crop) +
+      tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
+    tm_layout(title = species, title.size = .5,
+              frame = FALSE,
+              legend.text.size = .35,
+              legend.title.size = .45,
+              # title.position = 'TOP', 
+              legend.outside = FALSE,
+              legend.position = c('right', 'bottom'),
+              legend.bg.color = '#f0f8ff',
+              legend.bg.alpha = .5,
+              attr.outside = TRUE)
+
+  spp_map_list[[i]] <- spp_map
+
+  save_tmap(spp_map, file.path(dir_fig, sprintf('spp_maps/spp_map_%s_%s.png', 
                                                 spp_quad, 
-                                tolower(str_replace(spp_quads[i], ' ', '_'))))
-  ggsave(plotfile, height = 4.5, width = 8.7, units = 'cm')
+                                                str_replace(tolower(species), ' ', '_'))), 
+            width = 9.4, height = 6.5, units = 'cm', dpi = 600)
+
 }
+
+# multiplot <- grid.arrange(spp_map_list[[2]],
+#                           spp_map_list[[1]],
+#                           spp_map_list[[4]],
+#                           spp_map_list[[3]],
+#                           ncol = 2, main = "Main title")
 

### 2016-09-01 manuscript-main.Rmd
incorporating Halley's suggestions; updated figs to match PLOS-ONE requirements; bringing in Jamie's supplemental figures

@@
 
+Short title: "Aligning marine species range data"
 <!-- Full title	250 characters	Specific, descriptive, concise, and comprehensible to readers outside the field	 -->
@@
 <!--   SODIS and Childhood Diarrhoea -->
-
-_Author affiliation:_ <!--Include department, institution, and complete address, with the ZIP/postal code, for each author.
-Use superscripts to match authors with institutions.
-Authors are encouraged to supply their ORCID ID at submission-->
 

### 2016-09-01 ms_1_intro.Rmd
incorporating Halley's suggestions; updated figs to match PLOS-ONE requirements; bringing in Jamie's supplemental figures

@@
 
-Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation, among many others.  This knowledge provides foundational information for 
- understanding species ranges and diversity,
- predicting species responses to human impacts and climate change,
- and managing and protecting species effectively.
-<!--___A rich literature tackles the many dimensions of these questions.___-->
+Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation.
+This body of science has resulted in various compiled databases of species distribution maps which provide foundational information for understanding species ranges and diversity, predicting species responses to human impacts and climate change, and managing and protecting species effectively.
+No species range dataset can claim to represent the "truth" of any species' spatial distribution, but rather each offers its own distinct and often divergent understanding of the "true" distribution.
+Despite the uncertainty and variation among species range datasets, conservation managers and policy makers must act on conclusions drawn from these data.
+It is critical to understand the differences in spatial range datasets and the implications of these differences on conservation decisions.
 
-One major outcome of this body of science is the various compiled databases of species distribution maps.
-The two most comprehensive, widely-used global-scale repositories that predict marine species ranges throughout the world's oceans are AquaMaps [@kaschner_aquamaps:_2015] and range data from the International Union for Conservation of Nature (IUCN) [@iucn_iucn_2015].
-Neither dataset claims to represent the "truth" of a species' spatial distribution, but rather each offers an understanding of the truth, with IUCN ranges relying primarily on expert opinion and AquaMaps on model predictions.
+<!--implications of choosing one versus the other of these datasets is not evaluated or discussed, yet strong conservation and management conclusions are drawn from the results of the studies.-->
+
+The two most comprehensive, widely-used global-scale repositories that predict marine species ranges throughout the world's oceans are AquaMaps, which rely primarily on model predictions [@kaschner_aquamaps:_2015], and range data from the International Union for Conservation of Nature (IUCN), which rely primarily on expert opinion [@iucn_iucn_2015].
 Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: geographic range data such as IUCN range maps frequently introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps will likely introduce fewer commission errors at the expense of more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
@@
 In most of these cases, the implications of choosing one versus the other of these datasets is not evaluated or discussed, yet strong conservation and management conclusions are drawn from the results of the studies.
-<!-- For example, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.-->
 
@@
 Close alignment between two independent range maps can therefore indicate higher confidence in species presence or absence than a single map.
-For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align in both distribution and overall area; from the results we determine methodological issues that can introduce commission errors, and suggest methods to improve confidence in species range predictions.
-In particular, we show how solving a simple problem in coral ranges in the IUCN range data dramatically ___improves agreement with Aquamaps data.___ - _does this make it sound like agreement is the goal?
-make it clear that agreement implies better confidence_
+For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align in both distribution and overall area.
+
+Using this alignment framework, we explore methodological issues that can introduce errors, and suggest methods to improve confidence in species range predictions.
+In particular, we show how explicitly using depth as a constraint on coral ranges in IUCN range data dramatically improves concurrence with AquaMaps data, ___and highlight a side effect of AquaMaps methodology that results in surprising species range boundaries.___
 
@@
 
-We then reexamine a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as a case study to explore the implications of prioritizing one data set over the other.
+As a case study, we then reexamine a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as a case study to explore the implications of prioritizing one data set over the other.
 The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
@@
 While we have identified some potential issues in each dataset, we cannot simply recommend one data set over the other.
-Instead, we recommend simple methods to reduce incidence of errors in each dataset, improving confidence in species range predictions to better inform conservation management and policy decisions.
+Instead, we recommend simple methods to account for incidence of errors in each dataset, improving confidence in species range predictions to better inform conservation management and policy decisions.

### 2016-09-01 ms_2_methods.Rmd
incorporating Halley's suggestions; updated figs to match PLOS-ONE requirements; bringing in Jamie's supplemental figures

@@
 The IUCN publishes species range maps developed by species experts.
-These experts outline spatial boundaries of a given species' "limits of distribution" [@iucn_iucn_2015], based on observation records and informed by expert understanding of species' range and habitat preferences.
-___Fig SXXX___
+These experts outline spatial boundaries of a given species' "limits of distribution" [@iucn_iucn_2015], based on observation records and informed by expert understanding of species' range and habitat preferences (Fig S1A)
 
-In contrast, AquaMaps models species distribution based on environmental preferences, such as temperature, depth, and salinity, deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a global 0.5° grid to determine suitable habitat, resulting in a "probability of occurrence" for each species.
-Of these, 1296 (5.7%) have been further refined through an expert review process [@kaschner_mapping_2006; @ready_predicting_2010].
-___Fig SXXX___
+In contrast, AquaMaps models species distribution based on envelopes of environmental preference, such as temperature, depth, and salinity; these preference envelopes are deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a global 0.5° grid to determine suitable habitat, resulting in a "probability of occurrence" for each species (Fig S1B).
+To roughly constrain species ranges to appropriate georegions, the AquaMaps model uses Food and Agriculture Organization of the United Nations (FAO) Major Fishing Area [@REF] boundaries.
+Of the resulting maps, 1296 (5.7%) have been further refined through an expert review process [@kaschner_mapping_2006; @ready_predicting_2010].
 
@@
 To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source, and determined the proportion of each class represented in each dataset.
-To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; we determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell  ___(Fig SXXX)___.
-For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps ___(Fig SXXX)___.
+To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; we determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell  ___(Fig S2)___.
+For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.
 
-__Comparison of paired maps__:
-Using genus and species binomials as a matching key, we identified "paired map" species - the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.
+__Alignment framework for paired maps__: OR __Comparison of paired maps__:
+Using genus and species binomials as a matching key, we identified "paired map" species, i.e.
+the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.
 We used the taxize package [@REF] in R [@REF] to standardize species names and synonyms; for species with separate subpopulation maps in IUCN, we combined all subpopulations to create a single global population.
@@
 $A_{small \cap large}$ represents the amount of overlapping area between the two datasets.
+Distribution alignment uses overlapping predictions of presence as means of identifying areas of higher confidence.
+Area ratio provides a proxy for frequency of commission and/or omission errors.
 
-Distribution alignment uses overlapping predictions of presence as means of identifying areas of higher confidence.  Area ratio provides a proxy for frequency of commission and/or omission errors.
-
-To examine errors of commission related to depth, we selected corals as a case study due to their importance in supporting biodiversity as well as their dependence on photosynthesis.
-Extracting data from the IUCN API [@REF?] we identified the depth limitations of each of the coral species mapped in the IUCN dataset.
-We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] ___(Fig SXXX)___ and masked our IUCN coral rasters to identify mapped coral presence below 200 m.
+__(some heading here? to separate from alignment)__
+Corals provide a useful case study to examine differences in species range alignment.
+They are an intensely studied taxonomic group whose importance in supporting biodiversity is undisputed; their dependence on photosynthesis provides habitat-limiting life history traits, and they don't tend to move.
+Extracting data from the IUCN API [@REF?] we identified depth limitations of each coral species mapped in the IUCN dataset to verify that none are found below the photosynthetic depth limit of 200 m.
+Using the same methodology as shown in Fig 2, we created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] and masked our IUCN coral rasters to identify mapped coral presence below 200 m.
 The resulting maps were again compared to the AquaMaps rasters to examine distribution alignment and area ratio.
 
-AquaMaps uses Food and Agriculture Organization of the United Nations (FAO) Major Fishing Area [@REF] boundaries as a rough method of constraining species ranges to appropriate georegions.
-To explore the impact of these artificial constraints on species range predictions, we chose to focus on boundaries defined by longitude, as they seem less likely to relate to ecological conditions than boundaries defined by latitude.
+AquaMaps' reliance upon FAO zones to constrain species ranges seems likely to result in odd species range boundaries where ecologically suitable habitat intersects with human-defined borders.
+To explore the impact of these artificial constraints on species range predictions, we chose to focus on boundaries defined by longitude, as they seem less likely to coincide with ecological differences than boundaries defined by latitude.
 As an example, we identified all AquaMaps species whose eastern or western range limit between between 5° S and 25° N latitudes coincided with the large vertical boundary between FAO regions 71 and 77 at 175° W longitude.
 
-To identify data-poor species, we used the ___robis???___ and ___rgbif???___ packages [@REFS] in R to identify known occurrences of each species; occurrences were averaged between the two occurrence databases as a proxy for data richness.
+The oceans abound with rare and poorly-understood species who offer sparse information on behavior and habitat preferences to inform models and expert opinion.
+To identify such data-poor species within the set of paired map species, we used the ___robis???___ and ___rgbif???___ packages [@REFS] in R to identify known occurrences of each species; occurrences were averaged between the two occurrence databases as a proxy for data richness.
 

### 2016-09-01 ms_3_res_disc_concl.Rmd
incorporating Halley's suggestions; updated figs to match PLOS-ONE requirements; bringing in Jamie's supplemental figures

@@
 AquaMaps encompasses a broader range of taxa than IUCN, as IUCN spatial data files are only available for select taxonomic groups that have been comprehensively assessed.
-While species numbers in both datasets peak in tropical latitudes near the equator, species counts for IUCN maps drop quickly beyond 30°N and 30°S, while species counts for AquaMaps remain robust well into temperate latitudes.
+While species numbers in both datasets peak in tropical latitudes near the equator, species counts for IUCN maps drop quickly beyond 30°N and 30°S, while AquaMaps includes distribution of species well into temperate latitudes.
 The longitude frequency plots show a slight shift in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
@@
 
-[discussion] To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area ratio (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig 2A).
+[discussion] To explore differences in species distribution and range between the two datasets, we assessed the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) relative to the area ratio (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig 2A).
 Where two independent methods predict presence (or absence) of a species in the same location, we can assume a higher confidence in the accuracy of that prediction.
@@
 Membership in the "well-aligned" quadrant jumped from 22.4% to 76.2%, with a corresponding decrease in all other quadrants.
-See SXXX to examine the shifts of individual species among the quadrants.
+See Fig SXXX to examine the shifts of individual species among the quadrants.
 

### 2016-09-01 ms_6_supplemental.Rmd
incorporating Halley's suggestions; updated figs to match PLOS-ONE requirements; bringing in Jamie's supplemental figures

@@
 
-from PNAS http://www.pnas.org/site/authors/preparation.xhtml:
+-->
 
-The main text of the paper must stand on its own without the SI.
-Refer to SI in the manuscript at an appropriate point in the text.
-Number supporting figures and tables starting with S1, S2, etc.
-Authors are limited to no more than 10 SI files, not including movie files.
+methods:
+IUCN raw -> grid -> raster -> final
+AM raw -> presence
+bathymetry poly -> raster?
 
-Authors who place detailed materials and methods in SI must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the online methods.
-If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
+results/discussion:
+quad plot with risk and four species (?)
+four representative maps
 
-Because PNAS edits SI and composes it into a single PDF, authors must provide the following file formats only:
+coral quadrant shifts
 
-Text: Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).
-Figures: Provide a brief legend for each supporting figure after the supporting text.
-Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text.
-When saving TIFF files, use only LZW compression; do not use JPEG compression.
-Do not save figure numbers, legends, or author names as part of the image.
-Composite figures must be preassembled.
-3D figures: Supply a composable U3D or PRC file so that it may be edited and composed.
-Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.
-Tables: Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file.
-Do not use tabs or spaces to separate columns in Word tables.
-Datasets: Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
-Movies: Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file.
-All movies should be submitted at the desired reproduction size and length.
-Movies should be no more than 10 MB in size.
-Still images: Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.
-Appendices: PNAS prefers that authors submit individual source files to ensure readability.
-If this is not possible, supply a single PDF file that contains all of the SI associated with the paper.
-This file type will be published in raw format and will not be edited or composed.
--->
-<!--
-# Data sources and methods
+AM thresholds? not mentioned in body of paper as of yet
 
-All code is publicly available and freely downloadable [here on GitHub](link to github).
+## Preparing AquaMaps and IUCN data layers
 
-## AquaMaps
+__Fig S1.__  Examples of AquaMaps and IUCN raw species range data.
+(A) AquaMaps distribution of _Thunnus alalunga_ (Albacore Tuna) showing global range with varying probabilities of occurrence assigned to 0.5° grid cells.
+(B) IUCN distribution of _T.
+alalunga_ represented as extent of occurrence polygons.
 
-AquaMaps data was provided by AquaMaps directly via ftp.
-Three SQL files were sent that contained (1) list of all 22,889 modeled species and their taxonomic information, (2) information for all (259,200) half degree cells used for range mapping and (3) the probability of occurrence per cell for all 22,889 species.
-These were extracted into .csv form using R Statistical Software.
-
-## IUCN
-
-As of December 2015, IUCN had published species distribution maps for XXX marine species across 24 taxonomic groups.
-For this analysis, we did not consider IUCN range maps for bird species, as those data are hosted separately by BirdLife International.
-
-Each shapefile contains spatial polygons outlining limits of distribution for each species, indicating regions of presence/absence; additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
--->
+__Fig S2.__ Rasterizing shapefiles provided by IUCN.
+A portion of the _T.
+alalunga_ range map is used to exemplify the rasterization process.
+To enable direct comparison of IUCN species ranges to AquaMaps species ranges, the raw IUCN polygon (A) is overlaid with a 0.5° degree grid matching the AquaMaps grid (B).
+Each cell is assigned a value of "present" if the cell overlaps any portion of the polygon (C).
+The resulting raster (D).
 
@@
 
-![](figures/si_four_maps.png)
-
->  __Fig.
-S1.__ Representative species maps to illustrate each quadrant from Fig.
-2A.
-Note that B-E are arranged to match quadrants in Fig.
-S1A.
-(A) Alternative version of Figure 2A that includes extinction risk categories for each species.
-Four representative species, one from each quadrant, are mapped in (B-E).
+__Fig S3.__ Representative species maps to illustrate each quadrant from Fig 2A.  Each map is positioned to match its quadrant in Fig 2A.
+(A) Distribution-matched: Conus episcopatus, the dignified cone snail.  Ranges shows excellent overlap in the western Pacific, though IUCN range extends beyond the bounds of the AquaMaps range.
 (B) Well-matched: Kajikia albida, the Atlantic white marlin.  Ranges from each data set show nearly complete overlap, and very similar area of coverage.
-(C) Distribution-matched: Conus episcopatus, the dignified cone snail.  Ranges shows excellent overlap in the western Pacific, though IUCN range extends beyond the bounds of the AquaMaps range.
-(D) Area-matched: Conus magnificus, the magnificent cone snail.  Ranges overlap in the southern Pacific, but align poorly elsewhere.  The total area for each range is similar.
-(E) Poorly-matched: Acanthurus nigroris, the blue-lined surgeonfish.
+(C) Poorly-matched: Acanthurus nigroris, the blue-lined surgeonfish.
 IUCN predicts species range only near the Hawaiian islands; AquaMaps predicts extensive range throughout the central and western Pacific Ocean.
 The datasets align in neither distribution nor area of coverage.
+(D) Area-matched: Conus magnificus, the magnificent cone snail.  Ranges overlap in the southern Pacific, but align poorly elsewhere.  The total area for each range is similar.
 
@@
 
->  __Fig.
-S2.__ Breakdown of quadrants (Fig.
-S1A) by IUCN extinction risk categories.
+>  __Fig S2.__ Breakdown of quadrants (Fig S1A) by IUCN extinction risk categories.
 IUCN and Aquamaps ranges were in better agreement (i.e., had a higher probability of being in the upper-right, "well-aligned" quadrant, Figure 2A/Figure S1A) for species at greater risk according to IUCN risk categories (p < 0.001, generalized linear model, family = binomial, Table S1).
@@
 
->  __Fig.
-S3.__ IUCN-defined range of _Oculina varicosa_, a typical photosynthetic coral, plotted against a 200 m bathymetry contour.
+>  __Fig S3.__ IUCN-defined range of _Oculina varicosa_, a typical photosynthetic coral, plotted against a 200 m bathymetry contour.
 According to IUCN, _O.
@@
 
->  __Fig.
-S4.__ AquaMaps distribution map extent remaining after applying a presence threshold.
+>  __Fig S4.__ AquaMaps distribution map extent remaining after applying a presence threshold.
 A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.

### 2016-09-02 data_explore/data_explore_quadmaps.Rmd
updated figs, etc

@@
 
+  world_crop     <- crop(wrld_simpl2, extent(spp_extents[[i]]))
   fao_polys_crop <- crop(fao_polys,   extent(spp_extents[[i]]))
-  world_crop     <- crop(wrld_simpl2, extent(spp_extents[[i]]))
 
@@
     spp_map <- tm_shape(spp_rast) +
-      tm_raster(palette = c('#76A94C', '#9168a9', '#e18258'),
+      tm_raster(palette = c('#4dac26', '#d01c8b', '#4d4dac'),
                 title  = 'Dataset',
@@
     spp_map <- tm_shape(spp_rast) +
-      tm_raster(palette = c('#76A94C', '#9168a9', '#e18258'),
+      tm_raster(palette = c('#4dac26', '#d01c8b', '#4d4dac'),
                 legend.show = FALSE,
@@
       tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
-    tm_layout(title = species, title.size = .5,
+    tm_layout(# title = species, title.size = .75,
+              # title.position = c('left', 'bottom'),
               frame = FALSE,
-              legend.text.size = .35,
-              legend.title.size = .45,
-              # title.position = 'TOP', 
+              legend.text.size = .5,
+              legend.title.size = .6,
               legend.outside = FALSE,

### 2016-09-02 ms_1_intro.Rmd
updated figs, etc

@@
 Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation.
-This body of science has resulted in various compiled databases of species distribution maps which provide foundational information for understanding species ranges and diversity, predicting species responses to human impacts and climate change, and managing and protecting species effectively.
-No species range dataset can claim to represent the "truth" of any species' spatial distribution, but rather each offers its own distinct and often divergent understanding of the "true" distribution.
-Despite the uncertainty and variation among species range datasets, conservation managers and policy makers must act on conclusions drawn from these data.
-It is critical to understand the differences in spatial range datasets and the implications of these differences on conservation decisions.
-
-<!--implications of choosing one versus the other of these datasets is not evaluated or discussed, yet strong conservation and management conclusions are drawn from the results of the studies.-->
+This body of science has resulted in various compiled databases of species distribution and range maps which provide foundational information for understanding species diversity and extinction risk, predicting species responses to human impacts and climate change, and managing and protecting species effectively.
+No species range dataset can claim to represent the "truth" of any species' spatial distribution, but rather each offers its own distinct understanding of that latent distribution.
+These varying predictions of species presence and absence, driven by intent and methodology, can result in very different predictions of species presence; however, conservation managers and policy makers must base their actions on conclusions drawn from these imperfect datasets.
+It is critical to understand the differences in spatial range datasets and the implications of these differences for conservation research and decision-making.
 
 The two most comprehensive, widely-used global-scale repositories that predict marine species ranges throughout the world's oceans are AquaMaps, which rely primarily on model predictions [@kaschner_aquamaps:_2015], and range data from the International Union for Conservation of Nature (IUCN), which rely primarily on expert opinion [@iucn_iucn_2015].
-Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: geographic range data such as IUCN range maps frequently introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps will likely introduce fewer commission errors at the expense of more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
+The two datasets are designed to serve different purposes:  IUCN range maps are drawn to estimate extent of occurrence (EOO) [@GastonXXX 1994], communicating range size as an [indicator/factor/criterion] for extinction risk assessments [@IUCNXXX]; AquaMaps essentially communicate the area of occupancy (AOO) [@REFXXX] of a species based on habitat suitability [@AquaMapsXXX].
+
+The fundamental differences between the datasets (EOO versus AOO, measure of range size versus distribution) suggest that the choice of one over the other should be carefully matched to the purpose for which it is to be used [@GastonXXX 2008, XXX Other refs], and yet these datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
+In most of these cases, the implications of choosing one versus the other of these datasets is not evaluated or discussed, yet strong conservation and management conclusions are drawn from the results of the studies.
+
+<!--
+_The fundamental differences between the intent of the two datasets precludes casual interchangeability between the two [@GastonXXX 2008], and yet they_
+
+_"The two measures (EOO and AOO) serve different purposes, and should not be regarded as alternatives that simply differ in accuracy" - Gaston et al 2004 - literature suggests that this distinction precludes casual interchangeability between the two [@gastonXXX 2008]; they have specific purposes.
+And yet they get used, with no explicit acknowldegment of the difference, for a wide range of conservation studies.
+It's important to understand the implications and measures to help adapt them to specific uses._
+-->
+
+Here we focus not on how the two datasets should ideally be used; instead, in recognizing that each provides value for many aspects of conservation research, we highlight hidden assumptions and sources of error within each dataset, and explore the implications of choosing one over the other.
+
+Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: extent of occurrence range data such as IUCN range maps will inevitably introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps are expected to introduce fewer commission errors at the expense of more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
 Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006; @jetz_ecological_2008].
-
-These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
-In most of these cases, the implications of choosing one versus the other of these datasets is not evaluated or discussed, yet strong conservation and management conclusions are drawn from the results of the studies.
 
@@
 Although relatively small in number, these overlap species present a unique opportunity to evaluate the two datasets overall.
-As understanding of a given species improves, we expect that range maps, regardless of method, would become more accurate in predicting species presence and absence.
+As understanding of a given species improves, we expect that range maps, regardless of method, would more closely approach the true distribution in predicting species presence and absence.
 Close alignment between two independent range maps can therefore indicate higher confidence in species presence or absence than a single map.
-For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align in both distribution and overall area.
+For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align in both spatial distribution and overall area.
 
 Using this alignment framework, we explore methodological issues that can introduce errors, and suggest methods to improve confidence in species range predictions.
-In particular, we show how explicitly using depth as a constraint on coral ranges in IUCN range data dramatically improves concurrence with AquaMaps data, ___and highlight a side effect of AquaMaps methodology that results in surprising species range boundaries.___
+In particular, we show how explicitly using depth as a constraint on coral ranges in IUCN range data dramatically mitigates a source of commission error, and highlight a side effect of AquaMaps methodology that results in potentially unexpected species range boundaries.
 
@@
 
-As a case study, we then reexamine a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as a case study to explore the implications of prioritizing one data set over the other.
-The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
+As a case study, we then reexamine a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] as a case study to explore the implications of selecting one data set over the other.
+Repeating the study using IUCN range data rather than AquaMaps data increases the apparent proportion of gap species from 1.4% to 6.4%, while also increasing the proportion of well-protected species from 1.5% to 2.9%.
+This shift highlights the implications of different data use decisions on our understanding of marine biodiversity status and protection.
 

### 2016-09-02 case_study_mpa/3_mpa_spp_analysis.Rmd

@@
        filename = file.path(dir_fig, 'barchart_mpa_gap.png'),
-       width = 11.4, height = 5, units = 'cm', dpi = 600)
+       width = 19.05, height = 5, units = 'cm', dpi = 600)
 # ggsave(plot = barchart_mpa_gap,

### 2016-09-02 clip_fao/fao_clip_pac.Rmd

@@
                             layer = 'fao_rgns2')
-rast_ext  <- extent(c(85, 220, -59, 50))
+
+rast_ext  <- extent(c(-30, 330, -75, 75))
 fao_polys <- crop(fao_polys, rast_ext)
 
-World1 <- crop(wrld_simpl, rast_ext)
+# wrld_simpl_shift <- wrld_simpl %>%
+#   raster::shift(360)
+# 
+# World1 <- rgeos::gUnion(wrld_simpl, wrld_simpl_shift, byid = FALSE) %>%
+#   crop(rast_ext) %>%
+#   SpatialPolygonsDataFrame(data.frame(x = 1))
+# 
+# maptools::writePolyShape(World1, 'clip_fao/fao_rgns/world1')
+
+World1 <- maptools::readShapePoly('clip_fao/fao_rgns/world1')
+
+rast_spp_shift <- rast_spp %>%
+  raster::shift(360)
+
+rast_spp_e <- merge(rast_spp, rast_spp_shift)
 
@@
 ### group affected spp by loiczid and sum
-spp_cells_clipped <- am_spp_cells %>%
-  filter(am_sid %in% spp_examples_e$am_sid) %>%
-  group_by(loiczid) %>%
-  summarize(n_spp = n())
+
+# spp_cells_clipped <- am_spp_cells %>%
+#   filter(am_sid %in% spp_examples_e$am_sid) %>%
+#   group_by(loiczid) %>%
+#   summarize(n_spp = n())
+# write_csv(spp_cells_clipped, file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv'))
+
+spp_cells_clipped <- read_csv(file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv'))
 
@@
 # values(rast_spp_e)[is.na(values(rast_spp_e))] <- 0
-rast_spp_e  <- crop(rast_spp_e, extent(c(85, 220, -59, 50)))
+rast_spp_e  <- crop(rast_spp_e, rast_ext)
 
@@
     tm_borders(col = 'grey40', lwd = 0.25) +
-    tm_text('F_CODE_1',
+    tm_text('F_AREA_1',
             col = 'orangered4',
@@
             size = .35) +
-  tm_layout(basemaps = "Esri.WorldTopoMap", 
-            frame = FALSE,
+  tm_layout(# basemaps = "Esri.WorldTopoMap", 
+            frame = TRUE,
             legend.text.size = .35,
@@
             # title.position = 'TOP', 
-            legend.outside = TRUE,
-            # legend.position = c('right', 'bottom'),
-            # legend.bg.color = 'white',
-            # legend.bg.alpha = .5,
+            legend.outside = FALSE,
+            legend.position = c('right', 'top'),
+            legend.bg.color = 'white',
+            legend.bg.alpha = .8,
             attr.outside = TRUE)
@@
 
-save_tmap(rast_spp_map_e, file.path(dir_fig, 'map_am_edge_map.png'), 
-          width = 9, height = 6, units = 'cm', dpi = 600)
+save_tmap(rast_spp_map_e, 
+          file.path(dir_fig, 'map_am_edge_map.png'), 
+          width = 19, 
+          # height = 6, 
+          units = 'cm', dpi = 600)
 

### 2016-09-02 data_explore/data_explore.Rmd

@@
                  position = position_dodge(width = 0.5)) +
-  scale_size_manual(values = c('IUCN' = .25, 'AquaMaps' = .25, 'Paired' = 2.5)) +
-  scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey60', 'Paired' = 'black')) +
+  scale_size_manual(values = c('IUCN' = .5, 'AquaMaps' = .5, 'Paired' = 3)) +
+  scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey70', 'Paired' = 'black')) +
   geom_text(aes(label = n_label, y = 102),

### 2016-09-02 ms_2_methods.Rmd

@@
 The IUCN publishes species range maps developed by species experts.
-These experts outline spatial boundaries of a given species' "limits of distribution" [@iucn_iucn_2015], based on observation records and informed by expert understanding of species' range and habitat preferences (Fig S1A)
+These experts outline spatial boundaries that define the "limits of distribution" of a given species, based on observation records and informed by expert understanding of species' range and habitat preferences (Fig S1A).
+<!--The result is essentially an extent of occurrence (EOO) map for the species-->
 
@@
 Of the resulting maps, 1296 (5.7%) have been further refined through an expert review process [@kaschner_mapping_2006; @ready_predicting_2010].
+
+<!--In addition to these methodological differences, the two datasets also differ in intent.
+IUCN maps essentially communicate the extent of occurrence (EOO) while AquaMaps uses suitable habitat as an indicator of probability of occurrence, -->
 
@@
 Area ratio provides a proxy for frequency of commission and/or omission errors.
+[range area is one critical feature/use of IUCN used as a criterion to help define extinction risk; can AquaMaps provide useful information here as well?]
 

### 2016-09-03 ms_3_res_disc_concl.Rmd

@@
 
-<!-- need discussion here? [discussion] Each dataset offers spatial information on a -->
+[ideas for discussion here?
+each provides information unavailable in the other; in addition to considerations about the nature of each dataset, the species and regions covered are an important consideration in which dataset to use]
 
@@
 Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
-It is not surprising that range maps based upon fewer observations bear greater uncertainty.
+In addition, the AquaMaps dataset offers a quality metric based on the number of unique cells containing valid occurrences; for this quadrant, the median "occurcells" is 11 compared to a median of 40 across the other three quadrants.
+Care should be taken when using distribution and range maps based upon fewer observations, as they clearly bear greater uncertainty; AquaMaps explicitly recommends using any of its maps generated with an "occurcells" count fewer than 10 [@AquaMapsXXX].
 
@@
 
-[discussion] We cannot know for certain the true distribution of each of these corals; much of the apparent over-expansion of coral ranges may be due to experts taking a precautionary approach.
-But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported ranges.
-Note that this change applies just as well to the IUCN coral maps that are not included in the paired map analysis, and likely to other reef-associated flora and fauna.
+[discussion] We cannot know for certain the true distribution of each of these corals; certainly some commission errors result from satisfying the connectivity requirements of an EOO, and others may be due to experts taking a precautionary approach.
+But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported distributions for most purposes.
+Note that this change applies just as readily to the IUCN coral maps that are not included in the paired map analysis, and likely to other reef-associated flora and fauna.
 While species depth preferences are an easy and consistent means of constraining range predictions, other conditions such as salinity and temperature could be cautiously used to refine the results of expert opinions.
@@
 
-[discussion] FAO Major Fishing Area boundaries provide a readily available method to roughly contrain AquaMaps predictions to appropriate ocean basins, enabling rapid modeling of thousands of species ranges.
-However, these boundaries are defined for statistical purposes based on economic and political considerations rather than ecological considerations, and can result in odd species range predictions where otherwise suitable habitat encounters border between regions.
-While such an odd boundary would likely be obvious when inspecting ranges of individual species, the distinction is likely to be obscured when many species ranges are aggregated as is typical for biodiversity or conservation studies.
-Applying an additional spatial constraint based upon ecological or environmental boundaries, such as Marine Ecoregions of the World [@REF???], may help boost confidence in these predicted ranges; expert review, though time-consuming, is certain to improve confidence in these predicted ranges.
+[discussion] FAO Major Fishing Area boundaries provide a readily available method to roughly contrain AquaMaps predictions to appropriate ocean basins, thus eliminating a large source of potential commission errors and enabling rapid modeling of thousands of species ranges.
+However, these boundaries are defined for statistical purposes based on economic and political considerations rather than ecological considerations, and can result in odd species range predictions where otherwise suitable habitat is omitted where it encounters border between regions.
+While such an odd boundary would likely be obvious when inspecting the distribution of an individual species, the distinction is likely to be obscured when aggregating many species ranges as is typical for biodiversity or conservation studies.
+Applying an ecoregional spatial constraint, such as Marine Ecoregions of the World [@REF???], may reduce commission errors on these questionable margins, boosting confidence in these predicted distributions; expert review, though time-consuming, is the most certain route to improving the quality of these maps.
 
@@
 
+[can this wrap up the results/discussion?] To achieve more comprehensive global coverage of species ranges these two datasets can be used together, understanding that the underlying differences complicate such direct comparisons.
+Major considerations include aligning the criteria used to determine "presence" (e.g.
+selecting an appropriate presence threshold for AquaMaps; _a presence threshold of 0% most closely approximates the "limits of distribution" criterion described by IUCN range maps_ ___- not really discussed or backed up here___) and addressing mechanisms that introduce errors of commission and omission.
+___as it stands, this paper really doesn't go into combining datasets and how to do it; omit this paragraph?___
+
 # Conclusions
@@
 
-For IUCN range data, clipping ranges to known depth limits improves confidence in predicted ranges for coral species; similarly, depth may be a useful characteristic to consider for other reef-associated organisms.
-Other parameters such as salinity and temperature may provide useful constraints on other taxa to bolster and refine expert opinion.
-Simple but conservative rules of thumb will likely reduce commission errors without introducing substantial omission errors.
-
-For AquaMaps range data, dependent primarily on environmental and physical preferences and conditions, implementing area restrictions based on biogeographical criteria such as Marine Ecoregions of the World [@spalding_marine_2007], rather than political and economic criteria such as FAO Major Fishing Areas, would likely decrease commission errors and improve predictive power, especially for data-poor species.
-
-For either data set, maps based on few occurrences are more likely to bear high uncertainty in range predictions.
-Occurrence counts from external sources such as OBIS or GBIF can help identify relatively data-poor species; additionally, the "occurcells" attribute in the AquaMaps data set, which counts the number of half-degree cells used to generate the environmental envelope for each species, can be used in a similar manner.
-The only certain remedy for data-poor species, of course, is further research.
-
-<!-- ___OMIT:___ _Many studies using AquaMaps data test the sensitivity of results by varying the probability of occurrence threshold to determine presence.
-This decision ultimately represents a tradeoff between errors of commission (low threshold) and omission (high threshold).
-Using AquaMaps in conjunction with IUCN can mitigate potential errors, while also increasing the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized.
-In this case, a presence threshold of 0% most closely approximates the "limits of distribution" criterion defined by IUCN data providers._ -->
-
-To achieve more comprehensive global coverage of species ranges these two datasets can be used together, understanding that the underlying differences complicate such direct comparisons.
-Major considerations include aligning the criteria used to determine "presence" (e.g.
-selecting an appropriate presence threshold for AquaMaps; _a presence threshold of 0% most closely approximates the "limits of distribution" criterion described by IUCN range maps_ ___- not really discussed or backed up here___) and addressing mechanisms that introduce errors of commission and omission.
-___as it stands, this paper really doesn't go into combining datasets and how to do it; omit this paragraph?___
-
 Effective management and protection of marine species depends on a robust understanding of where species exist and where they do not; without this knowledge we risk wasting resources protecting low-value regions while missing opportunities to protect critical ones.
-By identifying the differences between these two fundamental marine species range datasets and understanding the likely mechanisms causing these discrepancies, we improve our ability to develop strategic and effective conservation policy that supports a resilient ocean ecosystem.
+By identifying the differences between these two important and fundamentally different marine species range datasets and understanding the likely mechanisms causing these discrepancies, we improve our ability to develop strategic and effective conservation policy that supports a resilient ocean ecosystem.

### 2016-09-03 clip_depth/coral_maps.Rmd
dammit - need to redo central limit calcs

@@
 
+loiczid_raster_file  <- file.path(dir_git, 'shiny_am_iucn/data/loiczid_raster.tif')
+loiczid_raster       <- raster(loiczid_raster_file)
+names(loiczid_raster) <- 'loiczid'
+
 coral_cells_iucn <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv'))
@@
 coral_comp <- coral_cells_iucn %>%
-  select(-iucn_sid) %>%
+  dplyr::select(-iucn_sid) %>%
   mutate(deep = deep * 200) %>%
-  distinct()
+  distinct() %>%
+  as.data.frame()
 
-rast_coral_comp <- subs(x = loiczid_raster, y = coral_comp %>% select(loiczid, deep), 
+rast_coral_comp <- raster::subs(x = loiczid_raster, y = coral_comp %>% select(loiczid, deep), 
                      by = 'loiczid', which = 'deep', 
@@
   tm_layout(basemaps = "Esri.WorldTopoMap", 
-            legend.text.size = .35,
-            legend.title.size = .45,
+            legend.text.size = .5,
+            legend.title.size = .6,
             # title.position = 'TOP', 

### 2016-09-03 clip_fao/fao_clip_pac.Rmd
dammit - need to redo central limit calcs

@@
 ``` {r central_band}
-### 25 S to 5 N  - look for spikes at 120W (240E), 175W (185E), 80E (and 77E)
+### 25 S to 20 N  - look for spikes at 120W (240E), 175W (185E), 80E (and 77E)
 am_spp_central <- am_spp_latlong %>%
-  filter(centerlat <= 5 & centerlat >= -25)
+  filter(centerlat <= 20 & centerlat >= -25)
 
@@
 
-spp_cells_clipped <- read_csv(file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv'))
+spp_cells_clipped <- read_csv(file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv')) %>%
+  as.data.frame()
 
@@
 
+### boundary line
+fao_boundary <- Line(matrix(c(185, 185, -25, 5), ncol = 2)) %>%
+  Lines(ID = 'a') %>%
+  list() %>%
+  SpatialLines()
+
 rast_spp_map_e <- 
@@
             fontfamily = 'Helvetica',
-            size = .35) +
+            size = .5) +
+  tm_shape(fao_boundary) +
+    tm_lines(col = 'blue', lwd = 1.5) +
   tm_layout(# basemaps = "Esri.WorldTopoMap", 
             frame = TRUE,
-            legend.text.size = .35,
-            legend.title.size = .45,
+            legend.text.size = .5,
+            legend.title.size = .6,
             # title.position = 'TOP', 
@@
             legend.bg.color = 'white',
-            legend.bg.alpha = .8,
+            legend.bg.alpha = .9,
             attr.outside = TRUE)
@@
 save_tmap(rast_spp_map_e, 
-          file.path(dir_fig, 'map_am_edge_map.png'), 
+          file.path(dir_fig, 'map_am_edge.png'), 
           width = 19, 
@@
   group_by(loiczid) %>%
-  summarize(n_spp = n())
+  summarize(n_spp = n()) %>%
+  as.data.frame()
 

### 2016-09-03 clip_fao/fao_clip_pac.Rmd
updated figures

@@
 source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
-### When knitting this, it automatically sets WD to be this directory...
-### the 'setwd()' is there for running by chunk
+
 ```
@@
          wlimitadj = ifelse(wlimit < 0, wlimit + 360, wlimit),
-         elimitadj = ifelse(elimit < 0, elimit + 360, elimit)) %>%
-  filter(clongadj >= 60 & clongadj <= 240)
+         elimitadj = ifelse(elimit < 0, elimit + 360, elimit))
 
-### filter out species from Atlantic... rough pass:
-### - filter out everything west of 30 E and east of 70 W
 message('binding hcaf to am_spp_latlong')
@@
 
-``` {r northmost_band, eval = FALSE}
-### 5N north     - look for spike at 175 W (185E)
-am_spp_north <- am_spp_latlong %>%
-  filter(centerlat >= 5)
+``` {r atlantic_spp} 
+### identify species that appear in the mid-atlantic - focus just on pacific/indopacific spp only
 
-am_spp_north_lims <- am_spp_north %>%
-  group_by(am_sid) %>%
-  summarize(w_lim = min(wlimitadj),
-            e_lim = max(elimitadj),
-            n_w_lim = sum(wlimitadj == w_lim),
-            n_e_lim = sum(elimitadj == e_lim),
-            n_tot = n()) %>%
-  gather(lim, long, w_lim:e_lim) %>%
-  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
-  filter(str_detect(lim_n, lim)) %>%
-  select(-lim_n) %>%
-  filter(!(long <= 70 | long >= 200))
-    ### delete artificially imposed limits
+hcaf_atl <- hcaf %>%
+  filter(centerlong >= -60 & centerlong <= 10)
 
-tmp <- am_spp_north_lims %>%
-  group_by(long) %>%
-  summarize(n_cells = sum(ncells),
-            n_spp   = n())
+am_spp_atl <- hcaf_atl %>%
+  left_join(am_spp_cells, by = 'loiczid')
 
-ggplot(tmp, aes(x = long, y = n_spp)) +
-  geom_point() +
-  geom_vline(xintercept = c(77, 185), color = 'red')
-
+spp_atl <- am_spp_atl %>%
+  .$am_sid %>%
+  unique()
 ```
@@
 ``` {r central_band}
-### 25 S to 20 N  - look for spikes at 120W (240E), 175W (185E), 80E (and 77E)
+### 25 S to 20 N  - look for spikes at 120W (240E), 80E (and 77E), and especially 175W (185E)
 am_spp_central <- am_spp_latlong %>%
@@
 write_csv(am_spp_central_lims, file.path(dir_git, 'clip_fao/am_spp_central_lims.csv'))
-
+### NOTE: these still include spp that appear in Atlantic, so global -
+### "east limit" and "west limit" are not exactly accurate
 tmp <- am_spp_central_lims %>%
@@
 ``` {r what_species_are_these}
-
+### here, the `filter(!am_sid %in% spp_atl)` line eliminates species with a border
+### on 185E who also appear in the Atlantic... global species have no eastern/
+### western border...
 clipped_spp <- am_spp_central_lims %>%
-  # bind_rows(am_spp_north_lims) %>%
-  # filter(long %in% c(77, 185)) %>%
+  filter(!am_sid %in% spp_atl) %>%
   filter(long == 185) %>%
@@
 
-loiczid_raster <- raster('data/rasters/loiczid_raster.tif')
+loiczid_raster <- raster(file.path(dir_git, 'data/rasters/loiczid_raster.tif'))
+
 fao_polys <- rgdal::readOGR(dsn = file.path(path.expand(dir_git), 'clip_fao/fao_rgns'),
@@
 
-assemble_map <- function(map_rast, spp) {
-  message('...mapping ', spp)
-  map_obj <- tm_shape(map_rast) +
-    tm_raster(palette = 'Reds',
-              colorNA = NULL,
-              title = spp,
-              alpha = 1) +
-    tm_shape(World1) +
-      tm_polygons() + 
-    tm_layout(basemaps = "Esri.WorldTopoMap", 
-              # title.position = 'TOP', 
-              legend.outside = TRUE, attr.outside = TRUE)
-
-  return(map_obj)
-}
 ```
@@
 
-# spp_cells_clipped <- am_spp_cells %>%
-#   filter(am_sid %in% spp_examples_e$am_sid) %>%
-#   group_by(loiczid) %>%
-#   summarize(n_spp = n())
-# write_csv(spp_cells_clipped, file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv'))
+spp_cells_clipped <- am_spp_cells %>%
+  filter(am_sid %in% spp_examples_e$am_sid) %>%
+  group_by(loiczid) %>%
+  summarize(n_spp = n())
+write_csv(spp_cells_clipped, file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv'))
 
-spp_cells_clipped <- read_csv(file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv')) %>%
-  as.data.frame()
+# spp_cells_clipped <- read_csv(file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv')) %>%
+#   as.data.frame()
 
@@
 ### boundary line
-fao_boundary <- Line(matrix(c(185, 185, -25, 5), ncol = 2)) %>%
+fao_boundary <- Line(matrix(c(185, 185, -25, 20), ncol = 2)) %>%
   Lines(ID = 'a') %>%
@@
     tm_raster(palette = 'Reds',
-              breaks = c(1, 15, 100, 200, 300, Inf),
+              breaks = c(1, 15, 100, 150, 200, Inf),
               colorNA = 'lightblue1',
@@
     tm_text('F_AREA_1',
-            col = 'orangered4',
+            col = 'royalblue4',
             fontface = 'bold.italic',
@@
             legend.bg.color = 'white',
-            legend.bg.alpha = .9,
+            legend.bg.alpha = 1,
             attr.outside = TRUE)
@@
 ### group affected spp by loiczid and sum
+
 spp_cells_clipped <- am_spp_cells %>%
@@
   group_by(loiczid) %>%
-  summarize(n_spp = n()) %>%
-  as.data.frame()
+  summarize(n_spp = n())
+write_csv(spp_cells_clipped, file.path(dir_git, 'clip_fao', 'spp_cells_clipped_w.csv'))
+
+# spp_cells_clipped <- read_csv(file.path(dir_git, 'clip_fao', 'spp_cells_clipped_w.csv')) %>%
+#   as.data.frame()
 
@@
 # values(rast_spp_w)[is.na(values(rast_spp_w))] <- 0
-rast_spp_w <- crop(rast_spp_w, extent(c(85, 220, -59, 50)))
+rast_spp_w  <- crop(rast_spp_w, rast_ext)
+
+### boundary line
+fao_boundary <- Line(matrix(c(185, 185, -25, 20), ncol = 2)) %>%
+  Lines(ID = 'a') %>%
+  list() %>%
+  SpatialLines()
 
@@
     tm_raster(palette = 'Reds',
-              breaks = c(1, 50, 100, 500, 1000, 2000, Inf),
+              breaks = c(1, 15, 100, 500, 1000, 2000, Inf),
               colorNA = 'lightblue1',
@@
     tm_borders(col = 'grey40', lwd = 0.25) +
-    tm_text('F_CODE_1',
-            col = 'orangered4',
+    tm_text('F_AREA_1',
+            col = 'royalblue4',
             fontface = 'bold.italic',
             fontfamily = 'Helvetica',
-            size = .35) +
-  tm_layout(basemaps = "Esri.WorldTopoMap", 
-            frame = FALSE,
-            legend.text.size = .35,
-            legend.title.size = .45,
+            size = .5) +
+  tm_shape(fao_boundary) +
+    tm_lines(col = 'blue', lwd = 1.5) +
+  tm_layout(# basemaps = "Esri.WorldTopoMap", 
+            frame = TRUE,
+            legend.text.size = .5,
+            legend.title.size = .6,
             # title.position = 'TOP', 
-            legend.outside = TRUE,
-            # legend.position = c('right', 'bottom'),
-            # legend.bg.color = 'white',
-            # legend.bg.alpha = .5,
+            legend.outside = FALSE,
+            legend.position = c('right', 'top'),
+            legend.bg.color = 'white',
+            legend.bg.alpha = 1,
             attr.outside = TRUE)
@@
 
-save_tmap(rast_spp_map_w, file.path(dir_fig, 'map_am_edge_w.png'), 
-          width = 9, height = 6, units = 'cm', dpi = 600)
+save_tmap(rast_spp_map_w, 
+          file.path(dir_fig, 'map_am_edge_w.png'), 
+          width = 19, 
+          # height = 6, 
+          units = 'cm', dpi = 600)
 
@@
     tm_polygons() + 
-  tm_shape(fao_polys_e) +
+  tm_shape(fao_polys) +
     tm_borders() +

### 2016-09-03 ms_2_methods.Rmd
updated figures

@@
 
-AquaMaps' reliance upon FAO zones to constrain species ranges seems likely to result in odd species range boundaries where ecologically suitable habitat intersects with human-defined borders.
+AquaMaps' reliance upon FAO zones to constrain species ranges seems likely to result in abrupt species range boundaries where ecologically suitable habitat intersects with human-defined borders.
 To explore the impact of these artificial constraints on species range predictions, we chose to focus on boundaries defined by longitude, as they seem less likely to coincide with ecological differences than boundaries defined by latitude.
-As an example, we identified all AquaMaps species whose eastern or western range limit between between 5° S and 25° N latitudes coincided with the large vertical boundary between FAO regions 71 and 77 at 175° W longitude.
+As an example, we identified all AquaMaps species whose eastern or western range limit between between 25° S and 20° N latitudes coincided with the large vertical boundary between FAO regions 71 and 77 at 175° W longitude.
 

### 2016-09-03 ms_3_res_disc_concl.Rmd
updated figures

@@
 
-[results] We identified 4,168 species whose equatorial ranges (between 5° S and 25° N) encounter a range limit at 175° W.
-This longitudinal boundary created a western limit for 512 species, and eastern limit for 3,656 species.
-Plotting the entire ranges of the 512 west-limited species (Fig 4) results in a very clear exclusion zone that matches perfectly with FAO region 71; while we did not focus on the boundary for region 57, a similar effect is readily apparent.
+[results] From the entirety of the AquaMaps dataset, we identified 3,577 Indo-Pacific species whose equatorial distributions (between 25° S and 20° N) encounter a range limit at 175° W.
+This longitudinal boundary created a western limit for 369 species, and an eastern limit for 3,208 species.
+A clear discontinuity in species distributions of the 3,208 west-limited species (Fig 4) matches perfectly with FAO region 77; other discontinuities are apparent at other FAO boundaries, despite these boundaries not being actively studied in the analysis.
+Similar discontinuities were clear when mapping the west-limited species.
 
-> __Fig 4.__ Composite map showing AquaMaps predicted ranges for 512 species whose equatorial range (between 5° S and 25° N) encounters a western limit exactly at 175° W, the boundary between FAO Major Fishing Areas 71 and 77.
-The range maps for these species conform exactly to the bounds of Area 71, as a result of AquaMaps using FAO areas as a rough georegional constraint on species ranges.
-The boundary of Area 57 creates an additional clear exclusion zone for most of these species, despite the fact that this was not used as a selection criteria.
+> __Fig 4.__ Composite map showing AquaMaps predicted ranges for 3,208 species whose equatorial distribution encounters an eastern discontinuity exactly at 175° W, the boundary between FAO Major Fishing Areas 71 and 77 (shown in blue).
+Other FAO area boundaries create additional clear discontinuities.
 
 [discussion] FAO Major Fishing Area boundaries provide a readily available method to roughly contrain AquaMaps predictions to appropriate ocean basins, thus eliminating a large source of potential commission errors and enabling rapid modeling of thousands of species ranges.
-However, these boundaries are defined for statistical purposes based on economic and political considerations rather than ecological considerations, and can result in odd species range predictions where otherwise suitable habitat is omitted where it encounters border between regions.
-While such an odd boundary would likely be obvious when inspecting the distribution of an individual species, the distinction is likely to be obscured when aggregating many species ranges as is typical for biodiversity or conservation studies.
-Applying an ecoregional spatial constraint, such as Marine Ecoregions of the World [@REF???], may reduce commission errors on these questionable margins, boosting confidence in these predicted distributions; expert review, though time-consuming, is the most certain route to improving the quality of these maps.
+However, these boundaries are defined for statistical purposes based on economic and political considerations rather than ecological considerations, and can result in odd discontinuities in species range predictions where otherwise suitable habitat is excluded.
+While such a discontinuous boundary would likely be obvious when inspecting the distribution of an individual species, the distinction is likely to be obscured when aggregating many species ranges as is typical for biodiversity or conservation studies.
+Ecoregional spatial zones such as Marine Ecoregions of the World [@REF???] may provide an additional useful constraint on species ranges; expert review, though time-consuming, is the most certain route to boosting confidence in these predicted distributions.
 

### 2016-09-03 ms_5_figures.Rmd
updated figures

@@
 
+___NOTE:___ These images are all based on .png outputs, mostly from .psd files; submit the .tif versions for publication.
+
 ### Fig. 1
@@
 
-![](figures/bar_by_taxa_all.png)
-![](figures/all_aquamaps_map.png)
-![](figures/all_iucn_map.png)
+![](figures/Fig_1.png)
 
@@
 
-![](figures/scatter_quadmap.png)
-![](figures/barchart_spp_gp_quads.png)
+![](figures/Fig_2.png)
 
@@
 
-![coral area change due to depth clipping](figures/coral_map.png)
-
-![coral quadmap breakdown](figures/barchart_coral_quads.png)
+![coral area subject to depth clipping](figures/Fig_3.png)
 
@@
 
-![aquamaps edge comparison](figures/am_edge_map.png)
+![aquamaps edge comparison](figures/Fig_4.png)
 
@@
 
-![Change in MPA coverage](figures/barchart_mpa_gap.png)
+![Change in MPA coverage](figures/Fig_5.png)
 

### 2016-09-04 ms_2_methods.Rmd
updated figures incl pngs, reworked some of the results/discussion based on Halley's suggestions

@@
 __Alignment framework for paired maps__: OR __Comparison of paired maps__:
-Using genus and species binomials as a matching key, we identified "paired map" species, i.e.
-the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.
+Species range maps communicate a prediction of both the distribution of a species and the size of a species range.
+To enable direct comparison of these measures between the two datasets, we identified "paired map" species, i.e.
+the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution, using genus and species binomials as a matching key.
 We used the taxize package [@REF] in R [@REF] to standardize species names and synonyms; for species with separate subpopulation maps in IUCN, we combined all subpopulations to create a single global population.
@@
 $A_{small \cap large}$ represents the amount of overlapping area between the two datasets.
-Distribution alignment uses overlapping predictions of presence as means of identifying areas of higher confidence.
-Area ratio provides a proxy for frequency of commission and/or omission errors.
-[range area is one critical feature/use of IUCN used as a criterion to help define extinction risk; can AquaMaps provide useful information here as well?]
+Distribution alignment uses overlapping predictions of presence as a measure of concurrence between the two datasets.
+Area ratio compares range size, used by IUCN as a criterion to help define extinction risk; it also provides a proxy for frequency of commission and/or omission errors.
 

### 2016-09-09 ms_1_intro.Rmd
asdf

@@
 
-The fundamental differences between the datasets (EOO versus AOO, measure of range size versus distribution) suggest that the choice of one over the other should be carefully matched to the purpose for which it is to be used [@GastonXXX 2008, XXX Other refs], and yet these datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
+_paragraph describing IUCN and AM as EOO and AOO respectively, with basic defs and implication that AM will generally be smaller than IUCN; possibly refer to Table S1 for defs and details?_ _one reference: http://www.biodiversitya-z.org/content/species-extent-of-occurrence-eoo "The intent behind this parameter is to measure the degree to which risks from threatening factors are spread spatially across the taxon’s geographical distribution.
+It is not intended to be an estimate of the amount of occupied or potential habitat, or a general measure of the taxon’s range" and http://www.biodiversitya-z.org/content/species-area-of-occupancy-aoo "The measure reflects the fact that a taxon will not usually occur throughout the area of its extent of occurrence, which may, for example, contain unsuitable habitats"_
+
+The fundamental differences between the datasets suggest that the choice of one over the other should be carefully matched to the purpose for which it is to be used [@GastonXXX 2008, XXX Other refs], and yet these datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
 In most of these cases, the implications of choosing one versus the other of these datasets is not evaluated or discussed, yet strong conservation and management conclusions are drawn from the results of the studies.
@@
 Here we focus not on how the two datasets should ideally be used; instead, in recognizing that each provides value for many aspects of conservation research, we highlight hidden assumptions and sources of error within each dataset, and explore the implications of choosing one over the other.
+_You science with the data you have, not with the data you wish you had.
+Often, the ideal type of spatial data is not always available, and non-ideal data is better than no data at all.
+In such a case, it is crucial to acknowledge, and where possible address, the resulting tradeoffs._
 
@@
 
+___maybe here, based on EOO/AOO defs, we expect AM to generally fall within, and describe a smaller range than, IUCN; where the two disagree, that's where it gets interesting___
+
 Using this alignment framework, we explore methodological issues that can introduce errors, and suggest methods to improve confidence in species range predictions.
@@
 As a case study, we then reexamine a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] as a case study to explore the implications of selecting one data set over the other.
-Repeating the study using IUCN range data rather than AquaMaps data increases the apparent proportion of gap species from 1.4% to 6.4%, while also increasing the proportion of well-protected species from 1.5% to 2.9%.
+Repeating the study using IUCN range data rather than AquaMaps data increases the apparent proportion of gap species from 1.4% to 6.4%, _while also increasing the proportion of well-protected species from 1.5% to 2.9%_ ___verify this second part!!!___.
 This shift highlights the implications of different data use decisions on our understanding of marine biodiversity status and protection.

### 2016-09-09 ms_3_res_disc_concl.Rmd

@@
 # Results and Discussion
+
+[brief overview of results/discussion?]
 
@@
 
-## Distribution and range area alignment
+## Distribution and range size alignment
 
-[discussion] To explore differences in species distribution and range between the two datasets, we assessed the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) relative to the area ratio (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig 2A).
+[discussion] [results] Comparing distribution alignment and area ratio for the `r format(n_spp_both, big.mark = ',')` paired map species (Fig 2A), a weak negative linear pattern appears to emerge, suggesting that increasing similarity in range area correlates very slightly with decreasing distribution alignment (R^2 = .016).
+The pattern itself is unimportant, and emerges simply due to the nature of the analysis and the datasets:  AquaMaps tends to extrapolate species ranges into suitable areas beyond known occurrences more frequently than IUCN maps, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
+For species with dissimilar range areas, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
+For species with increasingly similar range areas, differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.
+
+The mean distribution alignment for species included in both datasets was `r dist_align_mean`%; the mean area alignment was `r area_align_mean`%.
+By dividing the paired map species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
+Representative maps from each category are provided in the supporting materials (Fig S1).
+
 Where two independent methods predict presence (or absence) of a species in the same location, we can assume a higher confidence in the accuracy of that prediction.
 Where one method predicts presence while the other predicts absence, we can investigate potential sources of commission and omission errors to refine our understanding of the species' range.
-Understanding these sources of error can provide guidance to improve predictions of species range even when only one source of spatial data is available.
+Understanding and addressing these sources of error provides guidance to improve predictions of species range even when only one source of spatial data is available.
 
@@
 (B) Alignment quadrant breakdown of species by taxonomic group.
-
-[discussion] This analysis revealed a weak negative linear pattern, suggesting that increasing similarity in range area correlates with decreasing distribution alignment.
- AquaMaps tends to extrapolate species ranges into suitable areas beyond known occurrences, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
-For species with dissimilar range areas, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
-For species with increasingly similar range areas, differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.
-
-[results] The mean distribution alignment for species included in both datasets was `r dist_align_mean`%; the mean area alignment was `r area_align_mean`%.
-By dividing the paired map species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
-Representative maps from each category are provided in the supporting materials (Fig S1).
 
@@
 Corals offer an excellent case study, due to their foundational role in biodiverse habitats, as well as their lack of mobility and reliance on photosynthesis.
-Ocean depth preference is explicitly included in AquaMaps models.
+Ocean depth preference is explicitly included as a parameter in AquaMaps models.
 While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." [@REF http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources]), it is not presented as a requirement, so we cannot take its inclusion for granted.

### 2016-09-09 ms_5_figures.Rmd

@@
 
-___NOTE:___ These images are all based on .png outputs, mostly from .psd files; submit the .tif versions for publication.
+___NOTE:___ These images are all based on .png outputs, exported from .psd files; submit the .tif versions for publication.
 
@@
 
-![](figures/Fig_1.png)
+![Taxonomic and geographic representation](figures/Fig_1.png)
 
@@
 
-![](figures/Fig_2.png)
+![Quadrant plot and spp breakdown by quadrant](figures/Fig_2.png)
 

### 2016-09-09 ms_6_supplemental.Rmd

@@
 AM raw -> presence
-bathymetry poly -> raster?
 
@@
 __Fig S3.__ Representative species maps to illustrate each quadrant from Fig 2A.  Each map is positioned to match its quadrant in Fig 2A.
-(A) Distribution-matched: Conus episcopatus, the dignified cone snail.  Ranges shows excellent overlap in the western Pacific, though IUCN range extends beyond the bounds of the AquaMaps range.
-(B) Well-matched: Kajikia albida, the Atlantic white marlin.  Ranges from each data set show nearly complete overlap, and very similar area of coverage.
+(A) Distribution-matched: Conus episcopatus, the dignified cone snail.
+Distributions shows excellent overlap in the western Pacific, though IUCN range extends well beyond the bounds of the AquaMaps range.
+(B) Well-matched: Kajikia albida, the Atlantic white marlin.  Distributions from each data set show nearly complete overlap, and very similar range size.
 (C) Poorly-matched: Acanthurus nigroris, the blue-lined surgeonfish.
-IUCN predicts species range only near the Hawaiian islands; AquaMaps predicts extensive range throughout the central and western Pacific Ocean.
-The datasets align in neither distribution nor area of coverage.
-(D) Area-matched: Conus magnificus, the magnificent cone snail.  Ranges overlap in the southern Pacific, but align poorly elsewhere.  The total area for each range is similar.
+IUCN predicts species distribution only near the Hawaiian islands; AquaMaps predicts extensive distribution throughout the central and western Pacific Ocean.
+The datasets align in neither distribution nor range size.
+(D) Area-matched: Conus magnificus, the magnificent cone snail.  Distributions overlap in the southern Pacific, but align poorly elsewhere.  The range sizes are similar.
 

### 2016-09-11 case_study_mpa/3_mpa_spp_analysis.Rmd
all text updated, including references.  Still need to complete abstract.

@@
   mutate(cat = factor(cat, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
-  mutate(trial = factor(trial, levels = c('IUCN 2015', 'AM 2015 (0%) +\nIUCN 2015', 'AM 2015 (0%)', 'AM 2015 (50%)', 'AM 2014 (50%)'))) %>%
+  filter(trial != 'AM 2015 (0%) +\nIUCN 2015') %>%
+  mutate(trial = factor(trial, levels = c('IUCN 2015', 
+                                          # 'AM 2015 (0%) +\nIUCN 2015', 
+                                          'AM 2015 (0%)', 'AM 2015 (50%)', 'AM 2014 (50%)'))) %>%
   mutate(p_cat = p_cat * 100) %>%
@@
             color = 'grey30') +
-  scale_x_discrete(labels = c('5. IUCN 2015',
-                              '4. AM 2015 (0%) +\nIUCN 2015', 
+  scale_x_discrete(labels = c('4. IUCN 2015',
+                              # '4. AM 2015 (0%) +\nIUCN 2015', 
                               '3. AM 2015 (0%)', 

### 2016-09-11 ms_1_intro.Rmd
all text updated, including references.  Still need to complete abstract.

--- 2016-09-02 ms_1_intro.Rmd
@@
 
-The two most comprehensive, widely-used global-scale repositories that predict marine species ranges throughout the world's oceans are AquaMaps, which rely primarily on model predictions [@kaschner_aquamaps:_2015], and range data from the International Union for Conservation of Nature (IUCN), which rely primarily on expert opinion [@iucn_iucn_2015].
-The two datasets are designed to serve different purposes:  IUCN range maps are drawn to estimate extent of occurrence (EOO) [@GastonXXX 1994], communicating range size as an [indicator/factor/criterion] for extinction risk assessments [@IUCNXXX]; AquaMaps essentially communicate the area of occupancy (AOO) [@REFXXX] of a species based on habitat suitability [@AquaMapsXXX].
+The two most comprehensive, widely-used global-scale repositories that predict marine species ranges throughout the world's oceans are AquaMaps, which rely primarily on model predictions to communicate the distribution of a species based on habitat suitability [@kaschner_aquamaps:_2015], and range data from the International Union for Conservation of Nature (IUCN), which rely primarily on expert opinion to communicate range size as a criterion for extinction risk assessments [@iucn_iucn_2015].
 
-The fundamental differences between the datasets (EOO versus AOO, measure of range size versus distribution) suggest that the choice of one over the other should be carefully matched to the purpose for which it is to be used [@GastonXXX 2008, XXX Other refs], and yet these datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
+While the two datasets ostensibly describe the same information, i.e.
+where can a particular species be found, they communicate fundamentally distinct concepts.
+The IUCN range map for a given species describes the "extent of occurrence" for that species, a contiguous region that encloses all known occurrences and connecting regions, with the explicit caveat that this "does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon" [@iucn_iucn_2015].
+The AquaMaps map for a species, on the other hand, seems to better be defined as "area of occupancy": the subset of the extent of occurrence in which the species is actually likely to be found, for example by excluding unsuitable habitats from the extent of occurrence [@iucn_iucn_2015].
+The distinctions between extent of occurrence and area of occupancy, including methodology, limitations, and appropriate usage, are well documented [@rondinini_tradeoffs_2006; @gaston_sizes_2009].
+___include reference to a supplemental table for more details?___
+
+<!--
+_paragraph describing IUCN and AM as extent of occurrence and area of occupancy respectively, with basic defs and implication that AM will generally be smaller than IUCN; possibly refer to Table S1 for defs and details?_ _one reference: http://www.biodiversitya-z.org/content/species-extent-of-occurrence-eoo "The intent behind this parameter is to measure the degree to which risks from threatening factors are spread spatially across the taxon’s geographical distribution.
+It is not intended to be an estimate of the amount of occupied or potential habitat, or a general measure of the taxon’s range" and http://www.biodiversitya-z.org/content/species-area-of-occupancy-aoo "The measure reflects the fact that a taxon will not usually occur throughout the area of its extent of occurrence, which may, for example, contain unsuitable habitats"_
+-->
+
+The fundamental differences between the datasets suggest that the choice of one over the other should be carefully matched to the purpose for which it is to be used [@gaston_sizes_2009], and yet these datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
 In most of these cases, the implications of choosing one versus the other of these datasets is not evaluated or discussed, yet strong conservation and management conclusions are drawn from the results of the studies.
 
-<!--
-_The fundamental differences between the intent of the two datasets precludes casual interchangeability between the two [@GastonXXX 2008], and yet they_
+Here we focus not on how the two datasets should ideally be used; instead, recognizing that each provides distinct value for conservation research, we explore hidden assumptions and sources of error within each dataset, and test the implications of choosing one over the other.
+You do science with the data you have, not with the data you wish you had; often, the ideal type of spatial data is not available, and non-ideal data is better than no data at all.
+In such a case, it is crucial to acknowledge, and where possible address, the resulting tradeoffs.
 
-_"The two measures (EOO and AOO) serve different purposes, and should not be regarded as alternatives that simply differ in accuracy" - Gaston et al 2004 - literature suggests that this distinction precludes casual interchangeability between the two [@gastonXXX 2008]; they have specific purposes.
-And yet they get used, with no explicit acknowldegment of the difference, for a wide range of conservation studies.
-It's important to understand the implications and measures to help adapt them to specific uses._
--->
+___include as table, in body or in supplemental?
+It really kind of doesn't fit here any more___ _Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: extent of occurrence range data such as IUCN range maps will inevitably introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps are expected to introduce fewer commission errors at the expense of more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
+Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006; @jetz_ecological_2008]._
 
-Here we focus not on how the two datasets should ideally be used; instead, in recognizing that each provides value for many aspects of conservation research, we highlight hidden assumptions and sources of error within each dataset, and explore the implications of choosing one over the other.
-
-Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: extent of occurrence range data such as IUCN range maps will inevitably introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps are expected to introduce fewer commission errors at the expense of more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
-Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006; @jetz_ecological_2008].
-
-To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of the `r format(n_spp_all, big.mark = ',')` marine species mapped by one or both datasets.
+Data availability is a crucial factor in deciding which dataset to use.
+We compare how each data source represents the global spatial and taxonomic distribution of the `r format(n_spp_all, big.mark = ',')` marine species mapped by one or both datasets.
 Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).
@@
 Although relatively small in number, these overlap species present a unique opportunity to evaluate the two datasets overall.
-As understanding of a given species improves, we expect that range maps, regardless of method, would more closely approach the true distribution in predicting species presence and absence.
-Close alignment between two independent range maps can therefore indicate higher confidence in species presence or absence than a single map.
 For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align in both spatial distribution and overall area.
+Based upon the definitions of extent of occurrence and area of occupancy, we expect that for a given species, the AquaMaps predicted distribution will fall within, and describe a smaller range than, the IUCN predicted distribution.
+Where these expectations seem to fail, we explore methodological issues that can introduce errors.
+In particular, we show how explicitly using depth as a constraint on coral ranges in IUCN range data dramatically mitigates a source of commission error, and highlight a side effect of AquaMaps methodology that results in discontinuous species range boundaries.
 
-Using this alignment framework, we explore methodological issues that can introduce errors, and suggest methods to improve confidence in species range predictions.
-In particular, we show how explicitly using depth as a constraint on coral ranges in IUCN range data dramatically mitigates a source of commission error, and highlight a side effect of AquaMaps methodology that results in potentially unexpected species range boundaries.
+<!--As understanding of a given species improves, we expect that range maps, regardless of method, would more closely approach the true distribution in predicting species presence and absence.
+Close alignment between two independent range maps can therefore indicate higher confidence in species presence or absence than a single map.-->
 
@@
 As a case study, we then reexamine a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] as a case study to explore the implications of selecting one data set over the other.
-Repeating the study using IUCN range data rather than AquaMaps data increases the apparent proportion of gap species from 1.4% to 6.4%, while also increasing the proportion of well-protected species from 1.5% to 2.9%.
+Repeating the study using IUCN range data rather than AquaMaps data quintuples the apparent proportion of "gap" species, i.e.
+those species whose ranges remain completely unprotected, from 1.2% to 6.4%.
 This shift highlights the implications of different data use decisions on our understanding of marine biodiversity status and protection.
 
-While we have identified some potential issues in each dataset, we cannot simply recommend one data set over the other.
-Instead, we recommend simple methods to account for incidence of errors in each dataset, improving confidence in species range predictions to better inform conservation management and policy decisions.
+Due to the variation in georegional and taxonomic coverage, and fundamental distinctions in method and intent of these two spatial range datasets, there is no way to simply recommend one dataset over the other.
+<!--Instead, we recommend simple methods to account for incidence of errors in each dataset, accounting for and reconciling differences in species range predictions to better inform conservation management and policy decisions.-->  Instead, by making explicit the distinctions between the two, and offering methods to identify and reconcile potential errors in each, we hope to reinforce the effective use of these datasets to better inform conservation management and policy decisions.

### 2016-09-11 ms_2_methods.Rmd
all text updated, including references.  Still need to complete abstract.

@@
 These experts outline spatial boundaries that define the "limits of distribution" of a given species, based on observation records and informed by expert understanding of species' range and habitat preferences (Fig S1A).
-<!--The result is essentially an extent of occurrence (EOO) map for the species-->
+___some note about IUCN directions for creating the maps???___.
+For the purposes of this analysis, we did not use IUCN data for bird species, which are available separately through Bird Life International [@birdlife_international_and_natureserve_bird_2015].
 
@@
 The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a global 0.5° grid to determine suitable habitat, resulting in a "probability of occurrence" for each species (Fig S1B).
-To roughly constrain species ranges to appropriate georegions, the AquaMaps model uses Food and Agriculture Organization of the United Nations (FAO) Major Fishing Area [@REF] boundaries.
+To roughly constrain species ranges to appropriate georegions, the AquaMaps model uses Food and Agriculture Organization of the United Nations (FAO) Major Fishing Area [@fao_fisheries_and_aquaculture_department_cwp_2015] boundaries.
 Of the resulting maps, 1296 (5.7%) have been further refined through an expert review process [@kaschner_mapping_2006; @ready_predicting_2010].
-
-<!--In addition to these methodological differences, the two datasets also differ in intent.
-IUCN maps essentially communicate the extent of occurrence (EOO) while AquaMaps uses suitable habitat as an indicator of probability of occurrence, -->
 
@@
 
-__Comparison of taxonomic and regional distribution__:
+__Comparison of taxonomic and regional distribution:__
 To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source, and determined the proportion of each class represented in each dataset.
@@
 
-__Alignment framework for paired maps__: OR __Comparison of paired maps__:
+__Comparison of paired maps:__
 Species range maps communicate a prediction of both the distribution of a species and the size of a species range.
@@
 the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution, using genus and species binomials as a matching key.
-We used the taxize package [@REF] in R [@REF] to standardize species names and synonyms; for species with separate subpopulation maps in IUCN, we combined all subpopulations to create a single global population.
+We used the taxize package [@scott_chamberlain_taxize_2013] in R [@r_core_team_r:_2016] to standardize species names and synonyms; for species with separate subpopulation maps in IUCN, we combined all subpopulations to create a single global population.
 For each of these paired map species, we determined species presence within each spatial cell for each dataset using the same criteria outlined above.
@@
 
-__(some heading here? to separate from alignment)__
+__Examining issues in paired map alignment:__
 Corals provide a useful case study to examine differences in species range alignment.
-They are an intensely studied taxonomic group whose importance in supporting biodiversity is undisputed; their dependence on photosynthesis provides habitat-limiting life history traits, and they don't tend to move.
-Extracting data from the IUCN API [@REF?] we identified depth limitations of each coral species mapped in the IUCN dataset to verify that none are found below the photosynthetic depth limit of 200 m.
-Using the same methodology as shown in Fig 2, we created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] and masked our IUCN coral rasters to identify mapped coral presence below 200 m.
+Corals are an intensely studied taxonomic group whose importance in supporting biodiversity is undisputed; their dependence on photosynthesis provides habitat-limiting life history traits, and they don't tend to move.
+Extracting data from the IUCN API [@iucn_iucn_2015] we identified depth limitations of each coral species mapped in the IUCN dataset to verify that none are found below the photosynthetic depth limit of 200 m.
+Using the same methodology as shown in Fig 2, we created a 200 m bathymetry raster from a 200 m bathymetry spatial dataset (public domain; available from Natural Earth, www.naturalearthdata.com) and masked our IUCN coral rasters to identify mapped coral presence below 200 m.
 The resulting maps were again compared to the AquaMaps rasters to examine distribution alignment and area ratio.
@@
 
-The oceans abound with rare and poorly-understood species who offer sparse information on behavior and habitat preferences to inform models and expert opinion.
-To identify such data-poor species within the set of paired map species, we used the ___robis???___ and ___rgbif???___ packages [@REFS] in R to identify known occurrences of each species; occurrences were averaged between the two occurrence databases as a proxy for data richness.
+The oceans abound with rare and poorly-understood species which offer sparse information on behavior and habitat preferences to inform models and expert opinion.
+To identify such data-poor species within the set of paired map species, we used R clients for OBIS [@obis_data_2016] and GBIF [@gbif_global_2010] to identify known occurrences of each species; occurrences were averaged between the two occurrence databases as a proxy for data richness.
 
@@
 To assess the effectiveness of MPAs in protecting biodiversity, Klein et al.
-[-@klein_shortfalls_2015] compared the coverage of the global MPA network presented by the World Database on Protected Areas (WDPA) [@iucn_world_2014] to the species ranges described in the 2014 AquaMaps dataset [@aquamaps_XXX].
+[-@klein_shortfalls_2015] compared the coverage of the global MPA network presented by the World Database on Protected Areas (WDPA) [@iucn_world_2014] to the species ranges described in the 2014 AquaMaps dataset [@kaschner_aquamaps:_2015].
 For the primary analysis, the researchers defined species presence as 50% or greater probability of occurrence.

### 2016-09-11 ms_3_res_disc_concl.Rmd
all text updated, including references.  Still need to complete abstract.

@@
 
-[brief overview of results/discussion?]
+[brief overview of results/discussion?] In comparing the IUCN and AquaMaps datasets, it is again important to emphasize that the two differ in both methodology and intent.
+As area of occupancy is generally defined as a subset of a species' extent of occurrence, AquaMaps ranges (effectively area of occupancy) should generally fall within the bounds of IUCN ranges (effectively extent of occurrence), capturing a smaller total area.
+However, AquaMaps range maps are created independently from IUCN data and therefore exceptions are certain to arise.
+Here we are looking for systematic deviations from our expectations that might highlight implications of data use decisions.
 
@@
 The margin frequency plots show relative species count per cell at each latitude and longitude.
-
-[ideas for discussion here?
-each provides information unavailable in the other; in addition to considerations about the nature of each dataset, the species and regions covered are an important consideration in which dataset to use]
 
@@
 [discussion] [results] Comparing distribution alignment and area ratio for the `r format(n_spp_both, big.mark = ',')` paired map species (Fig 2A), a weak negative linear pattern appears to emerge, suggesting that increasing similarity in range area correlates very slightly with decreasing distribution alignment (R^2 = .016).
-The pattern itself is unimportant, and emerges simply due to the nature of the analysis and the datasets:  AquaMaps tends to extrapolate species ranges into suitable areas beyond known occurrences more frequently than IUCN maps, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
+The pattern itself is not particularly important, and emerges simply due to the nature of the analysis and the datasets:  The AquaMaps model tends to extrapolate species ranges into suitable areas beyond known occurrences more frequently than IUCN maps, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
 For species with dissimilar range areas, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
@@
 By dividing the paired map species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
-Representative maps from each category are provided in the supporting materials (Fig S1).
+Representative maps from each category are provided in the supporting materials (Fig SXXX).
 
-Where two independent methods predict presence (or absence) of a species in the same location, we can assume a higher confidence in the accuracy of that prediction.
+<!--
+Where two independent methods predict presence (or absence) of a species in the same location, we assume a higher confidence in the accuracy of that prediction.
 Where one method predicts presence while the other predicts absence, we can investigate potential sources of commission and omission errors to refine our understanding of the species' range.
 Understanding and addressing these sources of error provides guidance to improve predictions of species range even when only one source of spatial data is available.
+-->
 
 > __Fig 2.__ (A) Distribution alignment (overlap of smaller range within larger) versus area ratio (the ratio of smaller range area to the larger range area) for `r format(n_spp_both, big.mark = ',')` species included in both IUCN and AquaMaps datasets.
-The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the extent of described ranges (n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
+The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than mean value) in both spatial distribution and the extent of described ranges (n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
 The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in area
@@
 These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig 2B).
-This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their ranges span nearly the entire map.
+This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their extent of occurrence and area of occupancy both span nearly the entire map.
 
-[results/discussion] The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where one range is notably larger than the other (For `r round(n_q2_iucn_larger/n_q2 * 100, 1)`% of the species in this quadrant, the IUCN range is an average of 2.57 times larger than the AquaMaps range).
+[results/discussion] The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where one range is notably larger than the other.
+For `r round(n_q2_iucn_larger/n_q2 * 100, 1)`% of the species in this quadrant, the IUCN range is larger (on average, 2.57 times as large) than the AquaMaps range, as expected when comparing an extent of occurrence to an area of occupancy map.
 This suggests a high rate of commission and/or omission errors by one or both datasets; further analysis is required to disentangle the source and type of error contributing to poor area alignment.
@@
 Consequently, area ratios are close to 100%, though the poor distribution alignment indicates that one or both datasets are introducing significant errors.
+In this quadrant, the IUCN range is the larger for only 56.5% of species, defying our expectations based on definitions of area of occupancy and extent of occurrence.
 As above, further analysis is required to disentangle the causes of error.
 
+<!--
+   quad am_sm iucn_sm tot_n
+  <chr> <int>   <int> <int>
+1    q1   422     105   527
+2    q2   624      85   709
+3    q3   359     276   635
+4    q4   109     334   443
+-->
+
 [results] The lower left quadrant includes species (n = `r n_q4`) where alignment is poor in both dimensions.
+In this quadrant, the IUCN range is larger for only 24.6% of species; this again defies our expectations for extent of occurrence and area of occupancy range predictions.
 Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
-In addition, the AquaMaps dataset offers a quality metric based on the number of unique cells containing valid occurrences; for this quadrant, the median "occurcells" is 11 compared to a median of 40 across the other three quadrants.
-Care should be taken when using distribution and range maps based upon fewer observations, as they clearly bear greater uncertainty; AquaMaps explicitly recommends using any of its maps generated with an "occurcells" count fewer than 10 [@AquaMapsXXX].
+The AquaMaps dataset offers its own quality metric based on the number of unique cells containing valid occurrences; for this quadrant, the median "occurcells" is 11 compared to a median of 40 across the other three quadrants.
+Care should be taken when using distribution and range maps based upon fewer observations, as they clearly bear greater uncertainty; AquaMaps explicitly warns against using any of its maps generated with an "occurcells" count fewer than 10 [@kaschner_aquamaps:_2015].
 
@@
 Ocean depth preference is explicitly included as a parameter in AquaMaps models.
-While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." [@REF http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources]), it is not presented as a requirement, so we cannot take its inclusion for granted.
+While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." [@iucn_iucn_2015, from website:  http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources]), it is not presented as a requirement, so we cannot take its inclusion for granted.
 
@@
 
-[results] In constraining coral ranges to shallow waters, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps.
+[results] In constraining coral ranges to appropriately shallow waters, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps.
 Using the original quadrant definitions from Fig 2B, we see in Fig 3B a massive shift in 354 paired-map coral species due to constraining coral depths to 200 m.
 Membership in the "well-aligned" quadrant jumped from 22.4% to 76.2%, with a corresponding decrease in all other quadrants.
+By excluding the unsuitable areas from IUCN's extent of occurrence, we eliminate preventable commission errors and more closely approximate the area of occupancy described by AquaMaps.
 See Fig SXXX to examine the shifts of individual species among the quadrants.
 
-[discussion] We cannot know for certain the true distribution of each of these corals; certainly some commission errors result from satisfying the connectivity requirements of an EOO, and others may be due to experts taking a precautionary approach.
+[discussion] We cannot know for certain the true distribution of each of these corals; certainly some commission errors result from satisfying the connectivity requirements of an extent of occurrence, and others may be due to experts taking a precautionary approach.
 But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported distributions for most purposes.
@@
 While such a discontinuous boundary would likely be obvious when inspecting the distribution of an individual species, the distinction is likely to be obscured when aggregating many species ranges as is typical for biodiversity or conservation studies.
-Ecoregional spatial zones such as Marine Ecoregions of the World [@REF???] may provide an additional useful constraint on species ranges; expert review, though time-consuming, is the most certain route to boosting confidence in these predicted distributions.
+Ecoregional spatial zones such as Marine Ecoregions of the World [@spalding_marine_2007] may provide an additional useful constraint on species ranges; expert review, though time-consuming, is the most certain route to boosting confidence in these predicted distributions.
 
@@
 [-@klein_shortfalls_2015] compare the global distribution of species to the global distribution of marine protected areas to assess how well current MPAs overlap with species ranges and identify which species fall through gaps in protection.
-The study relied on the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
+The study relied on the 2014 version of the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
 They found that the global MPA network leaves 90.5% of marine species with less than 5% of their overall range represented within MPAs, and 1.4% of species have no protection at all (i.e., "gap" species).
 
-We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using the most recent AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig4'>Fig 4</a>).
+We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using the 2015 AquaMaps data and a 0% threshold to more closely approximate the extent of occurrence represented by IUCN data, <a href = '#fig4'>Fig 4</a>).
 We found a five-fold increase in the proportion of gap species (6.4% of species vs.
@@
 Scenario 3, using 2015 AquaMaps data, drops the presence threshold to zero, showing an expected decrease in gap species, but also a decrease in species with 5% or greater protected range.
-Scenario 4 adds an additional 1745 species unique to IUCN, resulting in increases in gap species and species with less than 2% coverage.
-Scenario 5 examines species MPA coverage using only the IUCN dataset.
+Scenario 4 examines species MPA coverage using only the IUCN dataset.
 
-[discussion?]
-
-[can this wrap up the results/discussion?] To achieve more comprehensive global coverage of species ranges these two datasets can be used together, understanding that the underlying differences complicate such direct comparisons.
-Major considerations include aligning the criteria used to determine "presence" (e.g.
-selecting an appropriate presence threshold for AquaMaps; _a presence threshold of 0% most closely approximates the "limits of distribution" criterion described by IUCN range maps_ ___- not really discussed or backed up here___) and addressing mechanisms that introduce errors of commission and omission.
-___as it stands, this paper really doesn't go into combining datasets and how to do it; omit this paragraph?___
+To achieve more comprehensive global coverage of species ranges it may be desirable to use these two datasets together, as the Ocean Health Index (OHI) does to calculate scores for its Species goal [@halpern_index_2012; @halpern_patterns_2015] ___should I even mention this here, since I do not provide details?___  Trimming unsuitable habitat from the IUCN's extent of occurrence, for example by explicitly clipping map to appropriate depths for corals, would result in maps that more closely approximate areas of occupancy.
+Conversely, including all AquaMaps cells with a non-zero "probability of occurrence" (rather than using a probability threshold to determine presence, e.g.
+greater than 40% for the Ocean Health Index [@halpern_index_2012; @halpern_patterns_2015] or 50% for the MPA gap analysis [@klein_shortfalls_2015]),  would result in maps that more closely, though still imperfectly, approximate extents of occurrence.
 
@@
 No spatial dataset can ever claim to know the "truth" of the whole of marine biodiversity.
-AquaMaps and IUCN range maps show strong agreement for many well-studied species, but substantial differences illustrate uncertainty in our understanding of spatial distribution for many others.
-Although many other approaches exist for species distribution modeling, these two are the largest approaches applied globally across a broad range of marine taxa.
-Method-driven differences in commission and omission errors produce clear and significant disagreement in species range descriptions between AquaMaps and IUCN datasets.
-Conclusions drawn from each of these datasets would paint dramatically different pictures of global marine biodiversity or the effectiveness of conservation management decisions.
-Identifying and addressing differences in these datasets will increase their utility for research and conservation actions.
+AquaMaps and IUCN range maps show strong agreement for many well-studied species, but for many others, substantial differences arise from differences in methodology and intent of each dataset.
+Although other approaches exist for predicting species distribution, these two are the largest approaches applied globally across a broad range of marine taxa.
+Recognizing and acknowledging the advantages and intended purposes of the range maps presented by these datasets will increase their utility for research and conservation actions.
+Conclusions drawn from each of these datasets could paint dramatically different pictures of global marine biodiversity or the effectiveness of conservation management decisions.
 
 Effective management and protection of marine species depends on a robust understanding of where species exist and where they do not; without this knowledge we risk wasting resources protecting low-value regions while missing opportunities to protect critical ones.
-By identifying the differences between these two important and fundamentally different marine species range datasets and understanding the likely mechanisms causing these discrepancies, we improve our ability to develop strategic and effective conservation policy that supports a resilient ocean ecosystem.
+By understanding the distinctions between these two important and fundamentally different marine species range datasets, we improve our ability to inform strategic and effective conservation policy that supports a resilient ocean ecosystem.

### 2016-09-11 ms_6_supplemental.Rmd
all text updated, including references.  Still need to complete abstract.

@@
 __Fig S1.__  Examples of AquaMaps and IUCN raw species range data.
-(A) AquaMaps distribution of _Thunnus alalunga_ (Albacore Tuna) showing global range with varying probabilities of occurrence assigned to 0.5° grid cells.
-(B) IUCN distribution of _T.
-alalunga_ represented as extent of occurrence polygons.
+(A, B) AquaMaps distribution of _Thunnus alalunga_ (Albacore Tuna) [@kaschner_aquamaps:_2015] showing global range with varying probabilities of occurrence assigned to 0.5° grid cells (A) and presence within 0.5° grid cells (B).
+(C, D) IUCN distribution of _T.
+alalunga_ [@@iucn_iucn_2015] represented as extent of occurrence polygons (C) and presence within 0.5° grid cells (D).
+
+<!-- 	
+Fis-22832, Thunnus alalunga, 21856, citation general IUCN; 
+Reviewed distribution maps for Thunnus alalunga (Albacore), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+www.aquamaps.org, version of Aug.
+2013.
+Web.
+Accessed 11 Sep.
+2016.
+-->
 
@@
 
-__Fig S3.__ Representative species maps to illustrate each quadrant from Fig 2A.  Each map is positioned to match its quadrant in Fig 2A.
-(A) Distribution-matched: Conus episcopatus, the dignified cone snail.
-Distributions shows excellent overlap in the western Pacific, though IUCN range extends well beyond the bounds of the AquaMaps range.
-(B) Well-matched: Kajikia albida, the Atlantic white marlin.  Distributions from each data set show nearly complete overlap, and very similar range size.
-(C) Poorly-matched: Acanthurus nigroris, the blue-lined surgeonfish.
+__Fig S3.__ Representative species maps [@kaschner_aquamaps:_2015; @iucn_iucn_2015] to illustrate each quadrant from Fig 2A.  Each map is positioned to match its quadrant in Fig 2A.
+(A) Distribution-matched: _Conus episcopatus_, the dignified cone snail.
+Distributions show excellent overlap in the western Pacific, though IUCN range extends well beyond the bounds of the AquaMaps range.
+(B) Well-matched: _Kajikia albida_, the Atlantic white marlin.  Distributions from each data set show nearly complete overlap, and very similar range size.
+(C) Poorly-matched: _Acanthurus nigroris_, the blue-lined surgeonfish.
 IUCN predicts species distribution only near the Hawaiian islands; AquaMaps predicts extensive distribution throughout the central and western Pacific Ocean.
 The datasets align in neither distribution nor range size.
-(D) Area-matched: Conus magnificus, the magnificent cone snail.  Distributions overlap in the southern Pacific, but align poorly elsewhere.  The range sizes are similar.
+(D) Area-matched: _Conus magnificus_, the magnificent cone snail.  Distributions overlap in the southern Pacific, but align poorly elsewhere.  The range sizes are similar.
 
+<!--
+W-Msc-215455, Conus episcopatus, 192664 citation Conch Books, Hackenheim, Germany; Computer generated distribution maps for Conus episcopatus, with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+www.aquamaps.org, version of Aug.
+2013.
+Web.
+Accessed 11 Sep.
+2016.
+
+W-Msc-215551, Conus magnificus, 192573 citation Conch Books, Hackenheim, Germany; Computer generated distribution maps for Conus magnificus, with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+www.aquamaps.org, version of Aug.
+2013.
+Web.
+Accessed 11 Sep.
+2016.
+
+Fis-31079, Acanthurus nigroris, 177999 citation IUCN; Reviewed distribution maps for Acanthurus nigroris (Bluelined surgeonfish), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+www.aquamaps.org, version of Aug.
+2013.
+Web.
+Accessed 11 Sep.
+2016.
+
+Fis-164253, Kajikia albida, 170322 citation IUCN; Computer generated distribution maps for Kajikia albida (Atlantic white marlin), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+www.aquamaps.org, version of Aug.
+2013.
+Web.
+Accessed 11 Sep.
+2016.
+-->
+
+<!--
 ## Risk by quadrant
@@
 Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity of apparent species range to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.
+-->
 
@@
 -->
-
-<!--
-#### Caveats
-For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes, but are missing economically important subgroups including salmon, rockfish, and clupeids.
-However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.
--->

### 2016-09-11 ms_6_supplemental.Rmd
updated refs for supplemental

@@
 __Fig S1.__  Examples of AquaMaps and IUCN raw species range data.
-(A, B) AquaMaps distribution of _Thunnus alalunga_ (Albacore Tuna) [@kaschner_aquamaps:_2015] showing global range with varying probabilities of occurrence assigned to 0.5° grid cells (A) and presence within 0.5° grid cells (B).
+(A, B) AquaMaps distribution of _Thunnus alalunga_ (Albacore Tuna) (AquaMaps 2013e) showing global range with varying probabilities of occurrence assigned to 0.5° grid cells (A) and presence within 0.5° grid cells (B).
 (C, D) IUCN distribution of _T.
-alalunga_ [@@iucn_iucn_2015] represented as extent of occurrence polygons (C) and presence within 0.5° grid cells (D).
+alalunga_ (IUCN 2011b) represented as extent of occurrence polygons (C) and presence within 0.5° grid cells (D).
 
@@
 
-__Fig S3.__ Representative species maps [@kaschner_aquamaps:_2015; @iucn_iucn_2015] to illustrate each quadrant from Fig 2A.  Each map is positioned to match its quadrant in Fig 2A.
+__Fig S3.__ Representative species maps (AquaMaps 2013abcd; Conch 2013ab; IUCN 2011a; IUCN 2012) to illustrate each quadrant from Fig 2A.  Each map is positioned to match its quadrant in Fig 2A.
 (A) Distribution-matched: _Conus episcopatus_, the dignified cone snail.
@@
 -->
+
+## References for supplemental information
+
+Conch Books, Hackenheim, Germany 2013. Conus episcopatus. The IUCN Red List of Threatened Species. Version 2016-2
+
+Conch Books, Hackenheim, Germany 2013. Conus magnificus. The IUCN Red List of Threatened Species. Version 2016-2
+
+International Union for Conservation of Nature (IUCN) 2011. Kajikia albida. The IUCN Red List of Threatened Species. Version 2016-2
+
+International Union for Conservation of Nature (IUCN) 2011. Thunnus alalunga. The IUCN Red List of Threatened Species. Version 2016-2
+
+International Union for Conservation of Nature (IUCN) 2012. Acanthurus nigroris. The IUCN Red List of Threatened Species. Version 2016-2
+
+www.aquamaps.org, version of Aug.
+2013.
+Computer generated distribution maps for Conus episcopatus, with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+Web.
+Accessed 11 Sep.
+2016.
+
+www.aquamaps.org, version of Aug.
+2013.
+Computer generated distribution maps for Conus magnificus, with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+Web.
+Accessed 11 Sep.
+2016.
+
+www.aquamaps.org, version of Aug.
+2013.
+Computer generated distribution maps for Kajikia albida (Atlantic white marlin), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+Web.
+Accessed 11 Sep.
+2016.
+
+www.aquamaps.org, version of Aug.
+2013.
+Reviewed distribution maps for Acanthurus nigroris (Bluelined surgeonfish), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+Web.
+Accessed 11 Sep.
+2016.
+
+www.aquamaps.org, version of Aug.
+2013.
+Reviewed distribution maps for Thunnus alalunga (Albacore), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+Web.
+Accessed 11 Sep.
+2016.

### 2016-09-12 ms_0_abst.Rmd
argh this fucking abstract

@@
 -->
+from intro:
+Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation.
+This body of science has resulted in various compiled databases of species distribution and range maps which provide foundational information for understanding species diversity and extinction risk, predicting species responses to human impacts and climate change, and managing and protecting species effectively.
+No species range dataset can claim to represent the "truth" of any species' spatial distribution, but rather each offers its own distinct understanding of that latent distribution.
+These varying predictions of species presence and absence, driven by intent and methodology, can result in very different predictions of species presence; however, conservation managers and policy makers must base their actions on conclusions drawn from these imperfect datasets.
+It is critical to understand the differences in spatial range datasets and the implications of these differences for conservation research and decision-making.
 
-Species distribution data provide the foundation for a wide range of ecological research studies and conservation management decisions, yet most species ranges remain unknown, and existing range maps often suffer from data limitations and inconsistencies.
-AquaMaps and the International Union for Conservation of Nature (IUCN) are two distinct efforts to map marine species distributions at a global scale.
+Species distribution data provide the foundation for a wide range of ecological research studies and conservation management decisions.
+Two major efforts to provide marine species distributions at a global scale are the International Union for Conservation of Nature (IUCN), which provides expert-generated range maps that outline the complete extent of a species' distribution; and AquaMaps, which provides model-generated species distribution maps that predict areas occupied by the species.
+Neither dataset claims to represent the "truth" of species' spatial distributions, but rather each communicates a distinct understanding of the underlying distribution, with a distinct purpose in mind.
+
+Distinctions in intent and methodology can result in very different predictions of species distributions; these differences bear important implications for scientists and decision makers who rely on these datasets to inform conservation policy and management actions.
+
 Together these databases represent `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` shared species.
-Here we examine differences in predicted species ranges between the two datasets and find that these misalignments mainly result from divergent methodologies that introduce differing frequencies of commission and omission errors.
-We illustrate the scientific and management implications of these differences by repeating a global analysis of gaps in coverage of marine protected areas, and find significantly different results depending on how the two datasets are used.
+Ideally, the decision to use one dataset one over the other should carefully consider the purpose for which it was designed, but this small overlap between the two suggests that data availability is likely a more important factor.
+
+Comparing these overlapping predicted species ranges, we highlight several important differences between the two datasets, tradeoffs, blah blah
+
+We illustrate the scientific and management implications of these differences/tradeoffs by repeating a global analysis of gaps in coverage of marine protected areas, and find significantly different results depending on how the two datasets are used.
+
 While we suggest methods to address specific errors in each spatial dataset, it remains essential to understand the implications of dataset differences for conservation planning and decision-making.
+
+FRAME as tradeoffs rather than differences; this could be a way to bring the two datasets together for the last paragraph of the results and discussion - minimizing tradeoffs rather than fixing differences?

### 2016-09-12 ms_1_intro.Rmd
argh this fucking abstract

@@
 where can a particular species be found, they communicate fundamentally distinct concepts.
-The IUCN range map for a given species describes the "extent of occurrence" for that species, a contiguous region that encloses all known occurrences and connecting regions, with the explicit caveat that this "does not mean that it is distributed equally within that polygon or occurs everywhere within that polygon" [@iucn_iucn_2015].
+The IUCN range map for a given species describes the "extent of occurrence" for that species, a contiguous region that encloses all known occurrences and connecting regions, with the explicit caveat that this "does not mean that [the species] is distributed equally within that polygon or occurs everywhere within that polygon" [@iucn_iucn_2015].
 The AquaMaps map for a species, on the other hand, seems to better be defined as "area of occupancy": the subset of the extent of occurrence in which the species is actually likely to be found, for example by excluding unsuitable habitats from the extent of occurrence [@iucn_iucn_2015].
@@
 We compare how each data source represents the global spatial and taxonomic distribution of the `r format(n_spp_all, big.mark = ',')` marine species mapped by one or both datasets.
-Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).
-Yet the small overlap in mapped species between the two datasets means each provides unique value in taxonomic and geographic coverage.
+Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps.
+IUCN range map data exist for `r format(n_spp_iucn, big.mark = ',')` unique marine species, with only `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets.
+This small overlap means each dataset provides unique value in taxonomic and geographic coverage, but also suggests that data availability likely plays a larger role in selecting a dataset than does the intended purpose of the data.
 
 Although relatively small in number, these overlap species present a unique opportunity to evaluate the two datasets overall.
-For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align in both spatial distribution and overall area.
+For the species included in both datasets, we examine how well the maps align in both spatial distribution and overall area.
 Based upon the definitions of extent of occurrence and area of occupancy, we expect that for a given species, the AquaMaps predicted distribution will fall within, and describe a smaller range than, the IUCN predicted distribution.

### 2016-09-12 manuscript-main.Rmd
finished up draft; working on suppl figs

@@
 
-# library(ggplot2)
-# library(maps)
-
-# library(raster)
-# library(data.table)
 library(tidyr)
@@
 library(readr)
-
-# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-#            'Darwin'  = '/Volumes/data_edit',
-#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
-
-# dir_anx  <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
 

### 2016-09-12 ms_0_abst.Rmd
finished up draft; working on suppl figs

@@
 -->
-from intro:
-Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation.
+
+<!-- from intro: -->
+<!-- Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation.
 This body of science has resulted in various compiled databases of species distribution and range maps which provide foundational information for understanding species diversity and extinction risk, predicting species responses to human impacts and climate change, and managing and protecting species effectively.
@@
 It is critical to understand the differences in spatial range datasets and the implications of these differences for conservation research and decision-making.
+-->
 
@@
 
-Distinctions in intent and methodology can result in very different predictions of species distributions; these differences bear important implications for scientists and decision makers who rely on these datasets to inform conservation policy and management actions.
+Together these databases represent `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` shared species.
+Ideally, the decision to use one dataset one over the other should carefully consider the purpose for which it was designed, but this small overlap between the two suggests that data availability is likely an important factor as well.
 
-Together these databases represent `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` shared species.
-Ideally, the decision to use one dataset one over the other should carefully consider the purpose for which it was designed, but this small overlap between the two suggests that data availability is likely a more important factor.
+Differences in intent and methodology can result in very different predictions of species distributions, which bear important implications for scientists and decision makers who rely upon these datasets to inform conservation policy and management actions.
+Comparing distributions for the small subset of species with maps in both datasets, we highlight several differences resulting from differences in methodology and intent.
+In particular, we find that IUCN maps greatly overpredict coral presence into unsuitably deep waters, and we demonstrate that AquaMaps methodology can produce odd discontinuities at the extremes of a species range.
+We illustrate the scientific and management implications of these tradeoffs by repeating a global analysis of gaps in coverage of marine protected areas, and find significantly different results depending on how the two datasets are used.
 
-Comparing these overlapping predicted species ranges, we highlight several important differences between the two datasets, tradeoffs, blah blah
-
-We illustrate the scientific and management implications of these differences/tradeoffs by repeating a global analysis of gaps in coverage of marine protected areas, and find significantly different results depending on how the two datasets are used.
-
-While we suggest methods to address specific errors in each spatial dataset, it remains essential to understand the implications of dataset differences for conservation planning and decision-making.
-
-FRAME as tradeoffs rather than differences; this could be a way to bring the two datasets together for the last paragraph of the results and discussion - minimizing tradeoffs rather than fixing differences?
+It remains essential to recognizes the tradeoffs between these two valuable marine datasets, and understand the implications of these tradeoffs for conservation research and decision-making.

### 2016-09-12 ms_1_intro.Rmd
finished up draft; working on suppl figs

@@
 The IUCN range map for a given species describes the "extent of occurrence" for that species, a contiguous region that encloses all known occurrences and connecting regions, with the explicit caveat that this "does not mean that [the species] is distributed equally within that polygon or occurs everywhere within that polygon" [@iucn_iucn_2015].
-The AquaMaps map for a species, on the other hand, seems to better be defined as "area of occupancy": the subset of the extent of occurrence in which the species is actually likely to be found, for example by excluding unsuitable habitats from the extent of occurrence [@iucn_iucn_2015].
-The distinctions between extent of occurrence and area of occupancy, including methodology, limitations, and appropriate usage, are well documented [@rondinini_tradeoffs_2006; @gaston_sizes_2009].
-___include reference to a supplemental table for more details?___
+The AquaMaps map for a species, on the other hand, excludes unsuitable habitats and thus seems to better be defined as "area of occupancy": the subset of the extent of occurrence in which the species is actually likely to be found [@iucn_iucn_2015].
+<!--The distinctions between extent of occurrence and area of occupancy, including methodology, limitations, and intended purpose, are well documented [@rondinini_tradeoffs_2006; @gaston_sizes_2009].
+___include table for more details?___ -->
+
+<!--
+_Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: extent of occurrence range data such as IUCN range maps will inevitably introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps are expected to introduce fewer commission errors at the expense of more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
+Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006; @jetz_ecological_2008]._
+
+| ... | Extent of Occurrence | Area of Occupancy |
+|:--- |:-----                |:------            |
+| Definition | Extent of occurrence is defined as the area contained within the shortest continuous imaginary boundary which can be drawn to encompass all the known, inferred or projected sites of present occurrence of a taxon, excluding cases of vagrancy (see Figure 2).
+This measure may exclude discontinuities or disjunctions within the overall distributions of taxa (e.g.
+large areas of obviously unsuitable habitat) (but see 'area of occupancy', point 10 below).
+[@iucn] | Area of occupancy is defined as the area within its 'extent of occurrence' (see point 9 above) which is occupied by a taxon, excluding cases of vagrancy.
+The measure reflects the fact that a taxon will not usually occur throughout the area of its extent of occurrence, which may contain unsuitable or unoccupied habitats.
+[@iucn] |
+| Purpose | 
+| 
+-->
 
@@
 In such a case, it is crucial to acknowledge, and where possible address, the resulting tradeoffs.
-
-___include as table, in body or in supplemental?
-It really kind of doesn't fit here any more___ _Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: extent of occurrence range data such as IUCN range maps will inevitably introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps are expected to introduce fewer commission errors at the expense of more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
-Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006; @jetz_ecological_2008]._
 

### 2016-09-12 ms_2_methods.Rmd
finished up draft; working on suppl figs

@@
 These experts outline spatial boundaries that define the "limits of distribution" of a given species, based on observation records and informed by expert understanding of species' range and habitat preferences (Fig S1A).
-___some note about IUCN directions for creating the maps???___.
-For the purposes of this analysis, we did not use IUCN data for bird species, which are available separately through Bird Life International [@birdlife_international_and_natureserve_bird_2015].
+IUCN guidelines generally recommend that boundaries be drawn as a "minimum convex polygon": "the smallest polygon in which no internal angle exceeds 180 degrees and which contains all the sites of occurrence" [@iucn_iucn_2015].
+For the purposes of this analysis, we elected not to use IUCN data for bird species, which are available separately through Bird Life International [@birdlife_international_and_natureserve_bird_2015].
 
@@
 <!--
-___OMIT: Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., previously, the Ocean Health Index defined presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
-In the 2016 OHI assessment, the presence threshold was eliminated, to better approximate the "extent of occurrence" generally indicated by IUCN maps, since the two datasets are used together.___
+OMIT: Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., previously, the Ocean Health Index defined presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
+In the 2016 OHI assessment, the presence threshold was eliminated, to better approximate the "extent of occurrence" generally indicated by IUCN maps, since the two datasets are used together
 -->
 
-__Comparison of taxonomic and regional distribution:__
+__Comparison of taxonomic and regional distribution__:
 To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source, and determined the proportion of each class represented in each dataset.
-To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; we determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell  ___(Fig S2)___.
-For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.
+To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; we determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell  (Fig S2)  For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.
 
-__Comparison of paired maps:__
+__Comparison of paired maps__:
 Species range maps communicate a prediction of both the distribution of a species and the size of a species range.
@@
 
-__Examining issues in paired map alignment:__
+__Examining issues in paired map alignment__:
 Corals provide a useful case study to examine differences in species range alignment.

### 2016-09-12 ms_3_res_disc_concl.Rmd
argh this fucking abstract
finished up draft; working on suppl figs

@@
 
-[brief overview of results/discussion?] In comparing the IUCN and AquaMaps datasets, it is again important to emphasize that the two differ in both methodology and intent.
+In comparing the IUCN and AquaMaps datasets, it is again important to emphasize that the two differ in both methodology and intent.
 As area of occupancy is generally defined as a subset of a species' extent of occurrence, AquaMaps ranges (effectively area of occupancy) should generally fall within the bounds of IUCN ranges (effectively extent of occurrence), capturing a smaller total area.
@@
 
-[results] The two datasets have notably different taxonomic (Fig 1A) and regional (Figs 1B, 1C) coverage.
+The two datasets have notably different taxonomic (Fig 1A) and regional (Figs 1B, 1C) coverage.
 AquaMaps encompasses a broader range of taxa than IUCN, as IUCN spatial data files are only available for select taxonomic groups that have been comprehensively assessed.
@@
 The longitude frequency plots show a slight shift in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
+
+![](figures/Fig_1.png)
 
@@
 
-[discussion] [results] Comparing distribution alignment and area ratio for the `r format(n_spp_both, big.mark = ',')` paired map species (Fig 2A), a weak negative linear pattern appears to emerge, suggesting that increasing similarity in range area correlates very slightly with decreasing distribution alignment (R^2 = .016).
+Comparing distribution alignment and area ratio for the `r format(n_spp_both, big.mark = ',')` paired map species (Fig 2A), a weak negative linear pattern appears to emerge, suggesting that increasing similarity in range area correlates very slightly with decreasing distribution alignment (R^2 = .016).
 The pattern itself is not particularly important, and emerges simply due to the nature of the analysis and the datasets:  The AquaMaps model tends to extrapolate species ranges into suitable areas beyond known occurrences more frequently than IUCN maps, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
@@
 By dividing the paired map species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
-Representative maps from each category are provided in the supporting materials (Fig SXXX).
+Representative maps from each category are provided in the supporting materials (Fig S3).
 
@@
 -->
+
+![](figures/Fig_2.png)
 
@@
 
-[results/discussion] The upper right quadrant includes the species (n = `r n_q1`) whose described ranges are above average in alignment of both spatial distribution and area.
+The upper right quadrant includes the species (n = `r n_q1`) whose described ranges are above average in alignment of both spatial distribution and area.
 These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig 2B).
@@
 
-[results/discussion] The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where one range is notably larger than the other.
+The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where one range is notably larger than the other.
 For `r round(n_q2_iucn_larger/n_q2 * 100, 1)`% of the species in this quadrant, the IUCN range is larger (on average, 2.57 times as large) than the AquaMaps range, as expected when comparing an extent of occurrence to an area of occupancy map.
@@
 
-[results/discussion] Species found in the lower right quadrant (n = `r n_q3`) seem to represent cases of "two wrongs make a right."  For these species, IUCN and AquaMaps both predict ranges extending far beyond the overlapping region, but the methodological differences result in very different extrapolations.
+Species found in the lower right quadrant (n = `r n_q3`) seem to represent cases of "two wrongs make a right."  For these species, IUCN and AquaMaps both predict ranges extending far beyond the overlapping region, but the methodological differences result in very different extrapolations.
 Consequently, area ratios are close to 100%, though the poor distribution alignment indicates that one or both datasets are introducing significant errors.
@@
 
-[results] The lower left quadrant includes species (n = `r n_q4`) where alignment is poor in both dimensions.
+The lower left quadrant includes species (n = `r n_q4`) where alignment is poor in both dimensions.
 In this quadrant, the IUCN range is larger for only 24.6% of species; this again defies our expectations for extent of occurrence and area of occupancy range predictions.
@@
 
-[discussion] Noting from Fig 2B that corals dominate the upper-left "distribution-aligned" quadrant of Fig 2A (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), we chose to examine the effect of explicitly restricting IUCN ranges to depths based on species' life histories.
+Noting from Fig 2B that corals dominate the upper-left "distribution-aligned" quadrant of Fig 2A (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), we chose to examine the effect of explicitly restricting IUCN ranges to depths based on species' life histories.
 Corals offer an excellent case study, due to their foundational role in biodiverse habitats, as well as their lack of mobility and reliance on photosynthesis.
@@
 
-[results] Fig 3A shows aggregated ranges of the 463 coral species mapped in the IUCN dataset, with their ranges broken into proportional area deeper and shallower than 200 m.
+Fig 3A shows aggregated ranges of the 463 coral species mapped in the IUCN dataset, with their ranges broken into proportional area deeper and shallower than 200 m.
 According to IUCN descriptions, none of these species is indicated to occur deeper than 200 m, and 94% are confined to waters shallower than 50 m; seven of the mapped species had no reported depth information.
 Clipping coral ranges to shallower than 200 m eliminated an average of 47.6% of the total predicted area while still allowing for a generous estimate of suitable habitat.
+
+![](figures/Fig_3.png)
 
@@
 
-[results] In constraining coral ranges to appropriately shallow waters, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps.
+In constraining coral ranges to appropriately shallow waters, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps.
 Using the original quadrant definitions from Fig 2B, we see in Fig 3B a massive shift in 354 paired-map coral species due to constraining coral depths to 200 m.
@@
 By excluding the unsuitable areas from IUCN's extent of occurrence, we eliminate preventable commission errors and more closely approximate the area of occupancy described by AquaMaps.
-See Fig SXXX to examine the shifts of individual species among the quadrants.
+See Fig S4 to examine the shifts of individual species among the quadrants.
 
-[discussion] We cannot know for certain the true distribution of each of these corals; certainly some commission errors result from satisfying the connectivity requirements of an extent of occurrence, and others may be due to experts taking a precautionary approach.
+We cannot know for certain the true distribution of each of these corals; certainly some commission errors result from satisfying the connectivity requirements of an extent of occurrence, and others may be due to experts taking a precautionary approach.
 But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported distributions for most purposes.
@@
 
-[results] From the entirety of the AquaMaps dataset, we identified 3,577 Indo-Pacific species whose equatorial distributions (between 25° S and 20° N) encounter a range limit at 175° W.
+From the entirety of the AquaMaps dataset, we identified 3,577 Indo-Pacific species whose equatorial distributions (between 25° S and 20° N) encounter a range limit at 175° W.
 This longitudinal boundary created a western limit for 369 species, and an eastern limit for 3,208 species.
-A clear discontinuity in species distributions of the 3,208 west-limited species (Fig 4) matches perfectly with FAO region 77; other discontinuities are apparent at other FAO boundaries, despite these boundaries not being actively studied in the analysis.
+A clear discontinuity in species distributions of the 3,208 west-limited species (Fig 4) matches perfectly with FAO region 77 [@fao_fisheries_and_aquaculture_department_cwp_2015]; other discontinuities are apparent at other FAO boundaries, despite these boundaries not being actively studied in the analysis.
 Similar discontinuities were clear when mapping the west-limited species.
+
+![](figures/Fig_4.png)
 
@@
 
-[discussion] FAO Major Fishing Area boundaries provide a readily available method to roughly contrain AquaMaps predictions to appropriate ocean basins, thus eliminating a large source of potential commission errors and enabling rapid modeling of thousands of species ranges.
+FAO Major Fishing Area boundaries provide a readily available method to roughly contrain AquaMaps predictions to appropriate ocean basins, thus eliminating a large source of potential commission errors and enabling rapid modeling of thousands of species ranges.
 However, these boundaries are defined for statistical purposes based on economic and political considerations rather than ecological considerations, and can result in odd discontinuities in species range predictions where otherwise suitable habitat is excluded.
@@
 
+![](figures/Fig_5.png)
+
 > __Fig 5.__  Percent of species range covered by MPAs based upon methods in Klein et al.
@@
 
-To achieve more comprehensive global coverage of species ranges it may be desirable to use these two datasets together, as the Ocean Health Index (OHI) does to calculate scores for its Species goal [@halpern_index_2012; @halpern_patterns_2015] ___should I even mention this here, since I do not provide details?___  Trimming unsuitable habitat from the IUCN's extent of occurrence, for example by explicitly clipping map to appropriate depths for corals, would result in maps that more closely approximate areas of occupancy.
+To achieve more comprehensive global coverage of species ranges it may be desirable to use these two datasets together, as the Ocean Health Index (OHI) does to calculate scores for its Species goal [@halpern_index_2012; @halpern_patterns_2015].
+While it may be unrealistic to "fix" one data set to match the other, we may be able to reduce the impact of the tradeoffs inherent in each.
+Trimming unsuitable habitat from the IUCN's extent of occurrence maps, for example by explicitly clipping them to appropriate depths for corals and reef-associated species, would result in maps that more closely approximate areas of occupancy.
 Conversely, including all AquaMaps cells with a non-zero "probability of occurrence" (rather than using a probability threshold to determine presence, e.g.
-greater than 40% for the Ocean Health Index [@halpern_index_2012; @halpern_patterns_2015] or 50% for the MPA gap analysis [@klein_shortfalls_2015]),  would result in maps that more closely, though still imperfectly, approximate extents of occurrence.
+greater than 40% for the Ocean Health Index [@halpern_index_2012; @halpern_patterns_2015] or 50% for the MPA gap analysis [@klein_shortfalls_2015]) would allow for the most generous inclusion of species range, resulting in maps that more closely, though still imperfectly, approximate extents of occurrence.
 

### 2016-09-12 ms_4_refs.Rmd
updated formats incl. references, figures, etc; split SI and figures into separate docs.  Also added bit about OBIS/GBIF
finished up draft; working on suppl figs

--- 2016-03-27 ms_5_refs.Rmd
@@
 # Acknowledgments
+
 <!-- Acknowledgments. List acknowledgments and funding sources. PNAS collaborates with FundRef to use a standard taxonomy of funder names. Dedications are rarely allowed. -->
+
+We are thankful to Melanie Frazier, Julia Stewart Lowndes, Halley Froelich, Claire Runge, and two anonymous reviewers for their insightful comments on earlier versions of this manuscript, and to the National Center for Ecological Analysis and Synthesis for computation support.
+We acknowledge financial support from the Gordon and Betty Moore Foundation.
+
+-----
 

### 2016-09-12 ms_6_supplemental.Rmd
finished up draft; working on suppl figs

@@
 
-methods:
-IUCN raw -> grid -> raster -> final
-AM raw -> presence
+<!-- methods: -->
+<!-- IUCN raw -> grid -> raster -> final -->
+<!-- AM raw -> presence -->
 
-results/discussion:
-quad plot with risk and four species (?)
-four representative maps
+<!-- results/discussion: -->
+<!-- quad plot with risk and four species (?) -->
+<!-- four representative maps -->
 
-coral quadrant shifts
+<!-- coral quadrant shifts -->
 
-AM thresholds? not mentioned in body of paper as of yet
+<!-- AM thresholds? not mentioned in body of paper as of yet -->
 
@@
 
-__Fig S1.__  Examples of AquaMaps and IUCN raw species range data.
+![](figures/Fig_S1ab.png)
+
+![](figures/Fig_S1cd.png)
+
+> __Fig S1.__  Examples of AquaMaps and IUCN raw species range data.
 (A, B) AquaMaps distribution of _Thunnus alalunga_ (Albacore Tuna) (AquaMaps 2013e) showing global range with varying probabilities of occurrence assigned to 0.5° grid cells (A) and presence within 0.5° grid cells (B).
@@
 
-<!-- 	
-Fis-22832, Thunnus alalunga, 21856, citation general IUCN; 
-Reviewed distribution maps for Thunnus alalunga (Albacore), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
-www.aquamaps.org, version of Aug.
-2013.
-Web.
-Accessed 11 Sep.
-2016.
--->
+<!-- Fis-22832, Thunnus alalunga, 21856 -->
 
-__Fig S2.__ Rasterizing shapefiles provided by IUCN.
+![](figures/Fig_S2a.png)
+
+![](figures/Fig_S2b.png)
+
+![](figures/Fig_S2c.png)
+
+![](figures/Fig_S2d.png)
+
+> __Fig S2.__ Rasterizing shapefiles provided by IUCN.
 A portion of the _T.
@@
 
-__Fig S3.__ Representative species maps (AquaMaps 2013abcd; Conch 2013ab; IUCN 2011a; IUCN 2012) to illustrate each quadrant from Fig 2A.  Each map is positioned to match its quadrant in Fig 2A.
+![](figures/Fig_S3.png)
+
+> __Fig S3.__ Representative species maps (AquaMaps 2013abcd; Conch 2013ab; IUCN 2011a; IUCN 2012) to illustrate each quadrant from Fig 2A.
+Each map is positioned to match its quadrant in Fig 2A.
+FAO Major Fishing Area boundaries [@fao_fisheries_and_aquaculture_department_cwp_2015] are outlined in light grey.
 (A) Distribution-matched: _Conus episcopatus_, the dignified cone snail.
@@
 <!--
-W-Msc-215455, Conus episcopatus, 192664 citation Conch Books, Hackenheim, Germany; Computer generated distribution maps for Conus episcopatus, with modelled year 2100 native range map based on IPCC A2 emissions scenario.
-www.aquamaps.org, version of Aug.
-2013.
-Web.
-Accessed 11 Sep.
-2016.
+W-Msc-215455, Conus episcopatus, 192664
+W-Msc-215551, Conus magnificus, 192573
+Fis-31079, Acanthurus nigroris, 177999
+Fis-164253, Kajikia albida, 170322
+-->
 
-W-Msc-215551, Conus magnificus, 192573 citation Conch Books, Hackenheim, Germany; Computer generated distribution maps for Conus magnificus, with modelled year 2100 native range map based on IPCC A2 emissions scenario.
-www.aquamaps.org, version of Aug.
-2013.
-Web.
-Accessed 11 Sep.
-2016.
+## Shifts in coral species range map alignment due to depth clipping
 
-Fis-31079, Acanthurus nigroris, 177999 citation IUCN; Reviewed distribution maps for Acanthurus nigroris (Bluelined surgeonfish), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
-www.aquamaps.org, version of Aug.
-2013.
-Web.
-Accessed 11 Sep.
-2016.
+![](figures/Fig_S4.png)
 
-Fis-164253, Kajikia albida, 170322 citation IUCN; Computer generated distribution maps for Kajikia albida (Atlantic white marlin), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
-www.aquamaps.org, version of Aug.
-2013.
-Web.
-Accessed 11 Sep.
-2016.
--->
+> __Fig.
+S4.__  Shift in alignment of paired-map coral species due to clipping IUCN ranges to areas shallower than 200 m.
+The grey lines represent the change in apparent alignment for a single species.
+Most coral species shift rightward from the upper left quadrant to the upper right, with little if any loss in distribution alignment, since in general, only unsuitable habitat has been removed.
+Leftward shifts can be seen in species whose larger original range is represented in AquaMaps; by trimming IUCN ranges, the area ratio becomes smaller.
 

### 2016-09-12 figures/si_figs_JA/supp_figs_JA.R

@@ -1,0 +1,130 @@
+library(raster)
+# library(maps)
+library(maptools)
+library(tmap)
+library(readr)
+
+source('~/github/ohiprep/src/R/common.R')
+
+
+dir_git <- '~/github/IUCN-AquaMaps'
+
+data("wrld_simpl", package = "maptools")
+data(World)
+
+ohi  <- readOGR(dsn = file.path(dir_M, 'git-annex/globalprep/spatial/d2014/data'), layer = 'regions_gcs')
+land <- ohi %>%
+  subset(rgn_typ == 'land')
+
+sp <- "Thunnus alalunga"
+setwd('shiny_am_iucn') ### setwd because there are some relative paths within server_fxns.R
+source('server_fxns.R')
+setwd('../')
+
+spp_list <- read_csv('shiny_am_iucn/data/spp_list.csv')
+
+# get rasters for tuna from IUCN and AM
+
+df     <- get_spp_map_df(sp)
+r_am   <- get_rast(df, type = 'am')
+r_iucn <- get_rast(df, type = 'iucn')
+
+plot(r_iucn)
+#playing with grids
+grid(nx = 720, ny = 360, col = 'lightgray')
+
+#crop so we can zoom in and show grid cells
+crop_ext    <- c(-80, -50,30,50)
+r_iucn_crop <- crop(r_iucn, crop_ext)
+r_am_crop   <- crop(r_am, crop_ext)
+land_crop   <- crop(land, crop_ext) %>%
+  subset(rgn_nam %in% c('Canada', 'United States'))
+plot(land_crop)
+
+
+basemap_simpl <- crop(wrld_simpl, crop_ext)
+basemap       <- crop(World, crop_ext)
+
+#pretty up with tmap
+tuna_iucn_raster_grid <- tm_shape(r_iucn_crop) +
+  tm_raster(palette = 'aquamarine3',
+            colorNA = NULL,
+            alpha = .8,
+            legend.show = FALSE) +
+  tm_layout(basemaps = "Esri.WorldTopoMap", 
+            # title.position = 'TOP', 
+            legend.outside = TRUE, attr.outside = TRUE)+
+  tm_grid(n.x = 60, n.y = 80, projection = 'longlat', col = 'gray', 
+          labels.inside.frame = FALSE,labels.size=0)+
+  tm_shape(land_crop) +
[…]

### 2016-09-12 figures/supp_figs_JA.R

@@ -1,128 +1,0 @@
-# plot for supplement
-
-library(raster)
-library(rgdal)
-library(maps)
-library(readr)
-library(maptools)
-library(tmap)
-
-data("wrld_simpl", package="maptools")
-data(World)
-ohi  <- readOGR(dsn = file.path(dir_M,'git-annex/globalprep/spatial/d2014/data'), layer = 'regions_gcs')
-land <- ohi%>%subset(rgn_typ=='land')
-
-
-source('~/github/ohiprep/src/R/common.R')
-
-sp <- "Thunnus alalunga"
-setwd('shiny_am_iucn')
-source('server_fxns.R')
-setwd('../')
-
-spp_list <- read.csv('shiny_am_iucn/data/spp_list.csv')
-
-# get rasters for tuna from IUCN and AM
-
-df     <- get_spp_map_df(sp)
-r_am   <- get_rast(df,type = 'am')
-r_iucn <- get_rast(df,type = 'iucn')
-
-plot(r_iucn)
-#playing with grids
-grid(nx = 720, ny = 360, col = 'lightgray')
-
-#crop so we can zoom in and show grid cells
-crop_ext    <- c(-80, -50,30,50)
-r_iucn_crop <- crop(r_iucn, crop_ext)
-r_am_crop   <- crop(r_am, crop_ext)
-land_crop   <- crop(land, crop_ext)%>%subset(rgn_nam %in% c('Canada','United States'))
-plot(land_crop)
-
-
-basemap_simpl <- crop(wrld_simpl, crop_ext)
-basemap       <- crop(World, crop_ext)
-
-#pretty up with tmap
-tuna_iucn_raster_grid <- tm_shape(r_iucn_crop) +
-                              tm_raster(palette = 'aquamarine3',
-                                        colorNA = NULL,
-                                        alpha = .8,
-                                        legend.show = FALSE) +
-                              tm_layout(basemaps = "Esri.WorldTopoMap", 
-                                        # title.position = 'TOP', 
-                                        legend.outside = TRUE, attr.outside = TRUE)+
-                              tm_grid(n.x = 60, n.y = 80, projection = 'longlat', col = 'gray', 
-                                      labels.inside.frame = FALSE,labels.size=0)+
-                                  tm_shape(land_crop) +
-                                  tm_polygons() +
-                                  tm_fill(col = 'gray')
-
[…]

### 2016-09-14 ms_1_intro.Rmd
adding SI figs from Jamie

@@
 
-The fundamental differences between the datasets suggest that the choice of one over the other should be carefully matched to the purpose for which it is to be used [@gaston_sizes_2009], and yet these datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
+The fundamental differences between the datasets suggest that the choice of one over the other should be carefully matched to the purpose for which it is to be used [@gaston_sizes_2009], and yet these datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
 In most of these cases, the implications of choosing one versus the other of these datasets is not evaluated or discussed, yet strong conservation and management conclusions are drawn from the results of the studies.

### 2016-09-14 ms_3_res_disc_concl.Rmd
adding SI figs from Jamie

@@
 AquaMaps encompasses a broader range of taxa than IUCN, as IUCN spatial data files are only available for select taxonomic groups that have been comprehensively assessed.
+Of the `r format(n_spp_all, big.mark = ',')` species mapped within these datasets, only `r format(n_spp_both, big.mark = ',')` (`r p_both_all`%) are mapped within both, with many taxa completely unrepresented in one dataset or the other.
 While species numbers in both datasets peak in tropical latitudes near the equator, species counts for IUCN maps drop quickly beyond 30°N and 30°S, while AquaMaps includes distribution of species well into temperate latitudes.
 The longitude frequency plots show a slight shift in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
+Together, the limitations of spatial and taxonomic coverage are likely to drive a researcher's choice of dataset far more strongly than the quality, format, or intended purpose of the dataset.
 
@@
 No spatial dataset can ever claim to know the "truth" of the whole of marine biodiversity.
-AquaMaps and IUCN range maps show strong agreement for many well-studied species, but for many others, substantial differences arise from differences in methodology and intent of each dataset.
-Although other approaches exist for predicting species distribution, these two are the largest approaches applied globally across a broad range of marine taxa.
-Recognizing and acknowledging the advantages and intended purposes of the range maps presented by these datasets will increase their utility for research and conservation actions.
+Although other approaches exist for predicting species distribution, AquaMaps and IUCN range maps are the largest approaches applied globally across a broad range of marine taxa.
+These datasets show strong agreement for many well-studied species, but for many others, substantial differences arise from differences in methodology and intent of each dataset.
+While the decision of which dataset to use should ideally be driven by the intended purpose for which it was created, the fact is that taxonomic coverage will likely be a more important factor.
+Recognizing and acknowledging the advantages and differences of the range maps presented by these datasets will increase their utility for research and conservation actions.
 Conclusions drawn from each of these datasets could paint dramatically different pictures of global marine biodiversity or the effectiveness of conservation management decisions.
@@
 Effective management and protection of marine species depends on a robust understanding of where species exist and where they do not; without this knowledge we risk wasting resources protecting low-value regions while missing opportunities to protect critical ones.
-By understanding the distinctions between these two important and fundamentally different marine species range datasets, we improve our ability to inform strategic and effective conservation policy that supports a resilient ocean ecosystem.
+By highlighting the distinctions between these two important and fundamentally different marine species range datasets, we improve our ability to inform strategic and effective conservation policy that supports a resilient ocean ecosystem.

### 2016-09-14 data_explore/mapping_aquamaps.R

@@ -1,0 +1,87 @@
+#load data for species map function, function, and plotting a species
+
+#JAfflerbach
+#9.11.2014
+
+## SETUP ##
+
+#libraries
+
+library(data.table)
+library(raster)
+library(rgdal)
+library(dplyr)
+library(RColorBrewer)
+library(maps)
+
+
+#increase R memory size to deal with reading in the large files
+
+memory.limit(size=200000000)
+
+
+dir_N = c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
+          'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
+
+dir_aquamaps = file.path(dir_N,"git-annex/globalprep/SPP_ICO/raw/aquamaps_2014")
+dir_cells    = file.path(dir_N,"git-annex/Global/NCEAS-SpatialFishCatch_v2014/raw/ohi_spp/data")
+
+#read in data
+#fread comes from data.table package and is much more efficient at reading in the larger .csv files
+
+spp       = fread(input=file.path(dir_aquamaps,"tables/ohi_speciesoccursum.csv"),header=T)%>%
+              mutate(scientific = paste(Genus,Species,sep=" "))
+
+cells     = read.csv(file.path(dir_cells,'cells.csv'))%>%
+              dplyr::select(cid,csquarecod,faoaream)
+    names(cells)<-c("cid","csquare_code","faoaream")
+
+spp_cells = fread(input=file.path(dir_aquamaps,"tables/ohi_hcaf_species_native.csv"),header=T,
+                  colClasses=c('NULL','character',NA,'numeric','NULL','NULL')) #1 min 49 sec
+
+setnames(spp_cells,c('SpeciesID','CsquareCode','probability'),c("species_id","csquare_code","probability"))
+
+
+#convert spp_cells, which is both a 'data.table' and a 'data.frame' to just a 'data.frame' to allow for plotting. A weird
+# caveat of fread() is that the output is both class 'data.table' and 'data.frame'
+
+spp_cells = as.data.frame(spp_cells)
+
+
+### Species Map Function ###
+# This function takes a single species scientific name as input, then grabs all occurrence cells and associated probability per cell
+
+sp_map_fun <- function(species){
+  
+  sp_id  = filter(spp,scientific==species)$SPECIESID
+  sp_map = filter(spp_cells,species_id==sp_id)%>%
[…]

### 2016-09-14 figures/si_figs_JA/supp_figs_JA1.R

@@ -1,0 +1,229 @@
+library(raster)
+library(rgdal)
+# library(maps)
+library(maptools)
+library(tmap)
+library(readr)
+
+source('~/github/ohiprep/src/R/common.R')
+
+
+dir_git <- '~/github/IUCN-AquaMaps'
+
+data("wrld_simpl", package = "maptools")
+data(World)
+
+
+sp <- "Thunnus alalunga"
+setwd('shiny_am_iucn') ### setwd because there are some relative paths within server_fxns.R
+source('server_fxns.R')
+setwd('../')
+
+spp_list <- read_csv('shiny_am_iucn/data/spp_list.csv')
+
+### get rasters for tuna from IUCN and AM
+
+df     <- get_spp_map_df(sp)
+tuna_rast_am   <- get_rast(df, type = 'am')
+tuna_rast_iucn <- get_rast(df, type = 'iucn')
+
+
+### crop so we can zoom in and show grid cells
+crop_ext    <- c(-80, -50, 30, 50)
+tuna_rast_iucn_crop <- crop(tuna_rast_iucn, crop_ext)
+tuna_rast_am_crop   <- crop(tuna_rast_am, crop_ext)
+
+# ohi  <- readOGR(dsn = file.path(dir_M, 'git-annex/globalprep/spatial/d2014/data'), layer = 'regions_gcs')
+# land <- ohi %>%
+#   subset(rgn_typ == 'land')
+# land_crop   <- crop(land, crop_ext) %>%
+#   subset(rgn_nam %in% c('Canada', 'United States'))
+# writeOGR(land_crop, dsn = 'figures/si_figs_JA', layer = 'land_crop', driver = 'ESRI Shapefile')
+land_crop      <- readOGR(dsn = 'figures/si_figs_JA', layer = 'land_crop')
+
+# tuna_poly <- readOGR(dsn = file.path(dir_M, 'git-annex/globalprep/_raw_data/iucn_spp/d2015/iucn_shp'),
+#                      layer = 'TUNAS_BILLFISHES')
+# tuna_poly <- tuna_poly %>% 
+#   subset(binomial == 'Thunnus alalunga')
+# tuna_poly_crop <- crop(tuna_poly, crop_ext)
+# extent(tuna_poly_crop) <- crop_ext
+# writeOGR(tuna_poly, dsn = 'figures/si_figs_JA', layer = 'tuna_poly', driver = 'ESRI Shapefile')
[…]

### 2016-09-14 clip_depth/coral_maps.Rmd
bumping up text size on figures

@@
     tm_raster(breaks = c(0, 200, Inf),
-              palette = c('lightblue', 'royalblue4'),
+              palette = c('lightblue', 'cyan4'),
               colorNA = NULL,
@@
   tm_layout(basemaps = "Esri.WorldTopoMap", 
-            legend.text.size = .5,
-            legend.title.size = .6,
+            legend.text.size = .8,
+            legend.title.size = .9,
             # title.position = 'TOP', 

### 2016-09-14 figures/si_figs_JA/supp_figs_JA1.R

@@ -63,8 +63,8 @@ tuna_iucn_raster <- tm_shape(tuna_rast_iucn_crop) +
             legend.show = FALSE) +
   tm_shape(land_crop) +
   tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
-  tm_layout(legend.text.size = .5,
-            legend.title.size = .6,
+  tm_layout(legend.text.size = .8,
+            legend.title.size = .9,
             # title.position = 'TOP', 
             legend.outside = FALSE, 
             legend.position = c('left', 'bottom'),
@@ -209,8 +209,8 @@ tuna_am_combine <- tm_shape(rast_am_prob_half) +
             alpha = .8,
             legend.show = TRUE,
             title = "Probability of Occurrence") +
-  tm_layout(legend.text.size = .5, #Casey's default from coral maps was .5 - pretty small for this map; Jamie changed to .8
-            legend.title.size = .6, #default from coral maps was .6; Jamie changed to .9
+  tm_layout(legend.text.size = .8,
+            legend.title.size = .9,
             # title.position = 'TOP', 
             legend.outside = FALSE, 
             legend.position = c('left', 'bottom'),

### 2016-09-14 clip_depth/coral_maps.Rmd
updated figs for text size, etc

@@
   scale_y_continuous(expand = c(0, 0), 
-                     limits = c(0, 1.1),
+                     limits = c(0, 1),
                      breaks = break_nums/100,

### 2016-09-14 clip_fao/fao_clip_pac.Rmd
updated figs for text size, etc

@@
             frame = TRUE,
-            legend.text.size = .5,
-            legend.title.size = .6,
+            legend.text.size = .6,
+            legend.title.size = .7,
             # title.position = 'TOP', 
@@
     tm_raster(palette = 'Reds',
-              breaks = c(1, 15, 100, 500, 1000, 2000, Inf),
-              colorNA = 'lightblue1',
+              breaks = c(1, 15, 100, 1000, 2000, Inf),
+              colorNA = '#ddeeff',
               textNA = 'Zero',
@@
             legend.bg.color = 'white',
-            legend.bg.alpha = 1,
+            legend.bg.alpha = .95,
             attr.outside = TRUE)
@@
           width = 19, 
-          # height = 6, 
           units = 'cm', dpi = 600)

### 2016-09-14 data_explore/data_explore.Rmd
all text updated, including references.  Still need to complete abstract.
updated figs for text size, etc

@@
 
-if(basename(getwd()) != 'data_explore') setwd('data_explore')
-source('data_explore_fxns.R')
+# if(basename(getwd()) != 'data_explore') setwd('data_explore')
+source(file.path(dir_git, 'data_explore', 'data_explore_fxns.R'))
 ### When knitting this, it automatically sets WD to be this directory...
@@
                  position = position_dodge(width = 0.5)) +
-  scale_size_manual(values = c('IUCN' = .5, 'AquaMaps' = .5, 'Paired' = 3)) +
+  scale_size_manual(values = c('IUCN' = .5, 'AquaMaps' = .5, 'Paired' = 2.9)) +
   scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey70', 'Paired' = 'black')) +
   geom_text(aes(label = n_label, y = 102),
-            size = 2,
+            size = 2.5,
             angle = 90, hjust = 0, color = 'grey30') +
@@
   scale_y_continuous(expand = c(0, 0),
-                     limits = c(0, 130),
+                     limits = c(0, 140),
                      breaks = break_nums,
@@
 
-ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
-       width = 19.05, height = 9, units = 'cm', dpi = 600)
+# ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
+#        width = 19.05, height = 9, units = 'cm', dpi = 600)
 
@@
             aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
-            size = 2, 
+            size = 2.5, 
             color = 'grey30') +
@@
 
-ggsave(plot = barchart_spp_gp_quads,
-       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
-       width = 19.05, height = 7, units = 'cm', dpi = 600)
+# ggsave(plot = barchart_spp_gp_quads,
+#        filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
+#        width = 19.05, height = 7, units = 'cm', dpi = 600)
 

### 2016-09-14 data_explore/data_explore_quadmaps.Rmd
updated figs for text size, etc

@@
     spp_map <- tm_shape(spp_rast) +
-      tm_raster(palette = c('#4dac26', '#d01c8b', '#4d4dac'),
+      tm_raster(palette = c('#4dac26', '#d01c8b', '#88ccff'),
                 title  = 'Dataset',
@@
                 showNA = FALSE,
-                alpha = 1)
+                alpha = .8)
   } else {
     spp_map <- tm_shape(spp_rast) +
-      tm_raster(palette = c('#4dac26', '#d01c8b', '#4d4dac'),
+      tm_raster(palette = c('#4dac26', '#d01c8b', '#aaccff'),
                 legend.show = FALSE,
                 colorNA = '#f0f8ff',
-                alpha = 1)
+                alpha = .8)
   }
@@
               frame = FALSE,
-              legend.text.size = .5,
-              legend.title.size = .6,
+              legend.text.size = .6,
+              legend.title.size = .7,
               legend.outside = FALSE,
@@
               legend.bg.color = '#f0f8ff',
-              legend.bg.alpha = .5,
+              legend.bg.alpha = .9,
               attr.outside = TRUE)

### 2016-09-14 data_explore/data_explore_fxns.R

@@ -4,7 +4,7 @@ library(ggplot2)
 
 ### generic theme for all plots
 ggtheme_basic <- theme(axis.ticks = element_blank(),
-        text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
+        text = element_text(family = 'Helvetica', color = 'gray30', size = 9),
         plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
         panel.background = element_blank(),
         legend.position = 'right')
@@ -14,6 +14,7 @@ ggtheme_plot <- ggtheme_basic +
         panel.grid.minor = element_blank(), 
         # panel.grid.major = element_line(colour = 'grey90', size = .25),
         panel.grid.major = element_blank(),
+        legend.key = element_rect(colour = NA, fill = NA),
         axis.line = element_line(colour = "grey30", size = .5))
 
   

### 2016-09-20 case_study_mpa/3_mpa_spp_analysis.Rmd
Created maps for a few species with suspicious borders; these include circles indicating known occurrences via obis.

@@
   geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
-  scale_y_continuous(expand = c(0, 0), limits = c(0, 125), breaks = break_nums,
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 112), breaks = break_nums,
                      labels = sprintf('%s%%', break_nums)) + 
   geom_text(aes(label = sprintf('n = %s', n_trial), y = 101), hjust = 0, 
-            size = 2, 
+            size = 2.5, 
             color = 'grey30') +
@@
        width = 19.05, height = 5, units = 'cm', dpi = 600)
-# ggsave(plot = barchart_mpa_gap,
-#        filename = file.path(dir_fig, 'barchart_mpa_gap.tif'), device = 'tiff',
-#        width = 11.4, height = 5, units = 'cm', dpi = 600)
 

### 2016-09-20 clip_fao/fao_clip_pac.Rmd
Created maps for a few species with suspicious borders; these include circles indicating known occurrences via obis.

@@
 
+``` {r plot_several_spp_w, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))
+
+spp_examples_w <- clipped_spp %>%
+  filter(lim == 'e_lim')
+
+spp_w_cells <- am_spp_cells %>%
+  filter(am_sid %in% spp_examples_w$am_sid)
+
+spp_w_all <- spp_w_cells %>%
+  group_by(am_sid) %>%
+  summarize(n_cells = n())
+
+spp_w_samples <- spp_w_all %>%
+  filter(n_cells > 9000) %>% ### focus on large-range species
+  sample_n(20) %>%
+  .$am_sid
+
+print(spp_w_all %>% filter(am_sid %in% spp_w_samples))
+### some good candidates:
+#     am_sid n_cells
+# Fis-143854   23422
+#  Fis-34036   13796
+# URM-2-3539   16664
+#  Fis-33989   15715
+spp_plotted <- read_csv('data/spp_all.csv') %>%
+  filter(am_sid %in% spp_w_samples) %>%
+  left_join(read_csv('data/speciesoccursum.csv') %>%
+              select(am_sid = speciesid, occurcells, fbname),
+            by = 'am_sid')
+
+for(spp in spp_w_samples) {
+  # spp <- spp_w_samples[1]
+  message('Processing map for ', spp)
+  message('... filtering cells for ', spp)
+  spp_cells <- spp_w_cells %>%
+    filter(am_sid == spp) %>%
+    mutate(presence = 1)
+  rast_spp  <-  subs(loiczid_raster, 
+                     spp_cells[ , c('loiczid', 'presence')], 
+                     by = 'loiczid', 
+                     which = 'presence', 
+                     subsWithNA = TRUE)
+
+  message('... shifting raster and setting 0 to NA for ', spp)
+  rast_spp_shift <- rast_spp %>%
+    raster::shift(360)
+
+  rast_spp_w <- merge(rast_spp, rast_spp_shift)
+
+  values(rast_spp_w)[values(rast_spp_w) == 0] <- NA
+  # values(rast_spp_w)[is.na(values(rast_spp_w))] <- 0
+  rast_spp_w  <- crop(rast_spp_w, rast_ext)
+
+  ### boundary line
+  fao_boundary <- Line(matrix(c(185, 185, -25, 20), ncol = 2)) %>%
+    Lines(ID = 'a') %>%
+    list() %>%
+    SpatialLines()
+
+  message('... generating map for ', spp)
+
+  rast_spp_map_w <- 
+    tm_shape(rast_spp_w) +
+      tm_raster(palette = 'Reds',
+                colorNA = '#ddeeff',
+                textNA = 'not present',
+                title = spp,
+                alpha = 1) +
+    tm_shape(World1) +
+      tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
+    tm_shape(fao_polys) +
+      tm_borders(col = 'grey40', lwd = 0.25) +
+      tm_text('F_AREA_1',
+              col = 'royalblue4',
+              fontface = 'bold.italic',
+              fontfamily = 'Helvetica',
+              size = .5) +
+    tm_shape(fao_boundary) +
+      tm_lines(col = 'blue', lwd = 1.5) +
+    tm_layout(# basemaps = "Esri.WorldTopoMap", 
+              frame = TRUE,
+              legend.text.size = .6,
+              legend.title.size = .7,
+              # title.position = 'TOP', 
+              legend.outside = FALSE,
+              legend.position = c('right', 'top'),
+              legend.bg.color = 'white',
+              legend.bg.alpha = .95,
+              attr.outside = TRUE)
+
+  print(rast_spp_map_w)
+}
+
+# save_tmap(rast_spp_map_w, 
+#           file.path(dir_fig, 'map_am_edge_w.png'), 
+#           width = 19, 
+#           units = 'cm', dpi = 600)
+
+```
+
+``` {r plot_example_spp, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+spp_ex <- c('Hoplichthys regani' = 'Fis-34036', 
+            'Ipnops meadi'       = 'Fis-143854', 
+            'Culeolus herdmani'  = 'URM-2-3539', 
+            'Chrionema chlorotaenia' = 'Fis-33989')
+
+# spp_ex_cells <- am_spp_cells %>%
+#   filter(am_sid %in% spp_ex)
+# 
+# spp_ex_info <- read_csv('data/spp_all.csv') %>%
+#   filter(am_sid %in% spp_ex)
+
+for(i in 1:length(spp_ex)) {
+  spp <- spp_ex[i]
+  spp_name <- names(spp_ex)[i]
+
+  message('Processing map for ', spp)
+
+  message('... loading occurrences for ', spp)
+  spp_occur <- read_csv(paste0('clip_fao/obs_for_', tolower(spp) %>% 
+                                 str_replace_all('-', '_'), '.csv')) %>%
+    select(id, long = decimalLongitude, lat = decimalLatitude) %>%
+    mutate(long = ifelse(long < 0, long + 360, long))
+  spp_occur_pts <- SpatialPoints(spp_occur[ , 2:3])
+
+  message('... creating raster for ', spp)
+  spp_cells <- spp_ex_cells %>%
+    filter(am_sid == spp) %>%
+    mutate(presence = 1)
+  rast_spp  <-  subs(loiczid_raster, 
+                     spp_cells[ , c('loiczid', 'presence')], 
+                     by = 'loiczid', 
+                     which = 'presence', 
+                     subsWithNA = TRUE)
+
+  message('... shifting raster and setting 0 to NA for ', spp)
+  rast_spp_shift <- rast_spp %>%
+    raster::shift(360)
+
+  rast_spp <- merge(rast_spp, rast_spp_shift)
+
+  values(rast_spp)[values(rast_spp_w) == 0] <- NA
+  # values(rast_spp_w)[is.na(values(rast_spp_w))] <- 0
+  rast_spp  <- crop(rast_spp, rast_ext)
+
+  ### boundary line
+  fao_boundary <- Line(matrix(c(185, 185, -25, 20), ncol = 2)) %>%
+    Lines(ID = 'a') %>%
+    list() %>%
+    SpatialLines()
+
+  message('... generating map for ', spp)
+
+  rast_spp_map <- 
+    tm_shape(rast_spp) +
+      tm_raster(palette = 'Reds',
+                colorNA = '#ddeeff',
+                textNA  = 'not present',
+                title   = spp_name,
+                legend.show = FALSE,
+                showNA  = FALSE,
+                alpha   = 1) +
+    tm_shape(World1) +
+      tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
+    tm_shape(fao_polys) +
+      tm_borders(col = 'grey40', lwd = 0.25) +
+      tm_text('F_AREA_1',
+              col = 'royalblue4',
+              fontface = 'bold.italic',
+              fontfamily = 'Helvetica',
+              size = .5) +
+    tm_shape(fao_boundary) +
+      tm_lines(col = 'blue', lwd = 1.5) +
+    tm_shape(spp_occur_pts) + 
+      tm_bubbles(col = NA,
+                 border.col = 'red',
+                 border.lwd = 0.25,
+                 border.alpha = 1,
+                 size = 1,
+                 scale = .3,
+                 alpha = .2) +
+      # tm_dots(col = 'red') +
+    tm_layout(title = spp_name,
+              title.position = c('left', 'bottom'),
+              fontface = 'bold.italic',
+              fontfamily = 'Helvetica',
+              title.size = .8,
+              frame = TRUE,
+              legend.text.size = .6,
+              legend.title.size = .7,
+              legend.outside = FALSE,
+              legend.position = c('right', 'top'),
+              legend.bg.color = 'white',
+              legend.bg.alpha = .95,
+              attr.outside = TRUE)
+
+  print(rast_spp_map)
+
+  message('... saving map')
+  save_tmap(rast_spp_map,
+            paste0('clip_fao/map_fao_clip_', tolower(spp_name) %>% 
+                     str_replace_all(' ', '_'), '.png'),
+            width = 19,
+            units = 'cm', dpi = 600)
+
+}
+
+# save_tmap(rast_spp_map_w, 
+#           file.path(dir_fig, 'map_am_edge_w.png'), 
+#           width = 19, 
+#           units = 'cm', dpi = 600)
+
+```
+
 ``` {r combine_rasts}

### 2016-09-20 ms_3_res_disc_concl.Rmd
Created maps for a few species with suspicious borders; these include circles indicating known occurrences via obis.

@@
 
-Comparing distribution alignment and area ratio for the `r format(n_spp_both, big.mark = ',')` paired map species (Fig 2A), a weak negative linear pattern appears to emerge, suggesting that increasing similarity in range area correlates very slightly with decreasing distribution alignment (R^2 = .016).
+Comparing distribution alignment and area ratio for the `r format(n_spp_both, big.mark = ',')` paired map species (Fig 2A), a weak negative linear pattern appears to emerge, suggesting that increasing similarity in range area correlates very slightly with decreasing distribution alignment (R^2^ = .016).
 The pattern itself is not particularly important, and emerges simply due to the nature of the analysis and the datasets:  The AquaMaps model tends to extrapolate species ranges into suitable areas beyond known occurrences more frequently than IUCN maps, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.

### 2016-09-20 clip_fao/obis_occurrences.R

@@ -1,0 +1,134 @@
+### support functions for obis_count()
+library(httr)
+library(jsonlite)
+
+obis_url <- function() {
+  getOption("robis_url", "http://api.iobis.org/")
+}
+
+max_characters <- function() {
+  getOption("robis_max_characters", 50000)
+}
+
+handle_date <- function(date) {
+  if(!is.null(date) && class(date) == "Date") {
+    as.character(date)
+  } else {
+    date
+  }
+}
+
+handle_vector <- function(x) {
+  if(!is.null(x)) {
+    paste0(x, collapse = ",")
+  } else {
+    x
+  }
+}
+
+log_request <- function(result) {
+  cat("\n", paste(result$request$method, result$request$url, result$status_code, result$headers$age, result$times[["total"]]), "\n", sep="")
+}
+
+log_progress <- function(total, count) {
+  cat("\rRetrieved ", total, " records of ", count, " (", floor(total/count*100),"%)", sep="")
+}
+
+http_request <- function(method, path, query) {
+  if (method == "GET") {
+    httr::GET(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
+              path = path, query = query)
+  } else if (method == "POST") {
+    httr::POST(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
+               path = path, body = query)
+  }
+}
+
+obis_occurrence <- function(
+  scientificname = NULL,
+  year = NULL,
+  obisid = NULL,
+  aphiaid = NULL,
+  resourceid = NULL,
+  startdate = NULL,
+  enddate = NULL,
+  startdepth = NULL,
+  enddepth = NULL,
+  geometry = NULL,
+  qc = NULL,
+  fields = NULL,
+  verbose = FALSE) {
+  
+  
+  offset <- 0
+  i <- 1
+  lastpage <- FALSE
+  total <- 0
+  datalist <- list()
+  
+  t <- proc.time()
+  
+  while (!lastpage) {
+    query <- list(scientificname = scientificname,
+                  year = year,
+                  obisid = obisid,
+                  aphiaid = aphiaid,
+                  resourceid = resourceid,
+                  # startdate = handle_date(startdate),
+                  # enddate = handle_date(enddate),
+                  startdepth = startdepth,
+                  enddepth = enddepth,
+                  geometry = geometry
+                  # qc = handle_vector(qc),
+                  # fields = handle_vector(fields),
[…]

### 2016-09-23 clip_depth/coral_maps.Rmd
wrapping up edits from JA, BH, CS updating fig captions

@@
   mutate(deep = 1) %>%
-  select(-weight)
+  dplyr::select(-weight)
 
@@
   bind_rows() %>%
-  select(iucn_sid, loiczid = LOICZID) %>%
+  dplyr::select(iucn_sid, loiczid = LOICZID) %>%
   # filter(iucn_sid %in% spp_depth$iucn_sid[spp_depth$paired_map]) %>%
@@
   spp_depth <- spp_narr %>%
-    select(iucn_sid = species_id, habitat) %>%
+    dplyr::select(iucn_sid = species_id, habitat) %>%
     mutate(paired_map = (iucn_sid %in% spp_map_pairs$iucn_sid))
@@
 
-rast_coral_comp <- raster::subs(x = loiczid_raster, y = coral_comp %>% select(loiczid, deep), 
+rast_coral_comp <- raster::subs(x = loiczid_raster, y = coral_comp %>% dplyr::select(loiczid, deep), 
                      by = 'loiczid', which = 'deep', 
@@
 hcaf <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
-  select(loiczid, cellarea, oceanarea)
+  dplyr::select(loiczid, cellarea, oceanarea)
 
@@
     filter(iucn_sid %in% coral_cells_iucn_shallow$iucn_sid) %>%
-    select(am_sid, sciname, iucn_sid)
+    dplyr::select(am_sid, sciname, iucn_sid)
 
@@
     sids <- spp_coralmaps %>% 
-      select(am_sid, iucn_sid, sciname) %>%
+      dplyr::select(am_sid, iucn_sid, sciname) %>%
       filter(iucn_sid == sid)
@@
 
-    r_am_spp <- subs(x = loiczid_raster, y = map_am %>% select(loiczid, am_pres), 
+    r_am_spp <- subs(x = loiczid_raster, y = map_am %>% dplyr::select(loiczid, am_pres), 
                      by = 'loiczid', which = 'am_pres', 
@@
     if(TRUE %in% duplicated(map_iucn$loiczid)) stop('Duplicated cells in IUCN map')
-    r_iucn_spp <- subs(loiczid_raster, map_iucn %>% select(loiczid, iucn_pres), 
+    r_iucn_spp <- subs(loiczid_raster, map_iucn %>% dplyr::select(loiczid, iucn_pres), 
                        by = 'loiczid', which = 'iucn_pres', 
@@
              ### use largest data set as approx for total ocean range
-    select(-lg_area, -sm_area)
+    dplyr::select(-lg_area, -sm_area)
 
@@
   spp_coralmaps <- spp_coralmaps %>%
-    select(iucn_sid, am_sid, sciname, area_am, 
+    dplyr::select(iucn_sid, am_sid, sciname, area_am, 
            area_iucn_shallow = area_iucn, area_overlap_shallow = area_overlap, 
@@
     left_join(read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
-                select(am_sid, iucn_sid, 
+                dplyr::select(am_sid, iucn_sid, 
                        area_iucn_raw = area_iucn, area_overlap_raw = area_overlap, 
@@
 spp_coral_clip <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_areas.csv')) %>%
-  select(iucn_sid, shallow_pct, deep_pct, totarea, shallow_area = area) %>%
+  dplyr::select(iucn_sid, shallow_pct, deep_pct, totarea, shallow_area = area) %>%
   mutate(shallow_pct = 100 * shallow_pct,
@@
 iucn_coral_quads <- iucn_coral_fixed %>%
-  select(iucn_sid, sciname, d_perc, a_ratio, method) %>% ### ditch am_sid, b/c duplicates
+  dplyr::select(iucn_sid, sciname, d_perc, a_ratio, method) %>% ### ditch am_sid, b/c duplicates
   distinct() %>%
@@
 quad_names <- data.frame('quad' = c('q1', 'q2', 'q3', 'q4'),
-                         'quad_name' = factor(c('well-matched', 'dist-matched', 
-                                                'area-matched', 'poorly matched'),
+                         'quad_name' = factor(c('well-aligned', 'dist-aligned', 
+                                                'area-aligned', 'poorly aligned'),
                                              ordered = TRUE))
@@
 iucn_coral_quad_sum <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_quads.csv')) %>%
-  select(method, quad, quad_name, pct_quad) %>%
+  dplyr::select(method, quad, quad_name, pct_quad) %>%
   distinct()
@@
   coord_flip() +
-  labs(x = 'Taxonomic Group', 
+  labs(x = 'Depth limit', 
        y = 'Percent of corals by quadrant', 
@@
 
-ggsave(plot = barchart_coral_quads,
-       filename = file.path(dir_fig, 'barchart_coral_quads.png'),
-       width = 19, height = 4, units = 'cm', dpi = 600)
+# ggsave(plot = barchart_coral_quads,
+#        filename = file.path(dir_fig, 'barchart_coral_quads.png'),
+#        width = 19, height = 4, units = 'cm', dpi = 600)
 

### 2016-09-23 clip_fao/fao_clip_pac.Rmd
wrapping up edits from JA, BH, CS updating fig captions

--- 2016-09-14 clip_fao/fao_clip_pac.Rmd
@@
 spp_quad_list <- read_csv(file.path(dir_git, 'data', 'spp_list_quads.csv')) %>%
-  select(sciname, am_sid, iucn_sid, reviewed, occurcells, quad) %>%
+  dplyr::select(sciname, am_sid, iucn_sid, reviewed, occurcells, quad) %>%
   filter(quad %in% c('q3', 'q4'))
@@
   filter(!duplicated(am_sid)) %>%
-  select(am_sid, sciname, reviewed) %>%
+  dplyr::select(am_sid, sciname, reviewed) %>%
   distinct()
@@
                            'csv/hcaf_truncated.csv')) %>%
-  select(-csquarecode, -cellarea, -oceanarea) %>%
+  dplyr::select(-csquarecode, -cellarea, -oceanarea) %>%
   mutate(clongadj  = ifelse(centerlong < 0, centerlong + 360, centerlong),
@@
 ``` {r atlantic_spp} 
-### identify species that appear in the mid-atlantic - focus just on pacific/indopacific spp only
+### identify (to cut) species that appear in the mid-atlantic - focus just on pacific/indopacific spp only
 
@@
   filter(str_detect(lim_n, lim)) %>%
-  select(-lim_n) %>%
+  dplyr::select(-lim_n) %>%
   filter(!(long <= 70 | long >= 200))
@@
   filter(long == 185) %>%
-  select(am_sid, lim) %>%
+  dplyr::select(am_sid, lim) %>%
   distinct() %>%
@@
 
-rast_ext  <- extent(c(-30, 330, -75, 75))
+rast_ext  <- extent(c(-30, 330, -63, 75)) ### was -75 but changed to -63 to cut more from bottom
 fao_polys <- crop(fao_polys, rast_ext)
@@
 World1 <- maptools::readShapePoly('clip_fao/fao_rgns/world1')
-
-rast_spp_shift <- rast_spp %>%
-  raster::shift(360)
-
-rast_spp_e <- merge(rast_spp, rast_spp_shift)
 
@@
 
-save_tmap(rast_spp_map_w, 
-          file.path(dir_fig, 'map_am_edge_w.png'), 
+# save_tmap(rast_spp_map_w, 
+#           file.path(dir_fig, 'map_am_edge_w.png'), 
+#           width = 19, 
+#           units = 'cm', dpi = 600)
+
+```
+
+``` {r plot_several_spp_w, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))
+
+spp_examples_w <- clipped_spp %>%
+  filter(lim == 'e_lim')
+
+spp_w_cells <- am_spp_cells %>%
+  filter(am_sid %in% spp_examples_w$am_sid)
+
+spp_w_all <- spp_w_cells %>%
+  group_by(am_sid) %>%
+  summarize(n_cells = n())
+
+spp_w_samples <- spp_w_all %>%
+  filter(n_cells > 9000) %>% ### focus on large-range species
+  sample_n(20) %>%
+  .$am_sid
+
+print(spp_w_all %>% filter(am_sid %in% spp_w_samples))
+### some good candidates:
+#     am_sid n_cells
+# Fis-143854   23422
+#  Fis-34036   13796
+# URM-2-3539   16664
+#  Fis-33989   15715
+spp_plotted <- read_csv('data/spp_all.csv') %>%
+  filter(am_sid %in% spp_w_samples) %>%
+  left_join(read_csv('data/speciesoccursum.csv') %>%
+              dplyr::select(am_sid = speciesid, occurcells, fbname),
+            by = 'am_sid')
+
+for(spp in spp_w_samples) {
+  # spp <- spp_w_samples[1]
+  message('Processing map for ', spp)
+  message('... filtering cells for ', spp)
+  spp_cells <- spp_w_cells %>%
+    filter(am_sid == spp) %>%
+    mutate(presence = 1)
+  rast_spp  <-  subs(loiczid_raster, 
+                     spp_cells[ , c('loiczid', 'presence')], 
+                     by = 'loiczid', 
+                     which = 'presence', 
+                     subsWithNA = TRUE)
+
+  message('... shifting raster and setting 0 to NA for ', spp)
+  rast_spp_shift <- rast_spp %>%
+    raster::shift(360)
+
+  rast_spp_w <- merge(rast_spp, rast_spp_shift)
+
+  values(rast_spp_w)[values(rast_spp_w) == 0] <- NA
+  # values(rast_spp_w)[is.na(values(rast_spp_w))] <- 0
+  rast_spp_w  <- crop(rast_spp_w, rast_ext)
+
+  ### boundary line
+  fao_boundary <- Line(matrix(c(185, 185, -25, 20), ncol = 2)) %>%
+    Lines(ID = 'a') %>%
+    list() %>%
+    SpatialLines()
+
+  message('... generating map for ', spp)
+
+  rast_spp_map_w <- 
+    tm_shape(rast_spp_w) +
+      tm_raster(palette = 'Reds',
+                colorNA = '#ddeeff',
+                textNA = 'not present',
+                title = spp,
+                alpha = 1) +
+    tm_shape(World1) +
+      tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
+    tm_shape(fao_polys) +
+      tm_borders(col = 'grey40', lwd = 0.25) +
+      tm_text('F_AREA_1',
+              col = 'royalblue4',
+              fontface = 'bold.italic',
+              fontfamily = 'Helvetica',
+              size = .5) +
+    tm_shape(fao_boundary) +
+      tm_lines(col = 'blue', lwd = 1.5) +
+    tm_layout(# basemaps = "Esri.WorldTopoMap", 
+              frame = TRUE,
+              legend.text.size = .6,
+              legend.title.size = .7,
+              # title.position = 'TOP', 
+              legend.outside = FALSE,
+              legend.position = c('right', 'top'),
+              legend.bg.color = 'white',
+              legend.bg.alpha = .95,
+              attr.outside = TRUE)
+
+  print(rast_spp_map_w)
+}
+
+# save_tmap(rast_spp_map_w, 
+#           file.path(dir_fig, 'map_am_edge_w.png'), 
+#           width = 19, 
+#           units = 'cm', dpi = 600)
+
+```
+
+``` {r plot_selected_spp, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+### Select four example species
+spp_ex <- c('Hoplichthys regani' = 'Fis-34036', 
+            'Ipnops meadi'       = 'Fis-143854', 
+            'Culeolus herdmani'  = 'URM-2-3539', 
+            'Chrionema chlorotaenia' = 'Fis-33989')
+
+spp_ex_cells <- am_spp_cells %>%
+  filter(am_sid %in% spp_ex)
+
+spp_ex_info <- read_csv('data/spp_all.csv') %>%
+  filter(am_sid %in% spp_ex)
+
+for(i in 1:length(spp_ex)) {
+  # i <- 1
+  spp <- spp_ex[i]
+  spp_name <- names(spp_ex)[i]
+
+  message('Processing map for ', spp)
+
+  message('... loading occurrences for ', spp)
+  spp_occur <- read_csv(paste0('clip_fao/obs_for_', tolower(spp) %>% 
+                                 str_replace_all('-', '_'), '.csv')) %>%
+    mutate(long = ifelse(long < 0, long + 360, long),
+           obs = toupper(obs),
+           obs = ifelse(obs == 'BOTH', 'Both', obs),
+           obs = factor(obs, levels = c('OBIS', 'GBIF', 'Both')))
+  spp_occur_pts <- SpatialPointsDataFrame(spp_occur[ , 1:2], data = spp_occur[ , 3])
+
+  message('... creating raster for ', spp)
+  spp_cells <- spp_ex_cells %>%
+    filter(am_sid == spp) %>%
+    mutate(presence = 1)
+  rast_spp  <-  subs(loiczid_raster, 
+                     spp_cells[ , c('loiczid', 'presence')], 
+                     by = 'loiczid', 
+                     which = 'presence', 
+                     subsWithNA = TRUE)
+
+  message('... shifting raster and setting 0 to NA for ', spp)
+  rast_spp_shift <- rast_spp %>%
+    raster::shift(360)
+
+  rast_spp <- merge(rast_spp, rast_spp_shift)
+
+  values(rast_spp)[values(rast_spp) == 0] <- NA
+  rast_spp  <- crop(rast_spp, rast_ext)
+
+  ### boundary line
+  fao_boundary <- Line(matrix(c(185, 185, -25, 20), ncol = 2)) %>%
+    Lines(ID = 'a') %>%
+    list() %>%
+    SpatialLines()
+
+  message('... generating map for ', spp)
+
+  rast_spp_map <- 
+    tm_shape(rast_spp) +
+      tm_raster(palette = 'Reds',
+                colorNA = '#ddeeff',
+                textNA  = 'not present',
+                # title   = spp_name,
+                legend.show = FALSE,
+                showNA  = FALSE,
+                alpha   = 1) +
+    tm_shape(World1) +
+      tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
+    tm_shape(fao_polys) +
+      tm_borders(col = 'grey40', lwd = 0.25) +
+      tm_text('F_AREA_1',
+              col = 'royalblue4',
+              fontface = 'bold.italic',
+              fontfamily = 'Helvetica',
+              size = .5) +
+    tm_shape(fao_boundary) +
+      tm_lines(col = 'blue', lwd = 1.5)
+
+  rast_map <- rast_spp_map +
+    tm_shape(spp_occur_pts) + 
+      tm_bubbles(col = 'obs',
+                 title.col = 'Occurrences ',
+                 palette = c('red', 'blue', 'green'),
+                 border.col = 'red',
+                 border.lwd = 0.25,
+                 border.alpha = 1,
+                 size = 1,
+                 scale = .3,
+                 alpha = .3) +
+      # tm_dots(col = 'red') +
+    tm_layout(# title = spp_name,
+              title.position = c('left', 'bottom'),
+              fontface = 'bold.italic',
+              fontfamily = 'Helvetica',
+              title.size = .8,
+              frame = TRUE,
+              legend.text.size = .6,
+              legend.title.size = .7,
+              legend.outside = FALSE,
+              legend.position = c('right', 'top'),
+              legend.bg.color = 'white',
+              legend.bg.alpha = .95,
+              attr.outside = TRUE)
+
+  print(rast_map)
+
+  message('... saving map')
+  save_tmap(rast_map,
+            paste0('clip_fao/map_fao_clip_', tolower(spp_name) %>% 
+                     str_replace_all(' ', '_'), '.png'),
             width = 19,
             units = 'cm', dpi = 600)
+
+}
+
+# save_tmap(rast_spp_map_w, 
+#           file.path(dir_fig, 'map_am_edge_w.png'), 
+#           width = 19, 
+#           units = 'cm', dpi = 600)
+
+```
+
+``` {r quad_spp_w, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
+                                   'v2016/int/am_cells_spp_prob0.csv'))
+
+am_spp_list <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>%
+  dplyr::select(am_sid = speciesid, genus, species, reviewed, occurcells) %>%
+  mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' ')) %>%
+  dplyr::select(-genus, -species) %>%
+  distinct()
+
+hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
+  dplyr::select(loiczid, cellarea, oceanarea)
+
+am_spp_cells2 <- hcaf %>%
+  left_join(am_spp_cells, by = 'loiczid')
+
+clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv')) %>%
+  filter(lim == 'e_lim') %>%
+  dplyr::select(am_sid)
+
+spp_quad_list <- read_csv(file.path(dir_data, 'spp_list_quads.csv'))
+
+spp_area_df <- am_spp_cells2 %>%
+  group_by(am_sid) %>%
+  summarize(cellarea = sum(cellarea),
+            oceanarea = sum(oceanarea))
+
+spp_area_df <- spp_area_df %>%
+  inner_join(am_spp_list, by = 'am_sid') %>%
+  mutate(area_per_cells = oceanarea/occurcells)
+
+head(spp_area_df)
+#                                     am_sid    cellarea   oceanarea reviewed occurcells                   sciname
+#                                      <chr>       <dbl>       <dbl>    <chr>      <int>                     <chr>
+# 1                                Biv-32126   949988.02   642706.64     null          4        Corbicula japonica
+# 2                                Biv-32832  9698932.54  7589672.98     null         23          Polymesoda erosa
+# 3                               Bra-235673    16143.75    13895.96        1         10       Dyscritosia secreta
+# 4 Chn-013ca97e-46a8-4c5e-b398-8bb4f340f88b 11288439.81 10048833.94     null         76          Lenisquilla lata
+# 5 Chn-015453c0-3ff9-477d-ab51-c68d85e65af5 15078672.39 12760166.60     null         13        Clorida albolitura
+# 6 Chn-0721ac0c-7d19-46fa-9dcc-02da54310867 35256268.05 33846838.22     null         23 Leiogalathea laevirostris
+summary(spp_area_df$oceanarea)
+   # Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
+   # 1806    946600   2921000  10210000   7875000 501300000 
+summary(spp_area_df$occurcells)
+   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+   # 0.00    6.00   16.00   57.13   49.00 7663.00 
+summary(spp_area_df %>% filter(occurcells != 0) %>% .$area_per_cells)
+    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
+    #  190    50540   158000   485300   439100 44700000 
+nrow(spp_area_df %>% filter(!str_detect(reviewed, 'null')))
+   # 1296 out of 22889 = 5.66%
+
+spp_area_clipped <- spp_area_df %>%
+  filter(am_sid %in% clipped_spp$am_sid)
+nrow(spp_area_clipped %>% filter(!str_detect(reviewed, 'null')))
+   # 22 out of 3208 = .686%
+
+summary(spp_area_clipped$oceanarea)
+   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
+   # 134900   4646000   7032000   8960000   9690000 111300000 
+summary(spp_area_clipped$occurcells)
+   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+   # 3.00    6.00   15.00   26.86   33.00  494.00 
+summary(spp_area_clipped$area_per_cells)
+    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
+    # 6578   202900   452400   827200   923300 26090000
+
+# quad_names <- data.frame('quad' = c('q1', 'q2', 'q3', 'q4'),
+#                          'quad_name' = factor(c('well-matched', 'dist-matched', 
+#                                                 'area-matched', 'poorly matched'),
+#                                              ordered = TRUE))
+# 
+# clipped_spp_quads <- clipped_spp_quads %>%
+#   left_join(quad_names, by = 'quad') %>%
+#   mutate(method = 'FAO clipped')
+# 
+# break_nums <- seq(0, 100, 20)
+
+### Plot the bar chart
+# barchart_clipped_spp_quads <- ggplot(clipped_spp_quads, 
+#                                 aes(x = method, fill = quad, weight = pct_quad)) +
+#   ggtheme_plot + 
+#   theme(panel.grid.major.y = element_blank()) +
+#   geom_bar(stat = 'count', alpha = 1) +
+#   scale_fill_manual(values = c('q4' = '#d01c8b', 
+#                                'q3' = '#f1b6da', 
+#                                'q2' = '#b8e186',
+#                                'q1' = '#4dac26'),
+#                     labels = quad_names$quad_name) +
+#   scale_y_continuous(expand = c(0, 0), 
+#                      limits = c(0, 1),
+#                      breaks = break_nums/100,
+#                      labels = sprintf('%s%%', break_nums)) + 
+#   ### add grid lines; horizontal but then get flipped
+#   geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
+#   coord_flip() +
+#   labs(x = '', 
+#        y = 'Percent of species by quadrant', 
+#        fill = 'Alignment')
+# 
+# print(barchart_clipped_spp_quads)
 
@@
   tm_layout(basemaps = "Esri.WorldTopoMap", 
-            frame = FALSE,
+            frame = TRUE,
             # title.position = 'TOP', 

### 2016-09-23 data_explore/data_explore.Rmd
wrapping up edits from JA, BH, CS updating fig captions

@@
             by = 'popn_category') %>%
-  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
-         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
-         quad = str_replace(quad, 'q3', 'q3 area-matched'),
-         quad = str_replace(quad, 'q4', 'q4 poorly matched')) %>%
+  mutate(quad = str_replace(quad, 'q1', 'q1 well-aligned'),
+         quad = str_replace(quad, 'q2', 'q2 dist-aligned'),
+         quad = str_replace(quad, 'q3', 'q3 area-aligned'),
+         quad = str_replace(quad, 'q4', 'q4 poorly aligned')) %>%
   left_join(spp_list %>%
@@
   geom_bar(stat = 'count', alpha = 1) +
-  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
-                               'q2 dist-matched'   = '#b8e186', 
-                               'q3 area-matched'   = '#f1b6da', 
-                               'q4 poorly matched' = '#d01c8b')) +
+  scale_fill_manual(values = c('q1 well-aligned'   = '#4dac26', 
+                               'q2 dist-aligned'   = '#b8e186', 
+                               'q3 area-aligned'   = '#f1b6da', 
+                               'q4 poorly aligned' = '#d01c8b')) +
   scale_y_continuous(expand = c(0, 0), 

### 2016-09-23 manuscript-main.Rmd
wrapping up edits from JA, BH, CS updating fig captions

@@
 
-``` {r load data}
-### use data_explore.Rmd to create this file, if it does not yet exist.  This
-### filters out presence = Extinct from IUCN cells.
-
-data_file <- 'data/spp_list_w_area_trimmed.csv'
-spp_map_pairs <- read_csv(data_file)
-### NOTE: as of Jan 31, includes all map pairs, even ones where IUCN area == 0
-
-### generate other lists
-spp_all_file <- 'data/spp_all.csv'
-spp_all <- read_csv(spp_all_file) ### NOTE: contains duped scinames
-
-# iucn_marine_file <- 'data/spp_iucn_marine.csv'
-# spp_iucn_marine <- read_csv(iucn_marine_file)
+``` {r load_dataset_specific_info}
 
@@
 
+``` {r load_spp_list_by_quad, echo = FALSE}
+spp_list_quads <- read_csv('data/spp_list_quads.csv')
+
+area_align_mean <- round(mean(spp_list_quads$area_ratio, na.rm = TRUE), 1)
+dist_align_mean <- round(mean(spp_list_quads$sm_perc, na.rm = TRUE), 1)
+
+n_q1 <- nrow(spp_list_quads %>% filter(quad == 'q1'))
+n_q2 <- nrow(spp_list_quads %>% filter(quad == 'q2'))
+n_q3 <- nrow(spp_list_quads %>% filter(quad == 'q3'))
+n_q4 <- nrow(spp_list_quads %>% filter(quad == 'q4'))
+
+n_q2_iucn_larger <- nrow(spp_list_quads %>% filter(sm_range == 'AM' & quad == 'q2'))
+n_q2_corals <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral') & quad == 'q2'))
+n_corals    <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral')))
+
+spp_quads_occ <- read_csv(file.path(dir_data, 'spp_quads_occ.csv'))
+med_occ_q4    <- round(median(spp_quads_occ$mean_occ[spp_quads_occ$quad == 'q4']))
+med_occ_q123  <- round(median(spp_quads_occ$mean_occ[spp_quads_occ$quad %in% c('q1', 'q2', 'q3')]))
+
+```
+
 ``` {r calc species numbers}
@@
 n_spp_iucn   <- length(unique(spp_iucn_maps$iucn_sid))
-n_spp_both   <- length(unique(spp_map_pairs$iucn_sid))
+n_spp_both   <- length(unique(spp_list_quads$iucn_sid))
 n_spp_all    <- n_spp_am + n_spp_iucn - n_spp_both
@@
 
-spp_list <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
-area_align_mean <- round(mean(spp_list$area_ratio, na.rm = TRUE), 1)
-dist_align_mean <- round(mean(spp_list$sm_perc, na.rm = TRUE), 1)
-
 ```
@@
 ``` {r captions_for_barchart_by_taxa, echo = FALSE}
-n_spp <- spp_map_pairs %>%
+n_spp <- spp_list_quads %>%
   mutate(spp_group = ifelse(str_detect(spp_group, 'CORAL'), 'CORAL', spp_group)) %>%
@@
 
-n_corals   <- nrow(spp_map_pairs %>% 
+n_corals   <- nrow(spp_list_quads %>% 
                      filter(str_detect(spp_group, 'CORAL')))
-n_bonyfish <- nrow(spp_map_pairs %>% 
+n_bonyfish <- nrow(spp_list_quads %>% 
                      filter(spp_group %in% c('WRASSES', 'BUTTERFLYFISH', 'PARROTFISH',
@@
                                              'DAMSELFISH', 'BONEFISH_TARPONS')))
-```
-
-``` {r captions_for_tally_quad_counts, echo = FALSE}
-dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
-area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
-q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
-q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
-q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
-q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
 ```

### 2016-09-23 ms_0_abst.Rmd
wrapping up edits from JA, BH, CS updating fig captions

@@
 Two major efforts to provide marine species distributions at a global scale are the International Union for Conservation of Nature (IUCN), which provides expert-generated range maps that outline the complete extent of a species' distribution; and AquaMaps, which provides model-generated species distribution maps that predict areas occupied by the species.
-Neither dataset claims to represent the "truth" of species' spatial distributions, but rather each communicates a distinct understanding of the underlying distribution, with a distinct purpose in mind.
-
 Together these databases represent `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` shared species.
-Ideally, the decision to use one dataset one over the other should carefully consider the purpose for which it was designed, but this small overlap between the two suggests that data availability is likely an important factor as well.
-
-Differences in intent and methodology can result in very different predictions of species distributions, which bear important implications for scientists and decision makers who rely upon these datasets to inform conservation policy and management actions.
-Comparing distributions for the small subset of species with maps in both datasets, we highlight several differences resulting from differences in methodology and intent.
+Differences in intent and methodology can result in very different predictions of species distributions, which bear important implications for scientists and decision makers who rely upon these datasets when conducting research or informing conservation policy and management actions.
+Comparing distributions for the small subset of species with maps in both datasets, we highlight several key examples in which introduced errors drive differences in predicted species ranges.
 In particular, we find that IUCN maps greatly overpredict coral presence into unsuitably deep waters, and we demonstrate that AquaMaps methodology can produce odd discontinuities at the extremes of a species range.
 We illustrate the scientific and management implications of these tradeoffs by repeating a global analysis of gaps in coverage of marine protected areas, and find significantly different results depending on how the two datasets are used.
-
-It remains essential to recognizes the tradeoffs between these two valuable marine datasets, and understand the implications of these tradeoffs for conservation research and decision-making.
+Efforts to understand tradeoffs between the two datasets, and ideally to enable the use of both when doing research that requires information on species distributions, will greatly improve the science and policy recommendations around understanding, managing, and protecting marine biodiversity.

### 2016-09-23 ms_1_intro.Rmd
wrapping up edits from JA, BH, CS updating fig captions

--- 2016-09-12 ms_1_intro.Rmd
@@
 No species range dataset can claim to represent the "truth" of any species' spatial distribution, but rather each offers its own distinct understanding of that latent distribution.
-These varying predictions of species presence and absence, driven by intent and methodology, can result in very different predictions of species presence; however, conservation managers and policy makers must base their actions on conclusions drawn from these imperfect datasets.
+These varying predictions of species presence and absence, driven by intent and methodology, can result in very different predictions of species' ranges.
+However, conservation managers and policy makers must base their actions on conclusions drawn from these imperfect datasets.
 It is critical to understand the differences in spatial range datasets and the implications of these differences for conservation research and decision-making.
 
+A broad literature exists on methods and results for predicting individual species distributions, but relatively few have been applied to broad suites of taxa globally, particularly in marine contexts.
 The two most comprehensive, widely-used global-scale repositories that predict marine species ranges throughout the world's oceans are AquaMaps, which rely primarily on model predictions to communicate the distribution of a species based on habitat suitability [@kaschner_aquamaps:_2015], and range data from the International Union for Conservation of Nature (IUCN), which rely primarily on expert opinion to communicate range size as a criterion for extinction risk assessments [@iucn_iucn_2015].
 
-While the two datasets ostensibly describe the same information, i.e.
-where can a particular species be found, they communicate fundamentally distinct concepts.
+While the two datasets ostensibly describe the same information, i.e., where can a particular species be found, they communicate fundamentally distinct concepts.
 The IUCN range map for a given species describes the "extent of occurrence" for that species, a contiguous region that encloses all known occurrences and connecting regions, with the explicit caveat that this "does not mean that [the species] is distributed equally within that polygon or occurs everywhere within that polygon" [@iucn_iucn_2015].
-The AquaMaps map for a species, on the other hand, seems to better be defined as "area of occupancy": the subset of the extent of occurrence in which the species is actually likely to be found, for example by excluding unsuitable habitats from the extent of occurrence [@iucn_iucn_2015].
-The distinctions between extent of occurrence and area of occupancy, including methodology, limitations, and appropriate usage, are well documented [@rondinini_tradeoffs_2006; @gaston_sizes_2009].
-___include reference to a supplemental table for more details?___
+The AquaMaps map for a species, on the other hand, excludes unsuitable habitats and thus seems to better be defined as "area of occupancy": the subset of the extent of occurrence in which the species is actually likely to be found [@kaschner_aquamaps:_2015].
 
-<!--
-_paragraph describing IUCN and AM as extent of occurrence and area of occupancy respectively, with basic defs and implication that AM will generally be smaller than IUCN; possibly refer to Table S1 for defs and details?_ _one reference: http://www.biodiversitya-z.org/content/species-extent-of-occurrence-eoo "The intent behind this parameter is to measure the degree to which risks from threatening factors are spread spatially across the taxon’s geographical distribution.
-It is not intended to be an estimate of the amount of occupied or potential habitat, or a general measure of the taxon’s range" and http://www.biodiversitya-z.org/content/species-area-of-occupancy-aoo "The measure reflects the fact that a taxon will not usually occur throughout the area of its extent of occurrence, which may, for example, contain unsuitable habitats"_
--->
+<!--The distinctions between extent of occurrence and area of occupancy, including methodology, limitations, and intended purpose, are well documented [@rondinini_tradeoffs_2006; @gaston_sizes_2009].
+___include table for more details?___ -->
 
-The fundamental differences between the datasets suggest that the choice of one over the other should be carefully matched to the purpose for which it is to be used [@gaston_sizes_2009], and yet these datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
-In most of these cases, the implications of choosing one versus the other of these datasets is not evaluated or discussed, yet strong conservation and management conclusions are drawn from the results of the studies.
+The fundamental differences between the datasets suggest that the choice of one over the other should be carefully matched to the purpose for which it is to be used [@gaston_sizes_2009], and yet these datasets have been used in hundreds of studies and applications for a wide range of purposes, typically without directly addressing the limitations of the chosen dataset or comparing outcomes between the two.
+Example uses of the data include assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].
 
-Here we focus not on how the two datasets should ideally be used; instead, recognizing that each provides distinct value for conservation research, we explore hidden assumptions and sources of error within each dataset, and test the implications of choosing one over the other.
-You do science with the data you have, not with the data you wish you had; often, the ideal type of spatial data is not available, and non-ideal data is better than no data at all.
-In such a case, it is crucial to acknowledge, and where possible address, the resulting tradeoffs.
-
-___include as table, in body or in supplemental?
-It really kind of doesn't fit here any more___ _Uncertainties inherent in method and intent inevitably drive differences in the range predictions made by each dataset: extent of occurrence range data such as IUCN range maps will inevitably introduce commission errors (inaccurate indication of presence), while species distribution models such as AquaMaps are expected to introduce fewer commission errors at the expense of more omission errors (inaccurate indications of absence) [@rondinini_tradeoffs_2006].
-Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range [@rondinini_tradeoffs_2006; @jetz_ecological_2008]._
-
-Data availability is a crucial factor in deciding which dataset to use.
-We compare how each data source represents the global spatial and taxonomic distribution of the `r format(n_spp_all, big.mark = ',')` marine species mapped by one or both datasets.
-Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps.
-IUCN range map data exist for `r format(n_spp_iucn, big.mark = ',')` unique marine species, with only `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets.
-This small overlap means each dataset provides unique value in taxonomic and geographic coverage, but also suggests that data availability likely plays a larger role in selecting a dataset than does the intended purpose of the data.
-
-Although relatively small in number, these overlap species present a unique opportunity to evaluate the two datasets overall.
-For the species included in both datasets, we examine how well the maps align in both spatial distribution and overall area.
-Based upon the definitions of extent of occurrence and area of occupancy, we expect that for a given species, the AquaMaps predicted distribution will fall within, and describe a smaller range than, the IUCN predicted distribution.
-Where these expectations seem to fail, we explore methodological issues that can introduce errors.
-In particular, we show how explicitly using depth as a constraint on coral ranges in IUCN range data dramatically mitigates a source of commission error, and highlight a side effect of AquaMaps methodology that results in discontinuous species range boundaries.
-
-<!--As understanding of a given species improves, we expect that range maps, regardless of method, would more closely approach the true distribution in predicting species presence and absence.
-Close alignment between two independent range maps can therefore indicate higher confidence in species presence or absence than a single map.-->
-
-<!--
-We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (OHI) [@halpern_index_2012; @halpern_patterns_2015] and a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as case studies to explore the implications of prioritizing one data set over the other.
-The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.
--->
-
-As a case study, we then reexamine a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] as a case study to explore the implications of selecting one data set over the other.
-Repeating the study using IUCN range data rather than AquaMaps data quintuples the apparent proportion of "gap" species, i.e.
-those species whose ranges remain completely unprotected, from 1.2% to 6.4%.
-This shift highlights the implications of different data use decisions on our understanding of marine biodiversity status and protection.
-
-Due to the variation in georegional and taxonomic coverage, and fundamental distinctions in method and intent of these two spatial range datasets, there is no way to simply recommend one dataset over the other.
-<!--Instead, we recommend simple methods to account for incidence of errors in each dataset, accounting for and reconciling differences in species range predictions to better inform conservation management and policy decisions.-->  Instead, by making explicit the distinctions between the two, and offering methods to identify and reconcile potential errors in each, we hope to reinforce the effective use of these datasets to better inform conservation management and policy decisions.
+We recognize that each dataset provides distinct value for conservation research, and so here we focus not on how the two datasets should ideally be used, but instead, on hidden assumptions and sources of error within each dataset.
+Importantly, we test the implications of choosing one over the other, and use overlapping species mapped within both datasets to better understand the causes of differences in results from using one or the other.
+We also repeat a recent analysis of gaps in protection afforded by marine protected areas (MPAs) to examine the implications of these differences for global conservation and management priorities.
+Because these datasets are so widely used, it is crucial to understand and acknowledge, and where possible address, data limitations.

### 2016-09-23 ms_2_methods.Rmd
wrapping up edits from JA, BH, CS updating fig captions

@@
 
-The IUCN publishes species range maps developed by species experts.
-These experts outline spatial boundaries that define the "limits of distribution" of a given species, based on observation records and informed by expert understanding of species' range and habitat preferences (Fig S1A).
-IUCN guidelines generally recommend that boundaries be drawn as a "minimum convex polygon": "the smallest polygon in which no internal angle exceeds 180 degrees and which contains all the sites of occurrence" [@iucn_iucn_2015].
+The IUCN has published species range maps developed by species experts for `r format(n_spp_iucn, big.mark = ',')` unique marine species.
+Experts outline spatial boundaries that define the "limits of distribution" of a given species, based on observation records and informed by expert understanding of species' range and habitat preferences (Fig S1A FIX THIS REF OR THE FIGURE).
+IUCN guidelines generally recommend that boundaries be drawn as a "minimum convex polygon", i.e.
+"the smallest polygon in which no internal angle exceeds 180 degrees and which contains all the sites of occurrence" [@iucn_iucn_2015].
+
+In contrast, AquaMaps has modeled species distribution for `r format(n_spp_am, big.mark = ',')` species based on envelopes of environmental preference, such as temperature, depth, and salinity.
+These preference envelopes are deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a global 0.5° grid to determine suitable habitat, resulting in a "probability of occurrence" for each species (Fig S1B FIX THIS REF OR THE FIGURE).
+To roughly constrain species ranges to appropriate georegions, the AquaMaps model uses Food and Agriculture Organization of the United Nations (FAO) Major Fishing Area [@fao_fisheries_and_aquaculture_department_cwp_2015] boundaries.
+Of the resulting maps, 1,296 (5.7% of the full dataset) have been further refined through an expert review process [@kaschner_mapping_2006; @ready_predicting_2010].
+
+In total, the two datasets provide range maps for `r format(n_spp_all, big.mark = ',')` species, with a small subset of species mapped in both datasets.
 For the purposes of this analysis, we elected not to use IUCN data for bird species, which are available separately through Bird Life International [@birdlife_international_and_natureserve_bird_2015].
-
-In contrast, AquaMaps models species distribution based on envelopes of environmental preference, such as temperature, depth, and salinity; these preference envelopes are deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a global 0.5° grid to determine suitable habitat, resulting in a "probability of occurrence" for each species (Fig S1B).
-To roughly constrain species ranges to appropriate georegions, the AquaMaps model uses Food and Agriculture Organization of the United Nations (FAO) Major Fishing Area [@fao_fisheries_and_aquaculture_department_cwp_2015] boundaries.
-Of the resulting maps, 1296 (5.7%) have been further refined through an expert review process [@kaschner_mapping_2006; @ready_predicting_2010].
 
@@
 To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source, and determined the proportion of each class represented in each dataset.
-To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps; we determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell  (Fig S2)  For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence, to best approximate the "extent of occurrence" generally indicated by IUCN maps.
+To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps.
+We determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell (Fig S2).
+For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence to best approximate the "extent of occurrence" generally indicated by IUCN maps.
 
 __Comparison of paired maps__:
-Species range maps communicate a prediction of both the distribution of a species and the size of a species range.
-To enable direct comparison of these measures between the two datasets, we identified "paired map" species, i.e.
-the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution, using genus and species binomials as a matching key.
+Although relatively small in number, overlap species present a unique opportunity to evaluate the two datasets overall.
+For the species included in both datasets, we examine how well the maps align in both spatial distribution and overall area.
+Based upon the definitions of extent of occurrence and area of occupancy, we expect that for a given species, the AquaMaps predicted distribution will fall within, and describe a smaller range than, the IUCN predicted distribution.
+Where these expectations seem to fail, we explore methodological issues that can introduce errors.
+
+We identified "paired map" species using genus and species binomials as a matching key.
 We used the taxize package [@scott_chamberlain_taxize_2013] in R [@r_core_team_r:_2016] to standardize species names and synonyms; for species with separate subpopulation maps in IUCN, we combined all subpopulations to create a single global population.
@@
 
-For each paired map species, $A_{small}$ and $A_{large}$ indicate the smaller and larger range representation (regardless of which dataset).
+For each paired map species, $A_{small}$ and $A_{large}$ indicate the smaller and larger range representation (regardless of which dataset) in km^2^.
 $A_{small \cap large}$ represents the amount of overlapping area between the two datasets.
 Distribution alignment uses overlapping predictions of presence as a measure of concurrence between the two datasets.
-Area ratio compares range size, used by IUCN as a criterion to help define extinction risk; it also provides a proxy for frequency of commission and/or omission errors.
+Area ratio compares range size, used by IUCN as a criterion to help define extinction risk; it also provides an indicator for frequency of commission and/or omission errors.
 
 __Examining issues in paired map alignment__:
-Corals provide a useful case study to examine differences in species range alignment.
-Corals are an intensely studied taxonomic group whose importance in supporting biodiversity is undisputed; their dependence on photosynthesis provides habitat-limiting life history traits, and they don't tend to move.
-Extracting data from the IUCN API [@iucn_iucn_2015] we identified depth limitations of each coral species mapped in the IUCN dataset to verify that none are found below the photosynthetic depth limit of 200 m.
-Using the same methodology as shown in Fig 2, we created a 200 m bathymetry raster from a 200 m bathymetry spatial dataset (public domain; available from Natural Earth, www.naturalearthdata.com) and masked our IUCN coral rasters to identify mapped coral presence below 200 m.
-The resulting maps were again compared to the AquaMaps rasters to examine distribution alignment and area ratio.
+Given the wide variation in the alignment of predicted range for the paired map species, we examined two potential drivers of error, one for each dataset.
+For IUCN data, we explored the assumed consideration of depth as a criterion for range predictions of coral species, a large portion (14%) of the IUCN dataset and an intensely studied taxonomic group whose importance in supporting biodiversity is undisputed.
+Extracting data from the IUCN API [@iucn_iucn_2015] we identified depth limitations of each coral species mapped in the IUCN dataset to verify that none is found below the photosynthetic depth limit of 200 m.
+Using the same methodology as shown in Fig S2, we created a 200 m bathymetry raster from a bathymetry spatial dataset (public domain; available from Natural Earth, www.naturalearthdata.com) and masked our IUCN coral rasters to identify mapped coral presence below 200 m.
+The resulting maps were again compared to the AquaMaps ranges to examine distribution alignment and area ratio.
 
-AquaMaps' reliance upon FAO zones to constrain species ranges seems likely to result in abrupt species range boundaries where ecologically suitable habitat intersects with human-defined borders.
-To explore the impact of these artificial constraints on species range predictions, we chose to focus on boundaries defined by longitude, as they seem less likely to coincide with ecological differences than boundaries defined by latitude.
-As an example, we identified all AquaMaps species whose eastern or western range limit between between 25° S and 20° N latitudes coincided with the large vertical boundary between FAO regions 71 and 77 at 175° W longitude.
+For AquaMaps data, whose reliance upon FAO Major Fishing Areas to constrain species ranges seems likely to result in abrupt discontinuities in species range where ecologically suitable habitat intersects with human-defined borders, we identified FAO boundaries where such discontinuities are likely to occur.
+Cropping species distributions to narrow latitudinal bands, we determined species whose eastern or western range limit coincided exactly with a defined FAO boundary.
+As an illustrative example, we cropped AquaMaps range maps for Indo-Pacific species to a band between between 25° S and 20° N latitudes, then identified all species whose eastern range limit within this band coincided with 175° W longitude, the vertical boundary between FAO regions 71 and 77.
 
-The oceans abound with rare and poorly-understood species which offer sparse information on behavior and habitat preferences to inform models and expert opinion.
-To identify such data-poor species within the set of paired map species, we used R clients for OBIS [@obis_data_2016] and GBIF [@gbif_global_2010] to identify known occurrences of each species; occurrences were averaged between the two occurrence databases as a proxy for data richness.
-
-<!--
-
-## Methods for Ocean Health Index (OHI) case study
-
-The global Ocean Health Index (OHI) [@halpern_index_2012; @halpern_patterns_2015], a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data to calculate biodiversity status for each of the world's 221 exclusive economic zones.
-
-The Species component of the OHI Biodiversity goal (abbreviated as SPP) uses IUCN Red List extinction risk as a proxy for a species' health.
-Each species is assigned a risk score based on IUCN Red List extinction risk category, from Least Concern (0.0) to Extinct (1.0) (Data Deficient and unassessed species are excluded).
-Species range is determined from its IUCN range map, where available, or its AquaMaps range map if an IUCN map is unavailable.
-For each assessed OHI region, an area-weighted mean extinction risk across all species is calculated, and rescaled to calculate a status score from 0 (mean extinction risk = 0.75) to 100 (mean extinction risk = 0) [@halpern_index_2012; @halpern_patterns_2015].
-
-_To highlight the differences in the datasets, we recalculated the OHI SPP status values using only one complete dataset at a time, and compared to the values resulting from the combined datasets.
-For the AquaMaps dataset, we used a "probability of occurrence" threshold of 0% to determine presence, to align with the OHI 2016 assessment._
-
--->
+Range maps for data-poor species are more likely to introduce errors, whether by commission (false indication of presence) or omission (false indication of absence).
+This is likely to be evident as poor alignment between range maps for paired-map species.
+To explore the impact of empirical measures of species occurrence on range predictions, we used R clients for OBIS [@obis_data_2016] and GBIF [@gbif_global_2010] to count occurrence records for each paired-map species; these counts were averaged for each species as a proxy for data richness.
 
@@
 The WDPA polygons and marine polygons were rasterized to 0.01° and then aggregated to AquaMaps native 0.5° cells, to calculate proportion of marine protected area within each cell.
-After verifying our results using the 2014 AquaMaps dataset, we updated the analysis to use the 2015 AquaMaps dataset, at a presence threshold of 50% (to compare to Klein et al.
+After verifying our method using the 2014 AquaMaps dataset, we updated the analysis to use the 2015 AquaMaps dataset, at a presence threshold of 50% (to compare to Klein et al.
 directly) and 0% (to better compare with IUCN spatial data).
 To analyze MPA coverage against IUCN spatial data, we extracted IUCN polygon weights per 0.5° cell for each species and compared against the protected area raster.
-Finally, we combined AquaMaps data (at 0% threshold) and IUCN data, using AquaMaps for the `r format(n_spp_both, big.mark = ',')` overlapping species and again compared against the protected area raster.
 

### 2016-09-23 ms_3_res_disc_concl.Rmd
wrapping up edits from JA, BH, CS updating fig captions

@@
 ---
-output: pdf_document
+output: word_document
 ---
@@
 The two datasets have notably different taxonomic (Fig 1A) and regional (Figs 1B, 1C) coverage.
-AquaMaps encompasses a broader range of taxa than IUCN, as IUCN spatial data files are only available for select taxonomic groups that have been comprehensively assessed.
+AquaMaps encompasses a broader range of taxa than IUCN, as IUCN spatial data are only available for select taxonomic groups that have been comprehensively assessed.
 Of the `r format(n_spp_all, big.mark = ',')` species mapped within these datasets, only `r format(n_spp_both, big.mark = ',')` (`r p_both_all`%) are mapped within both, with many taxa completely unrepresented in one dataset or the other.
 While species numbers in both datasets peak in tropical latitudes near the equator, species counts for IUCN maps drop quickly beyond 30°N and 30°S, while AquaMaps includes distribution of species well into temperate latitudes.
-The longitude frequency plots show a slight shift in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.
 Together, the limitations of spatial and taxonomic coverage are likely to drive a researcher's choice of dataset far more strongly than the quality, format, or intended purpose of the dataset.
@@
 
-``` {r load_spp_list_by_quad, echo = FALSE}
-spp_list_quads <- read.csv('data/spp_list_quads.csv', stringsAsFactors = FALSE)
-n_q1 <- nrow(spp_list_quads %>% filter(quad == 'q1'))
-n_q2 <- nrow(spp_list_quads %>% filter(quad == 'q2'))
-n_q2_iucn_larger <- nrow(spp_list_quads %>% filter(sm_range == 'AM' & quad == 'q2'))
-n_q2_corals <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral') & quad == 'q2'))
-n_corals    <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral')))
-n_q3 <- nrow(spp_list_quads %>% filter(quad == 'q3'))
-n_q4 <- nrow(spp_list_quads %>% filter(quad == 'q4'))
-
-# n_q4_datapoor <- nrow(spp_list_quads %>% filter(quad == 'q4' & occurcells < 10))
-# p_q4_datapoor <- round(n_q4_datapoor/n_q4 * 100, 1)
-# n_q123_datapoor <- nrow(spp_list_quads %>% filter(quad != 'q4' & occurcells < 10))
-# p_q123_datapoor <- round(n_q123_datapoor/n_spp_both * 100, 1)
-
-# n_q4_rev <- nrow(spp_list_quads %>% filter(quad == 'q4' & !is.na(reviewed)))
-# p_q4_rev <- round(n_q4_rev/n_q4 * 100, 1)
-# n_q123_rev <- nrow(spp_list_quads %>% filter(quad != 'q4' &  !is.na(reviewed)))
-# p_q123_rev <- round(n_q123_rev/n_spp_both * 100, 1)
-
-# quad_count_means <- read_csv(file.path(dir_data, 'quads_data_poor.csv')) %>%
-#   arrange(quad)
-# dp_med_am_q4    <- quad_count_means$med_occurcells[4]
-# dp_med_am_q123  <- quad_count_means$med_occurcells_total[1]
-# dp_med_occ_q4   <- quad_count_means$med_occ[4]
-# dp_med_occ_q123 <- quad_count_means$med_occ_total[1]
-
-spp_quads_occ <- read_csv(file.path(dir_data, 'spp_quads_occ.csv'))
-med_occ_q4   <- round(median(spp_quads_occ$mean_occ[spp_quads_occ$quad == 'q4']))
-med_occ_q123 <- round(median(spp_quads_occ$mean_occ[spp_quads_occ$quad %in% c('q1', 'q2', 'q3')]))
-
-```
-
 ## Distribution and range size alignment
@@
 Comparing distribution alignment and area ratio for the `r format(n_spp_both, big.mark = ',')` paired map species (Fig 2A), a weak negative linear pattern appears to emerge, suggesting that increasing similarity in range area correlates very slightly with decreasing distribution alignment (R^2^ = .016).
-The pattern itself is not particularly important, and emerges simply due to the nature of the analysis and the datasets:  The AquaMaps model tends to extrapolate species ranges into suitable areas beyond known occurrences more frequently than IUCN maps, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
+The pattern itself is not particularly important, and emerges simply due to the nature of the analysis and the datasets.
+In particular, the AquaMaps model tends to extrapolate species ranges into suitable areas beyond known occurrences more frequently than IUCN maps, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology.
 For species with dissimilar range areas, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range.
-For species with increasingly similar range areas, differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.
+For species with increasingly similar range areas, differences in methodology become more difficult to "hide," and the distribution alignment generally becomes slightly poorer.
 
 The mean distribution alignment for species included in both datasets was `r dist_align_mean`%; the mean area alignment was `r area_align_mean`%.
-By dividing the paired map species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern.
+By dividing the paired map species into quadrants based on these means, we highlight categories of relationships to identify patterns in alignment differences.
 Representative maps from each category are provided in the supporting materials (Fig S3).
-
-<!--
-Where two independent methods predict presence (or absence) of a species in the same location, we assume a higher confidence in the accuracy of that prediction.
-Where one method predicts presence while the other predicts absence, we can investigate potential sources of commission and omission errors to refine our understanding of the species' range.
-Understanding and addressing these sources of error provides guidance to improve predictions of species range even when only one source of spatial data is available.
--->
 
@@
 > __Fig 2.__ (A) Distribution alignment (overlap of smaller range within larger) versus area ratio (the ratio of smaller range area to the larger range area) for `r format(n_spp_both, big.mark = ',')` species included in both IUCN and AquaMaps datasets.
-The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than mean value) in both spatial distribution and the extent of described ranges (n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %).
+The upper right quadrant (quadrant 1) comprises species whose maps largely agree in both spatial distribution and the extent of described ranges (n = `r n_q1`; `r round(n_q1/nrow(spp_list_quads) * 100, 1)` %).
 The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in area
-(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %).
+(n = `r n_q2`; `r round(n_q2/nrow(spp_list_quads) * 100, 1)` %).
 The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range area, but disagree on where those ranges occur
-(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %).
+(n = `r n_q3`; `r round(n_q3/nrow(spp_list_quads) * 100, 1)` %).
 The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
-(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %).
+(n = `r n_q4`; `r round(n_q4/nrow(spp_list_quads) * 100, 1)` %).
 (B) Alignment quadrant breakdown of species by taxonomic group.
@@
 The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where one range is notably larger than the other.
-For `r round(n_q2_iucn_larger/n_q2 * 100, 1)`% of the species in this quadrant, the IUCN range is larger (on average, 2.57 times as large) than the AquaMaps range, as expected when comparing an extent of occurrence to an area of occupancy map.
-This suggests a high rate of commission and/or omission errors by one or both datasets; further analysis is required to disentangle the source and type of error contributing to poor area alignment.
-The results of the coral analysis (described below) provide some insight.
+For `r round(n_q2_iucn_larger/n_q2 * 100, 1)`% of the species in this quadrant, the IUCN range is larger than the AquaMaps range, as expected when comparing an extent of occurrence to an area of occupancy map.
+However, the low area ratio suggests a high rate of commission and/or omission errors by one or both datasets.
+Below we explore one underlying source of commission errors.
 
-Species found in the lower right quadrant (n = `r n_q3`) seem to represent cases of "two wrongs make a right."  For these species, IUCN and AquaMaps both predict ranges extending far beyond the overlapping region, but the methodological differences result in very different extrapolations.
-Consequently, area ratios are close to 100%, though the poor distribution alignment indicates that one or both datasets are introducing significant errors.
+Species found in the lower right quadrant (n = `r n_q3`) represent cases of "two wrongs make a right."  For these species, IUCN and AquaMaps both predict ranges extending far beyond the overlapping region, but the methodological differences result in very different extrapolations.
+Consequently, area ratios are high, though the poor distribution alignment indicates that one or both datasets are introducing significant errors.
 In this quadrant, the IUCN range is the larger for only 56.5% of species, defying our expectations based on definitions of area of occupancy and extent of occurrence.
-As above, further analysis is required to disentangle the causes of error.
 
@@
 Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
-The AquaMaps dataset offers its own quality metric based on the number of unique cells containing valid occurrences; for this quadrant, the median "occurcells" is 11 compared to a median of 40 across the other three quadrants.
+The AquaMaps dataset offers its own quality metric based on the number of unique 0.5° cells containing valid occurrences; for this quadrant, the median "occurcells" is 11 compared to a median of 40 across the other three quadrants.
 Care should be taken when using distribution and range maps based upon fewer observations, as they clearly bear greater uncertainty; AquaMaps explicitly warns against using any of its maps generated with an "occurcells" count fewer than 10 [@kaschner_aquamaps:_2015].
@@
 
-Noting from Fig 2B that corals dominate the upper-left "distribution-aligned" quadrant of Fig 2A (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), we chose to examine the effect of explicitly restricting IUCN ranges to depths based on species' life histories.
-Corals offer an excellent case study, due to their foundational role in biodiverse habitats, as well as their lack of mobility and reliance on photosynthesis.
-Ocean depth preference is explicitly included as a parameter in AquaMaps models.
+Because corals dominate the upper-left "distribution-aligned" quadrant of Fig 2A (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), we explored implications of explicitly restricting IUCN ranges to depths based on species' life histories.
+This adjustment was not necessary for AquaMaps data because models explicitly include ocean depth preference as a parameter.
 While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." [@iucn_iucn_2015, from website:  http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources]), it is not presented as a requirement, so we cannot take its inclusion for granted.
@@
 
-In constraining coral ranges to appropriately shallow waters, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps.
-Using the original quadrant definitions from Fig 2B, we see in Fig 3B a massive shift in 354 paired-map coral species due to constraining coral depths to 200 m.
+In constraining coral ranges to appropriate depths, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps (Fig 3B).
 Membership in the "well-aligned" quadrant jumped from 22.4% to 76.2%, with a corresponding decrease in all other quadrants.
@@
 
-We cannot know for certain the true distribution of each of these corals; certainly some commission errors result from satisfying the connectivity requirements of an extent of occurrence, and others may be due to experts taking a precautionary approach.
-But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported distributions for most purposes.
-Note that this change applies just as readily to the IUCN coral maps that are not included in the paired map analysis, and likely to other reef-associated flora and fauna.
+The true distribution of each of these corals remains imperfectly known.
+Certainly some commission errors result from extent of occurrence boundaries including areas where species do not actually exist (as recognized by the definition of this method), and others may be due to experts taking a precautionary (i.e., generous) approach to likely occurrence.
+Yet a simple and sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported distributions for most purposes.
+This change applies just as readily to the IUCN coral maps that are not included in the paired map analysis, and likely to other reef-associated flora and fauna.
 While species depth preferences are an easy and consistent means of constraining range predictions, other conditions such as salinity and temperature could be cautiously used to refine the results of expert opinions.
@@
 
-From the entirety of the AquaMaps dataset, we identified 3,577 Indo-Pacific species whose equatorial distributions (between 25° S and 20° N) encounter a range limit at 175° W.
-This longitudinal boundary created a western limit for 369 species, and an eastern limit for 3,208 species.
-A clear discontinuity in species distributions of the 3,208 west-limited species (Fig 4) matches perfectly with FAO region 77 [@fao_fisheries_and_aquaculture_department_cwp_2015]; other discontinuities are apparent at other FAO boundaries, despite these boundaries not being actively studied in the analysis.
-Similar discontinuities were clear when mapping the west-limited species.
+From the entirety of the AquaMaps dataset, we identified 3,208 Indo-Pacific species whose equatorial distributions (between 25° S and 20° N) encounter an eastern range limit at 175° W.
+A clear discontinuity in species distributions of these species (Figs 4A, 4B) matches perfectly with FAO region 77 [@fao_fisheries_and_aquaculture_department_cwp_2015]; other discontinuities are apparent at other FAO boundaries, despite these boundaries not being actively studied in the analysis.
 
@@
 
-> __Fig 4.__ Composite map showing AquaMaps predicted ranges for 3,208 species whose equatorial distribution encounters an eastern discontinuity exactly at 175° W, the boundary between FAO Major Fishing Areas 71 and 77 (shown in blue).
+> __Fig 4.__ (A) AquaMaps species distribution of _Hoplichthys regani_, the ghost flathead, with known occurrence records.
+(B) Aggregated AquaMaps predicted ranges for 3,208 species whose equatorial distribution encounters an eastern discontinuity exactly at 175° W, the boundary between FAO Major Fishing Areas 71 and 77 (shown in blue).
 Other FAO area boundaries create additional clear discontinuities.
 
-FAO Major Fishing Area boundaries provide a readily available method to roughly contrain AquaMaps predictions to appropriate ocean basins, thus eliminating a large source of potential commission errors and enabling rapid modeling of thousands of species ranges.
+FAO Major Fishing Area boundaries provide a readily available method to roughly constrain AquaMaps predictions to appropriate ocean basins, thus eliminating a large source of potential commission errors and enabling rapid modeling of thousands of species ranges.
 However, these boundaries are defined for statistical purposes based on economic and political considerations rather than ecological considerations, and can result in odd discontinuities in species range predictions where otherwise suitable habitat is excluded.
 While such a discontinuous boundary would likely be obvious when inspecting the distribution of an individual species, the distinction is likely to be obscured when aggregating many species ranges as is typical for biodiversity or conservation studies.
-Ecoregional spatial zones such as Marine Ecoregions of the World [@spalding_marine_2007] may provide an additional useful constraint on species ranges; expert review, though time-consuming, is the most certain route to boosting confidence in these predicted distributions.
 
-<!-- ## Case Studies
-
-To examine the implications of differences between the AquaMaps and IUCN predicted species ranges, we repeated two recent studies, varying only the selection of one data set over the other.
-
--->
-
-<!--
-### Case Study: The Ocean Health Index
-
-``` {r calc_ohi_numbers, echo = FALSE, message = FALSE, warning = FALSE}
-spp_ohi <- spp_all %>%
-  filter(!is.na(spatial_source) & !is.na(category_score)) %>%
-  select(am_sid, iucn_sid, spatial_source) %>%
-  unique()
-
-n_spp_am_ohi <- spp_ohi %>% 
-                    filter(str_detect(spatial_source, 'am')) %>%
-                    select(am_sid) %>%
-                    unique() %>%
-                    nrow()
-n_spp_iucn_ohi <- spp_ohi %>%
-                    filter(str_detect(spp_ohi$spatial_source, 'iucn')) %>%
-                    filter(is.na(am_sid)) %>%
-                    unique() %>%
-                    nrow()
-n_spp_both_ohi <- spp_ohi %>% 
-                    filter(!is.na(am_sid)) %>%
-                    filter(str_detect(spatial_source, 'iucn')) %>%
-                    select(iucn_sid) %>%
-                    unique() %>%
-                    nrow()
-
-n_spp_ohi <- n_spp_iucn_ohi + n_spp_am_ohi + n_spp_both_ohi
-
-p_spp_both_ohi <- round(n_spp_both_ohi/n_spp_ohi * 100, 1)
-
-```
-
-___THIS SECTION NEEDS TO BE REWRITTEN if we choose not to omit it___
-_We calculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig3'>Fig 3</a>).
-Reducing the threshold increases the apparent range of a species, so a decrease in a region’s score under scenario 1 would indicate increased spatial representation of threatened species for that region.
-Shifting priority from IUCN to AquaMaps may increase or decrease a species’ apparent range, so a decrease in region score for scenario 2 may indicate a decrease in spatial representation of low-risk species, an increase in spatial representation of high-risk species, or a combination of both.
-Scenario three combines effects of scenarios 1 and 2._
-
-_Given that only `r p_spp_both_ohi`% of Red List-assessed species are included in both datasets, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in substantial country-level score shifts as seen in scenarios 2 and 3.
-While the mean global score did not vary significantly from scenario to scenario, select countries gained up to 7.5 points while others dropped as many as 5 points.
-This result indicates that especially on a national or regional scale, an arbitrary change in how the two datasets are combined can result in a different assessment of species conservation status._
-
--->
+The 3,208 species range maps included in Fig 4 tend to extrapolate farther based on limited data: a median predicted range of 452,000 km^2^ per "occurcell" compared to a median predicted range of 158,000 km^2^ per "occurcell" for the overall AquaMaps dataset.
+This suggests that the FAO boundaries may not be sufficient to adequately constrain computer-generated ranges.
+To reduce the incidence of commission errors due to aggressive extrapolation, it may be desirable to fine-tune the computer model output with additional filters, including ecosystem-based constraints such as Marine Ecoregions of the World [@spalding_marine_2007] or methods such as inverse distance weighting to enforce proximity to known observations.
+Expert review, though time-consuming, is the most certain route to boosting confidence in these predicted distributions.
 
@@
 
-We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using the 2015 AquaMaps data and a 0% threshold to more closely approximate the extent of occurrence represented by IUCN data, <a href = '#fig4'>Fig 4</a>).
-We found a five-fold increase in the proportion of gap species (6.4% of species vs.
+We recalculated the amount of under-protected and gap species using all available IUCN species ranges, as well as the 2015 AquaMaps data at a 50% threshold to replicate the original methods and a 0% threshold to more closely approximate the extent of occurrence represented by IUCN data (Fig 5).
+Comparing the IUCN results to the AquaMaps 2015 results (at 0% threshold) we found a five-fold increase in the proportion of gap species (6.4% of species vs.
 1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% of species vs.
@@
 
-To achieve more comprehensive global coverage of species ranges it may be desirable to use these two datasets together, as the Ocean Health Index (OHI) does to calculate scores for its Species goal [@halpern_index_2012; @halpern_patterns_2015].
+To achieve more comprehensive taxonomic coverage of species ranges it may be desirable to use these two datasets together.
+The Ocean Health Index (OHI) is one major analysis that merges IUCN and AquaMaps spatial data, in conjunction with extinction risk categories from the IUCN Red List of Threatened Species, to calculate area-weighted extinction risks for the Species component of the index.
+To combine these data sets, OHI uses all available IUCN range maps, and then supplements with AquaMaps data (with "presence" determined by 40% or greater probability of occurrence) for species whose ranges are not represented within IUCN; species without valid Red List extinction risk categories are excluded from analysis [@halpern_index_2012; @halpern_patterns_2015].
+Such a combination of two inherently different datasets increases the number of species represented within the analysis, but the differences in range representations between the datasets is likely to distort the results.
+
 While it may be unrealistic to "fix" one data set to match the other, we may be able to reduce the impact of the tradeoffs inherent in each.
@@
 Conversely, including all AquaMaps cells with a non-zero "probability of occurrence" (rather than using a probability threshold to determine presence, e.g.
-greater than 40% for the Ocean Health Index [@halpern_index_2012; @halpern_patterns_2015] or 50% for the MPA gap analysis [@klein_shortfalls_2015]) would allow for the most generous inclusion of species range, resulting in maps that more closely, though still imperfectly, approximate extents of occurrence.
+greater than 40% for the Ocean Health Index [@halpern_index_2012; @halpern_patterns_2015] or 50% for the MPA gap analysis [@klein_shortfalls_2015]) would allow for the most generous inclusion of species range, resulting in maps that more closely, though still imperfectly, approach the IUCN's extents of occurrence.
 
@@
 
-No spatial dataset can ever claim to know the "truth" of the whole of marine biodiversity.
-Although other approaches exist for predicting species distribution, AquaMaps and IUCN range maps are the largest approaches applied globally across a broad range of marine taxa.
-These datasets show strong agreement for many well-studied species, but for many others, substantial differences arise from differences in methodology and intent of each dataset.
-While the decision of which dataset to use should ideally be driven by the intended purpose for which it was created, the fact is that taxonomic coverage will likely be a more important factor.
+No dataset can ever claim to know the "truth" of the location and extent of marine biodiversity.
+AquaMaps and IUCN range maps show strong agreement for many well-studied species, but for many others, substantial differences arise from differences in methodology and intent of each dataset.
+While the decision of which dataset to use should ideally be driven by the intended purpose for which it was created, the fact is that geographic and taxonomic coverage will likely be a more important factor in determining which dataset is used.
 Recognizing and acknowledging the advantages and differences of the range maps presented by these datasets will increase their utility for research and conservation actions.
 Conclusions drawn from each of these datasets could paint dramatically different pictures of global marine biodiversity or the effectiveness of conservation management decisions.
-
-Effective management and protection of marine species depends on a robust understanding of where species exist and where they do not; without this knowledge we risk wasting resources protecting low-value regions while missing opportunities to protect critical ones.
 By highlighting the distinctions between these two important and fundamentally different marine species range datasets, we improve our ability to inform strategic and effective conservation policy that supports a resilient ocean ecosystem.

### 2016-09-23 clip_fao/obis_occurrences.R

@@ -1,134 +1,0 @@
-### support functions for obis_count()
-library(httr)
-library(jsonlite)
-
-obis_url <- function() {
-  getOption("robis_url", "http://api.iobis.org/")
-}
-
-max_characters <- function() {
-  getOption("robis_max_characters", 50000)
-}
-
-handle_date <- function(date) {
-  if(!is.null(date) && class(date) == "Date") {
-    as.character(date)
-  } else {
-    date
-  }
-}
-
-handle_vector <- function(x) {
-  if(!is.null(x)) {
-    paste0(x, collapse = ",")
-  } else {
-    x
-  }
-}
-
-log_request <- function(result) {
-  cat("\n", paste(result$request$method, result$request$url, result$status_code, result$headers$age, result$times[["total"]]), "\n", sep="")
-}
-
-log_progress <- function(total, count) {
-  cat("\rRetrieved ", total, " records of ", count, " (", floor(total/count*100),"%)", sep="")
-}
-
-http_request <- function(method, path, query) {
-  if (method == "GET") {
-    httr::GET(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
-              path = path, query = query)
-  } else if (method == "POST") {
-    httr::POST(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
-               path = path, body = query)
-  }
-}
-
-obis_occurrence <- function(
-  scientificname = NULL,
-  year = NULL,
-  obisid = NULL,
-  aphiaid = NULL,
-  resourceid = NULL,
-  startdate = NULL,
-  enddate = NULL,
-  startdepth = NULL,
-  enddepth = NULL,
-  geometry = NULL,
-  qc = NULL,
-  fields = NULL,
-  verbose = FALSE) {
-  
-  
-  offset <- 0
-  i <- 1
-  lastpage <- FALSE
-  total <- 0
-  datalist <- list()
-  
-  t <- proc.time()
-  
-  while (!lastpage) {
-    query <- list(scientificname = scientificname,
-                  year = year,
-                  obisid = obisid,
-                  aphiaid = aphiaid,
-                  resourceid = resourceid,
-                  # startdate = handle_date(startdate),
-                  # enddate = handle_date(enddate),
-                  startdepth = startdepth,
-                  enddepth = enddepth,
-                  geometry = geometry
-                  # qc = handle_vector(qc),
-                  # fields = handle_vector(fields),
[…]

### 2016-09-23 clip_fao/fao_clip_pac.Rmd
Added explanation for data extrapolation rate. STILL NEED TO FIX REFS TO S1 and S2 FIGURES!!!!!!!

@@
 
-``` {r quad_spp_w, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+``` {r area_to_data_ratio, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 
@@
 
-# quad_names <- data.frame('quad' = c('q1', 'q2', 'q3', 'q4'),
-#                          'quad_name' = factor(c('well-matched', 'dist-matched', 
-#                                                 'area-matched', 'poorly matched'),
-#                                              ordered = TRUE))
-# 
-# clipped_spp_quads <- clipped_spp_quads %>%
-#   left_join(quad_names, by = 'quad') %>%
-#   mutate(method = 'FAO clipped')
-# 
-# break_nums <- seq(0, 100, 20)
-
-### Plot the bar chart
-# barchart_clipped_spp_quads <- ggplot(clipped_spp_quads, 
-#                                 aes(x = method, fill = quad, weight = pct_quad)) +
-#   ggtheme_plot + 
-#   theme(panel.grid.major.y = element_blank()) +
-#   geom_bar(stat = 'count', alpha = 1) +
-#   scale_fill_manual(values = c('q4' = '#d01c8b', 
-#                                'q3' = '#f1b6da', 
-#                                'q2' = '#b8e186',
-#                                'q1' = '#4dac26'),
-#                     labels = quad_names$quad_name) +
-#   scale_y_continuous(expand = c(0, 0), 
-#                      limits = c(0, 1),
-#                      breaks = break_nums/100,
-#                      labels = sprintf('%s%%', break_nums)) + 
-#   ### add grid lines; horizontal but then get flipped
-#   geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
-#   coord_flip() +
-#   labs(x = '', 
-#        y = 'Percent of species by quadrant', 
-#        fill = 'Alignment')
-# 
-# print(barchart_clipped_spp_quads)
-
 ```
-
-``` {r combine_rasts}
-
-stack_plot_both <- 
-  tm_shape(rast_spp_w) +
-    tm_raster(palette = 'Greens',
-              colorNA = NULL,
-              title = 'n_spp',
-              alpha = 1) +
-  tm_shape(rast_spp_e) +
-    tm_raster(palette = 'Reds',
-              colorNA = NULL,
-              title = 'n_spp') +
-  tm_shape(World1) +
-    tm_polygons() + 
-  tm_shape(fao_polys) +
-    tm_borders() +
-  tm_layout(basemaps = "Esri.WorldTopoMap", 
-            frame = TRUE,
-            # title.position = 'TOP', 
-            legend.outside = TRUE, attr.outside = TRUE)
-
-print(stack_plot_both)
-```

### 2016-09-23 ms_2_methods.Rmd
Added explanation for data extrapolation rate. STILL NEED TO FIX REFS TO S1 and S2 FIGURES!!!!!!!

@@
 The IUCN has published species range maps developed by species experts for `r format(n_spp_iucn, big.mark = ',')` unique marine species.
-Experts outline spatial boundaries that define the "limits of distribution" of a given species, based on observation records and informed by expert understanding of species' range and habitat preferences (Fig S1A FIX THIS REF OR THE FIGURE).
-IUCN guidelines generally recommend that boundaries be drawn as a "minimum convex polygon", i.e.
+Experts outline spatial boundaries that define the "limits of distribution" of a given species, based on observation records and informed by expert understanding of species' range and habitat preferences ___(Fig S1A FIX THIS REF OR THE FIGURE)___.
+IUCN guidelines recommend that boundaries be drawn as a "minimum convex polygon", i.e.
 "the smallest polygon in which no internal angle exceeds 180 degrees and which contains all the sites of occurrence" [@iucn_iucn_2015].
@@
 These preference envelopes are deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a global 0.5° grid to determine suitable habitat, resulting in a "probability of occurrence" for each species (Fig S1B FIX THIS REF OR THE FIGURE).
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a global 0.5° grid to determine suitable habitat, resulting in a "probability of occurrence" for each species ___(Fig S1B FIX THIS REF OR THE FIGURE)___.
 To roughly constrain species ranges to appropriate georegions, the AquaMaps model uses Food and Agriculture Organization of the United Nations (FAO) Major Fishing Area [@fao_fisheries_and_aquaculture_department_cwp_2015] boundaries.
@@
 Distribution alignment uses overlapping predictions of presence as a measure of concurrence between the two datasets.
-Area ratio compares range size, used by IUCN as a criterion to help define extinction risk; it also provides an indicator for frequency of commission and/or omission errors.
+Area ratio compares range size, used by IUCN as a criterion to help define extinction risk; it also provides an indicator for frequency of errors of commission (false indication of presence) or omission (false indication of absence).
 
@@
 
-For AquaMaps data, whose reliance upon FAO Major Fishing Areas to constrain species ranges seems likely to result in abrupt discontinuities in species range where ecologically suitable habitat intersects with human-defined borders, we identified FAO boundaries where such discontinuities are likely to occur.
+For AquaMaps data, whose reliance upon FAO Major Fishing Areas [@fao_fisheries_and_aquaculture_department_cwp_2015] to constrain species ranges seems likely to result in abrupt discontinuities in species range where ecologically suitable habitat intersects with human-defined borders, we identified FAO boundaries where such discontinuities are likely to occur.
 Cropping species distributions to narrow latitudinal bands, we determined species whose eastern or western range limit coincided exactly with a defined FAO boundary.
-As an illustrative example, we cropped AquaMaps range maps for Indo-Pacific species to a band between between 25° S and 20° N latitudes, then identified all species whose eastern range limit within this band coincided with 175° W longitude, the vertical boundary between FAO regions 71 and 77.
-
-Range maps for data-poor species are more likely to introduce errors, whether by commission (false indication of presence) or omission (false indication of absence).
-This is likely to be evident as poor alignment between range maps for paired-map species.
-To explore the impact of empirical measures of species occurrence on range predictions, we used R clients for OBIS [@obis_data_2016] and GBIF [@gbif_global_2010] to count occurrence records for each paired-map species; these counts were averaged for each species as a proxy for data richness.
+As an illustrative example, we cropped AquaMaps range maps for all Pacific and Indian Ocean species to a band between between 25° S and 20° N latitudes, then identified all species whose eastern range limit within this band coincided with 175° W longitude, the vertical boundary between two prominent FAO regions.
 

### 2016-09-23 ms_3_res_disc_concl.Rmd
Added explanation for data extrapolation rate. STILL NEED TO FIX REFS TO S1 and S2 FIGURES!!!!!!!

@@
 
-The 3,208 species range maps included in Fig 4 tend to extrapolate farther based on limited data: a median predicted range of 452,000 km^2^ per "occurcell" compared to a median predicted range of 158,000 km^2^ per "occurcell" for the overall AquaMaps dataset.
+The ratio of the total predicted range for a species to the number of "occurcells" used to generate the map provides a measure of the degree to which AquaMaps extrapolates range area from limited data.
+For example, AquaMaps predicts a total range of 5.4 million km^2^ for both the round ray _Rajella fyllae_ and the brittle star _Ophiothrix plana_; but the map for _R.
+fyllae_ is generated using 116 "occurcells" (for a data extrapolation rate of 46,800 km^2^ per cell) while the map for _O.
+plana_ is generated using only four (for an extrapolation rate of 1,360,000 km^2^ per cell).
+
+By this measure, the 3,208 species range maps included in Fig 4 tend to extrapolate farther based on limited data: a median predicted range of 452,000 km^2^ per "occurcell" compared to a median predicted range of 158,000 km^2^ per "occurcell" for the overall AquaMaps dataset.
 This suggests that the FAO boundaries may not be sufficient to adequately constrain computer-generated ranges.
@@
 To achieve more comprehensive taxonomic coverage of species ranges it may be desirable to use these two datasets together.
-The Ocean Health Index (OHI) is one major analysis that merges IUCN and AquaMaps spatial data, in conjunction with extinction risk categories from the IUCN Red List of Threatened Species, to calculate area-weighted extinction risks for the Species component of the index.
+The Ocean Health Index (OHI) is one analysis that uses both IUCN and AquaMaps spatial data, in conjunction with extinction risk categories from the IUCN Red List of Threatened Species, to calculate area-weighted extinction risks for the Species component of the index.
 To combine these data sets, OHI uses all available IUCN range maps, and then supplements with AquaMaps data (with "presence" determined by 40% or greater probability of occurrence) for species whose ranges are not represented within IUCN; species without valid Red List extinction risk categories are excluded from analysis [@halpern_index_2012; @halpern_patterns_2015].

### 2016-09-24 clip_fao/fao_clip_pac.Rmd
unbolded legend for map

@@
             'Culeolus herdmani'  = 'URM-2-3539', 
-            'Chrionema chlorotaenia' = 'Fis-33989')
+            'Chrionema chlorotaenia' = 'Fis-33989')[1]
 
@@
                  border.col = 'red',
-                 border.lwd = 0.25,
+                 border.lwd = 0.5,
                  border.alpha = 1,
                  size = 1,
-                 scale = .3,
+                 scale = .4,
                  alpha = .3) +
       # tm_dots(col = 'red') +
-    tm_layout(# title = spp_name,
-              title.position = c('left', 'bottom'),
-              fontface = 'bold.italic',
-              fontfamily = 'Helvetica',
-              title.size = .8,
-              frame = TRUE,
+    tm_layout(frame = TRUE,
               legend.text.size = .6,
@@
 }
-
-# save_tmap(rast_spp_map_w, 
-#           file.path(dir_fig, 'map_am_edge_w.png'), 
-#           width = 19, 
-#           units = 'cm', dpi = 600)
 

### 2016-09-24 ms_3_res_disc_concl.Rmd
updated figures; captions in main body and supplemental figs; references to supplemental figures from body

@@
 
-> __Fig 1.__ (A) Number and proportion of species by taxa included in each dataset.
+> __Fig 1.
+Taxonomic and geographic coverage of AquaMaps and IUCN range data.__ (A) Number and proportion of species by taxa included in each dataset.
 Overlapping species are dominated by bony fishes (`r format(n_bonyfish, big.mark = ',')` species, primarily tropical taxa) and corals (`r format(n_corals, big.mark = ',')` species).
@@
 
-> __Fig 2.__ (A) Distribution alignment (overlap of smaller range within larger) versus area ratio (the ratio of smaller range area to the larger range area) for `r format(n_spp_both, big.mark = ',')` species included in both IUCN and AquaMaps datasets.
+> __Fig 2.
+Comparison of alignment between AquaMaps and IUCN range data.__ (A) Distribution alignment (overlap of smaller range within larger) versus area ratio (the ratio of smaller range area to the larger range area) for `r format(n_spp_both, big.mark = ',')` species included in both IUCN and AquaMaps datasets.
 The upper right quadrant (quadrant 1) comprises species whose maps largely agree in both spatial distribution and the extent of described ranges (n = `r n_q1`; `r round(n_q1/nrow(spp_list_quads) * 100, 1)` %).
@@
 
-> __Fig 3.__ (A) Aggregate map combining ranges of the 463 coral species mapped in the IUCN dataset, showing raw ranges and ranges clipped to 200 m depth.
+> __Fig 3.
+Effect of 200 m depth constraint on IUCN range maps for coral species.__ (A) Aggregate map combining ranges of the 463 coral species mapped in the IUCN dataset, showing raw ranges and ranges clipped to 200 m depth.
 (B) Alignment quadrant breakdown of paired map coral species using original data from IUCN and AquaMaps (as in Fig 2B) and the same species with IUCN ranges clipped to 200 m depth.
@@
 From the entirety of the AquaMaps dataset, we identified 3,208 Indo-Pacific species whose equatorial distributions (between 25° S and 20° N) encounter an eastern range limit at 175° W.
-A clear discontinuity in species distributions of these species (Figs 4A, 4B) matches perfectly with FAO region 77 [@fao_fisheries_and_aquaculture_department_cwp_2015]; other discontinuities are apparent at other FAO boundaries, despite these boundaries not being actively studied in the analysis.
+A clear discontinuity in species distributions of a single example species (Fig 4A) and all 3,208 species in aggregate (Fig 4B) matches perfectly with FAO region 77 [@fao_fisheries_and_aquaculture_department_cwp_2015]; other discontinuities are apparent at other FAO boundaries, despite these boundaries not being actively studied in this analysis.
 
@@
 
-> __Fig 4.__ (A) AquaMaps species distribution of _Hoplichthys regani_, the ghost flathead, with known occurrence records.
+> __Fig 4.
+Effect of FAO Major Fishing Area constraints on AquaMaps distributions.__ (A) AquaMaps species distribution of _Hoplichthys regani_, the ghost flathead, with known occurrence records.
 (B) Aggregated AquaMaps predicted ranges for 3,208 species whose equatorial distribution encounters an eastern discontinuity exactly at 175° W, the boundary between FAO Major Fishing Areas 71 and 77 (shown in blue).
@@
 
-> __Fig 5.__  Percent of species range covered by MPAs based upon methods in Klein et al.
+> __Fig 5.
+MPA gap analysis results based upon alternate choices of datasets.__  Percent of species range covered by MPAs based upon methods in Klein et al.
 (2015).

### 2016-09-24 ms_6_supplemental.Rmd
ok, figs are done.  DONE.  Do not look at them again.
updated figures; captions in main body and supplemental figs; references to supplemental figures from body

@@
 ---
-title: "Supplemental Information: Aligning marine species range data to better serve science and conservation"
+title: "Supplemental Figures: Aligning marine species range data to better serve science and conservation"
 author: "Casey O'Hara, Jamie C. Afflerbach, Courtney Scarborough, Benjamin S. Halpern"
@@
 
-![](figures/Fig_S1ab.png)
+![](figures/Fig_S1.png)
 
-![](figures/Fig_S1cd.png)
-
-> __Fig S1.__  Examples of AquaMaps and IUCN raw species range data.
-(A, B) AquaMaps distribution of _Thunnus alalunga_ (Albacore Tuna) (AquaMaps 2013e) showing global range with varying probabilities of occurrence assigned to 0.5° grid cells (A) and presence within 0.5° grid cells (B).
+> __Fig S1.
+Examples of AquaMaps and IUCN raw species range data.__ (A, B) AquaMaps distribution of _Thunnus alalunga_ (Albacore Tuna) (AquaMaps 2013e) showing (A) global range with varying probabilities of occurrence assigned to 0.5° grid cells and (B) presence within 0.5° grid cells.
 (C, D) IUCN distribution of _T.
-alalunga_ (IUCN 2011b) represented as extent of occurrence polygons (C) and presence within 0.5° grid cells (D).
+alalunga_ (IUCN 2011b) represented as (C) extent of occurrence polygons and (D) presence within 0.5° grid cells.
 
@@
 
-![](figures/Fig_S2a.png)
+![](figures/Fig_S2.png)
 
-![](figures/Fig_S2b.png)
-
-![](figures/Fig_S2c.png)
-
-![](figures/Fig_S2d.png)
-
-> __Fig S2.__ Rasterizing shapefiles provided by IUCN.
-A portion of the _T.
+> __Fig S2.
+Rasterizing shapefiles provided by IUCN.__ A portion of the _T.
 alalunga_ range map is used to exemplify the rasterization process.
@@
 
-> __Fig S3.__ Representative species maps (AquaMaps 2013abcd; Conch 2013ab; IUCN 2011a; IUCN 2012) to illustrate each quadrant from Fig 2A.
-Each map is positioned to match its quadrant in Fig 2A.
+> __Fig S3.
+Representative species maps to illustrate each quadrant from Fig 2A.__ (AquaMaps 2013abcd; Conch 2013ab; IUCN 2011a; IUCN 2012) Each map is positioned to match its quadrant in Fig 2A.
 FAO Major Fishing Area boundaries [@fao_fisheries_and_aquaculture_department_cwp_2015] are outlined in light grey.
@@
 > __Fig.
-S4.__  Shift in alignment of paired-map coral species due to clipping IUCN ranges to areas shallower than 200 m.
-The grey lines represent the change in apparent alignment for a single species.
+S4.
+Shift in alignment of paired-map coral species due to clipping IUCN ranges to areas shallower than 200 m.__  The grey lines represent the change in apparent alignment for a single species.
 Most coral species shift rightward from the upper left quadrant to the upper right, with little if any loss in distribution alignment, since in general, only unsuitable habitat has been removed.
@@
 
-<!--
-## Risk by quadrant
-
-<a id = 'S2'></a>
-
-![](figures/si_barchart_spp_risk_quads.png)
-
->  __Fig S2.__ Breakdown of quadrants (Fig S1A) by IUCN extinction risk categories.
-IUCN and Aquamaps ranges were in better agreement (i.e., had a higher probability of being in the upper-right, "well-aligned" quadrant, Figure 2A/Figure S1A) for species at greater risk according to IUCN risk categories (p < 0.001, generalized linear model, family = binomial, Table S1).
-This suggests at-risk species may receive more attention in terms of data collection or model refinement.
-
-> __Table S1.__  Generalized linear model (family = binomial) describing how the agreement between IUCN and Aquamap range maps (i.e., probability of occurring in quadrant 1, Figure 2A/S1A) increases for a species as its IUCN risk status increases.
-For this analysis, we excluded all species that had "data deficient" IUCN status (n = 255).
-We analyzed the IUCN risk categories as a continuous variable (LC = 0.0, NT = 0.2, VU = 0.4, EN = 0.6, CR = 0.8, EX = 1.0).
-
-_Model: glm(quadrant 1 membership ~ IUCN category, family = binomial) n = 2,059_
-
-| Coefficients:  | Estimate | Std. Error | z value | p value      |
-| :-----         |   -----: |     -----: |  -----: |       -----: |
-| Intercept      | -1.32933 |   0.05804  | -22.902 | < 2e-16 ***  |
-| IUCN category  |  1.18672 |   0.33469  |   3.546 | 0.000391 *** |
-
-### Coral range vs. depth
-
-<a id = 'S3'></a>
-
-![oculina varicosa](figures/si_coralmap.png)
-
->  __Fig S3.__ IUCN-defined range of _Oculina varicosa_, a typical photosynthetic coral, plotted against a 200 m bathymetry contour.
-According to IUCN, _O.
-varicosa_ can be found to depths of 152 m.
-Maps of many other coral species and other reef-associated organisms followed similar patterns.
-
-### AquaMaps threshold examination
-
-<a id = 'S4'></a>
-
-AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value as indicated by IUCN maps.
-Many studies convert this AquaMaps probability to a simple presence value by assigning a threshold value.
-
-To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0% to 100% and calculated the average species range relative to a zero threshold.
-
-A. 
-![](figures/si_am_threshold.png)
-
-B. 
-![](figures/si_ribbon_am_threshold.png)
-
->  __Fig S4.__ AquaMaps distribution map extent remaining after applying a presence threshold.
-A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.
-(A) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of 28% of total range compared to a 0% threshold, with a wide distribution in which some species lose nearly all of their apparent range.
-(B) Median remaining extent at increments of presence threshold.
-Dark green ribbon includes 25% to 75% quantiles of remaining species range, while lighter ribbon includes 5% to 95% quantiles.
-Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity of apparent species range to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.
--->
-
-<!--
-## MPA gap analysis - explanation
-
-___why so many gaps and 10+ species show up with IUCN.  I will write up a paragraph once I've had a little more time to examine maps___
-
-- check which species are gap species; check maps
-- check which species are 10%+ species and check maps
-- where do these species fall georegionally? is there anything odd about their ranges? any reason why IUCN would be 
-- hypothesis to explain more gap species: Aquamaps more diffuse than IUCN, catching individual cells that could be suitable; a single cell (even low probability) moves a species from "gap" to non-gap status.
-IUCN more clustered by polygon boundaries; so no scattered cells to accidentally fall into an MPA.
-    - introducing errors: AM = likely commission (overestimates species with tiny bits of range in MPAs) vs IUCN = possible omission (misses possible real habitat that falls within MPAs)
-- hypothesis to explain more 10%+ species - ten random samples most have IUCN polygons just east of Australia where a large MPA is indicated - Great Barrier Reef I assume - a single large poly falling in GBR gets a high score, while for the same species, many of the AM ranges (also in same area) scatter suitable area over larger extent, so lower percent of species range is protected
-    - introducing errors: IUCN = likely omission (skipping possible real habitat outside of MPAs) and likely commission (possibly overcounting real habitat in MPAs e.g.
-GBR); AquaMaps = probably a little commission (possibly underestimating, if it overpredicts lots of range outside MPAs, thus reducing the proportion of protected range)
--->
-
-## References for supplemental information
+## References for supplemental figures
 

### 2016-09-24 clip_fao/fao_clip_pac.Rmd
added t-test for mean area-per-cells

@@
 
-am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
-                                   'v2016/int/am_cells_spp_prob0.csv'))
+# am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
+#                                    'v2016/int/am_cells_spp_prob0.csv')) 
+# 
+# am_spp_list <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>%
+#   dplyr::select(am_sid = speciesid, genus, species, reviewed, occurcells) %>%
+#   mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' ')) %>%
+#   dplyr::select(-genus, -species) %>%
+#   distinct()
+# 
+# hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
+#   dplyr::select(loiczid, cellarea, oceanarea)
+# 
+# am_spp_cells2 <- hcaf %>%
+#   left_join(am_spp_cells, by = 'loiczid')
+# 
+# 
+# spp_area_df <- am_spp_cells2 %>%
+#   group_by(am_sid) %>%
+#   summarize(cellarea = sum(cellarea),
+#             oceanarea = sum(oceanarea))
+# 
+# spp_area_df <- spp_area_df %>%
+#   inner_join(am_spp_list, by = 'am_sid') %>%
+#   mutate(area_per_cells = oceanarea/occurcells)
+# 
+# clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv')) %>%
+#   filter(lim == 'e_lim') %>%
+#   dplyr::select(am_sid)
+# 
+# spp_area_df <- spp_area_df %>%
+#   mutate(log_area_cells = log(area_per_cells + 1),
+#          clipped = am_sid %in% clipped_spp$am_sid)
 
-am_spp_list <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>%
-  dplyr::select(am_sid = speciesid, genus, species, reviewed, occurcells) %>%
-  mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' ')) %>%
-  dplyr::select(-genus, -species) %>%
-  distinct()
+# write_csv(spp_area_df, 'clip_fao/spp_area_per_cell.csv')
 
-hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
-  dplyr::select(loiczid, cellarea, oceanarea)
+spp_area_df <- read_csv('clip_fao/spp_area_per_cell.csv')
 
-am_spp_cells2 <- hcaf %>%
-  left_join(am_spp_cells, by = 'loiczid')
+spp_area_clipped <- spp_area_df %>%
+  filter(clipped)
 
-clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv')) %>%
-  filter(lim == 'e_lim') %>%
-  dplyr::select(am_sid)
-
-spp_quad_list <- read_csv(file.path(dir_data, 'spp_list_quads.csv'))
-
-spp_area_df <- am_spp_cells2 %>%
-  group_by(am_sid) %>%
-  summarize(cellarea = sum(cellarea),
-            oceanarea = sum(oceanarea))
-
-spp_area_df <- spp_area_df %>%
-  inner_join(am_spp_list, by = 'am_sid') %>%
-  mutate(area_per_cells = oceanarea/occurcells)
-
-head(spp_area_df)
-#                                     am_sid    cellarea   oceanarea reviewed occurcells                   sciname
-#                                      <chr>       <dbl>       <dbl>    <chr>      <int>                     <chr>
-# 1                                Biv-32126   949988.02   642706.64     null          4        Corbicula japonica
-# 2                                Biv-32832  9698932.54  7589672.98     null         23          Polymesoda erosa
-# 3                               Bra-235673    16143.75    13895.96        1         10       Dyscritosia secreta
-# 4 Chn-013ca97e-46a8-4c5e-b398-8bb4f340f88b 11288439.81 10048833.94     null         76          Lenisquilla lata
-# 5 Chn-015453c0-3ff9-477d-ab51-c68d85e65af5 15078672.39 12760166.60     null         13        Clorida albolitura
-# 6 Chn-0721ac0c-7d19-46fa-9dcc-02da54310867 35256268.05 33846838.22     null         23 Leiogalathea laevirostris
+### compare ocean areas
 summary(spp_area_df$oceanarea)
@@
    #   1806    946600   2921000  10210000   7875000 501300000 
+summary(spp_area_clipped$oceanarea)
+   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
+   # 134900   4646000   7032000   8960000   9690000 111300000
+
+### compare occurcells
 summary(spp_area_df$occurcells)
@@
    # 0.00    6.00   16.00   57.13   49.00 7663.00 
+summary(spp_area_clipped$occurcells)
+   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+   # 3.00    6.00   15.00   26.86   33.00  494.00
+
+### compare area_per_cells
 summary(spp_area_df %>% filter(occurcells != 0) %>% .$area_per_cells)
@@
     #  190    50540   158000   485300   439100 44700000 
-nrow(spp_area_df %>% filter(!str_detect(reviewed, 'null')))
-   # 1296 out of 22889 = 5.66%
-
-spp_area_clipped <- spp_area_df %>%
-  filter(am_sid %in% clipped_spp$am_sid)
-nrow(spp_area_clipped %>% filter(!str_detect(reviewed, 'null')))
-   # 22 out of 3208 = .686%
-
-summary(spp_area_clipped$oceanarea)
-   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
-   # 134900   4646000   7032000   8960000   9690000 111300000 
-summary(spp_area_clipped$occurcells)
-   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-   # 3.00    6.00   15.00   26.86   33.00  494.00 
 summary(spp_area_clipped$area_per_cells)
@@
 
+### compare reviewed species
+nrow(spp_area_df %>% filter(!str_detect(reviewed, 'null')))
+   # 1296 out of 22889 = 5.66%
+nrow(spp_area_clipped %>% filter(!str_detect(reviewed, 'null')))
+   # 22 out of 3208 = .686%
+
+hist(spp_area_clipped$area_per_cells)
+hist(spp_area_clipped$log_area_cells)
+hist(spp_area_df$log_area_cells)
+
+t.test(x = spp_area_clipped$log_area_cells,
+       y = spp_area_df %>% filter(occurcells != 0) %>% .$log_area_cells,
+       alternative = 'greater') ### clipped sample mean greater than overall
+# t = 49.082, df = 5244.6, p-value < 2.2e-16
+# alternative hypothesis: true difference in means is greater than 0
+# 95 percent confidence interval:
+#  1.058579      Inf
+# sample estimates:
+# mean of x mean of y 
+#  12.97963  11.88434
+
 ```

### 2016-09-24 ms_3_res_disc_concl.Rmd
added t-test for mean area-per-cells

--- 2016-09-23 ms_3_res_disc_concl.Rmd
@@
 
-By this measure, the 3,208 species range maps included in Fig 4 tend to extrapolate farther based on limited data: a median predicted range of 452,000 km^2^ per "occurcell" compared to a median predicted range of 158,000 km^2^ per "occurcell" for the overall AquaMaps dataset.
+By this measure, the 3,208 species range maps included in Fig 4 tend to extrapolate farther based on limited data: a mean predicted range of 827,200 km^2^ per "occurcell" compared to a mean predicted range of 485,300 km^2^ per "occurcell" for the overall AquaMaps dataset.
 This suggests that the FAO boundaries may not be sufficient to adequately constrain computer-generated ranges.
-To reduce the incidence of commission errors due to aggressive extrapolation, it may be desirable to fine-tune the computer model output with additional filters, including ecosystem-based constraints such as Marine Ecoregions of the World [@spalding_marine_2007] or methods such as inverse distance weighting to enforce proximity to known observations.
+To reduce the incidence of commission errors due to aggressive extrapolation, it may be desirable to fine-tune the computer model output with additional filters, such as ecoregional constraints, e.g.
+Marine Ecoregions of the World [@spalding_marine_2007], or distance-based methods, e.g.
+inverse distance weighting to enforce proximity to known observations.
 Expert review, though time-consuming, is the most certain route to boosting confidence in these predicted distributions.

### 2016-09-24 ms_2_methods.Rmd

@@
 The IUCN has published species range maps developed by species experts for `r format(n_spp_iucn, big.mark = ',')` unique marine species.
-Experts outline spatial boundaries that define the "limits of distribution" of a given species, based on observation records and informed by expert understanding of species' range and habitat preferences ___(Fig S1A FIX THIS REF OR THE FIGURE)___.
+Experts outline spatial boundaries that define the "limits of distribution" of a given species, based on observation records and informed by expert understanding of species' range and habitat preferences (Fig S1A).
 IUCN guidelines recommend that boundaries be drawn as a "minimum convex polygon", i.e.
@@
 These preference envelopes are deduced from occurrence records, published species databases such as FishBase, and expert knowledge.
-The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a global 0.5° grid to determine suitable habitat, resulting in a "probability of occurrence" for each species ___(Fig S1B FIX THIS REF OR THE FIGURE)___.
+The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a global 0.5° grid to determine suitable habitat, resulting in a "probability of occurrence" for each species (Fig S1B).
 To roughly constrain species ranges to appropriate georegions, the AquaMaps model uses Food and Agriculture Organization of the United Nations (FAO) Major Fishing Area [@fao_fisheries_and_aquaculture_department_cwp_2015] boundaries.
@@
 
-<!--
-OMIT: Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., previously, the Ocean Health Index defined presence as 40% or greater probability of occurrence [@halpern_index_2012; @halpern_patterns_2015].
-In the 2016 OHI assessment, the presence threshold was eliminated, to better approximate the "extent of occurrence" generally indicated by IUCN maps, since the two datasets are used together
--->
-
 __Comparison of taxonomic and regional distribution__:
 To examine the overall taxonomic distribution across the spatial datasets, we grouped species by taxonomic class and data source, and determined the proportion of each class represented in each dataset.
-To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps.
-We determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell (Fig S2).
-For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence to best approximate the "extent of occurrence" generally indicated by IUCN maps.
+To compare the spatial representation of the two datasets directly, we rasterized the IUCN species polygons to the same 0.5° grid as the AquaMaps species maps (Fig S2).
+We determined species presence within a grid cell as any non-zero overlap of a species polygon with the cell (Fig S1C).
+For the AquaMaps dataset, we determined per-cell species count by including all species with non-zero probability of occurrence to best approximate the "extent of occurrence" generally indicated by IUCN maps (Fig S1D).
 

### 2016-09-25 ms_0_abst.Rmd
cleaning up and preparing to make public for submit

@@
 ## Abstract
-
-<!-- 
-The Abstract comes after the title page in the manuscript file. The abstract text is also entered in a separate field in the submission system.
-
-The Abstract should:
-  Describe the main objective(s) of the study
-  Explain how the study was done, including any model organisms used, without methodological detail
-  Summarize the most important results and their significance
-  Not exceed 300 words
-Abstracts should not include:
-  Citations
-  Abbreviations, if possible
--->
-
-<!-- from intro: -->
-<!-- Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation.
-This body of science has resulted in various compiled databases of species distribution and range maps which provide foundational information for understanding species diversity and extinction risk, predicting species responses to human impacts and climate change, and managing and protecting species effectively.
-No species range dataset can claim to represent the "truth" of any species' spatial distribution, but rather each offers its own distinct understanding of that latent distribution.
-These varying predictions of species presence and absence, driven by intent and methodology, can result in very different predictions of species presence; however, conservation managers and policy makers must base their actions on conclusions drawn from these imperfect datasets.
-It is critical to understand the differences in spatial range datasets and the implications of these differences for conservation research and decision-making.
--->
 

### 2016-09-25 ms_1_intro.Rmd
cleaning up and preparing to make public for submit

@@
 ---
-
-<!--
-The introduction should:
-  Provide background that puts the manuscript into context and allows readers outside the field to understand the purpose and significance of the study
-  Define the problem addressed and why it is important
-  Include a brief review of the key literature
-  Note any relevant controversies or disagreements in the field
-  Conclude with a brief statement of the overall aim of the work and a comment about whether that aim was achieved
--->
 

### 2016-09-25 ms_3_res_disc_concl.Rmd
cleaning up and preparing to make public for submit

@@
 
-![](figures/Fig_1.png)
+![](figures/fig1.png)
 
@@
 
-![](figures/Fig_2.png)
+![](figures/fig2.png)
 
@@
 Comparison of alignment between AquaMaps and IUCN range data.__ (A) Distribution alignment (overlap of smaller range within larger) versus area ratio (the ratio of smaller range area to the larger range area) for `r format(n_spp_both, big.mark = ',')` species included in both IUCN and AquaMaps datasets.
-The upper right quadrant (quadrant 1) comprises species whose maps largely agree in both spatial distribution and the extent of described ranges (n = `r n_q1`; `r round(n_q1/nrow(spp_list_quads) * 100, 1)` %).
-The upper left quadrant (quadrant 2) comprises species whose maps agree well in distribution, but disagree in area
+The upper right quadrant comprises species whose maps largely agree in both spatial distribution and the extent of described ranges (n = `r n_q1`; `r round(n_q1/nrow(spp_list_quads) * 100, 1)` %).
+The upper left quadrant comprises species whose maps agree well in distribution, but disagree in area
 (n = `r n_q2`; `r round(n_q2/nrow(spp_list_quads) * 100, 1)` %).
-The lower right quadrant (quadrant 3) includes species for which the paired maps generally agree in range area, but disagree on where those ranges occur
+The lower right quadrant includes species for which the paired maps generally agree in range area, but disagree on where those ranges occur
 (n = `r n_q3`; `r round(n_q3/nrow(spp_list_quads) * 100, 1)` %).
-The lower left quadrant (quadrant 4) indicates species for which the map pairs agree poorly in both area and distribution
+The lower left quadrant indicates species for which the map pairs agree poorly in both area and distribution
 (n = `r n_q4`; `r round(n_q4/nrow(spp_list_quads) * 100, 1)` %).
@@
 
-![](figures/Fig_3.png)
+![](figures/fig3.png)
 
@@
 
-![](figures/Fig_4.png)
+![](figures/fig4.png)
 
@@
 
-![](figures/Fig_5.png)
+![](figures/fig5.png)
 

### 2016-09-25 ms_4_refs.Rmd
cleaning up and preparing to make public for submit

@@
 # Acknowledgments
-
-<!-- Acknowledgments. List acknowledgments and funding sources. PNAS collaborates with FundRef to use a standard taxonomy of funder names. Dedications are rarely allowed. -->
 
@@
 
-<!--
-References.
-References must be in PNAS style.
-Only published or in-press papers and books may be cited in the reference list.
-For references to in-press articles, please confirm with the cited journal that the article is in press and include a DOI number and online publication date.
-Unpublished abstracts of papers presented at meetings or references to "data not shown" are not permitted.
-References should be cited in numerical order as they appear in text, and all references, including for the SI, should be included in the main manuscript file.
-Because tables and figures will be inserted in the text where first cited, references in these sections should be numbered accordingly.
-Include the full title for each cited article.
-All authors (unless there are more than five) should be named in the citation.
-If there are more than five, list the first author's name followed by et al.
-Provide inclusive page ranges for journal articles and book chapters.
-Cite databases in the text or as footnotes.
-
-Journal articles are cited as follows:
-
-10.
-Neuhaus J-M, Sitcher L, Meins F, Jr, Boller T (1991) A short C-terminal sequence is necessary and sufficient for the targeting of chitinases to the plant vacuole.
-Proc Natl Acad Sci USA 88(22):10362–10366.
--->
+<!-- references will be automatically added from Zotero bib when knitting the document from manuscript-main.Rmd -->

### 2016-09-25 ms_5_figures.Rmd
cleaning up and preparing to make public for submit

@@
 
-![Taxonomic and geographic representation](figures/Fig_1.png)
+![Taxonomic and geographic representation](figures/fig1.png)
 
@@
 
-![Quadrant plot and spp breakdown by quadrant](figures/Fig_2.png)
+![Quadrant plot and spp breakdown by quadrant](figures/fig2.png)
 
@@
 
-![coral area subject to depth clipping](figures/Fig_3.png)
+![coral area subject to depth clipping](figures/fig3.png)
 
@@
 
-![aquamaps edge comparison](figures/Fig_4.png)
+![aquamaps edge comparison](figures/fig4.png)
 
@@
 
-![Change in MPA coverage](figures/Fig_5.png)
-
------
-
-``` {r figure_legends, child = 'ms_5_fig_legends.Rmd', eval = FALSE}
-```
+![Change in MPA coverage](figures/fig5.png)

### 2016-09-25 ms_6_supplemental.Rmd
cleaning up and preparing to make public for submit

@@
 title: "Supplemental Figures: Aligning marine species range data to better serve science and conservation"
-author: "Casey O'Hara, Jamie C. Afflerbach, Courtney Scarborough, Benjamin S. Halpern"
 output:
@@
 
-<!--
+## Supplemental figure captions
 
--->
-
-<!-- methods: -->
-<!-- IUCN raw -> grid -> raster -> final -->
-<!-- AM raw -> presence -->
-
-<!-- results/discussion: -->
-<!-- quad plot with risk and four species (?) -->
-<!-- four representative maps -->
-
-<!-- coral quadrant shifts -->
-
-<!-- AM thresholds? not mentioned in body of paper as of yet -->
-
-## Preparing AquaMaps and IUCN data layers
-
-![](figures/Fig_S1.png)
+![](figures/S1_fig.png)
 
 > __Fig S1.
-Examples of AquaMaps and IUCN raw species range data.__ (A, B) AquaMaps distribution of _Thunnus alalunga_ (Albacore Tuna) (AquaMaps 2013e) showing (A) global range with varying probabilities of occurrence assigned to 0.5° grid cells and (B) presence within 0.5° grid cells.
-(C, D) IUCN distribution of _T.
-alalunga_ (IUCN 2011b) represented as (C) extent of occurrence polygons and (D) presence within 0.5° grid cells.
+Examples of AquaMaps and IUCN raw species range data.__ (A, B) AquaMaps distribution of Thunnus alalunga (Albacore Tuna) [23] showing (A) global range with varying probabilities of occurrence assigned to 0.5° grid cells and (B) presence within 0.5° grid cells.
+(C, D) IUCN distribution of T.
+alalunga [24] represented as (C) extent of occurrence polygons and (D) presence within 0.5° grid cells.
 
@@
 
-![](figures/Fig_S2.png)
+![](figures/S2_fig.png)
 
 > __Fig S2.
-Rasterizing shapefiles provided by IUCN.__ A portion of the _T.
-alalunga_ range map is used to exemplify the rasterization process.
+Rasterizing shapefiles provided by IUCN.__ A portion of the T.
+alalunga range map [24] is used to exemplify the rasterization process.
 To enable direct comparison of IUCN species ranges to AquaMaps species ranges, the raw IUCN polygon (A) is overlaid with a 0.5° degree grid matching the AquaMaps grid (B).
@@
 
-## Representative maps from each quadrant
-
-![](figures/Fig_S3.png)
+![](figures/S3_fig.png)
 
 > __Fig S3.
-Representative species maps to illustrate each quadrant from Fig 2A.__ (AquaMaps 2013abcd; Conch 2013ab; IUCN 2011a; IUCN 2012) Each map is positioned to match its quadrant in Fig 2A.
-FAO Major Fishing Area boundaries [@fao_fisheries_and_aquaculture_department_cwp_2015] are outlined in light grey.
-(A) Distribution-matched: _Conus episcopatus_, the dignified cone snail.
+Representative species maps to illustrate each quadrant from Fig 2A.__ Each map is positioned to match its quadrant in Fig 2A.
+FAO Major Fishing Area boundaries [11] are outlined in light grey.
+(A) Distribution-aligned: Conus episcopatus, the dignified cone snail [25, 26].
 Distributions show excellent overlap in the western Pacific, though IUCN range extends well beyond the bounds of the AquaMaps range.
-(B) Well-matched: _Kajikia albida_, the Atlantic white marlin.  Distributions from each data set show nearly complete overlap, and very similar range size.
-(C) Poorly-matched: _Acanthurus nigroris_, the blue-lined surgeonfish.
+(B) Well-aligned: Kajikia albida, the Atlantic white marlin [27, 28].
+Distributions from each data set show nearly complete overlap, and very similar range size.
+(C) Poorly aligned: Acanthurus nigroris, the blue-lined surgeonfish [29, 30].
 IUCN predicts species distribution only near the Hawaiian islands; AquaMaps predicts extensive distribution throughout the central and western Pacific Ocean.
 The datasets align in neither distribution nor range size.
-(D) Area-matched: _Conus magnificus_, the magnificent cone snail.  Distributions overlap in the southern Pacific, but align poorly elsewhere.  The range sizes are similar.
+(D) Area-aligned: Conus magnificus, the magnificent cone snail [31, 32].
+Distributions overlap in the southern Pacific, but align poorly elsewhere.
+The range sizes are similar.
 
@@
 
-## Shifts in coral species range map alignment due to depth clipping
+![](figures/S4_fig.png)
 
-![](figures/Fig_S4.png)
-
-> __Fig.
-S4.
+> __Fig S4.
 Shift in alignment of paired-map coral species due to clipping IUCN ranges to areas shallower than 200 m.__ The grey lines represent the change in apparent alignment for a single species.
@@
 
-Conch Books, Hackenheim, Germany 2013. Conus episcopatus. The IUCN Red List of Threatened Species. Version 2016-2
+23.
+www.aquamaps.org, version of Aug.
+2013.
+Reviewed distribution maps for Thunnus alalunga (Albacore), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+Web.
+Accessed 11 Sep.
+2016.
 
-Conch Books, Hackenheim, Germany 2013. Conus magnificus. The IUCN Red List of Threatened Species. Version 2016-2
+24. International Union for Conservation of Nature (IUCN) 2011. Thunnus alalunga. The IUCN Red List of Threatened Species. Version 2016-2
 
-International Union for Conservation of Nature (IUCN) 2011. Kajikia albida. The IUCN Red List of Threatened Species. Version 2016-2
-
-International Union for Conservation of Nature (IUCN) 2011. Thunnus alalunga. The IUCN Red List of Threatened Species. Version 2016-2
-
-International Union for Conservation of Nature (IUCN) 2012. Acanthurus nigroris. The IUCN Red List of Threatened Species. Version 2016-2
-
+25.
 www.aquamaps.org, version of Aug.
@@
 
+26. Conch Books, Hackenheim, Germany 2013. Conus episcopatus. The IUCN Red List of Threatened Species. Version 2016-2.
+
+27.
+www.aquamaps.org, version of Aug.
+2013.
+Computer generated distribution maps for Kajikia albida (Atlantic white marlin), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+Web.
+Accessed 11 Sep.
+2016.
+
+28. International Union for Conservation of Nature (IUCN) 2011. Kajikia albida. The IUCN Red List of Threatened Species. Version 2016-2.
+
+29.
+www.aquamaps.org, version of Aug.
+2013.
+Reviewed distribution maps for Acanthurus nigroris (Bluelined surgeonfish), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
+Web.
+Accessed 11 Sep.
+2016.
+
+30. International Union for Conservation of Nature (IUCN) 2012. Acanthurus nigroris. The IUCN Red List of Threatened Species. Version 2016-2.
+
+31.
 www.aquamaps.org, version of Aug.
@@
 
-www.aquamaps.org, version of Aug.
-2013.
-Computer generated distribution maps for Kajikia albida (Atlantic white marlin), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
-Web.
-Accessed 11 Sep.
-2016.
-
-www.aquamaps.org, version of Aug.
-2013.
-Reviewed distribution maps for Acanthurus nigroris (Bluelined surgeonfish), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
-Web.
-Accessed 11 Sep.
-2016.
-
-www.aquamaps.org, version of Aug.
-2013.
-Reviewed distribution maps for Thunnus alalunga (Albacore), with modelled year 2100 native range map based on IPCC A2 emissions scenario.
-Web.
-Accessed 11 Sep.
-2016.
+32. Conch Books, Hackenheim, Germany 2013. Conus magnificus. The IUCN Red List of Threatened Species. Version 2016-2.

### 2016-09-25 case_study_ohi/boxplots.R
moved OHI case study to .zip

@@ -1,130 +1,0 @@
-### Create box plot for each scenario
-
-# get list of files (for all scenarios) - status (then eventually trend)
-# define base status as IUCN pref, 0.4 aquamaps thresh
-
-# smash together into a data frame of:
-# rgn_id | base status | scenario 1 status | scen 2 st | scen 3 st |
-
-# then box plot those bastards
-
-library(dplyr)
-library(tidyr)
-library(stringr)
-library(ggplot2)
-library(readr)
-
-dir_data  <- '~/github/ohiprep/globalprep/spp_ico/vAM_IUCN'
-dir_fig <- '~/github/IUCN-AquaMaps/figures'
-
-### generic theme for all plots
-ggtheme_basic <- theme(axis.ticks = element_blank(),
-                       text = element_text(family = 'Helvetica', color = 'gray30', size = 8),
-                       plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
-                       legend.position = 'none')
-
-ggtheme_plot <- ggtheme_basic + 
-  theme(panel.border = element_blank(),
-        panel.grid.minor = element_blank(), 
-        panel.grid.major = element_line(colour = 'grey90'),
-        panel.background = element_blank(),
-        axis.line = element_line(colour = "grey30"))
-
-
-st_base  <- read_csv(file.path(dir_data, 'data/spp_status_global_IUCNpref_prob0.4.csv'))
-st_scen1 <- read_csv(file.path(dir_data, 'data/spp_status_global_IUCNpref_prob0.csv'))
-st_scen2 <- read_csv(file.path(dir_data, 'data/spp_status_global_AMpref_prob0.4.csv'))
-st_scen3 <- read_csv(file.path(dir_data, 'data/spp_status_global_AMpref_prob0.csv'))
-
-### scenario 1: prefer IUCN data; drop AM probability threshold to 0.01
-### scenario 2: prefer AM data; keep AM prob thresh at 0.40
-### scenario 3: prefer AM data; drop AM prob thresh to 0.01
-
-st_all <- st_base %>% rename(st_b = score) %>%
-  left_join(st_scen1 %>% rename(st_1 = score), by = 'rgn_id') %>%
-  left_join(st_scen2 %>% rename(st_2 = score), by = 'rgn_id') %>%
-  left_join(st_scen3 %>% rename(st_3 = score), by = 'rgn_id')
-  
[…]

### 2016-09-28 ms_4_refs.Rmd
the submitted version

@@
 
-We are thankful to Melanie Frazier, Julia Stewart Lowndes, Halley Froelich, Claire Runge, and two anonymous reviewers for their insightful comments on earlier versions of this manuscript, and to the National Center for Ecological Analysis and Synthesis for computation support.
+We are thankful to Melanie Frazier, Julia Stewart Lowndes, Halley Froelich, and Claire Runge for their insightful comments on this manuscript, and to the National Center for Ecological Analysis and Synthesis for computation support.
 We acknowledge financial support from the Gordon and Betty Moore Foundation.

### 2016-12-19 data_explore/data_explore.Rmd
fixing the too many mammals issue

@@
 
-spp_am_file <- file.path(dir_data, 'spp_am.csv')
-spp_am <- read_csv(spp_am_file)
-
-spp_am <- spp_am %>%
+spp_am <- read_csv(file.path(dir_data, 'spp_am.csv')) %>%
   left_join(read_csv(file.path(dir_M, 'git-annex/globalprep/_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>% 
@@
 
-iucn_maps_class <- spp_iucn_maps %>% 
+iucn_maps_class_all <- spp_iucn_maps %>% 
   left_join(sciname_lookup %>% 
@@
               unique(), 
-            by = 'sciname')
+            by = 'sciname') %>%
+  select(-iucn_subpop, -spp_group) %>%
+  distinct() %>% ### first, get rid of dupes due to subpopulations
+  mutate(class    = str_to_title(class),
+         am_match = sciname %in% spp_am$sciname)
 
-taxa_iucn_all <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
+### next, identify spp scinames with matches in AM; if an iucn_sid has a match,
+### keep it; if there are multiple matches, keep the one with the fewest
+### instances of sciname (to keep more specific names...?)
+iucn_maps_class <- iucn_maps_class_all %>% 
+  group_by(sciname) %>%
+  mutate(sciname_count = n()) %>% ### for each sciname, figure out how many times it appears
+  group_by(iucn_sid) %>%
+  arrange(desc(am_match), sciname_count) %>% ### for each iucn_spp, put the AM matches on top; then...
+  filter(sciname == first(sciname)) %>% ### just keep the first instance of each iucn_spp
+  ungroup()
+
+taxa_iucn_all <- iucn_maps_class %>%
   filter(!is.na(class)) %>%           ### don't count the unknowns
-  mutate(class = str_to_title(class),
-         n_iucn_tot = n()) %>%        ### length of whole maps list
+  mutate(n_iucn_tot = n()) %>%        ### length of whole maps list
   group_by(class) %>%
@@
 
-taxa_iucn_solo <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
+taxa_iucn_solo <- iucn_maps_class %>%
   filter(!is.na(class)) %>%                 ### don't count the unknowns
   filter(!sciname %in% spp_am$sciname) %>%  ### filter list to IUCN only
-  mutate(class = str_to_title(class),
-         n_iucn_solo = n()) %>%             ### length of IUCN only list
+  mutate(n_iucn_solo = n()) %>%             ### length of IUCN only list
   group_by(class) %>%
@@
   filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  mutate(class = str_to_title(class),
-         n_am_tot = n()) %>%        ### length of whole AM maps list
+  mutate(n_am_tot = n()) %>%        ### length of whole AM maps list
   group_by(class) %>%
@@
   filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
-  mutate(class = str_to_title(class),
-         n_am_solo = n()) %>%       ### length of AM only list
+  mutate(n_am_solo = n()) %>%       ### length of AM only list
   group_by(class) %>%
@@
   filter(!class %in% c('Not assigned', 'null')) %>%
-  mutate(class = str_to_title(class),
-         n_both_tot = n()) %>%
+  mutate(n_both_tot = n()) %>%
   group_by(class) %>%
@@
 
+taxa_omitted <- taxa_am_all %>%
+  full_join(taxa_am_solo,   by = 'class') %>%
+  full_join(taxa_iucn_all,  by = 'class') %>%
+  full_join(taxa_iucn_solo, by = 'class') %>%
+  full_join(taxa_both,      by = 'class') %>%
+  filter(is.na(n_iucn_class_all) & p_am_class_all < 0.5)
+
 # knitr::kable(taxa_df %>%
@@
 #                              Both = n_class_both))
+
+write_csv(taxa_df, file.path(dir_data, 'taxa_df.csv'))
+write_csv(taxa_omitted, file.path(dir_data, 'taxa_omitted.csv'))
+
 ```
@@
 ### bar charts comparing number of species in each class for each data set
+
+taxa_df <- read_csv(file.path(dir_data, 'taxa_df.csv'))
 
@@
               arrange(desc(n_total))) %>%
+  bind_rows(taxa_omitted %>%
+              summarize(class = 'other (47 classes)',
+                        n_total = sum(n_am_class_all),
+                        dataset = 'AquaMaps',
+                        start = 0,
+                        end = 100)) %>%
   mutate(class = factor(class, levels = unique(.$class), ordered = TRUE),
@@
          dataset = factor(dataset, levels = c('AquaMaps', 'Paired', 'IUCN'), ordered = TRUE))
+
+### cut the labels for all but the first instance
+taxa_df2$n_label[duplicated(taxa_df2$class)] <- NA
 
@@
                  position = position_dodge(width = 0.5)) +
-  scale_size_manual(values = c('IUCN' = .5, 'AquaMaps' = .5, 'Paired' = 2.9)) +
+  scale_size_manual(values = c('IUCN' = .5, 'AquaMaps' = .5, 'Paired' = 2.8)) +
   scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey70', 'Paired' = 'black')) +
   geom_text(aes(label = n_label, y = 102),
-            size = 2.5,
+            size = 2.4,
             angle = 90, hjust = 0, color = 'grey30') +
-  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
+  theme(axis.text.x = element_text(angle = 75, hjust = 1),
+        axis.title.y = element_text(hjust = 0)) +
   scale_y_continuous(expand = c(0, 0),
@@
   labs(x = 'Class (in order of decreasing number of included species)',
-       y = 'Percent of mapped species',
+       y = '% of mapped species',
        fill = 'Dataset')
 
-print(bar_taxa)
+# print(bar_taxa)
 ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
@@
 
-![](../figures/bar2_by_taxa_all.png)
+![](../figures/bar_by_taxa_all.png)
 

### 2016-12-19 data_explore/data_explore.Rmd
revisiting analysis based on new matching algorithm

@@
 iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps.csv')
-spp_iucn_maps <- read_csv(iucn_maps_file) %>%
-  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
-  filter(sciname %in% spp_iucn$sciname)   ### ditch freshwater spp
+spp_iucn_maps <- read_csv(iucn_maps_file)
 
@@
 ``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
-
-sciname_lookup <- read_csv(file.path(dir_data, 'sciname_lookups/spp_iucn_scinames.csv'))
 
@@
 iucn_maps_class_all <- spp_iucn_maps %>% 
-  left_join(sciname_lookup %>% 
-              dplyr::select(sciname, class) %>%
-              unique(), 
-            by = 'sciname') %>%
   select(-iucn_subpop, -spp_group) %>%
@@
   full_join(taxa_both,      by = 'class') %>%
-  filter(is.na(n_iucn_class_all) & p_am_class_all < 0.5)
+  filter(is.na(n_iucn_class_all) & p_am_class_all < 0.5) %>%
+  left_join(read_csv(file.path(dir_data, 'sciname_lookups/scinames.csv')) %>%
+              select(class, kingdom, phylum) %>%
+              mutate(class = str_to_title(class)),
+            by = 'class') %>%
+  mutate(class = paste('other', str_to_title(kingdom), sep = ' ')) %>%
+  select(-kingdom, -phylum) %>%
+  distinct()
 
@@
               arrange(desc(n_total))) %>%
-  bind_rows(taxa_omitted %>%
-              summarize(class = 'other (47 classes)',
-                        n_total = sum(n_am_class_all),
+  bind_rows(x <- taxa_omitted %>%
+              group_by(class) %>%
+              summarize(n_total = sum(n_am_class_all),
                         dataset = 'AquaMaps',
@@
                  position = position_dodge(width = 0.5)) +
-  scale_size_manual(values = c('IUCN' = .5, 'AquaMaps' = .5, 'Paired' = 2.8)) +
+  scale_size_manual(values = c('IUCN' = .5, 'AquaMaps' = .5, 'Paired' = 2.6)) +
   scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey70', 'Paired' = 'black')) +

### 2016-12-19 data_explore/data_prep.Rmd
revisiting analysis based on new matching algorithm

@@
 
-dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
-           'Darwin'  = '/Volumes/data_edit',
+dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/data_edit',
+           'Darwin'  = '/Volumes/ohi',
            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
@@
 
-dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
+dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico')
 dir_fig <- file.path(dir_git, 'figures')
 dir_data <- file.path(dir_git, 'data')
-dir_data_remote <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
+dir_data_remote <- file.path(dir_M, 'git-annex/globalprep/_raw_data')
 scenario <- 'v2016'
@@
 
-Need to:
+```{r get_spp_all, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
-gather spp_all, spp_am, spp_iucn, spp_iucn_maps, and spp_am_raw into local data folder.
+spp_iucn_maps <- read_csv(file.path(dir_data, 'spp_iucn_maps.csv'))
 
-``` {r pull_files_to_local, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
-### pull spp_iucn_marine.csv from git-annex and bring to local repo
-spp_iucn <- read_csv(file.path(dir_anx, scenario, 'int/spp_iucn_cleaned.csv'))
-spp_iucn <- spp_iucn %>%
-  dplyr::select(sciname, iucn_sid, iucn_category, 
-                popn_trend, parent_sid, subpop_sid) %>%
-  unique()
-write_csv(spp_iucn, file.path(dir_data, 'spp_iucn.csv'))
+spp_am_maps <- read_csv(file.path(dir_data, 'speciesoccursum.csv')) %>%
+  mutate(sciname = paste(str_trim(genus), str_trim(species)),
+         iucn_id = as.integer(iucn_id)) %>%
+  dplyr::select(am_sid = speciesid, sciname, reviewed, iucn_id)
 
-### pull spp_iucn_maps_all.csv from git-annex and bring to local repo
-spp_iucn_maps <- read_csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv')) %>%
-  rename(iucn_sid = id_no)
+spp_all <- spp_iucn_maps %>%
+  full_join(spp_am_maps, by = 'sciname') %>%
+  distinct()
 
-spp_iucn_maps <- spp_iucn_maps %>%
-  filter(!str_detect(spp_group, 'MARINEFISH')) %>%
-  unique() %>%
-  arrange(iucn_sid)
+### some confusion due to non-matching scinames and IDs...
+spp_all <- spp_all %>%
+  mutate(iucn_match = iucn_sid == iucn_id,
+         iucn_wrong_map = ifelse(!iucn_match & iucn_id %in% spp_all$iucn_sid, TRUE, FALSE)) %>%
+  mutate(iucn_sid = ifelse(iucn_wrong_map & !is.na(iucn_wrong_map), iucn_id, iucn_sid)) %>%
+  select(-iucn_match, -iucn_wrong_map, -iucn_id, -sciname_count) %>%
+  distinct()
 
-### At this point, still some duplicated iucn_sids, due to several
-### names matching a single ID.  Not sure which names might be in
-### AquaMaps though, so leave 'em in.
+length(unique(x$iucn_sid)) ###  4013, incl NA
+length(unique(x$am_sid))   ### 22890, incl NA
 
-write_csv(spp_iucn_maps, file.path(dir_data, 'spp_iucn_maps.csv'))
-
-### pull cleaned spp_am data from git-annex and bring to local repo
-spp_am <- read_csv(file.path(dir_anx, scenario, 'int/spp_am_cleaned.csv'))
-write_csv(spp_am, file.path(dir_data, 'spp_am.csv'))
-
-### pull raw aquamaps species list from git-annex and bring to local repo
-spp_am_raw <- read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'))
-write_csv(spp_am_raw, file.path(dir_data, 'speciesoccursum.csv'))
-
-```
-
-```{r get_spp_all, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
-### get spp_all_complete from git-annex, bring to local repo.  The spp_all_complete
-### file was created in ohiprep/globalprep/SPP_ICO/data_prep_SPP.R, running the
-### function create_spp_master_lookup(source_pref = 'iucn', fn_tag = '_complete', reload = TRUE)
-spp_all_remote <- read_csv(file.path(dir_anx, scenario, 'int/spp_all.csv'))
-spp_all_remote_clean <- spp_all_remote %>%
-  filter(!str_detect(spatial_source, 'subpop')) %>%
-  filter(!str_detect(spatial_source,  'alias'))
-### At this point, plenty of duplicates:
-### * This code returns 984 duplicated names:
-###     length(spp_all_remote_clean$sciname) - length(unique(spp_all_remote_clean$sciname))
-### * some are AM subpops which were ignored in processing the remote spp_all file.
-###   Note: All subpops have an identified parent, so can filter them out without
-###   losing any species.
-###   code to verify this:
-###     z_parent <- spp_all_remote_clean %>% 
-###       filter(!is.na(subpop_sid) & spatial_source == 'am')
-###     z_subpop <- spp_all_remote_clean %>% 
-###       filter(!is.na(parent_sid) & spatial_source == 'am')
-###     nrow(z_subpop %>% filter(!sciname %in% z_parent$sciname))
-### * some are duplicated IUCN parents (e.g. Dermochelys coriacea) that will be 
-###   eliminated when select(-iucn_subpop) column and then unique()
-
-### filter out spatial_source == 'am' that also have parent_sid (i.e.
-### these are AM-mapped subpopulations); then: 
-### select away the iucn_subpop, parent_sid, and subpop_sid (among others),
-### these are creating false unique rows.
-spp_all <- spp_all_remote_clean %>%
-  filter(!(spatial_source == 'am' & !is.na(parent_sid))) %>%
-  dplyr::select(am_sid, sciname, reviewed, iucn_sid, popn_trend,
-                popn_category, category_score, spp_group, 
-                id_no, spatial_source, name_verified) %>%
-  unique()
-### At this point, still some duped scinames - these are ones with multiple IUCN
-### SIDs that are not designated as parents or subpops.  We will brute-force
-### this by just eliminating the duplicated rows, keeping only the first instance.
-### By arranging by iucn_sid, we will keep the lowest value, which seems to
-### generally correspond to the parent...
-spp_all <- spp_all %>%
-  arrange(iucn_sid)
-spp_all <- spp_all[!duplicated(spp_all$sciname), ]
-
-length(spp_all$sciname) - length(unique(spp_all_remote_clean$sciname))
-### no duped scinames - same length as original (cleaned) list
-
-### But, some of the extracted IUCN maps don't appear on this list! add 'em in.
-spp_iucn_maps_new <- read_csv(file.path(dir_data, 'spp_iucn_maps.csv')) %>%
-  filter(!sciname %in% spp_all$sciname) %>%
-  filter(!str_detect(sciname, 'subpop')) %>% ### a Hawaiian subpop... ditch it!
-  select(-iucn_subpop)
-
-spp_all1 <- spp_all %>%
-  full_join(spp_iucn_maps_new)
-
-### Also, some of the AM listings have disappeared due to name conflicts
-### - aliases, etc.
-spp_am_new <- read_csv(file.path(dir_data, 'speciesoccursum.csv')) %>%
-  mutate(sciname = paste(str_trim(genus), str_trim(species)),
-         iucn_sid = as.integer(iucn_id)) %>%
-  dplyr::select(am_sid = speciesid, sciname, reviewed, iucn_sid,
-                popn_category = iucn_code) %>%
-  filter(!am_sid %in% spp_all$am_sid) %>%
-  mutate(spatial_source = 'am',
-         popn_category = ifelse(popn_category == 'LR/lc', 'LC', popn_category),
-         popn_category = ifelse(popn_category == 'LR/nt', 'NT', popn_category),
-         popn_category = ifelse(popn_category == 'N.E.',  'NE', popn_category))
-
-### 545 new spp_am not in spp_all from remote source, by am_sid.
-### 24 of these, however, list iucn_sids that match existing iucn_sid.
-### * for species with iucn_sid that match, and no am_sid, add am_sid to 
-###   existing record (use existing values for other fields; drop new values).
-### * for species with iucn_sid that match, and with a different am_sid, no action.
-### * for species with non-matching or NA iucn_sid, just add to list.
-spp_am_new_iucnmatch <- spp_am_new %>% 
-  filter(!is.na(iucn_sid) & iucn_sid %in% spp_all$iucn_sid) %>%
-  dplyr::select(iucn_sid, am_sid_new = am_sid, reviewed_new = reviewed)
-spp_all2 <- spp_all1 %>%
-  left_join(spp_am_new_iucnmatch, by = 'iucn_sid') %>%
-  mutate(am_sid   = ifelse(is.na(am_sid) & !is.na(am_sid_new), am_sid_new, am_sid),
-         reviewed = ifelse(is.na(am_sid) & !is.na(am_sid_new), reviewed_new, reviewed)) %>%
-  select(-am_sid_new, -reviewed_new)
-spp_all3 <- spp_all2 %>%
-  full_join(spp_am_new %>%
-              filter(!iucn_sid %in% spp_am_new_iucnmatch$iucn_sid),
-            by = c("am_sid", "sciname", "reviewed", "iucn_sid", "popn_category", "spatial_source"))
-
-write_csv(spp_all3, file.path(dir_data, 'spp_all.csv'))
+write_csv(spp_all, file.path(dir_data, 'spp_all_maps.csv'))
 
@@
 ### that also have a non-NA AquaMaps ID code.
-spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))
+spp_all <- read_csv(file.path(dir_data, 'spp_all_maps.csv'))
 
 spp_am_v_iucn <- spp_all %>%
-  filter(str_detect(spatial_source, 'iucn') & !is.na(am_sid)) %>%
-  filter(!is.na(id_no)) %>%
-  filter(!str_detect(spp_group, 'MARINEFISH')) %>%
+  select(-iucn_subpop) %>%
+  filter(!is.na(iucn_sid) & !is.na(am_sid)) %>%
   unique()
-# length(unique(spp_am_v_iucn$am_sid)); length(unique(spp_am_v_iucn$iucn_sid))
-### NOTE: 2345 unique scientific names; but 2283 unique am_sid and 2249 unique iucn_sid.
-### Some unique species (as identified by am_sid and iucn_sid) have multiple
-### scinames.  At this point, not really important what the sciname is?
-spp_am_v_iucn1 <- spp_am_v_iucn %>%
+
+# length(unique(spp_am_v_iucn$sciname)); length(unique(spp_am_v_iucn$am_sid)); length(unique(spp_am_v_iucn$iucn_sid))
+### NOTE: 2314 unique scientific names; but 2314 unique am_sid and 2320 unique iucn_sid.
+### Seasnakes are included in Reptiles (IUCN).  Two "mangroves" that are actually seagrasses.
+spp_am_v_iucn <- spp_am_v_iucn %>%
   group_by(am_sid, iucn_sid) %>%
-  mutate(dupes = n() - 1,
-         first_instance = sciname == first(sciname)) %>%
-  filter(!(dupes > 0 & !first_instance)) %>%
-  select(-dupes, -first_instance) %>%
-  ungroup()
-# length(unique(spp_am_v_iucn1$am_sid)); length(unique(spp_am_v_iucn1$iucn_sid))
+  mutate(dupes = n() - 1) %>%
+  ungroup() %>%
+  filter(!(dupes == 1 & spp_group == 'MANGROVES')) %>%
+  filter(spp_group != 'SEASNAKES') %>%
+  select(-dupes)
+# length(unique(spp_am_v_iucn$sciname)); length(unique(spp_am_v_iucn$am_sid)); length(unique(spp_am_v_iucn$iucn_sid))
 
-### What species are represented by the duplicated IUCN ID numbers?  These
-### are species that IUCN experts lumped together (via synonyms etc) 
-### but are treated as separate by AquaMaps (the reverse does not seem
-### to happen)...  There are 64 AM species IDs and only 30 IUCN species IDs.  
-### How to treat in the mapping?
-### * lump AM species together and plot against a single IUCN species map?
-###   That would occur in the area determining function - group by iucn_sid
-###   and map all am_sids against that iucn_sid
-x <- spp_am_v_iucn1 %>%
-  group_by(iucn_sid) %>%
+sciname_dupes <- spp_am_v_iucn %>%
+  group_by(sciname) %>%
   filter(n() > 1)
-# length(unique(x$am_sid)); length(unique(x$iucn_sid))
+### these seem to be spp with IUCN maps for synonyms.  Keep 'em in.
 
-write_csv(spp_am_v_iucn1, file.path(dir_data, 'spp_am_v_iucn.csv'))
+write_csv(spp_am_v_iucn, file.path(dir_data, 'spp_am_v_iucn.csv'))
+
 ```
@@
 
-### This file is generated in globalprep/SPP_ICO:
-iucn_spp_cells_all <- read_csv(file.path(dir_data_remote, 'iucn_spp/iucn_cells_v2016.csv'),
-                            col_types = 'cddddc')
+# iucn_spp_cells_files <- list.files(file.path(dir_anx, 'v2016/iucn_intersections'), full.names = TRUE)
+# iucn_spp_cells_all <- lapply(iucn_spp_cells_files, FUN = function(x) {
+#   read_csv(x) %>% 
+#     mutate(id_no = as.integer(id_no)) %>%
+#     rename(iucn_sid = id_no)
+#   }) %>%
+#   bind_rows()
+# 
+# iucn_spp_cells <- iucn_spp_cells_all %>%
+#   filter(iucn_sid %in% spp_am_v_iucn$iucn_sid) %>%
+#   filter(presence != 5)
 
-iucn_spp_cells <- iucn_spp_cells_all %>%
-  rename(iucn_sid = id_no) %>%
-  filter(iucn_sid %in% spp_am_v_iucn$iucn_sid) %>%
-  filter(presence != 5)
-### length(unique(iucn_spp_cells$iucn_sid)) results in 2279 species found...
-### some species with no cells reported.  2279 listed in the am v iucn file.
-write_csv(iucn_spp_cells, file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'))
+### length(unique(iucn_spp_cells$iucn_sid)) results in 2271 species found...
+###   some species with no cells reported.  
+### spp_missing <- spp_am_v_iucn %>% filter(!iucn_sid %in% iucn_spp_cells$iucn_sid)
+###   49 spp with no cells reported.  Shapefile errors?
+# write_csv(iucn_spp_cells, file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'))
 
 message('Loading AquaMaps species to cell table.  This might take a few minutes.')
-am_spp_cells_all <- fread(file.path(dir_data_remote, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv'), 
+am_spp_cells_all <- fread(file.path(dir_data_remote, 'aquamaps/d2015/csv/hcaf_sp_native_trunc.csv'), 
                          showProgress = TRUE)
@@
   filter(am_sid %in% spp_am_v_iucn$am_sid)
-### length(unique(am_spp_cells$am_sid)) results in 2314 species found - great!
+### length(unique(am_spp_cells$am_sid)) results in 2314 species found
 write_csv(am_spp_cells, file.path(dir_anx, 'vAM_IUCN/am_spp_cells_2015.csv'))
@@
   if(!exists('iucn_spp_cells'))
-    iucn_spp_cells <- read_csv(file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'),
-                               col_types = 'cddddc')
+    iucn_spp_cells <- read_csv(file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'))
+
   ### species info list - scinames, IUCN species ID, AM species ID, category etc

### 2016-12-27 data_explore/data_explore.Rmd
pushing recent updates to analysis and documentation based on reviewer comments

@@
 data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
-spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
-  filter(!is.na(popn_category)) %>% ### filter out unassessed species
-  filter(!is.na(sm_perc))           ### filter out NA maps
+spp_map_pairs <- read_csv(data_file)
 
@@
             by = 'spp_group_text') %>%
-  left_join(quad_names, by = 'quad')
+  left_join(quad_names, by = 'quad') %>% 
+  mutate(quad = factor(quad, levels = c('q4', 'q3', 'q2', 'q1'))) %>%
+  transform(spp_group_text = reorder(spp_group_text, pct_q1))
 
@@
 ### Plot the bar chart
-barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
-                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
+barchart_spp_gp_quads <- ggplot(spp_gp_quadrants, 
                                 aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
   ggtheme_plot + 
-  theme(panel.grid.major.y = element_blank()) +
+  # theme(panel.grid.major.y = element_blank()) +
   geom_bar(stat = 'count', alpha = 1) +
-  scale_fill_manual(values = c('q4' = '#d01c8b', 
+  scale_fill_manual(values = c('q1' = '#4dac26', 
+                               'q2' = '#b8e186', 
                                'q3' = '#f1b6da',
-                               'q2' = '#b8e186',
-                               'q1' = '#4dac26'),
+                               'q4' = '#d01c8b'),
                     labels = quad_names$quad_name) +
@@
 # ggsave(plot = barchart_spp_gp_quads,
-#        filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
-#        width = 19.05, height = 7, units = 'cm', dpi = 600)
+       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
+       width = 19.05, height = 7, units = 'cm', dpi = 600)
 
@@
 # Plots for SI
+
+## quad plot experts only
+
+``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+
+spp_list_experts <- spp_list %>%
+  filter(reviewed == 1)
+
+### means for both x and y axes - use original means
+area_align_mean <- mean(spp_list$area_ratio, na.rm = TRUE)
+dist_align_mean <- mean(spp_list$sm_perc, na.rm = TRUE)
+
+area_align_mean_ex <- mean(spp_list_experts$area_ratio, na.rm = TRUE)
+dist_align_mean_ex <- mean(spp_list_experts$sm_perc, na.rm = TRUE)
+
+### define windows for labels
+q1 <- c('x1' = 84, 'x2' = 98, 'y1' = 92.5, 'y2' = 97.5)
+q2 <- c('x1' =  2, 'x2' = 22, 'y1' = 92.5, 'y2' = 97.5)
+q3 <- c('x1' = 84, 'x2' = 98, 'y1' =  2.5, 'y2' = 7.5)
+q4 <- c('x1' =  2, 'x2' = 18, 'y1' =  2.5, 'y2' = 7.5)
+
+spp_list_experts_labs <- spp_list_experts %>%
+  rename(x = area_ratio, y = sm_perc) %>%
+  mutate(fade = FALSE,
+         fade = ifelse((x > q1[1] & x < q1[2] & y > q1[3] & y < q1[4]), TRUE, fade),
+         fade = ifelse((x > q2[1] & x < q2[2] & y > q2[3] & y < q2[4]), TRUE, fade),
+         fade = ifelse((x > q3[1] & x < q3[2] & y > q3[3] & y < q3[4]), TRUE, fade),
+         fade = ifelse((x > q4[1] & x < q4[2] & y > q4[3] & y < q4[4]), TRUE, fade))
+
+break_nums <- seq(0, 100, 20)
+
+### mongo plot time
+scatter_quadplot_ex <- ggplot(spp_list_experts_labs, aes(x = x, y = y)) +
+  ggtheme_plot + 
+  ### color the quadrant backgrounds:
+  annotate("rect", xmin = area_align_mean, xmax = 100, ymin = dist_align_mean, ymax = 100, 
+           alpha = .3, fill= "#4dac26")  + 
+  annotate("rect", xmax = area_align_mean, xmin =   0, ymin = dist_align_mean, ymax = 100, 
+           alpha = .3, fill= "#b8e186") + 
+  annotate("rect", xmin = area_align_mean, xmax = 100, ymax = dist_align_mean, ymin =   0, 
+           alpha = .3, fill= "#f1b6da") + 
+  annotate("rect", xmax = area_align_mean, xmin =   0, ymax = dist_align_mean, ymin =   0, 
+           alpha = .3, fill= "#d01c8b") +
+  ### add grid lines
+  geom_vline(xintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
+  geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5)
+
+### plot the data points!
+scatter_quadplot_ex <- scatter_quadplot_ex +
+  # geom_point(color = '#4d4dac', aes(alpha = fade), show.legend = FALSE) +
+  # geom_point(data = spp_list_experts_labs %>%
+  #              filter(str_detect(spp_group_text, 'Corals')),
+  #            color = 'red', aes(alpha = fade), 
+  #            shape = 4, stroke = .5, 
+  #            size = 1, show.legend = FALSE) +
+  # scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = 1)) +
+  geom_point(color = 'black', aes(alpha = fade), size = .1, show.legend = FALSE) +
+  scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = 1))
+
+### label the quadrants:
+scatter_quadplot_ex <- scatter_quadplot_ex +
+  annotate("text", x = 97, y = 95, hjust = 1, vjust = .5, size = 3, color = 'grey20', 
+           fontface = 'bold.italic', label = "Well-aligned") + 
+  annotate("text", x =  3, y = 95, hjust = 0, vjust = .5, size = 3, color = 'grey20',
+           fontface = 'bold.italic', label = "Distribution-aligned") + 
+  annotate("text", x = 97, y =  5, hjust = 1, vjust = .5, size = 3, color = 'grey20',
+           fontface = 'bold.italic', label = "Area-aligned") + 
+  annotate("text", x =  3, y =  5, hjust = 0, vjust = .5, size = 3, color = 'grey20',
+           fontface = 'bold.italic', label = "Poorly aligned")
+
+### Manage axis scales 
+scatter_quadplot_ex <- scatter_quadplot_ex +
+  scale_x_continuous(expand = c(0, 0), limits = c(-1, 101),
+                     breaks = break_nums,
+                     labels = sprintf('%s%%', break_nums)) + 
+  scale_y_continuous(expand = c(0, 0), limits = c(-1, 101),
+                     breaks = break_nums,
+                     labels = sprintf('%s%%', break_nums))
+
+### here are mean labels:
+scatter_quadplot_ex <- scatter_quadplot_ex +
+  annotate(geom = 'text', x = area_align_mean, y = 5,
+           hjust = 0, vjust = 0, color = 'grey30', size = 2,
+           fontface = 'bold.italic', angle = 90,
+           label = sprintf('Mean (all) = %s%%', round(area_align_mean, 1))) +
+  annotate(geom = 'text', x = 5, y = dist_align_mean,
+           hjust = 0, vjust = 0, color = 'grey30', size = 2,
+           fontface = 'bold.italic', angle = 0,
+           label = sprintf('Mean (all) = %s%%', round(dist_align_mean, 1))) +
+  geom_hline(yintercept = dist_align_mean_ex, color = 'grey40', linetype = 2, size = 0.25) + 
+  geom_vline(xintercept = area_align_mean_ex, color = 'grey40', linetype = 2, size = 0.25) + 
+  annotate(geom = 'text', x = area_align_mean_ex - .25, y = 5,
+           hjust = 0, vjust = 0, color = 'grey30', size = 2,
+           fontface = 'bold.italic', angle = 90,
+           label = sprintf('Mean (reviewed) = %s%%', round(area_align_mean_ex, 1))) +
+  annotate(geom = 'text', x = 5, y = dist_align_mean_ex + .25,
+           hjust = 0, vjust = 0, color = 'grey30', size = 2,
+           fontface = 'bold.italic', angle = 0,
+           label = sprintf('Mean (reviewed) = %s%%', round(dist_align_mean_ex, 1)))
+
+scatter_quadplot_ex <- scatter_quadplot_ex +
+  labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
+       y = bquote('Distribution alignment '*alpha['dist']*' (%)'))
+
+# print(scatter_quadplot_ex)
+
+# ggsave(file.path(dir_fig, 'scatter_quadmap_expert.png'),
+#        width = 19.05, height = 9, units = 'cm', dpi = 600)
+
+```
+
+![](../figures/scatter_quadmap.png)
+
+## Break down quads by species group for expert reviewed maps
+
+``` {r examine_spp_gp_quartiles, echo = FALSE, eval = TRUE}
+
+### means for both x and y axes - use original means
+area_align_mean <- mean(spp_list$area_ratio, na.rm = TRUE)
+dist_align_mean <- mean(spp_list$sm_perc, na.rm = TRUE)
+
+spp_list_experts1 <- spp_list_experts %>%
+  group_by(spp_group_text) %>%
+  mutate(n_spp = n()) %>%
+  ungroup()
+
+### Create species lists by quadrant
+spp_list_experts_q1 <- spp_list_experts1 %>%
+  filter(area_ratio >= area_align_mean & sm_perc >= dist_align_mean)
+spp_list_experts_q2 <- spp_list_experts1 %>%
+  filter(area_ratio <  area_align_mean & sm_perc >= dist_align_mean)
+spp_list_experts_q3 <- spp_list_experts1 %>%
+  filter(area_ratio >= area_align_mean & sm_perc < dist_align_mean)
+spp_list_experts_q4 <- spp_list_experts1 %>%
+  filter(area_ratio <  area_align_mean & sm_perc < dist_align_mean)
+
+spp_list_experts_quads <- spp_list_experts1 %>%
+  mutate(quad = NA,
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_experts_q1$sciname, 'q1', quad),
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_experts_q2$sciname, 'q2', quad),
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_experts_q3$sciname, 'q3', quad),
+         quad = ifelse(is.na(quad) & sciname %in% spp_list_experts_q4$sciname, 'q4', quad))
+write_csv(spp_list_experts_quads, file.path(dir_data, 'spp_list_experts_quads.csv'))
+
+### Summarize quadrant species list by species group (e.g. CORALS1)
+spp_gp_q1_ex <- spp_list_experts_q1 %>%
+  group_by(spp_group_text) %>%
+  summarize(n_spp = first(n_spp), n_spp_q1 = n())
+spp_gp_q2_ex <- spp_list_experts_q2 %>%
+  group_by(spp_group_text) %>%
+  summarize(n_spp = first(n_spp), n_spp_q2 = n())
+spp_gp_q3_ex <- spp_list_experts_q3 %>%
+  group_by(spp_group_text) %>%
+  summarize(n_spp = first(n_spp), n_spp_q3 = n())
+spp_gp_q4_ex <- spp_list_experts_q4 %>%
+  group_by(spp_group_text) %>%
+  summarize(n_spp = first(n_spp), n_spp_q4 = n())
+
+### Join species group by quadrant into a single dataframe
+spp_gp_quads_ex <- spp_gp_q1_ex %>%
+  full_join(spp_gp_q2_ex, by = c('spp_group_text', 'n_spp')) %>%
+  full_join(spp_gp_q3_ex, by = c('spp_group_text', 'n_spp')) %>%
+  full_join(spp_gp_q4_ex, by = c('spp_group_text', 'n_spp')) %>%
+  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
+  mutate(quad = str_replace(quad, 'n_spp_', ''),
+         pct_quad = n_quad/n_spp,
+         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
+         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))
+
+quad_names <- data.frame('quad' = c('q1', 'q2', 'q3', 'q4'),
+                         'quad_name' = factor(c('well-aligned', 'dist-aligned', 
+                                                'area-aligned', 'poorly aligned'),
+                                             ordered = TRUE))
+
+spp_gp_quads_ex <- spp_gp_quads_ex %>%
+  left_join(spp_gp_quads_ex %>% 
+              filter(quad == 'q1') %>%
+              dplyr::select(spp_group_text, pct_q1 = pct_quad),
+            by = 'spp_group_text') %>%
+  left_join(quad_names, by = 'quad') %>% 
+  mutate(quad = factor(quad, levels = c('q4', 'q3', 'q2', 'q1'))) %>%
+  transform(spp_group_text = reorder(spp_group_text, pct_q1))
+
+break_nums <- seq(0, 100, 20)
+
+### Plot the bar chart
+bar_spp_gp_quads_ex <- ggplot(spp_gp_quads_ex, 
+                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
+  ggtheme_plot + 
+  # theme(panel.grid.major.y = element_blank()) +
+  geom_bar(stat = 'count', alpha = 1) +
+  scale_fill_manual(values = c('q1' = '#4dac26', 
+                               'q2' = '#b8e186', 
+                               'q3' = '#f1b6da',
+                               'q4' = '#d01c8b'),
+                    labels = quad_names$quad_name) +
+  scale_y_continuous(expand = c(0, 0), 
+                     limits = c(0, 1.1),
+                     breaks = break_nums/100,
+                     labels = sprintf('%s%%', break_nums)) + 
+  ### add grid lines; horizontal but then get flipped
+  geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
+  ### add text
+  geom_text(# data = spp_gp_quadrants %>%
+              # select(spp_group_text, n_spp) %>%
+              # distinct(),
+            aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
+            size = 2.5, 
+            color = 'grey30') +
+  coord_flip() +
+  labs(x = 'Taxonomic Group', 
+       y = 'Percent of species by quadrant', 
+       fill = 'Alignment')
+
+print(bar_spp_gp_quads_ex)
+
+# ggsave(plot = bar_spp_gp_quads_ex,
+       filename = file.path(dir_fig, 'barchart_spp_gp_quads_ex.png'),
+       width = 19.05, height = 7, units = 'cm', dpi = 600)
+
+write_csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads_ex.csv'))
+
+```
+
+![](../figures/barchart_spp_gp_quads.png)
 

### 2016-12-27 data_explore/data_explore_am_thresh.Rmd
pushing recent updates to analysis and documentation based on reviewer comments

@@
 spp_cells <- read_csv(file.path(dir_aquamaps, 'csv/hcaf_sp_native_trunc.csv'),
-                      col_types = 'ccd', progress = TRUE)
+                      progress = TRUE)
 
@@
 
+```
+
+``` {r inspect_extrapolation_ratios}
+spp_list <- read_csv(file.path(dir_git, 'data/spp_list_w_area_trimmed.csv'))
+hcaf <- read_csv(file.path(dir_aquamaps, 'csv/hcaf_truncated.csv'))
+
+spp_cell_matches <- spp_cells %>%
+  rename(am_sid = speciesid) %>%
+  filter(am_sid %in% spp_list$am_sid | am_sid %in% c('Fis-131802', 'W-Oph-245456')) %>%
+  left_join(hcaf, by = 'loiczid') %>%
+  mutate(probability = as.numeric(probability)) %>%
+  select(-csquarecode, -contains('limit'), -contains('center'))
+
+spp_cell_sums <- spp_cell_matches %>%
+  group_by(am_sid) %>%
+  summarize(n_cells = n(),
+            area    = sum(oceanarea, na.rm = TRUE),
+            area_wt = sum(oceanarea * probability, na.rm = TRUE)) %>%
+  left_join(spp, by = c('am_sid' = 'speciesid')) %>%
+  select(am_sid, n_cells, area, area_wt, sciname, reviewed, occurcells) %>%
+  mutate(geog_space_extrapolate = area / occurcells,
+         env_space_extrapolate = area_wt / occurcells)
+
+spp_cell_sums_Oplana <- spp_cells %>%
+  filter(speciesid == 'W-Oph-245456')
+spp_cell_sums_Rfyllae <- spp_cells %>%
+  filter(speciesid == 'Fis-131802')
+
+hist(spp_cell_sums_Oplana$probability)
+hist(spp_cell_sums_Rfyllae$probability)
+
+summary(spp_cell_sums_Oplana$probability)
+summary(spp_cell_sums_Rfyllae$probability)
+### Very similar distributions
+# > summary(spp_cell_sums_Oplana$probability)
+#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+#  0.0100  0.2300  0.6000  0.5725  0.9600  1.0000 
+# > summary(spp_cell_sums_Rfyllae$probability)
+#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+#  0.0100  0.2300  0.6500  0.5981  1.0000  1.0000
+
+### How do prob-weighted areas compare to non-prob-weighted areas?  consistent 
+### ratio would indicate that the uncertainty due to # of occurcells is not
+### being considered in the predictions
+
+spp_cell_sums1 <- spp_cell_sums %>%
+  mutate(a_to_a_wt = area / area_wt,
+         gspace_to_espace = geog_space_extrapolate / env_space_extrapolate) %>%
+  arrange(desc(a_to_a_wt)) %>%
+  filter(a_to_a_wt < 50) ### get rid of Zalophus wollbaeki outlier
+
+summary(spp_cell_sums1$a_to_a_wt)
+  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+  # 1.023   1.258   1.371   1.495   1.571   8.062 
+summary(spp_cell_sums1$gspace_to_espace)
+### haha the same, duh
+### but fairly consistent ratios seem to indicate no relationship
+
+summary(lm(a_to_a_wt ~ occurcells, data = spp_cell_sums1))
+# Adjusted R-squared:  0.01058 - essentially no explanatory power
 ```

### 2016-12-27 data_explore/data_prep.Rmd
pushing recent updates to analysis and documentation based on reviewer comments

@@
            'Darwin'  = '/Volumes/ohi',
-           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
+           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]
 
@@
 
-loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
+loiczid_raster_file  <- file.path(dir_git, 'data/rasters/loiczid_raster.tif')
 loiczid_raster       <- raster::raster(loiczid_raster_file)
@@
   spp_list_base <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))
+  create_spp_list2(spp_list_base, data_file, am_spp_cells, iucn_spp_cells, loiczid_raster)
 
-  create_spp_list(spp_list_base, data_file, am_spp_cells, iucn_spp_cells, loiczid_raster)
 }
@@
             by = 'spp_group') %>%
-  left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>% 
+  left_join(read_csv(file.path(dir_data_remote, 'aquamaps/d2015/csv/speciesoccursum.csv')) %>% 
               select(am_sid = speciesid, occurcells),
@@
 
-cat_score_df <- data.frame(popn_category  = c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE'),
-                           category_score = c( 0.0,  0.2,  0.4,  0.6,  0.8,  1.0,   NA,  NA))
-spp_list <- spp_list %>%
-  left_join(cat_score_df, by = 'popn_category')
+# cat_score_df <- data.frame(popn_category  = c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE'),
+#                            category_score = c( 0.0,  0.2,  0.4,  0.6,  0.8,  1.0,   NA,  NA))
+# spp_list <- spp_list %>%
+#   left_join(cat_score_df, by = 'popn_category')
 
@@
 ``` {r counts_spp_all_observations, echo = FALSE}
+
 spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))
@@
 
-totes_occur <- obis_occur %>%
-  left_join(gbif_occur) %>%
-  rowwise() %>%
-  mutate(mean_occ = mean(c(gbif_occ, obis_occ), na.rm = TRUE))
-
 ```
-
-__Possible "data poor" indicators:__
-
-* (of matched species) how does bottom decile get distributed across quadrants
-* use mean of OBIS and GBIF, and just report average (or median?) occurrences across all species in each quadrant
-* also use similar method for AM occurcells
-
-``` {r data_poor_analysis, echo = FALSE, eval = FALSE}
-spp_quads <- read_csv(file.path(dir_data, 'spp_list_quads.csv')) %>% 
-  select(am_sid, iucn_sid, occurcells, quad) %>%
-  unique()
-### This contains all mapped species, discarding those with no actual IUCN maps
-
-spp_occ_counts <- spp_quads %>%
-  left_join(totes_occur) %>%
-  group_by(iucn_sid, am_sid, quad) %>%
-  summarize(occurcells = mean(occurcells, na.rm = TRUE), occ = mean(mean_occ, na.rm = TRUE)) %>%
-  group_by(iucn_sid, quad) %>%
-  summarize(occurcells = sum(occurcells, na.rm = TRUE), occ = sum(occ, na.rm = TRUE)) %>%
-  ungroup()
-
-btm_dec_occurcells <- spp_occ_counts$occurcells %>% quantile(c(.1, .25))
-btm_dec_occ <- spp_occ_counts$occ %>% quantile(c(.1, .25))
-
-quad_count_means <- spp_occ_counts %>%
-  mutate(quad = ifelse(quad == 'q4', 'q4', 'q123')) %>%
-  group_by(quad) %>%
-  summarize(mean_occurcells = mean(occurcells),
-            mean_occ        = mean(occ),
-            med_occurcells  = median(occurcells),
-            med_occ         = median(occ),
-            dp_am_10 = sum(occurcells <= btm_dec_occurcells[1])/n(),
-            dp_am_25 = sum(occurcells <= btm_dec_occurcells[2])/n(),
-            dp_gb_10 = sum(occ <= btm_dec_occ[1])/n(),
-            dp_gb_25 = sum(occ <= btm_dec_occ[2])/n())
-
-write_csv(quad_count_means, file.path(dir_data, 'quads_data_poor.csv'))
-
-```

### 2016-12-27 data_explore/data_explore_fxns.R

@@ -3,139 +3,25 @@
 library(ggplot2)
 
 ### generic theme for all plots
-ggtheme_basic <- theme(axis.ticks = element_blank(),
-        text = element_text(family = 'Helvetica', color = 'gray30', size = 9),
-        plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
-        panel.background = element_blank(),
-        legend.position = 'right')
+ggtheme_plot <- theme(axis.ticks = element_blank(),
+                      text = element_text(family = 'Helvetica', color = 'gray30', size = 9),
+                      plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
+                      panel.background = element_blank(),
+                      legend.position = 'right',
+                      panel.border     = element_blank(),
+                      panel.grid.minor = element_blank(), 
+                      # panel.grid.major = element_line(colour = 'grey90', size = .25),
+                      panel.grid.major = element_blank(),
+                      legend.key = element_rect(colour = NA, fill = NA),
+                      axis.line = element_blank()) # element_line(colour = "grey30", size = .5))
 
-ggtheme_plot <- ggtheme_basic + 
-  theme(panel.border     = element_blank(),
-        panel.grid.minor = element_blank(), 
-        # panel.grid.major = element_line(colour = 'grey90', size = .25),
-        panel.grid.major = element_blank(),
-        legend.key = element_rect(colour = NA, fill = NA),
-        axis.line = element_line(colour = "grey30", size = .5))
 
-  
 ### theme for species range maps - based on generic plot theme
 ggtheme_map <- ggtheme_basic +
   theme(axis.text = element_blank(),
         axis.title = element_blank())
 
 
-#################################################################=
-### Function to create a species list from scratch... use when new data is available
-create_spp_list <- function(spp_list_base, data_file, am_spp_cells, 
-                            iucn_spp_cells, loiczid_raster) {
-  spp_list_build <- spp_list_base %>%
-    mutate(area_total = NA,  area_am = NA,  area_iucn = NA,
[…]

### 2016-12-27 clip_fao/fao_clip_pac.Rmd
pushing recent updates to analysis and documentation based on reviewer comments
added environmental space analysis to fao_clip_pac.Rmd; now to put it into the manuscript.

@@
 write_csv(clipped_spp, file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))
+
+### how many of these are expert-reviewed?
+clipped_spp_rev <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv')) %>%
+  filter(reviewed == '1')
 
@@
 
-# am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
-#                                    'v2016/int/am_cells_spp_prob0.csv')) 
+# am_spp_cells <- read_csv(file.path(dir_anx, '_raw_data/aquamaps',
+#                                    'd2015/csv/hcaf_sp_native_trunc.csv'))
 # 
@@
 # 
-# 
 # spp_area_df <- am_spp_cells2 %>%
+#   rename(am_sid = speciesid) %>%
 #   group_by(am_sid) %>%
-#   summarize(cellarea = sum(cellarea),
-#             oceanarea = sum(oceanarea))
+#   summarize(env_area  = sum(oceanarea * probability),
+#             geom_area = sum(oceanarea))
 # 
@@
 #   inner_join(am_spp_list, by = 'am_sid') %>%
-#   mutate(area_per_cells = oceanarea/occurcells)
+#   mutate(geom_area_per_cells = geom_area/occurcells,
+#          env_area_per_cells  = env_area/occurcells)
 # 
@@
 # spp_area_df <- spp_area_df %>%
-#   mutate(log_area_cells = log(area_per_cells + 1),
-#          clipped = am_sid %in% clipped_spp$am_sid)
+#   mutate(clipped = am_sid %in% clipped_spp$am_sid)
 
@@
 
-### compare ocean areas
-summary(spp_area_df$oceanarea)
+### compare geometric space areas
+summary(spp_area_df$geom_area)
    #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
    #   1806    946600   2921000  10210000   7875000 501300000 
-summary(spp_area_clipped$oceanarea)
+summary(spp_area_clipped$geom_area)
    #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
@@
 
-### compare occurcells
+### compare environmental space areas
+summary(spp_area_df$env_area)
+   # Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
+   #   90    592200   1843000   6772000   5506000 402200000 
+summary(spp_area_clipped$env_area)
+   #  Min.  1st Qu.    Median      Mean   3rd Qu.      Max. 
+   # 75930  2994000   4909000   5989000   7156000  89480000
+
+### compare occurcells by quadrant
 summary(spp_area_df$occurcells)
@@
 
-### compare area_per_cells
-summary(spp_area_df %>% filter(occurcells != 0) %>% .$area_per_cells)
+### compare geometric area_per_cells
+summary(spp_area_df %>% filter(occurcells != 0) %>% .$geom_area_per_cells)
     # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
     #  190    50540   158000   485300   439100 44700000 
-summary(spp_area_clipped$area_per_cells)
+summary(spp_area_clipped$geom_area_per_cells)
     # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
     # 6578   202900   452400   827200   923300 26090000 
+### compare environmental area_per_cells
+summary(spp_area_df %>% filter(occurcells != 0) %>% .$env_area_per_cells)
+    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
+    #   30    32270   101600   303700   287400 30100000 
+summary(spp_area_clipped$env_area_per_cells)
+    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
+    # 3352   140000   311100   511300   619800 12480000
+
+### sample species
+### Rajella fyllae: Fis-131802
+spp_area_df %>% filter(am_sid == 'Fis-131802') %>% select(geom_area, env_area, occurcells, geom_area_per_cells, env_area_per_cells)
+    #   geom_area env_area occurcells geom_area_per_cells env_area_per_cells
+    #       <dbl>    <dbl>      <int>               <dbl>              <dbl>
+    # 1   5425916  3386906        116            46775.14           29197.47
+### Ophiotrix plana: W-Oph-245456
+spp_area_df %>% filter(am_sid == 'W-Oph-245456') %>% select(geom_area, env_area, occurcells, geom_area_per_cells, env_area_per_cells)
+    #   geom_area env_area occurcells geom_area_per_cells env_area_per_cells
+    #       <dbl>    <dbl>      <int>               <dbl>              <dbl>
+    # 1   5426931  3053926          4             1356733           763481.5
 

### 2016-12-29 ms_0_abst.Rmd
ditching large summary files; pushing edits back to Mazu to try knitting there (pandoc errors on laptop)

@@
 Comparing distributions for the small subset of species with maps in both datasets, we highlight several key examples in which introduced errors drive differences in predicted species ranges.
-In particular, we find that IUCN maps greatly overpredict coral presence into unsuitably deep waters, and we demonstrate that AquaMaps methodology can produce odd discontinuities at the extremes of a species range.
+In particular, we find that IUCN maps greatly overpredict coral presence into unsuitably deep waters, and we show that AquaMaps computer-generated maps can produce odd discontinuities at the extremes of a species range.
 We illustrate the scientific and management implications of these tradeoffs by repeating a global analysis of gaps in coverage of marine protected areas, and find significantly different results depending on how the two datasets are used.

### 2016-12-29 ms_1_intro.Rmd
ditching large summary files; pushing edits back to Mazu to try knitting there (pandoc errors on laptop)

@@
 
-While the two datasets ostensibly describe the same information, i.e., where can a particular species be found, they communicate fundamentally distinct concepts.
-The IUCN range map for a given species describes the "extent of occurrence" for that species, a contiguous region that encloses all known occurrences and connecting regions, with the explicit caveat that this "does not mean that [the species] is distributed equally within that polygon or occurs everywhere within that polygon" [@iucn_iucn_2015].
-The AquaMaps map for a species, on the other hand, excludes unsuitable habitats and thus seems to better be defined as "area of occupancy": the subset of the extent of occurrence in which the species is actually likely to be found [@kaschner_aquamaps:_2015].
+While the two datasets ostensibly describe the same information, i.e., where can a particular species be found, they communicate the information in very different ways.
+The IUCN range map for a given species describes the "extent of occurrence" for that species, an undifferentiated contiguous region that encloses all known occurrences and connecting regions, with the explicit caveat that this "does not mean that [the species] is distributed equally within that polygon or occurs everywhere within that polygon" [@iucn_iucn_2015].
+The AquaMaps map for a species, on the other hand, provides a more nuanced understanding of species distribution by communicating the “probability of occurrence” of a species throughout its predicted range [1].
+By considering any non-zero probability of occurrence, we can effectively consider AquaMaps as an extent of occurrence, though bounded by modeled environmental conditions rather than by expert opinion [@kaschner_aquamaps:_2015].
 
@@
 
-The fundamental differences between the datasets suggest that the choice of one over the other should be carefully matched to the purpose for which it is to be used [@gaston_sizes_2009], and yet these datasets have been used in hundreds of studies and applications for a wide range of purposes, typically without directly addressing the limitations of the chosen dataset or comparing outcomes between the two.
+The fundamental differences between the datasets suggest that the choice of one over the other (and for AquaMaps, consideration of a presence threshold) should be carefully matched to the purpose for which it is to be used [@gaston_sizes_2009], and yet these datasets have been used in many studies and applications for a wide range of purposes, typically without directly addressing the limitations of the chosen dataset or comparing outcomes between the two.
 Example uses of the data include assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].

### 2016-12-29 ms_2_methods.Rmd
ditching large summary files; pushing edits back to Mazu to try knitting there (pandoc errors on laptop)

@@
 For the species included in both datasets, we examine how well the maps align in both spatial distribution and overall area.
-Based upon the definitions of extent of occurrence and area of occupancy, we expect that for a given species, the AquaMaps predicted distribution will fall within, and describe a smaller range than, the IUCN predicted distribution.
+If each dataset is communicating its own prediction of extent of occurrence, we expect that for a given species, the two predicted distributions will largely overlap, with similar total range sizes.
 Where these expectations seem to fail, we explore methodological issues that can introduce errors.
@@
 To assess the effectiveness of MPAs in protecting biodiversity, Klein et al.
-[-@klein_shortfalls_2015] compared the coverage of the global MPA network presented by the World Database on Protected Areas (WDPA) [@iucn_world_2014] to the species ranges described in the 2014 AquaMaps dataset [@kaschner_aquamaps:_2015].
+[-@klein_shortfalls_2015] compared the coverage of the global MPA network presented by the World Database on Protected Areas (WDPA) [@iucn_world_2014] to the species ranges described in the AquaMaps dataset, version 08/2013 [@kaschner_aquamaps:_2013].
 For the primary analysis, the researchers defined species presence as 50% or greater probability of occurrence.
@@
 The WDPA polygons and marine polygons were rasterized to 0.01° and then aggregated to AquaMaps native 0.5° cells, to calculate proportion of marine protected area within each cell.
-After verifying our method using the 2014 AquaMaps dataset, we updated the analysis to use the 2015 AquaMaps dataset, at a presence threshold of 50% (to compare to Klein et al.
+After verifying our method using the 08/2013 AquaMaps dataset, we updated the analysis to use the 2015 AquaMaps dataset [@kaschner_aquamaps:_2015], at a presence threshold of 50% (to compare to Klein et al.
 directly) and 0% (to better compare with IUCN spatial data).

### 2016-12-29 ms_3_res_disc_concl.Rmd
ditching large summary files; pushing edits back to Mazu to try knitting there (pandoc errors on laptop)

@@
 In comparing the IUCN and AquaMaps datasets, it is again important to emphasize that the two differ in both methodology and intent.
-As area of occupancy is generally defined as a subset of a species' extent of occurrence, AquaMaps ranges (effectively area of occupancy) should generally fall within the bounds of IUCN ranges (effectively extent of occurrence), capturing a smaller total area.
+For any given species, the IUCN range map and AquaMaps distribution (including all non-zero probability of occurrence) both effectively represent a prediction of extent of occurrence; therefore, we should expect the two maps to show significant overlap in predicted range, and to capture a similar total area.
 However, AquaMaps range maps are created independently from IUCN data and therefore exceptions are certain to arise.
@@
 The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where one range is notably larger than the other.
-For `r round(n_q2_iucn_larger/n_q2 * 100, 1)`% of the species in this quadrant, the IUCN range is larger than the AquaMaps range, as expected when comparing an extent of occurrence to an area of occupancy map.
-However, the low area ratio suggests a high rate of commission and/or omission errors by one or both datasets.
+For `r round(n_q2_iucn_larger/n_q2 * 100, 1)`% of the species in this quadrant, the IUCN range is larger than the AquaMaps range, which suggests a systematic introduction of commission errors by IUCN and/or omission errors by AquaMaps.
 Below we explore one underlying source of commission errors.
@@
 Consequently, area ratios are high, though the poor distribution alignment indicates that one or both datasets are introducing significant errors.
-In this quadrant, the IUCN range is the larger for only 56.5% of species, defying our expectations based on definitions of area of occupancy and extent of occurrence.
+In this quadrant, the IUCN range is the larger for only 56.5% of species, which does not seem to imply a systematic introduction of errors.
 
@@
 The lower left quadrant includes species (n = `r n_q4`) where alignment is poor in both dimensions.
-In this quadrant, the IUCN range is larger for only 24.6% of species; this again defies our expectations for extent of occurrence and area of occupancy range predictions.
+In this quadrant, the IUCN range is larger for only 24.6% of species; suggesting a systematic introduction of commission errors by AquaMaps and/or omission errors by IUCN.
 Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging occurrences from the Ocean Biogeographic Information System (OBIS) [@obis_data_2016] and the Global Biodiversity Information Facility (GBIF) [@gbif_global_2010]) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.
@@
 This adjustment was not necessary for AquaMaps data because models explicitly include ocean depth preference as a parameter.
-While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." [@iucn_iucn_2015, from website:  http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources]), it is not presented as a requirement, so we cannot take its inclusion for granted.
+While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." [REF IS THIS IN THE TRAINING MATERIALS?
+@iucn_iucn_2015, from website:  http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources]), it is not presented as a requirement, so we cannot take its inclusion for granted.
+Additionally, IUCN Red List mapping standards formerly required, and still allow, a 50 km buffer around the coastline for coastal species [iucn_iucn_2015-1]; such a buffer directly conflicts with habitat limitations (such as depth) and distorts our understanding of species distribution.
 
@@
 Membership in the "well-aligned" quadrant jumped from 22.4% to 76.2%, with a corresponding decrease in all other quadrants.
-By excluding the unsuitable areas from IUCN's extent of occurrence, we eliminate preventable commission errors and more closely approximate the area of occupancy described by AquaMaps.
+By excluding the unsuitable areas from IUCN's predicted range, we eliminate preventable commission errors and more closely approximate the range described by AquaMaps.
 See Fig S4 to examine the shifts of individual species among the quadrants.
@@
 The true distribution of each of these corals remains imperfectly known.
-Certainly some commission errors result from extent of occurrence boundaries including areas where species do not actually exist (as recognized by the definition of this method), and others may be due to experts taking a precautionary (i.e., generous) approach to likely occurrence.
+Certainly some commission errors result from from IUCN Red List mapping standards including coastline buffers and “minimum convex polygons”, and others may be due to experts taking a precautionary (i.e., generous) approach to likely occurrence.
 Yet a simple and sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported distributions for most purposes.
-This change applies just as readily to the IUCN coral maps that are not included in the paired map analysis, and likely to other reef-associated flora and fauna.
-While species depth preferences are an easy and consistent means of constraining range predictions, other conditions such as salinity and temperature could be cautiously used to refine the results of expert opinions.
+This change applies just as readily to the IUCN coral maps that are not included in the paired map analysis, and likely to other coastal and reef-associated flora and fauna.
+While species depth preferences are an easy and consistent means of constraining range predictions, other conditions such as salinity and temperature could be cautiously used to refine the results of expert opinions, much as AquaMaps models use such conditions to predict suitable habitat.
 
@@
 
-The ratio of the total predicted range for a species to the number of "occurcells" used to generate the map provides a measure of the degree to which AquaMaps extrapolates range area from limited data.
+The ratio of the total predicted range for a species to the number of "occurcells" used to generate the map provides a measure of the degree to which AquaMaps extrapolates geographic range area from limited data.
 For example, AquaMaps predicts a total range of 5.4 million km^2^ for both the round ray _Rajella fyllae_ and the brittle star _Ophiothrix plana_; but the map for _R.
-fyllae_ is generated using 116 "occurcells" (for a data extrapolation rate of 46,800 km^2^ per cell) while the map for _O.
+fyllae_ is generated using 116 "occurcells" (for a “geometric space” extrapolation rate of 46,800 km^2^ per cell) while the map for _O.
 plana_ is generated using only four (for an extrapolation rate of 1,360,000 km^2^ per cell).
+To estimate the rate at which AquaMaps extrapolates into “environmental space”, we weighted each cell’s geographic area by its environmental suitability (i.e., probability of occurrence), and see a similar pattern: R.
+fyllae’s “environmental space” extrapolation rate is 29,000 km^2^ of suitability-weighted area per occurcell, compared to 763,000 km^2^ of suitability-weighted area per cell for O.
+plana.
 
-By this measure, the 3,208 species range maps included in Fig 4 tend to extrapolate farther based on limited data: a mean predicted range of 827,200 km^2^ per "occurcell" compared to a mean predicted range of 485,300 km^2^ per "occurcell" for the overall AquaMaps dataset.
+By these measures, the 3,208 species range maps included in Fig 4 tend to extrapolate farther based on limited data: into geographic space, a mean predicted range of 827,200 km^2^ per "occurcell" compared to a mean predicted range of 485,300 km^2^ per "occurcell" for the overall AquaMaps dataset; into environmental space, the mean rate is 511,300 km^2^ per cell compared to 303,700 km^2^ per cell.
 This suggests that the FAO boundaries may not be sufficient to adequately constrain computer-generated ranges.
@@
 [-@klein_shortfalls_2015] compare the global distribution of species to the global distribution of marine protected areas to assess how well current MPAs overlap with species ranges and identify which species fall through gaps in protection.
-The study relied on the 2014 version of the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
+The study relied on the 08/2013 version of the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.
 They found that the global MPA network leaves 90.5% of marine species with less than 5% of their overall range represented within MPAs, and 1.4% of species have no protection at all (i.e., "gap" species).
+But what if the researchers had chosen to use IUCN data for their analysis rather than AquaMaps?
 
@@
 (2015).
-Scenario 1 replicates the original results, measuring protected range of species in AquaMaps 2014 dataset, with a 50% presence threshold, against the 2014 World Database of Protected Areas, filtered for IUCN categories I-IV that overlap marine areas.
-Scenario 2 updates the results using AquaMaps 2015, showing very small changes despite the inclusion of an additional 5,545 species.
-Scenario 3, using 2015 AquaMaps data, drops the presence threshold to zero, showing an expected decrease in gap species, but also a decrease in species with 5% or greater protected range.
+Scenario 1 replicates the original results, measuring protected range of species in AquaMaps version 08/2013 dataset, with a 50% presence threshold, against the 2014 World Database of Protected Areas, filtered for IUCN categories I-IV that overlap marine areas.
+Scenario 2 updates the results using AquaMaps version 08/2015, showing very small changes despite the inclusion of an additional 5,545 species.
+Scenario 3, still using 2015 AquaMaps data, drops the presence threshold to zero, showing an expected decrease in gap species, but also a decrease in species with 5% or greater protected range.
 Scenario 4 examines species MPA coverage using only the IUCN dataset.
+
+In performing this analysis, our intent is not to call into question the assumptions, methodology, or results of the original MPA gap analysis.
+Rather we intend to illustrate how differences between these two datasets could significantly influence conservation research outcomes (e.g.
+an apparent fivefold increase in gap species) and resulting conclusions about conservation efforts (e.g.
+the effectiveness of MPA policy).
 
@@
 While it may be unrealistic to "fix" one data set to match the other, we may be able to reduce the impact of the tradeoffs inherent in each.
-Trimming unsuitable habitat from the IUCN's extent of occurrence maps, for example by explicitly clipping them to appropriate depths for corals and reef-associated species, would result in maps that more closely approximate areas of occupancy.
+Trimming unsuitable habitat from the IUCN's extent of occurrence maps, for example by explicitly clipping them to appropriate depths for corals and reef-associated species, would reduce commission errors introduced by adherence to IUCN mapping standards, particularly the legacy 50 km coastline buffer.
 Conversely, including all AquaMaps cells with a non-zero "probability of occurrence" (rather than using a probability threshold to determine presence, e.g.
-greater than 40% for the Ocean Health Index [@halpern_index_2012; @halpern_patterns_2015] or 50% for the MPA gap analysis [@klein_shortfalls_2015]) would allow for the most generous inclusion of species range, resulting in maps that more closely, though still imperfectly, approach the IUCN's extents of occurrence.
+greater than 40% for the Ocean Health Index [@halpern_index_2012; @halpern_patterns_2015] or 50% for the MPA gap analysis [@klein_shortfalls_2015]) would allow for the most generous inclusion of species range, resulting in maps that more closely align with the intent of the IUCN's extents of occurrence.
 
@@
 Conclusions drawn from each of these datasets could paint dramatically different pictures of global marine biodiversity or the effectiveness of conservation management decisions.
-By highlighting the distinctions between these two important and fundamentally different marine species range datasets, we improve our ability to inform strategic and effective conservation policy that supports a resilient ocean ecosystem.
+By highlighting the distinctions between these two important and fundamentally different marine species range datasets, we hope to encourage stronger collaboration between IUCN experts and AquaMaps to benefit both datasets.
+A more transparent and reproducible approach for describing species ranges, incorporating the best aspects of both distribution modeling and expert opinion, would greatly improve  our ability to inform strategic and effective conservation policy that supports a resilient ocean ecosystem.

### 2016-12-30 data_explore/data_explore.Rmd
updated the figures and checked the new reference numbers...

@@
 taxa_df <- read_csv(file.path(dir_data, 'taxa_df.csv'))
+taxa_omitted <- read_csv(file.path(dir_data, 'taxa_omitted.csv'))
+
+n_spp_omitted <- taxa_omitted$n_am_class_all %>% sum()
 
@@
   summarize(class = 'Total',
-            AquaMaps_total = sum(AquaMaps) + sum(Paired),
+            AquaMaps_total = sum(AquaMaps)  + n_spp_omitted + sum(Paired),
             IUCN_total = sum(IUCN) + sum(Paired),
-            n_total  = sum(n_total),
+            n_total  = sum(n_total) + n_spp_omitted,
             IUCN_end = 100 * IUCN_total/n_total,
@@
               arrange(desc(n_total))) %>%
-  bind_rows(x <- taxa_omitted %>%
+  bind_rows(taxa_omitted %>%
               group_by(class) %>%

### 2016-12-30 manuscript_main.Rmd
updated the figures and checked the new reference numbers...

--- 2016-09-23 manuscript-main.Rmd
@@
 <!--
-_Keywords:_  Marine spatial planning, species distribution model, extent of occurrence, spatial ecology
+_Keywords:_  Marine spatial planning, species distribution model, extent of occurrence, area of occupancy, spatial ecology
 -->
@@
 
------
-
 ``` {r abstract, child = 'ms_0_abst.Rmd'}
@@
 -->
-
------
 
@@
 ```
-
-<!--
-| Figure  |  w x h (px) | w x h (cm)  | char # |
-| :------ |     ------: |     ------: |------: |
-| 1 A - C | 2102 x 1536 | 17.8 x 13.0 |   4680 |
-| caption |             |             |    367 | 
-| 2 A - B | 2102 x 1940 | 17.8 x 16.5 |   5940 |
-| caption |             |             |    934 | 
-|       3 | 1027 x  531 |  8.7 x  4.5 |    810 |
-| caption |             |             |    744 | 
-|       4 | 1346 x  590 | 11.4 x  5.0 |   1800 |
-| caption |             |             |    870 | 
-|  total  |             |             |  16145 |
--->

### 2016-12-30 ms_3_res_disc_concl.Rmd
updated the figures and checked the new reference numbers...

@@
 This adjustment was not necessary for AquaMaps data because models explicitly include ocean depth preference as a parameter.
-While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." [REF IS THIS IN THE TRAINING MATERIALS?
-@iucn_iucn_2015, from website:  http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources]), it is not presented as a requirement, so we cannot take its inclusion for granted.
-Additionally, IUCN Red List mapping standards formerly required, and still allow, a 50 km buffer around the coastline for coastal species [iucn_iucn_2015-1]; such a buffer directly conflicts with habitat limitations (such as depth) and distorts our understanding of species distribution.
+While depth is recommended by the IUCN as a criterion for providers of range maps ("The limits of distribution can be determined by using known occurrences of the species, along with the knowledge of habitat preferences, remaining suitable habitat, elevation limited, and other expert knowledge of the species and its range." REF IS THIS IN THE TRAINING MATERIALS?
+[@iucn_iucn_2015]), it is not presented as a requirement, so we cannot take its inclusion for granted.
+Additionally, IUCN Red List mapping standards formerly required, and still allow, a 50 km buffer around the coastline for coastal species [@iucn_red_2015]; such a buffer directly conflicts with habitat limitations (such as depth) and distorts our understanding of species distribution.
 

### 2017-01-04 data_explore/data_explore.Rmd
updates to manuscript

@@
 
-quad_names <- data.frame('quad' = c('q1', 'q2', 'q3', 'q4'),
-                         'quad_name' = factor(c('well-aligned', 'dist-aligned', 
-                                                'area-aligned', 'poorly aligned'),
+quad_names <- data.frame('quad' = c('q4', 'q3', 'q2', 'q1'),
+                         'quad_name' = factor(c('poorly aligned', 
+                                                'area-aligned', 
+                                                'dist-aligned',
+                                                'well-aligned'),
                                              ordered = TRUE))
@@
   left_join(quad_names, by = 'quad') %>% 
-  mutate(quad = factor(quad, levels = c('q4', 'q3', 'q2', 'q1'))) %>%
+  mutate(quad = factor(quad, levels = c('q4', 'q3', 'q2', 'q1'), ordered = TRUE)) %>%
   transform(spp_group_text = reorder(spp_group_text, pct_q1))
@@
 barchart_spp_gp_quads <- ggplot(spp_gp_quadrants, 
-                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
+                                aes(x = spp_group_text, 
+                                    y = pct_quad,
+                                    fill = quad, 
+                                    weight = pct_quad)) +
   ggtheme_plot + 
-  # theme(panel.grid.major.y = element_blank()) +
-  geom_bar(stat = 'count', alpha = 1) +
+  geom_bar(stat = 'identity', alpha = 1) +
   scale_fill_manual(values = c('q1' = '#4dac26',
@@
                                'q4' = '#d01c8b'),
-                    labels = quad_names$quad_name) +
+                    labels = quad_names$quad_name,
+                    guide = guide_legend(reverse = TRUE)) +
   scale_y_continuous(expand = c(0, 0), 
@@
   ### add text
-  geom_text(# data = spp_gp_quadrants %>%
-              # select(spp_group_text, n_spp) %>%
-              # distinct(),
-            aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
+  geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
             size = 2.5, 
@@
   # scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = 1)) +
-  geom_point(color = 'black', aes(alpha = fade), size = .1, show.legend = FALSE) +
+  geom_point(data = spp_list, aes(x = area_ratio, y = sm_perc), 
+             color = 'black', alpha = .15, size = .1, show.legend = FALSE) +
+  geom_point(color = 'black', aes(alpha = fade), size = .25, show.legend = FALSE) +
   scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = 1))
@@
 
-![](../figures/scatter_quadmap.png)
+![](../figures/scatter_quadmap_expert.png)
 
@@
 dist_align_mean <- mean(spp_list$sm_perc, na.rm = TRUE)
+
+spp_list_experts <- spp_list %>%
+  filter(reviewed == 1)
 

### 2017-01-04 ms_3_res_disc_concl.Rmd
updates to manuscript

@@
 
+It is unsurprising that species with expert-reviewed AquaMaps fare far better in this paired-map analysis; for expert-reviewed maps, the mean distribution alignment improved to 75.7% (compared to 63.1% for the full set of paired maps), while the mean area alignment improved 60.8% (compared to 54.7%).
+See Fig S4 for a version of Fig 2A and 2B focused on expert-reviewed species.
+
 ## Coral depth exploration
@@
 By excluding the unsuitable areas from IUCN's predicted range, we eliminate preventable commission errors and more closely approximate the range described by AquaMaps.
-See Fig S4 to examine the shifts of individual species among the quadrants.
+See Fig S5 to examine the shifts of individual species among the quadrants.
 

### 2017-01-04 ms_6_supplemental.Rmd
updates to manuscript

@@
 > __Fig S1.
-Examples of AquaMaps and IUCN raw species range data.__ (A, B) AquaMaps distribution of Thunnus alalunga (Albacore Tuna) [23] showing (A) global range with varying probabilities of occurrence assigned to 0.5° grid cells and (B) presence within 0.5° grid cells.
+Examples of AquaMaps and IUCN raw species range data.__ (A, B) AquaMaps distribution of Thunnus alalunga (Albacore Tuna) [25] showing (A) global range with varying probabilities of occurrence assigned to 0.5° grid cells and (B) presence within 0.5° grid cells.
 (C, D) IUCN distribution of T.
-alalunga [24] represented as (C) extent of occurrence polygons and (D) presence within 0.5° grid cells.
+alalunga [26] represented as (C) extent of occurrence polygons and (D) presence within 0.5° grid cells.
 
@@
 Rasterizing shapefiles provided by IUCN.__ A portion of the T.
-alalunga range map [24] is used to exemplify the rasterization process.
+alalunga range map [26] is used to exemplify the rasterization process.
 To enable direct comparison of IUCN species ranges to AquaMaps species ranges, the raw IUCN polygon (A) is overlaid with a 0.5° degree grid matching the AquaMaps grid (B).
@@
 FAO Major Fishing Area boundaries [11] are outlined in light grey.
-(A) Distribution-aligned: Conus episcopatus, the dignified cone snail [25, 26].
+(A) Distribution-aligned: Conus episcopatus, the dignified cone snail [27, 28].
 Distributions show excellent overlap in the western Pacific, though IUCN range extends well beyond the bounds of the AquaMaps range.
-(B) Well-aligned: Kajikia albida, the Atlantic white marlin [27, 28].
+(B) Well-aligned: Kajikia albida, the Atlantic white marlin [29, 30].
 Distributions from each data set show nearly complete overlap, and very similar range size.
-(C) Poorly aligned: Acanthurus nigroris, the blue-lined surgeonfish [29, 30].
+(C) Poorly aligned: Acanthurus nigroris, the blue-lined surgeonfish [31, 32].
 IUCN predicts species distribution only near the Hawaiian islands; AquaMaps predicts extensive distribution throughout the central and western Pacific Ocean.
 The datasets align in neither distribution nor range size.
-(D) Area-aligned: Conus magnificus, the magnificent cone snail [31, 32].
+(D) Area-aligned: Conus magnificus, the magnificent cone snail [33, 34].
 Distributions overlap in the southern Pacific, but align poorly elsewhere.
@@
 > __Fig S4.
+Improvement in alignment due to expert review of AquaMaps.__ (A) Modification of Fig 2A to highlight species with expert-reviewed AquaMaps shows that the mean distribution alignment and mean area ratio both improve.
+(B) Including only expert-reviewed species in each quadrant shows increased membership in the well-aligned and distribution-aligned quadrants relative to Fig 2B.
+
+![](figures/S5_fig.png)
+
+> __Fig S5.
 Shift in alignment of paired-map coral species due to clipping IUCN ranges to areas shallower than 200 m.__ The grey lines represent the change in apparent alignment for a single species.
@@
 
-23.
+25.
 www.aquamaps.org, version of Aug.
@@
 
-24. International Union for Conservation of Nature (IUCN) 2011. Thunnus alalunga. The IUCN Red List of Threatened Species. Version 2016-2
+26. International Union for Conservation of Nature (IUCN) 2011. Thunnus alalunga. The IUCN Red List of Threatened Species. Version 2016-2
 
-25.
+27.
 www.aquamaps.org, version of Aug.
@@
 
-26. Conch Books, Hackenheim, Germany 2013. Conus episcopatus. The IUCN Red List of Threatened Species. Version 2016-2.
+28. Conch Books, Hackenheim, Germany 2013. Conus episcopatus. The IUCN Red List of Threatened Species. Version 2016-2.
 
-27.
+29.
 www.aquamaps.org, version of Aug.
@@
 
-28. International Union for Conservation of Nature (IUCN) 2011. Kajikia albida. The IUCN Red List of Threatened Species. Version 2016-2.
+30. International Union for Conservation of Nature (IUCN) 2011. Kajikia albida. The IUCN Red List of Threatened Species. Version 2016-2.
 
-29.
+31.
 www.aquamaps.org, version of Aug.
@@
 
-30. International Union for Conservation of Nature (IUCN) 2012. Acanthurus nigroris. The IUCN Red List of Threatened Species. Version 2016-2.
+32. International Union for Conservation of Nature (IUCN) 2012. Acanthurus nigroris. The IUCN Red List of Threatened Species. Version 2016-2.
 
-31.
+33.
 www.aquamaps.org, version of Aug.
@@
 
-32. Conch Books, Hackenheim, Germany 2013. Conus magnificus. The IUCN Red List of Threatened Species. Version 2016-2.
+34. Conch Books, Hackenheim, Germany 2013. Conus magnificus. The IUCN Red List of Threatened Species. Version 2016-2.

### 2017-03-08 data_explore/data_prep.Rmd
checking the name matching

@@
   mutate(iucn_sid = ifelse(iucn_wrong_map & !is.na(iucn_wrong_map), iucn_id, iucn_sid)) %>%
-  select(-iucn_match, -iucn_wrong_map, -iucn_id, -sciname_count) %>%
+  # select(-iucn_match, -iucn_wrong_map, -iucn_id, -sciname_count) %>%
   distinct()
 
-length(unique(x$iucn_sid)) ###  4013, incl NA
-length(unique(x$am_sid))   ### 22890, incl NA
+# length(unique(spp_all$iucn_sid)) ###  4013, incl NA
+# length(unique(spp_all$am_sid))   ### 22890, incl NA
+# qwer <- spp_all %>% filter(!is.na(iucn_sid) & !is.na(am_sid))
+# length(unique(qwer$iucn_sid)) ### 2320
+# length(unique(qwer$am_sid)) ### 2314
+# length(qwer %>% filter(iucn_match) %>% .$iucn_sid %>% unique()) ### 1852
+# 
+# x <- spp_iucn_maps %>%
+#   rename(iucn_sciname = sciname) %>%
+#   full_join(spp_am_maps %>% 
+#               rename(am_sciname = sciname), 
+#             by = c('iucn_sid' = 'iucn_id')) %>%
+#   select(iucn_sid, iucn_sciname, am_sid, am_sciname) %>%
+#   mutate(name_match = iucn_sciname == am_sciname) %>%
+#   distinct()
+# asdf <- x %>% filter(!is.na(iucn_sid) & !is.na(am_sid)) %>%
+#   filter(!is.na(iucn_sciname) & !is.na(am_sciname))
+# length(unique(asdf$iucn_sid))
+# length(unique(asdf$am_sid))
 

### 2017-03-08 data_explore/data_explore_fxns.R

@@ -17,8 +17,8 @@ ggtheme_plot <- theme(axis.ticks = element_blank(),
 
 
 ### theme for species range maps - based on generic plot theme
-ggtheme_map <- ggtheme_basic +
-  theme(axis.text = element_blank(),
+ggtheme_map <- ggtheme_plot +
+  theme_update(axis.text = element_blank(),
         axis.title = element_blank())
 
 

### 2017-03-12 clip_depth/coral_maps.Rmd
updated the matching algorithm (again) and reran analysis

@@
 ### load iucn spp_cells
-coral_files <- list.files(file.path(dir_anx, 'v2016', 'iucn_intersections'), 
+coral_files <- list.files(file.path(dir_anx, 'vAM_IUCN', 'iucn_intersections'), 
                           pattern = 'CORAL*.*',
@@
 length(unique(coral_cells_iucn$iucn_sid))
-### NOTE: only 473 coral species with IUCN maps
+### NOTE: 562 coral species with IUCN maps
+
 ```
@@
   spp_map_pairs <- read_csv(paired_maps_file) %>%
-    filter(!is.na(popn_category)) %>% ### filter out unassessed species
     filter(!is.na(sm_perc))           ### filter out NA maps
@@
 
-} else 
+} else {
   spp_depth <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))
+}
 
@@
 spp_map_pairs <- read_csv(data_file) %>%
-  filter(!is.na(popn_category)) %>%   ### filter out unassessed species
   filter(!is.na(sm_perc))             ### filter out NA maps
@@
                     aes(x = a_ratio, y = d_perc, group = iucn_sid)) +
-  ggtheme_plot + 
+  theme(panel.grid.major.x = element_blank()) +
   ### color the quadrant backgrounds:
@@
 
-area_align_mean <- 54.53334
-dist_align_mean <- 62.98212
+spp_list <- read_csv(file.path(dir_git, 'data/spp_list_w_area_trimmed.csv'))
+area_align_mean <- mean(spp_list$area_ratio, na.rm = TRUE)
+dist_align_mean <- mean(spp_list$sm_perc, na.rm = TRUE)
 
@@
                                 aes(x = method, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
   theme(panel.grid.major.y = element_blank()) +
@@
 
-# ggsave(plot = barchart_coral_quads,
-#        filename = file.path(dir_fig, 'barchart_coral_quads.png'),
-#        width = 19, height = 4, units = 'cm', dpi = 600)
+ggsave(plot = barchart_coral_quads,
+       filename = file.path(dir_fig, 'barchart_coral_quads.png'),
+       width = 19, height = 4, units = 'cm', dpi = 600)
 

### 2017-03-12 data_explore/data_explore.Rmd
updated the matching algorithm (again) and reran analysis

@@
 message('Getting IUCN maps list')
-iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps.csv')
+iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps_verified.csv')
 spp_iucn_maps <- read_csv(iucn_maps_file)
@@
               unique(), 
-            by = 'am_sid')
+            by = 'am_sid') %>%
+  mutate(paired_map = am_sid %in% spp_map_pairs$am_sid)
+### nrow(spp_am %>% filter(paired_map)) ### 2330 spp
 
-iucn_maps_class_all <- spp_iucn_maps %>% 
-  select(-iucn_subpop, -spp_group) %>%
-  distinct() %>% ### first, get rid of dupes due to subpopulations
+iucn_maps_class <- spp_iucn_maps %>% 
+  select(-iucn_subpop, -name_verified, -spp_group) %>%
+  distinct() %>%
   mutate(class      = str_to_title(class),
-         am_match = sciname %in% spp_am$sciname)
-
-### next, identify spp scinames with matches in AM; if an iucn_sid has a match,
-### keep it; if there are multiple matches, keep the one with the fewest
-### instances of sciname (to keep more specific names...?)
-iucn_maps_class <- iucn_maps_class_all %>% 
-  group_by(sciname) %>%
-  mutate(sciname_count = n()) %>% ### for each sciname, figure out how many times it appears
-  group_by(iucn_sid) %>%
-  arrange(desc(am_match), sciname_count) %>% ### for each iucn_spp, put the AM matches on top; then...
-  filter(sciname == first(sciname)) %>% ### just keep the first instance of each iucn_spp
-  ungroup()
+         paired_map = iucn_sid %in% spp_map_pairs$iucn_sid)
+### nrow(iucn_maps_class %>% filter(paired_map)) ### 2330 spp
 
@@
   filter(!is.na(class)) %>%                 ### don't count the unknowns
-  filter(!sciname %in% spp_am$sciname) %>%  ### filter list to IUCN only
+  filter(!paired_map) %>%  ### filter list to IUCN only
   mutate(n_iucn_solo = n()) %>%             ### length of IUCN only list
@@
   filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
+  filter(!paired_map) %>%  ### filter list to AM only
   mutate(n_am_solo = n()) %>%       ### length of AM only list
@@
 taxa_both <- spp_am %>%
-  filter(sciname %in% spp_iucn_maps$sciname) %>%
+  filter(paired_map) %>%
   filter(!class %in% c('Not assigned', 'null')) %>%
@@
   bind_rows(taxa_all_df %>%
-              mutate(class = ' ',
+              mutate(class = '',
                      start = 0,
                      end   = 0,
-                     n_total = NA)) %>%
+                     n_total = 0)) %>%
   bind_rows(taxa_df1 %>%
@@
   mutate(class = factor(class, levels = unique(.$class), ordered = TRUE),
-         n_label = ifelse(!is.na(n_total), paste0('n = ', n_total), ''),
+         n_label = ifelse(n_total > 0, paste0('n = ', n_total), ''),
          dataset = factor(dataset, levels = c('AquaMaps', 'Paired', 'IUCN'), ordered = TRUE))
@@
 ### cut the labels for all but the first instance
-taxa_df2$n_label[duplicated(taxa_df2$class)] <- NA
+taxa_df2$n_label[duplicated(taxa_df2$class)] <- ''
 
@@
 bar_taxa <- ggplot(taxa_df2, aes(x = class, color = dataset)) +
-  ggtheme_plot + 
-  theme(panel.grid.major.x = element_blank()) +
+  theme(panel.grid.major.x = element_blank(),
+        axis.text.x  = element_text(angle = 75, hjust = 1),
+        axis.title.y = element_text(hjust = 0)) +
   geom_linerange(aes(ymin = start, ymax = end, size = dataset),
@@
             size = 2.4,
-            angle = 90, hjust = 0, color = 'grey30') +
-  theme(axis.text.x = element_text(angle = 75, hjust = 1),
-        axis.title.y = element_text(hjust = 0)) +
+            angle = 90,
+            hjust = 0,
+            color = 'grey30') +
   scale_y_continuous(expand = c(0, 0),
@@
 scatter_quadplot <- ggplot(spp_list_labs, aes(x = x, y = y)) +
-  ggtheme_plot + 
+  theme(panel.grid.major.x = element_blank()) +
         ### color the quadrant backgrounds:
@@
 
-# ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
-#        width = 19.05, height = 9, units = 'cm', dpi = 600)
+ggsave(file.path(dir_fig, 'scatter_quadmap.png'),
+       width = 19.05, height = 9, units = 'cm', dpi = 600)
 
@@
                                     weight = pct_quad)) +
-  ggtheme_plot + 
+  theme(panel.grid.major.x = element_blank()) +
   geom_bar(stat = 'identity', alpha = 1) +
@@
 scatter_quadplot_ex <- ggplot(spp_list_experts_labs, aes(x = x, y = y)) +
-  ggtheme_plot + 
+  theme(panel.grid.major.x = element_blank()) +
   ### color the quadrant backgrounds:
@@
                                 aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
-  ggtheme_plot + 
-  # theme(panel.grid.major.y = element_blank()) +
+  theme(panel.grid.major.x = element_blank()) +
   geom_bar(stat = 'count', alpha = 1) +

### 2017-03-12 data_explore/data_prep.Rmd
updated the matching algorithm (again) and reran analysis

--- 2016-12-27 data_explore/data_prep.Rmd
@@
 
+```{r verify_spp_names_in_shps, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+
+spp_iucn_maps <- read_csv(file.path(dir_data, 'spp_iucn_maps.csv'))
+
+# x <- spp_iucn_maps %>%
+#   filter(is.na(iucn_subpop) & !is.na(iucn_sid)) %>%
+#   filter(! spp_group %in% c('REPTILES', 'SEASNAKES', 'MANGROVES', 'SEAGRASSES')) %>%
+#   filter(sciname %in% .$sciname[duplicated(.$sciname)])
+### 37 instances of duped scinames with different IDs
+
+api_key <- scan(file.path(dir_anx, 'api_key.csv'), what = 'character')
+
+spp_id_url <- 'http://apiv3.iucnredlist.org/api/v3/species/id/%s?token=%s'
+
+sid_list <- spp_iucn_maps %>%
+  filter(!is.na(iucn_sid)) %>% 
+  .$iucn_sid %>%
+  unique() %>%
+  sort()
+
+ptm <- proc.time()
+spp_names_df <- mc_get_from_api(spp_id_url, sid_list, api_key, delay = 1, cores = 16) %>%
+  select(iucn_sid = taxonid, sciname2 = scientific_name)
+proc.time() - ptm
+
+spp_iucn_maps1 <- spp_iucn_maps %>%
+  left_join(spp_names_df, by = 'iucn_sid') %>% ### here, both the shapefile name and API name
+  select(-sciname) %>%
+  rename(sciname = sciname2) ### now just the API name by ID number
+
+write_csv(spp_iucn_maps1, file.path(dir_data, 'spp_iucn_maps_verified.csv'))
+
+```
+
 ```{r get_spp_all, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
 
-spp_iucn_maps <- read_csv(file.path(dir_data, 'spp_iucn_maps.csv'))
+spp_iucn_maps <- read_csv(file.path(dir_data, 'spp_iucn_maps_verified.csv'))
 
@@
 spp_all <- spp_iucn_maps %>%
-  full_join(spp_am_maps, by = 'sciname') %>%
+  filter(iucn_sid %in% spp_am_maps$iucn_id) %>%
+  left_join(spp_am_maps %>% 
+              rename(am_sciname = sciname) %>%
+              mutate(iucn_sid = iucn_id), ### keep the IUCN noted ID separate from AM-noted IUCN id
+            by = c('iucn_sid')) %>%
+  bind_rows(x <- spp_iucn_maps %>%
+              filter(!iucn_sid %in% spp_am_maps$iucn_id) %>%
+              full_join(spp_am_maps %>%
+                          mutate(am_sciname = sciname) %>%
+                          filter(!iucn_id %in% spp_iucn_maps$iucn_sid | is.na(iucn_id)),
+                        by = c('sciname'))) %>%
   distinct()
 
-### some confusion due to non-matching scinames and IDs...
-spp_all <- spp_all %>%
-  mutate(iucn_match = iucn_sid == iucn_id,
-         iucn_wrong_map = ifelse(!iucn_match & iucn_id %in% spp_all$iucn_sid, TRUE, FALSE)) %>%
-  mutate(iucn_sid = ifelse(iucn_wrong_map & !is.na(iucn_wrong_map), iucn_id, iucn_sid)) %>%
-  select(-iucn_match, -iucn_wrong_map, -iucn_id, -sciname_count) %>%
-  distinct()
+# x <- spp_all %>% 
+#   filter(is.na(iucn_subpop) & !is.na(iucn_sid)) %>%
+#   filter(iucn_sid %in% .$iucn_sid[duplicated(.$iucn_sid)])
+### just seasnakes/reptiles and mangroves/seagrasses
 
-length(unique(x$iucn_sid)) ###  4013, incl NA
-length(unique(x$am_sid))   ### 22890, incl NA
+# length(unique(spp_all$iucn_sid)) ###  4028, incl NA
+# length(unique(spp_all$am_sid))   ### 22890, incl NA
+# qwer <- spp_all %>% filter(!is.na(iucn_sid) & !is.na(am_sid))
+# length(unique(qwer$iucn_sid)) ### 2343
+# length(unique(qwer$am_sid)) ### 2337
+
+# 
+# x <- spp_iucn_maps %>%
+#   rename(iucn_sciname = sciname) %>%
+#   full_join(spp_am_maps %>%
+#               rename(am_sciname = sciname),
+#             by = c('iucn_sid' = 'iucn_id')) %>%
+#   select(iucn_sid, iucn_sciname, am_sid, am_sciname) %>%
+#   mutate(name_match = iucn_sciname == am_sciname) %>%
+#   distinct()
+# asdf <- x %>% filter(!is.na(iucn_sid) & !is.na(am_sid)) %>%
+#   filter(!is.na(iucn_sciname) & !is.na(am_sciname))
+# length(unique(asdf$iucn_sid)) # 1880
+# length(unique(asdf$am_sid)) # 1880
+### 1880 species matched if only using ID number...
 
@@
 # length(unique(spp_am_v_iucn$sciname)); length(unique(spp_am_v_iucn$am_sid)); length(unique(spp_am_v_iucn$iucn_sid))
-### NOTE: 2314 unique scientific names; but 2314 unique am_sid and 2320 unique iucn_sid.
+### NOTE: 2332 unique scientific names, unique am_sid and unique iucn_sid.
 ### Seasnakes are included in Reptiles (IUCN).  Two "mangroves" that are actually seagrasses.
@@
 # length(unique(spp_am_v_iucn$sciname)); length(unique(spp_am_v_iucn$am_sid)); length(unique(spp_am_v_iucn$iucn_sid))
+### still same scinames and am_sid
 
@@
 
-spp_am_v_iucn <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))
+# spp_am_v_iucn <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))
 
@@
 
-# iucn_spp_cells_files <- list.files(file.path(dir_anx, 'v2016/iucn_intersections'), full.names = TRUE)
+# iucn_spp_cells_files <- list.files(file.path(dir_anx, 'vAM_IUCN/iucn_intersections'), full.names = TRUE)
 # iucn_spp_cells_all <- lapply(iucn_spp_cells_files, FUN = function(x) {
 #   read_csv(x) %>%
-#     mutate(id_no = as.integer(id_no)) %>%
-#     rename(iucn_sid = id_no)
+#     mutate(iucn_sid = as.integer(iucn_sid))
 #   }) %>%
-#   bind_rows()
+#   bind_rows() %>%
+#   setNames(tolower(names(.)))
 # 
-# iucn_spp_cells <- iucn_spp_cells_all %>%
+# spp_iucn_maps1 <- read_csv(file.path(dir_data, 'spp_iucn_maps_verified.csv')) %>%
+#   select(sciname, iucn_sid_ver = iucn_sid) %>%
+#   distinct()
+
+### because the 2015 shapefiles for reptiles and mammals were modified
+### to attach subpop IDs rather than parent IDs to subpops, we will undo
+### this here by overwriting the IDs based on parent IDs.
+# iucn_spp_cells_verified <- iucn_spp_cells_all %>%
+#   left_join(spp_iucn_maps1, by = 'sciname') %>%
+#   mutate(iucn_sid = iucn_sid_ver)
+# 
+# iucn_spp_cells <- iucn_spp_cells_verified %>%
 #   filter(iucn_sid %in% spp_am_v_iucn$iucn_sid) %>%
-#   filter(presence != 5)
+#   filter(presence != 5) %>%
+#   select(-iucn_sid_ver)
 
-### length(unique(iucn_spp_cells$iucn_sid)) results in 2271 species found...
-###   some species with no cells reported.  
-### spp_missing <- spp_am_v_iucn %>% filter(!iucn_sid %in% iucn_spp_cells$iucn_sid)
-###   49 spp with no cells reported.  Shapefile errors?
+## length(unique(iucn_spp_cells$iucn_sid)) #  2330 species found...
+##   some species with no cells reported.
+## spp_missing <- spp_am_v_iucn %>% filter(!iucn_sid %in% iucn_spp_cells$iucn_sid)
+##   2 spp with no cells reported.  Shapefile errors?
 # write_csv(iucn_spp_cells, file.path(dir_anx, 'vAM_IUCN/iucn_spp_cells_2015.csv'))
-
-message('Loading AquaMaps species to cell table.  This might take a few minutes.')
-am_spp_cells_all <- fread(file.path(dir_data_remote, 'aquamaps/d2015/csv/hcaf_sp_native_trunc.csv'), 
-                         showProgress = TRUE)
-am_spp_cells <- am_spp_cells_all %>%
-  rename(am_sid = speciesid, prob = probability) %>%
-  filter(am_sid %in% spp_am_v_iucn$am_sid)
-### length(unique(am_spp_cells$am_sid)) results in 2314 species found
-write_csv(am_spp_cells, file.path(dir_anx, 'vAM_IUCN/am_spp_cells_2015.csv'))
+# 
+# message('Loading AquaMaps species to cell table.  This might take a few minutes.')
+# am_spp_cells_all <- fread(file.path(dir_data_remote, 'aquamaps/d2015/csv/hcaf_sp_native_trunc.csv'),
+#                          showProgress = TRUE)
+# am_spp_cells <- am_spp_cells_all %>%
+#   rename(am_sid = speciesid, prob = probability) %>%
+#   filter(am_sid %in% spp_am_v_iucn$am_sid)
+# ### length(unique(am_spp_cells$am_sid)) # results in 2332 species found
+# write_csv(am_spp_cells, file.path(dir_anx, 'vAM_IUCN/am_spp_cells_2015.csv'))
 
@@
 
-```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
+``` {r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
 ### Create files for species list with area info, both a complete list and a trimmed list
@@
 data_file <- file.path(dir_data, 'spp_list_w_area_all.csv')
-reload <- FALSE
+reload <- TRUE
 if (!file.exists(data_file) | reload) {
@@
   spp_list_base <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))
-  create_spp_list2(spp_list_base, data_file, am_spp_cells, iucn_spp_cells, loiczid_raster)
+  create_spp_list(spp_list_base, data_file, am_spp_cells, iucn_spp_cells, loiczid_raster)
 
@@
   ### will ditch the ones with no IUCN area - formerly caused by DD species
-  ### being filtered out.  Now, none suffer this consequence.
+  ### being filtered out.  Now, only 2 suffer this consequence.
 

### 2017-03-12 data_explore/data_explore_fxns.R

@@ -3,7 +3,8 @@
 library(ggplot2)
 
 ### generic theme for all plots
-ggtheme_plot <- theme(axis.ticks = element_blank(),
+theme_set(theme_bw())
+ggtheme_plot <- theme_update(axis.ticks = element_blank(),
                       text = element_text(family = 'Helvetica', color = 'gray30', size = 9),
                       plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
                       panel.background = element_blank(),
@@ -16,12 +17,6 @@ ggtheme_plot <- theme(axis.ticks = element_blank(),
                       axis.line = element_blank()) # element_line(colour = "grey30", size = .5))
 
 
-### theme for species range maps - based on generic plot theme
-ggtheme_map <- ggtheme_plot +
-  theme_update(axis.text = element_blank(),
-        axis.title = element_blank())
-
-
 
 #################################################################=
 ### Species Map Function ###
@@ -32,6 +27,8 @@ ggtheme_map <- ggtheme_plot +
 get_spp_map <- function(sid, spp_cells) { 
   data_type <- ifelse('am_sid' %in% names(spp_cells), 'am', 'iucn')
   
+  
+  
   if(data_type == 'am') {
     ### if AquaMaps, process it thusly:
     spp_map <- spp_cells %>%
@@ -41,6 +38,10 @@ get_spp_map <- function(sid, spp_cells) {
       message('no AquaMaps cells found')
   } else {
     ### if IUCN, process it thusly:
+    if(!'prop_area' %in% names(spp_cells)) {
+      spp_cells <- spp_cells %>%
+        mutate(prop_area = 1)
+    }
     spp_map <- spp_cells %>%
       filter(iucn_sid %in% sid)
     if(nrow(spp_map) == 0) {
@@ -101,7 +102,6 @@ plot_rangemap <- function(spp) {
   colvec <- c('AquaMaps' = '#76A94C', 'Both' = '#e18258', 'IUCN' = '#9168a9')
   
   spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
-    ggtheme_map + 
     geom_raster(aes(fill = presence), alpha = .8) +
     scale_fill_manual(values = colvec) +
     borders('world', color='gray30', fill='gray40', size = .1) +  # create a layer of borders
@@ -144,6 +144,7 @@ create_spp_list <- function(spp_list_base, data_file, am_spp_cells,
[…]

### 2017-03-15 data_explore/data_explore.Rmd
updating based on WoRMS database.  Turns out the method doesn't improve upon the most recent round of species matches, but it is probably easier for future reference!
updated figures for manuscript

@@
 
-spp_am <- read_csv(file.path(dir_data, 'spp_am.csv')) %>%
+spp_match_key <- spp_map_pairs %>%
+  select(iucn_sid, am_sid)
+
+am_maps_class <- read_csv(file.path(dir_data, 'spp_am.csv')) %>%
+  select(am_sid, sciname) %>%
   left_join(read_csv(file.path(dir_M, 'git-annex/globalprep/_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>% 
-              dplyr::select(am_sid = speciesid, class) %>%
+              dplyr::select(am_sid = speciesid, am_class = class) %>%
               unique(), 
             by = 'am_sid') %>%
-  mutate(paired_map = am_sid %in% spp_map_pairs$am_sid)
+  left_join(spp_match_key, by = 'am_sid')
 ### nrow(spp_am %>% filter(paired_map)) ### 2330 spp
@@
 iucn_maps_class <- spp_iucn_maps %>% 
-  select(-iucn_subpop, -name_verified, -spp_group) %>%
+  select(iucn_sid, sciname_iucn = sciname, iucn_class = class) %>%
   distinct() %>%
-  mutate(class      = str_to_title(class),
-         paired_map = iucn_sid %in% spp_map_pairs$iucn_sid)
+  mutate(iucn_class = str_to_title(iucn_class))
 ### nrow(iucn_maps_class %>% filter(paired_map)) ### 2330 spp
 
-taxa_iucn_all <- iucn_maps_class %>%
-  filter(!is.na(class)) %>%           ### don't count the unknowns
+spp_class_all <- am_maps_class %>%
+  full_join(iucn_maps_class, by = 'iucn_sid') %>%
+  mutate(am_class = ifelse(am_class %in% c('null', 'not assigned'), NA, am_class),
+         class    = if_else(is.na(am_class), iucn_class, am_class))
+
+### Fix some missing class values
+# spp_class_all1 %>% filter(is.na(class))
+#       am_sid                   sciname am_class iucn_sid sciname_iucn iucn_class class
+# 1  Fis-22442 Metapenaeopsis tarawensis     <NA>       NA         <NA>       <NA>  <NA> # Malacostraca
+# 2 Fis-170350          Oenopota eriopis     <NA>       NA         <NA>       <NA>  <NA> # Gastropoda
+# 3 Fis-170316          Yarica hyalosoma     <NA>       NA         <NA>       <NA>  <NA> # Actinopterygii
+spp_class_all <- spp_class_all %>%
+  mutate(class = case_when(
+    is.na(.$class) & .$am_sid == 'Fis-22442' ~ 'Malacostraca',
+    is.na(.$class) & .$am_sid == 'Fis-170350' ~ 'Gastropoda',
+    is.na(.$class) & .$am_sid == 'Fis-170316' ~ 'Actinopterygii',
+    TRUE ~ .$class))
+
+write_csv(spp_class_all, file.path(dir_data, 'spp_class_all.csv'))
+
+```
+
+``` {r break down by taxa and data source}
+
+spp_class_all <- read_csv(file.path(dir_data, 'spp_class_all.csv'))
+
+taxa_iucn_all <- spp_class_all %>%
+  filter(!is.na(iucn_sid)) %>%
   mutate(n_iucn_tot = n()) %>%        ### length of whole maps list
@@
 
-taxa_iucn_solo <- iucn_maps_class %>%
-  filter(!is.na(class)) %>%                 ### don't count the unknowns
-  filter(!paired_map) %>%  ### filter list to IUCN only
+taxa_iucn_solo <- spp_class_all %>%
+  filter(!is.na(iucn_sid) & is.na(am_sid)) %>%  ### filter list to IUCN only
   mutate(n_iucn_solo = n()) %>%             ### length of IUCN only list
@@
 
-taxa_am_all <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
+taxa_am_all <- spp_class_all %>%
+  filter(!is.na(am_sid)) %>%
   mutate(n_am_tot = n()) %>%        ### length of whole AM maps list
@@
 
-taxa_am_solo <- spp_am %>%
-  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
-  filter(!paired_map) %>%  ### filter list to AM only
+taxa_am_solo <- spp_class_all %>%
+  filter(!is.na(am_sid) & is.na(iucn_sid)) %>%
   mutate(n_am_solo = n()) %>%       ### length of AM only list
@@
 
-taxa_both <- spp_am %>%
-  filter(paired_map) %>%
-  filter(!class %in% c('Not assigned', 'null')) %>%
+taxa_both <- spp_class_all %>%
+  filter(!is.na(am_sid) & !is.na(iucn_sid)) %>%
   mutate(n_both_tot = n()) %>%
@@
 
-taxa_df <- taxa_am_all %>%
+taxa_all <- taxa_am_all %>%
   full_join(taxa_am_solo,   by = 'class') %>%
@@
   full_join(taxa_both,      by = 'class') %>%
-  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))
+  mutate(small_class = (is.na(n_iucn_class_all) & p_am_class_all < 0.5))
 
-taxa_omitted <- taxa_am_all %>%
-  full_join(taxa_am_solo,   by = 'class') %>%
-  full_join(taxa_iucn_all,  by = 'class') %>%
-  full_join(taxa_iucn_solo, by = 'class') %>%
-  full_join(taxa_both,      by = 'class') %>%
-  filter(is.na(n_iucn_class_all) & p_am_class_all < 0.5) %>%
-  left_join(read_csv(file.path(dir_data, 'sciname_lookups/scinames.csv')) %>%
+taxa_df <- taxa_all %>%
+  filter(!small_class) %>%
+  select(-small_class)
+
+class_to_kingdom <- read_csv(file.path(dir_M, 'git-annex/globalprep/_raw_data',
+                                       'aquamaps/d2015/csv/speciesoccursum.csv')) %>%
   select(class, kingdom, phylum) %>%
-              mutate(class = str_to_title(class)),
+  mutate(class = str_to_title(class)) %>%
+  distinct()
+
+taxa_other <- taxa_all %>%
+  filter(small_class) %>%
+  select(-small_class) %>%
+  left_join(class_to_kingdom,
             by = 'class') %>%
@@
   select(-kingdom, -phylum) %>%
-  distinct()
-
-# knitr::kable(taxa_df %>%
-#                dplyr::select(class, 
-#                              AquaMaps = n_am_class_all, 
-#                              IUCN = n_iucn_class_all, 
-#                              Both = n_class_both))
+  group_by(class) %>%
+  summarize(n_am_class_all = sum(n_am_class_all),
+            p_am_class_all = sum(p_am_class_all),
+            n_am_class_solo = sum(n_am_class_solo),
+            p_am_class_solo = sum(p_am_class_solo))
 
 write_csv(taxa_df, file.path(dir_data, 'taxa_df.csv'))
-write_csv(taxa_omitted, file.path(dir_data, 'taxa_omitted.csv'))
+write_csv(taxa_other, file.path(dir_data, 'taxa_other.csv'))
 
@@
 taxa_df <- read_csv(file.path(dir_data, 'taxa_df.csv'))
-taxa_omitted <- read_csv(file.path(dir_data, 'taxa_omitted.csv'))
+taxa_other <- read_csv(file.path(dir_data, 'taxa_other.csv'))
 
-n_spp_omitted <- taxa_omitted$n_am_class_all %>% sum()
+n_spp_other <- taxa_other$n_am_class_all %>% sum()
 
@@
 
-taxa_all_df <- taxa_df1 %>%
+taxa_total_df <- taxa_df1 %>%
   group_by(dataset) %>%
   summarize(class = 'Total',
-            AquaMaps_total = sum(AquaMaps)  + n_spp_omitted + sum(Paired),
+            AquaMaps_total = sum(AquaMaps)  + n_spp_other + sum(Paired),
             IUCN_total = sum(IUCN) + sum(Paired),
-            n_total  = sum(n_total) + n_spp_omitted,
+            n_total  = sum(AquaMaps) + sum(IUCN) + n_spp_other + sum(Paired),
             IUCN_end = 100 * IUCN_total/n_total,
@@
 
-taxa_df2 <- taxa_all_df %>%
-  bind_rows(taxa_all_df %>%
+taxa_df2 <- taxa_total_df %>%
+  bind_rows(taxa_total_df %>%
               mutate(class = '',
@@
               arrange(desc(n_total))) %>%
-  bind_rows(taxa_omitted %>%
+  bind_rows(taxa_other %>%
               group_by(class) %>%
@@
 
+spp_list_quads <- read_csv(file.path(dir_data, 'spp_list_quads.csv'))
+
 gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence.csv'))

### 2017-03-15 data/sciname_lookups/process_scinames.R

@@ -1,34 +1,0 @@
-library(tidyr)
-library(dplyr)
-library(stringr)
-
-files_list <- list.files('sciname_lookups')
-
-all_df <- data.frame()
-
-for (i in 1:length(files_list)) { # i = 1
-  x <- scan(file.path('sciname_lookups', files_list[i]), what = 'character', sep = '\n', skip = 3)
-  
-  df <- data.frame('raw' = str_trim(x), stringsAsFactors = FALSE) %>%
-    mutate(raw = str_replace(raw, '  ', ' ')) %>%
-    separate(raw, c('level', 'name'), sep = ' ', remove = FALSE, extra = 'drop') %>%
-    mutate(level = tolower(str_replace(level, ':', '')),
-           name = tolower(name)) %>%
-    mutate(sub_infra = str_detect(level, '^sub')  | str_detect(level, '^infra')) %>%
-    mutate(kingdom = ifelse(str_detect(level, 'kingdom'), name, NA),
-           phylum  = ifelse(str_detect(level, 'phylum'),  name, NA),
-           class   = ifelse(str_detect(level, 'class'),   name, NA),
-           order   = ifelse(str_detect(level, 'order'),   name, NA),
-           family  = ifelse(str_detect(level, 'family'),  name, NA),
-           genus   = ifelse(str_detect(level, 'genus'),   name, NA))
-  
-  df1 <- df %>%
-    filter(!sub_infra) %>% 
-    fill(kingdom:family) %>%
-    filter(!is.na(genus)) %>%
-    select(kingdom:genus)
-  
-  all_df <- rbind(all_df, df1)
-}
-
-write.csv(all_df, 'sciname_lookups/scinames.csv', row.names = FALSE)

### 2017-03-16 data_explore/data_explore.Rmd
done once and for all with this?

@@
   full_join(iucn_maps_class, by = 'iucn_sid') %>%
-  mutate(am_class = ifelse(am_class %in% c('null', 'not assigned'), NA, am_class),
-         class    = if_else(is.na(am_class), iucn_class, am_class))
+  mutate(am_class = ifelse(tolower(am_class) %in% c('null', 'not assigned'), NA, am_class),
+         class    = ifelse(identical(iucn_class, am_class), am_class, NA),
+         class    = ifelse(is.na(class & !is.na(iucn_class)), iucn_class, am_class))
 
-### Fix some missing class values
-# spp_class_all1 %>% filter(is.na(class))
-#       am_sid                   sciname am_class iucn_sid sciname_iucn iucn_class class
-# 1  Fis-22442 Metapenaeopsis tarawensis     <NA>       NA         <NA>       <NA>  <NA> # Malacostraca
-# 2 Fis-170350          Oenopota eriopis     <NA>       NA         <NA>       <NA>  <NA> # Gastropoda
-# 3 Fis-170316          Yarica hyalosoma     <NA>       NA         <NA>       <NA>  <NA> # Actinopterygii
+### NAs due to null or not assigned Aquamaps species.
+# spp_class_na <- spp_class_all %>%
+#   filter(is.na(class)) %>%
+#   select(am_sid, sciname)
+# 
+# spp_class_from_worrms <- data.table()
+# for (i in seq_along(spp_class_na$sciname)) { ## i <- 1
+#   am_sid  = spp_class_na$am_sid[i]
+#   sciname = spp_class_na$sciname[i]
+#   print(paste(i, sciname, sep = ' '))
+#   x <- try(worrms::wm_records_taxamatch(name = spp_class_na$sciname[i]))
+#   if(class(x) == 'try-error') {
+#     x <- data.frame('am_sid'  = am_sid,
+#                     'sciname' = sciname,
+#                     'result'  = 'try-error')
+#   } else {
+#     x <- x %>%
+#       as.data.frame() %>%
+#       select(scientificname, valid_name, class, phylum, kingdom) %>%
+#       mutate(sciname = sciname,
+#              am_sid  = am_sid)
+#   }
+#   spp_class_from_worrms <- bind_rows(spp_class_from_worrms, x)
+# }
+# 
+# ### still a few NAs:
+# spp_class_from_worrms %>% filter(result == 'try-error') %>% select(am_sid, sciname)
+# # Fis-170350	Oenopota eriopis			
+# # Fis-170316	Yarica hyalosoma			
+# # ITS-156736	Loxosoma murmanica
+# spp_class_from_worrms1 <- spp_class_from_worrms %>%
+#   mutate(class = case_when(
+#     is.na(.$class) & .$am_sid == 'Fis-170350' ~ 'Gastropoda', ### from ITIS
+#     is.na(.$class) & .$am_sid == 'Fis-170316' ~ 'Actinopterygii', ### from FishBase: cardinalfish
+#     is.na(.$class) & .$am_sid == 'ITS-156736' ~ 'Entoprocta', ### from ITIS; Loxosomella murmanica
+#     TRUE ~ .$class))
+# write_csv(spp_class_from_worrms, file.path(dir_data, 'sciname_lookups/spp_class_from_worrms.csv'))
+
+# spp_class_from_itis <- data.table()
+# for (i in seq_along(spp_class_na$sciname)) { ## i <- 3
+#   am_sid  = spp_class_na$am_sid[i]
+#   sciname = spp_class_na$sciname[i]
+#   print(paste(i, sciname, sep = ' '))
+#   x <- taxize::classification(spp_class_na$sciname[i], 'itis')[[1]]
+#   if(is.na(x)) {
+#     y <- data.frame(am_sid  = am_sid,
+#                   sciname = sciname,
+#                   result = 'error')
+#   } else {
+#     class <- x$name[x$rank == 'class']
+#     phylum <- x$name[x$rank == 'phylum']
+#     kingdom <- x$name[x$rank == 'kingdom']
+#     y <- data.frame(am_sid  = am_sid,
+#                     sciname = sciname,
+#                     class_itis = ifelse(length(class > 0), x$name[x$rank == 'class'], NA),
+#                     phylum_itis = ifelse(length(phylum > 0), x$name[x$rank == 'phylum'], NA),
+#                     kingdom_itis = ifelse(length(kingdom > 0), x$name[x$rank == 'kingdom'], NA))
+#   }
+#   spp_class_from_itis <- bind_rows(spp_class_from_itis, y)
+# }
+# write_csv(spp_class_from_itis, file.path(dir_data, 'sciname_lookups/spp_class_from_itis.csv'))
+
+spp_class_fill <- read_csv(file.path(dir_data, 'sciname_lookups/spp_class_from_worrms.csv')) %>%
+  select(-result) %>%
+  full_join(read_csv(file.path(dir_data, 'sciname_lookups/spp_class_from_itis.csv')) %>%
+                       select(-result),
+            by = c('sciname', 'am_sid')) %>%
+  mutate(class = ifelse(!is.na(class_worrms), class_worrms, class_itis)) %>%
+  select(am_sid, sciname, class)
+### NOTE: still some NAs; these are all class Animalia
+
 spp_class_all <- spp_class_all %>%
-  mutate(class = case_when(
-    is.na(.$class) & .$am_sid == 'Fis-22442' ~ 'Malacostraca',
-    is.na(.$class) & .$am_sid == 'Fis-170350' ~ 'Gastropoda',
-    is.na(.$class) & .$am_sid == 'Fis-170316' ~ 'Actinopterygii',
-    TRUE ~ .$class))
+  filter(!is.na(class)) %>%
+  bind_rows(spp_class_fill)
 
@@
   full_join(taxa_both,      by = 'class') %>%
-  mutate(small_class = (is.na(n_iucn_class_all) & p_am_class_all < 0.5))
+  mutate(small_class = (is.na(n_iucn_class_all) & p_am_class_all < 0.5) | is.na(class))
 
@@
   select(class, kingdom, phylum) %>%
+  bind_rows(read_csv(file.path(dir_data, 'sciname_lookups/spp_class_from_worrms.csv')) %>%
+              select(class = class_worrms, kingdom = kingdom_worrms, phylum = phylum_worrms)) %>%
+  bind_rows(read_csv(file.path(dir_data, 'sciname_lookups/spp_class_from_itis.csv')) %>%
+              select(class = class_itis, kingdom = kingdom_itis, phylum = phylum_itis)) %>%
   mutate(class = str_to_title(class)) %>%
-  distinct()
+  select(-phylum) %>% ### avoid duplicates
+  filter(!is.na(kingdom)) %>% ### NA class doesn't match with kingdom or phylum
+  distinct() ### All remaining NAs are Animalia
 
@@
             by = 'class') %>%
-  mutate(class = paste('other', str_to_title(kingdom), sep = ' ')) %>%
-  select(-kingdom, -phylum) %>%
+  mutate(kingdom = ifelse(is.na(class), 'Animalia', kingdom),
+         class = paste('other', str_to_title(kingdom), sep = ' ')) %>%
+  select(-kingdom) %>%
   group_by(class) %>%
@@
             p_am_class_solo = sum(p_am_class_solo))
+
+# taxa_other$n_am_class_all %>% sum(na.rm = TRUE) ### 968
+# taxa_df$n_am_class_all %>% sum(na.rm = TRUE) ### 21921
+### All AM species accounted for!
 
@@
 
-# print(bar_taxa)
+print(bar_taxa)
 ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 

### 2017-04-10 shiny_am_iucn/server_fxns.R
adding in leaflet

@@ -2,6 +2,7 @@
 
 library(raster)
 library(maps)
+library(leaflet)
 library(tmap)
 data(World)
 library(RColorBrewer)
@@ -91,6 +92,18 @@ get_rast <- function(spp_map_df, type) {
   
   return(rast_obj)
 }
+
+#function to use leaflet for species maps
+assemble_map_leaflet <- function(map_rast,spp){
+  message('in assemble_map_leaflet()')
+  pal <- colorFactor(c("#0C2C84", "#41B6C4","#FFFFCC"), values(rast),
+                      na.color = "transparent")
+  
+  leaflet() %>% addTiles() %>%
+    addRasterImage(resampleBy(rast, 2), colors = pal,opacity = 1) %>%
+    addLegend(pal = 'Spectral', values = values(rast))
+}
+
 
 assemble_map <- function(map_rast, spp) {
   message('in assemble_map()')

### 2017-04-10 shiny_am_iucn/server_fxns.R
changed up some leaflet stuff, but not using it so shouldn't impact app

@@ -93,15 +93,18 @@ get_rast <- function(spp_map_df, type) {
   return(rast_obj)
 }
 
-#function to use leaflet for species maps
+#function to use leaflet for species maps (we tested this but are not using it due to weird zoom issues)
 assemble_map_leaflet <- function(map_rast,spp){
   message('in assemble_map_leaflet()')
-  pal <- colorFactor(c("#0C2C84", "#41B6C4","#FFFFCC"), values(rast),
+  pal <- colorFactor(c("#FFAEB9", "#41B6C4","#0C2C84"), values(rast),
                       na.color = "transparent")
   
   leaflet() %>% addTiles() %>%
-    addRasterImage(resampleBy(rast, 2), colors = pal,opacity = 1) %>%
-    addLegend(pal = 'Spectral', values = values(rast))
+    addRasterImage(rast, colors = pal,opacity = 1) %>%
+    addLegend(colors = c("#FFAEB9", "#41B6C4","#0C2C84"), 
+              labels = c("Aquamaps","IUCN","Both"),
+              title = "Dataset",
+              opacity = 1)
 }
 
 

### 2017-04-10 data_explore/data_explore.Rmd
bar chart now appears under quad plot

@@
 # ggsave(plot = barchart_spp_gp_quads,
-       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
-       width = 19.05, height = 7, units = 'cm', dpi = 600)
+       # filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
+       # width = 19.05, height = 7, units = 'cm', dpi = 600)
 
@@
 # ggsave(plot = bar_spp_gp_quads_ex,
-       filename = file.path(dir_fig, 'barchart_spp_gp_quads_ex.png'),
-       width = 19.05, height = 7, units = 'cm', dpi = 600)
+       # filename = file.path(dir_fig, 'barchart_spp_gp_quads_ex.png'),
+       # width = 19.05, height = 7, units = 'cm', dpi = 600)
 
-write_csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads_ex.csv'))
+write_csv(spp_gp_quads_ex, file.path(dir_data, 'spp_gp_quads_ex.csv'))
 

### 2017-04-10 shiny_am_iucn/server.R

@@ -55,6 +55,11 @@ server <- shinyServer(function(input, output, session) {
     create_quadplot(input$taxa_quad, input$expert_rev)
   })
   
+  output$barchart <- renderPlot({
+    create_barchart(input$expert_rev)
+  })
+  
+  
   output$mini_quad <- renderPlot({
     create_miniquad(input$species)
   })

### 2017-04-10 shiny_am_iucn/server_fxns.R

@@ -14,6 +14,7 @@ loiczid_raster       <- raster(loiczid_raster_file)
 names(loiczid_raster) <- 'loiczid'
 
 ### read in spp cell files
+### this may be slowing down the initial display of the app... load elsewhere?
 iucn_spp_cells <- read_csv('data/iucn_cells.csv', col_types = 'dd')
 am_spp_cells   <- read_csv('data/am_cells.csv')
 
@@ -32,6 +33,11 @@ spp_coralmaps <- read_csv('data/coral_spp_areas.csv') %>%
   filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
   mutate(method = ifelse(str_detect(dist, 'clip'), '200m clip', 'all depth'))
 
+quad_gp_list <- read_csv('data/spp_gp_quads.csv') %>%
+  mutate(expert = FALSE) %>%
+  bind_rows(read_csv('data/spp_gp_quads_ex.csv') %>%
+              mutate(expert = TRUE))
+  
 ### generic theme for all plots
 theme_set(theme_bw())
 ggtheme_plot <- theme_update(axis.ticks = element_blank(),
@@ -130,6 +136,53 @@ assemble_map <- function(map_rast, spp) {
 
   return(map_obj)
 }
+
+create_barchart <- function(expt_rev) {
+  
+  if(expt_rev != 'all') {
+    spp_gp_quadrants <- quad_gp_list %>%
+      filter(expert)
+  } else {
+    spp_gp_quadrants <- quad_gp_list %>%
+      filter(!expert)
+  }
+  
+  spp_gp_quadrants <- spp_gp_quadrants %>% 
+    mutate(quad = factor(quad, levels = c('q4', 'q3', 'q2', 'q1'))) %>%
+    transform(spp_group_text = reorder(spp_group_text, pct_q1))
+  
+  barchart_spp_gp_quads <- ggplot(spp_gp_quadrants, 
+                                  aes(x = spp_group_text, 
+                                      y = pct_quad,
+                                      fill = quad, 
+                                      weight = pct_quad)) +
+    theme(panel.grid.major.x = element_blank()) +
+    geom_bar(stat = 'identity', alpha = 1) +
+    scale_fill_manual(values = c('q1' = '#4dac26',
+                                 'q2' = '#b8e186', 
+                                 'q3' = '#f1b6da',
+                                 'q4' = '#d01c8b'),
+                      labels = quad_names$quad_name,
[…]

### 2017-04-10 shiny_am_iucn/ui.R

@@ -59,10 +59,6 @@ ui <- navbarPage('Aligning marine species range data to better serve science and
   ),
   
   tabPanel('Map alignment',
-  ### This tab shows side bar with controls: expert-reviewed vs. all.
-  ### The main panel shows the quadrant plot with all vs expert-reviewed
-  ### map pairs, and the bar chart below it to show the membership in various quads.
-  ### Could there be a selection menu to select by taxa? yes, there could.
     fluidRow(
       column(2, 
         selectInput('taxa_quad', 'Choose a taxon:',
@@ -76,7 +72,7 @@ ui <- navbarPage('Aligning marine species range data to better serve science and
       column(10,
         plotlyOutput('quad_plot'),
         hr(),
-        h5('Bar chart goes here')
+        plotOutput('barchart')
       )
     )
   ),
@@ -87,7 +83,7 @@ ui <- navbarPage('Aligning marine species range data to better serve science and
   ### species falls relative to all the others (this species is a red dot).
   ### There will be buttons to show only IUCN, only AM, or both;
   ### there will also be a slider for AquaMaps threshold?
-  ### And a toggle to turn on/off the FAO boundaries on the main map.
+  ### Include FAO boundaries on the main map - no need for toggle on/off.
   ### The main panel will show the species map for the given species,
   ### with the given parameters.
     fluidRow(
@@ -105,7 +101,6 @@ ui <- navbarPage('Aligning marine species range data to better serve science and
                      selected = 'both'),
         sliderInput('am_cutoff', 'Probability of Occurrence cutoff (Aquamaps):',
                     min = 0, max = 1, value = 0, step = 0.1),
-        checkboxInput('show_fao', 'Show FAO boundaries?', value = TRUE),
         hr(),
         plotOutput('mini_quad')
       ),

### 2017-04-10 shiny_am_iucn/server.R
removed leaflet package, added in base plot and read in new land shapefile

@@ -41,7 +41,7 @@ server <- shinyServer(function(input, output, session) {
     {
       message('observed change in spp_map$df or input$show_maps; creating map')
       map_rast <- get_rast(spp_map$df, type = input$show_maps)
-      map_obj  <- assemble_map(map_rast, spp = input$species)
+      map_obj  <- assemble_map_base(map_rast, spp = input$species)
       
       spp_map$map <- map_obj
     }

### 2017-04-10 shiny_am_iucn/server_fxns.R

@@ -1,11 +1,13 @@
 ### server_fxns.R
-
+library(stringr)
 library(raster)
 library(maps)
-library(leaflet)
 library(tmap)
 data(World)
 library(RColorBrewer)
+library(rgdal)
+library(ggplot2)
+
 
 
 ### read data for raster (used to plot species)
@@ -13,6 +15,31 @@ message('Reading in raster info...\n')
 loiczid_raster_file  <- 'data/loiczid_raster.tif'
 loiczid_raster       <- raster(loiczid_raster_file)
 names(loiczid_raster) <- 'loiczid'
+
+land <- readOGR(dsn = 'data', layer = "ne_110m_land")
+
+# convert to dataframe
+land_df <- fortify(land)
+
+# create a blank ggplot theme
+theme_opts <- list(theme(panel.grid.minor = element_blank(),
+                         panel.grid.major = element_blank(),
+                         panel.background = element_blank(),
+                         plot.background = element_rect(fill="#e6e8ed"),
+                         panel.border = element_blank(),
+                         axis.line = element_blank(),
+                         axis.text.x = element_blank(),
+                         axis.text.y = element_blank(),
+                         axis.ticks = element_blank(),
+                         axis.title.x = element_blank(),
+                         axis.title.y = element_blank(),
+                         plot.title = element_text(size=22)))
+
+# plot map
+ggplot(land_df, aes(long,lat, group=group)) + 
+  geom_polygon()  + 
+  coord_equal() + 
+  theme_opts
 
 ### read in spp cell files
 ### this may be slowing down the initial display of the app... load elsewhere?
@@ -113,6 +140,15 @@ assemble_map_leaflet <- function(map_rast,spp){
               opacity = 1)
 }
 
+assemble_map_base <- function(map_rast,spp){
+  
+  plot(land,col='darkgray')
+  plot(map_rast,axes=F,col = c("#FFAEB9", "#41B6C4","#0C2C84"),
+       title = spp, add=T,box=F,legend=F)
+  legend("topright", inset=c(0,-.05),legend=c("Aquamaps","IUCN","Both"),
+                      col=c("#FFAEB9","#41B6C4","#0C2C84"),pch = 15,cex=0.7,bty='n')
+  
+}
 
 assemble_map <- function(map_rast, spp) {
   message('in assemble_map()')

### 2017-04-10 shiny_am_iucn/server.R
changing the coral quadplot to mini plot

@@ -19,7 +19,8 @@ server <- shinyServer(function(input, output, session) {
       spp_choices <- spp_list %>%
         filter(spp_group_text == input$spp_group) %>%
         distinct() %>%
-        .$sciname
+        .$sciname %>%
+        sort()
       updateSelectInput(session, inputId = "species",
                         choices = spp_choices)
     }
@@ -64,8 +65,8 @@ server <- shinyServer(function(input, output, session) {
     create_miniquad(input$species)
   })
 
-  output$coral_plot <- renderPlotly({
-    create_coralplot()
+  output$coral_quad <- renderPlot({
+    create_coralquad()
   })
   
 })

### 2017-04-10 shiny_am_iucn/server_fxns.R

@@ -10,9 +10,8 @@ library(RColorBrewer)
 
 ### read data for raster (used to plot species)
 message('Reading in raster info...\n')
-loiczid_raster_file  <- 'data/loiczid_raster.tif'
-loiczid_raster       <- raster(loiczid_raster_file)
-names(loiczid_raster) <- 'loiczid'
+loiczid_raster       <- raster('data/loiczid_raster.tif') %>%
+  setNames('loiczid')
 
 ### read in spp cell files
 ### this may be slowing down the initial display of the app... load elsewhere?
@@ -25,6 +24,7 @@ quad_list <- read_csv('data/spp_list_quads.csv') %>%
 
 area_align_mean <- mean(quad_list$area_ratio, na.rm = TRUE)
 dist_align_mean <- mean(quad_list$dist_align, na.rm = TRUE)
+
 
 spp_coralmaps <- read_csv('data/coral_spp_areas.csv') %>%
   rename(dist_align_raw = sm_perc_raw, dist_align_clipped = sm_perc_clipped) %>%
@@ -91,7 +91,7 @@ get_spp_map_df <- function(species, am_cutoff = 0) { ### species <- spp_list$sci
 
 get_rast <- function(spp_map_df, type) {
   message('in get_rast()')
-  rast_obj  <-  subs(loiczid_raster, spp_map_df, 
+  rast_obj  <-  raster::subs(loiczid_raster, spp_map_df, 
                      by    = 'loiczid', 
                      which = paste0(type, '_pres'), 
                      subsWithNA = TRUE)
@@ -172,7 +172,8 @@ create_barchart <- function(expt_rev) {
                                       y = pct_quad,
                                       fill = quad, 
                                       weight = pct_quad)) +
-    theme(panel.grid.major.x = element_blank()) +
+    theme(panel.grid.major.x = element_blank(),
+          text = element_text(size = 14)) +
     geom_bar(stat = 'identity', alpha = 1) +
     scale_fill_manual(values = c('q1' = '#4dac26',
                                  'q2' = '#b8e186', 
@@ -188,7 +189,7 @@ create_barchart <- function(expt_rev) {
     geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
     ### add text
     geom_text(aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
-              size = 2.5, 
+              size = 3,
               color = 'grey30') +
     coord_flip() +
     labs(x = 'Taxonomic Group', 
[…]

### 2017-04-10 shiny_am_iucn/ui.R

@@ -92,18 +92,16 @@ ui <- navbarPage('Aligning marine species range data to better serve science and
         h6('Select species group, then
           select species'),
         selectInput('spp_group', 'Choose a taxon:',
-                    choices = unique(spp_list$spp_group_text)),
+                    choices = unique(spp_list$spp_group_text) %>% sort()),
         selectInput('species', 'Choose a species:',
-                    choices = unique(spp_list$sciname)),
+                    choices = unique(spp_list$sciname) %>% sort()),
         radioButtons('show_maps', label = h3('Maps'),
                      choices = list('AquaMaps' = 'am',
                                     'IUCN'     = 'iucn', 
                                     'Both'     = 'both'),
                      selected = 'both'),
-        sliderInput('am_cutoff', 'Probability of Occurrence cutoff (Aquamaps):',
-                    min = 0, max = 1, value = 0, step = 0.1),
         hr(),
-        plotOutput('mini_quad')
+        plotOutput('mini_quad', height = '150px')
       ),
       column(10,
         plotOutput('compare_map')
@@ -119,12 +117,15 @@ ui <- navbarPage('Aligning marine species range data to better serve science and
   ### damselfish and butterflyfish? or is that going out on a limb...
   ### if we did that, use the same process as corals to determine depths
     fluidRow(
-      column(2, 
-        h5('controls?'),
-        h6('Toggle for bathymetry poly')
+      column(2,
+        h6('Select coral species'),
+        selectInput('coral_species', 'Choose a species:',
+                    choices = unique(spp_coralmaps$sciname) %>% sort()),
+        hr(),
+        plotOutput('coral_quad', height = '150px')
       ),
       column(10,
-        plotlyOutput('coral_plot'),
+        # plotlyOutput('coral_plot'),
         hr(),
         h5('Bar chart goes here')
       )

### 2017-04-10 shiny_am_iucn/server.R
bar chart works again; coral mini-quad works

@@ -14,7 +14,8 @@ server <- shinyServer(function(input, output, session) {
   ### This function takes a single species scientific name as input, then grabs 
   ### all occurrence cells and associated probability per cell
   
-  observeEvent(input$spp_group, ### select a group, then update species list choices
+  ### for Species Map tab: upon selecting a group, update species list choices
+  observeEvent(input$spp_group, 
     {
       spp_choices <- spp_list %>%
         filter(spp_group_text == input$spp_group) %>%
@@ -26,15 +27,18 @@ server <- shinyServer(function(input, output, session) {
     }
   )
   
+  ### For Species map tab:  upon selecting a species, get map cells
   observeEvent(
-    {input$species; input$am_cutoff}, ### change in species selector regenerates both maps
+    {input$species}, 
     {
-      message('observed change in input$species or input$am_cutoff; getting spp_map dataframe')
-      spp_map$df <- get_spp_map_df(input$species, input$am_cutoff)
+      message('observed change in input$species; getting spp_map dataframe')
+      spp_map$df <- get_spp_map_df(input$species)
       print(head(spp_map$df))
     }
   )
   
+  ### For Species map tab:  upon a change in the spp_map$df OR the input$show_maps,
+  ### create the map to show AM, IUCN, or Both
   create_map <- observeEvent(
     {spp_map$df; input$show_maps}, 
       ### triggered by change in spp_map$am_rast (which may also include 
@@ -52,6 +56,7 @@ server <- shinyServer(function(input, output, session) {
     spp_map$map
   }, width = 800, height = 600) 
   
+  ### For Map Alignment tab
   output$quad_plot <- renderPlotly({
     create_quadplot(input$taxa_quad, input$expert_rev)
   })
@@ -60,13 +65,13 @@ server <- shinyServer(function(input, output, session) {
     create_barchart(input$expert_rev)
   })
   
-  
   output$mini_quad <- renderPlot({
     create_miniquad(input$species)
   })
 
+  ### For Coral Depth tab
   output$coral_quad <- renderPlot({
-    create_coralquad()
+    create_coralquad(input$coral_species)
   })
   
 })

### 2017-04-10 shiny_am_iucn/ui.R

@@ -17,7 +17,7 @@ ui <- navbarPage('Aligning marine species range data to better serve science and
     ),
     mainPanel(
       includeMarkdown('pages/abstract.md')
-      )
+    )
   ),
   
 
@@ -63,7 +63,7 @@ ui <- navbarPage('Aligning marine species range data to better serve science and
     fluidRow(
       column(2, 
         selectInput('taxa_quad', 'Choose a taxon:',
-                    choices = c('all', unique(spp_list$spp_group_text)),
+                    choices = c('all', unique(spp_list$spp_group_text) %>% sort()),
                     selected = 'all'),
         radioButtons('expert_rev', label = 'AquaMaps review status',
                      choices = list('All'           = 'all',
@@ -120,7 +120,10 @@ ui <- navbarPage('Aligning marine species range data to better serve science and
       column(2,
         h6('Select coral species'),
         selectInput('coral_species', 'Choose a species:',
-                    choices = unique(spp_coralmaps$sciname) %>% sort()),
+                    choices = unique(spp_list %>% 
+                                       filter(spp_group_text == 'Corals') %>% 
+                                       .$sciname %>% 
+                                       sort())),
         hr(),
         plotOutput('coral_quad', height = '150px')
       ),

### 2017-04-10 shiny_am_iucn/ui_setup.R

@@ -1,4 +1,6 @@
-library(tidyverse)
+library(raster)
+library(tidyverse) ### after raster
+library(stringr)
 library(plotly)
 
 library(shinythemes)

### 2017-04-10 data_explore/data_explore.Rmd
changing the coral quadplot to mini plot
using tmap for maps; figured out the headspace problem using aspect ratio and inner/outer margin.

@@
 
-### define 25%, 50%, and 75% quartiles for both x and y axes
+### define means for both x and y axes
 area_align_mean <- mean(spp_list$area_ratio, na.rm = TRUE)

### 2017-04-10 shiny_am_iucn/server.R

@@ -45,8 +45,9 @@ server <- shinyServer(function(input, output, session) {
       ### change in iucn_rast) or a change in the selected maps to display
     {
       message('observed change in spp_map$df or input$show_maps; creating map')
+      message('input$show_maps = ', input$show_maps, '; input$species = ', input$species)
       map_rast <- get_rast(spp_map$df, type = input$show_maps)
-      map_obj  <- assemble_map_base(map_rast, spp = input$species)
+      map_obj  <- assemble_map_tmap(map_rast, spp = input$species)
       
       spp_map$map <- map_obj
     }
@@ -54,7 +55,7 @@ server <- shinyServer(function(input, output, session) {
   
   output$compare_map <- renderPlot({
     spp_map$map
-  }, width = 800, height = 600) 
+  }) 
   
   ### For Map Alignment tab
   output$quad_plot <- renderPlotly({

### 2017-04-10 shiny_am_iucn/server_fxns.R

@@ -10,7 +10,8 @@ message('Reading in raster info...\n')
 loiczid_raster <- raster('data/loiczid_raster.tif') %>%
   setNames('loiczid')
 
-land <- readOGR(dsn = 'data', layer = "ne_110m_land")
+land <- readOGR(dsn = 'data', layer = "ne_110m_land") %>%
+  crop(extent(-180, 180, -84, 85))
 
 # convert to dataframe
 land_df <- fortify(land)
@@ -120,7 +121,7 @@ get_spp_map_df <- function(species) { ### species <- spp_list$sciname[1]
 get_rast <- function(spp_map_df, type) {
   message('in get_rast(), spp_map_df is of class ', class(spp_map_df))
   print(head(spp_map_df))
-  rast_obj  <-  raster::subs(loiczid_raster, spp_map_df, 
+  rast_obj <- raster::subs(loiczid_raster, spp_map_df, 
                      by    = 'loiczid', 
                      which = paste0(type, '_pres'), 
                      subsWithNA = TRUE)
@@ -129,39 +130,68 @@ get_rast <- function(spp_map_df, type) {
 }
 
 #function to use leaflet for species maps (we tested this but are not using it due to weird zoom issues)
-assemble_map_leaflet <- function(map_rast,spp){
-  message('in assemble_map_leaflet()')
-  pal <- colorFactor(c("#FFAEB9", "#41B6C4","#0C2C84"), values(rast),
-                      na.color = "transparent")
-  
-  leaflet() %>% addTiles() %>%
-    addRasterImage(rast, colors = pal,opacity = 1) %>%
-    addLegend(colors = c("#FFAEB9", "#41B6C4","#0C2C84"), 
-              labels = c("Aquamaps","IUCN","Both"),
-              title = "Dataset",
-              opacity = 1)
-}
+# assemble_map_leaflet <- function(map_rast,spp) {
+#   message('in assemble_map_leaflet()')
+#   pal <- colorFactor(c("#FFAEB9", "#41B6C4","#0C2C84"), values(rast),
+#                       na.color = "transparent")
+#   
+#   leaflet() %>% addTiles() %>%
+#     addRasterImage(rast, colors = pal,opacity = 1) %>%
+#     addLegend(colors = c("#FFAEB9", "#41B6C4","#0C2C84"), 
+#               labels = c("Aquamaps","IUCN","Both"),
[…]

### 2017-04-11 shiny_am_iucn/data_prep.Rmd
setting up to get a better sample of species for maps

---
title: 'Data prep for Shiny app'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
  pdf_document:
    toc: true
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(maps)
library(data.table)
library(raster)
library(tidyverse)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git   <- '~/github/IUCN-AquaMaps'
dir_shiny <- file.path(dir_git, 'shiny_am_iucn')

dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico/v2016')

# if(basename(getwd()) != 'data_explore') setwd('data_explore')
source(file.path(dir_git, 'data_explore', 'data_explore_fxns.R'))
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

# Set up IUCN and AquaMaps species lists and cell lists

Things to consider when choosing species to include in the app:

* The pull down menus get unwieldy when long.
* Maps with more area (i.e. more cells) look better.
* We want a good cross section of species from each quadrant.
* We want a good cross section of species that are expert reviewed?

``` {r choose_species_to_map}

### copy and trim the list of overlapping species
spp_list_quads <- read_csv(file.path(dir_git, 'data', 'spp_list_quads.csv')) %>%
  select(sciname, iucn_sid, am_sid, 
         area_iucn, area_am,
         dist_align = sm_perc, area_ratio, 
         spp_group_text, 
         reviewed, occurcells, 
         n_spp, quad) %>%
  mutate(reviewed = ifelse(!is.na(reviewed), TRUE, FALSE))
write_csv(spp_list_quads, file.path(dir_shiny, 'data', 'spp_list_quads_app.csv'))

### copy the barchart files too
file.copy(file.path(dir_git, 'data', 'spp_gp_quads.csv'), file.path(dir_shiny, 'data', 'spp_gp_quads_app.csv'))
file.copy(file.path(dir_git, 'data', 'spp_gp_quads_ex.csv'), file.path(dir_shiny, 'data', 'spp_gp_quads_ex_app.csv'))

### Because we're using newer cell data (2016), and the quads list is based
### on 2015 data, cut any spp whose IDs have changed or otherwise gone awol.
if(!exists('iucn_cells')) {
  iucn_cells <- read_csv(file.path(dir_anx, 'int', 'iucn_cells_spp.csv'), col_types = 'ciid__')
}

if(!exists('am_cells')) {
  am_cells <- read_csv(file.path(dir_anx, 'int', 'am_cells_spp_prob0.csv'), col_types = 'cd')
}

spp_list_quads2016 <- spp_list_quads %>%
  filter(iucn_sid %in% iucn_cells$iucn_sid & am_sid %in% am_cells$am_sid)

### determine species list from IUCN area and AquaMaps area, as well as quadrant.
### Knock out anything below first quartile for both; also knock out the really huge ones?
# summary(spp_list_quads$area_iucn)
     # Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
     # 8383   1308000   4427000  13050000  16260000 355400000 
# summary(spp_list_quads$area_am)
     # Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
     # 2940   1425000   4495000  10960000  10100000 364700000
spp_list_med_areas <- spp_list_quads2016 %>%
  filter(area_iucn > quantile(area_iucn, .25)) %>%
  filter(area_am > quantile(area_am, .25)) %>%
  filter(area_am < quantile(area_am, .75) | area_iucn < quantile(area_iucn, .75))

### Start with a sample of 15% of the medium-area species; this is about 170 spp.
### Then group by species group and quadrant, and for each, ensure a minimum
### representation by selecting the first two (or one, if group is only 1)
spp_list_samples <- spp_list_med_areas %>%
  mutate(keep_spp = ifelse(runif(n()) <= .15, TRUE, FALSE)) %>%
  group_by(spp_group_text, quad) %>%
  mutate(n_spp_gp = n(),
         keep_spp = ifelse(iucn_sid == first(iucn_sid), TRUE, keep_spp),
         keep_spp = ifelse(iucn_sid == nth(iucn_sid, 2) & n_spp_gp >= 2, TRUE, keep_spp),
         kept     = sum(keep_spp)) %>%
  ungroup() %>%
  filter(keep_spp) %>%
  select(iucn_sid, am_sid, sciname, spp_group_text) %>%
  distinct()

write_csv(spp_list_samples, file.path(dir_shiny, 'data', 'spp_list_app.csv'))

```

``` {r prep_cells_files}

spp_list <- read_csv(file.path(dir_shiny, 'data', 'spp_list_app_stable.csv'))

### read IUCN cells from spp_ico/v2016, trim to just overlapping species cells
if(!exists('iucn_cells')) {
  iucn_cells <- read_csv(file.path(dir_anx, 'int', 'iucn_cells_spp.csv'), col_types = 'ciid__')
}

iucn_cells_map <- iucn_cells %>%
  filter(presence != 5) %>%
  filter(iucn_sid %in% spp_list$iucn_sid) %>%
  select(-sciname, -presence) %>%
  distinct()

write_csv(iucn_cells_map, file.path(dir_shiny, 'data', 'iucn_cells_app.csv'))

if(!exists('am_cells')) {
  am_cells <- read_csv(file.path(dir_anx, 'int', 'am_cells_spp_prob0.csv'), col_types = 'cd')
}

am_cells_map <- am_cells %>%
  filter(am_sid %in% spp_list$am_sid) %>%
  distinct()

write_csv(am_cells_map, file.path(dir_shiny, 'data', 'am_cells_app.csv'))

```

## Coral species

Get coral species file and bring it over here; clean it up a bit for easier processing.

``` {r get_coral_spp_info}
spp_coralmaps <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))

### recalc areas and overlaps
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, dist_align, c(sm_perc_raw, sm_perc_shallow)) %>%
  gather(area, area_ratio, c(area_ratio_raw, area_ratio_shallow)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'shallow') & !(str_detect(area, 'shallow')))) %>%
  mutate(method = ifelse(str_detect(dist, 'shallow'), '200 m clip', 'all depth')) %>%
  select(iucn_sid, am_sid, sciname, dist_align, area_ratio, method) %>%
  distinct()

write_csv(iucn_coral_fixed, file.path(dir_shiny, 'data', 'coral_spp_areas_app.csv'))

```

***

``` {r prov_footer, results = 'asis'}
prov_wrapup(commit_outputs = FALSE)
```

### 2017-04-11 shiny_am_iucn/server.R

@@ -2,10 +2,6 @@
 
 source('server_fxns.R')
 
-# am_prob_cols <- rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)) # rainbow color scheme
-# am_pres_cols <- colorRampPalette(brewer.pal(9, 'Oranges'))(255) # orange color scheme
-# iucn_cols    <- colorRampPalette(brewer.pal(9, 'Purples'))(255) # purple color scheme
-
 server <- shinyServer(function(input, output, session) {
   
   ### Species Map Function ###

### 2017-04-11 shiny_am_iucn/server_fxns.R

@@ -1,92 +1,60 @@
 ### server_fxns.R
-library(maps)
+library(raster)
+library(stringr)
+# library(maps)
 library(tmap)
-data(World)
+# data(World)
 library(RColorBrewer)
 library(rgdal)
 
 ### read data for raster (used to plot species)
-message('Reading in raster info...\n')
+message('Reading in raster info...')
 loiczid_raster <- raster('data/loiczid_raster.tif') %>%
   setNames('loiczid')
 
 land <- readOGR(dsn = 'data', layer = "ne_110m_land") %>%
   crop(extent(-180, 180, -84, 85))
 
-# convert to dataframe
-land_df <- fortify(land)
-
 # create a blank ggplot theme
-theme_opts <- list(theme(panel.grid.minor = element_blank(),
-                         panel.grid.major = element_blank(),
-                         panel.background = element_blank(),
-                         plot.background = element_rect(fill="#e6e8ed"),
-                         panel.border = element_blank(),
-                         axis.line = element_blank(),
-                         axis.text.x = element_blank(),
-                         axis.text.y = element_blank(),
-                         axis.ticks = element_blank(),
-                         axis.title.x = element_blank(),
-                         axis.title.y = element_blank(),
-                         plot.title = element_text(size=22)))
-
-# plot map
-# ggplot(land_df, aes(long,lat, group=group)) + 
-#   geom_polygon()  + 
-#   coord_equal() + 
-#   theme_opts
+ggtheme_basic <- function(textsize = 10) {
+  theme_bw() +
+  theme_update(panel.grid.minor = element_blank(),
+          panel.grid.major = element_blank(),
+          panel.background = element_blank(),
+          plot.background = element_blank(),
+          panel.border = element_blank(),
+          axis.line = element_blank(),
+          axis.ticks = element_blank(),
+          text = element_text(size = textsize),
+          plot.title = element_text(size = textsize * 1.5))
+}
 
 ### read in spp cell files
 ### this may be slowing down the initial display of the app... load elsewhere?
-iucn_spp_cells <- read_csv('data/iucn_cells.csv', col_types = 'dd')
[…]

### 2017-04-11 shiny_am_iucn/setup.R

@@ -1,43 +1,0 @@
-### shiny app setup: get proper files into place
-library(readr)
-library(dplyr)
-library(tidyr)
-
-dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
-           'Darwin'  = '/Volumes/ohi',
-           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]
-
-dir_anx      <- file.path(dir_M, 'git-annex/globalprep/spp_ico')
-
-
-### Set up the species cell files.  
-### These will be 200 randomly sampled species, with loiczid tables
-### for both IUCN and AM.  Also the species info file.
-if(FALSE) {
-  cat('Reading in IUCN species info...\n')
-  iucn_spp_cells_all <- read_csv(file.path(dir_anx, 'v2016/int/iucn_cells_spp.csv'),
-                             col_types = 'ciiddc', progress = TRUE)
-  ### sciname | iucn_sid | presence | loiczid | prop_area | subpop
-  
-  cat('Reading in Aquamaps species info...\n')
-  am_spp_cells_all   <- read_csv(file.path(dir_anx, 'v2016/int/am_cells_spp_prob0.csv'), progress = TRUE)
-  
-  spp_list <- read_csv(file.path(dir_anx, 'v2016/int/spp_all_cleaned.csv')) %>%
-    filter(!is.na(spp_group) &!is.na(am_sid)) %>%
-    filter(is.na(iucn_subpop)) %>%
-    sample_n(200) %>%
-    left_join(read_csv('../data/sciname_lookups/spp_groupnames.csv'),
-              by = 'spp_group')
-  
-  iucn_spp_cells <- iucn_spp_cells_all %>%
-    dplyr::select(iucn_sid, loiczid) %>%
-    filter(iucn_sid %in% spp_list$iucn_sid)
-  am_spp_cells <- am_spp_cells_all %>%
-    filter(am_sid %in% spp_list$am_sid)
-  
-  write_csv(spp_list, 'data/spp_list.csv')
-  write_csv(iucn_spp_cells, 'data/iucn_cells.csv')
-  write_csv(am_spp_cells, 'data/am_cells.csv')
-}
-
-### copy quad list file and quad plot

### 2017-04-11 shiny_am_iucn/ui_setup.R

@@ -1,8 +1,5 @@
-library(raster)
-library(tidyverse) ### after raster
-library(stringr)
+library(tidyverse) 
 library(plotly)
-
 library(shinythemes)
 
-spp_list <- read_csv('data/spp_list.csv')
+spp_list <- read_csv('data/spp_list_app_stable.csv')

### 2017-04-14 shiny_am_iucn/data_prep.Rmd
added coral maps functionality and lots of explanatory text

@@
 
-dir_git   <- '~/github/IUCN-AquaMaps'
+dir_git   <- '~/github/IUCN-AquaMaps' %>% path.expand()
 dir_shiny <- file.path(dir_git, 'shiny_am_iucn')
@@
 
+## Get FAO regions
+
+``` {r copy_fao_rgns}
+
+# fao_files <- list.files(file.path(dir_git, 'clip_fao/fao_rgns'), pattern = 'fao_rgns')
+# 
+# file.copy(from = file.path(dir_git, 'clip_fao/fao_rgns', fao_files),
+#           to   = file.path(dir_shiny, 'data', fao_files %>%
+#                              str_replace('\\.', '_app.')))
+# fao_rgn <- rgdal::readOGR(dsn = 'data', layer = "fao_rgns2_app") %>%
+#   crop(extent(-180, 180, -90, 90))
+# rgdal::writeOGR(fao_rgn, dsn = 'data', layer = "fao_rgns2_app", driver = 'ESRI Shapefile', overwrite.layer = TRUE)
+
+```
+
 ## Coral species
 
-Get coral species file and bring it over here; clean it up a bit for easier processing.
+Get coral species alignment file and bring it over here; clean it up a bit for easier processing.
 
@@
 
-write_csv(iucn_coral_fixed, file.path(dir_shiny, 'data', 'coral_spp_areas_app.csv'))
+write_csv(iucn_coral_fixed, file.path(dir_shiny, 'data', 'coral_align_app.csv'))
+
+```
+
+### Get coral cells and info for mapping
+
+``` {r get_coral_info}
+
+### load iucn spp_cells from main analysis; clip to just those in app
+spp_list <- read_csv(file.path(dir_shiny, 'data', 'spp_list_app_stable.csv'))
+
+coral_cells_app <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv')) %>%
+  filter(iucn_sid %in% spp_list$iucn_sid)
+# length(unique(coral_cells_app$iucn_sid))
+write_csv(coral_cells_app, file.path(dir_shiny, 'data', 'coral_cells_app.csv'))
+
+coral_spp_list  <- spp_list %>%
+  select(iucn_sid, sciname) %>%
+  filter(iucn_sid %in% coral_cells_app$iucn_sid) %>%
+  distinct()
+write_csv(spp_coral_maps, file.path(dir_shiny, 'data', 'coral_spp_list.csv'))
+
+coral_quad_sum <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_quad_sum.csv'))
+write_csv(coral_quad_sum, file.path(dir_shiny, 'data', 'coral_quads_app.csv'))
+
+```
+
+``` {r simplify_bathymetry}
+library(maptools)
+library(tmap)
+
+bathy_shp_orig <- rgdal::readOGR(dsn = file.path(dir_git, 'clip_depth/ne_10m_bathymetry_K_200'),
+                            layer = 'ne_10m_bathymetry_K_200') %>%
+  rgeos::gUnaryUnion()
+
+keep_df <- data.frame(
+  id   = 1:length(bathy_shp_orig@polygons[[1]]@Polygons),
+  area = lapply(bathy_shp_orig@polygons, function(x) sapply(x@Polygons, function(y) y@area)) %>%
+    unlist(),
+  hole = lapply(bathy_shp_orig@polygons, function(x) sapply(x@Polygons, function(y) y@hole)) %>%
+    unlist()
+)
+quantile(keep_df$area)
+#           0%          25%          50%          75%         100% 
+# 2.319330e-10 2.026033e-02 3.717621e-02 9.453268e-02 4.108813e+04
+
+### ditch the small polys.  Tolerance? 6e-2 (.06) is .25° x .25°; can go bigger,
+### perhaps up to a single .5° cell, or .25?
+hole_thresh <- .0625
+feat_thresh <- 2
+
+keep_df1 <- keep_df %>%
+  mutate(keep = TRUE,
+         # keep = ifelse(area < hole_thresh, FALSE, keep))
+         keep = ifelse(hole == TRUE  & area < hole_thresh, FALSE, keep),
+         keep = ifelse(hole == FALSE & area < feat_thresh, FALSE, keep))
+
+keep_ids <- keep_df1$id[keep_df1$keep]
+
+bathy_shp <- bathy_shp_orig
+
+bathy_shp@polygons[[1]]@Polygons <- bathy_shp@polygons[[1]]@Polygons[keep_ids]
+bathy_shp@polygons[[1]]@plotOrder <- 1:length(bathy_shp@polygons[[1]]@Polygons)
+bathy_shp@plotOrder <- 1:length(bathy_shp@polygons)
+
+### need to repair comments... ugh
+slot(bathy_shp, "polygons") <- lapply(slot(bathy_shp, "polygons"), checkPolygonsHoles)
+
+# # will assign comments, seen by: 
+# lapply(slot(bathy_shp, "polygons"), comment) 
+# 
+# # or remove them by: 
+# slot(bathy_shp, "polygons") <- lapply(slot(bathy_shp, "polygons"), 
+#    "comment<-", NULL)
+
+### convert to spatial polygons dataframe
+bathy_shp1 <- bathy_shp %>%
+  SpatialPolygonsDataFrame(data = data.frame('depth' = 200))
+
+### check new map against original map
+tm_shape(bathy_shp_orig) + 
+  tm_polygons(col = 'blue', lwd = 0) + 
+  tm_shape(bathy_shp1) + 
+  tm_polygons(col = 'red', 
+              border.col = 'red3', 
+              lwd = 0, alpha = .8)
+
+writePolyShape(bathy_shp1,
+               file.path(dir_shiny, 'data/ne_10m_bathymetry_K_200/bathy_app_simple'))
+file.copy(from = file.path(dir_git, 'clip_depth/ne_10m_bathymetry_K_200/ne_10m_bathymetry_K_200.prj'),
+          to   = file.path(dir_shiny, 'data/ne_10m_bathymetry_K_200/bathy_app_simple.prj'),
+          overwrite = FALSE)
+# rgdal::writeOGR(bathy_shp,
+#                 dsn    = path.expand(file.path(dir_shiny, 'data/ne_10m_bathymetry_K_200')),
+#                 layer  = 'bathy_app_simple',
+#                 driver = 'ESRI Shapefile',
+#                 overwrite_layer = TRUE)
 

### 2017-04-14 shiny_am_iucn/global.R

@@ -1,3 +1,0 @@
-# global.R
-
-

### 2017-04-14 shiny_am_iucn/server.R

@@ -68,7 +68,16 @@ server <- shinyServer(function(input, output, session) {
 
   ### For Coral Depth tab
   output$coral_quad <- renderPlot({
-    create_coralquad(input$coral_species)
+    create_coralquad(input$coral_spp)
   })
+  
+  output$coral_map <- renderPlot({
+    create_coral_map(input$coral_spp)
+  })
+  
+  output$coral_barchart <- renderPlot({
+    create_coral_barchart()
+  })
+  
   
 })

### 2017-04-14 shiny_am_iucn/server_fxns.R

@@ -5,15 +5,26 @@ library(stringr)
 library(tmap)
 # data(World)
 library(RColorBrewer)
-library(rgdal)
 
 ### read data for raster (used to plot species)
 message('Reading in raster info...')
 loiczid_raster <- raster('data/loiczid_raster.tif') %>%
   setNames('loiczid')
 
-land <- readOGR(dsn = 'data', layer = "ne_110m_land") %>%
-  crop(extent(-180, 180, -84, 85))
+display_extents <- extent(-180, 180, -84, 85)
+
+land <- rgdal::readOGR(dsn   = 'data/ne_110m_land', 
+                       layer = 'ne_110m_land') %>%
+  crop(display_extents)
+
+bathy_rast <- subs(loiczid_raster, 
+                   read_csv('data/bathy_cells.csv'),
+                   by = 'loiczid', which = 'deep') %>%
+  crop(display_extents)
+
+fao_rgn <- rgdal::readOGR(dsn   = 'data/fao_rgns', 
+                          layer = 'fao_rgns_app') %>%
+  crop(display_extents)
 
 # create a blank ggplot theme
 ggtheme_basic <- function(textsize = 10) {
@@ -41,7 +52,8 @@ area_align_mean <- mean(quad_list$area_ratio, na.rm = TRUE)
 dist_align_mean <- mean(quad_list$dist_align, na.rm = TRUE)
 
 ### read in coral species pre- and post-clipping
-spp_coralmaps <- read_csv('data/coral_spp_areas_app.csv')
+spp_coral_align <- read_csv('data/coral_align_app.csv')
+spp_coral_cells <- read_csv(file.path('data', 'coral_cells_app.csv'))
 
 quad_gp_list <- read_csv('data/spp_gp_quads_app.csv') %>%
   mutate(expert = FALSE) %>%
@@ -84,12 +96,15 @@ get_spp_map_df <- function(species) { ### species <- spp_list$sciname[1]
   return(spp_map_df)
 }
 
-get_rast <- function(spp_map_df, type) {
+get_rast <- function(spp_map_df, type, legend_classes = c(1, 2, 3)) {
+  ### legend classes is to force the legend to include all, even if
+  ### some classes are missing. 
+  ### e.g. c(1, 2, 3) for AM/IUCN/Both, c(1, 2) for shallow/deep corals
   message('in get_rast()')
   message('... rasterizing to type = ', type)
 
[…]

### 2017-04-14 shiny_am_iucn/ui.R

@@ -20,92 +20,73 @@ ui <- navbarPage('Aligning marine species range data to better serve science and
     )
   ),
   
-
   tabPanel('Figures',
     # Abstract of paper; sidebar is authors, publication info, etc
-    fluidRow(
-      column(2, 
-        h5('sidebar text'),
-        h6('This panel could be a table of contents of figures, ',
-           'with each a link to the figure that would navigate the ',
-           'main panel to the appropriate figure.'),
-        h6('OR: Thumbnails of figures in side bar, which pull up ',
-           'the full figure with caption in the main window - ',
-           'each figure gets its own .html file')
-      ),
-      column(10,
-        h4('Figures from manuscript'),
-        includeMarkdown('pages/fig1.md'),
-        hr(),
-        includeMarkdown('pages/fig2.md'),
-        hr(),
-        includeMarkdown('pages/fig3.md'),
-        hr(),
-        includeMarkdown('pages/fig4.md'),
-        hr(),
-        includeMarkdown('pages/fig5.md'),
-        hr(),
-        h4('Supporting figures'),
-        includeMarkdown('pages/s1fig.md'),
-        hr(),
-        includeMarkdown('pages/s2fig.md'),
-        hr(),
-        includeMarkdown('pages/s3fig.md'),
-        hr(),
-        includeMarkdown('pages/s4fig.md'),
-        hr(),
-        includeMarkdown('pages/s5fig.md')
-      )
+    sidebarPanel(
+      includeMarkdown('pages/author_list.md')
+    ),
+    mainPanel(
+      h4('Figures from manuscript'),
+      h5('Figure 1:'),
+      includeMarkdown('pages/fig1.md'),
+      hr(),
+      h5('Figure 2:'),
+      includeMarkdown('pages/fig2.md'),
+      hr(),
+      h5('Figure 3:'),
+      includeMarkdown('pages/fig3.md'),
+      hr(),
+      h5('Figure 4:'),
+      includeMarkdown('pages/fig4.md'),
+      hr(),
+      h5('Figure 5:'),
+      includeMarkdown('pages/fig5.md')
     )
   ),
   
   tabPanel('Map alignment',
-    fluidRow(
-      column(2, 
-        selectInput('taxa_quad', 'Choose a taxon:',
-                    choices = c('all', unique(spp_list$spp_group_text) %>% sort()),
-                    selected = 'all'),
-        radioButtons('expert_rev', label = 'AquaMaps review status',
[…]

### 2017-04-14 shiny_am_iucn/ui_setup.R

@@ -3,3 +3,4 @@ library(plotly)
 library(shinythemes)
 
 spp_list <- read_csv('data/spp_list_app_stable.csv')
+coral_spp_list <- read_csv('data/coral_spp_list.csv')

### 2017-04-21 case_study_mpa/1_wdpa_poly_to_raster.Rmd
initialize dummy branch

--- 2016-03-01 mpa_analysis/wdpa_prep.Rmd
@@
 ``` {r rasterize_that_shit}
-dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
+dir_rast <- file.path(dir_anx, 'vAM_IUCN/mpa_rasters')
 
@@
 ### import WDPA database polygons from saved .shp
-dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
+dir_rast <- file.path(dir_anx, 'vAM_IUCN/mpa_rasters')
 dir_wdpa   <- file.path(dir_rast, 'WDPA2014_mar_klein')

### 2017-04-21 case_study_mpa/2_wdpa_rast_to_csv.Rmd
initialize dummy branch

--- 2016-03-03 mpa_analysis/2_wdpa_rast_to_csv.Rmd
@@
 
-dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
+dir_rast <- file.path(dir_anx, 'vAM_IUCN/mpa_rasters')
 rast_mpa_mar_file <- file.path(dir_rast, 'rast_mpa_mar_i-iv.tif')

### 2017-04-21 case_study_mpa/x_mpa_gap_in_time.Rmd
initialize dummy branch

---
title: 'MPA Gap Analysis: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(raster)
library(readr)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')
dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')

if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)

### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

``` {r Analysis using WDPA 2015 and AM 2015, echo = FALSE, eval = FALSE}
### Use cell_df_des to capture only DESIGNATED protected areas.

area_mpa <- cell_df_des %>%
  summarize(marine_area  = sum(marine_area), 
            mpa_tot_1990 = sum(y1990_area), 
            mpa_tot_1995 = sum(y1995_area), 
            mpa_tot_2000 = sum(y2000_area), 
            mpa_tot_2005 = sum(y2005_area), 
            mpa_tot_2010 = sum(y2010_area), 
            mpa_tot_2014 = sum(y2014_area), 
            mpa_tot_2015 = sum(y2015_area))

### Assemble the AM species-cell df with info on marine area and 
### protected area per cell
mpa_spp_2015 <- am_spp_cells %>%
  left_join(cell_df_des, by = 'loiczid')

### filter to the appropriate threshold...
thresh <- 0.50

mpa_spp_2015 <- mpa_spp_2015 %>%
  filter(probability >= thresh)

### group by speciesid, then summarize sum(marine area), 
### sum(protected area), ratio
mpa_spp_2015_sum <- mpa_spp_2015 %>%
  group_by(speciesid) %>%
  summarize(cellarea_tot = sum(cellarea), 
            ocean_range  = sum(oceanarea), 
            marine_range = sum(marine_area),
            mpa1990 = sum(y1990_area),
            mpa1995 = sum(y1995_area),
            mpa2000 = sum(y2000_area),
            mpa2005 = sum(y2005_area),
            mpa2010 = sum(y2010_area),
            mpa2014 = sum(y2014_area),
            mpa2015 = sum(y2015_area),
            pct1990 = mpa1990/marine_range,
            pct1995 = mpa1995/marine_range,
            pct2000 = mpa2000/marine_range,
            pct2005 = mpa2005/marine_range,
            pct2010 = mpa2010/marine_range,
            pct2014 = mpa2014/marine_range,
            pct2015 = mpa2015/marine_range)

mpa_spp_2015_sum <- mpa_spp_2015_sum %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))

mpa_spp_2015_sum2 <- mpa_spp_2015_sum %>%
  group_by(cat2015) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2015 = factor(cat2015, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2015)

knitr::kable(mpa_spp_2015_sum2)

# |cat2015 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     | 22885|   802|     3.5|
# |0-2%    | 22885| 11035|    48.2|
# |2-5%    | 22885|  8703|    38.0|
# |5-10%   | 22885|  1656|     7.2|
# |>10%    | 22885|   689|     3.0|

```

| study            | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :--------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015 | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, wdpa2014 (recreate Klein) | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| am2015, wdpa2014 (update Klein)   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.7% | 45.7% | 42.7% |  7.2% | 2.7% |
| am2015, wdpa2015 | IUCN I-IV, Sp. Mar |   2,889,758  | 3.5% | 48.2% | 38.0% |  7.2% | 3.0% |

So: for analysis, just keep using the Klein WDPA shapefile... avoid that particular variable!

``` {r check results for 1990 through 2010 at 50%, echo = FALSE, eval = FALSE}

### Save this for the time series analysis...

mpa_spp_2015_sum <- mpa_spp_2015_sum %>%
  mutate(cat1990 = ifelse(pct1990 == 0,                       'gap', NA),
         cat1990 = ifelse(pct1990 > 0    & pct1990 <= 0.02,  '0-2%', cat1990),
         cat1990 = ifelse(pct1990 > 0.02 & pct1990 <= 0.05,  '2-5%', cat1990),
         cat1990 = ifelse(pct1990 > 0.05 & pct1990 <= 0.10, '5-10%', cat1990),
         cat1990 = ifelse(pct1990 > 0.10,                    '>10%', cat1990),
         cat1995 = ifelse(pct1995 == 0,                       'gap', NA),
         cat1995 = ifelse(pct1995 > 0    & pct1995 <= 0.02,  '0-2%', cat1995),
         cat1995 = ifelse(pct1995 > 0.02 & pct1995 <= 0.05,  '2-5%', cat1995),
         cat1995 = ifelse(pct1995 > 0.05 & pct1995 <= 0.10, '5-10%', cat1995),
         cat1995 = ifelse(pct1995 > 0.10,                    '>10%', cat1995),
         cat2000 = ifelse(pct2000 == 0,                       'gap', NA),
         cat2000 = ifelse(pct2000 > 0    & pct2000 <= 0.02,  '0-2%', cat2000),
         cat2000 = ifelse(pct2000 > 0.02 & pct2000 <= 0.05,  '2-5%', cat2000),
         cat2000 = ifelse(pct2000 > 0.05 & pct2000 <= 0.10, '5-10%', cat2000),
         cat2000 = ifelse(pct2000 > 0.10,                    '>10%', cat2000),
         cat2005 = ifelse(pct2005 == 0,                       'gap', NA),
         cat2005 = ifelse(pct2005 > 0    & pct2005 <= 0.02,  '0-2%', cat2005),
         cat2005 = ifelse(pct2005 > 0.02 & pct2005 <= 0.05,  '2-5%', cat2005),
         cat2005 = ifelse(pct2005 > 0.05 & pct2005 <= 0.10, '5-10%', cat2005),
         cat2005 = ifelse(pct2005 > 0.10,                    '>10%', cat2005),
         cat2010 = ifelse(pct2010 == 0,                       'gap', NA),
         cat2010 = ifelse(pct2010 > 0    & pct2010 <= 0.02,  '0-2%', cat2010),
         cat2010 = ifelse(pct2010 > 0.02 & pct2010 <= 0.05,  '2-5%', cat2010),
         cat2010 = ifelse(pct2010 > 0.05 & pct2010 <= 0.10, '5-10%', cat2010),
         cat2010 = ifelse(pct2010 > 0.10,                    '>10%', cat2010),
         cat2014 = ifelse(pct2014 == 0,                       'gap', NA),
         cat2014 = ifelse(pct2014 > 0    & pct2014 <= 0.02,  '0-2%', cat2014),
         cat2014 = ifelse(pct2014 > 0.02 & pct2014 <= 0.05,  '2-5%', cat2014),
         cat2014 = ifelse(pct2014 > 0.05 & pct2014 <= 0.10, '5-10%', cat2014),
         cat2014 = ifelse(pct2014 > 0.10,                    '>10%', cat2014))

mpa_spp_1990 <- mpa_spp_2015_sum %>%
  group_by(cat1990) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat1990 = factor(cat1990, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat1990)

mpa_spp_1995 <- mpa_spp_2015_sum %>%
  group_by(cat1995) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat1995 = factor(cat1995, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat1995)

mpa_spp_2000 <- mpa_spp_2015_sum %>%
  group_by(cat2000) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2000 = factor(cat2000, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2000)

mpa_spp_2005 <- mpa_spp_2015_sum %>%
  group_by(cat2005) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2005 = factor(cat2005, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2005)

mpa_spp_2010 <- mpa_spp_2015_sum %>%
  group_by(cat2010) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2010 = factor(cat2010, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2010)

mpa_spp_2014 <- mpa_spp_2015_sum %>%
  group_by(cat2014) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2014 = factor(cat2014, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2014)

#   cat1990 n_tot n_cat pct_cat
#       gap 22889   925     4.0
#      0-2% 22889 21569    94.2
#      2-5% 22889   376     1.6
#     5-10% 22889    18     0.1
#      >10% 22889     1     0.0
# 
#   cat1995 n_tot n_cat pct_cat
#       gap 22889   878     3.8
#      0-2% 22889 21111    92.2
#      2-5% 22889   817     3.6
#     5-10% 22889    70     0.3
#      >10% 22889    13     0.1
#
#   cat2000 n_tot n_cat pct_cat
#       gap 22889   801     3.5
#      0-2% 22889 20973    91.6
#      2-5% 22889  1004     4.4
#     5-10% 22889    92     0.4
#      >10% 22889    19     0.1
#
#   cat2005 n_tot n_cat pct_cat
#       gap 22889   608     2.7
#      0-2% 22889 19828    86.6
#      2-5% 22889  2174     9.5
#     5-10% 22889   205     0.9
#      >10% 22889    74     0.3
# 
#   cat2010 n_tot n_cat pct_cat
#       gap 22889   578     2.5
#      0-2% 22889 18670    81.6
#      2-5% 22889  3278    14.3
#     5-10% 22889   288     1.3
#      >10% 22889    75     0.3

argh <- am_spp_mpas_filt %>%
  select(-(mpa1990:mpa2015)) %>%
  gather(key = year, value = pct, pct1990:pct2015) %>%
  mutate(year = str_replace(year, 'pct', '')) %>%
  group_by(year)

hist_plot <- ggplot(data = argh, aes(x = pct, color = year)) + 
  geom_density() +
  scale_x_continuous(limits = c(0, .05))
#  scale_x_log10()

hist_plot
ggsave(file.path(dir_git, 'figures/mpa_analysis_dist.png'))
```

### Original study (Klein 2015) vs new analyses

| study           | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :-------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015| IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, mpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| .               | .                  | .            | .    | .     | .     | .     | .    |
| am2015, mpa2015 | IUCN I-IV, Sp. Mar |   2,889,758  | 2.4% | 50.3% | 39.1% |  6.4% | 1.7% |
| am2015, mpa2014 | IUCN I-IV, Sp. Mar |   2,885,465  | 2.4% | 50.4% | 39.1% |  6.4% | 1.7% |
| am2015, mpa2010 | IUCN I-IV, Sp. Mar |   1,137,142  | 2.5% | 81.6% | 14.3% |  1.3% | 0.3% |
| am2015, mpa2005 | IUCN I-IV, Sp. Mar |     932,817  | 2.7% | 86.6% |  9.5% |  0.9% | 0.3% |
| am2015, mpa2000 | IUCN I-IV, Sp. Mar |     748,172  | 3.5% | 91.6% |  4.4% |  0.4% | 0.1% |
| am2015, mpa1995 | IUCN I-IV, Sp. Mar |     532,098  | 3.8% | 92.2% |  3.6% |  0.3% | 0.1% |
| am2015, mpa1990 | IUCN I-IV, Sp. Mar |     351,592  | 4.0% | 94.2% |  1.6% |  0.1% | 0.0% |

max protected species = Fis-160882

### 2017-04-21 clip_depth/coral_maps.Rmd
initialize dummy branch

@@
 
+ggsave(plot = clipped_quads,
+       filename = file.path(dir_fig, 'si_scatter_quadmap_corals.tif'), device = 'tiff',
+       width = 19, height = 9, units = 'cm', dpi = 600)
+
 ```
@@
 
-quad_names <- data.frame('quad' = c('q1', 'q2', 'q3', 'q4'),
-                         'quad_name' = factor(c('well-aligned', 'dist-aligned', 
-                                                'area-aligned', 'poorly aligned'),
+quad_names <- data.frame('quad' = c('q4', 'q3', 'q2', 'q1'),
+                         'quad_name' = factor(c('poorly aligned', 
+                                                'area-aligned', 
+                                                'dist-aligned',
+                                                'well-aligned'),
                                              ordered = TRUE))
 
-iucn_coral_quads <- iucn_coral_quads %>%
-  left_join(quad_names, by = 'quad')
+iucn_coral_quads1 <- iucn_coral_quads %>%
+  left_join(quad_names, by = 'quad') %>%
+  mutate(quad = factor(quad, levels = c('q4', 'q3', 'q2', 'q1'), ordered = TRUE))
 
-write_csv(iucn_coral_quads, file.path(dir_git, 'clip_depth', 'iucn_coral_quads.csv'))
+write_csv(iucn_coral_quads1, file.path(dir_git, 'clip_depth', 'iucn_coral_quads.csv'))
 
@@
 
-iucn_coral_quad_sum <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_quads.csv')) %>%
+iucn_coral_quad_sum <- iucn_coral_quads1 %>%
   dplyr::select(method, quad, quad_name, pct_quad) %>%
@@
                                'q1' = '#4dac26'),
-                    labels = quad_names$quad_name) +
+                    labels = quad_names$quad_name,
+                    guide = guide_legend(reverse = TRUE)) +
   scale_y_continuous(expand = c(0, 0), 

### 2017-04-21 data_explore/data_explore.Rmd
initialize dummy branch

@@
 bar_taxa <- ggplot(taxa_df2, aes(x = class, color = dataset)) +
+  ggtheme_basic() +
   theme(panel.grid.major.x = element_blank(),
@@
 print(bar_taxa)
-ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
-       width = 19.05, height = 7, units = 'cm', dpi = 600)
+# ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
+#        width = 19.05, height = 7, units = 'cm', dpi = 600)
 

### 2017-04-21 data_explore/data_explore_gapmaps.Rmd
initialize dummy branch

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(maps)
library(readr)
library(data.table)
library(raster)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

if(basename(getwd()) != 'data_explore') setwd('data_explore')
source('data_explore_fxns.R')
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
### this file created in data_prep
spp_list <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))

loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
loiczid_raster       <- raster::raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'

```

``` {r plot_spp_iucn_gaps_tenplus, eval = TRUE, echo = FALSE, warning = FALSE}

iucn_gap10     <- read_csv(file.path(dir_data, 'iucn_gaps_tenplus.csv'))
iucn_gap10_spp <- unique(iucn_gap10[iucn_gap10$iucn_sid %in% spp_list$iucn_sid, ])

iucn_gapmaps <- sample_n(iucn_gap10_spp %>% filter(cat == 'gap'), 10)
iucn_10maps  <- sample_n(iucn_gap10_spp %>% filter(cat != 'gap'), 10)

if(!exists('iucn_spp_cells')) {
  iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'),
                             col_types = 'cddddc')
  am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'))
}

am_thresh = 0

message('Printing gap maps')
for (i in 1:nrow(iucn_gapmaps)) { ### i <- 1
  spp <- iucn_gapmaps[i, ]

  spp_plot <- plot_rangemap(spp)
  spp_plot <- spp_plot +
    labs(title = sprintf('%s: %s', spp$sciname, spp$spp_group))

  print(spp_plot)
}

message('Printing >10% maps')

for (i in 1:nrow(iucn_10maps)) { ### i <- 1
  spp <- iucn_10maps[i, ]
  spp_plot <- plot_rangemap(spp)
  spp_plot <- spp_plot +
    labs(title = sprintf('%s: %s', spp$sciname, spp$spp_group))

  print(spp_plot)
}

```

### 2017-04-21 data_explore/data_explore_global_maps.Rmd
initialize dummy branch

@@
 all_am_rast_file   <- file.path(dir_data, 'rasters/am_all_rast.tif')
-dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')
+dir_am_data <- file.path(dir_M, 'git-annex/globalprep/_raw_data/aquamaps/d2015/csv')
 
@@
 png(file.path(dir_fig, 'all_aquamaps_map.png'),
-    width = 8.7, height = 6, units = 'in', res = 300)
+    width = 8.7, height = 6, units = 'cm', res = 600)
 print(am_levelplot1)
@@
 all_iucn_rast_file   <- file.path(dir_data, 'rasters/iucn_all_rast.tif')
-dir_data_iucn <- file.path(dir_N, 'git-annex/globalprep/_raw_data/iucn_spp')
+dir_data_iucn <- file.path(dir_M, 'git-annex/globalprep/_raw_data/iucn_spp')
 
@@
 png(file.path(dir_fig, 'all_iucn_map.png'),
-    width = 8.7, height = 6, units = 'in', res = 300)
+    width = 8.7, height = 6, units = 'cm', res = 600)
 print(iucn_levelplot1)

### 2017-04-21 data_explore/data_explore_quadmaps.Rmd
initialize dummy branch

@@
 
-for (i in 1:length(spp_quads)) { # i = 1
+for (i in 1:length(spp_quads)) { # i = 3
   species = spp_quads[i]
@@
     spp_map <- tm_shape(spp_rast) +
-      tm_raster(palette = c('#4dac26', '#d01c8b', '#88ccff'),
+      # tm_raster(palette = c('#8da0cb', '#fc8d62', '#66c2a5'),
+      tm_raster(palette = c('#660000', 'yellowgreen', 'firebrick1'),
                 title  = 'Dataset',
@@
     spp_map <- tm_shape(spp_rast) +
-      tm_raster(palette = c('#4dac26', '#d01c8b', '#aaccff'),
+      tm_raster(palette = c('#660000', 'yellowgreen', 'firebrick1'),
                 legend.show = FALSE,
@@
     tm_shape(world_crop) +
-      tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
+      tm_polygons(border.col = 'grey50', col = 'grey40', lwd = 0.25) + 
     tm_layout(# title = species, title.size = .75,

### 2017-04-21 data_explore/data_prep.Rmd
initialize dummy branch

@@
 ```{r verify_spp_names_in_shps, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
+### Some species names in the shapefile polygons do not match the IUCN database names.  Fix this!
 
@@
 write_csv(spp_iucn_maps1, file.path(dir_data, 'spp_iucn_maps_verified.csv'))
+
+```
+
+``` {r test_worrms_matching}
+
+library(worrms)
+spp_map_pairs <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
+
+spp_names <- spp_map_pairs$sciname
+
+iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps_verified.csv')
+spp_iucn_maps <- read_csv(iucn_maps_file)
+
+spp_am_maps <- read_csv(file.path(dir_data, 'speciesoccursum.csv')) %>%
+  mutate(sciname = paste(str_trim(genus), str_trim(species)),
+         iucn_id = as.integer(iucn_id)) %>%
+  dplyr::select(am_sid = speciesid, sciname, reviewed, iucn_id)
+
+spp_am_ver <- lapply(spp_am_maps$sciname, FUN = function(x) {
+    y <- try(wm_records_names(x))
+    if (class(y) != 'try-error') {
+      return(as.data.frame(y))
+    } else {
+      return(data.frame(status = 'error'))
+    }
+  }) %>%
+  setNames(spp_am_maps$sciname) %>%
+  bind_rows(.id = 'sciname_orig')
+
+spp_am_ver1 <- spp_am_maps %>%
+  select(am_sid, sciname, iucn_id) %>%
+  left_join(spp_am_ver %>%
+    select(sciname_orig, sciname_worrms = scientificname, status, sciname_accepted = valid_name),
+    by = c('sciname' = 'sciname_orig')) %>%
+  distinct() %>%
+  mutate(am_name_ver = sciname == sciname_accepted)
+
+write_csv(spp_am_ver1, file.path(dir_git, 'data/sciname_lookups/am_worrms_verified.csv'))
+
+spp_iucn_ver <- lapply(spp_iucn_maps$sciname, FUN = function(x) {
+    y <- try(wm_records_names(x))
+    if (class(y) != 'try-error') {
+      return(as.data.frame(y))
+    } else {
+      return(data.frame(status = 'error'))
+    }
+  }) %>%
+  setNames(spp_iucn_maps$sciname) %>%
+  bind_rows(.id = 'sciname_orig')
+
+spp_iucn_ver1 <- spp_iucn_maps %>%
+  select(iucn_sid, sciname, spp_group) %>%
+  left_join(spp_iucn_ver %>%
+    select(sciname_orig, sciname_worrms = scientificname, status, sciname_accepted = valid_name),
+    by = c('sciname' = 'sciname_orig')) %>%
+  distinct() %>%
+  mutate(iucn_name_ver = sciname == sciname_accepted)
+
+write_csv(spp_iucn_ver1, file.path(dir_git, 'data/sciname_lookups/iucn_worrms_verified.csv'))
+
+### match by id first
+id_match <- inner_join(spp_iucn_ver1 %>%
+                         select(iucn_sid, iucn_sciname = sciname, iucn_sciname_accepted = sciname_accepted), 
+                       spp_am_ver1 %>%
+                         select(am_sid, iucn_id, am_sciname = sciname, am_sciname_accepted = sciname_accepted),
+                       by = c('iucn_sid' = 'iucn_id')) %>%
+  distinct()
+### 1960 matches; all distinct?
+# length(id_match$iucn_sid %>% unique()); length(id_match$am_sid %>% unique())
+# 1880 for both (same as prior method, which makes sense)
+
+### then match the rest by verified accepted name
+iucn_id_miss <- spp_iucn_ver1 %>%
+  filter(!iucn_sid %in% id_match$iucn_sid)
+am_id_miss <- spp_am_ver1 %>%
+  filter(!am_sid %in% id_match$am_sid)
+
+name_match_base <- inner_join(iucn_id_miss, am_id_miss, by = 'sciname')
+# length(name_match_base$iucn_sid %>% unique()); length(name_match_base$am_sid %>% unique())
+### additional 463; 452 are unique, for total of 2332, which matches old method
+name_match_worms <- inner_join(iucn_id_miss %>%
+                                 filter(!is.na(sciname_accepted)) %>%
+                                 rename(iucn_sciname = sciname), 
+                               am_id_miss %>%
+                                 filter(!is.na(sciname_accepted)) %>%
+                                 rename(am_sciname = sciname), 
+                               by = 'sciname_accepted') %>%
+  filter(status.x != 'unaccepted' & status.y != 'unaccepted') ### both unaccepted? too much uncertainty - eliminate
+
+# length(name_match_worms$iucn_sid %>% unique()); length(name_match_worms$am_sid %>% unique())
+### additional 480; 467 IUCN are unique, and 464 AquaMaps, for total of 2332, which makes sense
+worms_no_basename <- name_match_worms %>%
+  filter(!iucn_sid %in% name_match_base$iucn_sid) %>%
+  select(iucn_sid, am_sid, sciname = sciname_accepted)
+### matched by WoRMS names but not by basic names from original databases
+
+basename_no_worms <- name_match_base %>%
+  filter(!am_sid %in% name_match_worms$am_sid) %>%
+  select(iucn_sid, am_sid, sciname)
+### these are all ones with no WoRMS match for either database, but matched by original names
+
+id_match_cleaned <- id_match %>%
+  select(iucn_sid, am_sid, sciname = iucn_sciname) %>%
+  distinct()
+
+species_id_lookup <- id_match_cleaned %>%
+  mutate(match = 'id') %>%
+  bind_rows(name_match_base %>%
+              select(iucn_sid, am_sid, sciname) %>%
+              mutate(match = 'basic name')) %>%
+  bind_rows(worms_no_basename %>%
+              mutate(match = 'worms')) %>%
+  distinct()
+
+### 2351 species on the list
+x <- species_id_lookup %>%
+  filter(iucn_sid %in% species_id_lookup$iucn_sid[duplicated(species_id_lookup$iucn_sid)])
+y <- species_id_lookup %>%
+  filter(am_sid %in% species_id_lookup$am_sid[duplicated(species_id_lookup$am_sid)])
+
+write_csv(species_id_lookup, file.path(dir_git, 'data/species_id_lookup.csv'))
+# length(species_id_lookup$iucn_sid %>% unique()); length(species_id_lookup$am_sid %>% unique())
+### 2335 iucn_sid, 2335 am_sid
 
@@
 
+species_id_lookup <- read_csv(file.path(dir_git, 'data/species_id_lookup.csv'))
+
 spp_am_v_iucn <- spp_all %>%
   select(-iucn_subpop) %>%
-  filter(!is.na(iucn_sid) & !is.na(am_sid)) %>%
+  filter(iucn_sid %in% species_id_lookup$iucn_sid & am_sid %in% species_id_lookup$am_sid) %>% #  & !is.na(am_sid)) %>%
   unique()
@@
 # length(unique(spp_am_v_iucn$sciname)); length(unique(spp_am_v_iucn$am_sid)); length(unique(spp_am_v_iucn$iucn_sid))
-### NOTE: 2332 unique scientific names, unique am_sid and unique iucn_sid.
+### NOTE: 23325 unique scientific names, unique am_sid and unique iucn_sid.
 ### Seasnakes are included in Reptiles (IUCN).  Two "mangroves" that are actually seagrasses.
@@
 sciname_dupes <- spp_am_v_iucn %>%
-  group_by(sciname) %>%
+  group_by(am_sid) %>%
   filter(n() > 1)
@@
 
-spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))
+spp_map_pairs <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
 
-# library(rgbif)
+library(rgbif)
 
 # system.time({
-#   x <- occ_search(scientificName = spp_all$sciname[1:100], fields = 'minimal', limit = 1)
+#   x <- occ_search(scientificName = spp_map_pairs$sciname, fields = 'minimal', limit = 1)
 #   ### seems linear - 10 names ~ 2.5 sec, 100 names ~ 25 sec, 1000 names, ~250 sec
@@
 # })
-# y <- sapply(x, function(x) x$meta$count)
+# y <- sapply(x, function(x) x$meta$count) %>%
+#   unlist()
 # gbif_occur <- data.frame(sciname = names(y), gbif_occ = y)
 # 
-# gbif_occur <- gbif_occur %>% 
-#   left_join(spp_all %>%
-#               select(iucn_sid, am_sid, sciname),
+# gbif_occur <- spp_map_pairs %>%
+#   select(iucn_sid, am_sid, sciname) %>%
+#   left_join(gbif_occur,
 #             by = 'sciname') %>%
-#   left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
+#   left_join(read_csv(file.path(dir_data_remote, 'aquamaps/d2015/csv/speciesoccursum.csv')) %>%
 #                        select(am_sid = speciesid, occurcells),
 #                      by = 'am_sid')
-
-# plot(ln_obs ~ ln_occurcells, data = gbif_occur)
-# plot(obs ~ occurcells, data = gbif_occur)
-# print(summary(lm(obs ~ occurcells, data = gbif_occur)))
-
+# 
+# # plot(gbif_occ ~ occurcells, data = gbif_occur)
+# # print(summary(lm(obs ~ occurcells, data = gbif_occur)))
+# 
 # write_csv(gbif_occur, file.path(dir_data, 'gbif_occurrence.csv'))
-gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence.csv'))
+# # gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence.csv'))
 
 # source('obis_count.R')
-# scinames <- spp_all$sciname
+# scinames <- spp_map_pairs$sciname
 # 
-# obis_occur <- obis_count(scinames)
-# obis_occur <- obis_count(scinames[26500:26600], verbose = TRUE)
+# obis_occur <- obis_count(scinames, verbose = TRUE)
+# # obis_occur <- obis_count(scinames[1:10], verbose = TRUE)
 # 
-# obis_occur <- obis_occur %>% 
-#   left_join(spp_all %>%
-#               select(iucn_sid, am_sid, sciname),
+# obis_occur <- spp_map_pairs %>%
+#   select(iucn_sid, am_sid, sciname) %>%
+#   left_join(obis_occur,
 #             by = 'sciname') %>%
-#   left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
+#   left_join(read_csv(file.path(dir_data_remote, 'aquamaps/d2015/csv/speciesoccursum.csv')) %>%
 #                        select(am_sid = speciesid, occurcells),
@@
 # write_csv(obis_occur, file.path(dir_data, 'obis_occurrence.csv'))
-obis_occur <- read_csv(file.path(dir_data, 'obis_occurrence.csv'))
+# # obis_occur <- read_csv(file.path(dir_data, 'obis_occurrence.csv'))
 

### 2017-04-21 data_explore/data_prep_ingest_iucn_shps.Rmd
initialize dummy branch

---
title: 'OHI: Species: ingest IUCN shapefiles'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohiprep/src/templates/ohi_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(rgdal)
library(maptools)
library(raster)
library(tidyverse)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/ohi',
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico/vAM_IUCN')
dir_data_remote <- file.path(dir_M, 'git-annex/globalprep/_raw_data')

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))

```

# Methods

## Ingest shapefiles from IUCN into half-degree cells

``` {r define_paths}

iucn_shp_dir <- file.path(dir_data_remote, 'iucn_spp', 'd2015', 'iucn_shp')

cache_dir <- (file.path(dir_anx, 'iucn_intersections'))

rast_file <- file.path(dir_anx, '../rgns/loiczid_raster.tif')
```

``` {r define_functions}

extract_from_shp <- function(shp, cache_dir, rast, 
                             fn_tag   = NULL,
                             reload   = FALSE) {
  ### will extract a shapefile to a raster, and save the .csv as a
  ### file in cache_dir with a matching name (e.g. BOTW.shp -> BOTW.csv).
  ### NOTE: any filtering of shapefiles should be done before passing them
  ### to this function; e.g. filter TERRESTRIAL_MAMMALS.shp to just those
  ### that occur in marine habitats.
  ### * shp must include .shp extension and must include full path name.

  if(!file.exists(shp)) {
    message('Shapefile ', shp, ' does not exist.  shp argument must contain ',
            'the full path name, including .shp extension.')

    return(data.frame(msg = 'pathname error'))
  }

  ### determine species group (the shapefile base name without extension)
  spp_gp <- basename(shp) %>% str_replace('.shp', '')

  cache_file <- file.path(cache_dir, 
                          paste0(spp_gp, 
                                 ifelse(is.null(fn_tag), '', fn_tag),
                                 '.csv'))

  ### if reload == FALSE, and the file exists, don't waste your friggin' time here, move on to next group.
  if(file.exists(cache_file) & reload == FALSE) {
    message(sprintf('IUCN <-< LOICZID lookup file already exists for species group %s;',
                    ' file location:\n  %s', spp_gp, cache_file))
    return(data.frame(msg = paste(cache_file, ' already exists, not reprocessed')))
  } else {
    ptm <- proc.time()
    fsize <- round(file.size(shp)/1e6, 2)
    message(sprintf('Reading species group shapefile %s, %.2f MB\n  %s', spp_gp, fsize, shp))

    ### Because the IUCN metadata states that shapefiles are unprojected lat-long with WGS84,
    ### use readShapePoly (rather than readOGR) and manually tell it the projection...
    spp_shp <- maptools::readShapePoly(fn = shp,
                                       proj4string = CRS('+init=epsg:4326'),
                                       delete_null_obj = TRUE)
    # spp_shp <- rgdal::readOGR(dsn = dirname(shp),
    #                           layer = basename(shp) %>% str_replace('.shp', ''),
    #                           stringsAsFactors = FALSE)
    message(sprintf('Elapsed read time: %.2f seconds', (proc.time() - ptm)[3]))

    ### standardize some variable names; this is mostly for BOTW
    names(spp_shp@data) <- tolower(names(spp_shp@data))
    names(spp_shp@data)[names(spp_shp@data) %in% c('id_no', 'sisid')] <- 'iucn_sid'
    names(spp_shp@data)[names(spp_shp@data) %in% c('binomial')]       <- 'sciname'
    names(spp_shp@data)[names(spp_shp@data) %in% c('presenc')]        <- 'presence'
    ### corals 1 (and others?) doesn't have subpop variable; add NAs
    if(!'subpop' %in% names(spp_shp@data))
      spp_shp@data$subpop <- NA

    message(sprintf('... processing %s species',
                    length(unique(spp_shp@data$iucn_sid))))

    message('Extracting polygons to LOICZID cells...')
    ptm <- proc.time()
    spp_shp_prop <- raster::extract(rast, spp_shp, 
                                    weights = TRUE, normalizeWeights = FALSE, 
                                    progress = 'text')
    message(sprintf('Elapsed process time: %.2f minutes', (proc.time() - ptm)[3]/60))

    ### combine sciname, iucn_sid, presence, and subpop code for a single unique identifier
    shp_id <- data.frame('sciname'  = spp_shp@data$sciname,
                         'iucn_sid' = spp_shp@data$iucn_sid,
                         'presence' = spp_shp@data$presence,
                         'subpop'   = spp_shp@data$subpop) %>%
      unite(name_id, sciname, iucn_sid, presence, subpop, sep = '_')

    names(spp_shp_prop) <- shp_id$name_id

    ### convert list to data frame.
    spp_shp_prop_df <- plyr::ldply(spp_shp_prop, rbind)
    spp_shp_prop_df <- spp_shp_prop_df %>%
      rename(name_id   = .id,
             LOICZID   = value,
             prop_area = weight) %>%
      separate(name_id, c('sciname', 'iucn_sid', 'presence', 'subpop'), sep = '_') %>%
      distinct()

    ### save .csv for this species group
    message(sprintf('%s: %s species maps, %s total cells in output file',
                    spp_gp, length(unique(spp_shp_prop_df$iucn_sid)),
                    nrow(spp_shp_prop_df)))
    message(sprintf('Writing IUCN<->LOICZID intersection file for %s to:\n  %s', spp_gp, cache_file))
    write_csv(spp_shp_prop_df, cache_file)
  }

  return(invisible(spp_shp_prop_df))
}

```

### Extract IUCN shapefiles to csvs.

All shapefiles within the IUCN shapefile folder are extracted against the LOICZID raster from AquaMaps.  The resulting dataframes are saved as a CSV for each species group (each shapefile).

NOTE: reload is disabled.  To re-extract a shapefile, simply delete the existing .csv extracted from .shp

### available shape files:

``` {r display_shps, results = 'asis'}

shp_files <- c(list.files(iucn_shp_dir,
                        pattern = '.shp$',
                        full.names = TRUE))

csv_files <- list.files(file.path(dir_anx, 'iucn_intersections'),
                        pattern = '.csv$') %>%
  str_replace('.csv$', '')

shp_files_to_do <- shp_files[!(basename(shp_files) %>% str_replace('.shp$', '') %in% csv_files)]

shp_info <- lapply(shp_files, file.info) %>%
  bind_rows() %>%
  mutate(shp_file = shp_files %>% str_replace('/home/shares/ohi/git-annex/globalprep/', ''),
         size_mb  = round(size / 1e6, 2)) %>%
  dplyr::select(shp_file, size_mb, mod_time = mtime, user = uname, group = grname) %>%
  mutate(to_process = basename(shp_file) %in% basename(shp_files_to_do))

DT::datatable(shp_info)

```

``` {r extract_shps_to_csv_series}

spp_list <- read_csv(file.path(dir_git, 'data/spp_iucn_maps_verified.csv'))

rast <- raster(rast_file)

reload <- FALSE

shps_done <- list.files(cache_dir) %>%
  str_replace('.csv', '.shp') %>%
  paste(collapse = '|') ### converts multiple strings into single regex pattern

if(shps_done == '') {
  shps_remain <- shp_files
} else {
  shps_remain <- shp_files[!str_detect(shp_files, shps_done)]
}

for (shp in shps_remain) {
  # shp <- shps_remain[1]
  shp_csv <- extract_from_shp (shp,
                               cache_dir = cache_dir,
                               rast = rast,
                               reload = reload)
}

```

``` {r check_that_all_files_processed}
### run this bit to check whether all files processed properly.  Errors
### seem to typically knock a few out of the running for some reason.
shps_done <- list.files(cache_dir) %>%
  str_replace('.csv', '.shp') %>%
  paste(collapse = '|') 
    ### converts multiple strings into single regex pattern so str_detect can
    ### test for patternA OR patternB OR patternC etc at one swoop
shps_remain <- shp_files[!str_detect(shp_files, shps_done)]

if (length(shps_remain > 0)) {
  message('Shapefiles not processed: \n  ', paste(shps_remain, collapse = '\n  '))
  message('... Reprocessing these files!')
  shp_df_list <- mclapply(shps_remain, 
                            FUN = extract_from_shp,
                                    cache_dir = cache_dir,
                                    rast = rast,
                                    reload = reload,
                          mc.cores = 12)

} else {
  message('All shapefiles processed! yay!')
}
```

### available CSV files:

``` {r display_csvs, results = 'asis'}

csv_files <- list.files(cache_dir, pattern = '.csv$', full.names = TRUE)

cat('All .csv files located in ', cache_dir, '\n')

csv_info <- lapply(csv_files, file.info) %>%
  bind_rows() %>%
  mutate(basename = basename(csv_files),
         size_mb  = round(size / 1e6, 2)) %>%
  dplyr::select(basename, size_mb, mod_time = mtime, user = uname, group = grname)

DT::datatable(csv_info)

```

## Combine all extracted CSVs

``` {r combine_all_extracted_csvs}

spp_list <- read_csv(file.path(dir_git, 'data/spp_iucn_maps_verified.csv'))

csv_list <- list.files(cache_dir, pattern = '.csv$', full.names = TRUE)

spp_cells_df <- lapply(csv_list, read_csv) %>%
  bind_rows()

names(spp_cells_df) <- tolower(names(spp_cells_df))

spp_missing <- spp_list %>%
  filter(!iucn_sid %in% spp_cells_df$iucn_sid) %>%
  distinct()
### none missing!

spp_extra <- spp_cells_df %>%
  filter(!iucn_sid %in% spp_list$iucn_sid) %>%
  dplyr::select(-loiczid, -prop_area, -presence) %>%
  distinct()
### no extras!

spp_cells_file <- file.path(dir_anx, 'int', 'iucn_spp_cells_d2015.csv')
write_csv(spp_cells_df, spp_cells_file)

spp_cell_file_info <- file.info(spp_cells_file) %>% 
  mutate(size_mb = size / 1e6 %>% round(2)) %>%
  dplyr::select(size_mb, mod_time = mtime, user = uname, group = grname)

```

File saved at: `r spp_cells_file`

`r knitr::kable(spp_cell_file_info)`

Extra species?  `r nrow(spp_extra)`

Missing species?  `r nrow(spp_missing)`

-----

### 2017-04-21 shiny_am_iucn/data_prep.Rmd
initialize dummy branch

@@
 ### copy and trim the list of overlapping species
+comnames <- read_csv(file.path(dir_shiny, 'data/sci_to_common.csv'))
+
 spp_list_quads <- read_csv(file.path(dir_git, 'data', 'spp_list_quads.csv')) %>%
@@
          n_spp, quad) %>%
-  mutate(reviewed = ifelse(!is.na(reviewed), TRUE, FALSE))
+  left_join(comnames, by = 'sciname') %>%
+  mutate(reviewed = ifelse(!is.na(reviewed), TRUE, FALSE),
+         name = ifelse(is.na(common) | common == 'Null', sciname, common)) %>%
+  select(-fbname)
 write_csv(spp_list_quads, file.path(dir_shiny, 'data', 'spp_list_quads_app.csv'))
@@
 
+comnames <- read_csv(file.path(dir_shiny, 'data', 'spp_list_quads_app.csv')) %>%
+  dplyr::select(iucn_sid, sciname, name)
+
 ### recalc areas and overlaps
@@
   select(iucn_sid, am_sid, sciname, dist_align, area_ratio, method) %>%
-  distinct()
+  distinct() %>%
+  left_join(comnames, by = c('iucn_sid', 'sciname'))
 
@@
 coral_spp_list  <- spp_list %>%
-  select(iucn_sid, sciname) %>%
+  select(iucn_sid, sciname, name) %>%
   filter(iucn_sid %in% coral_cells_app$iucn_sid) %>%
   distinct()
-write_csv(spp_coral_maps, file.path(dir_shiny, 'data', 'coral_spp_list.csv'))
+write_csv(coral_spp_list, file.path(dir_shiny, 'data', 'coral_spp_list.csv'))
 

### 2017-04-21 clip_fao/point_observations.R

@@ -1,0 +1,160 @@
+library(tidyr)
+library(dplyr)
+library(stringr)
+library(readr)
+library(rgbif)
+
+
+### support functions for obis_count()
+library(httr)
+library(jsonlite)
+
+obis_url <- function() {
+  getOption("robis_url", "http://api.iobis.org/")
+}
+
+max_characters <- function() {
+  getOption("robis_max_characters", 50000)
+}
+
+http_request <- function(method, path, query) {
+  if (method == "GET") {
+    httr::GET(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
+              path = path, query = query)
+  } else if (method == "POST") {
+    httr::POST(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
+               path = path, body = query)
+  }
+}
+
+obis_occurrence <- function(scientificname,  verbose = FALSE) {
+  
+  offset <- 0
+  i <- 1
+  lastpage <- FALSE
+  total <- 0
+  datalist <- list()
+  
+  t <- proc.time()
+  
+  while (!lastpage) {
+    query <- list(scientificname = scientificname)
+    
+    result <- http_request("GET", "occurrence", query)
+    
+    stop_for_status(result)
+    text <- content(result, "text", encoding="UTF-8")
+    res <- fromJSON(text, simplifyVector=TRUE)
+    
+    if(!is.null(res$message)) {
+      lastpage = TRUE
+      warning(res$message)
+    } else {
+      limit <- res$limit
+      offset <- offset + limit
+      lastpage <- res$lastpage
+      if(res$count > 0) {
+        datalist[[i]] <- res$results
+        total <- total + nrow(res$results)
+        if(verbose) message("Retrieved ", total, " records of ", res$count, " (", floor(total/res$count*100),"%)", sep="")
+        i <- i + 1
+      }
+    }
+  }
+
+  if (verbose) {
+    message("Total time:", (proc.time() - t)[["elapsed"]], "seconds")
+  }
+  
+  data <- bind_rows(datalist)
+  
+  return(data)
+}
+
+### spp_ex copied from fao_clip_pac script - these are four species with interesting maps
+spp_ex <- c('Hoplichthys regani' = 'Fis-34036', 
[…]

### 2017-04-21 data_explore/data_explore_fxns.R

@@ -1,0 +1,317 @@
+### data_explore_fxns.R
+
+library(ggplot2)
+
+### generic theme for all plots
+theme_set(theme_bw())
+ggtheme_plot <- theme_update(axis.ticks = element_blank(),
+                      text = element_text(family = 'Helvetica', color = 'gray30', size = 9),
+                      plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
+                      panel.background = element_blank(),
+                      legend.position = 'right',
+                      panel.border     = element_blank(),
+                      panel.grid.minor = element_blank(), 
+                      # panel.grid.major = element_line(colour = 'grey90', size = .25),
+                      panel.grid.major = element_blank(),
+                      legend.key = element_rect(colour = NA, fill = NA),
+                      axis.line = element_blank()) # element_line(colour = "grey30", size = .5))
+
+
+
+#################################################################=
+### Species Map Function ###
+# This function takes a dataframe of loiczid per species ID, and a target species ID 
+### (as single or vector).  Return a df of cells ready to be rasterized.  
+### If AquaMaps, returned df includes aquamaps probability; if IUCN, returned
+### df includes proportional area.
+get_spp_map <- function(sid, spp_cells) { 
+  data_type <- ifelse('am_sid' %in% names(spp_cells), 'am', 'iucn')
+  
+  
+  
+  if(data_type == 'am') {
+    ### if AquaMaps, process it thusly:
+    spp_map <- spp_cells %>%
+      filter(am_sid %in% sid) %>%
+      rename(am_prob = prob)
+    if(nrow(spp_map) == 0)
+      message('no AquaMaps cells found')
+  } else {
+    ### if IUCN, process it thusly:
+    if(!'prop_area' %in% names(spp_cells)) {
+      spp_cells <- spp_cells %>%
+        mutate(prop_area = 1)
+    }
+    spp_map <- spp_cells %>%
+      filter(iucn_sid %in% sid)
+    if(nrow(spp_map) == 0) {
+      message('no IUCN cells found')
+      spp_map <- spp_map %>%
+        as.data.frame() %>%
+        select(iucn_sid, loiczid) %>%
[…]

### 2017-04-21 data_explore/map_am_v_iucn.R

@@ -1,0 +1,173 @@
+#load data for species map function, function, and plotting a species
+
+#JAfflerbach
+#9.11.2014
+
+## SETUP ##
+
+#libraries
+
+library(data.table)
+library(readr)
+library(raster)
+library(rgdal)
+library(dplyr)
+library(RColorBrewer)
+library(maps)
+
+
+#increase R memory size to deal with reading in the large files
+
+memory.limit(size=200000000)
+
+
+dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
+           'Darwin'  = '/Volumes/data_edit',
+           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]
+
+dir_anx      <- file.path(dir_N, 'git-annex/globalprep/SpeciesDiversity')
+dir_aquamaps <- file.path(dir_anx, 'raw/aquamaps_2014')
+dir_iucn     <- file.path(dir_anx, 'iucn_intersections')
+dir_cells    <- file.path(dir_N, 'git-annex/Global/NCEAS-SpatialFishCatch_v2014/raw/ohi_spp/data')
+
+#read in data
+#fread comes from data.table package and is much more efficient at reading in the larger .csv files
+
+spp_list <- read.csv('spp_am_v_iucn.csv', stringsAsFactors = FALSE)
+### list of species with both IUCN and Aquamaps spatial data
+spp_list_ltd <- spp_list %>% filter(!is.na(parent_sid) | !is.na(subpop_sid))
+### truncated list
+
+# am_spp   <- fread(input = file.path(dir_aquamaps, 'tables/ohi_speciesoccursum.csv'), header = TRUE) %>%
+#               mutate(sciname = paste(Genus, Species, sep = ' '))
+
+
+### get cell ID lookup table
+# cells    <- read.csv(file.path(dir_cells, 'cells.csv')) %>%
+# cells <- cells %>%
+#   dplyr::select(cid, csq = csquarecod, loiczid, faoaream)
+# 
+# cells2 <- read.csv(file.path(dir_anx, 'raw/aquamaps_2014/tables/hcaf.csv'))
+# cells2 <- cells2 %>% 
+#   dplyr::select(csq = CsquareCode, loiczid = LOICZID, faoaream = FAOAreaM)
+# write_csv(cells,  file.path(dir_anx, 'explore_am_v_iucn/cells.csv'))
[…]

### 2017-04-21 data_explore/obis_count.R

@@ -1,0 +1,96 @@
+### support functions for obis_count()
+library(httr)
+
+obis_url <- function() {
+  getOption("robis_url", "http://api.iobis.org/")
+}
+
+max_characters <- function() {
+  getOption("robis_max_characters", 50000)
+}
+
+handle_date <- function(date) {
+  if(!is.null(date) && class(date) == "Date") {
+    as.character(date)
+  } else {
+    date
+  }
+}
+
+handle_vector <- function(x) {
+  if(!is.null(x)) {
+    paste0(x, collapse = ",")
+  } else {
+    x
+  }
+}
+
+log_request <- function(result) {
+  cat("\n", paste(result$request$method, result$request$url, result$status_code, result$headers$age, result$times[["total"]]), "\n", sep="")
+}
+
+log_progress <- function(total, count) {
+  cat("\rRetrieved ", total, " records of ", count, " (", floor(total/count*100),"%)", sep="")
+}
+
+http_request <- function(method, path, query) {
+  if (method == "GET") {
+    httr::GET(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
+        path = path, query = query)
+  } else if (method == "POST") {
+    httr::POST(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
+         path = path, body = query)
+  }
+}
+
+##########################################################################
+
+obis_count <- function(scinames, verbose = FALSE) {
+
+  offset <- 0
+  i <- 0
+  occ_list <- vector('list', length = length(scinames)) ### initialize occurrence count list
+  ptm_total <- proc.time()
+ 
+  for (sciname in scinames) {
+    i <- i + 1
+    lastpage <- FALSE
+    ptm <- proc.time()
+    
+    query <- list(scientificname = sciname,
+                  offset = format(offset, scientific = FALSE))
+    
+    result <- http_request("GET", "occurrence", query)
+    
+    stop_for_status(result)
+    res <- content(result)
+    
+    if(!is.null(res$message)) {
+      lastpage = TRUE
+      warning(res$message)
+      occ_count <- NA
+    } else {
+      occ_count <- res$count
+    }
+    tmp_df <- data.frame('sciname' = sciname, 'obis_occ' = occ_count, stringsAsFactors = FALSE)
+    
+    if(verbose) {
+      ptm <- round((proc.time() - ptm)[["elapsed"]], 2)
[…]

### 2017-04-21 figures/si_figs_JA/supp_figs_JA.R

@@ -1,0 +1,228 @@
+library(raster)
+# library(maps)
+library(maptools)
+library(tmap)
+library(rgdal)
+library(readr)
+
+source('~/github/ohiprep/src/R/common.R')
+
+
+dir_git <- '~/github/IUCN-AquaMaps'
+
+data("wrld_simpl", package = "maptools")
+data(World)
+
+ohi  <- readOGR(dsn = file.path(dir_M, 'git-annex/globalprep/spatial/d2014/data'), layer = 'regions_gcs')
+land <- ohi %>%
+  subset(rgn_typ == 'land')
+
+sp <- "Thunnus alalunga"
+setwd('shiny_am_iucn') ### setwd because there are some relative paths within server_fxns.R
+source('server_fxns.R')
+setwd('../')
+
+spp_list <- read_csv('shiny_am_iucn/data/spp_list.csv')
+
+# get rasters for tuna from IUCN and AM
+
+df     <- get_spp_map_df(sp)
+r_am   <- get_rast(df, type = 'am')
+r_iucn <- get_rast(df, type = 'iucn')
+
+plot(r_iucn)
+#playing with grids
+grid(nx = 720, ny = 360, col = 'lightgray')
+
+#crop so we can zoom in and show grid cells
+crop_ext    <- c(-80, -50,30,50)
+r_iucn_crop <- crop(r_iucn, crop_ext)
+r_am_crop   <- crop(r_am, crop_ext)
+land_crop   <- crop(land, crop_ext) %>%
+  subset(rgn_nam %in% c('Canada', 'United States'))
+plot(land_crop)
+
+
+basemap_simpl <- crop(wrld_simpl, crop_ext)
+basemap       <- crop(World, crop_ext)
+
+#pretty up with tmap
+tuna_iucn_raster_grid <- tm_shape(r_iucn_crop) +
+  tm_raster(palette = 'aquamarine3',
+            colorNA = NULL,
+            alpha = .8,
+            legend.show = FALSE) +
+  tm_layout(basemaps = "Esri.WorldTopoMap", 
+            # title.position = 'TOP', 
+            legend.outside = TRUE, attr.outside = TRUE)+
+  tm_grid(n.x = 60, n.y = 40, projection = 'longlat', col = 'gray', 
+          labels.inside.frame = FALSE,labels.size=0)+
+  tm_shape(land_crop) +
+  tm_polygons() +
+  tm_fill(col = 'gray')
+
[…]

### 2017-04-21 figures/si_figs_JA/supp_figs_JA1.R

@@ -1,0 +1,229 @@
+library(raster)
+library(rgdal)
+# library(maps)
+library(maptools)
+library(tmap)
+library(readr)
+
+source('~/github/ohiprep/src/R/common.R')
+
+
+dir_git <- '~/github/IUCN-AquaMaps'
+
+data("wrld_simpl", package = "maptools")
+data(World)
+
+
+sp <- "Thunnus alalunga"
+setwd('shiny_am_iucn') ### setwd because there are some relative paths within server_fxns.R
+source('server_fxns.R')
+setwd('../')
+
+spp_list <- read_csv('shiny_am_iucn/data/spp_list.csv')
+
+### get rasters for tuna from IUCN and AM
+
+df     <- get_spp_map_df(sp)
+tuna_rast_am   <- get_rast(df, type = 'am')
+tuna_rast_iucn <- get_rast(df, type = 'iucn')
+
+
+### crop so we can zoom in and show grid cells
+crop_ext    <- c(-80, -50, 30, 50)
+tuna_rast_iucn_crop <- crop(tuna_rast_iucn, crop_ext)
+tuna_rast_am_crop   <- crop(tuna_rast_am, crop_ext)
+
+# ohi  <- readOGR(dsn = file.path(dir_M, 'git-annex/globalprep/spatial/d2014/data'), layer = 'regions_gcs')
+# land <- ohi %>%
+#   subset(rgn_typ == 'land')
+# land_crop   <- crop(land, crop_ext) %>%
+#   subset(rgn_nam %in% c('Canada', 'United States'))
+# writeOGR(land_crop, dsn = 'figures/si_figs_JA', layer = 'land_crop', driver = 'ESRI Shapefile')
+land_crop      <- readOGR(dsn = 'figures/si_figs_JA', layer = 'land_crop')
+
+# tuna_poly <- readOGR(dsn = file.path(dir_M, 'git-annex/globalprep/_raw_data/iucn_spp/d2015/iucn_shp'),
+#                      layer = 'TUNAS_BILLFISHES')
+# tuna_poly <- tuna_poly %>% 
+#   subset(binomial == 'Thunnus alalunga')
+# tuna_poly_crop <- crop(tuna_poly, crop_ext)
+# extent(tuna_poly_crop) <- crop_ext
+# writeOGR(tuna_poly, dsn = 'figures/si_figs_JA', layer = 'tuna_poly', driver = 'ESRI Shapefile')
[…]

### 2017-04-21 shiny_am_iucn/server.R

@@ -1,0 +1,84 @@
+### server.R
+
+source('server_fxns.R')
+
+server <- shinyServer(function(input, output, session) {
+  
+  ### Species Map Function ###
+  spp_map   <- reactiveValues()
+  coral_map <- reactiveValues()
+  ### This function takes a single species scientific name as input, then grabs 
+  ### all occurrence cells and associated probability per cell
+  
+  ### for Species Map tab: upon selecting a group, update species list choices
+  observeEvent(input$spp_group, 
+    {
+      message('Observed change in spp_group; updating select input')
+      spp_choices <- spp_list %>%
+        filter(spp_group_text == input$spp_group) %>%
+        distinct() %>%
+        .$name %>%
+        sort()
+      updateSelectInput(session, inputId = "species",
+                        choices = spp_choices)
+    }
+  )
+  
+  ### For Species map tab:  upon selecting a species, get map cells
+  observeEvent(
+    {input$species}, 
+    {
+      message('observed change in input$species; getting spp_map dataframe')
+      spp_map$df <- get_spp_map_df(input$species)
+      print(head(spp_map$df))
+    }
+  )
+  
+  ### For Species map tab:  upon a change in the spp_map$df OR the input$show_maps,
+  ### create the map to show AM, IUCN, or Both
+  create_map <- observeEvent(
+    {spp_map$df; input$show_maps}, 
+      ### triggered by change in spp_map$am_rast (which may also include 
+      ### change in iucn_rast) or a change in the selected maps to display
+    {
+      message('observed change in spp_map$df or input$show_maps; creating map')
+      message('input$show_maps = ', input$show_maps, '; input$species = ', input$species)
+      map_rast <- get_rast(spp_map$df, type = input$show_maps)
+      map_obj  <- assemble_map_tmap(map_rast, spp = input$species)
+      
+      spp_map$map <- map_obj
+    }
+  )
+  
+  output$compare_map <- renderPlot({
+    spp_map$map
+  }) 
+  
+  ### For Map Alignment tab
+  output$quad_plot <- renderPlotly({
+    create_quadplot(input$taxa_quad, input$expert_rev)
+  })
+  
+  output$barchart <- renderPlot({
+    create_barchart(input$expert_rev)
+  })
+  
+  output$mini_quad <- renderPlot({
+    create_miniquad(input$species)
+  })
+
+  ### For Coral Depth tab
[…]

### 2017-04-21 shiny_am_iucn/server_fxns.R

@@ -1,0 +1,519 @@
+### server_fxns.R
+library(raster)
+library(stringr)
+# library(maps)
+library(tmap)
+# data(World)
+library(RColorBrewer)
+
+### read data for raster (used to plot species)
+message('Reading in raster info...')
+loiczid_raster <- raster('data/loiczid_raster.tif') %>%
+  setNames('loiczid')
+
+display_extents <- extent(-180, 180, -84, 85)
+
+land <- rgdal::readOGR(dsn   = 'data/ne_110m_land', 
+                       layer = 'ne_110m_land') %>%
+  crop(display_extents)
+
+bathy_rast <- subs(loiczid_raster, 
+                   read_csv('data/bathy_cells.csv'),
+                   by = 'loiczid', which = 'deep') %>%
+  crop(display_extents)
+
+fao_rgn <- rgdal::readOGR(dsn   = 'data/fao_rgns', 
+                          layer = 'fao_rgns_app') %>%
+  crop(display_extents)
+
+# create a blank ggplot theme
+ggtheme_basic <- function(textsize = 10) {
+  theme_bw() +
+  theme_update(panel.grid.minor = element_blank(),
+          panel.grid.major = element_blank(),
+          panel.background = element_blank(),
+          plot.background = element_blank(),
+          panel.border = element_blank(),
+          axis.line = element_blank(),
+          axis.ticks = element_blank(),
+          text = element_text(size = textsize),
+          plot.title = element_text(size = textsize * 1.5))
+}
+
+### read in spp cell files
+### this may be slowing down the initial display of the app... load elsewhere?
+iucn_spp_cells <- read_csv('data/iucn_cells_app.csv', col_types = 'dd')
+am_spp_cells   <- read_csv('data/am_cells_app.csv',   col_types = 'cd')
+
+### read in quad plot files and establish means
+quad_list <- read_csv('data/spp_list_quads_app.csv')
+
+area_align_mean <- mean(quad_list$area_ratio, na.rm = TRUE)
+dist_align_mean <- mean(quad_list$dist_align, na.rm = TRUE)
+
+### read in coral species pre- and post-clipping
+spp_coral_align <- read_csv('data/coral_align_app.csv')
+spp_coral_cells <- read_csv(file.path('data', 'coral_cells_app.csv'))
+
+quad_gp_list <- read_csv('data/spp_gp_quads_app.csv') %>%
+  mutate(expert = FALSE) %>%
[…]

### 2017-04-21 shiny_am_iucn/ui.R

@@ -1,0 +1,134 @@
+#ui.r
+
+source('ui_setup.R')
+
+ui <- navbarPage(title = '',
+                 theme = shinytheme('cerulean'),
+                 
+  tabPanel('Abstract',
+    sidebarPanel(
+      includeMarkdown('pages/article_info.md')
+    ),
+    mainPanel(
+      includeMarkdown('pages/abstract.md')
+    )
+  ),
+  
+  tabPanel('Figures',
+    sidebarPanel(
+      includeMarkdown('pages/article_info.md')
+    ),
+    mainPanel(
+      h4('Figures from manuscript'),
+      h5('Figure 1:'),
+      includeMarkdown('pages/fig1.md'),
+      hr(),
+      h5('Figure 2:'),
+      includeMarkdown('pages/fig2.md'),
+      hr(),
+      h5('Figure 3:'),
+      includeMarkdown('pages/fig3.md'),
+      hr(),
+      h5('Figure 4:'),
+      includeMarkdown('pages/fig4.md'),
+      hr(),
+      h5('Figure 5:'),
+      includeMarkdown('pages/fig5.md')
+    )
+  ),
+  
+  tabPanel('Map alignment',
+    sidebarPanel(
+      includeMarkdown('pages/align_tab_side1.md'),
+      selectInput('taxa_quad', 'Choose a taxon:',
+                  choices = c('all', unique(spp_list$spp_group_text) %>% sort()),
+                  selected = 'all'),
+      radioButtons('expert_rev', label = 'AquaMaps review status',
+                   choices = list('All'             = 'all',
+                                  'Expert reviewed' = 'expert'),
+                   selected = 'all'),
+      includeMarkdown('pages/footer_sidebar.md')
+    ),
+    mainPanel(
+      includeMarkdown('pages/align_tab_main1.md'),
+      plotlyOutput('quad_plot', height = '300px'),
+      hr(),
+      includeMarkdown('pages/align_tab_main2.md'),
+      plotOutput('barchart', height = '300px')
+    )
+  ),
+  
+  tabPanel('Species maps',
+    sidebarPanel(
+      includeMarkdown('pages/map_tab_side1.md'),
+      selectInput('spp_group', 'Select a taxonomic group:',
+                  choices = unique(spp_list$spp_group_text) %>% sort()),
+      selectInput('species', 'Then select a species:',
+                  choices = unique(spp_list$name) %>% sort()),
[…]

### 2017-04-21 case_study_mpa/model_wdpa2014_i-iv.py
updating app with ohirepos commit 0.1.0

@@ -1,100 +1,0 @@
-# File: ohiprep/globalprep/WDPA_MPA/v2015/model.py
-#
-# Modified March 2015 by Casey O'Hara
-#
-# Filters by 
-# * "STATUS" == 'Designated':  includes only officially designated areas, not
-#   areas still in planning or proposal phases.
-# * "MANG_PLAN" NOT LIKE 'Non-MPA%': eliminates all areas specifically designated
-#   as Non-MPA Programmatic Management Plans (code only used for U.S.).  Eliminates
-#   groundfish management areas, '25 mph manatee' zones, and the like.
-
-import arcpy, os, subprocess, csv, sys, socket
-from arcpy.sa import *
-
-# configuration based on machine name
-dirs =  {'git'    :'C:/Users/ohara/Documents/github/IUCN-AquaMaps', 
-         'neptune':'N:',
-         'tmp'    :'C:/Users/ohara/tmp',
-         }
-print 'Setting pathnames...'
-print dirs
-
-# paths
-prod     = 'mpa_analysis/WDPA_MPA'          # name of product
-dir_git  = '%s/%s' % (dirs['git'], prod)    # local github directory inside IUCN-AquaMaps
-dir_tmp  = '%s/%s' % (dirs['tmp'], prod)    # local temp directory
-dir_anx  = '%s/git-annex/Global/WDPA-MPA_v2014' % (dirs['neptune']) # git annex directory on neptune
-gdb      = '%s/geodb.gdb' % dir_tmp         # file geodatabase
-print 'dir_git = ' + dir_git
-print 'dir_tmp = ' + dir_tmp
-print 'dir_anx = ' + dir_anx
-
-# inputs
-poly_wdpa  = '%s/raw/WDPA_Apr2014_Public1/WDPA_Apr2014.gdb/WDPA_poly_Apr2014' % dir_anx
-dir_rgn    = '%s/git-annex/globalprep/spatial/v2015/data' % dirs['neptune']
-
-msk = '%s/rgn_marine.tif' % dir_git   # .tif of 500 m marine raster? will be huge - put on git-annex
-
-print 'input poly_wdpa = ' + poly_wdpa
-print 'input dir_rgn = ' + dir_rgn
-print 'input rgn_off_3nm = ' + rgn_off_3nm
-print 'input rgn_inl_1km = ' + rgn_inl_1km
-print 'input msk = ' + msk
-
-
-# outputs
[…]

### 2017-04-21 clip_fao/point_observations.R

@@ -1,160 +1,0 @@
-library(tidyr)
-library(dplyr)
-library(stringr)
-library(readr)
-library(rgbif)
-
-
-### support functions for obis_count()
-library(httr)
-library(jsonlite)
-
-obis_url <- function() {
-  getOption("robis_url", "http://api.iobis.org/")
-}
-
-max_characters <- function() {
-  getOption("robis_max_characters", 50000)
-}
-
-http_request <- function(method, path, query) {
-  if (method == "GET") {
-    httr::GET(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
-              path = path, query = query)
-  } else if (method == "POST") {
-    httr::POST(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
-               path = path, body = query)
-  }
-}
-
-obis_occurrence <- function(scientificname,  verbose = FALSE) {
-  
-  offset <- 0
-  i <- 1
-  lastpage <- FALSE
-  total <- 0
-  datalist <- list()
-  
-  t <- proc.time()
-  
-  while (!lastpage) {
-    query <- list(scientificname = scientificname)
-    
-    result <- http_request("GET", "occurrence", query)
-    
-    stop_for_status(result)
-    text <- content(result, "text", encoding="UTF-8")
-    res <- fromJSON(text, simplifyVector=TRUE)
-    
-    if(!is.null(res$message)) {
-      lastpage = TRUE
-      warning(res$message)
-    } else {
-      limit <- res$limit
-      offset <- offset + limit
-      lastpage <- res$lastpage
-      if(res$count > 0) {
-        datalist[[i]] <- res$results
-        total <- total + nrow(res$results)
-        if(verbose) message("Retrieved ", total, " records of ", res$count, " (", floor(total/res$count*100),"%)", sep="")
-        i <- i + 1
-      }
-    }
-  }
-
-  if (verbose) {
-    message("Total time:", (proc.time() - t)[["elapsed"]], "seconds")
-  }
-  
-  data <- bind_rows(datalist)
-  
-  return(data)
-}
-
-### spp_ex copied from fao_clip_pac script - these are four species with interesting maps
-spp_ex <- c('Hoplichthys regani' = 'Fis-34036', 
[…]

### 2017-04-21 data_explore/data_explore_fxns.R

@@ -1,317 +1,0 @@
-### data_explore_fxns.R
-
-library(ggplot2)
-
-### generic theme for all plots
-theme_set(theme_bw())
-ggtheme_plot <- theme_update(axis.ticks = element_blank(),
-                      text = element_text(family = 'Helvetica', color = 'gray30', size = 9),
-                      plot.title = element_text(size = rel(1.25), hjust = 0, face = 'bold'),
-                      panel.background = element_blank(),
-                      legend.position = 'right',
-                      panel.border     = element_blank(),
-                      panel.grid.minor = element_blank(), 
-                      # panel.grid.major = element_line(colour = 'grey90', size = .25),
-                      panel.grid.major = element_blank(),
-                      legend.key = element_rect(colour = NA, fill = NA),
-                      axis.line = element_blank()) # element_line(colour = "grey30", size = .5))
-
-
-
-#################################################################=
-### Species Map Function ###
-# This function takes a dataframe of loiczid per species ID, and a target species ID 
-### (as single or vector).  Return a df of cells ready to be rasterized.  
-### If AquaMaps, returned df includes aquamaps probability; if IUCN, returned
-### df includes proportional area.
-get_spp_map <- function(sid, spp_cells) { 
-  data_type <- ifelse('am_sid' %in% names(spp_cells), 'am', 'iucn')
-  
-  
-  
-  if(data_type == 'am') {
-    ### if AquaMaps, process it thusly:
-    spp_map <- spp_cells %>%
-      filter(am_sid %in% sid) %>%
-      rename(am_prob = prob)
-    if(nrow(spp_map) == 0)
-      message('no AquaMaps cells found')
-  } else {
-    ### if IUCN, process it thusly:
-    if(!'prop_area' %in% names(spp_cells)) {
-      spp_cells <- spp_cells %>%
-        mutate(prop_area = 1)
-    }
-    spp_map <- spp_cells %>%
-      filter(iucn_sid %in% sid)
-    if(nrow(spp_map) == 0) {
-      message('no IUCN cells found')
-      spp_map <- spp_map %>%
-        as.data.frame() %>%
-        select(iucn_sid, loiczid) %>%
[…]

### 2017-04-21 data_explore/obis_count.R

@@ -1,96 +1,0 @@
-### support functions for obis_count()
-library(httr)
-
-obis_url <- function() {
-  getOption("robis_url", "http://api.iobis.org/")
-}
-
-max_characters <- function() {
-  getOption("robis_max_characters", 50000)
-}
-
-handle_date <- function(date) {
-  if(!is.null(date) && class(date) == "Date") {
-    as.character(date)
-  } else {
-    date
-  }
-}
-
-handle_vector <- function(x) {
-  if(!is.null(x)) {
-    paste0(x, collapse = ",")
-  } else {
-    x
-  }
-}
-
-log_request <- function(result) {
-  cat("\n", paste(result$request$method, result$request$url, result$status_code, result$headers$age, result$times[["total"]]), "\n", sep="")
-}
-
-log_progress <- function(total, count) {
-  cat("\rRetrieved ", total, " records of ", count, " (", floor(total/count*100),"%)", sep="")
-}
-
-http_request <- function(method, path, query) {
-  if (method == "GET") {
-    httr::GET(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
-        path = path, query = query)
-  } else if (method == "POST") {
-    httr::POST(obis_url(), user_agent("robis - https://github.com/iobis/robis"),
-         path = path, body = query)
-  }
-}
-
-##########################################################################
-
-obis_count <- function(scinames, verbose = FALSE) {
-
-  offset <- 0
-  i <- 0
-  occ_list <- vector('list', length = length(scinames)) ### initialize occurrence count list
-  ptm_total <- proc.time()
- 
-  for (sciname in scinames) {
-    i <- i + 1
-    lastpage <- FALSE
-    ptm <- proc.time()
-    
-    query <- list(scientificname = sciname,
-                  offset = format(offset, scientific = FALSE))
-    
-    result <- http_request("GET", "occurrence", query)
-    
-    stop_for_status(result)
-    res <- content(result)
-    
-    if(!is.null(res$message)) {
-      lastpage = TRUE
-      warning(res$message)
-      occ_count <- NA
-    } else {
-      occ_count <- res$count
-    }
-    tmp_df <- data.frame('sciname' = sciname, 'obis_occ' = occ_count, stringsAsFactors = FALSE)
-    
-    if(verbose) {
-      ptm <- round((proc.time() - ptm)[["elapsed"]], 2)
[…]

### 2017-04-21 figures/si_figs_JA/supp_figs_JA.R

@@ -1,228 +1,0 @@
-library(raster)
-# library(maps)
-library(maptools)
-library(tmap)
-library(rgdal)
-library(readr)
-
-source('~/github/ohiprep/src/R/common.R')
-
-
-dir_git <- '~/github/IUCN-AquaMaps'
-
-data("wrld_simpl", package = "maptools")
-data(World)
-
-ohi  <- readOGR(dsn = file.path(dir_M, 'git-annex/globalprep/spatial/d2014/data'), layer = 'regions_gcs')
-land <- ohi %>%
-  subset(rgn_typ == 'land')
-
-sp <- "Thunnus alalunga"
-setwd('shiny_am_iucn') ### setwd because there are some relative paths within server_fxns.R
-source('server_fxns.R')
-setwd('../')
-
-spp_list <- read_csv('shiny_am_iucn/data/spp_list.csv')
-
-# get rasters for tuna from IUCN and AM
-
-df     <- get_spp_map_df(sp)
-r_am   <- get_rast(df, type = 'am')
-r_iucn <- get_rast(df, type = 'iucn')
-
-plot(r_iucn)
-#playing with grids
-grid(nx = 720, ny = 360, col = 'lightgray')
-
-#crop so we can zoom in and show grid cells
-crop_ext    <- c(-80, -50,30,50)
-r_iucn_crop <- crop(r_iucn, crop_ext)
-r_am_crop   <- crop(r_am, crop_ext)
-land_crop   <- crop(land, crop_ext) %>%
-  subset(rgn_nam %in% c('Canada', 'United States'))
-plot(land_crop)
-
-
-basemap_simpl <- crop(wrld_simpl, crop_ext)
-basemap       <- crop(World, crop_ext)
-
-#pretty up with tmap
-tuna_iucn_raster_grid <- tm_shape(r_iucn_crop) +
-  tm_raster(palette = 'aquamarine3',
-            colorNA = NULL,
-            alpha = .8,
-            legend.show = FALSE) +
-  tm_layout(basemaps = "Esri.WorldTopoMap", 
-            # title.position = 'TOP', 
-            legend.outside = TRUE, attr.outside = TRUE)+
-  tm_grid(n.x = 60, n.y = 40, projection = 'longlat', col = 'gray', 
-          labels.inside.frame = FALSE,labels.size=0)+
-  tm_shape(land_crop) +
-  tm_polygons() +
-  tm_fill(col = 'gray')
-
[…]

### 2017-04-21 figures/si_figs_JA/supp_figs_JA1.R

@@ -1,229 +1,0 @@
-library(raster)
-library(rgdal)
-# library(maps)
-library(maptools)
-library(tmap)
-library(readr)
-
-source('~/github/ohiprep/src/R/common.R')
-
-
-dir_git <- '~/github/IUCN-AquaMaps'
-
-data("wrld_simpl", package = "maptools")
-data(World)
-
-
-sp <- "Thunnus alalunga"
-setwd('shiny_am_iucn') ### setwd because there are some relative paths within server_fxns.R
-source('server_fxns.R')
-setwd('../')
-
-spp_list <- read_csv('shiny_am_iucn/data/spp_list.csv')
-
-### get rasters for tuna from IUCN and AM
-
-df     <- get_spp_map_df(sp)
-tuna_rast_am   <- get_rast(df, type = 'am')
-tuna_rast_iucn <- get_rast(df, type = 'iucn')
-
-
-### crop so we can zoom in and show grid cells
-crop_ext    <- c(-80, -50, 30, 50)
-tuna_rast_iucn_crop <- crop(tuna_rast_iucn, crop_ext)
-tuna_rast_am_crop   <- crop(tuna_rast_am, crop_ext)
-
-# ohi  <- readOGR(dsn = file.path(dir_M, 'git-annex/globalprep/spatial/d2014/data'), layer = 'regions_gcs')
-# land <- ohi %>%
-#   subset(rgn_typ == 'land')
-# land_crop   <- crop(land, crop_ext) %>%
-#   subset(rgn_nam %in% c('Canada', 'United States'))
-# writeOGR(land_crop, dsn = 'figures/si_figs_JA', layer = 'land_crop', driver = 'ESRI Shapefile')
-land_crop      <- readOGR(dsn = 'figures/si_figs_JA', layer = 'land_crop')
-
-# tuna_poly <- readOGR(dsn = file.path(dir_M, 'git-annex/globalprep/_raw_data/iucn_spp/d2015/iucn_shp'),
-#                      layer = 'TUNAS_BILLFISHES')
-# tuna_poly <- tuna_poly %>% 
-#   subset(binomial == 'Thunnus alalunga')
-# tuna_poly_crop <- crop(tuna_poly, crop_ext)
-# extent(tuna_poly_crop) <- crop_ext
-# writeOGR(tuna_poly, dsn = 'figures/si_figs_JA', layer = 'tuna_poly', driver = 'ESRI Shapefile')
[…]

### 2017-04-21 global.R

@@ -1,0 +1,461 @@
+# packages ----
+
+# refresh fitz.nceas.ucsb.edu after `sudo R` to gain write permissions on R library path:
+#   remove.packages(c('leaflet','ohicore','htmlwidgets','aster','sunburstR'))
+#   update.packages(ask=F)
+#   # then run up to and including `install_packages(pkgs_df)` below
+
+if (!'devtools'  %in% installed.packages()[,1]) install.packages('devtools')
+if (!'tidyverse' %in% installed.packages()[,1]) install.packages('tidyverse')
+# TODO: check that all packages below are actually needed, including specific tidyverse packages
+library(devtools , quietly=T)
+library(tidyverse, quietly=T)
+
+pkgs_df = tibble::tribble(
+  ~package   ,     ~location,                                                    ~install_args, ~version_min,
+  'dplyr',              'CRAN',                                                               '',           '',
+  'tidyr',              'CRAN',                                                               '',           '',
+  'readr',              'CRAN',                                                               '',           '',
+  'stringr',            'CRAN',                                                               '',           '',
+  'lubridate',          'CRAN',                                                               '',           '',
+  'rgdal',              'CRAN',                                                               '',           '',
+  'shiny',              'CRAN',                                                               '',           '',
+  'shinydashboard',     'CRAN',                                                               '',           '',
+  'htmltools',          'CRAN',                                                               '',           '',
+  'markdown',           'CRAN',                                                               '',           '',
+  'geojsonio',          'CRAN',                                                               '',           '',
+  'jsonlite',           'CRAN',                                                               '',           '',
[…]

### 2017-04-21 sci_to_common.R

@@ -1,49 +1,0 @@
-# creating a sciname -> common name lookup table. 
-
-## Note: there are 15 species that don't have common names. Many of which are corals, I've saved their sci names as common names
-## so we can just use the common name column
-
-library(readr)
-library(dplyr)
-
-#get the species information from aquamaps
-spp_occur <- read_csv('data/speciesoccursum.csv')%>%
-              mutate(sciname = paste(genus,species,sep=" "))%>%
-              select(sciname, fbname)
-
-#read in the data used for the quad map, which contains all sci names for our species of interest for the app
-data <- read_csv('shiny_am_iucn/data/spp_list_quads_app.csv')%>%
-        select(sciname)%>%
-        left_join(spp_occur)
-
-#get species that don't have a match
-nas <- filter(data,is.na(fbname))
-
-#create a dataframe manually that adds in these common names
-missing <- data_frame(sciname = c("Neomonachus schauinslandi", "Centropyge loricula", "Centropyge interruptus", 
-"Centropyge nigriocellus", 
-"Palinustus mossamicus", 
-"Justitia longimanus", 
-"Hydrophis major", 
-"Zebrasoma veliferum", 
-"Parastichopus californicus", 
-"Stereomastis sculpta", 
-"Symphodus trutta", 
-"Xyrichtys koteamea", 
-"Halichoeres raisneri", 
-"Xyrichtys pastellus", 
-"Pseudolabrus mortonii", 
-"Novaculichthys macrolepidotus", 
-"Xyrichtys halsteadi", 
-"Conus chiangi"),
-common = c("Hawaiian monk seal","Flame angelfish", "Japanese angelfish", "Blackspot angelfish", 
-           "Buffalo Blunthorn Lobster", "Gibbon Furrow Lobster", "Olive-headed sea snake", "Sailfin tang", "Giant california sea cucumber",
-           "Flatback lobster", "Emerald wrasse", "Red razorfish", "Cape wrasse", "Pastel razorfish", "Rosy parrotfish",
-           "Seagrass wrasse", "Halstead's razorfish", "Chiang's cone"))
-
-#combine the two dataframes and save
-out <- data%>%
-      mutate(common = ifelse(is.na(fbname),missing$common[match(sciname,missing$sciname)],fbname))%>%
-      mutate(common = ifelse(is.na(common),sciname,common))
-
-write.csv(out,file='shiny_am_iucn/data/sci_to_common.csv')

### 2017-04-21 server.R

@@ -1,0 +1,704 @@
+shinyServer(function(input, output, session) {
+
+  # hide compare ----
+  observeEvent(input$sidebarmenu, {
+    if (length(y$scenario_dirs) == 1){
+      shinyjs::hide(selector = "a[data-value=='compare']")
+    }
+  })
+
+
+  # get_scenario ----
+
+  # write remote git commit hash every 5 seconds to compare with remote
+  log_gitsha <- observe({
+    # Invalidate this observer every 2 seconds (2000 milliseconds)
+    invalidateLater(2000, session)
+
+    # write remote git commit sha
+    gh_write_remote(y$gh_slug, y$gh_branch_data, remote_sha_txt)
+  })
+
+  # monitor file for changes every 1 seconds (1000 milliseconds)
+  fileReaderData <- reactiveFileReader(1000, session, remote_sha_txt, readLines)
+
+  # get_scenario ----
+  get_scenario = function(env=.GlobalEnv){
+
+    switch(
+      input$sel_type,
+
+      # case: output
+      output = {
+        req(input$sel_output_goal)
+        req(input$sel_output_goal_dimension)
+
+        list(
+          data = env$scores %>%
+            filter(
+              goal      == input$sel_output_goal,
+              dimension == input$sel_output_goal_dimension) %>%
+            select(
+              rgn_id = region_id,
+              value  = score) %>%
+            select(rgn_id, value),
+          label = sprintf('%s - %s', input$sel_output_goal, input$sel_output_goal_dimension),
+          description = env$dims %>%
+            filter(dimension == input$sel_output_goal_dimension)  %>%
+            markdownToHTML(text = .$description, fragment.only=T)) },
+
+      # case: input
+      input = {
+        req(input$sel_input_target_layer)
+
+        env$fld_category = filter(env$layers, layer==input$sel_input_target_layer) %>% .$fld_category
+        env$fld_year     = filter(env$layers, layer==input$sel_input_target_layer) %>% .$fld_year
+
+        # get data
+        env$data = env$d_lyrs %>%
+          filter(layer == input$sel_input_target_layer) %>%
+          mutate(
+            rgn_id = fld_id_num,
+            value  = fld_val_num)
+
+        # if layer has category, filter
+        if (!is.na(fld_category)){
+          req(input$sel_input_target_layer_category)
+          env$data = env$data %>%
[…]

### 2017-04-21 shiny_am_iucn/server.R

@@ -1,84 +1,0 @@
-### server.R
-
-source('server_fxns.R')
-
-server <- shinyServer(function(input, output, session) {
-  
-  ### Species Map Function ###
-  spp_map   <- reactiveValues()
-  coral_map <- reactiveValues()
-  ### This function takes a single species scientific name as input, then grabs 
-  ### all occurrence cells and associated probability per cell
-  
-  ### for Species Map tab: upon selecting a group, update species list choices
-  observeEvent(input$spp_group, 
-    {
-      message('Observed change in spp_group; updating select input')
-      spp_choices <- spp_list %>%
-        filter(spp_group_text == input$spp_group) %>%
-        distinct() %>%
-        .$name %>%
-        sort()
-      updateSelectInput(session, inputId = "species",
-                        choices = spp_choices)
-    }
-  )
-  
-  ### For Species map tab:  upon selecting a species, get map cells
-  observeEvent(
-    {input$species}, 
-    {
-      message('observed change in input$species; getting spp_map dataframe')
-      spp_map$df <- get_spp_map_df(input$species)
-      print(head(spp_map$df))
-    }
-  )
-  
-  ### For Species map tab:  upon a change in the spp_map$df OR the input$show_maps,
-  ### create the map to show AM, IUCN, or Both
-  create_map <- observeEvent(
-    {spp_map$df; input$show_maps}, 
-      ### triggered by change in spp_map$am_rast (which may also include 
-      ### change in iucn_rast) or a change in the selected maps to display
-    {
-      message('observed change in spp_map$df or input$show_maps; creating map')
-      message('input$show_maps = ', input$show_maps, '; input$species = ', input$species)
-      map_rast <- get_rast(spp_map$df, type = input$show_maps)
-      map_obj  <- assemble_map_tmap(map_rast, spp = input$species)
-      
-      spp_map$map <- map_obj
-    }
-  )
-  
-  output$compare_map <- renderPlot({
-    spp_map$map
-  }) 
-  
-  ### For Map Alignment tab
-  output$quad_plot <- renderPlotly({
-    create_quadplot(input$taxa_quad, input$expert_rev)
-  })
-  
-  output$barchart <- renderPlot({
-    create_barchart(input$expert_rev)
-  })
-  
-  output$mini_quad <- renderPlot({
-    create_miniquad(input$species)
-  })
-
-  ### For Coral Depth tab
[…]

### 2017-04-21 shiny_am_iucn/server_fxns.R

@@ -1,519 +1,0 @@
-### server_fxns.R
-library(raster)
-library(stringr)
-# library(maps)
-library(tmap)
-# data(World)
-library(RColorBrewer)
-
-### read data for raster (used to plot species)
-message('Reading in raster info...')
-loiczid_raster <- raster('data/loiczid_raster.tif') %>%
-  setNames('loiczid')
-
-display_extents <- extent(-180, 180, -84, 85)
-
-land <- rgdal::readOGR(dsn   = 'data/ne_110m_land', 
-                       layer = 'ne_110m_land') %>%
-  crop(display_extents)
-
-bathy_rast <- subs(loiczid_raster, 
-                   read_csv('data/bathy_cells.csv'),
-                   by = 'loiczid', which = 'deep') %>%
-  crop(display_extents)
-
-fao_rgn <- rgdal::readOGR(dsn   = 'data/fao_rgns', 
-                          layer = 'fao_rgns_app') %>%
-  crop(display_extents)
-
-# create a blank ggplot theme
-ggtheme_basic <- function(textsize = 10) {
-  theme_bw() +
-  theme_update(panel.grid.minor = element_blank(),
-          panel.grid.major = element_blank(),
-          panel.background = element_blank(),
-          plot.background = element_blank(),
-          panel.border = element_blank(),
-          axis.line = element_blank(),
-          axis.ticks = element_blank(),
-          text = element_text(size = textsize),
-          plot.title = element_text(size = textsize * 1.5))
-}
-
-### read in spp cell files
-### this may be slowing down the initial display of the app... load elsewhere?
-iucn_spp_cells <- read_csv('data/iucn_cells_app.csv', col_types = 'dd')
-am_spp_cells   <- read_csv('data/am_cells_app.csv',   col_types = 'cd')
-
-### read in quad plot files and establish means
-quad_list <- read_csv('data/spp_list_quads_app.csv')
-
-area_align_mean <- mean(quad_list$area_ratio, na.rm = TRUE)
-dist_align_mean <- mean(quad_list$dist_align, na.rm = TRUE)
-
-### read in coral species pre- and post-clipping
-spp_coral_align <- read_csv('data/coral_align_app.csv')
-spp_coral_cells <- read_csv(file.path('data', 'coral_cells_app.csv'))
-
-quad_gp_list <- read_csv('data/spp_gp_quads_app.csv') %>%
-  mutate(expert = FALSE) %>%
[…]

### 2017-04-21 shiny_am_iucn/ui.R

@@ -1,134 +1,0 @@
-#ui.r
-
-source('ui_setup.R')
-
-ui <- navbarPage(title = '',
-                 theme = shinytheme('cerulean'),
-                 
-  tabPanel('Abstract',
-    sidebarPanel(
-      includeMarkdown('pages/article_info.md')
-    ),
-    mainPanel(
-      includeMarkdown('pages/abstract.md')
-    )
-  ),
-  
-  tabPanel('Figures',
-    sidebarPanel(
-      includeMarkdown('pages/article_info.md')
-    ),
-    mainPanel(
-      h4('Figures from manuscript'),
-      h5('Figure 1:'),
-      includeMarkdown('pages/fig1.md'),
-      hr(),
-      h5('Figure 2:'),
-      includeMarkdown('pages/fig2.md'),
-      hr(),
-      h5('Figure 3:'),
-      includeMarkdown('pages/fig3.md'),
-      hr(),
-      h5('Figure 4:'),
-      includeMarkdown('pages/fig4.md'),
-      hr(),
-      h5('Figure 5:'),
-      includeMarkdown('pages/fig5.md')
-    )
-  ),
-  
-  tabPanel('Map alignment',
-    sidebarPanel(
-      includeMarkdown('pages/align_tab_side1.md'),
-      selectInput('taxa_quad', 'Choose a taxon:',
-                  choices = c('all', unique(spp_list$spp_group_text) %>% sort()),
-                  selected = 'all'),
-      radioButtons('expert_rev', label = 'AquaMaps review status',
-                   choices = list('All'             = 'all',
-                                  'Expert reviewed' = 'expert'),
-                   selected = 'all'),
-      includeMarkdown('pages/footer_sidebar.md')
-    ),
-    mainPanel(
-      includeMarkdown('pages/align_tab_main1.md'),
-      plotlyOutput('quad_plot', height = '300px'),
-      hr(),
-      includeMarkdown('pages/align_tab_main2.md'),
-      plotOutput('barchart', height = '300px')
-    )
-  ),
-  
-  tabPanel('Species maps',
-    sidebarPanel(
-      includeMarkdown('pages/map_tab_side1.md'),
-      selectInput('spp_group', 'Select a taxonomic group:',
-                  choices = unique(spp_list$spp_group_text) %>% sort()),
-      selectInput('species', 'Then select a species:',
-                  choices = unique(spp_list$name) %>% sort()),
[…]

### 2017-04-21 ui.R

@@ -1,0 +1,121 @@
+dashboardPage(
+
+  dashboardHeader(title="Ocean Health Index"),
+
+  dashboardSidebar(
+
+    sidebarMenu(
+      id = 'sidebarmenu',
+
+      htmlOutput('ui_commit'),
+
+      menuItem("Explore", tabName='explore',icon=icon("globe",lib='font-awesome'), selected=T),
+
+      if(length(y$scenario_dirs) > 1){
+        menuItem("Compare", tabName='compare',icon=icon("exchange",lib='font-awesome'), selected=F)
+      } else {
+        span()
+      },
+
+      selectInput(
+        'sel_scenario',
+        label    = '0. Choose scenario:',
+        choices  = sort(y$scenario_dirs),
+        selected = y$scenario_dirs[1]),
+
+      conditionalPanel(
+        "input.sidebarmenu == 'compare'",
+
+        uiOutput('ui_sel_scenario_b')),
+
+        selectInput(
+          'sel_type',
+          label='1. Choose type:',
+          choices=c('Input Layer'='input', 'Output Score'='output'),
+          selected='output'),
+
+        conditionalPanel(
+          condition = "input.sel_type == 'output'",
+
+          selectInput(
+            'sel_output_goal',
+            label    = '2. Choose goal:',
+            choices  = output_goals,
+            selected = 'Index'),
+
+          uiOutput('ui_sel_output')),
+
+        conditionalPanel(
+          condition = "input.sel_type == 'input'",
+
+          selectInput(
+            'sel_input_target',
+            label    = '2. Choose target:',
+            choices  = with(layer_targets, setNames(target, target_label))),
+
+          uiOutput('ui_sel_input')),
+
+        selectInput(
+          'sel_rgn',
+          label    = 'Zoom to region:',
+          choices  = c(y$app_title,sort(as.character(rgns@data$rgn_name)))),
+
+        htmlOutput('var_description', class='shiny-input-container') )),
+  #),
+
+  dashboardBody(
+    tabItems(
+
+      tabItem(
+        tabName='explore',
+
+        fluidRow(
+          tabBox(width=12, selected='Map',
+
+            tabPanel(
+              'Map', #title    = 'Map', status='primary', collapsible=T,
+              width=12,
+
+              div(
+                position = 'relative',
+
+                # leaflet map
+                leafletOutput('map1', height = 550),
+
+                # hover text showing info on hover area
+                absolutePanel(
[…]
